### 2014-04-02 Paper.tex
Initial commit.

\documentclass[9pt,oneside]{amsart}
%\usepackage{tweaklist}
\usepackage{url}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{natbib}
\usepackage{float}
\usepackage{pdflscape}

\newcommand*\eg{e.g.\@\xspace}
\newcommand*\Eg{e.g.\@\xspace}
\newcommand*\ie{i.e.\@\xspace}
%\renewcommand{\itemhook}{\setlength{\topsep}{0pt}  \setlength{\itemsep}{0pt}\setlength{\leftmargin}{15pt}}

\title{Ethereum: A Secure Decentralised Compute Platform}
\author{Gavin Wood, Vitalik Buterin}

\begin{document}

\begin{abstract}
The block-chain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin. Each such project can be seen as a simple application on a decentralised, but singleton, compute resource. We can call this paradigm a transactional singleton machine with shared-state.

Ethereum implements this paradigm in a generalised manner. Furthermore it provides a plurality of such resources, each with a distinct state and operating code but able to interact through a message-passing framework with others. We discuss its design, implementation issues, the opportunities it provides and the future hurdles we envisage.
\end{abstract}

\maketitle

\setlength{\columnsep}{20pt}
\begin{multicols}{2}

\section{Introduction}\label{sec:introduction}

With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap. Technological-rooted movements like Bitcoin have demonstrated through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use. This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine. Follow-up systems such as namecoin adapted this original ``currency application'' of the technology into other applications allbeit rather simplistic ones.

Ethereum is a project that attempts to build the generalised technology; technology on which all transaction-based state machine concepts may be built. Moreover it aims to provide to the end-developer a tightly integrated end-to-end system for building software on a hitherto unexplored compute paradigm in the mainstream: a trustful object messaging compute framework.

\subsection{Driving Factors} \label{ch:driving}

Transparency. Open state. User can be guaranteed that backend will do what front-end says it will. Parties can analyse and be wholly confident in outcomes given events.

\subsection{Previous Work} \label{ch:previous}

E language. Smart contracts. Bitcoin, namecoin \&c. Mastercoin.

\section{Proposal} \label{ch:proposal}

\subsection{Overview} \label{ch:overview}

Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. Mathematically:

\begin{equation}
S_{t+1} := f(S_t, T)
\end{equation}

and $f$ is the Ethereum state transition function. In Ethereum, $f$, together with $S$ are considerably more powerful then any existing comparable system; $f$ allows components to carry out arbitrary computation to be done while $S$ allows components to store arbitrary state between transactions.

%where

%\begin{eqnarray}
%S_{t+1} := F(S_t, \mathbb{T}) \\
%\mathbb{T} = [ T_0, T_1, ... ] \\
%F(S, \mathbb{T}) := g(f(f(S, T_0), T_1) ...)
%\end{eqnarray}

Transactions are collated into blocks; we then chain blocks together using a cryptographic hash as a means of reference. Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).

Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, there must be a method of selecting a single block-chain from the block-tree that results. If there is ever a disagreement between parties as to which root-to-leaf path down the block-tree is the `best', then a \textit{fork} occurs and there is a co-existance of multiple states of the system; this is to be avoided at all costs. In order to generate consensus we specify an algorithm to determine the canonical `best' block-chain given the block-tree; we use a simplified version of the GHOST algorithm \cite{ghost}.

\subsection{Blocks, State and Transactions} \label{ch:bst}

\subsubsection{World State} \label{ch:state}

The world-state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP). Though not stored on the block chain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}). The trie requires a simple database back-end that maintains a mapping of bytearrays to bytearrays; we name this underlaying database the state database. This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state. Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly. Since we store all such root hashes in the block chain, we are able to trivially revert to old states.

The account state comprises the first two, and potentially the last two, of the following fields:

\begin{description}
\item[nonce] A scalar value equal to the number of transactions sent from this address, or, in the case of contract accounts, the number of contract-creations made by this account.
\item[balance] A scalar value equal to the number of wei owned by this address.
\item[stateRoot] A 256-bit hash equal to the root node of a further trie structure that encodes the storage contents of the contract. Though a separate data structure, this trie is still stored using the same underlying state database. This trie takes the form as a simple mapping between domains of 256-bit values.
\item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address reveive a call; it is immutable and thus, unlike all other fields, cannot be changed after construction. All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
\end{description}

If the latter two fields are missing, the node represents a simple (non-contract) account: it maintains no state and executes no code on receipt of a transaction.

\subsubsection{Transaction} \label{ch:transaction}

A transaction is a single cryptographically signed instruction sent by an actor external to Ethereum. e.g. This might be from a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server. There are two types of transactions: those which result in message calls and those which result in the creation of new contracts. Both types specify a number of common fields:

\begin{description}
\item[nonce] A scalar value equal to the number of transactions sent by the sender.
\item[value] A scalar value equal to the number of wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created contract's account.
\item[gasPrice] A scalar value equal to the number of wei to be paid per unit of gas for all computation costs incurred as a result of the execution of this transaction.
\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction. This is paid up-front, before any computation is done and may not be topped-up later.
\item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction.
\item[v, r, s] Three scalar values corresponding to the signature of the transaction and used to determine the sender of the transaction. See Appendix \ref{app:signing}.
\end{description}

Additionally, a contract-creation transaction contains:

\begin{description}
\item[body] An unlimited size byte array specifying the EVM-code for the contract.
\item[init] An unlimited size byte array specifying the EVM-code for the contract initialisation procedure.
\end{description}

Both \textbf{body} and \textbf{init} are EVM-code fragments; \textbf{body} is executed each time the contract account receives a message call (either through a transaction or due to the internal execution of code). \textbf{init} is a piece of code executed only once and at contract creation. It gets discarded immediately.

Whereas a message call transaction contains:

\begin{description}
\item[data] An unlimited size byte array specifying the input data of the message call.
\end{description}

\subsubsection{Block} \label{ch:block}

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}) together with a set of transactions and a set of other block headers that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}). The block header contains several pieces of information:

\begin{description}
\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety.
\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block.
\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred.
\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied.
\item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block.
\item[difficulty] A scalar value corresonding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp.
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception.
\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block.
\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block.
\end{description}

The difficulty, $d$, is defined as:

\begin{equation}
d' :=  \begin{cases}
2^{22} & \quad \text{if genesis block}\\
d' + \frac{d'}{1024} & \quad \text{if $t' < t + 42$}\\
d' - \frac{d'}{1024} & \quad \text{otherwise}\\
\end{cases}
\end{equation}

Where $d'$ is the previous block's difficulty, $t'$ is the prvious block's timestamp and $t$ is the current block's timestamp.

The nonce, $n$, must satisfy:

\begin{equation}
\mathcal{P}(n) < \frac{d}{2^{256}}
\end{equation}

Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}). The other two parts to a block are simply a list of the transactions (as defined previously) and a list of uncle headers (of the same format as above).

\subsection{Gas and Payment} \label{ch:payment}

In order to avoid issues of network abuse and to sidestep the inevitable questions stemming from Turing completeness, all programmable computation in Ethereum is subject to fees. The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation). Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing contract storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.

Every transaction has a specific amount of gas associated with it: \textbf{gasLimit}. This is the amount of gas which is implicitly purchased from the sender's account balance. The purchase happens at the according \textbf{gasPrice}, also specified in the transaction. The transaction is considered invalid if the account balance cannot support such a purchase. It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender account. Gas does not exist outside of the execution of a transaction. Thus for trusted contracts, a relatively high gas limit may be set and left alone.

Any ether paid for gas by the sender that is not refunded is delivered to the miner. Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose. A higher gas price on a transaction will therefore cost the sender more in terms of ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners. Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer. Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.

%\subsubsection{Determining Computation Costs}

\subsection{Transaction Execution} \label{ch:execution}

The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $f$ as specified before. It is assumed that any transactions executed first pass the initial tests of intrinsic validity. These include:

\begin{enumerate}
\item The transaction signature is valid;
\item the transaction nonce is valid (equivalent to the sender account's current nonce);
\item the gas limit is no smaller that the intrinsic gas, $i$, used by the transaction;
\item the sender account balance contains at least the cost, $c$, (in wei) required in up-front payment;
\end{enumerate}

where:

\begin{eqnarray}
i \equiv & \begin{cases}
  |\mathbf{d}|.G_{data} + G_{call} & \quad \text{call} \\
  (|\mathbf{i}| + |\mathbf{b}|).G_{data} + G_{create} & \quad \text{create} \\
\end{cases} \\
c \equiv & l.p + v
\end{eqnarray}

where $|\mathbf{d}|$, $|\mathbf{i}|$ and $|\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation \& body EVM-code and $G$ is defined in Appendix \ref{app:fees}.

The execution of a valid transaction begins with an irrevocable change made to the state: the sender's nonce is incremented by one and the balance is reduced by the cost, $c$. The gas available for the proceeding computation, $g$, is defined as $l - i$. The proceeding computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate, thus we increase the sender's balance by $g.p$.

\subsubsection{Create} \label{ch:create}

There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($o$), available gas ($g$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.

The address of the new contract is defined as being the leftmost 160 bits of the SHA3 hash of RLP encoding of the structure \texttt{[ sender, nonce ]}. In in the unlikely event that the addresses is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at, thus:

\begin{eqnarray}
a' := \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
a := \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \emptyset
\end{eqnarray}

where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the leftmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\emptyset$ if none exists.

The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the new state becomes $\mathbb{S}'$:

\begin{equation}
\mathbb{S}'[r] := ( 0, v, \empty, \mathcal{S}(\mathbf{b}) )
\end{equation}

The state database is altered to include the pair $(\mathcal{S}(\mathbf{b}), \mathbf{b})$. Finally, the contract is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}). Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made.

Code execution depletes gas; thus it may exit before the code has come to a natural halting state. In this exceptional case we say an Out-of-Gas exception has occured and the state database is reverted to the point immediately prior to the account initialisation. All available gas is used and no gas is refunded to the originator. If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.

\subsubsection{Message Call} \label{ch:call}

In the case of executing a message call, several parameters are required: sender ($s$), recipient ($r$), available gas ($g$), value ($v$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call. Message calls also have intrinsic return data; the call's output byte array $\mathbf{r}$. This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case the return data is used.

In executing a message call, the account balance of the recipient is increased by the value passed, thus $\mathbb{S}$ becomes replaced with $\mathbb{S}'$:

\begin{equation}
\mathbb{S}'[r]_{balance} := \mathbb{S}[r]_{balance} + v
\end{equation}

In the case that the recipient contains code to be executed (i.e. that the account is a contract), then the contract's body code, to be found in the state database with key $\mathbb{S}[r]_{code}$, is executed according to the execution model (see section \ref{ch:model}). Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution. Unlike with contract creation,

\subsection{Execution Model} \label{ch:model}

The original caller (i.e. the original transaction's txsender) pre-pays for a specific amount of 'GAS'. What was baseFee now describes the ETH -> GAS price. Gas is the internal currency of the transaction. Execution operations deplete the GAS supply, as per the original ratios. Once the transaction is finished, any remaining GAS is reconverted back to ETH and refunded.

Storage remains 256-bit address, 256-bit data.

Word size: 256-bit (i.e. stack item size 256-bit).

Memory is byte array (256-bit address, 8-bit data); memoryFee is proportional to greatest index that has ever been referenced in this instance and thus it's highly unlikely addresses will ever go above 32-bit. Whenever a higher memory index is used, the fee difference to take it to the higher usage from the original (lower) usage is charged. MEMSIZE is initially just the size of the input data. Note: MSTORE and MLOAD increase the highest-accessed index to their target index + 31.

Given an instruction, it is possible to calculate the gas cost of executing it as follows:

\begin{itemize}
\item Unless covered by another rule below, an operation costs < STEPGAS > gas
\item SHA3 costs $G_{sha3}$ gas
\item SLOAD costs $G_{sload}$ gas
\item BALANCE costs $G_{balance}$ gas
\item SSTORE costs $d.G{sstore}$ gas where:
\begin{itemize}
\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
\item $d = 1$ otherwise.
\end{itemize}
\item CALL costs $G_{call} + g$, where $g$ is the quantity of gas provided; some gas may later be refunded.
\item CREATE costs $G_{create} + g$ gas, where $g$ is the quantity of gas provided.
\item When you read or write memory with MSTORE, MLOAD, RETURN, SHA3, CALLDATA or CALL, enlarge the memory so that all bytes now fit in it (memory must always be a whole number of 32-byte words). Suppose that the highest previously accessed memory index is $m$, and the new index is $n$. If $(n + 31) / 32 > (m + 31) / 32$, add an additional $((n + 31) / 32 - (m + 31) / 32) * G_{memory}$ gas to the cost.
\end{itemize}

\begin{enumerate}
\item Let $\mathbb{S}' := \mathbb{S}$, the current state.
\item Let $PC := 0$, the program counter.
Let it be known that all storage operations operate on TO's state.
Note that the memory is empty, thus (last used index + 1) = 0.
\item Set TXDATA to the first INSIZE bytes of memory starting from INOFFSET in the caller memory.
\item Repeat
\begin{itemize}
\item Calculate the cost of the current instruction (see below), set to $C$.
\item If the instruction is invalid or STOP, goto step 6.
\item If $GAS < C$ then $GAS := 0$; $S := ROLLBACK$ and evaluation finishes, returning false. 
\item $GAS := GAS - C$
\item Apply the instruction.
\end{itemize}
Until an operation execution error has occured or the instruction is STOP, RETURN or SUICIDE.
\item If the output data of the call (R) is specified (through RETURN), then let the OUTSIZE bytes of caller memory beginning at OUTOFFSET.
\item Returns true.
\end{enumerate}

When a contract address receives a transaction, a virtual machine is initiated with the contract's state.

\subsubsection{Terms}

There exists a stack of variable size that stores 256-bit values at each location. (Note: most instructions operate solely on the stack altering its state in some way.)

T'[i] is the ith item counting down from the top of the pre-stack (i.e. the stack immediately after instruction execution), with the top of the stack corresponding to i == 0.

T[i] is the ith item counting down from the top of the post-stack (i.e. the stack immediately prior to instruction execution), with the top of the stack corresponding to i == 0.

The exists a permanent store addressable by a 256-bit value that stores 256-bit values at each location.

S'[i] is the permanent store (sometimes refered to as 'state' or 'store') of the VM at index i counting from zero PRIOR to instruction execution.

S[i] is the permanent store (sometimes refered to as 'state' or 'store') of the VM at index i counting from zero AFTER to instruction execution.

The exists a temporary store addressable by a 256-bit value that stores 256-bit values at each location.

M'[i] is the temporary store (sometimes refered to as 'memory') of the VM at index i counting from zero PRIOR to instruction execution.

M[i] is the temporary store (sometimes refered to as 'memory') of the VM at index i counting from zero AFTER instruction execution.

C is the code, a byte sequence.

PC' is the program counter PRIOR to instruction execution.

PC is the program counter AFTER instruction execution.

FEE(I, S', P', D) is the fee associated with the execution of instruction I with a machine of stack S', permanent store P' and which has already completed D operations.

It is defined as F where:

TODO: No it's not!

IF I == SSTORE AND P[ S'[0] ] != 0 AND S'[1] == 0 THEN
    F = S + dataFee - memoryFee
IF I == SSTORE AND P[ S'[0] ] == 0 AND S'[1] != 0 THEN
    F = S + dataFee + memoryFee
IF I == SLOAD
    F = S + dataFee
IF I == EXTRO OR I == BALANCE
    F = S + extroFee
IF I == MKTX
    F = S + txFee
IF I == SHA256 OR I == SHA3 OR I == RIPEMD160 OR I == ECMUL OR I == ECADD
    OR I == ECSIGN OR I == ECRECOVER OR I == ECVALID THEN
    F = S + cryptoFee

Where:

B[ A ] is the balance of the address given by A, with A interpreted as an address.

ADDRESS is the address of the present contract.

10. b. Initial Operation

STEPSDONE := 0
PC' := 0
FOR ALL i: M' is initialised to the empty byte sequence.
T' is initialised such that its cardinality is zero (i.e. the stack starts empty).
S' must be equal to the value of P when the previous execution halted.

10. c. General Operation

The given steps are looped:
1. Execution halts if B[ ADDRESS ] < F( C[PC'], S', P', STEPSDONE )
2. B[ ADDRESS ] := B[ ADDRESS ] - F( P'[PC'], S', P', STEPSDONE )
3. The operation given by P'[PC'] determines PC, P, T, S.
4. PC' := PC; P' := P; T' := T; S' := S; STEPSDONE := STEPSDONE + 1

10. d. VM Operations

Summary line:
<Op-code>: <Mnemonic> -<R> +<A>

If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are PUSH, JUMP and JUMPI.

The cardinality of T (i.e. size of the stack) is altered by A - R between T \& T', by adding or removing items as necessary from the front.

Where:
R: The minimal cardinality of the stack for this instruction to proceed. If this is not achieved then the machine halts with a stack underflow exception. (Note: In some cases of some implementations, this is also the number of values "popped" from the implementation's stack during the course of instruction execution.)
(A - R): The net change in cardinality of the stack over the course of instruction execution.

$\forall$ i: if T[i] is not defined explicitly, then it must be assumed T[i] := T'[i + R - A] where i + R >= A.
$\forall$ i: if M[i] is not defined explicitly, then it must be assumed M[i] := M'[i].
$\forall$ i: if S[i] is not defined explicitly, then it must be assumed S[i] := S'[i].

The expression (COND ? ONE : ZERO), where COND is an expression and ONE and ZERO are both value placeholders, evaluates to ONE if COND is true, and ZERO if COND is false. This is similar to the C-style ternary operator.

When a 32-byte machine datum is interpreted as a 160-bit address or hash, the rightwards 20 bytes are taken (i.e. the low-order bytes when interpreting the data as Big-Endian).

++ is the concatenation operator; all operands are byte arrays (mostly 32-byte arrays here, since that's the size of the VM's data \& address types).

LEFT\_BYTES(A, n) returns the array bytes comprising the first (leftmost) n bytes of the 32 byte array A, which can be considered equivalent to a single value in the VM.

\subsection{Block Finalisation} \label{ch:finalisation}

\subsection{Mining Proof-of-Work} \label{ch:pow}

\subsection{Block-tree to Block-chain} \label{ch:ghost}

\section{User Interface}

TODO

\section{Discussion} \label{ch:discussion}

\subsection{Examples} \label{ch:examples}

\section{Conclusion} \label{ch:conclusion}

\subsection{Further Work} \label{ch:further}

\bibliography{Biblio}
\bibliographystyle{plainnat}

\appendix

\section{Terminology}

\begin{description}
\item[External Actor] A person or other entity able to interface to an Ethereum node, but external to the world of Ethereum. It can interact with Ethereum through depositing signed Transactions and inspecting the block-chain and associated state. Has one (or more) intrinsic Accounts.

\item[Address] A 160-bit code used for identifying Accounts.

\item[Account] Accounts have an intrinsic balance and transaction count maintained as part of the Ethereum state. They are owned either by External Actors or intrinsically (as an identity) an Autonomous Object within Ethereum. If an Account identifies an Autonomous Object, then Ethereum will also maintain a Storage State particular to that Account. Each Account has a single Address that identifies it.

\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the block-chain.

\item[Autonomous Object] A virtual object existant only within the hypothetical state of Ethereum.  Has an intrinsic address. Incorporated only as the state of the storage component of the VM.

\item[Storage State] The information particular to a given Autonomous Object that is maintained between the times that it runs.

\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts in a perfectly trusted way, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.

\item[Message Call] The act of passing a message from one Account to another. If the destination account is an Autonomous Object, then the VM will be started with the state of said Object and the Message acted upon. If the message sender is an Autonomous Object, then the Call passes any data returned from the VM operation.

\item[Gas] The fundamental network cost unit. Paid for exclusively by Ether (as of PoC-4), which is converted freely to and from Gas as required. Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and miners are free to ignore Transactions whose Gas price is too low.

\item[Contract] Synonym for Autonomous Object used for non-technical audiences.

\item[Object] Synonym for Autonomous Object.

\item[App] An end-user-visible application hosted in the Ethereum Browser.

\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (ala Chrome) that is able to host sandboxed applications whose back-end is purely on the Ethereum protocol.

\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for a contract's program code.

\item[EVM Code] The bytecode that the EVM can natively execute.

\item[EVM Assembly] The human-readable form of EVM-code.

\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts.

\end{description}

\section{Recursive Length Prefix}\label{app:rlp}

\section{Modified Merkle Patricia Tree}\label{app:trie}

\section{Hex-Prefix Encoding}\label{app:hexprefix}

\section{Formal Specification of Structures}

Though RLP is data-agnostic, it does specify a canonical representation for integer quantities. It is big-endian with no leading zero-bytes. Thus for elements than feasibly can be stored as integers, it becomes important to specify whether they should adhere to this canonical representation or be left in some other (perhaps more 'native') format.

In the case of counts, balances, fees and amounts of wei, the canon-integer form must be used when storing in RLP. We call these INTs.

In the case of hashes (256-bit or 160-bit), user-visible strings and specialised byte-arrays (e.g. hex-prefix notation from the trie), they should be stored as unformatted byte-array data and not altered into some other form. We call these BINs.

When interpreting RLP data, clients are required to consider non-canonical INT fields in the same way as otherwise invalid RLP data and dismiss it completely.

Specifically:

for the Block header:
\begin{verbatim}
[
  parentHash: BIN (256-bit),
  unclesHash: BIN (256-bit),
  coinbase: BIN (160-bit),
  stateRoot: BIN (256-bit),
  transactionsHash: BIN (256-bit),
  difficulty: INT,
  timestamp: INT,
  extraData: BIN (typically 256-bit),
  nonce: BIN (256-bit)
]
\end{verbatim}

(note: 'nonce', the last element, refers to a hash here and so is binary)

for entries in the State trie for normal addresses:
\begin{verbatim}
[
  balance: INT,
  nonce: INT
]
\end{verbatim}

and for contract addresses:
\begin{verbatim}
[
  balance: INT,
  nonce: INT,
  storageRoot: BIN (256-bit),
  codeHash: BIN (256-bit)
]
\end{verbatim}

(note: 'nonce', the second element, refers to a tx-count here and so is integer)

for message call transactions:

\begin{verbatim}
[
  nonce: INT,
  value: INT,
  gasPrice: INT,
  gas: INT,
  recvAddr: BIN (160-bit),
  data: BIN (arbitrary length),
  v: INT,
  r: INT,
  s: INT
]
\end{verbatim}

or, for contract creation transactions,

\begin{verbatim}
[
  nonce: INT,
  value: INT,
  gasPrice: INT,
  gas: INT,
  0: BIN (160-bit),
  code: BIN (arbitrary length),
  init: BIN (arbitrary length),
  v: INT,
  r: INT,
  s: INT
]
\end{verbatim}

(note: 'nonce', the first element, refers to a tx-count here and so is integer)

The nonce in the transaction refers to the total amount of transactions sent from the address up until that moment in time.

for blocks, there are no immediate data field, but lists:

\begin{verbatim}
[
  blockHeader: [...]
  uncleList: [ [...], [...], ... ]
  txList: [ [...], [...], ... ]
]
\end{verbatim}

Uncle-blocks contain only the uncle's header.

\section{The State DB}\label{app:state}

\section{Signing Transactions}\label{app:signing}

\section{Virtual Machine Specification}\label{app:vm}

0s: arithmetic operations

0x00: STOP -0 +0
Halts execution.
Any gas left over gets returned to caller (or in the case of the top-level call, the sender converted back to ETH).
0x01: ADD -2 +1
S[0] := S'[0] + S'[1]
0x02: MUL -2 +1
S[0] := S'[0] * S'[1]
0x03: SUB -2 +1
S[0] := S'[0] - S'[1]
0x04: DIV -2 +1
S[0] := S'[0] / S'[1]
0x05: SDIV -2 +1
S[0] := S'[0] / S'[1]
S'[0] \& S'[1] are interpreted as signed 256-bit values for the purposes of this operation.
0x06: MOD -2 +1
S[0] := S'[0] % S'[1]
0x07: SMOD -2 +1
S[0] := S'[0] % S'[1]
S'[0] \& S'[1] are interpreted as signed 256-bit values for the purposes of this operation.
0x08: EXP -2 +1
S[0] := S'[0] + S'[1]
0x09: NEG -1 +1
S[0] := -S'[0]
0x0a: LT -2 +1
S[0] := S'[0] < S'[1] ? 1 : 0
0x0b: GT -2 +1
S[0] := S'[0] > S'[1] ? 1 : 0
0x0c: EQ -2 +1
S[0] := S'[0] == S'[1] ? 1 : 0
0x0d: NOT -1 +1
S[0] := S'[0] == 0 ? 1 : 0

10s: bit operations

0x10: AND -2 +1
S[0] := S'[0] AND S'[1]
0x11: OR -2 +1
S[0] := S'[0] OR S'[1]
0x12: XOR -2 +1
S[0] := S'[0] XOR S'[1]
0x13: BYTE -2 +1
S[0] := S'[0]th byte of S'[1]
if S'[0] < 32
S[0] := 0
otherwise
for Xth byte, we count from left - 0th is therefore the leftmost (most significant in BE) byte.

20s: crypto opcodes

0x20: SHA3 -2 +1
S[0] := SHA3( T'[ S'[0] ++ ... ++ (S'[0] + S'[1]) ])

30s: closure state opcodes

0x30: ADDRESS -0 +1
S[0] := ADDRESS
i.e. the address of this closure.
0x31: BALANCE -1 +1
S[0] := B[ S'[0] ]
i.e. the balance of this closure.
0x32: ORIGIN -0 +1
S[0] := A
Where A is the address of the account that made the original transaction leading to the current closure and is paying the fees
0x33: CALLER -0 +1
S[0] := A
Where A is the address of the object that made this call.
0x34: CALLVALUE -0 +1
S[0] := V
Where V is the value attached to this call.
0x35: CALLDATALOAD -1 +0
S[0] := D[ S'[0] ... (S'[0] + 31) ]
Where D is the data attached to this call (as a byte array).
Any bytes that are out of bounds of the data are defined as zero.
0x36: CALLDATASIZE -0 +1
S[0] := DS
Where DS is the number of bytes of data attached to this call.
0x37: GASPRICE -0 +1
S[0] := V
Where V is the current gas price (né fee multiplier).

40s: block operations

0x40: PREVHASH -0 +1
S[0] := H
Where H is the SHA3 hash of the previous block.
0x41: COINBASE -0 +1
S[0] := A
Where A is the coinbase address of the current block.
0x42: TIMESTAMP -0 +1
S[0] := T
Where T is the timestamp of the current block (given as the Unix time\_t when this block began its existence).
0x43: NUMBER -0 +1
S[0] := N
Where N is the block number of the current block (counting upwards from genesis block which has N == 0).
0x44: DIFFICULTY -0 +1
S[0] := D
Where D is the difficulty of the current block.
0x45: GASLIMIT -0 +1
S[0] := L
Where L is the total gas limit of the current block. Always $10^6$.

50s: stack, memory, storage and execution path operations

0x51: POP -1 +0
0x52: DUP -1 +2
S[0] := S'[0]
0x53: SWAP -2 +2
S[0] := S'[1]
S[1] := S'[0]
0x54: MLOAD -1 +1
S[0] := T'[ S'[0] ... S'[0] + 31 ]
0x55: MSTORE -2 +0
T[ S'[0] ... S'[0] + 31 ] := S'[1]
0x56: MSTORE8 -2 +0
T[ S'[0] ... S'[0] + 31 ] := S'[1] \& 0xff
0x57: SLOAD -1 +1
S[0] := P'[ S'[0] ]
0x58: SSTORE -2 +0
P[ S'[0] ] := S'[1]
0x59: JUMP -1 +0
PC := S'[0]
0x5a: JUMPI -2 +0
PC := S'[1] == 0 ? PC' : S'[0]
0x:5b PC -0 +1
S[0] := PC
0x5c: MSIZE -0 +1
S[0] = sizeof(T)
0x5d: GAS
S[0] := G
Where G is the amount of gas remaining after executing the opcode.

60s \& 70s: push

0x60: PUSH1 0 +1
S[0] := C[ PC' + 1 ]
PC := PC' + 2
0x61: PUSH2 0 +1
S[0] := C[ PC' + 1 ] ++ C[ PC' + 2 ]
PC := PC' + 3
...
0x7f: PUSH32 0 +1
S[0] := C[ PC' + 1 ] ++ C[ PC' + 2 ] ++ ... ++ C[ PC' + 32 ]
PC := PC' + 33

f0s: closure-level operations

0xf0: CREATE -5 +1
Immediately creates a contract where:
The endowment is given by S'[0]
The body code of the eventual closure is given by T'[ S'[1] ... ( S'[1] + S'[2] - 1 ) ]
The initialisation code of the eventual closure is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
(Thus the total number of bytes of the transaction data is given by S'[2] + S'[4].)
S[0] = A
where A is the address of the created contract or 0 if the creation failed.
Fees are deducted from the sender balance to pay for enough gas to complete the operation (i.e. contract creation fee + initial storage fee). If there was not enough gas to complete the operation, then all gas will be deducted and the operation fails.
0xf1: CALL -7 +1
Immediately executes a call where:
The recipient is given by S'[0], when interpreted as an address.
The value is given by S'[1]
The gas is given by S'[2]
The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
(Thus the number of bytes of the transaction is given by S'[4].)
The output data of the call is given by T[ S'[5] ... ( S'[5] + MIN( S'[6], S[0] ) - 1 ) ]
If 0 gas is specified, transfer all gas from the caller to callee gas balance, otherwise transfer only the amount given. If there isn't enough gas in the caller balance, operation fails.
If the value is less than the amount in the caller's balance then nothing is executed and the S'[2] gas gets refunded to the caller.
See Appendix A for execution.
Add any remaining callee gas to the caller's gas.
S[0] = R
where R = 1 when the instrinsic return code of the call is true, R = 0 otherwise.
NOT YET: POST -5 +1
Registers for delayed execution a call where:
The recipient is given by S'[0], when interpreted as an address.
The value is given by S'[1]
The gas to supply the transaction with is given by S'[2] (paid for from the current gas balance)
The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
(Thus the number of bytes of the transaction is given by S'[4].)
Contract pays for itself to run at a defered time from its own GAS supply. The miner has no choice but to execute.
NOT YET: ALARM -6 +1
Registers for delayed execution a call where:
The recipient is given by S'[0], when interpreted as an address.
The value is given by S'[1]
The gas (to convert from ETH at the later time) is given by S'[2]
The number of blocks to wait before executing is S'[3]
The input data of the call is given by T'[ S'[4] ... ( S'[4] + S'[5] - 1 ) ]
(Thus the number of bytes of the transaction is given by S'[5].)
Total gas used now is S'[3] * S'[5] * deferFee.
Contract pays for itself to run at the defered time converting given amount of gas from its ETH balance; if it cannot pay it terminates as a bad transaction. TODO: include baseFee and allow miner freedom to determine whether to execute or not. If not, the next miner will have the chance.
0xf2: RETURN -2 +0
Halts execution.
R := T[ S'[0] ... ( S'[0] + S'[1] - 1 ) ]
Where the output data of the call is specified as R.
Any gas left over gets returned to caller (or in the case of the top-level call, the sender converted back to ETH).
0xff: SUICIDE -1 +0
Halts execution.
FOR ALL i: IF P[i] NOT EQUAL TO 0 THEN B[ S'[0] ] := B[ S'[0] ] + memoryFee
B[ S'[0] ] := B[ S'[0] ] + B[ ADDRESS ]
Removes all contract-related information from the Ethereum system.

\section{Low-level Lisp-like Language}\label{app:lll}

\section{Wire Protocol}\label{app:wire}

\section{Peer Strategy}\label{app:peers}

\section{Genesis Block}\label{app:genesis}

The header of the genesis block is 9 items, and is specified thus:

$[z_{256}, \mathcal{S}(\mathcal{R}([])), z_{160}, stateRoot, \mathcal{S}(\mathcal{R}([])), 2^{22}, 0, \text{`'}, 42]$

Where:

$z_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes.

$z_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes.

$2^{22}$ refers to the difficulty.

0 refers to the timestamp (the Unix epoch).

`' refers to the extradata, an empty byte array.

$\mathcal{S}(\mathcal{R}([]))$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.

$stateRoot$ refers to the state root.

\section{Fee Schedule}\label{app:fees}

Constants:

* STEPGAS = 1

* SHA3GAS = 20

* SLOADGAS = 20

* SSTOREGAS = 100

* BALANCEGAS = 20

* CREATEGAS = 100

* CALLGAS = 20

* MEMORYGAS = 1

* TXDATAGAS = 5 [not used in the VM]

\end{multicols}

\end{document}

### 2014-04-02 Paper.tex
More stuff.

@@
 
+\subsubsection{Currency}
+
+Ether, wei.
+
 \subsubsection{World State} \label{ch:state}
@@
 
-Where $d'$ is the previous block's difficulty, $t'$ is the prvious block's timestamp and $t$ is the current block's timestamp.
+Where $d'$ is the previous block's difficulty, $t'$ is the prvious block's timestamp and $t$ is the current block's timestamp. This results in an expected period between blocks of roughly 60 seconds.
 
@@
 \begin{eqnarray}
-i \equiv & \begin{cases}
+i & \equiv & \begin{cases}
   |\mathbf{d}|.G_{data} + G_{call} & \quad \text{call} \\
@@
 \end{cases} \\
-c \equiv & l.p + v
+c & \equiv & l.p + v
 \end{eqnarray}
@@
 \begin{eqnarray}
-a' := \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
-a := \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \emptyset
+a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
+a & \equiv & \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \emptyset
 \end{eqnarray}
@@
 
+\subsection{Block-tree to Block-chain} \label{ch:ghost}
+
+The canonical block chain is a path from root to leaf through the entire block tree.
+In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
+Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
+The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
+This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
+
+This scheme notably ignores so-called \textit{stale} blocks: valid, mined blocks, which were propogated too late into the network and thus were beaten to network consensus by a sibling block (one with the same parent).
+Such blocks become more common as the network propogation time approaches the ideal inter-block time.
+However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to butress the more popular block-chain making it a stronger choice over less popular (though potentially longer) competitors.
+
+This increases overall network security by making it much harder for an adversary to silently mine a canonical block chain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of reversing existing blocks and the transactions within.
+
+In order to validate the extra computation, a given block $B$ may include the block headers from any known uncle blocks (i.e.
+blocks whose parent is equivalent to the grandparent of $B$).
+Since a block header includes the nonce, a proof-of-work, then the header alone is enough to validate the computation done.
+Any such blocks contribute toward the total computation or \textit{total difficulty} of a chain that includes them.
+To incentivise computation and inclusion, a reward is given both to the miner of the stale block and the miner of the block that references it.
+
+Thus we define the total difficulty of block $B$ recursively as:
+
+\begin{eqnarray}
+B_{td} & \equiv & P_{td} + B_d + \sum\limits_{U \in B_\mathbb{U}} U_d \\
+P & \equiv & B_P
+\end{eqnarray}
+
+As such given a block $B$, $B_td$ is its total difficulty, $B_P$ is its parent block, $B_d$ is its difficulty and $B_\mathbb{U}$ is its set of uncle blocks.
+
 \subsection{Block Finalisation} \label{ch:finalisation}
+
+Validate uncles.
+
+Reward everything.
+
+Compute proof-of-work.
 
@@
 
-\subsection{Block-tree to Block-chain} \label{ch:ghost}
+The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
+It is utilised to enforce the block chain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
+However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
+
+For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
+The requirement of or reward from specialised and uncommon hardware should be minimised.
+This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to ether at roughly the same rate for anyone around the world.
+
+Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
+Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
+
+One plague of the Bitcoin world is ASICs.
+These are specialised pieces of compute hardware that exist only to do a single task.
+In Bitcoin's case the task is the SHA256 hash function.
+While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
+Because of this, a proof-of-work function that is ASIC-resistant (i.e.
+difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
+
+Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
+engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
+
+More formally, the proof-of-work function takes the form of $\mathcal{P}$:
+
+\begin{equation}
+\mathcal{P}(H, n) < \frac{H_d}{2^{256}}
+\end{equation}
+
+Where $H$ is the new block's header \textit{without} the nonce entry; $H_d$ is the difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
+
+As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
+
+\subsubsection{PoC Series}
+
+For the proof-of-concept series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder.
+
+It is defined as $\mathcal{P}$:
+
+\begin{equation}
+\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \bigoplus n))
+\end{equation}
+
+where:
+$\mathcal{R}(H)$ is the RLP encoding of the $H$;
+$\mathcal{S}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
+$n$ is the nonce, a series of 32 bytes;
+$\bigoplus$ is the series concatenation operator;
+$\mathcal{B}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
+
+\subsubsection{Release Series}
+
+For the release series, we use a more complex proof-of-work. This has yet to be formally defined, but involves two components; firstly,
 

### 2014-04-02 Paper.tex
More beginning stuff.

@@
 
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. Mathematically:
+Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. Formally:
 
@@
 and $f$ is the Ethereum state transition function.
-In Ethereum, $f$, together with $S$ are considerably more powerful then any existing comparable system; $f$ allows components to carry out arbitrary computation to be done while $S$ allows components to store arbitrary state between transactions.
-
-%where
-
-%\begin{eqnarray}
-%S_{t+1} := F(S_t, \mathbb{T}) \\
-%\mathbb{T} = [ T_0, T_1, ... ] \\
-%F(S, \mathbb{T}) := g(f(f(S, T_0), T_1) ...)
-%\end{eqnarray}
+In Ethereum, $f$, together with $S$ are considerably more powerful then any existing comparable system; $f$ allows components to carry out arbitrary computation to be done, while $S$ allows components to store arbitrary state between transactions.
 
@@
 Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
+They also punctuate the transaction series with incentives for nodes to \textit{mine}.
+This incentivisation takes places as a state-transition function, adding value to a nominated account.
 
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, there must be a method of selecting a single block-chain from the block-tree that results.
-If there is ever a disagreement between parties as to which root-to-leaf path down the block-tree is the `best', then a \textit{fork} occurs and there is a co-existance of multiple states of the system; this is to be avoided at all costs.
-In order to generate consensus we specify an algorithm to determine the canonical `best' block-chain given the block-tree; we use a simplified version of the GHOST algorithm \cite{ghost}.
+Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other through a cryptographically secure proof.
+This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
+
+Formally, we expand to:
+
+\begin{eqnarray}
+S_{t+1} & \equiv & F(S_t, \mathbb{T}) \\
+\mathbb{T} & \equiv & [ T_0, T_1, ... ] \\
+F(S, \mathbb{T}) & \equiv & g(f(f(S, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $g$ is the block-finalisation state transition function (a function that rewards a nominated party); $\mathbb{T}$ is the series of transactions included in this block; and $F$ is the block-level state-transition function.
+
+\subsubsection{Value}
+
+In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
+To address this issue, Ethereum has an intrinsic currency, Ether, known also as ETH.
+The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted is the Wei.
+One Ether is defined as being $10^{18}$ Wei.
+There exist other subdenominations of Ether:
+
+\par\bigskip
+\begin{center}
+\begin{tabular}{lr}
+\toprule
+Multiplier & Name \\
+\midrule
+$10^0$ & Wei \\
+$10^{12}$ & Szabo \\
+$10^{15}$ & Finney \\
+$10^{18}$ & Turing \\
+\bottomrule
+\end{tabular}
+\end{center}
+\par
+
+Throughout the document, any reference to value, in the context of Ether, currency, a balance or a payment, should assumed to be counted in Wei.
+
+\subsubsection{Which History?}
+
+Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block the resultant structure is necessarily a tree of blocks.
+In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the block-chain, there must be an agreed-upon scheme.
+If there is ever a disagreement between nodes as to which root-to-leaf path down the block-tree is the `best' block-chain, then a \textit{fork} occurs.
+
+This would mean that past a given point in time (block), there is a temporal co-existance of multiple states of the system: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
+This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.
+
+The scheme we use in order to generate consensus is a simplified version of the GHOST algorithm \cite{ghost}. This process is described in detail in section \ref{ch:ghost}.
 
@@
 
-\subsubsection{Currency}
-
-Ether, wei.
+Having introduced the basic concepts behind Ethereum, we will discuss the meaning of a transaction, a block and the state in more detail.
 
@@
 
-For the proof-of-concept series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder.
+For the proof-of-concept series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder. It utilises the SHA3 hash function to
 
-It is defined as $\mathcal{P}$:
+It is formally defined as $\mathcal{P}$:
 

### 2014-04-02 Paper.tex
More.

@@
 
-The difficulty, $d$, is defined as:
+The difficulty of a block $H$, is defined as $H_d$:
 
 \begin{equation}
-d' :=  \begin{cases}
+H_d :=  \begin{cases}
 2^{22} & \quad \text{if genesis block}\\
-d' + \frac{d'}{1024} & \quad \text{if $t' < t + 42$}\\
-d' - \frac{d'}{1024} & \quad \text{otherwise}\\
+H'_d + \frac{H'_d}{1024} & \quad \text{if $H_t < H'_t + 42$}\\
+H'_d - \frac{H'_d}{1024} & \quad \text{otherwise}\\
 \end{cases}
@@
 
-Where $d'$ is the previous block's difficulty, $t'$ is the prvious block's timestamp and $t$ is the current block's timestamp. This results in an expected period between blocks of roughly 60 seconds.
-
-The nonce, $n$, must satisfy:
+Where $H'$ is the previous block, $H_t$ is the timestamp of block $H$.
+
+This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
+Conversely, is the period is too large the difficulty, and expected time to the next block, is reduced.
+
+The nonce, $n$, must satisfy the relation:
 
 \begin{equation}
-\mathcal{P}(n) < \frac{d}{2^{256}}
+\mathcal{P}(H, n) < \frac{2^{256}}{H_d}
 \end{equation}
 
-Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}).
-The other two parts to a block are simply a list of the transactions (as defined previously) and a list of uncle headers (of the same format as above).
+Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $n$.
+Given an approximately uniform distribution in the range $[0, 2^{256})$, then the expected time to find a solution is proportional to the difficulty, $H_d$.
+
+This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propogate newly created blocks that would otherwise overwrite (``rewrite'') history.
+Because the nonce must satisfy this requirement, and because satisfaction of this requirement both depends on the contents of the block and
+
+The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle headers (of the same format as above).
 
@@
 \begin{equation}
-\mathcal{P}(H, n) < \frac{H_d}{2^{256}}
+\mathcal{P}(H, n) < \frac{2^{256}}{H_d}
 \end{equation}
@@
 
-For the proof-of-concept series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder. It utilises the SHA3 hash function to
+For the PoC series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder. It utilises the SHA3 hash function to
 
@@
 
-TODO
+Browser-style.
+
+Displays ETH balance (i.e. functions as a wallet).
+
+Tabs or perhaps MacOS style app icons.
+
+Display taken up with a single app.
+
+Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
+
+Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
+
+Preconfigured with home app that is the Eth app store.
+
+Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
+
+\section{Implementing Contracts}
+
+\subsection{Data Feeds}
+
+External server runs node.
+Creates \& signs a transaction every minute containing new data.
+Sends to contract which knows to accept data only from this address.
+Allows polling of data from all other contracts (possibly for a microfee).
+
+\subsection{Random Numbers}
+
+Hash the block timestamp; for a series of pseudorandoms, take the previous, add some constant and hash the result.
+
+For a more secure pseudo-random offering both parties agree on a number of random data feed contracts; these are concatenated along with the block timestamp and hashed to produce the first number in the series.
 

### 2014-04-02 Paper.tex
Maths.

@@
 \usepackage{pdflscape}
-
+\usepackage{mathtools}
+
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
 \newcommand*\eg{e.g.\@\xspace}
@@
 \title{Ethereum: A Secure Decentralised Compute Platform}
-\author{Gavin Wood, Vitalik Buterin}
-
+\author{
+    Gavin Wood\\
+    CTO, Ethereum Project\\
+    gavin@ethereum.org
+}
 \begin{document}
@@
 
-\subsection{Overview} \label{ch:overview}
-
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state. Formally:
+\subsection{The Block Chain Paradigm} \label{ch:overview}
+
+Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
+It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
+The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissable.
+Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
+Invalid state changes might, \eg be things such as reducing an account balance without an equal and opposite increase elsewhere.
+We define a valid state transitions as one which comes about through a transaction.
+Formally:
 
 \begin{equation}
-S_{t+1} := f(S_t, T)
+S_{t+1} \equiv f(S_t, T)
 \end{equation}
@@
 
-Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other through a cryptographically secure proof.
+Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block.
+It is achieved thanks to a cryptographically secure proof.
 This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
@@
 Where $g$ is the block-finalisation state transition function (a function that rewards a nominated party); $\mathbb{T}$ is the series of transactions included in this block; and $F$ is the block-level state-transition function.
+
+This is the basis of the block-chain paradigm, a model that forms the backbone of Ethereum.
 
@@
 $10^{15}$ & Finney \\
-$10^{18}$ & Turing \\
+$10^{18}$ & Ether \\
 \bottomrule
@@
 The transaction is considered invalid if the account balance cannot support such a purchase.
-It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender account.
+It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender's account.
 Gas does not exist outside of the execution of a transaction.
@@
 
-Any ether paid for gas by the sender that is not refunded is delivered to the miner.
+Any ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under control of the miner.
 Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
+A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
 Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
@@
 
-The original caller (i.e.
-the original transaction's txsender) pre-pays for a specific amount of 'GAS'.
-What was baseFee now describes the ETH -> GAS price.
-Gas is the internal currency of the transaction.
-Execution operations deplete the GAS supply, as per the original ratios.
-Once the transaction is finished, any remaining GAS is reconverted back to ETH and refunded.
-
-Storage remains 256-bit address, 256-bit data.
-
-Word size: 256-bit (i.e. stack item size 256-bit).
-
-Memory is byte array (256-bit address, 8-bit data); memoryFee is proportional to greatest index that has ever been referenced in this instance and thus it's highly unlikely addresses will ever go above 32-bit.
-Whenever a higher memory index is used, the fee difference to take it to the higher usage from the original (lower) usage is charged.
-MEMSIZE is initially just the size of the input data.
-Note: MSTORE and MLOAD increase the highest-accessed index to their target index + 31.
-
-Given an instruction, it is possible to calculate the gas cost of executing it as follows:
+The execution model specifies how the system state is altered given a series of bytecode instructions and a small tuple of environmental data.
+This is specified through a formal model of a virtual state machine, known as the Ethereum Virtual Machine (EVM).
+It is a \textit{quasi-}Turing-complete machine; the \textit{quasi} qualification comes from the fact that the computation is intrinsically bounded through a parameter, \textit{gas}, which limits the total amount of computation done.
+
+\subsubsection{Basics}
+
+The EVM is a simple stack-based architecture.
+The word size of the machine (and thus size of stack item) is 256-bit.
+This was chosen to facilitate the SHA3-256 hash scheme.
+The memory model is a simple word-addressed byte array.
+The stack has an unlimited size.
+The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
+Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state in the state database.
+
+The machine can have exceptional execution for several reasons, including stack underflows, invalid instructions and divides by zero.
+These unambiguously and validly result in immediate halting of the machine with all state changes left intact.
+The one piece of exceptional execution that does not leave state changes intact is the out-of-gas (OOG) exception.
+Here, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.
+
+\subsubsection{Fees Overview}
+
+Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
+The first and most common is the fee intrinsic to the computation of the operation.
+Most operations require a single gas fee to be paid for their execution; exceptions include \textsc{SSTORE}, \textsc{SLOAD}, \textsc{CALL}, \textsc{CREATE}, \textsc{BALANCE} and \textsc{SHA3}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for \textsc{CREATE} and \textsc{CALL}.
+Finally, gas may be paid due to an increase in the usage of the memory.
+
+Over a contract's execution, the total fee for memory usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
+This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result is an addition memory usage fee.
+
+More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
 
 \begin{itemize}
-\item Unless covered by another rule below, an operation costs < STEPGAS > gas
 \item SHA3 costs $G_{sha3}$ gas
@@
 \item BALANCE costs $G_{balance}$ gas
-\item SSTORE costs $d.G{sstore}$ gas where:
+\item SSTORE costs $d.G_{sstore}$ gas where:
 \begin{itemize}
@@
 \item CREATE costs $G_{create} + g$ gas, where $g$ is the quantity of gas provided.
-\item When you read or write memory with MSTORE, MLOAD, RETURN, SHA3, CALLDATA or CALL, enlarge the memory so that all bytes now fit in it (memory must always be a whole number of 32-byte words).
-Suppose that the highest previously accessed memory index is $m$, and the new index is $n$.
-If $(n + 31) / 32 > (m + 31) / 32$, add an additional $((n + 31) / 32 - (m + 31) / 32) * G_{memory}$ gas to the cost.
+\item All other operations cost $G_{step}$ gas.
 \end{itemize}
+
+Additionally, when memory is accessed with MSTORE, MLOAD, RETURN, SHA3, CALLDATA or CALL, enlarge the memory so that all bytes now fit in it (memory must always be a whole number of 32-byte words).
+Suppose that the highest previously accessed memory index is $i$, and the new index is $i'$; an additional $a$ gas must be paid:
+
+\begin{equation}
+a \equiv \max(0, \ceil[\Bigg]{\frac{i'+1}{32}} - \ceil[\Bigg]{\frac{i+1}{32}}) G_{memory}
+\end{equation}
+
+Originally, memory being empty, the highest accessed index, $i$, is defined as $-1$.
+
+memoryFee is proportional to greatest index that has ever been referenced in this instance and thus it's highly unlikely addresses will ever go above 32-bit.
+Whenever a higher memory index is used, the fee difference to take it to the higher usage from the original (lower) usage is charged.
+MEMSIZE is initially just 0.
+Notably, because \textsc{MSTORE} and \textsc{MLOAD} operate on word lengths, they increase the highest-accessed index to 32 greater than their target index.
 

### 2014-04-02 Paper.tex
Execution stuff.

@@
 Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state in the state database.
+All locations in both storage and memory are well-defined initially as zero.
+
+The machine does not follow the standard von Neumann architecture.
+Rather than storing program code in memory or storage, it is stored separately in a virtual ROM.
+It cannot be read directly; instead it exists only as a model for determining the next instruction to execute.
 
@@
 The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include \textsc{SSTORE}, \textsc{SLOAD}, \textsc{CALL}, \textsc{CREATE}, \textsc{BALANCE} and \textsc{SHA3}.
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for \textsc{CREATE} and \textsc{CALL}.
+Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small SHA3}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE} and {\small CALL}.
 Finally, gas may be paid due to an increase in the usage of the memory.
 
-Over a contract's execution, the total fee for memory usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
+Over a contract's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
 This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result is an addition memory usage fee.
+Due to this fee it's highly unlikely addresses will ever go above 32-bit bounds since at the present price of ether and default gas price, that would cost around US\$20M for the memory fee alone.
+
+Storage fees have a slightly nuanced behaviour---to incentivise contracts to minimise storage use (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is waived; in fact, it is effectively paid up-front since the initial usage of a storage location costs twice as much as the normal usage.
 
@@
 \begin{itemize}
-\item SHA3 costs $G_{sha3}$ gas
-\item SLOAD costs $G_{sload}$ gas
-\item BALANCE costs $G_{balance}$ gas
-\item SSTORE costs $d.G_{sstore}$ gas where:
+\item {\small SHA3} costs $G_{sha3}$ gas
+\item {\small SLOAD} costs $G_{sload}$ gas
+\item {\small BALANCE} costs $G_{balance}$ gas
+\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
 \begin{itemize}
@@
 \end{itemize}
-\item CALL costs $G_{call} + g$, where $g$ is the quantity of gas provided; some gas may later be refunded.
-\item CREATE costs $G_{create} + g$ gas, where $g$ is the quantity of gas provided.
+\item {\small CALL} costs $G_{call} + g$, where $g$ is the quantity of gas provided; some gas may later be refunded.
+\item {\small CREATE} costs $G_{create} + g$ gas, where $g$ is the quantity of gas provided.
 \item All other operations cost $G_{step}$ gas.
@@
 
-Additionally, when memory is accessed with MSTORE, MLOAD, RETURN, SHA3, CALLDATA or CALL, enlarge the memory so that all bytes now fit in it (memory must always be a whole number of 32-byte words).
-Suppose that the highest previously accessed memory index is $i$, and the new index is $i'$; an additional $a$ gas must be paid:
+Additionally, when memory is accessed with {\small MSTORE}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CALLDATA} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+The memory size must always be a whole number of 32-byte words.
+Suppose that the highest previously accessed memory index is $i$, and the new index is $i'$; an additional $G_{m}(i, i')$ gas must be paid:
 
 \begin{equation}
-a \equiv \max(0, \ceil[\Bigg]{\frac{i'+1}{32}} - \ceil[\Bigg]{\frac{i+1}{32}}) G_{memory}
+G_{m}(i, i') \equiv \max(0, \ceil[\Bigg]{\frac{i'+1}{32}} - \ceil[\Bigg]{\frac{i+1}{32}}) G_{memory}
 \end{equation}
 
-Originally, memory being empty, the highest accessed index, $i$, is defined as $-1$.
-
-memoryFee is proportional to greatest index that has ever been referenced in this instance and thus it's highly unlikely addresses will ever go above 32-bit.
-Whenever a higher memory index is used, the fee difference to take it to the higher usage from the original (lower) usage is charged.
-MEMSIZE is initially just 0.
-Notably, because \textsc{MSTORE} and \textsc{MLOAD} operate on word lengths, they increase the highest-accessed index to 32 greater than their target index.
+Originally, with no accesses taken place, the highest accessed index of memory $i$ is defined as $-1$.
+
+%Whenever a higher memory index is referenced, the fee difference to take it to the higher usage from the original (lower) usage is charged.
+Notably, because {\small MSTORE} and {\small MLOAD} operate on word lengths, they implicitly increase the highest-accessed index to 31 greater than their target index.
+
+\subsubsection{Execution Environment}
+
+In addition to the system state $\mathbb{S}$, there are several pieces of important, immutable, information used in the execution environment that the execution agent must provide:
+
+\begin{itemize}
+\item $a$, the address of the account which own the code that is executing.
+\item $o$, the sender address of the transaction that originated this execution.
+\item $p$, the price of gas in the transaction that originated this execution.
+\item $D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $H$, the block header of the previous block.
+\item $H'$, the block header of the current block.
+\item $n$, the block number of the current block.
+\end{itemize}
+
+The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$, given the required information:
+
+\begin{equation}
+\mathbb{S}' \equiv X(\mathbb{S}, a, o, p, D, s, v, H, H', n)
+\end{equation}
+
+\subsubsection{Execution Overview}
 

### 2014-04-02 Paper.tex
Begin work on formal, non-sequential definition of execution model.

@@
 
-In addition to the system state $\mathbb{S}$, there are several pieces of important, immutable, information used in the execution environment that the execution agent must provide:
+In addition to the system state $\mathbb{S}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide:
 
 \begin{itemize}
-\item $a$, the address of the account which own the code that is executing.
+\item $a$, the address of the account which owns the code that is executing.
 \item $o$, the sender address of the transaction that originated this execution.
@@
 
-The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$, given the required information:
+These are assumed to be defined as such throughout.
+The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$ and the remaining gas $g'$, given these definitions:
 
 \begin{equation}
-\mathbb{S}' \equiv X(\mathbb{S}, a, o, p, D, s, v, H, H', n)
+(\mathbb{S}', g') \equiv \Xi(\mathbb{S}, g)
 \end{equation}
@@
 \subsubsection{Execution Overview}
+
+\begin{eqnarray}
+\Xi(\mathbb{S}, g) & \equiv & X(\mathbb{S}, \mathbb{M}) \\
+\mathbb{M}_g & \equiv & g \\
+\mathbb{M}_P & \equiv & 0 \\
+\mathbb{M}_M & \equiv & [0, 0,...] \\
+\mathbb{M}_i & \equiv & -1 \\
+\mathbb{M}_T & \equiv & [] \\
+X(\mathbb{S}, \mathbb{M}) & \equiv & \begin{cases}
+(\mathbb{S}, \mathbb{M}) & \text{if} \quad term(\mathbb{S}, \mathbb{M})\\
+X(O(\mathbb{S}, \mathbb{M})) & \text{otherwise}\\
+\end{cases}
+\end{eqnarray}
 

### 2014-04-03 Paper.tex
Execution model.

--- 2014-04-02 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Compute Platform}
+\title{Ethereum: A Secure Decentralised Transaction Progression Platform}
 \author{
@@
 
-In addition to the system state $\mathbb{S}$, there are several pieces of important, immutable, information used in the execution environment that the execution agent must provide:
+In addition to the system state $\mathbb{S}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
 
 \begin{itemize}
-\item $a$, the address of the account which own the code that is executing.
-\item $o$, the sender address of the transaction that originated this execution.
-\item $p$, the price of gas in the transaction that originated this execution.
-\item $D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $H$, the block header of the previous block.
-\item $H'$, the block header of the current block.
-\item $n$, the block number of the current block.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_H$, the block header of the previous block.
+\item $I_{H'}$, the block header of the current block.
+\item $I_n$, the block number of the current block.
+\item $I_C$, the byte array that is the machine code to be executed.
 \end{itemize}
 
-The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$, given the required information:
+The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$ and the remaining gas $g'$, given these definitions:
 
 \begin{equation}
-\mathbb{S}' \equiv X(\mathbb{S}, a, o, p, D, s, v, H, H', n)
+(\mathbb{S}', g') \equiv \Xi(\mathbb{S}, g, I)
 \end{equation}
@@
 
-\begin{enumerate}
-\item Let $\mathbb{S}' := \mathbb{S}$, the current state.
-\item Let $PC := 0$, the program counter.
-Let it be known that all storage operations operate on TO's state.
-Note that the memory is empty, thus (last used index + 1) = 0.
-\item Set TXDATA to the first INSIZE bytes of memory starting from INOFFSET in the caller memory.
-\item Repeat
-\begin{itemize}
-\item Calculate the cost of the current instruction (see below), set to $C$.
-\item If the instruction is invalid or STOP, goto step 6.
-\item If $GAS < C$ then $GAS := 0$; $S := ROLLBACK$ and evaluation finishes, returning false. 
-\item $GAS := GAS - C$
-\item Apply the instruction.
-\end{itemize}
-Until an operation execution error has occured or the instruction is STOP, RETURN or SUICIDE.
-\item If the output data of the call (R) is specified (through RETURN), then let the OUTSIZE bytes of caller memory beginning at OUTOFFSET.
-\item Returns true.
-\end{enumerate}
-
-When a contract address receives a transaction, a virtual machine is initiated with the contract's state.
+We must now define the $\Xi$ function.
+In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\mathbb{S}$ and the machine state, $\mathbb{M}$.
+Formally, we define it recursively with a function $X$.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying if the present state is a normal halting state of the machine:
+
+\begin{eqnarray}
+\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
+\mathbb{M}_g & \equiv & g \\
+\mathbb{M}_{pc} & \equiv & 0 \\
+\mathbb{M}_M & \equiv & [0, 0, ...] \\
+\mathbb{M}_i & \equiv & -1 \\
+\mathbb{M}_T & \equiv & [] \\
+X(\mathbb{S}, \mathbb{M}, I) & \equiv & \begin{cases}
+(\mathbb{S}, \mathbb{M}, I) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
+O(\mathbb{S}, \mathbb{M}, I) & \text{if} \quad H(\mathbb{S}, \mathbb{M}, I)\\
+X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
+\end{cases}
+\end{eqnarray}
+
+The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the highest used memory index, and the stack contents.
+The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.
+
+For the ease of reading, the instruction mnemonics, written in smallcaps (\eg {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
+
+\subsubsection{Exceptional Halting}
+
+For the purposes of defining $H$ and $O$, we define $o$ as the current operation to be executed:
+
+\begin{equation}
+o \equiv \begin{cases} I_C[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_C \rVert \\
+\text{\small STOP} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack movement items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
+We also institute a function
+
+For calculating the exceptional halting function $Z$, we assume a secondary, instruction-specific halting function $Y$, which identifies edge cases specific to particular instructions (such as the division by zero or initiating a message call with insufficient funds).
+Thus $Z$ is defined as:
+
+\begin{equation}
+\begin{array} {lll}
+Z(\mathbb{S}, \mathbb{M}, I) & \equiv & Y(\mathbb{S}, \mathbb{M}, I) \\
+& & \vee \quad \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \\
+& & \vee \quad \lVert\mathbb{M}_T\rVert < \mathbf{\delta}_o
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is an instruction-specific issue, if there is insufficient gas or if there are insufficient stack items.
+
+\subsubsection{The Execution Cycle}
+
+Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
+
+\begin{eqnarray}
+O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
+\Delta & \equiv & \mathbf{\alpha}_o - \mathbf{\delta}_o \\
+\lVert\mathbb{M}'_{T}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
+\quad \forall i \in [\mathbf{\alpha}_o, \lVert\mathbb{M}'_T\rVert): \mathbb{M}'_{T_i} & \equiv & \mathbb{M}_{T_{i+\Delta}}
+\end{eqnarray}
+
+The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$ which evaluates to the according value:
+
+\begin{eqnarray}
+\mathbb{M}'_{g} \equiv \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
+\mathbb{M}'_{pc} \equiv \begin{cases}
+J(\mathbb{M}) & \text{if} \quad o \in \{ \text{\small JUMP, JUMPI} \} \\
+\mathbb{M}_{pc} + p & \text{if} \quad o \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
+\mathbb{M}_{pc} + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+where $p$ is the byte size of the push instruction, defined as:
+
+\begin{equation}
+p \equiv o - \text{\small PUSH1} + 2
+\end{equation}
+
+In general, we assume the memory and system state don't change:
+
+\begin{eqnarray}
+\mathbb{M}'_M & \equiv & \mathbb{M}_M \\
+\mathbb{M}'_i & \equiv & \mathbb{M}_i \\
+\mathbb{S}' & \equiv & \mathbb{S}
+\end{eqnarray}
+
+Certain instructions do; alter them---such changes are noted in Appendix \ref{app:vm}.
 
 \subsubsection{Terms}
-
-There exists a stack of variable size that stores 256-bit values at each location. (Note: most instructions operate solely on the stack altering its state in some way.)
-
-T'[i] is the ith item counting down from the top of the pre-stack (i.e. the stack immediately after instruction execution), with the top of the stack corresponding to i == 0.
-
-T[i] is the ith item counting down from the top of the post-stack (i.e. the stack immediately prior to instruction execution), with the top of the stack corresponding to i == 0.
-
-The exists a permanent store addressable by a 256-bit value that stores 256-bit values at each location.
-
-S'[i] is the permanent store (sometimes refered to as 'state' or 'store') of the VM at index i counting from zero PRIOR to instruction execution.
-
-S[i] is the permanent store (sometimes refered to as 'state' or 'store') of the VM at index i counting from zero AFTER to instruction execution.
-
-The exists a temporary store addressable by a 256-bit value that stores 256-bit values at each location.
-
-M'[i] is the temporary store (sometimes refered to as 'memory') of the VM at index i counting from zero PRIOR to instruction execution.
-
-M[i] is the temporary store (sometimes refered to as 'memory') of the VM at index i counting from zero AFTER instruction execution.
-
-C is the code, a byte sequence.
-
-PC' is the program counter PRIOR to instruction execution.
-
-PC is the program counter AFTER instruction execution.
-
-FEE(I, S', P', D) is the fee associated with the execution of instruction I with a machine of stack S', permanent store P' and which has already completed D operations.
-
-It is defined as F where:
-
-TODO: No it's not!
-
-IF I == SSTORE AND P[ S'[0] ] != 0 AND S'[1] == 0 THEN
-    F = S + dataFee - memoryFee
-IF I == SSTORE AND P[ S'[0] ] == 0 AND S'[1] != 0 THEN
-    F = S + dataFee + memoryFee
-IF I == SLOAD
-    F = S + dataFee
-IF I == EXTRO OR I == BALANCE
-    F = S + extroFee
-IF I == MKTX
-    F = S + txFee
-IF I == SHA256 OR I == SHA3 OR I == RIPEMD160 OR I == ECMUL OR I == ECADD
-    OR I == ECSIGN OR I == ECRECOVER OR I == ECVALID THEN
-    F = S + cryptoFee
-
-Where:
-
-B[ A ] is the balance of the address given by A, with A interpreted as an address.
-
-ADDRESS is the address of the present contract.
-
-10. b. Initial Operation
-
-STEPSDONE := 0
-PC' := 0
-FOR ALL i: M' is initialised to the empty byte sequence.
-T' is initialised such that its cardinality is zero (i.e. the stack starts empty).
-S' must be equal to the value of P when the previous execution halted.
-
-10. c. General Operation
-
-The given steps are looped:
-1. Execution halts if B[ ADDRESS ] < F( C[PC'], S', P', STEPSDONE )
-2. B[ ADDRESS ] := B[ ADDRESS ] - F( P'[PC'], S', P', STEPSDONE )
-3. The operation given by P'[PC'] determines PC, P, T, S.
-4. PC' := PC; P' := P; T' := T; S' := S; STEPSDONE := STEPSDONE + 1
-
-10. d. VM Operations
 
@@
 
-If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are PUSH, JUMP and JUMPI.
+If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are .
 

### 2014-04-04 Paper.tex
Most of instruction set documented.

@@
 \usepackage{mathtools}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{afterpage}
+
+\definecolor{lightyellow}{rgb}{1,0.98,0.9}
 
@@
 
-\title{Ethereum: A Secure Decentralised Transaction Progression Platform}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger}
 \author{
-    Gavin Wood\\
+    Dr. Gavin Wood\\
     CTO, Ethereum Project\\
@@
 \begin{document}
+
+\pagecolor{lightyellow}
 
@@
 \begin{equation}
-H_d :=  \begin{cases}
+H_d \equiv  \begin{cases}
 2^{22} & \quad \text{if genesis block}\\
@@
 a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
-a & \equiv & \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \emptyset
+a & \equiv & \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \varnothing
 \end{eqnarray}
 
-where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the leftmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\emptyset$ if none exists.
+where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the leftmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 
@@
 \begin{equation}
-\mathbb{S}'[r] := ( 0, v, \empty, \mathcal{S}(\mathbf{b}) )
+\mathbb{S}'[r] \equiv ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
 \end{equation}
@@
 \begin{equation}
-\mathbb{S}'[r]_{balance} := \mathbb{S}[r]_{balance} + v
+\mathbb{S}'[r]_{balance} \equiv \mathbb{S}[r]_{balance} + v
 \end{equation}
@@
 \item $I_H$, the block header of the previous block.
-\item $I_{H'}$, the block header of the current block.
 \item $I_n$, the block number of the current block.
@@
 
-Certain instructions do; alter them---such changes are noted in Appendix \ref{app:vm}.
-
-\subsubsection{Terms}
-
-Summary line:
-<Op-code>: <Mnemonic> -<R> +<A>
-
-If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are .
-
-The cardinality of T (i.e. size of the stack) is altered by A - R between T \& T', by adding or removing items as necessary from the front.
-
-Where:
-R: The minimal cardinality of the stack for this instruction to proceed.
-If this is not achieved then the machine halts with a stack underflow exception.
-(Note: In some cases of some implementations, this is also the number of values "popped" from the implementation's stack during the course of instruction execution.)
-(A - R): The net change in cardinality of the stack over the course of instruction execution.
-
-$\forall$ i: if T[i] is not defined explicitly, then it must be assumed T[i] := T'[i + R - A] where i + R >= A.
-$\forall$ i: if M[i] is not defined explicitly, then it must be assumed M[i] := M'[i].
-$\forall$ i: if S[i] is not defined explicitly, then it must be assumed S[i] := S'[i].
-
-The expression (COND ?
-ONE : ZERO), where COND is an expression and ONE and ZERO are both value placeholders, evaluates to ONE if COND is true, and ZERO if COND is false.
-This is similar to the C-style ternary operator.
-
-When a 32-byte machine datum is interpreted as a 160-bit address or hash, the rightwards 20 bytes are taken (i.e. the low-order bytes when interpreting the data as Big-Endian).
-
-++ is the concatenation operator; all operands are byte arrays (mostly 32-byte arrays here, since that's the size of the VM's data \& address types).
-
-LEFT\_BYTES(A, n) returns the array bytes comprising the first (leftmost) n bytes of the 32 byte array A, which can be considered equivalent to a single value in the VM.
+However, instructions do typically alter one or several components of these values. Altered components, on an listed by instruction are noted in Appendix \ref{app:vm}.
 
@@
 engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
+The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
 
@@
 \begin{equation}
-\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \bigoplus n))
+\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \circ n))
 \end{equation}
@@
 $n$ is the nonce, a series of 32 bytes;
-$\bigoplus$ is the series concatenation operator;
+$\circ$ is the series concatenation operator;
 $\mathcal{B}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
@@
 
-For the release series, we use a more complex proof-of-work. This has yet to be formally defined, but involves two components; firstly,
+For the release series, we use a more complex proof-of-work.
+This has yet to be formally defined, but involves two components; firstly that it concerns contract processing, and by extension the evaluation of programs on the EVM.
+Secondly that it concerns the utilisation of either the block-chain or the full state trie.
+
+As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
+A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
+$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligable amount of processing by the EVM.
+Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
+Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
+swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
+This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
+
+Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
 
@@
 
-\section{Discussion} \label{ch:discussion}
-
-\subsection{Examples} \label{ch:examples}
-
 \section{Conclusion} \label{ch:conclusion}
 
-\subsection{Further Work} \label{ch:further}
+This paper has introduced, discussed and formally defined the protocol of Ethereum.
 
@@
 \bibliographystyle{plainnat}
+
+\end{multicols}
 
@@
 \end{description}
+
+\section{Fee Schedule}\label{app:fees}
+
+\begin{tabular*}{\columnwidth}[h]{rll}
+\toprule
+Name & Value & Description* \\
+\midrule
+$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
+$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
+$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
+$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
+$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled of waived sometimes). \\
+$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
+$G_{create}$ & 100 & Paid for a {\small CREATE} operation and for a contract creation transaction. \\
+$G_{call}$ & 20 & Paid for a {\small CALL} operation and for a message call transaction. \\
+$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
+$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
+\bottomrule
+\end{tabular*}
 
@@
 
-0s: arithmetic operations
-
-0x00: STOP -0 +0
-Halts execution.
-Any gas left over gets returned to caller (or in the case of the top-level call, the sender converted back to ETH).
-0x01: ADD -2 +1
-S[0] := S'[0] + S'[1]
-0x02: MUL -2 +1
-S[0] := S'[0] * S'[1]
-0x03: SUB -2 +1
-S[0] := S'[0] - S'[1]
-0x04: DIV -2 +1
-S[0] := S'[0] / S'[1]
-0x05: SDIV -2 +1
-S[0] := S'[0] / S'[1]
-S'[0] \& S'[1] are interpreted as signed 256-bit values for the purposes of this operation.
-0x06: MOD -2 +1
-S[0] := S'[0] % S'[1]
-0x07: SMOD -2 +1
-S[0] := S'[0] % S'[1]
-S'[0] \& S'[1] are interpreted as signed 256-bit values for the purposes of this operation.
-0x08: EXP -2 +1
-S[0] := S'[0] + S'[1]
-0x09: NEG -1 +1
-S[0] := -S'[0]
-0x0a: LT -2 +1
-S[0] := S'[0] < S'[1] ? 1 : 0
-0x0b: GT -2 +1
-S[0] := S'[0] > S'[1] ? 1 : 0
-0x0c: EQ -2 +1
-S[0] := S'[0] == S'[1] ? 1 : 0
-0x0d: NOT -1 +1
-S[0] := S'[0] == 0 ? 1 : 0
-
-10s: bit operations
-
-0x10: AND -2 +1
-S[0] := S'[0] AND S'[1]
-0x11: OR -2 +1
-S[0] := S'[0] OR S'[1]
-0x12: XOR -2 +1
-S[0] := S'[0] XOR S'[1]
-0x13: BYTE -2 +1
-S[0] := S'[0]th byte of S'[1]
-if S'[0] < 32
-S[0] := 0
-otherwise
-for Xth byte, we count from left - 0th is therefore the leftmost (most significant in BE) byte.
-
-20s: crypto opcodes
-
-0x20: SHA3 -2 +1
-S[0] := SHA3( T'[ S'[0] ++ ... ++ (S'[0] + S'[1]) ])
-
-30s: closure state opcodes
-
-0x30: ADDRESS -0 +1
-S[0] := ADDRESS
-i.e. the address of this closure.
-0x31: BALANCE -1 +1
-S[0] := B[ S'[0] ]
-i.e. the balance of this closure.
-0x32: ORIGIN -0 +1
-S[0] := A
-Where A is the address of the account that made the original transaction leading to the current closure and is paying the fees
-0x33: CALLER -0 +1
-S[0] := A
-Where A is the address of the object that made this call.
-0x34: CALLVALUE -0 +1
-S[0] := V
-Where V is the value attached to this call.
-0x35: CALLDATALOAD -1 +0
-S[0] := D[ S'[0] ... (S'[0] + 31) ]
-Where D is the data attached to this call (as a byte array).
-Any bytes that are out of bounds of the data are defined as zero.
-0x36: CALLDATASIZE -0 +1
-S[0] := DS
-Where DS is the number of bytes of data attached to this call.
-0x37: GASPRICE -0 +1
-S[0] := V
-Where V is the current gas price (né fee multiplier).
-
-40s: block operations
-
-0x40: PREVHASH -0 +1
-S[0] := H
-Where H is the SHA3 hash of the previous block.
-0x41: COINBASE -0 +1
-S[0] := A
-Where A is the coinbase address of the current block.
-0x42: TIMESTAMP -0 +1
-S[0] := T
-Where T is the timestamp of the current block (given as the Unix time\_t when this block began its existence).
-0x43: NUMBER -0 +1
-S[0] := N
-Where N is the block number of the current block (counting upwards from genesis block which has N == 0).
-0x44: DIFFICULTY -0 +1
-S[0] := D
-Where D is the difficulty of the current block.
-0x45: GASLIMIT -0 +1
-S[0] := L
-Where L is the total gas limit of the current block. Always $10^6$.
-
-50s: stack, memory, storage and execution path operations
-
-0x51: POP -1 +0
-0x52: DUP -1 +2
-S[0] := S'[0]
-0x53: SWAP -2 +2
-S[0] := S'[1]
-S[1] := S'[0]
-0x54: MLOAD -1 +1
-S[0] := T'[ S'[0] ... S'[0] + 31 ]
-0x55: MSTORE -2 +0
-T[ S'[0] ... S'[0] + 31 ] := S'[1]
-0x56: MSTORE8 -2 +0
-T[ S'[0] ... S'[0] + 31 ] := S'[1] \& 0xff
-0x57: SLOAD -1 +1
-S[0] := P'[ S'[0] ]
-0x58: SSTORE -2 +0
-P[ S'[0] ] := S'[1]
-0x59: JUMP -1 +0
-PC := S'[0]
-0x5a: JUMPI -2 +0
-PC := S'[1] == 0 ? PC' : S'[0]
-0x:5b PC -0 +1
-S[0] := PC
-0x5c: MSIZE -0 +1
-S[0] = sizeof(T)
-0x5d: GAS
-S[0] := G
-Where G is the amount of gas remaining after executing the opcode.
-
-60s \& 70s: push
-
-0x60: PUSH1 0 +1
-S[0] := C[ PC' + 1 ]
-PC := PC' + 2
-0x61: PUSH2 0 +1
-S[0] := C[ PC' + 1 ] ++ C[ PC' + 2 ]
-PC := PC' + 3
-...
-0x7f: PUSH32 0 +1
-S[0] := C[ PC' + 1 ] ++ C[ PC' + 2 ] ++ ... ++ C[ PC' + 32 ]
-PC := PC' + 33
-
-f0s: closure-level operations
-
-0xf0: CREATE -5 +1
-Immediately creates a contract where:
-The endowment is given by S'[0]
-The body code of the eventual closure is given by T'[ S'[1] ... ( S'[1] + S'[2] - 1 ) ]
-The initialisation code of the eventual closure is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
-(Thus the total number of bytes of the transaction data is given by S'[2] + S'[4].)
-S[0] = A
-where A is the address of the created contract or 0 if the creation failed.
-Fees are deducted from the sender balance to pay for enough gas to complete the operation (i.e.
+As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
+In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $\mathbb{S}'$ $\mathbb{M}'$ such that:
+
+\begin{equation}
+O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I)
+\end{equation}
+
+When interpreting 256-bit binary values as integers the representation is big-endian.
+
+When a 32-byte machine datum is converted to and from as a 160-bit address or hash, the leftwards (high-order for BE) 20 bytes are used and the right most 12 are discarded or filled with zeroes.
+
+TODO: Specify minimum memory limits for a general cost definition.
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+&&&& $\mathbb{M}'_R = []$ \\
+\midrule
+0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] + \mathbb{M}_T[1]$ \\
+\midrule
+0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \times \mathbb{M}_T[1]$ \\
+\midrule
+0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] - \mathbb{M}_T[1]$ \\
+\midrule
+0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
+\midrule
+0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
+&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+\midrule
+0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
+\midrule
+0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
+&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+\midrule
+0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] ^ {\mathbb{M}_T[1] }$ \\
+\midrule
+0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
+&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_T[0]$ \\
+&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+\midrule
+0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] < \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] > \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
+\multicolumn{5}{l}{$\mathbb{M}_T[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_T[0]$} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \wedge \mathbb{M}_T[1]_i$ \\
+\midrule
+0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \vee \mathbb{M}_T[1]_i$ \\
+\midrule
+0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \oplus \mathbb{M}_T[1]_i$ \\
+\midrule
+0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
+&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_T[1]_{(i + 8\mathbb{M}_T[0])} & \text{if} \quad \mathbb{M}_T[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_a$ \\
+\midrule
+0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_balance$ \\
+\midrule
+0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_o$ \\
+&&&& This is the sender of original transaction; it's never a contract. \\
+\midrule
+0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_c$ \\
+&&&& This is the address of the account that is directly responsible for this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_v$ \\
+\midrule
+0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_D[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
+&&&& $\mathbb{M}'_T[0] \equiv \lVert \mathcal{I}_D \rVert$ \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+\midrule
+0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_p$ \\
+&&&& This is gas price specified by the originating transaction.\\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_{H_{prevHash}}$ \\
+&&&& $I_{H_{prevHash}}$ is the previous block's hash. \\
+\midrule
+0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_{H_{coinbase}}$ \\
+\midrule
+0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_{H_{timestamp}}$ \\
+\midrule
+0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_{H_{number}}$ \\
+&&&& $I_{H_{number}}$ is the total number of ancestor blocks of the current block. \\
+\midrule
+0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_{H_{difficulty}}$ \\
+\midrule
+0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
+&&&& $\mathbb{M}'_T[0] \equiv 10^6$ \\
+&&&& Presently, the gas limit is hard-coded in the protocol. \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
+\midrule
+0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0]$ \\
+\midrule
+0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[1]$ \\
+&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_T[0]$ \\
+\midrule
+0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ]$ \\
+\midrule
+0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
+&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ] \equiv \mathbb{M}_T[1]$ \\
+\midrule
+0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
+&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] ] \equiv (\mathbb{M}_T[1] \mod 256) $ \\
+\midrule
+0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[\mathbb{M}_T[0]]$ \\
+\midrule
+0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\mathbb{M}'_S[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
+\midrule
+0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+&&&& $J(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
+&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$ \\
+\midrule
+0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& $J(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$ \\
+\midrule
+0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
+\midrule
+0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
+\midrule
+0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_C[\mathbb{M}_{pc} + 1]$ \\
+&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
+\midrule
+0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_C[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
+&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+\midrule
+\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
+\midrule
+0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
+&&&& $\mathbb{M}'_T[0] \equiv I_C[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0xf0 & {\small CREATE} & 5 & 1 & Create a new contract. \\
+&&&&  \\
+\midrule
+0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
+&&&&  \\
+\midrule
+0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $\mathbb{M}'_R = \mathbb{M}_M[ \mathbb{M}_T[0] \dots ( \mathbb{M}_T[0] + \mathbb{M}_T[1] - 1 ) ]$ \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
+&&&& $\mathbb{S}'[\mathbb{M}_T[0]]_{balance} \equiv \mathbb{S}[\mathbb{M}_T[0]]_{balance} + \mathbb{S}[I_a]_{balance}$ \\
+&&&& $\mathbb{S}'[I_a] \equiv \varnothing$ \\
+\bottomrule
+\end{tabular*}
+
+%0xf0: CREATE -5 +1
+%Immediately creates a contract where:
+%The endowment is given by S'[0]
+%The body code of the eventual closure is given by T'[ S'[1] ... ( S'[1] + S'[2] - 1 ) ]
+%The initialisation code of the eventual closure is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
+%(Thus the total number of bytes of the transaction data is given by S'[2] + S'[4].)
+%S[0] = A
+%where A is the address of the created contract or 0 if the creation failed.
+%Fees are deducted from the sender balance to pay for enough gas to complete the operation (i.e.
 contract creation fee + initial storage fee).
 If there was not enough gas to complete the operation, then all gas will be deducted and the operation fails.
-0xf1: CALL -7 +1
-Immediately executes a call where:
-The recipient is given by S'[0], when interpreted as an address.
-The value is given by S'[1]
-The gas is given by S'[2]
-The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
-(Thus the number of bytes of the transaction is given by S'[4].)
-The output data of the call is given by T[ S'[5] ... ( S'[5] + MIN( S'[6], S[0] ) - 1 ) ]
-If 0 gas is specified, transfer all gas from the caller to callee gas balance, otherwise transfer only the amount given. If there isn't enough gas in the caller balance, operation fails.
-If the value is less than the amount in the caller's balance then nothing is executed and the S'[2] gas gets refunded to the caller.
-See Appendix A for execution.
-Add any remaining callee gas to the caller's gas.
-S[0] = R
-where R = 1 when the instrinsic return code of the call is true, R = 0 otherwise.
-NOT YET: POST -5 +1
-Registers for delayed execution a call where:
-The recipient is given by S'[0], when interpreted as an address.
-The value is given by S'[1]
-The gas to supply the transaction with is given by S'[2] (paid for from the current gas balance)
-The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
-(Thus the number of bytes of the transaction is given by S'[4].)
-Contract pays for itself to run at a defered time from its own GAS supply. The miner has no choice but to execute.
-NOT YET: ALARM -6 +1
-Registers for delayed execution a call where:
-The recipient is given by S'[0], when interpreted as an address.
-The value is given by S'[1]
-The gas (to convert from ETH at the later time) is given by S'[2]
-The number of blocks to wait before executing is S'[3]
-The input data of the call is given by T'[ S'[4] ... ( S'[4] + S'[5] - 1 ) ]
-(Thus the number of bytes of the transaction is given by S'[5].)
-Total gas used now is S'[3] * S'[5] * deferFee.
-Contract pays for itself to run at the defered time converting given amount of gas from its ETH balance; if it cannot pay it terminates as a bad transaction.
-TODO: include baseFee and allow miner freedom to determine whether to execute or not.
-If not, the next miner will have the chance.
-0xf2: RETURN -2 +0
-Halts execution.
-R := T[ S'[0] ... ( S'[0] + S'[1] - 1 ) ]
-Where the output data of the call is specified as R.
-Any gas left over gets returned to caller (or in the case of the top-level call, the sender converted back to ETH).
-0xff: SUICIDE -1 +0
-Halts execution.
-FOR ALL i: IF P[i] NOT EQUAL TO 0 THEN B[ S'[0] ] := B[ S'[0] ] + memoryFee
-B[ S'[0] ] := B[ S'[0] ] + B[ ADDRESS ]
-Removes all contract-related information from the Ethereum system.
+%0xf1: CALL -7 +1
+%Immediately executes a call where:
+%The recipient is given by S'[0], when interpreted as an address.
+%The value is given by S'[1]
+%The gas is given by S'[2]
+%The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
+%(Thus the number of bytes of the transaction is given by S'[4].)
+%The output data of the call is given by T[ S'[5] ... ( S'[5] + MIN( S'[6], S[0] ) - 1 ) ]
+%If 0 gas is specified, transfer all gas from the caller to callee gas balance, otherwise transfer only the amount given. If there isn't enough gas in the caller balance, operation fails.
+%If the value is less than the amount in the caller's balance then nothing is executed and the S'[2] gas gets refunded to the caller.
+%See Appendix A for execution.
+%Add any remaining callee gas to the caller's gas.
+%S[0] = R
+%where R = 1 when the instrinsic return code of the call is true, R = 0 otherwise.
 
@@
 
-\section{Fee Schedule}\label{app:fees}
-
-Constants:
-
-* STEPGAS = 1
-
-* SHA3GAS = 20
-
-* SLOADGAS = 20
-
-* SSTOREGAS = 100
-
-* BALANCEGAS = 20
-
-* CREATEGAS = 100
-
-* CALLGAS = 20
-
-* MEMORYGAS = 1
-
-* TXDATAGAS = 5 [not used in the VM]
-
-\end{multicols}
-
 \end{document}

### 2014-04-05 Paper.tex
Fixes from Vitalik.

@@
 \begin{equation}
-\mathcal{P}(H, n) < \frac{2^{256}}{H_d}
+\mathcal{P}(H, n) \leqslant \frac{2^{256}}{H_d}
 \end{equation}
@@
 
-The address of the new contract is defined as being the leftmost 160 bits of the SHA3 hash of RLP encoding of the structure \texttt{[ sender, nonce ]}.
+The address of the new contract is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure \texttt{[ sender, nonce ]}.
 In in the unlikely event that the addresses is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at, thus:
@@
 
-where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the leftmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the rightmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 
@@
 
+This is a method for encoding arbitrarily structured binary data (byte arrays).
+
 \section{Modified Merkle Patricia Tree}\label{app:trie}
 
+The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
+It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
+
 \section{Hex-Prefix Encoding}\label{app:hexprefix}
+
+Hex-prefix encoding is an efficient method of encoding a arbitrary number of nibbles as a byte array.
+It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
 
@@
 
+The state database is a mapping between SHA3-256 hashes (i.e. 256-bit binary fragments) and arbitrary byte data.
+
 \section{Signing Transactions}\label{app:signing}
+
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve.
 
@@
 
+The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
 \section{Wire Protocol}\label{app:wire}
 
+The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transfering information such as blocks \& transactions
+
 \section{Peer Strategy}\label{app:peers}

### 2014-04-06 Paper.tex
Work on execution model.

@@
 
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $f$ as specified before.
+The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function as specified before.
 It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
@@
 \item the transaction nonce is valid (equivalent to the sender account's current nonce);
-\item the gas limit is no smaller that the intrinsic gas, $i$, used by the transaction;
+\item the gas limit is no smaller that the intrinsic gas, $g_i$, used by the transaction;
 \item the sender account balance contains at least the cost, $c$, (in wei) required in up-front payment;
@@
 \begin{eqnarray}
-i & \equiv & \begin{cases}
+g_i & \equiv & \begin{cases}
   |\mathbf{d}|.G_{data} + G_{call} & \quad \text{call} \\
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the sender's nonce is incremented by one and the balance is reduced by the cost, $c$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender ($s$) is incremented by one and the balance is reduced by the cost, $c$.
 The gas available for the proceeding computation, $g$, is defined as $l - i$.
@@
 
-After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate, thus we increase the sender's balance by $g.p$.
+TODO: formal description of reduction of sender's balance by $c$, then re-addition later on by $g'.p$
+
+We define the rollback state $\mathbb{S}_R$:
+
+\begin{eqnarray}
+\mathbb{S}_R & \equiv & \mathbb{S} \quad \text{except:} \\
+\mathbb{S}_R[s]_{balance} & \equiv & \mathbb{S}[s]_{balance} - c \\
+\mathbb{S}_R[s]_{nonce} & \equiv & \mathbb{S}[s]_{nonce} + 1
+\end{eqnarray}
+
+After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate, thus we increase the sender's balance by the refunded amount.
+So we define the final state $\mathbb{S}'$ in terms of the pre-final state $\mathbb{S}_C$.
+
+\begin{eqnarray}
+\mathbb{S}' & \equiv & \mathbb{S}_C \quad \text{except:} \\
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_C[s]_{balance} + g'.p
+\end{eqnarray}
+
+Evaluating $\mathbb{S}_C$ from $\mathbb{S}_R$ depends on the transaction type; either contract-creation or message-call; I will describe and define these separately.
 
@@
 
-There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($o$), available gas ($g$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.
+There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_C$ and additional state database entries $\mathbb{D}$, to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, \mathbb{D}', g')$:
+
+\begin{equation}
+(\mathbb{S}_P, \mathbb{D}', g') \equiv \Lambda(\mathbb{S}_C, \mathbb{D}, s, o, g, p, v, \mathbf{i}, \mathbf{b})
+\end{equation}
+
+TODO: Move rest to own section after execution model.
+
+TODO: Transaction structure layout.
 
 The address of the new contract is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure \texttt{[ sender, nonce ]}.
-In in the unlikely event that the addresses is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at, thus:
-
-\begin{eqnarray}
-a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o])))\\
-a & \equiv & \arg \min_x : x \geq a' \wedge \mathbb{S}[x] = \varnothing
-\end{eqnarray}
-
-where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the rightmost $y$ bits of the binary data $X$ and $\mathbb{S}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the new state becomes $\mathbb{S}'$:
-
-\begin{equation}
-\mathbb{S}'[r] \equiv ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
-\end{equation}
-
-The state database is altered to include the pair $(\mathcal{S}(\mathbf{b}), \mathbf{b})$.
+In in the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
+Thus we define the creation address function $A$:
+
+\begin{eqnarray}
+A(s, o) & \equiv & a \quad \text{where:} \\
+a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o]))) \\
+a & \equiv & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_C[x] = \varnothing
+\end{eqnarray}
+
+where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the rightmost $y$ bits of the binary data $X$ and $\mathbb{S}_C[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\mathbb{S}^*$:
+
+\begin{eqnarray}
+\mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
+\mathbb{S}^*[A(s, o)] & \equiv & ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
+\end{eqnarray}
+
+The mutated state database $\mathbb{D}'$, is defined as to include the pair $(\mathcal{S}(\mathbf{b}), \mathbf{b})$:
+
+\begin{eqnarray}
+\mathbb{D}' & \equiv & \mathbb{D} \quad \text{except:} \\
+\mathbb{D}'[\mathcal{S}(\mathbf{b})] & \equiv & \mathbf{b}
+\end{eqnarray}
+
 Finally, the contract is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
 Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}'$, additional database entries $\mathbb{D}'$ and available gas remaining $g'$.
 
 Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this exceptional case we say an Out-of-Gas exception has occured and the state database is reverted to the point immediately prior to the account initialisation.
-All available gas is used and no gas is refunded to the originator.
+In this exceptional case we say an Out-of-Gas exception has occured.
+In this case, the evaluated state is defined as being the empty set $\varnothing$.
+In this circumstance, the entire create operation should have no effect on the state/database; the gas remaining should of course be zero.
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the new state, database entries and gas as $\mathbb{S}'$, $\mathbb{D}', g'$ where:
+
+\begin{eqnarray}
+(\mathbb{S}^{**}, \mathbb{D}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, \mathbb{D}^*, g, I) \\
+(\mathbb{S}_P, \mathbb{D}') & \equiv & \begin{cases}
+(\mathbb{S}_C, \mathbb{D}) & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
+(\mathbb{S}^{**}, \mathbb{D}^{**}) & \text{otherwise}
+\end{cases} \\
+I'_a & \equiv & a \\
+I'_o & \equiv & s \\
+I'_p & \equiv & p \\
+I'_D & \equiv & [] \\
+I'_s & \equiv & s \\
+I'_v & \equiv & v \\
+I'_C & \equiv & \mathbf{i}
+\end{eqnarray}
+
+$I'_D$ evaluates to the empty array. $I'_H$ and $I'_n$ have no special treatment and are determined from the block chain.
 
@@
 
-In the case of executing a message call, several parameters are required: sender ($s$), recipient ($r$), available gas ($g$), value ($v$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call.
-Message calls also have intrinsic return data; the call's output byte array $\mathbf{r}$.
-This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case the return data is used.
-
-In executing a message call, the account balance of the recipient is increased by the value passed, thus $\mathbb{S}$ becomes replaced with $\mathbb{S}'$:
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call.
+Aside from evaluating to a new state and additional database entries, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
 
 \begin{equation}
-\mathbb{S}'[r]_{balance} \equiv \mathbb{S}[r]_{balance} + v
+(\mathbb{S}_P, \mathbb{D}', g', \mathbf{o}) \equiv \Theta(\mathbb{S}_C, \mathbb{D}, s, o, r, g, v, p, \mathbf{d})
 \end{equation}
 
+TODO: Move rest to own section after execution model.
+
+TODO: Transaction structure layout.
+
 In the case that the recipient contains code to be executed (i.e.
-that the account is a contract), then the contract's body code, to be found in the state database with key $\mathbb{S}[r]_{code}$, is executed according to the execution model (see section \ref{ch:model}).
+that the account is a contract), then the contract's body code, $\mathbb{D}[\mathbb{S}[r]_{code}]$, is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
-Unlike with contract creation,
+
+\begin{eqnarray}
+(\mathbb{S}^{**}, \mathbb{D}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, \mathbb{D}^*, g, I) \\
+(\mathbb{S}', \mathbb{D}') & \equiv & \begin{cases}
+(\mathbb{S}^*, \mathbb{D}) & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
+(\mathbb{S}^{**}, \mathbb{D}^{**}) & \text{otherwise}
+\end{cases} \\
+I'_a & \equiv & a \\
+I'_o & \equiv & o \\
+I'_p & \equiv & p \\
+I'_D & \equiv & \mathbf{d} \\
+I'_s & \equiv & s \\
+I'_v & \equiv & v \\
+I'_C & \equiv & \mathbb{D}[\mathbb{S}[r]_{code}]
+\end{eqnarray}
 
@@
 The word size of the machine (and thus size of stack item) is 256-bit.
-This was chosen to facilitate the SHA3-256 hash scheme.
+This was chosen to facilitate the SHA3-256 hash scheme and ECC.
 The memory model is a simple word-addressed byte array.
@@
 \item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_C$, the byte array that is the machine code to be executed.
 \item $I_H$, the block header of the previous block.
 \item $I_n$, the block number of the current block.
-\item $I_C$, the byte array that is the machine code to be executed.
 \end{itemize}
 
-The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$ and the remaining gas $g'$, given these definitions:
+The execution model defines the function $\Xi$, which can compute the resultant state $\mathbb{S}'$ and the remaining gas $g'$, given these definitions:
 
@@
 
-\begin{tabular*}{\columnwidth}[h]{rll}
+\begin{tabular*}{\columnwidth}[h]{lrl}
 \toprule
@@
 0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_balance$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
 \midrule
@@
 0xf0 & {\small CREATE} & 5 & 1 & Create a new contract. \\
-&&&&  \\
+&&&& $\mathbf{b} \equiv \mathbb{M}_M[ \mathbb{M}_S[1] \dots (\mathbb{M}_S[1] + \mathbb{M}_S[2] - 1) ]$ \\
+&&&& $\mathbf{i} \equiv \mathbb{M}_M[ \mathbb{M}_S[3] \dots (\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) ]$ \\
+&&&& $(\mathbb{S}', \mathbb{D}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}, \mathbb{D}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
 \midrule
 0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
-&&&&  \\
+&&&& $\mathbf{i} \equiv \mathbb{M}_M[ \mathbb{M}_S[3] \dots (\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) ]$ \\
+&&&& $\mathbf{o} \equiv \mathbb{M}_M[ \mathbb{M}_S[5] \dots (\mathbb{M}_S[5] + \mathbb{M}_S[6] - 1) ]$ \\
+&&&& $(\mathbb{S}', \mathbb{D}', \mathbb{M}'_g, \mathbf{o}) \equiv \Theta(\mathbb{S}, \mathbb{D}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbb{M}_S[1], \mathbb{M}_S[2], \mathbf{i})$ \\
 \midrule

### 2014-04-06 Paper.tex
trivial changes only - typos and obvious grammar correction - consistent standard orthography: block-chain, block chain -> blockchain, back-end -> backend - minor reorg - . deleted for multiplication in equations - a few convoluted sentences simplified

@@
 \begin{abstract}
-The block-chain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 and $f$ is the Ethereum state transition function.
-In Ethereum, $f$, together with $S$ are considerably more powerful then any existing comparable system; $f$ allows components to carry out arbitrary computation to be done, while $S$ allows components to store arbitrary state between transactions.
+In Ethereum, $f$, together with $S$ are considerably more powerful then any existing comparable system; $f$ allows components to carry out arbitrary computation, while $S$ allows components to store arbitrary state between transactions.
 
@@
 
-This is the basis of the block-chain paradigm, a model that forms the backbone of Ethereum.
+This is the basis of the blockchain paradigm, a model that forms the backbone of Ethereum.
 
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the block-chain, there must be an agreed-upon scheme.
-If there is ever a disagreement between nodes as to which root-to-leaf path down the block-tree is the `best' block-chain, then a \textit{fork} occurs.
-
-This would mean that past a given point in time (block), there is a temporal co-existance of multiple states of the system: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
+In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
+
+This would mean that past a given point in time (block), multiple states of the system may coexist: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
 This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.
@@
 
-The world-state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP).
+The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP).
 Though not stored on the block chain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database back-end that maintains a mapping of bytearrays to bytearrays; we name this underlaying database the state database.
+The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
 This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
@@
 A transaction is a single cryptographically signed instruction sent by an actor external to Ethereum.
-e.g.
-This might be from a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
+An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
 There are two types of transactions: those which result in message calls and those which result in the creation of new contracts.
@@
 Both \textbf{body} and \textbf{init} are EVM-code fragments; \textbf{body} is executed each time the contract account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is a piece of code executed only once and at contract creation.
-It gets discarded immediately.
+\textbf{init} is a piece of code that is executed only once at contract creation and that gets discarded immediately thereafter.
 
@@
 This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, is the period is too large the difficulty, and expected time to the next block, is reduced.
+Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
 
@@
 Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $n$.
-Given an approximately uniform distribution in the range $[0, 2^{256})$, then the expected time to find a solution is proportional to the difficulty, $H_d$.
+Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 
-The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle headers (of the same format as above).
+The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle block headers (of the same format as above).
 
@@
 
-Any ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under control of the miner.
+Any ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under the control of the miner.
 Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
@@
 \item the transaction nonce is valid (equivalent to the sender account's current nonce);
-\item the gas limit is no smaller that the intrinsic gas, $g_i$, used by the transaction;
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
 \item the sender account balance contains at least the cost, $c$, (in wei) required in up-front payment;
@@
 
-where:
-
-\begin{eqnarray}
-g_i & \equiv & \begin{cases}
-  |\mathbf{d}|.G_{data} + G_{call} & \quad \text{call} \\
-  (|\mathbf{i}| + |\mathbf{b}|).G_{data} + G_{create} & \quad \text{create} \\
-\end{cases} \\
-c & \equiv & l.p + v
-\end{eqnarray}
-
-where $|\mathbf{d}|$, $|\mathbf{i}|$ and $|\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation \& body EVM-code and $G$ is defined in Appendix \ref{app:fees}.
+Intrinsic gas $g_0$ is calculated as follows:
+
+\begin{eqnarray}
+g_0 & \equiv & \begin{cases}
+  |\mathbf{d}| G_{data} + G_{call} & \quad \text{call} \\
+  (|\mathbf{i}| + |\mathbf{b}|) G_{data} + G_{create} & \quad \text{create} \\
+\end{cases}
+\end{eqnarray}
+
+where $|\mathbf{d}|$, $|\mathbf{i}|$ and $|\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
+
+The up-front cost $c$ is calculated as:
+
+\begin{eqnarray}
+c & \equiv & l p + v
+\end{eqnarray}
+
+where $l$ is gasLimit, $p$ is gasPrice and $v$ is value as defined parts of the transaction.
 
 The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender ($s$) is incremented by one and the balance is reduced by the cost, $c$.
-The gas available for the proceeding computation, $g$, is defined as $l - i$.
+The gas available for the proceeding computation, $g$, is defined as $l - g_0$.
 The proceeding computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \mathbb{S}' & \equiv & \mathbb{S}_C \quad \text{except:} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_C[s]_{balance} + g'.p
-\end{eqnarray}
-
-Evaluating $\mathbb{S}_C$ from $\mathbb{S}_R$ depends on the transaction type; either contract-creation or message-call; I will describe and define these separately.
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_C[s]_{balance} + r p
+\end{eqnarray}
+
+Evaluating $\mathbb{S}_C$ from $\mathbb{S}_R$ depends on the transaction type; either contract creation or message call; I will describe and define these separately.
 
@@
 The address of the new contract is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure \texttt{[ sender, nonce ]}.
-In in the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
+In the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
 Thus we define the creation address function $A$:
@@
 
-However, instructions do typically alter one or several components of these values. Altered components, on an listed by instruction are noted in Appendix \ref{app:vm}.
-
-\subsection{Block-tree to Block-chain} \label{ch:ghost}
+However, instructions do typically alter one or several components of these values. Altered components listed by instruction are noted in Appendix \ref{app:vm}.
+
+\subsection{Blocktree to Blockchain} \label{ch:ghost}
 
@@
 Such blocks become more common as the network propogation time approaches the ideal inter-block time.
-However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to butress the more popular block-chain making it a stronger choice over less popular (though potentially longer) competitors.
+However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to butress the more popular blockchain making it a stronger choice over less popular (though potentially longer) competitors.
 
@@
 
-As such given a block $B$, $B_td$ is its total difficulty, $B_P$ is its parent block, $B_d$ is its difficulty and $B_\mathbb{U}$ is its set of uncle blocks.
+As such given a block $B$, $B_{td}$ is its total difficulty, $B_P$ is its parent block, $B_d$ is its difficulty and $B_\mathbb{U}$ is its set of uncle blocks.
 
@@
 This has yet to be formally defined, but involves two components; firstly that it concerns contract processing, and by extension the evaluation of programs on the EVM.
-Secondly that it concerns the utilisation of either the block-chain or the full state trie.
+Secondly that it concerns the utilisation of either the blockchain or the full state trie.
 
@@
 \item[External Actor] A person or other entity able to interface to an Ethereum node, but external to the world of Ethereum.
-It can interact with Ethereum through depositing signed Transactions and inspecting the block-chain and associated state.
+It can interact with Ethereum through depositing signed Transactions and inspecting the blockchain and associated state.
 Has one (or more) intrinsic Accounts.
@@
 
-\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the block-chain.
+\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the blockchain.
 
@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (ala Chrome) that is able to host sandboxed applications whose back-end is purely on the Ethereum protocol.
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (ala Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
 

### 2014-04-06 Paper.tex
g_0 and block chain -> blockchain

@@
 
-\subsection{The Block Chain Paradigm} \label{ch:overview}
+\subsection{The Blockchain Paradigm} \label{ch:overview}
 
@@
 The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP).
-Though not stored on the block chain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
+Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
 The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
@@
 Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly.
-Since we store all such root hashes in the block chain, we are able to trivially revert to old states.
+Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.
 
@@
 \mathbb{S}' & \equiv & \mathbb{S}_C \quad \text{except:} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_C[s]_{balance} + r p
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_C[s]_{balance} + g' p
 \end{eqnarray}
@@
 
-$I'_D$ evaluates to the empty array. $I'_H$ and $I'_n$ have no special treatment and are determined from the block chain.
+$I'_D$ evaluates to the empty array. $I'_H$ and $I'_n$ have no special treatment and are determined from the blockchain.
 
@@
 
-The canonical block chain is a path from root to leaf through the entire block tree.
+The canonical blockchain is a path from root to leaf through the entire block tree.
 In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
@@
 
-This increases overall network security by making it much harder for an adversary to silently mine a canonical block chain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of reversing existing blocks and the transactions within.
+This increases overall network security by making it much harder for an adversary to silently mine a canonical blockchain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of reversing existing blocks and the transactions within.
 
@@
 The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the block chain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
+It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
 However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.

### 2014-04-06 Paper.tex
Better spec for code.
Transaction spec.

@@
 \begin{equation}
-S_{t+1} \equiv f(S_t, T)
+\mathbb{S}_{t+1} \equiv \Upsilon(\mathbb{S}_t, T)
 \end{equation}
@@
 and $f$ is the Ethereum state transition function.
-In Ethereum, $f$, together with $S$ are considerably more powerful then any existing comparable system; $f$ allows components to carry out arbitrary computation, while $S$ allows components to store arbitrary state between transactions.
+In Ethereum, $\Upsilon$, together with $\mathbb{S}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\mathbb{S}$ allows components to store arbitrary state between transactions.
 
@@
 \begin{eqnarray}
-S_{t+1} & \equiv & F(S_t, \mathbb{T}) \\
-\mathbb{T} & \equiv & [ T_0, T_1, ... ] \\
-F(S, \mathbb{T}) & \equiv & g(f(f(S, T_0), T_1) ...)
+\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, \mathbf{T}) \\
+\mathbf{T} & \equiv & [ T_0, T_1, ... ] \\
+\Pi(\mathbb{S}, \mathbf{T}) & \equiv & \Omega(\Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
 \end{eqnarray}
 
-Where $g$ is the block-finalisation state transition function (a function that rewards a nominated party); $\mathbb{T}$ is the series of transactions included in this block; and $F$ is the block-level state-transition function.
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $\mathbf{T}$ is the series of transactions included in this block; and $\Pi$ is the block-level state-transition function.
 
@@
 
-A transaction is a single cryptographically signed instruction sent by an actor external to Ethereum.
+A transaction (formally, $T$) is a single cryptographically signed instruction sent by an actor external to Ethereum.
 An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender.
-\item[value] A scalar value equal to the number of wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created contract's account.
-\item[gasPrice] A scalar value equal to the number of wei to be paid per unit of gas for all computation costs incurred as a result of the execution of this transaction.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[value] A scalar value equal to the number of wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created contract's account; formally $T_v$.
+\item[gasPrice] A scalar value equal to the number of wei to be paid per unit of gas for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be topped-up later.
-\item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction.
-\item[v, r, s] Three scalar values corresponding to the signature of the transaction and used to determine the sender of the transaction. See Appendix \ref{app:signing}.
+This is paid up-front, before any computation is done and may not be topped-up later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction; formally $T_r$.
+\item[v, r, s] Three scalar values corresponding to the signature of the transaction and used to determine the sender of the transaction.
 \end{description}
@@
 \begin{description}
-\item[body] An unlimited size byte array specifying the EVM-code for the contract.
-\item[init] An unlimited size byte array specifying the EVM-code for the contract initialisation procedure.
+\item[body] An unlimited size byte array specifying the EVM-code for the contract, formally $T_\mathbf{b}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the contract initialisation procedure, formally $T_\mathbf{i}$.
 \end{description}
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
 \end{description}
 
+Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDMA of the SECP-256k1 curve, using the hash of the transaction as the datum to sign.
+For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
+
 \subsubsection{Block} \label{ch:block}
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}) together with a set of transactions and a set of other block headers that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions and a set of other block headers that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety.
-\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block.
-\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred.
-\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied.
-\item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block.
-\item[difficulty] A scalar value corresonding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception.
-\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block.
-\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block.
+\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety; formally $H_p$.
+\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block; formally $H_u$.
+\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_b$.
+\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block; formally $H_t$.
+\item[difficulty] A scalar value corresonding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block; formally $H_x$.
+\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 
-The nonce, $n$, must satisfy the relation:
-
-\begin{equation}
-\mathcal{P}(H, n) \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $n$.
+The nonce, $H_n$, must satisfy the relation:
+
+\begin{equation}
+\mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
 Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 
-\subsection{Transaction Execution} \label{ch:execution}
-
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function as specified before.
+\subsection{Transaction Execution} \label{ch:transactions}
+
+The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
 It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
@@
 
-Intrinsic gas $g_0$ is calculated as follows:
+Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\mathbb{S}$ the state:
+
+\begin{equation}
+\mathbb{S}' = \Upsilon(\mathbb{S}, T)
+\end{equation}
+
+Thus $\mathbb{S}'$ is the post-transactional state.
+
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 
@@
 g_0 & \equiv & \begin{cases}
-  |\mathbf{d}| G_{data} + G_{call} & \quad \text{call} \\
-  (|\mathbf{i}| + |\mathbf{b}|) G_{data} + G_{create} & \quad \text{create} \\
+  |T_\mathbf{d}| G_{data} + G_{call} & \quad \text{call} \\
+  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \quad \text{create} \\
 \end{cases}
@@
 
-where $|\mathbf{d}|$, $|\mathbf{i}|$ and $|\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
-
-The up-front cost $c$ is calculated as:
+where $|T_\mathbf{d}|$, $|T_\mathbf{i}|$ and $|T_\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
+
+The up-front cost $v_0$ is calculated as:
+
+\begin{equation}
+v_0 \equiv (T_g + g_0) T_p + T_v
+\end{equation}
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender ($S(T)$) is incremented by one and the balance is reduced by the up-front cost, $v_0$.
+The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
+
+We define the rollback state $\mathbb{S}_0$:
 
 \begin{eqnarray}
-c & \equiv & l p + v
+\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
+\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - c \\
+\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
 \end{eqnarray}
 
-where $l$ is gasLimit, $p$ is gasPrice and $v$ is value as defined parts of the transaction.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender ($s$) is incremented by one and the balance is reduced by the cost, $c$.
-The gas available for the proceeding computation, $g$, is defined as $l - g_0$.
-The proceeding computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
-
-TODO: formal description of reduction of sender's balance by $c$, then re-addition later on by $g'.p$
-
-We define the rollback state $\mathbb{S}_R$:
+Evaluating $\mathbb{S}_C$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
+
+\begin{equation}
+(\mathbb{S}_P, g') \equiv \begin{cases}
+\Lambda(\mathbb{S}_0, S(T), T_o, T_g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
+\Theta_{0,1}(\mathbb{S}_0, S(T), T_o, T_r, T_g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Note we use $\Theta_{0,1}$ to take only the first two components of the function's value; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+
+After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate.
+So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$.
 
 \begin{eqnarray}
-\mathbb{S}_R & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}_R[s]_{balance} & \equiv & \mathbb{S}[s]_{balance} - c \\
-\mathbb{S}_R[s]_{nonce} & \equiv & \mathbb{S}[s]_{nonce} + 1
+\mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' p
 \end{eqnarray}
 
-After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate, thus we increase the sender's balance by the refunded amount.
-So we define the final state $\mathbb{S}'$ in terms of the pre-final state $\mathbb{S}_C$.
-
-\begin{eqnarray}
-\mathbb{S}' & \equiv & \mathbb{S}_C \quad \text{except:} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_C[s]_{balance} + g' p
-\end{eqnarray}
-
-Evaluating $\mathbb{S}_C$ from $\mathbb{S}_R$ depends on the transaction type; either contract creation or message call; I will describe and define these separately.
-
-\subsubsection{Create} \label{ch:create}
+\subsection{Contract Creation} \label{ch:create}
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_C$ and additional state database entries $\mathbb{D}$, to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, \mathbb{D}', g')$:
-
-\begin{equation}
-(\mathbb{S}_P, \mathbb{D}', g') \equiv \Lambda(\mathbb{S}_C, \mathbb{D}, s, o, g, p, v, \mathbf{i}, \mathbf{b})
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_C$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
+
+\begin{equation}
+(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_C, s, o, g, p, v, \mathbf{i}, \mathbf{b})
 \end{equation}
@@
 
-The mutated state database $\mathbb{D}'$, is defined as to include the pair $(\mathcal{S}(\mathbf{b}), \mathbf{b})$:
-
-\begin{eqnarray}
-\mathbb{D}' & \equiv & \mathbb{D} \quad \text{except:} \\
-\mathbb{D}'[\mathcal{S}(\mathbf{b})] & \equiv & \mathbf{b}
-\end{eqnarray}
+It is asserted that the state database will also change such that it defines the pair $(\mathcal{S}(\mathbf{b}), \mathbf{b})$.
 
@@
 Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}'$, additional database entries $\mathbb{D}'$ and available gas remaining $g'$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}'$ and available gas remaining $g'$.
 
 Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this exceptional case we say an Out-of-Gas exception has occured.
-In this case, the evaluated state is defined as being the empty set $\varnothing$.
-In this circumstance, the entire create operation should have no effect on the state/database; the gas remaining should of course be zero.
+In this exceptional case we say an Out-of-Gas exception has occured: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should of course be zero.
+If the creation was conducted as the manifestation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the new state, database entries and gas as $\mathbb{S}'$, $\mathbb{D}', g'$ where:
+Thus formally, we may specify the new state and gas as $\mathbb{S}'$, $g'$ where:
 
 \begin{eqnarray}
-(\mathbb{S}^{**}, \mathbb{D}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, \mathbb{D}^*, g, I) \\
-(\mathbb{S}_P, \mathbb{D}') & \equiv & \begin{cases}
-(\mathbb{S}_C, \mathbb{D}) & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
-(\mathbb{S}^{**}, \mathbb{D}^{**}) & \text{otherwise}
+(\mathbb{S}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, g, I) \\
+\mathbb{S}_P & \equiv & \begin{cases}
+\mathbb{S}_C & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
+\mathbb{S}^{**} & \text{otherwise}
 \end{cases} \\
@@
 
-\subsubsection{Message Call} \label{ch:call}
+\subsection{Message Call} \label{ch:call}
 
@@
 \begin{equation}
-(\mathbb{S}_P, \mathbb{D}', g', \mathbf{o}) \equiv \Theta(\mathbb{S}_C, \mathbb{D}, s, o, r, g, v, p, \mathbf{d})
-\end{equation}
-
-TODO: Move rest to own section after execution model.
-
-TODO: Transaction structure layout.
+(\mathbb{S}_P, g', \mathbf{o}) \equiv \Theta(\mathbb{S}_C, s, o, r, g, p, v, \mathbf{d})
+\end{equation}
 
 In the case that the recipient contains code to be executed (i.e.
-that the account is a contract), then the contract's body code, $\mathbb{D}[\mathbb{S}[r]_{code}]$, is executed according to the execution model (see section \ref{ch:model}).
+that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_{code}]$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
@@
 \begin{eqnarray}
-(\mathbb{S}^{**}, \mathbb{D}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, \mathbb{D}^*, g, I) \\
-(\mathbb{S}', \mathbb{D}') & \equiv & \begin{cases}
-(\mathbb{S}^*, \mathbb{D}) & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
-(\mathbb{S}^{**}, \mathbb{D}^{**}) & \text{otherwise}
+(\mathbb{S}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, g, I) \\
+\mathbb{S}' & \equiv & \begin{cases}
+\mathbb{S}^* & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
+\mathbb{S}^{**} & \text{otherwise}
 \end{cases} \\
@@
 I'_v & \equiv & v \\
-I'_C & \equiv & \mathbb{D}[\mathbb{S}[r]_{code}]
+\text{Let} \mathcal{S}(I'_C) & = & \mathbb{S}[r]_{code}
 \end{eqnarray}
+
+We assert that the state database contains the entry $(\mathcal{S}(I'_C), I'_C)$.
 
@@
 &&&& $\mathbf{i} \equiv \mathbb{M}_M[ \mathbb{M}_S[3] \dots (\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) ]$ \\
-&&&& $(\mathbb{S}', \mathbb{D}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}, \mathbb{D}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
+&&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
 \midrule
@@
 &&&& $\mathbf{o} \equiv \mathbb{M}_M[ \mathbb{M}_S[5] \dots (\mathbb{M}_S[5] + \mathbb{M}_S[6] - 1) ]$ \\
-&&&& $(\mathbb{S}', \mathbb{D}', \mathbb{M}'_g, \mathbf{o}) \equiv \Theta(\mathbb{S}, \mathbb{D}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbb{M}_S[1], \mathbb{M}_S[2], \mathbf{i})$ \\
+&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \mathbb{M}_S[2] = 0 \\ \mathbb{M}_g - \min(\mathbb{M}_g, \mathbb{M}_S[2]) & \text{otherwise} \end{cases}$ \\
+&&&& $(\mathbb{S}', g', \mathbf{o}) \equiv \Theta(\mathbb{S}, I_a, I_o, \mathbb{M}_S[0], g, I_p, \mathbb{M}_g - g_k, \mathbb{M}_S[1], \mathbf{i})$ \\
+&&&& $\mathbb{M}'_g \equiv g' + g_k$ \\
 \midrule

### 2014-04-06 Paper.tex
Complete execution model.

@@
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $c$, (in wei) required in up-front payment;
+\item the sender account balance contains at least the cost, $v_0$, (in wei) required in up-front payment;
 \end{enumerate}
@@
 g_0 & \equiv & \begin{cases}
-  |T_\mathbf{d}| G_{data} + G_{call} & \quad \text{call} \\
-  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \quad \text{create} \\
+  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \text{if} \quad T_r = 0 \\
+  |T_\mathbf{d}| G_{data} + G_{call} & \text{otherwise} \\
 \end{cases}
@@
 \begin{equation}
-v_0 \equiv (T_g + g_0) T_p + T_v
-\end{equation}
+v_0 \equiv T_g T_p + T_v
+\end{equation}
+
+the validity is determined as:
+
+\begin{eqnarray}
+S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_{nonce} \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_{balance}
+\end{eqnarray}
 
@@
 
-Evaluating $\mathbb{S}_C$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
+Evaluating $\mathbb{S}_0$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
 
@@
 (\mathbb{S}_P, g') \equiv \begin{cases}
-\Lambda(\mathbb{S}_0, S(T), T_o, T_g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
-\Theta_{0,1}(\mathbb{S}_0, S(T), T_o, T_r, T_g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
+\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
+\Theta^*(\mathbb{S}_0, S(T), T_o, T_r, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{0,1}$ to take only the first two components of the function's value; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+where
+
+\begin{equation}
+g \equiv T_g - g_0
+\end{equation}
+
+Note we use $\Theta^*$ to take only the first two components of the function's value; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' p
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p
 \end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_C$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
-
-\begin{equation}
-(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_C, s, o, g, p, v, \mathbf{i}, \mathbf{b})
-\end{equation}
-
-TODO: Move rest to own section after execution model.
-
-TODO: Transaction structure layout.
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
+
+\begin{equation}
+(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_0, s, o, g, p, v, \mathbf{i}, \mathbf{b})
+\end{equation}
 
@@
 A(s, o) & \equiv & a \quad \text{where:} \\
-a' & \equiv & \mathcal{L}_{160}(\mathcal{S}(\mathcal{R}([s, o]))) \\
-a & \equiv & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_C[x] = \varnothing
-\end{eqnarray}
-
-where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{L}_y(X)$ evaluates to the rightmost $y$ bits of the binary data $X$ and $\mathbb{S}_C[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a' & \equiv & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, o]))) \\
+a & \equiv & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing
+\end{eqnarray}
+
+where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\mathbb{S}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 
@@
 \mathbb{S}_P & \equiv & \begin{cases}
-\mathbb{S}_C & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
+\mathbb{S}_0 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
 \mathbb{S}^{**} & \text{otherwise}
 \end{cases} \\
-I'_a & \equiv & a \\
-I'_o & \equiv & s \\
-I'_p & \equiv & p \\
-I'_D & \equiv & [] \\
-I'_s & \equiv & s \\
-I'_v & \equiv & v \\
-I'_C & \equiv & \mathbf{i}
-\end{eqnarray}
-
-$I'_D$ evaluates to the empty array. $I'_H$ and $I'_n$ have no special treatment and are determined from the blockchain.
+I_a & \equiv & a \\
+I_o & \equiv & s \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & [] \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_\mathbf{b} & \equiv & \mathbf{i}
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty array. $I_H$ and $I_n$ have no special treatment and are determined from the blockchain.
 
@@
 \begin{equation}
-(\mathbb{S}_P, g', \mathbf{o}) \equiv \Theta(\mathbb{S}_C, s, o, r, g, p, v, \mathbf{d})
-\end{equation}
-
+(\mathbb{S}_P, g', \mathbf{o}) \equiv \Theta(\mathbb{S}, s, o, r, g, p, v, \mathbf{d})
+\end{equation}
+
+We define $\mathbb{S}_1$, the rollback state as the original state but with the value transferred to the recipient:
+
+\begin{eqnarray}
+\mathbb{S}_1 & \equiv & \mathbb{S} \quad \text{except:} \\
+\mathbb{S}_1[r]_{balance} & \equiv & \mathbb{S}_1[r]_{balance} + v
+\end{eqnarray}
+
+If the recipient account contains no code to be executed (i.e.
+it's not a contract), then we define the final state as being equivalent to the rollback state.
 In the case that the recipient contains code to be executed (i.e.
@@
 \begin{eqnarray}
-(\mathbb{S}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, g, I) \\
 \mathbb{S}' & \equiv & \begin{cases}
-\mathbb{S}^* & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
+\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_{code} = \varnothing \\
+\mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
 \mathbb{S}^{**} & \text{otherwise}
 \end{cases} \\
-I'_a & \equiv & a \\
-I'_o & \equiv & o \\
-I'_p & \equiv & p \\
-I'_D & \equiv & \mathbf{d} \\
-I'_s & \equiv & s \\
-I'_v & \equiv & v \\
-\text{Let} \mathcal{S}(I'_C) & = & \mathbb{S}[r]_{code}
-\end{eqnarray}
-
-We assert that the state database contains the entry $(\mathcal{S}(I'_C), I'_C)$.
+(\mathbb{S}^{**}, g', \mathbf{o}) & \equiv & \Xi(\mathbb{S}_1, g, I) \\
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+\text{Let} \quad \mathcal{S}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
+\end{eqnarray}
+
+We assert that the state database contains the entry $(\mathcal{S}(I_\mathbf{b}), I_\mathbf{b})$.
 
@@
 The memory size must always be a whole number of 32-byte words.
-Suppose that the highest previously accessed memory index is $i$, and the new index is $i'$; an additional $G_{m}(i, i')$ gas must be paid:
+Suppose that the highest previously accessed memory word index is $i$, and the new word index is $i'$; an additional $G_{m}(i, i')$ gas must be paid:
 
@@
 \item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
 \item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_C$, the byte array that is the machine code to be executed.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
 \item $I_H$, the block header of the previous block.
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying if the present state is a normal halting state of the machine:
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+We consider the empty sequence (denoted $[]$ to be non-equal to the empty set, denoted $\varnothing$).
 
@@
 \mathbb{M}_i & \equiv & -1 \\
-\mathbb{M}_T & \equiv & [] \\
-X(\mathbb{S}, \mathbb{M}, I) & \equiv & \begin{cases}
-(\mathbb{S}, \mathbb{M}, I) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
-O(\mathbb{S}, \mathbb{M}, I) & \text{if} \quad H(\mathbb{S}, \mathbb{M}, I)\\
+\mathbb{M}_T & \equiv & []
+\end{eqnarray}
+\begin{equation}
+X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
+(\mathbb{S}, \mathbb{M}, I, []) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
+(O(\mathbb{S}, \mathbb{M}, I), \mathbf{o}) & \text{if} \quad \mathbf{o} \neq \varnothing\\
 X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
 \end{cases}
-\end{eqnarray}
+\end{equation}
+
+where
+\begin{equation}
+\mathbf{o} \equiv H(\mathbb{S}, \mathbb{M}, I)
+\end{equation}
 
@@
 
-For the ease of reading, the instruction mnemonics, written in smallcaps (\eg {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
-
-\subsubsection{Exceptional Halting}
-
-For the purposes of defining $H$ and $O$, we define $o$ as the current operation to be executed:
-
-\begin{equation}
-o \equiv \begin{cases} I_C[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_C \rVert \\
+For the ease of reading, the instruction mnemonics, written in smallcaps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
+
+For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
+
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack movement items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-We also institute a function
-
-For calculating the exceptional halting function $Z$, we assume a secondary, instruction-specific halting function $Y$, which identifies edge cases specific to particular instructions (such as the division by zero or initiating a message call with insufficient funds).
-Thus $Z$ is defined as:
-
-\begin{equation}
-\begin{array} {lll}
-Z(\mathbb{S}, \mathbb{M}, I) & \equiv & Y(\mathbb{S}, \mathbb{M}, I) \\
-& & \vee \quad \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \\
-& & \vee \quad \lVert\mathbb{M}_T\rVert < \mathbf{\delta}_o
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is an instruction-specific issue, if there is insufficient gas or if there are insufficient stack items.
+
+\subsubsection{Exceptional Halting}
+
+The exceptional halting function $Z$ is defined as:
+\begin{equation}
+Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \lVert\mathbb{M}_T\rVert < \mathbf{\delta}_w
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas or if there are insufficient stack items.
+The astute read will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+
+\subsubsection{Normal Halting}
+
+We define the normal halting function $H$:
+\begin{equation}
+H(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
+H_{\text{\tiny RETURN}}(\mathbb{M}) & \text{if} \quad w = \text{\small RETURN} \\
+[] & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
+\varnothing & \text{otherwise}
+\end{cases}
+\end{equation}
+
+The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
-
 \begin{eqnarray}
 O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
-\Delta & \equiv & \mathbf{\alpha}_o - \mathbf{\delta}_o \\
+\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
 \lVert\mathbb{M}'_{T}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
-\quad \forall i \in [\mathbf{\alpha}_o, \lVert\mathbb{M}'_T\rVert): \mathbb{M}'_{T_i} & \equiv & \mathbb{M}_{T_{i+\Delta}}
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_T\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
 \end{eqnarray}
@@
 The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$ which evaluates to the according value:
-
-\begin{eqnarray}
-\mathbb{M}'_{g} \equiv \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
-\mathbb{M}'_{pc} \equiv \begin{cases}
-J(\mathbb{M}) & \text{if} \quad o \in \{ \text{\small JUMP, JUMPI} \} \\
-\mathbb{M}_{pc} + p & \text{if} \quad o \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
+\begin{eqnarray}
+\mathbb{M}'_{g} & \equiv & \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
+\mathbb{M}'_{pc} & \equiv & \begin{cases}
+J_o(\mathbb{M}) & \text{if} \quad w \in \{ \text{\small JUMP, JUMPI} \} \\
+\mathbb{M}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
 \mathbb{M}_{pc} + 1 & \text{otherwise}
@@
 where $p$ is the byte size of the push instruction, defined as:
-
-\begin{equation}
-p \equiv o - \text{\small PUSH1} + 2
+\begin{equation}
+p \equiv w - \text{\small PUSH1} + 2
 \end{equation}
@@
 In general, we assume the memory and system state don't change:
-
 \begin{eqnarray}
@@
 
-However, instructions do typically alter one or several components of these values. Altered components listed by instruction are noted in Appendix \ref{app:vm}.
+However, instructions do typically alter one or several components of these values.
+Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$.
 
@@
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_D[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lVert \mathcal{I}_D \rVert$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \lVert \mathcal{I}_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
-&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$ \\
+&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
+&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
 \midrule
 0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$ \\
+&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_C[\mathbb{M}_{pc} + 1]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{b}[\mathbb{M}_{pc} + 1]$ \\
 &&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_C[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_C[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
 \bottomrule
@@
 0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $\mathbb{M}'_R = \mathbb{M}_M[ \mathbb{M}_T[0] \dots ( \mathbb{M}_T[0] + \mathbb{M}_T[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\mathbb{M}) \equiv \mathbb{M}_M[ \mathbb{M}_T[0] \dots ( \mathbb{M}_T[0] + \mathbb{M}_T[1] - 1 ) ]$ \\
+&&&& This has the effect of halting the execution at this point with output defined.\\
+&&&& See section \ref{ch:model}. \\
 \midrule
@@
 
-%0xf0: CREATE -5 +1
-%Immediately creates a contract where:
-%The endowment is given by S'[0]
-%The body code of the eventual closure is given by T'[ S'[1] ... ( S'[1] + S'[2] - 1 ) ]
-%The initialisation code of the eventual closure is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
-%(Thus the total number of bytes of the transaction data is given by S'[2] + S'[4].)
-%S[0] = A
-%where A is the address of the created contract or 0 if the creation failed.
-%Fees are deducted from the sender balance to pay for enough gas to complete the operation (i.e.
-contract creation fee + initial storage fee).
-If there was not enough gas to complete the operation, then all gas will be deducted and the operation fails.
-%0xf1: CALL -7 +1
-%Immediately executes a call where:
-%The recipient is given by S'[0], when interpreted as an address.
-%The value is given by S'[1]
-%The gas is given by S'[2]
-%The input data of the call is given by T'[ S'[3] ... ( S'[3] + S'[4] - 1 ) ]
-%(Thus the number of bytes of the transaction is given by S'[4].)
-%The output data of the call is given by T[ S'[5] ... ( S'[5] + MIN( S'[6], S[0] ) - 1 ) ]
-%If 0 gas is specified, transfer all gas from the caller to callee gas balance, otherwise transfer only the amount given. If there isn't enough gas in the caller balance, operation fails.
-%If the value is less than the amount in the caller's balance then nothing is executed and the S'[2] gas gets refunded to the caller.
-%See Appendix A for execution.
-%Add any remaining callee gas to the caller's gas.
-%S[0] = R
-%where R = 1 when the instrinsic return code of the call is true, R = 0 otherwise.
-
 \section{Low-level Lisp-like Language}\label{app:lll}

### 2014-04-06 Paper.tex
Execution model fixes.

@@
 &&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
+&&&& $\mathbb{M}'_T[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
+&&&& $x=A(I_a, \mathbb{S}[I_a]_{nonce})$, the address of the newly created contract, otherwise. \\
 \midrule
@@
 &&&& $\mathbb{M}'_g \equiv g' + g_k$ \\
+&&&& $\mathbb{M}'_T[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
+&&&& $x=1$ otherwise. \\
 \midrule

### 2014-04-07 Paper.tex
Uncles and what not.

@@
 
-\section{Proposal} \label{ch:proposal}
-
-\subsection{The Blockchain Paradigm} \label{ch:overview}
+\section{The Blockchain Paradigm} \label{ch:overview}
 
@@
 \mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, \mathbf{T}) \\
-\mathbf{T} & \equiv & [ T_0, T_1, ... ] \\
-\Pi(\mathbb{S}, \mathbf{T}) & \equiv & \Omega(\Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $\mathbf{T}$ is the series of transactions included in this block; and $\Pi$ is the block-level state-transition function.
+B & \equiv & (..., [ T_0, T_1, ... ]) \\
+\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 
-\subsubsection{Value}
+\subsection{Value}
 
@@
 
-\subsubsection{Which History?}
+\subsection{Which History?}
 
@@
 
-\subsection{Blocks, State and Transactions} \label{ch:bst}
+\section{Blocks, State and Transactions} \label{ch:bst}
 
@@
 
-\subsubsection{World State} \label{ch:state}
+\subsection{World State} \label{ch:state}
 
@@
 
-\subsubsection{Transaction} \label{ch:transaction}
+\subsection{The Transaction} \label{ch:transaction}
 
@@
 
-\subsubsection{Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions and a set of other block headers that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
+\subsection{The Block} \label{ch:block}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
 The block header contains several pieces of information:
@@
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block; formally $H_x$.
+\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block. Not formally used.
 \item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
@@
 
-The difficulty of a block $H$, is defined as $H_d$:
-
-\begin{equation}
-H_d \equiv  \begin{cases}
+The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle block headers (of the same format as above).
+Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (H, \mathbf{T}, \mathbf{U})
+\end{equation}
+
+We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes; the given transactions $\mathbf{T}$, when executed in order on the base state $\mathbb{S}$ result in a new state of the identity $H_r$:
+\begin{eqnarray}
+& & \mathcal{S}(\mathcal{R}(\mathbf{U})) \equiv H_u \\
+& \wedge & \mathcal{S}(\mathcal{R}(\mathbf{T})) \equiv H_t \\
+& \wedge & r(\Pi(\mathbb{S}, B)) \equiv H_r
+\end{eqnarray}
+
+where $r(\mathbb{S})$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of $\mathbb{S}$ with values encoded using RLP.
+
+\subsubsection{Block Header Validity}
+
+We define $P(B_H)$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': S(R(B')) = H_p
+\end{equation}
+
+The canonical difficulty of a block of header $H$, is defined as $D(H)$:
+\begin{equation}
+D(H) \equiv \begin{cases}
 2^{22} & \quad \text{if genesis block}\\
-H'_d + \frac{H'_d}{1024} & \quad \text{if $H_t < H'_t + 42$}\\
-H'_d - \frac{H'_d}{1024} & \quad \text{otherwise}\\
+{P(H)_H}_d + \frac{{P(H)_H}_d}{1024} & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
+{P(H)_H}_d - \frac{{P(H)_H}_d}{1024} & \quad \text{otherwise}\\
 \end{cases}
@@
 
-Where $H'$ is the previous block, $H_t$ is the timestamp of block $H$.
+$H_t$ is the timestamp of block $H$, must fulfill the relation:
+\begin{equation}
+H_t > {P(H)_H}_t
+\end{equation}
 
@@
 The nonce, $H_n$, must satisfy the relation:
-
 \begin{equation}
@@
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propogate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because satisfaction of this requirement both depends on the contents of the block and
-
-The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle block headers (of the same format as above).
-
-\subsection{Gas and Payment} \label{ch:payment}
+Because the nonce must satisfy this requirement, and because satisfaction of this requirement depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult, and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+
+Thus we are able to define the block header validity function $V(H)$:
+\begin{eqnarray}
+V(H) & \equiv & \mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
+& \wedge & H_d = D(H) \\
+& \wedge & H_t > {P(H)_H}_t
+\end{eqnarray}
+
+\section{Gas and Payment} \label{ch:payment}
 
@@
 
-\subsection{Transaction Execution} \label{ch:transactions}
+\section{Transaction Execution} \label{ch:transactions}
 
@@
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, (in wei) required in up-front payment;
+\item the sender account balance contains at least the cost, $v_0$, required in up-front payment;
 \end{enumerate}
@@
 
+The transaction must not require more gas to be utilised in this block than the global block limit of gas, specified at $10^6$.
+
 The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender ($S(T)$) is incremented by one and the balance is reduced by the up-front cost, $v_0$.
@@
 
-After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate.
-So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$.
+After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate and the Ether for the gas actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
+So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
 
@@
 \mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p
-\end{eqnarray}
-
-\subsection{Contract Creation} \label{ch:create}
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p \\
+\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_o[m]_{balance} + (g - g') T_p \\
+m & \equiv & {B_H}_b
+\end{eqnarray}
+
+\section{Contract Creation} \label{ch:create}
 
@@
 A(s, o) & \equiv & a \quad \text{where:} \\
-a' & \equiv & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, o]))) \\
-a & \equiv & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing
+a & = & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing \\
+a' & = & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, o])))
 \end{eqnarray}
@@
 
-\subsection{Message Call} \label{ch:call}
+\section{Message Call} \label{ch:call}
 
@@
 
-\subsection{Execution Model} \label{ch:model}
+\section{Execution Model} \label{ch:model}
 
@@
 
-\subsubsection{Basics}
+\subsection{Basics}
 
@@
 
-\subsubsection{Fees Overview}
+\subsection{Fees Overview}
 
@@
 
-\subsubsection{Execution Environment}
+\subsection{Execution Environment}
 
@@
 
-\subsubsection{Execution Overview}
+\subsection{Execution Overview}
 
@@
 
-\subsubsection{The Execution Cycle}
+\subsection{The Execution Cycle}
 
@@
 
-The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$ which evaluates to the according value:
+The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
 \begin{eqnarray}
@@
 
-\subsection{Blocktree to Blockchain} \label{ch:ghost}
+\section{Blocktree to Blockchain} \label{ch:ghost}
 
@@
 \begin{eqnarray}
-B_{td} & \equiv & P_{td} + B_d + \sum\limits_{U \in B_\mathbb{U}} U_d \\
-P & \equiv & B_P
-\end{eqnarray}
-
-As such given a block $B$, $B_{td}$ is its total difficulty, $B_P$ is its parent block, $B_d$ is its difficulty and $B_\mathbb{U}$ is its set of uncle blocks.
-
-\subsection{Block Finalisation} \label{ch:finalisation}
-
-Validate uncles.
-
-Reward everything.
-
-Compute proof-of-work.
+B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
+B' & \equiv & P(H) \\
+H & \equiv & B_H \\
+\end{eqnarray}
+
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
+
+\section{Block Finalisation} \label{ch:finalisation}
+
+The process of finalising a block involves three stages:
+
+\begin{enumerate}
+\item Validate (or, if mining, determine) uncles.
+\item Apply rewards.
+\item Verify (or, if mining, compute a valid) nonce.
+\end{enumerate}
+
+\subsubsection{Uncle Validation}
+
+The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:
+
+\begin{equation}
+\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; P(U) = P(P(B_H)) \; \wedge \; P(B_H) \neq B
+\end{equation}
+
+\subsubsection{Reward Application}
+
+The application of rewards to a block involves little more
 
@@
 
-\section{User Interface}
-
-Browser-style.
-
-Displays ETH balance (i.e. functions as a wallet).
-
-Tabs or perhaps MacOS style app icons.
-
-Display taken up with a single app.
-
-Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
-
-Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
-
-Preconfigured with home app that is the Eth app store.
-
-Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
+%\section{User Interface}
+%
+%Browser-style.
+%Displays ETH balance (i.e. functions as a wallet).
+%Tabs or perhaps MacOS style app icons.
+%Display taken up with a single app.
+%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
+%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
+%Preconfigured with home app that is the Eth app store.
+%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
 
@@
 For a more secure pseudo-random offering both parties agree on a number of random data feed contracts; these are concatenated along with the block timestamp and hashed to produce the first number in the series.
+
+\section{Future Directions} \label{ch:future}
+
+Rether than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.
 
@@
 
-TODO: Specify minimum memory limits for a general cost definition.
+The general gas cost function is defined as:
+
+\begin{equation}
+C(\mathbb{S}, \mathbb{M}) \equiv C_M(\mathbb{S}, \mathbb{M}) + \begin{cases}
+c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
+G_{call} & \text{if} \quad o = \text{\small CALL}\\
+G_{create} & \text{if} \quad o = \text{\small CREATE}\\
+G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
+G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
+G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
+G_{stop} & \text{if} \quad o = \text{\small STOP}\\
+G_{step} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory.
 
@@
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lVert \mathcal{I}_\mathbf{d} \rVert$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_p$ \\
+&&&& $\mathbb{M}'_T[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_{H_{prevHash}}$ \\
-&&&& $I_{H_{prevHash}}$ is the previous block's hash. \\
+&&&& $\mathbb{M}'_T[0] \equiv {I_H}_p$ \\
+&&&& $I_{H_p}$ is the previous block's hash. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_{H_{coinbase}}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv {I_H}_b$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_{H_{timestamp}}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv {I_H}_t$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_{H_{number}}$ \\
-&&&& $I_{H_{number}}$ is the total number of ancestor blocks of the current block. \\
+&&&& $\mathbb{M}'_T[0] \equiv n({I_H})$ \\
+&&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_{H_{difficulty}}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv {I_H}_d$ \\
 \midrule
@@
 0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[\mathbb{M}_T[0]]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_T[0]]$ \\
 \midrule
 0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\mathbb{M}'_S[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
+&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
+&&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
+2G_{sstore} & \text{if} \quad \mathbb{M}_T[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] = 0 \\
+0 & \text{if} \quad \mathbb{M}_T[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] \neq 0 \\
+G_{sstore} & \text{otherwise}
+\end{cases}$ \\
 \midrule

### 2014-04-07 Paper.tex
EVM gas cost for memory.

@@
 \end{itemize}
-\item {\small CALL} costs $G_{call} + g$, where $g$ is the quantity of gas provided; some gas may later be refunded.
-\item {\small CREATE} costs $G_{create} + g$ gas, where $g$ is the quantity of gas provided.
+\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of contract body code.
+\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of contract initialisation code.
 \item All other operations cost $G_{step}$ gas.
@@
 
-Additionally, when memory is accessed with {\small MSTORE}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CALLDATA} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
-The memory size must always be a whole number of 32-byte words.
-Suppose that the highest previously accessed memory word index is $i$, and the new word index is $i'$; an additional $G_{m}(i, i')$ gas must be paid:
-
-\begin{equation}
-G_{m}(i, i') \equiv \max(0, \ceil[\Bigg]{\frac{i'+1}{32}} - \ceil[\Bigg]{\frac{i+1}{32}}) G_{memory}
-\end{equation}
-
-Originally, with no accesses taken place, the highest accessed index of memory $i$ is defined as $-1$.
+Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+See Appendix \ref{app:vm} for a rigourous definition of the EVM gas cost.
 
@@
 \mathbb{M}_M & \equiv & [0, 0, ...] \\
-\mathbb{M}_i & \equiv & -1 \\
-\mathbb{M}_T & \equiv & []
+\mathbb{M}_i & \equiv & 0 \\
+\mathbb{M}_S & \equiv & []
 \end{eqnarray}
@@
 
-The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the highest used memory index, and the stack contents.
+The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.
@@
 \begin{equation}
-Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \lVert\mathbb{M}_T\rVert < \mathbf{\delta}_w
+Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
 \end{equation}
@@
 
-The application of rewards to a block involves little more
+The application of rewards to a block involves placing in the account of the miner-determined coinbase address additional Ether according to
 
@@
 \begin{equation}
-C(\mathbb{S}, \mathbb{M}) \equiv C_M(\mathbb{S}, \mathbb{M}) + \begin{cases}
+C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
 c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
@@
 
-where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory.
+where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] + \mathbb{M}_T[1]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \times \mathbb{M}_T[1]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] - \mathbb{M}_T[1]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor$ \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_T[0] \div \mathbb{M}_T[1]\rfloor$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor$ \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \mod{ \mathbb{M}_S[1] }$ \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] \mod{ \mathbb{M}_T[1] }$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \mod{ \mathbb{M}_S[1] }$ \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0] ^ {\mathbb{M}_T[1] }$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
 \midrule
 0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_T[0]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_S[0]$ \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] < \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] > \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = \mathbb{M}_T[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_T[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \end{tabular*}
@@
 \multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
-\multicolumn{5}{l}{$\mathbb{M}_T[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_T[0]$} \vspace{5pt} \\
+\multicolumn{5}{l}{$\mathbb{M}_S[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_S[0]$} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
 0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \wedge \mathbb{M}_T[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
 \midrule
 0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \vee \mathbb{M}_T[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
 \midrule
 0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_T[0]_i \oplus \mathbb{M}_T[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
 \midrule
 0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_T[1]_{(i + 8\mathbb{M}_T[0])} & \text{if} \quad \mathbb{M}_T[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 &&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) \div 32 })$ \\
 \bottomrule
@@
 0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0]$ \\
 \midrule
 0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[1]$ \\
-&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_T[0]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_S[0]$ \\
 \midrule
 0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 31) \div 32 })$ \\
 \midrule
 0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ] \equiv \mathbb{M}_T[1]$ \\
+&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 31) \div 32 })$ \\
 \midrule
 0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] ] \equiv (\mathbb{M}_T[1] \mod 256) $ \\
+&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] ] \equiv (\mathbb{M}_S[1] \mod 256) $ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \mathbb{M}_S[0] \div 32 })$ \\
 \midrule
 0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_T[0]]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
 \midrule
 0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
+&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_S[0] ] \equiv \mathbb{M}_S[1] $ \\
 &&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
-2G_{sstore} & \text{if} \quad \mathbb{M}_T[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] = 0 \\
-0 & \text{if} \quad \mathbb{M}_T[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] \neq 0 \\
+2G_{sstore} & \text{if} \quad \mathbb{M}_S[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] = 0 \\
+0 & \text{if} \quad \mathbb{M}_S[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] \neq 0 \\
 G_{sstore} & \text{otherwise}
@@
 0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
+&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_S[0] $ \\
 &&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
@@
 0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
@@
 &&&& $x=A(I_a, \mathbb{S}[I_a]_{nonce})$, the address of the newly created contract, otherwise. \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[1] + \mathbb{M}_S[2] - 1, \mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) \div 32 })$ \\
 \midrule
@@
 &&&& $x=1$ otherwise. \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1, \mathbb{M}_S[5] + \mathbb{M}_S[6] - 1) \div 32 })$ \\
 \midrule
 0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\mathbb{M}) \equiv \mathbb{M}_M[ \mathbb{M}_T[0] \dots ( \mathbb{M}_T[0] + \mathbb{M}_T[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\mathbb{M}) \equiv \mathbb{M}_M[ \mathbb{M}_S[0] \dots ( \mathbb{M}_S[0] + \mathbb{M}_S[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) \div 32 })$ \\
 \midrule
 0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
-&&&& $\mathbb{S}'[\mathbb{M}_T[0]]_{balance} \equiv \mathbb{S}[\mathbb{M}_T[0]]_{balance} + \mathbb{S}[I_a]_{balance}$ \\
+&&&& $\mathbb{S}'[\mathbb{M}_S[0]]_{balance} \equiv \mathbb{S}[\mathbb{M}_S[0]]_{balance} + \mathbb{S}[I_a]_{balance}$ \\
 &&&& $\mathbb{S}'[I_a] \equiv \varnothing$ \\

### 2014-04-07 Paper.tex
Formal stuff mostly complete.

@@
 \definecolor{lightyellow}{rgb}{1,0.98,0.9}
+\definecolor{lightpink}{rgb}{1,0.94,0.95}
 
@@
 \pagecolor{lightyellow}
+%\pagecolor{lightpink}
 
@@
 
-and $f$ is the Ethereum state transition function.
+where $\Upsilon$ is the Ethereum state transition function.
 In Ethereum, $\Upsilon$, together with $\mathbb{S}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\mathbb{S}$ allows components to store arbitrary state between transactions.
@@
 \begin{eqnarray}
-\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, \mathbf{T}) \\
+\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, B) \\
 B & \equiv & (..., [ T_0, T_1, ... ]) \\
@@
 
-This is the basis of the blockchain paradigm, a model that forms the backbone of Ethereum.
+This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
 
@@
 In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as ETH.
-The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted is the Wei.
+To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH}.
+The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
 One Ether is defined as being $10^{18}$ Wei.
@@
 \begin{center}
-\begin{tabular}{lr}
+\begin{tabular}{rl}
 \toprule
@@
 
-Throughout the document, any reference to value, in the context of Ether, currency, a balance or a payment, should assumed to be counted in Wei.
+Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.
 
@@
 
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block the resultant structure is necessarily a tree of blocks.
+Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
 In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
@@
 
-The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP).
+The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
 Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
@@
 Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\mathbb{S}$ the state:
-
 \begin{equation}
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-
-\begin{eqnarray}
-g_0 & \equiv & \begin{cases}
+\begin{equation}
+g_0 \equiv \begin{cases}
   (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \text{if} \quad T_r = 0 \\
@@
 \end{cases}
-\end{eqnarray}
+\end{equation}
 
@@
 The up-front cost $v_0$ is calculated as:
-
 \begin{equation}
@@
 the validity is determined as:
-
 \begin{eqnarray}
@@
 We define the rollback state $\mathbb{S}_0$:
-
 \begin{eqnarray}
@@
 Evaluating $\mathbb{S}_0$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
-
 \begin{equation}
@@
 where
-
 \begin{equation}
@@
 
-Note we use $\Theta^*$ to take only the first two components of the function's value; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
-
-After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate and the Ether for the gas actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
+Note we use $\Theta^*$ to denote the fact that only the first two components of the function's value are taken; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+
+After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate.
+The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
 So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
-
-\begin{eqnarray}
-\mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p \\
-\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_o[m]_{balance} + (g - g') T_p \\
+\begin{eqnarray}
+\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
+\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (g - g') T_p \\
 m & \equiv & {B_H}_b
+\end{eqnarray}
+
+In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+\begin{eqnarray}
+\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
 \end{eqnarray}
@@
 The word size of the machine (and thus size of stack item) is 256-bit.
-This was chosen to facilitate the SHA3-256 hash scheme and ECC.
+This was chosen to facilitate the SHA3-256 hash scheme and elliptic-curve computations.
 The memory model is a simple word-addressed byte array.
@@
 \item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of contract initialisation code.
+\item {\small STOP} costs $G_{stop}$ gas
 \item All other operations cost $G_{step}$ gas.
@@
 However, instructions do typically alter one or several components of these values.
-Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$.
+Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and formal a description of the gas requirements.
 
@@
 Thus we define the total difficulty of block $B$ recursively as:
-
 \begin{eqnarray}
@@
 \item Apply rewards.
-\item Verify (or, if mining, compute a valid) nonce.
+\item Verify (or, if mining, compute a valid) state and nonce.
 \end{enumerate}
@@
 The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:
-
 \begin{equation}
@@
 
-The application of rewards to a block involves placing in the account of the miner-determined coinbase address additional Ether according to
+The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
+We raise the block's coinbase account by $R_b$, the block reward, and the coinbase of each uncle by $\frac{7}{8}$ of that.
+Formally we define the function $\Omega$:
+\begin{eqnarray}
+\Omega(B, \mathbb{S}) & \equiv & \mathbb{S}': \mathbb{S}' = \mathbb{S} \quad \text{except:} \\
+\mathbb{S}'[{B_H}_b] & = & \mathbb{S}[{B_H}_b] + R_b \\
+\forall U \in B_\mathbf{U}: \quad \mathbb{S}'[U_b] & = & \mathbb{S}[U_b] + \frac{7}{8}R_b
+\end{eqnarray}
+
+We define the block reward as 1500 Finney:
+\begin{equation}
+\text{Let} \quad R_b = 1.5^{16}
+\end{equation}
+
+\subsubsection{State \& Nonce Validation}
+
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+\begin{equation}
+\Gamma(B) \equiv \begin{cases}
+\mathbb{S}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\mathbb{S}_i: r(\mathbb{S}_i) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $r(\mathbb{S}_i)$ means the hash of the root node of a trie of state $\mathbb{S}_i$; this is stored in the state database since the trie is a mutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\begin{eqnarray}
+\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
+B'_n & = & n: \quad \mathcal{P}(B^*, n) < \frac{2^{256}}{H_d} \\
+B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
+\end{eqnarray}
+
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined:
+\begin{eqnarray}
+\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...) \\
+B & \equiv & (H, [ T_0, T_1, ... ], \mathbf{U})
+\end{eqnarray}
+
+Thus the complete block-transition mechanism, less $\mathcal{P}$, the proof-of-work function is defined.
 
@@
 
-For the PoC series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder. It utilises the SHA3 hash function to
+For the PoC series, we use a simplified proof-of-work.
+This is not ASIC resistant and is meant merely as a placeholder.
+It utilises the bare SHA3 hash function to secure the block chain by requiring the SHA3 hash of the concatenation of the nonce and the header's SHA3 hash to be sufficiently low.
 
 It is formally defined as $\mathcal{P}$:
-
 \begin{equation}
@@
 Rether than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.
+
+State database won't keep everything; maintains an age and eventually throws out nodes that are not recent enough or checkpoints.
+
+Throw out nodes in state that haven't sent/received a transaction in over X blocks.
 

### 2014-04-07 Paper.tex
Fixes.

@@
 
-Transparency. Open state. User can be guaranteed that backend will do what front-end says it will. Parties can analyse and be wholly confident in outcomes given events.
+Incorruptability of judgement. Transparency. Open state. User can be guaranteed that backend will do what front-end says it will. Parties can analyse and be wholly confident in outcomes given events.
 
@@
 
-The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle block headers (of the same format as above).
+The other two components in the block are simply a list of uncle block headers (of the same format as above) and a list of the transactions (as defined previously).
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (H, \mathbf{T}, \mathbf{U})
+B \equiv (H, \mathbf{U}, \mathbf{T})
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': S(R(B')) = H_p
+P(H) \equiv B': \mathcal{S}(\mathcal{R}(B')) = H_p
 \end{equation}
@@
 We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
-
 \begin{equation}
@@
 Thus we define the creation address function $A$:
-
-\begin{eqnarray}
-A(s, o) & \equiv & a \quad \text{where:} \\
+\begin{eqnarray}
+A(s, n) & \equiv & a \quad \text{where:} \\
 a & = & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing \\
-a' & = & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, o])))
+a' & = & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, n])))
 \end{eqnarray}
@@
 where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\mathbb{S}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
 The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\mathbb{S}^*$:
-
 \begin{eqnarray}
 \mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}^*[A(s, o)] & \equiv & ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
+\mathbb{S}^*[A(s, \mathbb{S}[s]_{nonce})] & \equiv & ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
 \end{eqnarray}
@@
 Thus formally, we may specify the new state and gas as $\mathbb{S}'$, $g'$ where:
-
 \begin{eqnarray}
@@
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
-
 \begin{equation}
@@
 We define $\mathbb{S}_1$, the rollback state as the original state but with the value transferred to the recipient:
-
 \begin{eqnarray}
@@
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
-
 \begin{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_b = 1.5^{16}
+\text{Let} \quad R_b = 1.5 * 10^{16}
 \end{equation}
@@
 &&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
 \bottomrule
@@
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ]$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 31) \div 32 })$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ]$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
 \midrule
 0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 31) \div 32 })$ \\
+&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ] \equiv \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
 \midrule
@@
 &&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] ] \equiv (\mathbb{M}_S[1] \mod 256) $ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \mathbb{M}_S[0] \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 1) \div 32 })$ \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \mathbb{M}_M[ \mathbb{M}_S[3] \dots (\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) ]$ \\
-&&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
+&&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}^*, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
+&&&& $\mathbb{S}^* \equiv \mathbb{S} \quad \text{except} \quad \mathbb{S}^*[I_a]_{nonce} = \mathbb{S}[I_a]_{nonce} + 1$ \\
 &&&& $\mathbb{M}'_T[0] \equiv x$ \\
@@
 &&&& $x=A(I_a, \mathbb{S}[I_a]_{nonce})$, the address of the newly created contract, otherwise. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[1] + \mathbb{M}_S[2] - 1, \mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[1] + \mathbb{M}_S[2], \mathbb{M}_S[3] + \mathbb{M}_S[4]) \div 32 })$ \\
 \midrule
@@
 &&&& $x=1$ otherwise. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1, \mathbb{M}_S[5] + \mathbb{M}_S[6] - 1) \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[3] + \mathbb{M}_S[4], \mathbb{M}_S[5] + \mathbb{M}_S[6]) \div 32 })$ \\
 \midrule
@@
 &&&& See section \ref{ch:model}. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
 \midrule
@@
 
-$[z_{256}, \mathcal{S}(\mathcal{R}([])), z_{160}, stateRoot, \mathcal{S}(\mathcal{R}([])), 2^{22}, 0, \text{`'}, 42]$
+$\langle 0_{256}, \mathcal{S}(\mathcal{R}([])), 0_{160}, 0_{256}, \mathcal{S}(\mathcal{R}([])), 2^{22}, 0, [], 42 \rangle$
 
@@
 
-$z_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes.
-
-$z_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes.
+$0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes.
+
+$0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes.
 
@@
 
-`' refers to the extradata, an empty byte array.
+$[]$ refers to the extradata, an empty byte array.
 
@@
 
-$stateRoot$ refers to the state root.
-
 \end{document}

### 2014-04-07 Paper.tex
Cleanups and additional text.

@@
 \usepackage{url}
+\usepackage{cancel}
 \usepackage{xspace}
@@
 \usepackage{afterpage}
+\usepackage{tikz}
+
+\newcommand{\hcancel}[1]{%
+    \tikz[baseline=(tocancel.base)]{
+        \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
+        \draw[black] (tocancel.south west) -- (tocancel.north east);
+    }%
+}%
 
@@
 
-Incorruptability of judgement. Transparency. Open state. User can be guaranteed that backend will do what front-end says it will. Parties can analyse and be wholly confident in outcomes given events.
-
-\subsection{Previous Work} \label{ch:previous}
-
-E language. Smart contracts. Bitcoin, namecoin \&c. Mastercoin.
+There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
+This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetance, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
+By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
+
+Dealings in this proposed system would have several attributes not often found in the real word.
+The incorruptibility of judgement, often difficult to find, comes naturally from a disinterested algorithmic interpreter.
+Transparency, through being about to see exactly how a state or judgement came about through the transaction log and rules or instructional codes never happens perfectly in human-based systems.
+
+Overall, we wish to provide a system such that the user can be guaranteed that any other individuals, systems or organisations that they interact with, they can do so with absolute confidence in the possible outcomes and how they might come about.
+
+%\subsection{Previous Work} \label{ch:previous}
+%E language. Smart contracts. Bitcoin, namecoin \&c. Mastercoin.
 
@@
 \begin{eqnarray}
-& & \mathcal{S}(\mathcal{R}(\mathbf{U})) \equiv H_u \\
-& \wedge & \mathcal{S}(\mathcal{R}(\mathbf{T})) \equiv H_t \\
+& & \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(\mathbf{U})) \equiv H_u \\
+& \wedge & \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(\mathbf{T})) \equiv H_t \\
 & \wedge & r(\Pi(\mathbb{S}, B)) \equiv H_r
@@
 \begin{equation}
-P(H) \equiv B': \mathcal{S}(\mathcal{R}(B')) = H_p
+P(H) \equiv B': \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(B')) = H_p
 \end{equation}
@@
 \begin{equation}
-\mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $\mathcal{P}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
+\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
 Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
+V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
 & \wedge & H_d = D(H) \\
@@
 a & = & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing \\
-a' & = & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, n])))
-\end{eqnarray}
-
-where $\mathcal{S}$ is the SHA3 256-bit hash function, $\mathcal{R}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\mathbb{S}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;\langle s, n\rangle\;\big)\Big)\Big)
+\end{eqnarray}
+
+where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\mathbb{S}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}^*[A(s, \mathbb{S}[s]_{nonce})] & \equiv & ( 0, v, \varnothing, \mathcal{S}(\mathbf{b}) )
-\end{eqnarray}
-
-It is asserted that the state database will also change such that it defines the pair $(\mathcal{S}(\mathbf{b}), \mathbf{b})$.
+\mathbb{S}^*[A(s, \mathbb{S}[s]_{nonce})] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
+\end{eqnarray}
+
+It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
 
@@
 I_v & \equiv & v \\
-\text{Let} \quad \mathcal{S}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
-\end{eqnarray}
-
-We assert that the state database contains the entry $(\mathcal{S}(I_\mathbf{b}), I_\mathbf{b})$.
+\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
+\end{eqnarray}
+
+We assert that the state database contains the entry $(\mathtt{\tiny SHA3}(I_\mathbf{b}), I_\mathbf{b})$.
 
@@
 B' & \equiv & P(H) \\
-H & \equiv & B_H \\
+H & \equiv & B_H
 \end{eqnarray}
@@
 
-\subsubsection{Uncle Validation}
+\subsection{Uncle Validation}
 
@@
 
-\subsubsection{Reward Application}
+\subsection{Reward Application}
 
@@
 
-\subsubsection{State \& Nonce Validation}
+\subsection{State \& Nonce Validation}
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathcal{P}(B^*, n) < \frac{2^{256}}{H_d} \\
+B'_n & = & n: \quad \mathtt{PoW}(B^*, n) < \frac{2^{256}}{H_d} \\
 B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
@@
 
-Thus the complete block-transition mechanism, less $\mathcal{P}$, the proof-of-work function is defined.
+Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
 
@@
 
-More formally, the proof-of-work function takes the form of $\mathcal{P}$:
-
-\begin{equation}
-\mathcal{P}(H, n) < \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H$ is the new block's header \textit{without} the nonce entry; $H_d$ is the difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
+More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
+\begin{equation}
+\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
 
@@
 
-It is formally defined as $\mathcal{P}$:
-\begin{equation}
-\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \circ n))
+It is formally defined as $\mathtt{PoW}$:
+\begin{equation}
+\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny SHA3}(\mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
 \end{equation}
@@
 where:
-$\mathcal{R}(H)$ is the RLP encoding of the $H$;
-$\mathcal{S}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
+$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
+$\mathtt{\tiny SHA3}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
 $n$ is the nonce, a series of 32 bytes;
 $\circ$ is the series concatenation operator;
-$\mathcal{B}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
+$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
 
@@
 
-Rether than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.
+Rather than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.
 
@@
 
-Throw out nodes in state that haven't sent/received a transaction in over X blocks.
+Nodes in state trie that haven't sent/received a transaction in over X blocks could be thrown out, reducing Ether-leakage and reducing the growth of the state database.
 
@@
 \section{Recursive Length Prefix}\label{app:rlp}
-
 This is a method for encoding arbitrarily structured binary data (byte arrays).
@@
 \section{Modified Merkle Patricia Tree}\label{app:trie}
-
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
@@
 \section{Hex-Prefix Encoding}\label{app:hexprefix}
-
 Hex-prefix encoding is an efficient method of encoding a arbitrary number of nibbles as a byte array.
@@
 \section{Formal Specification of Structures}
-
 Though RLP is data-agnostic, it does specify a canonical representation for integer quantities.
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathcal{S}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
 &&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
@@
 
-\section{Low-level Lisp-like Language}\label{app:lll}
-
-The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
 
 \section{Wire Protocol}\label{app:wire}
-
-The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transfering information such as blocks \& transactions
-
-\section{Peer Strategy}\label{app:peers}
+The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transfering information such as blocks \& transactions.
+
+%\section{Peer Strategy}\label{app:peers}
 
@@
 
-$\langle 0_{256}, \mathcal{S}(\mathcal{R}([])), 0_{160}, 0_{256}, \mathcal{S}(\mathcal{R}([])), 2^{22}, 0, [], 42 \rangle$
+$\langle 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 2^{22}, 0, [], 42 \rangle$
 
@@
 
-$\mathcal{S}(\mathcal{R}([]))$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.
+$\mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([]))$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.
 

### 2014-04-08 Paper.tex
Changed Div &c. to allow doc/0.

@@
 This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
-Follow-up systems such as namecoin adapted this original ``currency application'' of the technology into other applications allbeit rather simplistic ones.
+Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications allbeit rather simplistic ones.
 
@@
 
-The machine can have exceptional execution for several reasons, including stack underflows, invalid instructions and divides by zero.
+The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
 These unambiguously and validly result in immediate halting of the machine with all state changes left intact.
@@
 This states that the execution is in an exceptional halting state if there is insufficient gas or if there are insufficient stack items.
-The astute read will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
 
@@
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \mod{ \mathbb{M}_S[1] }$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \mod{ \mathbb{M}_S[1] }$ \\
+&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\

### 2014-04-08 Paper.tex
Tweaks.

@@
 \begin{equation}
-Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas or if there are insufficient stack items.
+Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined) or if there are insufficient stack items.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.

### 2014-04-15 Paper.tex
Tweaks.

@@
     Dr. Gavin Wood\\
-    CTO, Ethereum Project\\
+    Co-Founder \& CTO, Ethereum Project\\
     gavin@ethereum.org
@@
 There exist other subdenominations of Ether:
-
-\par\bigskip
+\par`
 \begin{center}
@@
 2^{22} & \quad \text{if genesis block}\\
-{P(H)_H}_d + \frac{{P(H)_H}_d}{1024} & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
-{P(H)_H}_d - \frac{{P(H)_H}_d}{1024} & \quad \text{otherwise}\\
+{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
+{P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{otherwise}\\
 \end{cases}
@@
 g_0 \equiv \begin{cases}
-  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \text{if} \quad T_r = 0 \\
-  |T_\mathbf{d}| G_{data} + G_{call} & \text{otherwise} \\
+g_d + G_{create} & \text{if} \quad T_r = 0 \\
+g_d + G_{call} & \text{otherwise} \\
 \end{cases}
@@
 
-where $|T_\mathbf{d}|$, $|T_\mathbf{i}|$ and $|T_\mathbf{b}|$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
+where
+
+\begin{equation}
+g_d \equiv \begin{cases} (\lVert T_\mathbf{i}\rVert + \lVert T_\mathbf{b}\rVert) G_{data} & \text{if} \quad T_r = 0 \\
+ \lVert T_\mathbf{d}\rVert G_{data} & \text{otherwise}
+ \end{cases}
+\end{equation}
+
+where $\lVert T_\mathbf{d}\rVert$, $\lVert T_\mathbf{i}\rVert$ and $\lVert T_\mathbf{b}\rVert$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
+The portion of $g_0$ that will not be paid to the miner is $g_d$, and is removed from the system.
+
+% Explain g_d reason?
 
@@
 \mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - c \\
+\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - v_0 \\
 \mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
@@
 \mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
-\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (g - g') T_p \\
+\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (T_g - g' - g_d) T_p \\
 m & \equiv & {B_H}_b
@@
 
-In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
-\begin{eqnarray}
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
-\end{eqnarray}
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
+%\end{eqnarray}
 
@@
 
-There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.
+There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.
 
@@
 \begin{equation}
-(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_0, s, o, g, p, v, \mathbf{i}, \mathbf{b})
+(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_0, s, n, g, p, v, \mathbf{i}, \mathbf{b})
 \end{equation}
@@
 \begin{equation}
-Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
+Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
 \end{equation}
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\mathbb{M}'_{T}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_T\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
+\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
 \end{eqnarray}
@@
 This paper has introduced, discussed and formally defined the protocol of Ethereum.
+
+\section{Acknowledgements}
+
+Vitalik Buterin first proposed in the original Ethereum whitepaper (first distributed in December 2013) moving from a single state-transaction function with trivial state to something much more general and expansive.
+Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan
 
@@
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
 \midrule
 0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
-&&&& $\mathbb{M}'_T[0] \equiv -\mathbb{M}_S[0]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv -\mathbb{M}_S[0]$ \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
-&&&& $\mathbb{M}'_T[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \end{tabular*}
@@
 0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
 \midrule
 0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
 \midrule
 0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_T[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
 \midrule
 0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..7]: \mathbb{M}'_T[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..7]: \mathbb{M}'_S[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
 &&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_a$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it's never a contract. \\
+&&&& $\mathbb{M}'_S[0] \equiv I_o$ \\
+&&&& This is the sender of original transaction; it is never a contract. \\
 \midrule
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_c$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_c$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_v$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_v$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_p$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
-&&&& $\mathbb{M}'_T[0] \equiv {I_H}_p$ \\
+&&&& $\mathbb{M}'_S[0] \equiv {I_H}_p$ \\
 &&&& $I_{H_p}$ is the previous block's hash. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\mathbb{M}'_T[0] \equiv {I_H}_b$ \\
+&&&& $\mathbb{M}'_S[0] \equiv {I_H}_b$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\mathbb{M}'_T[0] \equiv {I_H}_t$ \\
+&&&& $\mathbb{M}'_S[0] \equiv {I_H}_t$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\mathbb{M}'_T[0] \equiv n({I_H})$ \\
+&&&& $\mathbb{M}'_S[0] \equiv n({I_H})$ \\
 &&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
@@
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\mathbb{M}'_T[0] \equiv {I_H}_d$ \\
+&&&& $\mathbb{M}'_S[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\mathbb{M}'_T[0] \equiv 10^6$ \\
+&&&& $\mathbb{M}'_S[0] \equiv 10^6$ \\
 &&&& Presently, the gas limit is hard-coded in the protocol. \\
@@
 0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[0]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0]$ \\
 \midrule
 0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[1]$ \\
-&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_S[0]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_S[1] \equiv \mathbb{M}_S[0]$ \\
 \midrule
 0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
 &&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
@@
 0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 32) ] \equiv \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
 &&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
@@
 0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
 \midrule
@@
 0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[1] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
@@
 0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{pc}$ \\
 \midrule
 0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{i}$ \\
 \midrule
 0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
-&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
+&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{g}$ \\
 \bottomrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{b}[\mathbb{M}_{pc} + 1]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{b}[\mathbb{M}_{pc} + 1]$ \\
 &&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
+&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
 \bottomrule
@@
 &&&& $\mathbb{S}^* \equiv \mathbb{S} \quad \text{except} \quad \mathbb{S}^*[I_a]_{nonce} = \mathbb{S}[I_a]_{nonce} + 1$ \\
-&&&& $\mathbb{M}'_T[0] \equiv x$ \\
+&&&& $\mathbb{M}'_S[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
@@
 &&&& $\mathbb{M}'_g \equiv g' + g_k$ \\
-&&&& $\mathbb{M}'_T[0] \equiv x$ \\
+&&&& $\mathbb{M}'_S[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\

### 2014-04-15 Paper.tex
Switch Tx and Uncles

@@
 \begin{equation}
-B \equiv (H, \mathbf{U}, \mathbf{T})
+B \equiv (H, \mathbf{T}, \mathbf{U})
 \end{equation}
@@
 \section{Recursive Length Prefix}\label{app:rlp}
-This is a method for encoding arbitrarily structured binary data (byte arrays).
+This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
+Within RLP, integers are specified as the byte array equal to the big-endian representation of the value.
+Thus we may define the notation of a tree structure $\mathbf{t}$:
+
+\begin{equation}
+\mathbf{t}
+\end{equation}
+
+We define the function $\mathtt{\tiny RLP}$ that maps from an arbitrarily structured tree whose leaves are byte arrays into a single byte array.
 
@@
   blockHeader: [...]
+  txList: [ [...], [...], ... ]
   uncleList: [ [...], [...], ... ]
-  txList: [ [...], [...], ... ]
 ]

### 2014-04-17 Paper.tex
Incorrect timestamp reference.

@@
 2^{22} & \quad \text{if genesis block}\\
-{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
+{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{if $H_s < {P(H)_H}_s + 42$}\\
 {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{otherwise}\\
@@
 
-$H_t$ is the timestamp of block $H$, must fulfill the relation:
-\begin{equation}
-H_t > {P(H)_H}_t
+$H_s$ is the timestamp of block $H$, must fulfill the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 \section{Recursive Length Prefix}\label{app:rlp}
-This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
-Within RLP, integers are specified as the byte array equal to the big-endian representation of the value.
-Thus we may define the notation of a tree structure $\mathbf{t}$:
-
-\begin{equation}
-\mathbf{t}
-\end{equation}
-
-We define the function $\mathtt{\tiny RLP}$ that maps from an arbitrarily structured tree whose leaves are byte arrays into a single byte array.
+This is a serialisation method for encoding arbitrarily structured binary data (byte arrays). Within RLP, integers are specified as the byte array equal to the big-endian representation of the value.
+
+%Thus we define the set of possible structures $\mathbb{T}$:
+
+%\begin{eqnarray}
+%\text{Let} \mathbf{t} & = & \langle \mathbf{t}_0, \mathbf{t}_1, ... \rangle
+%\vee \mathbf{t} & = & 
+%\end{eqnarray}
+
+%We define the function $\mathtt{\tiny RLP}$ that maps from an arbitrarily structured tree whose leaves are byte arrays into a single byte array.
 

### 2014-04-17 Paper.tex
More timestamp fixes.
Fixes for Marko's bugs.

@@
 & \wedge & H_d = D(H) \\
-& \wedge & H_t > {P(H)_H}_t
+& \wedge & H_s > {P(H)_H}_s
 \end{eqnarray}
@@
 
-Evaluating $\mathbb{S}_0$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
+Evaluating $\mathbb{S}_P$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
 \begin{equation}
@@
 Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}'$ and available gas remaining $g'$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}^{**}$ and available gas remaining $g'$.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the new state and gas as $\mathbb{S}'$, $g'$ where:
+Thus formally, we may specify the resultant state and gas as $\mathbb{S}_P$, $g'$ where:
 \begin{eqnarray}
@@
 \begin{eqnarray}
-\mathbb{S}' & \equiv & \begin{cases}
+\mathbb{S}_P & \equiv & \begin{cases}
 \mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_{code} = \varnothing \\
@@
 
-The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, S)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_{T_x} & \equiv & \mathbb{M}_{T_{x+\Delta}}
+\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_T\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_S[x] & \equiv & \mathbb{M}_S[x+\Delta]
 \end{eqnarray}
@@
 C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
-c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
+C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad o = \text{\small SSTORE} \\
 G_{call} & \text{if} \quad o = \text{\small CALL}\\
@@
 
-where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+
+\subsection{Instruction Set}
+
+Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
+For each instruction, also specify is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack.
 
@@
 &&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_S[0] ] \equiv \mathbb{M}_S[1] $ \\
-&&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
+&&&& $C_{\text{\tiny SSTORE}}(\mathbb{S}, \mathbb{M}) \equiv \begin{cases}
 2G_{sstore} & \text{if} \quad \mathbb{M}_S[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] = 0 \\

### 2014-04-17 Paper.tex
Fixed citation.

@@
 
-The scheme we use in order to generate consensus is a simplified version of the GHOST algorithm \cite{ghost}. This process is described in detail in section \ref{ch:ghost}.
+The scheme we use in order to generate consensus is a simplified version of the GHOST protocol \cite{cryptoeprint:2013:881}. This process is described in detail in section \ref{ch:ghost}.
 
@@
 
-Vitalik Buterin first proposed in the original Ethereum whitepaper (first distributed in December 2013) moving from a single state-transaction function with trivial state to something much more general and expansive.
-Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan
+Vitalik Buterin first proposed in the original Ethereum whitepaper (first distributed in December 2013) moving from a suffocating state-transaction function with trivial state to something much more general and expansive.
+Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Nick Savers and Marko Simovic.
 
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve.
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve. This has yet to be formally specified.
 

### 2014-04-17 Paper.tex
Other minor fixes.

@@
 This trie takes the form as a simple mapping between domains of 256-bit values.
-\item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address reveive a call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
+\item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
@@
 \item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction; formally $T_r$.
-\item[v, r, s] Three scalar values corresponding to the signature of the transaction and used to determine the sender of the transaction.
+\item[v, r, s] A scalar and complex value corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_v$ and $T_{rs}$.
+This is expanded in Appendix \ref{app:signing}.
 \end{description}
@@
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block. Not formally used.
+\item[extraData] An arbitrary byte array containing data relavant to this block. With the exception of the genesis block, this must be 32 bytes or fewer. Not formally used.
 \item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.

### 2014-04-17 Paper.tex
Conventions.

@@
 \mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, B) \\
-B & \equiv & (..., [ T_0, T_1, ... ]) \\
+B & \equiv & (..., (T_0, T_1, ...) ) \\
 \Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
@@
 
+\section{Conventions}\label{ch:conventions}
+
+I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
+
+Highly structured, `top-level', state values, are denoted with black-board type.
+They fall into two main sets, those of world-state, which are denoted $\mathbb{S}$ (or a variant thereupon) and those of machine-state, $\mathbb{M}$.
+
+Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+
+For most functions, an uppercase letter is used, e.g.
+$C$, the general cost function.
+These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the SHA3 hash function is denoted $\texttt{SHA3}$.
+
+Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
+This symbol may, if accoringly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction.
+The form of the subscript is used to denote its type; \eg bold subscripts refer to sequences
+
+Scalars and fixed-size byte vectors are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Those with a partcularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
+For particularly important values, an bold uppercase letter may be used.
+
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\mathbb{M}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\mathbb{M}_\mathbf{m}[0..31]$ denotes the first 32-byte word of the machine's memory.
+
+In the case of the global state $\mathbb{S}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
+
+When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
+
 \section{Blocks, State and Transactions} \label{ch:bst}
@@
 \item[nonce] A scalar value equal to the number of transactions sent from this address, or, in the case of contract accounts, the number of contract-creations made by this account.
-\item[balance] A scalar value equal to the number of wei owned by this address.
+For account of address $a$ in state $\mathbb{S}$, this would be formally denoted $\mathbb{S}[a]_n$.
+\item[balance] A scalar value equal to the number of wei owned by this address. Formally denoted $\mathbb{S}[a]_b$.
 \item[stateRoot] A 256-bit hash equal to the root node of a further trie structure that encodes the storage contents of the contract.
@@
 This trie takes the form as a simple mapping between domains of 256-bit values.
+The storage, modelled as a $2^{256}$ size sequence of 256-bit values is formally denoted $\mathbb{S}[a]_\mathbf{s}$.
 \item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
+This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code is denoted as $c$, given that $c = \texttt{\small SHA3}(\mathbb{S}[a]_c)$.
 \end{description}
@@
 \begin{eqnarray}
-S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_{nonce} \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_{balance}
+S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_n \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_b
 \end{eqnarray}
@@
 \mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - v_0 \\
-\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
+\mathbb{S}_0[S(T)]_b & \equiv & \mathbb{S}[S(T)]_b - v_0 \\
+\mathbb{S}_0[S(T)]_n & \equiv & \mathbb{S}[S(T)]_n + 1
 \end{eqnarray}
@@
 \mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
-\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
-\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (T_g - g' - g_d) T_p \\
+\mathbb{S}'[s]_b & \equiv & \mathbb{S}_P[s]_b + g' T_p \\
+\mathbb{S}'[m]_b & \equiv & \mathbb{S}_P[m]_b + (T_g - g' - g_d) T_p \\
 m & \equiv & {B_H}_b
@@
 %\begin{eqnarray}
-%\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
+%\mathbb{S}'[s]_b & \equiv & \mathbb{S}_P[s]_b + g
 %\end{eqnarray}
@@
 \mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}^*[A(s, \mathbb{S}[s]_{nonce})] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
+\mathbb{S}^*[A(s, \mathbb{S}[s]_n)] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
 \end{eqnarray}
@@
 \mathbb{S}_1 & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}_1[r]_{balance} & \equiv & \mathbb{S}_1[r]_{balance} + v
+\mathbb{S}_1[r]_b & \equiv & \mathbb{S}_1[r]_b + v
 \end{eqnarray}
@@
 In the case that the recipient contains code to be executed (i.e.
-that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_{code}]$) is executed according to the execution model (see section \ref{ch:model}).
+that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
@@
 \mathbb{S}_P & \equiv & \begin{cases}
-\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_{code} = \varnothing \\
+\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_c = \varnothing \\
 \mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
@@
 I_v & \equiv & v \\
-\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_{code}
+\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_c
 \end{eqnarray}
@@
 This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-We consider the empty sequence (denoted $[]$ to be non-equal to the empty set, denoted $\varnothing$).
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$.
 
@@
 \mathbb{M}_{pc} & \equiv & 0 \\
-\mathbb{M}_M & \equiv & [0, 0, ...] \\
+\mathbb{M}_\mathbf{m} & \equiv & (0, 0, ...) \\
 \mathbb{M}_i & \equiv & 0 \\
-\mathbb{M}_S & \equiv & []
+\mathbb{M}_\mathbf{s} & \equiv & ()
 \end{eqnarray}
@@
 X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
-(\mathbb{S}, \mathbb{M}, I, []) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
-(O(\mathbb{S}, \mathbb{M}, I), \mathbf{o}) & \text{if} \quad \mathbf{o} \neq \varnothing\\
+(\mathbb{S}, \mathbb{M}, I, ()) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
+O(\mathbb{S}, \mathbb{M}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
 X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
@@
 where
-\begin{equation}
-\mathbf{o} \equiv H(\mathbb{S}, \mathbb{M}, I)
-\end{equation}
-
-The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, S)$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.
+
+\begin{eqnarray}
+\mathbf{o} & \equiv & H(\mathbb{M}, I) \\
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
+\end{eqnarray}
+
+The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The memory contents $\mathbb{M}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
 
@@
 
-\begin{equation}
+\begin{equation}\label{eq:currentoperation}
 w \equiv \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
@@
 \begin{equation}
-Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_S\rVert < \mathbf{\delta}_w
+Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_\mathbf{s}\rVert < \mathbf{\delta}_w
 \end{equation}
@@
 \begin{equation}
-H(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
+H(\mathbb{M}, I) \equiv \begin{cases}
 H_{\text{\tiny RETURN}}(\mathbb{M}) & \text{if} \quad w = \text{\small RETURN} \\
-[] & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
 \varnothing & \text{otherwise}
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\mathbb{M}'_{S}\rVert & \equiv & \lVert\mathbb{M}_T\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_S\rVert): \mathbb{M}'_S[x] & \equiv & \mathbb{M}_S[x+\Delta]
+\lVert\mathbb{M}'_\mathbf{s}\rVert & \equiv & \lVert\mathbb{M}_\mathbf{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_\mathbf{s}\rVert): \mathbb{M}'_\mathbf{s}[x] & \equiv & \mathbb{M}_\mathbf{s}[x+\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathbb{M}'_M & \equiv & \mathbb{M}_M \\
+\mathbb{M}'_\mathbf{m} & \equiv & \mathbb{M}_\mathbf{m} \\
 \mathbb{M}'_i & \equiv & \mathbb{M}_i \\
@@
 \Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...) \\
-B & \equiv & (H, [ T_0, T_1, ... ], \mathbf{U})
+B & \equiv & (H, (T_0, T_1, ...), \mathbf{U})
 \end{eqnarray}
@@
 
+The fee schedule $G$ is a tuple of 10 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
+
 \begin{tabular*}{\columnwidth}[h]{lrl}
@@
 $G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
-$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled of waived sometimes). \\
+$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled or waived sometimes). \\
 $G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
@@
 $G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
+$G_{txdata}$ & 5 & Burnt (discarded from the system) for every byte of data or code for a transaction. \\
 \bottomrule
@@
 
-As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
-In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $\mathbb{S}'$ $\mathbb{M}'$ such that:
-
-\begin{equation}
-O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I)
-\end{equation}
-
-When interpreting 256-bit binary values as integers the representation is big-endian.
-
-When a 32-byte machine datum is converted to and from as a 160-bit address or hash, the leftwards (high-order for BE) 20 bytes are used and the right most 12 are discarded or filled with zeroes.
-
-The general gas cost function is defined as:
-
-\begin{equation}
-C(\mathbb{S}, \mathbb{M}) \equiv G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
-C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad o = \text{\small SSTORE} \\
-G_{call} & \text{if} \quad o = \text{\small CALL}\\
-G_{create} & \text{if} \quad o = \text{\small CREATE}\\
-G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
-G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
-G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
-G_{stop} & \text{if} \quad o = \text{\small STOP}\\
+When interpreting 256-bit binary values as integers, the representation is big-endian.
+
+When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the left most 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
+
+\subsection{Gas Cost}
+
+The general gas cost function, $C$, is defined as:
+
+\begin{eqnarray}
+C(\mathbb{S}, \mathbb{M}, I) & \equiv & G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
+C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad w = \text{\small SSTORE} \\
+G_{call} & \text{if} \quad w = \text{\small CALL}\\
+G_{create} & \text{if} \quad w = \text{\small CREATE}\\
+G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
+G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{stop} & \text{if} \quad w = \text{\small STOP}\\
 G_{step} & \text{otherwise}
+\end{cases} \\
+w & \equiv & \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
+\text{\small STOP} & \text{otherwise}
 \end{cases}
-\end{equation}
+\end{eqnarray}
 
@@
 
+As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
+In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\mathbb{S}', \mathbb{M}')$ such that:
+
+\begin{equation}
+O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I) \quad \text{with exceptions, as noted}
+\end{equation}
+
 Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specify is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack.
+For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
 
@@
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] + \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] \times \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] \times \mathbb{M}_\mathbf{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] - \mathbb{M}_S[1]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] - \mathbb{M}_\mathbf{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \lfloor\mathbb{M}_\mathbf{s}[0] \div \mathbb{M}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \lfloor\mathbb{M}_S[0] \div \mathbb{M}_S[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \lfloor\mathbb{M}_\mathbf{s}[0] \div \mathbb{M}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \mathbb{M}_\mathbf{s}[0] \bmod \mathbb{M}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_S[1] = 0\\ \mathbb{M}_S[0] \bmod \mathbb{M}_S[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \mathbb{M}_\mathbf{s}[0] \bmod \mathbb{M}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0] ^ {\mathbb{M}_S[1] }$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] ^ {\mathbb{M}_\mathbf{s}[1] }$ \\
 \midrule
 0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
-&&&& $\mathbb{M}'_S[0] \equiv -\mathbb{M}_S[0]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv -\mathbb{M}_\mathbf{s}[0]$ \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] < \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] < \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] > \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] > \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = \mathbb{M}_S[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] = \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
-&&&& $\mathbb{M}'_S[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_S[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \end{tabular*}
@@
 \multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
-\multicolumn{5}{l}{$\mathbb{M}_S[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_S[0]$} \vspace{5pt} \\
+\multicolumn{5}{l}{$\mathbb{M}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_\mathbf{s}[0]$} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
 0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \wedge \mathbb{M}_S[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \wedge \mathbb{M}_\mathbf{s}[1]_i$ \\
 \midrule
 0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \vee \mathbb{M}_S[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \vee \mathbb{M}_\mathbf{s}[1]_i$ \\
 \midrule
 0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_S[0]_i \equiv \mathbb{M}_S[0]_i \oplus \mathbb{M}_S[1]_i$ \\
+&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \oplus \mathbb{M}_\mathbf{s}[1]_i$ \\
 \midrule
 0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..7]: \mathbb{M}'_S[0]_i \equiv \begin{cases} \mathbb{M}_S[1]_{(i + 8\mathbb{M}_S[0])} & \text{if} \quad \mathbb{M}_S[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..7]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \begin{cases} \mathbb{M}_\mathbf{s}[1]_{(i + 8\mathbb{M}_\mathbf{s}[0])} & \text{if} \quad \mathbb{M}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + \mathbb{M}_S[1] - 1) ])$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]) \div 32 })$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_a$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{S}[I_a]_b$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_o$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_o$ \\
 &&&& This is the sender of original transaction; it is never a contract. \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_c$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_c$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_v$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_v$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + 31) ]$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\mathbb{M}'_S[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_p$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
-&&&& $\mathbb{M}'_S[0] \equiv {I_H}_p$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_p$ \\
 &&&& $I_{H_p}$ is the previous block's hash. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\mathbb{M}'_S[0] \equiv {I_H}_b$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_b$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\mathbb{M}'_S[0] \equiv {I_H}_t$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_t$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\mathbb{M}'_S[0] \equiv n({I_H})$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv n({I_H})$ \\
 &&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
@@
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\mathbb{M}'_S[0] \equiv {I_H}_d$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\mathbb{M}'_S[0] \equiv 10^6$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv 10^6$ \\
 &&&& Presently, the gas limit is hard-coded in the protocol. \\
@@
 0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[0]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0]$ \\
 \midrule
 0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_S[1]$ \\
-&&&& $\mathbb{M}'_S[1] \equiv \mathbb{M}_S[0]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[1]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[1] \equiv \mathbb{M}_\mathbf{s}[0]$ \\
 \midrule
 0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_M[\mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{m}[\mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 32) \div 32 })$ \\
 \midrule
 0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ] \equiv \mathbb{M}_S[1]$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 32) \div 32 })$ \\
+&&&& $\mathbb{M}'_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + 31) ] \equiv \mathbb{M}_\mathbf{s}[1]$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 32) \div 32 })$ \\
 \midrule
 0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\mathbb{M}'_M[ \mathbb{M}_S[0] ] \equiv (\mathbb{M}_S[1] \mod 256) $ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + 1) \div 32 })$ \\
+&&&& $\mathbb{M}'_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] ] \equiv (\mathbb{M}_\mathbf{s}[1] \mod 256) $ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 1) \div 32 })$ \\
 \midrule
 0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{S}[I_a]_\mathbf{s}[\mathbb{M}_\mathbf{s}[0]]$ \\
 \midrule
 0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_S[0] ] \equiv \mathbb{M}_S[1] $ \\
+&&&& $\mathbb{S}'[I_a]_\mathbf{s}[ \mathbb{M}_\mathbf{s}[0] ] \equiv \mathbb{M}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\mathbb{S}, \mathbb{M}) \equiv \begin{cases}
-2G_{sstore} & \text{if} \quad \mathbb{M}_S[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] = 0 \\
-0 & \text{if} \quad \mathbb{M}_S[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_S[0]] \neq 0 \\
+2G_{sstore} & \text{if} \quad \mathbb{M}_\mathbf{s}[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_\mathbf{s}[\mathbb{M}_\mathbf{s}[0]] = 0 \\
+0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0 \; \wedge \; \mathbb{S}[I_a]_\mathbf{s}[\mathbb{M}_\mathbf{s}[0]] \neq 0 \\
 G_{sstore} & \text{otherwise}
@@
 0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_S[0] $ \\
+&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_\mathbf{s}[0] $ \\
 &&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
@@
 0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_S[0] & \text{if} \quad \mathbb{M}_S[1] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_\mathbf{s}[0] & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
@@
 0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{pc}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_{pc}$ \\
 \midrule
 0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{i}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_{i}$ \\
 \midrule
 0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
-&&&& $\mathbb{M}'_S[0] \equiv \mathbb{M}_{g}$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_{g}$ \\
 \bottomrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{b}[\mathbb{M}_{pc} + 1]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{b}[\mathbb{M}_{pc} + 1]$ \\
 &&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\mathbb{M}'_S[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
 \bottomrule
@@
 0xf0 & {\small CREATE} & 5 & 1 & Create a new contract. \\
-&&&& $\mathbf{b} \equiv \mathbb{M}_M[ \mathbb{M}_S[1] \dots (\mathbb{M}_S[1] + \mathbb{M}_S[2] - 1) ]$ \\
-&&&& $\mathbf{i} \equiv \mathbb{M}_M[ \mathbb{M}_S[3] \dots (\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) ]$ \\
-&&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}^*, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_S[0], \mathbf{i}, \mathbf{b})$ \\
-&&&& $\mathbb{S}^* \equiv \mathbb{S} \quad \text{except} \quad \mathbb{S}^*[I_a]_{nonce} = \mathbb{S}[I_a]_{nonce} + 1$ \\
-&&&& $\mathbb{M}'_S[0] \equiv x$ \\
+&&&& $\mathbf{b} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[1] \dots (\mathbb{M}_\mathbf{s}[1] + \mathbb{M}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $\mathbf{i} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[3] \dots (\mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}^*, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
+&&&& $\mathbb{S}^* \equiv \mathbb{S} \quad \text{except} \quad \mathbb{S}^*[I_a]_n = \mathbb{S}[I_a]_n + 1$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
-&&&& $x=A(I_a, \mathbb{S}[I_a]_{nonce})$, the address of the newly created contract, otherwise. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[1] + \mathbb{M}_S[2], \mathbb{M}_S[3] + \mathbb{M}_S[4]) \div 32 })$ \\
+&&&& $x=A(I_a, \mathbb{S}[I_a]_n)$, the address of the newly created contract, otherwise. \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_\mathbf{s}[1] + \mathbb{M}_\mathbf{s}[2], \mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4]) \div 32 })$ \\
 \midrule
 0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
-&&&& $\mathbf{i} \equiv \mathbb{M}_M[ \mathbb{M}_S[3] \dots (\mathbb{M}_S[3] + \mathbb{M}_S[4] - 1) ]$ \\
-&&&& $\mathbf{o} \equiv \mathbb{M}_M[ \mathbb{M}_S[5] \dots (\mathbb{M}_S[5] + \mathbb{M}_S[6] - 1) ]$ \\
-&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \mathbb{M}_S[2] = 0 \\ \mathbb{M}_g - \min(\mathbb{M}_g, \mathbb{M}_S[2]) & \text{otherwise} \end{cases}$ \\
-&&&& $(\mathbb{S}', g', \mathbf{o}) \equiv \Theta(\mathbb{S}, I_a, I_o, \mathbb{M}_S[0], g, I_p, \mathbb{M}_g - g_k, \mathbb{M}_S[1], \mathbf{i})$ \\
+&&&& $\mathbf{i} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[3] \dots (\mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $\mathbf{o} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[5] \dots (\mathbb{M}_\mathbf{s}[5] + \mathbb{M}_\mathbf{s}[6] - 1) ]$ \\
+&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[2] = 0 \\ \mathbb{M}_g - \min(\mathbb{M}_g, \mathbb{M}_\mathbf{s}[2]) & \text{otherwise} \end{cases}$ \\
+&&&& $(\mathbb{S}', g', \mathbf{o}) \equiv \Theta(\mathbb{S}, I_a, I_o, \mathbb{M}_\mathbf{s}[0], g, I_p, \mathbb{M}_g - g_k, \mathbb{M}_\mathbf{s}[1], \mathbf{i})$ \\
 &&&& $\mathbb{M}'_g \equiv g' + g_k$ \\
-&&&& $\mathbb{M}'_S[0] \equiv x$ \\
+&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
 &&&& $x=1$ otherwise. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_S[3] + \mathbb{M}_S[4], \mathbb{M}_S[5] + \mathbb{M}_S[6]) \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4], \mathbb{M}_\mathbf{s}[5] + \mathbb{M}_\mathbf{s}[6]) \div 32 })$ \\
 \midrule
 0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\mathbb{M}) \equiv \mathbb{M}_M[ \mathbb{M}_S[0] \dots ( \mathbb{M}_S[0] + \mathbb{M}_S[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\mathbb{M}) \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] \dots ( \mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_S[0] + \mathbb{M}_S[1]) \div 32 })$ \\
+&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]) \div 32 })$ \\
 \midrule
 0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
-&&&& $\mathbb{S}'[\mathbb{M}_S[0]]_{balance} \equiv \mathbb{S}[\mathbb{M}_S[0]]_{balance} + \mathbb{S}[I_a]_{balance}$ \\
+&&&& $\mathbb{S}'[\mathbb{M}_\mathbf{s}[0]]_b \equiv \mathbb{S}[\mathbb{M}_\mathbf{s}[0]]_b + \mathbb{S}[I_a]_b$ \\
 &&&& $\mathbb{S}'[I_a] \equiv \varnothing$ \\
@@
 
-$\langle 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 2^{22}, 0, [], 42 \rangle$
+$\langle 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 2^{22}, 0, (), 42 \rangle$
 
@@
 
-$[]$ refers to the extradata, an empty byte array.
+$()$ refers to the extradata, an empty byte array.
 

### 2014-04-17 Paper.tex
RLP.

@@
 
-%Thus we define the set of possible structures $\mathbb{T}$:
-
-%\begin{eqnarray}
-%\text{Let} \mathbf{t} & = & \langle \mathbf{t}_0, \mathbf{t}_1, ... \rangle
-%\vee \mathbf{t} & = & 
-%\end{eqnarray}
-
-%We define the function $\mathtt{\tiny RLP}$ that maps from an arbitrarily structured tree whose leaves are byte arrays into a single byte array.
+We define the set of possible structures $\mathbb{T}$:
+\begin{eqnarray}
+\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+
+Where $\mathbb{Y}$ is the set of bytes.
+Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
+
+We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\end{equation}
+
+If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+
+\begin{itemize}
+\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
+\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
+\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
+\end{itemize}
+
+Formally, we define $R_b$:
+\begin{eqnarray}
+R_b(\mathbf{x}) & \equiv & \begin{cases}
+\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
+(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
+\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\end{cases} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum^{n = 0}_{n < \lVert \mathbf{b} \rVert} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
+\end{eqnarray}
+
+Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+
+If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+
+\begin{itemize}
+\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
+\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\end{itemize}
+
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
+(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
+\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+\end{cases} \\
+s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
+\end{eqnarray}
+
+\section{Hex-Prefix Encoding}\label{app:hexprefix}
+Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array.
+It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
+
+It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{E}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
+
+\begin{eqnarray}
+\mathtt{\tiny HP}(\mathbf{x}, t) & \equiv & ...
+\end{eqnarray}
 
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-
-\section{Hex-Prefix Encoding}\label{app:hexprefix}
-Hex-prefix encoding is an efficient method of encoding a arbitrary number of nibbles as a byte array.
-It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
 

### 2014-04-18 Paper.tex
Tweaks.

@@
 
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol \cite{cryptoeprint:2013:881}. This process is described in detail in section \ref{ch:ghost}.
+The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
+This process is described in detail in section \ref{ch:ghost}.
 
@@
 This symbol may, if accoringly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction.
-The form of the subscript is used to denote its type; \eg bold subscripts refer to sequences
+The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
 Scalars and fixed-size byte vectors are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a partcularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
 
 Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
-For particularly important values, an bold uppercase letter may be used.
+For particularly important values, a bold uppercase letter may be used.
 
 Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\mathbb{M}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\mathbb{M}_\mathbf{m}[0..31]$ denotes the first 32-byte word of the machine's memory.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\mathbb{M}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code is denoted as $c$, given that $c = \texttt{\small SHA3}(\mathbb{S}[a]_c)$.
+This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code is denoted as $b$, given that $\texttt{\small SHA3}(b) = \mathbb{S}[a]_c$.
 \end{description}
@@
 
-We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes; the given transactions $\mathbf{T}$, when executed in order on the base state $\mathbb{S}$ result in a new state of the identity $H_r$:
+We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $\mathbf{T}$, when executed in order on the base state $\mathbb{S}$, result in a new state of the identity $H_r$:
 \begin{eqnarray}
@@
 
-The canonical difficulty of a block of header $H$, is defined as $D(H)$:
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
@@
 
-$H_s$ is the timestamp of block $H$, must fulfill the relation:
+$H_s$ is the timestamp of block $H$ and must fulfill the relation:
 \begin{equation}
@@
 a & = & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing \\
-a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;\langle s, n\rangle\;\big)\Big)\Big)
+a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
 \end{eqnarray}

### 2014-04-19 Paper.tex
Modified Merkle Patricia Trie.

@@
 The incorruptibility of judgement, often difficult to find, comes naturally from a disinterested algorithmic interpreter.
-Transparency, through being about to see exactly how a state or judgement came about through the transaction log and rules or instructional codes never happens perfectly in human-based systems.
-
-Overall, we wish to provide a system such that the user can be guaranteed that any other individuals, systems or organisations that they interact with, they can do so with absolute confidence in the possible outcomes and how they might come about.
-
-%\subsection{Previous Work} \label{ch:previous}
-%E language. Smart contracts. Bitcoin, namecoin \&c. Mastercoin.
+Transparency, or being able to see exactly how a state or judgement came about through the transaction log and rules or instructional codes, never happens perfectly in human-based systems since natural language is necessarily vague, information is often lacking, and plain old prejudices are difficult to shake.
+
+Overall, I wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
+
+\subsection{Previous Work} \label{ch:previous}
+
+\cite{buterin2013ethereum} first proposed the kernel of this work in late November, 2013.
+Though now evolved in many ways, the key functionality of a block-chain with a Turing-complete language and an effectively unlimited inter-transaction storage capability remains unchanged.
+
+Hashcash, introduced by \cite{laurie2004proof}, provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
+Though not widely adopted, the work was later utilised and expanded upon by \cite{nakamoto2008bitcoin} in order to devise a cryptographically secure mechanism for coming to a decentralised social consensus over the order and contents of a series of cryptographically signed financial transactions.
+The fruits of this project, Bitcoin, provided a first glimpse into a decentralised transaction ledger.
+
+Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol.
+The best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
+Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.
+
+Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
+The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol though utilisation of a number of auxilliary parts to the core protocol.
+The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transactions in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet'' protocol aware piece of software.
+
+Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system.
+It has demonstrate that high efficiency gains can be made if the decentralisation premise is discarded.
+
+Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
+Around the 1990s it became clear that algorithm enforcement of agreements could become a significant force in human cooperation.
+Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
+In this light, Ethereum may be seen as a general implementation of such a \textit{crypto-law} system.
+
+%E language?
 
@@
 Invalid state changes might, \eg be things such as reducing an account balance without an equal and opposite increase elsewhere.
-We define a valid state transitions as one which comes about through a transaction.
+A valid state transition is one which comes about through a transaction.
 Formally:
@@
 
-Transactions are collated into blocks; we then chain blocks together using a cryptographic hash as a means of reference.
+Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
 Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
@@
 In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH}.
+To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes refered to by the Old English \DH{}.
 The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
@@
 
-We define the rollback state $\mathbb{S}_0$:
+We define the checkpoint state $\mathbb{S}_0$:
 \begin{eqnarray}
@@
 
-We define $\mathbb{S}_1$, the rollback state as the original state but with the value transferred to the recipient:
+We define $\mathbb{S}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
 \begin{eqnarray}
@@
 If the recipient account contains no code to be executed (i.e.
-it's not a contract), then we define the final state as being equivalent to the rollback state.
+it's not a contract), then we define the final state as being equivalent to the checkpoint state.
 In the case that the recipient contains code to be executed (i.e.
@@
 
-State database won't keep everything; maintains an age and eventually throws out nodes that are not recent enough or checkpoints.
-
-Nodes in state trie that haven't sent/received a transaction in over X blocks could be thrown out, reducing Ether-leakage and reducing the growth of the state database.
+The state database won't be forced to maintain all past state trie structures in to the future.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+
+Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10000th block) could be maintained by the peer network, effectively recasting the genesis block.
+This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
+
+Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in over some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
+
+\subsection{Scalability}
+
+Scalability remains an eternal concern.
+With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy.
+Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger.
+However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.
+
+Some form of hierarchical structure, either through consolidating smaller lighter-weight chains into the main block or through building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
+Parallelism could also come from a prioritised set of parallel blockchains, consolidated each block and with duplicate or invalid transactions thrown out accordingly.
+
+Verifyable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
 
@@
 
-This paper has introduced, discussed and formally defined the protocol of Ethereum.
+I have introduced, discussed and formally defined the protocol of Ethereum.
+Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
+Contracts may be authored in order to algorithmically specify and autonomously enforce rules of interaction.
 
@@
 
-Vitalik Buterin first proposed in the original Ethereum whitepaper (first distributed in December 2013) moving from a suffocating state-transaction function with trivial state to something much more general and expansive.
-Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Nick Savers and Marko Simovic.
+Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
 
@@
 \section{Recursive Length Prefix}\label{app:rlp}
-This is a serialisation method for encoding arbitrarily structured binary data (byte arrays). Within RLP, integers are specified as the byte array equal to the big-endian representation of the value.
+This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
 
@@
 
+If RLP is used to encode a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+Thus the RLP of some positive integer $i$ is defined as:
+\begin{equation}
+\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\end{equation}
+
+There is no specific canonical encoding format for signed or floating-point values.
+
 \section{Hex-Prefix Encoding}\label{app:hexprefix}
@@
 \begin{eqnarray}
-\mathtt{\tiny HP}(\mathbf{x}, t) & \equiv & ...
-\end{eqnarray}
+\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{E} & \equiv & \begin{cases}
+(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
+\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
+(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
+\text{otherwise}
+\end{cases} \\
+f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \\ 0 & \text{otherwise} \end{cases}
+\end{eqnarray}
+
+Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$.
+The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number.
+All remaining nibbles (now an even number) fit properly into the remaining bytes.
 
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
+The core of the trie, and its sole rquirement in terms of the protocol specification is to provide a single 32-byte value that identifies a given set of key-value pairs.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
+
+Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
+\begin{equation}
+\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
+\end{equation}
+
+When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
+\begin{equation}
+\forall_{I: \in \mathfrak{I}} I \equiv (I_0, I_1)
+\end{equation}
+
+Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
+\begin{eqnarray}
+y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
+\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+\begin{cases}
+\lfloor \mathbf{k}_n[\lfloor i \div 2 \rfloor] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure.
+The empty trie is defined as being a 256-bit hash of zeroes, $0_{256}$:`
+\begin{equation}
+\texttt{\small TRIE}(\mathfrak{I}) \equiv \begin{cases}
+0_{256} & \text{if} \quad \mathfrak{I} = \varnothing \\
+n(y(\mathfrak{I}), 0) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+We assume a function $n$, the trie's node cap function.
+When composing a node, we use RLP to encode the structure.
+As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose SHA3 hash evaluates to our reference.
+Thus we define in terms of $c$, the node composition function:
+\begin{equation}
+n(\mathfrak{I}, i) \equiv \begin{cases}
+() & \text{if} \quad \mathfrak{I} = \varnothing \\
+c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
+\texttt{\small SHA3}(c(\mathfrak{I}, i)) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair.
+The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
+Unlike a radix tree, in the case of multiple keys shared the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
+Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
+There are three kinds of nodes in the trie:
+\begin{description}
+\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
+\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of nibbles keys and branches as traversed from the root.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
+\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
+The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
+\end{description}
+
+A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
+\begin{equation}
+c(\mathfrak{I}, i) \equiv \begin{cases}
+\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
+(\lVert I_0\rVert - 1)], I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{\small RLP}\Big(\big(\texttt{\small HP}(I_0[i ..
+(j - 1)]), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(x - 1)] = \mathbf{l} \\
+\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
+u(n) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = n \}, i + 1) \\
+v & = & \begin{cases}
+I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
+() & \text{otherwise}
+\end{cases}
+\end{array}
+\end{cases}
+\end{equation}
+
+\subsection{Trie Database}
+Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision.
+In reality, a sensible implemention will not fully recompute the trie root hash for each set.
+
+A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
+This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
+Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
 
@@
 
-\section{The State DB}\label{app:state}
-
-The state database is a mapping between SHA3-256 hashes (i.e. 256-bit binary fragments) and arbitrary byte data.
-
 \section{Signing Transactions}\label{app:signing}
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve. This has yet to be formally specified.
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}. This has yet to be formally specified.
 
@@
 0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\mathbb{M}'_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] ] \equiv (\mathbb{M}_\mathbf{s}[1] \mod 256) $ \\
+&&&& $\mathbb{M}'_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] ] \equiv (\mathbb{M}_\mathbf{s}[1] \bmod 256) $ \\
 &&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 1) \div 32 })$ \\
@@
 
-The header of the genesis block is 9 items, and is specified thus:
-
-$\langle 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([])), 2^{22}, 0, (), 42 \rangle$
+The genesis block is 9 items, and is specified thus:
+
+$\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 2^{22}, 0, (), 42 \big), (), () \big)$
 
@@
 
-$()$ refers to the extradata, an empty byte array.
-
-$\mathtt{\tiny SHA3}(\mathtt{\tiny RLP}([]))$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.
+$()$ refers to the extradata and the sequences of both uncles and transactions, all empty.
+
+$\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.
 

### 2014-04-19 Paper.tex
Trie & serialisation format folded into main definition.

@@
 
-Scalars and fixed-size byte vectors are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
 Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
+The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$.
+This is formally defined in section \ref{ch:block}.
+
 Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\mathbb{M}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
@@
 
+When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
+It is formally defined in chapter \ref{ch:block}.
+
 \section{Blocks, State and Transactions} \label{ch:bst}
@@
 
-The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
+The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
 Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
@@
 For account of address $a$ in state $\mathbb{S}$, this would be formally denoted $\mathbb{S}[a]_n$.
-\item[balance] A scalar value equal to the number of wei owned by this address. Formally denoted $\mathbb{S}[a]_b$.
-\item[stateRoot] A 256-bit hash equal to the root node of a further trie structure that encodes the storage contents of the contract.
-Though a separate data structure, this trie is still stored using the same underlying state database.
-This trie takes the form as a simple mapping between domains of 256-bit values.
-The storage, modelled as a $2^{256}$ size sequence of 256-bit values is formally denoted $\mathbb{S}[a]_\mathbf{s}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\mathbb{S}[a]_b$.
+\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the contract, itself a simple mapping from byte arrays of size 32 to arbitrary length byte arrays.
+The hash is formally denoted $\mathbb{S}[a]_s$.
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence $\texttt{\small TRIE}(\mathbb{S}[a]_\mathbf{s}) \equiv \mathbb{S}[a]_s$.
+It shall be understood that $\mathbb{S}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 \item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code is denoted as $b$, given that $\texttt{\small SHA3}(b) = \mathbb{S}[a]_c$.
+This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small SHA3}(\mathbf{b}) = \mathbb{S}[a]_c$.
 \end{description}
@@
 If the latter two fields are missing, the node represents a simple (non-contract) account: it maintains no state and executes no code on receipt of a transaction.
+
+Thus we may define a world-state collapse function $L_S$:
+\begin{eqnarray}
+L_S(\mathbb{S}) & \equiv & \{ p(a): \mathbb{S}[a] \neq \varnothing \} \\
+p(a) & \equiv & \big(a, \texttt{\small RLP}\big( (\mathbb{S}[a]_n, \mathbb{S}[a]_b, \mathbb{S}[a]_s, \mathbb{S}[a]_c) \big) \big)
+\end{eqnarray}
+
+This function, $L$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+\begin{equation}
+\forall a: \mathbb{S}[a] = \varnothing \vee (a \in \mathbb{B}_{20} \wedge v(\mathbb{S}[a]))
+\end{equation}
+
+Where $v$ is the account validity function:
+\begin{eqnarray}
+v(x) & \equiv & x_n \in \mathbb{P} \wedge x_p \in \mathbb{P} \wedge (x_r = x_c = \varnothing \vee w(x)) \\
+w(x) & \equiv & x_r \in \mathbb{B}_{32} \wedge \quad x_c \in \mathbb{B}_{32}
+\end{eqnarray}
 
@@
 This is paid up-front, before any computation is done and may not be topped-up later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction; formally $T_r$.
-\item[v, r, s] A scalar and complex value corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_v$ and $T_{rs}$.
+\item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction; formally $T_t$.
+\item[v, r, s] A scalar and complex value corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
 This is expanded in Appendix \ref{app:signing}.
@@
 For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
+
+The function $L_T$ prepares a transaction for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_T(T) \equiv \begin{cases}
+(T_n, T_v, T_p, T_g, T_t, T_\mathbf{b}, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
+(T_n, T_v, T_p, T_g, T_t, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{b}$, $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
+\begin{equation}
+\begin{array}{lclclcl}
+T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge & T_g \in \mathbb{P} \\
+& \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge & T_s \in \mathbb{P} \\
+& \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{b} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} \\
+& \wedge & T_t \in \mathbb{B}_{20} &&\\
+\end{array}
+\end{equation}
 
@@
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relavant to this block. With the exception of the genesis block, this must be 32 bytes or fewer. Not formally used.
+\item[extraData] An arbitrary byte array containing data relavant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
 \item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
@@
 \begin{equation}
-B \equiv (H, \mathbf{T}, \mathbf{U})
-\end{equation}
-
-We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $\mathbf{T}$, when executed in order on the base state $\mathbb{S}$, result in a new state of the identity $H_r$:
-\begin{eqnarray}
-& & \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(\mathbf{U})) \equiv H_u \\
-& \wedge & \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(\mathbf{T})) \equiv H_t \\
-& \wedge & r(\Pi(\mathbb{S}, B)) \equiv H_r
-\end{eqnarray}
-
-where $r(\mathbb{S})$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of $\mathbb{S}$ with values encoded using RLP.
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\mathbb{S}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\begin{eqnarray}
+H_u & \equiv & \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) \\
+\wedge \quad H_t & \equiv & \mathtt{\small SHA3}(\mathtt{\small RLP}(L_T^*(B_\mathbf{T}))) \\
+\wedge \quad H_r & \equiv & \mathtt{\small TRIE}(L_S(\Pi(\mathbb{S}, B)))
+\end{eqnarray}
+
+where 
+\begin{equation}
+\mathtt{\small TRIE}(L_S(\mathbb{S})) = {P(B_H)_H}_r
+\end{equation}
+
+this $\texttt{\small TRIE}(L_S(\mathbb{S}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\mathbb{S}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+Much like the transaction preparation function $L_T$, we assert the types and order of the structure for when the RLP transformation is required:
+
+\begin{eqnarray}
+\quad L_H(H) & \equiv & (H_p, H_u, H_b, H_r, H_t, H_d, H_s, H_x, H_n) \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+\end{equation}
+
+The component types are defined thus:
+\begin{equation}
+\begin{array}[t]{lclclcl}
+H_p \in \mathbb{B}_{32} & \wedge & H_u \in \mathbb{B}_{32} & \wedge & H_b \in \mathbb{B}_{20} & \wedge \\
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
+H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge & H_b \in \mathbb{B}_{32} &&
+\end{array}
+\end{equation}
+
+where
+\begin{equation}
+\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\end{equation}
+
+We now have a rogourous specification for the construction of a formal block structure.
+The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
 
@@
 g_0 \equiv \begin{cases}
-g_d + G_{create} & \text{if} \quad T_r = 0 \\
+g_d + G_{create} & \text{if} \quad T_t = 0 \\
 g_d + G_{call} & \text{otherwise} \\
@@
 \begin{equation}
-g_d \equiv \begin{cases} (\lVert T_\mathbf{i}\rVert + \lVert T_\mathbf{b}\rVert) G_{data} & \text{if} \quad T_r = 0 \\
+g_d \equiv \begin{cases} (\lVert T_\mathbf{i}\rVert + \lVert T_\mathbf{b}\rVert) G_{data} & \text{if} \quad T_t = 0 \\
  \lVert T_\mathbf{d}\rVert G_{data} & \text{otherwise}
@@
 (\mathbb{S}_P, g') \equiv \begin{cases}
-\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
-\Theta^*(\mathbb{S}_0, S(T), T_o, T_r, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
+\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_t = 0 \\
+\Theta^*(\mathbb{S}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-Z(\mathbb{S}, \mathbb{M}, I) \equiv \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \vee \mathbf{\delta}_w = \varnothing \vee \lVert\mathbb{M}_\mathbf{s}\rVert < \mathbf{\delta}_w
+Z(\mathbb{S}, \mathbb{M}, I) \equiv
+\begin{array}[t]{rl}
+& \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \\
+\vee & \mathbf{\delta}_w = \varnothing \\
+\vee & \lVert\mathbb{M}_\mathbf{s}\rVert < \mathbf{\delta}_w
+\end{array}
 \end{equation}
@@
 
-If RLP is used to encode a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 \end{equation}
+
+When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.
 
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(\lVert I_0\rVert - 1)], I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{\small RLP}\Big(\big(\texttt{\small HP}(I_0[i ..
-(j - 1)]), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
 \texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(n) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = n \}, i + 1) \\
+u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
 v & = & \begin{cases}
@@
 Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
-
-\section{Formal Specification of Structures}
-Though RLP is data-agnostic, it does specify a canonical representation for integer quantities.
-It is big-endian with no leading zero-bytes.
-Thus for elements than feasibly can be stored as integers, it becomes important to specify whether they should adhere to this canonical representation or be left in some other (perhaps more 'native') format.
-
-In the case of counts, balances, fees and amounts of wei, the canon-integer form must be used when storing in RLP. We call these INTs.
-
-In the case of hashes (256-bit or 160-bit), user-visible strings and specialised byte-arrays (e.g.
-hex-prefix notation from the trie), they should be stored as unformatted byte-array data and not altered into some other form.
-We call these BINs.
-
-When interpreting RLP data, clients are required to consider non-canonical INT fields in the same way as otherwise invalid RLP data and dismiss it completely.
-
-Specifically:
-
-for the Block header:
-\begin{verbatim}
-[
-  parentHash: BIN (256-bit),
-  unclesHash: BIN (256-bit),
-  coinbase: BIN (160-bit),
-  stateRoot: BIN (256-bit),
-  transactionsHash: BIN (256-bit),
-  difficulty: INT,
-  timestamp: INT,
-  extraData: BIN (typically 256-bit),
-  nonce: BIN (256-bit)
-]
-\end{verbatim}
-
-(note: 'nonce', the last element, refers to a hash here and so is binary)
-
-for entries in the State trie for normal addresses:
-\begin{verbatim}
-[
-  balance: INT,
-  nonce: INT
-]
-\end{verbatim}
-
-and for contract addresses:
-\begin{verbatim}
-[
-  balance: INT,
-  nonce: INT,
-  storageRoot: BIN (256-bit),
-  codeHash: BIN (256-bit)
-]
-\end{verbatim}
-
-(note: 'nonce', the second element, refers to a tx-count here and so is integer)
-
-for message call transactions:
-
-\begin{verbatim}
-[
-  nonce: INT,
-  value: INT,
-  gasPrice: INT,
-  gas: INT,
-  recvAddr: BIN (160-bit),
-  data: BIN (arbitrary length),
-  v: INT,
-  r: INT,
-  s: INT
-]
-\end{verbatim}
-
-or, for contract creation transactions,
-
-\begin{verbatim}
-[
-  nonce: INT,
-  value: INT,
-  gasPrice: INT,
-  gas: INT,
-  0: BIN (160-bit),
-  code: BIN (arbitrary length),
-  init: BIN (arbitrary length),
-  v: INT,
-  r: INT,
-  s: INT
-]
-\end{verbatim}
-
-(note: 'nonce', the first element, refers to a tx-count here and so is integer)
-
-The nonce in the transaction refers to the total amount of transactions sent from the address up until that moment in time.
-
-for blocks, there are no immediate data field, but lists:
-
-\begin{verbatim}
-[
-  blockHeader: [...]
-  txList: [ [...], [...], ... ]
-  uncleList: [ [...], [...], ... ]
-]
-\end{verbatim}
-
-Uncle-blocks contain only the uncle's header.
 

### 2014-04-21 Paper.tex
Minor genesis block fix.

@@
 \section{Wire Protocol}\label{app:wire}
-The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transfering information such as blocks \& transactions.
+The wire-protocol specifies a network-level protocol for how two peers can communicate.
+It includes handshake procedures and the means for transfering information such as peers, blocks and transactions.
+Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.
+
+Ethereum nodes may connect to each other over TCP only.
+Peers are free to advertise and accept connections on any port(s) they wish, however, a default port on which the connection may be listened and made will be 30303.
+
+Though TCP provides a connection-oriented medium, Ethereum nodes communicate in terms of packets.
+These packets are formed as a 4-byte synchronisation token (0x22400891), a 4-byte "payload size", to be interpreted as a big-endian integer and finally an N-byte RLP-serialised data structure, where N is the aforementioned "payload size".
+To be clear, the payload size specifies the number of bytes in the packet ''following'' the first 8.
+
+There are a number of different types of payload that may be encoded within the RLP. This ''type'' is always determined by the first entry of the RLP, represented as a scalar:
+
+\begin{tabular*}{\columnwidth}[h]{rlll}
+\toprule
+\multicolumn{4}{c}{\textbf{00s: Session control}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
+0x00 & {\small Hello} & {\small Hello}? & $(\text{0x}00, v \in \mathbb{P}, n \in \mathbb{P}, \mathbf{i} \in \mathbb{B}, c \in \mathbb{P}, p \in \mathbb{P}, i \in \mathbb{B}_{64})$ \\
+\midrule
+0x01 & {\small Disconnect} & {\small Disconnect}? & $(\text{0x}01, r \in \mathbb{P})$ \\
+\midrule
+0x02 & {\small Ping} & {\small Pong} & $(\text{0x}02)$ \\
+\midrule
+0x03 & {\small Pong} && $(\text{0x}03)$ \\
+\midrule
+
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlll}
+\toprule
+\multicolumn{4}{c}{\textbf{00s: Information}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
+0x10 & {\small GetPeers} & {\small Peers} & $(\text{0x}10)$ \\
+\midrule
+0x11 & {\small Peers} & & $(\text{0x}11, (a_1 \in \mathbb{B}_4, a_1 \in \mathbb{P}), (), ...)$ \\
+\midrule
+0x12 & {\small Transactions} & {\small Pong} & $(\text{0x}12)$ \\
+\midrule
+0x13 & {\small Blocks} && $(\text{0x}13)$ \\
+\midrule
+0x14 & {\small GetChain} & {\small Chain} & $(\text{0x}14)$ \\
+\midrule
+0x15 & {\small NotInChain} && $(\text{0x}15)$ \\
+\midrule
+0x16 & {\small GetTransactions} & {\small Transactions} & $(\text{0x}16)$ \\
+\bottomrule
+\end{tabular*}
+
+%(0x00, 
+%First packet sent over the connection, and sent once by both sides. No other messages may be sent until a Hello is received.
+%PROTOCOL_VERSION is one of:
+%0x00 for PoC-1;
+%0x01 for PoC-2;
+%0x07 for PoC-3;
+%0x09 for PoC-4.
+%NETWORK_ID should be 0.
+%CLIENT_ID Specifies the client software identity, as a human-readable string (e.g. "Ethereum(++)/1.0.0").
+%LISTEN_PORT specifies the port that the client is listening on (on the interface that the present connection traverses). If 0 it indicates the client is not listening.
+%CAPABILITIES specifies the capabilities of the client as a set of flags; presently three bits are used: 0x01 for peers discovery, 0x02 for transaction relaying, 0x04 for block-chain querying.
+%NODE_ID is optional and specifies a 512-bit hash, (potentially to be used as public key) that identifies this node.
+%Disconnect
+%
+%[0x01, REASON]
+%Inform the peer that a disconnection is imminent; if received, a peer should disconnect immediately.
+When sending, well-behaved hosts give their peers a fighting chance (read: wait 2 seconds) to disconnect to before disconnecting themselves.
+%REASON is an optional integer specifying one of a number of reasons for disconnect:
+%0x00 Disconnect requested;
+%0x01 TCP sub-system error;
+%0x02 Bad protocol;
+%0x03 Useless peer;
+%0x04 Too many peers;
+%0x05 Already connected;
+%0x06 Wrong genesis block;
+%0x07 Incompatible network protocols;
+%0x08 Client quitting.
+%Ping
+%
+%[0x02]
+%Requests an immediate reply of Pong from the peer.
+%Pong
+%
+%[0x03]
+%Reply to peer's Ping packet.
+%GetPeers
+%
+%[0x10]
+%Request the peer to enumerate some known peers for us to connect to. This should include the peer itself.
+%Peers
+%
+%[0x11, [IP1, Port1, Id1], [IP2, Port2, Id2], ... ]
+%Specifies a number of known peers.
+IP is a 4-byte array 'ABCD' that should be interpreted as the IP address A.B.C.D.
+Port is a 2-byte array that should be interpreted as a 16-bit big-endian integer.
+Id is the 512-bit hash that acts as the unique identifier of the node.
+%Transactions
+%
+%[0x12, [nonce, receiving_address, value, ... ], ... ]
+%Specify (a) transaction(s) that the peer should make sure is included on its transaction queue.
+The items in the list (following the first item 0x12) are transactions in the format described in the main Ethereum specification.
+%Blocks
+%
+%[0x13, [block_header, transaction_list, uncle_list], ... ]
+%Specify (a) block(s) that the peer should know about. The items in the list (following the first item, 0x13) are blocks in the format described in the main Ethereum specification.
+%GetChain
+%
+%[0x14, Parent1, Parent2, ..., ParentN, Count]
+%Request the peer to send Count (to be interpreted as an integer) blocks in the current canonical block chain that are children of Parent1 (to be interpreted as a SHA3 block hash).
+If Parent1 is not present in the block chain, it should instead act as if the request were for Parent2 &c.
+through to ParentN.
+If the designated parent is the present block chain head, an empty reply should be sent.
+If none of the parents are in the current canonical block chain, then NotInChain should be sent along with ParentN (i.e.
+the last Parent in the parents list).
+If no parents are passed, then a reply need not be made.
+%NotInChain
+%
+%[0x15, Hash]
+%Tell the peer that the given hash was not found in its block chain.
+%GetTransactions
+%
+%[0x16]
+%Request the peer to send all transactions currently in the queue. See Transactions.
 
@@
 
-$\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 2^{22}, 0, (), 42 \big), (), () \big)$
+$\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 2^{22}, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)$
 
@@
 
+$\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.
+
 $\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.

### 2014-04-21 Paper.tex
Wire protocol.

--- 2014-04-19 Paper.tex
@@
 \section{Wire Protocol}\label{app:wire}
-The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transfering information such as blocks \& transactions.
-
-%\section{Peer Strategy}\label{app:peers}
+The wire-protocol specifies a network-level protocol for how two peers can communicate.
+It includes handshake procedures and the means for transfering information such as peers, blocks and transactions.
+Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.
+
+Ethereum nodes may connect to each other over TCP only.
+Peers are free to advertise and accept connections on any port(s) they wish, however, a default port on which the connection may be listened and made will be 30303.
+
+Though TCP provides a connection-oriented medium, Ethereum nodes communicate in terms of packets.
+These packets are formed as a 4-byte synchronisation token (0x22400891), a 4-byte "payload size", to be interpreted as a big-endian integer and finally an N-byte \textbf{RLP-serialised} data structure, where N is the aforementioned "payload size".
+To be clear, the payload size specifies the number of bytes in the packet ''following'' the first 8.
+
+There are a number of different types of message that may be sent.
+This ``type'' is always determined by the first entry of the structure, represented as a scalar.
+The structure of each message type is described below.
+
+\begin{tabular*}{\columnwidth}[h]{rlll}
+\toprule
+\multicolumn{4}{c}{\textbf{00s: Session control}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
+0x00 & \textsc{Hello} & & $(\text{0x}00, v \in \mathbb{P}, n \in \mathbb{P}, \mathbf{i} \in \mathbb{B}, c \in \mathbb{P}, p \in \mathbb{P}, u \in \mathbb{B}_{64})$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{
+This is the first packet sent over the connection, and sent once by both sides. No other messages may be sent until a \textsc{Hello} is received.
+\begin{itemize}
+\item $v$ is the Protocol Version. See the latest documentation for which version is current.
+\item $n$ is the Network Id should be 0.
+\item $\mathbf{i}$ is the Client Id and specifies the client software identity as a human-readable string (e.g. ``Ethereum(++)/1.0.0'').
+\item $c$ is the client's Capabilities and specifies the capabilities of the client as a set of flags; presently three bits are used:
+\begin{description}
+\item[0x01] Client provides peer discovery service;
+\item[0x02] Client provides transaction relaying service;
+\item[0x04] Client provides block-chain querying service.
+\end{description}
+\item $p$ is the Listen Port and specifies the port that the client is listening on (on the interface that the present connection traverses). If 0 it indicates the client is not listening.
+\item $u$ is the Unique Identity of the node and specifies a 512-bit hash that identifies this node.
+\end{itemize}
+}\\
+\midrule
+0x01 & \textsc{Disconnect} && $(\text{0x}01, r \in \mathbb{P})$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{
+Inform the peer that a disconnection is imminent; if received, a peer should disconnect immediately.
+When sending, well-behaved hosts give their peers a fighting chance (read: wait 2 seconds) to disconnect to before disconnecting themselves.
+\begin{itemize}
+\item $r$ is an integer specifying one of a number of reasons for disconnect:
+\begin{description}
+\item[0x00] Disconnect requested;
+\item[0x01] TCP sub-system error;
+\item[0x02] Bad protocol;
+\item[0x03] Useless peer;
+\item[0x04] Too many peers;
+\item[0x05] Already connected;
+\item[0x06] Wrong genesis block;
+\item[0x07] Incompatible network protocols;
+\item[0x08] Client quitting.
+\end{description}
+\end{itemize}
+}\\
+\midrule
+0x02 & \textsc{Ping} & \textsc{Pong} & $(\text{0x}02)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{Requests an immediate reply of \textsc{Pong} from the peer.}\\
+\midrule
+0x03 & \textsc{Pong} && $(\text{0x}03)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{Reply to peer's \textsc{Ping} packet.}\\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlll}
+\toprule
+\multicolumn{4}{c}{\textbf{10s: Information}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
+0x10 & \textsc{GetPeers} & \textsc{Peers} & $(\text{0x}10)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to enumerate some known peers for us to connect to. This should include the peer itself.}\\
+\midrule
+0x11 & \textsc{Peers} & & $(\text{0x}11, (a_0 \in \mathbb{B}_4, p_0 \in \mathbb{P}, i_0 \in \mathbb{B}_{64}), (a_1 \in \mathbb{B}_4, p_1 \in \mathbb{P}, i_1 \in \mathbb{B}_{64}), ...)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{
+Specifies a number of known peers. 
+\begin{itemize}
+\item $a_0$, $a_1$, ... is the node's IPv4 address, a 4-byte array that should be interpreted as the IP address $a_0[0]$.$a_0[1]$.$a_0[2]$.$a_0[3]$.
+\item $p_0$, $p_1$, ... is the node's Port and is an integer.
+\item $i_0$, $i_1$, ... is the node's Unique Identifier and is the 512-bit hash that serves to identify the node.
+\end{itemize}
+}\\
+\midrule
+0x12 & \textsc{Transactions} & & $(\text{0x}12, L_T(T_0), L_T(T_1), ...)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{
+where $L_T$ is the transaction preparation function, as specified in section \ref{ch:block}.
+
+Specify a transaction or transactions that the peer should make sure is included on its transaction queue.
+The items in the list (following the first item 0x12) are transactions in the format described in the main Ethereum specification.
+\begin{itemize}
+\item $T_0$, $T_1$, ... are the transactions that should be assimilated.
+\end{itemize}
+}\\
+\midrule
+0x13 & \textsc{Blocks} && $(\text{0x}13, L_B(b_0), L_B(b_1), ...)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{
+Where $L_B$ is the block preparation function, as specified in section \ref{ch:block}.
+
+Specify a block or blocks that the peer should know about. The items in the list (following the first item, 0x13) are blocks as described in the format described in the main specification.
+\begin{itemize}
+\item $b_0$, $b_1$, ... are the blocks that should be assimilated.
+\end{itemize}
+}\\
+\midrule
+0x14 & \textsc{GetChain} & \textsc{Chain} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{
+Request the peer to send $c$ blocks in the current canonical block chain that are children of one of a number of given blocks, according to a preferential order with $p_0$ being the most prefered.
+If the designated parent is the present block chain head, an empty reply should be sent.
+If none of the parents are in the current canonical block chain, then a NotInChain message should be sent along with $p_n$, the least preferential parent.
+If no parents are passed, then a reply need not be made.
+\begin{itemize}
+\item $p_0$, $p_1$, ...
+are the SHA3 hashes of the parents of blocks that we should be informed of with a \textsc{Blocks} reply.
+Typically, these will be specificed in increasing age (or decreasing block number).
+\item $c$ is the number of children blocks of the most prefered parent that we should be informed of through the corresponding \textsc{Blocks} reply.
+\end{itemize}
+}\\
+\midrule
+0x15 & \textsc{NotInChain} && $(\text{0x}15, p \in \mathbb{B}_{32})$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{Inform the peer that a particular block was not found in its block chain.
+\begin{itemize}
+\item $p$ is the SHA3 hash of the block that was not found in the block chain. Typically, this will be the least preferential (oldest) block hash given in a previous \textsc{GetChain} message.
+\end{itemize}
+}\\
+\midrule
+0x16 & \textsc{GetTransactions} & \textsc{Transactions} & $(\text{0x}16)$ \\
+& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to send all transactions currently in the queue. See \textsc{Transactions}.}\\
+\bottomrule
+\end{tabular*}
 
@@
 The genesis block is 9 items, and is specified thus:
-
-$\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 2^{22}, 0, (), 42 \big), (), () \big)$
-
-Where:
-
-$0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes.
-
-$0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes.
-
-$2^{22}$ refers to the difficulty.
-
-0 refers to the timestamp (the Unix epoch).
-
-$()$ refers to the extradata and the sequences of both uncles and transactions, all empty.
-
+\begin{equation}
+\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 2^{22}, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent and state root hashes, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); $()$ refers to the extradata and the sequences of both uncles and transactions, all empty.
+$\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.
 $\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.
 
+The proof-of-concept series include a development premine, making the state root hash some value other than $0_{256}$. The latest documentation should be consulted for the value of the state root.
+
 \end{document}

### 2014-04-22 Paper.tex
Signitures.

@@
 
-We now have a rogourous specification for the construction of a formal block structure.
+We now have a rigourous specification for the construction of a formal block structure.
 The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}. This has yet to be formally specified.
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer in the range $(0, 2^{256})$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARESTORE}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
+It is assumed that $v$ is the `recovery id', a 2-bit value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
+
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the SHA3 hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny SHA3}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the SHA3 hash of the transaction without the latter three signiture components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_v, T_p, T_g, T_t, T_\mathbf{b}, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_v, T_p, T_g, T_t, T_\mathbf{d}) & \text{otherwise} \end{cases} \\
+h(T) & \equiv & \mathtt{\small SHA3}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
+
+We may then define the sender function $S$ of the transaction as:
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny SHA3}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
+\end{equation}
 

### 2014-04-22 Paper.tex
Various tweaks and fixes.

@@
 With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap.
-Technological-rooted movements like Bitcoin have demonstrated through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
+Technology-rooted movements like Bitcoin have demonstrated, through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
 This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
-Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications allbeit rather simplistic ones.
-
-Ethereum is a project that attempts to build the generalised technology; technology on which all transaction-based state machine concepts may be built.
+Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications albeit rather simplistic ones.
+
+Ethereum is a project which attempts to build the generalised technology; technology on which all transaction-based state machine concepts may be built.
 Moreover it aims to provide to the end-developer a tightly integrated end-to-end system for building software on a hitherto unexplored compute paradigm in the mainstream: a trustful object messaging compute framework.
@@
 Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
-The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol though utilisation of a number of auxilliary parts to the core protocol.
-The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transactions in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet'' protocol aware piece of software.
+The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxilliary parts to the core protocol.
+The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.
 
 Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system.
-It has demonstrate that high efficiency gains can be made if the decentralisation premise is discarded.
+It has demonstrated that high efficiency gains can be made if the decentralisation premise is discarded.
 
 Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s it became clear that algorithm enforcement of agreements could become a significant force in human cooperation.
+Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
 Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
@@
 Formally:
-
-\begin{equation}
-\mathbb{S}_{t+1} \equiv \Upsilon(\mathbb{S}_t, T)
+\begin{equation}
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\mathbb{S}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\mathbb{S}$ allows components to store arbitrary state between transactions.
+In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
 
@@
 Formally, we expand to:
-
-\begin{eqnarray}
-\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, B) \\
+\begin{eqnarray}
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
 \end{eqnarray}
@@
 
-Highly structured, `top-level', state values, are denoted with black-board type.
-They fall into two main sets, those of world-state, which are denoted $\mathbb{S}$ (or a variant thereupon) and those of machine-state, $\mathbb{M}$.
+The two sets of highly structured, `top-level', state values, are denoted with bold lowercase Greek letters.
+They fall into those of world-state, which are denoted $\boldsymbol{\sigma}$ (or a variant thereupon) and those of machine-state, $\boldsymbol{\mu}$.
 
@@
 
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\mathbb{M}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\mathbb{M}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
-
-In the case of the global state $\mathbb{S}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
+
+In the case of the global state $\boldsymbol{\sigma}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
 
@@
 When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in chapter \ref{ch:block}.
+It is formally defined in section \ref{ch:block}.
 
@@
 \item[nonce] A scalar value equal to the number of transactions sent from this address, or, in the case of contract accounts, the number of contract-creations made by this account.
-For account of address $a$ in state $\mathbb{S}$, this would be formally denoted $\mathbb{S}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\mathbb{S}[a]_b$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
 \item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the contract, itself a simple mapping from byte arrays of size 32 to arbitrary length byte arrays.
-The hash is formally denoted $\mathbb{S}[a]_s$.
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence $\texttt{\small TRIE}(\mathbb{S}[a]_\mathbf{s}) \equiv \mathbb{S}[a]_s$.
-It shall be understood that $\mathbb{S}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence $\texttt{\small TRIE}(\boldsymbol{\sigma}[a]_\mathbf{s}) \equiv \boldsymbol{\sigma}[a]_s$.
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 \item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\mathbb{S}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small SHA3}(\mathbf{b}) = \mathbb{S}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small SHA3}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
@@
 \begin{eqnarray}
-L_S(\mathbb{S}) & \equiv & \{ p(a): \mathbb{S}[a] \neq \varnothing \} \\
-p(a) & \equiv & \big(a, \texttt{\small RLP}\big( (\mathbb{S}[a]_n, \mathbb{S}[a]_b, \mathbb{S}[a]_s, \mathbb{S}[a]_c) \big) \big)
-\end{eqnarray}
-
-This function, $L$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
-\begin{equation}
-\forall a: \mathbb{S}[a] = \varnothing \vee (a \in \mathbb{B}_{20} \wedge v(\mathbb{S}[a]))
+L_S(\boldsymbol{\sigma}) & \equiv & \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \} \\
+p(a) & \equiv & \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+\end{eqnarray}
+
+This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+\begin{equation}
+\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
 \end{equation}
@@
 \begin{eqnarray}
-v(x) & \equiv & x_n \in \mathbb{P} \wedge x_p \in \mathbb{P} \wedge (x_r = x_c = \varnothing \vee w(x)) \\
-w(x) & \equiv & x_r \in \mathbb{B}_{32} \wedge \quad x_c \in \mathbb{B}_{32}
+\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_p \in \mathbb{P} \wedge (x_r = x_c = \varnothing \vee w(x)) \\
+\quad w(x) & \equiv & x_r \in \mathbb{B}_{32} \quad \wedge \quad x_c \in \mathbb{B}_{32}
 \end{eqnarray}
@@
 \item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[value] A scalar value equal to the number of wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created contract's account; formally $T_v$.
-\item[gasPrice] A scalar value equal to the number of wei to be paid per unit of gas for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created contract's account; formally $T_v$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be topped-up later; formally $T_g$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
 \item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction; formally $T_t$.
-\item[v, r, s] A scalar and complex value corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
 This is expanded in Appendix \ref{app:signing}.
@@
 
-Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDMA of the SECP-256k1 curve, using the hash of the transaction as the datum to sign.
+Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signiture fields) as the datum to sign.
 For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
@@
 \begin{equation}
-\begin{array}{lclclcl}
-T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge & T_g \in \mathbb{P} \\
-& \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge & T_s \in \mathbb{P} \\
-& \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{b} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} \\
-& \wedge & T_t \in \mathbb{B}_{20} &&\\
+\begin{array}[t]{lclclc}
+T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge \\
+T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
+T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{b} \in \mathbb{B} & \wedge \\
+T_\mathbf{i} \in \mathbb{B} & \wedge & T_t \in \mathbb{B}_{20} &&\\
 \end{array}
@@
 
-We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\mathbb{S}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
-\begin{eqnarray}
-H_u & \equiv & \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) \\
-\wedge \quad H_t & \equiv & \mathtt{\small SHA3}(\mathtt{\small RLP}(L_T^*(B_\mathbf{T}))) \\
-\wedge \quad H_r & \equiv & \mathtt{\small TRIE}(L_S(\Pi(\mathbb{S}, B)))
-\end{eqnarray}
-
-where 
-\begin{equation}
-\mathtt{\small TRIE}(L_S(\mathbb{S})) = {P(B_H)_H}_r
-\end{equation}
-
-this $\texttt{\small TRIE}(L_S(\mathbb{S}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\mathbb{S}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\begin{equation}
+\begin{array}[t]{lclc}
+H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_T^*(B_\mathbf{T}))) & \wedge \\
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
+\end{array}
+\end{equation}
+
+where:
+\begin{equation}
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\end{equation}
+
+thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
 
@@
 Much like the transaction preparation function $L_T$, we assert the types and order of the structure for when the RLP transformation is required:
-
 \begin{eqnarray}
@@
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propogate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because satisfaction of this requirement depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult, and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+Because the nonce must satisfy this requirement, and because satisfaction of this requirement depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
 
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
-& \wedge & H_d = D(H) \\
-& \wedge & H_s > {P(H)_H}_s
+V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
+& & H_d = D(H) \quad \wedge \\
+& & H_s > {P(H)_H}_s
 \end{eqnarray}
@@
 
-Any ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under the control of the miner.
+In general, Ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under the control of the miner.
 Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
@@
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment;
+\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
 \end{enumerate}
 
-Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\mathbb{S}$ the state:
-\begin{equation}
-\mathbb{S}' = \Upsilon(\mathbb{S}, T)
-\end{equation}
-
-Thus $\mathbb{S}'$ is the post-transactional state.
+Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
+\begin{equation}
+\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
+\end{equation}
+
+Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
 
@@
 the validity is determined as:
-\begin{eqnarray}
-S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_n \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_b
-\end{eqnarray}
+\begin{equation}
+\begin{array}[t]{rcl}
+S(T) & \neq & \varnothing \quad \wedge \\
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leq & T_g \quad \wedge \\ 
+v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b
+\end{array}
+\end{equation}
 
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender ($S(T)$) is incremented by one and the balance is reduced by the up-front cost, $v_0$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by the up-front cost, $v_0$.
 The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
@@
 
-We define the checkpoint state $\mathbb{S}_0$:
-\begin{eqnarray}
-\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}_0[S(T)]_b & \equiv & \mathbb{S}[S(T)]_b - v_0 \\
-\mathbb{S}_0[S(T)]_n & \equiv & \mathbb{S}[S(T)]_n + 1
-\end{eqnarray}
-
-Evaluating $\mathbb{S}_P$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
-\begin{equation}
-(\mathbb{S}_P, g') \equiv \begin{cases}
-\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_t = 0 \\
-\Theta^*(\mathbb{S}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
+We define the checkpoint state $\boldsymbol{\sigma}_0$:
+\begin{eqnarray}
+\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - v_0 \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\boldsymbol{\sigma}_P$) and remaining gas ($g'$):
+\begin{equation}
+(\boldsymbol{\sigma}_P, g') \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_t = 0 \\
+\Theta_{0,1}(\boldsymbol{\sigma}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta^*$ to denote the fact that only the first two components of the function's value are taken; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{0,1}$ to denote the fact that only the first two components of the function's value are taken; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
-So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
-\begin{eqnarray}
-\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
-\mathbb{S}'[s]_b & \equiv & \mathbb{S}_P[s]_b + g' T_p \\
-\mathbb{S}'[m]_b & \equiv & \mathbb{S}_P[m]_b + (T_g - g' - g_d) T_p \\
+So we define the final state $\boldsymbol{\sigma}'$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g' T_p \\
+\boldsymbol{\sigma}'[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g' - g_d) T_p \\
 m & \equiv & {B_H}_b
@@
 %\begin{eqnarray}
-%\mathbb{S}'[s]_b & \equiv & \mathbb{S}_P[s]_b + g
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
 %\end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\mathbb{S}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\mathbb{S}_P, g')$, as in \ref{ch:transactions}:
-\begin{equation}
-(\mathbb{S}_P, g') \equiv \Lambda(\mathbb{S}_0, s, n, g, p, v, \mathbf{i}, \mathbf{b})
-\end{equation}
-
-The address of the new contract is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure \texttt{[ sender, nonce ]}.
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\boldsymbol{\sigma}_P, g')$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g') \equiv \Lambda(\boldsymbol{\sigma}_0, s, n, g, p, v, \mathbf{i}, \mathbf{b})
+\end{equation}
+
+The address of the new contract is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure containing only the sender and the nonce.
 In the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
@@
 A(s, n) & \equiv & a \quad \text{where:} \\
-a & = & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing \\
+a & = & \arg \min_x : x \geqslant a' \wedge \boldsymbol{\sigma}_0[x] = \varnothing \\
 a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
@@
 
-where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\mathbb{S}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\mathbb{S}^*$:
-\begin{eqnarray}
-\mathbb{S}^* & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}^*[A(s, \mathbb{S}[s]_n)] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
+\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n)] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
 \end{eqnarray}
@@
 Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\mathbb{S}^{**}$ and available gas remaining $g'$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$ and available gas remaining $g'$.
 
@@
 In this exceptional case we say an Out-of-Gas exception has occured: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-The gas remaining should of course be zero.
+The gas remaining should be zero.
 If the creation was conducted as the manifestation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state and gas as $\mathbb{S}_P$, $g'$ where:
-\begin{eqnarray}
-(\mathbb{S}^{**}, g') & \equiv & \Xi(\mathbb{S}^*, g, I) \\
-\mathbb{S}_P & \equiv & \begin{cases}
-\mathbb{S}_0 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
-\mathbb{S}^{**} & \text{otherwise}
+Thus formally, we may specify the resultant state and gas as $(\boldsymbol{\sigma}_P, g')$ where:
+\begin{eqnarray}
+(\boldsymbol{\sigma}^{**}, g') & \equiv & \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\boldsymbol{\sigma}_P & \equiv & \begin{cases}
+\boldsymbol{\sigma}_0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{otherwise}
 \end{cases} \\
@@
 I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & [] \\
+I_\mathbf{d} & \equiv & () \\
 I_s & \equiv & s \\
@@
 
-$I_\mathbf{d}$ evaluates to the empty array. $I_H$ and $I_n$ have no special treatment and are determined from the blockchain.
+$I_\mathbf{d}$ evaluates to the empty tuple. $I_H$ and $I_n$ have no special treatment and are determined from the blockchain.
 
@@
 \begin{equation}
-(\mathbb{S}_P, g', \mathbf{o}) \equiv \Theta(\mathbb{S}, s, o, r, g, p, v, \mathbf{d})
-\end{equation}
-
-We define $\mathbb{S}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
-\begin{eqnarray}
-\mathbb{S}_1 & \equiv & \mathbb{S} \quad \text{except:} \\
-\mathbb{S}_1[r]_b & \equiv & \mathbb{S}_1[r]_b + v
+(\boldsymbol{\sigma}_P, g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
+\end{equation}
+
+We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
+\begin{eqnarray}
+\boldsymbol{\sigma}_1 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
+\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}_1[r]_b + v
 \end{eqnarray}
@@
 If the recipient account contains no code to be executed (i.e.
-it's not a contract), then we define the final state as being equivalent to the checkpoint state.
+it is not a contract), then we define the final state as being equivalent to the checkpoint state.
 In the case that the recipient contains code to be executed (i.e.
-that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\mathbb{S}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
+that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
 \begin{eqnarray}
-\mathbb{S}_P & \equiv & \begin{cases}
-\mathbb{S}_1 & \text{if} \quad \mathbb{S}_1[r]_c = \varnothing \\
-\mathbb{S}_1 & \text{if} \quad \mathbb{S}^{**} = \varnothing \\
-\mathbb{S}^{**} & \text{otherwise}
+\boldsymbol{\sigma}_P & \equiv & \begin{cases}
+\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}_1[r]_c = \varnothing \\
+\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{otherwise}
 \end{cases} \\
-(\mathbb{S}^{**}, g', \mathbf{o}) & \equiv & \Xi(\mathbb{S}_1, g, I) \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{o}) & \equiv & \Xi(\boldsymbol{\sigma}_1, g, I) \\
 I_a & \equiv & a \\
@@
 I_v & \equiv & v \\
-\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \mathbb{S}[r]_c
-\end{eqnarray}
-
-We assert that the state database contains the entry $(\mathtt{\tiny SHA3}(I_\mathbf{b}), I_\mathbf{b})$.
+\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny SHA3}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
-Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state in the state database.
+Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state.
 All locations in both storage and memory are well-defined initially as zero.
@@
 Over a contract's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
-This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result is an addition memory usage fee.
-Due to this fee it's highly unlikely addresses will ever go above 32-bit bounds since at the present price of ether and default gas price, that would cost around US\$20M for the memory fee alone.
+This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
+Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds since at the present price of Ether and default gas price, that would cost around US\$20M for the memory fee alone.
 
@@
 
-In addition to the system state $\mathbb{S}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
+In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
 
@@
 \item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\mathbb{S}'$ and the remaining gas $g'$, given these definitions:
-
-\begin{equation}
-(\mathbb{S}', g') \equiv \Xi(\mathbb{S}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$ and the remaining gas $g'$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g') \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 We must now define the $\Xi$ function.
-In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\mathbb{S}$ and the machine state, $\mathbb{M}$.
+In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$.
 Formally, we define it recursively with a function $X$.
@@
 The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$.
-
-\begin{eqnarray}
-\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
-\mathbb{M}_g & \equiv & g \\
-\mathbb{M}_{pc} & \equiv & 0 \\
-\mathbb{M}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\mathbb{M}_i & \equiv & 0 \\
-\mathbb{M}_\mathbf{s} & \equiv & ()
-\end{eqnarray}
-\begin{equation}
-X(\mathbb{S}, \mathbb{M}, I) \equiv \begin{cases}
-(\mathbb{S}, \mathbb{M}, I, ()) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
-O(\mathbb{S}, \mathbb{M}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\boldsymbol{\mu}_g & \equiv & g \\
+\boldsymbol{\mu}_{pc} & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
+X(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \begin{cases}
+(\boldsymbol{\sigma}, \boldsymbol{\mu}, I, ()) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)) & \text{otherwise}\\
 \end{cases}
@@
 where
-
-\begin{eqnarray}
-\mathbf{o} & \equiv & H(\mathbb{M}, I) \\
+\begin{eqnarray}
+\mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
 (a, b, c) \cdot d & \equiv & (a, b, c, d)
@@
 
-The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\mathbb{M}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
 
@@
 For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
-
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 
-We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack movement items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
+We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
 
@@
 \begin{equation}
-Z(\mathbb{S}, \mathbb{M}, I) \equiv
-\begin{array}[t]{rl}
-& \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \\
-\vee & \mathbf{\delta}_w = \varnothing \\
-\vee & \lVert\mathbb{M}_\mathbf{s}\rVert < \mathbf{\delta}_w
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+\begin{array}[t]{l}
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w
 \end{array}
@@
 
-We define the normal halting function $H$:
-\begin{equation}
-H(\mathbb{M}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\mathbb{M}) & \text{if} \quad w = \text{\small RETURN} \\
+The normal halting function $H$ is defined:
+\begin{equation}
+H(\boldsymbol{\mu}, I) \equiv \begin{cases}
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
 () & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
@@
 \begin{eqnarray}
-O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', I) \\
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\mathbb{M}'_\mathbf{s}\rVert & \equiv & \lVert\mathbb{M}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\mathbb{M}'_\mathbf{s}\rVert): \mathbb{M}'_\mathbf{s}[x] & \equiv & \mathbb{M}_\mathbf{s}[x+\Delta]
+\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathbb{M}'_{g} & \equiv & \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
-\mathbb{M}'_{pc} & \equiv & \begin{cases}
-J_o(\mathbb{M}) & \text{if} \quad w \in \{ \text{\small JUMP, JUMPI} \} \\
-\mathbb{M}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
-\mathbb{M}_{pc} + 1 & \text{otherwise}
+\boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+J_o(\boldsymbol{\mu}) & \text{if} \quad w \in \{ \text{\small JUMP, JUMPI} \} \\
+\boldsymbol{\mu}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
+\boldsymbol{\mu}_{pc} + 1 & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-\mathbb{M}'_\mathbf{m} & \equiv & \mathbb{M}_\mathbf{m} \\
-\mathbb{M}'_i & \equiv & \mathbb{M}_i \\
-\mathbb{S}' & \equiv & \mathbb{S}
+\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
+\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
 \end{eqnarray}
@@
 However, instructions do typically alter one or several components of these values.
-Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and formal a description of the gas requirements.
+Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and a formal description of the gas requirements.
 
@@
 
-This increases overall network security by making it much harder for an adversary to silently mine a canonical blockchain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of reversing existing blocks and the transactions within.
+This increases overall network security by making it much harder for an adversary to silently mine a canonical blockchain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of overriding existing blocks and reversing the transactions within.
 
@@
 B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
-B' & \equiv & P(H) \\
-H & \equiv & B_H
+B' & \equiv & P(B_H)
 \end{eqnarray}
@@
 \begin{enumerate}
-\item Validate (or, if mining, determine) uncles.
-\item Apply rewards.
-\item Verify (or, if mining, compute a valid) state and nonce.
+\item Validate (or, if mining, determine) uncles;
+\item apply rewards;
+\item verify (or, if mining, compute a valid) state and nonce.
 \end{enumerate}
@@
 \begin{eqnarray}
-\Omega(B, \mathbb{S}) & \equiv & \mathbb{S}': \mathbb{S}' = \mathbb{S} \quad \text{except:} \\
-\mathbb{S}'[{B_H}_b] & = & \mathbb{S}[{B_H}_b] + R_b \\
-\forall U \in B_\mathbf{U}: \quad \mathbb{S}'[U_b] & = & \mathbb{S}[U_b] + \frac{7}{8}R_b
+\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
+\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + R_b \\
+\forall U \in B_\mathbf{U}: \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{7}{8}R_b
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\mathbb{S}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\mathbb{S}_i: r(\mathbb{S}_i) = {P(B_H)_H}_r & \text{otherwise}
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
 \end{cases}
@@
 
-Here, $r(\mathbb{S}_i)$ means the hash of the root node of a trie of state $\mathbb{S}_i$; this is stored in the state database since the trie is a mutable data structure.
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature a mutable data structure.
 
@@
 \begin{eqnarray}
-\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...) \\
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...) \\
 B & \equiv & (H, (T_0, T_1, ...), \mathbf{U})
@@
 For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of or reward from specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to ether at roughly the same rate for anyone around the world.
+The requirement of, or reward from, specialised and uncommon hardware should be minimised.
+This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
 
@@
 
-The state database won't be forced to maintain all past state trie structures in to the future.
+The state database won't be forced to maintain all past state trie structures into the future.
 It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
@@
 
-Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in over some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
+Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
 
@@
 
-Some form of hierarchical structure, either through consolidating smaller lighter-weight chains into the main block or through building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
+Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
 Parallelism could also come from a prioritised set of parallel blockchains, consolidated each block and with duplicate or invalid transactions thrown out accordingly.
 
-Verifyable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
+Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
 
@@
 \end{eqnarray}
-
+gc
 We may then define the sender function $S$ of the transaction as:
@@
 \begin{eqnarray}
-C(\mathbb{S}, \mathbb{M}, I) & \equiv & G_{memory}(\mathbb{M}'_i - \mathbb{M}_i) + \begin{cases}
-C_\text{\tiny SSTORE}(\mathbb{S}, \mathbb{M}) & \text{if} \quad w = \text{\small SSTORE} \\
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & G_{memory}(\boldsymbol{\mu}'_i - \boldsymbol{\mu}_i) + \begin{cases}
+C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
 G_{call} & \text{if} \quad w = \text{\small CALL}\\
@@
 \end{cases} \\
-w & \equiv & \begin{cases} I_\mathbf{b}[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_\mathbf{b} \rVert \\
+w & \equiv & \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write; $\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
-In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\mathbb{S}', \mathbb{M}')$ such that:
-
-\begin{equation}
-O(\mathbb{S}, \mathbb{M}, I) \equiv (\mathbb{S}', \mathbb{M}', I) \quad \text{with exceptions, as noted}
+In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
+
+\begin{equation}
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', I) \quad \text{with exceptions, as noted}
 \end{equation}
@@
 0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
-&&&& $\mathbb{M}'_R = []$ \\
+&&&& $\boldsymbol{\mu}'_R = []$ \\
 \midrule
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] \times \mathbb{M}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] - \mathbb{M}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \lfloor\mathbb{M}_\mathbf{s}[0] \div \mathbb{M}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \lfloor\mathbb{M}_\mathbf{s}[0] \div \mathbb{M}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \mathbb{M}_\mathbf{s}[0] \bmod \mathbb{M}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0\\ \mathbb{M}_\mathbf{s}[0] \bmod \mathbb{M}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0] ^ {\mathbb{M}_\mathbf{s}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
 \midrule
 0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv -\mathbb{M}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv -\boldsymbol{\mu}_\mathbf{s}[0]$ \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] < \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] > \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] = \mathbb{M}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \mathbb{M}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \end{tabular*}
@@
 \multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
-\multicolumn{5}{l}{$\mathbb{M}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\mathbb{M}_\mathbf{s}[0]$} \vspace{5pt} \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[0]$} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
 0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \wedge \mathbb{M}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \vee \mathbb{M}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \mathbb{M}_\mathbf{s}[0]_i \oplus \mathbb{M}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..7]: \mathbb{M}'_\mathbf{s}[0]_i \equiv \begin{cases} \mathbb{M}_\mathbf{s}[1]_{(i + 8\mathbb{M}_\mathbf{s}[0])} & \text{if} \quad \mathbb{M}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..7]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathtt{\tiny SHA3}(\mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny SHA3}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{S}[I_a]_b$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_b$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
 &&&& This is the sender of original transaction; it is never a contract. \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_c$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_p$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_p$ \\
 &&&& $I_{H_p}$ is the previous block's hash. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_b$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_b$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_t$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_t$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv n({I_H})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv n({I_H})$ \\
 &&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
@@
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv 10^6$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 10^6$ \\
 &&&& Presently, the gas limit is hard-coded in the protocol. \\
@@
 0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
 0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{s}[1]$ \\
-&&&& $\mathbb{M}'_\mathbf{s}[1] \equiv \mathbb{M}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
 0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_\mathbf{m}[\mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
 \midrule
 0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\mathbb{M}'_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] \dots (\mathbb{M}_\mathbf{s}[0] + 31) ] \equiv \mathbb{M}_\mathbf{s}[1]$ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
 \midrule
 0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\mathbb{M}'_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] ] \equiv (\mathbb{M}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
 \midrule
 0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{S}[I_a]_\mathbf{s}[\mathbb{M}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\mathbb{S}'[I_a]_\mathbf{s}[ \mathbb{M}_\mathbf{s}[0] ] \equiv \mathbb{M}_\mathbf{s}[1] $ \\
-&&&& $C_{\text{\tiny SSTORE}}(\mathbb{S}, \mathbb{M}) \equiv \begin{cases}
-2G_{sstore} & \text{if} \quad \mathbb{M}_\mathbf{s}[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_\mathbf{s}[\mathbb{M}_\mathbf{s}[0]] = 0 \\
-0 & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0 \; \wedge \; \mathbb{S}[I_a]_\mathbf{s}[\mathbb{M}_\mathbf{s}[0]] \neq 0 \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+2G_{sstore} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 G_{sstore} & \text{otherwise}
@@
 0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
 0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_\mathbf{s}[0] & \text{if} \quad \mathbb{M}_\mathbf{s}[1] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
 0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{i}$ \\
 \midrule
 0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv \mathbb{M}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \bottomrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{b}[\mathbb{M}_{pc} + 1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[\boldsymbol{\mu}_{pc} + 1]$ \\
 &&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 2) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) ]$ \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\mathbb{M}_{pc} + 1) \dots (\mathbb{M}_{pc} + 32) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) ]$ \\
 \bottomrule
@@
 0xf0 & {\small CREATE} & 5 & 1 & Create a new contract. \\
-&&&& $\mathbf{b} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[1] \dots (\mathbb{M}_\mathbf{s}[1] + \mathbb{M}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $\mathbf{i} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[3] \dots (\mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\mathbb{S}', \mathbb{M}'_g) \equiv \Lambda(\mathbb{S}^*, I_a, I_o, \mathbb{M}_g, I_p, \mathbb{M}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
-&&&& $\mathbb{S}^* \equiv \mathbb{S} \quad \text{except} \quad \mathbb{S}^*[I_a]_n = \mathbb{S}[I_a]_n + 1$ \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{b} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
-&&&& $x=A(I_a, \mathbb{S}[I_a]_n)$, the address of the newly created contract, otherwise. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_\mathbf{s}[1] + \mathbb{M}_\mathbf{s}[2], \mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4]) \div 32 })$ \\
+&&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created contract, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4]) \div 32 })$ \\
 \midrule
 0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
-&&&& $\mathbf{i} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[3] \dots (\mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $\mathbf{o} \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[5] \dots (\mathbb{M}_\mathbf{s}[5] + \mathbb{M}_\mathbf{s}[6] - 1) ]$ \\
-&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \mathbb{M}_\mathbf{s}[2] = 0 \\ \mathbb{M}_g - \min(\mathbb{M}_g, \mathbb{M}_\mathbf{s}[2]) & \text{otherwise} \end{cases}$ \\
-&&&& $(\mathbb{S}', g', \mathbf{o}) \equiv \Theta(\mathbb{S}, I_a, I_o, \mathbb{M}_\mathbf{s}[0], g, I_p, \mathbb{M}_g - g_k, \mathbb{M}_\mathbf{s}[1], \mathbf{i})$ \\
-&&&& $\mathbb{M}'_g \equiv g' + g_k$ \\
-&&&& $\mathbb{M}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
+&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0 \\ \boldsymbol{\mu}_g - \min(\boldsymbol{\mu}_g, \boldsymbol{\mu}_\mathbf{s}[2]) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[0], g, I_p, \boldsymbol{\mu}_g - g_k, \boldsymbol{\mu}_\mathbf{s}[1], \mathbf{i})$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv g' + g_k$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
 &&&& $x=1$ otherwise. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ \max(\mathbb{M}_\mathbf{s}[3] + \mathbb{M}_\mathbf{s}[4], \mathbb{M}_\mathbf{s}[5] + \mathbb{M}_\mathbf{s}[6]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\
 \midrule
 0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\mathbb{M}) \equiv \mathbb{M}_\mathbf{m}[ \mathbb{M}_\mathbf{s}[0] \dots ( \mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\mathbb{M}'_i \equiv \max(\mathbb{M}_i, \ceil{ (\mathbb{M}_\mathbf{s}[0] + \mathbb{M}_\mathbf{s}[1]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
 \midrule
 0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
-&&&& $\mathbb{S}'[\mathbb{M}_\mathbf{s}[0]]_b \equiv \mathbb{S}[\mathbb{M}_\mathbf{s}[0]]_b + \mathbb{S}[I_a]_b$ \\
-&&&& $\mathbb{S}'[I_a] \equiv \varnothing$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a] \equiv \varnothing$ \\
 \bottomrule

### 2014-04-22 Paper.tex
Nick's corrections.
More discussion

@@
 Thus we may define a world-state collapse function $L_S$:
-\begin{eqnarray}
-L_S(\boldsymbol{\sigma}) & \equiv & \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \} \\
-p(a) & \equiv & \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
-\end{eqnarray}
-
+\begin{equation}
+L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+\end{equation}
+where
+\begin{equation}
+p(a) \equiv \begin{cases}
+\big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b) \big) \big) & \text{if} \; \boldsymbol{\sigma}[a]_s = \varnothing\\
+\big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big) & \text{otherwise}
+\end{cases}
+\end{equation}
 This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
@@
 \begin{eqnarray}
-\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_p \in \mathbb{P} \wedge (x_r = x_c = \varnothing \vee w(x)) \\
-\quad w(x) & \equiv & x_r \in \mathbb{B}_{32} \quad \wedge \quad x_c \in \mathbb{B}_{32}
+\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge (x_s = x_c = \varnothing \vee w(x)) \\
+\quad w(x) & \equiv & x_s \in \mathbb{B}_{32} \quad \wedge \quad x_c \in \mathbb{B}_{32}
 \end{eqnarray}
@@
 
+There are several patterns of contracts engineering that allow particular useful behaviours; two of these that I will briefly discuss are data feeds and random numbers.
+
 \subsection{Data Feeds}
-
-External server runs node.
-Creates \& signs a transaction every minute containing new data.
-Sends to contract which knows to accept data only from this address.
-Allows polling of data from all other contracts (possibly for a microfee).
+A data feed contract is one which provides a single service: it gives access to information from the external world within Ethereum.
+The accuracy and timeliness of this information is not guaranteed and it is the task of a secondary contract author---the contract that utilises the data feed---to determine how much trust can be placed in any single data feed.
+
+The general pattern involves a single contract within Ethereum which, when given a message call, replies with some timely information concerning an external phenomenon.
+An example might be the local temperature of New York City.
+This would be implemented as a contract that returned that value of some known point in storage.
+Of course this point in storage must be maintained with the correct such temperature, and thus the second part of the pattern would be for an external server to run an Ethereum node, and immediately on discovery of a new block, creates a new valid transaction, sent to the contract, updating said value in storage.
+The contract's code would accept such updates only from the identity contained on said server.
 
 \subsection{Random Numbers}
-
-Hash the block timestamp; for a series of pseudorandoms, take the previous, add some constant and hash the result.
-
-For a more secure pseudo-random offering both parties agree on a number of random data feed contracts; these are concatenated along with the block timestamp and hashed to produce the first number in the series.
+Providing random numbers within a deterministic system is, naturally, an impossible task.
+However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
+Such data might include the block's hash, the block's timestamp and the block's coinbase address.
+For a series of such numbers, a trivial solution would be to work from the previous pseudo-random number, adding some constant amount and hashing the result.
+
+This strategy does have a downside: a miner with sufficient power could alter any of the above values in order to deliver a seed in order to alter the outcome of the pseudorandom-based executions.
+For a more secure pseudo-random offering, all involved parties could agree on a number of data feed contracts; these could be combined along with the block timestamp and hashed to produce the first number in the series.
+By spreading the inputs and thus the trust between numerous parties the likelihood of a malicious miner altering the outcome becomes increasingly less likely.
 

### 2014-04-23 Paper.tex
Switch transaction order for PoC-4.
Wire protocol tweak.

@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_v, T_p, T_g, T_t, T_\mathbf{b}, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
-(T_n, T_v, T_p, T_g, T_t, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
+(T_n, T_v, T_t, T_p, T_g, T_\mathbf{b}, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
+(T_n, T_v, T_t, T_p, T_g, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
 \end{cases}
@@
 \midrule
-0x14 & \textsc{GetChain} & \textsc{Chain} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
+0x14 & \textsc{GetChain} & \textsc{Blocks} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
 & \multicolumn{3}{p{0.8\columnwidth}}{

### 2014-04-29 Paper.tex
PoC-4
More fixes.

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept V}}}
 \author{
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-J_o(\boldsymbol{\mu}) & \text{if} \quad w \in \{ \text{\small JUMP, JUMPI} \} \\
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
+J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
 \boldsymbol{\mu}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\

### 2014-05-02 Paper.tex

@@
 There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
-This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetance, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
+This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
 By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
@@
 Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
-The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxilliary parts to the core protocol.
+The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxiliary parts to the core protocol.
 The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.
@@
 It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
-The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissable.
+The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
 Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
@@
 In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes refered to by the Old English \DH{}.
+To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
 The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accoringly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 
-Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signiture fields) as the datum to sign.
+Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
 For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
@@
 \item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block; formally $H_t$.
-\item[difficulty] A scalar value corresonding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relavant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
 \item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
@@
 
-We now have a rigourous specification for the construction of a formal block structure.
+We now have a rigorous specification for the construction of a formal block structure.
 The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
@@
 
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propogate newly created blocks that would otherwise overwrite (``rewrite'') history.
+This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
 Because the nonce must satisfy this requirement, and because satisfaction of this requirement depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
@@
 Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this exceptional case we say an Out-of-Gas exception has occured: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this exceptional case we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 The gas remaining should be zero.
@@
 Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
-See Appendix \ref{app:vm} for a rigourous definition of the EVM gas cost.
+See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
 
@@
 
-This scheme notably ignores so-called \textit{stale} blocks: valid, mined blocks, which were propogated too late into the network and thus were beaten to network consensus by a sibling block (one with the same parent).
-Such blocks become more common as the network propogation time approaches the ideal inter-block time.
-However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to butress the more popular blockchain making it a stronger choice over less popular (though potentially longer) competitors.
+This scheme notably ignores so-called \textit{stale} blocks: valid, mined blocks, which were propagated too late into the network and thus were beaten to network consensus by a sibling block (one with the same parent).
+Such blocks become more common as the network propagation time approaches the ideal inter-block time.
+However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to buttress the more popular blockchain making it a stronger choice over less popular (though potentially longer) competitors.
 
@@
 A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
-$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligable amount of processing by the EVM.
+$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
 Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
@@
 
-\item[Autonomous Object] A virtual object existant only within the hypothetical state of Ethereum.  Has an intrinsic address. Incorporated only as the state of the storage component of the VM.
+\item[Autonomous Object] A virtual object existent only within the hypothetical state of Ethereum.  Has an intrinsic address. Incorporated only as the state of the storage component of the VM.
 
@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (ala Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
 
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole rquirement in terms of the protocol specification is to provide a single 32-byte value that identifies a given set of key-value pairs.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single 32-byte value that identifies a given set of key-value pairs.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
@@
 Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision.
-In reality, a sensible implemention will not fully recompute the trie root hash for each set.
+In reality, a sensible implementation will not fully recompute the trie root hash for each set.
 
@@
 
-The message hash, $h(T)$, to be signed is the SHA3 hash of the transaction without the latter three signiture components, formally described as $T_r$, $T_s$ and $T_w$:
+The message hash, $h(T)$, to be signed is the SHA3 hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
@@
 The wire-protocol specifies a network-level protocol for how two peers can communicate.
-It includes handshake procedures and the means for transfering information such as peers, blocks and transactions.
+It includes handshake procedures and the means for transferring information such as peers, blocks and transactions.
 Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.
@@
 are the SHA3 hashes of the parents of blocks that we should be informed of with a \textsc{Blocks} reply.
-Typically, these will be specificed in increasing age (or decreasing block number).
-\item $c$ is the number of children blocks of the most prefered parent that we should be informed of through the corresponding \textsc{Blocks} reply.
+Typically, these will be specified in increasing age (or decreasing block number).
+\item $c$ is the number of children blocks of the most preferred parent that we should be informed of through the corresponding \textsc{Blocks} reply.
 \end{itemize}

### 2014-05-02 Paper.tex
Additional instructions for PoC-5.

@@
 \midrule
-0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
+0x37 & {\small CALLDATACOPY} & 0 & 3 & Copy input data in current environment to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+\midrule
+0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
+\midrule
+0x39 & {\small CODECOPY} & 0 & 3 & Copy code running in current environment to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\midrule
+0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\

### 2014-05-02 Paper.tex
Most of PoC-5.

@@
 
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+
+\begin{equation}
+\ell(\mathbf{x}) \equiv \mathbf{x}[\lVert \mathbf{x} \rVert - 1]
+\end{equation}
+
 \section{Blocks, State and Transactions} \label{ch:bst}
@@
 
-If the latter two fields are missing, the node represents a simple (non-contract) account: it maintains no state and executes no code on receipt of a transaction.
+If the latter field is the SHA3 hash of the empty string, i.e.
+$\boldsymbol{\sigma}[a]_c = \texttt{\small SHA3}\big(()\big)$, then the node represents a simple account, sometimes refered to as a ``non-contract'' account.
 
@@
 \begin{eqnarray}
-\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge (x_s = x_c = \varnothing \vee w(x)) \\
-\quad w(x) & \equiv & x_s \in \mathbb{B}_{32} \quad \wedge \quad x_c \in \mathbb{B}_{32}
+\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
 \end{eqnarray}
@@
 \begin{description}
-\item[body] An unlimited size byte array specifying the EVM-code for the contract, formally $T_\mathbf{b}$.
 \item[init] An unlimited size byte array specifying the EVM-code for the contract initialisation procedure, formally $T_\mathbf{i}$.
@@
 
-Both \textbf{body} and \textbf{init} are EVM-code fragments; \textbf{body} is executed each time the contract account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is a piece of code that is executed only once at contract creation and that gets discarded immediately thereafter.
+\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executed each time the account receives a message call (either through a transaction or due to the internal execution of code).
+\textbf{init} is executed only once at contract creation and that gets discarded immediately thereafter.
 
@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_v, T_t, T_p, T_g, T_\mathbf{b}, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
+(T_n, T_v, T_t, T_p, T_g, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
 (T_n, T_v, T_t, T_p, T_g, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
@@
 T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
-T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{b} \in \mathbb{B} & \wedge \\
-T_\mathbf{i} \in \mathbb{B} & \wedge & T_t \in \mathbb{B}_{20} &&\\
+T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} & \wedge \\
+T_t \in \mathbb{B}_{20} &&&&\\
 \end{array}
@@
 \item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block; formally $H_t$.
+\item[transactionsTrie] The SHA3 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[minGasPrice] A scalar value equal to the minimum price of gas a transaction must have provided in order to be sufficient for inclusion by this miner in this block; formally $H_m$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_u$.
 \item[extraData] An arbitrary byte array containing data relevant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
@@
 
-The other two components in the block are simply a list of uncle block headers (of the same format as above) and a list of the transactions (as defined previously).
+The other two components in the block are simply a list of uncle block headers (of the same format as above) and a list of the transaction information (as defined previously).
 Formally, we can refer to a block $B$:
@@
 H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_T^*(B_\mathbf{T}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert: (i, B_\mathbf{T}[i])\}) & \wedge \\
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
@@
 
-where:
+where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+% TODO: formal.
+
+Furthermore:
+
 \begin{equation}
@@
 
-thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+
+TODO: Transaction manifest more than just transactions: $(T[i], s[i], g[i])$
+
+TODO: Remove extrameous references to number.
 
@@
 D(H) \equiv \begin{cases}
-2^{22} & \quad \text{if genesis block}\\
-{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{if $H_s < {P(H)_H}_s + 42$}\\
-{P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \quad \text{otherwise}\\
+2^{22} & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 42\\
+{P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{otherwise}\\
 \end{cases}
 \end{equation}
+
+The canonical gas limit of a block of header $H$ is defined as $L(H)$:
+\begin{eqnarray}
+L(H) & \equiv & \begin{cases}
+10^6 & \text{if} \quad H_i = 0\\
+L'(H) & \text{otherwise}
+\end{cases}\\
+L'(H) & \equiv & \big\lfloor \frac{1023 {P(H)_H}_l + \lfloor \frac{6}{5}{P(H)_H}_u \rfloor}{1024}\big\rfloor
+\end{eqnarray}
 
@@
 & & H_d = D(H) \quad \wedge \\
-& & H_s > {P(H)_H}_s
-\end{eqnarray}
+& & H_l = L(H) \quad \wedge \\
+& & H_s > {P(H)_H}_s \quad \wedge \\
+& & \lVert H_x \rVert < 1024
+\end{eqnarray}
+
+Noting additionally that \textbf{extraData} must be at most 1024 bytes.
 
@@
 
-Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
+Thus $\boldsymbol{\sigma}'$ is the post-transactional state. We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, to be defined later.
 
@@
 g_0 \equiv \begin{cases}
-g_d + G_{create} & \text{if} \quad T_t = 0 \\
-g_d + G_{call} & \text{otherwise} \\
+\lVert T_\mathbf{i}\rVert G_{data} + G_{transaction} & \text{if} \quad T_t = 0 \\
+\lVert T_\mathbf{d}\rVert G_{data} + G_{transaction} & \text{otherwise} \\
 \end{cases}
@@
 
-where
-\begin{equation}
-g_d \equiv \begin{cases} (\lVert T_\mathbf{i}\rVert + \lVert T_\mathbf{b}\rVert) G_{data} & \text{if} \quad T_t = 0 \\
- \lVert T_\mathbf{d}\rVert G_{data} & \text{otherwise}
- \end{cases}
-\end{equation}
-
-where $\lVert T_\mathbf{d}\rVert$, $\lVert T_\mathbf{i}\rVert$ and $\lVert T_\mathbf{b}\rVert$ are the sizes, in bytes, of the transaction's associated data, initialisation and body EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
-The portion of $g_0$ that will not be paid to the miner is $g_d$, and is removed from the system.
+where $\lVert T_\mathbf{d}\rVert$ and $\lVert T_\mathbf{i}\rVert$ are the sizes, in bytes, of the transaction's associated data and initialisation EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
 
@@
 
-the validity is determined as:
+The validity is determined as:
 \begin{equation}
@@
 
-The transaction must not require more gas to be utilised in this block than the global block limit of gas, specified at $10^6$.
+The sum of the transaction gas limit and the gas utilised in this block prior must be no greater than the block's \textbf{gasLimit}, $H_l$.
+
+% TODO: specify formally.
 
@@
 \boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g' T_p \\
-\boldsymbol{\sigma}'[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g' - g_d) T_p \\
+\boldsymbol{\sigma}'[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g') T_p \\
 m & \equiv & {B_H}_b
 \end{eqnarray}
+
+And finally specify $\Upsilon^g$, the total gas used in this transaction:
+\begin{equation}
+\Upsilon^g(\boldsymbol{\sigma}, T) \equiv g_0 + g'
+\end{equation}
 
@@
 
-There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with the two arbitrary length byte arrays, $\mathbf{i}$ for the initialisation EVM code and $\mathbf{b}$ for the EVM code of the body.
+There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}_P, g') \equiv \Lambda(\boldsymbol{\sigma}_0, s, n, g, p, v, \mathbf{i}, \mathbf{b})
+(\boldsymbol{\sigma}_P, g') \equiv \Lambda(\boldsymbol{\sigma}_0, s, n, g, p, v, \mathbf{i})
 \end{equation}
@@
 
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the code, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the empty string, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
 \begin{eqnarray}
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n)] & \equiv & ( 0, v, \varnothing, \mathtt{\tiny SHA3}(\mathbf{b}) )
+\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n)] & \equiv & \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
 \end{eqnarray}
@@
 Thus formally, we may specify the resultant state and gas as $(\boldsymbol{\sigma}_P, g')$ where:
-\begin{eqnarray}
-(\boldsymbol{\sigma}^{**}, g') & \equiv & \Xi(\boldsymbol{\sigma}^*, g, I) \\
-\boldsymbol{\sigma}_P & \equiv & \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{otherwise}
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad \boldsymbol{\sigma}_P & \equiv & \begin{cases}
+\boldsymbol{\sigma}_0 & \text{if} \quad \mathbf{o} = () \\
+\begin{array}[b]{l}
+\boldsymbol{\sigma}^{**} \quad \text{except:} \\
+\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}} = \mathbf{o}
+\end{array} & \text{otherwise}
 \end{cases} \\
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple. $I_H$ and $I_n$ have no special treatment and are determined from the blockchain.
+$I_\mathbf{d}$ evaluates to the empty tuple.
+$I_H$ and $I_n$ have no special treatment and are determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}_P$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created contract's account, with $\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+\subsection{Subtleties}
+Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code.
+Thus any message call received by it during this time causes no code to be executed.
+If the initialisation execution ends with a {\small SUICIDE} instruction or normal {\small STOP} code, or if the code returned is otherwise empty, then the state changes are effectively reverted, the final state being equivalent to $\boldsymbol{\sigma}_0$, the pre-execution checkpoint state.
 
@@
 
-If the recipient account contains no code to be executed (i.e.
-it is not a contract), then we define the final state as being equivalent to the checkpoint state.
-In the case that the recipient contains code to be executed (i.e.
-that the account is a contract), then the contract's body code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
+The contract's body code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
@@
 \item {\small STOP} costs $G_{stop}$ gas
+\item {\small SUICIDE} costs $G_{suicide}$ gas
 \item All other operations cost $G_{step}$ gas.
@@
 
-Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
 See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
@@
 
-The process of finalising a block involves three stages:
+The process of finalising a block involves four stages:
 
@@
 \item Validate (or, if mining, determine) uncles;
+\item validate (or, if mining, determine) transactions;
 \item apply rewards;
@@
 
+\subsection{Transaction Validation}
+
+The given \textbf{gasUsed} and \textbf{minGasLimit} must correspond faithfully to the transactions listed.
+In the case of ${B_H}_m$, the minimum gas limit, all transactions included in the transaction manifest must have a gas price, ${T_0}_p$, that is at least this value:
+\begin{equation}
+\forall_{T \in B_\mathbf{T}} {T_0}_p \geq {B_H}_m
+\end{equation}
+
+In the case of ${B_H}_u$, the total gas used in the block, it must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_u = \ell(\mathbf{T})_1
+\end{equation}
+
 \subsection{Reward Application}
@@
 \boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + R_b \\
-\forall U \in B_\mathbf{U}: \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{7}{8}R_b
+\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{7}{8}R_b
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined:
-\begin{eqnarray}
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...) \\
-B & \equiv & (H, (T_0, T_1, ...), \mathbf{U})
-\end{eqnarray}
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+We begin the definition by supposing we have a series of transactions $\mathbf{T}$.
+
+\begin{equation}
+\text{Let} \quad \mathbf{T} = (\mathbf{T}[0], \mathbf{T}[1], ...)
+\end{equation}
+
+These are the transactions that comprise the block, thus:
+\begin{eqnarray}
+\quad \quad B & \equiv & \Big(H, \big( \mathbf{M}[0], \mathbf{M}[1], ..., \mathbf{M}[\lVert \mathbf{T}\rVert - 1]\big), \mathbf{U}\Big) \\
+\quad \quad \mathbf{M}[n] & \equiv & (\mathbf{T}[n], \mathbf{s}[n], \mathbf{g}[n])
+\end{eqnarray}
+
+Here, $\mathbf{s}$ and $\mathbf{g}$ are the corresponding sequences of states and cumulative gas used for each transaction.
+$\mathbf{s}$ is defined simply as the state resulting from applying the corrresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{s}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{s}[n - 1], \mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluatng the corresponding transaction summed with the previous item (or zero, if it's the first), giving us a running total:
+\begin{equation}
+\mathbf{g}[n] = \begin{cases} 0 & \text{if} \quad n < 0 \\ \Upsilon^g(\mathbf{s}[n - 1], \mathbf{T}[n]) + \mathbf{g}[n-1] & \text{otherwise} \end{cases}
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{T})_1$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{T})_1)
+\end{equation}
 
@@
 $G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
+$G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
 $G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
@@
 $G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
-$G_{create}$ & 100 & Paid for a {\small CREATE} operation and for a contract creation transaction. \\
-$G_{call}$ & 20 & Paid for a {\small CALL} operation and for a message call transaction. \\
+$G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
+$G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
 $G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdata}$ & 5 & Burnt (discarded from the system) for every byte of data or code for a transaction. \\
+$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
+$G_{transaction}$ & 500 & Paid for every transaction. \\
 \bottomrule
@@
 G_{stop} & \text{if} \quad w = \text{\small STOP}\\
+G_{suicide} & \text{if} \quad w = \text{\small SUICIDE}\\
 G_{step} & \text{otherwise}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xf0 & {\small CREATE} & 5 & 1 & Create a new contract. \\
-&&&& $\mathbf{b} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+0xf0 & {\small CREATE} & 3 & 1 & Create a new contract. \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
 &&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
@@
 &&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created contract, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
 \midrule
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 2^{22}, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent and state root hashes, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); $()$ refers to the extradata and the sequences of both uncles and transactions, all empty.
+\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, 0_{256}, 2^{22}, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent and state and transaction root hashes, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); $()$ refers to the extradata and the sequences of both uncles and transactions, all empty.
 $\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.

### 2014-05-03 Paper.tex
PoC-5 cleanups.

@@
 
-When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
 On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
@@
 \begin{equation}
-p(a) \equiv \begin{cases}
-\big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b) \big) \big) & \text{if} \; \boldsymbol{\sigma}[a]_s = \varnothing\\
-\big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big) & \text{otherwise}
-\end{cases}
-\end{equation}
+p(a) \equiv  \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+\end{equation}
+
 This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
@@
 \end{equation}
-
-Where $v$ is the account validity function:
+where $v$ is the account validity function:
 \begin{eqnarray}
@@
 
-A transaction (formally, $T$) is a single cryptographically signed instruction sent by an actor external to Ethereum.
+A transaction (formally, $T$) is a single cryptographically-signed instruction sent by an actor external to Ethereum.
 An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
@@
 
-Whereas a message call transaction contains:
+However, a message call transaction contains:
 
@@
 
-Appendix \ref{app:signing} specifies the function mapping transactions to the sender $S$, which happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
+Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
 For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
@@
 T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} & \wedge \\
-T_t \in \mathbb{B}_{20} &&&&\\
+T_t \in \mathbb{B}_{20} &&&&
 \end{array}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{R}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
 The block header contains several pieces of information:
@@
 
-The other two components in the block are simply a list of uncle block headers (of the same format as above) and a list of the transaction information (as defined previously).
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
+The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transaction \textit{receipts}. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{R}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Manifest}
+
+The transaction receipt is a tuple of three items comprising the transaction, $R_T$, together with the post-transaction state, $R_{\boldsymbol{\sigma}}$, and the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$:
+\begin{equation}
+R \equiv (R_T, R_{\boldsymbol{\sigma}}, R_u)
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (L_T(R_T), \mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u)
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the second item.
+
+We assert $R_u$, the cumulative gas used is a positive integer:
+\begin{equation}
+R_u \in \mathbb{P}
+\end{equation}
+
+For convenience, I also name $B_\mathbf{T}$ the series of transactions encoded by the transaction receipts $B_\mathbf{R}$:
+\begin{equation}
+B_\mathbf{T} \equiv (B_\mathbf{R}[0]_T, B_\mathbf{R}[1]_T, ...)
+\end{equation}
+
+Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+\subsubsection{Holistic Validity}
 
@@
 H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert: (i, B_\mathbf{T}[i])\}) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert: (i, B_\mathbf{R}[i])\}) & \wedge \\
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
@@
 
-where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-% TODO: formal.
-
 Furthermore:
-
 \begin{equation}
@@
 
-TODO: Transaction manifest more than just transactions: $(T[i], s[i], g[i])$
-
 TODO: Remove extrameous references to number.
 
+\subsubsection{Serialisation}
+
 The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction preparation function $L_T$, we assert the types and order of the structure for when the RLP transformation is required:
+Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
 \quad L_H(H) & \equiv & (H_p, H_u, H_b, H_r, H_t, H_d, H_s, H_x, H_n) \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
 \end{eqnarray}
@@
 g_0 & \leq & T_g \quad \wedge \\ 
-v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b
+v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b \\
+T_l \leq {B_H}_l - \ell(B_\mathbf{T})_2
 \end{array}
@@
 
-The sum of the transaction gas limit and the gas utilised in this block prior must be no greater than the block's \textbf{gasLimit}, $H_l$.
-
-% TODO: specify formally.
+Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{T})_2$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$:
 
@@
 The gas remaining should be zero.
-If the creation was conducted as the manifestation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+If the creation was conducted as the receiptation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
 
@@
 
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
 The given \textbf{gasUsed} and \textbf{minGasLimit} must correspond faithfully to the transactions listed.
-In the case of ${B_H}_m$, the minimum gas limit, all transactions included in the transaction manifest must have a gas price, ${T_0}_p$, that is at least this value:
-\begin{equation}
-\forall_{T \in B_\mathbf{T}} {T_0}_p \geq {B_H}_m
+In the case of ${B_H}_m$, the minimum gas limit, all transactions included in the transaction receipt must have a gas price, ${T_0}_p$, that is at least this value:
+\begin{equation}
+\forall_{T \in B_\mathbf{T}} T_p \geq {B_H}_m
 \end{equation}
@@
 \begin{equation}
-{B_H}_u = \ell(\mathbf{T})_1
+{B_H}_u = \ell(\mathbf{R})_u
 \end{equation}
@@
 As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-We begin the definition by supposing we have a series of transactions $\mathbf{T}$.
-
-\begin{equation}
-\text{Let} \quad \mathbf{T} = (\mathbf{T}[0], \mathbf{T}[1], ...)
-\end{equation}
-
-These are the transactions that comprise the block, thus:
-\begin{eqnarray}
-\quad \quad B & \equiv & \Big(H, \big( \mathbf{M}[0], \mathbf{M}[1], ..., \mathbf{M}[\lVert \mathbf{T}\rVert - 1]\big), \mathbf{U}\Big) \\
-\quad \quad \mathbf{M}[n] & \equiv & (\mathbf{T}[n], \mathbf{s}[n], \mathbf{g}[n])
-\end{eqnarray}
-
-Here, $\mathbf{s}$ and $\mathbf{g}$ are the corresponding sequences of states and cumulative gas used for each transaction.
-$\mathbf{s}$ is defined simply as the state resulting from applying the corrresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\mathbf{s}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{s}[n - 1], \mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluatng the corresponding transaction summed with the previous item (or zero, if it's the first), giving us a running total:
-\begin{equation}
-\mathbf{g}[n] = \begin{cases} 0 & \text{if} \quad n < 0 \\ \Upsilon^g(\mathbf{s}[n - 1], \mathbf{T}[n]) + \mathbf{g}[n-1] & \text{otherwise} \end{cases}
+
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states and cumulative gas used after each transaction.
+The former is defined simply as the state resulting from applying the corrresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+B_\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+B_\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\begin{array}[b]{l}
+\Upsilon^g(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + B_\mathbf{R}[n-1]_u
+\end{array}
+ & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{T})_1)
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{T})_{\boldsymbol{\sigma}})
 \end{equation}
@@
 \section{Future Directions} \label{ch:future}
-
-Rather than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.
 
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{call} & \text{if} \quad w = \text{\small CALL}\\
+G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL}\\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& Let $g_k = \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0 \\ \boldsymbol{\mu}_g - \min(\boldsymbol{\mu}_g, \boldsymbol{\mu}_\mathbf{s}[2]) & \text{otherwise} \end{cases}$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[0], g, I_p, \boldsymbol{\mu}_g - g_k, \boldsymbol{\mu}_\mathbf{s}[1], \mathbf{i})$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv g' + g_k$ \\
+&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i})$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 
+\section{Javascript API}\label{app:jsapi}
+
+Yet TODO.
 \end{document}

### 2014-05-03 Paper.tex
Removed references to to number.

@@
 
-TODO: Remove extrameous references to number.
-
 \subsubsection{Serialisation}
@@
 $I_\mathbf{d}$ evaluates to the empty tuple.
-$I_H$ and $I_n$ have no special treatment and are determined from the blockchain.
+$I_H$ has no special treatment and are determined from the blockchain.
 The exception in the determination of $\boldsymbol{\sigma}_P$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created contract's account, with $\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
@@
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the previous block.
-\item $I_n$, the block number of the current block.
+\item $I_H$, the block header of the present block.
 \end{itemize}
@@
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv n({I_H})$ \\
-&&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
@@
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 10^6$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 &&&& Presently, the gas limit is hard-coded in the protocol. \\

### 2014-05-03 Paper.tex
Start o JS API

@@
 
-Yet TODO.
+To make an \DH{}App work with on Ethereum, you'll need to know about the Ethereum Javascript bindings, or, if you like, magic Javascript objects.
+
+There is currently only one such object; the \texttt{eth} object.
+
+Parameters
+
+Parameters are, in general, strings which are interpreted according to the required type.
+When converting between integers and hashes/byte-arrays, it is interpreted as a big-endian as is standard for Ethereum.
+The following forms are allowed; they are all interpreted in the same way:
+%
+%"ABC"
+%"0x414243"
+%"4276803"
+%In each case, they are interpreted as the number 4276803. The first two values may be alternated between with bin() and unbin().
+%
+%Such byte-arrays may be concatenated with the + operator as is normal for strings.
+%
+%Strings also have a number of additional methods to help with conversion and alignment when switching between addresses, 256-bit integers and free-form byte-arrays for transaction data:
+%
+%bin(): Converts the string to binary format.
+%pad(a, b): Converts the string to binary format (ready for data parameters) and pads with zeroes on the left size until it is of width a.
+Then pads with zeroes on the right side until it has grown to size b.
+If b is less that the width of the string, it is resized accordingly.
+%pad(l): Converts the string to binary format (ready for data parameters) and pads with zeroes until it is of width l.
+Will pad to the left if the original string is numeric, or to the right if binary.
+If l is less than the width of the string, it is resized accordingly.
+%unbin(): Converts the string from binary format to hex format.
+%unpad(): Converts the string from binary format to hex format, first removing any zeroes from the right side.
+%eth
+%
+%Properties: For each such item, there is also an asynchronous method, taking a parameter of the callback function, itself taking a single parameter of the property's return value and of the same name but prefixed with get and recapitalised, e.g.
+getCoinbase(_fn).
+%
+%coinbase Returns the coinbase address of the client.
+%isListening Returns true if and only if the client is actively listening for network connections.
+%isMining Returns true if and only if the client is actively mining new blocks.
+%gasPrice Returns the special 256-bit number equal to the hard-coded testnet price of gas.
+%key Returns the special key-pair object corresponding to the preferred account owned by the client.
+%keys Returns a list of the special key-pair objects corresponding to all accounts owned by the client.
+%peerCount Returns the number of peers currently connected to the client.
+%Synchronous Getters: For each such item, there is also an asynchronous method, taking an additional parameter of the callback function, itself taking a single parameter of the synchronous method's return value and of the same name but prefixed with get and recapitalised, e.g.
+getBalanceAt(_a, _fn).
+%
+%balanceAt(_a) Returns the balance of the account of address given by the address _a.
+%storageAt(_a, _x) Returns the value in storage at position given by the number _x of the account of address given by the address _a.
+%txCountAt(_a) Returns the number of transactions send from the account of address given by _a.
+%isContractAt(_a) Returns true if the account of address given by _a is a contract-account.
+%Transactions
+%
+%create(_sec, _xEndowment, _bCode, _xGas, _xGasPrice, _fn) Creates a new contract-creation transaction, given parameters:
+%_sec, the secret-key for the sender;
+%_xEndowment, the number equal to the contract's endowment;
+%_bCode, the byte-array EVM-bytecode for the initialisation of the contract;
+%_xGas, the number equal to the amount of gas to purchase for the transaction (unused gas is refunded); and
+%_xGasPrice, the number equal to the price of gas for this transaction. Returns the special address object representing the new contract's account.
+%_fn, the callback function, called on completion of the transaction.
+%transact(_sec, _xValue, _aDest, _bData, _xGas, _xGasPrice, _fn) Creates a new message-call transaction, given parameters:
+%_sec, the secret-key for the sender;
+%_xValue, the value transferred for the transaction (in Wei);
+%_aDest, the address representing the destination address of the message;
+%_bData, the byte array containing the associated data of the message;
+%_xGas, the amount of gas to purchase for the transaction (unused gas is refunded); and
+%_xGasPrice, the price of gas for this transaction.
+%_fn, the callback function, called on completion of the transaction.
+%Events
+%
+%watch(_a, _fn): Registers _fn as a callback for whenever anything about the given address's state changes, and also on the initial load.
+%watch(_a, _x, _fn): Registers _fn as a callback for whenever the given storage location of the given address changes, and also on the initial load.
+%newBlock(_fn): Registers _fn as a callback for whenever the state changes, and also on the initial load.
+%Misc
+%
+%secretToAddress(_a): Determines the address from the secret key _a.
+
 \end{document}

### 2014-05-03 Paper.tex
JS API.

--- 2014-05-03 Paper.tex
@@
 
-Yet TODO.
+The JavaScript API provides a consistent API across multiple scenarios including each of the clients' web-based in-process \DH{}App frameworks and the out-of-process RPC-based infrastructure.
+All key access takes places though the special \texttt{eth} object, part of the global namespace.
+
+\subsection{Values}
+There are no special object types in the API; all values are strings. As strings, values may be of several forms, and are interpreted by the API according to a series of rules:
+
+\begin{enumerate}
+\item If the string contains only digits from 0-9, then it is interpreted as a decimal integer;
+\item if the string begins with the characters \texttt{0x}, then it is interpreted as a hexadecimal integer;
+\item it is interpreted as a binary string otherwise.
+\end{enumerate}
+
+The only exception to this are for parameters that expect a binary string; in this case the string is always interpreted as such.
+
+Values are implicitly converted between integers and hashes/byte-arrays; when this happens, integers are interpreted as big-endian as is standard for Ethereum.
+The following forms are allowed; they are all interpreted in the same way:
+
+\begin{enumerate}
+\item \texttt{"4276803"}
+\item \texttt{"0x414243"}
+\item \texttt{"ABC"}
+\end{enumerate}
+
+In each case, they are interpreted as the number 4276803. The first two values may be alternated between with the additional String methods \texttt{bin()} and \texttt{unbin()}.
+
+As byte arrays, values may be concatenated with the \texttt{+} operator as is normal for strings.
+
+Strings also have a number of additional methods to help with conversion and alignment when switching between addresses, 256-bit integers and free-form byte-arrays for transaction data:
+
+\begin{itemize}
+\item \texttt{bin()}: Converts the string to binary format.
+\item \texttt{pad(l)}: Converts the string to binary format (ready for data parameters) and pads with zeroes until it is of width \texttt{l}.
+Will pad to the left if the original string is numeric, or to the right if binary.
+If \texttt{l} is less than the width of the string, it is resized accordingly.
+\item \texttt{pad(a, b)}: Converts the string to binary format (ready for data parameters) and pads with zeroes on the left size until it is of width \texttt{a}.
+Then pads with zeroes on the right side until it has grown to size \texttt{b}.
+If \texttt{b} is less hat the width of the string, it is resized accordingly.
+\item \texttt{unbin()}: Converts the string from binary format to hex format.
+\item \texttt{unpad()}: Converts the string from binary format to hex format, first removing any zeroes from the right side.
+\end{itemize}
+
+\subsection{The \texttt{eth} object}
+
+\subsubsection{Properties}
+
+For each such item, there is also an asynchronous method, taking a parameter of the callback function, itself taking a single parameter of the property's return value and of the same name but prefixed with get and recapitalised, e.g.
+\texttt{getCoinbase(fn)}.
+
+\begin{itemize}
+\item \texttt{coinbase} Returns the coinbase address of the client.
+\item \texttt{isListening} Returns true if and only if the client is actively listening for network connections.
+\item \texttt{isMining} Returns true if and only if the client is actively mining new blocks.
+\item \texttt{gasPrice} Returns the client's present price of gas.
+\item \texttt{key} Returns the special key-pair object corresponding to the preferred account owned by the client.
+\item \texttt{keys} Returns a list of the special key-pair objects corresponding to all accounts owned by the client.
+\item \texttt{peerCount} Returns the number of peers currently connected to the client.
+\end{itemize}
+
+\subsubsection{Synchronous Getters}
+For each such item, there is also an asynchronous method, taking an additional parameter of the callback function, itself taking a single parameter of the synchronous method's return value and of the same name but prefixed with get and recapitalised, e.g.
+\texttt{getBalanceAt(a, fn)}.
+
+\begin{itemize}
+\item \texttt{balanceAt(a)} Returns the balance of the account of address given by the address \texttt{a}.
+\item \texttt{storageAt(a, x)} Returns the value in storage at position given by the number x of the account of address given by the address \texttt{a}.
+\item \texttt{txCountAt(a)} Returns the number of transactions send from the account of address given by \texttt{a}.
+\item \texttt{isContractAt(a)} Returns true if the account of address given by \texttt{a} has associated code.
+\end{itemize}
+
+\subsubsection{Transactions}
+
+\begin{itemize}
+\item \texttt{create(sec, xEndowment, bCode, xGas, xGasPrice, fn)} Creates a new contract-creation transaction, given parameters:
+\begin{itemize}
+\item \texttt{sec}, the secret-key for the sender;
+\item \texttt{xEndowment}, the number equal to the contract's endowment;
+\item \texttt{bCode}, the binary string (byte array) of EVM-bytecode for the initialisation of the contract;
+\item \texttt{xGas}, the number equal to the amount of gas to purchase for the transaction (unused gas is refunded);
+\item \texttt{xGasPrice}, the number equal to the price of gas for this transaction. Returns the special address object representing the new contract's account; and
+\item \texttt{fn}, the callback function, called on completion of the transaction.
+\end{itemize}
+\item \texttt{transact(sec, xValue, aDest, bData, xGas, xGasPrice, fn)} Creates a new message-call transaction, given parameters:
+\begin{itemize}
+\item \texttt{sec}, the secret-key for the sender;
+\item \texttt{xValue}, the value transferred for the transaction (in Wei);
+\item \texttt{aDest}, the address representing the destination address of the message;
+\item \texttt{bData}, the binary string (byte array), containing the associated data of the message;
+\item \texttt{xGas}, the amount of gas to purchase for the transaction (unused gas is refunded);
+\item \texttt{xGasPrice}, the price of gas for this transaction; and
+\item \texttt{fn}, the callback function, called on completion of the transaction.
+\end{itemize}
+\end{itemize}
+
+\subsubsection{Events}
+
+\begin{itemize}
+\item \texttt{watch(a, fn)}: Registers \texttt{fn} as a callback for whenever anything about the state of the account at address \texttt{a} changes, and also on the initial load.
+\item \texttt{watch(a, x, fn)}: Registers \texttt{fn} as a callback for whenever the storage location \texttt{x} of the account at address \texttt{a} changes, and also on the initial load.
+\item \texttt{newBlock(fn)}: Registers \texttt{fn} as a callback for whenever the state changes, and also on the initial load.
+\end{itemize}
+
+\subsubsection{Misc}
+
+\begin{itemize}
+\item \texttt{secretToAddress(a)}: Determines the address from the secret key \texttt{a}.
+\end{itemize}
+
 \end{document}

### 2014-05-06 Paper.tex
Add dec() API.
Receipt/Transaction confusion.

@@
 g_0 & \leq & T_g \quad \wedge \\ 
-v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b \\
-T_l \leq {B_H}_l - \ell(B_\mathbf{T})_2
+v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_l & \leq & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{T})_2$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$:
+Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$:
 
@@
 \begin{equation}
-\forall_{T \in B_\mathbf{T}} T_p \geq {B_H}_m
+\forall \; T \in B_\mathbf{T}: T_p \geq {B_H}_m
 \end{equation}
@@
 
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{T})_1$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{T})_{\boldsymbol{\sigma}})
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}
@@
 \item \texttt{unpad()}: Converts the string from binary format to hex format, first removing any zeroes from the right side.
+\item \texttt{dec()}: Converts the string to decimal format (typically from hex).
 \end{itemize}

### 2014-05-06 Paper.tex
Minor PoC-5 protocol changes.

@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (H_p, H_u, H_b, H_r, H_t, H_d, H_s, H_x, H_n) \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d, H_s,\\ H_i, H_m, H_l, H_u, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n)] & \equiv & \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
-\end{eqnarray}
-
-It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
+\boldsymbol{\sigma}^*[a] & \equiv & \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
+\end{eqnarray}
+
+where $a$ is the address of the new account, as defined above. It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
 
@@
 \quad \boldsymbol{\sigma}_P & \equiv & \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad \mathbf{o} = () \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad \boldsymbol{\sigma}^{**}[a] = \varnothing \\
 \begin{array}[b]{l}
 \boldsymbol{\sigma}^{**} \quad \text{except:} \\
-\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}} = \mathbf{o}
+\boldsymbol{\sigma}_P[a]_{\mathbf{b}} = \mathbf{o}
 \end{array} & \text{otherwise}
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SUICIDE} instruction or normal {\small STOP} code, or if the code returned is otherwise empty, then the state changes are effectively reverted, the final state being equivalent to $\boldsymbol{\sigma}_0$, the pre-execution checkpoint state.
+If the initialisation execution ends with a {\small SUICIDE} instruction, no code is set.
+For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
 
@@
 The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in memory or storage, it is stored separately in a virtual ROM.
-It cannot be read directly; instead it exists only as a model for determining the next instruction to execute.
+Rather than storing program code in generally-accessible memory or storage, it is stored in a separately in a virtual ROM interactable only though a specialised instruction.
 

### 2014-05-06 Paper.tex
Signed comparisons.

@@
 \midrule
-0x07 & {\small SMOD} & 2 & 1 & Modulo remainder operation. \\
+0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x0c & {\small EQ} & 2 & 1 & Equality comparision. \\
+0x0c & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+\midrule
+0x0d & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+\midrule
+0x0e & {\small EQ} & 2 & 1 & Equality comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x0d & {\small NOT} & 1 & 1 & Simple not operator. \\
+0x0f & {\small NOT} & 1 & 1 & Simple not operator. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\

### 2014-05-12 Paper.tex
PoC-5 Genesis block updates.

@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
-&&&& Presently, the gas limit is hard-coded in the protocol. \\
 \bottomrule
@@
 
-The genesis block is 9 items, and is specified thus:
-\begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, 0_{256}, 0_{256}, 2^{22}, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
+The genesis block is 13 items, and is specified thus:
+\begin{equation}
+\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0_{256}, 2^{22}, 0, 0, 1000000, 0, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
 \end{equation}
@@
 
-The proof-of-concept series include a development premine, making the state root hash some value other than $0_{256}$. The latest documentation should be consulted for the value of the state root.
+The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$. The latest documentation should be consulted for the value of the state root.
 

### 2014-05-12 Paper.tex
Fixes #182.

@@
 \midrule
-0x37 & {\small CALLDATACOPY} & 0 & 3 & Copy input data in current environment to memory. \\
+0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
@@
 \midrule
-0x39 & {\small CODECOPY} & 0 & 3 & Copy code running in current environment to memory. \\
+0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv

### 2014-05-15 Paper.tex
Move timestamp to correct position.
Minor correction & clarification.
Update transaction component ordering.
Remove old reference to body.

@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_v, T_t, T_p, T_g, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
-(T_n, T_v, T_t, T_p, T_g, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
 \end{cases}
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{b}$, $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
 \begin{equation}
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d, H_s,\\ H_i, H_m, H_l, H_u, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d,\\ H_i, H_m, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
@@
 $\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ values refer to the hashes of the transaction and uncle lists in RLP, both empty.
+$\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
 

### 2014-05-19 Paper.tex
Clarify minor point on hash collisions. Fixed #7.

@@
 \subsection{Trie Database}
-Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision.
+Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the SHA3 hash's collision resistance.
 In reality, a sensible implementation will not fully recompute the trie root hash for each set.

### 2014-05-19 Paper.tex
Fixed #6.

@@
 
-It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{E}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{E} & \equiv & \begin{cases}
+It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
+
+\begin{eqnarray}
+\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
 (16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &

### 2014-05-20 Paper.tex
manifest -> receipt
Min gas limit of 10000.
Use canon order in non-formal text.
Rid ourselves of 32-zero trie roots.

@@
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
@@
 \item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_u$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
 \item[extraData] An arbitrary byte array containing data relevant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
@@
 
-\subsubsection{Transaction Manifest}
+\subsubsection{Transaction Receipt}
 
@@
 10^6 & \text{if} \quad H_i = 0\\
+10^4 & \text{if} \quad L'(H) < 10^4\\
 L'(H) & \text{otherwise}
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single 32-byte value that identifies a given set of key-value pairs.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
@@
 We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure.
-The empty trie is defined as being a 256-bit hash of zeroes, $0_{256}$:`
+The empty trie is defined as being a the empty byte sequence, $()$:
 \begin{equation}
 \texttt{\small TRIE}(\mathfrak{I}) \equiv \begin{cases}
-0_{256} & \text{if} \quad \mathfrak{I} = \varnothing \\
-n(y(\mathfrak{I}), 0) & \text{otherwise}
+() & \text{if} \quad \mathfrak{I} = \varnothing \\
+\texttt{\small SHA3}(c(\mathfrak{I}, 0)) & \text{otherwise}
 \end{cases}
@@
 
-We assume a function $n$, the trie's node cap function.
+We also assume a function $n$, the trie's node cap function.
 When composing a node, we use RLP to encode the structure.
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0_{256}, 2^{22}, 0, 0, 1000000, 0, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent and state and transaction root hashes, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); $()$ refers to the extradata and the sequences of both uncles and transactions, all empty.
+\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, (), 2^{22}, 0, 0, 1000000, 0, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); $()$ refers to the transaction trie root, extradata and the sequences of both uncles and transactions, all empty.
 $\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.

### 2014-05-20 Paper.tex
Minor fix () is the empty list/sequence, not the empty byte array - replaced with 0 where appropriate.

@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, (), 2^{22}, 0, 0, 1000000, 0, 0, (), \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); $()$ refers to the transaction trie root, extradata and the sequences of both uncles and transactions, all empty.
+\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
+The sequences of both uncles and transactions are empty and represented by $()$.
 $\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.

### 2014-05-21 Paper.tex
byte arrays are not arb length.

@@
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the contract, itself a simple mapping from byte arrays of size 32 to arbitrary length byte arrays.
+\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the contract, itself a simple mapping between byte arrays of size 32.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.

### 2014-05-22 Paper.tex
Correction to block reward.

@@
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence $\texttt{\small TRIE}(\boldsymbol{\sigma}[a]_\mathbf{s}) \equiv \boldsymbol{\sigma}[a]_s$.
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equi	valence $\texttt{\small TRIE}(\boldsymbol{\sigma}[a]_\mathbf{s}) \equiv \boldsymbol{\sigma}[a]_s$.
 It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
@@
 \begin{equation}
-\text{Let} \quad R_b = 1.5 * 10^{16}
+\text{Let} \quad R_b = 1.5 \times 10^{18}
 \end{equation}

### 2014-05-22 Paper.tex
Transaction signing.

@@
 
-The function $L_T$ prepares a transaction for being transformed into an RLP-serialised byte array:
 \begin{equation}
@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_v, T_p, T_g, T_t, T_\mathbf{b}, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_v, T_p, T_g, T_t, T_\mathbf{d}) & \text{otherwise} \end{cases} \\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
+\end{cases} \\
 h(T) & \equiv & \mathtt{\small SHA3}( L_S(T) )

### 2014-05-31 Paper.tex
JUMPI definition minor fix.

@@
 0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
@@
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
+&&&& Thus the operand order is: gas, input offset, input size. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
 \midrule

### 2014-06-09 Paper.tex
Balance fix.
Update Paper.tex
spelling fix

@@
 
-Dealings in this proposed system would have several attributes not often found in the real word.
+Dealings in this proposed system would have several attributes not often found in the real world.
 The incorruptibility of judgement, often difficult to find, comes naturally from a disinterested algorithmic interpreter.
@@
 \midrule
-0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
+0x31 & {\small BALANCE} & 1 & 1 & Get balance of currently executing account. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_b$ \\

### 2014-06-09 Paper.tex
Typo.

--- 2014-05-31 Paper.tex
@@
 The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in generally-accessible memory or storage, it is stored in a separately in a virtual ROM interactable only though a specialised instruction.
+Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only though a specialised instruction.
 

### 2014-06-16 Paper.tex
Minor fix to naming.
Various terminology modifications and cleanups.

@@
 
-Hashcash, introduced by \cite{laurie2004proof}, provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
+Hashcash, introduced by \cite{back2002hashcash} (in a five-year retrospective), provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
 Though not widely adopted, the work was later utilised and expanded upon by \cite{nakamoto2008bitcoin} in order to devise a cryptographically secure mechanism for coming to a decentralised social consensus over the order and contents of a series of cryptographically signed financial transactions.
@@
 Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol.
-The best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
+Some of the best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
 Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address, or, in the case of contract accounts, the number of contract-creations made by this account.
+\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the contract, itself a simple mapping between byte arrays of size 32.
+\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the account, itself a simple mapping between byte arrays of size 32.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
@@
 It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
-\item[codeHash] The hash of the EVM code of this contract---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
+\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
@@
 If the latter field is the SHA3 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small SHA3}\big(()\big)$, then the node represents a simple account, sometimes refered to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_c = \texttt{\small SHA3}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
-There are two types of transactions: those which result in message calls and those which result in the creation of new contracts.
+There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
 Both types specify a number of common fields:
@@
 \item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created contract's account; formally $T_v$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
@@
 This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or the zero address for a contract-creation transaction; formally $T_t$.
+\item[to] The 160-bit address of the message call's recipient or the zero address for a contract creation transaction; formally $T_t$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
@@
 
-Additionally, a contract-creation transaction contains:
+Additionally, a contract creation transaction contains:
 
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the contract initialisation procedure, formally $T_\mathbf{i}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
 \end{description}
@@
 \textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executed each time the account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is executed only once at contract creation and that gets discarded immediately thereafter.
+\textbf{init} is executed only once at account creation and gets discarded immediately thereafter.
 
@@
 The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation).
-Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing contract storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.
+Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing account storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.
 
@@
 Gas does not exist outside of the execution of a transaction.
-Thus for trusted contracts, a relatively high gas limit may be set and left alone.
+Thus for accounts with trusted code associated, a relatively high gas limit may be set and left alone.
 
@@
 g_0 \equiv \begin{cases}
-\lVert T_\mathbf{i}\rVert G_{data} + G_{transaction} & \text{if} \quad T_t = 0 \\
-\lVert T_\mathbf{d}\rVert G_{data} + G_{transaction} & \text{otherwise} \\
+\lVert T_\mathbf{i}\rVert G_{txdata} + G_{transaction} & \text{if} \quad T_t = 0 \\
+\lVert T_\mathbf{d}\rVert G_{txdata} + G_{transaction} & \text{otherwise} \\
 \end{cases}
@@
 
-There are number of intrinsic parameters used when creating a contract: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
+There are number of intrinsic parameters used when creating an account: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
 
@@
 
-The address of the new contract is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure containing only the sender and the nonce.
 In the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
@@
 
-Finally, the contract is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
-Code execution can effect several events that are not internal to the execution state: the contract's storage can be altered, further contracts can be created and further message calls can be made.
+Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
+Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
 As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$ and available gas remaining $g'$.
@@
 $I_H$ has no special treatment and are determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}_P$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created contract's account, with $\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+The exception in the determination of $\boldsymbol{\sigma}_P$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
 
@@
 
-The contract's body code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
@@
 
-Over a contract's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
+Over a account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
 This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
@@
 
-Storage fees have a slightly nuanced behaviour---to incentivise contracts to minimise storage use (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is waived; in fact, it is effectively paid up-front since the initial usage of a storage location costs twice as much as the normal usage.
+Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is waived; in fact, it is effectively paid up-front since the initial usage of a storage location costs twice as much as the normal usage.
 
@@
 \end{itemize}
-\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of contract body code.
-\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of contract initialisation code.
+\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
+\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
 \item {\small STOP} costs $G_{stop}$ gas
@@
 For the release series, we use a more complex proof-of-work.
-This has yet to be formally defined, but involves two components; firstly that it concerns contract processing, and by extension the evaluation of programs on the EVM.
+This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
 Secondly that it concerns the utilisation of either the blockchain or the full state trie.
@@
 \item[Account] Accounts have an intrinsic balance and transaction count maintained as part of the Ethereum state.
-They are owned either by External Actors or intrinsically (as an identity) an Autonomous Object within Ethereum.
-If an Account identifies an Autonomous Object, then Ethereum will also maintain a Storage State particular to that Account.
+They also have some (possibly empty) EVM Code and a (possibly empty) Storage State associated with them.
+Though homogenous, it makes sense to distinguish between two practical types of account: those with empty associated EVM Code (thus the account balance is controlled, if at all, by some external entity) and those with non-empty associated EVM Code (thus the account represents an Autonomous Object).
 Each Account has a single Address that identifies it.
@@
 
-\item[Autonomous Object] A virtual object existent only within the hypothetical state of Ethereum.  Has an intrinsic address. Incorporated only as the state of the storage component of the VM.
-
-\item[Storage State] The information particular to a given Autonomous Object that is maintained between the times that it runs.
-
-\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts in a perfectly trusted way, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.
+\item[Autonomous Object] A notional object existent only within the hypothetical state of Ethereum.
+Has an intrinsic address and thus an associated account; the account will have non-empty associated EVM Code.
+Incorporated only as the Storage State of that account.
+
+\item[Storage State] The information particular to a given Account that is maintained between the times that the Account's associated EVM Code runs.
+
+\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.
 
 \item[Message Call] The act of passing a message from one Account to another.
-If the destination account is an Autonomous Object, then the VM will be started with the state of said Object and the Message acted upon.
+If the destination account is associated with non-empty EVM Code, then the VM will be started with the state of said Object and the Message acted upon.
 If the message sender is an Autonomous Object, then the Call passes any data returned from the VM operation.
@@
 
-\item[Contract] Synonym for Autonomous Object used for non-technical audiences.
+\item[Contract] Informal term used to mean both a piece of EVM Code that may be associated with an Account or an Autonomous Object.
 
@@
 
-\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for a contract's program code.
-
-\item[EVM Code] The bytecode that the EVM can natively execute.
+\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for an Account's associated EVM Code.
+
+\item[EVM Code] The bytecode that the EVM can natively execute. Used to formally specify the meaning and ramifications of a message to an Account.
 
@@
 
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts.
+\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-piling to.
 
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never a contract. \\
+&&&& This is the sender of original transaction; it is never an account with non-empty associated code. \\
 \midrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xf0 & {\small CREATE} & 3 & 1 & Create a new contract. \\
+0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
-&&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created contract, otherwise. \\
+&&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
@@
 \midrule
-0xf1 & {\small CALL} & 7 & 1 & Message-call into contract. \\
+0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
@@
 \item \texttt{sec}, the secret-key for the sender;
-\item \texttt{xEndowment}, the number equal to the contract's endowment;
-\item \texttt{bCode}, the binary string (byte array) of EVM-bytecode for the initialisation of the contract;
+\item \texttt{xEndowment}, the number equal to the account's endowment;
+\item \texttt{bCode}, the binary string (byte array) of EVM-bytecode for the initialisation of the account;
 \item \texttt{xGas}, the number equal to the amount of gas to purchase for the transaction (unused gas is refunded);
-\item \texttt{xGasPrice}, the number equal to the price of gas for this transaction. Returns the special address object representing the new contract's account; and
+\item \texttt{xGasPrice}, the number equal to the price of gas for this transaction. Returns the special address object representing the new account; and
 \item \texttt{fn}, the callback function, called on completion of the transaction.

### 2014-07-02 Paper.tex
Update Paper.tex
VM spec fix.

@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b})$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, \boldsymbol{\mu}_g) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
-&&&& $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created account, otherwise. \\
+&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$; $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
-&&&& Thus the operand order is: gas, input offset, input size. \\
+&&&& Thus the operand order is: value, input offset, input size. \\
 \midrule
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i})$ \\
+&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas;\\
-&&&& $x=1$ otherwise. \\
+&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$; $x=1$ otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\

### 2014-07-17 Paper.tex
Balance fixed.

@@
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equi	valence $\texttt{\small TRIE}(\boldsymbol{\sigma}[a]_\mathbf{s}) \equiv \boldsymbol{\sigma}[a]_s$.
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence $\texttt{\small TRIE}(\boldsymbol{\sigma}[a]_\mathbf{s}) \equiv \boldsymbol{\sigma}[a]_s$.
 It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
@@
 
-\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executed each time the account receives a message call (either through a transaction or due to the internal execution of code).
+\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executes each time the account receives a message call (either through a transaction or due to the internal execution of code).
 \textbf{init} is executed only once at account creation and gets discarded immediately thereafter.
@@
 The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only though a specialised instruction.
+Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only through a specialised instruction.
 
@@
 \midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get balance of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_b$ \\
+0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{when} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule

### 2014-07-21 Paper.tex
Updated gas limit.

--- 2014-07-02 Paper.tex
@@
 10^6 & \text{if} \quad H_i = 0\\
-10^4 & \text{if} \quad L'(H) < 10^4\\
+125000 & \text{if} \quad L'(H) < 125000\\
 L'(H) & \text{otherwise}
@@
 
+\item[Ethereum Runtime Environment] (aka ERE) The environment which is provided to an Autonomous Object executing in the EVM.
+Includes the EVM but also the structure of the world state on which the EVM relies for certain I/O instructions including CALL \& CREATE.
+
 \item[EVM Code] The bytecode that the EVM can natively execute. Used to formally specify the meaning and ramifications of a message to an Account.

### 2014-07-21 Paper.tex
Explicit definition of RLP in transaction trie.

--- 2014-07-17 Paper.tex
@@
 \item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
@@
 \item[to] The 160-bit address of the message call's recipient or the zero address for a contract creation transaction; formally $T_t$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
@@
 H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert: (i, B_\mathbf{R}[i])\}) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i])\})) & \wedge \\
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
 \end{array}
+\end{equation}
+
+where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
+
+\begin{equation}
+P(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
 \end{equation}
@@
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
-H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge & H_b \in \mathbb{B}_{32} &&
+H_i \in \mathbb{P} & \wedge & H_m \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
+H_u \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge \\
+H_n \in \mathbb{B}_{32} &&&&
 \end{array}
@@
 
-$H_s$ is the timestamp of block $H$ and must fulfill the relation:
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
 \begin{equation}
@@
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because satisfaction of this requirement depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
 
@@
 
-The given \textbf{gasUsed} and \textbf{minGasLimit} must correspond faithfully to the transactions listed.
-In the case of ${B_H}_m$, the minimum gas limit, all transactions included in the transaction receipt must have a gas price, ${T_0}_p$, that is at least this value:
+The given \textbf{gasUsed} and \textbf{minGasPrice} must correspond faithfully to the transactions listed.
+In the case of ${B_H}_m$, the minimum gas price, all transactions included in the transaction receipt must have a gas price, ${T_0}_p$, that is at least this value:
 \begin{equation}
@@
 \item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for an Account's associated EVM Code.
+
+\item[Ethereum Runtime Environment] (aka ERE) The environment which is provided to an Autonomous Object executing in the EVM.
+Includes the EVM but also the structure of the world state on which the EVM relies for certain I/O instructions including CALL \& CREATE.
 

### 2014-07-21 Paper.tex
Formalise stateTrie mapping.

@@
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the account, itself a simple mapping between byte arrays of size 32.
+\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from a byte array of size 32 to an RLP-encoded integer.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence $\texttt{\small TRIE}(\boldsymbol{\sigma}[a]_\mathbf{s}) \equiv \boldsymbol{\sigma}[a]_s$.
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
@@
 
-If the latter field is the SHA3 hash of the empty string, i.e.
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
+\begin{equation}
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\begin{equation}
+L_I\big( (k, v) \big) \equiv \big(k, \texttt{\small RLP}(v)\big)
+\end{equation}
+
+where:
+\begin{equation}
+k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
+\end{equation}
+
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+
+If the \textbf{codeHash} field is the SHA3 hash of the empty string, i.e.
 $\boldsymbol{\sigma}[a]_c = \texttt{\small SHA3}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
@@
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
-
 \begin{equation}

### 2014-08-03 Paper.tex
Age-old bugs in the transaction algebra.

@@
     Dr. Gavin Wood\\
-    Co-Founder \& CTO, Ethereum Project\\
+    Co-Founder \& Lead, Ethereum Project\\
     gavin@ethereum.org
@@
 (\boldsymbol{\sigma}_P, g') \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_t = 0 \\
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}) & \text{if} \quad T_t = 0 \\
 \Theta_{0,1}(\boldsymbol{\sigma}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
@@
 
-There are number of intrinsic parameters used when creating an account: sender ($s$), nonce ($n$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}_0$ to the tuple containing the new state, additional database entries and remaining gas $(\boldsymbol{\sigma}_P, g')$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g') \equiv \Lambda(\boldsymbol{\sigma}_0, s, n, g, p, v, \mathbf{i})
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, additional database entries and remaining gas $(\boldsymbol{\sigma}', g')$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g') \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i})
 \end{equation}
@@
 A(s, n) & \equiv & a \quad \text{where:} \\
-a & = & \arg \min_x : x \geqslant a' \wedge \boldsymbol{\sigma}_0[x] = \varnothing \\
+a & = & \arg \min_x : x \geqslant a' \wedge \boldsymbol{\sigma}[x] = \varnothing \\
 a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
@@
 
-where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}_0[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the empty string, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}^*[a] & \equiv & \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
-\end{eqnarray}
-
-where $a$ is the address of the new account, as defined above. It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
+\begin{equation}
+\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except:}
+\end{equation}
+\begin{equation}
+\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
+\end{equation}
+
+where $a$ is the address of the new account, as defined above.
+% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
 
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$ and available gas remaining $g'$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$ and the body code of the account $\mathbf{b}$.
 
@@
 The gas remaining should be zero.
-If the creation was conducted as the receiptation of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
 
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state and gas as $(\boldsymbol{\sigma}_P, g')$ where:
+Thus formally, we may specify the resultant state and gas as $(\boldsymbol{\sigma}', g')$ where:
 \begin{equation}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\sigma}_P & \equiv & \begin{cases}
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
 \boldsymbol{\sigma}^{**} & \text{if} \quad \boldsymbol{\sigma}^{**}[a] = \varnothing \\
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} \\
-\boldsymbol{\sigma}_P[a]_{\mathbf{b}} = \mathbf{o}
+\boldsymbol{\sigma}'[a]_{\mathbf{b}} = \mathbf{o}
 \end{array} & \text{otherwise}
@@
 I_a & \equiv & a \\
-I_o & \equiv & s \\
+I_o & \equiv & o \\
 I_p & \equiv & p \\
@@
 $I_H$ has no special treatment and are determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}_P$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}_P[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}_P, g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
+(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_P & \equiv & \begin{cases}
+\boldsymbol{\sigma}' & \equiv & \begin{cases}
 \boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}_1[r]_c = \varnothing \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, \mathbf{b}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, \boldsymbol{\mu}_g) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, \boldsymbol{\mu}_g) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\

### 2014-08-03 Paper.tex
Typo.

@@
 As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states and cumulative gas used after each transaction.
-The former is defined simply as the state resulting from applying the corrresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}

### 2014-08-03 Paper.tex
PoC-5 final.

@@
 
-% Explain g_d reason?
+%todo Explain g_d reason?
 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\boldsymbol{\sigma}_P$) and remaining gas ($g'$):
-\begin{equation}
-(\boldsymbol{\sigma}_P, g') \equiv \begin{cases}
+%todo explain suicide list
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and suicide list $\mathbf{s}$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', \mathbf{s}) \equiv \begin{cases}
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}) & \text{if} \quad T_t = 0 \\
-\Theta_{0,1}(\boldsymbol{\sigma}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{0,1}$ to denote the fact that only the first two components of the function's value are taken; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
-So we define the final state $\boldsymbol{\sigma}'$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g' T_p \\
-\boldsymbol{\sigma}'[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g') T_p \\
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g' T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g') T_p \\
 m & \equiv & {B_H}_b
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
+\begin{eqnarray}
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
+\forall i \in \mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g') \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i})
+(\boldsymbol{\sigma}', g', \mathbf{s}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i})
 \end{equation}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$ and the body code of the account $\mathbf{b}$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$, the suicide list $\mathbf{s}$ and the body code of the account $\mathbf{b}$.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state and gas as $(\boldsymbol{\sigma}', g')$ where:
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+Thus formally, we may specify the resultant state, gas and suicide list as $(\boldsymbol{\sigma}', g', \mathbf{s})$ where:
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
 \end{equation}
@@
 $I_\mathbf{d}$ evaluates to the empty tuple.
-$I_H$ has no special treatment and are determined from the blockchain.
+$I_H$ has no special treatment and is determined from the blockchain.
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SUICIDE} instruction, no code is set.
+If the initialisation execution ends with a {\small SUICIDE} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call.
-Aside from evaluating to a new state and additional database entries, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+Aside from evaluating to a new state and suicide list, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
 \begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
 \end{equation}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{o}) & \equiv & \Xi(\boldsymbol{\sigma}_1, g, I) \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \Xi(\boldsymbol{\sigma}_1, g, I) \\
 I_a & \equiv & a \\
@@
 Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds since at the present price of Ether and default gas price, that would cost around US\$20M for the memory fee alone.
+That said, implementations must be able to manage this eventuality.
 
@@
 
-More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
-
-\begin{itemize}
-\item {\small SHA3} costs $G_{sha3}$ gas
-\item {\small SLOAD} costs $G_{sload}$ gas
-\item {\small BALANCE} costs $G_{balance}$ gas
-\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
-\begin{itemize}
-\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
-\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
-\item $d = 1$ otherwise.
-\end{itemize}
-\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
-\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
-\item {\small STOP} costs $G_{stop}$ gas
-\item {\small SUICIDE} costs $G_{suicide}$ gas
-\item All other operations cost $G_{step}$ gas.
-\end{itemize}
-
-Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+%More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
+%
+%\begin{itemize}
+%\item {\small SHA3} costs $G_{sha3}$ gas
+%\item {\small SLOAD} costs $G_{sload}$ gas
+%\item {\small BALANCE} costs $G_{balance}$ gas
+%\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
+%\begin{itemize}
+%\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
+%\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
+%\item $d = 1$ otherwise.
+%\end{itemize}
+%\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
+%\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
+%\item {\small STOP} costs $G_{stop}$ gas
+%\item {\small SUICIDE} costs $G_{suicide}$ gas
+%\item All other operations cost $G_{step}$ gas.
+%\end{itemize}
+%
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+
 See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$ and the remaining gas $g'$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g') \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \emptyset, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \begin{equation}
-X(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \begin{cases}
-(\boldsymbol{\sigma}, \boldsymbol{\mu}, I, ()) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)) & \text{otherwise}\\
+X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \big) \equiv \begin{cases}
+\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, \emptyset, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 
+Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
+
 The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
@@
 
-For the ease of reading, the instruction mnemonics, written in smallcaps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
+For the ease of reading, the instruction mnemonics, written in small-caps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
 
@@
 \begin{eqnarray}
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', I) \\
+O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', \mathbf{s}', I) \\
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
@@
 
-In general, we assume the memory and system state don't change:
+In general, we assume the memory, suicide list and system state don't change:
 \begin{eqnarray}
@@
 \boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\mathbf{s}' & \equiv & \mathbf{s} \\
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, \boldsymbol{\mu}_g) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \emptyset\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \mathbf{s}^+$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \emptyset, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \mathbf{s}^+$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
@@
 \midrule
-0xff & {\small SUICIDE} & 1 & 0 & Halt execution and obliterate account. \\
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a] \equiv \varnothing$ \\
-\bottomrule
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\\bottomrule
 \end{tabular*}

### 2014-08-15 Paper.tex
Version bump to PoC-6.
Block reward update.

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept V}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept VI}}}
 \author{
@@
 The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$, the block reward, and the coinbase of each uncle by $\frac{7}{8}$ of that.
+We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{8}$ of the block reward and the coinbase of the uncle by $\frac{3}{4}$ of the reward.
 Formally we define the function $\Omega$:
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + R_b \\
-\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{7}{8}R_b
+\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + (1 + \frac{|B_\mathbf{U}|}{8})R_b \\
+\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{3}{4}R_b
 \end{eqnarray}

### 2014-09-14 Paper.tex
Minor fix.
Small fixes.

@@
 
-However, a message call transaction contains:
+In contrast, a message call transaction contains:
 
@@
 
-We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
@@
 \begin{equation}
-P(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
+p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
 \end{equation}
@@
 \boldsymbol{\sigma}_1 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}_1[r]_b + v
+\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}[r]_b + v
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \Xi(\boldsymbol{\sigma}_1, g, I) \\
-I_a & \equiv & a \\
+I_a & \equiv & r \\
 I_o & \equiv & o \\
@@
 
-Over a account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
+Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
 This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \emptyset, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \varnothing, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \big) \equiv \begin{cases}
-\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, \emptyset, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, \varnothing, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-piling to.
+\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for transpiling to.
 
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum^{n = 0}_{n < \lVert \mathbf{b} \rVert} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs; this value may either be a 32 byte sequence or the empty byte sequence.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
 
@@
 \begin{equation}
-\forall_{I: \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 \end{eqnarray}
-gc
+
 We may then define the sender function $S$ of the transaction as:
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty associated code. \\
+&&&& This is the sender of original transaction; it is never an account with non-empty \\
+&&&& associated code. \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \emptyset\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \emptyset, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\

### 2014-10-04 JS.tex
Moved out old JS and Wire APIs.

--- 2014-08-03 Paper.tex
@@
-\documentclass[9pt,oneside]{amsart}
-%\usepackage{tweaklist}
-\usepackage{url}
-\usepackage{cancel}
-\usepackage{xspace}
-\usepackage{graphicx}
-\usepackage{multicol}
-\usepackage{subfig}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
-\usepackage{booktabs}
-\usepackage{array}
-\usepackage{verbatim}
-\usepackage{caption}
-\usepackage{natbib}
-\usepackage{float}
-\usepackage{pdflscape}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{afterpage}
-\usepackage{tikz}
-
-\newcommand{\hcancel}[1]{%
-    \tikz[baseline=(tocancel.base)]{
-        \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
-        \draw[black] (tocancel.south west) -- (tocancel.north east);
-    }%
-}%
-
-\definecolor{lightyellow}{rgb}{1,0.98,0.9}
-\definecolor{lightpink}{rgb}{1,0.94,0.95}
-
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\newcommand*\eg{e.g.\@\xspace}
-\newcommand*\Eg{e.g.\@\xspace}
-\newcommand*\ie{i.e.\@\xspace}
-%\renewcommand{\itemhook}{\setlength{\topsep}{0pt}  \setlength{\itemsep}{0pt}\setlength{\leftmargin}{15pt}}
-
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept V}}}
-\author{
-    Dr. Gavin Wood\\
-    Co-Founder \& Lead, Ethereum Project\\
-    gavin@ethereum.org
-}
-\begin{document}
-
-\pagecolor{lightyellow}
-%\pagecolor{lightpink}
-
-\begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
-Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
-We can call this paradigm a transactional singleton machine with shared-state.
-
-Ethereum implements this paradigm in a generalised manner.
-Furthermore it provides a plurality of such resources, each with a distinct state and operating code but able to interact through a message-passing framework with others.
-We discuss its design, implementation issues, the opportunities it provides and the future hurdles we envisage.
-\end{abstract}
-
-\maketitle
-
-\setlength{\columnsep}{20pt}
-\begin{multicols}{2}
-
-\section{Introduction}\label{sec:introduction}
-
-With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap.
-Technology-rooted movements like Bitcoin have demonstrated, through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
-This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
-Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications albeit rather simplistic ones.
-
-Ethereum is a project which attempts to build the generalised technology; technology on which all transaction-based state machine concepts may be built.
-Moreover it aims to provide to the end-developer a tightly integrated end-to-end system for building software on a hitherto unexplored compute paradigm in the mainstream: a trustful object messaging compute framework.
-
-\subsection{Driving Factors} \label{ch:driving}
-
-There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
-This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
-By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
-
-Dealings in this proposed system would have several attributes not often found in the real world.
-The incorruptibility of judgement, often difficult to find, comes naturally from a disinterested algorithmic interpreter.
-Transparency, or being able to see exactly how a state or judgement came about through the transaction log and rules or instructional codes, never happens perfectly in human-based systems since natural language is necessarily vague, information is often lacking, and plain old prejudices are difficult to shake.
-
-Overall, I wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
-
-\subsection{Previous Work} \label{ch:previous}
-
-\cite{buterin2013ethereum} first proposed the kernel of this work in late November, 2013.
-Though now evolved in many ways, the key functionality of a block-chain with a Turing-complete language and an effectively unlimited inter-transaction storage capability remains unchanged.
-
-Hashcash, introduced by \cite{back2002hashcash} (in a five-year retrospective), provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
-Though not widely adopted, the work was later utilised and expanded upon by \cite{nakamoto2008bitcoin} in order to devise a cryptographically secure mechanism for coming to a decentralised social consensus over the order and contents of a series of cryptographically signed financial transactions.
-The fruits of this project, Bitcoin, provided a first glimpse into a decentralised transaction ledger.
-
-Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol.
-Some of the best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
-Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.
-
-Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
-The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxiliary parts to the core protocol.
-The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.
-
-Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system.
-It has demonstrated that high efficiency gains can be made if the decentralisation premise is discarded.
-
-Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
-Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
-In this light, Ethereum may be seen as a general implementation of such a \textit{crypto-law} system.
-
-%E language?
-
-\section{The Blockchain Paradigm} \label{ch:overview}
-
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
-It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
-The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
-Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg be things such as reducing an account balance without an equal and opposite increase elsewhere.
-A valid state transition is one which comes about through a transaction.
-Formally:
-\begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
-\end{equation}
-
-where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
-
-Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
-They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes places as a state-transition function, adding value to a nominated account.
-
-Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block.
-It is achieved thanks to a cryptographically secure proof.
-This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
-
-Formally, we expand to:
-\begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
-B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
-
-This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
-
-\subsection{Value}
-
-In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
-The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
-One Ether is defined as being $10^{18}$ Wei.
-There exist other subdenominations of Ether:
-\par`
-\begin{center}
-\begin{tabular}{rl}
-\toprule
-Multiplier & Name \\
-\midrule
-$10^0$ & Wei \\
-$10^{12}$ & Szabo \\
-$10^{15}$ & Finney \\
-$10^{18}$ & Ether \\
-\bottomrule
-\end{tabular}
-\end{center}
-\par
-
-Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.
-
-\subsection{Which History?}
-
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
-If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
-
-This would mean that past a given point in time (block), multiple states of the system may coexist: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
-This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.
-
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
-This process is described in detail in section \ref{ch:ghost}.
-
-\section{Conventions}\label{ch:conventions}
-
-I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
-
-The two sets of highly structured, `top-level', state values, are denoted with bold lowercase Greek letters.
-They fall into those of world-state, which are denoted $\boldsymbol{\sigma}$ (or a variant thereupon) and those of machine-state, $\boldsymbol{\mu}$.
-
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
-
-For most functions, an uppercase letter is used, e.g.
-$C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the SHA3 hash function is denoted $\texttt{SHA3}$.
-
-Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction.
-The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
-
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
-For particularly important values, a bold uppercase letter may be used.
-
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
-The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$.
-This is formally defined in section \ref{ch:block}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
-
-In the case of the global state $\boldsymbol{\sigma}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
-
-When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
-
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
-
-\begin{equation}
-\ell(\mathbf{x}) \equiv \mathbf{x}[\lVert \mathbf{x} \rVert - 1]
-\end{equation}
-
-\section{Blocks, State and Transactions} \label{ch:bst}
-
-Having introduced the basic concepts behind Ethereum, we will discuss the meaning of a transaction, a block and the state in more detail.
-
-\subsection{World State} \label{ch:state}
-
-The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
-Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
-This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
-Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly.
-Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.
-
-The account state comprises the first two, and potentially the last two, of the following fields:
-
-\begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from a byte array of size 32 to an RLP-encoded integer.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
-\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
-All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small SHA3}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
-\end{description}
-
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
-\begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
-\begin{equation}
-L_I\big( (k, v) \big) \equiv \big(k, \texttt{\small RLP}(v)\big)
-\end{equation}
-
-where:
-\begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
-\end{equation}
-
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
-
-If the \textbf{codeHash} field is the SHA3 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small SHA3}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
-\begin{equation}
-L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
-\end{equation}
-where
-\begin{equation}
-p(a) \equiv  \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
-\end{equation}
-
-This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
-\begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
-\end{equation}
-where $v$ is the account validity function:
-\begin{eqnarray}
-\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{eqnarray}
-
-\subsection{The Transaction} \label{ch:transaction}
-
-A transaction (formally, $T$) is a single cryptographically-signed instruction sent by an actor external to Ethereum.
-An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
-There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
-
-\begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or the zero address for a contract creation transaction; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
-This is expanded in Appendix \ref{app:signing}.
-\end{description}
-
-Additionally, a contract creation transaction contains:
-
-\begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
-\end{description}
-
-\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executes each time the account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is executed only once at account creation and gets discarded immediately thereafter.
-
-However, a message call transaction contains:
-
-\begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
-\end{description}
-
-Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
-For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
-
-\begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
-\begin{equation}
-\begin{array}[t]{lclclc}
-T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge \\
-T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
-T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} & \wedge \\
-T_t \in \mathbb{B}_{20} &&&&
-\end{array}
-\end{equation}
-
-\subsection{The Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{R}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
-The block header contains several pieces of information:
-
-\begin{description}
-\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety; formally $H_p$.
-\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block; formally $H_u$.
-\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_b$.
-\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsTrie] The SHA3 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[minGasPrice] A scalar value equal to the minimum price of gas a transaction must have provided in order to be sufficient for inclusion by this miner in this block; formally $H_m$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_u$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
-\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
-\end{description}
-
-The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transaction \textit{receipts}. Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{R}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}
-
-The transaction receipt is a tuple of three items comprising the transaction, $R_T$, together with the post-transaction state, $R_{\boldsymbol{\sigma}}$, and the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$:
-\begin{equation}
-R \equiv (R_T, R_{\boldsymbol{\sigma}}, R_u)
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (L_T(R_T), \mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u)
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the second item.
-
-We assert $R_u$, the cumulative gas used is a positive integer:
-\begin{equation}
-R_u \in \mathbb{P}
-\end{equation}
-
-For convenience, I also name $B_\mathbf{T}$ the series of transactions encoded by the transaction receipts $B_\mathbf{R}$:
-\begin{equation}
-B_\mathbf{T} \equiv (B_\mathbf{R}[0]_T, B_\mathbf{R}[1]_T, ...)
-\end{equation}
-
-Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-\subsubsection{Holistic Validity}
-
-We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
-\begin{equation}
-\begin{array}[t]{lclc}
-H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i])\})) & \wedge \\
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
-\end{array}
-\end{equation}
-
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
-\begin{equation}
-P(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
-\end{equation}
-
-Furthermore:
-\begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
-
-\subsubsection{Serialisation}
-
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d,\\ H_i, H_m, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
-
-The component types are defined thus:
-\begin{equation}
-\begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_u \in \mathbb{B}_{32} & \wedge & H_b \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
-H_i \in \mathbb{P} & \wedge & H_m \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
-H_u \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge \\
-H_n \in \mathbb{B}_{32} &&&&
-\end{array}
-\end{equation}
-
-where
-\begin{equation}
-\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
-\end{equation}
-
-We now have a rigorous specification for the construction of a formal block structure.
-The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
-
-\subsubsection{Block Header Validity}
-
-We define $P(B_H)$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(B')) = H_p
-\end{equation}
-
-The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{cases}
-2^{22} & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 42\\
-{P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{otherwise}\\
-\end{cases}
-\end{equation}
-
-The canonical gas limit of a block of header $H$ is defined as $L(H)$:
-\begin{eqnarray}
-L(H) & \equiv & \begin{cases}
-10^6 & \text{if} \quad H_i = 0\\
-125000 & \text{if} \quad L'(H) < 125000\\
-L'(H) & \text{otherwise}
-\end{cases}\\
-L'(H) & \equiv & \big\lfloor \frac{1023 {P(H)_H}_l + \lfloor \frac{6}{5}{P(H)_H}_u \rfloor}{1024}\big\rfloor
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
-
-The nonce, $H_n$, must satisfy the relation:
-\begin{equation}
-\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
-Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
-
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_l = L(H) \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & \lVert H_x \rVert < 1024
-\end{eqnarray}
-
-Noting additionally that \textbf{extraData} must be at most 1024 bytes.
-
-\section{Gas and Payment} \label{ch:payment}
-
-In order to avoid issues of network abuse and to sidestep the inevitable questions stemming from Turing completeness, all programmable computation in Ethereum is subject to fees.
-The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation).
-Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing account storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.
-
-Every transaction has a specific amount of gas associated with it: \textbf{gasLimit}.
-This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according \textbf{gasPrice}, also specified in the transaction.
-The transaction is considered invalid if the account balance cannot support such a purchase.
-It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender's account.
-Gas does not exist outside of the execution of a transaction.
-Thus for accounts with trusted code associated, a relatively high gas limit may be set and left alone.
-
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under the control of the miner.
-Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
-Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
-Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.
-
-%\subsubsection{Determining Computation Costs}
-
-\section{Transaction Execution} \label{ch:transactions}
-
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
-It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
-These include:
-
-\begin{enumerate}
-\item The transaction signature is valid;
-\item the transaction nonce is valid (equivalent to the sender account's current nonce);
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
-\end{enumerate}
-
-Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
-\begin{equation}
-\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
-\end{equation}
-
-Thus $\boldsymbol{\sigma}'$ is the post-transactional state. We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, to be defined later.
-
-We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{equation}
-g_0 \equiv \begin{cases}
-\lVert T_\mathbf{i}\rVert G_{txdata} + G_{transaction} & \text{if} \quad T_t = 0 \\
-\lVert T_\mathbf{d}\rVert G_{txdata} + G_{transaction} & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where $\lVert T_\mathbf{d}\rVert$ and $\lVert T_\mathbf{i}\rVert$ are the sizes, in bytes, of the transaction's associated data and initialisation EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
-
-%todo Explain g_d reason?
-
-The up-front cost $v_0$ is calculated as:
-\begin{equation}
-v_0 \equiv T_g T_p + T_v
-\end{equation}
-
-The validity is determined as:
-\begin{equation}
-\begin{array}[t]{rcl}
-S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leq & T_g \quad \wedge \\ 
-v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_l & \leq & {B_H}_l - \ell(B_\mathbf{R})_u
-\end{array}
-\end{equation}
-
-Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$:
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by the up-front cost, $v_0$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
-The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
-
-We define the checkpoint state $\boldsymbol{\sigma}_0$:
-\begin{eqnarray}
-\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - v_0 \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
-\end{eqnarray}
-
-%todo explain suicide list
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and suicide list $\mathbf{s}$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', \mathbf{s}) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}) & \text{if} \quad T_t = 0 \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where
-\begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
-
-After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate.
-The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g' T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g') T_p \\
-m & \equiv & {B_H}_b
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in \mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-And finally specify $\Upsilon^g$, the total gas used in this transaction:
-\begin{equation}
-\Upsilon^g(\boldsymbol{\sigma}, T) \equiv g_0 + g'
-\end{equation}
-
-%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
-%\begin{eqnarray}
-%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
-%\end{eqnarray}
-
-\section{Contract Creation} \label{ch:create}
-
-There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, additional database entries and remaining gas $(\boldsymbol{\sigma}', g')$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i})
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure containing only the sender and the nonce.
-In the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
-Thus we define the creation address function $A$:
-\begin{eqnarray}
-A(s, n) & \equiv & a \quad \text{where:} \\
-a & = & \arg \min_x : x \geqslant a' \wedge \boldsymbol{\sigma}[x] = \varnothing \\
-a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
-\end{eqnarray}
-
-where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
-
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the empty string, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
-\begin{equation}
-\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except:}
-\end{equation}
-\begin{equation}
-\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
-\end{equation}
-
-where $a$ is the address of the new account, as defined above.
-% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
-
-Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
-Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$, the suicide list $\mathbf{s}$ and the body code of the account $\mathbf{b}$.
-
-Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this exceptional case we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-The gas remaining should be zero.
-If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state, gas and suicide list as $(\boldsymbol{\sigma}', g', \mathbf{s})$ where:
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
-\end{equation}
-\begin{eqnarray}
-\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}^{**} & \text{if} \quad \boldsymbol{\sigma}^{**}[a] = \varnothing \\
-\begin{array}[b]{l}
-\boldsymbol{\sigma}^{**} \quad \text{except:} \\
-\boldsymbol{\sigma}'[a]_{\mathbf{b}} = \mathbf{o}
-\end{array} & \text{otherwise}
-\end{cases} \\
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i}
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple.
-$I_H$ has no special treatment and is determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
-
-\subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code.
-Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SUICIDE} instruction, the matter is moot since the account will be deleted before the transaction is completed.
-For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
-
-\section{Message Call} \label{ch:call}
-
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call.
-Aside from evaluating to a new state and suicide list, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
-This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
-\end{equation}
-
-We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
-\begin{eqnarray}
-\boldsymbol{\sigma}_1 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}_1[r]_b + v
-\end{eqnarray}
-
-The account's associated code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}_1[r]_c = \varnothing \\
-\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \Xi(\boldsymbol{\sigma}_1, g, I) \\
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
-\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny SHA3}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
-
-\section{Execution Model} \label{ch:model}
-
-The execution model specifies how the system state is altered given a series of bytecode instructions and a small tuple of environmental data.
-This is specified through a formal model of a virtual state machine, known as the Ethereum Virtual Machine (EVM).
-It is a \textit{quasi-}Turing-complete machine; the \textit{quasi} qualification comes from the fact that the computation is intrinsically bounded through a parameter, \textit{gas}, which limits the total amount of computation done.
-
-\subsection{Basics}
-
-The EVM is a simple stack-based architecture.
-The word size of the machine (and thus size of stack item) is 256-bit.
-This was chosen to facilitate the SHA3-256 hash scheme and elliptic-curve computations.
-The memory model is a simple word-addressed byte array.
-The stack has an unlimited size.
-The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
-Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state.
-All locations in both storage and memory are well-defined initially as zero.
-
-The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only through a specialised instruction.
-
-The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-These unambiguously and validly result in immediate halting of the machine with all state changes left intact.
-The one piece of exceptional execution that does not leave state changes intact is the out-of-gas (OOG) exception.
-Here, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.
-
-\subsection{Fees Overview}
-
-Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small SHA3}.
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE} and {\small CALL}.
-Finally, gas may be paid due to an increase in the usage of the memory.
-
-Over a account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
-This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
-Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds since at the present price of Ether and default gas price, that would cost around US\$20M for the memory fee alone.
-That said, implementations must be able to manage this eventuality.
-
-Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is waived; in fact, it is effectively paid up-front since the initial usage of a storage location costs twice as much as the normal usage.
-
-%More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
-%
-%\begin{itemize}
-%\item {\small SHA3} costs $G_{sha3}$ gas
-%\item {\small SLOAD} costs $G_{sload}$ gas
-%\item {\small BALANCE} costs $G_{balance}$ gas
-%\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
-%\begin{itemize}
-%\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
-%\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
-%\item $d = 1$ otherwise.
-%\end{itemize}
-%\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
-%\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
-%\item {\small STOP} costs $G_{stop}$ gas
-%\item {\small SUICIDE} costs $G_{suicide}$ gas
-%\item All other operations cost $G_{step}$ gas.
-%\end{itemize}
-%
-%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
-
-See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
-
-%Whenever a higher memory index is referenced, the fee difference to take it to the higher usage from the original (lower) usage is charged.
-Notably, because {\small MSTORE} and {\small MLOAD} operate on word lengths, they implicitly increase the highest-accessed index to 31 greater than their target index.
-
-\subsection{Execution Environment}
-
-In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
-
-\begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\end{itemize}
-
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-\subsection{Execution Overview}
-
-We must now define the $\Xi$ function.
-In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$.
-Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \emptyset, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
-\end{eqnarray}
-\begin{equation}
-X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \big) \equiv \begin{cases}
-\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, \emptyset, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I)\big) & \text{otherwise}\\
-\end{cases}
-\end{equation}
-
-where
-\begin{eqnarray}
-\mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
-\end{eqnarray}
-
-Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
-
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
-
-For the ease of reading, the instruction mnemonics, written in small-caps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
-
-For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
-\begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
-\text{\small STOP} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
-\subsubsection{Exceptional Halting}
-
-The exceptional halting function $Z$ is defined as:
-\begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
-\begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined) or if there are insufficient stack items.
-The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-
-\subsubsection{Normal Halting}
-
-The normal halting function $H$ is defined:
-\begin{equation}
-H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
-\varnothing & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
-
-\subsection{The Execution Cycle}
-
-Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
-\begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', \mathbf{s}', I) \\
-\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
-\end{eqnarray}
-
-The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
-\begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-\boldsymbol{\mu}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
-\boldsymbol{\mu}_{pc} + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-where $p$ is the byte size of the push instruction, defined as:
-\begin{equation}
-p \equiv w - \text{\small PUSH1} + 2
-\end{equation}
-
-In general, we assume the memory, suicide list and system state don't change:
-\begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
-\mathbf{s}' & \equiv & \mathbf{s} \\
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
-\end{eqnarray}
-
-However, instructions do typically alter one or several components of these values.
-Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and a formal description of the gas requirements.
-
-\section{Blocktree to Blockchain} \label{ch:ghost}
-
-The canonical blockchain is a path from root to leaf through the entire block tree.
-In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
-The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
-This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
-
-This scheme notably ignores so-called \textit{stale} blocks: valid, mined blocks, which were propagated too late into the network and thus were beaten to network consensus by a sibling block (one with the same parent).
-Such blocks become more common as the network propagation time approaches the ideal inter-block time.
-However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to buttress the more popular blockchain making it a stronger choice over less popular (though potentially longer) competitors.
-
-This increases overall network security by making it much harder for an adversary to silently mine a canonical blockchain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of overriding existing blocks and reversing the transactions within.
-
-In order to validate the extra computation, a given block $B$ may include the block headers from any known uncle blocks (i.e.
-blocks whose parent is equivalent to the grandparent of $B$).
-Since a block header includes the nonce, a proof-of-work, then the header alone is enough to validate the computation done.
-Any such blocks contribute toward the total computation or \textit{total difficulty} of a chain that includes them.
-To incentivise computation and inclusion, a reward is given both to the miner of the stale block and the miner of the block that references it.
-
-Thus we define the total difficulty of block $B$ recursively as:
-\begin{eqnarray}
-B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
-B' & \equiv & P(B_H)
-\end{eqnarray}
-
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
-
-\section{Block Finalisation} \label{ch:finalisation}
-
-The process of finalising a block involves four stages:
-
-\begin{enumerate}
-\item Validate (or, if mining, determine) uncles;
-\item validate (or, if mining, determine) transactions;
-\item apply rewards;
-\item verify (or, if mining, compute a valid) state and nonce.
-\end{enumerate}
-
-\subsection{Uncle Validation}
-
-The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:
-\begin{equation}
-\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; P(U) = P(P(B_H)) \; \wedge \; P(B_H) \neq B
-\end{equation}
-
-\subsection{Transaction Validation}
-
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} and \textbf{minGasPrice} must correspond faithfully to the transactions listed.
-In the case of ${B_H}_m$, the minimum gas price, all transactions included in the transaction receipt must have a gas price, ${T_0}_p$, that is at least this value:
-\begin{equation}
-\forall \; T \in B_\mathbf{T}: T_p \geq {B_H}_m
-\end{equation}
-
-In the case of ${B_H}_u$, the total gas used in the block, it must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_u = \ell(\mathbf{R})_u
-\end{equation}
-
-\subsection{Reward Application}
-
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$, the block reward, and the coinbase of each uncle by $\frac{7}{8}$ of that.
-Formally we define the function $\Omega$:
-\begin{eqnarray}
-\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + R_b \\
-\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{7}{8}R_b
-\end{eqnarray}
-
-We define the block reward as 1500 Finney:
-\begin{equation}
-\text{Let} \quad R_b = 1.5 \times 10^{18}
-\end{equation}
-
-\subsection{State \& Nonce Validation}
-
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
-\begin{equation}
-\Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature a mutable data structure.
-
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
-\begin{eqnarray}
-\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathtt{PoW}(B^*, n) < \frac{2^{256}}{H_d} \\
-B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
-\end{eqnarray}
-
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states and cumulative gas used after each transaction.
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-B_\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-B_\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
-\begin{array}[b]{l}
-\Upsilon^g(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + B_\mathbf{R}[n-1]_u
-\end{array}
- & \text{otherwise} \end{cases}
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{R})_{\boldsymbol{\sigma}})
-\end{equation}
-
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
-
-\subsection{Mining Proof-of-Work} \label{ch:pow}
-
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
-However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
-
-For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of, or reward from, specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
-
-Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
-Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
-
-One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task.
-In Bitcoin's case the task is the SHA256 hash function.
-While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
-Because of this, a proof-of-work function that is ASIC-resistant (i.e.
-difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
-
-Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-
-More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
-\begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
-
-As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
-
-\subsubsection{PoC Series}
-
-For the PoC series, we use a simplified proof-of-work.
-This is not ASIC resistant and is meant merely as a placeholder.
-It utilises the bare SHA3 hash function to secure the block chain by requiring the SHA3 hash of the concatenation of the nonce and the header's SHA3 hash to be sufficiently low.
-
-It is formally defined as $\mathtt{PoW}$:
-\begin{equation}
-\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny SHA3}(\mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
-\end{equation}
-
-where:
-$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
-$\mathtt{\tiny SHA3}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
-$n$ is the nonce, a series of 32 bytes;
-$\circ$ is the series concatenation operator;
-$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
-
-\subsubsection{Release Series}
-
-For the release series, we use a more complex proof-of-work.
-This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
-Secondly that it concerns the utilisation of either the blockchain or the full state trie.
-
-As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
-A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
-$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
-Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
-Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
-swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
-This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
-
-Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
-
-%\section{User Interface}
-%
-%Browser-style.
-%Displays ETH balance (i.e. functions as a wallet).
-%Tabs or perhaps MacOS style app icons.
-%Display taken up with a single app.
-%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
-%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
-%Preconfigured with home app that is the Eth app store.
-%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
-
-\section{Implementing Contracts}
-
-There are several patterns of contracts engineering that allow particular useful behaviours; two of these that I will briefly discuss are data feeds and random numbers.
-
-\subsection{Data Feeds}
-A data feed contract is one which provides a single service: it gives access to information from the external world within Ethereum.
-The accuracy and timeliness of this information is not guaranteed and it is the task of a secondary contract author---the contract that utilises the data feed---to determine how much trust can be placed in any single data feed.
-
-The general pattern involves a single contract within Ethereum which, when given a message call, replies with some timely information concerning an external phenomenon.
-An example might be the local temperature of New York City.
-This would be implemented as a contract that returned that value of some known point in storage.
-Of course this point in storage must be maintained with the correct such temperature, and thus the second part of the pattern would be for an external server to run an Ethereum node, and immediately on discovery of a new block, creates a new valid transaction, sent to the contract, updating said value in storage.
-The contract's code would accept such updates only from the identity contained on said server.
-
-\subsection{Random Numbers}
-Providing random numbers within a deterministic system is, naturally, an impossible task.
-However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's coinbase address.
-For a series of such numbers, a trivial solution would be to work from the previous pseudo-random number, adding some constant amount and hashing the result.
-
-This strategy does have a downside: a miner with sufficient power could alter any of the above values in order to deliver a seed in order to alter the outcome of the pseudorandom-based executions.
-For a more secure pseudo-random offering, all involved parties could agree on a number of data feed contracts; these could be combined along with the block timestamp and hashed to produce the first number in the series.
-By spreading the inputs and thus the trust between numerous parties the likelihood of a malicious miner altering the outcome becomes increasingly less likely.
-
-\section{Future Directions} \label{ch:future}
-
-The state database won't be forced to maintain all past state trie structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
-
-Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10000th block) could be maintained by the peer network, effectively recasting the genesis block.
-This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
-
-Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
-
-\subsection{Scalability}
-
-Scalability remains an eternal concern.
-With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy.
-Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger.
-However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.
-
-Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidated each block and with duplicate or invalid transactions thrown out accordingly.
-
-Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
-
-\section{Conclusion} \label{ch:conclusion}
-
-I have introduced, discussed and formally defined the protocol of Ethereum.
-Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
-Contracts may be authored in order to algorithmically specify and autonomously enforce rules of interaction.
-
-\section{Acknowledgements}
-
-Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
-
-\bibliography{Biblio}
-\bibliographystyle{plainnat}
-
-\end{multicols}
-
-\appendix
-
-\section{Terminology}
-
-\begin{description}
-\item[External Actor] A person or other entity able to interface to an Ethereum node, but external to the world of Ethereum.
-It can interact with Ethereum through depositing signed Transactions and inspecting the blockchain and associated state.
-Has one (or more) intrinsic Accounts.
-
-\item[Address] A 160-bit code used for identifying Accounts.
-
-\item[Account] Accounts have an intrinsic balance and transaction count maintained as part of the Ethereum state.
-They also have some (possibly empty) EVM Code and a (possibly empty) Storage State associated with them.
-Though homogenous, it makes sense to distinguish between two practical types of account: those with empty associated EVM Code (thus the account balance is controlled, if at all, by some external entity) and those with non-empty associated EVM Code (thus the account represents an Autonomous Object).
-Each Account has a single Address that identifies it.
-
-\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the blockchain.
-
-\item[Autonomous Object] A notional object existent only within the hypothetical state of Ethereum.
-Has an intrinsic address and thus an associated account; the account will have non-empty associated EVM Code.
-Incorporated only as the Storage State of that account.
-
-\item[Storage State] The information particular to a given Account that is maintained between the times that the Account's associated EVM Code runs.
-
-\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.
-
-\item[Message Call] The act of passing a message from one Account to another.
-If the destination account is associated with non-empty EVM Code, then the VM will be started with the state of said Object and the Message acted upon.
-If the message sender is an Autonomous Object, then the Call passes any data returned from the VM operation.
-
-\item[Gas] The fundamental network cost unit.
-Paid for exclusively by Ether (as of PoC-4), which is converted freely to and from Gas as required.
-Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and miners are free to ignore Transactions whose Gas price is too low.
-
-\item[Contract] Informal term used to mean both a piece of EVM Code that may be associated with an Account or an Autonomous Object.
-
-\item[Object] Synonym for Autonomous Object.
-
-\item[App] An end-user-visible application hosted in the Ethereum Browser.
-
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
-
-\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for an Account's associated EVM Code.
-
-\item[Ethereum Runtime Environment] (aka ERE) The environment which is provided to an Autonomous Object executing in the EVM.
-Includes the EVM but also the structure of the world state on which the EVM relies for certain I/O instructions including CALL \& CREATE.
-
-\item[EVM Code] The bytecode that the EVM can natively execute. Used to formally specify the meaning and ramifications of a message to an Account.
-
-\item[EVM Assembly] The human-readable form of EVM-code.
-
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-piling to.
-
-\end{description}
-
-\section{Fee Schedule}\label{app:fees}
-
-The fee schedule $G$ is a tuple of 10 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
-
-\begin{tabular*}{\columnwidth}[h]{lrl}
-\toprule
-Name & Value & Description* \\
-\midrule
-$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
-$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
-$G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
-$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
-$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
-$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled or waived sometimes). \\
-$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
-$G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
-$G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
-$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
-$G_{transaction}$ & 500 & Paid for every transaction. \\
-\bottomrule
-\end{tabular*}
-
-\section{Recursive Length Prefix}\label{app:rlp}
-This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
-
-We define the set of possible structures $\mathbb{T}$:
-\begin{eqnarray}
-\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
-\end{eqnarray}
-
-Where $\mathbb{Y}$ is the set of bytes.
-Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
-
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
-
-\begin{itemize}
-\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
-\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
-\end{itemize}
-
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
-\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
-(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
-\end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum^{n = 0}_{n < \lVert \mathbf{b} \rVert} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
-\end{eqnarray}
-
-Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
-
-If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
-
-\begin{itemize}
-\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
-\end{itemize}
-
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
-(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
-\end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
-\end{eqnarray}
-
-If RLP is used to encode a scalar, defined only as a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
-Thus the RLP of some positive integer $i$ is defined as:
-\begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
-\end{equation}
-
-When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.
-
-There is no specific canonical encoding format for signed or floating-point values.
-
-\section{Hex-Prefix Encoding}\label{app:hexprefix}
-Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array.
-It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
-
-It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
-(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
-\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
-(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
-\text{otherwise}
-\end{cases} \\
-f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \\ 0 & \text{otherwise} \end{cases}
-\end{eqnarray}
-
-Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$.
-The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number.
-All remaining nibbles (now an even number) fit properly into the remaining bytes.
-
-\section{Modified Merkle Patricia Tree}\label{app:trie}
-The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
-It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
-
-Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
-\begin{equation}
-\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
-\end{equation}
-
-When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
-\begin{equation}
-\forall_{I: \in \mathfrak{I}} I \equiv (I_0, I_1)
-\end{equation}
-
-Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
-\begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
-\begin{cases}
-\lfloor \mathbf{k}_n[\lfloor i \div 2 \rfloor] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure.
-The empty trie is defined as being a the empty byte sequence, $()$:
-\begin{equation}
-\texttt{\small TRIE}(\mathfrak{I}) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-\texttt{\small SHA3}(c(\mathfrak{I}, 0)) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-We also assume a function $n$, the trie's node cap function.
-When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose SHA3 hash evaluates to our reference.
-Thus we define in terms of $c$, the node composition function:
-\begin{equation}
-n(\mathfrak{I}, i) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{\small SHA3}(c(\mathfrak{I}, i)) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair.
-The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
-Unlike a radix tree, in the case of multiple keys shared the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
-Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
-There are three kinds of nodes in the trie:
-\begin{description}
-\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
-\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of nibbles keys and branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
-\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
-The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
-\end{description}
-
-A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
-\begin{equation}
-c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
-(x - 1)] = \mathbf{l} \\
-\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
-v & = & \begin{cases}
-I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
-() & \text{otherwise}
-\end{cases}
-\end{array}
-\end{cases}
-\end{equation}
-
-\subsection{Trie Database}
-Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the SHA3 hash's collision resistance.
-In reality, a sensible implementation will not fully recompute the trie root hash for each set.
-
-A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
-This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
-Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
-
-It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer in the range $(0, 2^{256})$ represented as a byte array of length 32 in big-endian form.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARESTORE}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
-It is assumed that $v$ is the `recovery id', a 2-bit value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
-
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the SHA3 hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny SHA3}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the SHA3 hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
-\begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small SHA3}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}
-gc
-We may then define the sender function $S$ of the transaction as:
-\begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny SHA3}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
-\end{equation}
-
-The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
-\begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
-\end{equation}
-
-\section{Virtual Machine Specification}\label{app:vm}
-
-When interpreting 256-bit binary values as integers, the representation is big-endian.
-
-When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the left most 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
-
-\subsection{Gas Cost}
-
-The general gas cost function, $C$, is defined as:
-
-\begin{eqnarray}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & G_{memory}(\boldsymbol{\mu}'_i - \boldsymbol{\mu}_i) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL}\\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{stop} & \text{if} \quad w = \text{\small STOP}\\
-G_{suicide} & \text{if} \quad w = \text{\small SUICIDE}\\
-G_{step} & \text{otherwise}
-\end{cases} \\
-w & \equiv & \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
-\text{\small STOP} & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write; $\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
-
-\subsection{Instruction Set}
-
-As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
-In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
-
-\begin{equation}
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', I) \quad \text{with exceptions, as noted}
-\end{equation}
-
-Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
-&&&& $\boldsymbol{\mu}'_R = []$ \\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-\midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
-\midrule
-0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
-\midrule
-0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
-\midrule
-0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
-\midrule
-0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv -\boldsymbol{\mu}_\mathbf{s}[0]$ \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
-\midrule
-0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x0c & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
-\midrule
-0x0d & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
-\midrule
-0x0e & {\small EQ} & 2 & 1 & Equality comparision. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x0f & {\small NOT} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[0]$} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..7]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny SHA3}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
-\midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{when} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
-\midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty associated code. \\
-\midrule
-0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_c$ \\
-&&&& This is the address of the account that is directly responsible for this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
-\midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
-\midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-\midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
-&&&& This is gas price specified by the originating transaction.\\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_p$ \\
-&&&& $I_{H_p}$ is the previous block's hash. \\
-\midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_b$ \\
-\midrule
-0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_t$ \\
-\midrule
-0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
-\midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
-\midrule
-0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
-\midrule
-0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-\midrule
-0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-\midrule
-0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-\midrule
-0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
-\midrule
-0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-2G_{sstore} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
-G_{sstore} & \text{otherwise}
-\end{cases}$ \\
-\midrule
-0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
-\midrule
-0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{i}$ \\
-\midrule
-0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[\boldsymbol{\mu}_{pc} + 1]$ \\
-&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) ]$ \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) ]$ \\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \emptyset\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \mathbf{s}^+$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$; $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created account, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
-&&&& Thus the operand order is: value, input offset, input size. \\
-\midrule
-0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \emptyset, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \mathbf{s}^+$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$; $x=1$ otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-\midrule
-0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
-&&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
-\midrule
-0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\\bottomrule
-\end{tabular*}
-
-%\section{Low-level Lisp-like Language}\label{app:lll}
-%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-
-\section{Wire Protocol}\label{app:wire}
-The wire-protocol specifies a network-level protocol for how two peers can communicate.
-It includes handshake procedures and the means for transferring information such as peers, blocks and transactions.
-Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.
-
-Ethereum nodes may connect to each other over TCP only.
-Peers are free to advertise and accept connections on any port(s) they wish, however, a default port on which the connection may be listened and made will be 30303.
-
-Though TCP provides a connection-oriented medium, Ethereum nodes communicate in terms of packets.
-These packets are formed as a 4-byte synchronisation token (0x22400891), a 4-byte "payload size", to be interpreted as a big-endian integer and finally an N-byte \textbf{RLP-serialised} data structure, where N is the aforementioned "payload size".
-To be clear, the payload size specifies the number of bytes in the packet ''following'' the first 8.
-
-There are a number of different types of message that may be sent.
-This ``type'' is always determined by the first entry of the structure, represented as a scalar.
-The structure of each message type is described below.
-
-\begin{tabular*}{\columnwidth}[h]{rlll}
-\toprule
-\multicolumn{4}{c}{\textbf{00s: Session control}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
-0x00 & \textsc{Hello} & & $(\text{0x}00, v \in \mathbb{P}, n \in \mathbb{P}, \mathbf{i} \in \mathbb{B}, c \in \mathbb{P}, p \in \mathbb{P}, u \in \mathbb{B}_{64})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-This is the first packet sent over the connection, and sent once by both sides. No other messages may be sent until a \textsc{Hello} is received.
-\begin{itemize}
-\item $v$ is the Protocol Version. See the latest documentation for which version is current.
-\item $n$ is the Network Id should be 0.
-\item $\mathbf{i}$ is the Client Id and specifies the client software identity as a human-readable string (e.g. ``Ethereum(++)/1.0.0'').
-\item $c$ is the client's Capabilities and specifies the capabilities of the client as a set of flags; presently three bits are used:
-\begin{description}
-\item[0x01] Client provides peer discovery service;
-\item[0x02] Client provides transaction relaying service;
-\item[0x04] Client provides block-chain querying service.
-\end{description}
-\item $p$ is the Listen Port and specifies the port that the client is listening on (on the interface that the present connection traverses). If 0 it indicates the client is not listening.
-\item $u$ is the Unique Identity of the node and specifies a 512-bit hash that identifies this node.
-\end{itemize}
-}\\
-\midrule
-0x01 & \textsc{Disconnect} && $(\text{0x}01, r \in \mathbb{P})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Inform the peer that a disconnection is imminent; if received, a peer should disconnect immediately.
-When sending, well-behaved hosts give their peers a fighting chance (read: wait 2 seconds) to disconnect to before disconnecting themselves.
-\begin{itemize}
-\item $r$ is an integer specifying one of a number of reasons for disconnect:
-\begin{description}
-\item[0x00] Disconnect requested;
-\item[0x01] TCP sub-system error;
-\item[0x02] Bad protocol;
-\item[0x03] Useless peer;
-\item[0x04] Too many peers;
-\item[0x05] Already connected;
-\item[0x06] Wrong genesis block;
-\item[0x07] Incompatible network protocols;
-\item[0x08] Client quitting.
-\end{description}
-\end{itemize}
-}\\
-\midrule
-0x02 & \textsc{Ping} & \textsc{Pong} & $(\text{0x}02)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Requests an immediate reply of \textsc{Pong} from the peer.}\\
-\midrule
-0x03 & \textsc{Pong} && $(\text{0x}03)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Reply to peer's \textsc{Ping} packet.}\\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlll}
-\toprule
-\multicolumn{4}{c}{\textbf{10s: Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
-0x10 & \textsc{GetPeers} & \textsc{Peers} & $(\text{0x}10)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to enumerate some known peers for us to connect to. This should include the peer itself.}\\
-\midrule
-0x11 & \textsc{Peers} & & $(\text{0x}11, (a_0 \in \mathbb{B}_4, p_0 \in \mathbb{P}, i_0 \in \mathbb{B}_{64}), (a_1 \in \mathbb{B}_4, p_1 \in \mathbb{P}, i_1 \in \mathbb{B}_{64}), ...)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Specifies a number of known peers. 
-\begin{itemize}
-\item $a_0$, $a_1$, ... is the node's IPv4 address, a 4-byte array that should be interpreted as the IP address $a_0[0]$.$a_0[1]$.$a_0[2]$.$a_0[3]$.
-\item $p_0$, $p_1$, ... is the node's Port and is an integer.
-\item $i_0$, $i_1$, ... is the node's Unique Identifier and is the 512-bit hash that serves to identify the node.
-\end{itemize}
-}\\
-\midrule
-0x12 & \textsc{Transactions} & & $(\text{0x}12, L_T(T_0), L_T(T_1), ...)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-where $L_T$ is the transaction preparation function, as specified in section \ref{ch:block}.
-
-Specify a transaction or transactions that the peer should make sure is included on its transaction queue.
-The items in the list (following the first item 0x12) are transactions in the format described in the main Ethereum specification.
-\begin{itemize}
-\item $T_0$, $T_1$, ... are the transactions that should be assimilated.
-\end{itemize}
-}\\
-\midrule
-0x13 & \textsc{Blocks} && $(\text{0x}13, L_B(b_0), L_B(b_1), ...)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Where $L_B$ is the block preparation function, as specified in section \ref{ch:block}.
-
-Specify a block or blocks that the peer should know about. The items in the list (following the first item, 0x13) are blocks as described in the format described in the main specification.
-\begin{itemize}
-\item $b_0$, $b_1$, ... are the blocks that should be assimilated.
-\end{itemize}
-}\\
-\midrule
-0x14 & \textsc{GetChain} & \textsc{Blocks} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Request the peer to send $c$ blocks in the current canonical block chain that are children of one of a number of given blocks, according to a preferential order with $p_0$ being the most prefered.
-If the designated parent is the present block chain head, an empty reply should be sent.
-If none of the parents are in the current canonical block chain, then a NotInChain message should be sent along with $p_n$, the least preferential parent.
-If no parents are passed, then a reply need not be made.
-\begin{itemize}
-\item $p_0$, $p_1$, ...
-are the SHA3 hashes of the parents of blocks that we should be informed of with a \textsc{Blocks} reply.
-Typically, these will be specified in increasing age (or decreasing block number).
-\item $c$ is the number of children blocks of the most preferred parent that we should be informed of through the corresponding \textsc{Blocks} reply.
-\end{itemize}
-}\\
-\midrule
-0x15 & \textsc{NotInChain} && $(\text{0x}15, p \in \mathbb{B}_{32})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Inform the peer that a particular block was not found in its block chain.
-\begin{itemize}
-\item $p$ is the SHA3 hash of the block that was not found in the block chain. Typically, this will be the least preferential (oldest) block hash given in a previous \textsc{GetChain} message.
-\end{itemize}
-}\\
-\midrule
-0x16 & \textsc{GetTransactions} & \textsc{Transactions} & $(\text{0x}16)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to send all transactions currently in the queue. See \textsc{Transactions}.}\\
-\bottomrule
-\end{tabular*}
-
-\section{Genesis Block}\label{app:genesis}
-
-The genesis block is 13 items, and is specified thus:
-\begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
-The sequences of both uncles and transactions are empty and represented by $()$.
-$\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
-
-The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$. The latest documentation should be consulted for the value of the state root.
-
 \section{Javascript API}\label{app:jsapi}
@@
 \end{itemize}
-
-\end{document}

### 2014-10-04 Paper.tex
MSIZE is bytes.
Moved out old JS and Wire APIs.

--- 2014-08-15 Paper.tex
@@
 \midrule
-0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{i}$ \\
+0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 
-\section{Wire Protocol}\label{app:wire}
-The wire-protocol specifies a network-level protocol for how two peers can communicate.
-It includes handshake procedures and the means for transferring information such as peers, blocks and transactions.
-Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.
-
-Ethereum nodes may connect to each other over TCP only.
-Peers are free to advertise and accept connections on any port(s) they wish, however, a default port on which the connection may be listened and made will be 30303.
-
-Though TCP provides a connection-oriented medium, Ethereum nodes communicate in terms of packets.
-These packets are formed as a 4-byte synchronisation token (0x22400891), a 4-byte "payload size", to be interpreted as a big-endian integer and finally an N-byte \textbf{RLP-serialised} data structure, where N is the aforementioned "payload size".
-To be clear, the payload size specifies the number of bytes in the packet ''following'' the first 8.
-
-There are a number of different types of message that may be sent.
-This ``type'' is always determined by the first entry of the structure, represented as a scalar.
-The structure of each message type is described below.
-
-\begin{tabular*}{\columnwidth}[h]{rlll}
-\toprule
-\multicolumn{4}{c}{\textbf{00s: Session control}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
-0x00 & \textsc{Hello} & & $(\text{0x}00, v \in \mathbb{P}, n \in \mathbb{P}, \mathbf{i} \in \mathbb{B}, c \in \mathbb{P}, p \in \mathbb{P}, u \in \mathbb{B}_{64})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-This is the first packet sent over the connection, and sent once by both sides. No other messages may be sent until a \textsc{Hello} is received.
-\begin{itemize}
-\item $v$ is the Protocol Version. See the latest documentation for which version is current.
-\item $n$ is the Network Id should be 0.
-\item $\mathbf{i}$ is the Client Id and specifies the client software identity as a human-readable string (e.g. ``Ethereum(++)/1.0.0'').
-\item $c$ is the client's Capabilities and specifies the capabilities of the client as a set of flags; presently three bits are used:
-\begin{description}
-\item[0x01] Client provides peer discovery service;
-\item[0x02] Client provides transaction relaying service;
-\item[0x04] Client provides block-chain querying service.
-\end{description}
-\item $p$ is the Listen Port and specifies the port that the client is listening on (on the interface that the present connection traverses). If 0 it indicates the client is not listening.
-\item $u$ is the Unique Identity of the node and specifies a 512-bit hash that identifies this node.
-\end{itemize}
-}\\
-\midrule
-0x01 & \textsc{Disconnect} && $(\text{0x}01, r \in \mathbb{P})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Inform the peer that a disconnection is imminent; if received, a peer should disconnect immediately.
-When sending, well-behaved hosts give their peers a fighting chance (read: wait 2 seconds) to disconnect to before disconnecting themselves.
-\begin{itemize}
-\item $r$ is an integer specifying one of a number of reasons for disconnect:
-\begin{description}
-\item[0x00] Disconnect requested;
-\item[0x01] TCP sub-system error;
-\item[0x02] Bad protocol;
-\item[0x03] Useless peer;
-\item[0x04] Too many peers;
-\item[0x05] Already connected;
-\item[0x06] Wrong genesis block;
-\item[0x07] Incompatible network protocols;
-\item[0x08] Client quitting.
-\end{description}
-\end{itemize}
-}\\
-\midrule
-0x02 & \textsc{Ping} & \textsc{Pong} & $(\text{0x}02)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Requests an immediate reply of \textsc{Pong} from the peer.}\\
-\midrule
-0x03 & \textsc{Pong} && $(\text{0x}03)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Reply to peer's \textsc{Ping} packet.}\\
-\bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlll}
-\toprule
-\multicolumn{4}{c}{\textbf{10s: Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
-0x10 & \textsc{GetPeers} & \textsc{Peers} & $(\text{0x}10)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to enumerate some known peers for us to connect to. This should include the peer itself.}\\
-\midrule
-0x11 & \textsc{Peers} & & $(\text{0x}11, (a_0 \in \mathbb{B}_4, p_0 \in \mathbb{P}, i_0 \in \mathbb{B}_{64}), (a_1 \in \mathbb{B}_4, p_1 \in \mathbb{P}, i_1 \in \mathbb{B}_{64}), ...)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Specifies a number of known peers. 
-\begin{itemize}
-\item $a_0$, $a_1$, ... is the node's IPv4 address, a 4-byte array that should be interpreted as the IP address $a_0[0]$.$a_0[1]$.$a_0[2]$.$a_0[3]$.
-\item $p_0$, $p_1$, ... is the node's Port and is an integer.
-\item $i_0$, $i_1$, ... is the node's Unique Identifier and is the 512-bit hash that serves to identify the node.
-\end{itemize}
-}\\
-\midrule
-0x12 & \textsc{Transactions} & & $(\text{0x}12, L_T(T_0), L_T(T_1), ...)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-where $L_T$ is the transaction preparation function, as specified in section \ref{ch:block}.
-
-Specify a transaction or transactions that the peer should make sure is included on its transaction queue.
-The items in the list (following the first item 0x12) are transactions in the format described in the main Ethereum specification.
-\begin{itemize}
-\item $T_0$, $T_1$, ... are the transactions that should be assimilated.
-\end{itemize}
-}\\
-\midrule
-0x13 & \textsc{Blocks} && $(\text{0x}13, L_B(b_0), L_B(b_1), ...)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Where $L_B$ is the block preparation function, as specified in section \ref{ch:block}.
-
-Specify a block or blocks that the peer should know about. The items in the list (following the first item, 0x13) are blocks as described in the format described in the main specification.
-\begin{itemize}
-\item $b_0$, $b_1$, ... are the blocks that should be assimilated.
-\end{itemize}
-}\\
-\midrule
-0x14 & \textsc{GetChain} & \textsc{Blocks} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{
-Request the peer to send $c$ blocks in the current canonical block chain that are children of one of a number of given blocks, according to a preferential order with $p_0$ being the most prefered.
-If the designated parent is the present block chain head, an empty reply should be sent.
-If none of the parents are in the current canonical block chain, then a NotInChain message should be sent along with $p_n$, the least preferential parent.
-If no parents are passed, then a reply need not be made.
-\begin{itemize}
-\item $p_0$, $p_1$, ...
-are the SHA3 hashes of the parents of blocks that we should be informed of with a \textsc{Blocks} reply.
-Typically, these will be specified in increasing age (or decreasing block number).
-\item $c$ is the number of children blocks of the most preferred parent that we should be informed of through the corresponding \textsc{Blocks} reply.
-\end{itemize}
-}\\
-\midrule
-0x15 & \textsc{NotInChain} && $(\text{0x}15, p \in \mathbb{B}_{32})$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Inform the peer that a particular block was not found in its block chain.
-\begin{itemize}
-\item $p$ is the SHA3 hash of the block that was not found in the block chain. Typically, this will be the least preferential (oldest) block hash given in a previous \textsc{GetChain} message.
-\end{itemize}
-}\\
-\midrule
-0x16 & \textsc{GetTransactions} & \textsc{Transactions} & $(\text{0x}16)$ \\
-& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to send all transactions currently in the queue. See \textsc{Transactions}.}\\
-\bottomrule
-\end{tabular*}
-
 \section{Genesis Block}\label{app:genesis}
@@
 
-\section{Javascript API}\label{app:jsapi}
-
-The JavaScript API provides a consistent API across multiple scenarios including each of the clients' web-based in-process \DH{}App frameworks and the out-of-process RPC-based infrastructure.
-All key access takes places though the special \texttt{eth} object, part of the global namespace.
-
-\subsection{Values}
-There are no special object types in the API; all values are strings. As strings, values may be of several forms, and are interpreted by the API according to a series of rules:
-
-\begin{enumerate}
-\item If the string contains only digits from 0-9, then it is interpreted as a decimal integer;
-\item if the string begins with the characters \texttt{0x}, then it is interpreted as a hexadecimal integer;
-\item it is interpreted as a binary string otherwise.
-\end{enumerate}
-
-The only exception to this are for parameters that expect a binary string; in this case the string is always interpreted as such.
-
-Values are implicitly converted between integers and hashes/byte-arrays; when this happens, integers are interpreted as big-endian as is standard for Ethereum.
-The following forms are allowed; they are all interpreted in the same way:
-
-\begin{enumerate}
-\item \texttt{"4276803"}
-\item \texttt{"0x414243"}
-\item \texttt{"ABC"}
-\end{enumerate}
-
-In each case, they are interpreted as the number 4276803. The first two values may be alternated between with the additional String methods \texttt{bin()} and \texttt{unbin()}.
-
-As byte arrays, values may be concatenated with the \texttt{+} operator as is normal for strings.
-
-Strings also have a number of additional methods to help with conversion and alignment when switching between addresses, 256-bit integers and free-form byte-arrays for transaction data:
-
-\begin{itemize}
-\item \texttt{bin()}: Converts the string to binary format.
-\item \texttt{pad(l)}: Converts the string to binary format (ready for data parameters) and pads with zeroes until it is of width \texttt{l}.
-Will pad to the left if the original string is numeric, or to the right if binary.
-If \texttt{l} is less than the width of the string, it is resized accordingly.
-\item \texttt{pad(a, b)}: Converts the string to binary format (ready for data parameters) and pads with zeroes on the left size until it is of width \texttt{a}.
-Then pads with zeroes on the right side until it has grown to size \texttt{b}.
-If \texttt{b} is less hat the width of the string, it is resized accordingly.
-\item \texttt{unbin()}: Converts the string from binary format to hex format.
-\item \texttt{unpad()}: Converts the string from binary format to hex format, first removing any zeroes from the right side.
-\item \texttt{dec()}: Converts the string to decimal format (typically from hex).
-\end{itemize}
-
-\subsection{The \texttt{eth} object}
-
-\subsubsection{Properties}
-
-For each such item, there is also an asynchronous method, taking a parameter of the callback function, itself taking a single parameter of the property's return value and of the same name but prefixed with get and recapitalised, e.g.
-\texttt{getCoinbase(fn)}.
-
-\begin{itemize}
-\item \texttt{coinbase} Returns the coinbase address of the client.
-\item \texttt{isListening} Returns true if and only if the client is actively listening for network connections.
-\item \texttt{isMining} Returns true if and only if the client is actively mining new blocks.
-\item \texttt{gasPrice} Returns the client's present price of gas.
-\item \texttt{key} Returns the special key-pair object corresponding to the preferred account owned by the client.
-\item \texttt{keys} Returns a list of the special key-pair objects corresponding to all accounts owned by the client.
-\item \texttt{peerCount} Returns the number of peers currently connected to the client.
-\end{itemize}
-
-\subsubsection{Synchronous Getters}
-For each such item, there is also an asynchronous method, taking an additional parameter of the callback function, itself taking a single parameter of the synchronous method's return value and of the same name but prefixed with get and recapitalised, e.g.
-\texttt{getBalanceAt(a, fn)}.
-
-\begin{itemize}
-\item \texttt{balanceAt(a)} Returns the balance of the account of address given by the address \texttt{a}.
-\item \texttt{storageAt(a, x)} Returns the value in storage at position given by the number x of the account of address given by the address \texttt{a}.
-\item \texttt{txCountAt(a)} Returns the number of transactions send from the account of address given by \texttt{a}.
-\item \texttt{isContractAt(a)} Returns true if the account of address given by \texttt{a} has associated code.
-\end{itemize}
-
-\subsubsection{Transactions}
-
-\begin{itemize}
-\item \texttt{create(sec, xEndowment, bCode, xGas, xGasPrice, fn)} Creates a new contract-creation transaction, given parameters:
-\begin{itemize}
-\item \texttt{sec}, the secret-key for the sender;
-\item \texttt{xEndowment}, the number equal to the account's endowment;
-\item \texttt{bCode}, the binary string (byte array) of EVM-bytecode for the initialisation of the account;
-\item \texttt{xGas}, the number equal to the amount of gas to purchase for the transaction (unused gas is refunded);
-\item \texttt{xGasPrice}, the number equal to the price of gas for this transaction. Returns the special address object representing the new account; and
-\item \texttt{fn}, the callback function, called on completion of the transaction.
-\end{itemize}
-\item \texttt{transact(sec, xValue, aDest, bData, xGas, xGasPrice, fn)} Creates a new message-call transaction, given parameters:
-\begin{itemize}
-\item \texttt{sec}, the secret-key for the sender;
-\item \texttt{xValue}, the value transferred for the transaction (in Wei);
-\item \texttt{aDest}, the address representing the destination address of the message;
-\item \texttt{bData}, the binary string (byte array), containing the associated data of the message;
-\item \texttt{xGas}, the amount of gas to purchase for the transaction (unused gas is refunded);
-\item \texttt{xGasPrice}, the price of gas for this transaction; and
-\item \texttt{fn}, the callback function, called on completion of the transaction.
-\end{itemize}
-\end{itemize}
-
-\subsubsection{Events}
-
-\begin{itemize}
-\item \texttt{watch(a, fn)}: Registers \texttt{fn} as a callback for whenever anything about the state of the account at address \texttt{a} changes, and also on the initial load.
-\item \texttt{watch(a, x, fn)}: Registers \texttt{fn} as a callback for whenever the storage location \texttt{x} of the account at address \texttt{a} changes, and also on the initial load.
-\item \texttt{newBlock(fn)}: Registers \texttt{fn} as a callback for whenever the state changes, and also on the initial load.
-\end{itemize}
-
-\subsubsection{Misc}
-
-\begin{itemize}
-\item \texttt{secretToAddress(a)}: Determines the address from the secret key \texttt{a}.
-\end{itemize}
-
 \end{document}

### 2014-10-04 Wire.tex
Moved out old JS and Wire APIs.

\section{Wire Protocol}\label{app:wire}
The wire-protocol specifies a network-level protocol for how two peers can communicate. It includes handshake procedures and the means for transferring information such as peers, blocks and transactions. Peer-to-peer communications between nodes running Ethereum clients are designed to be governed by a simple wire-protocol making use of existing Ethereum technologies and standards such as RLP wherever practical.

Ethereum nodes may connect to each other over TCP only. Peers are free to advertise and accept connections on any port(s) they wish, however, a default port on which the connection may be listened and made will be 30303.

Though TCP provides a connection-oriented medium, Ethereum nodes communicate in terms of packets. These packets are formed as a 4-byte synchronisation token (0x22400891), a 4-byte "payload size", to be interpreted as a big-endian integer and finally an N-byte \textbf{RLP-serialised} data structure, where N is the aforementioned "payload size". To be clear, the payload size specifies the number of bytes in the packet ''following'' the first 8.

There are a number of different types of message that may be sent. This ``type'' is always determined by the first entry of the structure, represented as a scalar. The structure of each message type is described below.

\begin{tabular*}{\columnwidth}[h]{rlll}
\toprule
\multicolumn{4}{c}{\textbf{00s: Session control}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
0x00 & \textsc{Hello} & & $(\text{0x}00, v \in \mathbb{P}, n \in \mathbb{P}, \mathbf{i} \in \mathbb{B}, c \in \mathbb{P}, p \in \mathbb{P}, u \in \mathbb{B}_{64})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
This is the first packet sent over the connection, and sent once by both sides. No other messages may be sent until a \textsc{Hello} is received.
\begin{itemize}
\item $v$ is the Protocol Version. See the latest documentation for which version is current.
\item $n$ is the Network Id should be 0.
\item $\mathbf{i}$ is the Client Id and specifies the client software identity as a human-readable string (e.g. ``Ethereum(++)/1.0.0'').
\item $c$ is the client's Capabilities and specifies the capabilities of the client as a set of flags; presently three bits are used:
\begin{description}
\item[0x01] Client provides peer discovery service;
\item[0x02] Client provides transaction relaying service;
\item[0x04] Client provides block-chain querying service.
\end{description}
\item $p$ is the Listen Port and specifies the port that the client is listening on (on the interface that the present connection traverses). If 0 it indicates the client is not listening.
\item $u$ is the Unique Identity of the node and specifies a 512-bit hash that identifies this node.
\end{itemize}
}\\
\midrule
0x01 & \textsc{Disconnect} && $(\text{0x}01, r \in \mathbb{P})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Inform the peer that a disconnection is imminent; if received, a peer should disconnect immediately. When sending, well-behaved hosts give their peers a fighting chance (read: wait 2 seconds) to disconnect to before disconnecting themselves.
\begin{itemize}
\item $r$ is an integer specifying one of a number of reasons for disconnect:
\begin{description}
\item[0x00] Disconnect requested;
\item[0x01] TCP sub-system error;
\item[0x02] Bad protocol;
\item[0x03] Useless peer;
\item[0x04] Too many peers;
\item[0x05] Already connected;
\item[0x06] Incompatible network protocols;
\item[0x07] Client quitting.
\end{description}
\end{itemize}
}\\
\midrule
0x02 & \textsc{Ping} & \textsc{Pong} & $(\text{0x}02)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Requests an immediate reply of \textsc{Pong} from the peer.}\\
\midrule
0x03 & \textsc{Pong} && $(\text{0x}03)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Reply to peer's \textsc{Ping} packet.}\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlll}
\toprule
\multicolumn{4}{c}{\textbf{10s: Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & \textbf{Expected Reply} & \textbf{Packet Format} \vspace{5pt} \\
0x10 & \textsc{GetPeers} & \textsc{Peers} & $(\text{0x}10)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to enumerate some known peers for us to connect to. This should include the peer itself.}\\
\midrule
0x11 & \textsc{Peers} & & $(\text{0x}11, (a_0 \in \mathbb{B}_4, p_0 \in \mathbb{P}, i_0 \in \mathbb{B}_{64}), (a_1 \in \mathbb{B}_4, p_1 \in \mathbb{P}, i_1 \in \mathbb{B}_{64}), ...)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Specifies a number of known peers. 
\begin{itemize}
\item $a_0$, $a_1$, ... is the node's IPv4 address, a 4-byte array that should be interpreted as the IP address $a_0[0]$.$a_0[1]$.$a_0[2]$.$a_0[3]$.
\item $p_0$, $p_1$, ... is the node's Port and is an integer.
\item $i_0$, $i_1$, ... is the node's Unique Identifier and is the 512-bit hash that serves to identify the node.
\end{itemize}
}\\
\midrule
0x12 & \textsc{Transactions} & & $(\text{0x}12, L_T(T_0), L_T(T_1), ...)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
where $L_T$ is the transaction preparation function, as specified in section \ref{ch:block}.

Specify a transaction or transactions that the peer should make sure is included on its transaction queue. The items in the list (following the first item 0x12) are transactions in the format described in the main Ethereum specification.
\begin{itemize}
\item $T_0$, $T_1$, ... are the transactions that should be assimilated.
\end{itemize}
}\\
\midrule
0x13 & \textsc{Blocks} && $(\text{0x}13, L_B(b_0), L_B(b_1), ...)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Where $L_B$ is the block preparation function, as specified in section \ref{ch:block}.

Specify a block or blocks that the peer should know about. The items in the list (following the first item, 0x13) are blocks as described in the format described in the main specification.
\begin{itemize}
\item $b_0$, $b_1$, ... are the blocks that should be assimilated.
\end{itemize}
}\\
\midrule
0x14 & \textsc{GetChain} & \textsc{Blocks} or \textsc{NotInChain} & $(\text{0x}14, p_0 \in \mathbb{B}_{32}, p_1 \in \mathbb{B}_{32}, ..., c \in \mathbb{P})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{
Request the peer to send $c$ blocks in the current canonical block chain that are children of one of a number of given blocks, according to a preferential order with $p_0$ being the most prefered. If the designated parent is the present block chain head, an empty reply should be sent. If none of the parents are in the current canonical block chain, then a NotInChain message should be sent along with $p_n$, the least preferential parent. If no parents are passed, then a reply need not be made.
\begin{itemize}
\item $p_0$, $p_1$, ... are the SHA3 hashes of the parents of blocks that we should be informed of with a \textsc{Blocks} reply. Typically, these will be specified in increasing age (or decreasing block number).
\item $c$ is the number of children blocks of the most preferred parent that we should be informed of through the corresponding \textsc{Blocks} reply.
\end{itemize}
}\\
\midrule
0x15 & \textsc{NotInChain} && $(\text{0x}15, p \in \mathbb{B}_{32})$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Inform the peer that a particular block was not found in its block chain.
\begin{itemize}
\item $p$ is the SHA3 hash of the block that was not found in the block chain. Typically, this will be the least preferential (oldest) block hash given in a previous \textsc{GetChain} message.
\end{itemize}
}\\
\midrule
0x16 & \textsc{GetTransactions} & \textsc{Transactions} & $(\text{0x}16)$ \\
& \multicolumn{3}{p{0.8\columnwidth}}{Request the peer to send all transactions currently in the queue. See \textsc{Transactions}.}\\
\bottomrule
\end{tabular*}

### 2014-10-06 Paper.tex
We don't increase the nonce until the address is valid for creation any more.

@@
 \begin{eqnarray}
-A(s, n) & \equiv & a \quad \text{where:} \\
-a & = & \arg \min_x : x \geqslant a' \wedge \boldsymbol{\sigma}[x] = \varnothing \\
-a' & = & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
+A(s, n) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
 \end{eqnarray}
@@
 where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+Note, as defined directly, we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 

### 2014-10-23 Paper.tex
Bytes past the end in a PUSH are 0.

@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[\boldsymbol{\mu}_{pc} + 1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& The bytes are read in line from the program code's bytes array. \\
+&&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
 &&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& where $c$ is defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{b}[(\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& where $c$ is defined as above. \\
+&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
 \bottomrule

### 2014-12-02 Paper.tex
Added DUP* and SWAP*. Prepped LOG* and CALLCODE. Removed DUP & SWAP.

@@
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
@@
 0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
-\midrule
-0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
@@
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{80s: Duplication Operations}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+\midrule
+0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+\midrule
+\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
+\midrule
+0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{90s: Exchange Operations}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+\midrule
+0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+\midrule
+\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
+\midrule
+0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+\bottomrule
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
+\midrule
+0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
+\midrule
+0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
 \bottomrule

### 2014-12-02 Paper.tex
More spec for CALLCODE.

@@
 \midrule
-0xf2 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
+&&&& Same as CALL to this account, except that the code is exchanged for $\boldsymbol{\sigma}[\boldsymbol{\mu}_s[1]]_c$. \\
+\midrule
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\

### 2014-12-02 Paper.tex
EXTCODE*

@@
 &&&& This is gas price specified by the originating transaction.\\
+\midrule
+0x3b & {\small EXTCODESIZE} & 0 & 1 & Get size of an account's code. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0]]_c \rVert$ \\
+\midrule
+0x3c & {\small EXTCODECOPY} & 0 & 4 & Copy an account's code to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0]]_c$ \\
 \bottomrule
@@
 \midrule
-0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
+0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
@@
 \midrule
-0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
+0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
@@
 \midrule
-0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
+0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
@@
 \midrule
-0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
+0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
-0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
 &&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
@@
 \midrule
-0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
@@
 \midrule
-0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
@@
 \midrule
-0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
+0x58 & {\small PC} & 0 & 1 & Get the program counter. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
-0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
+0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
-0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+\midrule
+0x5b & {\small JUMPDEST} & 0 & 1 & Mark a valid destination for jumps. \\
+&&&& TODO \\
 \bottomrule
@@
 0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
+&&&& TODO \\
 \midrule
 0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
+&&&& TODO \\
 \midrule
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
+&&&& TODO \\
 \bottomrule
@@
 &&&& Same as CALL to this account, except that the code is exchanged for $\boldsymbol{\sigma}[\boldsymbol{\mu}_s[1]]_c$. \\
+&&&& TODO \\
 \midrule

### 2014-12-02 Paper.tex
Instruction reorg and extra AL ops.

@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 \midrule
-0x08 & {\small EXP} & 2 & 1 & Exponential operation. \\
+0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+\midrule
+0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+\midrule
+0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
 \midrule
-0x09 & {\small NEG} & 1 & 1 & Negation operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv -\boldsymbol{\mu}_\mathbf{s}[0]$ \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
-\midrule
-0x0a & {\small LT} & 2 & 1 & Less-than comparision. \\
+0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny SIGNEXTEND}(\boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1]) $ \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+\multicolumn{5}{c}{\textbf{10s: Comparison \& Bitwise Logic Operations}} \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[0]$} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x10 & {\small LT} & 2 & 1 & Less-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x0b & {\small GT} & 2 & 1 & Greater-than comparision. \\
+0x11 & {\small GT} & 2 & 1 & Greater-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x0c & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
+0x12 & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x0d & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
+0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x0e & {\small EQ} & 2 & 1 & Equality comparision. \\
+0x14 & {\small EQ} & 2 & 1 & Equality comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x0f & {\small NOT} & 1 & 1 & Simple not operator. \\
+0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-\multicolumn{5}{c}{\textbf{10s: Bitwise Logic Operations}} \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[0]$} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
+\midrule
+0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
-0x11 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
+0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
-0x12 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
+0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
-0x13 & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
+0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
 &&&& $\forall i \in [0..7]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\

### 2014-12-06 Paper.tex
More specific over SHA-3

@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the SHA3 hash function is denoted $\texttt{SHA3}$.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 
@@
 
-The account state comprises the first two, and potentially the last two, of the following fields:
+The account state comprises the following four fields:
 
@@
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small SHA3}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
@@
 
-If the \textbf{codeHash} field is the SHA3 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small SHA3}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+If the \textbf{codeHash} field is the Keccak hash of the empty string, i.e.
+$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{description}
-\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety; formally $H_p$.
-\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block; formally $H_u$.
+\item[parentHash] The Keccak 256-bit hash of the parent block, in its entirety; formally $H_p$.
+\item[unclesHash] The Keccak 256-bit hash of the uncles list portion of this block; formally $H_u$.
 \item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_b$.
-\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsTrie] The SHA3 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsTrie] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
@@
 \begin{array}[t]{lclc}
-H_u &\equiv& \mathtt{\small SHA3}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_u &\equiv& \mathtt{\small Keccak}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i])\})) & \wedge \\
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(B')) = H_p
+P(H) \equiv B': \mathtt{\tiny Keccak}(\mathtt{\tiny RLP}(B')) = H_p
 \end{equation}
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the SHA3 hash of RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 In the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
@@
 \begin{eqnarray}
-A(s, n) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{\tiny SHA3}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
-\end{eqnarray}
-
-where $\mathtt{\tiny SHA3}$ is the SHA3 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+A(s, n) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{\tiny Keccak}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
+\end{eqnarray}
+
+where $\mathtt{\tiny Keccak}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note, as defined directly, we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the SHA3 256-bit hash of the empty string, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
 \begin{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny SHA3}\big(()\big) \big)
+\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny Keccak}\big(()\big) \big)
 \end{equation}
@@
 where $a$ is the address of the new account, as defined above.
-% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny SHA3}(\mathbf{b}), \mathbf{b})$.
+% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny Keccak}(\mathbf{b}), \mathbf{b})$.
 
@@
 
-The account's associated code (identified as the fragment whose SHA3 hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
@@
 I_v & \equiv & v \\
-\text{Let} \quad \mathtt{\tiny SHA3}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny SHA3}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \quad \mathtt{\tiny Keccak}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny Keccak}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 The word size of the machine (and thus size of stack item) is 256-bit.
-This was chosen to facilitate the SHA3-256 hash scheme and elliptic-curve computations.
+This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations.
 The memory model is a simple word-addressed byte array.
@@
 The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small SHA3}.
+Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small Keccak}.
 Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE} and {\small CALL}.
@@
 %\begin{itemize}
-%\item {\small SHA3} costs $G_{sha3}$ gas
+%\item {\small Keccak} costs $G_{sha3}$ gas
 %\item {\small SLOAD} costs $G_{sload}$ gas
@@
 %
-%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small Keccak}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
 
@@
 This is not ASIC resistant and is meant merely as a placeholder.
-It utilises the bare SHA3 hash function to secure the block chain by requiring the SHA3 hash of the concatenation of the nonce and the header's SHA3 hash to be sufficiently low.
+It utilises the bare Keccak hash function to secure the block chain by requiring the Keccak hash of the concatenation of the nonce and the header's Keccak hash to be sufficiently low.
 
@@
 \begin{equation}
-\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny SHA3}(\mathtt{\tiny SHA3}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
+\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny Keccak}(\mathtt{\tiny Keccak}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
 \end{equation}
@@
 $\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
-$\mathtt{\tiny SHA3}$ is the SHA3 hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
+$\mathtt{\tiny Keccak}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
 $n$ is the nonce, a series of 32 bytes;
@@
 $G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
-$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
+$G_{sha3}$ & 20 & Paid for a {\small Keccak} operation. \\
 $G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
@@
 () & \text{if} \quad \mathfrak{I} = \varnothing \\
-\texttt{\small SHA3}(c(\mathfrak{I}, 0)) & \text{otherwise}
+\texttt{\small KEC}(c(\mathfrak{I}, 0)) & \text{otherwise}
 \end{cases}
@@
 When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose SHA3 hash evaluates to our reference.
+As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
 Thus we define in terms of $c$, the node composition function:
@@
 c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{\small SHA3}(c(\mathfrak{I}, i)) & \text{otherwise}
+\texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
 \end{cases}
@@
 \subsection{Trie Database}
-Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the SHA3 hash's collision resistance.
+Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the Keccak hash's collision resistance.
 In reality, a sensible implementation will not fully recompute the trie root hash for each set.
@@
 
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the SHA3 hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny SHA3}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the SHA3 hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small SHA3}( L_S(T) )
+h(T) & \equiv & \mathtt{\small Keccak}( L_S(T) )
 \end{eqnarray}
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny SHA3}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
 \end{equation}
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
+G_{sha3} & \text{if} \quad w = \text{\small Keccak}\\
 G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
@@
 \toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{20s: Keccak}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute SHA3-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny SHA3}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+0x20 & {\small Keccak} & 2 & 1 & Compute Keccak-256 hash. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny SHA3}\big( (42) \big) \big), (), () \big)
+\big( \big( 0_{256}, \mathtt{\tiny Keccak}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny Keccak}\big( (42) \big) \big), (), () \big)
 \end{equation}
@@
 The sequences of both uncles and transactions are empty and represented by $()$.
-$\mathtt{\tiny SHA3}\big( (42) \big)$ refers to the SHA3 hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny SHA3}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
+$\mathtt{\tiny Keccak}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
+$\mathtt{\tiny Keccak}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
 

### 2014-12-06 Paper.tex
No minimum gas price.

@@
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[minGasPrice] A scalar value equal to the minimum price of gas a transaction must have provided in order to be sufficient for inclusion by this miner in this block; formally $H_m$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d,\\ H_i, H_m, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d,\\ H_i, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
@@
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
-H_i \in \mathbb{P} & \wedge & H_m \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
-H_u \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge \\
-H_n \in \mathbb{B}_{32} &&&&
+H_i \in \mathbb{P} & \wedge & H_u \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
+H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge & H_n \in \mathbb{B}_{32}
 \end{array}
@@
 
-The given \textbf{gasUsed} and \textbf{minGasPrice} must correspond faithfully to the transactions listed.
-In the case of ${B_H}_m$, the minimum gas price, all transactions included in the transaction receipt must have a gas price, ${T_0}_p$, that is at least this value:
-\begin{equation}
-\forall \; T \in B_\mathbf{T}: T_p \geq {B_H}_m
-\end{equation}
-
-In the case of ${B_H}_u$, the total gas used in the block, it must be equal to the accumulated gas used according to the final transaction:
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_u$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
 \begin{equation}

### 2014-12-06 Paper.tex
Zero ranges don't increase memory size.

@@
 $G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
+$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
 $G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
+$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
 $G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
-$G_{sha3}$ & 20 & Paid for a {\small Keccak} operation. \\
 $G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
-$G_{sstore}$ & 100 & Paid for a normal {\small SSTORE} operation (doubled or waived sometimes). \\
-$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
+$G_{sset}$ & 300 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 100 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
+$G_{sclear}$ & 0 & Nothing paid for an {\small SSTORE} operation when the storage value is set to zero from non-zero. \\
+$R_{sclear}$ & 150 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
 $G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
 $G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
+$G_{exp}$ & 1 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 1 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdata}$ & 5 & Paid for every byte of data or code for a transaction. \\
+$G_{txdatazero}$ & 1 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 5 & Paid for every non-zero byte of data or code for a transaction. \\
 $G_{transaction}$ & 500 & Paid for every transaction. \\
+$G_{log}$ & 1 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 1 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 1 & Paid for each topic of a {\small LOG} operation. \\
+$G_{copy}$ & 1 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3} & \text{if} \quad w = \text{\small Keccak}\\
+G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
 G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
@@
 where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write; $\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+Such accesses must be for non-zero number of bytes.
+Referencing a zero length range (e.g.
+by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
+$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+
+While defining the instruction set, we defined the memory-expansion for range function, $M$, thus:
+
+\begin{equation}
+M(s, f, l) \equiv \begin{cases}
+s & \text{if} \quad l = 0 \\
+\max(s, \ceil{ (f + l) \div 32 }) & \text{otherwise}
+\end{cases}
+\end{equation}
 
@@
 \toprule
-\multicolumn{5}{c}{\textbf{20s: Keccak}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small Keccak} & 2 & 1 & Compute Keccak-256 hash. \\
+0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$; $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created account, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$; $x=1$ otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ \max(\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule

### 2014-12-06 Paper.tex
Prices *COPY, EXP.

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Proof-of-Concept VI}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - WiP}}}
 \author{
@@
 \end{description}
+
+\section{Recursive Length Prefix}\label{app:rlp}
+This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
+
+We define the set of possible structures $\mathbb{T}$:
+\begin{eqnarray}
+\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+
+Where $\mathbb{Y}$ is the set of bytes.
+Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
+
+We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\end{equation}
+
+If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+
+\begin{itemize}
+\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
+\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
+\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
+\end{itemize}
+
+Formally, we define $R_b$:
+\begin{eqnarray}
+R_b(\mathbf{x}) & \equiv & \begin{cases}
+\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
+(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
+\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\end{cases} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum^{n = 0}_{n < \lVert \mathbf{b} \rVert} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
+\end{eqnarray}
+
+Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+
+If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+
+\begin{itemize}
+\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
+\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\end{itemize}
+
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
+(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
+\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+\end{cases} \\
+s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
+\end{eqnarray}
+
+If RLP is used to encode a scalar, defined only as a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+Thus the RLP of some positive integer $i$ is defined as:
+\begin{equation}
+\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\end{equation}
+
+When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.
+
+There is no specific canonical encoding format for signed or floating-point values.
+
+\section{Hex-Prefix Encoding}\label{app:hexprefix}
+Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array.
+It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
+
+It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
+
+\begin{eqnarray}
+\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
+(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
+\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
+(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
+\text{otherwise}
+\end{cases} \\
+f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \\ 0 & \text{otherwise} \end{cases}
+\end{eqnarray}
+
+Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$.
+The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number.
+All remaining nibbles (now an even number) fit properly into the remaining bytes.
+
+\section{Modified Merkle Patricia Tree}\label{app:trie}
+The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
+It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
+
+Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
+\begin{equation}
+\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
+\end{equation}
+
+When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
+\begin{equation}
+\forall_{I: \in \mathfrak{I}} I \equiv (I_0, I_1)
+\end{equation}
+
+Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
+\begin{eqnarray}
+y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
+\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+\begin{cases}
+\lfloor \mathbf{k}_n[\lfloor i \div 2 \rfloor] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure.
+The empty trie is defined as being a the empty byte sequence, $()$:
+\begin{equation}
+\texttt{\small TRIE}(\mathfrak{I}) \equiv \begin{cases}
+() & \text{if} \quad \mathfrak{I} = \varnothing \\
+\texttt{\small KEC}(c(\mathfrak{I}, 0)) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+We also assume a function $n$, the trie's node cap function.
+When composing a node, we use RLP to encode the structure.
+As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
+Thus we define in terms of $c$, the node composition function:
+\begin{equation}
+n(\mathfrak{I}, i) \equiv \begin{cases}
+() & \text{if} \quad \mathfrak{I} = \varnothing \\
+c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
+\texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair.
+The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
+Unlike a radix tree, in the case of multiple keys shared the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
+Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
+There are three kinds of nodes in the trie:
+\begin{description}
+\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
+\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of nibbles keys and branches as traversed from the root.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
+\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
+The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
+\end{description}
+
+A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
+\begin{equation}
+c(\mathfrak{I}, i) \equiv \begin{cases}
+\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
+(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(x - 1)] = \mathbf{l} \\
+\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
+u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
+v & = & \begin{cases}
+I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
+() & \text{otherwise}
+\end{cases}
+\end{array}
+\end{cases}
+\end{equation}
+
+\subsection{Trie Database}
+Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the Keccak hash's collision resistance.
+In reality, a sensible implementation will not fully recompute the trie root hash for each set.
+
+A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
+This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
+Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
+
+\section{Signing Transactions}\label{app:signing}
+
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer in the range $(0, 2^{256})$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARESTORE}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
+It is assumed that $v$ is the `recovery id', a 2-bit value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
+
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
+\end{cases} \\
+h(T) & \equiv & \mathtt{\small Keccak}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
+gc
+We may then define the sender function $S$ of the transaction as:
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
+\end{equation}
 
@@
 
-\section{Recursive Length Prefix}\label{app:rlp}
-This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
-
-We define the set of possible structures $\mathbb{T}$:
-\begin{eqnarray}
-\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
-\end{eqnarray}
-
-Where $\mathbb{Y}$ is the set of bytes.
-Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
-
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
-
-\begin{itemize}
-\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
-\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
-\end{itemize}
-
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
-\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
-(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
-\end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum^{n = 0}_{n < \lVert \mathbf{b} \rVert} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
-\end{eqnarray}
-
-Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
-
-If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
-
-\begin{itemize}
-\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
-\end{itemize}
-
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
-(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
-\end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
-\end{eqnarray}
-
-If RLP is used to encode a scalar, defined only as a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
-Thus the RLP of some positive integer $i$ is defined as:
-\begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
-\end{equation}
-
-When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.
-
-There is no specific canonical encoding format for signed or floating-point values.
-
-\section{Hex-Prefix Encoding}\label{app:hexprefix}
-Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array.
-It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
-
-It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
-(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
-\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
-(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
-\text{otherwise}
-\end{cases} \\
-f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \\ 0 & \text{otherwise} \end{cases}
-\end{eqnarray}
-
-Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$.
-The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number.
-All remaining nibbles (now an even number) fit properly into the remaining bytes.
-
-\section{Modified Merkle Patricia Tree}\label{app:trie}
-The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
-It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
-
-Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
-\begin{equation}
-\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
-\end{equation}
-
-When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
-\begin{equation}
-\forall_{I: \in \mathfrak{I}} I \equiv (I_0, I_1)
-\end{equation}
-
-Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
-\begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
-\begin{cases}
-\lfloor \mathbf{k}_n[\lfloor i \div 2 \rfloor] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure.
-The empty trie is defined as being a the empty byte sequence, $()$:
-\begin{equation}
-\texttt{\small TRIE}(\mathfrak{I}) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-\texttt{\small KEC}(c(\mathfrak{I}, 0)) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-We also assume a function $n$, the trie's node cap function.
-When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
-Thus we define in terms of $c$, the node composition function:
-\begin{equation}
-n(\mathfrak{I}, i) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair.
-The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
-Unlike a radix tree, in the case of multiple keys shared the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
-Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
-There are three kinds of nodes in the trie:
-\begin{description}
-\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
-\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of nibbles keys and branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
-\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
-The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
-\end{description}
-
-A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
-\begin{equation}
-c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
-(x - 1)] = \mathbf{l} \\
-\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
-v & = & \begin{cases}
-I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
-() & \text{otherwise}
-\end{cases}
-\end{array}
-\end{cases}
-\end{equation}
-
-\subsection{Trie Database}
-Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the Keccak hash's collision resistance.
-In reality, a sensible implementation will not fully recompute the trie root hash for each set.
-
-A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
-This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
-Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
-
-It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer in the range $(0, 2^{256})$ represented as a byte array of length 32 in big-endian form.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARESTORE}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
-It is assumed that $v$ is the `recovery id', a 2-bit value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
-
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
-\begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small Keccak}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}
-gc
-We may then define the sender function $S$ of the transaction as:
-\begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
-\end{equation}
-
-The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
-\begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
-\end{equation}
-
 \section{Virtual Machine Specification}\label{app:vm}
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
+C_\text{\tiny EXP}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small EXP} \\
+G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[2] & \text{if} \quad w = \text{\small CALLDATACOPY} \\
+G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[2] & \text{if} \quad w = \text{\small CODECOPY} \\
+G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[3] & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL}\\
@@
 where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.

### 2014-12-06 Paper.tex
Empty Trie is the SHA3(rlp('')), not empty string.

@@
 
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure.
-The empty trie is defined as being a the empty byte sequence, $()$:
-\begin{equation}
-\texttt{\small TRIE}(\mathfrak{I}) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-\texttt{\small KEC}(c(\mathfrak{I}, 0)) & \text{otherwise}
-\end{cases}
+We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
+\begin{equation}
+\texttt{\small TRIE}(\mathfrak{I}) \equiv \texttt{\small KEC}(c(\mathfrak{I}, 0))
 \end{equation}

### 2014-12-06 Paper.tex
New uncle rewards.
Minor fix.

@@
 The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{8}$ of the block reward and the coinbase of the uncle by $\frac{3}{4}$ of the reward.
+We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle by $\frac{15}{16}$ of the reward.
 Formally we define the function $\Omega$:
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + (1 + \frac{|B_\mathbf{U}|}{8})R_b \\
-\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{3}{4}R_b
-\end{eqnarray}
+\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
+\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{15}{16}R_b
+\end{eqnarray}
+
+If there are collisions of the coinbase addresses between uncles and the block (i.e.
+two uncles with the same coinbase address or an uncle with the same coinbase address as the present block), additions are applied cumulatively.
 

### 2014-12-06 Paper.tex
Modified GHOST: allow multiple gnerations of uncles.

@@
 
-The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:
-\begin{equation}
-\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; P(U) = P(P(B_H)) \; \wedge \; P(B_H) \neq B
+The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of $N$th-generation uncle to the present block where $N \leq 6$.
+Formally:
+\begin{equation}
+\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H), 6)
+\end{equation}
+
+where $k$ denotes the ``is-kin'' property:
+\begin{equation}
+k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\ 
+s(U, H) &\\
+\quad \vee \; u(U, P(H), n - 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+and $s$ denotes the ``is-sibling'' property:
+\begin{equation}
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U)
 \end{equation}

### 2014-12-08 Paper.tex
More explicit about account creation.

@@
 
-We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient:
-\begin{eqnarray}
-\boldsymbol{\sigma}_1 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_1[r]_b & \equiv & \boldsymbol{\sigma}[r]_b + v
-\end{eqnarray}
+We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient.
+\begin{equation}
+\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v
+\end{equation}
+
+Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and nonce.
+Thus the previous equation should be taken to mean:
+\begin{equation}
+\boldsymbol{\sigma}_1 \equiv \boldsymbol{\sigma} \quad \text{except:} \\
+\end{equation}
+\begin{equation}
+\begin{cases}
+\boldsymbol{\sigma}_1[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
+\end{equation}
 
@@
 \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}_1[r]_c = \varnothing \\
 \boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
@@
 In order to validate the extra computation, a given block $B$ may include the block headers from any known uncle blocks (i.e.
-blocks whose parent is equivalent to the grandparent of $B$).
+blocks whose parent is equivalent to the $N$th generation grandparent of $B$, for a limited set of $N$).
 Since a block header includes the nonce, a proof-of-work, then the header alone is enough to validate the computation done.

### 2014-12-16 Paper.tex
Alterations to the transaction execution description; unfinished but better.

--- 2014-12-08 Paper.tex
@@
 
-If the \textbf{codeHash} field is the Keccak hash of the empty string, i.e.
+If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
 $\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
@@
 This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or the zero address for a contract creation transaction; formally $T_t$.
+\item[to] The 160-bit address of the message call's recipient or $\varnothing$ for a contract creation transaction; formally $T_t$.
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
 (T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$ and the address hash $T_\mathbf{t}$:
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
 \begin{equation}
@@
 T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
-T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B} & \wedge \\
-T_t \in \mathbb{B}_{20} &&&&
+T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
+\end{equation}
+
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty string and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_t \in \begin{cases} \mathbb{B}_{0} & \text{if} \quad T_t = \varnothing \\
+\mathbb{B}_{20} & \text{otherwise}\end{cases}
 \end{equation}
@@
 \begin{array}[t]{lclc}
-H_u &\equiv& \mathtt{\small Keccak}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_u &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i])\})) & \wedge \\
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny Keccak}(\mathtt{\tiny RLP}(B')) = H_p
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
 \end{equation}
@@
 \begin{equation}
+%\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
 \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
@@
 \begin{enumerate}
-\item The transaction signature is valid;
+\item The transaction is well-formed RLP, with no additional trailing bytes;
+\item the transaction signature is valid;
 \item the transaction nonce is valid (equivalent to the sender account's current nonce);
@@
 
+\subsection{Substate}
+Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
+We call this \textit{transaction substate}, and represent it as $A$, which is a tuple:
+\begin{equation}
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Though not immediately refunded, it is allowed to partially offset the total execution costs.
+
+\subsection{Execution}
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{equation}
-g_0 \equiv \begin{cases}
-\lVert T_\mathbf{i}\rVert G_{txdata} + G_{transaction} & \text{if} \quad T_t = 0 \\
-\lVert T_\mathbf{d}\rVert G_{txdata} + G_{transaction} & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where $\lVert T_\mathbf{d}\rVert$ and $\lVert T_\mathbf{i}\rVert$ are the sizes, in bytes, of the transaction's associated data and initialisation EVM-code, respectively and $G$ is defined in Appendix \ref{app:fees}.
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
+
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G$ is defined in Appendix \ref{app:fees}.
 
@@
 T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leq & T_g \quad \wedge \\ 
-v_0 & \leq & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_l & \leq & {B_H}_l - \ell(B_\mathbf{R})_u
+g_0 & \leqslant & T_g \quad \wedge \\ 
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_l & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$:
+Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
 
@@
 
-%todo explain suicide list
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and suicide list $\mathbf{s}$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', \mathbf{s}) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}) & \text{if} \quad T_t = 0 \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, T_t, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-where
+where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
 \begin{equation}
@@
 
-After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate.
-The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$.
+After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
+\begin{equation}
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+
+The Ether for the gas is given to the miner, whose address is specified as the coinbase of the present block $B$.
 So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g' T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g') T_p \\
+\boldsymbol{\sigma}^*[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
 m & \equiv & {B_H}_b
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in \mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-And finally specify $\Upsilon^g$, the total gas used in this transaction:
-\begin{equation}
-\Upsilon^g(\boldsymbol{\sigma}, T) \equiv g_0 + g'
-\end{equation}
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & g_0 + g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
 
@@
 
-There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code.
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, additional database entries and remaining gas $(\boldsymbol{\sigma}', g')$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i})
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($d$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, d)
 \end{equation}
@@
 The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
-In the unlikely event that the address is already in use, it is treated as a big-endian integer and incremented by one until an unused address is arrived at.
 Thus we define the creation address function $A$:
 \begin{eqnarray}
-A(s, n) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{\tiny Keccak}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
-\end{eqnarray}
-
-where $\mathtt{\tiny Keccak}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+A(s, n) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
+\end{eqnarray}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note, as defined directly, we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{equation}
-\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny Keccak}\big(()\big) \big)
+\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big)
+\end{equation}
+
+where $v'$ is the account's pre-existing value, in the event it was previously in existance:
+
+\begin{equation}
+v' \equiv 
 \end{equation}
@@
 where $a$ is the address of the new account, as defined above.
-% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny Keccak}(\mathbf{b}), \mathbf{b})$.
+% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
 
@@
 Thus formally, we may specify the resultant state, gas and suicide list as $(\boldsymbol{\sigma}', g', \mathbf{s})$ where:
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
-\end{equation}
-\begin{eqnarray}
+
+\textit{TODO: rollback when $\mathbf{o} = \varnothing$ or is it sigma? clear this up, formally define and normalise into message-call and execution overview.}
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+g^{**} & \text{if} \quad g^{**} < c \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
 \quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}^{**} & \text{if} \quad \boldsymbol{\sigma}^{**}[a] = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad a \notin \mathbf{s} \vee g^{**} < c \\
 \begin{array}[b]{l}
@@
 
+where $c$ is the code-deposit cost:
+\begin{equation}
+c \equiv G_{createdata} \times |\mathbf{o}|
+\end{equation}
+
 $I_\mathbf{d}$ evaluates to the empty tuple.
@@
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the newly created account was suicided or if the gas does not cover the additional per-byte contract deposit fee.
 
@@
 \section{Message Call} \label{ch:call}
-
 In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call.
@@
 \begin{cases}
-\boldsymbol{\sigma}_1[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
 \boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
@@
 \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}_1 & \text{if} \quad \mathbf{o} = \varnothing \\
 \boldsymbol{\sigma}^{**} & \text{otherwise}
@@
 I_v & \equiv & v \\
-\text{Let} \quad \mathtt{\tiny Keccak}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny Keccak}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \quad \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small Keccak}.
+Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small SHA3}.
 Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE} and {\small CALL}.
@@
 
-Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is waived; in fact, it is effectively paid up-front since the initial usage of a storage location costs twice as much as the normal usage.
+Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is not only waived, a qualified refund is given; in fact, this refund is effectively paid up-front since the initial usage of a storage location costs substantially more than normal usage.
 
@@
 %\begin{itemize}
-%\item {\small Keccak} costs $G_{sha3}$ gas
+%\item {\small SHA3} costs $G_{sha3}$ gas
 %\item {\small SLOAD} costs $G_{sload}$ gas
@@
 %
-%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small Keccak}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
 
@@
 \end{equation}
+
+TODO: MUCH MORE DISCUSSION OF THIS
 
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+
+\textit{TODO: JUMP not onto valid JUMPDEST.}
 
@@
 \begin{equation}
-\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny Keccak}(\mathtt{\tiny Keccak}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
+\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny KEC}(\mathtt{\tiny KEC}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
 \end{equation}
@@
 $\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
-$\mathtt{\tiny Keccak}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
+$\mathtt{\tiny KEC}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
 $n$ is the nonce, a series of 32 bytes;
@@
 \begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
 \end{equation}
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small Keccak}( L_S(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
 \end{eqnarray}
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny Keccak}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
 \end{equation}
@@
 $G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
-$G_{sha3}$ & 20 & Paid for a {\small SHA3} operation. \\
 $G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
@@
 $G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
+$G_{createdata}$ & 5 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
 $G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
@@
 $G_{logtopic}$ & 1 & Paid for each topic of a {\small LOG} operation. \\
+$G_{sha3}$ & 10 & Paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 10 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
 $G_{copy}$ & 1 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3} & \text{if} \quad w = \text{\small SHA3}\\
+G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
 G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
@@
 0x5b & {\small JUMPDEST} & 0 & 1 & Mark a valid destination for jumps. \\
-&&&& TODO \\
+&&&& \textit{TODO} \\
 \bottomrule
@@
 0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
-&&&& TODO \\
+&&&& \textit{TODO} \\
 \midrule
 0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& TODO \\
+&&&& \textit{TODO} \\
 \midrule
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& TODO \\
+&&&& \textit{TODO} \\
 \bottomrule
@@
 &&&& Same as CALL to this account, except that the code is exchanged for $\boldsymbol{\sigma}[\boldsymbol{\mu}_s[1]]_c$. \\
-&&&& TODO \\
+&&&& \textit{TODO} \\
 \midrule
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny Keccak}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny Keccak}\big( (42) \big) \big), (), () \big)
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
 \end{equation}
@@
 The sequences of both uncles and transactions are empty and represented by $()$.
-$\mathtt{\tiny Keccak}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny Keccak}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
 

### 2014-12-16 Paper.tex
1024 max call depth.

@@
 
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
+\end{equation}
+
 \subsection{Execution}
@@
 
-There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($d$).
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, d)
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
-Thus we define the creation address function $A$:
-\begin{eqnarray}
-A(s, n) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, n)\;\big)\Big)\Big)
-\end{eqnarray}
+Thus we define the resultant address for the new account $a$:
+\begin{equation}
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
 
 where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note, as defined directly, we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 \begin{equation}
-\boldsymbol{\sigma}^*[A(s, \boldsymbol{\sigma}[s]_n - 1)] \equiv \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big)
-\end{equation}
-
-where $v'$ is the account's pre-existing value, in the event it was previously in existance:
-
-\begin{equation}
-v' \equiv 
-\end{equation}
-
-where $a$ is the address of the new account, as defined above.
-% It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
+\boldsymbol{\sigma}^*[a] \equiv \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big)
+\end{equation}
+
+where $v'$ is the account's pre-existing value, in the event it was previously in existence:
+\begin{equation}
+v' \equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
+\boldsymbol{\sigma}[a]_b & \text{otherwise}
+\end{cases}
+\end{equation}
+
+%It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
 
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$, the suicide list $\mathbf{s}$ and the body code of the account $\mathbf{b}$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
 
 Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this exceptional case we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-The gas remaining should be zero.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
 If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state, gas and suicide list as $(\boldsymbol{\sigma}', g', \mathbf{s})$ where:
-
-\textit{TODO: rollback when $\mathbf{o} = \varnothing$ or is it sigma? clear this up, formally define and normalise into message-call and execution overview.}
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
 \end{equation}
@@
 \quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{if} \quad g^{**} < c \\
@@
 \quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma}^* & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} & \text{if} \quad a \notin \mathbf{s} \vee g^{**} < c \\
@@
 I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i}
+I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_e & \equiv & e
 \end{eqnarray}
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
 $I_H$ has no special treatment and is determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[A(s, \boldsymbol{\sigma}[s]_n)]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
-
-No code is deposited in the state if the newly created account was suicided or if the gas does not cover the additional per-byte contract deposit fee.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee.
 
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call.
-Aside from evaluating to a new state and suicide list, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
+
+Aside from evaluating to a new state and transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
 \begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d})
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d}, e)
 \end{equation}
@@
 The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts due to an exhausted gas supply, then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution.
+Just as with contract creation, if the execution halts due in an exceptional fashion (i.e.
+due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution (i.e.
+$\boldsymbol{\sigma}_1$).
+
 \begin{eqnarray}
@@
 I_p & \equiv & p \\
+I_d & \equiv & d \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
@@
 I_v & \equiv & v \\
+I_e & \equiv & e \\
 \text{Let} \quad \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
@@
 \item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
@@
 This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \varnothing, I)\big) \\
+
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \begin{equation}
-X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \big) \equiv \begin{cases}
-\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, \varnothing, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I)\big) & \text{otherwise}\\
+X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
 \end{equation}
-
-TODO: MUCH MORE DISCUSSION OF THIS
 
@@
 
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+
+\subsubsection{Machine State}
 The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
@@
 \mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+w = \text{\small JUMP} \wedge Z_\text{JUMP}(\boldsymbol{\mu}) \quad \vee \\
+w = \text{\small JUMPI} \wedge Z_\text{JUMPI}(\boldsymbol{\mu}) \quad \vee \\
+( w \in \{ \text{\small CALL}, \text{\small CALLCODE}, \text{\small CREATE} \} \quad \wedge \\
+\quad I_e = 1024 )
 \end{array}
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined) or if there are insufficient stack items.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid ($Z_\text{JUMP}$/$Z_\text{JUMPI}$ are the test functions, defined in the instruction set appendix) or if a {\small CALL}, {\small CALLCODE} or {\small CREATE} instruction is executed when the call stack limit of 1024 is reached.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-
-\textit{TODO: JUMP not onto valid JUMPDEST.}
 
@@
 \boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
-\forall_{U \in B_\mathbf{U}} \quad \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{15}{16}R_b
+\forall_{U \in B_\mathbf{U}} \; \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{15}{16}R_b
 \end{eqnarray}

### 2014-12-16 Paper.tex
Refunds for SSTORE.

@@
 \begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, \mathbf{s}, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', \mathbf{s}', I) \\
+O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
@@
 \boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
-\mathbf{s}' & \equiv & \mathbf{s} \\
+A' & \equiv & A \\
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
@@
 \begin{equation}
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', I) \quad \text{with exceptions, as noted}
-\end{equation}
-
-Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \quad \text{with exceptions, as noted}
+\end{equation}
+
+Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$, $Z$ and $C$.
 For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
@@
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-2G_{sstore} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
-G_{sstore} & \text{otherwise}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+G_{sreset} & \text{otherwise}
+\end{cases}$ \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+0 & \text{otherwise}
 \end{cases}$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\

### 2014-12-17 Paper.tex
Jump destinations defined.

@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-w = \text{\small JUMP} \wedge Z_\text{JUMP}(\boldsymbol{\mu}) \quad \vee \\
-w = \text{\small JUMPI} \wedge Z_\text{JUMPI}(\boldsymbol{\mu}) \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w \in \{ \text{\small CALL}, \text{\small CALLCODE}, \text{\small CREATE} \} \quad \wedge \\
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid ($Z_\text{JUMP}$/$Z_\text{JUMPI}$ are the test functions, defined in the instruction set appendix) or if a {\small CALL}, {\small CALLCODE} or {\small CREATE} instruction is executed when the call stack limit of 1024 is reached.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or if a {\small CALL}, {\small CALLCODE} or {\small CREATE} instruction is executed when the call stack limit of 1024 is reached.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+
+\subsubsection{Jump Destination Validity}
+
+We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run. We define this as:
+
+\begin{itemize}
+\item Any position in the code occupied by a {\small JUMPDEST} instruction.
+\item Any position $p$ in the code such that there exists a {\small PUSH} of the value $p$ immediately followed by a {\small JUMP} or {\small JUMPI}.
+\end{itemize}
+
+All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined {\small STOP} operations that trail it).
+
+Formally:
+\begin{equation}
+D(\mathbf{c}) \equiv (D_J(\mathbf{c, 0}) \cup D_P(\mathbf{c, 0})) \cap Y(\mathbf{c}, 0)
+\end{equation}
+
+where:
+\begin{equation}
+Y(\mathbf{c}, i) \equiv \begin{cases}
+\{\} & \text{if} \quad i \geqslant |\mathbf{c}| \\
+\{ i \} \cup Y(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\end{cases}
+\end{equation}
+\begin{equation}
+D_J(\mathbf{c}, i) \equiv \begin{cases}
+\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
+\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\end{cases}
+\end{equation}
+\begin{equation}
+D_P(\mathbf{c}, i) \equiv \begin{cases}
+\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
+\{ p \} \cup D_P(\mathbf{c}, n) & \mathbf{c}[i] \in [\text{\small PUSH1}, \text{\small PUSH32}] \quad \wedge \\ & \mathbf{c}[n] \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \\
+D_P(\mathbf{c}, n) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+where, for the purposes of the final equation, $n$ is the next instruction index and $p$ is the {\small PUSH}$x$ instruction's associated data (the following $x$ bytes interpreted as a big endian unsigned):
+\begin{equation}
+n \equiv N(i, \mathbf{c}[i]) \\
+\end{equation}
+\begin{equation}
+p \equiv \begin{cases}
+\mathbf{c}[i + 1] & \text{if}\quad \mathbf{c}[i] = \text{\small PUSH1} \\
+2^8\mathbf{c}[i + 1] + \mathbf{c}[i + 2] & \text{if}\quad \mathbf{c}[i] = \text{\small PUSH2} \\
+\quad\quad\vdots & \quad\quad\vdots \\
+2^{240}\mathbf{c}[i + 1] + ... + \mathbf{c}[i + 32] & \text{if}\quad \mathbf{c}[i] = \text{\small PUSH32} \\
+\end{cases}
+\end{equation}
+
+and $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
+\begin{equation}
+N(i, w) \equiv \begin{cases}
+i + w - \text{\small PUSH1} + 2 & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
+i + 1 & \text{otherwise} \end{cases}
+\end{equation}
 
@@
 J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-\boldsymbol{\mu}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
-\boldsymbol{\mu}_{pc} + 1 & \text{otherwise}
+N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
 \end{cases}
 \end{eqnarray}
-
-where $p$ is the byte size of the push instruction, defined as:
-\begin{equation}
-p \equiv w - \text{\small PUSH1} + 2
-\end{equation}
 
@@
 
-Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$, $Z$ and $C$.
+Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
 For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
@@
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& $Z_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) $ \\
 \midrule
@@
 \midrule
-0x5b & {\small JUMPDEST} & 0 & 1 & Mark a valid destination for jumps. \\
-&&&& \textit{TODO} \\
+0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
+&&&& This operation has no effect on machine state during execution. \\
 \bottomrule

### 2014-12-17 Paper.tex
Usage of substate in the VM spec.
Call code.

@@
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, g, p, v, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \mathbf{d}, e)
 \end{equation}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[r]_c$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts due in an exceptional fashion (i.e.
@@
 I_e & \equiv & e \\
-\text{Let} \quad \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[r]_c
+\text{Let} \quad \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
 \end{eqnarray}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, \mathbf{s}^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \mathbf{s}^+$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', \mathbf{s}^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& $\mathbf{s}' \equiv \mathbf{s} \cup \mathbf{s}^+$ \\
+&&&& $A' \equiv A \Cup A^+$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
@@
 0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
-&&&& Same as CALL to this account, except that the code is exchanged for $\boldsymbol{\sigma}[\boldsymbol{\mu}_s[1]]_c$. \\
-&&&& \textit{TODO} \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule

### 2014-12-17 Paper.tex
Additional fields in block.

@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{R}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
 The block header contains several pieces of information:
 
+\textit{TODO: Introduce logs}
+
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block, in its entirety; formally $H_p$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
 \item[unclesHash] The Keccak 256-bit hash of the uncles list portion of this block; formally $H_u$.
-\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_b$.
+\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsTrie] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_b, H_r, H_t, H_d,\\ H_i, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_t, H_e, H_b, H_t, H_d,\\ H_i, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_u \in \mathbb{B}_{32} & \wedge & H_b \in \mathbb{B}_{20} & \wedge \\
+H_p \in \mathbb{B}_{32} & \wedge & H_u \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
@@
 2^{22} & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 42\\
+{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
 {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{otherwise}\\
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_b]_b & = & \boldsymbol{\sigma}[{B_H}_b]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
-\forall_{U \in B_\mathbf{U}} \; \boldsymbol{\sigma}'[U_b]_b & = & \boldsymbol{\sigma}[U_b]_b + \frac{15}{16}R_b
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
+\forall_{U \in B_\mathbf{U}} \; \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + \frac{15}{16}R_b
 \end{eqnarray}
@@
 
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature a mutable data structure.
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 

### 2014-12-17 Paper.tex
Introduction of receipts.

@@
 
-\textit{TODO: Introduce logs}
+%\textit{TODO: Introduce logs}
 
@@
 
-The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transaction \textit{receipts}. Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{R}, B_\mathbf{U})
+The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 
-The transaction receipt is a tuple of three items comprising the transaction, $R_T$, together with the post-transaction state, $R_{\boldsymbol{\sigma}}$, and the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$:
-\begin{equation}
-R \equiv (R_T, R_{\boldsymbol{\sigma}}, R_u)
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (L_T(R_T), \mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u)
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the second item.
-
-We assert $R_u$, the cumulative gas used is a positive integer:
-\begin{equation}
-R_u \in \mathbb{P}
-\end{equation}
-
-For convenience, I also name $B_\mathbf{T}$ the series of transactions encoded by the transaction receipts $B_\mathbf{R}$:
-\begin{equation}
-B_\mathbf{T} \equiv (B_\mathbf{R}[0]_T, B_\mathbf{R}[1]_T, ...)
-\end{equation}
-
-Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 64 bytes:
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{32}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+\end{equation}
 
@@
 
-We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}

### 2014-12-18 Paper.tex
Specify Blooms formally.

@@
 
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 64 bytes:
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{32}
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 512 bits (64 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{64}
 \end{equation}
@@
 \end{equation}
+
+We define the Bloom filter function, $M$, to reduce a log entry include a single 64-byte hash:
+\begin{equation}
+M(O) \equiv h: m(O_a) \; \vee \; \bigvee_{t \in O_\mathbf{t}} ( M_{3:512}(t) )
+\end{equation}
+
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 512, given an arbitrary byte series.
+It does this through taking the low-order 9 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
+\begin{eqnarray}
+M_{3:512}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{64} \quad \text{where:}\\
+\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
+\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
+m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 512
+\end{eqnarray}
+
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 

### 2014-12-18 Paper.tex
Additional cleanup of Blooms.
Formal definition of LOG* instructions.
Transaction validity strictness.

@@
 \begin{equation}
-M(O) \equiv h: m(O_a) \; \vee \; \bigvee_{t \in O_\mathbf{t}} ( M_{3:512}(t) )
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:512}(t) \big)
 \end{equation}
@@
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARESTORE}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
-It is assumed that $v$ is the `recovery id', a 2-bit value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
+
+We declare that a signature is invalid unless the following are true:
+\begin{eqnarray}
+v &<& 27\\
+v &>& 28\\
+r &>=& \mathtt{\tiny secp256k1n}\\
+s &>=& \mathtt{\tiny secp256k1p}\\
+\end{eqnarray}
+
+where:
+\begin{eqnarray}
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337\\
+\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 \begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except} \\
+G(T, p_r) \equiv T \quad \text{except:} \\
 (T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
-gc
+
 We may then define the sender function $S$ of the transaction as:
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARESTORE}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
 \end{equation}
@@
 \multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
+\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
 0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
-&&&& \textit{TODO} \\
+&&&& $\mathbf{t} \equiv ()$ \\
 \midrule
 0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& \textit{TODO} \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
+\midrule
+\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
 \midrule
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& \textit{TODO} \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
 \bottomrule

### 2014-12-18 Paper.tex
corrected refund from 150 to 100
refund gas.

@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \Xi(\boldsymbol{\sigma}_1, g, I) \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 3 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & a \\
@@
 I_e & \equiv & e \\
-\text{Let} \quad \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
 \end{eqnarray}
@@
 It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are three exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are three so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The three contracts in addresses 1, 2 and 3 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme and the RIPEMD 160-bit hash scheme respectively.
+
+Their full formal definition is in Appendix \ref{app:precompiled}.
 
@@
 
+\section{Precompiled Contracts}\label{app:precompiled}
+
 \section{Signing Transactions}\label{app:signing}
@@
 $G_{sclear}$ & 0 & Nothing paid for an {\small SSTORE} operation when the storage value is set to zero from non-zero. \\
-$R_{sclear}$ & 150 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{sclear}$ & 100 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
 $G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
@@
 In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
-
 \begin{equation}

### 2014-12-19 Paper.tex
Premined contract functions.

@@
 \section{Precompiled Contracts}\label{app:precompiled}
+
+For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
+
+\begin{eqnarray}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) & \equiv & \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases} \\
+\end{eqnarray}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+
+For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
+
+\begin{eqnarray}
+\Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
+g_r &=& 500\\
+|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+\text{if} \quad |\mathbf{o}| = 32: &&\\
+\mathbf{o}[0..11] &=& 0 \\
+\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
+\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
+h &=& \mathbf{d}[0..31]\\
+v &=& \mathbf{d}[32..63]\\
+r &=& \mathbf{d}[64..95]\\
+s &=& \mathbf{d}[96..127]
+\end{eqnarray}
+
+The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
+Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
+
+\begin{eqnarray}
+\Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
+g_r &=& 50 + 50\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+\Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
+g_r &=& 50 + 50\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..11] &=& 0 \\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\end{eqnarray}
+
+For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
+
+\begin{eqnarray}
+\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
+\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\end{eqnarray}
 

### 2014-12-19 Paper.tex
Minor changes. Fixed #20.

@@
 
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-piling to.
+\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-compiling to.
 
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum^{n = 0}_{n < \lVert \mathbf{b} \rVert} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert - 1} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)

### 2014-12-19 Paper.tex
Receipts formally defined.

@@
 This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or $\varnothing$ for a contract creation transaction; formally $T_t$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty string and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_t \in \begin{cases} \mathbb{B}_{0} & \text{if} \quad T_t = \varnothing \\
-\mathbb{B}_{20} & \text{otherwise}\end{cases}
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_t \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+\mathbb{B}_{0} & \text{otherwise}\end{cases}
 \end{equation}
@@
 
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
 The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
@@
 A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-
 \begin{equation}
@@
 \begin{array}[t]{lclc}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \\
 H_u &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i])\})) & \wedge \\
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) &
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
@@
 Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_t, H_e, H_b, H_t, H_d,\\ H_i, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_R^*(B_\mathbf{R}), L_H^*(B_\mathbf{U}) \big)
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
 \end{eqnarray}
@@
 
-Thus $\boldsymbol{\sigma}'$ is the post-transactional state. We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, to be defined later.
+Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
 
@@
 \boldsymbol{\sigma}^* & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad a \notin \mathbf{s} \vee g^{**} < c \\
-\begin{array}[b]{l}
-\boldsymbol{\sigma}^{**} \quad \text{except:} \\
-\boldsymbol{\sigma}'[a]_{\mathbf{b}} = \mathbf{o}
-\end{array} & \text{otherwise}
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_{\mathbf{b}} = \mathbf{o} & \text{otherwise}
 \end{cases} \\
@@
 
-\subsection{State \& Nonce Validation}
+\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
 
@@
 
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states and cumulative gas used after each transaction.
+As previously detailed, $B_\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $B_\mathbf{R}[n]_\mathbf{l}$ and $B_\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($B_\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $\mathbf{g}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
  & \text{otherwise} \end{cases}
+\end{equation}
+
+For $B_\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+B_\mathbf{R}[n]_u = 
+\Upsilon^\mathbf{l}(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
 \end{equation}

### 2014-12-27 Paper.tex
Feature complete.
Typos and suggestions

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - WiP}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
@@
 Moreover it aims to provide to the end-developer a tightly integrated end-to-end system for building software on a hitherto unexplored compute paradigm in the mainstream: a trustful object messaging compute framework.
+
+\subsection{Structure of the paper} \label{ch:structure}
+We start with...
+Then go on with ...
+Please note the appendix contains definitions of frequently used terms, ....
 
@@
 
-The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
+The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as Recursive Length Prefix Encoding or short-hand RLP, see Appendix \ref{app:rlp}).
 Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
@@
 \end{equation}
+% What's the indecx I, what's the function L?
+The notation is confusing.
+Put the \equiv in quotation marks...
+or say by an abuse of notation, cause it's technically not an quiivalence.
+Also, what's TRIE?
+First time use.
 

### 2014-12-27 Paper.tex
Some more

@@
 First time use.
-
+%What is L? Is this a notation for a general transfromation?
 The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
@@
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+Note the deviating index $\cdot_w$ instead of $\cdot_v$ which is already in use for the value, see previous definition above.
 This is expanded in Appendix \ref{app:signing}.
 \end{description}
-
+% Why do you change notation here to \mathbf cor the 
 Additionally, a contract creation transaction contains:
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
+The address hash $T_\mathbf{t}$ 
+% no $\mathbf$?
+is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
 \begin{equation}
@@
 \begin{equation}
+\lable{eq:PoW}
 %\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
@@
 
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+Identity \eqref{eq:PoW} is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
+Because the nonce must satisfy requirement \eqref{eq:PoW}, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
 
@@
 \end{eqnarray}
-
+%Could you remind of the definiton of I or introduce it.
 where $c$ is the code-deposit cost:
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $\cdot_d$ dentotes a block's difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
 
@@
 
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-compiling to.
-
 \end{description}
@@
 
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+We define the Recursive Lentth Prefix (RLP) function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
 \begin{equation}
@@
 The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
 
@@
 
-The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
+The two hash functions RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
@@
 \begin{eqnarray}
-v &<& 27\\
-v &>& 28\\
+v &<& 27 \text{ or } v &>& 28\\
 r &>=& \mathtt{\tiny secp256k1n}\\

### 2015-01-02 Paper.tex

--- 2014-12-19 Paper.tex
@@
 
-We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run. We define this as:
-
-\begin{itemize}
-\item Any position in the code occupied by a {\small JUMPDEST} instruction.
-\item Any position $p$ in the code such that there exists a {\small PUSH} of the value $p$ immediately followed by a {\small JUMP} or {\small JUMPI}.
-\end{itemize}
+We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run.
+We define this as any position in the code occupied by a {\small JUMPDEST} instruction.
 
@@
 \begin{equation}
-D(\mathbf{c}) \equiv (D_J(\mathbf{c, 0}) \cup D_P(\mathbf{c, 0})) \cap Y(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_J(\mathbf{c, 0}) \cap Y(\mathbf{c}, 0)
 \end{equation}
@@
 \end{equation}
-\begin{equation}
-D_P(\mathbf{c}, i) \equiv \begin{cases}
-\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ p \} \cup D_P(\mathbf{c}, n) & \mathbf{c}[i] \in [\text{\small PUSH1}, \text{\small PUSH32}] \quad \wedge \\ & \mathbf{c}[n] \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \\
-D_P(\mathbf{c}, n) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where, for the purposes of the final equation, $n$ is the next instruction index and $p$ is the {\small PUSH}$x$ instruction's associated data (the following $x$ bytes interpreted as a big endian unsigned):
-\begin{equation}
-n \equiv N(i, \mathbf{c}[i]) \\
-\end{equation}
-\begin{equation}
-p \equiv \begin{cases}
-\mathbf{c}[i + 1] & \text{if}\quad \mathbf{c}[i] = \text{\small PUSH1} \\
-2^8\mathbf{c}[i + 1] + \mathbf{c}[i + 2] & \text{if}\quad \mathbf{c}[i] = \text{\small PUSH2} \\
-\quad\quad\vdots & \quad\quad\vdots \\
-2^{240}\mathbf{c}[i + 1] + ... + \mathbf{c}[i + 32] & \text{if}\quad \mathbf{c}[i] = \text{\small PUSH32} \\
-\end{cases}
-\end{equation}
-
-and $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
+
+where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
 \begin{equation}

### 2015-01-02 Paper.tex

@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-C_\text{\tiny EXP}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small EXP} \\
+G_{exp} + G_{expbyte}\times\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1]) & \text{if} \quad w = \text{\small EXP} \\
 G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[2] & \text{if} \quad w = \text{\small CALLDATACOPY} \\

### 2015-01-03 Paper.tex
New PREVHASH.

--- 2015-01-02 Paper.tex
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_p$ \\
-&&&& $I_{H_p}$ is the previous block's hash. \\
+0x40 & {\small BLOCKHASH} & 10 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv A(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $A$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& For a number greater than or equal to, or less than 256 below, the present block's\\
+&&&& number, 0 is left on the stack.\\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_n \vee a = 256 \vee h = 0 \\ H_h & \text{if} \quad n = H_n \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined from the hash given: $h = H_h$ \\
 \midrule

### 2015-01-03 Paper.tex
renamed gasUsed abbreviation in Block header
clearer definition of transaction

@@
 
-A transaction (formally, $T$) is a single cryptographically-signed instruction sent by an actor external to Ethereum.
-An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
+A transaction (formally, $T$) is a single cryptographically-signed instruction sent by a node in the Ethereum network.
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').

### 2015-01-05 Paper.tex
present price of ether is unknown
clearification of the stateRoot

--- 2015-01-03 Paper.tex
@@
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[stateRoot] A 256-bit hash of the root node of a trie structure that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from a byte array of size 32 to an RLP-encoded integer.
+\item[stateRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the 256-bit integer keys to the RLP-encoded 256-bit integer values.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.

### 2015-01-05 Paper.tex
typo blockhash input parameters

--- 2015-01-03 Paper.tex
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 10 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv A(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\

### 2015-01-05 Paper.tex
Fourth precompiled contract -> identity function.

--- 2015-01-03 Paper.tex
@@
 
-A transaction (formally, $T$) is a single cryptographically-signed instruction sent by a node in the Ethereum network.
+A transaction (formally, $T$) is a single cryptographically-signed instruction sent by an actor external to Ethereum.
+An external actor can be a person (via a mobile device or desktop computer) or could be from a piece of automated software running on a server.
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
@@
 \Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 4 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\end{eqnarray}
+
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+
+\begin{eqnarray}
+\Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
+g_r &=& 1 + \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
 \end{eqnarray}

### 2015-01-05 Paper.tex
CALL/CREATE return 0 on call depth limit reached (rather than bombing out the caller with OOG).

@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w \in \{ \text{\small CALL}, \text{\small CALLCODE}, \text{\small CREATE} \} \quad \wedge \\
-\quad I_e = 1024 )
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad
 \end{array}
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$; $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly created account, otherwise. \\
+&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$; $x=1$ otherwise. \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\

### 2015-01-05 Paper.tex
Fix for EXP.

--- 2015-01-05 Paper.tex
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-C_\text{\tiny EXP}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small EXP} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
 G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[2] & \text{if} \quad w = \text{\small CALLDATACOPY} \\

### 2015-01-05 Paper.tex
Fix gas used, B_R->R as receipts are outside the block.

@@
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_u$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_t, H_e, H_b, H_t, H_d,\\ H_i, H_l, H_u, H_s, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_t, H_e, H_b, H_t, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
@@
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
-H_i \in \mathbb{P} & \wedge & H_u \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
+H_i \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
 H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge & H_n \in \mathbb{B}_{32}
@@
 \begin{equation}
-{B_H}_u = \ell(\mathbf{R})_u
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 
-As previously detailed, $B_\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $B_\mathbf{R}[n]_\mathbf{l}$ and $B_\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($B_\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-B_\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{equation}
-B_\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + B_\mathbf{R}[n-1]_u
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $B_\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-B_\mathbf{R}[n]_u = 
-\Upsilon^\mathbf{l}(B_\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_u = 
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
 \end{equation}
@@
 \begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(B_\mathbf{R})_{\boldsymbol{\sigma}})
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}

### 2015-01-05 Paper.tex
Fixes #30.

@@
 
-A transaction (formally, $T$) is a single cryptographically-signed instruction sent by a node in the Ethereum network.
+A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
+It is assumed that the external actor will be human in nature, though software tools will be used in its construction and dissemination.
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').

### 2015-01-05 Paper.tex

@@
 This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
-Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds since at the present price of Ether and default gas price, that would cost around US\$20M for the memory fee alone.
+Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds.
 That said, implementations must be able to manage this eventuality.

### 2015-01-05 Paper.tex

@@
 
+\subsection{Structure of the paper} \label{ch:structure}
+We start with...
+Then go on with ...
+Please note the appendix contains definitions of frequently used terms, ....
+
 \subsection{Driving Factors} \label{ch:driving}
@@
 \end{equation}
-
+% What's the indecx I, what's the function L?
+The notation is confusing.
+Put the \equiv in quotation marks...
+or say by an abuse of notation, cause it's technically not an quiivalence.
+Also, what's TRIE?
+First time use.
+%What is L? Is this a notation for a general transfromation?
 The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:

### 2015-01-05 Paper.tex
Fix for new depth semantics.
Clarification/correction for sigs.

--- 2015-01-05 Paper.tex
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; unlike some implementations which keep this value in the range of $[27, 30]$, we subtract the lower bound thus reducing it to a pure 2-bit value in the range $[0, 3]$.
-
-We declare that a signature is invalid unless the following are true:
-\begin{eqnarray}
-v &<& 27\\
-v &>& 28\\
-r &>=& \mathtt{\tiny secp256k1n}\\
-s &>=& \mathtt{\tiny secp256k1p}\\
-\end{eqnarray}
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{equation}
+v \in \{27,28\} \quad\wedge\quad r \geqslant \mathtt{\tiny secp256k1n} \quad\wedge\quad s \geqslant \mathtt{\tiny secp256k1p}
+\end{equation}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[1],\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, \boldsymbol{\mu}_\mathbf{s}[1],\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\

### 2015-01-06 Paper.tex
typo output parameter

@@
 \midrule
-0x3b & {\small EXTCODESIZE} & 0 & 1 & Get size of an account's code. \\
+0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0]]_c \rVert$ \\
 \midrule
-0x3c & {\small EXTCODECOPY} & 0 & 4 & Copy an account's code to memory. \\
+0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv

### 2015-01-08 Paper.tex
Definition of SIGNEXTEND
Clarification to use two's complement for signed numbers.

@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as signed 256-bit integers for the purposes of this operation. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule

### 2015-01-19 Paper.tex
Minor fix to SIGNEXTEND.
fix mulmod addmod mod 0
typo in block gas limit calculation
rename stateRoot to storageRoot

@@
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[stateRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the 256-bit integer keys to the RLP-encoded 256-bit integer values.
+\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the 256-bit integer keys to the RLP-encoded 256-bit integer values.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny SIGNEXTEND}(\boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1]) $ \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv
+\begin{cases}
+    \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where } t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1)$\\
+    \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise}
+\end{cases} 
 \end{tabular*}

### 2015-01-21 Paper.tex
replace timestamp of transaction example
block header serialisation
complete definition of block header component types
changed block time from 8 to 12 seconds
extra data is allowed to have 1024 bytes or fewer
All addresses are to be interpreted as mod 2^160.
Exponential memory cost.

--- 2015-01-08 Paper.tex
@@
 \end{cases}\\
-L'(H) & \equiv & \big\lfloor \frac{1023 {P(H)_H}_l + \lfloor \frac{6}{5}{P(H)_H}_u \rfloor}{1024}\big\rfloor
+L'(H) & \equiv & \big\lfloor \frac{1023 {P(H)_H}_l + \lfloor \frac{6}{5}{P(H)_H}_g \rfloor}{1024}\big\rfloor
 \end{eqnarray}
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-
-\begin{eqnarray}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) & \equiv & \begin{cases}
+\begin{equation}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
 (\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases} \\
-\end{eqnarray}
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
 
@@
 Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
-
 \begin{eqnarray}
@@
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
-
 \begin{eqnarray}
@@
 For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
-
 \begin{eqnarray}
@@
 Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
-
 \begin{eqnarray}
@@
 
-\begin{eqnarray}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) & \equiv & G_{memory}(\boldsymbol{\mu}'_i - \boldsymbol{\mu}_i) + \begin{cases}
+\begin{equation}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{memory}(\boldsymbol{\mu}'_i)-C_{memory}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 G_{step} & \text{otherwise}
-\end{cases} \\
-w & \equiv & \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
 \end{cases}
-\end{eqnarray}
-
-where $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
+\end{equation}
+
+where:
+\begin{equation}
+C_{memory}(a) \equiv G_{memory}(a + \Big\lfloor \dfrac{a^2}{1024} \Big\rfloor)
+\end{equation}
+
+and $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
+
 Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
+
 Referencing a zero length range (e.g.
@@
 $\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+
+Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 32KB of memory used, after which it costs substantially more.
 
@@
 For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
+
+TODO: Whenever a stack item is used as an address, it should be assumed it is the stack item modulo $2^{160}$.
 
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny SIGNEXTEND}(\boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1]) $ \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$
 \end{tabular*}
@@
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{when} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{when} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0]]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0]]_c$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 \bottomrule
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[1],\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, t, t,\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 &&&& $A' \equiv A \Cup A^+$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, \boldsymbol{\mu}_\mathbf{s}[1],\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0]]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
 &&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\\bottomrule

### 2015-01-22 Paper.tex
restricted transaciton parameters to u256
double mentioning if to-address in post provisional creation state
defintion of T_o was missing

@@
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
 This is formally defined in section \ref{ch:block}.
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge \\
-T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
-T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_5 & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_{256} & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
+\end{equation}
+where
+\begin{equation}
+\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}

### 2015-01-22 Paper.tex

@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
 \midrule

### 2015-01-22 Paper.tex

@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_s$, denotes the timestamp of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. With the exception of the genesis block, this must be 32 bytes or fewer; formally $H_x$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 1024 bytes or fewer; formally $H_x$.
 \item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_t, H_e, H_b, H_t, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
@@
 H_p \in \mathbb{B}_{32} & \wedge & H_u \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_d \in \mathbb{P} & \wedge \\
-H_i \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_l \in \mathbb{P} & \wedge \\
-H_s \in \mathbb{P} & \wedge & H_x \in \mathbb{B} & \wedge & H_n \in \mathbb{B}_{32}
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
+H_b \in \mathbb{B}_{64} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_n \in \mathbb{B}_{32}
 \end{array}
@@
 & & H_s > {P(H)_H}_s \quad \wedge \\
-& & \lVert H_x \rVert < 1024
+& & \lVert H_x \rVert \le 1024
 \end{eqnarray}

### 2015-01-22 Paper.tex
Fix BYTE def.

--- 2015-01-22 Paper.tex
@@
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$.
 This is formally defined in section \ref{ch:block}.
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_5 & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_{256} & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_n \in \mathbb{P} & \wedge & T_v \in \mathbb{P} & \wedge & T_p \in \mathbb{P} & \wedge \\
+T_g \in \mathbb{P} & \wedge & T_w \in \mathbb{P} & \wedge & T_r \in \mathbb{P} & \wedge \\
+T_s \in \mathbb{P} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
-\end{equation}
-where
-\begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..7]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\

### 2015-01-22 Paper.tex

--- 2015-01-22 Paper.tex
@@
 \end{equation}
+and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 

### 2015-02-04 Paper.tex
fixing typo
Minor typo.
replace body code of an account by the KECCAK hash, clear definition and avoids doulbe use of sigma_b

@@
 There exist other subdenominations of Ether:
-\par`
+\par
 \begin{center}
@@
 \begin{eqnarray}
-\quad v(x) & \equiv & x_n \in \mathbb{P} \wedge x_b \in \mathbb{P} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\quad v(x) & \equiv & x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
 \end{eqnarray}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathbf{b}} = \mathbf{o} & \text{otherwise}
-\end{cases} \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+Where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
+\begin{eqnarray}
 I_a & \equiv & a \\
@@
 $I_H$ has no special treatment and is determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
 

### 2015-02-04 Paper.tex
Minor T_l typo.

--- 2015-01-22 Paper.tex
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_l & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_l$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
 
@@
 
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_{memory}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 32KB of memory used, after which it costs substantially more.

### 2015-02-05 Paper.tex
Minor typo.
(65) correction.
Sender fix.
Uncle typo.
Typo R_l, not _n.
Bold l.
Text fix.

@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
 \boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_b
+m & \equiv & {B_H}_c
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & g_0 + g' \\
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-These unambiguously and validly result in immediate halting of the machine with all state changes left intact.
-The one piece of exceptional execution that does not leave state changes intact is the out-of-gas (OOG) exception.
-Here, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.
+Like the out-of-gas (OOG) exception, they not leave state changes intact.
+Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.
 
@@
 s(U, H) &\\
-\quad \vee \; u(U, P(H), n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H), n - 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\mathbf{R}[n]_u = 
+\mathbf{R}[n]_\mathbf{l} = 
 \Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])

### 2015-02-05 Paper.tex
reference block finalisation in order to clearify what internal consisteny with transaction and uncle list hash exactly means

@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}

### 2015-02-21 Paper.tex
verify block number consistency
lower bound difficulty
Add block header validation rule for block number increment
Minor typo.
Add call depth limit condition to exception handling function
Improve 0x40 BLOCKHASH function definition
* Correct reference to the P helper function
* Correct reference to block header number Hi
* Clarify the condition for when 0 is left on the stack
* Use h instead of Hh when refering to the current block hash

--- 2015-02-05 Paper.tex
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_5 & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_{256} & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
 T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
@@
 P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+\end{equation}
+
+The block number is the parent's block number incremented by one:
+\begin{equation}
+H_i \equiv {{P(H)_H}_i} + 1
 \end{equation}
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv A(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $A$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& For a number greater than or equal to, or less than 256 below, the present block's\\
-&&&& number, 0 is left on the stack.\\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_n \vee a = 256 \vee h = 0 \\ H_h & \text{if} \quad n = H_n \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from the hash given: $h = H_h$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule

### 2015-02-21 Paper.tex
Correct CALLER, COINBASE and TIMESTAMP reference to block header fields

--- 2015-02-05 Paper.tex
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_b$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_t$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule

### 2015-02-21 Paper.tex
Set gas cost multiplier for copy op codes as ceiling of number of words

@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[2] & \text{if} \quad w = \text{\small CALLDATACOPY} \\
-G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[2] & \text{if} \quad w = \text{\small CODECOPY} \\
-G_{step} + G_{copy}\times\boldsymbol{\mu}_\mathbf{s}[3] & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{step} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \\
+G_{step} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CODECOPY} \\
+G_{step} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_c$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_b$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_t$ \\
 \midrule

### 2015-02-21 Paper.tex
Add missing wedges, ifs and use if instead of else/when

--- 2015-02-21 Paper.tex
@@
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \\
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
 H_u &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
@@
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \mathbf{c}[i] = \text{\small JUMPDEST} \\
+\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
 D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
@@
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
-(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
+(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert < 56 \\
 \big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
@@
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{when} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, t, t,\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, t, t,\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{when} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\

### 2015-02-23 Paper.tex

@@
 & & H_s > {P(H)_H}_s \quad \wedge \\
+& & H_i = {P(H)_H}_i +1 \quad \wedge \\
 & & \lVert H_x \rVert \le 1024

### 2015-02-23 Paper.tex
Add back else to imply ordering of conditionals

--- 2015-02-23 Paper.tex
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$, when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \\
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
 H_u &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g'$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-Where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
-\begin{eqnarray}
+\quad\boldsymbol{\sigma}'[a]_{\mathbf{b}} = \mathbf{o} & \text{otherwise}
+\end{cases} \\
 I_a & \equiv & a \\
@@
 $I_H$ has no special treatment and is determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_{\mathbf{b}}$ being the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
 

### 2015-02-23 Paper.tex
More contributors.

--- 2015-02-23 Paper.tex
@@
 
-Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
+Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Christoph Jentzsch, Pawe\l{} Bylica, Jutta Steiner, Gustav Simonsson, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
 
@@
 \begin{cases}
-\lfloor \mathbf{k}_n[\lfloor i \div 2 \rfloor] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\lfloor \mathbf{k}_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
 \mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
@@
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-&&&& $Z_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) $ \\
 \midrule

### 2015-02-24 Paper.tex
message call to account with no code
add identity function

@@
 
-As can be seen, there are three exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are three so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The three contracts in addresses 1, 2 and 3 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme and the RIPEMD 160-bit hash scheme respectively.
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 

### 2015-02-25 Paper.tex
add CALLCODE and others to fee exceptions

--- 2015-02-23 Paper.tex
@@
 The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE} and {\small SHA3}.
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE} and {\small CALL}.
+Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE}, {\small EXP}, {\small LOG} and {\small SHA3}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
 Finally, gas may be paid due to an increase in the usage of the memory.

### 2015-02-25 Paper.tex
add log series and refunds to output of execution function

--- 2015-02-23 Paper.tex
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}

### 2015-02-26 Paper.tex
Min 1024 on difficulty. Fixes #65

@@
 {P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
-{P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{otherwise}\\
+\max \{ 1024, x \} & \text{otherwise}\\
 \end{cases}
+\end{equation}
+where:
+\begin{equation}
+x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor
 \end{equation}
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}

### 2015-03-05 Paper.tex
197 statement test change
Update Paper.tex
typo
Update Paper.tex
halts due in -> halts in

@@
 The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts due in an exceptional fashion (i.e.
+Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
 due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution (i.e.
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas (OOG) exception, they not leave state changes intact.
+Like the out-of-gas (OOG) exception, they do not leave state changes intact.
 Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.

### 2015-03-13 Paper.tex
Fix typo in revert-state-upon-exception equation 91
This equation previously read:

    sigma' = sigma1 if o==emptyset, else sigma**

I think that "o" is never the empty set: my reading of eqs 92, 103, 109, 110, 113, and 118 leads me to believe that o will be:

* () if a Z=True "exceptional halt" occurs (out-of-gas, bad-instruction,
  stack-underflow, bad-jumpdest, call-stack-overflow)
* () if STOP or SUICIDE instructions are executed
* () if RETURN is executed with length-to-return = 0
* (stuff) is RETURN is executed with length-to-return > 0

And I think that the revert-to-sigma1 behavior is

--- 2015-02-26 Paper.tex
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operations are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operations are not subject to the $2^{256}$ modulo. \\
 \midrule

### 2015-03-13 Paper.tex
Minor typo.

--- 2015-02-26 Paper.tex
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1] \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule

### 2015-03-17 Paper.tex
Wording in intro to execution model

@@
 \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}_1 & \text{if} \quad \mathbf{o} = \varnothing \\
+\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} & \text{otherwise}
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas (OOG) exception, they not leave state changes intact.
-Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.
+Like the out-of-gas (OOG) exception, they do not leave state changes intact.
+Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
 
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
@@
 \begin{equation}
-v \in \{27,28\} \quad\wedge\quad r \geqslant \mathtt{\tiny secp256k1n} \quad\wedge\quad s \geqslant \mathtt{\tiny secp256k1p}
+v \in \{27,28\} \quad\wedge\quad r < \mathtt{\tiny secp256k1n} \quad\wedge\quad s < \mathtt{\tiny secp256k1p}
 \end{equation}
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
+&&&& Note the overflow semantic when $-2^{255}$ is negated.\\
 \midrule

### 2015-03-17 Paper.tex
Clarify definition of PC and GAS instructions

--- 2015-03-17 Paper.tex
@@
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the 256-bit integer keys to the RLP-encoded 256-bit integer values.
+\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
@@
 \begin{equation}
-L_I\big( (k, v) \big) \equiv \big(k, \texttt{\small RLP}(v)\big)
+L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 \begin{equation}
-p(a) \equiv  \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
 \end{equation}
@@
 where $v$ is the account validity function:
-\begin{eqnarray}
-\quad v(x) & \equiv & x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{eqnarray}
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
 
@@
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-It is assumed that the external actor will be human in nature, though software tools will be used in its construction and dissemination.
+While is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+\eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[unclesHash] The Keccak 256-bit hash of the uncles list portion of this block; formally $H_u$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
 \item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
@@
 
-The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
@@
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_u &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_u, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_u \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
@@
 D(H) \equiv \begin{cases}
-2^{22} & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{1024}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
-\max \{ 1024, x \} & \text{otherwise}\\
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{2048}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
+\max \{ 131072, x \} & \text{otherwise}\\
 \end{cases}
@@
 \begin{equation}
-x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor
+x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{2048}\rfloor
 \end{equation}
@@
 & & H_l = L(H) \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
 & & H_s > {P(H)_H}_s \quad \wedge \\
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - v_0 \\
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
 \boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
@@
 
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string, thus the mutated state becomes $\boldsymbol{\sigma}^*$:
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-\boldsymbol{\sigma}^*[a] \equiv \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big)
-\end{equation}
+\begin{eqnarray}
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}^*[s]_b - v
+\end{eqnarray}
 
@@
 \quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}^* & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
@@
 \begin{equation}
-c \equiv G_{createdata} \times |\mathbf{o}|
+c \equiv G_{codedeposit} \times |\mathbf{o}|
 \end{equation}
@@
 
-No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee.
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 
-We define $\boldsymbol{\sigma}_1$, the checkpoint state as the original state but with the value transferred to the recipient.
-\begin{equation}
-\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v
+We define $\boldsymbol{\sigma}_1$, the first transitional state as the original state but with the value transferred from sender to recipient:
+\begin{equation}
+\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1 \equiv \boldsymbol{\sigma} \quad \text{except:} \\
+\boldsymbol{\sigma}_1 \equiv \boldsymbol{\sigma}_1' \quad \text{except:} \\
+\end{equation}
+\begin{equation}
+\boldsymbol{\sigma}_1'[s]_b \equiv \boldsymbol{\sigma}_1[s]_b - v
+\end{equation}
+\begin{equation}
+\text{and}\quad \boldsymbol{\sigma}_1' \equiv \boldsymbol{\sigma} \quad \text{except:} \\
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
 \end{cases}
@@
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
-due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to code execution (i.e.
-$\boldsymbol{\sigma}_1$).
+due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
+$\boldsymbol{\sigma}$).
 
@@
 \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma}_1 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} & \text{otherwise}
@@
 
-In order to validate the extra computation, a given block $B$ may include the block headers from any known uncle blocks (i.e.
+In order to validate the extra computation, a given block $B$ may include the block headers from any known ommer blocks (i.e.
 blocks whose parent is equivalent to the $N$th generation grandparent of $B$, for a limited set of $N$).
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of ommer blocks.
 
@@
 \begin{enumerate}
-\item Validate (or, if mining, determine) uncles;
+\item Validate (or, if mining, determine) ommers;
 \item validate (or, if mining, determine) transactions;
@@
 
-\subsection{Uncle Validation}
-
-The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of $N$th-generation uncle to the present block where $N \leq 6$.
+\subsection{Ommer Validation}
+
+The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
 Formally:
@@
 
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle by $\frac{15}{16}$ of the reward.
+The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each ommer by a certain amount.
+We raise the block's coinbase account by $R_b$; for each ommer, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 \boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
-\forall_{U \in B_\mathbf{U}} \; \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + \frac{15}{16}R_b
-\end{eqnarray}
-
-If there are collisions of the coinbase addresses between uncles and the block (i.e.
-two uncles with the same coinbase address or an uncle with the same coinbase address as the present block), additions are applied cumulatively.
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 - \frac{1}{8} (U_i - {B_H}_i)) R_b 
+\end{eqnarray}
+
+If there are collisions of the coinbase addresses between ommers and the block (i.e.
+two ommers with the same coinbase address or an ommer with the same coinbase address as the present block), additions are applied cumulatively.
 
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 \midrule
-$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
-$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
-$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
-$G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
-$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
-$G_{sset}$ & 300 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 100 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
-$G_{sclear}$ & 0 & Nothing paid for an {\small SSTORE} operation when the storage value is set to zero from non-zero. \\
-$R_{sclear}$ & 100 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
-$G_{createdata}$ & 5 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
-$G_{exp}$ & 1 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 1 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdatazero}$ & 1 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 5 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 500 & Paid for every transaction. \\
-$G_{log}$ & 1 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 1 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 1 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 10 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 10 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 1 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
+$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$. Only applicable when the transferred value is non-zero. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
+$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{copy}$ & 2 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{step} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \\
-G_{step} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CODECOPY} \\
-G_{step} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{ext} + G_{ext}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL}\\
+G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
 G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
 G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{stop} & \text{if} \quad w = \text{\small STOP}\\
-G_{suicide} & \text{if} \quad w = \text{\small SUICIDE}\\
-G_{step} & \text{otherwise}
+G_{zero} & \text{if} \quad w \in W_{zero}\\
+G_{base} & \text{if} \quad w \in W_{base}\\
+G_{verylow} & \text{if} \quad w \in W_{verylow}\\
+G_{low} & \text{if} \quad w \in W_{low}\\
+G_{mid} & \text{if} \quad w \in W_{mid}\\
+G_{high} & \text{if} \quad w \in W_{high}\\
+G_{ext} & \text{if} \quad w \in W_{ext}
 \end{cases}
@@
 
-and $C_\text{\tiny SSTORE}$ is specified in the appropriate section below.
+and $C_\text{\tiny SSTORE}$ is specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{high}$ = \{{\small JUMPI}\}
+
+$W_{ext}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small BLOCKHASH}\}
 
@@
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 G_{sreset} & \text{otherwise}
@@
 \midrule
-0x58 & {\small PC} & 0 & 1 & Get the program counter. \\
+0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{before} it's incremented by this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
@@
 \midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
+0x5a & {\small GAS} & 0 & 1 & Get the amount of gas available after execution of this instruction.  \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
@@
 &&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1 \wedge \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, t, t,\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_b = \boldsymbol{\sigma}[I_a]_b - \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 1000000, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
 \end{equation}
@@
 Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
-The sequences of both uncles and transactions are empty and represented by $()$.
+The sequences of both ommers and transactions are empty and represented by $()$.
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 

### 2015-03-17 Paper.tex

--- 2015-03-17 Paper.tex
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter, the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.

### 2015-03-18 Paper.tex
Clarify clarification to avoid erroneous implication

@@
 Like the out-of-gas (OOG) exception, they do not leave state changes intact.
-Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) and which will deal with it separately.
+Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
 
@@
 \midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, taking into account the cost of this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\

### 2015-04-01 Paper.tex
uncles block reward 15/16 -> (1-1/8*diffBlockNumber)

@@
 The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle by $\frac{15}{16}$ of the reward.
+We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 \boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
-\forall_{U \in B_\mathbf{U}} \; \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + \frac{15}{16}R_b
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 - \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}

### 2015-04-02 Paper.tex
update genesisGasLimit
Rename uncle to aunt to reduce androcentrism

--- 2015-03-24 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{aunts}).
 The block header contains several pieces of information:
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[unclesHash] The Keccak 256-bit hash of the uncles list portion of this block; formally $H_u$.
+\item[auntsHash] The Keccak 256-bit hash of the aunts list portion of this block; formally $H_u$.
 \item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
@@
 
-The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+The other two components in the block are simply a list of aunt block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the aunt and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
@@
 
-In order to validate the extra computation, a given block $B$ may include the block headers from any known uncle blocks (i.e.
+In order to validate the extra computation, a given block $B$ may include the block headers from any known aunt blocks (i.e.
 blocks whose parent is equivalent to the $N$th generation grandparent of $B$, for a limited set of $N$).
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of aunt blocks.
 
@@
 \begin{enumerate}
-\item Validate (or, if mining, determine) uncles;
+\item Validate (or, if mining, determine) aunts;
 \item validate (or, if mining, determine) transactions;
@@
 
-\subsection{Uncle Validation}
-
-The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of $N$th-generation uncle to the present block where $N \leq 6$.
+\subsection{Aunt Validation}
+
+The validation of aunt headers means nothing more than verifying that each aunt header is both a valid header and satisfies the relation of $N$th-generation aunt to the present block where $N \leq 6$.
 Formally:
@@
 
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle by $\frac{15}{16}$ of the reward.
+The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each aunt by a certain amount.
+We raise the block's coinbase account by $R_b$; for each aunt, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the aunt by $\frac{15}{16}$ of the reward.
 Formally we define the function $\Omega$:
@@
 
-If there are collisions of the coinbase addresses between uncles and the block (i.e.
-two uncles with the same coinbase address or an uncle with the same coinbase address as the present block), additions are applied cumulatively.
+If there are collisions of the coinbase addresses between aunts and the block (i.e.
+two aunts with the same coinbase address or an aunt with the same coinbase address as the present block), additions are applied cumulatively.
 
@@
 Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
-The sequences of both uncles and transactions are empty and represented by $()$.
+The sequences of both aunts and transactions are empty and represented by $()$.
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the aunt lists in RLP, both empty lists.
 

### 2015-04-02 Paper.tex
Change block header gas limit validation to new algo

--- 2015-04-01 Paper.tex
@@
 \begin{equation}
-x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{1024}\rfloor
-\end{equation}
-
-The canonical gas limit of a block of header $H$ is defined as $L(H)$:
-\begin{eqnarray}
-L(H) & \equiv & \begin{cases}
-10^6 & \text{if} \quad H_i = 0\\
-125000 & \text{if} \quad L'(H) < 125000\\
-L'(H) & \text{otherwise}
-\end{cases}\\
-L'(H) & \equiv & \big\lfloor \frac{1023 {P(H)_H}_l + \lfloor \frac{6}{5}{P(H)_H}_g \rfloor}{1024}\big\rfloor
+x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{2048}\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l > 12499
 \end{eqnarray}
@@
 & & H_d = D(H) \quad \wedge \\
-& & H_l = L(H) \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l > 12499  \quad \wedge \\
 & & H_s > {P(H)_H}_s \quad \wedge \\
@@
 
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle gets rewarded depending on the block number.
+The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each ommer by a certain amount.
+We raise the block's coinbase account by $R_b$; for each ommer, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 
-If there are collisions of the coinbase addresses between uncles and the block (i.e.
-two uncles with the same coinbase address or an uncle with the same coinbase address as the present block), additions are applied cumulatively.
+If there are collisions of the coinbase addresses between ommers and the block (i.e.
+two ommers with the same coinbase address or an ommer with the same coinbase address as the present block), additions are applied cumulatively.
 

### 2015-04-02 Paper.tex
new gas costs

--- 2015-04-01 Paper.tex
@@
 The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
-We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle gets rewarded depending on the block number.
+We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle by $\frac{15}{16}$ of the reward.
 Formally we define the function $\Omega$:
@@
 \boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 - \frac{1}{8} (U_i - {B_H}_i)) R_b 
+\forall_{U \in B_\mathbf{U}} \; \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + \frac{15}{16}R_b
 \end{eqnarray}
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer, it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 \midrule
-$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
-$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
-$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
-$G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
-$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
-$G_{sset}$ & 300 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 100 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
-$G_{sclear}$ & 0 & Nothing paid for an {\small SSTORE} operation when the storage value is set to zero from non-zero. \\
-$R_{sclear}$ & 100 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
-$G_{createdata}$ & 5 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
-$G_{exp}$ & 1 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 1 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdatazero}$ & 1 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 5 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 500 & Paid for every transaction. \\
-$G_{log}$ & 1 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 1 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 1 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 10 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 10 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 1 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
+$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+$G_{createdata}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$. Only applicable when the transferred value is non-zero. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
+$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{copy}$ & 2 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
 

### 2015-04-04 Paper.tex
Clear up any uncertainties over how P is to be encoded in RLP.

--- 2015-04-02 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{aunts}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}).
 The block header contains several pieces of information:
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[auntsHash] The Keccak 256-bit hash of the aunts list portion of this block; formally $H_u$.
+\item[unclesHash] The Keccak 256-bit hash of the uncles list portion of this block; formally $H_u$.
 \item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
@@
 
-The other two components in the block are simply a list of aunt block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+The other two components in the block are simply a list of uncle block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the aunt and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the uncle and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
@@
 
-In order to validate the extra computation, a given block $B$ may include the block headers from any known aunt blocks (i.e.
+In order to validate the extra computation, a given block $B$ may include the block headers from any known uncle blocks (i.e.
 blocks whose parent is equivalent to the $N$th generation grandparent of $B$, for a limited set of $N$).
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of aunt blocks.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.
 
@@
 \begin{enumerate}
-\item Validate (or, if mining, determine) aunts;
+\item Validate (or, if mining, determine) uncles;
 \item validate (or, if mining, determine) transactions;
@@
 
-\subsection{Aunt Validation}
-
-The validation of aunt headers means nothing more than verifying that each aunt header is both a valid header and satisfies the relation of $N$th-generation aunt to the present block where $N \leq 6$.
+\subsection{Uncle Validation}
+
+The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of $N$th-generation uncle to the present block where $N \leq 6$.
 Formally:
@@
 
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each aunt by a certain amount.
-We raise the block's coinbase account by $R_b$; for each aunt, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the aunt by $\frac{15}{16}$ of the reward.
+The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount.
+We raise the block's coinbase account by $R_b$; for each uncle, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the uncle by $\frac{15}{16}$ of the reward.
 Formally we define the function $\Omega$:
@@
 
-If there are collisions of the coinbase addresses between aunts and the block (i.e.
-two aunts with the same coinbase address or an aunt with the same coinbase address as the present block), additions are applied cumulatively.
+If there are collisions of the coinbase addresses between uncles and the block (i.e.
+two uncles with the same coinbase address or an uncle with the same coinbase address as the present block), additions are applied cumulatively.
 
@@
 Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
-The sequences of both aunts and transactions are empty and represented by $()$.
+The sequences of both uncles and transactions are empty and represented by $()$.
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the aunt lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the uncle lists in RLP, both empty lists.
 

### 2015-04-06 Paper.tex
gasUsed must be smaller than gasLimit in blockHeader

--- 2015-04-02 Paper.tex
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}\}
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
 

### 2015-04-07 Paper.tex
secure trie for storage and state trie

--- 2015-04-04 Paper.tex
@@
 \begin{eqnarray}
-\quad v(x) & \equiv & x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \\ \nonumber
+\wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
 \end{eqnarray}

### 2015-04-07 Paper.tex
Gender neutral terming for siblings of parents. Fixes #96

@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{aunts}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[auntsHash] The Keccak 256-bit hash of the aunts list portion of this block; formally $H_u$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_u$.
 \item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
@@
 
-The other two components in the block are simply a list of aunt block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the aunt and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
@@
 
-In order to validate the extra computation, a given block $B$ may include the block headers from any known aunt blocks (i.e.
+In order to validate the extra computation, a given block $B$ may include the block headers from any known ommer blocks (i.e.
 blocks whose parent is equivalent to the $N$th generation grandparent of $B$, for a limited set of $N$).
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of aunt blocks.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of ommer blocks.
 
@@
 \begin{enumerate}
-\item Validate (or, if mining, determine) aunts;
+\item Validate (or, if mining, determine) ommers;
 \item validate (or, if mining, determine) transactions;
@@
 
-\subsection{Aunt Validation}
-
-The validation of aunt headers means nothing more than verifying that each aunt header is both a valid header and satisfies the relation of $N$th-generation aunt to the present block where $N \leq 6$.
+\subsection{Ommer Validation}
+
+The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
 Formally:
@@
 
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each aunt by a certain amount.
-We raise the block's coinbase account by $R_b$; for each aunt, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the aunt by $\frac{15}{16}$ of the reward.
+The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each ommer by a certain amount.
+We raise the block's coinbase account by $R_b$; for each ommer, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the ommer by $\frac{15}{16}$ of the reward.
 Formally we define the function $\Omega$:
@@
 
-If there are collisions of the coinbase addresses between aunts and the block (i.e.
-two aunts with the same coinbase address or an aunt with the same coinbase address as the present block), additions are applied cumulatively.
+If there are collisions of the coinbase addresses between ommers and the block (i.e.
+two ommers with the same coinbase address or an ommer with the same coinbase address as the present block), additions are applied cumulatively.
 
@@
 Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
-The sequences of both aunts and transactions are empty and represented by $()$.
+The sequences of both ommers and transactions are empty and represented by $()$.
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the aunt lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 

### 2015-04-07 Paper.tex
createdata -> codedeposit

--- 2015-04-07 Paper.tex
@@
 The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each ommer by a certain amount.
-We raise the block's coinbase account by $R_b$; for each ommer, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the ommer by $\frac{15}{16}$ of the reward.
+We raise the block's coinbase account by $R_b$; for each ommer, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:

### 2015-04-07 Paper.tex
Reduce andro-centricism.

--- 2015-04-07 Paper.tex
@@
 \item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
+\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the 256-bit integer keys to the RLP-encoded 256-bit integer values.
 The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
@@
 \begin{equation}
-L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+L_I\big( (k, v) \big) \equiv \big(k, \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+p(a) \equiv  \big(a, \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
 \end{equation}
@@
 where $v$ is the account validity function:
-\begin{eqnarray}
-v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \\ \nonumber
-\wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{eqnarray}
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
 
@@
 & & H_l = L(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
 & & H_s > {P(H)_H}_s \quad \wedge \\

### 2015-04-07 Paper.tex
Adjust symbol used for ommer hash to name change.

@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_u$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
 \item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.

### 2015-04-07 Paper.tex
update difficulty parameters

--- 2015-04-07 Paper.tex
@@
 \midrule
-$G_{step}$ & 1 & Default amount of gas to pay for execution cycle. \\
-$G_{balance}$ & 20 & Paid for a {\small BALANCE} operation. \\
-$G_{stop}$ & 0 & Nothing paid for the {\small STOP} operation. \\
-$G_{suicide}$ & 0 & Nothing paid for the {\small SUICIDE} operation. \\
-$G_{sload}$ & 20 & Paid for a {\small SLOAD} operation. \\
-$G_{sset}$ & 300 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 100 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
-$G_{sclear}$ & 0 & Nothing paid for an {\small SSTORE} operation when the storage value is set to zero from non-zero. \\
-$R_{sclear}$ & 100 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$G_{create}$ & 100 & Paid for a {\small CREATE} operation. \\
-$G_{createdata}$ & 5 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 20 & Paid for a {\small CALL} operation. \\
-$G_{exp}$ & 1 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 1 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 1 & Paid for every additional word when expanding memory. \\
-$G_{txdatazero}$ & 1 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 5 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 500 & Paid for every transaction. \\
-$G_{log}$ & 1 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 1 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 1 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 10 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 10 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 1 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
+$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$. Only applicable when the transferred value is non-zero. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
+$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{copy}$ & 2 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
 

### 2015-04-07 Paper.tex
clearify G_{sreset}

@@
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\

### 2015-04-07 Paper.tex
Use geqslant instead of greater-than for gas limit validation
Off by one (order of magnitude) error

@@
 \midrule
-0x58 & {\small PC} & 0 & 1 & Get the program counter. \\
+0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
+&&&&  corresponding to this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
@@
 \midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
+&&&& for the cost of this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\

### 2015-04-08 Paper.tex

@@
 
-The canonical gas limit of a block of header $H$ is defined as $L(H)$:
-\begin{eqnarray}
-L(H) & \equiv & \begin{cases}
-10^6 & \text{if} \quad H_i = 0\\
-125000 & \text{if} \quad L'(H) < 125000\\
-L'(H) & \text{otherwise}
-\end{cases}\\
-L'(H) & \equiv & \big\lfloor \frac{1023 {P(H)_H}_l + \lfloor \frac{6}{5}{P(H)_H}_g \rfloor}{1024}\big\rfloor
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor \quad \wedge \\
+& & H_l \geqslant 125000
 \end{eqnarray}
@@
 & & H_d = D(H) \quad \wedge \\
-& & H_l = L(H) \quad \wedge \\
 & & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l \geqslant 125000  \quad \wedge \\
 & & H_s > {P(H)_H}_s \quad \wedge \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 &&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLVALUE}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEWACCOUNT}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLVALUE}}(\boldsymbol{\mu}) \equiv \begin{cases}
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny CALLNEWACCOUNT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1]] = \varnothing \\
+0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
 \midrule

### 2015-04-08 Paper.tex
Minor alterations to naming.

@@
 T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
-\end{array}
+\end{array}\dfrac{num}{den}
 \end{equation}
@@
 
-Useful corrections and suggestions were provided by a number of others from the Ethereum community including Aeron Buchanan, Christoph Jentzsch, Pawe\l{} Bylica, Jutta Steiner, Gustav Simonsson, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
 
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLVALUE}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEWACCOUNT}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLVALUE}}(\boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
 G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny CALLNEWACCOUNT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1]] = \varnothing \\

### 2015-04-08 Paper.tex
max 2 ommers

@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or if a {\small CALL}, {\small CALLCODE} or {\small CREATE} instruction is executed when the call stack limit of 1024 is reached.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items or a {\small JUMP}/{\small JUMPI} destination is invalid.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
+The maximum of ommer headers is two.
 Formally:
 \begin{equation}
-\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H), 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H), 6)
 \end{equation}

### 2015-04-08 Paper.tex
bloom size 64 -> 256 bytes

--- 2015-04-08 Paper.tex
@@
 
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 512 bits (64 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{64}
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
 \end{equation}
@@
 
-We define the Bloom filter function, $M$, to reduce a log entry include a single 64-byte hash:
-\begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:512}(t) \big)
-\end{equation}
-
-where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 512, given an arbitrary byte series.
+We define the Bloom filter function, $M$, to reduce a log entry include a single 256-byte hash:
+\begin{equation}
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+\end{equation}
+
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
 It does this through taking the low-order 9 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
@@
 \begin{eqnarray}
-M_{3:512}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{64} \quad \text{where:}\\
+M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
 \forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
-m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 512
+m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
 \end{eqnarray}
@@
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{64} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
 H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\

### 2015-04-09 Paper.tex
add suicide refund

--- 2015-04-08 Paper.tex
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
@@
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$. Only applicable when the transferred value is non-zero. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
 $G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
@@
 &&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\\bottomrule
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+0 & \text{otherwise}
+\end{cases}$ \\\bottomrule
 \end{tabular*}

### 2015-04-09 Paper.tex
update genesis difficulty

@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
@@
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$. Only applicable when the transferred value is non-zero. \\
 $G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
@@
 &&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
-0 & \text{otherwise}
-\end{cases}$ \\\bottomrule
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\\bottomrule
 \end{tabular*}
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 131072, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $131072$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.

### 2015-04-09 Paper.tex

--- 2015-04-09 Paper.tex
@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or if a {\small CALL}, {\small CALLCODE} or {\small CREATE} instruction is executed when the call stack limit of 1024 is reached.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.

### 2015-04-09 Paper.tex
Minor fix for blooms definition.
style

--- 2015-04-09 Paper.tex
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 131072, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $131072$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{17}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{17}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.

### 2015-04-09 Paper.tex

--- 2015-04-09 Paper.tex
@@
 where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
-It does this through taking the low-order 9 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
 Formally:
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
 \midrule
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\

### 2015-04-09 Paper.tex
ad mixHash to blockHeader

@@
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 1024 bytes or fewer; formally $H_x$.
-\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mixHash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
@@
 H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_n \in \mathbb{B}_{32}
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
 \midrule
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\

### 2015-04-09 Paper.tex
Correct ranges for ECDSA privkey and signature r, s values

--- 2015-04-09 Paper.tex
@@
 
-It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer in the range $(0, 2^{256})$ represented as a byte array of length 32 in big-endian form.
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
 
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
 It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
@@
 We declare that a signature is invalid unless the following is true:
-\begin{equation}
-v \in \{27,28\} \quad\wedge\quad r < \mathtt{\tiny secp256k1n} \quad\wedge\quad s < \mathtt{\tiny secp256k1p}
-\end{equation}
+\begin{eqnarray}
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
 
@@
 \begin{eqnarray}
-\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337\\
-\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
 \end{eqnarray}

### 2015-04-09 Paper.tex
ethash I - block header validity

--- 2015-04-09 Paper.tex
@@
 \begin{equation}
-%\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
-\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
+\mathtt{PoW}(H, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+With $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
+V(H) & \equiv & \mathtt{PoW}(H, H_n, \mathbf{d})[0] = H_m \quad \wedge \\
+& & \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
 & & H_d = D(H) \quad \wedge \\

### 2015-04-09 Paper.tex
introducing ethash

--- 2015-04-09 Paper.tex
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
-
-As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
-
-\subsubsection{PoC Series}
-
-For the PoC series, we use a simplified proof-of-work.
-This is not ASIC resistant and is meant merely as a placeholder.
-It utilises the bare Keccak hash function to secure the block chain by requiring the Keccak hash of the concatenation of the nonce and the header's Keccak hash to be sufficiently low.
-
-It is formally defined as $\mathtt{PoW}$:
-\begin{equation}
-\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny KEC}(\mathtt{\tiny KEC}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
-\end{equation}
-
-where:
-$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
-$\mathtt{\tiny KEC}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
-$n$ is the nonce, a series of 32 bytes;
-$\circ$ is the series concatenation operator;
-$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
-
-\subsubsection{Release Series}
-
-For the release series, we use a more complex proof-of-work.
-This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
-Secondly that it concerns the utilisation of either the blockchain or the full state trie.
-
-As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
-A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
-$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
-Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
-Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
-swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
-This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
-
-Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
-
-%\section{User Interface}
-%
-%Browser-style.
-%Displays ETH balance (i.e. functions as a wallet).
-%Tabs or perhaps MacOS style app icons.
-%Display taken up with a single app.
-%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
-%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
-%Preconfigured with home app that is the Eth app store.
-%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
+\mathtt{PoW}(H, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H$ is the new block's header; $H_n$ is the nonce of the header; $\mathbf{d}$ is the DAG (a large data set neede to compute the mixHash) and $H_d$ is the new block's difficulty value (i.e.
+the block difficulty from section \ref{ch:ghost}).
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+The underlying algorithm is called Ethash and is described below.
+
+\subsubsection{Ethash}
+Ethash is the planned PoW algorithm for Ethereum 1.0.
+It is the latest version of Dagger-Hashimoto, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+The general route that the algorithm takes is as follows:
+There exists a seed which can be computed for each block by scanning through the block headers up until that point.
+From the seed, one can compute a pseudorandom cache, $J_{cacheInit}$ bytes in intial size. Light clients store the cache.
+From the cache, we can generate a dataset, $J_{datasetInit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+Full clients and miners store the dataset.
+The dataset grows linearly with time.
+Mining involves grabbing random slices of the dataset and hashing them together.
+Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 
@@
 
-It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
-% to avoid line break in range definition
-
-in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer in the range $(0, 2^{256})$ represented as a byte array of length 32 in big-endian form.
 
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer $< 2^{256}$).
 It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
@@
 We declare that a signature is invalid unless the following is true:
-\begin{eqnarray}
-0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-v \in \{27,28\}
-\end{eqnarray}
+\begin{equation}
+v \in \{27,28\} \quad\wedge\quad r < \mathtt{\tiny secp256k1n} \quad\wedge\quad s < \mathtt{\tiny secp256k1p}
+\end{equation}
 
@@
 \begin{eqnarray}
-\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337\\
+\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
 \end{eqnarray}

### 2015-04-11 Paper.tex
nonce validation
add Ethash overview in block finalisation section

@@
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
+For Ethereum 1.0 we have choosen the first path.
 
@@
 
-Where $H$ is the new block's header; $H_n$ is the nonce of the header; $\mathbf{d}$ is the DAG (a large data set neede to compute the mixHash) and $H_d$ is the new block's difficulty value (i.e.
+Where $H$ is the new block's header; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 The underlying algorithm is called Ethash and is described below.
-
 \subsubsection{Ethash}
 Ethash is the planned PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
 The general route that the algorithm takes is as follows:
+
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheInit}$ bytes in intial size. Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetInit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
+Light clients store the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
 The dataset grows linearly with time.
+
 Mining involves grabbing random slices of the dataset and hashing them together.
@@
 The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
 
@@
 
+\section{Ethash}\label{app:ethash}
+
 \end{document}

### 2015-04-11 Paper.tex
style
Full DAG creation

--- 2015-04-09 Paper.tex
@@
 \begin{equation}
-\mathtt{PoW}(H, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-With $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+%\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
+\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
 Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H, H_n, \mathbf{d})[0] = H_m \quad \wedge \\
-& & \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
+V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
 & & H_d = D(H) \quad \wedge \\
@@
 
+\section{Ethash}\label{app:ethash}
+\subsection{Definitions}
+We employ the following definitions:
+
+\begin{tabular*}{\columnwidth}[h]{lrl}
+\toprule
+Name & Value & Description \\
+\midrule
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Cache growth per epoch. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{cachemultiplier}$ & 1024 & Size of the dataset relative to the cache. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{datasetparents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of parents of each dataset element. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\bottomrule
+\end{tabular*}
+
+\subsection{Size of dataset and cache}
+The size for Ethash's cache and dataset depend on the epoch, which depends on the block number.
+\begin{equation}
+ C_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$, and the size of the cache by $J_{cachegrowth}$, every epoch.
+In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $\mathbf{d}_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ \mathbf{d}_{size} = C_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot C_{epoch}, J_{mixbytes})
+\end{equation}
+The size of the cache, $\mathbf{c}_{size}$, is calculated using
+\begin{equation}
+ \mathbf{c}_{size} = C_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot C_{epoch}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ C_{prime}(x, y) = \begin{cases}
+x & \text{if} \quad x \in \mathbb{P} \\
+C_{prime}(x - y, y) & \text{otherwise}
+\end{cases}
+\end{equation}
+\subsection{Dataset generation}
+In order the generate the dataset we need the cache $\mathbf{c}$ which is an array of bytes. It depends on the cache size  $\mathbf{c}_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
+\subsubsection{Seed hash}
+The seed hash is different for every epoch.
+For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
+For every other epoch it is always the Keccak-256 hash of the previous seed hash:
+\begin{equation}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad C_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+\end{cases}
+\end{equation}
+With $\mathbf{0}_{32}$ being 32 bytes of zeros.
+
+\subsubsection{Cache}
+The cache production process involves using the seed hash to first sequentially filling up $\mathbf{c}_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of \cite{lerner2014randmemohash} RandMemoHash algorithm:
+
+The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows:
+
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
+\begin{equation}
+ \mathbf{c}_{i} = \begin{cases}
+\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
+\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
+\end{cases}
+\end{equation}
+Therefore $ \mathbf{c'}$ can be defined as 
+\begin{equation}
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{\mathbf{c}_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+\mathbf{x} & \text{if} \quad y = 0 \quad  \\
+E_{rmh}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
+E_{cacherounds}(\mathbf{x}, y -1 ) & \text{otherwise}
+\end{cases}
+\end{equation}
+Where a single round modifies each subset of the cache as follows:
+\begin{equation}
+ E_{rmh}(\mathbf{x}) = \lbrace E_{rmhSingle}(\mathbf{x}, 0), E_{rmhSingle}(\mathbf{x}, 1), ... , E_{rmhSingle}(\mathbf{x}, n) \rbrace
+\end{equation}
+\begin{equation}
+  E_{rmhSingle}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n) 
+\end{equation}
+\subsubsection{Full dataset calculation}
+Essentially, we combine data from $J_{datasetparents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by $\mathbf{d}_{size}$ items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\mathbf{d}_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
+\begin{equation}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot \mathrm{0x01000193} \oplus \mathbf{y}) \mod 2^{22}
+\end{equation}
+The single item of the dataset can now be calculated as:
+\begin{equation}
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, 0, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E'_{mix}) & \text{if} \quad p < J_{datasetparents} \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E'_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[(E_\text{\tiny FNV}(i \oplus p), \mathbf{m}[ \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod \mathbf{c}_{size})
+\end{equation}
+
 \end{document}
+
+C_\text{\tiny SSTORE}

### 2015-04-11 Paper.tex
corrections

@@
 $J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{cachemultiplier}$ & 1024 & Size of the dataset relative to the cache. \\
 $J_{epoch}$ & 30000 & Blocks per epoch. \\
@@
 $J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{datasetparents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of parents of each dataset element. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
 $J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
@@
 \subsection{Size of dataset and cache}
-The size for Ethash's cache and dataset depend on the epoch, which depends on the block number.
-\begin{equation}
- C_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$, and the size of the cache by $J_{cachegrowth}$, every epoch.
+The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which depends on the block number.
+\begin{equation}
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
 Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $\mathbf{d}_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- \mathbf{d}_{size} = C_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot C_{epoch}, J_{mixbytes})
-\end{equation}
-The size of the cache, $\mathbf{c}_{size}$, is calculated using
-\begin{equation}
- \mathbf{c}_{size} = C_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot C_{epoch}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- C_{prime}(x, y) = \begin{cases}
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
 x & \text{if} \quad x \in \mathbb{P} \\
-C_{prime}(x - y, y) & \text{otherwise}
+E_{prime}(x - y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$ which is an array of bytes. It depends on the cache size  $\mathbf{c}_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
 \subsubsection{Seed hash}
@@
  C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad C_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
 \texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $\mathbf{c}_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of \cite{lerner2014randmemohash} RandMemoHash algorithm:
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}:
 
@@
 \begin{equation}
- n = \left\lfloor\frac{\mathbf{c}_{size}}{J_{hashbytes}}\right\rfloor
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 E_{rmh}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(\mathbf{x}, y -1 ) & \text{otherwise}
+E_{cacherounds}(E_{rmh}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Full dataset calculation}
-Essentially, we combine data from $J_{datasetparents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by $\mathbf{d}_{size}$ items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\mathbf{d}_{size}}{J_{hashbytes}}\right\rfloor
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot \mathrm{0x01000193} \oplus \mathbf{y}) \mod 2^{22}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{22}
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, 0, \varnothing)
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
 \end{equation}
@@
   E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E'_{mix}) & \text{if} \quad p < J_{datasetparents} \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E'_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[(E_\text{\tiny FNV}(i \oplus p), \mathbf{m}[ \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod \mathbf{c}_{size})
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[(E_\text{\tiny FNV}(i \oplus p), \mathbf{m}[ \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}) & \text{otherwise}
+\end{cases}
 \end{equation}

### 2015-04-13 Paper.tex
corrections

@@
 \subsection{Size of dataset and cache}
-The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which depends on the block number.
+The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
 \begin{equation}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}:
-
-The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows:
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
 
@@
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
-E_{rmh}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_{rmh}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{rmh}(\mathbf{x}) = \lbrace E_{rmhSingle}(\mathbf{x}, 0), E_{rmhSingle}(\mathbf{x}, 1), ... , E_{rmhSingle}(\mathbf{x}, n) \rbrace
-\end{equation}
-\begin{equation}
-  E_{rmhSingle}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n) 
+ E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n) \rbrace
+\end{equation}
+\begin{equation}
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n]) 
 \end{equation}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[(E_\text{\tiny FNV}(i \oplus p), \mathbf{m}[ \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}) & \text{otherwise}
-\end{cases}
-\end{equation}
+E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Proof-of-work function}
 

### 2015-04-15 Paper.tex
PoW function complete

@@
 $J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Cache growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
 $J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
@@
 \subsection{Proof-of-work function}
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the fnv function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the nonce is valid.
+Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
+It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
+
+In order to formalize this algorithm, we first define a seed hash $\mathbf{s}_h$:
+\begin{equation}
+ \mathbf{s}_h = \texttt{KEC512}(\mathbf{h} + \mathbf{n}_r)
+\end{equation}
+With $\mathbf{h}$ being the hash of the header without the nonce, and $\mathbf{n}_r$ the reverted bytes sequence of the nonce:
+\begin{equation}
+ \mathbf{n}_r[i] =  \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+\end{equation}
+with $\mathbf{n}$ being the nonce of the header.
+
+The mix $\mathbf{m}$ consists of $n_{mix}$ replicated sequences of $\mathbf{s}_h$.
+\begin{equation}
+ \mathbf{m}[i] = \mathbf{s}_h \forall 0 \leqslant i < n_{mix}
+\end{equation}
+\begin{equation}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+Now we add random dataset nodes to the mix:
+\begin{equation}
+ \mathbf{m} = E_{accesses}(\mathbf{m}, \mathbf{s}_h, -1)
+\end{equation}
+\begin{equation}
+ E_{accesses}(\mathbf{m}, \mathbf{s}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixDataset}(\mathbf{m}), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ (E_{mixDataset}(\mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{m}, \mathbf{s}, i))
+\end{equation}
+\begin{equation}
+ E_{newdata}(\mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j]
+\end{equation}
+The mix $\mathbf{m}$ is now compressed as follows:
+\begin{equation}
+ \mathbf{m}_c = E_{compress}(\mathbf{m}, -4)
+\end{equation}
+\begin{equation}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i], \mathbf{m}[i + 1]), \mathbf{m}[i + 2]), \mathbf{m}[i + 3]), i + 4) & \text{otherwise}
+\end{cases}
+\end{equation}
+Finally the PoW-function can be defined as an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
+\begin{equation}
+ \mathtt{PoW}(H, H_n, \mathbf{d}) = \lbrace  \mathbf{m}_c, \texttt{KEC}(\mathbf{s} + \mathbf{m}_c) \rbrace
+\end{equation}
 

### 2015-04-16 Paper.tex
revert order of functions - start with PoW function

--- 2015-04-13 Paper.tex
@@
 \end{equation}
-\subsubsection{Full dataset calculation}
+\subsubsection{Full dataset calculation} \label{dataset} 
 Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
@@
 \subsection{Proof-of-work function}
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the nonce is valid.
+Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
+It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
+
+The PoW-function can be defined as an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
+\begin{equation}
+ \mathtt{PoW}(H, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(H, H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(H, H_n) + \mathbf{m}_c(H, H_n, \mathbf{d})) \rbrace
+\end{equation}
+The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(H, H_n, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(H, H_n), \mathbf{s}_h(H, H_n), -1), -4)
+\end{equation}
+
+The seed hash being:
+\begin{equation}
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+
+With $\mathbf{h}$ being the hash of the header without the nonce, and $E_{revert}(\mathbf{n})$ the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+\end{equation}
+
+The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
+
+The number of replicated sequences in the mix is:
+\begin{equation}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j]
+\end{equation}
+The mix is compressed as follows:
+\begin{equation}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i], \mathbf{m}[i + 1]), \mathbf{m}[i + 2]), \mathbf{m}[i + 3]), i + 4) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Mining}
 
 \end{document}
-
-C_\text{\tiny SSTORE}

### 2015-04-16 Paper.tex
hash of header without nonce

@@
 
-The PoW-function can be defined as an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
-\begin{equation}
- \mathtt{PoW}(H, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(H, H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(H, H_n) + \mathbf{m}_c(H, H_n, \mathbf{d})) \rbrace
-\end{equation}
-The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(H, H_n, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(H, H_n), \mathbf{s}_h(H, H_n), -1), -4)
+The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{h}), -1), -4)
 \end{equation}
@@
 \end{equation}
-
-With $\mathbf{h}$ being the hash of the header without the nonce, and $E_{revert}(\mathbf{n})$ the reverted bytes sequence of the nonce $\mathbf{n}$:
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
 \begin{equation}
@@
 \end{equation}
+We note that the ``$+$''-opertor between two byte sequences results in the concatenation of both sequences.
 
@@
 \subsection{Mining}
+$$\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}})))$$
 

### 2015-04-16 Paper.tex
mining
header without nonce

--- 2015-04-15 Paper.tex
@@
 \begin{equation}
-%\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
-\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
+\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
+V(H) & \equiv & \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \\
+& & \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
 & & H_d = D(H) \quad \wedge \\
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathtt{PoW}(B^*, n) < \frac{2^{256}}{H_d} \\
+B'_n & = & n: \quad \mathtt{PoW}(B^*, n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \\
 B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
@@
 
-\section{Ethash}\label{app:ethash}
-\subsection{Definitions}
-We employ the following definitions:
-
-\begin{tabular*}{\columnwidth}[h]{lrl}
-\toprule
-Name & Value & Description \\
-\midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
-\bottomrule
-\end{tabular*}
-
-\subsection{Size of dataset and cache}
-The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
-\begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
-In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x \in \mathbb{P} \\
-E_{prime}(x - y, y) & \text{otherwise}
-\end{cases}
-\end{equation}
-\subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
-\subsubsection{Seed hash}
-The seed hash is different for every epoch.
-For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
-For every other epoch it is always the Keccak-256 hash of the previous seed hash:
-\begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
-\end{cases}
-\end{equation}
-With $\mathbf{0}_{32}$ being 32 bytes of zeros.
-
-\subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
-The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
-
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
-\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
-\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
-\end{cases}
-\end{equation}
-Therefore $ \mathbf{c'}$ can be defined as 
-\begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
-\mathbf{x} & \text{if} \quad y = 0 \quad  \\
-E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
-\end{cases}
-\end{equation}
-Where a single round modifies each subset of the cache as follows:
-\begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n) \rbrace
-\end{equation}
-\begin{equation}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n]) 
-\end{equation}
-\subsubsection{Full dataset calculation}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
-\begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{22}
-\end{equation}
-The single item of the dataset can now be calculated as:
-\begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\subsection{Proof-of-work function}
-Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the fnv function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the nonce is valid.
-Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
-It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
-
-In order to formalize this algorithm, we first define a seed hash $\mathbf{s}_h$:
-\begin{equation}
- \mathbf{s}_h = \texttt{KEC512}(\mathbf{h} + \mathbf{n}_r)
-\end{equation}
-With $\mathbf{h}$ being the hash of the header without the nonce, and $\mathbf{n}_r$ the reverted bytes sequence of the nonce:
-\begin{equation}
- \mathbf{n}_r[i] =  \mathbf{n}[\lVert \mathbf{n} \rVert -i]
-\end{equation}
-with $\mathbf{n}$ being the nonce of the header.
-
-The mix $\mathbf{m}$ consists of $n_{mix}$ replicated sequences of $\mathbf{s}_h$.
-\begin{equation}
- \mathbf{m}[i] = \mathbf{s}_h \forall 0 \leqslant i < n_{mix}
-\end{equation}
-\begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-Now we add random dataset nodes to the mix:
-\begin{equation}
- \mathbf{m} = E_{accesses}(\mathbf{m}, \mathbf{s}_h, -1)
-\end{equation}
-\begin{equation}
- E_{accesses}(\mathbf{m}, \mathbf{s}, i) = \begin{cases}
-\mathbf{m} & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixDataset}(\mathbf{m}), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- (E_{mixDataset}(\mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{m}, \mathbf{s}, i))
-\end{equation}
-\begin{equation}
- E_{newdata}(\mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j]
-\end{equation}
-The mix $\mathbf{m}$ is now compressed as follows:
-\begin{equation}
- \mathbf{m}_c = E_{compress}(\mathbf{m}, -4)
-\end{equation}
-\begin{equation}
- E_{compress}(\mathbf{m}, i) = \begin{cases}
-\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i], \mathbf{m}[i + 1]), \mathbf{m}[i + 2]), \mathbf{m}[i + 3]), i + 4) & \text{otherwise}
-\end{cases}
-\end{equation}
-Finally the PoW-function can be defined as an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
-\begin{equation}
- \mathtt{PoW}(H, H_n, \mathbf{d}) = \lbrace  \mathbf{m}_c, \texttt{KEC}(\mathbf{s} + \mathbf{m}_c) \rbrace
-\end{equation}
-
 \end{document}
-
-C_\text{\tiny SSTORE}

### 2015-04-16 Paper.tex
header without nonce

@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+%\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
+\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
 Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \\
-& & \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
+V(H) & \equiv & \mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
 & & H_d = D(H) \quad \wedge \\
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathtt{PoW}(B^*, n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \\
+B'_n & = & n: \quad \mathtt{PoW}(B^*, n) < \frac{2^{256}}{H_d} \\
 B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
+For Ethereum 1.0 we have choosen the first path.
 
@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
-
-As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
-
-\subsubsection{PoC Series}
-
-For the PoC series, we use a simplified proof-of-work.
-This is not ASIC resistant and is meant merely as a placeholder.
-It utilises the bare Keccak hash function to secure the block chain by requiring the Keccak hash of the concatenation of the nonce and the header's Keccak hash to be sufficiently low.
-
-It is formally defined as $\mathtt{PoW}$:
-\begin{equation}
-\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny KEC}(\mathtt{\tiny KEC}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
-\end{equation}
-
-where:
-$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
-$\mathtt{\tiny KEC}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
-$n$ is the nonce, a series of 32 bytes;
-$\circ$ is the series concatenation operator;
-$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
-
-\subsubsection{Release Series}
-
-For the release series, we use a more complex proof-of-work.
-This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
-Secondly that it concerns the utilisation of either the blockchain or the full state trie.
-
-As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
-A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
-$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
-Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
-Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
-swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
-This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
-
-Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
-
-%\section{User Interface}
-%
-%Browser-style.
-%Displays ETH balance (i.e. functions as a wallet).
-%Tabs or perhaps MacOS style app icons.
-%Display taken up with a single app.
-%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
-%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
-%Preconfigured with home app that is the Eth app store.
-%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
+\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce component; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+the block difficulty from section \ref{ch:ghost}).
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+The underlying algorithm is called Ethash and is described below.
+\subsubsection{Ethash}
+Ethash is the planned PoW algorithm for Ethereum 1.0.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+The general route that the algorithm takes is as follows:
+
+There exists a seed which can be computed for each block by scanning through the block headers up until that point.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
+Light clients store the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+Full clients and miners store the dataset.
+The dataset grows linearly with time.
+
+Mining involves grabbing random slices of the dataset and hashing them together.
+Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
 

### 2015-04-16 Paper.tex

@@
 
+\section{Ethash}\label{app:ethash}
+\subsection{Definitions}
+We employ the following definitions:
+
+\begin{tabular*}{\columnwidth}[h]{lrl}
+\toprule
+Name & Value & Description \\
+\midrule
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\bottomrule
+\end{tabular*}
+
+\subsection{Size of dataset and cache}
+The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
+\begin{equation}
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
+x & \text{if} \quad x \in \mathbb{P} \\
+E_{prime}(x - y, y) & \text{otherwise}
+\end{cases}
+\end{equation}
+\subsection{Dataset generation}
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
+\subsubsection{Seed hash}
+The seed hash is different for every epoch.
+For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
+For every other epoch it is always the Keccak-256 hash of the previous seed hash:
+\begin{equation}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+\end{cases}
+\end{equation}
+With $\mathbf{0}_{32}$ being 32 bytes of zeros.
+
+\subsubsection{Cache}
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
+
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
+\begin{equation}
+ \mathbf{c}_{i} = \begin{cases}
+\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
+\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
+\end{cases}
+\end{equation}
+Therefore $ \mathbf{c'}$ can be defined as 
+\begin{equation}
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+\mathbf{x} & \text{if} \quad y = 0 \quad  \\
+E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+\end{cases}
+\end{equation}
+Where a single round modifies each subset of the cache as follows:
+\begin{equation}
+ E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n) \rbrace
+\end{equation}
+\begin{equation}
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n]) 
+\end{equation}
+\subsubsection{Full dataset calculation} \label{dataset} 
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
+\begin{equation}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{22}
+\end{equation}
+The single item of the dataset can now be calculated as:
+\begin{equation}
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Proof-of-work function}
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the nonce is valid.
+Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
+It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
+
+The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{h}), -1), -4)
+\end{equation}
+
+The seed hash being:
+\begin{equation}
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+\end{equation}
+We note that the ``$+$''-opertor between two byte sequences results in the concatenation of both sequences.
+
+The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
+
+The number of replicated sequences in the mix is:
+\begin{equation}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j]
+\end{equation}
+The mix is compressed as follows:
+\begin{equation}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i], \mathbf{m}[i + 1]), \mathbf{m}[i + 2]), \mathbf{m}[i + 3]), i + 4) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2015-04-16 Paper.tex
fix block transition function

--- 2015-04-16 Paper.tex
@@
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 1024 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mixHash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_n \; )\end{array} \\
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
@@
 H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+H_x \in \mathbb{B} & \wedge & H_n \in \mathbb{B}_{32}
 \end{array}
@@
 
-The nonce, $H_n$, must satisfy the relation:
-\begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
 \end{equation}
@@
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathtt{PoW}(B^*, n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \\
+B'_n & = & n: \quad \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \\
 B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))

### 2015-04-16 Paper.tex
add ref to DAG creation in block transition function
update size calculation of DAG and cache

--- 2015-04-16 Paper.tex
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the nonce requires a lot of memory and that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have choosen the first path.
 
@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce component; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
-\subsubsection{Ethash}
-Ethash is the planned PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
+\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
+
+As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
+
+\subsubsection{PoC Series}
+
+For the PoC series, we use a simplified proof-of-work.
+This is not ASIC resistant and is meant merely as a placeholder.
+It utilises the bare Keccak hash function to secure the block chain by requiring the Keccak hash of the concatenation of the nonce and the header's Keccak hash to be sufficiently low.
+
+It is formally defined as $\mathtt{PoW}$:
+\begin{equation}
+\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny KEC}(\mathtt{\tiny KEC}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
+\end{equation}
+
+where:
+$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
+$\mathtt{\tiny KEC}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
+$n$ is the nonce, a series of 32 bytes;
+$\circ$ is the series concatenation operator;
+$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
+
+\subsubsection{Release Series}
+
+For the release series, we use a more complex proof-of-work.
+This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
+Secondly that it concerns the utilisation of either the blockchain or the full state trie.
+
+As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
+A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
+$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
+Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
+Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
+swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
+This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
+
+Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
+
+%\section{User Interface}
+%
+%Browser-style.
+%Displays ETH balance (i.e. functions as a wallet).
+%Tabs or perhaps MacOS style app icons.
+%Display taken up with a single app.
+%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
+%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
+%Preconfigured with home app that is the Eth app store.
+%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
 
@@
 \begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch}, J_{mixbytes})
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch}, J_{hashbytes})
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
 \end{equation}
@@
  E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x \in \mathbb{P} \\
-E_{prime}(x - y, y) & \text{otherwise}
+x & \text{if} \quad x / y \in \mathbb{P} \\
+E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}

### 2015-04-16 Paper.tex
cache generation fix
add Keccak 512 hash function

@@
 For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
 
@@
 \begin{equation}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n]) 
-\end{equation}
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \quad \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+\end{equation}
+
 \subsubsection{Full dataset calculation} \label{dataset} 

### 2015-04-16 Paper.tex
typos and corrections

--- 2015-04-16 Paper.tex
@@
 \end{equation}
-\begin{equation}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x}[(i - 1 + n) \mod n] \oplus \mathbf{x}[\mathbf{x}[i][0] \mod n]) 
-\end{equation}
+\begin{multline}
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+\end{multline}
+
 \subsubsection{Full dataset calculation} \label{dataset} 
@@
 \begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{22}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{h}), -1), -4)
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
  E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-\mathbf{m} & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}), \mathbf{s}, i + 1) & \text{otherwise}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \end{equation}
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j]
+$E_{newdata}$ returns an array with $n_{mix}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}
@@
 \mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i], \mathbf{m}[i + 1]), \mathbf{m}[i + 2]), \mathbf{m}[i + 3]), i + 4) & \text{otherwise}
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
 \end{equation}

### 2015-04-16 Paper.tex

--- 2015-04-16 Paper.tex
@@
 
-The nonce, $H_n$, must satisfy the relation:
-\begin{equation}
-%\mathtt{PoW}(H, H_n) \dfrac{\leqslant}{den} \frac{2^{256}}{H_d}
-\mathtt{PoW}(H, H_n) \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an pseudo-random number cryptographically dependent on the parameters $H$ and $H_n$.
-Given an approximately uniform distribution in the range $[0, 2^{256})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathtt{PoW}(B^*, n) < \frac{2^{256}}{H_d} \\
+B'_n & = & n: \quad \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \\
 B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
+With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:dataset}.
 
@@
 \begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch}, J_{hashbytes})
 \end{equation}
@@
  E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+x & \text{if} \quad x \in \mathbb{P} \\
+E_{prime}(x - y, y) & \text{otherwise}
 \end{cases}

### 2015-04-16 Paper.tex
add parentHash formalization

--- 2015-04-16 Paper.tex
@@
 For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
-Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
 
@@
 H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_p &\equiv& \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(P(H))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
@@
 
-\section{Ethash}\label{app:ethash}
-\subsection{Definitions}
-We employ the following definitions:
-
-\begin{tabular*}{\columnwidth}[h]{lrl}
-\toprule
-Name & Value & Description \\
-\midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
-\bottomrule
-\end{tabular*}
-
-\subsection{Size of dataset and cache}
-The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
-\begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
-In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
-\end{cases}
-\end{equation}
-\subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
-\subsubsection{Seed hash}
-The seed hash is different for every epoch.
-For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
-For every other epoch it is always the Keccak-256 hash of the previous seed hash:
-\begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
-\end{cases}
-\end{equation}
-With $\mathbf{0}_{32}$ being 32 bytes of zeros.
-
-\subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
-The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
-
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
-\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
-\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
-\end{cases}
-\end{equation}
-Therefore $ \mathbf{c'}$ can be defined as 
-\begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
-\mathbf{x} & \text{if} \quad y = 0 \quad  \\
-E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
-\end{cases}
-\end{equation}
-Where a single round modifies each subset of the cache as follows:
-\begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n) \rbrace
-\end{equation}
-\begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
-\end{multline}
-
-\subsubsection{Full dataset calculation} \label{dataset} 
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
-\begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
-\end{equation}
-The single item of the dataset can now be calculated as:
-\begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\subsection{Proof-of-work function}
-Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the nonce is valid.
-Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
-It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
-
-The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
-\begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
-\end{equation}
-
-The seed hash being:
-\begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
-\end{equation}
-We note that the ``$+$''-opertor between two byte sequences results in the concatenation of both sequences.
-
-The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
-
-The number of replicated sequences in the mix is:
-\begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
-\end{equation}
-The mix is compressed as follows:
-\begin{equation}
- E_{compress}(\mathbf{m}, i) = \begin{cases}
-\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\subsection{Mining}
-In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
-\begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
-\end{equation}
-$H_d$ being the current difficulty of the block.
 \end{document}

### 2015-04-16 Paper.tex
update number of parameters

@@
 H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_p &\equiv& \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(P(H))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
@@
 
-The fee schedule $G$ is a tuple of 10 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
+The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
 

### 2015-04-16 Paper.tex
ceiling style

@@
 131072 & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{2048}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
 \max \{ 131072, x \} & \text{otherwise}\\
@@
 \begin{equation}
-x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{2048}\rfloor
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
 \end{equation}
@@
 \begin{eqnarray}
-& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
 & & H_l \geqslant 125000
@@
 & & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
 & & H_l \geqslant 125000  \quad \wedge \\
@@
 
-The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
+The fee schedule $G$ is a tuple of 10 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
 

### 2015-04-20 Paper.tex

--- 2015-04-16 Paper.tex
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{17}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{17}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.

### 2015-04-27 Paper.tex
mixHash -> mix-hash

@@
 \item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mixHash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mixHash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mixHash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.

### 2015-04-28 Paper.tex
list as return value of PoW function
update PoW return value

--- 2015-04-20 Paper.tex
@@
 131072 & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
+{P(H)_H}_d + \lfloor\frac{{P(H)_H}_d}{2048}\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
 \max \{ 131072, x \} & \text{otherwise}\\
@@
 \begin{equation}
-x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+x = {P(H)_H}_d - \lfloor\frac{{P(H)_H}_d}{2048}\rfloor
 \end{equation}
@@
 \begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor \quad \wedge \\
 & & H_l \geqslant 125000
@@
 & & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \lfloor\frac{{P(H)_H}_l}{1024}\rfloor  \quad \wedge \\
 & & H_l \geqslant 125000  \quad \wedge \\
@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
-
-As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
-
-\subsubsection{PoC Series}
-
-For the PoC series, we use a simplified proof-of-work.
-This is not ASIC resistant and is meant merely as a placeholder.
-It utilises the bare Keccak hash function to secure the block chain by requiring the Keccak hash of the concatenation of the nonce and the header's Keccak hash to be sufficiently low.
-
-It is formally defined as $\mathtt{PoW}$:
-\begin{equation}
-\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny KEC}(\mathtt{\tiny KEC}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
-\end{equation}
-
-where:
-$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
-$\mathtt{\tiny KEC}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
-$n$ is the nonce, a series of 32 bytes;
-$\circ$ is the series concatenation operator;
-$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
-
-\subsubsection{Release Series}
-
-For the release series, we use a more complex proof-of-work.
-This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
-Secondly that it concerns the utilisation of either the blockchain or the full state trie.
-
-As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
-A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
-$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
-Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
-Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
-swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
-This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
-
-Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
-
-%\section{User Interface}
-%
-%Browser-style.
-%Displays ETH balance (i.e. functions as a wallet).
-%Tabs or perhaps MacOS style app icons.
-%Display taken up with a single app.
-%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
-%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
-%Preconfigured with home app that is the Eth app store.
-%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
+m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+the block difficulty from section \ref{ch:ghost}).
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+The underlying algorithm is called Ethash and is described below.
+\subsubsection{Ethash}
+Ethash is the planned PoW algorithm for Ethereum 1.0.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+The general route that the algorithm takes is as follows:
+
+There exists a seed which can be computed for each block by scanning through the block headers up until that point.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
+Light clients store the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+Full clients and miners store the dataset.
+The dataset grows linearly with time.
+
+Mining involves grabbing random slices of the dataset and hashing them together.
+Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
 
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{17}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{17}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{22}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{22}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.

### 2015-04-28 Paper.tex

--- 2015-04-27 Paper.tex
@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
-\end{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv & \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \\
-& & \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \quad \wedge \\
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
 & & H_d = D(H) \quad \wedge \\
@@
 \end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d} \\
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
@@
 
+\section{Ethash}\label{app:ethash}
+\subsection{Definitions}
+We employ the following definitions:
+
+\begin{tabular*}{\columnwidth}[h]{lrl}
+\toprule
+Name & Value & Description \\
+\midrule
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\bottomrule
+\end{tabular*}
+
+\subsection{Size of dataset and cache}
+The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
+\begin{equation}
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
+x & \text{if} \quad x / y \in \mathbb{P} \\
+E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+\end{cases}
+\end{equation}
+\subsection{Dataset generation}
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
+\subsubsection{Seed hash}
+The seed hash is different for every epoch.
+For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
+For every other epoch it is always the Keccak-256 hash of the previous seed hash:
+\begin{equation}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+\end{cases}
+\end{equation}
+With $\mathbf{0}_{32}$ being 32 bytes of zeros.
+
+\subsubsection{Cache}
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
+
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
+\begin{equation}
+ \mathbf{c}_{i} = \begin{cases}
+\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
+\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
+\end{cases}
+\end{equation}
+Therefore $ \mathbf{c'}$ can be defined as 
+\begin{equation}
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+\mathbf{x} & \text{if} \quad y = 0 \quad  \\
+E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+\end{cases}
+\end{equation}
+Where a single round modifies each subset of the cache as follows:
+\begin{equation}
+ E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n) \rbrace
+\end{equation}
+\begin{multline}
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+\end{multline}
+
+\subsubsection{Full dataset calculation} \label{dataset} 
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
+\begin{equation}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
+\end{equation}
+The single item of the dataset can now be calculated as:
+\begin{equation}
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Proof-of-work function}
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the nonce is valid.
+Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
+It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
+
+The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+\end{equation}
+
+The seed hash being:
+\begin{equation}
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+\end{equation}
+We note that the ``$+$''-opertor between two byte sequences results in the concatenation of both sequences.
+
+The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
+
+The number of replicated sequences in the mix is:
+\begin{equation}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{newdata}$ returns an array with $n_{mix}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+\end{equation}
+The mix is compressed as follows:
+\begin{equation}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2015-04-28 Paper.tex

@@
 \begin{equation}
-\mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[0] = H_m \quad \wedge \quad \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce component; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).

### 2015-04-29 Paper.tex
remove ommer diff contribution to total diff

--- 2015-04-28 Paper.tex
@@
 
-This scheme notably ignores so-called \textit{stale} blocks: valid, mined blocks, which were propagated too late into the network and thus were beaten to network consensus by a sibling block (one with the same parent).
-Such blocks become more common as the network propagation time approaches the ideal inter-block time.
-However, by counting the computation work of stale block headers, we are able to do better: we can utilise this otherwise wasted computation and put it to use in helping to buttress the more popular blockchain making it a stronger choice over less popular (though potentially longer) competitors.
-
-This increases overall network security by making it much harder for an adversary to silently mine a canonical blockchain (which, it is assumed, would contain different transactions to the current consensus) and dump it on the network with the effect of overriding existing blocks and reversing the transactions within.
-
-In order to validate the extra computation, a given block $B$ may include the block headers from any known ommer blocks (i.e.
-blocks whose parent is equivalent to the $N$th generation grandparent of $B$, for a limited set of $N$).
-Since a block header includes the nonce, a proof-of-work, then the header alone is enough to validate the computation done.
-Any such blocks contribute toward the total computation or \textit{total difficulty} of a chain that includes them.
-To incentivise computation and inclusion, a reward is given both to the miner of the stale block and the miner of the block that references it.
+Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.
 
@@
 \begin{eqnarray}
-B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
+B_t & \equiv & B'_t + B_d \\
 B' & \equiv & P(B_H)
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of ommer blocks.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
 
@@
 \begin{equation}
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
-\subsubsection{Ethash}
-Ethash is the planned PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
+\mathtt{PoW}(H_{\hcancel{n}}, n) < \frac{2^{256}}{H_d}
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce component; $H_d$ is the new block's difficulty value (i.e. the block difficulty from section \ref{ch:ghost}).
+
+As of the proof-of-concept (PoC) series of the Ethereum software, the proof-of-work function is simplistic and does not attempt to secure these goals. It will be described here for completeness.
+
+\subsubsection{PoC Series}
+
+For the PoC series, we use a simplified proof-of-work.
+This is not ASIC resistant and is meant merely as a placeholder.
+It utilises the bare Keccak hash function to secure the block chain by requiring the Keccak hash of the concatenation of the nonce and the header's Keccak hash to be sufficiently low.
+
+It is formally defined as $\mathtt{PoW}$:
+\begin{equation}
+\mathtt{PoW}(H, n) \equiv \mathtt{BE}(\mathtt{\tiny KEC}(\mathtt{\tiny KEC}(\mathtt{\tiny RLP}(H_{\hcancel{n}})) \circ n))
+\end{equation}
+
+where:
+$\mathtt{\tiny RLP}(H_{\hcancel{n}})$ is the RLP encoding of the block header $H$, not including the final nonce component;
+$\mathtt{\tiny KEC}$ is the Keccak hash function accepting an arbitrary length series of bytes and evaluating to a series of 32 bytes (i.e. 256-bit);
+$n$ is the nonce, a series of 32 bytes;
+$\circ$ is the series concatenation operator;
+$\mathtt{BE}(X)$ evaluates to the value equal to X when interpreted as a big-endian-encoded integer.
+
+\subsubsection{Release Series}
+
+For the release series, we use a more complex proof-of-work.
+This has yet to be formally defined, but involves two components; firstly that it concerns the evaluation of programs on the EVM.
+Secondly that it concerns the utilisation of either the blockchain or the full state trie.
+
+As an overview, the output of the function is based upon the system state, defined as the hash of the root node of the state trie.
+A set of transactions, pseudo-randomly determined from the nonce value and selected from the last $N$ blocks is taken.
+$N$ is large enough and the selection criteria are such that execution of the transactions requires some non-negligible amount of processing by the EVM.
+Whenever code is executed on the EVM, it is pseudo-randomly (seeded again by the nonce) corrupted before alteration.
+Corruption could involve switching addresses with other transactions or rotating them through in the state trie (perhaps to the next address with the same order of magnitude of funds), rotating through instructions that have equivalent stack behaviour (e.g.
+swapping {\small ADD} for {\small SUB} or {\small GT} for {\small EQ}), or more destructive techniques such as randomly changing opcodes.
+This results in a problem that both require generalised computation hardware and is sequentially memory (and perhaps even disk) hard.
+
+Any specialised hardware to perform this task could also be leveraged to speed up (and thus drive down costs) of general Ethereum transaction processing.
+
+%\section{User Interface}
+%
+%Browser-style.
+%Displays ETH balance (i.e. functions as a wallet).
+%Tabs or perhaps MacOS style app icons.
+%Display taken up with a single app.
+%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
+%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
+%Preconfigured with home app that is the Eth app store.
+%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.
 

### 2015-05-12 Paper.tex
bugfixes
max stack size is 1024
more complex gas fees

@@
 Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE}, {\small EXP}, {\small LOG} and {\small SHA3}.
+The first and most common is the fee intrinsic to the computation of the operatio (see Appendix \ref{app:fees}).
 Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.

### 2015-05-13 Paper.tex
typo
fix block reward calculation
style
take the BLOCKHASH random number strategy into account

--- 2015-04-29 Paper.tex
@@
 Such data might include the block's hash, the block's timestamp and the block's coinbase address.
-For a series of such numbers, a trivial solution would be to work from the previous pseudo-random number, adding some constant amount and hashing the result.
-
-This strategy does have a downside: a miner with sufficient power could alter any of the above values in order to deliver a seed in order to alter the outcome of the pseudorandom-based executions.
-For a more secure pseudo-random offering, all involved parties could agree on a number of data feed contracts; these could be combined along with the block timestamp and hashed to produce the first number in the series.
-By spreading the inputs and thus the trust between numerous parties the likelihood of a malicious miner altering the outcome becomes increasingly less likely.
+In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
 

### 2015-05-13 Paper.tex
clearification
forgot to mention C_CALL as exception
µ_r is not defined and not needed. this is handeld in eq. 126 in sec. 9.4.4
whenever a stack item is used as an address, it should be assumed is is the stack item modulo 2^160

--- 2015-04-29 Paper.tex
@@
 
-TODO: Whenever a stack item is used as an address, it should be assumed it is the stack item modulo $2^{160}$.
-
 \begin{tabular*}{\columnwidth}[h]{rlrrl}
@@
 0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
-&&&& $\boldsymbol{\mu}'_R = []$ \\
 \midrule
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \end{tabular*}
@@
 \multicolumn{5}{c}{\textbf{10s: Comparison \& Bitwise Logic Operations}} \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[0]$} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
@@
 &&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1]] = \varnothing \\
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2015-05-18 Paper.tex

--- 2015-05-12 Paper.tex
@@
 Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operatio (see Appendix \ref{app:fees}).
+The first and most common is the fee intrinsic to the computation of the operation.
+Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE}, {\small EXP}, {\small LOG} and {\small SHA3}.
 Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\

### 2015-05-18 Paper.tex
expand lack of gas to exceptional halting

@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to lack of gas or if  \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
@@
 
+\section{Ethash}\label{app:ethash}
+\subsection{Definitions}
+We employ the following definitions:
+
+\begin{tabular*}{\columnwidth}[h]{lrl}
+\toprule
+Name & Value & Description \\
+\midrule
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\bottomrule
+\end{tabular*}
+
+\subsection{Size of dataset and cache}
+The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
+\begin{equation}
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
+x & \text{if} \quad x / y \in \mathbb{P} \\
+E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
+\end{cases}
+\end{equation}
+\subsection{Dataset generation}
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
+\subsubsection{Seed hash}
+The seed hash is different for every epoch.
+For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
+For every other epoch it is always the Keccak-256 hash of the previous seed hash:
+\begin{equation}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+\end{cases}
+\end{equation}
+With $\mathbf{0}_{32}$ being 32 bytes of zeros.
+
+\subsubsection{Cache}
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
+
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
+\begin{equation}
+ \mathbf{c}_{i} = \begin{cases}
+\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
+\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
+\end{cases}
+\end{equation}
+Therefore $ \mathbf{c'}$ can be defined as 
+\begin{equation}
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+\mathbf{x} & \text{if} \quad y = 0 \quad  \\
+E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+\end{cases}
+\end{equation}
+Where a single round modifies each subset of the cache as follows:
+\begin{equation}
+ E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \rbrace
+\end{equation}
+\begin{multline}
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+\end{multline}
+
+\subsubsection{Full dataset calculation} \label{dataset} 
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
+\begin{equation}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
+\end{equation}
+The single item of the dataset can now be calculated as:
+\begin{equation}
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Proof-of-work function}
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the nonce is valid.
+Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
+It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
+
+The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+\end{equation}
+
+The seed hash being:
+\begin{equation}
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+\end{equation}
+We note that the ``$+$''-opertor between two byte sequences results in the concatenation of both sequences.
+
+The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
+
+The number of replicated sequences in the mix is:
+\begin{equation}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{newdata}$ returns an array with $n_{mix}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+\end{equation}
+The mix is compressed as follows:
+\begin{equation}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
+\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2015-05-18 Paper.tex

@@
 The memory model is a simple word-addressed byte array.
-The stack has an unlimited size.
+The stack has a maximum size of $1024$.
 The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
@@
 Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operation.
-Most operations require a single gas fee to be paid for their execution; exceptions include {\small SSTORE}, {\small SLOAD}, {\small CALL}, {\small CREATE}, {\small BALANCE}, {\small EXP}, {\small LOG} and {\small SHA3}.
+The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
 Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.

### 2015-05-18 Paper.tex

@@
 T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
-\end{array}\dfrac{num}{den}
+\end{array}
 \end{equation}
@@
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 - \frac{1}{8} (U_i - {B_H}_i)) R_b 
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert - 1} b_n256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 \end{cases} \\
-f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \\ 0 & \text{otherwise} \end{cases}
+f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \neq 0 \\ 0 & \text{otherwise} \end{cases}
 \end{eqnarray}

### 2015-05-18 Paper.tex

@@
 
-and $C_\text{\tiny SSTORE}$ is specified in the appropriate section below. We define the following subsets of instructions:
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
 
@@
 For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
-
-TODO: Whenever a stack item is used as an address, it should be assumed it is the stack item modulo $2^{160}$.
 
@@
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\

### 2015-06-01 Paper.tex
Clarify CALL's output behaviour.
correct reference og ethash

--- 2015-05-18 Paper.tex
@@
 \end{eqnarray}
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:dataset}.
+With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $\mathbf{o} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + \boldsymbol{\mu}_\mathbf{s}[6] - 1) ]$ \\
 &&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\

### 2015-06-04 Paper.tex
update memory gas cost to PoC9
correct linear memory cost limit

@@
 \end{eqnarray}
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
+With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:dataset}.
 
@@
 \begin{equation}
-C_{memory}(a) \equiv G_{memory}(a + \Big\lfloor \dfrac{a^2}{1024} \Big\rfloor)
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}
@@
 Note also that $C_{memory}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 32KB of memory used, after which it costs substantially more.
+It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
 

### 2015-06-04 Paper.tex
fix ambigious PUSH

--- 2015-06-01 Paper.tex
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
-&&&& where $c$ is defined as above. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
-&&&& where $c$ is defined as above. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& where $\boldsymbol{c}$ is defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\

### 2015-06-04 Paper.tex
forbid uncles to be reused
truncated SDIV

--- 2015-06-01 Paper.tex
@@
 \midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\

### 2015-06-04 Paper.tex
update genesis block

--- 2015-06-04 Paper.tex
@@
 \begin{equation}
-C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{memory}(a) \equiv G_{memory}(a + \Big\lfloor \dfrac{a^2}{1024} \Big\rfloor)
 \end{equation}
@@
 Note also that $C_{memory}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
+It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 32KB of memory used, after which it costs substantially more.
 
@@
 
-The genesis block is 13 items, and is specified thus:
-\begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{17}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{17}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
+The genesis block is 15 items, and is specified thus:
+\begin{equation}
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
 $\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
-The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$. The latest documentation should be consulted for the value of the state root.
+The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
+Also $time$ will be set to the intial timestamp of the genesis block.
+The latest documentation should be consulted for those values.
 

### 2015-06-04 Paper.tex
correct limits for copying data/code

--- 2015-06-04 Paper.tex
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
 \midrule
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
-&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& where $c$ is defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
-&&&& where $\boldsymbol{c}$ is defined as above. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& where $c$ is defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\

### 2015-06-04 Paper.tex
specify calldataload out of bounds behaviour
add boundaries to block parameter size due to vm opcodes

--- 2015-06-04 Paper.tex
@@
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P}_{256} & \wedge & H_i \in \mathbb{P}_{256} & \wedge \\
+H_l \in \mathbb{P}_{256} & \wedge & H_g \in \mathbb{P}_{256} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_x \in \mathbb{B}_{\leqslant 1024} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
@@
 
-The genesis block is 15 items, and is specified thus:
-\begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+The genesis block is 13 items, and is specified thus:
+\begin{equation}
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 2^{17}, 0, 0, 3141592, 0, 0, 0, \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $2^{17}$ refers to the difficulty; 0 refers to the timestamp (the Unix epoch); the transaction trie root and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42.
 $\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
-The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the intial timestamp of the genesis block.
-The latest documentation should be consulted for those values.
+The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$. The latest documentation should be consulted for the value of the state root.
 

### 2015-06-11 Paper.tex

@@
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P}_{256} & \wedge & H_i \in \mathbb{P}_{256} & \wedge \\
-H_l \in \mathbb{P}_{256} & \wedge & H_g \in \mathbb{P}_{256} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B}_{\leqslant 1024} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}

### 2015-06-11 Paper.tex

@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U)
-\end{equation}
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
+\end{equation}
+where $B(H)$ is the block of the corresponding header $H$.
 

### 2015-06-11 Paper.tex

@@
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\

### 2015-06-23 Paper.tex
update precompiled contracts gas cost

@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 500\\
+g_r &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 50 + 50\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 50 + 50\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 1 + \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}

### 2015-07-07 Paper.tex
32 bytes in extradata only.

--- 2015-06-11 Paper.tex
@@
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 1024 bytes or fewer; formally $H_x$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
 \item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
@@
 & & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 1024
+& & \lVert H_x \rVert \le 32
 \end{eqnarray}
@@
 
-Noting additionally that \textbf{extraData} must be at most 1024 bytes.
+Noting additionally that \textbf{extraData} must be at most 32 bytes.
 
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
-\end{equation}
-where $B(H)$ is the block of the corresponding header $H$.
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U)
+\end{equation}
 
@@
 0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+&&&& $\boldsymbol{\mu}'_R = []$ \\
 \midrule
@@
 \midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$
 \end{tabular*}
@@
 \multicolumn{5}{c}{\textbf{10s: Comparison \& Bitwise Logic Operations}} \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[0]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[0]$} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
@@
 &&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1]] = \varnothing \\
 0 & \text{otherwise}

### 2015-07-25 Paper.tex
Update YP to new constants. Fixes #144.

--- 2015-06-23 Paper.tex
@@
 $G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 2 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
 
@@
 G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{ext} + G_{ext}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\

### 2015-07-26 Paper.tex
Minor terminology alteration.

@@
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
@@
 
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{coinbase} address, the address of an account typically under the control of the miner.
+In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
 Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
@@
 
-The Ether for the gas is given to the miner, whose address is specified as the coinbase of the present block $B$.
+The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
 So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
@@
 
-The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each ommer by a certain amount.
-We raise the block's coinbase account by $R_b$; for each ommer, we raise the block's coinbase by an additional $\frac{1}{32}$ of the block reward and the coinbase of the ommer gets rewarded depending on the block number.
+The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 
-If there are collisions of the coinbase addresses between ommers and the block (i.e.
-two ommers with the same coinbase address or an ommer with the same coinbase address as the present block), additions are applied cumulatively.
+If there are collisions of the beneficiary addresses between ommers and the block (i.e.
+two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
 
@@
 However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's coinbase address.
+Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
 In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
@@
 \midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
+0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
@@
 
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the coinbase address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.

### 2015-09-08 Paper.tex
update block reward to 5 ether
set minimal gasLimit to 5000
Timestamps now bounded to < 2^256.
Typo fix, thanks David G Anderson.

@@
 H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
 H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
@@
 131072 & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 8\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
 \max \{ 131072, x \} & \text{otherwise}\\
@@
 & & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 125000
+& & H_l \geqslant 5000
 \end{eqnarray}
@@
 
-We define the block reward as 1500 Finney:
-\begin{equation}
-\text{Let} \quad R_b = 1.5 \times 10^{18}
+We define the block reward as 5 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 5 \times 10^{18}
 \end{equation}
@@
 \end{equation}
-We note that the ``$+$''-opertor between two byte sequences results in the concatenation of both sequences.
+We note that the ``$+$''-operator between two byte sequences results in the concatenation of both sequences.
 
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}

### 2016-01-08 Paper.tex
Correct secp256k1n constant.
Remove redundant clauses.

@@
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_J(\mathbf{c, 0}) \cap Y(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
 \end{equation}
@@
 where:
-\begin{equation}
-Y(\mathbf{c}, i) \equiv \begin{cases}
-\{\} & \text{if} \quad i \geqslant |\mathbf{c}| \\
-\{ i \} \cup Y(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
-\end{cases}
-\end{equation}
 \begin{equation}

### 2016-01-21 Paper.tex
Update Paper.tex
My first commit
careful replacement of SHA3 with KECCAK
Initial step for https://github.com/ethereum/EIPs/issues/59

@@
 %\begin{itemize}
-%\item {\small SHA3} costs $G_{sha3}$ gas
+%\item {\small KECCAK} costs $G_{keccak}$ gas
 %\item {\small SLOAD} costs $G_{sload}$ gas
@@
 %
-%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small KECCAK}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
 
@@
 $G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{keccak}$ & 30 & Paid for each {\small KECCAK} operation. \\
+$G_{keccakword}$ & 6 & Paid for each word (rounded up) for input data to a {\small KECCAK} operation. \\
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{keccak}+G_{keccakword} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK}\\
 G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
@@
 \toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{20s: KECCAK}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
+0x20 & {\small KECCAK} & 2 & 1 & Compute Keccak-256 hash. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\

### 2016-02-12 Paper.tex
Fix two minor typos

--- 2016-01-08 Paper.tex
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+L_R(R) \equiv (\mathtt{TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
-m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
+m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
 \end{eqnarray}
@@
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_r &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
@@
 \begin{equation}
-p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
+p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
 \end{equation}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\mathtt{TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_H)) = H_p
 \end{equation}
@@
 & & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
+& & H_l \geqslant 125000
 \end{eqnarray}
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by the up-front cost, $v_0$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
 The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}^*[s]_b - v
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{TRIE}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{KEC}(()), \mathtt{TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
 \boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
@@
 (\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad a = 4 \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_a & \equiv & a \\
+I_a & \equiv & r \\
 I_o & \equiv & o \\
 I_p & \equiv & p \\
-I_d & \equiv & d \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
@@
 I_e & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \; \mathtt{KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H), 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H), n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_u$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
 \begin{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{|B_\mathbf{U}|}{32})R_b \\
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_i: \mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have choosen the first path.
+For Ethereum 1.0 we have chosen the first path.
 
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
 Light clients store the cache.
@@
 
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+Thus $\mathtt{BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
 
@@
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
 \end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
+s(\mathbf{x}) & \equiv & \mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) ...
 \end{eqnarray}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\mathtt{RLP}(i : i \in \mathbb{P}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
 \end{equation}
@@
 
-It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
+It is defined as the function $\mathtt{HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
+
+\begin{eqnarray}
+\mathtt{HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
 (16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
@@
 For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
+Importantly in the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
 \begin{eqnarray}
@@
 g_r &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
 \text{if} \quad |\mathbf{o}| = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
 \mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
@@
 g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
@@
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_\mathbf{d})\\
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
+\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
 \end{eqnarray}
@@
 
-in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+in the range $(1, \mathtt{secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{ECDSASIGN}$, $\mathtt{ECDSARESTORE}$ and $\mathtt{ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 \begin{eqnarray}
-0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
 v \in \{27,28\}
@@
 \begin{eqnarray}
-\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
 \end{eqnarray}
@@
 \begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_r) \big) \big)
 \end{equation}
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+h(T) & \equiv & \mathtt{KEC}( L_S(T) )
 \end{eqnarray}
@@
 G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+(T_w, T_r, T_s) = \mathtt{ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
 \end{equation}
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
 \end{equation}
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
@@
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
+E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \lbrace E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \rbrace
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
 \end{equation}
@@
 Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by $d_{size}$ items, each $J_{hashbytes}$ bytes in size:
+The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
 \begin{equation}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
 \end{equation}

### 2016-02-15 Paper.tex
Fix two minor typos
Various typos fixed thanks to 看K线图.

--- 2016-01-08 Paper.tex
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}^*[s]_b - v
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)

### 2016-02-21 Paper.tex
Updated difficulty update equation for Homestead from EIP 2
Updated signature variable conditions for Homestead from EIP 2

@@
 
+\newcommand{\firsthomesteadblock}{\ensuremath{\mathit{TBA}}}
+
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
-% to avoid line break in range definition
-
-in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$(1, \mathtt{\tiny secp256k1n} - 1)$}.
 
@@
 
-We declare that a signature is invalid unless the following is true:
-\begin{eqnarray}
-0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-v \in \{27,28\}
-\end{eqnarray}
-
+\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
+
+We declare that a signature is invalid unless all the following conditions are true:
+\begin{align}
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \begin{dcases}
+\hphantom{\lfloor} \mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
+\left\lfloor \mathtt{\tiny secp256k1n} \div 2 \right\rfloor & \text{otherwise} \\
+\end{dcases} \\
+ v &\in \{27,28\} 
+\end{align}
 where:
-\begin{eqnarray}
-\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
-\end{eqnarray}
+\begin{align}
+\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
+\end{align}
 

### 2016-02-21 Paper.tex
Added ref command to unflagged occurance of 'TBD' in relation to "homestead block number"
Added new command for referencing "firsthomesteadblock" number
Removed incorrect appearance of 'floor' operator on secp256k1n div 2 in eq. 207
Corrected range brackets on signature variable 'r'

@@
 
+\newcommand{\mindifficulty}{D_0}
+\newcommand{\frontiermod}{\ensuremath{\varsigma_1}}
+\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
+\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
+\newcommand{\diffadjustment}{x}
+
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{cases}
-131072 & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
-\max \{ 131072, x \} & \text{otherwise}\\
-\end{cases}
+D(H) \equiv \begin{dcases}
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\firsthomesteadblock\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\end{dcases}
 \end{equation}
@@
 \begin{equation}
-x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\mindifficulty \equiv 131072
+\end{equation}
+\begin{equation}
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\frontiermod \equiv \begin{cases}
+1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
+-1 & \text{otherwise} \\
+\end{cases}
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
 \end{equation}
@@
 
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$(1, \mathtt{\tiny secp256k1n} - 1)$}.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
 
@@
 0 < s &< \begin{dcases}
-\hphantom{\lfloor} \mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\left\lfloor \mathtt{\tiny secp256k1n} \div 2 \right\rfloor & \text{otherwise} \\
+\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
+\mathtt{\tiny secp256k1n} \div 2 & \text{otherwise} \\
 \end{dcases} \\

### 2016-02-21 Paper.tex
Added the EIP 2 increase in gas cost for contract-creation transactions

--- 2016-02-15 Paper.tex
@@
 
+\newcommand{\mindifficulty}{D_0}
+\newcommand{\frontiermod}{\ensuremath{\varsigma_1}}
+\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
+\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
+\newcommand{\diffadjustment}{x}
+
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{cases}
-131072 & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
-\max \{ 131072, x \} & \text{otherwise}\\
-\end{cases}
+D(H) \equiv \begin{dcases}
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\mathit{TBD}\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\end{dcases}
 \end{equation}
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{equation}
-g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
-\end{equation}
+\begin{align}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge T_o = S(T) \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
+\end{align}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G$ is defined in Appendix \ref{app:fees}.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}.
+$G$ is fully defined in Appendix \ref{app:fees}.
 
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\

### 2016-02-21 Paper.tex
Changed the contract-creation process to not having any transfer of value if the contract wasn't created

@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge T_o = S(T) \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}.
-$G$ is fully defined in Appendix \ref{app:fees}.
+$G$ is defined in Appendix \ref{app:fees}.
 
@@
 
-Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-The gas remaining should be zero in any such exceptional condition.
-If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
 \begin{equation}
@@
 \end{equation}
-\begin{eqnarray}
-\quad g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{if} \quad g^{**} < c \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-Where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
 \begin{eqnarray}
@@
 
-where $c$ is the code-deposit cost:
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+
+Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+
+If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
 \begin{equation}
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
-$I_H$ has no special treatment and is determined from the blockchain.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception, and discard all the effects on the state as described above.
+
+The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
+However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas, whichever way it came about.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
 
-No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
+Note that the intention here is to have either the result of a successfully created new contract with its endowment, or no new contract with no transfer of value.
 
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\

### 2016-02-21 Paper.tex
Added pre-Homestead contract-creation quirks

@@
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 g^{**} - c & \text{otherwise} \\
@@
 \boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{*} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 
-Note that the intention here is to have either the result of a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Note that the intention from block \firsthomesteadblock onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Before {\it Homestead}, if there is not enough gas to pay $c$, no contract is created but the value is transferred.
 

### 2016-02-21 Paper.tex
Formatting and corrections

--- 2016-02-21 Paper.tex
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
 
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception, and discard all the effects on the state as described above.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception.
 
 The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas, whichever way it came about.
+However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas.
 
@@
 
-\begin{eqnarray}
-\quad g' & \equiv & \begin{cases}
+\begin{align}
+\quad g' &\equiv \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 g^{**} - c & \text{otherwise} \\
 \end{cases} \\
-\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
 \boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \end{cases}
-\end{eqnarray}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
-
-Note that the intention here is to have either the result of a successfully created new contract with its endowment, or no new contract with no transfer of value.
+\end{align}
+
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
+
+Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
 

### 2016-02-22 Paper.tex
trial

--- 2016-02-21 Paper.tex
@@
 
-\newcommand{\firsthomesteadblock}{\ensuremath{\mathit{TBA}}}
-
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
 \begin{equation}
@@
 \end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**} < c \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+Where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
 \begin{eqnarray}
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+where $c$ is the code-deposit cost:
 \begin{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception, and discard all the effects on the state as described above.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas, whichever way it came about.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
-\begin{eqnarray}
-\quad g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
 
-Note that the intention here is to have either the result of a successfully created new contract with its endowment, or no new contract with no transfer of value.
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 The exceptional halting function $Z$ is defined as:
-\begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
-\begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
-\end{array}
-\end{equation}
+%\begin{equation}
+%Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+%\begin{array}[t]{l}
+%\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+%\mathbf{\delta}_w = \varnothing \quad \vee \\
+%\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+%( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+%\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+%\end{array}
+%\end{equation}
+
+\begin{gather} 
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\ 
+\mathbf{\delta}_w = \varnothing \quad \vee   \nonumber \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \nonumber \\
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \nonumber
+\end{gather}
 

### 2016-02-22 Paper.tex
Removed superfluous condition

@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{equation}
-g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
-\end{equation}
+\begin{align}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
+\end{align}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G$ is defined in Appendix \ref{app:fees}.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}.
+$G$ is fully defined in Appendix \ref{app:fees}.
 
@@
 The exceptional halting function $Z$ is defined as:
-%\begin{equation}
-%Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
-%\begin{array}[t]{l}
-%\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-%\mathbf{\delta}_w = \varnothing \quad \vee \\
-%\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-%( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-%\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
-%\end{array}
-%\end{equation}
-
-\begin{gather} 
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\ 
-\mathbf{\delta}_w = \varnothing \quad \vee   \nonumber \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \nonumber \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \nonumber
-\end{gather}
+\begin{equation}
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+\begin{array}[t]{l}
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+\end{array}
+\end{equation}
 
@@
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
+For Ethereum 1.0 we have choosen the first path.
 
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
 Light clients store the cache.

### 2016-02-24 Paper.tex
Fix minor typo

@@
 
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\frontiermod}{\ensuremath{\varsigma_1}}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\mathit{TBD}\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
 \end{equation}
@@
 \begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\frontiermod \equiv \begin{cases}
-1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
--1 & \text{otherwise} \\
-\end{cases}
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
 \end{equation}
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}.
-$G$ is fully defined in Appendix \ref{app:fees}.
+$G$ is defined in Appendix \ref{app:fees}.
 
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{TRIE}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)

### 2016-02-24 Paper.tex
Fix minor typo

--- 2016-02-22 Paper.tex
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:

### 2016-02-26 Paper.tex

--- 2016-02-22 Paper.tex
@@
 \mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\mathit{TBD}\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\firsthomesteadblock\\
 \text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
@@
 
-It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
-% to avoid line break in range definition
-
-in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
 
@@
 
-We declare that a signature is invalid unless the following is true:
-\begin{eqnarray}
-0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-v \in \{27,28\}
-\end{eqnarray}
-
+\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
+
+We declare that a signature is invalid unless all the following conditions are true:
+\begin{align}
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \begin{dcases}
+\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
+\mathtt{\tiny secp256k1n} \div 2 & \text{otherwise} \\
+\end{dcases} \\
+ v &\in \{27,28\} 
+\end{align}
 where:
-\begin{eqnarray}
-\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
-\end{eqnarray}
+\begin{align}
+\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
+\end{align}
 

### 2016-02-26 Paper.tex

@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
-
-Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-The gas remaining should be zero in any such exceptional condition.
-If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
 
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
-$I_H$ has no special treatment and is determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
-
-No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception.
+
+The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
+However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{align}
+\quad g' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{align}
+
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
+
+Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
 

### 2016-02-26 Paper.tex
Alter name to Homestead.
Addition of DELEGATECALL opcode with inclusion of an extra argument to the message call evaluation function \Theta

--- 2016-02-24 Paper.tex
@@
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \mathbf{d}, e)
-\end{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+\end{equation}
+Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
 
@@
 I_s & \equiv & s \\
-I_v & \equiv & v \\
+I_v & \equiv & \tilde{v} \\
 I_e & \equiv & e \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overridden altered.\\
+&&&& same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\midrule
+0xf4 & {\small DELEGATECALL} & 7 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+&&&& the current values for {\it sender} and {\it value}. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
+&&&& parameters to the call $\Theta$.\\
+&&&& This means that the recipient is in fact the same account as at present, simply\\
+&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
 \midrule
@@
 0 & \text{otherwise}
-\end{cases}$ \\\bottomrule
+\end{cases}$ \\
+\bottomrule
 \end{tabular*}

### 2016-03-01 Paper.tex

--- 2016-02-26 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead Draft}}}
 \author{

### 2016-03-10 Paper.tex
first round of edits -- typos, and inconsistency

@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\qquad (\boldsymbol{\sigma}^{**}, g', A, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the transaction substate $A$, and the resultant output $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 \begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
+O\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\

### 2016-03-17 Paper.tex
typo, v --> w in transaction signature information

@@
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+\item[w, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
 This is expanded in Appendix \ref{app:signing}.

### 2016-04-06 Paper.tex
Added paragraph about the Homestead transition and defined the transition block number

--- 2016-03-01 Paper.tex
@@
 
-\newcommand{\firsthomesteadblock}{\ensuremath{\mathit{TBA}}}
+\newcommand{\firsthomesteadblock}{\ensuremath{N_H}}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead Draft}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
@@
 \end{equation}
+
+\subsection{Homestead} \label{ch:homestead}
+
+A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
+\begin{equation}
+\firsthomesteadblock \equiv 1,\! 150,\! 000
+\end{equation}
+The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
 
@@
 \mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\firsthomesteadblock\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\mathit{TBD}\\
 \text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
@@
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
 
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
-\end{equation}
-Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \mathbf{d}, e)
+\end{equation}
 
@@
 I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
+I_v & \equiv & v \\
 I_e & \equiv & e \\
@@
 
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
 
@@
 
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28\} 
-\end{align}
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 7 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
-&&&& parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
 \midrule
@@
 0 & \text{otherwise}
-\end{cases}$ \\
-\bottomrule
+\end{cases}$ \\\bottomrule
 \end{tabular*}

### 2016-04-23 Paper.tex

--- 2016-03-10 Paper.tex
@@
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g', A, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the transaction substate $A$, and the resultant output $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 \begin{eqnarray}
-O\big(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
+O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\

### 2016-05-30 Paper.tex
Minor typo
Added a missing verb in a sentence

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead Draft}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}}
 \author{
     Dr. Gavin Wood\\
-    Co-Founder \& Lead, Ethereum Project\\
-    gavin@ethereum.org
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path (known as blockchain) from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure is to be considered canonical, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

### 2016-05-30 Paper.tex
Added a missing verb in a sentence

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead Draft}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes places as a state-transition function, adding value to a nominated account.
+This incentivisation takes place as a state-transition function, adding value to a nominated account.
 

### 2016-05-31 Paper.tex
Typo, missing verb

--- 2016-04-23 Paper.tex
@@
 
-\newcommand{\firsthomesteadblock}{\ensuremath{N_H}}
-
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead Draft}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
     Dr. Gavin Wood\\
-    Co-Founder \& Lead, Ethereum Project\\
-    gavin@ethereum.org
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 \end{equation}
-
-\subsection{Homestead} \label{ch:homestead}
-
-A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
-\begin{equation}
-\firsthomesteadblock \equiv 1,\! 150,\! 000
-\end{equation}
-The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
 
@@
 
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\frontiermod}{\ensuremath{\varsigma_1}}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\firsthomesteadblock\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
 \end{equation}
@@
 
-There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
+I_v & \equiv & v \\
 I_e & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \; \mathtt{KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
@@
 
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{ECDSASIGN}$, $\mathtt{ECDSARESTORE}$ and $\mathtt{ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28\} 
-\end{align}
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously existing account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\

### 2016-05-31 Paper.tex
Typo, missing verb

--- 2016-05-30 Paper.tex
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path (known as blockchain) from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure is to be considered canonical, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

### 2016-05-31 Paper.tex
Added for clarity a reference to the structural composition function c that gets only defined later

@@
 
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
+We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure (see also formal definition \ref{eq:structuralcomposition} for the structural composition function $c$):
 \begin{equation}
@@
 A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
-\begin{equation}
+\begin{equation} \label{eq:structuralcomposition}
 c(\mathfrak{I}, i) \equiv \begin{cases}

### 2016-05-31 Paper.tex
Changed the letter to represent the set of bytes from Y to O, since Y is already used nearby for the set of nibble sequences.

--- 2016-05-31 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
@@
 \begin{equation}
-\boldsymbol{\sigma}_1'[s]_b \equiv \boldsymbol{\sigma}_1[s]_b - v
+\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
 \end{equation}
@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
-\end{eqnarray}
-
-Where $\mathbb{Y}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
+\end{eqnarray}
+
+Where $\mathbb{O}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
 
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
-Unlike a radix tree, in the case of multiple keys shared the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
+Unlike a radix tree, in the case of multiple keys sharing the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
 Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.

### 2016-05-31 Paper.tex
Changed the letter to represent the set of bytes from Y to O, since Y is already used nearby for the set of nibble sequences.

--- 2016-05-31 Paper.tex
@@
 They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes place as a state-transition function, adding value to a nominated account.
+This incentivisation takes places as a state-transition function, adding value to a nominated account.
 
@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
-\end{eqnarray}
-
-Where $\mathbb{Y}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
+\end{eqnarray}
+
+Where $\mathbb{O}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.

### 2016-05-31 Paper.tex
Minor typo

@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+
+Where $\mathbb{Y}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.

### 2016-06-01 Paper.tex
Typo (missed update) on bloom size

@@
 
-Hashcash, introduced by \cite{back2002hashcash} (in a five-year retrospective), provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
-Though not widely adopted, the work was later utilised and expanded upon by \cite{nakamoto2008bitcoin} in order to devise a cryptographically secure mechanism for coming to a decentralised social consensus over the order and contents of a series of cryptographically signed financial transactions.
-The fruits of this project, Bitcoin, provided a first glimpse into a decentralised transaction ledger.
+\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet.
+The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}.
+\cite{back2002hashcash} later produced a system in a similar vein.
+
+The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
+In this instance, the token was used to keep peer-to-peer file trading in check, ensuring ``consumers'' be able to make micro-payments to ``suppliers'' for their services.
+The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
+Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope.
+The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.
 
@@
 
+Sometimes, a path follows a new protocol from a particular height.
+This document describes one version of the protocol.
+In order to follow back the history of a path, one might need to reference multiple versions of this document.
+
 \section{Conventions}\label{ch:conventions}
@@
 
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
 For particularly important values, a bold uppercase letter may be used.
@@
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
 \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_t \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
@@
 
-We define the Bloom filter function, $M$, to reduce a log entry include a single 256-byte hash:
+We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
 \begin{equation}
@@
 
-where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
 Formally:
@@
 These are used to help define the transaction receipt, discussed later.
-
-%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
-%\begin{eqnarray}
-%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
-%\end{eqnarray}
 
@@
 \end{equation}
+unless $s = r$.
 
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
-\end{eqnarray}
-
-Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \subsubsection{Ethash}
-Ethash is the planned PoW algorithm for Ethereum 1.0.
+Ethash is the PoW algorithm for Ethereum 1.0.
 It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
@@
 
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
+
+\section{Availability}
+
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 
@@
 The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
 
@@
 
-The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
+The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
 \begin{equation}
@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparision. \\
+0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparision. \\
+0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
+0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
+0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparision. \\
+0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 \multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{and to update the memory consumption counter:}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 \midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the intial timestamp of the genesis block.
+Also $time$ will be set to the initial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
-The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
-
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
+The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
+
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
 \begin{equation}
@@
 
-\subsection{Mining}
-In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
-\begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
-\end{equation}
-$H_d$ being the current difficulty of the block.
 \end{document}

### 2016-06-02 Paper.tex
Typo (missed update) on bloom size
minor grammar fix "a number of"

--- 2016-05-31 Paper.tex
@@
 
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure (see also formal definition \ref{eq:structuralcomposition} for the structural composition function $c$):
+We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
 \begin{equation}
@@
 The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
-Unlike a radix tree, in the case of multiple keys sharing the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
+Unlike a radix tree, in the case of multiple keys shared the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
 Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
@@
 A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
-\begin{equation} \label{eq:structuralcomposition}
+\begin{equation}
 c(\mathfrak{I}, i) \equiv \begin{cases}

### 2016-06-02 Paper.tex
minor grammar fix "a number of"

--- 2016-05-31 Paper.tex
@@
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously existing account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\

### 2016-06-07 Paper.tex
Update my details.

--- 2016-06-02 Paper.tex
@@
 
-\newcommand{\firsthomesteadblock}{\ensuremath{N_H}}
+\newcommand{\firsthomesteadblock}{\ensuremath{\mathit{TBA}}}
 
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 

### 2016-06-09 Paper.tex

@@
 
-where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
 It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.

### 2016-06-22 Paper.tex
"takes place" instead of "takes places"

@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
 It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1'[s]_b \equiv \boldsymbol{\sigma}_1[s]_b - v
 \end{equation}

### 2016-07-09 Paper.tex
Fix δ for DELEGATECALL

@@
 
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
+We define the Bloom filter function, $M$, to reduce a log entry include a single 256-byte hash:
 \begin{equation}
@@
 \midrule
-0xf4 & {\small DELEGATECALL} & 7 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
 &&&& the current values for {\it sender} and {\it value}. \\

### 2016-07-22 Paper.tex
Correct definition of C_callgas

@@
 \boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-0 & \text{otherwise}
+\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
 \end{cases}$ \\
 \midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\

### 2016-07-27 Paper.tex
Correct definition of C_callgas
Add delegatecall to VM specs

--- 2016-07-12 Paper.tex
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\

### 2016-08-05 Paper.tex
Fix grammatical typo

@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 \midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\

### 2016-09-09 Paper.tex
Remove the modulo 2^{256} effect in the memory size computation
This commit adds notes to MLOAD, MSTORE and MSTORE8 specification
saying the modulo is not considered when computing the new memory
size.

Before this change, since the beginning of the table contains
`All arithmetic is modulo $2^{256}$ unless otherwise noted.`,
the memory size computation was affected by the modulo computation too.
For instance, MLOAD with the biggest uint256 didn't change the memory size
most of the time.

@@
 \midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\

### 2016-09-09 Paper.tex
Remove the modulo 2^{256} effect in the memory size computation
This commit adds notes to MLOAD, MSTORE and MSTORE8 specification
saying the modulo is not considered when computing the new memory
size.

Before this change, since the beginning of the table contains
`All arithmetic is modulo $2^{256}$ unless otherwise noted.`,
the memory size computation was affected by the modulo computation too.
For instance, MLOAD with the biggest uint256 didn't change the memory size
most of the time.

@@
 \midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule

### 2016-09-22 Paper.tex

@@
 \boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-0 & \text{otherwise}
+\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
 \end{cases}$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\

### 2016-09-23 Paper.tex
Correct the historical record.

--- 2016-09-22 Paper.tex
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
-&&&& parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
 \midrule

### 2016-09-23 Paper.tex
Correct the historical record.

--- 2016-09-22 Paper.tex
@@
 \boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
+0 & \text{otherwise}
 \end{cases}$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 \midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+0xf4 & {\small DELEGATECALL} & 7 & 1 & Message-call into this account with an alternative account's code, but persisting\\
 &&&& the current values for {\it sender} and {\it value}. \\

### 2016-09-29 Paper.tex
Fix DELEGATECALL semantics
Before this commit, the semantics of DELEGATECALL referred to
`\mu_s[2]`.  However, this argument is omitted from DELEGATECALL.
Instead, the semantics should refer to `I_v`, which is the current
"value".

@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
-&&&& parameters to the call $\Theta$.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second,\\
+&&&& eighth and ninth parameters to the call $\Theta$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\

### 2016-09-29 Paper.tex
Explain the fewer arguments of DELEGATECALL

@@
 &&&& the current values for {\it sender} and {\it value}. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& Otherwise exactly equivalent to {\small CALL} except: \\
 &&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second,\\
-&&&& eighth and ninth parameters to the call $\Theta$.\\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
+&&&& and ninth parameters to the call $\Theta$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\

### 2016-09-29 Paper.tex
Nitpicking equation (100)
Before this commit, equation (100) was not satisfiable when s and r
were equal and v + v was not zero.  Though this equation is further
complemented in equations (101) through (104), I thought (100) needed
a condition s != r.

--- 2016-09-29 Paper.tex
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second,\\
-&&&& eighth and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
 \midrule

### 2016-09-29 Paper.tex
Nitpicking equation (100)
Before this commit, equation (100) was not satisfiable when s and r
were equal and v + v was not zero.  Though this equation is further
complemented in equations (101) through (104), I thought (100) needed
a condition s != r.

--- 2016-09-29 Paper.tex
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second,\\
-&&&& eighth and ninth parameters to the call $\Theta$.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
+&&&& parameters to the call $\Theta$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\

### 2016-10-05 Paper.tex
Add a parentheses in SMOD definition
This solves #190.  The other way of putting parentheses
did not make sense becasue `sign(x) * |x|` can immediately
be optimized into `x`.

Also, I checked one EVM implementation's behavior
```
evm --debug --code "60601960401907"
```
and saw SMOD returning a negative-looking number.

--- 2016-09-29 Paper.tex
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule

### 2016-10-05 Paper.tex
Add a parentheses in SMOD definition
This solves #190.  The other way of putting parentheses
did not make sense becasue `sign(x) * |x|` can immediately
be optimized into `x`.

Also, I checked one EVM implementation's behavior
```
evm --debug --code "60601960401907"
```
and saw SMOD returning a negative-looking number.

--- 2016-09-29 Paper.tex
@@
 
+\input{Version.tex}
+
 \newcommand{\hcancel}[1]{%
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 
-The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
@@
 
-For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
 \begin{equation}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
 
 The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas.
+However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
 
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SUICIDE} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas (OOG) exception, they do not leave state changes intact.
+Like the out-of-gas exception, they do not leave state changes intact.
 Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
@@
 
-%More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
-%
-%\begin{itemize}
-%\item {\small SHA3} costs $G_{sha3}$ gas
-%\item {\small SLOAD} costs $G_{sload}$ gas
-%\item {\small BALANCE} costs $G_{balance}$ gas
-%\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
-%\begin{itemize}
-%\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
-%\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
-%\item $d = 1$ otherwise.
-%\end{itemize}
-%\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
-%\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
-%\item {\small STOP} costs $G_{stop}$ gas
-%\item {\small SUICIDE} costs $G_{suicide}$ gas
-%\item All other operations cost $G_{step}$ gas.
-%\end{itemize}
-%
-%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
-
 See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
-
-%Whenever a higher memory index is referenced, the fee difference to take it to the higher usage from the original (lower) usage is charged.
-Notably, because {\small MSTORE} and {\small MLOAD} operate on word lengths, they implicitly increase the highest-accessed index to 31 greater than their target index.
 
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
 \begin{equation}
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
 \varnothing & \text{otherwise}
@@
 
-In general, we assume the memory, suicide list and system state don't change:
+In general, we assume the memory, self-destruct set and system state don't change:
 \begin{eqnarray}
@@
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10000th block) could be maintained by the peer network, effectively recasting the genesis block.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
+Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
+
+\section{Availability}
+
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 
@@
 \mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 & \text{otherwise} \\
+\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
 \end{dcases} \\
@@
 $G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
-$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
+$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
 $G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously existing account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{memory}(\boldsymbol{\mu}'_i)-C_{memory}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 G_{high} & \text{if} \quad w \in W_{high}\\
-G_{ext} & \text{if} \quad w \in W_{ext}
+G_{extcode} & \text{if} \quad w \in W_{extcode}\\
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
 \end{cases}
@@
 \begin{equation}
-C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
+C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
 
@@
 
-$W_{ext}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small BLOCKHASH}\}
+$W_{extcode}$ = \{{\small EXTCODESIZE}\}
 
@@
 
-Note also that $C_{memory}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 \end{cases}
+\end{equation}
+
+Another useful function is ``all but one 64th'' function~$L$ defined as:
+
+\begin{equation}
+L(n) \equiv n - \lfloor n / 64 \rfloor
 \end{equation}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\end{cases}$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
 G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
-\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
 \end{cases}$ \\
@@
 &&&& the current values for {\it sender} and {\it value}. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
-&&&& parameters to the call $\Theta$.\\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& Otherwise exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
+&&&& and ninth parameters to the call $\Theta$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\
@@
 \midrule
-0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
+\midrule
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2016-10-12 Paper.tex
CODECOPY updates the memory usage

--- 2016-10-05 Paper.tex
@@
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+While is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
 \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
@@
 These are used to help define the transaction receipt, discussed later.
+
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
 
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
+Ethash is the planned PoW algorithm for Ethereum 1.0.
 It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
@@
 
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
 
@@
 
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
 \begin{equation}
@@
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
+0x10 & {\small LT} & 2 & 1 & Less-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
+0x11 & {\small GT} & 2 & 1 & Greater-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
+0x12 & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
+0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
+0x14 & {\small EQ} & 2 & 1 & Equality comparision. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
@@
 \multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
-\multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the initial timestamp of the genesis block.
+Also $time$ will be set to the intial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
-The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
-
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
+The intial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
+
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the intial cache:
 \begin{equation}

### 2016-10-12 Paper.tex
CODECOPY updates the memory usage

--- 2016-10-05 Paper.tex
@@
 
-\input{Version.tex}
-
 \newcommand{\hcancel}[1]{%
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}}
 \author{
@@
 
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
@@
 
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
 \begin{equation}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception.
 
 The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
+However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas.
 
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+If the initialisation execution ends with a {\small SUICIDE} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 \end{equation}
-unless $s = r$.
 
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas exception, they do not leave state changes intact.
+Like the out-of-gas (OOG) exception, they do not leave state changes intact.
 Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
@@
 
+%More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
+%
+%\begin{itemize}
+%\item {\small SHA3} costs $G_{sha3}$ gas
+%\item {\small SLOAD} costs $G_{sload}$ gas
+%\item {\small BALANCE} costs $G_{balance}$ gas
+%\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
+%\begin{itemize}
+%\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
+%\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
+%\item $d = 1$ otherwise.
+%\end{itemize}
+%\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
+%\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
+%\item {\small STOP} costs $G_{stop}$ gas
+%\item {\small SUICIDE} costs $G_{suicide}$ gas
+%\item All other operations cost $G_{step}$ gas.
+%\end{itemize}
+%
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+
 See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
+
+%Whenever a higher memory index is referenced, the fee difference to take it to the higher usage from the original (lower) usage is charged.
+Notably, because {\small MSTORE} and {\small MLOAD} operate on word lengths, they implicitly increase the highest-accessed index to 31 greater than their target index.
 
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
 \begin{equation}
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
 \varnothing & \text{otherwise}
@@
 
-In general, we assume the memory, self-destruct set and system state don't change:
+In general, we assume the memory, suicide list and system state don't change:
 \begin{eqnarray}
@@
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10000th block) could be maintained by the peer network, effectively recasting the genesis block.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 
-Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
 
@@
 \mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
+\mathtt{\tiny secp256k1n} \div 2 & \text{otherwise} \\
 \end{dcases} \\
@@
 $G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
 $G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously existing account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{memory}(\boldsymbol{\mu}'_i)-C_{memory}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+G_{ext} & \text{if} \quad w \in W_{ext}
 \end{cases}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
 
@@
 
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
+$W_{ext}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small BLOCKHASH}\}
 
@@
 
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_{memory}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 \end{cases}
-\end{equation}
-
-Another useful function is ``all but one 64th'' function~$L$ defined as:
-
-\begin{equation}
-L(n) \equiv n - \lfloor n / 64 \rfloor
 \end{equation}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
 G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
 \end{cases}$ \\
@@
 &&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second and eighth\\ 
+&&&& parameters to the call $\Theta$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\
@@
 \midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}

### 2016-10-12 Paper.tex
CODECOPY and CALLDATACOPY update the memory usage

@@
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 \bottomrule

### 2016-10-12 Paper.tex
Add link to README

@@
 
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
 

### 2016-10-12 Paper.tex
Add link to README

@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
 \midrule

### 2016-10-12 Paper.tex
Split the instruction class W_{ext} into W_{extcode} and W_{extbal}

--- 2016-10-12 Paper.tex
@@
 $G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
+$G_{extbal}$ & 400 & Amount of gas to pay for operations of the set {\small $W_{extbal}$}. \\
 $G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
@@
 G_{high} & \text{if} \quad w \in W_{high}\\
-G_{ext} & \text{if} \quad w \in W_{ext}
+G_{extcode} & \text{if} \quad w \in W_{extcode}\\
+G_{extbal} & \text{if} \quad w \in W_{extbal}
 \end{cases}
@@
 
-$W_{ext}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small BLOCKHASH}\}
+$W_{extcode}$ = \{{\small EXTCODESIZE}, {\small BLOCKHASH}\}
+
+$W_{extbal}$ = \{{\small BALANCE}\}
 

### 2016-10-12 Paper.tex
Change the basic gas cost of CALL, CALLCODE and DELEGATECALL to 700

@@
 $G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{extbal}$ & 400 & Amount of gas to pay for operations of the set {\small $W_{extbal}$}. \\
-$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
+$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
 $G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
@@
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
@@
 G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{extbal} & \text{if} \quad w \in W_{extbal}
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}
 \end{cases}
@@
 $W_{extcode}$ = \{{\small EXTCODESIZE}, {\small BLOCKHASH}\}
-
-$W_{extbal}$ = \{{\small BALANCE}\}
 

### 2016-10-12 Paper.tex
Increase the gas cost of SUICIDE to 5000

@@
 $R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$G_{suicide}$ & 5000 & Amount of gas to pay for a {\small SUICIDE} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+G_{suicide} & \text{if} \quad w = \text{\small SUICIDE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
 

### 2016-10-12 Paper.tex
Refactor the definition of BALANCE gas cost
When CALL gets a too big gas argument, it tries to cap the argument
The gas cost of CALL is capped at all but one 64th
CREATE only provides all but one 64th of the parent gas

@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-G_{suicide} & \text{if} \quad w = \text{\small SUICIDE} \\
+C_\text{\tiny SUICIDE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SUICIDE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 
-with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SUICIDE}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
 
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
+C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+L(\boldsymbol{\mu}_g - C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})) & \text{if} \quad L(\boldsymbol{\mu}_g - C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})) < \boldsymbol{\mu}_{\mathbf{s}}[0] \;\wedge \\
+&\quad\quad \boldsymbol{\mu}_g \ge C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\end{cases}$\\
+&&&& $C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
@@
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
-\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
 \end{cases}$ \\
@@
 0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{suicide} + G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{suicide} & \text{otherwise}
 \end{cases}$ \\

### 2016-10-15 Paper.tex
Fixes and improvements to EIP-150 spec
- Condense and simplify the `CALL` gas costs.
- Correct `BLOCKHASH` gas cost.

@@
 
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
+Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
 
@@
 G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
 \end{cases}
@@
 \begin{equation}
-C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}
@@
 
-$W_{extcode}$ = \{{\small EXTCODESIZE}, {\small BLOCKHASH}\}
+$W_{extcode}$ = \{{\small EXTCODESIZE}\}
 
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAPPED}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-L(\boldsymbol{\mu}_g - C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})) & \text{if} \quad L(\boldsymbol{\mu}_g - C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})) < \boldsymbol{\mu}_{\mathbf{s}}[0] \;\wedge \\
-&\quad\quad \boldsymbol{\mu}_g \ge C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny CALLEXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
 G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\

### 2016-10-15 Paper.tex
Cleanup `SUICIDE` instruction.

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Homestead revision}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}}}
 \author{
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously existing account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SUICIDE} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{suicide} + G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
-G_{suicide} & \text{otherwise}
+&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{suicide} + \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+0 & \text{otherwise}
 \end{cases}$ \\

### 2016-11-25 Paper.tex
Update the fee schedule following EIP 160
https://github.com/ethereum/EIPs/issues/160

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
+
+\section{Availability}
+
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
-$G_{suicide}$ & 5000 & Amount of gas to pay for a {\small SUICIDE} operation. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SUICIDE} operation which creates an account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SUICIDE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SUICIDE} \\
+C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SUICIDE}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
 
@@
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{suicide} + \begin{cases}
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\

### 2016-11-25 Paper.tex
Update the fee schedule following EIP 160
https://github.com/ethereum/EIPs/issues/160

@@
 Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg be things such as reducing an account balance without an equal and opposite increase elsewhere.
+Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
 A valid state transition is one which comes about through a transaction.
@@
 
-%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
-%\begin{eqnarray}
-%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
-%\end{eqnarray}
-
 \section{Contract Creation} \label{ch:create}
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will defined it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}

### 2016-11-25 Paper.tex
Update the fee schedule following EIP 160
https://github.com/ethereum/EIPs/issues/160

@@
 \newcommand{\mindifficulty}{D_0}
-\newcommand{\frontiermod}{\ensuremath{\varsigma_1}}
 \newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
@@
 \mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\firsthomesteadblock\\
 \text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
@@
 \diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\frontiermod \equiv \begin{cases}
-1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
--1 & \text{otherwise} \\
-\end{cases}
 \end{equation}
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 g^{**} - c & \text{otherwise} \\
@@
 \boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 
-Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
+Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
 v &\in \{27,28\}
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2016-12-01 Paper.tex
EIP170: max code size
EIP170
small typo

--- 2016-11-25 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}}}
 \author{
@@
 Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$G_{suicide}$ & 5000 & Amount of gas to pay for a {\small SUICIDE} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SUICIDE} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+C_\text{\tiny SUICIDE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SUICIDE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SUICIDE}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
 
@@
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{suicide} + \begin{cases}
 G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\

### 2016-12-02 Paper.tex
small typo (#196)

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
@@
 
-Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
 
@@
 $R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
-$G_{suicide}$ & 5000 & Amount of gas to pay for a {\small SUICIDE} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SUICIDE} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously existing account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SUICIDE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SUICIDE} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SUICIDE}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
 
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
 \midrule
@@
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{suicide} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\bottomrule
+\end{cases}$ \\\bottomrule
 \end{tabular*}

### 2016-12-02 Paper.tex
EIP155

--- 2016-11-25 Paper.tex
@@
 
+\newcommand{\firsthomesteadblock}{\ensuremath{N_H}}
+
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
-In order to follow back the history of a path, one might need to reference multiple versions of this document.
-
 \section{Conventions}\label{ch:conventions}
@@
 \newcommand{\mindifficulty}{D_0}
+\newcommand{\frontiermod}{\ensuremath{\varsigma_1}}
 \newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
@@
 \mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\frontiermod + \expdiffsymb \right) & \text{if} \quad H_i<\firsthomesteadblock\\
 \text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
@@
 \diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\frontiermod \equiv \begin{cases}
+1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
+-1 & \text{otherwise} \\
+\end{cases}
 \end{equation}
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \wedge H_i \geq \firsthomesteadblock \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code or before the {\it Homestead transition}.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
+\begin{align}
+c \equiv \begin{cases}
+  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24000 \\
+  G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}
+\end{cases}
+\end{align}
 
@@
 \boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
 
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
+\end{eqnarray}
+
+Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
 
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is either the `recovery id' or `chain id doubled plus 35 or 36'.
+The recovery id is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The chain id is a constant natural number.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
+0 < s &< \begin{dcases}
+\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
+\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
+\end{dcases} \\
+ v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+Two different flavors of signing schemes are available.
+One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+The other operates on nine elements:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0 \wedge v \in \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{if} \; T_t \neq 0 \wedge v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; T_t = 0 \wedge v \notin \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}
 \end{cases} \\
@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2016-12-02 Paper.tex
EIP155: replay protection

@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{align}
-c \equiv \begin{cases}
-  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24000 \\
-  G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}
-\end{cases}
-\end{align}
+\begin{equation}
+c \equiv G_{codedeposit} \times |\mathbf{o}|
+\end{equation}
 
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery id' or `chain id doubled plus 35 or 36'.
-The recovery id is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The chain id is a constant natural number.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The chain identifier is a constant natural number.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
 v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0 \wedge v \in \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{if} \; T_t \neq 0 \wedge v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; T_t = 0 \wedge v \notin \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; v \notin \{27, 28\}\\
+\end{cases} \\
+\nonumber \text{where} \\
+\nonumber \mathbf{p} & \equiv & \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
@@
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\

### 2016-12-02 Paper.tex
EIP155: replay protection

@@
 
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg $\Upsilon$, the Ethereum state transition function.
 
@@
 Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
+Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
+
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
 For particularly important values, a bold uppercase letter may be used.
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
 Formally:
@@
 & & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 125000  \quad \wedge \\
+& & H_l \geqslant 5000  \quad \wedge \\
 & & H_s > {P(H)_H}_s \quad \wedge \\
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad F \\
 g^{**} - c & \text{otherwise} \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
 \boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
+For the present context, we will define it as:
 \begin{equation}
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger than 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
@@
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
+E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+Essentially, we maintain a ``mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
 $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.

### 2016-12-02 Paper.tex
EIP155

--- 2016-12-02 Paper.tex
@@
 Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
+Invalid state changes might, \eg be things such as reducing an account balance without an equal and opposite increase elsewhere.
 A valid state transition is one which comes about through a transaction.
@@
 
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
 \section{Contract Creation} \label{ch:create}
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}

### 2016-12-05 Paper.tex
EIP161a: change the initial nonce of an account

--- 2016-12-02 Paper.tex
@@
 
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
@@
 The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The chain identifier is a constant natural number.
+The chain identifier $\mathtt{\tiny chain\_id}$ is a constant natural number.
 
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; v \notin \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
-\end{equation}
+\begin{eqnarray}
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
+v_0 &\equiv& \begin{cases}
+T_w &\text{if}\ T_w\in\{27, 28\} \\
+28 - (T_w \bmod 2) &\text{otherwise}
+\end{cases}
+\end{eqnarray}
 

### 2016-12-05 Paper.tex
EIP161a: change the initial nonce of an account

--- 2016-12-05 Paper.tex
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery id' or `chain id doubled plus 35 or 36'.
-The recovery id is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The chain id is a constant natural number.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
-The other operates on nine elements:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0 \wedge v \in \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{if} \; T_t \neq 0 \wedge v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; T_t = 0 \wedge v \notin \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases} \\

### 2016-12-05 Paper.tex
EIP 161: Define when an account state is empty

@@
 
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \equiv \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+
 \subsection{The Transaction} \label{ch:transaction}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The chain identifier $\mathtt{\tiny chain\_id}$ is a constant natural number.
 
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+Two different flavors of signing schemes are available.
+One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+The other operates on nine elements:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+\end{cases} \\
+\nonumber \text{where} \\
+\nonumber \mathbf{p} & \equiv & \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\

### 2016-12-05 Paper.tex
Define when an account state is empty

--- 2016-12-05 Paper.tex
@@
 \quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+An account state is empty when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\equiv \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
+\wedge \boldsymbol{\sigma}[a]_n = 0
+\wedge \boldsymbol{\sigma}[a]_b = 0
 \end{equation}

### 2016-12-05 Paper.tex
EIP 161: Define when an account state is empty
EIP 161: Define when an account is dead
Define when an account is dead

@@
 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+
+An account is dead when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
+\equiv \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}

### 2016-12-05 Paper.tex
EIP161b: CALL and SUICIDE gas change

@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma},\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
 0 & \text{otherwise}

### 2016-12-05 Paper.tex
EIP161b: CALL and SUICIDE gas change

--- 2016-12-05 Paper.tex
@@
 
-An account state is empty when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\equiv \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
-\wedge \boldsymbol{\sigma}[a]_n = 0
-\wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-
-\subsection{Homestead} \label{ch:homestead}
-
-A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
-\begin{equation}
-\firsthomesteadblock \equiv 1,\! 150,\! 000
-\end{equation}
-The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \equiv \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \equiv \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
 
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
 0 & \text{otherwise}

### 2016-12-05 Paper.tex
EIP161.c

--- 2016-12-05 Paper.tex
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[s] &\equiv& \begin{cases}
+\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
+\mathbf{a}^* & \text{otherwise}
+\end{cases} \\
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
 \end{eqnarray}
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
+\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
+\mathbf{a}_1 &\text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
-\end{cases}
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
 \end{equation}
@@
 \begin{eqnarray}
+\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
@@
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
+\mathbf{a}' &\text{otherwise}
+\end{cases} \\
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
+(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
+\end{cases}$\\
+&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
+&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
 \end{cases}$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule

### 2016-12-05 Paper.tex
EIP161.c

--- 2016-12-05 Paper.tex
@@
 \end{equation}
+
+Here is a function that updates the balance of an account state but deletes the account state when it becomes empty:
+\begin{equation}
+\rho_b(\boldsymbol{\sigma}, a, x) \equiv
+\begin{cases}
+\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
+(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
+\end{cases}
+\end{equation}
+for $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+\boldsymbol{\sigma}^*[S(T)] & \equiv & \rho_b(\boldsymbol{\sigma}_P, S(T), \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p) \\
+\boldsymbol{\sigma}^*[m] & \equiv & \rho_b(\boldsymbol{\sigma}_P, m, \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p) \\
 m & \equiv & {B_H}_c
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[s] &\equiv& \rho_b(\boldsymbol{\sigma}, s, \boldsymbol{\sigma}[s]_b - v)
 \end{eqnarray}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1[s] \equiv \rho_b(\boldsymbol{\sigma}_1', s, \boldsymbol{\sigma}_1'[s]_b - v)
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + v) & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
+\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\\ \nonumber
+\boldsymbol{\sigma}'[{B_H}_c] & \equiv & \rho_b(\boldsymbol{\sigma}, {B_H}_c, \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b) \\
+\\ \nonumber
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+ \boldsymbol{\sigma}'[U_c] & \equiv & \rho_b(\boldsymbol{\sigma}, U_c, \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b)
 \end{eqnarray}
@@
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $\boldsymbol{\sigma}'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b)$ where $r \equiv \boldsymbol{\mu}_s[0]\mod 2^{160}$\\
+&&&& $\boldsymbol{\sigma}'[I_a] \equiv \rho_b(\boldsymbol{\sigma}, I_a, 0)$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}

### 2016-12-05 Paper.tex
EIP161.c

--- 2016-12-05 Paper.tex
@@
 
-Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
 
@@
 Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
+Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
 For particularly important values, a bold uppercase letter may be used.
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_t \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
 Formally:
@@
 & & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 5000  \quad \wedge \\
+& & H_l \geqslant 125000  \quad \wedge \\
 & & H_s > {P(H)_H}_s \quad \wedge \\
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} - c & \text{otherwise} \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\end{cases}
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
 \boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
@@
 The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will define it as:
+For the present context, we will defined it as:
 \begin{equation}
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger than 1024.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
 \end{eqnarray}
@@
 The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
 
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The chain identifier $\mathtt{\tiny chain\_id}$ is a constant natural number.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+v &\in \{27,28\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
-The other operates on nine elements:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases} \\
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_w &\text{if}\ T_w\in\{27, 28\} \\
-28 - (T_w \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
 
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
+E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
 $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.

### 2016-12-05 Paper.tex
Keep the touched accounts in the substate

--- 2016-12-05 Paper.tex
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
 \end{equation}
@@
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
@@
 
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), 0)
+For brevity, we define the initial substate $A^0$ to have no self-destructs, no logs, a set of touched accounts and a zero refund balance:
+\begin{equation}
+A^0(T) \equiv (\varnothing, (), T, 0) \quad\text{for}\quad T \subseteq \mathbb{B}_{20}
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
 \end{equation}
@@
 (\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
+T   & \equiv & \{s, r\} \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I, T)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0(T), I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}, A^0(\emptyset), I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 \begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
+(\varnothing, 0, A^0(\emptyset), ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0(T), \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\

### 2016-12-05 Paper.tex
EIP 161: Keep the touched accounts in the substate

--- 2016-12-05 Paper.tex
@@
 
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), 0)
+We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), \{\}, 0)
 \end{equation}
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
 \boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
@@
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
+T & \equiv & \{s, r\} \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, r)
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
@@
 \begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
 (\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\

### 2016-12-05 Paper.tex
Keep the touched accounts in the substate

--- 2016-12-05 Paper.tex
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I, T)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will defined it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
 \end{equation}

### 2016-12-06 Paper.tex
EIP161.d empty, touched accounts are deleted at the end of a transaction

@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I, T)
 \end{equation}

### 2016-12-06 Paper.tex
EIP161.d empty, touched accounts are deleted at the end of a transaction

--- 2016-12-05 Paper.tex
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
 &&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\

### 2016-12-06 Paper.tex
Precompiled contracts are touched even when out-of-gas
Change the revision

--- 2016-12-05 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-170 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0(\emptyset), ()) & \text{if} \quad g < g_r \\
+(\varnothing, 0, A^0(T), ()) & \text{if} \quad g < g_r \\
 (\boldsymbol\sigma, g - g_r, A^0(T), \mathbf{o}) & \text{otherwise}\end{cases}

### 2016-12-09 Paper.tex
Change the revision
Define the balance update function that also performs the cleanup

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-170 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
 \end{equation}
@@
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
@@
 
-For brevity, we define the initial substate $A^0$ to have no self-destructs, no logs, a set of touched accounts and a zero refund balance:
-\begin{equation}
-A^0(T) \equiv (\varnothing, (), T, 0) \quad\text{for}\quad T \subseteq \mathbb{B}_{20}
+For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)] & \equiv & \rho_b(\boldsymbol{\sigma}_P, S(T), \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p) \\
-\boldsymbol{\sigma}^*[m] & \equiv & \rho_b(\boldsymbol{\sigma}_P, m, \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p) \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
 m & \equiv & {B_H}_c
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
 \end{eqnarray}
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &\equiv& \rho_b(\boldsymbol{\sigma}, s, \boldsymbol{\sigma}[s]_b - v)
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
 \end{equation}
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \rho_b(\boldsymbol{\sigma}_1', s, \boldsymbol{\sigma}_1'[s]_b - v)
+\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + v) & \text{otherwise}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
 \end{cases}
@@
 (\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-T   & \equiv & \{s, r\} \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0(T), I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0(\emptyset), I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 \begin{eqnarray}
-\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\\ \nonumber
-\boldsymbol{\sigma}'[{B_H}_c] & \equiv & \rho_b(\boldsymbol{\sigma}, {B_H}_c, \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b) \\
-\\ \nonumber
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c] & \equiv & \rho_b(\boldsymbol{\sigma}, U_c, \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b)
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 \begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0(T), ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0(T), \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b)$ where $r \equiv \boldsymbol{\mu}_s[0]\mod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a] \equiv \rho_b(\boldsymbol{\sigma}, I_a, 0)$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}

### 2016-12-09 Paper.tex
EIP 161: Improve local layouts

--- 2016-12-09 Paper.tex
@@
 
-An account state is empty when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\equiv \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
-\wedge \boldsymbol{\sigma}[a]_n = 0
-\wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-
-An account is dead when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
-\equiv \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-Here is a function that updates the balance of an account state but deletes the account state when it becomes empty:
-\begin{equation}
-\rho_b(\boldsymbol{\sigma}, a, x) \equiv
-\begin{cases}
-\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
-(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
-\end{cases}
-\end{equation}
-for $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
-
-\subsection{Homestead} \label{ch:homestead}
-
-A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
-\begin{equation}
-\firsthomesteadblock \equiv 1,\! 150,\! 000
-\end{equation}
-The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
 
@@
 \begin{equation}
-\frontiermod \equiv \begin{cases}
-1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
--1 & \text{otherwise} \\
-\end{cases}
-\end{equation}
-\begin{equation}
 \homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+0 & \text{if} \quad F \\
 g^{**} - c & \text{otherwise} \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
-\end{cases}
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
 \end{equation}
@@
 g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\end{cases} \\ \nonumber
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
+\mathbf{t} & \equiv & \{s, r\} \\
+\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery id' or `chain id doubled plus 35 or 36'.
-The recovery id is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The chain id is a constant natural number.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The chain identifier $\mathtt{\tiny chain\_id}$ is a constant natural number.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0 \wedge v \in \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{if} \; T_t \neq 0 \wedge v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; T_t = 0 \wedge v \notin \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+\end{cases} \\
+\nonumber \text{where} \\
+\nonumber \mathbf{p} & \equiv & \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
 \end{cases}$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule

### 2016-12-09 Paper.tex
EIP 161: Improve local layouts

@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
 &&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\

### 2016-12-09 Paper.tex
Improve local layouts

--- 2016-12-09 Paper.tex
@@
 \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\equiv \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
+\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
 \wedge \boldsymbol{\sigma}[a]_n = 0
@@
 \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
-\equiv \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
+\mathbf{t}   & \equiv & \{s, r\} \\
+\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will defined it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
 \end{equation}
@@
 \begin{eqnarray}
+\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\\ \nonumber
+\boldsymbol{\sigma}'[{B_H}_c] & \equiv & \rho_b(\boldsymbol{\sigma}, {B_H}_c, \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b) \\
+\\ \nonumber
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
+ \boldsymbol{\sigma}'[U_c] & \equiv & \rho_b(\boldsymbol{\sigma}, U_c, \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b)
 \end{eqnarray}
@@
 \begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
+(\varnothing, 0, A^0(T), ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0(T), \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}

### 2016-12-09 Paper.tex
Improve local layouts

@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+\\ \nonumber
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-\mathbf{t}   & \equiv & \{s, r\} \\
+T   & \equiv & \{s, r\} \\
 \\ \nonumber

### 2016-12-15 Paper.tex
correct secp256k1n off-by-one for homestead blocks

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-170 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}}}
 \author{
@@
 
-An account state is empty when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
-\wedge \boldsymbol{\sigma}[a]_n = 0
-\wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-
-An account is dead when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
-\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-Here is a function that updates the balance of an account state but deletes the account state when it becomes empty:
-\begin{equation}
-\rho_b(\boldsymbol{\sigma}, a, x) \equiv
-\begin{cases}
-\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
-(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
-\end{cases}
-\end{equation}
-for $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
-
 \subsection{Homestead} \label{ch:homestead}
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
@@
 
-For brevity, we define the initial substate $A^0$ to have no self-destructs, no logs, a set of touched accounts and a zero refund balance:
-\begin{equation}
-A^0(T) \equiv (\varnothing, (), T, 0) \quad\text{for}\quad T \subseteq \mathbb{B}_{20}
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
 \end{eqnarray}
@@
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &\equiv& \rho_b(\boldsymbol{\sigma}, s, \boldsymbol{\sigma}[s]_b - v)
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{align}
-c \equiv \begin{cases}
-  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24000 \\
-  G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}
-\end{cases}
-\end{align}
-
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
+\begin{equation}
+c \equiv G_{codedeposit} \times |\mathbf{o}|
+\end{equation}
+
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an Out-of-Gas exception.
 
 The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
+However, the value of the transaction is not transferred to the aborted contract's address when we are Out-of-Gas.
 
@@
 \end{cases} \\
-\\ \nonumber
 (\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-T   & \equiv & \{s, r\} \\
-\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I, T)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 \begin{eqnarray}
-\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\\ \nonumber
-\boldsymbol{\sigma}'[{B_H}_c] & \equiv & \rho_b(\boldsymbol{\sigma}, {B_H}_c, \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b) \\
-\\ \nonumber
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c] & \equiv & \rho_b(\boldsymbol{\sigma}, U_c, \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b)
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 \end{dcases} \\
- v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
+ v &\in \{27,28\} 
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
-The other operates on nine elements:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0 \wedge v \in \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{if} \; T_t \neq 0 \wedge v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; T_t = 0 \wedge v \notin \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
 \end{cases} \\
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma},\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b)$ where $r \equiv \boldsymbol{\mu}_s[0]\mod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a] \equiv \rho_b(\boldsymbol{\sigma}, I_a, 0)$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
+&&&& $C_{\text{\tiny SUICIDE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{suicide} + \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2016-12-22 Paper.tex
Fix the memory consumption of CODECOPY

@@
 
-\subsection{Homestead} \label{ch:homestead}
-
-A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
-\begin{equation}
-\firsthomesteadblock \equiv 1,\! 150,\! 000
-\end{equation}
-The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
-
 \subsection{The Transaction} \label{ch:transaction}
@@
 
-Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
+
+\section{Availability}
+
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 
@@
 
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28\} 
-\end{align}
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
-\end{equation}
-
-The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
 \begin{equation}
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
 The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the intial timestamp of the genesis block.
+Also $time$ will be set to the initial timestamp of the genesis block.
 The latest documentation should be consulted for those values.

### 2016-12-22 Paper.tex
Fix the memory consumption of CODECOPY

--- 2016-12-15 Paper.tex
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SUICIDE} operation which creates an account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\

### 2016-12-22 Paper.tex
Decument the memory usage update in LOG operations

--- 2016-12-22 Paper.tex
@@
 Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 

### 2016-12-23 Paper.tex
CODECOPY read position calculation does not use modulo

--- 2016-12-22 Paper.tex
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SUICIDE} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\

### 2016-12-30 Paper.tex
The output from a CALL is considered as the empty string when the call fails

--- 2016-12-30 Paper.tex
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\

### 2016-12-30 Paper.tex
The output from a CALL is considered as the empty string when the call fails

--- 2016-12-30 Paper.tex
@@
 Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Yoichi Hirai, Giacomo Tazzari and, of course, Vitalik Buterin.
+
+\section{Availability}
+
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 

### 2016-12-30 Paper.tex
Also in CALLCODE, the output string is empty when the call fails
A runtime exception returns an empty state, which cannot be used in the caller

@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\boldsymbol{\sigma}' \equiv
+\begin{cases}
+\boldsymbol{\sigma} & \text{if}\quad \boldsymbol{\sigma}'' = \varnothing \\
+\boldsymbol{\sigma''} & \text{otherwise}
+\end{cases}$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2016-12-30 Paper.tex
An exceptional haulting can be detected with the empty state

@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
-&&&& otherwise. \\
+&&&& $\boldsymbol{\sigma}'' = \varnothing$ or if $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached);\\
+&&&& $x=1$ otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\

### 2017-01-11 travis_deploy.sh
Deploy pdf from travis.

@@ -1,0 +1,28 @@
+#!/usr/bin/env bash
+
+set -e
+
+SHA=`git rev-parse --verify HEAD`
+
+git config user.name "$COMMIT_AUTHOR"
+git config user.email "$COMMIT_AUTHOR_EMAIL"
+git checkout --orphan gh-pages
+git rm --cached -r .
+echo "# Automatic build" > README.md
+echo "Built pdf from \`$SHA\`. See https://github.com/ethereum/yellowpaper/ for details." >> README.md
+echo "The generated pdf is here: https://ethereum.github.io/yellowpaper/paper.pdf" >> README.md
+echo '<html><head><meta http-equiv="refresh" content="0; url=paper.pdf" /></head><body></body></html>' > index.html
+mv Paper.pdf paper.pdf
+git add -f README.md index.html paper.pdf
+git commit -m "Built pdf from {$SHA}."
+
+ENCRYPTED_KEY_VAR="encrypted_${ENCRYPTION_LABEL}_key"
+ENCRYPTED_IV_VAR="encrypted_${ENCRYPTION_LABEL}_iv"
+ENCRYPTED_KEY=${!ENCRYPTED_KEY_VAR}
+ENCRYPTED_IV=${!ENCRYPTED_IV_VAR}
+openssl aes-256-cbc -K $ENCRYPTED_KEY -iv $ENCRYPTED_IV -in deploykey.enc -out deploykey -d
+chmod 600 deploykey
+eval `ssh-agent -s`
+ssh-add deploykey
+
+git push -f "$PUSH_REPO" gh-pages

### 2017-01-12 build.sh
Include build script

@@ -1,0 +1,9 @@
+#!/usr/bin/env bash
+
+set -e
+
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
+bibtex Paper
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex

### 2017-01-12 Paper.tex
Include git hash in generated PDF

--- 2016-12-30 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}} \\ {\smaller {\YellowPaperVersionNumber{}}}}
 \author{
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& $\boldsymbol{\sigma}' \equiv
-\begin{cases}
-\boldsymbol{\sigma} & \text{if}\quad \boldsymbol{\sigma}'' = \varnothing \\
-\boldsymbol{\sigma''} & \text{otherwise}
-\end{cases}$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $\boldsymbol{\sigma}'' = \varnothing$ or if $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached);\\
-&&&& $x=1$ otherwise. \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
 \midrule

### 2017-01-12 build.sh
Include git hash in generated PDF

@@ -2,6 +2,10 @@
 
 set -e
 
+SHA=`git rev-parse --short --verify HEAD`
+
+echo "\newcommand{\YellowPaperVersionNumber}{$SHA}" > Version.tex
+
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
 bibtex Paper && \
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \

### 2017-01-12 Paper.tex
Ethash isn't planned, but is the reality for current Ethereum

--- 2017-01-12 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}} \\ {\smaller {\YellowPaperVersionNumber{}}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision}}}
 \author{
@@
 \midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\

### 2017-01-12 build.sh
Add build date

@@ -3,8 +3,9 @@
 set -e
 
 SHA=`git rev-parse --short --verify HEAD`
+DATE=`date "+%Y-%m-%d"`
 
-echo "\newcommand{\YellowPaperVersionNumber}{$SHA}" > Version.tex
+echo "\newcommand{\YellowPaperVersionNumber}{$SHA - $DATE}" > Version.tex
 
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
 bibtex Paper && \

### 2017-01-12 Paper.tex
Rename SUICIDE to SELFDESTRUCT (as per EIP6)

@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
-$G_{suicide}$ & 5000 & Amount of gas to pay for a {\small SUICIDE} operation. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
@@
 \midrule
-0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\

### 2017-01-12 Paper.tex
Rename other references of suicide to self-destruct

@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct list:
 \begin{eqnarray}
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
 \begin{equation}
@@
 
-In general, we assume the memory, suicide list and system state don't change:
+In general, we assume the memory, self-destruct list and system state don't change:
 \begin{eqnarray}
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
 $G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\

### 2017-01-12 Paper.tex
Use 'self-destruct set' consistently

@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct list:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
 \begin{equation}
@@
 
-In general, we assume the memory, self-destruct list and system state don't change:
+In general, we assume the memory, self-destruct set and system state don't change:
 \begin{eqnarray}

### 2017-01-12 Paper.tex
Remove obsolete commented out fee description (it has its own appendix now)

--- 2017-01-12 Paper.tex
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$G_{suicide}$ & 5000 & Amount of gas to pay for a {\small SUICIDE} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\

### 2017-01-12 build.sh
Proper error propagation in build script

@@ -2,8 +2,8 @@
 
 set -e
 
-pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
-bibtex Paper
-pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
-pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
+bibtex Paper && \
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
+pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex

### 2017-01-12 build.sh
Take date of last commit as build date

@@ -3,7 +3,7 @@
 set -e
 
 SHA=`git rev-parse --short --verify HEAD`
-DATE=`date "+%Y-%m-%d"`
+DATE=`git show -s --format="%cd" --date="format:%Y-%m-%d" HEAD`
 
 echo "\newcommand{\YellowPaperVersionNumber}{$SHA - $DATE}" > Version.tex
 

### 2017-01-12 build.sh
Use 'unknown revision' if building without git

@@ -2,10 +2,19 @@
 
 set -e
 
+if [ -d ".git" ]; then
+
 SHA=`git rev-parse --short --verify HEAD`
 DATE=`git show -s --format="%cd" --date="format:%Y-%m-%d" HEAD`
+REV="$SHA - $DATE"
 
-echo "\newcommand{\YellowPaperVersionNumber}{$SHA - $DATE}" > Version.tex
+else
+
+REV="unknown revision"
+
+fi
+
+echo "\newcommand{\YellowPaperVersionNumber}{$REV}" > Version.tex
 
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
 bibtex Paper && \

### 2017-01-12 Paper.tex
Use out-of-gas consistently

@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an out-of-gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 

### 2017-01-12 Paper.tex
Move the OOG shorthand to the first occurance of out-of-gas

@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 

### 2017-01-13 Paper.tex
Replace unbalanced bracket with comma

--- 2017-01-12 Paper.tex
@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
 

### 2017-01-17 build.sh
Use git short date as YYYY-MM-DD

@@ -5,7 +5,7 @@ set -e
 if [ -d ".git" ]; then
 
 SHA=`git rev-parse --short --verify HEAD`
-DATE=`git show -s --format="%cd" --date="format:%Y-%m-%d" HEAD`
+DATE=`git show -s --format="%cd" --date=short HEAD`
 REV="$SHA - $DATE"
 
 else

### 2017-01-17 Paper.tex
Update Paper.tex

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+where $c$ is the code-deposit cost:
 \begin{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\bottomrule
+\end{cases}$ \\\bottomrule
 \end{tabular*}

### 2017-02-17 Paper.tex
Change the delta and alpha of INVALID instruction into nothing.  This makes sure that the machine halts.
Add a new element I_w in the input tuple I

--- 2017-01-17 Paper.tex
@@
 With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap.
-Technology-rooted movements like Bitcoin have demonstrated, through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
+Technology-rooted movements like Bitcoin have demonstrated through the power of the default, consensus mechanisms, and voluntary respect of the social contract, that it is possible to use the internet to make a decentralised value-transfer system that can be shared across the world and virtually free to use.
 This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
@@
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
-Formally:
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e
+I_e & \equiv & e \\
+I_w & \equiv & \top
 \end{eqnarray}
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 I_e & \equiv & e \\
+I_w & \equiv & \top \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will define it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
 
@@
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
 \midrule
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}

### 2017-02-17 Paper.tex
Add a new element I_w in the input tuple I

--- 2017-01-17 Paper.tex
@@
 
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
 
@@
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will defined it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}

### 2017-02-17 Paper.tex
Add a new element I_w in the input tuple I

@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l})
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
 
@@
 \item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
+\item $I_e$, the depth of the present message-call or contract-creation
+      (i.e. the number of {\small CALL}s or {\small CREATE}s being
+      executed at present).
+\item $I_w$, the permission to write to the state.
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}

### 2017-02-17 Paper.tex
Add a "write permission" argument to Theta

--- 2017-02-17 Paper.tex
@@
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 I_e & \equiv & e \\
-I_w & \equiv & \top \\
+I_w & \equiv & w \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2017-02-17 Paper.tex
Add a "write permission" argument to Theta

@@
 \item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation
-      (i.e. the number of {\small CALL}s or {\small CREATE}s being
-      executed at present).
-\item $I_w$, the permission to write to the state.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will defined it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}

### 2017-02-17 Paper.tex
Add "write permission" argument to \Lambda function
Add an exceptional halt for state changes during static calls

@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 I_e & \equiv & e \\
-I_w & \equiv & \top
+I_w & \equiv & w
 \end{eqnarray}
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv
+\begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge
+\boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger then 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\

### 2017-02-17 Paper.tex
Add an exceptional halt for state changes during static calls

--- 2017-02-17 Paper.tex
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger then 1024 or state modification is attempted during a static call.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.

### 2017-02-17 Paper.tex
Specify STATICCALL

--- 2017-02-17 Paper.tex
@@
 \midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and
+     $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-02-20 Paper.tex
Define REVERT instruction
Fixes #232

--- 2017-02-17 Paper.tex
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
+I_e & \equiv & e
 \end{eqnarray}
@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
+\begin{eqnarray}
+ c & \equiv &
+ \begin{cases}
+  G_{codedeposit} \times |\mathbf{o}| & \text{if}\quad \boldsymbol{\sigma}^{**} \neq \varnothing \\
+  0 & \text{otherwise}
+ \end{cases}
+\end{eqnarray}
 
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge \mathbf{o} = \varnothing \\
+g^{**} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge \mathbf{o} \neq \varnothing \\
 g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 $\boldsymbol{\sigma}$).
+However, if the execution halted by a {\small REVERT} instruction, the remaining gas (if any) is paid back although the state is reverted.
 
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
+               \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 I_e & \equiv & e \\
-I_w & \equiv & w \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation
-      (i.e. the number of {\small CALL}s or {\small CREATE}s being
-      executed at present).
-\item $I_w$, the permission to write to the state.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction.
 
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv
-\begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge
-\boldsymbol{\mu}_\mathbf{s}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger then 1024 or state modification is attempted during a static call.
-The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 
@@
 
-\subsubsection{Normal Halting}
-
-The normal halting function $H$ is defined:
-\begin{equation}
+\subsubsection{Post-Execution Halting}
+
+The post-execution halting function $H$ is defined:
+\begin{equation} \label{eq:output}
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in
+                               \{\text{\small RETURN}, \text{\small REVERT}\}  \\
 () & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 \midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and
-     $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+&&&& The definition of $H$ mentions {\small REVERT} and that has the effect of halting the execution.\\
+&&&& See equation (\ref{eq:output}). \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\sigma' = \varnothing,\quad A' = A^0$ \\
 \midrule

### 2017-02-20 Paper.tex
The execution function X returns remaining gas instead of a state
Fixes #238

@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{eqnarray}
- c & \equiv &
- \begin{cases}
-  G_{codedeposit} \times |\mathbf{o}| & \text{if}\quad \boldsymbol{\sigma}^{**} \neq \varnothing \\
-  0 & \text{otherwise}
- \end{cases}
-\end{eqnarray}
+\begin{equation}
+c \equiv G_{codedeposit} \times |\mathbf{o}|
+\end{equation}
 
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge \mathbf{o} = \varnothing \\
-g^{**} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge \mathbf{o} \neq \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 $\boldsymbol{\sigma}$).
-However, if the execution halted by a {\small REVERT} instruction, the remaining gas (if any) is paid back although the state is reverted.
 
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
-               \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
 
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+\big(\varnothing, \boldsymbol{\mu}_g, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \boxplus \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
 X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
+(a, b, c, d) \boxplus e & \equiv & (a, b_g, c, d, e)
 \end{eqnarray}
@@
 This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
 
@@
 
-\subsubsection{Post-Execution Halting}
-
-The post-execution halting function $H$ is defined:
-\begin{equation} \label{eq:output}
+\subsubsection{Normal Halting}
+
+The normal halting function $H$ is defined:
+\begin{equation}
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in
-                               \{\text{\small RETURN}, \text{\small REVERT}\}  \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
 () & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
 
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-&&&& The definition of $H$ mentions {\small REVERT} and that has the effect of halting the execution.\\
-&&&& See equation (\ref{eq:output}). \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-&&&& $\sigma' = \varnothing,\quad A' = A^0$ \\
 \midrule

### 2017-02-20 Paper.tex
Make sure that message call consumes all gas when exceptional halting happens

--- 2017-02-20 Paper.tex
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
-               \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
@@
 \midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-&&&& The definition of $H$ mentions {\small REVERT} and that has the effect of halting the execution.\\
-&&&& See equation (\ref{eq:output}). \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-&&&& $\sigma' = \varnothing,\quad A' = A^0$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-02-20 Paper.tex
Make sure that message call consumes all gas when exceptional halting happens

@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\

### 2017-02-22 Paper.tex
Remove a left-over comment
Add more EIP numbers in the title
These EIPs were applied on the same block number, so it makes sense to list them all.

--- 2017-02-20 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-155/160/161/170 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 \end{equation}
+
+An account state is empty when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
+\wedge \boldsymbol{\sigma}[a]_n = 0
+\wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+
+An account is dead when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
+
+Here is a function that updates the balance of an account state but deletes the account state when it becomes empty:
+\begin{equation}
+\rho_b(\boldsymbol{\sigma}, a, x) \equiv
+\begin{cases}
+\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
+(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
+\end{cases}
+\end{equation}
+for $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
 
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s] &\equiv& \rho_b(\boldsymbol{\sigma}, s, \boldsymbol{\sigma}[s]_b - v)
 \end{eqnarray}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\\ \nonumber
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
+T   & \equiv & \{s, r\} \\
+\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \end{dcases} \\
- v &\in \{27,28\} 
+ v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+Two different flavors of signing schemes are available.
+One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+The other operates on nine elements:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0 \wedge v \in \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{if} \; T_t \neq 0 \wedge v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, \mathtt{\tiny chain\_id}, (), ()) & \text{if} \; T_t = 0 \wedge v \notin \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}
 \end{cases} \\

### 2017-02-22 Paper.tex
Add more EIP numbers in the title
These EIPs were applied on the same block number, so it makes sense to list them all.

@@
 \end{cases} \\
-\\ \nonumber
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-T   & \equiv & \{s, r\} \\
+\mathbf{t}   & \equiv & \{s, r\} \\
 \\ \nonumber

### 2017-02-24 Paper.tex
Update wording
Modify \Theta to return a flag indicating exceptional halt

--- 2017-02-22 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-155/160/161/170 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 \end{equation}
-
-An account state is empty when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
-\wedge \boldsymbol{\sigma}[a]_n = 0
-\wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-
-An account is dead when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
-\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-Here is a function that updates the balance of an account state but deletes the account state when it becomes empty:
-\begin{equation}
-\rho_b(\boldsymbol{\sigma}, a, x) \equiv
-\begin{cases}
-\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
-(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
-\end{cases}
-\end{equation}
-for $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
 
@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{align}
-c \equiv \begin{cases}
-  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24000 \\
-  G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}
-\end{cases}
-\end{align}
+\begin{eqnarray}
+ c & \equiv &
+ \begin{cases}
+  G_{codedeposit} \times |\mathbf{o}| & \text{if}\quad \boldsymbol{\sigma}^{**} \neq \varnothing \\
+  0 & \text{otherwise}
+ \end{cases}
+\end{eqnarray}
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}, e) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 \end{cases} \\
-\\ \nonumber
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
+e & \equiv & \begin{cases}
+\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\bot & \text{otherwise}
+\end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
+               \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I, T)
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
-The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
-&&&& otherwise. \\
+&&&& $e = \top$ or if $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); \\
+&&&& $x=1$ otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 \midrule
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+&&&& The definition of $H$ mentions {\small REVERT} and that has the effect of halting the execution.\\
+&&&& See equation (\ref{eq:output}). \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\sigma' = \varnothing,\quad A' = A^0$ \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2017-02-24 Paper.tex
\Lambda returns a flag indicating exceptional halt

@@
 \begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+(\boldsymbol{\sigma}_P, g', A, e) \equiv \begin{cases}
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
@@
 
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A, e) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+Thus formally, we may specify the resultant state, gas, substate and exception flag as $(\boldsymbol{\sigma}', g', A, e)$ where:
 
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases} \\
+\quad e &\equiv  \begin{cases}
+\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\bot & \text{otherwise}
 \end{cases}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, e) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing, \top\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& $e = \top$ or $I_e = 1024$ (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& (balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$,\\
+&&&& the address of the newly created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\

### 2017-02-28 Paper.tex
Fix wording

--- 2017-02-22 Paper.tex
@@
 
-Here is a function that updates the balance of an account state but deletes the account state when it becomes empty:
+We define the function that updates the balance of an account state but deletes the account state when it becomes empty:
 \begin{equation}
@@
 \end{equation}
-for $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
+where $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
 

### 2017-02-28 Paper.tex
Fix the code size limit in EIP170

@@
 c \equiv \begin{cases}
-  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24000 \\
+  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24576 \\
   G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}

### 2017-02-28 Paper.tex
Interpret the signature (chain_id, 0, 0) as signed by 2^{256} - 1
but only when the gas price, the nonce and the value are all zero.

--- 2017-02-24 Paper.tex
@@
 \begin{equation}
-(\boldsymbol{\sigma}_P, g', A, e) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
@@
 
-Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, e) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and exception flag as $(\boldsymbol{\sigma}', g', A, e)$ where:
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
 
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad e &\equiv  \begin{cases}
-\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\bot & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}, e) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 \end{cases} \\
-e & \equiv & \begin{cases}
-\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\bot & \text{otherwise}
-\end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
-               \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 \begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r) \quad \text{ where } (T_w, T_r, T_s) \text{ is a valid signature.}
+\end{equation}
+
+When a transaction has a particular invalid signature and it has zero gas price, zero nonce and zero value, the sender of the transaction is $2^{256} - 1$.
+\begin{equation}
+\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } \quad (T_w, T_r, T_s) = (\mathtt{chain\_id}, 0, 0) \,\wedge\, T_p = T_n = T_v = 0.
 \end{equation}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, e) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing, \top\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $e = \top$ or $I_e = 1024$ (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& (balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$,\\
-&&&& the address of the newly created account, otherwise. \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $e = \top$ or if $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); \\
-&&&& $x=1$ otherwise. \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2017-02-28 Paper.tex
Interpret the signature (0, 0, 0) as signed by 2^{256}

@@
 
-When a transaction has a particular invalid signature and it has zero gas price, zero nonce and zero value, the sender of the transaction is $2^{256} - 1$.
-\begin{equation}
-\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } \quad (T_w, T_r, T_s) = (\mathtt{chain\_id}, 0, 0) \,\wedge\, T_p = T_n = T_v = 0.
+When a transaction has a particular invalid signature $v = r = s = 0$, the sender of the transaction is $2^{256} - 1$.
+\begin{equation}
+\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } (T_w, T_r, T_s) = (0, 0, 0).
 \end{equation}

### 2017-02-28 Paper.tex
EIP86: change the address of newly created contracts

@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(0 \cdot \mathbf{i}\big)\Big)
+\end{equation}
+where $0$ denotes the zero element of $\mathbb{B}_{256}$, the operator $\cdot$ is the concatenation of binary data, $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data.
 

### 2017-02-28 Paper.tex
EIP86: preparation before adding CREATE_P2SH
The function \Lambda now takes the new address as a parameter

@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, a) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
@@
 and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+$a$ is the address of the created account:
+\begin{equation} \label{eq:address_calc}
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(0 \cdot \mathbf{i}\big)\Big)
+\end{equation}
+where $0$ denotes the zero element of $\mathbb{B}_{160}$, the operator $\cdot$ is the concatenation of binary data, $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(0 \cdot \mathbf{i}\big)\Big)
-\end{equation}
-where $0$ denotes the zero element of $\mathbb{B}_{256}$, the operator $\cdot$ is the concatenation of binary data, $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data.
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, a)
+\end{equation}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& where $a$ is defined as in (\ref{eq:address_calc}) \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\

### 2017-02-28 Paper.tex
EIP86: preparation before adding CREATE_P2SH
The function \Lambda now takes the new address as a parameter

@@
 and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-$a$ is the address of the created account:
-\begin{equation} \label{eq:address_calc}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(0 \cdot \mathbf{i}\big)\Big)
-\end{equation}
-where $0$ denotes the zero element of $\mathbb{B}_{160}$, the operator $\cdot$ is the concatenation of binary data, $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data.
+$a$ is the address of the created account, which is defined as the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+Thus we define the resultant address for the new account $a$:
+\begin{equation}
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}_0[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}_0[x]$ is the account state of $x$ or $\varnothing$ if none exists.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 
-When a transaction has a particular invalid signature $v = r = s = 0$, the sender of the transaction is $2^{256} - 1$.
-\begin{equation}
-\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } (T_w, T_r, T_s) = (0, 0, 0).
+When a transaction has a particular invalid signature and it has zero gas price, zero nonce and zero value, the sender of the transaction is $2^{256} - 1$.
+\begin{equation}
+\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } \quad (T_w, T_r, T_s) = (\mathtt{chain\_id}, 0, 0) \,\wedge\, T_p = T_n = T_v = 0.
 \end{equation}

### 2017-03-02 Paper.tex
EIP86: add CREATE_P2SH
Stop incrementing the nonce of the account at 2^{256}
Stop incrementing the nonce for transactions from the account at 2^{256} - 1
EIP98: remove state root from transaction receipts

--- 2017-02-28 Paper.tex
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(0 \cdot \mathbf{i}\big)\Big)
-\end{equation}
-where $0$ denotes the zero element of $\mathbb{B}_{256}$, the operator $\cdot$ is the concatenation of binary data, $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 \begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r) \quad \text{ where } (T_w, T_r, T_s) \text{ is a valid signature.}
-\end{equation}
-
-When a transaction has a particular invalid signature $v = r = s = 0$, the sender of the transaction is $2^{256} - 1$.
-\begin{equation}
-\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } (T_w, T_r, T_s) = (0, 0, 0).
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
 \end{equation}

### 2017-03-02 Paper.tex
EIP98: remove state root from transaction receipts

@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the self-destruct set $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will define it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}

### 2017-03-02 Paper.tex
Implement EIP-658 https://github.com/ethereum/EIPs/pull/658
Now transaction receipts contain one byte indicating success or failure.

@@
 
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
+
 \subsection{The Transaction} \label{ch:transaction}
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
 \end{equation}
@@
 where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+
+We assert that the status code $R_{s'}$ is a single byte.
+\begin{equation}
+R_{s'} \in \mathbb{B}_{8}
+\end{equation}
 
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
-Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), 0)
+We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), \varnothing, 0)
 \end{equation}
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
 \end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s] &=& \begin{cases}
+\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
+\mathbf{a}^* & \text{otherwise}
+\end{cases} \\
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
 \end{eqnarray}
@@
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e
+I_e & \equiv & e \\
+I_w & \equiv & w
 \end{eqnarray}
@@
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\quad s' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
+\end{cases}
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+s' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
+1 & \text{otherwise}
 \end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
+I_w & \equiv & w \\
+\mathbf{t} & \equiv & \{s, r\} \\
+\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
@@
 \end{eqnarray}
-\begin{equation}
+\begin{equation} \label{eq:X-def}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
+\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
+\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
 \end{eqnarray}
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger than 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
 () & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
@@
 
-The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 \begin{eqnarray}
+\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
+\mathbf{a}' &\text{otherwise}
+\end{cases} \\
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 \mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\end{equation}
+
+We define $\mathbf{R}[n]_{s'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{s'} =
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
 \end{equation}
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
+\begin{equation} \label{eq:pre}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
 (\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 \midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
 \end{cases}$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule
@@
 
+\section{Anomalies on the Main Network}
+
+\subsection{Deletion of an Account Dispite Out-of-gas}
+
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
+
 \end{document}

### 2017-03-02 Paper.tex
EIP100: change the difficulty adjustment factor so that it considers existence of ommers

--- 2017-03-02 Paper.tex
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation*}
+x \equiv \begin{cases}
+1 & \text{if} \, \lVert B_\mathbf{U}\rVert = 0 \\
+2 & \text{otherwise}
+\end{cases}
+\end{equation*}
 \begin{equation}

### 2017-03-02 Paper.tex
EIP-100: change the difficulty adjustment factor so that it considers existence of ommers

--- 2017-03-02 Paper.tex
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_{s'}$ is a single byte.
-\begin{equation}
-R_{s'} \in \mathbb{B}_{8}
-\end{equation}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\quad s' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases}
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\ \nonumber
-s' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 \mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}

### 2017-03-02 Paper.tex
EIP101: addition precompile

--- 2017-03-02 Paper.tex
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert B_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
 \begin{equation}
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 
-Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision addition on non-negative integers.
+The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
+The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
+
+\begin{eqnarray}
+\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\, X = Y = 0 \\
+\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract \\
+$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
 

### 2017-03-02 Paper.tex

--- 2017-03-02 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision addition on non-negative integers.
-The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
-The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
-
-\begin{eqnarray}
-\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\, X = Y = 0 \\
-\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract \\
-$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
 

### 2017-03-02 Paper.tex

@@
 \end{equation}
+
+An account state is empty when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
+\wedge \boldsymbol{\sigma}[a]_n = 0
+\wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+
+An account is dead when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
+
+We define the function that updates the balance of an account state but deletes the account state when it becomes empty:
+\begin{equation}
+\rho_b(\boldsymbol{\sigma}, a, x) \equiv
+\begin{cases}
+\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
+(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
+\end{cases}
+\end{equation}
+where $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
 
@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
+\begin{align}
+c \equiv \begin{cases}
+  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24576 \\
+  G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}
+\end{cases}
+\end{align}
 
@@
 \end{cases} \\
+\\ \nonumber
 (\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, T) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, T) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\

### 2017-03-02 Paper.tex
EIP101: add the subtraction precompile and the multiplication precompile

--- 2017-03-02 Paper.tex
@@
 \Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
+
+\begin{eqnarray}
+\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = Y \\
+\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The definition is similar to that of the addition contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = 0 \ \text{or}\  Y = 0 \\
+\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
 \end{cases}
@@
 $G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract \\
+$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract \\
 $G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
+$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 

### 2017-03-02 Paper.tex

@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision addition on non-negative integers.
-The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
-The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
-
-\begin{eqnarray}
-\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\, X = Y = 0 \\
-\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
-
-\begin{eqnarray}
-\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = Y \\
-\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The definition is similar to that of the addition contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = 0 \ \text{or}\  Y = 0 \\
-\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract \\
-$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract \\
-$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
-$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 

### 2017-03-02 Paper.tex

@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+e & \equiv & \begin{cases}
+\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\bot & \text{otherwise}
+\end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
+               \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+&&&& The definition of $H$ mentions {\small REVERT} and that has the effect of halting the execution.\\
+&&&& See equation (\ref{eq:output}). \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\sigma' = \varnothing,\quad A' = A^0$ \\
 \midrule

### 2017-03-03 Paper.tex
EIP 101: add division

--- 2017-03-02 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision addition on non-negative integers.
+The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
+The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
+
+\begin{eqnarray}
+\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\, X = Y = 0 \\
+\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
+
+\begin{eqnarray}
+\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = Y \\
+\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The input format is the same as that of the addition contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = 0 \ \text{or}\  Y = 0 \\
+\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The eigth contract performs arbitrary-precision division on non-negative integers.  The definition is similar to that of the multiplication contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{DIV}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  Y = 0 \\
+ \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ Y = 0 \ \text{or} \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
+\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract \\
+$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract \\
+$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
+$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 

### 2017-03-03 Paper.tex
EIP 101: Add EXPMOD precompile

--- 2017-03-02 Paper.tex
@@
 
-\subsection{Homestead} \label{ch:homestead}
-
-A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
-\begin{equation}
-\firsthomesteadblock \equiv 1,\! 150,\! 000
-\end{equation}
-The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}_P, g', A, e) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, e) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge \mathbf{o} = \varnothing \\
-g^{**} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge \mathbf{o} \neq \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+0 & \text{if} \quad F \\
 g^{**} - c & \text{otherwise} \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \end{cases} \\
-\quad e &\equiv  \begin{cases}
-\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\bot & \text{otherwise}
-\end{cases}
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}, e) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\
-e & \equiv & \begin{cases}
-\top & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\bot & \text{otherwise}
-\end{cases} \\
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \wedge
-               \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\end{cases} \\ \nonumber
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
 
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+These two words are followed by $B$ and $E$; and the rest of the input is interpreted as the third non-negative integer $M$.
+All non-negative integers $B$, $E$ $M$ are encoded as byte sequences in the big-endian way.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\ell_B &=& I_\mathbf{d}[0..31] \\
+B &=& I_\mathbf{d}[32..(31 + \ell_B)] \\
+\ell_E &=& I_\mathbf{d}[(32 + \ell_B)..(63 + \ell_B)] \\
+E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
+M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 32 + \ell_B \quad \vee \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
+g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  M = 0 \\
+ B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
+\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28\}
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{modexpbase}$ & 45 & Payment for the precompiled exponention under modulo. \\
+$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
+$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, e) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing, \top\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $e = \top$ or $I_e = 1024$ (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& (balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$,\\
-&&&& the address of the newly created account, otherwise. \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}, e) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, (), \top) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 \midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-&&&& The definition of $H$ mentions {\small REVERT} and that has the effect of halting the execution.\\
-&&&& See equation (\ref{eq:output}). \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-&&&& $\sigma' = \varnothing,\quad A' = A^0$ \\
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
 \midrule
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
 \end{cases}$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule

### 2017-03-03 Paper.tex
EIP 101: Add EXPMOD precompile
EIP 100: add EXPMOD precompile

--- 2017-03-03 Paper.tex
@@
 \Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 8 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 9 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \begin{cases}
-0 &\text{if}\ X = 0 \ \text{or}\  Y = 0 \\
+0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
 \lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
@@
 \begin{cases}
-0 &\text{if}\ Y = 0 \ \text{or} \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
+0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
 \lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\ell_B &=& I_\mathbf{d}[0..31] \\
+B &=& I_\mathbf{d}[32..(31 + \ell_B)] \\
+\ell_E &=& I_\mathbf{d}[(32 + \ell_B)..(63 + \ell_B)] \\
+E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
+M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 32 + \ell_B \quad \vee \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
+g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  M = 0 \\
+ B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
+\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
 \end{cases}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract \\
-$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract \\
+$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract. \\
+$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract. \\
+$G_{modexpbase}$ & 45 & Payment for the precompiled exponention under modulo. \\
 $G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.  $0 ^ 0$ is defined to be $1$.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\

### 2017-03-07 Paper.tex
EIP 101: simplify a formula A \/ B into B,
because A implies B.
EIP 101: simplification
EIP 166: Change the transaction validity condition on the nonce

@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+T_n \bmod 2^{64} & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+\big(\lfloor T_n / 2^{64} \rfloor = 0 & \vee & \lfloor T_n / 2^{64} \rfloor = \mathtt{\small chain\_id}\ \big) \quad \wedge \\
 g_0 & \leqslant & T_g \quad \wedge \\
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 8 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 9 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision addition on non-negative integers.
-The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
-The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
-
-\begin{eqnarray}
-\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\, X = Y = 0 \\
-\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
-
-\begin{eqnarray}
-\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = Y \\
-\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The input format is the same as that of the addition contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
-\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The eigth contract performs arbitrary-precision division on non-negative integers.  The definition is similar to that of the multiplication contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{DIV}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  Y = 0 \\
- \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
-\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\ell_B &=& I_\mathbf{d}[0..31] \\
-B &=& I_\mathbf{d}[32..(31 + \ell_B)] \\
-\ell_E &=& I_\mathbf{d}[(32 + \ell_B)..(63 + \ell_B)] \\
-E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
-M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 32 + \ell_B \quad \vee \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
-g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  M = 0 \\
- B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
-\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
-\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract. \\
-$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract. \\
-$G_{modexpbase}$ & 45 & Payment for the precompiled exponention under modulo. \\
-$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
-$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 
@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.  $0 ^ 0$ is defined to be $1$.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\

### 2017-03-07 Paper.tex
EIP-211: add the previous output buffer to the execution state

@@
 \begin{equation}
-\frontiermod \equiv \begin{cases}
-1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
--1 & \text{otherwise} \\
-\end{cases}
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation*}
+x \equiv \begin{cases}
+1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+2 & \text{otherwise}
+\end{cases}
+\end{equation*}
 \begin{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n \bmod 2^{64} & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-\big(\lfloor T_n / 2^{64} \rfloor = 0 & \vee & \lfloor T_n / 2^{64} \rfloor = \mathtt{\small chain\_id}\ \big) \quad \wedge \\
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
 g_0 & \leqslant & T_g \quad \wedge \\
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
-\end{eqnarray}
-\begin{equation}
+\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
+\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
+\end{eqnarray}
+\begin{equation} \label{eq:X-def}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
-\end{eqnarray}
-
-Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
+\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
+\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 
-Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
+Many thanks to Aeron Buchanan for authoring the \textit{Homestead} revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
 Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
@@
 The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
-\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of nibbles keys and branches as traversed from the root.
+\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
 The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+f(x) &\equiv& \begin{cases}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\end{cases}\\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
+8(\ell_E - 32) & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+g_r&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r &=& 500\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r &=& 40000\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}
@@
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\

### 2017-03-07 Paper.tex
EIP 211: add the previous output buffer to the execution state

--- 2017-03-07 Paper.tex
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
+\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
 \end{eqnarray}
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\

### 2017-03-07 Paper.tex
EIP 211: add the previous output buffer to the execution state
EIP 211: specify RETURNDATASIZE
EIP-211: specify RETURNDATASIZE
EIP 211: specify RETURNDATACOPY
EIP-211: specify RETURNDATACOPY
EIP 211: specify the gas costs for RETURNDATASIZE and RETURNDATACOPY

--- 2017-03-07 Paper.tex
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
@@
 
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+\midrule
+0x0d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
+\midrule
+0x0e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+\midrule
 \end{tabular*}

### 2017-03-13 Paper.tex
License the work under CC BY 4.0

--- 2017-03-07 Paper.tex
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
 \end{eqnarray}
@@
 
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/} and is licensed under the Creative Commons Attribution 4.0 International License.
+To view a copy of this license, visit \url{http://creativecommons.org/licenses/by/4.0/}.
+An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
@@
 
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\midrule
-0x0d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
-\midrule
-0x0e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-\midrule
 \end{tabular*}
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\

### 2017-03-14 Paper.tex

@@
 
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/} and is licensed under the Creative Commons Attribution 4.0 International License.
-To view a copy of this license, visit \url{http://creativecommons.org/licenses/by/4.0/}.
-An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
+The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
 

### 2017-03-21 Paper.tex
In evaluation of \Xi, drop the fourth element and extract the remaining gas from the machine state

--- 2017-03-14 Paper.tex
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, I'\!, \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 
-Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 

### 2017-03-21 Paper.tex
In evaluation of \Xi, drop the fourth element and extract the remaining gas from the machine state

@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \begin{dcases}
-\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
-\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
-\end{dcases} \\
- v &\in \{27,28\}
-\end{align}
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule

### 2017-03-21 Paper.tex
Instead of concatenating a triple with an element, now we need to concatenate a quadruple with an element

@@
 
-where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
 Formally:
@@
 \Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, I'\!, \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
 \end{eqnarray}

### 2017-03-21 Paper.tex
Instead of concatenating a triple with an element, now we need to concatenate a quadruple with an element
EIP-100: ommers of parent should be counted, not the direct ommers

--- 2017-03-21 Paper.tex
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\end{cases} \\ \nonumber
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, I'\!, \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
+\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
+\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
 \end{eqnarray}

### 2017-03-21 Paper.tex

@@
 With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap.
-Technology-rooted movements like Bitcoin have demonstrated through the power of the default, consensus mechanisms, and voluntary respect of the social contract, that it is possible to use the internet to make a decentralised value-transfer system that can be shared across the world and virtually free to use.
+Technology-rooted movements like Bitcoin have demonstrated, through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
 This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
@@
 
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
+\subsection{Homestead} \label{ch:homestead}
+
+A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
+\begin{equation}
+\firsthomesteadblock \equiv 1,\! 150,\! 000
+\end{equation}
+The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
 
@@
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
 \end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 g^{**} - c & \text{otherwise} \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\end{cases}
 \end{align}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-I_w & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
+\end{eqnarray}
+
+Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
 
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
 () & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
@@
 
-The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 \begin{eqnarray}
-\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
+\begin{equation}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
 (\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+0 < s &< \begin{dcases}
+\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
+\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
+\end{dcases} \\
+ v &\in \{27,28\}
 \end{align}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account. \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
@@
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Dispite Out-of-gas}
-
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
 \end{document}

### 2017-03-22 Paper.tex
EIP 211: CREATE resets the previous return data
This change is according to
https://github.com/ethereum/EIPs/pull/211#issuecomment-287728214

@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
 \begin{equation}
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\

### 2017-03-22 Paper.tex
EIP-211: CREATE resets the previous return data
This change is according to
https://github.com/ethereum/EIPs/pull/211#issuecomment-287728214

--- 2017-03-21 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+f(x) &\equiv& \begin{cases}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\end{cases}\\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
+8(\ell_E - 32) & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+g_r&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r &=& 500\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r &=& 40000\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\

### 2017-03-22 Paper.tex
EIP 161: Note precompiled contracts can have empty account states

--- 2017-03-21 Paper.tex
@@
 \begin{equation}
-\frontiermod \equiv \begin{cases}
-1 & \text{if} \quad H_s < {P(H)_H}_s + 13 \\
--1 & \text{otherwise} \\
-\end{cases}
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
 \begin{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s] &\equiv& \begin{cases}
+\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
+\mathbf{a}^* & \text{otherwise}
+\end{cases} \\
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
 \end{eqnarray}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\

### 2017-03-22 Paper.tex

@@
 
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+An account state is empty when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
+\wedge \boldsymbol{\sigma}[a]_n = 0
+\wedge \boldsymbol{\sigma}[a]_b = 0
 \end{equation}
@@
 
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
+An account is dead when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
+\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
+
+We define the function that updates the balance of an account state but deletes the account state when it becomes empty:
+\begin{equation}
+\rho_b(\boldsymbol{\sigma}, a, x) \equiv
+\begin{cases}
+\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
+(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
+\end{cases}
+\end{equation}
+where $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
+
+\subsection{Homestead} \label{ch:homestead}
+
+A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
+\begin{equation}
+\firsthomesteadblock \equiv 1,\! 150,\! 000
+\end{equation}
+The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
 
@@
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \{\}, 0)
+For brevity, we define the initial substate $A^0$ to have no self-destructs, no logs, a set of touched accounts and a zero refund balance:
+\begin{equation}
+A^0(T) \equiv (\varnothing, (), T, 0) \quad\text{for}\quad T \subseteq \mathbb{B}_{20}
 \end{equation}
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &\equiv& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\boldsymbol{\sigma}^*[s] &\equiv& \rho_b(\boldsymbol{\sigma}, s, \boldsymbol{\sigma}[s]_b - v)
 \end{eqnarray}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\boldsymbol{\sigma}_1[s] \equiv \rho_b(\boldsymbol{\sigma}_1', s, \boldsymbol{\sigma}_1'[s]_b - v)
 \end{equation}
@@
 \boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\boldsymbol{\sigma}_1'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + v) & \text{otherwise}
+\end{cases}
 \end{equation}
@@
 \end{cases} \\ \nonumber
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0(T), I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\\ \nonumber
+\boldsymbol{\sigma}'[{B_H}_c] & \equiv & \rho_b(\boldsymbol{\sigma}, {B_H}_c, \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b) \\
+\\ \nonumber
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+ \boldsymbol{\sigma}'[U_c] & \equiv & \rho_b(\boldsymbol{\sigma}, U_c, \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b)
 \end{eqnarray}
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+(\varnothing, 0, A^0(T), ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0(T), \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[r] \equiv \rho_b(\boldsymbol{\sigma}, r, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b)$ where $r \equiv \boldsymbol{\mu}_s[0]\mod 2^{160}$\\
+&&&& $\boldsymbol{\sigma}'[I_a] \equiv \rho_b(\boldsymbol{\sigma}, I_a, 0)$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0 \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule

### 2017-03-22 Paper.tex
EIP 101: change the layout of the EXPMOD arguments
This change is introduced in:
https://github.com/ethereum/EIPs/pull/198/commits/60fe6515378aca6c9d1acb4f84d2f0b2a180abe8

--- 2017-03-22 Paper.tex
@@
 \begin{equation}
-A^0 \equiv (\varnothing, (), \{\}, 0)
+A^0 \equiv (\varnothing, (), \varnothing, 0)
 \end{equation}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &\equiv& \begin{cases}
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s] &=& \begin{cases}
 \varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
@@
 I_e & \equiv & e \\
+I_w & \equiv & w \\
 \mathbf{t} & \equiv & \{s, r\} \\
@@
 \Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation}
+\begin{equation} \label{eq:pre}
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+These two words are followed by $B$ and $E$; and the rest of the input is interpreted as the third non-negative integer $M$.
+All non-negative integers $B$, $E$ $M$ are encoded as byte sequences in the big-endian way.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\ell_B &=& I_\mathbf{d}[0..31] \\
+\ell_E &=& I_\mathbf{d}[32..63] \\
+B &=& I_\mathbf{d}[64..(63 + \ell_B)] \\
+E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
+M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
+g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  M = 0 \\
+ B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
+\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}
@@
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The chain identifier $\mathtt{\tiny chain\_id}$ is a constant natural number.
+The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
 
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma},\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
 \midrule

### 2017-03-22 Paper.tex
EIP 101: change the layout of the EXPMOD arguments
This change is introduced in:
https://github.com/ethereum/EIPs/pull/198/commits/60fe6515378aca6c9d1acb4f84d2f0b2a180abe8

--- 2017-03-22 Paper.tex
@@
 \end{equation}
-
-An account state is empty when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)
-\wedge \boldsymbol{\sigma}[a]_n = 0
-\wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state.  This is because their account states do not usually contain the code describing its behavior.
-
-An account is dead when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a)
-\quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-We define the function that updates the balance of an account state but deletes the account state when it becomes empty:
-\begin{equation}
-\rho_b(\boldsymbol{\sigma}, a, x) \equiv
-\begin{cases}
-\varnothing \quad\text{if}\quad \boldsymbol{\sigma}[a]_n = 0 \wedge x = 0 \wedge \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \\
-(\boldsymbol{\sigma}[a]_n, x, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \quad \text{otherwise}
-\end{cases}
-\end{equation}
-where $a \in \mathbb{B}_{20}$ and $x \in \mathbb{P}_{256}$.
 
@@
 If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{align}
-c \equiv \begin{cases}
-  2^{256} - 1 & \text{if} \quad |\mathbf{o}| \le 24576 \\
-  G_{codedeposit} \times |\mathbf{o}| & \text{otherwise}
-\end{cases}
-\end{align}
+\begin{equation}
+c \equiv G_{codedeposit} \times |\mathbf{o}|
+\end{equation}
 
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 8 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 9 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-\mathbf{t}   & \equiv & \{s, r\} \\
-\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 
+The fifth contract performs arbitrary-precision addition on non-negative integers.
+The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
+The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
+
+\begin{eqnarray}
+\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\, X = Y = 0 \\
+\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
+
+\begin{eqnarray}
+\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = Y \\
+\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The input format is the same as that of the addition contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
+\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The eigth contract performs arbitrary-precision division on non-negative integers.  The definition is similar to that of the multiplication contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{DIV}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  Y = 0 \\
+ \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
+\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The ninth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\ell_B &=& I_\mathbf{d}[0..31] \\
+\ell_E &=& I_\mathbf{d}[32..63] \\
+B &=& I_\mathbf{d}[64..(63 + \ell_B)] \\
+E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
+M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
+g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  M = 0 \\
+ B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
+\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}
@@
 \end{dcases} \\
- v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
+ v &\in \{27,28\}
 \end{align}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract. \\
+$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract. \\
+$G_{modexpbase}$ & 45 & Payment for the precompiled exponention under modulo. \\
+$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
+$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 

### 2017-03-22 Paper.tex
EIP 101: the first byte of the output might not exist
https://github.com/ethereum/yellowpaper/pull/258#discussion_r105860607
EIP101: use BE function to encode natural numbers
This addresses
https://github.com/ethereum/yellowpaper/pull/258#discussion_r105863776
EIP101: clarify the argument parsing

@@
 The fifth contract performs arbitrary-precision addition on non-negative integers.
-The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
-The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
+The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.  The rest of the input is interpreted as $Y$.
+Both $X$ and $Y$ are interpreted as a natural number encoded as byte sequences in the big-endian way.
+The result is represented in the smallest possible number of bytes.  The first byte in the output, if exists, is never zero.
+The input and output formats are the same for $\Xi_{\mathtt{SUB}}$, $\Xi_{\mathtt{MUL}}$ and $\Xi_{\mathtt{DIV}}$ as well.
 
@@
 g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
+\mathbf{o} &=& \mathtt{\tiny BE}(X + Y) \\
 X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+l_X &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
 Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\, X = Y = 0 \\
-\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
+l_Y &=& |I_\mathbf{d}| - 32 - l_X
 \end{eqnarray}
@@
 g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
+\mathbf{o} &=& \mathtt{\tiny BE}(X - Y) \\
 X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+l_X &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
 Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = Y \\
-\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
+l_Y &=& |I_\mathbf{d}| - 32 - l_X
 \end{eqnarray}
@@
 \Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + l_X l_Y / G_{quaddivisor} \\
+\mathbf{o} &=& \mathtt{\tiny BE}(X \times Y) \\
 X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
+l_X &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
 Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
-\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
+l_Y &=& |I_\mathbf{d}| - 32 - l_X
 \end{eqnarray}
@@
 \Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + l_X l_Y / G_{quaddivisor} \\
 \mathbf{o} &=&
@@
  () & \text{if} \  Y = 0 \\
- \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
+ \mathtt{\tiny BE}\Big(\big\lfloor \dfrac X Y \big\rfloor\Big) & \text{otherwise}
 \end{cases} \\
 X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
+l_X &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
 Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
-\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
-\end{cases}
+l_Y &=& |I_\mathbf{d}| - 32 - l_X
 \end{eqnarray}
@@
 The ninth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+These two words are followed by $B$ and $E$; and the rest of the input is interpreted as the third non-negative integer $M$.
+All non-negative integers $B$, $E$ $M$ are encoded as byte sequences in the big-endian way.
+The output format is the same as the precompiled contract $\Xi_{\mathtt{ADD}}$.
+
 \begin{eqnarray}
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\ell_B &=& I_\mathbf{d}[0..31] \\
-\ell_E &=& I_\mathbf{d}[32..63] \\
-B &=& I_\mathbf{d}[64..(63 + \ell_B)] \\
-E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
-M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
+l_B &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
+l_E &=& I_\mathbf{d}[32..63] \in \mathbb{P}_{256} \\
+B &=& I_\mathbf{d}[64..(63 + l_B)] \\
+E &=& I_\mathbf{d}[(64 + l_B)..(63 + l_B + l_E)] \\
+M &=& I_\mathbf{d}[(64 + l_B + l_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + l_B + l_E \\
 g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
@@
  () & \text{if} \  M = 0 \\
- B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
-\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
+ \mathtt{\tiny BE}(B ^ E \bmod M) & \text{otherwise}
 \end{cases}

### 2017-03-23 Paper.tex
EIP101: clarify the argument parsing
EIP101: reuse symbols for the lengths of arguments
EIP101: EXPMOD throws if M <= E
This follows the change in
https://github.com/ethereum/EIPs/pull/198/commits/c4a771ed3421b1840c3a9d77fbac05b8b58ad2c4
Make sure only one version of the protocol is described

--- 2017-03-22 Paper.tex
@@
 \begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+L_R(R) \equiv (\mathtt{TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
 
@@
 \begin{eqnarray}
-\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
-\end{eqnarray}
-
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
+\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\end{eqnarray}
+
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
@@
 
-The fifth contract performs arbitrary-precision addition on non-negative integers.
-The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
-The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
-
-\begin{eqnarray}
-\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\, X = Y = 0 \\
-\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
-
-\begin{eqnarray}
-\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = Y \\
-\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The input format is the same as that of the addition contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
-\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The eigth contract performs arbitrary-precision division on non-negative integers.  The definition is similar to that of the multiplication contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{DIV}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  Y = 0 \\
- \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
-\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The ninth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\ell_B &=& I_\mathbf{d}[0..31] \\
-\ell_E &=& I_\mathbf{d}[32..63] \\
-B &=& I_\mathbf{d}[64..(63 + \ell_B)] \\
-E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
-M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
-g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  M = 0 \\
- B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
-\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
 \section{Signing Transactions}\label{app:signing}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract. \\
-$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract. \\
-$G_{modexpbase}$ & 45 & Payment for the precompiled exponention under modulo. \\
-$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
-$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
 

### 2017-03-23 Paper.tex
Use bold font for \mu_s
as in other occurrences.

@@
 0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+0 < s &< \begin{dcases}
+\mathtt{\tiny secp256k1n} & \text{if} \quad H_i < \firsthomesteadblock \\
+\mathtt{\tiny secp256k1n} \div 2 + 1 & \text{otherwise} \\
+\end{dcases} \\
  v &\in \{27,28\}

### 2017-03-23 travis_deploy.sh
Initial commit

@@ -1,8 +1,16 @@
+<<<<<<< HEAD
 #!/usr/bin/env bash
 
 set -e
 
 SHA=`git rev-parse --verify HEAD`
+=======
+#!/bin/bash
+
+set -e
+
+SHA=`git rev-parse --short --verify HEAD`
+>>>>>>> Deploy pdf from travis.
 
 git config user.name "$COMMIT_AUTHOR"
 git config user.email "$COMMIT_AUTHOR_EMAIL"
@@ -10,10 +18,15 @@ git checkout --orphan gh-pages
 git rm --cached -r .
 echo "# Automatic build" > README.md
 echo "Built pdf from \`$SHA\`. See https://github.com/ethereum/yellowpaper/ for details." >> README.md
+<<<<<<< HEAD
 echo "The generated pdf is here: https://ethereum.github.io/yellowpaper/paper.pdf" >> README.md
 echo '<html><head><meta http-equiv="refresh" content="0; url=paper.pdf" /></head><body></body></html>' > index.html
 mv Paper.pdf paper.pdf
 git add -f README.md index.html paper.pdf
+=======
+mv Paper.pdf paper.pdf
+git add -f paper.pdf
+>>>>>>> Deploy pdf from travis.
 git commit -m "Built pdf from {$SHA}."
 
 ENCRYPTED_KEY_VAR="encrypted_${ENCRYPTION_LABEL}_key"

### 2017-03-29 Paper.tex
\Xi does not return output on exceptional halt
This makes the exceptional halts easier to distinguish from reverts

--- 2017-03-23 Paper.tex
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 \midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-03-29 Paper.tex
\Xi does not return output on exceptional halt
This makes the exceptional halts easier to distinguish from reverts

--- 2017-03-23 Paper.tex
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing \\
 g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}

### 2017-03-29 Paper.tex
Specify REVERT
Following https://github.com/ethereum/EIPs/pull/206/files

@@
 \end{eqnarray}
-\begin{equation}
+\begin{equation} \label{eq:X-def}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
 \big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d)
+(a, b, c) \cdot d & \equiv & (a, b, c, d) \\
+\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
+\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
 \end{eqnarray}
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if  \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
@@
 \midrule
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-03-29 Paper.tex
Specify REVERT
Following https://github.com/ethereum/EIPs/pull/206/files

--- 2017-03-29 Paper.tex
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule

### 2017-03-30 Paper.tex
Turn an unused I' into underscore

--- 2017-03-29 Paper.tex
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, \_, \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 
-Another useful function is ``all but one 64th'' function~$L$ defined as:
-
-\begin{equation}
-L(n) \equiv n - \lfloor n / 64 \rfloor
-\end{equation}
-
 \subsection{Instruction Set}

### 2017-03-30 Paper.tex
Turn an unused I' into underscore

--- 2017-03-29 Paper.tex
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 \end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if  \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
@@
 \midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-03-30 Paper.tex
underscore becomes standard ellision operator

@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, \_, \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\

### 2017-04-01 Paper.tex
Insufficient gas to store code results in an exception
This fixes a mistake in specification of Homestead.
Fix remaining issues from PR #167, namely equation 105

@@
 
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
 \end{equation}
@@
 \begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
 \end{equation}
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
+$G$ is defined in Appendix \ref{app:fees}.
 
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
 \end{eqnarray}
@@
 
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
-\end{align}
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\bottomrule
+\end{cases}$ \\\bottomrule
 \end{tabular*}

### 2017-04-11 Paper.tex
EIP161: document the anomaly
The description is based on https://github.com/ethereum/go-ethereum/pull/3341#discussion_r89548318

@@
 
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
 For particularly important values, a bold uppercase letter may be used.
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
-\end{cases}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
 \end{equation}

### 2017-04-12 Paper.tex
Fix typo (the => that)

--- 2017-04-10 Paper.tex
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
 It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.

### 2017-04-12 Paper.tex

--- 2017-04-11 Paper.tex
@@
 
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
 \subsection{The Transaction} \label{ch:transaction}
@@
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \{\}, 0)
+For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &\equiv& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
 \end{equation}
@@
 g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
+\end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_e & \equiv & e \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
@@
 \begin{eqnarray}
-\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
@@
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule

### 2017-04-12 Paper.tex
Take comments

--- 2017-04-12 Paper.tex
@@
 \boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv &
+\begin{cases}
+\boldsymbol{\sigma}[S(T)]_n &\text{if}\ S(T) = 2^{256} - 1 \\
+\boldsymbol{\sigma}[S(T)]_n + 1 &\text{otherwise} \\
+\end{cases}
 \end{eqnarray}
@@
 and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+$a$ is the address of the created account, which is defined as the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+Thus we define the resultant address for the new account $a$:
+\begin{equation}
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}_0[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}_0[x]$ is the account state of $x$ or $\varnothing$ if none exists.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, a)
+\end{equation}
 
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 \midrule
+0xfb & {\small CREATE\_P2SH} & 3 & 1 & Create a new account with associated code on an address calculated using the \\
+&&&& creator's address. \\
+&&&& Exactly equivalent to {\small CREATE} except: \\
+&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a \cdot \mathtt{\tiny KEC}(\mathbf{i})\big)\Big)$ \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-04-13 Paper.tex
EIP-196, EIP-197: some basic mathematical descriptions

--- 2017-04-12 Paper.tex
@@
 Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
 
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation*}
+x \equiv \begin{cases}
+1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+2 & \text{otherwise}
+\end{cases}
+\end{equation*}
 \begin{equation}
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
-\end{cases} \\
+\end{cases} \\ \nonumber
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+f(x) &\equiv& \begin{cases}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\end{cases}\\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
+8(\ell_E - 32) & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
 \section{Signing Transactions}\label{app:signing}

### 2017-04-13 Paper.tex
EIP101: some basic mathematical descriptions

--- 2017-04-12 Paper.tex
@@
 \boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv &
-\begin{cases}
-\boldsymbol{\sigma}[S(T)]_n &\text{if}\ S(T) = 2^{256} - 1 \\
-\boldsymbol{\sigma}[S(T)]_n + 1 &\text{otherwise} \\
-\end{cases}
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, a) & \text{if} \quad T_t = \varnothing \\
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
@@
 and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-$a$ is the address of the created account, which is defined as the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}_0[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}_0[x]$ is the account state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, a)
-\end{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+Thus we define the resultant address for the new account $a$:
+\begin{equation}
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
 \section{Signing Transactions}\label{app:signing}
@@
 \begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r) \quad \text{ where } (T_w, T_r, T_s) \text{ is a valid signature.}
-\end{equation}
-
-When a transaction has a particular invalid signature and it has zero gas price, zero nonce and zero value, the sender of the transaction is $2^{256} - 1$.
-\begin{equation}
-\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } \quad (T_w, T_r, T_s) = (\mathtt{chain\_id}, 0, 0) \,\wedge\, T_p = T_n = T_v = 0.
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
 \end{equation}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& where $a$ is defined as in (\ref{eq:address_calc}) \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 \midrule
-0xfb & {\small CREATE\_P2SH} & 3 & 1 & Create a new account with associated code on an address calculated using the \\
-&&&& creator's address. \\
-&&&& Exactly equivalent to {\small CREATE} except: \\
-&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a \cdot \mathtt{\tiny KEC}(\mathbf{i})\big)\Big)$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-04-13 Paper.tex
EIP212: pairing checking precompiled contract
As described in https://github.com/ethereum/EIPs/pull/212

@@
 With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+g_r&=&?\\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 0\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
 

### 2017-04-13 Paper.tex
EIP-197: pairing checking precompiled contract
As described in https://github.com/ethereum/EIPs/pull/212

--- 2017-04-13 Paper.tex
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+f(x) &\equiv& \begin{cases}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\end{cases}\\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
+8(\ell_E - 32) & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
 \end{eqnarray}

### 2017-04-13 Paper.tex
EIP 213
Define addition precompiled contract on G_1

--- 2017-04-13 Paper.tex
@@
 b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
 \end{eqnarray}

### 2017-04-13 Paper.tex
EIP 213
Defining the precompiled contract for scalar multiplication on G_1

--- 2017-04-13 Paper.tex
@@
 \end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}

### 2017-04-13 Paper.tex
EIP 213
Defining the precompiled contract for scalar multiplication on G_1

@@
 
+\subsection{Homestead} \label{ch:homestead}
+
+A significant block number for compatibility with the public network is the block marking the transition between the {\it Frontier} and {\it Homestead} phases of the platform, which we denote with the symbol \firsthomesteadblock, defined thus
+\begin{equation}
+\firsthomesteadblock \equiv 1,\! 150,\! 000
+\end{equation}
+The protocol was upgraded at this block, so this symbol appears in some equations to account for the changes.
+
 \subsection{The Transaction} \label{ch:transaction}

### 2017-04-18 Paper.tex

--- 2017-04-13 Paper.tex
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-g_r&=&?\\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 0\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r&=&?\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r&=&?\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}

### 2017-04-18 Paper.tex

@@
 I_e & \equiv & e \\
+I_w & \equiv & w \\
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 \item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_e$, the depth of the present message-call or contract-creation
+      (i.e. the number of {\small CALL}s or {\small CREATE}s being
+      executed at present).
+\item $I_w$, the permission to write to the state.
 \end{itemize}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and
+     $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
 \midrule

### 2017-04-18 Paper.tex

@@
 \boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1 \quad \text{if} \, S(T) \neq 2^{256}
 \end{eqnarray}
@@
 \begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r) \quad \text{ where } (T_w, T_r, T_s) \text{ is a valid signature.}
+\end{equation}
+
+When a transaction has a particular invalid signature $v = r = s = 0$, the sender of the transaction is $2^{256} - 1$.
+\begin{equation}
+\forall T: S(T) \equiv 2^{256} - 1 \quad \text{ where } (T_w, T_r, T_s) = (0, 0, 0).
 \end{equation}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& where $a$ is defined as in (\ref{eq:address_calc}) \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ if $I_a \neq 2^{255}$. \\
 &&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
@@
 \midrule
+0xfb & {\small CREATE\_P2SH} & 3 & 1 & Create a new account with associated code on an address calculated using the \\
+&&&& creator's address. \\
+&&&& Exactly equivalent to {\small CREATE} except: \\
+&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a \cdot \mathbf{i}\big)\Big)$ \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-04-18 Paper.tex

--- 2017-04-18 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 8 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 9 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision addition on non-negative integers.
+The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
+The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
+
+\begin{eqnarray}
+\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\, X = Y = 0 \\
+\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
+
+\begin{eqnarray}
+\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
+g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
+\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = Y \\
+\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The input format is the same as that of the addition contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
+\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The eigth contract performs arbitrary-precision division on non-negative integers.  The definition is similar to that of the multiplication contract.
+
+\begin{eqnarray}
+\Xi_{\mathtt{DIV}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
+g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  Y = 0 \\
+ \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
+\ell_X &=& I_\mathbf{d}[0..31] \\
+Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
+\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
+\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+The ninth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+\ell_B &=& I_\mathbf{d}[0..31] \\
+B &=& I_\mathbf{d}[32..(31 + \ell_B)] \\
+\ell_E &=& I_\mathbf{d}[(32 + \ell_B)..(63 + \ell_B)] \\
+E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
+M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
+g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  M = 0 \\
+ B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
+\end{cases} \\
+\ell &=&
+\begin{cases}
+0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
+\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
+\end{cases}
 \end{eqnarray}

### 2017-04-18 Paper.tex
Clarify dependency of H on h in BLOCKHASH instruction

--- 2017-04-18 Paper.tex
@@
 %\usepackage{tweaklist}
-\usepackage{url}
 \usepackage{cancel}
@@
 \usepackage{tikz}
-
-\input{Version.tex}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+%,pagebackref=true
+
+\usepackage{tabu} %requires array.
+
+%This should be the last package before \input{Version.tex}
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
+% "hyperref loads the url package internally.
+Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
+This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
+% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
+Try loading the url package earlier than the package that requires it.
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
+
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\usepackage{seqsplit}
+
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+
+% Default rendering options
+\definecolor{pagecolor}{rgb}{1,0.98,0.9}
+\def\YellowPaperVersionNumber{unknown revision}
+\IfFileExists{Options.tex}{\input{Options.tex}}
 
@@
 
-\definecolor{lightyellow}{rgb}{1,0.98,0.9}
-\definecolor{lightpink}{rgb}{1,0.94,0.95}
-
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Founder, Ethereum \& Parity\\
+    gavin@parity.io
 }
@@
 
-\pagecolor{lightyellow}
-%\pagecolor{lightpink}
+\pagecolor{pagecolor}
 
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
-This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
+This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience, or corruption of existing legal systems.
 By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
@@
 
-Overall, I wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
+Overall, we wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
 
@@
 The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, ensuring ``consumers'' be able to make micro-payments to ``suppliers'' for their services.
+In this instance, the token was used to keep peer-to-peer file trading in check, providing ``consumers'' with the ability to make micro-payments to ``suppliers'' for their services.
 The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
@@
 %E language?
+For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (\hyperlink{Genesis_Block}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 
-I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
+We use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
 
@@
 
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the Ethereum state transition function.
 
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
+For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
+Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
 
@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \ref{ch:block}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
+This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
+
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
-
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
+
+When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
+It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
+
+We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
 
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
-\begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
-\begin{equation}
-L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
+\begin{equation}
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
+\begin{equation}
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
-\begin{equation}
-L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_{S}$:
+\begin{equation}
+L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
-\end{equation}
-
-This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
+\end{equation}
+
+This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
 \begin{equation}
@@
 \end{equation}
-where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{equation}
-
-\subsection{The Transaction} \label{ch:transaction}
+\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+\end{equation}
+
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
+\end{equation}
+Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\end{equation}
+
+\subsection{The Transaction} \label{subsec:transaction}
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
+\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathbf{i}}$.
 \end{description}
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathbf{d}}$.
 \end{description}
@@
 \begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+L_{T}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\subsection{The Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
+\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
+\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
+\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
+\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
+The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
+\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
-
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
+\end{equation}
+
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
+\end{equation}
+
+\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+\end{equation}
+where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\begin{equation}
+R_{\mathrm{z}} \in \mathbb{P}
+\end{equation}
+
+\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+
+The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
-Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
 \end{equation}
-
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
+
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\end{eqnarray}
+
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\begin{equation}
+\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
+\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 
-We define $P(B_H)$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
-\end{equation}
-
-The block number is the parent's block number incremented by one:
-\begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+We define $P(B_{H})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+\end{equation}
+
+\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
+\begin{equation}
+H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
 \end{equation}
@@
 
-The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
-
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 125000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation*}
+x \equiv \begin{cases}
+1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
+2 & \text{otherwise}
+\end{cases}
+\end{equation*}
+\begin{align}
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
+\end{align}
+
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000
+\end{eqnarray}
+
+\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
+\begin{equation}
+H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
 \end{equation}
@@
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
 
@@
 
-Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 125000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
+& & H_{\mathrm{d}} = D(H) \quad \wedge \\
+& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
+& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+& & \lVert H_{\mathrm{x}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
+The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon_state_transition}{$\Upsilon$}.
 It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
@@
 \item the transaction signature is valid;
-\item the transaction nonce is valid (equivalent to the sender account's current nonce);
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
+\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
 \item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
@@
 
-Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
+Formally, we consider the function \hyperlink{Upsilon_state_transition}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
 \begin{equation}
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
+We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
+These will be formally defined later.
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
 
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), 0)
+We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing,(), \varnothing, 0)
 \end{equation}
@@
 \subsection{Execution}
-We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
+g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
 \end{align}
 
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
 
-%todo Explain g_d reason?
-
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n \bmod 2^{64} & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-\big(\lfloor T_n / 2^{64} \rfloor = 0 & \vee & \lfloor T_n / 2^{64} \rfloor = \mathtt{\small chain\_id}\ \big) \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+g \equiv T_{\mathrm{g}} - g_0
+\end{equation}
+and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+
+Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_{H}}_{\mathrm{c}}
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
-\end{eqnarray}
-
-These are used to help define the transaction receipt, discussed later.
-
-\section{Contract Creation} \label{ch:create}
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
+
+\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s] &=& \begin{cases}
+\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
+\mathbf{a}^* & \text{otherwise}
+\end{cases} \\
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_b & \text{otherwise}
+\boldsymbol{\sigma}[a]_{\mathrm{b}} & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
-\end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
-\begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
+\end{equation}
+\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
+\begin{eqnarray}
+I_{\mathrm{a}} & \equiv & a \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathbf{d}} & \equiv & () \\
+I_{\mathrm{s}} & \equiv & s \\
+\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
+I_{\mathbf{b}} & \equiv & \mathbf{i} \\
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w
+\end{eqnarray}
+
+$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
 
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases} \\
+\quad z &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
+\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
+\mathbf{a}_1 &\text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+z & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
+1 & \text{otherwise}
 \end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
-
-As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+I_{\mathrm{a}} & \equiv & r \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathbf{d}} & \equiv & \mathbf{d} \\
+I_{\mathrm{s}} & \equiv & s \\
+I_{\mathrm{v}} & \equiv & \tilde{v} \\
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w \\
+\mathbf{t} & \equiv & \{s, r\} \\
+\\ \nonumber
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
+
+As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
 
@@
 
-\subsection{Execution Environment}
+\subsection{Execution Environment}\linkdest{exec_env}
 
@@
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
-\end{eqnarray}
-\begin{equation}
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
+
+\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
+\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
+\end{eqnarray}
+\begin{equation} \label{eq:X-def}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
+\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
+\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
@@
 The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
+The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
-\text{\small STOP} & \text{otherwise}
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert \\
+\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
 \end{cases}
@@
 
-\subsubsection{Exceptional Halting}
+\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
+ \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+ \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined {\small STOP} operations that trail it).
+All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).
 
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
 \end{equation}
@@
 \begin{equation}
-D_J(\mathbf{c}, i) \equiv \begin{cases}
+D_{J}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 
-\subsubsection{Normal Halting}
+\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}\linkdest{hhalt}
 
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
+() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
+Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
 
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
+\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
+\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
+\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
+\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
+N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
+\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
 A' & \equiv & A \\
@@
 In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
+Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{Genesis_Block}{genesis block}, in the path.
 The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
@@
 \begin{eqnarray}
-B_t & \equiv & B'_t + B_d \\
-B' & \equiv & P(B_H)
-\end{eqnarray}
-
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
+B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
+B' & \equiv & P(B_{H})
+\end{eqnarray}
+
+As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
 
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and nonce.
+\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
 \end{enumerate}
@@
 
-The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
+The validation of \hyperlink{ommer_block_headers_B__U}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
 The maximum of ommer headers is two.
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathbf{U}})
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
-\begin{eqnarray}
+We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+\begin{eqnarray}
+\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
+\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
+\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
+\mathbf{a}' &\text{otherwise}
+\end{cases} \\
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
 \end{eqnarray}
@@
 
-We define the block reward as 5 Ether:
-\begin{equation}
-\text{Let} \quad R_b = 5 \times 10^{18}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
 \end{equation}
@@
 
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+
+\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
-\end{eqnarray}
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+\end{eqnarray}
+
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
-\end{equation}
-
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_{\mathbf{l}} =
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\end{equation}
+
+We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{z}} =
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\end{equation}
+
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
+\end{equation}
+
+Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
 
@@
 \begin{equation}
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
 The general route that the algorithm takes is as follows:
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 
-There are several patterns of contracts engineering that allow particular useful behaviours; two of these that I will briefly discuss are data feeds and random numbers.
+There are several patterns of contracts engineering that allow particular useful behaviours; two of these that we will briefly discuss are data feeds and random numbers.
 
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
-The state database won't be forced to maintain all past state trie structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
+Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
 
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{Genesis_Block}{genesis block}.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidated each block and with duplicate or invalid transactions thrown out accordingly.
+Parallelism could also come from a prioritised set of parallel blockchains, consolidating each block and with duplicate or invalid transactions thrown out accordingly.
 
@@
 
-I have introduced, discussed and formally defined the protocol of Ethereum.
+We have introduced, discussed and formally defined the protocol of Ethereum.
 Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
@@
 
+\bibliographystyle{plainnat}
 \bibliography{Biblio}
-\bibliographystyle{plainnat}
 
@@
 
-\section{Terminology}
+\section{Terminology} \label{ch:Terminology}
 
@@
 
-\section{Recursive Length Prefix}\label{app:rlp}
+\section{Recursive Length Prefix}\label{app:rlp}\hypertarget{rlp}{}
 This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
+\end{equation}
+
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
+\begin{eqnarray}
+R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 
@@
 
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
+\begin{eqnarray}
+R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 
-\section{Modified Merkle Patricia Tree}\label{app:trie}
+\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
@@
 \begin{equation}
-\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_{\mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+\forall_{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_{n}\rVert} \quad \mathbf{k}_{n}'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
-
-For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
+\begin{equation} \label{eq:pre}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
+
+We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
+See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
+
 \begin{eqnarray}
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+g_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
-\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
+\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 
-The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
+We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively.
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
+
 \begin{eqnarray}
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
 \end{eqnarray}
@@
 For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
+
 \begin{eqnarray}
@@
 
-Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_{\mathbf{d}}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
+f(x) &\equiv& \begin{cases}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\end{cases}\\
+\ell'_{E} &=& \begin{cases}
+0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
+8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
+8(\ell_{E} - 32) & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
+\ell_{B} &\equiv& i[0..31] \\
+\ell_{E} &\equiv& i[32..63] \\
+\ell_{M} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{B})] \\
+E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
+M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
+\begin{equation}
+\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
+\end{eqnarray}
+
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
+k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
+g_{\mathrm{r}}&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_{\mathrm{r}} &=& 500\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_{\mathrm{r}} &=& 40000\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
-
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
+This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
+
+It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+
+We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
+These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
 
@@
 \begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
+0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
+0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
 \end{align}
@@
 
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
-\begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+\end{equation}
+
+\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+Two different flavors of signing schemes are available.
+One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
+The other operates on nine elements:
+\begin{eqnarray}
+L_{S}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\nonumber \text{where} \\
+\nonumber \mathbf{p} & \equiv & \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
+T_{\mathbf{d}} & \text{otherwise}
+\end{cases} \\
+h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
+\begin{eqnarray}
+G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+\end{eqnarray}
+
+\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
+\begin{eqnarray}
+\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
+\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
+\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
 \end{eqnarray}
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
-\end{equation}
+\begin{eqnarray}
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
+v_0 &\equiv& \begin{cases}
+T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
+28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
+\end{cases}
+\end{eqnarray}
 
@@
 \begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
+\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
 \end{equation}
@@
 $G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
 $G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
 $G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
+\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
 $G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
+\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+
+%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
 G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
+G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
 G_{extcode} & \text{if} \quad w \in W_{extcode}\\
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
+G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
 
@@
 
-$W_{high}$ = \{{\small JUMPI}\}
+$W_{\mathrm{high}}$ = \{{\small JUMPI}\}
 
@@
 
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod \boldsymbol{\mu}_{\mathbf{s}}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0]) (|\boldsymbol{\mu}_{\mathbf{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathbf{s}}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
+\midrule
 \end{tabular*}
@@
 0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+\midrule
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
+\midrule
+0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
 \bottomrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H(h)_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H(h)_{\mathrm{i}} \\ P(H(h)_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H(h)$ can be determined from its hash~$h$ unless $h$ is zero. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
-\midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
+\midrule
+\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
+&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \midrule
-0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}.\\
+\midrule
+\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
 \midrule
 0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
@@
 0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[15]$ \\
 \bottomrule
@@
 0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[1] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
 \midrule
 0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[2]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[2] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
 \midrule
@@
 0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[16]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[16] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
 \bottomrule
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
 \midrule
@@
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4], \boldsymbol{\mu}_{\mathbf{s}}[5])$ \\
 \bottomrule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
+&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
+&&&& created account. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
@@
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
 \midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
+\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \end{tabular*}
@@
 &&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 \midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
+\midrule
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we use the memory expansion function, \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
 \midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+\end{cases}$\\
+&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
 \end{cases}$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 
-%\section{Low-level Lisp-like Language}\label{app:lll}
-%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-
-\section{Genesis Block}\label{app:genesis}
+\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
 
@@
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
 
@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
+\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+ E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
 \end{equation}
@@
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
+\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+ c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
+E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+ n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
 \begin{equation}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_{\mathrm{hashbytes}}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+ E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
 $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+ \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{\mathrm{hashbytes}}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
+\begin{equation}
+ E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
+E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}
@@
 
+\section{Anomalies on the Main Network}
+
+\subsection{Deletion of an Account Despite Out-of-gas}
+
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
+
+\section{List of mathematical symbols} \label{app:symbols}
+\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
+\toprule
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+\bottomrule
+\end{tabu*}
 \end{document}

### 2017-04-20 Paper.tex
The nonce in a block header is not a hash.

@@
 \item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H(h)_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H(h)_{\mathrm{i}} \\ P(H(h)_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H(h)$ can be determined from its hash~$h$ unless $h$ is zero. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule

### 2017-04-20 Paper.tex
In many implementations, the used gas has refunds already considered
Fix #299 (see the issue for details)

--- 2017-04-18 Paper.tex
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g^* \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision addition on non-negative integers.
-The first word in the input specifies the number of bytes that the first non-negative integer $X$ occupies.
-The result is represented in the smallest possible number of bytes.  The first byte in the output is never zero.
-
-\begin{eqnarray}
-\Xi_{\mathtt{ADD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{ADD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X + Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\, X = Y = 0 \\
-\lfloor \log_8(X + Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The sixth contract performs arbitrary-precision subtraction on non-negative integers.  This is similar to the addition.  The subtraction contract halts exceptionally if the result would be negative.
-
-\begin{eqnarray}
-\Xi_{\mathtt{SUB}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-\Xi_{\mathtt{SUB}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad X < Y \\
-g_r &=& G_{addsubbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil \\
-\mathbf{o} &=& (X - Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = Y \\
-\lfloor \log_8(X - Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The seventh contract performs arbitrary-precision multiplication on non-negative integers.  The input format is the same as that of the addition contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{MUL}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{MUL}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=& (X \times Y) \in \mathbb{B}_\ell \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ X = 0 \ \vee\  Y = 0 \\
-\lfloor \log_8(X \times Y) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The eigth contract performs arbitrary-precision division on non-negative integers.  The definition is similar to that of the multiplication contract.
-
-\begin{eqnarray}
-\Xi_{\mathtt{DIV}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\Xi_{\mathtt{DIV}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| - 32 <  I_\mathbf{d}[0..31] \\
-g_r &=& G_{muldivbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + \ell_X \ell_Y / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  Y = 0 \\
- \Big\lfloor \dfrac X Y \Big\rfloor \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-X &=& I_\mathbf{d}[32..(32 + I_\mathbf{d}[0..31] - 1)] \\
-\ell_X &=& I_\mathbf{d}[0..31] \\
-Y &=& I_\mathbf{d}[(32 + I_\mathbf{d}[0..31])..(|I_\mathbf{d}| - 1)] \\
-\ell_Y &=& |I_\mathbf{d}| - 32 - \ell_X \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ Y = 0 \ \vee \ \Big\lfloor \dfrac X Y \Big\rfloor = 0 \\
-\lfloor \log_8(\Big\lfloor \dfrac X Y \Big\rfloor) \rfloor + 1 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The ninth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-\ell_B &=& I_\mathbf{d}[0..31] \\
-B &=& I_\mathbf{d}[32..(31 + \ell_B)] \\
-\ell_E &=& I_\mathbf{d}[(32 + \ell_B)..(63 + \ell_B)] \\
-E &=& I_\mathbf{d}[(64 + \ell_B)..(63 + \ell_B + \ell_E)] \\
-M &=& I_\mathbf{d}[(64 + \ell_B + \ell_E)..(|I_\mathbf{d}| - 1)] \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + \ell_B + \ell_E \\
-g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  M = 0 \\
- B ^ E \bmod M \in \mathbb{B}_\ell & \text{otherwise}
-\end{cases} \\
-\ell &=&
-\begin{cases}
-0 &\text{if}\ M = 0 \quad \vee \quad {B ^ E} \bmod M = 0 \\
-\lfloor \log_8(B ^ E \bmod M) \rfloor + 1 & \text{otherwise}
-\end{cases}
 \end{eqnarray}

### 2017-04-20 Paper.tex
After CREATE, the 1/64 of the original available gas is added again
Before this commit, after a successful CREATE, the available gas decreased at least by 1/64.
This problem was introduced in EIP-150 changes.  The gas passed on was reduced by 1/64 but this 1/64 was then forgotten.
Gas usage of CALL should first take C_CALLGAS into account
Before this change, the CALL instruction was receiving the remainder of the gas from the callee, but not losing the initial gas passed to the callee.
This change makes the caller to lose the initial gas passed to the callee before receiving the remainder.

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 
-\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet.
-The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}.
-\cite{back2002hashcash} later produced a system in a similar vein.
-
-The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, ensuring ``consumers'' be able to make micro-payments to ``suppliers'' for their services.
-The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
-Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope.
-The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.
+Hashcash, introduced by \cite{back2002hashcash} (in a five-year retrospective), provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
+Though not widely adopted, the work was later utilised and expanded upon by \cite{nakamoto2008bitcoin} in order to devise a cryptographically secure mechanism for coming to a decentralised social consensus over the order and contents of a series of cryptographically signed financial transactions.
+The fruits of this project, Bitcoin, provided a first glimpse into a decentralised transaction ledger.
 
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
 Formally:
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g^* \\
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
@@
 
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
 \section{Contract Creation} \label{ch:create}
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
 
@@
 
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
-\end{align}
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule
@@
 
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2017-04-24 Paper.tex
Remove add, sub, mul and div precompiles
Because https://github.com/ethereum/EIPs/pull/198 supersedes https://github.com/ethereum/EIPs/issues/101 and the new version does not contain these.

--- 2017-04-20 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+These two words are followed by $B$ and $E$; and the rest of the input is interpreted as the third non-negative integer $M$.
+All non-negative integers $B$, $E$ $M$ are encoded as byte sequences in the big-endian way.
+The output format is the same as the precompiled contract $\Xi_{\mathtt{ADD}}$.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
+l_B &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
+l_E &=& I_\mathbf{d}[32..63] \in \mathbb{P}_{256} \\
+B &=& I_\mathbf{d}[64..(63 + l_B)] \\
+E &=& I_\mathbf{d}[(64 + l_B)..(63 + l_B + l_E)] \\
+M &=& I_\mathbf{d}[(64 + l_B + l_E)..(|I_\mathbf{d}| - 1)] \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + l_B + l_E\,\vee\,M\le B \\
+g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
+\mathbf{o} &=&
+\begin{cases}
+ () & \text{if} \  M = 0 \\
+ \mathtt{\tiny BE}(B ^ E \bmod M) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}

### 2017-04-24 Paper.tex
Reformulate EXPMOD as https://github.com/ethereum/EIPs/pull/198 specifies

@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.  Here, $0 ^ 0$ is taken to be one.
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is ???.
 The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
 The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-These two words are followed by $B$ and $E$; and the rest of the input is interpreted as the third non-negative integer $M$.
-All non-negative integers $B$, $E$ $M$ are encoded as byte sequences in the big-endian way.
-The output format is the same as the precompiled contract $\Xi_{\mathtt{ADD}}$.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:}\\
-l_B &=& I_\mathbf{d}[0..31] \in \mathbb{P}_{256} \\
-l_E &=& I_\mathbf{d}[32..63] \in \mathbb{P}_{256} \\
-B &=& I_\mathbf{d}[64..(63 + l_B)] \\
-E &=& I_\mathbf{d}[(64 + l_B)..(63 + l_B + l_E)] \\
-M &=& I_\mathbf{d}[(64 + l_B + l_E)..(|I_\mathbf{d}| - 1)] \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}, g, I) &\equiv& (\varnothing, 0, A^0, ()) \quad \text{if} \quad |I_\mathbf{d}| < 64 + l_B + l_E\,\vee\,M\le B \\
-g_r &=& G_{modexpbase} + G_{arithword} \Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil + |M|^2 |E| / G_{quaddivisor} \\
-\mathbf{o} &=&
-\begin{cases}
- () & \text{if} \  M = 0 \\
- \mathtt{\tiny BE}(B ^ E \bmod M) & \text{otherwise}
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
 \end{cases}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{addsubbase}$ & 15 & Payment for the precompiled addition or subtraction contract. \\
-$G_{muldivbase}$ & 30 & Payment for the precompiled multiplication or division contract. \\
-$G_{modexpbase}$ & 45 & Payment for the precompiled exponention under modulo. \\
-$G_{arithword}$ & 6 & Paid for each word used in precompiled contracts for arbitrary precision arighmetics.\\
-$G_{quaddivisor}$ & 32 & The quadratic coefficient of the input sizes of multiplication and division precompiled contracts.  \\
+$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract.  \\
 

### 2017-04-24 Paper.tex
Reformulate EXPMOD as https://github.com/ethereum/EIPs/pull/198 specifies

--- 2017-04-22 Paper.tex
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 g^{**} & \text{otherwise}
-\end{cases} \\
+\end{cases} \\ \nonumber
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is ???.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract.  \\
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}'', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2017-04-24 Paper.tex
Specify modulo zero
EIP 86: the original CREATE instruction retains the original address computation

--- 2017-04-24 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is ???.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract.  \\
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(I_a, \boldsymbol{\sigma}[I_a]_n - 1)\;\big)\Big)\Big)$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\

### 2017-04-25 Paper.tex
Pull ommer's uniqueness test out of the "is-sibling" equation. Add a textual clarification for this test.

--- 2017-04-24 Paper.tex
@@
 The maximum of ommer headers is two.
+An ommer may be included only in one block in a blockchain.
 Formally:
@@
 k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
-s(U, H) &\\
-\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
+(s(U, H) &\\
+\quad \vee \; k(U, P(H)_H, n - 1)) &\\
+\quad \wedge \; U \notin B(H)_\mathbf{U} & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U)
 \end{equation}
@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is ???.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
 \section{Signing Transactions}\label{app:signing}

### 2017-04-25 Paper.tex
Normalizing the input of ECDSARECOVER function

@@
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 The maximum of ommer headers is two.
-An ommer may be included only in one block in a blockchain.
 Formally:
@@
 k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
-(s(U, H) &\\
-\quad \vee \; k(U, P(H)_H, n - 1)) &\\
-\quad \wedge \; U \notin B(H)_\mathbf{U} & \text{otherwise}
+s(U, H) &\\
+\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U)
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
 \end{equation}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The chain identifier is a constant natural number.
 
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35 , \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+Two different flavors of signing schemes are available.
+One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+The other operates on nine elements:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\}\\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}\\
+\end{cases} \\
+\nonumber \text{where} \\
+\nonumber \mathbf{p} & \equiv & \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\

### 2017-04-25 Paper.tex

@@
 \midrule
-$10^0$ & Wei \\
-$10^{12}$ & Szabo \\
-$10^{15}$ & Finney \\
-$10^{18}$ & Ether \\
+$10^{18}$ & Wei \\
+$10^{15}$ & Szabo \\
+$10^{12}$ & Finney \\
+$10^0$ & Ether \\
 \bottomrule

### 2017-04-28 Paper.tex
EIP98 option 1
Before this PR, the metropolis branch contained EIP98 option 2, but the consensus now is option 1.

@@
 \midrule
-$10^{18}$ & Wei \\
-$10^{15}$ & Szabo \\
-$10^{12}$ & Finney \\
-$10^0$ & Ether \\
+$10^0$ & Wei \\
+$10^{12}$ & Szabo \\
+$10^{15}$ & Finney \\
+$10^{18}$ & Ether \\
 \bottomrule
@@
 This process is described in detail in section \ref{ch:ghost}.
-
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
-In order to follow back the history of a path, one might need to reference multiple versions of this document.
 
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+L_R(R) \equiv (R_u, R_b, R_\mathbf{l})
+\end{equation}
 

### 2017-05-08 Paper.tex
EIP96: update the BLOCKHASH instruction

@@
 \begin{equation}
-L_R(R) \equiv (R_u, R_b, R_\mathbf{l})
-\end{equation}
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
 
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathbf{o}'[0..31]$ \\
+&&&& $\mathbf{o}'[i] \equiv \begin{cases}
+\mathbf{o}[i] & \text{if}\ 0 \le i < |\mathbf{o}| \\
+0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $(\ldots,\mathbf o) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, t, t, 10000000, I_p, 0, 0, \boldsymbol{\mu}_{\mathbf{s}}[0], I_e + 1)$ \\
+&&&& $t \equiv 240$ \\
+&&&& See Section~\ref{sec:blockhash-early-metropolis} for the special behavior of {\small BLOCKHASH} instruction during the first \\
+&&&& 256 blocks in the Metropolis phase. \\
 \midrule

### 2017-05-08 Paper.tex
EIP96: specify BLOCKHASH instruction during the first 256 blocks in Metropolis

@@
 
+\section{Anomalies on the Main Network}
+
+\subsection{Early behavior of BLOCKHASH instruction}
+\label{sec:blockhash-early-metropolis}
+
+During the first 256 blocks in the Metropolis phase, {\small BLOCKHASH} instruction behaves as specified below.
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
+\bottomrule
+\end{tabular*}
+
 \end{document}

### 2017-05-09 Paper.tex
EIP96: change the fee for BLOCKHASH
https://github.com/ethereum/EIPs/pull/210/files#diff-e02a92c2fb96c1a1bfb05e4c6e2ef5daR42
EIP96 [WIP] describe the setup of BLOCKHASH contract.
EIP86: specify CREATE2

--- 2017-04-28 Paper.tex
@@
 \midrule
+0xfb & {\small CREATE2} & 4 & 1 & Create a new account with associated code on a predictable address. \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[2]..(\boldsymbol{\mu}_\mathbf{s}[2] + \boldsymbol{\mu}_\mathbf{s}[3] - 1)]$ \\
+&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a\cdot \boldsymbol{\mu}_\mathbf{s}[1]\cdot \mathtt{\tiny KEC}(\mathbf i)\big)\Big)$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}_g', A^+) \equiv \begin{cases}
+\begin{array}{l}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \\ \phantom{\Lambda(}\boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a)\end{array} & \text{if}\begin{array}{l}(\boldsymbol{\sigma}^*[a]=\varnothing\vee\boldsymbol{\sigma}^*[a]_c=())\wedge\\\boldsymbol{\mu}_\mathbf{s}[0]\le\boldsymbol{\sigma}[I_a]_b\wedge I_e<1024\end{array}\\
+(\boldsymbol{\sigma},\boldsymbol{\mu}_g,\varnothing)&\text{otherwise}
+\end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^*\equiv\boldsymbol{\sigma}$ except $\boldsymbol{\sigma}^*[I_a]_n=\boldsymbol{\sigma}[I_a]_n+1$ \\
+&&&& $A'\equiv A\Cup A^+$ (with $\Cup$ defined in the definition of {\small CREATE}) \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x = 0$ if the code execution for this operation failed due to an exceptional \\
+&&&& halting $Z(\boldsymbol{\sigma}^*,\boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ (the maximum call depth limit is reached) or \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too low to fulfill the balance transfer) or \\
+&&&& $\boldsymbol{\sigma}[a]_c \neq ()$ (a contract already exists); otherwise $x = a$. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& Thus the operand order is: value, salt, input offset, input size. \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-05-09 Paper.tex
Remove a duplicate `otherwise`

--- 2017-05-08 Paper.tex
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathbf{o}'[0..31]$ \\
-&&&& $\mathbf{o}'[i] \equiv \begin{cases}
-\mathbf{o}[i] & \text{if}\ 0 \le i < |\mathbf{o}| \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $(\ldots,\mathbf o) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, t, t, 10000000, I_p, 0, 0, \boldsymbol{\mu}_{\mathbf{s}}[0], I_e + 1)$ \\
-&&&& $t \equiv 240$ \\
-&&&& See Section~\ref{sec:blockhash-early-metropolis} for the special behavior of {\small BLOCKHASH} instruction during the first \\
-&&&& 256 blocks in the Metropolis phase. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Early behavior of BLOCKHASH instruction}
-\label{sec:blockhash-early-metropolis}
-
-During the first 256 blocks in the Metropolis phase, {\small BLOCKHASH} instruction behaves as specified below.
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
-\bottomrule
-\end{tabular*}
-
 \end{document}

### 2017-05-09 Paper.tex
Move RETURNDATACOPY and RETURNDATASIZE to the 0x3. block

@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
-\end{cases} \\
+\end{cases} \\ \nonumber
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+\midrule
 \end{tabular*}
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+\midrule
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
+\midrule
+0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\

### 2017-05-09 Paper.tex
Move RETURNDATACOPY and RETURNDATASIZE to the 0x3. block

--- 2017-05-09 Paper.tex
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+\midrule
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
+\midrule
+0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account. \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\

### 2017-05-10 Paper.tex
EIP86 rule 5

@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \vee E(\boldsymbol{\sigma}, a) \\
 g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \vee E(\boldsymbol{\sigma}, a) \\
 \boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+E(\boldsymbol{\sigma}, a) &\equiv \big(\boldsymbol{\sigma}[a] \neq \varnothing \wedge (\boldsymbol{\sigma}[a]_c \neq () \vee \boldsymbol{\sigma}[a]_n \neq 0)\big)
 \end{align}
@@
 \multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\midrule
 \end{tabular*}
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\

### 2017-05-10 Paper.tex
EIP86: when CREATE(2) destination is occupied, the return value is zero

@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$, $E(\boldsymbol{\sigma}^*, a) = \top$ (the address is occupied) or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 &&&& where $x = 0$ if the code execution for this operation failed due to an exceptional \\
-&&&& halting $Z(\boldsymbol{\sigma}^*,\boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ (the maximum call depth limit is reached) or \\
+&&&& halting $Z(\boldsymbol{\sigma}^*,\boldsymbol{\mu}, I) = \top$, $E(\boldsymbol{\sigma}, a) = \top$ (the address is already occupied) or \\
+&&&& $I_e = 1024$ (the maximum call depth limit is reached), \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too low to fulfill the balance transfer) or \\

### 2017-05-10 Paper.tex
EIP 140: the output from REVERT during a contract creation
should be accessible from RETURNDATASIZE and RETURNDATACOPY, so \Lambda should return four values.

--- 2017-05-09 Paper.tex
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
@@
 
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
-\end{eqnarray}
-\begin{equation}
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation} \label{eq:X-def}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\

### 2017-05-10 Paper.tex
EIP 140: the output from REVERT during a contract creation
should be accessible from RETURNDATASIZE and RETURNDATACOPY, so \Lambda should return four values.

@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+0 & \text{if} \quad F \\
 g^{**} - c & \text{otherwise} \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d) \\
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 \midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
 0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes

### 2017-05-10 Paper.tex
Minor improvement of grammar, particularly comma use and that/which clause awkwardness.

--- 2017-05-10 Paper.tex
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
@@
 
-Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing \\
-g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} - c & \text{otherwise} \\
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\

### 2017-05-10 Paper.tex
EIP 101: remove spurious spaces

--- 2017-05-10 Paper.tex
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c) \cdot d & \equiv & (a, b, c, d) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
 \end{eqnarray}
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 

### 2017-05-10 Paper.tex
EIP 101: remove spurious spaces

--- 2017-05-10 Paper.tex
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
 \end{eqnarray}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule

### 2017-05-10 Paper.tex
Adjust white spaces

--- 2017-05-10 Paper.tex
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger then 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
 \end{eqnarray}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 

### 2017-05-10 Paper.tex
Remove a spurious newline

--- 2017-05-10 Paper.tex
@@
 \boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \begin{cases}
+\boldsymbol{\sigma}[S(T)]_n &\text{if}\ S(T) = 2^{256} - 1 \\
+\boldsymbol{\sigma}[S(T)]_n + 1 &\text{otherwise} \\
+\end{cases}
 \end{eqnarray}
@@
 \midrule
+0xfb & {\small CREATE2} & 4 & 1 & Create a new account with associated code on a predictable address. \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[2]..(\boldsymbol{\mu}_\mathbf{s}[2] + \boldsymbol{\mu}_\mathbf{s}[3] - 1)]$ \\
+&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a\cdot \boldsymbol{\mu}_\mathbf{s}[1]\cdot \mathtt{\tiny KEC}(\mathbf i)\big)\Big)$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}_g', A^+) \equiv \begin{cases}
+\begin{array}{l}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \\ \phantom{\Lambda(}\boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a)\end{array} & \text{if}\begin{array}{l}(\boldsymbol{\sigma}^*[a]=\varnothing\vee\boldsymbol{\sigma}^*[a]_c=())\wedge\\\boldsymbol{\mu}_\mathbf{s}[0]\le\boldsymbol{\sigma}[I_a]_b\wedge I_e<1024\end{array}\\
+(\boldsymbol{\sigma},\boldsymbol{\mu}_g,\varnothing)&\text{otherwise}
+\end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^*\equiv\boldsymbol{\sigma}$ except $\boldsymbol{\sigma}^*[I_a]_n=\boldsymbol{\sigma}[I_a]_n+1$ \\
+&&&& $A'\equiv A\Cup A^+$ (with $\Cup$ defined in the definition of {\small CREATE}) \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x = 0$ if the code execution for this operation failed due to an exceptional \\
+&&&& halting $Z(\boldsymbol{\sigma}^*,\boldsymbol{\mu}, I) = \top$, $E(\boldsymbol{\sigma}, a) = \top$ (the address is already occupied) or \\
+&&&& $I_e = 1024$ (the maximum call depth limit is reached), \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too low to fulfill the balance transfer) or \\
+&&&& $\boldsymbol{\sigma}[a]_c \neq ()$ (a contract already exists); otherwise $x = a$. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& Thus the operand order is: value, salt, input offset, input size. \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-05-10 Paper.tex
Small adjustments
Fix whitespaces

--- 2017-05-10 Paper.tex
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger then 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 \midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule

### 2017-05-10 Paper.tex
Remove spurious spaces

@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 \midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-05-10 Paper.tex
Remove spurious spaces

@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\end{cases}
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
 \boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}

### 2017-05-11 Paper.tex
Add spaces

@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\}\\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise}\\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\

### 2017-05-11 Paper.tex
Add a space

--- 2017-05-10 Paper.tex
@@
 g^{**} & \text{otherwise}
-\end{cases} \\
+\end{cases} \\ \nonumber
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\

### 2017-05-11 Paper.tex
A more uniform notation for the empty set
A partial overwriting is stated with = not with \equiv
In an account state, the storage root comes before the code
This has to be fixed in earlier versions as well
Format
The code is stored as its Keccak hash

--- 2017-05-11 Paper.tex
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\qquad\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\qquad\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \begin{cases}
+\boldsymbol{\sigma}[S(T)]_n &\text{if}\ S(T) = 2^{256} - 1 \\
+\boldsymbol{\sigma}[S(T)]_n + 1 &\text{otherwise} \\
+\end{cases}
 \end{eqnarray}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \vee E(\boldsymbol{\sigma}, a) \\
+g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 g^{**} - c & \text{otherwise} \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \vee E(\boldsymbol{\sigma}, a) \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+E(\boldsymbol{\sigma}, a) &\equiv \big(\boldsymbol{\sigma}[a] \neq \varnothing \wedge (\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}(()) \vee \boldsymbol{\sigma}[a]_n \neq 0)\big)
 \end{align}
@@
 \midrule
+0xfb & {\small CREATE2} & 4 & 1 & Create a new account with associated code on a predictable address. \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[2]..(\boldsymbol{\mu}_\mathbf{s}[2] + \boldsymbol{\mu}_\mathbf{s}[3] - 1)]$ \\
+&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a\cdot \boldsymbol{\mu}_\mathbf{s}[1]\cdot \mathtt{\tiny KEC}(\mathbf i)\big)\Big)$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}_g', A^+) \equiv \begin{cases}
+\begin{array}{l}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \\ \phantom{\Lambda(}\boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a)\end{array} & \text{if}\begin{array}{l}(\boldsymbol{\sigma}^*[a]=\varnothing\vee\boldsymbol{\sigma}^*[a]_c=())\wedge\\\boldsymbol{\mu}_\mathbf{s}[0]\le\boldsymbol{\sigma}[I_a]_b\wedge I_e<1024\end{array}\\
+(\boldsymbol{\sigma},\boldsymbol{\mu}_g,\varnothing)&\text{otherwise}
+\end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^*\equiv\boldsymbol{\sigma}$ except $\boldsymbol{\sigma}^*[I_a]_n=\boldsymbol{\sigma}[I_a]_n+1$ \\
+&&&& $A'\equiv A\Cup A^+$ (with $\Cup$ defined in the definition of {\small CREATE}) \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x = 0$ if the code execution for this operation failed due to an exceptional \\
+&&&& halting $Z(\boldsymbol{\sigma}^*,\boldsymbol{\mu}, I) = \top$, $E(\boldsymbol{\sigma}, a) = \top$ (the address is already occupied) or \\
+&&&& $I_e = 1024$ (the maximum call depth limit is reached), \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too low to fulfill the balance transfer); \\
+&&&& otherwise $x = a$. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& Thus the operand order is: value, salt, input offset, input size. \\
+\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\

### 2017-05-18 Paper.tex
An account state does not contain code, but it is associated with it

@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\qquad\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \begin{cases}
-\boldsymbol{\sigma}[S(T)]_n &\text{if}\ S(T) = 2^{256} - 1 \\
-\boldsymbol{\sigma}[S(T)]_n + 1 &\text{otherwise} \\
-\end{cases}
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
 \end{eqnarray}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \vee E(\boldsymbol{\sigma}, a) \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \vee E(\boldsymbol{\sigma}, a) \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-E(\boldsymbol{\sigma}, a) &\equiv \big(\boldsymbol{\sigma}[a] \neq \varnothing \wedge (\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}(()) \vee \boldsymbol{\sigma}[a]_n \neq 0)\big)
+\end{cases}
 \end{align}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{blockhash}$ & 800 & Payment for {\small BLOCKHASH} operation. \\
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(I_a, \boldsymbol{\sigma}[I_a]_n - 1)\;\big)\Big)\Big)$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$, $E(\boldsymbol{\sigma}^*, a) = \top$ (the address is occupied) or $I_e = 1024$ \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 \midrule
-0xfb & {\small CREATE2} & 4 & 1 & Create a new account with associated code on a predictable address. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[2]..(\boldsymbol{\mu}_\mathbf{s}[2] + \boldsymbol{\mu}_\mathbf{s}[3] - 1)]$ \\
-&&&& $a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\big(I_a\cdot \boldsymbol{\mu}_\mathbf{s}[1]\cdot \mathtt{\tiny KEC}(\mathbf i)\big)\Big)$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}_g', A^+) \equiv \begin{cases}
-\begin{array}{l}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \\ \phantom{\Lambda(}\boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, a)\end{array} & \text{if}\begin{array}{l}(\boldsymbol{\sigma}^*[a]=\varnothing\vee\boldsymbol{\sigma}^*[a]_c=())\wedge\\\boldsymbol{\mu}_\mathbf{s}[0]\le\boldsymbol{\sigma}[I_a]_b\wedge I_e<1024\end{array}\\
-(\boldsymbol{\sigma},\boldsymbol{\mu}_g,\varnothing)&\text{otherwise}
-\end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^*\equiv\boldsymbol{\sigma}$ except $\boldsymbol{\sigma}^*[I_a]_n=\boldsymbol{\sigma}[I_a]_n+1$ \\
-&&&& $A'\equiv A\Cup A^+$ (with $\Cup$ defined in the definition of {\small CREATE}) \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x = 0$ if the code execution for this operation failed due to an exceptional \\
-&&&& halting $Z(\boldsymbol{\sigma}^*,\boldsymbol{\mu}, I) = \top$, $E(\boldsymbol{\sigma}, a) = \top$ (the address is already occupied) or \\
-&&&& $I_e = 1024$ (the maximum call depth limit is reached), \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too low to fulfill the balance transfer); \\
-&&&& otherwise $x = a$. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& Thus the operand order is: value, salt, input offset, input size. \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
@@
 
+\section{Anomalies on the Main Network}
+
+\subsection{Configuration of BLOCKHASH contract}
+
+In the block XXX, before any transaction is executed, the account state under address 240 is modified so that it is associated with the code:
+\begin{verbatim}
+0x73fffffffffffffffffffffffffffffffffffffffe33141561006a576001430360003561010082
+0755610100810715156100455760003561010061010083050761010001555b620100008107151561
+0064576000356101006201000083050761020001555b5061013e565b436000351215156100845760
+0060405260206040f361013d565b61010060003543031315156100a8576101006000350754606052
+60206060f361013c565b6101006000350715156100c55762010000600035430313156100c8565b60
+005b156100ea576101006101006000350507610100015460805260206080f361013b565b62010000
+6000350715156101095763010000006000354303131561010c565b60005b1561012f576101006201
+00006000350507610200015460a052602060a0f361013a565b600060c052602060c0f35b5b5b5b5b
+\end{verbatim}
+This code is used in the {\small BLOCKHASH} instruction.
+
+\subsection{Early behavior of BLOCKHASH instruction}
+\label{sec:blockhash-early-metropolis}
+
+During the first 256 blocks in the Metropolis phase, {\small BLOCKHASH} instruction behaves as specified below.
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\toprule
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
+\bottomrule
+\end{tabular*}
+
 \end{document}

### 2017-05-22 Paper.tex
when processing a block, before processing any transactions execute a call
That sets up the previous block hash in the blockhash contract.
BLOCKHASH checks input
https://github.com/ethereum/EIPs/pull/210#issuecomment-300357695
Add some mappings from addresses to the new precompiled contracts

@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+g_r&=&?\\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 0\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 800 & Payment for {\small BLOCKHASH} operation. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
 
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Configuration of BLOCKHASH contract}
-
-In the block XXX, before any transaction is executed, the account state under address 240 is modified so that it is associated with the code:
-\begin{verbatim}
-0x73fffffffffffffffffffffffffffffffffffffffe33141561006a576001430360003561010082
-0755610100810715156100455760003561010061010083050761010001555b620100008107151561
-0064576000356101006201000083050761020001555b5061013e565b436000351215156100845760
-0060405260206040f361013d565b61010060003543031315156100a8576101006000350754606052
-60206060f361013c565b6101006000350715156100c55762010000600035430313156100c8565b60
-005b156100ea576101006101006000350507610100015460805260206080f361013b565b62010000
-6000350715156101095763010000006000354303131561010c565b60005b1561012f576101006201
-00006000350507610200015460a052602060a0f361013a565b600060c052602060c0f35b5b5b5b5b
-\end{verbatim}
-This code is used in the {\small BLOCKHASH} instruction.
-
-\subsection{Early behavior of BLOCKHASH instruction}
-\label{sec:blockhash-early-metropolis}
-
-During the first 256 blocks in the Metropolis phase, {\small BLOCKHASH} instruction behaves as specified below.
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\toprule
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
-\bottomrule
-\end{tabular*}
-
 \end{document}

### 2017-05-22 Paper.tex
Add some mappings from addresses to the new precompiled contracts

@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\

### 2017-05-24 Paper.tex
Specify gas for the snark verification precompiled contract
Update gas costs according to https://github.com/ethereum/EIPs/commit/7138d54

--- 2017-05-18 Paper.tex
@@
 \mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 256\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 256 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 800 & Payment for {\small BLOCKHASH} operation. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 

### 2017-05-24 Paper.tex
Update gas costs according to https://github.com/ethereum/EIPs/commit/7138d54

--- 2017-05-22 Paper.tex
@@
 \Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
+g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
 \ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
+0 & \text{if}\ \ell_E\le 256\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 256 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{otherwise} \\
 \end{cases} \\
@@
 \end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-g_r&=&?\\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 0\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r&=&?\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r&=&?\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule

### 2017-06-06 Paper.tex
Fill in a corner case
Fix the length constant from bits to bytes
Use the prepared state instead of the initial state

--- 2017-05-24 Paper.tex
@@
 
+We also define a function~$\pi$ that turns the initiation state into the prepared state:
+\begin{eqnarray}
+\pi(\boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}' \\
+(\boldsymbol{\sigma}',\ldots) & \equiv &\Theta(\boldsymbol{\sigma}, 2^{256}-2, 2^{256}-2, 240,\\
+&&\phantom{\Theta(} 240, 1000000, 0, 0, 0, {B_H}_p, \top)
+\end{eqnarray}
+
 And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
@@
 B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\pi(\Gamma(B)), B))
 \end{eqnarray}
@@
 As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's prepared state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \pi(\Gamma(B)) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 256\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 256 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
 \end{eqnarray}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+$G_{blockhash}$ & 800 & Payment for {\small BLOCKHASH} operation. \\
 
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}
+\mathbf{o}'[0..31] & \text{if}\ {I_H}_i \le \boldsymbol{\mu}_\mathbf{s}[0] + 256 < {I_H}_i + 256\ \text{where $+$ is not modulo} 2^{256} \\
+0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $\mathbf{o}'[i] \equiv \begin{cases}
+\mathbf{o}[i] & \text{if}\ 0 \le i < |\mathbf{o}| \\
+0 & \text{otherwise}
+\end{cases}$ \\
+&&&& $(\ldots,\mathbf o) \equiv \Theta(\boldsymbol{\sigma}, I_a, I_o, t, t, 10000000, I_p, 0, 0, \boldsymbol{\mu}_{\mathbf{s}}[0], I_e + 1)$ \\
+&&&& $t \equiv 240$ \\
+&&&& See Section~\ref{sec:blockhash-early-metropolis} for the special behavior of {\small BLOCKHASH} instruction during the first \\
+&&&& 256 blocks in the Metropolis phase. \\
 \midrule

### 2017-06-20 Paper.tex
The one-but-64th calculation should occur after the memory expansion fee is taken out

--- 2017-05-24 Paper.tex
@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{\max(\ell_M,\ell_B)^2\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 256\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 256 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
 \section{Signing Transactions}\label{app:signing}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(g^* - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad g^* \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
+&&&& $g^* \equiv \boldsymbol{\mu}_{g} - (C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i))$ \\
 &&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\

### 2017-06-20 Paper.tex
The one-but-64th calculation should occur after the memory expansion fee is taken out

@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, 0, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\

### 2017-06-21 Paper.tex
RETURNDATACOPY throws when it would overrun the return data buffer
even when the accessed length is zero.

@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, 0, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(g^* - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad g^* \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $g^* \equiv \boldsymbol{\mu}_{g} - (C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i))$ \\
 &&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\

### 2017-06-21 Paper.tex

--- 2017-06-20 Paper.tex
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad

### 2017-06-21 Paper.tex
Make sure that divisions and modulo operations are not subject to mod 2^{256}

--- 2017-06-20 Paper.tex
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& where division is not subject to modulo $2^{256}$. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers and division is not subject to modulo $2^{256}.$ \\
 &&&& Note the overflow semantic when $-2^{255}$ is negated.\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& where $\mod$ is not subject to modulo $2^{256}$. \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
+&&&& Where all values are treated as two's complement signed 256-bit integers and $\mod$ is not subject to modulo $2^{256}$. \\
+t to mod 2^{256}
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, 0, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(g^* - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad g^* \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $g^* \equiv \boldsymbol{\mu}_{g} - (C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i))$ \\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
 G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\

### 2017-06-21 Paper.tex
Make sure that divisions and modulo operations are not subject to mod 2^{256}

--- 2017-06-21 Paper.tex
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& where $\mod$ is not subject to modulo $2^{256}$. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers and $\mod$ is not subject to modulo $2^{256}$. \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule

### 2017-06-21 Paper.tex
An abbreviation abbreviates but does not imply

@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-&&&& where division is not subject to modulo $2^{256}$. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers and division is not subject to modulo $2^{256}.$ \\
+&&&& Where all values are treated as two's complement signed 256-bit integers. \\
 &&&& Note the overflow semantic when $-2^{255}$ is negated.\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
-&&&& where $\mod$ is not subject to modulo $2^{256}$. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers and $\mod$ is not subject to modulo $2^{256}$. \\
+&&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule

### 2017-06-27 Paper.tex

--- 2017-06-21 Paper.tex
@@
 
-\pagecolor{lightyellow}
+%\pagecolor{lightyellow}
 %\pagecolor{lightpink}

### 2017-07-05 Paper.tex
Fix off-by-one.
Should be strictly greater than the limit, the limit itself is acceptable.

@@
 
-%\pagecolor{lightyellow}
+\pagecolor{lightyellow}
 %\pagecolor{lightpink}
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| \gt 24576\big)
 \end{align}

### 2017-07-17 Paper.tex
Use actual latex.
Implement EIP-658 https://github.com/ethereum/EIPs/pull/658
Now transaction receipts contain one byte indicating success or failure.

@@
 
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+
+We assert that the status code $R_{s'}$ is a single byte.
+\begin{equation}
+R_{s'} \in \mathbb{B}_{8}
+\end{equation}
 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A, s') \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| \gt 24576\big)
+\quad s' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
+\end{cases}
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
+s' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
+1 & \text{otherwise}
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\

### 2017-07-24 Paper.tex
Update expmod gas schedule, according to https://github.com/ethereum/EIPs/commit/4d4d8fb

--- 2017-06-27 Paper.tex
@@
 
+The fifth contract performs arbitrary-precision exponentiation under modulo.
+Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
+The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
+The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
+The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
+These three words are followed by $B$, $E$ and $M$.
+The rest of the input is discarded.
+Whenever the input is too short, the missing bytes are considered to be zero.
+The output is encoded big-endian into the same format as $M$'s.
+
+\begin{eqnarray}
+\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+f(x) &\equiv& \begin{cases}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\end{cases}\\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
+8(\ell_E - 32) & \text{otherwise} \\
+\end{cases} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+i[x] &\equiv& \begin{cases}
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+0 &\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}

### 2017-07-24 Paper.tex

--- 2017-07-17 Paper.tex
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_{s'}$ is a single byte.
-\begin{equation}
-R_{s'} \in \mathbb{B}_{8}
-\end{equation}
 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
 \Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s') \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad s' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 \boldsymbol{\sigma}^{**} & \text{otherwise}
-\end{cases} \\
-s' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
 \end{cases} \\

### 2017-07-25 Paper.tex
Fix the pairing checker specification according to https://github.com/ethereum/EIPs/pull/212/commits/470c61bed3b94d67815557aca224641742badb56

@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
 \end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+g_r&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}

### 2017-07-25 Paper.tex
Fix the pairing checker specification according to https://github.com/ethereum/EIPs/pull/212/commits/470c61bed3b94d67815557aca224641742badb56

--- 2017-07-24 Paper.tex
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
 \Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
@@
 \end{cases}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+g_r&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r&=&?\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule

### 2017-07-30 Paper_Chinese.tex
修改build命令以支持中文，翻译标题

--- 2017-07-24 Paper.tex
@@
 \documentclass[9pt,oneside]{amsart}
-%\usepackage{tweaklist}
+\usepackage{xeCJK}
+\setCJKmainfont{Adobe Song Std}
+
 \usepackage{url}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{以太坊：一种安全去中心化的通用交易账本 \\ {\smaller \textbf{EIP-150版本  (\YellowPaperVersionNumber{})}}}
 \author{
-    Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    原文作者: Dr. Gavin Wood  gavin@ethcore.io\\
+    译者: 猿哥 yuange@cuiguangbin.com; 高天露 tianlu.jorden.gao@gmail.com
 }
@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
 \section{Signing Transactions}\label{app:signing}

### 2017-07-30 chinese_build.sh

@@ -1,0 +1,20 @@
+#!/usr/bin/env bash
+
+set -e
+
+if [ -d ".git" ]; then
+
+SHA=`git rev-parse --short --verify HEAD`
+DATE=`git show -s --format="%cd" --date=short HEAD`
+REV="$SHA - $DATE"
+
+else
+
+REV="unknown revision"
+
+fi
+
+echo "\newcommand{\YellowPaperVersionNumber}{$REV}" > Version.tex
+
+xelatex  -interaction=errorstopmode -halt-on-error Paper_Chinese.tex 
+

### 2017-08-04 Paper.tex
Creation on non-empty code or non-zero nonce throws an exception

@@
 \usepackage{tikz}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+%,pagebackref=true
+
+\usepackage{tabu} %requires array.
+
+%This should be the last package before \input{Version.tex}
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
+% "hyperref loads the url package internally.
+Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
+This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
+% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
+Try loading the url package earlier than the package that requires it.
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
+
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\usepackage{seqsplit}
+
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
 
@@
 \definecolor{lightyellow}{rgb}{1,0.98,0.9}
-\definecolor{lightpink}{rgb}{1,0.94,0.95}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Founder, Ethereum \& Parity\\
+    gavin@parity.io
 }
@@
 \pagecolor{lightyellow}
-%\pagecolor{lightpink}
 
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (\hyperlink{GenesisBlock}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
+Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-\subsubsection{Transaction Receipt}
-
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
+\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
 Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$, the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+
+\begin{equation}
+R \equiv (R_z, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+
+We assert that the status code $R_z$ is an integer.
+\begin{equation}
+R_z \in \mathbb{P}
+\end{equation}
 
@@
 
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
 \begin{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 \end{equation}
-
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 
-The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
@@
 \end{equation*}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
-
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{align}
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_i &\equiv \max(H_i - 3000000, 0)
+\end{align}
+
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}
@@
 
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+$H_s$ is the timestamp (in Unix time) of block $H$ at the block's inception, and must fulfil the relation:
 \begin{equation}
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 
-Thus we are able to define the block header validity function $V(H)$:
+\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
 \begin{eqnarray}
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items and $\Upsilon^z$ to evaluate to the status code resulting from the transaction.
+These will be formally defined later.
 
@@
 
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \subsection{Execution}
-We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
@@
 
-%todo Explain g_d reason?
-
 The up-front cost $v_0$ is calculated as:
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+\linkdest{transaction_nonce}{}T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
 g_0 & \leqslant & T_g \quad \wedge \\
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{3}$ and $\Lambda_{3}$ to denote the fact that only the first three components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
 \end{eqnarray}
@@
 
-\section{Contract Creation} \label{ch:create}
+\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
 \begin{eqnarray}
@@
 \end{cases} \\
+\quad z &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
+\end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F \equiv &\big(\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}\big(()\big) \ \vee\ \boldsymbol{\sigma}[a]_n \neq 0 \ \vee\ (\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \\\nonumber  & \vee\ g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\ \nonumber
+z & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+1 & \text{otherwise}
+\end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
@@
 
-As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
+As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
 
@@
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_a$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_o$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_p$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{} $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item\linkdest{Wei_value_exec}{} $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{} $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{} $I_H$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{} $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_w$, the permission to make modifications to the state.
 \end{itemize}
@@
 
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
 \begin{eqnarray}
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
+\hypertarget{Exceptional_Halting_function_Z}{}
 \subsubsection{Exceptional Halting}
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
  \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
@@
 
-\subsubsection{Normal Halting}
+\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}
 
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
 \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\qquad\forall_{\mathbf{U} \in B_{\mathbf{U}}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 
-We define the block reward as 5 Ether:
-\begin{equation}
-\text{Let} \quad R_b = 5 \times 10^{18}
+We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 3 \times 10^{18}
 \end{equation}
@@
 
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+As previously detailed, $\mathbf{R}[n]_z$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\end{equation}
+
+We define $\mathbf{R}[n]_z$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_z =
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
 \end{equation}
@@
 \end{description}
-
+\hypertarget{rlp}{}
 \section{Recursive Length Prefix}\label{app:rlp}
@@
 \end{itemize}
-
+\hypertarget{R b}{}
 Formally, we define $R_b$:
@@
 
-Thus we finish by formally defining $R_l$:
+\hypertarget{R l}{}Thus we finish by formally defining $R_l$:
 \begin{eqnarray}
@@
 All remaining nibbles (now an even number) fit properly into the remaining bytes.
-
+\hypertarget{trie}{}
 \section{Modified Merkle Patricia Tree}\label{app:trie}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{\mathbf{I} \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r&=&?\\
+g_r &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 \Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r&=&?\\
+g_r &=& 40000\\
 \mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
 \mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. 
+\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
 $G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
+\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
+\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+\linkdest{blockhash}{}0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
@@
 \midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
 &&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
@@
 \end{cases}$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+&&&& \linkdest{A r}{}$A'_{r} \equiv A_{r} + \begin{cases}
 R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
@@
 \midrule
-0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
+\midrule
+\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
@@
 \midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
+\midrule
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& For the gas calculation, we use the memory expansion function, \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
@@
 \midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad n \\
@@
 
-%\section{Low-level Lisp-like Language}\label{app:lll}
-%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-
+\hypertarget{GenesisBlock}{}
 \section{Genesis Block}\label{app:genesis}
@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
+\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
 \begin{equation}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
 $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
@@
 
-\subsection{Deletion of an Account Dispite Out-of-gas}
-
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\subsection{Deletion of an Account Despite Out-of-gas}
+
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\toprule
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
+\bottomrule
+\end{tabu*}
+
 \end{document}

### 2017-08-05 Paper_Chinese.tex
update for test
翻译完1. 简介 部

@@
 \author{
-    原文作者: Dr. Gavin Wood  gavin@ethcore.io\\
-    译者: 猿哥 yuange@cuiguangbin.com; 高天露 tianlu.jorden.gao@gmail.com
+    原文作者: Dr. Gavin Wood,  gavin@ethcore.io\\
+    译者: 猿哥, root@cxyym.com, 微信 yuange1024; 高天露, tianlu.jorden.gao@gmail.com
 }
@@
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
-Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
-We can call this paradigm a transactional singleton machine with shared-state.
-
-Ethereum implements this paradigm in a generalised manner.
-Furthermore it provides a plurality of such resources, each with a distinct state and operating code but able to interact through a message-passing framework with others.
-We discuss its design, implementation issues, the opportunities it provides and the future hurdles we envisage.
+区块链对交易数据加密以保证安全,已经通过一系列项目展示了它的实用性,尤其是比特币。每一个这个的项目都可以看作是一个基于去中心化的单实例且拥有计算资源的应用。我们称这种模式为可以共享状态的单例状态机。
+
+以太坊以更广义的方式实现了这种模式。 它提供了大量的资源,每一个资源都拥有独立的状态和操作码,并且可以通过消息传递方式和其它资源交互。我们讨论了它的设计、实现难题、它提供的机会以及以后可能有的一些问题。
 \end{abstract}
@@
 
-\section{Introduction}\label{sec:introduction}
-
-With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap.
-Technology-rooted movements like Bitcoin have demonstrated, through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
-This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
-Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications albeit rather simplistic ones.
-
-Ethereum is a project which attempts to build the generalised technology; technology on which all transaction-based state machine concepts may be built.
-Moreover it aims to provide to the end-developer a tightly integrated end-to-end system for building software on a hitherto unexplored compute paradigm in the mainstream: a trustful object messaging compute framework.
-
-\subsection{Driving Factors} \label{ch:driving}
-
-There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
-This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
-By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
-
-Dealings in this proposed system would have several attributes not often found in the real world.
-The incorruptibility of judgement, often difficult to find, comes naturally from a disinterested algorithmic interpreter.
-Transparency, or being able to see exactly how a state or judgement came about through the transaction log and rules or instructional codes, never happens perfectly in human-based systems since natural language is necessarily vague, information is often lacking, and plain old prejudices are difficult to shake.
-
-Overall, I wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
-
-\subsection{Previous Work} \label{ch:previous}
-
-\cite{buterin2013ethereum} first proposed the kernel of this work in late November, 2013.
-Though now evolved in many ways, the key functionality of a block-chain with a Turing-complete language and an effectively unlimited inter-transaction storage capability remains unchanged.
-
-\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet.
-The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}.
-\cite{back2002hashcash} later produced a system in a similar vein.
-
-The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, ensuring ``consumers'' be able to make micro-payments to ``suppliers'' for their services.
-The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
-Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope.
-The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.
-
-Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol.
-Some of the best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
-Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.
-
-Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
-The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxiliary parts to the core protocol.
-The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.
-
-Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system.
-It has demonstrated that high efficiency gains can be made if the decentralisation premise is discarded.
-
-Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
-Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
-In this light, Ethereum may be seen as a general implementation of such a \textit{crypto-law} system.
+\section{简介}\label{sec:introduction}
+
+随着互联网连接了世界上绝大多数地方，全球信息共享的成本越来越低。比特币网络通过共识机制、自愿遵守的社会合约,实现一个去中心化的价值转移系统且可以在全球范围内自由使用, 这样的技术改革展示了它的巨大力量。这样的系统可以说是加密安全、基于交易的状态机的一种具体应用。后续类似这样的系统,如域名币（Namecoin）, 从最原先的"货币应用"发展到了其它应用,虽然它只是其中很简单的一种应用。
+
+以太坊是一个尝试实现通用性技术的项目,所有基于交易的状态机都可以被构建。而且以太坊致力于为开发者提供主流一个紧凑的、整合的端到端系统, 这个系统提供了一种可信的消息传递计算框架让开发者以一种前所未有的范式来构建软件。
+
+\subsection{驱动因素} \label{ch:driving}
+
+这个项目有很多目标，其中最重要的目标是为了促成不信任对方的个体之间的交易。这些不信任可能是因为地理位置分离、接口对接难度,或者是不兼容、不称职、不情愿、支出、不确定、不方便,或现有法律系统的腐败。于是我们想用一个丰富且清晰的语言去实现一个状态变化的系统，期望协议可以自动被执行, 我们可以为此提供一种实现方式。
+
+这个提议系统中的交易, 有一些在现实世界中并不常见的属性。审判廉洁，在现实世界往往很难找到，但对公正的算法解释器是天然的; 透明，或者说通过交易日志和规则或代码指令能够清晰的看见到状态变化或者判决, 但因为人类语言的模糊性、信息的缺乏以及老的偏见难以撼动,导致基于人的系统中从来没有完美实现透明。
+
+总的来说，我希望能提供一个系统，能够保证用户无论是和其他个体、系统还是组织交互,都能对可能的结果及产生结果的过程完全信任。
+
+\subsection{前人工作} \label{ch:previous}
+
+ \cite{buterin2013ethereum} 在2013年9月下旬第一次提出了这种系统的核心机制。虽然现在发展出了多种方案, 但最关键的部分, 具备图灵完备语言且不受限制的内部交易存贮容量的区块链, 仍未变化。
+
+\cite{dwork92pricingvia} 提出了一种使用计算支出的密码学证明方式(proof-of-work, 工作量证明)在互联网上传递信号值。信号值用作阻挡垃圾邮件，而不是任何一种货币, 但展示了一个基本的数据通道可以承载强大经济信号的可能性,  允许接受者无需依赖信任而做出物理断言。\cite{back2002hashcash} 后来设计了一个类似的系统。
+
+\cite{vishnumurthy03karma:a}最早使用工作量证明作为强大的经济信号保证货币安全。 在这个案例中，代币用作检查点对点（peer-to-peer,p2p）文件交易，同时保证“消费者”能支付给为他们提供服务的“供应商”。这种通过工作量证明的安全模型逐步扩展, 包括使用电子签名和账本技术, 以保证历史记录不被篡改, 怀有恶意的用户不能进行欺诈支付或不公平的抱怨服务。五年后（2008年），中本聪\cite{nakamoto2008bitcoin} 介绍了另一种更广泛的工作量证明安全价值代币。这个项目的成果比特币,成为了第一个被全球广泛认可的去中心化交易账本。
+
+由于比特币的成功，竞争币(alt-coins)开始兴起,通过改变比特币的协议去创建了大量的其他数字货币。比较知名的有莱特币（Litecoin）和素数币（Primecoin）, 参见 \cite{sprankel2013technical} 。一些项目使用比特币的核心机制并重新改造以应用在其它领域,例如域名币（Namecoin）, 致力于提供一个去中心化的名字解析系统, 参见 \cite{aron2012bitcoin} 。
+
+其它在比特币网络之上构建的项目, 也是依赖巨大的系统价值和巨大的算力来保证共识机制。万事达币（Mastercoin）项目, 在比特币协议之上,通过一系列基于核心协议的的辅助插件, 构建一个包含许多高级功能的富协议, 参见 \cite{mastercoin2013willett}。彩色币（Coloured Coins, 参见 \cite{colouredcoins2012rosenfeld}, 采用了类似的但更简化的协议, 以实现比特币基础货币的可替代性, 并允许通过色度钱包（“chroma-wallet”）来创建和跟踪代币。
+
+其它一些工作通过放弃中心化来进行。瑞波币（Ripple）,参见 \cite{boutellier2014pirates}, 试图去创建一个货币兑换的联邦系统（“federated ”system）和一个新的金融清算系统。这个系统展示了放弃去中心化特性可以获得性能上的提升。
+
+\cite{szabo1997formalizing} 和 \cite{miller1997future} 进行了智能合约（smart contract）的早期工作。大约在上世纪90年代，人们逐渐认识到协议算法的执行可以成为人类合作的重要力量。虽然当时没有这样的系统, 但可以预见未来的法律将会受到这种系统的影响。基于此, 以太坊或许可以成为这种密码学-法律系统的通用实现。
 
@@
 
-\section{The Blockchain Paradigm} \label{ch:overview}
+\section{区块链} \label{ch:overview}
 

### 2017-08-07 Paper.tex
modify (136)

--- 2017-07-25 Paper.tex
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 7 \\
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
@@
 
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_r&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r&=&?\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r&=&?\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
-\end{eqnarray}
-
 \section{Signing Transactions}\label{app:signing}

### 2017-08-07 Paper_Chinese.tex
update 08-07

@@
 
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
-It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
-The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
-Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
-A valid state transition is one which comes about through a transaction.
-Formally:
+以太坊在整体上可以看成是一个基于交易的状态机：我们起始于一个创世块（Genesis）状态，然后随着交易的执行状态逐步改变一直到最终状态, 这个最终状态是以太坊世界的权威版本。状态中包含的信息有:账户余额、名誉度、信誉度、现实世界的附属数据等; 简而言之，能包含电脑可以描绘的任何信息。因此，交易是连接两个状态的有效桥梁；“有效”非常重要---因为无效的状态改变远超过有效的状态改变。例如：无效的状态改变可能是减少账号余额，但是没有在其它账户上加上同等的额度。一个有效的状态转换是通过交易进行的,表达式如下:
 \begin{equation}
@@
 
-where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
+$\Upsilon$ 是以太坊状态转换函数。在以太坊中, $\Upsilon$ 和 $\boldsymbol{\sigma}$ 比已有的任何比较系统都强; $\Upsilon$ 可以执行任意计算, 而 $\boldsymbol{\sigma}$  可以存贮交易中的任意状态。
 

### 2017-08-08 Paper_Chinese.tex
update
update 08-08

@@
 
-Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
-They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes place as a state-transition function, adding value to a nominated account.
-
-Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block.
-It is achieved thanks to a cryptographically secure proof.
-This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
-
-Formally, we expand to:
+区块中记录着交易信息；区块之间通过密码学哈希(hash)链接起来。区块链就像一个分类账,将一系列交易记录在一起,并且连接上一个区块及最终状态(并没有直接保存最终状态本身---否则整个区块链就太大了）。系统激励节点去挖矿, 挖矿激励时,有执行状态转移函数,增加挖矿者的账户余额。
+
+挖矿是和其它潜在区块竞争一系列交易(一个区块)的记账权。它是通过密码安全证明的方式来实现的。这个机制称为工作量证明, 会在  \ref{ch:pow} 详细讨论。
+
+公式如下:
 \begin{eqnarray}
@@
 
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
-
-This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
-
-\subsection{Value}
+其中 $\Omega$ 是区块定稿状态转换函数（这个函数奖励一个特定的账户）；$B$ 表示包含一系列交易的区块； $\Pi$ 是区块级的状态转换函数。
+
+上述是区块链的基本内容，这个模型不仅是以太坊的基础架构，还是迄今为止所有基于共识的去中心化交易系统的基础架构。
+
+\subsection{面值}
 

### 2017-08-09 Paper_Chinese.tex
update

@@
 
-上述是区块链的基本内容，这个模型不仅是以太坊的基础架构，还是迄今为止所有基于共识的去中心化交易系统的基础架构。
+上述是区块链的基本内容，这个模型不仅是以太坊的基础，还是迄今为止所有基于共识的去中心化交易系统的基础。
 
@@
 
-In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
-The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
-One Ether is defined as being $10^{18}$ Wei.
-There exist other subdenominations of Ether:
+为了激励网络中的计算，需要定义一种转账方法。以太坊设计了一个内置货币以太币(Ether)，ETH是大家所熟知的符号，有时用 \DH{} 表示。以太币最小的面额是Wei(伟)，所有货币值都以Wei的整数倍记录。一个以太币被定义成位。一个以太币等于 $10^{18}$ Wei 。不同的面值如下表:
+
 \par
@@
 \toprule
-Multiplier & Name \\
-\midrule
-$10^0$ & Wei \\
-$10^{12}$ & Szabo \\
-$10^{15}$ & Finney \\
-$10^{18}$ & Ether \\
+倍数 & 面值 \\
+\midrule
+$10^0$ & Wei(伟)) \\
+$10^{12}$ & Szabo(萨博) \\
+$10^{15}$ & Finney(芬尼) \\
+$10^{18}$ & Ether(以太) \\
 \bottomrule
@@
 
-Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.
+在整个工作中，任何涉及到价值、以太币相关的、货币、余额或者支付，都以Wei作为单位来计算。
 

### 2017-08-09 Paper_Chinese.tex
翻译完 2. The Blockchain Paradigm

@@
 
-总的来说，我希望能提供一个系统，能够保证用户无论是和其他个体、系统还是组织交互,都能对可能的结果及产生结果的过程完全信任。
+总的来说，我们希望能提供一个系统，能够保证用户无论是和其他个体、系统还是组织交互,都能对可能的结果及产生结果的过程完全信任。
 
@@
 
-\subsection{Which History?}
-
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
-If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
-
-This would mean that past a given point in time (block), multiple states of the system may coexist: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
-This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.
-
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
-This process is described in detail in section \ref{ch:ghost}.
-
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
-In order to follow back the history of a path, one might need to reference multiple versions of this document.
-
-\section{Conventions}\label{ch:conventions}
+\subsection{历史?}
+
+因为这是一个去中心化的系统, 所有人都有机会在之前的某一个区块创建新的区块并连接在其后, 这会行成一个树状的区块。为了能在这个树状结构上从根节点（创世块）到叶子节点（包含最新交易的区块）能形成一个一致的区块链,必须有一个共识方案。如果有人认为从根节点到叶子节点的路径不是"最佳"的区块链, 那这时候就会发生分叉。
+
+这个就意味着在一个给定的时间点, 系统中会有多个状态共存：一些节点相信一个区块是包含标准的交易，其他的节点则相信另外一些区块包含标准的交易，其中就包含彻底不同或者不兼容的交易。这一点必须要避免, 因为它会破坏整个系统信用。
+
+我们使用了一个简单GHOST协议版本来达成共识, 参见 \cite{cryptoeprint:2013:881} 。我们会在\ref{ch:ghost} 详细说明。
+
+有时会从一个特定的区块链高度启用新的协议。本文描述了协议的一个版本, 如果要跟踪历史区块链路径, 可能需要查看这份文档的历史版本。(译者注: 可以到 https://github.com/ethereum/yellowpaper 查看英文版历史版本)
+
+\section{约定}\label{ch:conventions}
 

### 2017-08-13 Paper_Chinese.tex
翻译完 4.2 交易

@@
 \midrule
-$10^0$ & Wei(伟)) \\
+$10^0$ & Wei(伟) \\
 $10^{12}$ & Szabo(萨博) \\
@@
 
-I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
-
-The two sets of highly structured, `top-level', state values, are denoted with bold lowercase Greek letters.
-They fall into those of world-state, which are denoted $\boldsymbol{\sigma}$ (or a variant thereupon) and those of machine-state, $\boldsymbol{\mu}$.
-
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
-
-For most functions, an uppercase letter is used, e.g.
-$C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
-Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
-
-Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
-The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
-
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
-For particularly important values, a bold uppercase letter may be used.
-
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
-The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \ref{ch:block}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
-
-In the case of the global state $\boldsymbol{\sigma}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
-
-When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
-
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+我用了大量的印刷约定表示公式中的符号，其中一些需要特别说明：
+
+有两个高度结构化的顶层状态值,使用粗体小写希腊字母: $\boldsymbol{\sigma}$  表示世界状态(world-state);  $\boldsymbol{\mu}$ 表示机器状态（machine-state）。
+
+作用在高度结构化值上的函数，使用大写的希腊字母，例如： $\Upsilon$ , 是以太坊中的状态转换函数。
+
+对于大部分函数来说，通常用一个大写的字母表示，例如：$C$, 表示费用函数。可能会使用下角标表示为一些特别的变量，例如：$C_\text{\tiny SSTORE}$，表示执行 {\tiny SSTORE} 操作的费用函数。对于一些可能是外部定义的函数，可能会使用打印机文字字体，例如：$\texttt{KEC512}$ 哈希函数（为赢得进入SHA-3竞赛） ,使用 $\texttt{KEC}$ 表示。$\texttt{KEC512}$  表示 Keccack-512 哈希函数。
+
+元组通常使用一个大写字母，例如：$T$，表示一个以太坊交易。使用下标可能会表示一个独立的变量,例如： $T_n$ ，表示交易中随机数。角标的形式用于表示它们的类型；例如：大写的下角标表示元组包含的下角标变量。
+
+标量和固定大小的字节序列（或数组）都使用小写字母来表示，例如： $n$ 在本文中表示交易随机数。小写的希腊字母一般表示一些特别的含义，例如：$\delta$ 表示在栈上一个给定操作需要的条目数量。
+
+任意长度的序列通常用加粗的小写字母表示，例如 $\mathbf{o}$ 表示消息调用中输出的数据字节序列。有时候，会对特别重要的值使用粗体。
+
+我们认为标量都是正整数且属于集合 $\mathbb{P}$。所有的字节序列属于集合 $\mathbb{B}$，附录 \ref{app:rlp} 给出了正式的定义。用下角标符号表示这样的序列集合限制在一定长度以内, 长度为32的字节序列使用 $\mathbb{B}_{32}$ 表示, 所有比 $2^{256}$ 小的正整数使用 $\mathbb{P}_{256}$ 表示。详细定义见 \ref{ch:block}。
+
+使用方括号表示序列中的一个元素或子序列,例如： $\boldsymbol{\mu}_\mathbf{s}[0]$ 表示计算机堆栈中的第一个条目。对于子序列来说，使用省略号表示一定的范围，且含头尾的限制，例如：$\boldsymbol{\mu}_\mathbf{m}[0..31]$ 表示计算机内存中的前32个条目。
+
+在全局状态的情况下，σ是一个含多个账号的序列，本身的数组，正方形括号被用作去表示一个单独的账号。
+
+以全局状态 $\boldsymbol{\sigma}$ 为例, 它表示一系列的账户, 它们自身的元组, 方括号用于表示一个独立的账户。
+
+当去考虑现有的变量时，我遵循在给定的范围内去定义的原则，我们使用占位符 $\Box$ 表示未修改的输入变量, 使用 $\Box'$ 表示修改的和可用的变量, $\Box^*$,  $\Box^{**}$ \&c 表示中间变量。在特殊情况下，为了提高可读性和清晰性，我可能会使用字母-数字下角标表示中间值。
+
+当使用去已有的函数时，给定一个函数 $f$,那么函数 $f^*$ 表示一个相似的、替换序列的函数映射。详细的定义见 \ref{ch:block}。
+
+整个过程中，我定义了大量的函数。一个常见的函数是 $\ell$，表示给定序列的最后一个条目：
 
@@
 
-\section{Blocks, State and Transactions} \label{ch:bst}
-
-Having introduced the basic concepts behind Ethereum, we will discuss the meaning of a transaction, a block and the state in more detail.
-
-\subsection{World State} \label{ch:state}
-
-The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
-Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
-This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
-Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly.
-Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.
-
-The account state comprises the following four fields:
-
+\section{块、状态和交易} \label{ch:bst}
+
+介绍了以太坊的基本概念后,我们将更详细地讨论交易、区块和状态的含义
+
+\subsection{世界状态} \label{ch:state}
+
+世界状态是在地址（160位的标志符）和账户状态（序列化为RLP的数据结构，详见附录 \ref{app:rlp}）的映射。虽然世界状态没有直接储存在区块链上，但会假定实施过程中会将这个映射维护在一个修改过的Merkle Patricia 树(简称trie，详见附录 \ref{app:trie}）。trie 需要一个简单的后端数据库去维护字节数组到字节数组的映射；我们称这个后端数据库为状态数据库。它有一系列的好处: 第一这个结构的根节点是加密的且依赖于所有的内部数据，且它的哈希可以作为整个系统状态的一个安全标志;第二，作为一个不变的数据结构，因此它允许任何一个之前状态（根部哈希已知的条件下）通过简单地改变根部哈希值而被召回。因为我们在区块链中储存了所以这样的根部哈希值，所以我们能恢复到指定的历史状态。
+
+账户状态包含以下四个字段：
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
-\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
-\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
-All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+\item[nonce,随机数] 这个值等于账户发出的交易数及这个账户创建的合约数量之和。 $\boldsymbol{\sigma}[a]_n$ 表示状态 $\boldsymbol{\sigma}$ 中的地址 $a$ 的nonce值。
+\item[balance,余额] $\boldsymbol{\sigma}[a]_b$, 表示这个账户拥有多少Wei。
+\item[storageRoot,存储跟节点] 保存账户内容的 Merkle Patricia 树根节点的256位哈希编码到trie中, 作为从 256位整数键值哈希的Keccak 256 位哈希到 256位整数的RLP-编码映射。这个哈希定义为 $\boldsymbol{\sigma}[a]_s$。
+\item[codeHash,代码哈希] 这个账户的EVM(Ethereum Virtual Machine,以太坊虚拟机)代码的哈希值---代码执行时,这个地址会接收一个消息调用; 它和其它字段不同, 创建后不可更改。状态数据库中包含所有像这样的代码片段的哈希, 以便后续使用。 这个哈希定义为 $\boldsymbol{\sigma}[a]_c$,$\mathbf{b}$ 表示代码,  $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
 
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
+因为我通常希望所指的并不是trie树的根哈希,而是所保存的键值对集合，我做了一个更方便的定义:
+
 \begin{equation}
@@
 
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+trie树中的键值对集合函数, $L_I^*$, 定义为基于基础函数 $L_I$ 的元素转换:
 \begin{equation}
@@
 
-where:
+其中:
 \begin{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
-
-If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
+需要说明的是,$\boldsymbol{\sigma}[a]_\mathbf{s}$ 不应算作这个账户的"物理"成员, 它不参与序列化。
+
+如果 \textbf{codeHash} 字段是一个空字符串的 Keccak-256 哈希, 例如 $\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, 则表示对应的节点表示一个简单账户, 有时简称"非合约"账户。
+
+因此我们可能定义一个世界状态的函数  $L_S$：
+
 \begin{equation}
@@
 
-This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+函数 $L_S$ 和 trie 函数是为了提供一个世界状态的简短身份（哈希）。我们假定：
 \begin{equation}
@@
 \end{equation}
-where $v$ is the account validity function:
+
+$v$ 是账户合法性验证函数:
 \begin{equation}
@@
 
-\subsection{The Transaction} \label{ch:transaction}
-
-A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
-\eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
-There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+\subsection{交易} \label{ch:transaction}
+
+交易（符号，$T$）是个单一的加密指令，通过以太坊中系统之外的操作者创建。我们假设外部的操作者是人,软件工具用于创建和传播\footnote{显著地，这样的“工具”可以从基于人类行为的初始化中移除---或者人类可能变成有原因的中立---可能有一点他们被视为自治的代理人。例如：合约可能会给人类好处,让人发送交易从而触发合约的执行。}。这里的交易类型有两种：一种是消息调用,另一种通过代码创建新的账户（称为“合约创建”）。两种类型的交易都有的共同字段如下:
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
-This is expanded in Appendix \ref{app:signing}.
+\item[nonce,随机数] $T_n$,账户发出的交易数量。
+\item[gasPrice,燃料价格] $T_p$, 为执行交易所需要的计算资源付的 \textit{gas} 价格, 以Wei为单位。
+\item[gasLimit,燃料上限] $T_g$, 用于执行交易的最大 gas 数量。 这个值须在交易前设置,且设定后不能再修改。
+\item[to, 接收者地址] 消息调用接收者的160位的地址。对与合约创建交易, 无需接收者地址, 使用 $\varnothing$ 表示, $\varnothing$ 是 $\mathbb{B}_0$ 的唯一成员。
+\item[value, 转账额度] $T_v$, 转到接收者账户的额度,以Wei为单位。对于合约创建,表示捐赠到合约地址的额度。
+\item[v, r, s] $T_w$, $T_r$ and $T_s$,和交易签名相关的变量, 用于确定交易的发送者。详见附录 \ref{app:signing}。
 \end{description}
 
-Additionally, a contract creation transaction contains:
+此外, 合约创建还包含以下字段:
 
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
+\item[init,初始化] $T_\mathbf{i}$, 一个不限制大小的字节数组,表示账户初始化程序的EVM代码。
 \end{description}
 
-\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executes each time the account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is executed only once at account creation and gets discarded immediately thereafter.
-
-In contrast, a message call transaction contains:
+\textbf{init} 是EVM代码片段; 执行 init后会返回另外一个代码片段, 每次合约接受消息调用(通过交易或内部调用)后都会执行这个代码片段。\textbf{init} 仅当合约账户创建的时候执行一次。
+
+相比之下, 一个消息调用的交易包括:
 
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
+\item[data,数据] $T_\mathbf{d}$, 一个不限制大小的字节数组, 表示消息调用的输入数据。
 \end{description}
 
-Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
-For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
+附录 \ref{app:signing} 详细描述了映射发送者交易的函数 $S$, 通过 SECP-256k1 的 ECDSA 曲线, 使用交易(除了最后的3个签名字段)作为数据来签名。目前我们先简单使用 $S(T)$ 表示发送者的指定交易 $T$ 。
 
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+在这里,我们假设所有变量都是 RLP 编码的整数, 除了2个任意长度的字节数组 $T_\mathbf{i}$ 和 $T_\mathbf{d}$。
+
 \begin{equation}
@@
 \end{equation}
-where
+其中
 \begin{equation}
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+地址哈希 $T_\mathbf{t}$ 稍微有些不同:它是一个20字节的地址哈希值, 但创建合约时它是RLP空字节系列, 表示为$\mathbb{B}_0$：
 \begin{equation}
@@
 
-\subsection{The Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
-The block header contains several pieces of information:
-
-%\textit{TODO: Introduce logs}
+\subsection{区块} \label{ch:block}
+
+在以太坊中, 区块是相关信息的集合。区块头 $H$, 与之想对应的交易信息 $\mathbf{T}$, 其它区块头数据的集合 $\mathbf{U}$，$\mathbf{U}$ 表示它的父级区块中有和当前区块的爷爷辈区块是相同的。（这样的区块称为\textit{ommers}\footnote{\textit{ommer} 的意思和自然界中的 "父母的兄弟姐妹"最相近, 详见 \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}, 译者注:不妨称之为叔链)。区块头包含的的信息如下：
 
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[parentHash, 父块哈希] $H_p$,父区块头的 Keccak 256位哈希。
+\item[ommersHash, 叔链哈希] $H_o$,当前区块的叔链列表 Keccak 256位哈希。
+\item[beneficiary, 受益者地址] $H_c$, 成功挖到这个区块的160位地址, 这个区块中的所有交易费用都会转到这个地址。
+\item[stateRoot,状态字典树根节点哈希] 状态字典树根节点的 Keccak 256位哈希, 交易打包到当前区块且区块定稿后可以生成这个值。
+\item[transactionsRoot,交易字典树根节点哈希] 交易字典树根节点的 Keccak 256位哈希,在交易字典树含有区块中的所有交易列表。
+\item[receiptsRoot,接受者字典树根节点哈希] 接受者字典树根节点的 Keccak 256位哈希,在接受者字典树含有区块中的所有交易信息中的接受者。
+\item[logsBloom,日志Bloom] $H_b$, 日记Bloom过滤器由可索引信息（日志地址和日志主题）组成，这个信息包含在每个日志入口, 来自交易列表中的每个交易的接受者。
+\item[difficulty,难度] $H_d$,表示当前区块的难度水平, 这个值根据前一个区块的难度水平和时间戳计算得到。
+\item[number,区块编号] $H_i$,等于当前区块的直系前辈区块数量。 创始区块的区块编号为 0。
+\item[gasLimit,燃料限制] $H_l$,目前每个区块的燃料消耗上限。
+\item[gasUsed, 燃料使用量] $H_g$,当前区块的所有交易使用燃料之和。
+\item[timestamp,时间戳] $H_s$,当前区块初始化时的Unix时间戳。
+\item[extraData,附加数据] $H_x$, 32字节以内的字节数组。
+\item[mixHash,混合哈希] $H_m$,与一个与随机数(nonce)相关的256位哈希计算, 用于证明针对当前区块已经完成了足够的计算。
+\item[nonce,随机数] $H_n$, 一个 64 位哈希, 和计算混合哈希相关, 用于证明针对当前区块已经完成了足够的计算。
 \end{description}
 
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+此外,当前区块还记录着这个区块的交易列表,以及2个叔链(ommer)的区块头列表。我们以 $B$ 表示一个区块:
 \begin{equation}
@@
 
-\subsubsection{Transaction Receipt}
+\subsubsection{交易收据}
 

### 2017-08-14 Paper.tex
fixed a typo in equation (118): (\sigma, \mu', ...) -> (\sigma', \mu', ...)
Update Paper.tex
fix a typo

@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\end{cases}
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
-\end{cases}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
 \end{equation}
@@
 
+\section{Anomalies on the Main Network}
+
+\subsection{Deletion of an Account Despite Out-of-gas}
+
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
+
 \end{document}

### 2017-08-14 Paper_Chinese.tex
翻译完 4.3.1. 交易收据

@@
 
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+为了让交易信息编码能有利于零知识证明、索引、搜索, 我们将每个包含一定信息的交易收据进行编码。以 $B_\mathbf{R}[i]$ 表示第 $i$ 个交易,保存在一个索引字典树中,$H_e$ 是这个字典树的根节点。
+为了去编译信息关于每一个有关系的交易而且可能是一个有用的工具去形成一个零知识证明（zero-knowledge proof），或者索引和搜索。我们编译每一个包含当前交易执行的某些信息为交易收据。每个收据（在第个交易中表示为）是被放置于一个带关键字索引的trie中和这个区块头中被记录下的根值，表示为。
+
+交易收据是一个包含四个条目的元组：交易后的状态,$R_{\boldsymbol{\sigma}}$；当前区块中交易累计燃料使用量, $R_u$, 交易发生后会立即更新这个值;交易执行过程中创建的日志集合,$R_\mathbf{l}$；和日志Bloom过滤器,$R_b$ ：
+
 \begin{equation}
@@
 
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+函数 $L_R$ 是一个将交易收据转换为 RLP 编码的预处理函数:
 \begin{equation}
@@
 \end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+
+交易后状态 $R_{\boldsymbol{\sigma}}$ 会编码到一个字典树中, 字典树的根节点组成了第一个条目。
+
+我们假定累计的燃料使用量 $R_u$ 是一个正整数, 日志Bloom $R_b$ 是2048位(256字节)的哈希:
 \begin{equation}
@@
 
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+$R_\mathbf{l}$ 是一系列的日志入口, 例如 $(O_0, O_1, ...)$。 一个日志入口 $O$ 是一个日志记录器的地址 $O_a$ 的元组, $O_\mathbf{t}$ 是一系列32字节的日志主题, $O_\mathbf{d}$ 是一些字节数据:
 \begin{equation}
@@
 
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
+我们定义 Bloom 过滤器函数 $M$ 将一个日志入口转换为一个 256 字节哈希:
 \begin{equation}
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
-Formally:
+其中 $M_{3:2048}$ 是一个特别的 Bloom 过滤器,针对任意一个字节序列,它舍弃这个字节序列2048位的前三位。它通过对一个字节序列的 Keccak-256 哈希的每一个前三对字节取其的低11位来实现:
 \begin{eqnarray}
@@
 
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
-
-\subsubsection{Holistic Validity}
+其中 $\mathcal{B}$ 是位引用函数, $\mathcal{B}_j(\mathbf{x})$ 等于字节数组 $\mathbf{x}$ 中的索引 j(从0开始索引) 的位。
+
+\subsubsection{整体有效性}
 

### 2017-08-15 Paper_Chinese.tex
update

@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+如果一个区块同时满足以下几个条件，我们才能认为这个区块是有效的：当从起始状态 $\boldsymbol{\sigma}$ (父块的最终状态) 按顺序执行完生成新的状态 $H_r$ 后, 在内部上要保持一致,包括叔链、交易区块哈希、给定的交易 $B_\mathbf{T}$（详细描述见 \ref{ch:finalisation}）:
+
 \begin{equation}
@@
 
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
+其中 $p(k, v)$ 是 RLP 的简单对转换, 在这个例子中, k 为这个区块中的交易索引,v 为交易收据:
 \begin{equation}
@@
 
-Furthermore:
+此外:
 \begin{equation}
@@
 
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
-
-\subsubsection{Serialisation}
+$\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ 是包含以RLP编码的状态 $\boldsymbol{\sigma}$ 键值对的 Merkle Patricia 树根节点哈希, $P(B_H)$ 是父节点。
+
+这些值根据交易计算产生，特别是交易收据 $B_\mathbf{R}$,这个通过交易状态累积函数 $\Pi$ 定义，在 \ref{sec:statenoncevalidation} 会详细说明。
+
+\subsubsection{序列化}
 

### 2017-08-17 Paper_Chinese.tex
翻译完 4.3.4 区块头验证

@@
 
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
+函数 $L_B$ 和 $L_H$ 分别是区块和区块头的准备函数。类似交易收据准备函数 $L_R$，当转换为 RLP 格式时,假设对应的类型、顺序及结构如下:
+
 \begin{eqnarray}
@@
 
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
-
-The component types are defined thus:
+其中 $L_T^*$ 和 $L_H^*$ 是元素序列转换函数,因此:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{对于任何函数} \; f
+\end{equation}
+
+元素类型定义如下:
 \begin{equation}
@@
 
-where
+其中
 \begin{equation}
@@
 
-We now have a rigorous specification for the construction of a formal block structure.
-The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
-
-\subsubsection{Block Header Validity}
-
-We define $P(B_H)$ to be the parent block of $B$, formally:
+我们现在有了一个严密正式的区块结构结构说明。RLP 函数（见附录 \ref{app:rlp}）提供了一个标准方法来把这个结构转换为一个字节序列。
+
+\subsubsection{区块头验证}
+
+我们定义 $P(B_H)$ 为 $B$ 的父区块::
 \begin{equation}
@@
 
-The block number is the parent's block number incremented by one:
+当前区块编号等于它的父块编号加1：
 \begin{equation}
@@
 
-The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+区块难度定义为 $D(H)$：
 \begin{equation}
@@
 \end{equation}
-where:
+其中:
 \begin{equation}
@@
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+区块的燃料限制 $H_l$ 需要满足下面条件:
 \begin{eqnarray}
@@
 
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+$H_s$ 是区块 $H$ 的时间戳,需满足下面条件:
 \begin{equation}
@@
 
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
-
-The nonce, $H_n$, must satisfy the relations:
+这个机制保证了区块间时间平衡；如果最近的两个区块时间间隔短,则会导致难度系数增加,因此需要额外的计算量，大概率会延长下个区块的出块时间。相反，如果最近的2个区块时间间隔时间过长，难度系数和下一个区块的出块预期时间也会减少。
+
+随机数 $H_n$，必须满足下面关系：
 \begin{equation}
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
-
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-Thus we are able to define the block header validity function $V(H)$:
+其中 $H_{\hcancel{n}}$ 是新区块的区块头，但不包含随机数和混合哈希值，$\mathbf{d}$  是当前的大数据集合 DAG(有向无环图)，需要去计算混合哈希,  $\mathtt{PoW}$ 是工作量证明函数（见 \ref{ch:pow}）：第一个元素用于计算混合哈希值，以证明使用了一个正确的 DAG, 第2个元素是伪随机数,依赖于 $H$ 及 $\mathbf{d}$ 。给定一个范围在$[0, 2^{64})$ 的均匀分布，则求解时间和难度$H_d$ 成比例。
+
+这就是区块链安全基础，这也是一个恶意节点不能用其新创建的区块中重写历史数据的重要原因。因为这个随机数必须满足这些条件，且因为条件依赖于这个区块的内容和相关交易, 创建新的合法的区块是困难的、耗时的, 需要超过所有诚实矿工的算力总和。
+
+因此,我们定义这个区块头的验证函数 $V(H)$ 为：
+
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
-
-Noting additionally that \textbf{extraData} must be at most 32 bytes.
-
-\section{Gas and Payment} \label{ch:payment}
+其中 $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+
+此外, \textbf{extraData} 最多 32 字节。
+
+\section{燃料和支付} \label{ch:payment}
 

### 2017-08-28 Paper.tex
fixed a typo
Fix the status code representation

@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+
+We assert that the status code $R_{s'}$ is an integer.
+\begin{equation}
+R_{s'} \in \mathbb{P}
+\end{equation}
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, s')$ where:
 
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
 \boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
 \end{cases} \\ \nonumber
+s' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+1 & \text{otherwise}
+\end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
 &&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
+&&&& Exactly equivalent to {\small CALL} except: \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The last argument of $\Theta$ is $\bot$. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
+&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
+&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 
-\subsection{Deletion of an Account Despite Out-of-gas}
+\subsection{Deletion of an Account Dispite Out-of-gas}
 

### 2017-08-28 Paper.tex

--- 2017-08-14 Paper.tex
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
 \end{eqnarray}
@@
 \boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases} \\
+\quad s' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
 \end{cases}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
+s' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+1 & \text{otherwise}
+\end{cases} \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
 \end{document}

### 2017-09-08 Paper.tex
Block reward to 3 Ether

--- 2017-08-28 Paper.tex
@@
 \begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
-\end{equation}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, s')$ where:
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
 

### 2017-09-08 Paper.tex
Block reward to 3 Ether

--- 2017-08-28 Paper.tex
@@
 where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
-\end{equation}
 

### 2017-09-16 Paper.tex
account state order fixed
fix typo
Footnote low-order 11 bits explanation

@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
-Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits = 2^2048.
+The modulo 2048 of the Keccak-256 hash is the remainder low-order 11 bits.} Formally:
 \begin{eqnarray}

### 2017-09-16 Paper.tex

@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits = 2^2048.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^2048$.
 The modulo 2048 of the Keccak-256 hash is the remainder low-order 11 bits.} Formally:

### 2017-09-16 Paper.tex
Footnote clarifying meaning of low-order 11 bits

--- 2017-09-16 Paper.tex
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits = 2^2048.
-The modulo 2048 of the Keccak-256 hash is the remainder low-order 11 bits.} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}

### 2017-09-16 Paper.tex
Rewording accumulation of nibbles keys and branches

@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
+Formally:
 \begin{eqnarray}

### 2017-09-16 Paper.tex
to proof [sic] that a correct DAG has been used

--- 2017-09-16 Paper.tex
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.

### 2017-09-16 Paper.tex
\usepackage{hyperref}
"Another way is to use special field url and make bibliography style recognise it.

url = "http://www.example.com"

You need to use \usepackage{url} in the first case or \usepackage{hyperref} in the second case."

Source: https://en.wikibooks.org/wiki/LaTeX/Bibliography_Management#Authors

--- 2017-09-16 Paper.tex
@@
 \usepackage{url}
+\usepackage{hyperref}
 \usepackage{cancel}
@@
 The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
-\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
+\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of nibbles keys and branches as traversed from the root.
 The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.

### 2017-09-17 Paper_Chinese.tex
翻译完正文，附录还未翻译

@@
     原文作者: Dr. Gavin Wood,  gavin@ethcore.io\\
-    译者: 猿哥, root@cxyym.com, 微信 yuange1024; 高天露, tianlu.jorden.gao@gmail.com
+    译者: 猿哥, root@cxyym.com; 高天露, tianlu.jorden.gao@gmail.com
 }
@@
 
-In order to avoid issues of network abuse and to sidestep the inevitable questions stemming from Turing completeness, all programmable computation in Ethereum is subject to fees.
-The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation).
-Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing account storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.
-
-Every transaction has a specific amount of gas associated with it: \textbf{gasLimit}.
-This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according \textbf{gasPrice}, also specified in the transaction.
-The transaction is considered invalid if the account balance cannot support such a purchase.
-It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender's account.
-Gas does not exist outside of the execution of a transaction.
-Thus for accounts with trusted code associated, a relatively high gas limit may be set and left alone.
-
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
-Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
-Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
-Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.
-
-%\subsubsection{Determining Computation Costs}
-
-\section{Transaction Execution} \label{ch:transactions}
-
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
-It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
-These include:
-
+为了避免网络滥用及回避由于图灵完整性而带来的一些不可避免的问题，在以太坊中所有的编程计算都需要费用。各种操作费用以 \textit{gas} (详见附录 \ref{app:fees} )为单位计算。任意的程序片段（包括合约创建、信息调回、利用及访问账户存储、在虚拟机上执行操作等）都可以根据规则计算出消耗的燃料。
+
+每一个交易都有一个燃料上限：\textbf{gasLimit} (燃料上限)。这些燃料从发送者的账户中扣除。具体从账户上扣除的额度和 \textbf{gasPrice}(燃料价格) 有关(译者注: 扣除额度 = \textbf{gasLimit} * \textbf{gasPrice}), 在执行交易前会指定燃料价格。如果这个账户不能支付起燃料费用，这个交易会被当作无效交易。之所以它被命名为燃料上限，是因为剩余的燃料在交易完成之后会被退回（以购买时的同样价格）到发送者账户。燃料不会被用在交易执行之外。因此对于可信任账户，应该设置一个相对较高的燃料上限。
+
+通常来说，以太币（Ether）用作去购买燃料, 未退回的那部分转到了区块受益人的地址，通常这个账户的地址是由矿工设定。交易者可以任意设定燃料价格，然而矿工也可以任意地忽略某个交易。在一个交易中，高价格的燃料将消费这个发送者更多的以太币,并转给矿工更多的以太币，因此这个交易会被更多的矿工选择。通常来说，矿工将会选择去通知这是他们执行交易最低燃料价格, 交易者们一般也会些选择一个高过燃料价格下限的价格。因此, 会有一个（加权的）最低燃料可接受价格分布，交易者们需要权衡降低燃料价格和交易快速被矿工打包。
+
+\section{交易执行} \label{ch:transactions}
+
+交易执行是以太坊协议中最复杂的部分：它定义了状态转换函数 $\Upsilon$。所有交易在执行时,都要先通过内部的有效性测试,这些包含：
 \begin{enumerate}
-\item The transaction is well-formed RLP, with no additional trailing bytes;
-\item the transaction signature is valid;
-\item the transaction nonce is valid (equivalent to the sender account's current nonce);
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
+\item 交易是 RLP 格式数据，没有多余的后缀字节；
+\item 交易的签名是有效的；
+\item 交易的随机数是有效的（等于发送者账户的当前随机数）;
+\item 燃料上限不小于实际交易过程中用的燃料 $g_0$;
+\item 发送者账户的余额至少大于费用 $v_0$,需要提前支付。
 \end{enumerate}
 
-Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
+$T$ 表示交易, $\boldsymbol{\sigma}$ 表示状态, 状态转移函数 $\Upsilon$ 如下:
 \begin{equation}
@@
 
-Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
-
-\subsection{Substate}
-Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
-We call this \textit{transaction substate}, and represent it as $A$, which is a tuple:
+$\boldsymbol{\sigma}'$ 是交易后的状态。我们定义 $\Upsilon^g$ 为交易执行所消耗的燃料量, $\Upsilon^\mathbf{l}$ 为交易过程中产生的日志记录, 都会在后文正式定义。
+
+\subsection{子状态}
+
+从交易执行过程来看，伴随交易会产生一些特定的信息, 我们称为交易子状态，用 $A$ 来表示:
 \begin{equation}
@@
 
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
-Though not immediately refunded, it is allowed to partially offset the total execution costs.
-
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+元组内容包含自毁集合 $A_\mathbf{s}$: 一个将在交易完成后被删除的账户集合。 $A_\mathbf{l}$ 是一系列的日志：在代码执行过程中是可归档, 可索引的检查点, 允许以太坊世界的外部旁观者(例如去中心化应用的前端)跟踪合约调用。$A_r$ 表示返回的余额，当使用 {\small SSTORE} 指令将非0的合约存贮空间置为0时,返回余额会增加。虽然不是立即返回余额，但可以部分抵消整个执行费用。
+
+我们定义空的子状态 $A^0$, 它没有自毁、没有日记、返回余额为0：
 \begin{equation}
@@
 
-\subsection{Execution}
-We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\subsection{执行}
+
+执行过程中需要的燃料需要在交易进行之前支付,定义 $g_0$ :
 \begin{align}
@@
 
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
-
-%todo Explain g_d reason?
-
-The up-front cost $v_0$ is calculated as:
+$T_\mathbf{i}$ 合约初始化EVM代码, $T_\mathbf{d}$ 是交易附带的关联数据, 取决于交易是合约创建还是消息调用。如果这个交易是合约创建,那么会增加 $G_\text{txcreate}$ , 其它情况则不增加。$G$ 的详细定义见附录 \ref{app:fees}。
+
+预支付的费用 $v_0$ 计算如下:
 \begin{equation}
@@
 
-The validity is determined as:
+需满足下面关系:
 \begin{equation}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
-The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
-
-We define the checkpoint state $\boldsymbol{\sigma}_0$:
+注意最后的那个条件: 交易燃料限制 $T_g$  <= 这个区块的燃料限制 ${B_H}_l$ - 这个区块之前的所有交易已使用的燃料 $\ell(B_\mathbf{R})_u$  。
+
+有效交易的执行起始于一个对状态不能撤回的改变：发送者账户 $S(T)$ 的随机数会加1, 账户余额扣减部分预付费用 $T_gT_p$。计算过程中有效的燃料 $g$ = $T_g - g_0$。无论是合约创建还是消息调用，得到的最终这状态（可能等于当前的状态），这个改变是确定的且永久无效:从这个观点看不会存在无效的交易。
+
+我们定义检查点状态 $\boldsymbol{\sigma}_0$：
 \begin{eqnarray}
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+从 $\boldsymbol{\sigma}_0$ 转变为 $\boldsymbol{\sigma}_P$ 和交易类型相关。不管它是创建合约还是消息调用,我们定义元组包括执行后的临时状态 $\boldsymbol{\sigma}_P$ , 剩余的燃料 $g'$ 和子状态 $A$ :
 \begin{equation}
@@
 
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+其中 $g$ 是燃料上限扣除已有的交易消耗的燃料：
 \begin{equation}
@@
 \end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
-
-After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
+
+$T_o$ 是原始执行者，与和合约创建或消息调用(直接由交易触发)不同, 可以通过执行 EVM 代码内部调用来执行。
+
+注意，我们使用 $\Theta_{3}$ 表示取函数值的前三个元素;最终结果是消息调用的输出（一个字节数组）, 最终结果并没有在交易的上下文使用。
+
+在消息调回或者创建合约被处理后，再确定退回的燃料后最终状态就确定了。 剩余的燃料 $g'$, 再加上一些补偿, 得到最终需要退回发送者的燃料 $g^*$:
+
 \begin{equation}
@@
 
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
-
-The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+终需要退回发送者的燃料 $g^*$ 等于当前剩余的燃料 $g'$, 再加一个补偿, 这个补偿是 $A_r$ 和总使用燃料量 $T_g - g'$ 的一半中的小者。
+
+燃料对应的以太币给了矿工, 矿工地址是当前区块 $B$ 的受益者地址。我们定义预备最终状态 $\boldsymbol{\sigma}^*$ , 临时状态 $\boldsymbol{\sigma}_P$:
+
 \begin{eqnarray}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
+删除所有出现在自毁集合中的账户后, 达到了最终状态 $\boldsymbol{\sigma}'$:
 \begin{eqnarray}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+最后, 我们定义这次交易中总共使用的燃料 $\Upsilon^g$,  这次交易中创建的日志 $\Upsilon^\mathbf{l}$:
 \begin{eqnarray}
@@
 
-These are used to help define the transaction receipt, discussed later.
-
-\section{Contract Creation} \label{ch:create}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+这些有助于后续讨论的交易收据。
+
+\section{合约创建} \label{ch:create}
+
+当创建一个账户时会有很多参数:发送者（$s$）、原始执行者（$o$）、可用的燃料（$g$）、燃料价格（$p$）、转账额度（$v$）、任意长度的字节数组 $\mathbf{i}$、EVM初始化代码、消息调用/合约创建的当前栈深度($e$) 。
+
+我们定义创建函数为函数 $\Lambda$，相关的变量有状态 $\boldsymbol{\sigma}$ , 包含新状态、剩余燃料及交易子状态 $(\boldsymbol{\sigma}', g', A)$ (详见第6小结) :：
 \begin{equation}
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
-Thus we define the resultant address for the new account $a$:
+这个新账户的地址被定义为包含发送者和随机数 RLP 编码的 Keccak 哈希的最边160位。因此我们定义新帐户 $a$ 的地址:
 \begin{equation}
@@
 
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
-
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
-Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
+其中 $\mathtt{\tiny KEC}$ 是Keccak 256 位哈希函数，$\mathtt{\tiny RLP}$ 是 RLP 编码函数，$\mathcal{B}_{a..b}(X)$ 表示取二进制数据 $X$ 位数为范围 $[a, b]$ 的值,$\boldsymbol{\sigma}[x]$ 是地址 $x$ 的状态，$\varnothing$ 表示地址不存在时的状态。注意,我们使用的是一个比发送者随机数要小的值；我们断言会在在这个合约创建是前会增加发送者账户的随机数，因此在刚开始在交易或VM操作中使用的随机数就是发送者的随机数。
+
+账户的随机数开始被定义为0，余额为传递的转账值，存储空间为空, 哈希代码为Keccak 256位的空字符串哈希值；发送者的余额会减去转账值。这个变化的状态变成 $\boldsymbol{\sigma}^*$:
 \begin{equation}
@@
 
-where $v'$ is the account's pre-existing value, in the event it was previously in existence:
+其中$v'$ 是账户在交易之前就有的余额:
 \begin{equation}
@@
 
-%It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
-
-Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
-Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
+最后，这个账户是通过执行初始化账户的 EVM 代码 $\mathbf{i}$ （执行模型详见小结 \ref{ch:model} ）来初始化的。代码执行可以影响一些事件：可以改变当前账户的存储，能创建更多的账户,执行更多的消息调用。代码执行函数 $\Xi$ 可以得到一个元组,包括结果状态 $\boldsymbol{\sigma}^{**}$, 还可以用的燃料剩余燃料 $g^{**}$, 子状态 $A$, 以及账户 $\mathbf{o}$ 的代码。    赋值到最后的状态的数组中，有效的剩余燃料，当前产生的子状态 $A$ 和账号代码本身 $\mathbf{o}$。
 
@@
 \end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+其中 $I$ 包含执行环境的相关参数, 这些参数在小结 \ref{ch:model} 中有详细定义:
 \begin{eqnarray}
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+如果没有输入数据,使用 $I_\mathbf{d}$ 表示空的元组。$I_H$ 没有什么特别的, 它由区块链决定。
+
+代码执行消耗燃料，且燃料不能低于 0，因此程序执行可能会在自然终止之前会退出。在这个（以及其他几个）异常情况, 我们称发生了燃料不足（out-of-gas, OOG）异常：演进的状态变成空集合 $\varnothing$, 整个创建操作对状态没有影响,状态就像尝试创建合作之前一样。
+
+如果这个初始化代码成功地执行完，那么对应的合约创建的也会被支付。代码保存费用  $c$ 和创建的合约代码大小成正比:
+
 \begin{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
+如果没有足够的燃料支付这些, 例如 $g^{**} < c$, 就会抛出燃料不足异常。
+
+发生这样的异常发生后, 剩余燃料将变为0。 如果合约创建是用于接受一个交易, 它不会影响合约创建内部消耗的燃料,无论如何都必须支付。然而，当燃料不足时, 并不会给这个合约地址转账。
+
+如果没有出现这样的异常,那么剩余的燃料会退会给最原始的发送者, 改变后的新状态也会永久保存。最终状态、燃料和子状态 $(\boldsymbol{\sigma}', g', A)$ 的关系如下:
 \begin{align}
@@
 
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-
-\subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code.
-Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
-For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
-
-\section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
-
-Aside from evaluating to a new state and transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
-This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
+上述 $\boldsymbol{\sigma}'$ 的公式,表明了执行代码初始化得到的最终字节序列 $\mathbf{o}$,就是新创建账户的代码体。
+
+当合约成功创建时, 如果有转账,对应的转账也会转到合约中;如果没有转账, 则仅是合约创建。
+
+\subsection{细微之处}
+有一种情况需要注意, 当初始化代码正在执行时，新创建的地址出现了,但还没有内部的代码时。在这个时间内,任意消息调用不会引起代码执行。如果这个初始化执行结束于一个 {\small SELFDESTRUCT} 指令，这个账号会在交易执行完前将被删除, 这个行为是无意义的。对于 一个正常的 {\small STOP} 指令代码，或者返回的代码是空的, 这时候会出现一个僵尸账户，而且账户中剩余的余额将被永远被锁定在这个僵尸账户中。
+
+\section{消息调用} \label{ch:call}
+当执行消息调用时需要多个参数：发送者（$s$）、交易发起人（$o$）、接受者（$r$）、执行代码的账户（$c$,通常就是接受者）、可用的燃料（$g$）、转账额度（$v$）、燃料价格（$p$）、函数调用的一个任意长度字节的数组 $\mathbf{d}$ 的输入数据以及消息调用/合约创建的当前栈($e$)深度。。
+
+除了转变到新的状态和交易子状态外，消息调用还有一个额外的元素 --- 用字节数组 $\mathbf{o}$ 表示的输出数据。执行交易时输出数据是被忽略的,但在消息调用时, 输出的数据由取决于虚拟机代码执行, 在这种情况下也使用了这些信息。
 \begin{equation}
@@
 \end{equation}
-Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
-
-We define $\boldsymbol{\sigma}_1$, the first transitional state as the original state but with the value transferred from sender to recipient:
+
+注意我们的是,当执行 {\small DELEGATECALL}  指令时,我们需要区别转账额度 $v$ 和执行上下文中的 $\tilde{v}$ 。
+
+我们定义，在原始状态基础上进行了发送者向接收者转账后为的状态第一个转变状态 $\boldsymbol{\sigma}_1$:
+
 \begin{equation}
@@
 
-Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and nonce.
-Thus the previous equation should be taken to mean:
+我们假设如果 $\boldsymbol{\sigma}_1[r]$ 还未定义, 则会创建一个没有代码或没有状态,且余额和随机数都为0的账户。我们改进上一个方程式:
 \begin{equation}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
-due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
-$\boldsymbol{\sigma}$).
+账户关联的代码（整个代码碎片, Keccak 哈希为  $\boldsymbol{\sigma}[c]_c$）依靠执行模式（见小结 \ref{ch:model}）执行。合约创建时，如果执行因为一个异常（例如：耗尽了燃料、堆栈溢出、无效的跳转目的地或者无效的指令）而被终止，燃料不会返回到调用者，且状态也会立即恢复到转账之前的状态(例如 $\boldsymbol{\sigma}$)。
 
@@
 
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
-
-As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
-
-Their full formal definition is in Appendix \ref{app:precompiled}.
-
-\section{Execution Model} \label{ch:model}
-
-The execution model specifies how the system state is altered given a series of bytecode instructions and a small tuple of environmental data.
-This is specified through a formal model of a virtual state machine, known as the Ethereum Virtual Machine (EVM).
-It is a \textit{quasi-}Turing-complete machine; the \textit{quasi} qualification comes from the fact that the computation is intrinsically bounded through a parameter, \textit{gas}, which limits the total amount of computation done.
-
-\subsection{Basics}
-
-The EVM is a simple stack-based architecture.
-The word size of the machine (and thus size of stack item) is 256-bit.
-This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations.
-The memory model is a simple word-addressed byte array.
-The stack has a maximum size of $1024$.
-The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
-Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state.
-All locations in both storage and memory are well-defined initially as zero.
-
-The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only through a specialised instruction.
-
-The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas exception, they do not leave state changes intact.
-Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
-
-\subsection{Fees Overview}
-
-Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
-Finally, gas may be paid due to an increase in the usage of the memory.
-
-Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
-This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
-Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds.
-That said, implementations must be able to manage this eventuality.
-
-Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is not only waived, a qualified refund is given; in fact, this refund is effectively paid up-front since the initial usage of a storage location costs substantially more than normal usage.
-
-See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
-
-\subsection{Execution Environment}
-
-In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
-
+我们假设客户端会先保存数据对 $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ , 以便更方便地根据 $I_\mathbf{b}$ 做出决定。
+
+如我们所见，消息调用执行框架 $\Xi$ 中有4个特例: 这四个是‘预编译’合约，作为最初架构中的一部分后续可能会变成本地扩展。这四个合约地址分别为1、2、3和4，分别是用来执行椭圆曲线公钥恢复函数、SHA2 256位哈希方案、RIPEMD 160位哈希方案和身份函数。
+
+这4个合约的详细定义见附录 \ref{app:precompiled}。
+
+\section{执行模型} \label{ch:model}
+执行模型具体说明怎么使用一系列字节代码指令和一个小的环境数据元组去改变这个系统状态。这些是通过以太坊虚拟机(Ethereum Virtual Machine - EVM)这个虚拟的状态来实现的。它是一个准图灵机,说"准"是因为计算会被燃料所限制。
+
+\subsection{基础}
+
+EVM 基于栈结构,机器的字大小（以及栈中数据的大小）是256位。主要是便于执行 Keccak-256 位哈希及椭圆曲线计算。内存模型基于字寻址的字节数据。栈的最大深度为1024。EVM也有一个独立的存储模型；类似内存但更新一个字节数组, 一个基于字寻址的字数组。不像易变的内存，存储是非易变的且是作为系统状态的一部分被维护。所有在内存和存储中的数据会初始化为0。
+
+EVM 不是标准的诺依曼结构。它通过一个特别的指令把程序代码保存在一个虚拟的可以交互的 ROM 中, 而不是保存在一般性可访问的内存或存储中。
+
+EVM 在某些情况下会有异常发生，包括堆栈溢出和非法指令。就像缺乏燃料异常那样, 不会改变状态。有异常时, EVM 立即停止并告知执行代理(交易的处理者，或执行环境), 代理会单独处理异常。
+
+\subsection{费用概述}
+
+在三个不同的情况下使用费用（命名为燃料），在这三种情况下，费用都是执行任何操作的必备条件。第一种情况也是最普通的情况就是计算操作费用（详见附录 \ref{app:fees}）。第二种情况，燃料可能因为一个子消息调用或者合约创建而被消耗,这是执行 {\small CREATE}, {\small CALL} and {\small CALLCODE} 费用中的一部分。第三种情况，燃料可能因为内存使用的增多而付费。
+
+对于一个账户的执行，内存总费用和最少需要的 32 字节倍数的内存量成正比, 所有的内存是按在一定范围中被包含的所有记忆索引（无论是读还是写）下要求的32位字节的最小倍数成比例的。这个为了 just-in-time (JIT) 的基础而去支付的；访问一个大于索引量 32 字节的地址的,就会需要额外的内存使用费。地址很容易超过 32 字节的限制, 但 EVM 不同。综上所述，必须能够去管理这些可能发生的事件。
+
+存储费用有一个微妙的行为 —— 为最小化存储费用（直接与一个在所有结点上的大型状态数据通信），清除存储的执行费用指令不仅仅免除,而且会返回一些费用；因为初始化的存储费用往往比实际实际使用的多,在清除存储空间后会有返回费用, 这个返回的费用是预先支付的费用中的一部分。
+
+详细的 EVM 燃料消耗定义见附录 \ref{app:vm}。
+
+\subsection{执行环境}
+
+除了系统状态$\boldsymbol{\sigma}$ 和 计算过程中剩余的燃料 $g$ 外, 还有一些在计算过程中需要提供的信息, 这些信息组成了元组 $I$:
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_a$, 拥有正在执行的代码的账户地址。
+\item $I_o$, 发起这次交易的发送者地址。
+\item $I_p$, 发起这次交易中的燃料价格。
+\item $I_\mathbf{d}$, 执行过程中的输入字节数组；如果这次执行是一个交易，这个就是交易数据。
+\item $I_s$, 触发代码执行的账户地址；如果这次执行是一个交易，则为交易发送者地址。
+\item $I_v$, 以 Wei 为单位的值,作为计算中的一部分传递给这个账户;如果这次执行是一个交易, 这个值为转账额度。
+\item $I_\mathbf{b}$, 机器代码字节数组, 会用来执行。
+\item $I_H$, 当前区块的区块头。
+\item $I_e$, 当前消息调用或合约创建的深度(例如: 当前 {\small CALL}s 或 {\small CREATE}s 被执行的次数)。
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
+执行模型定义了函数 $\Xi$, 用来计算结果状态 $\boldsymbol{\sigma}'$, 剩余的燃料 $g'$, 产生的子状态 $A$ 和结果输出 $\mathbf{o}$, 根据当前的上下文,我们定义:
 \begin{equation}
@@
 
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
-
+应该还记得 $A$, 这个产生的子状态定义为自毁集合 $\mathbf{s}$、 日志系列 $\mathbf{l}$ 和 回推金额 $r$ 的元组:
 \begin{equation}
@@
 
-\subsection{Execution Overview}
-
-We must now define the $\Xi$ function.
-In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$.
-Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\subsection{执行概述}
+
+我们现在必须去定义函数 $\Xi$。在大多实际执行过程中, 将会整个系统状态 $\boldsymbol{\sigma}$ 和机器状态 $\boldsymbol{\mu}$ 的迭代过程建模。我们定义一个递归函数 $X$, 它使用这个迭代函数 $O$(定义状态机中单循环的结果), 定义函数 $Z$ 用来表示当前状态是一个异常终止状态, 定义函数 $H$ 表示当前状态是正常终止时得到的结果数据。
+
+空序列使用 $()$ 表示,它不等于空的集合 $\varnothing$；这对理解 $H$ 的输出结果非常重要, 当 $H$ 的输出结果是 $\varnothing$ 时需要继续执行,但当 $H$ 的输出结果是时一个空的序列时执行应该终止。
+
 \begin{eqnarray}
@@
 
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
-
-\subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
-
-For the ease of reading, the instruction mnemonics, written in small-caps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
-
-For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
+需要注意的是, 我们在推导 $\Xi$ 时, 我们去掉了第4个元素 $I'$ 并从结果状态 $\boldsymbol{\mu}'$ 中提取了剩余的燃料 $\boldsymbol{\mu}'_g$
+
+$X$ 被循环调用（这里是递归，但是通常是去执行一个简单的迭代循环）直到 $Z$  变成 true,表示当前状态有异常并且需要终止, 并且所有的改动都会被舍弃;  或者直到 $H$ 变成了一个序列（不是空集合），表示达到了正常的终止状态。
+
+\subsubsection{机器状态}
+机器状态 $\boldsymbol{\mu}$ 定义为一个元组 $(g, pc, \mathbf{m}, i, \mathbf{s})$ ,包括可用的燃料 $g$, 程序计数器 $pc \in \mathbb{P}_{256}$ ,内存内容 $\mathbf{m}$ , 内存中激活的字数(从位置 0 开始连续计数) $i$,以及栈内容 $\mathbf{s}$。内存内容 $\boldsymbol{\mu}_\mathbf{m}$ 表示空间大小为 $2^{256}$ 但内容都是 0 的空间。
+
+为了提高可读性, 使用小型大写字母(例如 {\small ADD})的指令助记符都应该看作对应的数字方程式, 详细的指标表及定义见附录 \ref{app:vm} 。
+
+为了更好的定义 $Z$, $H$ 和 $O$, 我们定义 $w$ 为当前待执行的指令:
 \begin{equation}\label{eq:currentoperation}
@@
 
-We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
-\subsubsection{Exceptional Halting}
-
-The exceptional halting function $Z$ is defined as:
+我们使用 $\mathbf{\delta}$ 表示栈操作移除的元素数, 使用 $\mathbf{\alpha}$ 表示栈操作添加的元素数, 可以使用指令作为  $\mathbf{\delta}$ 、 $\mathbf{\alpha}$ 和指令费用函数 $C$ 的小标, $C$ 表示执行指定指令消耗的燃料。
+
+\subsubsection{异常终止}
+
+异常终止函数 $Z$ 定义如下:
 \begin{equation}
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
-The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-
-\subsubsection{Jump Destination Validity}
-
-We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run.
-We define this as any position in the code occupied by a {\small JUMPDEST} instruction.
-
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined {\small STOP} operations that trail it).
-
-Formally:
+以下情况都会发生异常而进入终止状态: 燃料不足、无效指令（无效指令的 $\delta$ 下角标不会进行定义）、缺少栈数据、指令 {\small JUMP}/{\small JUMPI} 的目标地址无效、新的栈大小会大于1024。聪明的读者将很容易发现,在整个执行过程中没有一个指令能显式触发异常而终止。
+
+\subsubsection{跳转地址验证}
+
+我们之前使用 $D$ 作为一个函数去定义正在运行的指定代码的有效跳转地址的集合。我们使用 JUMPDEST 指令表示代码可以跳转的地址。
+
+这些跳转地址都必须在有效指令边界内，并且需要在显示定义在代码中,而不是基于 {\small PUSH} 指令操作的数据。（也不是使用隐含定义的 {\small STOP} 指令去跟踪它）。
+
+正式定义如下:
 \begin{equation}
@@
 
-where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
+$N$ 是代码中下一个有效指令位置, 如果有 {\small PUSH} 指令,会跳过相关数据。
+
 \begin{equation}
@@
 
-\subsubsection{Normal Halting}
-
-The normal halting function $H$ is defined:
+\subsubsection{正常停止}
+
+正常停止的函数 $H$ 定义如下:
 \begin{equation}
@@
 
-The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
-
-\subsection{The Execution Cycle}
-
-Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
+数据返回停止操作 \text{\small RETURN}, 在附录 \ref{app:vm} 中有详细的定义,函数名字是 $H_{\text{\tiny RETURN}}$ 。
+
+\subsection{执行周期}
+序列的最左侧的、低端的数据被添加到栈或从栈中移除；所有其它数据都被保留不变:
+
 \begin{eqnarray}
@@
 
-The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
+对与大多数指令,燃料随着指令的执行而减少,每个周期程序计数器都会增加; 有三个特例，我们定义函数  $J$，下角标是二个指令中的一个，如下：
+
 \begin{eqnarray}
@@
 
-In general, we assume the memory, self-destruct set and system state don't change:
+在一般情况下，我们认为内存、自毁集合和系统状态不改变:
 \begin{eqnarray}
@@
 
-However, instructions do typically alter one or several components of these values.
-Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and a formal description of the gas requirements.
-
-\section{Blocktree to Blockchain} \label{ch:ghost}
-
-The canonical blockchain is a path from root to leaf through the entire block tree.
-In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
-The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
-This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
-
-Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.
-
-Thus we define the total difficulty of block $B$ recursively as:
+然而，指令通常会改变一个到几个元素的值。附录 \ref{app:vm} 列出了指令修改的元素,以及入栈数 $\alpha$ 、出栈数 $\delta$ 和燃料消耗。
+
+\section{区块树到区块链} \label{ch:ghost}
+
+权威的区块链是覆盖整个区块树的从根节点到叶子节点的路径。为了形成是路径共识，我们通过最大的计算量，或者说是最重的路径来识别它。识别最重路径的一个有效的因素是叶子节点的数量，它也是路径中区块的数量，不包含没有进行挖框的创世块(译者注:创始块通过硬编码写在了代码中及区块链中)。路径越长，到达最后一个叶子节点的代价越大。这和已有的比特币及其衍生协议类似。
+
+因为一个区块头包含难度信息，所以仅通过区块头就能验证已经做了挖矿计算。区块链中的任意一个区块都对总的计算或区块链难度有贡献。
+
+我们递归地定义区块 $B$ 的难度：
+
 \begin{eqnarray}
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
-
-\section{Block Finalisation} \label{ch:finalisation}
-
+给定区块 $B$, $B_t$ 是它的总难度, $B'$ 是它的父块难度, $B_d$ 是它的难度。
+
+\section{区块定稿} \label{ch:finalisation}
+
+区块定稿涉及到 4 个步骤:
 The process of finalising a block involves four stages:
@@
 \begin{enumerate}
-\item Validate (or, if mining, determine) ommers;
-\item validate (or, if mining, determine) transactions;
-\item apply rewards;
-\item verify (or, if mining, compute a valid) state and nonce.
+\item 验证叔链(如果是挖矿,确定叔链);
+\item 验证交易(如果是挖矿,确定交易);
+\item 应用奖励;
+\item 验证状态和随机数(如果是挖矿,计算有效的状态和随机数)。
 \end{enumerate}
 
-\subsection{Ommer Validation}
-
-The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
-The maximum of ommer headers is two.
-Formally:
+\subsection{叔链验证}
+
+验证叔链头意味着需要验证每个叔链头的有效性并且是当前区块父块的第 $N$ 代, $N$ ≤ 6, 最多只能有2个叔链头, 公式如下:
 \begin{equation}
@@
 
-where $k$ denotes the ``is-kin'' property:
+$k$ 表示直系关系:
 \begin{equation}
@@
 
-and $s$ denotes the ``is-sibling'' property:
+$s$ 表示兄弟关系:
 \begin{equation}
@@
 \end{equation}
-where $B(H)$ is the block of the corresponding header $H$.
-
-\subsection{Transaction Validation}
-
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+$B(H)$ 表示对应的区块头 $H$.
+
+\subsection{交易验证}
+
+使用的燃料数量一定要和交易列表中列出的一致：区块燃料总使用量 ${B_H}_g$ ，应该和这个区块中所有交易使用的燃料量之和相等：
+
 \begin{equation}
@@
 
-\subsection{Reward Application}
-
-The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+\subsection{奖励}
+对一个区块, 会奖励当前区块的受益账户及叔链账户。以 $R_b$ 表示当前区块受益者的固定奖励额度;对每个叔链, 奖励当前区块受益者固定奖励的 $\frac{1}{32}$, 叔链受益者的奖励额度取决于区块编号。 我们定义函数 $\Omega$ :
 \begin{eqnarray}
@@
 
-If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
-
-We define the block reward as 5 Ether:
+如果叔链和区块的受益者账号冲突（比如：两个相同的叔链受益者账户或者一个叔链受益者账户和当前区块的受益者账户相同），额外的奖励会累积。
+
+我们定义区块的奖励会 5 以太币:
 \begin{equation}
@@
 
-\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
-
+\subsection{状态和随机数验证}\label{sec:statenoncevalidation}
+
+我们定义函数 $\Gamma$, 映射到区块 $B$ 的初始状态:
 We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
@@
 
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+其中，$\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ 表示状态 $\boldsymbol{\sigma}_i$ 对应字典树根节点的哈希;执行时会将其保存在状态数据库中,字典树结构天生不易变,所以它是高效的。
+
+最后定义区块转换函数 $\Phi$，映射了一个不完整区块 $B$ 到一个完整的区块 $B'$：
 \begin{eqnarray}
@@
 \end{eqnarray}
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
-
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+其中 $\mathbf{d}$ 是一个数据集, 详见附录 \ref{app:ethash}。
+
+就像在本文开始定义的那样,$\Pi$ 是状态转换函数，通过区块定稿函数 $\Omega$ 和交易演变函数 $\Upsilon$ 去定义 $\Pi$。
+
+之前的定义中，$\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ 和 $\mathbf{R}[n]_u$，表示第 n 个对应的状态、日志和累计使用的燃料（在元组中的第四个元素 $\mathbf{R}[n]_b$，之前以日志条目去定义）。前者简单定义为通过对之前的交易状态执行交易得到新的结果状态（如果是第一个交易，之前的交易状态为创始块的状态）:
+
 \begin{equation}
@@
 
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+对于 $B_\mathbf{R}[n]_u$, 我们定义从之前的交易转变对应交易的累计使用燃料（如果是第一个,则为 0 ）：
+
 \begin{equation}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+对于 $\mathbf{R}[n]_\mathbf{l}$, 我们使用之前定义的交易执行函数 $\Upsilon^\mathbf{l}$。
 \begin{equation}
@@
 
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+最后,我们定义函数 $\Pi$ 是在交易结果状态 $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$ 之上执行了回报函数  $\Omega$ 后的新状态::
 \begin{equation}
@@
 
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
-
-\subsection{Mining Proof-of-Work} \label{ch:pow}
-
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
-However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
-
-For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of, or reward from, specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
-
-Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
-Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
-
-One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task.
-In Bitcoin's case the task is the SHA256 hash function.
-While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
-Because of this, a proof-of-work function that is ASIC-resistant (i.e.
-difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
-
-Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
-
-More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
+到此,我们完成了区块转换机制以及工作量证明 $\mathtt{PoW}$ 的定义。
+
+\subsection{挖矿和工作量证明} \label{ch:pow}
+
+工作量证明（PoW）通过加密安全随机数,来保证要获得要确定满足一定条件的 $n$,要花费一定量的计算资源。通过计算给定意义的数值和难度系数来增加区块链安全性。然而因为挖出一个新的区块会附带一些奖励，工作量证明不仅是保证区块链权威的安全函数，而且也是一个健康的分配机制。
+
+处于以上原因，工作量证明函数有两个重要的目标：首先，它应该是尽可能的被更多人去接受。对定制的特别硬件的需求或者回报，应该被减到最小。 这使得分配模型尽可能开放，理想情况是，通过挖矿消耗电力获得以太币, 在全世界各个地方都是一样的比例。
+
+第二，应该不允许获得超线性的收益，尤其是在有一个非常高的初始障碍条件下。如果允许这样, 一个资金充足的恶意者可以获得引起麻烦的网络挖矿算力, 并允许给他们获得超线性的汇报（按他们的医院改变收益分布）,并且会弱化网络的安全性。
+
+比特币的世界中一个灾难是 ASICs。有一些计算硬件仅仅是为了做一个简单的任务而存在。在比特币的案例中，这个任务就是 SHA256 哈希函数。当 ASICs 为了工作量证明函数而存在时，两个的目标会变得危险。因此，一个可抵抗 ASIC 的工作量证明函数（比如难以在专用硬件上执行,或者在专用硬件执行时并不划算）可以作为众所周知的银弹。
+
+防止 ASIC 漏洞的两个方向：第一是去让它变成有序列的内存困难，比如：设计一个函数,确定随机数需要大量的内存和带宽,以至于这些内存不能被并行地去计算随机数。第二个方向是让计算变得更普遍化; 对于这个普遍化的计算,使得特殊硬件和普通的桌面计算机计算起来都差不多。在以太坊 1.0中，我们选择了第一个方向。
+
+以太坊的工作量证明函数如下:
+
 \begin{equation}
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
+其中 $H_{\hcancel{n}}$ 是没有包含随机数和混合哈希元素的新区块头; $H_n$ 是区块头的随机数；$\mathbf{d}$ 是一个需要计算混合哈希值的数据集合;$H_d$ 是新区块的难度值（区块难度详见小结 \ref{ch:ghost} ）。$\mathtt{PoW}$ 是工作量证明函数，这个函数根据数组做计算, 这个数组中的第一个元素是混合哈希,第二个元素是依赖于 $H$ 和 $\mathbf{d}$ 的加密伪随机数。这个算法称为 Ethash , 下文会详细介绍。
+
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
-
-\section{Implementing Contracts}
-
-There are several patterns of contracts engineering that allow particular useful behaviours; two of these that I will briefly discuss are data feeds and random numbers.
-
-\subsection{Data Feeds}
-A data feed contract is one which provides a single service: it gives access to information from the external world within Ethereum.
-The accuracy and timeliness of this information is not guaranteed and it is the task of a secondary contract author---the contract that utilises the data feed---to determine how much trust can be placed in any single data feed.
-
-The general pattern involves a single contract within Ethereum which, when given a message call, replies with some timely information concerning an external phenomenon.
-An example might be the local temperature of New York City.
-This would be implemented as a contract that returned that value of some known point in storage.
-Of course this point in storage must be maintained with the correct such temperature, and thus the second part of the pattern would be for an external server to run an Ethereum node, and immediately on discovery of a new block, creates a new valid transaction, sent to the contract, updating said value in storage.
-The contract's code would accept such updates only from the identity contained on said server.
-
-\subsection{Random Numbers}
-Providing random numbers within a deterministic system is, naturally, an impossible task.
-However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
-For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
-
-\section{Future Directions} \label{ch:future}
-
-The state database won't be forced to maintain all past state trie structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
-
-Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
-This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
-
-Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
-
-\subsection{Scalability}
-
-Scalability remains an eternal concern.
-With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy.
-Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger.
-However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.
-
-Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidated each block and with duplicate or invalid transactions thrown out accordingly.
-
-Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
-
-\section{Conclusion} \label{ch:conclusion}
-
-I have introduced, discussed and formally defined the protocol of Ethereum.
-Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
-Contracts may be authored in order to algorithmically specify and autonomously enforce rules of interaction.
-
-\section{Acknowledgements}
-
-Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
+
+Ethash 是以太坊 1.0 的 PoW 算法。它是 Dagger-Hashimoto 的最后版本，Buterin
+ [2013b] 和 Dryja[2014] 介绍过。再被它为 Dagger-Hashimoto 已不大合适, 因为很多算法的最初特性已经被大量地修改。这个算法的概况如下：
+
+通过扫描区块头可以计算一个种子。根据种子，可以计算一个初始大小为 $J_{cacheinit}$ 字节位随机数缓存。轻节点客户端可以保存这个缓存。根据缓冲，我们能生成一个初始化为 $J_{datasetinit}$ 字节的数据集合，数据集合中每个元素依赖于缓冲中的一小部分数据。全节点客户和矿工保存着数据集合。数据集合随时间线性增长。
+
+挖矿涉及到组合数据集合的随机碎片并将他们哈希在一起。使用少量内存就可以验证, 可以借助缓存重新产生需要的数据片段, 所以仅需要保存缓存即可。大的数据集合每$J_{epoch}$ (译者注: 30000) 个区块更新一次，所以大多数矿工的工作都是读取这些数据集合，而不是尝试改变它。附录 \ref{app:ethash} 想 Ethash 的详细解释。
+
+\section{执行合约}
+有一些特别有用的合约模式；其中会讨论其中两个,分别是数据订阅和随机数。
+
+\subsection{数据订阅}
+一个数据订阅合约提供简单的服务：它允许外部的信息可以进入以太坊系统内。以太坊系统不会保证这个信息的精确度和及时性，这是第二方合约作者的任务, 第二方合约作者使用数据订阅并决定能否信任单个数据订阅服务。
+
+通常的模式是,通过消息调用触发单个以太坊中合约，得到一个关于外部现象的信息回应。一个例子可能是纽约城市的本地温度。这个会作为合约执行并返回在存储中的一些值。当然存储中的这些值必须比如温度需要维护以保证准确, 这样模式中的第二方才可以将其作为外部服务来运行以太坊节点,当发现一个新区块时立即创建一个新的有效交易发送到合约并更新存贮中的值。合约代码将仅仅接受包含指定服务的数据更新。
+
+\subsection{随机数}
+在确定系统中提供随机数字,显然是一个不可能实现的任务。然而我们可以利用根据交易时还不可知的数据来生成伪随机数。比如区块哈希值、区块时间戳、区块受益人的地址。为了避免被矿工难以控制这些随机数,建议使用 {\small BLOCKHASH} 操作获得最近 256 个区块的某个区块哈希作为伪随机数。 对于获得这样一系列这样的数字，一个方法就是去增加一个常量并对结果做哈希。
+
+\section{未来方向} \label{ch:future}
+未来状态数据库将不被强行维护所有之前的字典树状态。它应该对每个节点维护一定时间,并抛弃较久远的且不是检查点的节点；检查点或允许特别的区块状态字典树穿越的数据中的系列节点,可以用来最大限度的去替代一些计算, 来获得区块链中的任意一个状态。
+
+区块链合并可以用来减少作为全节点或挖矿节点客户端需要下载的区块数量。及时将给定节点（也许每10000个区块）对应的的字典树压缩, 并在点对点网络中维护，且有效地重塑创世块。下载单个归档时,可以减少下载的区块数量。
+
+最后，或许会引导压缩区块链：可以抛弃在一定量的区块总没有发送或接受交易的字典树状态节点,以便减少状态数据库的增长。
+
+\subsection{可扩展性}
+可扩展性仍然是一个永恒的顾虑。因为一个一般化的状态交易函数，使得切分及并行化交易难以使用分而治之的策略。仍未解决的是，系统的平均交易值在增加, 但系统的能力范围却是固定的,其中低价值的交易保存在主账本中已无经济意义,因而会被忽略。尽管如此，一些策略存有可能发展为可扩展性的协议来解决此问题。
+
+一些形式的层次结构，通过合并小的轻量的链到主区块或通过组合小的交易集合构建主区块, 或许可能实现交易组合和区块构建的并行化。并行化机制也可以这样实现: 针对有优先顺序的并列的区块链,通过合并或抛弃无效交易来合并区块。
+
+最后，可验证的计算，如果可用并且足够有效，可能提供一个方法允许工作量证明成为最后状态的验证。
+
+\section{结论} \label{ch:conclusion}
+我们已经介绍讨论并正式定义了以太坊协议。通过这个协议，读者可以太坊网络上部署一个节点并在一个去中心化的安全的开放系统中连接其它节点。合约可能会被授权, 以便计算性地确定和自动化执行交互规则。
+
+\section{致谢}
+
+非常感谢 Aeron Buchanan 编写家园版本, Christoph Jentzsch 编写 Ethash 算法,Yoichi Hirai 做了 EIP-150 大多数更新。以太坊开发组织和社区提供了重要的维护和大量的矫正及建议,其中包括Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari, 当然,还包含Vitalik Buterin。
+
+\section{可用性}
+本文的英文源文件维护在 \url{https://github.com/ethereum/yellowpaper}. 一个自动生成的英文 PDF 格式文件见 \url{https://ethereum.github.io/yellowpaper/paper.pdf}。
+
+本文的中文源文件维护在 \url{https://github.com/yuange1024/ethereum_yellowpaper}. 一个自动生成的中文 PDF 格式文件见 \url{https://github.com/yuange1024/ethereum_yellowpaper/blob/master/Paper_Chinese.pdf}。 (本段由译者增加)
 

### 2017-09-18 Paper_Chinese.tex
翻译校对

@@
 
-随着互联网连接了世界上绝大多数地方，全球信息共享的成本越来越低。比特币网络通过共识机制、自愿遵守的社会合约,实现一个去中心化的价值转移系统且可以在全球范围内自由使用, 这样的技术改革展示了它的巨大力量。这样的系统可以说是加密安全、基于交易的状态机的一种具体应用。后续类似这样的系统,如域名币（Namecoin）, 从最原先的"货币应用"发展到了其它应用,虽然它只是其中很简单的一种应用。
-
-以太坊是一个尝试实现通用性技术的项目,所有基于交易的状态机都可以被构建。而且以太坊致力于为开发者提供主流一个紧凑的、整合的端到端系统, 这个系统提供了一种可信的消息传递计算框架让开发者以一种前所未有的范式来构建软件。
+随着互联网连接了世界上绝大多数地方，全球信息共享的成本越来越低。比特币网络通过共识机制、自愿遵守的社会合约,实现一个去中心化的价值转移系统且可以在全球范围内自由使用, 这样的技术改革展示了它的巨大力量。这样的系统可以说是加密安全、基于交易的状态机的一种具体应用。后续类似这样的系统,如域名币（Namecoin）, 从最原先的货币应用发展到了其它应用,虽然它只是其中很简单的一种应用。
+
+以太坊是一个尝试达到通用性的技术项目,可以构建任何基于交易的状态机。而且以太坊致力于为开发者提供一个紧凑的、整合的端到端系统, 这个系统提供了一种可信的消息传递计算框架让开发者以一种前所未有的范式来构建软件。
 
@@
 
-这个提议系统中的交易, 有一些在现实世界中并不常见的属性。审判廉洁，在现实世界往往很难找到，但对公正的算法解释器是天然的; 透明，或者说通过交易日志和规则或代码指令能够清晰的看见到状态变化或者判决, 但因为人类语言的模糊性、信息的缺乏以及老的偏见难以撼动,导致基于人的系统中从来没有完美实现透明。
+这个提议系统中的交易, 有一些在现实世界中并不常见的属性。审判廉洁，在现实世界往往很难找到，但对公正的算法解释器是天然的; 透明，或者说通过交易日志和规则或代码指令能够清晰的看到状态变化或者判决, 但是因为人类语言的模糊性、信息的缺乏以及老的偏见难以撼动,导致基于人的系统中从来没有完美实现透明。
 
@@
 
-\cite{vishnumurthy03karma:a}最早使用工作量证明作为强大的经济信号保证货币安全。 在这个案例中，代币用作检查点对点（peer-to-peer,p2p）文件交易，同时保证“消费者”能支付给为他们提供服务的“供应商”。这种通过工作量证明的安全模型逐步扩展, 包括使用电子签名和账本技术, 以保证历史记录不被篡改, 怀有恶意的用户不能进行欺诈支付或不公平的抱怨服务。五年后（2008年），中本聪\cite{nakamoto2008bitcoin} 介绍了另一种更广泛的工作量证明安全价值代币。这个项目的成果比特币,成为了第一个被全球广泛认可的去中心化交易账本。
-
-由于比特币的成功，竞争币(alt-coins)开始兴起,通过改变比特币的协议去创建了大量的其他数字货币。比较知名的有莱特币（Litecoin）和素数币（Primecoin）, 参见 \cite{sprankel2013technical} 。一些项目使用比特币的核心机制并重新改造以应用在其它领域,例如域名币（Namecoin）, 致力于提供一个去中心化的名字解析系统, 参见 \cite{aron2012bitcoin} 。
-
-其它在比特币网络之上构建的项目, 也是依赖巨大的系统价值和巨大的算力来保证共识机制。万事达币（Mastercoin）项目, 在比特币协议之上,通过一系列基于核心协议的的辅助插件, 构建一个包含许多高级功能的富协议, 参见 \cite{mastercoin2013willett}。彩色币（Coloured Coins, 参见 \cite{colouredcoins2012rosenfeld}, 采用了类似的但更简化的协议, 以实现比特币基础货币的可替代性, 并允许通过色度钱包（“chroma-wallet”）来创建和跟踪代币。
-
-其它一些工作通过放弃中心化来进行。瑞波币（Ripple）,参见 \cite{boutellier2014pirates}, 试图去创建一个货币兑换的联邦系统（“federated ”system）和一个新的金融清算系统。这个系统展示了放弃去中心化特性可以获得性能上的提升。
+\cite{vishnumurthy03karma:a}最早使用工作量证明作为强大的经济信号保证货币安全。 在这个案例中，代币用作检查点对点（peer-to-peer,p2p）文件交易，同时保证“消费者”能支付给为他们提供服务的“供应商”。这种通过工作量证明的安全模型逐步扩展, 包括使用电子签名和账本技术, 以保证历史记录不被篡改, 怀有恶意的用户不能进行欺诈支付或不公平的抱怨。五年后（2008年），中本聪\cite{nakamoto2008bitcoin} 介绍了另一种更广泛的工作量证明安全价值代币。这个项目的成果就是比特币,比特币成为了第一个被全球广泛认可的去中心化交易账本。
+
+由于比特币的成功，竞争币(alt-coins)开始兴起,通过修改比特币的协议创建了大量的其它数字货币。比较知名的有莱特币（Litecoin）和素数币（Primecoin）, 参见 \cite{sprankel2013technical} 。一些项目使用比特币的核心机制并重新改造以应用在其它领域,例如域名币（Namecoin）致力于提供一个去中心化的名字解析系统, 参见 \cite{aron2012bitcoin} 。
+
+其它在比特币网络之上构建的项目, 也是依赖巨大的系统价值和巨大的算力来保证共识机制。万事达币（Mastercoin）项目是在比特币协议之上,通过一系列基于核心协议的的辅助插件, 构建一个包含许多高级功能的富协议, 参见 \cite{mastercoin2013willett}。彩色币（Coloured Coins, 参见 \cite{colouredcoins2012rosenfeld}, 采用了类似的但更简化的协议, 以实现比特币基础货币的可替代性, 并允许通过色度钱包（chroma-wallet）来创建和跟踪代币。
+
+其它一些工作通过放弃中心化来进行。瑞波币（Ripple）,参见 \cite{boutellier2014pirates}, 试图去创建一个货币兑换的联邦系统（“federated” system）和一个新的金融清算系统。这个系统展示了放弃去中心化特性可以获得性能上的提升。
 
@@
 
-以太坊在整体上可以看成是一个基于交易的状态机：我们起始于一个创世块（Genesis）状态，然后随着交易的执行状态逐步改变一直到最终状态, 这个最终状态是以太坊世界的权威版本。状态中包含的信息有:账户余额、名誉度、信誉度、现实世界的附属数据等; 简而言之，能包含电脑可以描绘的任何信息。因此，交易是连接两个状态的有效桥梁；“有效”非常重要---因为无效的状态改变远超过有效的状态改变。例如：无效的状态改变可能是减少账号余额，但是没有在其它账户上加上同等的额度。一个有效的状态转换是通过交易进行的,表达式如下:
+以太坊在整体上可以看作一个基于交易的状态机：起始于一个创世块（Genesis）状态，然后随着交易的执行状态逐步改变一直到最终状态, 这个最终状态是以太坊世界的权威版本。状态中包含的信息有:账户余额、名誉度、信誉度、现实世界的附属数据等; 简而言之，能包含电脑可以描绘的任何信息。因此，交易是连接两个状态的有效桥梁；“有效”非常重要---因为无效的状态改变远超过有效的状态改变。例如：无效的状态改变可能是减少账号余额，但是没有在其它账户上加上同等的额度。一个有效的状态转换是通过交易进行的,表达式如下:
 \begin{equation}
@@
 
-区块中记录着交易信息；区块之间通过密码学哈希(hash)链接起来。区块链就像一个分类账,将一系列交易记录在一起,并且连接上一个区块及最终状态(并没有直接保存最终状态本身---否则整个区块链就太大了）。系统激励节点去挖矿, 挖矿激励时,有执行状态转移函数,增加挖矿者的账户余额。
-
-挖矿是和其它潜在区块竞争一系列交易(一个区块)的记账权。它是通过密码安全证明的方式来实现的。这个机制称为工作量证明, 会在  \ref{ch:pow} 详细讨论。
+区块中记录着交易信息；区块之间通过密码学哈希(hash)链接起来。区块链就像一个分类账,将一系列交易记录在一起,并且连接上一个区块及最终状态(并没有直接保存最终状态本身---否则整个区块链就太大了）。系统激励节点去挖矿, 挖矿获得激励后,会执行状态转移函数,增加挖矿者的账户余额。
+
+挖矿是和其它潜在区块竞争一系列交易(一个区块)的记账权。它是通过密码安全证明的方式来实现。这个机制称为工作量证明, 会在  \ref{ch:pow} 详细讨论。
 
@@
 
-为了激励网络中的计算，需要定义一种转账方法。以太坊设计了一个内置货币以太币(Ether)，ETH是大家所熟知的符号，有时用 \DH{} 表示。以太币最小的面额是Wei(伟)，所有货币值都以Wei的整数倍记录。一个以太币被定义成位。一个以太币等于 $10^{18}$ Wei 。不同的面值如下表:
+为了激励网络中的计算，需要定义一种转账方法。以太坊设计了一种内置货币以太币(Ether)，ETH是大家所熟知的符号，有时用 \DH{} 表示。以太币最小的面额是Wei(伟)，所有货币值都以Wei的整数倍记录。一个以太币等于 $10^{18}$ Wei 。不同的面值如下表:
 
@@
 
-因为这是一个去中心化的系统, 所有人都有机会在之前的某一个区块创建新的区块并连接在其后, 这会行成一个树状的区块。为了能在这个树状结构上从根节点（创世块）到叶子节点（包含最新交易的区块）能形成一个一致的区块链,必须有一个共识方案。如果有人认为从根节点到叶子节点的路径不是"最佳"的区块链, 那这时候就会发生分叉。
-
-这个就意味着在一个给定的时间点, 系统中会有多个状态共存：一些节点相信一个区块是包含标准的交易，其他的节点则相信另外一些区块包含标准的交易，其中就包含彻底不同或者不兼容的交易。这一点必须要避免, 因为它会破坏整个系统信用。
-
-我们使用了一个简单GHOST协议版本来达成共识, 参见 \cite{cryptoeprint:2013:881} 。我们会在\ref{ch:ghost} 详细说明。
+因为这是一个去中心化的系统, 所有人都有机会在之前的某一个区块创建新的区块并连接在其后, 这会行成一个树状的区块。为了能在这个树状结构上从根节点（创世块）到叶子节点（包含最新交易的区块）能形成一个一致的区块链,必须有一个共识方案。如果有人认为从根节点到叶子节点的路径不是最佳的区块链, 那这时候就会发生分叉。
+
+这就意味着在一个给定的时间点, 系统中会有多个状态共存：一些节点相信一个区块是包含权威的交易，其它节点则相信另外一些区块包含权的交易，其中就包含彻底不同或者不兼容的交易。这一点必须要避免, 因为它会破坏整个系统信用。
+
+我们使用了一个简易 GHOST 协议版本来达成共识, 参见 \cite{cryptoeprint:2013:881} 。我们会在小结 \ref{ch:ghost} 详细说明。
 
@@
 
-我用了大量的印刷约定表示公式中的符号，其中一些需要特别说明：
+我们用了大量的印刷约定表示公式中的符号，其中一些需要特别说明：
 
@@
 
-作用在高度结构化值上的函数，使用大写的希腊字母，例如： $\Upsilon$ , 是以太坊中的状态转换函数。
-
-对于大部分函数来说，通常用一个大写的字母表示，例如：$C$, 表示费用函数。可能会使用下角标表示为一些特别的变量，例如：$C_\text{\tiny SSTORE}$，表示执行 {\tiny SSTORE} 操作的费用函数。对于一些可能是外部定义的函数，可能会使用打印机文字字体，例如：$\texttt{KEC512}$ 哈希函数（为赢得进入SHA-3竞赛） ,使用 $\texttt{KEC}$ 表示。$\texttt{KEC512}$  表示 Keccack-512 哈希函数。
-
-元组通常使用一个大写字母，例如：$T$，表示一个以太坊交易。使用下标可能会表示一个独立的变量,例如： $T_n$ ，表示交易中随机数。角标的形式用于表示它们的类型；例如：大写的下角标表示元组包含的下角标变量。
+作用在高度结构化数据上的函数，使用大写的希腊字母，例如： $\Upsilon$ , 是以太坊中的状态转换函数。
+
+对于大部分函数来说，通常用一个大写的字母表示，例如：$C$, 表示费用函数。可能会使用下角标表示一些特别的变量，例如：$C_\text{\tiny SSTORE}$，表示执行 {\tiny SSTORE} 操作的费用函数。对于一些可能是外部定义的函数，可能会使用打印机文字字体，例如：$\texttt{KEC512}$ 哈希函数,使用 $\texttt{KEC}$ 表示。$\texttt{KEC512}$  表示 Keccack-512 哈希函数。
+
+元组通常使用一个大写字母，例如：$T$，表示一个以太坊交易。使用下标可能会表示一个独立的变量,例如： $T_n$ ，表示交易中随机数。角标形式用于表示它们的类型；例如：大写的下角标表示元组包含的下角标变量。
 
@@
 
-我们认为标量都是正整数且属于集合 $\mathbb{P}$。所有的字节序列属于集合 $\mathbb{B}$，附录 \ref{app:rlp} 给出了正式的定义。用下角标符号表示这样的序列集合限制在一定长度以内, 长度为32的字节序列使用 $\mathbb{B}_{32}$ 表示, 所有比 $2^{256}$ 小的正整数使用 $\mathbb{P}_{256}$ 表示。详细定义见 \ref{ch:block}。
-
-使用方括号表示序列中的一个元素或子序列,例如： $\boldsymbol{\mu}_\mathbf{s}[0]$ 表示计算机堆栈中的第一个条目。对于子序列来说，使用省略号表示一定的范围，且含头尾的限制，例如：$\boldsymbol{\mu}_\mathbf{m}[0..31]$ 表示计算机内存中的前32个条目。
-
-在全局状态的情况下，σ是一个含多个账号的序列，本身的数组，正方形括号被用作去表示一个单独的账号。
+我们认为标量都是正整数且属于集合 $\mathbb{P}$。所有的字节序列属于集合 $\mathbb{B}$，附录 \ref{app:rlp} 给出了正式的定义。用下角标表示这样的序列集合限制在一定长度以内, 长度为 32 的字节序列使用 $\mathbb{B}_{32}$ 表示, 所有比 $2^{256}$ 小的正整数使用 $\mathbb{P}_{256}$ 表示。详细定义见 \ref{ch:block}。
+
+使用方括号表示序列中的一个元素或子序列,例如： $\boldsymbol{\mu}_\mathbf{s}[0]$ 表示栈中的第一个条目。对于子序列来说，使用省略号表示一定的范围，且含首尾限制，例如：$\boldsymbol{\mu}_\mathbf{m}[0..31]$ 表示计算机内存中的前32个条目。
+
+在全局状态的情况下，σ 是一个含多个账号的序列，本身的数组，方形括号被用作去表示一个单独的账号。
 
@@
 
-当去考虑现有的变量时，我遵循在给定的范围内去定义的原则，我们使用占位符 $\Box$ 表示未修改的输入变量, 使用 $\Box'$ 表示修改的和可用的变量, $\Box^*$,  $\Box^{**}$ \&c 表示中间变量。在特殊情况下，为了提高可读性和清晰性，我可能会使用字母-数字下角标表示中间值。
-
-当使用去已有的函数时，给定一个函数 $f$,那么函数 $f^*$ 表示一个相似的、替换序列的函数映射。详细的定义见 \ref{ch:block}。
-
-整个过程中，我定义了大量的函数。一个常见的函数是 $\ell$，表示给定序列的最后一个条目：
+当去考虑现有的变量时，我们遵循在给定的范围内去定义的原则，使用占位符 $\Box$ 表示未修改的输入变量, 使用 $\Box'$ 表示修改的和可用的变量, $\Box^*$,  $\Box^{**}$ \&c 表示中间变量。在特殊情况下，为了提高可读性和清晰性，我可能会使用字母-数字下角标表示中间值。
+
+当使用已有的函数时，给定一个函数 $f$,那么函数 $f^*$ 表示一个相似的、替换序列的函数映射。详细定义见 \ref{ch:block}。
+
+整个过程中，我们定义了大量的函数。一个常见的函数是 $\ell$，表示给定序列的最后一个条目：
 
@@
 
-介绍了以太坊的基本概念后,我们将更详细地讨论交易、区块和状态的含义
+介绍了以太坊的基本概念后,我们将详细地讨论交易、区块和状态的含义。
 
@@
 
-世界状态是在地址（160位的标志符）和账户状态（序列化为RLP的数据结构，详见附录 \ref{app:rlp}）的映射。虽然世界状态没有直接储存在区块链上，但会假定实施过程中会将这个映射维护在一个修改过的Merkle Patricia 树(简称trie，详见附录 \ref{app:trie}）。trie 需要一个简单的后端数据库去维护字节数组到字节数组的映射；我们称这个后端数据库为状态数据库。它有一系列的好处: 第一这个结构的根节点是加密的且依赖于所有的内部数据，且它的哈希可以作为整个系统状态的一个安全标志;第二，作为一个不变的数据结构，因此它允许任何一个之前状态（根部哈希已知的条件下）通过简单地改变根部哈希值而被召回。因为我们在区块链中储存了所以这样的根部哈希值，所以我们能恢复到指定的历史状态。
+世界状态是在地址（160 位的标志符）和账户状态（序列化为RLP的数据结构，详见附录 \ref{app:rlp}）的映射。虽然世界状态没有直接储存在区块链上，但会假定在实施过程中会将这个映射维护在一个修改过的 Merkle Patricia 树(简称 trie,字典树,详见附录 \ref{app:trie}）。字典树需要一个简单的后端数据库去维护字节数组到字节数组的映射；我们称这个后端数据库为状态数据库。它有一系列的好处: 第一,这个结构的根节点是加密的且依赖于所有的内部数据，它的哈希可以作为整个系统状态的一个安全标志;第二，作为一个不变的数据结构，因此它允许任何一个之前状态（根部哈希已知的条件下）通过简单地改变根部哈希值而被召回。因为我们在区块链中储存了所以这样的根部哈希值，所以我们能恢复到指定的历史状态。
 
@@
 \begin{description}
-\item[nonce,随机数] 这个值等于账户发出的交易数及这个账户创建的合约数量之和。 $\boldsymbol{\sigma}[a]_n$ 表示状态 $\boldsymbol{\sigma}$ 中的地址 $a$ 的nonce值。
-\item[balance,余额] $\boldsymbol{\sigma}[a]_b$, 表示这个账户拥有多少Wei。
-\item[storageRoot,存储跟节点] 保存账户内容的 Merkle Patricia 树根节点的256位哈希编码到trie中, 作为从 256位整数键值哈希的Keccak 256 位哈希到 256位整数的RLP-编码映射。这个哈希定义为 $\boldsymbol{\sigma}[a]_s$。
-\item[codeHash,代码哈希] 这个账户的EVM(Ethereum Virtual Machine,以太坊虚拟机)代码的哈希值---代码执行时,这个地址会接收一个消息调用; 它和其它字段不同, 创建后不可更改。状态数据库中包含所有像这样的代码片段的哈希, 以便后续使用。 这个哈希定义为 $\boldsymbol{\sigma}[a]_c$,$\mathbf{b}$ 表示代码,  $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+\item[nonce,随机数] 这个值等于账户发出的交易数及这个账户创建的合约数量之和。 $\boldsymbol{\sigma}[a]_n$ 表示状态 $\boldsymbol{\sigma}$ 中的地址 $a$ 的 nonce 值。
+\item[balance,余额] $\boldsymbol{\sigma}[a]_b$, 表示这个账户拥有多少 Wei。
+\item[storageRoot,存储根节点] 保存账户内容的 Merkle Patricia 树根节点的 256 位哈希编码到字典树中, 作为从 256 位整数键值哈希的Keccak 256 位哈希到 256 位整数的 RLP-编码映射。这个哈希定义为 $\boldsymbol{\sigma}[a]_s$。
+\item[codeHash,代码哈希] 这个账户的 EVM(Ethereum Virtual Machine,以太坊虚拟机)代码哈希值---代码执行时,这个地址会接收一个消息调用; 它和其它字段不同, 创建后不可更改。状态数据库中包含所有像这样的代码片段哈希, 以便后续使用。 这个哈希定义为 $\boldsymbol{\sigma}[a]_c$,$\mathbf{b}$ 表示代码,  $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
 
-因为我通常希望所指的并不是trie树的根哈希,而是所保存的键值对集合，我做了一个更方便的定义:
+因为我通常希望所指的并不是字典树树的根哈希,而是所保存的键值对集合，我做了一个更方便的定义:
 
@@
 
-trie树中的键值对集合函数, $L_I^*$, 定义为基于基础函数 $L_I$ 的元素转换:
+字典树中的键值对集合函数, $L_I^*$, 定义为基于基础函数 $L_I$ 的元素转换:
 \begin{equation}
@@
 
-如果 \textbf{codeHash} 字段是一个空字符串的 Keccak-256 哈希, 例如 $\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, 则表示对应的节点表示一个简单账户, 有时简称"非合约"账户。
-
-因此我们可能定义一个世界状态的函数  $L_S$：
+如果 \textbf{codeHash} 字段是一个空字符串的 Keccak-256 哈希, 例如 $\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, 则表示对应的节点表示一个简单账户, 有时简称非合约账户。
+
+因此我们可能定义一个世界状态函数 $L_S$：
 
@@
 
-函数 $L_S$ 和 trie 函数是为了提供一个世界状态的简短身份（哈希）。我们假定：
+函数 $L_S$ 和 字典树函数是为了提供一个世界状态的简短身份（哈希）。我们假定：
 \begin{equation}
@@
 \item[nonce,随机数] $T_n$,账户发出的交易数量。
-\item[gasPrice,燃料价格] $T_p$, 为执行交易所需要的计算资源付的 \textit{gas} 价格, 以Wei为单位。
+\item[gasPrice,燃料价格] $T_p$, 为执行交易所需要的计算资源付的 \textit{gas} 价格, 以 Wei 为单位。
 \item[gasLimit,燃料上限] $T_g$, 用于执行交易的最大 gas 数量。 这个值须在交易前设置,且设定后不能再修改。
-\item[to, 接收者地址] 消息调用接收者的160位的地址。对与合约创建交易, 无需接收者地址, 使用 $\varnothing$ 表示, $\varnothing$ 是 $\mathbb{B}_0$ 的唯一成员。
-\item[value, 转账额度] $T_v$, 转到接收者账户的额度,以Wei为单位。对于合约创建,表示捐赠到合约地址的额度。
+\item[to, 接收者地址] 消息调用接收者的 160 位地址。对与合约创建交易, 无需接收者地址, 使用 $\varnothing$ 表示, $\varnothing$ 是 $\mathbb{B}_0$ 的唯一成员。
+\item[value, 转账额度] $T_v$, 转到接收者账户的额度,以 Wei 为单位。对于合约创建,表示捐赠到合约地址的额度。
 \item[v, r, s] $T_w$, $T_r$ and $T_s$,和交易签名相关的变量, 用于确定交易的发送者。详见附录 \ref{app:signing}。
@@
 
-\textbf{init} 是EVM代码片段; 执行 init后会返回另外一个代码片段, 每次合约接受消息调用(通过交易或内部调用)后都会执行这个代码片段。\textbf{init} 仅当合约账户创建的时候执行一次。
+\textbf{init} 是EVM代码片段; 执行 init后会返回另外一个代码片段, 每次合约接受消息调用(通过交易或内部调用)后都会执行这个代码片段。 仅当合约账户创建时会执行一次 \textbf{init}。
 
@@
 
-地址哈希 $T_\mathbf{t}$ 稍微有些不同:它是一个20字节的地址哈希值, 但创建合约时它是RLP空字节系列, 表示为$\mathbb{B}_0$：
+地址哈希 $T_\mathbf{t}$ 稍微有些不同:它是一个 20 字节的地址哈希值, 但创建合约时它是RLP空字节系列, 使用 $\mathbb{B}_0$ 表示：
 \begin{equation}
@@
 
-在以太坊中, 区块是相关信息的集合。区块头 $H$, 与之想对应的交易信息 $\mathbf{T}$, 其它区块头数据的集合 $\mathbf{U}$，$\mathbf{U}$ 表示它的父级区块中有和当前区块的爷爷辈区块是相同的。（这样的区块称为\textit{ommers}\footnote{\textit{ommer} 的意思和自然界中的 "父母的兄弟姐妹"最相近, 详见 \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}, 译者注:不妨称之为叔链)。区块头包含的的信息如下：
+在以太坊中, 区块是相关信息的集合。与区块头 $H$ 想对应的交易信息为 $\mathbf{T}$, 其它区块头数据的集合 $\mathbf{U}$  表示它的父级区块中有和当前区块的爷爷辈区块是相同的。（这样的区块称为\textit{ommers}\footnote{\textit{ommer} 的意思和自然界中的父母的兄弟姐妹最相近, 详见 \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}, 译者注:不妨称之为叔链)。区块头包含的的信息如下：
 
@@
 为了让交易信息编码能有利于零知识证明、索引、搜索, 我们将每个包含一定信息的交易收据进行编码。以 $B_\mathbf{R}[i]$ 表示第 $i$ 个交易,保存在一个索引字典树中,$H_e$ 是这个字典树的根节点。
-为了去编译信息关于每一个有关系的交易而且可能是一个有用的工具去形成一个零知识证明（zero-knowledge proof），或者索引和搜索。我们编译每一个包含当前交易执行的某些信息为交易收据。每个收据（在第个交易中表示为）是被放置于一个带关键字索引的trie中和这个区块头中被记录下的根值，表示为。
-
-交易收据是一个包含四个条目的元组：交易后的状态,$R_{\boldsymbol{\sigma}}$；当前区块中交易累计燃料使用量, $R_u$, 交易发生后会立即更新这个值;交易执行过程中创建的日志集合,$R_\mathbf{l}$；和日志Bloom过滤器,$R_b$ ：
+
+交易收据是一个包含四个条目的元组：交易后的状态 $R_{\boldsymbol{\sigma}}$；当前区块中交易累计燃料使用量 $R_u$, 交易发生后会立即更新这个值;交易执行过程中创建的日志集合 $R_\mathbf{l}$；和日志 Bloom 过滤器 $R_b$ ：
 
@@
 
-其中 $M_{3:2048}$ 是一个特别的 Bloom 过滤器,针对任意一个字节序列,它舍弃这个字节序列2048位的前三位。它通过对一个字节序列的 Keccak-256 哈希的每一个前三对字节取其的低11位来实现:
+其中 $M_{3:2048}$ 是一个特别的 Bloom 过滤器,针对任意一个字节序列,它舍弃这个 2048 位字节序列的前三位。它通过对一个字节序列的 Keccak-256 哈希的每一个前三对字节取其的低11位来实现:
 \begin{eqnarray}
@@
 
-如果一个区块同时满足以下几个条件，我们才能认为这个区块是有效的：当从起始状态 $\boldsymbol{\sigma}$ (父块的最终状态) 按顺序执行完生成新的状态 $H_r$ 后, 在内部上要保持一致,包括叔链、交易区块哈希、给定的交易 $B_\mathbf{T}$（详细描述见 \ref{ch:finalisation}）:
+如果一个区块同时满足以下几个条件，我们才能认为这个区块是有效的：当从起始状态 $\boldsymbol{\sigma}$ (父块的最终状态) 按顺序执行完生成新的状态 $H_r$ 后, 在内部要保持一致,包括叔链、交易区块哈希、给定的交易 $B_\mathbf{T}$（详细描述见 \ref{ch:finalisation}）:
 
@@
 
-函数 $L_B$ 和 $L_H$ 分别是区块和区块头的准备函数。类似交易收据准备函数 $L_R$，当转换为 RLP 格式时,假设对应的类型、顺序及结构如下:
+函数 $L_B$ 和 $L_H$ 分别是区块和区块头的预备函数。类似交易收据预备函数 $L_R$，当转换为 RLP 格式时,假设对应的类型、顺序及结构如下:
 
@@
 
-随机数 $H_n$，必须满足下面关系：
+随机数 $H_n$ 需满足下面关系：
 \begin{equation}
@@
 
-其中 $H_{\hcancel{n}}$ 是新区块的区块头，但不包含随机数和混合哈希值，$\mathbf{d}$  是当前的大数据集合 DAG(有向无环图)，需要去计算混合哈希,  $\mathtt{PoW}$ 是工作量证明函数（见 \ref{ch:pow}）：第一个元素用于计算混合哈希值，以证明使用了一个正确的 DAG, 第2个元素是伪随机数,依赖于 $H$ 及 $\mathbf{d}$ 。给定一个范围在$[0, 2^{64})$ 的均匀分布，则求解时间和难度$H_d$ 成比例。
-
-这就是区块链安全基础，这也是一个恶意节点不能用其新创建的区块中重写历史数据的重要原因。因为这个随机数必须满足这些条件，且因为条件依赖于这个区块的内容和相关交易, 创建新的合法的区块是困难的、耗时的, 需要超过所有诚实矿工的算力总和。
+其中 $H_{\hcancel{n}}$ 是新区块的区块头，但不包含随机数和混合哈希值，$\mathbf{d}$  是当前的数据集合 DAG(有向无环图)，需要去计算混合哈希,  $\mathtt{PoW}$ 是工作量证明函数（见 \ref{ch:pow}）：第一个元素用于计算混合哈希值，以证明使用了一个正确的 DAG, 第 2 个元素是伪随机数,依赖于 $H$ 及 $\mathbf{d}$ 。给定一个范围在$[0, 2^{64})$ 的均匀分布，则求解时间和难度 $H_d$ 成比例。
+
+这就是区块链安全基础，这也是一个恶意节点不能用其新创建的区块中重写历史数据的重要原因。因为这个随机数必须满足这些条件，且因为条件依赖于这个区块的内容和相关交易, 创建新的合法的区块是困难且耗时的, 需要超过所有诚实矿工的算力总和。
 
@@
 
-为了避免网络滥用及回避由于图灵完整性而带来的一些不可避免的问题，在以太坊中所有的编程计算都需要费用。各种操作费用以 \textit{gas} (详见附录 \ref{app:fees} )为单位计算。任意的程序片段（包括合约创建、信息调回、利用及访问账户存储、在虚拟机上执行操作等）都可以根据规则计算出消耗的燃料。
-
-每一个交易都有一个燃料上限：\textbf{gasLimit} (燃料上限)。这些燃料从发送者的账户中扣除。具体从账户上扣除的额度和 \textbf{gasPrice}(燃料价格) 有关(译者注: 扣除额度 = \textbf{gasLimit} * \textbf{gasPrice}), 在执行交易前会指定燃料价格。如果这个账户不能支付起燃料费用，这个交易会被当作无效交易。之所以它被命名为燃料上限，是因为剩余的燃料在交易完成之后会被退回（以购买时的同样价格）到发送者账户。燃料不会被用在交易执行之外。因此对于可信任账户，应该设置一个相对较高的燃料上限。
-
-通常来说，以太币（Ether）用作去购买燃料, 未退回的那部分转到了区块受益人的地址，通常这个账户的地址是由矿工设定。交易者可以任意设定燃料价格，然而矿工也可以任意地忽略某个交易。在一个交易中，高价格的燃料将消费这个发送者更多的以太币,并转给矿工更多的以太币，因此这个交易会被更多的矿工选择。通常来说，矿工将会选择去通知这是他们执行交易最低燃料价格, 交易者们一般也会些选择一个高过燃料价格下限的价格。因此, 会有一个（加权的）最低燃料可接受价格分布，交易者们需要权衡降低燃料价格和交易快速被矿工打包。
+为了避免网络滥用及回避由于图灵完整性而带来的一些不可避免的问题，在以太坊中所有的程序执行都需要费用。各种操作费用以 \textit{gas} (详见附录 \ref{app:fees} )为单位计算。任意的程序片段（包括合约创建、信息调回、利用及访问账户存储、在虚拟机上执行操作等）都可以根据规则计算出消耗的燃料数量。
+
+每一个交易都有燃料上限：\textbf{gasLimit} (燃料上限)。这些燃料从发送者的账户中扣除。具体从账户上扣除的额度和 \textbf{gasPrice}(燃料价格) 有关(译者注: 扣除额度 = \textbf{gasLimit} * \textbf{gasPrice}), 在执行交易前会指定燃料价格。如果这个账户不能支付起燃料费用，交易会被当作无效交易。之所以命名为燃料上限，是因为剩余的燃料在交易完成之后会被退回（以购买时的同样价格）到发送者账户。燃料不会被用在交易执行之外。因此对于可信任账户，应该设置一个相对较高的燃料上限。
+
+通常来说，以太币（Ether）用作购买燃料, 未退回的部分转到了区块受益人的地址，通常这个账户的地址是由矿工设定。交易者可以任意设定燃料价格，然而矿工也可以任意地忽略某个交易。在一个交易中，高价格的燃料将消费这个发送者更多的以太币,并转给矿工更多的以太币，因此这个交易会被更多的矿工选择。通常来说，矿工会选择去通知这是他们执行交易最低燃料价格, 交易者们一般也会些选择一个高过燃料价格下限的价格。因此, 会有一个（加权的）最低燃料可接受价格分布，交易者们需要权衡降低燃料价格和交易快速被矿工打包。
 
@@
 
-$\boldsymbol{\sigma}'$ 是交易后的状态。我们定义 $\Upsilon^g$ 为交易执行所消耗的燃料量, $\Upsilon^\mathbf{l}$ 为交易过程中产生的日志记录, 都会在后文正式定义。
+$\boldsymbol{\sigma}'$ 是交易后的状态。我们定义 $\Upsilon^g$ 为交易执行所消耗的燃料量, $\Upsilon^\mathbf{l}$ 为交易过程中产生的日志记录, 这些都会在后文正式定义。
 
@@
 
-元组内容包含自毁集合 $A_\mathbf{s}$: 一个将在交易完成后被删除的账户集合。 $A_\mathbf{l}$ 是一系列的日志：在代码执行过程中是可归档, 可索引的检查点, 允许以太坊世界的外部旁观者(例如去中心化应用的前端)跟踪合约调用。$A_r$ 表示返回的余额，当使用 {\small SSTORE} 指令将非0的合约存贮空间置为0时,返回余额会增加。虽然不是立即返回余额，但可以部分抵消整个执行费用。
-
-我们定义空的子状态 $A^0$, 它没有自毁、没有日记、返回余额为0：
+元组内容包含自毁集合 $A_\mathbf{s}$: 一个在交易完成后被删除的账户集合。 $A_\mathbf{l}$ 是一系列的日志：在代码执行过程中是可归档, 可索引的检查点, 允许以太坊世界的外部旁观者(例如去中心化应用的前端)跟踪合约调用。$A_r$ 表示返回的余额，当使用 {\small SSTORE} 指令将非0的合约存贮空间置为0时,返回余额会增加。虽然不是立即返回余额，但可以部分抵消整个执行费用。
+
+我们定义空的子状态 $A^0$, 它没有自毁、没有日志、返回余额为 0：
 \begin{equation}
@@
 
-注意最后的那个条件: 交易燃料限制 $T_g$  <= 这个区块的燃料限制 ${B_H}_l$ - 这个区块之前的所有交易已使用的燃料 $\ell(B_\mathbf{R})_u$  。
-
-有效交易的执行起始于一个对状态不能撤回的改变：发送者账户 $S(T)$ 的随机数会加1, 账户余额扣减部分预付费用 $T_gT_p$。计算过程中有效的燃料 $g$ = $T_g - g_0$。无论是合约创建还是消息调用，得到的最终这状态（可能等于当前的状态），这个改变是确定的且永久无效:从这个观点看不会存在无效的交易。
+注意最后的条件: 交易燃料限制 $T_g$  <= 这个区块的燃料限制 ${B_H}_l$ - 这个区块之前的所有交易已使用的燃料 $\ell(B_\mathbf{R})_u$  。
+
+有效交易的执行起始于一个对状态不能撤回的改变：发送者账户 $S(T)$ 的随机数会加1, 账户余额扣减部分预付费用 $T_gT_p$。计算过程中有效的燃料 $g$ = $T_g - g_0$。无论是合约创建还是消息调用后的的最终状态（可能等于当前的状态），这个改变是确定的且从来没有无效的:从这个观点看不会存在无效的交易。
 
@@
 
-$T_o$ 是原始执行者，与和合约创建或消息调用(直接由交易触发)不同, 可以通过执行 EVM 代码内部调用来执行。
+$T_o$ 是原始执行者，与和合约创建或消息调用(直接由交易触发)不同, 可以通过 EVM 代码内部调用来执行。
 
@@
 
-终需要退回发送者的燃料 $g^*$ 等于当前剩余的燃料 $g'$, 再加一个补偿, 这个补偿是 $A_r$ 和总使用燃料量 $T_g - g'$ 的一半中的小者。
+最终要退回发送者的燃料 $g^*$ 等于当前剩余的燃料 $g'$, 再加一个补偿, 这个补偿是 $A_r$ 和总使用燃料量 $T_g - g'$ 的一半中的小者。
 
@@
 
-我们定义创建函数为函数 $\Lambda$，相关的变量有状态 $\boldsymbol{\sigma}$ , 包含新状态、剩余燃料及交易子状态 $(\boldsymbol{\sigma}', g', A)$ (详见第6小结) :：
+我们定义创建函数为函数 $\Lambda$，相关的变量有状态 $\boldsymbol{\sigma}$ , 包含新状态、剩余燃料及交易子状态 $(\boldsymbol{\sigma}', g', A)$ (详见第6小结):
 \begin{equation}
@@
 
-最后，这个账户是通过执行初始化账户的 EVM 代码 $\mathbf{i}$ （执行模型详见小结 \ref{ch:model} ）来初始化的。代码执行可以影响一些事件：可以改变当前账户的存储，能创建更多的账户,执行更多的消息调用。代码执行函数 $\Xi$ 可以得到一个元组,包括结果状态 $\boldsymbol{\sigma}^{**}$, 还可以用的燃料剩余燃料 $g^{**}$, 子状态 $A$, 以及账户 $\mathbf{o}$ 的代码。    赋值到最后的状态的数组中，有效的剩余燃料，当前产生的子状态 $A$ 和账号代码本身 $\mathbf{o}$。
+最后，这个账户是通过执行初始化账户的 EVM 代码 $\mathbf{i}$ （执行模型详见小结 \ref{ch:model} ）来初始化的。代码执行可以影响一些事件：可以改变当前账户的存储，能创建更多的账户,执行更多的消息调用。代码执行函数 $\Xi$ 可以得到一个元组,包括结果状态 $\boldsymbol{\sigma}^{**}$, 可用的剩余燃料 $g^{**}$, 子状态 $A$, 以及账户 $\mathbf{o}$ 的代码。    赋值到最后的状态的数组中，有效的剩余燃料，当前产生的子状态 $A$ 和账号代码本身 $\mathbf{o}$。
 
@@
 
-代码执行消耗燃料，且燃料不能低于 0，因此程序执行可能会在自然终止之前会退出。在这个（以及其他几个）异常情况, 我们称发生了燃料不足（out-of-gas, OOG）异常：演进的状态变成空集合 $\varnothing$, 整个创建操作对状态没有影响,状态就像尝试创建合作之前一样。
-
-如果这个初始化代码成功地执行完，那么对应的合约创建的也会被支付。代码保存费用  $c$ 和创建的合约代码大小成正比:
+代码执行会消耗燃料，且燃料不能低于 0，因此程序执行可能会在自然终止之前退出。在这个（以及其它几个）异常情况, 我们称发生了燃料不足（out-of-gas, OOG）异常：演进的状态变成空集合 $\varnothing$, 整个创建操作对状态没有影响,状态就像尝试创建合约之前一样。
+
+如果这个初始化代码成功地执行完，那么对应的合约创建也会被支付。代码保存费用 $c$ 和创建的合约代码大小成正比:
 
@@
 
-发生这样的异常发生后, 剩余燃料将变为0。 如果合约创建是用于接受一个交易, 它不会影响合约创建内部消耗的燃料,无论如何都必须支付。然而，当燃料不足时, 并不会给这个合约地址转账。
+发生这样的异常后, 剩余燃料将变为0。 如果合约创建是用于接受一个交易, 它不会影响合约创建内部消耗的燃料,无论如何都必须支付。然而，当燃料不足时, 并不会给这个合约地址转账。
 
@@
 \section{消息调用} \label{ch:call}
-当执行消息调用时需要多个参数：发送者（$s$）、交易发起人（$o$）、接受者（$r$）、执行代码的账户（$c$,通常就是接受者）、可用的燃料（$g$）、转账额度（$v$）、燃料价格（$p$）、函数调用的一个任意长度字节的数组 $\mathbf{d}$ 的输入数据以及消息调用/合约创建的当前栈($e$)深度。。
-
-除了转变到新的状态和交易子状态外，消息调用还有一个额外的元素 --- 用字节数组 $\mathbf{o}$ 表示的输出数据。执行交易时输出数据是被忽略的,但在消息调用时, 输出的数据由取决于虚拟机代码执行, 在这种情况下也使用了这些信息。
+当执行消息调用时需要多个参数：发送者（$s$）、交易发起人（$o$）、接受者（$r$）、执行代码的账户（$c$,通常就是接受者）、可用的燃料（$g$）、转账额度（$v$）、燃料价格（$p$）、函数调用的一个任意长度字节的数组 $\mathbf{d}$ 的输入数据以及消息调用/合约创建的当前栈($e$)深度。
+
+除了转变到新的状态和交易子状态外，消息调用还有一个额外的元素 --- 用字节数组 $\mathbf{o}$ 表示的输出数据。执行交易时输出数据是被忽略的,但在消息调用时, 输出的数据可以由虚拟机代码执行时进行初始化, 在这种情况下也使用了这些信息。
 \begin{equation}
@@
 
-我们定义，在原始状态基础上进行了发送者向接收者转账后为的状态第一个转变状态 $\boldsymbol{\sigma}_1$:
+我们定义在原始状态基础上进行了发送者向接收者转账后的状态为第一个转变状态 $\boldsymbol{\sigma}_1$:
 
@@
 
-我们假设如果 $\boldsymbol{\sigma}_1[r]$ 还未定义, 则会创建一个没有代码或没有状态,且余额和随机数都为0的账户。我们改进上一个方程式:
+我们假设如果 $\boldsymbol{\sigma}_1[r]$ 还未定义, 则会创建一个没有代码或没有状态且余额和随机数都为 0 的账户。我们改进上一个公式:
 \begin{equation}
@@
 
-账户关联的代码（整个代码碎片, Keccak 哈希为  $\boldsymbol{\sigma}[c]_c$）依靠执行模式（见小结 \ref{ch:model}）执行。合约创建时，如果执行因为一个异常（例如：耗尽了燃料、堆栈溢出、无效的跳转目的地或者无效的指令）而被终止，燃料不会返回到调用者，且状态也会立即恢复到转账之前的状态(例如 $\boldsymbol{\sigma}$)。
+账户关联的代码（整个代码碎片, Keccak 哈希为  $\boldsymbol{\sigma}[c]_c$）依靠执行模式（见小结 \ref{ch:model}）执行。合约创建时，如果执行因为一个异常（例如：耗尽了燃料、堆栈溢出、无效的跳转目的地或者无效的指令）而被终止，燃料不会返回给调用者，且状态也会立即恢复到转账之前的状态(例如 $\boldsymbol{\sigma}$)。
 
@@
 \section{执行模型} \label{ch:model}
-执行模型具体说明怎么使用一系列字节代码指令和一个小的环境数据元组去改变这个系统状态。这些是通过以太坊虚拟机(Ethereum Virtual Machine - EVM)这个虚拟的状态来实现的。它是一个准图灵机,说"准"是因为计算会被燃料所限制。
+执行模型具体说明怎么使用一系列字节代码指令和一个小的环境数据元组去改变这个系统状态。这些是通过以太坊虚拟机(Ethereum Virtual Machine - EVM),这个虚拟状态机来实现的。它是一个准图灵机,说“准”是因为计算会被燃料所限制。
 
@@
 
-EVM 基于栈结构,机器的字大小（以及栈中数据的大小）是256位。主要是便于执行 Keccak-256 位哈希及椭圆曲线计算。内存模型基于字寻址的字节数据。栈的最大深度为1024。EVM也有一个独立的存储模型；类似内存但更新一个字节数组, 一个基于字寻址的字数组。不像易变的内存，存储是非易变的且是作为系统状态的一部分被维护。所有在内存和存储中的数据会初始化为0。
+EVM 基于栈结构,机器的字大小（以及栈中数据的大小）是256位。主要是便于执行 Keccak-256 位哈希及椭圆曲线计算。内存模型基于字寻址的字节数据。栈的最大深度为1024。EVM也有一个独立的存储模型；类似内存但更像一个字节数组, 一个基于字寻址的字数组。不像易变的内存，存储是非易变的且是作为系统状态的一部分被维护。所有内存和存储中的数据会初始化为0。
 
@@
 
-EVM 在某些情况下会有异常发生，包括堆栈溢出和非法指令。就像缺乏燃料异常那样, 不会改变状态。有异常时, EVM 立即停止并告知执行代理(交易的处理者，或执行环境), 代理会单独处理异常。
+EVM 在某些情况下会有异常发生，包括堆栈溢出和非法指令。就像燃料缺乏异常那样, 不会改变状态。有异常时, EVM 立即停止并告知执行代理(交易的处理者，或执行环境), 代理会单独处理异常。
 
@@
 
-在三个不同的情况下使用费用（命名为燃料），在这三种情况下，费用都是执行任何操作的必备条件。第一种情况也是最普通的情况就是计算操作费用（详见附录 \ref{app:fees}）。第二种情况，燃料可能因为一个子消息调用或者合约创建而被消耗,这是执行 {\small CREATE}, {\small CALL} and {\small CALLCODE} 费用中的一部分。第三种情况，燃料可能因为内存使用的增多而付费。
-
-对于一个账户的执行，内存总费用和最少需要的 32 字节倍数的内存量成正比, 所有的内存是按在一定范围中被包含的所有记忆索引（无论是读还是写）下要求的32位字节的最小倍数成比例的。这个为了 just-in-time (JIT) 的基础而去支付的；访问一个大于索引量 32 字节的地址的,就会需要额外的内存使用费。地址很容易超过 32 字节的限制, 但 EVM 不同。综上所述，必须能够去管理这些可能发生的事件。
-
-存储费用有一个微妙的行为 —— 为最小化存储费用（直接与一个在所有结点上的大型状态数据通信），清除存储的执行费用指令不仅仅免除,而且会返回一些费用；因为初始化的存储费用往往比实际实际使用的多,在清除存储空间后会有返回费用, 这个返回的费用是预先支付的费用中的一部分。
+在三个不同的情况下使用费用（命名为燃料），在这三种情况下，费用都是执行任何操作的必备条件。第一种情况也是最普通的情况就是计算操作费用（详见附录 \ref{app:fees}）。第二种情况，可能因为一个子消息调用或者合约创建而消耗燃料,这是执行 {\small CREATE}, {\small CALL} and {\small CALLCODE} 费用中的一部分。第三种情况，可能因为内存使用的增多而为燃料付费。
+
+对于一个账户的执行，内存总费用和需要的 32 字节最小倍数的内存量成正比, 所有的内存是按在一定范围中被包含的所有记忆索引（无论是读还是写）下要求的32位字节的最小倍数成比例的。这个为了 just-in-time (JIT) 的基础而去支付的；访问一个大于索引量 32 字节的地址,就会需要额外的内存使用费。地址很容易超过 32 字节的限制, 但 EVM 不同。综上所述，必须能够去管理这些可能发生的事件。
+
+存储费用有一个微妙的行为 —— 为最小化存储费用（直接与一个在所有结点上的大型状态数据通信），清除存储的执行费用指令不仅仅免除,而且会返回一些费用；因为初始化的存储费用往往比实际使用的多,在清除存储空间后会有返回费用, 这个返回的费用是预先支付的费用中的一部分。
 
@@
 \begin{itemize}
-\item $I_a$, 拥有正在执行的代码的账户地址。
+\item $I_a$, 拥有正在执行代码的账户地址。
 \item $I_o$, 发起这次交易的发送者地址。
@@
 
-我们现在必须去定义函数 $\Xi$。在大多实际执行过程中, 将会整个系统状态 $\boldsymbol{\sigma}$ 和机器状态 $\boldsymbol{\mu}$ 的迭代过程建模。我们定义一个递归函数 $X$, 它使用这个迭代函数 $O$(定义状态机中单循环的结果), 定义函数 $Z$ 用来表示当前状态是一个异常终止状态, 定义函数 $H$ 表示当前状态是正常终止时得到的结果数据。
+我们现在必须去定义函数 $\Xi$。在大多实际执行过程中, 将整个系统状态 $\boldsymbol{\sigma}$ 和机器状态 $\boldsymbol{\mu}$ 的迭代过程建模。我们定义一个递归函数 $X$, 它使用这个迭代函数 $O$(定义状态机中单循环的结果), 定义函数 $Z$ 用来表示当前状态是一个异常终止状态, 定义函数 $H$ 表示当前状态是正常终止时得到的结果数据。
 
@@
 
-需要注意的是, 我们在推导 $\Xi$ 时, 我们去掉了第4个元素 $I'$ 并从结果状态 $\boldsymbol{\mu}'$ 中提取了剩余的燃料 $\boldsymbol{\mu}'_g$
+需要注意的是,在推导 $\Xi$ 时, 我们去掉了第4个元素 $I'$ 并从结果状态 $\boldsymbol{\mu}'$ 中提取了剩余的燃料 $\boldsymbol{\mu}'_g$
 
@@
 
-为了提高可读性, 使用小型大写字母(例如 {\small ADD})的指令助记符都应该看作对应的数字方程式, 详细的指标表及定义见附录 \ref{app:vm} 。
+为了提高可读性, 使用大写字母(例如 {\small ADD})的指令助记符都应该看作对应的数字方程式, 详细的指标表及定义见附录 \ref{app:vm} 。
 
@@
 
-我们使用 $\mathbf{\delta}$ 表示栈操作移除的元素数, 使用 $\mathbf{\alpha}$ 表示栈操作添加的元素数, 可以使用指令作为  $\mathbf{\delta}$ 、 $\mathbf{\alpha}$ 和指令费用函数 $C$ 的小标, $C$ 表示执行指定指令消耗的燃料。
+我们使用 $\mathbf{\delta}$ 表示栈操作移除的元素数, 使用 $\mathbf{\alpha}$ 表示栈操作添加的元素数, 可以使用指令作为  $\mathbf{\delta}$ 、 $\mathbf{\alpha}$ 和指令费用函数 $C$ 的下标, $C$ 表示执行指定指令消耗的燃料。
 
@@
 
-这些跳转地址都必须在有效指令边界内，并且需要在显示定义在代码中,而不是基于 {\small PUSH} 指令操作的数据。（也不是使用隐含定义的 {\small STOP} 指令去跟踪它）。
+这些跳转地址都必须在有效指令边界内，并且需要在显式定义在代码中,而不是基于 {\small PUSH} 指令操作的数据。（也不是使用隐含定义的 {\small STOP} 指令去跟踪它）。
 
@@
 
-权威的区块链是覆盖整个区块树的从根节点到叶子节点的路径。为了形成是路径共识，我们通过最大的计算量，或者说是最重的路径来识别它。识别最重路径的一个有效的因素是叶子节点的数量，它也是路径中区块的数量，不包含没有进行挖框的创世块(译者注:创始块通过硬编码写在了代码中及区块链中)。路径越长，到达最后一个叶子节点的代价越大。这和已有的比特币及其衍生协议类似。
+权威的区块链是覆盖整个区块树的从根节点到叶子节点的路径。为了形成路径共识，我们通过最大的计算量，或者说是最重的路径来识别它。识别最重路径的一个有效的因素是叶子节点的数量，它也是路径中区块的数量，不包含没有进行挖矿的创世块(译者注:创始块通过硬编码写在了代码中及区块链中)。路径越长，到达最后一个叶子节点的代价越大。这和已有的比特币及其衍生协议类似。
 
@@
 区块定稿涉及到 4 个步骤:
-The process of finalising a block involves four stages:
 
@@
 
-验证叔链头意味着需要验证每个叔链头的有效性并且是当前区块父块的第 $N$ 代, $N$ ≤ 6, 最多只能有2个叔链头, 公式如下:
+验证叔链头意味着需要验证每个叔链头的有效性并且是当前区块父块的第 $N$ 代, $N \leq 6$, 最多只能有 2 个叔链头, 公式如下:
 \begin{equation}
@@
 
-我们定义区块的奖励会 5 以太币:
+我们定义区块的奖励为 5 以太币:
 \begin{equation}
@@
 我们定义函数 $\Gamma$, 映射到区块 $B$ 的初始状态:
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
@@
 
-处于以上原因，工作量证明函数有两个重要的目标：首先，它应该是尽可能的被更多人去接受。对定制的特别硬件的需求或者回报，应该被减到最小。 这使得分配模型尽可能开放，理想情况是，通过挖矿消耗电力获得以太币, 在全世界各个地方都是一样的比例。
-
-第二，应该不允许获得超线性的收益，尤其是在有一个非常高的初始障碍条件下。如果允许这样, 一个资金充足的恶意者可以获得引起麻烦的网络挖矿算力, 并允许给他们获得超线性的汇报（按他们的医院改变收益分布）,并且会弱化网络的安全性。
-
-比特币的世界中一个灾难是 ASICs。有一些计算硬件仅仅是为了做一个简单的任务而存在。在比特币的案例中，这个任务就是 SHA256 哈希函数。当 ASICs 为了工作量证明函数而存在时，两个的目标会变得危险。因此，一个可抵抗 ASIC 的工作量证明函数（比如难以在专用硬件上执行,或者在专用硬件执行时并不划算）可以作为众所周知的银弹。
+出于以上原因，工作量证明函数有两个重要的目标：首先，它应该尽可能的被更多人去接受。对定制特别硬件的需求或者回报，应该被减到最小。 这使得分配模型尽可能开放，理想情况是，通过挖矿消耗电力获得以太币, 在全世界各个地方都是一样的比例。
+
+第二，应该不允许获得超线性的收益，尤其是在有一个非常高的初始障碍条件下。如果允许这样, 一个资金充足的恶意者可以获得引起麻烦的网络挖矿算力, 并允许给他们获得超线性的汇报（按他们的意愿改变收益分布）,并且会弱化网络的安全性。
+
+比特币世界中一个灾难是 ASICs。有一些计算硬件仅仅是为了做一个简单的任务而存在。在比特币的案例中，这个任务就是 SHA256 哈希函数。当 ASICs 为了工作量证明函数而存在时，两个的目标都会变得危险。因此，一个可抵抗 ASIC 的工作量证明函数（比如难以在专用硬件上执行,或者在专用硬件执行时并不划算）可以作为众所周知的银弹。
 
@@
 Ethash 是以太坊 1.0 的 PoW 算法。它是 Dagger-Hashimoto 的最后版本，Buterin
- [2013b] 和 Dryja[2014] 介绍过。再被它为 Dagger-Hashimoto 已不大合适, 因为很多算法的最初特性已经被大量地修改。这个算法的概况如下：
-
-通过扫描区块头可以计算一个种子。根据种子，可以计算一个初始大小为 $J_{cacheinit}$ 字节位随机数缓存。轻节点客户端可以保存这个缓存。根据缓冲，我们能生成一个初始化为 $J_{datasetinit}$ 字节的数据集合，数据集合中每个元素依赖于缓冲中的一小部分数据。全节点客户和矿工保存着数据集合。数据集合随时间线性增长。
+ [2013b] 和 Dryja[2014] 介绍过。再称它为 Dagger-Hashimoto 已不大合适, 因为很多算法的最初特性已经被大量地修改。这个算法的概况如下：
+
+通过扫描区块头得到一个种子。根据种子可以计算一个初始大小为 $J_{cacheinit}$ 字节位随机数缓存。轻节点客户端可以保存这个缓存。根据缓存，我们能生成一个初始化为 $J_{datasetinit}$ 字节的数据集合，数据集合中每个元素依赖于缓存中的一小部分数据。全节点客户和矿工保存着数据集合。数据集合随时间线性增长。
 
@@
 \section{执行合约}
-有一些特别有用的合约模式；其中会讨论其中两个,分别是数据订阅和随机数。
+有一些特别有用的合约模式；我们会讨论其中两个,分别是数据订阅和随机数。
 
 \subsection{数据订阅}
-一个数据订阅合约提供简单的服务：它允许外部的信息可以进入以太坊系统内。以太坊系统不会保证这个信息的精确度和及时性，这是第二方合约作者的任务, 第二方合约作者使用数据订阅并决定能否信任单个数据订阅服务。
-
-通常的模式是,通过消息调用触发单个以太坊中合约，得到一个关于外部现象的信息回应。一个例子可能是纽约城市的本地温度。这个会作为合约执行并返回在存储中的一些值。当然存储中的这些值必须比如温度需要维护以保证准确, 这样模式中的第二方才可以将其作为外部服务来运行以太坊节点,当发现一个新区块时立即创建一个新的有效交易发送到合约并更新存贮中的值。合约代码将仅仅接受包含指定服务的数据更新。
+一个数据订阅合约提供简单的服务：它允许外部的信息进入以太坊系统内。以太坊系统不会保证这个信息的精确度和及时性，这是第二方合约作者的任务, 第二方合约作者使用数据订阅并决定能否信任单个数据订阅服务。
+
+通常的模式是,通过消息调用触发单个以太坊中合约，得到一个关于外部现象的信息回应。一个例子可能是纽约当地温度。这个会作为合约执行并返回存储中的一些值。当然存储中的这些值比如温度需要维护以保证准确,这样模式中的第二方才可以将其作为外部服务来运行以太坊节点,当发现一个新区块时立即创建一个新的有效交易发送到合约并更新存储中的值。合约代码将仅仅接受包含指定服务的数据更新。
 
@@
 \section{可用性}
-本文的英文源文件维护在 \url{https://github.com/ethereum/yellowpaper}. 一个自动生成的英文 PDF 格式文件见 \url{https://ethereum.github.io/yellowpaper/paper.pdf}。
-
-本文的中文源文件维护在 \url{https://github.com/yuange1024/ethereum_yellowpaper}. 一个自动生成的中文 PDF 格式文件见 \url{https://github.com/yuange1024/ethereum_yellowpaper/blob/master/Paper_Chinese.pdf}。 (本段由译者增加)
+本文的英文源文件维护在 \url{https://github.com/ethereum/yellowpaper}。 一个自动生成的英文 PDF 格式文件见 \url{https://ethereum.github.io/yellowpaper/paper.pdf}。
+
+本文的中文源文件维护在 \url{https://github.com/yuange1024/ethereum_yellowpaper}。一个自动生成的中文 PDF 格式文件见 \url{https://github.com/yuange1024/ethereum_yellowpaper/blob/master/Paper_Chinese.pdf}。 (本段由译者增加)
 

### 2017-09-19 Paper.tex
%\usepackage{hyperref}
This does need to be added, but commenting out for now as I'm not sure how. Will look into it later.
prepend # with \ for Hyperref package

--- 2017-09-16 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
 The block header contains several pieces of information:

### 2017-09-19 Paper.tex
Footnote to explain why a 24576 byte limit exists for contract code o

--- 2017-09-19 Paper.tex
@@
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big\text{\footnotemark})
 \end{align}
+\footnotetext{The 24576 byte limit for output contract code data $\mathbf{o}$ exists because a contract creation call can trigger O(n) cost in terms of reading the code from disk, preprocessing the code for VM execution, and also adding O(n) data to the Merkle proof for the block's proof-of-validity.
+With higher gas limits that can be caused by dynamic gas limit rules, this is a greater concern, and is especially inconvenient with light clients verifying proofs of validity or invalidity.}
 

### 2017-09-19 Paper.tex
Moving footnote to body text for 24576 byte limit
To avoid complexity with adding a footnote in a math environment.

--- 2017-09-19 Paper.tex
@@
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big
 \end{align}
@@
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Note that the 24576 byte limit for $\mathbf{o}$ exists because a contract creation call can trigger O(n) cost in terms of reading the code from disk, preprocessing the code for VM execution, and also adding O(n) data to the Merkle proof for the block's proof-of-validity.
+With higher gas limits that can be caused by dynamic gas limit rules, this is a greater concern, and is especially inconvenient with light clients verifying proofs of validity or invalidity.
+
+Note that the intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
 

### 2017-09-19 Paper.tex
Grammar fixes

@@
 
-Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
+Over an account's execution, the total fee for memory-usage payable is proportional to the smallest multiple of 32 bytes that is required such that all memory indices (whether for read or write) are included in the range.
 This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
@@
 Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
+For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from the stack, as defined in section \ref{ch:model}.
 

### 2017-09-19 Paper.tex
Fill the determined gas costs

--- 2017-09-19 Paper.tex
@@
 \usepackage{url}
-\usepackage{hyperref}
 \usepackage{cancel}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big\text{\footnotemark})
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
-\footnotetext{The 24576 byte limit for output contract code data $\mathbf{o}$ exists because a contract creation call can trigger O(n) cost in terms of reading the code from disk, preprocessing the code for VM execution, and also adding O(n) data to the Merkle proof for the block's proof-of-validity.
-With higher gas limits that can be caused by dynamic gas limit rules, this is a greater concern, and is especially inconvenient with light clients verifying proofs of validity or invalidity.}
 

### 2017-09-19 Paper.tex

--- 2017-09-19 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**}<c \wedge H_i<\firsthomesteadblock \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big
+\end{cases}
 \end{align}
@@
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-Note that the 24576 byte limit for $\mathbf{o}$ exists because a contract creation call can trigger O(n) cost in terms of reading the code from disk, preprocessing the code for VM execution, and also adding O(n) data to the Merkle proof for the block's proof-of-validity.
-With higher gas limits that can be caused by dynamic gas limit rules, this is a greater concern, and is especially inconvenient with light clients verifying proofs of validity or invalidity.
-
-Note that the intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+
+Note that the intention from block \firsthomesteadblock\ onwards ({\it Homestead}) is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+Before {\it Homestead}, if there is not enough gas to pay $c$, an account at the new contract's address is created, along with all the initialisation side-effects, and the value is transferred, but no contract code is deployed.
 
@@
 
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+g_r&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r &=& 500\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r &=& 40000\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+\end{eqnarray}
+
 \section{Signing Transactions}\label{app:signing}

### 2017-09-21 Paper.tex
Accidentally omitted ) after \big, readded
Changed ECDSARESTORE to ECDSARECOVER
Formula 210 has ECDSARECOVER. For consistency, I think ECDSARESTORE should be changed to ECDSARECOVER. CTRL+F ECDSARECOVER returns 5 results, while ECDSARESTORE only returns one, where it is mentioned that it is asserted with "We assert the functions ECDSASIGN, ECDSARESTORE and ECDSAPUBKEY", but then ECDSASIGN, ECDSARECOVER and ECDSAPUBKEY are asserted.
l. 1769, reworded for clarification

--- 2017-09-19 Paper.tex
@@
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature.
 \begin{eqnarray}
@@
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash \\
 &&&& in the block following it. \\

### 2017-09-21 Paper.tex
fix small grammatical typo

@@
 The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-Note that the 24576 byte limit for $\mathbf{o}$ exists because a contract creation call can trigger O(n) cost in terms of reading the code from disk, preprocessing the code for VM execution, and also adding O(n) data to the Merkle proof for the block's proof-of-validity.
-With higher gas limits that can be caused by dynamic gas limit rules, this is a greater concern, and is especially inconvenient with light clients verifying proofs of validity or invalidity.
-
-Note that the intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+
+Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
 
@@
 
-Over an account's execution, the total fee for memory-usage payable is proportional to the smallest multiple of 32 bytes that is required such that all memory indices (whether for read or write) are included in the range.
+Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
 This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
@@
 
-We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
@@
 Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from the stack, as defined in section \ref{ch:model}.
+For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
 
@@
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
 &&&& in the block following it. \\

### 2017-09-22 Paper.tex
Opcode 0x00 cross-refs

--- 2017-09-21 Paper.tex
@@
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not the least being Bitcoin.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 
-%E language?
+%E language/glossary?
 
@@
 Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
+Blocks function as a journal or ledger, recording a series of transactions together with the previous block and an identifier for the final state (though blocks do not store the final state itself---that would be far too big).
 They also punctuate the transaction series with incentives for nodes to \textit{mine}.
@@
 
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+The empty sequence, denoted ,\hyperref[emptySequence]{''$()$''}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
 \begin{eqnarray}
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+() & \text{if} \quad w \in \{ \text{\hyperref[STOP]{''\small STOP''}}, \text{\small SELFDESTRUCT} \} \\
 \varnothing & \text{otherwise}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution, see \label{STOP} and \label{emptySequence}. \\
 \midrule

### 2017-09-22 Paper.tex
Moved \usepackage{hyperref} to last in loaded packages
As per (here)[https://mirror.aarnet.edu.au/pub/CTAN/macros/latex/contrib/hyperref/doc/manual.html#TBL-24], section 2.

--- 2017-09-21 Paper.tex
@@
 \usepackage{tikz}
+\usepackage{hyperref}
 

### 2017-09-22 Paper.tex
Signing transactions para. 1 citations

--- 2017-09-22 Paper.tex
@@
 \end{description}
-
-Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
-For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
 
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
+This method utilises the SECP-256k1 curve as described by \cite{secp256k1BitcoinWiki2016} and \cite{secp256k1StackExchange2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.
+9 of 15, para.
+3.
 

### 2017-09-22 Paper.tex
ECDSA function citations
Added footnote for ECDSAcerticom

@@
 
-We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$.
+These are formally defined in the literature, \eg by \cite{ECDSAcerticom} and \cite{ECDSAWikipedia}.
+\footnote{For the former citation, refer to section 6.2 for ECDSAPUBKEY, and section 7 for ECDSASIGN and ECDSARECOVER (assuming that the former is for signature generation and the latter is for signature verification).}
 \begin{eqnarray}

### 2017-09-22 Paper.tex
Instruction set: Grammar; 50s: the first word/byte
Prepended 'the' before countable definite articles, 'an' before indefinite articles, rewording to avoid stringing too many nouns in succession, etc. Added the top/first word/byte, e.g. instead of "Save word to memory", changed to "Save the first word to memory" (implying the first word of the stack as per the following formula.

@@
 \midrule
-0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
+0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
 &&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
+0x15 & {\small ISZERO} & 1 & 1 & This is the logical negation operation, also called the logical complement or the NOT operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
+0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+&&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
 \bottomrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
+0x20 & {\small SHA3} & 2 & 1 & Compute a Keccak-256 hash. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
+0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
+0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty \\
+&&&& This is the sender of the original transaction; it is never an account with non-empty \\
 &&&& associated code. \\
 \midrule
-0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
+0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
@@
 \midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
+0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
 \midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
+0x35 & {\small CALLDATALOAD} & 1 & 1 & Get the input data of the current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
@@
 \midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get the size of the input data in the current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
@@
 \midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
+0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy the input data in the current environment to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
@@
 \midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
+0x38 & {\small CODESIZE} & 0 & 1 & Get the size of code running in the current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
 \midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
+0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in the current environment to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
@@
 \midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
+0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
-&&&& This is gas price specified by the originating transaction.\\
-\midrule
-0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
+&&&& This is the gas price specified by the originating transaction.\\
+\midrule
+0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
-\midrule
-0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
+0x50 & {\small POP} & 1 & 0 & Remove the top, first item from the stack. \\
+\midrule
+0x51 & {\small MLOAD} & 1 & 1 & Load the first word from memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
@@
 \midrule
-0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
+0x52 & {\small MSTORE} & 2 & 0 & Save the first word (which is the first item in the stack) to memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
@@
 \midrule
-0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
+0x53 & {\small MSTORE8} & 2 & 0 & Save the first byte to memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
@@
 \midrule
-0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
+0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
 &&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
@@
 0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
-&&&& This operation has no effect on machine state during execution. \\
+&&&& This operation has no effect on the machine state during execution. \\
 \bottomrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
+0x60 & {\small PUSH1} & 0 & 1 & Place a 1 byte item on the stack. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
@@
 &&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
-&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
+&&&& The byte is right-aligned (\ie it takes the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
+\midrule
+0x61 & {\small PUSH2} & 0 & 1 & Place a 2-byte item on the stack. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
 \midrule
@@
 &&&& where $\boldsymbol{c}$ is defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
 \bottomrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
+0x80 & {\small DUP1} & 1 & 2 & Duplicate the 1st stack item. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
-0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
+0x81 & {\small DUP2} & 2 & 3 & Duplicate the 2nd stack item. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
@@
 \midrule
-0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
+0x8f & {\small DUP16} & 16 & 17 & Duplicate the 16th stack item. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\

### 2017-09-22 Paper.tex
Instruction set grammar
Complete sentences, NOT operation

@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution, see \label{STOP} and \label{emptySequence}. \\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
+0x00 & {\small STOP} & 0 & 0 & This operation halts execution, see \label{STOP} and \label{emptySequence} for further information. \\
+\midrule
+0x01 & {\small ADD} & 2 & 1 & This is the addition operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
-0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
+0x02 & {\small MUL} & 2 & 1 & This is the multiplication operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
-0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
+0x03 & {\small SUB} & 2 & 1 & This is the subtraction operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
-0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
+0x04 & {\small DIV} & 2 & 1 & This is the integer division operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
+0x05 & {\small SDIV} & 2 & 1 & This is the signed integer division operation (truncated). \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
+0x06 & {\small MOD} & 2 & 1 & This is the modulo remainder operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
-0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
+0x07 & {\small SMOD} & 2 & 1 & This is the signed modulo remainder operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
+0x08 & {\small ADDMOD} & 3 & 1 & This is the modulo addition operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
+0x09 & {\small MULMOD} & 3 & 1 & This is the modulo multiplication operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
+0x0a & {\small EXP} & 2 & 1 & This is the exponential operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
+0x10 & {\small LT} & 2 & 1 & This is the less-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
+0x11 & {\small GT} & 2 & 1 & This is the greater-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
+0x12 & {\small SLT} & 2 & 1 & This is the signed less-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
+0x13 & {\small SGT} & 2 & 1 & This is the signed greater-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
+0x14 & {\small EQ} & 2 & 1 & This is the equality comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
+0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
-0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
+0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
-0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
+0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
-0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
+0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
+0x90 & {\small SWAP1} & 2 & 2 & Exchange the 1st and the 2nd stack items. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
@@
 \midrule
-0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
+0x91 & {\small SWAP2} & 3 & 3 & Exchange the 1st and the 3rd stack items. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 \midrule
-0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
+0x9f & {\small SWAP16} & 17 & 17 & Exchange the 1st and the 17th stack items. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
+0xa0 & {\small LOG0} & 2 & 0 & Append the log record with no topics. \\
 &&&& $\mathbf{t} \equiv ()$ \\
 \midrule
-0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
+0xa1 & {\small LOG1} & 3 & 0 & Append the log record with one topic. \\
 &&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 \midrule
-0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
+0xa4 & {\small LOG4} & 6 & 0 & Append the log record with four topics. \\
 &&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
 &&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (the balance of the caller is  \\
+&&&& too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the  \\
+&&&& newly created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\

### 2017-09-23 Paper.tex
Add \usepackage{underscore}
To prevent builds failing that have underscores in URLs.
Fee Schedule grammar

@@
 \usepackage{tikz}
+\usepackage{underscore}
 \usepackage{hyperref}
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{zero}$ & 0 & Nothing is paid for operations of the set {\small $W_{zero}$}. \\
+$G_{base}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{low}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{mid}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{high}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{extcode}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
+$G_{balance}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 200 & This is paid for an {\small SLOAD} operation. \\
+$G_{jumpdest}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
+$G_{sset}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \\
+$R_{selfdestruct}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & This is the amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{create}$ & 32000 & This is paid for a {\small CREATE} operation. \\
+$G_{codedeposit}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
+$G_{call}$ & 700 & This is paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{callstipend}$ & 2300 & This is a stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+$G_{newaccount}$ & 25000 & This is paid for a {\small CALL} or for a {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{exp}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & This is paid for every additional word when expanding memory. \\
+$G_\text{txcreate}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_{txdatazero}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & This is paid for every transaction. \\
+$G_{log}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
+$G_{sha3}$ & 30 & This is paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{copy}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
+$G_{blockhash}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\
 

### 2017-09-23 Paper.tex
Ref to terminology
\usepackage[hyphens]{url}
To get URLs to break at the end of a column.
Appendix A ref.
Adding cross-refs

@@
 \documentclass[9pt,oneside]{amsart}
-%\usepackage{tweaklist}
-\usepackage{url}
 \usepackage{cancel}
@@
 \usepackage{hyperref}
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
+% "hyperref loads the url package internally.
+Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
+This avoids any package option clashes." Source: https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034.
+% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
+Try loading the url package earlier than the package that requires it.
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034
+% \usepackage{tweaklist}
+
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+%\PassOptionsToPackage{url}\g@addto@macro{\UrlBreaks}{\UrlOrds}
+% info on the above command: https://tex.stackexchange.com/posts/10419/revisions
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
 
@@
 %E language/glossary?
+
+For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the \hyperref[itm:nonce]{nonce} of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 
-\subsection{The Transaction} \label{ch:transaction}
+\subsection{The Transaction} \label{subsec:transaction}
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
 \item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
@@
 
-The empty sequence, denoted ,\hyperref[emptySequence]{''$()$''}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\hypertarget{emptySequence}
+The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
 \begin{eqnarray}
@@
 
-\subsubsection{Normal Halting}
+\subsubsection{Normal Halting} \label{subsec:Normal Halting}
 
 The normal halting function $H$ is defined:
-\begin{equation}
+\begin{equation} \label{eq:Normal Halting}
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\hyperref[STOP]{''\small STOP''}}, \text{\small SELFDESTRUCT} \} \\
+() & \text{if} \quad w \in \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
 \varnothing & \text{otherwise}
@@
 The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+Note also the difference between the empty sequence and the empty set as discussed \hyperlink{emptySequence}{here}.
 
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & This operation halts execution, see \label{STOP} and \label{emptySequence} for further information. \\
+0x00 & {\small STOP} & 0 & 0 & This operation halts execution, outputting the empty sequence as per equation \ref{eq:Normal Halting}.
+\\
 \midrule

### 2017-09-24 Paper.tex
nonce hyperlinks (block, account, transaction)

@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the \hyperref[itm:nonce]{nonce} of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the \hyperlink{transaction nonce}{nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
 Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 \begin{description}
+\hypertarget{account nonce}{}
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
@@
 \begin{description}
+\hypertarget{transaction nonce}{}
 \item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
@@
 \item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\hypertarget{block nonce}{}
 \item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
@@
 
-The nonce, $H_n$, must satisfy the relations:
+The \hyperlink{block nonce}{nonce of a block}, $H_n$, must satisfy the relations:
 \begin{equation}
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+Because the \hyperlink{block nonce}{block nonce} must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
 
@@
 \item the transaction signature is valid;
-\item the transaction nonce is valid (equivalent to the sender account's current nonce);
+\item the \hyperlink{transaction nonce}{transaction nonce} is valid (equivalent to the  \hyperlink{account nonce}{sender account's current nonce});
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
 The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
 Thus we define the resultant address for the new account $a$:
@@
 where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+Note we use one fewer than the \hyperlink{account nonce}{sender's nonce value}; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 
-Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and nonce.
+Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and \hyperlink{account nonce}{nonce}.
 Thus the previous equation should be taken to mean:
@@
 
-\hypertarget{emptySequence}
+\hypertarget{emptySequence}{}
 The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and nonce.
+\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{nonce}.
 \end{enumerate}
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the \hyperlink{block nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash components; $H_n$ is the \hyperlink{block nonce}{nonce} of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the \hyperlink{block nonce}{nonce}.
 $\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
@@
 
-If the output of this algorithm is below the desired target, then the nonce is valid.
+If the output of this algorithm is below the desired target, then the \hyperlink{block nonce}{nonce} is valid.
 Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
@@
 \end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+With $H_{\hcancel{n}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_c$ is obtained as follows:
 \begin{equation}
@@
 \end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
 \begin{equation}

### 2017-09-24 Paper.tex
Fixed overfull lines in tables with tabularx.

@@
 \usepackage{cancel}
+%\usepackage{tweaklist}
 \usepackage{xspace}
@@
 \usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
 \usepackage{hyperref}
+
+%\usepackage{longtable}
+% https://stackoverflow.com/a/1918465/7438857
+%tabularx seems better
+
+\usepackage{tabularx}
+%https://stackoverflow.com/a/1919447/7438857
+
 \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
@@
 Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034.
+This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
 % Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
 Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034
-% \usepackage{tweaklist}
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
 
@@
 %\PassOptionsToPackage{url}\g@addto@macro{\UrlBreaks}{\UrlOrds}
-% info on the above command: https://tex.stackexchange.com/posts/10419/revisions
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error.
 
@@
   \do\Y\do\Z}
+
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
 
@@
 
+\newpage
 \section{Fee Schedule}\label{app:fees}
@@
 
-\begin{tabular*}{\columnwidth}[h]{lrl}
+\begin{tabularx}{\textwidth}{lrX}
 \toprule
@@
 \bottomrule
-\end{tabular*}
+\end{tabularx}
 
@@
 
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+%\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of the original transaction; it is never an account with non-empty \\
-&&&& associated code. \\
+&&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
+0 is left on the stack if the looked for block number is greater than the current block number or more than 256 blocks behind the current block.
+\\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
 \midrule
@@
 \midrule
-0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
-&&&&  corresponding to this instruction. \\
+0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment corresponding to this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
@@
 \midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
-&&&& for the cost of this instruction. \\
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction for the cost of this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
@@
 &&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (the balance of the caller is  \\
-&&&& too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the  \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the  \\
 &&&& newly created account. \\
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persistingthe current values for {\it sender} and {\it value}.
+Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
+The omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$.
+Otherwise it is exactly equivalent to {\small CALL} except: \\
 &&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
+This means that the recipient is in fact the same account as at present, simplythat the code is overwritten {\it and} the context is almost entirely identical.\\
 \midrule
@@
 
-\begin{tabular*}{\columnwidth}[h]{lrl}
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule

### 2017-09-26 Paper.tex
Adding anchors and links plus minor edits

@@
 %\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage{hyperref}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [3 3]}]{hyperref}
+%,pagebackref=true
 
@@
 Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-The following command did fix this error.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
 
@@
   \do\Y\do\Z}
-
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
 %\sloppy
@@
 A valid state transition is one which comes about through a transaction.
-Formally:
+Formally:\hypertarget{Upsilon}{}
 \begin{equation}
@@
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case greek letter, \eg \hyperlink{Upsilon}{$\Upsilon$}, the Ethereum state transition function.
 
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \hypertarget{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other \hypertarget{ommerheaders}{block headers $\mathbf{U}$} that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, as it is not a well-known word) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{H e}{}
 \item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
@@
 
-\subsubsection{Transaction Receipt}
+\hypertarget{Transaction Receipt}{\subsubsection{Transaction Receipt}}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as \hyperlink{H e}{$H_e$}.
+
+The transaction receipt is a tuple of four items \hypertarget{Rsigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{Ru}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$}, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_b$}:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, \hyperlink{R b}{R_b}, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, \hyperlink{R b}{R_b}, R_\mathbf{l})
 \end{equation}
@@
 
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_b$}, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_b} \in \mathbb{B}_{256}
 \end{equation}
@@
 
-Thus we are able to define the block header validity function $V(H)$:
+Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
 \begin{eqnarray}
@@
 
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
+The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon}{$\Upsilon$}.
 It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
@@
 
-Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
+Formally, we consider the function \hyperlink{Upsilon}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
 \begin{equation}
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
+We also define \hyperlink{Upsilong}{$\Upsilon^g$} to evaluate to the amount of gas used in the execution of a transaction and \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} to evaluate to the transaction's accrued log items, both to be formally defined later.
 
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+T_g & \leqslant & {B_H}_l - \hyperlink{ell}{\ell}\hyperlink{Ru}{(\mathbf{R})_u}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
 
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\hypertarget{Upsilong}{And finally, we specify $\Upsilon^g$, the total gas used in this transaction}\hypertarget{Upsilonl}{ and $\Upsilon^\mathbf{l}$, the logs created by this transaction:}
 \begin{eqnarray}
@@
 
-These are used to help define the transaction receipt, discussed later.
+These are used to help define the \hyperlink{Transaction Receipt}{transaction receipt}.
 
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+If the initialisation execution ends with a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operation, \hyperlink{RETURN}{\text{\small RETURN}}, has a special function $H_{\text{\tiny RETURN}}$.
 Note also the difference between the empty sequence and the empty set as discussed \hyperlink{emptySequence}{here}.
@@
 \quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)  \\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\hypertarget{JUMP}{} \hypertarget{JUMPI}{} \label{eq:JUMP} \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
 J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
@@
 
-The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
+The validation of \hyperlink{ommerheaders}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
 The maximum of ommer headers is two.
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+{B_H}_g = \hyperlink{ell}{\ell}\hyperlink{Ru}{(\mathbf{R})_u}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+We raise the block's beneficiary account by \hyperlink{R b}{$R_b$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+\hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
 \begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})\hyperlink{R b}{R_b} \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) \hyperlink{R b}{R_b}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_b = 5 \times 10^{18}
+\text{Let} \quad \hyperlink{R b}{R_b} = 5 \times 10^{18}
 \end{equation}
@@
 
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
 \begin{equation}
 \mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
-\end{equation}
-
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
+\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{Rsigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
+\end{equation}
+
+Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
 
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_b}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-
-Formally, we define $R_b$:
+\hypertarget{R b}{
+Formally, we define $R_b$:}
 \begin{eqnarray}
@@
 These are formally defined in the literature, \eg by \cite{ECDSAcerticom} and \cite{ECDSAWikipedia}.
-\footnote{For the former citation, refer to section 6.2 for ECDSAPUBKEY, and section 7 for ECDSASIGN and ECDSARECOVER (assuming that the former is for signature generation and the latter is for signature verification).}
+\footnote{For the former citation, refer to section 6.2 for ECDSAPUBKEY, and section 7 for ECDSASIGN and ECDSARECOVER.} %(assuming that the former is for signature generation and the latter is for signature verification)
 \begin{eqnarray}
@@
 $G_{sreset}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & This is the amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
+$R_{selfdestruct}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
+$G_{selfdestruct}$ & 5000 & This is the amount of gas to pay for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & This is paid for a {\small CREATE} operation. \\
@@
 $G_{callvalue}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & This is a stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & This is paid for a {\small CALL} or for a {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callstipend}$ & 2300 & This is a stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
+$G_{newaccount}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+with \hyperlink{tiny CALL}{$C_\text{\tiny CALL}$}, \hyperlink{C tiny SELFDESTRUCT}{$C_\text{\tiny SELFDESTRUCT}$} and \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$} as specified in the appropriate section below.
+We define the following subsets of instructions:
 
@@
 0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
 \midrule
@@
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
 &&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
@@
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
 \midrule
@@
 &&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See section \ref{ch:model}. \\
+&&&& See equation \ref{eq:Normal Halting}. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
@@
 \midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persistingthe current values for {\it sender} and {\it value}.
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
 Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
@@
 &&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
-This means that the recipient is in fact the same account as at present, simplythat the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
+This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{SELFDESTRUCT}{}\\
 \midrule
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise} 
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
+&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{selfdestruct}} + \begin{cases}
+\hyperlink{Gnewaccount}{G_{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}

### 2017-09-26 Paper.tex
Fixing subscripts to /mathbf, cross-referencing.

@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_\mathbf{t}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_\mathbf{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the \hyperlink{transaction nonce}{nonce of said transaction}.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_\mathbf{n}$, denotes the \hyperlink{transaction nonce}{nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_\mathbf{n}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_\mathbf{b}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_\mathbf{s}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_\mathbf{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_\mathbf{c}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_\mathbf{s}
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_\mathbf{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_\mathbf{n}, \boldsymbol{\sigma}[a]_\mathbf{b}, \boldsymbol{\sigma}[a]_\mathbf{s}, \boldsymbol{\sigma}[a]_\mathbf{c}) \big) \big)
 \end{equation}
@@
 \end{equation}
-where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\hypertarget{v x}{where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_\mathbf{n} \in \mathbb{P}_{256} \wedge x_\mathbf{b} \in \mathbb{P}_{256} \wedge x_\mathbf{s} \in \mathbb{B}_{32} \wedge x_\mathbf{c} \in \mathbb{B}_{32}
+}
 \end{equation}
@@
 \begin{description}
-\hypertarget{transaction nonce}{}
-\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\hypertarget{T n}{
+\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_\mathbf{n}$.}\hypertarget{T p}{
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_\mathbf{p}$.}\hypertarget{T g}{
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_\mathbf{g}$.}\hypertarget{T t}{
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathbf{t}$.}\hypertarget{T v}{
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathbf{v}$.}
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hyperlink{T w}{$T_\mathbf{w}$}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+L_\mathbf{T}(T) \equiv \begin{cases}
+(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}, \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) & \text{if} \; T_\mathbf{t} = \varnothing\\
+(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}, \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_\mathbf{n} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_\mathbf{v}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{p} \in \mathbb{P}_{256} & \wedge \\
+T_\mathbf{g} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_\mathbf{w}} \in \mathbb{P}_5 & \wedge & T_\mathbf{r} \in \mathbb{P}_{256} & \wedge \\
+T_\mathbf{s} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{P}_\mathbf{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_\mathbf{t} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{H e}{}
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\hypertarget{block nonce}{}
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\begin{description}\hypertarget{H p}{
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathbf{p}$.}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathbf{o}}$.\hypertarget{H c}{
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_\mathbf{c}$.}
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarget{H d}{
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{H s}{$H_\mathbf{s}$}.\hypertarget{H x}{
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathbf{x}$.}\hypertarget{H m}{
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{m}$.}
+\hypertarget{block nonce}{\hypertarget{H n}{
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{n}$.}}
 \end{description}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as \hyperlink{H e}{$H_e$}.
-
-The transaction receipt is a tuple of four items \hypertarget{Rsigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{Ru}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$}, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_b$}:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, \hyperlink{R b}{R_b}, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as \hyperlink{H e}{$H_\mathbf{e}$}.
+
+The transaction receipt is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_\mathbf{u}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, \hyperlink{R b}{R_b}, R_\mathbf{l})
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathbf{u}}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
 \end{equation}
@@
 
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_b$}, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_b} \in \mathbb{B}_{256}
+We assert $\hyperlink{R u}{R_\mathbf{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_\mathbf{b}$}, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+\hyperlink{R u}{R_\mathbf{u}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_\mathbf{b}} \in \mathbb{B}_{256}
 \end{equation}
@@
 
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+The log entries, $\hyperlink{R l}{R_\mathbf{l}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_\mathbf{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_\mathbf{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_\mathbf{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
-\begin{equation}
-\begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathbf{r}$:
+\begin{equation}
+\begin{array}[t]{lclc}\hypertarget{H r}{
+H_\mathbf{r} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
+H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U})))} & \wedge \hypertarget{H t}{\\
+H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
+H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\})} & \wedge \hypertarget{H b}{\\
+H_\mathbf{b} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_\mathbf{r}
 \end{equation}
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_\mathbf{p}}, \hyperlink{H o}{H_\mathbf{o}}, \hyperlink{H c}{H_\mathbf{c}}, \hyperlink{H r}{H_\mathbf{r}}, \hyperlink{H t}{H_\mathbf{t}}, \hyperlink{H e}{H_\mathbf{e}}, \hyperlink{H b}{H_\mathbf{b}}, \hyperlink{H d}{H_\mathbf{d}},\\ \hyperlink{H i}{H_\mathbf{i}}, \hyperlink{H l}{H_\mathbf{l}}, \hyperlink{H g}{H_\mathbf{g}}, \hyperlink{H s}{H_\mathbf{s}}, \hyperlink{H x}{H_\mathbf{x}}, \hyperlink{H m}{H_\mathbf{m}}, \hyperlink{H n}{H_\mathbf{n}} \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathbf{T}^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_\mathbf{T}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+\hyperlink{H p}{H_\mathbf{p}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_\mathbf{o}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_\mathbf{c}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{H r}{H_\mathbf{r}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_\mathbf{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_\mathbf{e}} \in \mathbb{B}_{32} & \wedge \\
+H_\mathbf{b} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_\mathbf{d}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_\mathbf{i}} \in \mathbb{P} & \wedge \\
+\hyperlink{H l}{H_\mathbf{l}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_\mathbf{g}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_\mathbf{s}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{H x}{H_\mathbf{x}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_\mathbf{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_\mathbf{n}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_\mathbf{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_\mathbf{p}}
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+\hyperlink{H i}{H_\mathbf{i}} \equiv {{P(H)_H}_\mathbf{i}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad \hyperlink{H i}{H_\mathbf{i}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_\mathbf{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
-
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 125000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_\mathbf{d}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_\mathbf{s}} - {P(H)_H}_\mathbf{s}}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_\mathbf{i}} \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}
+
+The canonical gas limit $\hyperlink{H l}{H_\mathbf{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000
+\end{eqnarray}
+
+\hyperlink{H s}{$H_\mathbf{s}$} is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+\hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s}
 \end{equation}
@@
 
-The \hyperlink{block nonce}{nonce of a block}, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
-
+The nonce of a block, $\hyperlink{H n}{H_\mathbf{n}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad \mathbf{m} = \hyperlink{H m}{H_\mathbf{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$.
+
+\hypertarget{h cancel n}{}
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_\mathbf{d}}$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 125000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \wedge m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \\
+& & \hyperlink{H d}{H_\mathbf{d}} = D(H) \quad \wedge \\
+& & \hyperlink{H g}{H_\mathbf{g}} \le \hyperlink{H l}{H_\mathbf{l}}  \quad \wedge \\
+& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000  \quad \wedge \\
+& & \hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s} \quad \wedge \\
+& & \hyperlink{H i}{H_\mathbf{i}} = {P(H)_H}_\mathbf{i} +1 \quad \wedge \\
+& & \lVert \hyperlink{H x}{H_\mathbf{x}} \rVert \le 32
+\end{eqnarray}
+where $(\mathbf{n}, \mathbf{m}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{r})
 \end{equation}
@@
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Finally there is $A_\mathbf{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv T_\mathbf{g} T_\mathbf{p} + \hyperlink{T v}{T_\mathbf{v}}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \hyperlink{ell}{\ell}\hyperlink{Ru}{(\mathbf{R})_u}
+T_\mathbf{n} & = & \boldsymbol{\sigma}[S(T)]_\mathbf{n} \quad \wedge \\
+g_0 & \leqslant & T_\mathbf{g} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_\mathbf{b} \quad \wedge \\
+T_\mathbf{g} & \leqslant & {B_H}_\mathbf{l} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_\mathbf{R})_\mathbf{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_\mathbf{g}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathbf{R})_\mathbf{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_\mathbf{l}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_\mathbf{g}T_\mathbf{p}$.
+The gas available for the proceeding computation, $g$, is defined as $T_\mathbf{g} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\%
+\boldsymbol{\sigma}_0[S(T)]_\mathbf{n} & \equiv & \boldsymbol{\sigma}[S(T)]_\mathbf{n} + 1
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_\mathbf{o}, &\\ \quad\quad g, T_\mathbf{p}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}, 0) & \text{if} \quad T_\mathbf{t} = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_\mathbf{o}, &\\ \quad\quad T_\mathbf{t}, T_\mathbf{t}, g, T_\mathbf{p}, \hyperlink{T v}{T_\mathbf{v}}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_\mathbf{g} - g_0
+\end{equation}
+and $T_\mathbf{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_\mathbf{g} - g'}{2} \Big\rfloor, A_\mathbf{r} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_\mathbf{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_\mathbf{g} - g'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+\boldsymbol{\sigma}^*[S(T)]_\mathbf{b} & \equiv & \boldsymbol{\sigma}_P[S(T)]_\mathbf{b} + g^* T_\mathbf{p} \\
+\boldsymbol{\sigma}^*[m]_\mathbf{b} & \equiv & \boldsymbol{\sigma}_P[m]_\mathbf{b} + (T_\mathbf{g} - g^*) T_\mathbf{p} \\
+m & \equiv & {B_H}_\mathbf{c}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_\mathbf{g} - \mathbf{g}' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_\mathbf{n} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[s]_\mathbf{b} &\equiv& \boldsymbol{\sigma}[s]_\mathbf{b} - v
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_b & \text{otherwise}
+\boldsymbol{\sigma}[a]_\mathbf{b} & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_\mathbf{a} & \equiv & a \\
+I_\mathbf{o} & \equiv & o \\
+I_\mathbf{p} & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
+I_\mathbf{s} & \equiv & s \\
+\hyperlink{I v}{I_\mathbf{v}} & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e
+I_\mathbf{e} & \equiv & e
 \end{eqnarray}
@@
 0 & \text{if} \quad F \\
-g^{**} - c & \text{otherwise} \\
+g^{**} - \mathbf{c} & \text{otherwise} \\
 \end{cases} \\
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_\mathbf{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  \mathbf{g}^{**} < \mathbf{c} \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}_1[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_\mathbf{b} \equiv \boldsymbol{\sigma}[s]_\mathbf{b} - v
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1[s]_\mathbf{b} \equiv \boldsymbol{\sigma}_1'[s]_\mathbf{b} - v
 \end{equation}
@@
 \boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v & \text{otherwise}
 \end{cases}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_\mathbf{c}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_\mathbf{a} & \equiv & r \\
+I_\mathbf{o} & \equiv & o \\
+I_\mathbf{p} & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+I_\mathbf{s} & \equiv & s \\
+\hyperlink{I v}{I_\mathbf{v}} & \equiv & \tilde{v} \\
+I_\mathbf{e} & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_\mathbf{c}
 \end{eqnarray}
@@
 
-\begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\begin{itemize}\hypertarget{I a}{
+\item $I_\mathbf{a}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
+\item $I_\mathbf{o}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
+\item $I_\mathbf{p}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
+\item $I_\mathbf{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
+\item $I_\mathbf{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_H$, the block header of the present block.}\hypertarget{I e}{
+\item $I_\mathbf{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_\mathbf{g}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
+\boldsymbol{\mu}_\mathbf{g} & \equiv & \mathbf{g} \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{i} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & ()
@@
 
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_\mathbf{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_\mathbf{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
@@
 \boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\mu}'_\mathbf{i} & \equiv & \boldsymbol{\mu}_\mathbf{i} \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_t & \equiv & B'_t + B_d \\
+B_\mathbf{t} & \equiv & B'_\mathbf{t} + B_\mathbf{d} \\
 B' & \equiv & P(B_H)
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
+As such given a block $B$, $B_\mathbf{t}$ is its total difficulty, $B'$ is its parent block and $B_\mathbf{d}$ is its difficulty.
 
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \hyperlink{ell}{\ell}\hyperlink{Ru}{(\mathbf{R})_u}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_\mathbf{g}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_\mathbf{g} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathbf{u}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R b}{$R_b$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by \hyperlink{R b}{$R_\mathbf{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 \hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})\hyperlink{R b}{R_b} \\
+\boldsymbol{\sigma}'[{B_H}_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[{B_H}_\mathbf{c}]_\mathbf{b} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})\hyperlink{R b}{R_\mathbf{b}} \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) \hyperlink{R b}{R_b}
+ \boldsymbol{\sigma}'[U_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[U_\mathbf{c}]_\mathbf{b} + (1 + \frac{1}{8} (U_\mathbf{i} - {B_H}_\mathbf{i})) \hyperlink{R b}{R_\mathbf{b}}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad \hyperlink{R b}{R_b} = 5 \times 10^{18}
+\text{Let} \quad \hyperlink{R b}{R_\mathbf{b}} = 5 \times 10^{18}
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
 \end{cases}
@@
 
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\hypertarget{Phi}{
+And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+B'\mathbf{n} & = & \mathbf{n}: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
+B'_\mathbf{m} & = & \mathbf{m} \quad \text{with } (\mathbf{x}, \mathbf{m}) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, \mathbf{n}, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_\mathbf{r} = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_\mathbf{u}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_\mathbf{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+In the case of $B_\mathbf{R}[n]_\mathbf{u}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_\mathbf{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathbf{u}
 \end{array}
@@
 \begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{Rsigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
 \end{equation}
@@
 \begin{equation}
+m = H_\mathbf{m} \quad \wedge \quad \mathbf{n} \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (\mathbf{m}, \mathbf{n}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
 m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash components; $H_n$ is the \hyperlink{block nonce}{nonce} of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathbf{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathbf{d}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_b}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_\mathbf{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_\mathbf{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \hypertarget{R b}{
-Formally, we define $R_b$:}
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+Formally, we define $R_\mathbf{b}$:}
+\begin{eqnarray}
+R_\mathbf{b}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{itemize}
-
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R l}{
+Thus we finish by formally defining $R_\mathbf{l}$:
+\begin{eqnarray}
+R_\mathbf{l}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \end{eqnarray}
-
+}
 If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
@@
 \begin{equation}
-\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
+\mathfrak{I} = \{ (k_0 \in \mathbb{B}, v_0 \in \mathbb{B}), k_1 \in \mathbb{B}, v_1 \in \mathbb{B}), ... \}
 \end{equation}
@@
 \begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+y(\mathfrak{I}) & = & \{ (k_0' \in \mathbb{Y}, v_0 \in \mathbb{B}), (k_1' \in \mathbb{Y}, v_1 \in \mathbb{B}), ... \} \\
+\forall_n \quad \forall_{i: i < 2\lVert\ k_n\rVert} \quad k_n'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor k_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+k_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_\mathbf{r} \\
+(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_\mathbf{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_\mathbf{r}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+\mathbf{g}_\mathbf{r} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{g}_\mathbf{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{g}_\mathbf{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{g}_\mathbf{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range)\hypertarget{v}{.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28\}
+\hyperlink{v}{v} &\in \{27,28\}
 \end{align}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_\mathbf{r}$, $T_\mathbf{s}$ and $\hyperlink{T w}{T_\mathbf{w}}$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}) & \text{if} \; T_\mathbf{t} = 0\\
+(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}) & \text{otherwise}
 \end{cases} \\
@@
 
-The signed transaction $G(T, p_r)$ is defined as:
+\hypertarget{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
 \begin{eqnarray}
 G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}
+(\hyperlink{T w}{T_\mathbf{w}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}}
 
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) \big) \big)
 \end{equation}
@@
 
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+%extern u256 const \mathbf{c}_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathbf{i})-C_{mem}(\boldsymbol{\mu}_\mathbf{i}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_\mathbf{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_\mathbf{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \end{tabularx}
@@
 0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \vee \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{a}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_\mathbf{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{o}$ \\
 &&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{s}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \hyperlink{I v}{I_\mathbf{v}}$ \\
 \midrule
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{p}$ \\
 &&&& This is the gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{c} \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{c}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\hyperlink{H p}{H_\mathbf{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
@@
 \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_\mathbf{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_\mathbf{i}} \\ P(\hyperlink{H p}{H_\mathbf{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{c}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{s}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{i}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{d}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{l}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
+&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_\mathbf{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_\mathbf{g}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_\mathbf{a}, I_\mathbf{o}, L(\boldsymbol{\mu}_\mathbf{g}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_\mathbf{e} + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_\mathbf{a}]_\mathbf{n} = \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{n} + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the  \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_\mathbf{e} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_\mathbf{a}, \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{n})$, the address of the  \\
 &&&& newly created account. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_\mathbf{a}, I_\mathbf{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge \\ \quad\quad I_\mathbf{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{g} \equiv \boldsymbol{\mu}_\mathbf{g} + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ (not enough funds) or $I_\mathbf{e} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_\mathbf{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_\mathbf{s}[0] \} & \text{if} \quad \boldsymbol{\mu}_\mathbf{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
 \end{cases}$\\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{a}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\\ \quad\quad{}I_\mathbf{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_\mathbf{a}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
@@
 &&&& See equation \ref{eq:Normal Halting}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabularx}
@@
 Otherwise it is exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{s}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_\mathbf{p}, 0, \hyperlink{I v}{I_\mathbf{v}}, \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathbf{v}} \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_\mathbf{a} \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} + \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ \\
+&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{b} \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
 0 & \text{otherwise} 
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+ E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{J_{epoch}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+ \mathbf{c}_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
 \end{equation}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
+\end{equation}
+\begin{equation}
+ C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
  E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+ \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
 \end{equation}

### 2017-09-28 Paper.tex
11.3 and 11.4 cross refs

@@
 %\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [3 3]}]{hyperref}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 %,pagebackref=true
@@
 
+\hypertarget{Pi}{}
 Formally, we expand to:
@@
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathbf{o}}$.\hypertarget{H c}{
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_\mathbf{c}$.}
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathbf{c}$.}
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarget{H d}{
@@
 \end{description}
-
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\hypertarget{B U}{
+The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathbf{U}$, and a series of the transactions, $B_\mathbf{T}$.
+Formally, we can refer to a block $B$:
 \begin{equation}
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
+\end{equation}}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as \hyperlink{H e}{$H_\mathbf{e}$}.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_\mathbf{e}$}.
 
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathbf{u}}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathbf{u}}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.
 
@@
 \begin{array}[t]{lclc}\hypertarget{H r}{
-H_\mathbf{r} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
-H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U})))} & \wedge \hypertarget{H t}{\\
+H_\mathbf{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
+H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}})))} & \wedge \hypertarget{H t}{\\
 H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$,} are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_\mathbf{p}}, \hyperlink{H o}{H_\mathbf{o}}, \hyperlink{H c}{H_\mathbf{c}}, \hyperlink{H r}{H_\mathbf{r}}, \hyperlink{H t}{H_\mathbf{t}}, \hyperlink{H e}{H_\mathbf{e}}, \hyperlink{H b}{H_\mathbf{b}}, \hyperlink{H d}{H_\mathbf{d}},\\ \hyperlink{H i}{H_\mathbf{i}}, \hyperlink{H l}{H_\mathbf{l}}, \hyperlink{H g}{H_\mathbf{g}}, \hyperlink{H s}{H_\mathbf{s}}, \hyperlink{H x}{H_\mathbf{x}}, \hyperlink{H m}{H_\mathbf{m}}, \hyperlink{H n}{H_\mathbf{n}} \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathbf{T}^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathbf{T}^*(B_\mathbf{T}), L_H^*(\hyperlink{B U}{B_\mathbf{U}}) \big)
 \end{eqnarray}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s]_\mathbf{b} &\equiv& \boldsymbol{\sigma}[s]_\mathbf{b} - v
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
 \boldsymbol{\sigma}_1'[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v & \text{otherwise}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert \hyperlink{B U}{B_\mathbf{U}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_\mathbf{U}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_\mathbf{g}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_\mathbf{g} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathbf{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_\mathbf{g}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H g}{{B_H}_\mathbf{g}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathbf{u}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R b}{$R_\mathbf{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_\mathbf{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 \hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[{B_H}_\mathbf{c}]_\mathbf{b} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})\hyperlink{R b}{R_\mathbf{b}} \\
+\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_\mathbf{c}}]_\mathbf{b} & = & \boldsymbol{\sigma}[{B_H}_\mathbf{c}]_\mathbf{b} + (1 + \frac{\lVert \hyperlink{B U}{B_\mathbf{U}}\rVert}{32})\hyperlink{block reward R b}{R_\mathbf{b}} \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[U_\mathbf{c}]_\mathbf{b} + (1 + \frac{1}{8} (U_\mathbf{i} - {B_H}_\mathbf{i})) \hyperlink{R b}{R_\mathbf{b}}
+ \boldsymbol{\sigma}'[U_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[U_\mathbf{c}]_\mathbf{b} + (1 + \frac{1}{8} (U_\mathbf{i} - \hyperlink{H i}{{B_H}_\mathbf{i}})) \hyperlink{block reward R b}{R_\mathbf{b}}
 \end{eqnarray}
@@
 If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
+two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.\hypertarget{block reward R b}{}
 
@@
 \begin{equation}
-\text{Let} \quad \hyperlink{R b}{R_\mathbf{b}} = 5 \times 10^{18}
+\text{Let} \quad R_\mathbf{b} = 5 \times 10^{18}
 \end{equation}
@@
 
+\hypertarget{Gamma}{
 We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
+\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
 \end{cases}
 \end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+}
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 \hypertarget{Phi}{
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'\mathbf{n} & = & \mathbf{n}: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
-B'_\mathbf{m} & = & \mathbf{m} \quad \text{with } (\mathbf{x}, \mathbf{m}) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, \mathbf{n}, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_\mathbf{r} = r(\Pi(\Gamma(B), B))
+B'_\mathbf{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
+B'_\mathbf{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_\mathbf{r} = r(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$ is the state-transition function}, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
 
@@
 
-Finally, we define $\Pi$ as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
 \begin{equation}
@@
 m = H_\mathbf{m} \quad \wedge \quad \mathbf{n} \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (\mathbf{m}, \mathbf{n}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
 \end{equation}
@@
 
-The state database won't be forced to maintain all past state trie structures into the future.
+The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
 It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
@@
 
+\hypertarget{trie}{}
 \section{Modified Merkle Patricia Tree}\label{app:trie}

### 2017-09-29 Paper.tex
Cross references, grammar, maths in Appendix I and J, and section 11

@@
 %\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage [autostyle]{csquotes}
+\MakeOuterQuote{"}
 \usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
@@
 
+%\usepackage{glossaries}
+
 \input{Version.tex}
@@
 
-\hypertarget{Pi}{}
 Formally, we expand to:
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_\mathbf{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathbf{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (the \hyperlink{GenesisBlock}{genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
+\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{H s}{$H_\mathbf{s}$}.\hypertarget{H x}{
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}\hypertarget{H s}{
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathbf{s}$}.}\hypertarget{H x}{
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathbf{x}$.}\hypertarget{H m}{
@@
 
-The transaction receipt is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
+\hypertarget{R}{}
+The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
 \begin{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence".} Formally:
 \begin{eqnarray}
@@
 H_\mathbf{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
-H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}})))} & \wedge \hypertarget{H t}{\\
+H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}}))) & \wedge }\hypertarget{H t}{\\
 H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
-H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\})} & \wedge \hypertarget{H b}{\\
-H_\mathbf{b} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}
+H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
+H_\mathbf{b} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}}
 \end{array}
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$,} are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{stateR-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \begin{equation}
-n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad \mathbf{m} = \hyperlink{H m}{H_\mathbf{m}}
+n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathbf{m}}
 \end{equation}
@@
 \end{eqnarray}
-where $(\mathbf{n}, \mathbf{m}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$
+where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$
 
@@
 In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
+Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{GenesisBlock}{genesis block}, in the path.
 The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
@@
 B'_\mathbf{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_\mathbf{r} = r(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_\mathbf{r}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 
-In the case of $B_\mathbf{R}[n]_\mathbf{u}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+In the case of \hyperlink{Transaction Receipt}{${B_\mathbf{R}}[n]_\mathbf{u}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 
+\hypertarget{Pi}{
 Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
@@
 \Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
-\end{equation}
+\end{equation}}
 
@@
 One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task.
-In Bitcoin's case the task is the SHA256 hash function.
+These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}).
+In Bitcoin's case the task is the SHA256 hash function (\cite{ASICbitcoinwiki}).
 While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
@@
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
+For Ethereum \textit{Frontier} and \textit{Homestead} we have chosen the first path.
 
@@
 \begin{equation}
-m = H_\mathbf{m} \quad \wedge \quad \mathbf{n} \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (\mathbf{m}, \mathbf{n}) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
+m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
 \end{equation}
@@
 the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being the \hyperlink{block nonce}{block nonce}, a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 The underlying algorithm is called Ethash and is described below.
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+Ethash is the PoW algorithm for Ethereum \textit{Frontier} and \textit{Homestead}.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
 The general route that the algorithm takes is as follows:
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{cacheinit}$} bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{datasetinit}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every \hyperlink{Jepoch}{$J_{epoch}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{GenesisBlock}{genesis block}.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 
+\hypertarget{rlp}{}
 \section{Recursive Length Prefix}\label{app:rlp}
@@
 
+\hypertarget{GenesisBlock}{}
 \section{Genesis Block}\label{app:genesis}
@@
 
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the \hyperlink{H c}{beneficiary address}, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the \hyperlink{H b}{log bloom}, a 2048-bit sequence of all zeros; $2^{17}$ refers to the \hyperlink{H d}{difficulty}; the \hyperlink{H t}{transaction trie root}, \hyperlink{H e}{receipt trie root}, \hyperlink{H g}{gas used}, \hyperlink{H i}{block number} and \hyperlink{H x}{extradata} are all $0$ (listed in the order of decimal zeros in the above sequence), being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the \hyperlink{block nonce}{nonce}.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
-
-The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the initial timestamp of the genesis block.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in \hyperlink{rlp}{RLP}, both empty lists.
+
+The proof-of-concept series includes a development premine, making the state root hash some value $stateRoot$.
+Also \hyperlink{H s}{$time$} will be set to the initial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 
-\begin{tabularx}{\textwidth}{rlrrX}
+\begin{tabularx}{\textwidth}{rlX}
 \toprule
@@
 \midrule 
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\hypertarget{Jwordbytes}{
+$J_{wordbytes}$ & 4  & This is the bytes in a word. }\hypertarget{Jdatasetinit}{\\
+$J_{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
+$J_{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
+$J_{cacheinit}$ & $2^{24}$ & This is the bytes in the cache at genesis. }\hypertarget{Jcachegrowth}{\\
+$J_{cachegrowth}$ & $2^{17}$ & This is the cache growth per epoch. }\hypertarget{Jepoch}{\\
+$J_{epoch}$ & 30000 & This is the blocks per epoch. }\hypertarget{Jmixbytes}{\\
+$J_{mixbytes}$ & 128 & This is the mix length in bytes. }\hypertarget{Jhashbytes}{\\
+$J_{hashbytes}$ & 64 & This is the hash length in bytes. }\hypertarget{Jparents}{\\
+$J_{parents}$ & 256 & This is the number of parents of each dataset element. }\hypertarget{Jcacherounds}{\\
+$J_{cacherounds}$ & 3 & This is the number of rounds in cache production. }\hypertarget{Jaccesses}{\\
+$J_{accesses}$ & 64 & This is the number of accesses in a hashimoto loop. }\\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+ E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{epoch}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{cachegrowth}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{mixbytes}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{hashbytes}$} for the cache.
 Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
 \begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+ d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{datasetgrowth}} \cdot E_{epoch} - \hyperlink{Jmixbytes}{J_{mixbytes}}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- \mathbf{c}_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+ c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{cacheinit}} + \hyperlink{cachegrowth}{J_{cachegrowth}} \cdot E_{epoch} - \hyperlink{Jhashbytes}{J_{hashbytes}}, J_{hashbytes})
 \end{equation}
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $s \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - J_{epoch})) & \text{otherwise}
+\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{epoch}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fil up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{cacherounds}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+ n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{cacherounds}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{cacherounds}})
 \end{equation}
@@
 Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{hashbytes}$} bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
 \end{equation}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{hashbytes}} / \hyperlink{Jwordbytes}{J_{wordbytes}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+\hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+ n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{mixbytes}}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
 \end{equation}
@@
  E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{accesses}} -2 \\
 E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{mixbytes}}}{\hyperlink{Jwordbytes}{J_{wordbytes}}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / \hyperlink{Jhashbytes}{J_{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}
@@
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabularx}{\textwidth}{lcX}
+\toprule
+Symbol & Latex Command & Description \\
+\midrule
+\hypertarget{bigvee}{
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
+Thus it is the greatest element of such elements.
+More information is here: \cite{LatticeOrder}.}\\
+\bottomrule
+\end{tabularx}
+
 \end{document}

### 2017-09-29 Paper.tex
Made changes but now it fails building at the Ethash table, and I can't figure out why.

@@
 % https://stackoverflow.com/a/1918465/7438857
-%tabularx seems better
+%tabularx seems better, but we can include this if we need tables to break, as well as tabu and longtabu
 
@@
 %https://stackoverflow.com/a/1919447/7438857
+
+\usepackage{tabu}
 
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{cacheinit}$} bytes in initial size.
+From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{\rm{cacheinit}}$} bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{datasetinit}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{\rm{datasetinit}}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every \hyperlink{Jepoch}{$J_{epoch}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every \hyperlink{Jepoch}{$J_{\rm{epoch}}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 \begin{tabularx}{\textwidth}{rlrrX}
-\midrule
+\toprule
 0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
@@
 \subsection{Definitions}
+
 We employ the following definitions:
 
-\begin{tabularx}{\textwidth}{rlX}
+%\begin{tabularx}{\textwidth}{lrX}
+\begin{tabu}{X[,l]X[-r]X[-l]} 
 \toprule
@@
 \midrule 
-\hypertarget{Jwordbytes}{
-$J_{wordbytes}$ & 4  & This is the bytes in a word. }\hypertarget{Jdatasetinit}{\\
-$J_{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
-$J_{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
-$J_{cacheinit}$ & $2^{24}$ & This is the bytes in the cache at genesis. }\hypertarget{Jcachegrowth}{\\
-$J_{cachegrowth}$ & $2^{17}$ & This is the cache growth per epoch. }\hypertarget{Jepoch}{\\
-$J_{epoch}$ & 30000 & This is the blocks per epoch. }\hypertarget{Jmixbytes}{\\
-$J_{mixbytes}$ & 128 & This is the mix length in bytes. }\hypertarget{Jhashbytes}{\\
-$J_{hashbytes}$ & 64 & This is the hash length in bytes. }\hypertarget{Jparents}{\\
-$J_{parents}$ & 256 & This is the number of parents of each dataset element. }\hypertarget{Jcacherounds}{\\
-$J_{cacherounds}$ & 3 & This is the number of rounds in cache production. }\hypertarget{Jaccesses}{\\
-$J_{accesses}$ & 64 & This is the number of accesses in a hashimoto loop. }\\
+\hypertarget{Jwordbytes}{}
+$J_{\rm{wordbytes}}$ & 4  & This is the bytes in a word. \hypertarget{Jdatasetinit}{ \\
+$J_{\rm{{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
+$J_{\rm{{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
+$J_{\rm{{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis. }\hypertarget{Jcachegrowth}{\\
+$J_{\rm{{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch. }\hypertarget{Jepoch}{\\
+$J_{\rm{{epoch}}$ & 30000 & This is the blocks per epoch. }\hypertarget{Jmixbytes}{\\
+$J_{\rm{{mixbytes}}$ & 128 & This is the mix length in bytes. }\hypertarget{Jhashbytes}{\\
+$J_{\rm{{hashbytes}}$ & 64 & This is the hash length in bytes. }\hypertarget{Jparents}{\\
+$J_{\rm{\rm{{parents}}}$ & 256 & This is the number of parents of each dataset element. }\hypertarget{Jcacherounds}{\\
+$J_{\rm{{cacherounds}}$ & 3 & This is the number of rounds in cache production. }\hypertarget{Jaccesses}{\\
+$J_{\rm{{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop. }\\
 \bottomrule
-\end{tabularx}
+\end{tabu}
+%\end{tabularx}
 
@@
 \begin{equation}
- E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{epoch}}}\right\rfloor
-\end{equation}
-The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{cachegrowth}$} bytes, where the growth of each occurs every epoch.
+E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{\rm{epoch}}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\rm{{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_\rm{{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{mixbytes}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{hashbytes}$} for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{datasetgrowth}} \cdot E_{epoch} - \hyperlink{Jmixbytes}{J_{mixbytes}}, J_{mixbytes})
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_\rm{{hashbytes}}$} for the cache.\hypertarget{dsize}{
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
+\begin{equation}
+ d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\rm{{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\rm{{datasetgrowth}} \cdot E_\rm{{epoch}} - \hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}, J_\rm{{mixbytes}})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{cacheinit}} + \hyperlink{cachegrowth}{J_{cachegrowth}} \cdot E_{epoch} - \hyperlink{Jhashbytes}{J_{hashbytes}}, J_{hashbytes})
+ c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_\rm{{cacheinit}}} + \hyperlink\rm{{cachegrowth}}{J_\rm{{cachegrowth}}} \cdot E_\rm{{epoch}} - \hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}, J_\rm{{hashbytes}})
 \end{equation}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
+ s = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
 \end{equation}
@@
  C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{epoch}})) & \text{otherwise}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_\rm{{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_\rm{{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially fil up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{cacherounds}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_\rm{{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{cacherounds}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{cacherounds}})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing \hyperlink{Jcacherounds}{$J_\rm{{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_\rm{{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_\rm{{cacherounds}}})
+\end{equation}
+\begin{equation}
+ E_\rm{{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_\rm{{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \end{equation}
+\hypertarget{Ecacherounds}{}
 \begin{multline}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{hashbytes}$} bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
+Essentially, we combine data from $J_\rm{{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_\rm{{hashbytes}}$} bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+ E_{datasetitem}(\mathbf{c}, i) = E_\rm{{parents}}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_\rm{{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_\rm{{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_\rm{{parents}} -2 \\
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{hashbytes}} / \hyperlink{Jwordbytes}{J_{wordbytes}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_\rm{{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\rm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-\hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+\hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_\rm{{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{mixbytes}}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{accesses}} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+ n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_\rm{{accesses}}$ function is used:
+\begin{equation}
+ E_\rm{{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_\rm{{accesses}}} -2 \\
+E_\rm{{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{mixbytes}}}{\hyperlink{Jwordbytes}{J_{wordbytes}}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / \hyperlink{Jhashbytes}{J_{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\rm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}

### 2017-09-29 Paper.tex
Testing to fix passing build, for comparison

@@
 %\usepackage{longtable}
-% https://stackoverflow.com/a/1918465/7438857
-%tabularx seems better, but we can include this if we need tables to break, as well as tabu and longtabu
+%https://stackoverflow.com/questions/1918366/automated-line-breaks-in-latex-tables/1918465#1918465
+%tabularx seems better, but we can include this if we need tables to break across pages automatically, as well as tabu and longtabu
 
@@
 
+%was using tabularx, but found tabu which seems more flexible, and generally better.
 \usepackage{tabu}
@@
 \begin{tabularx}{\textwidth}{rlrrX}
-\toprule
+\midrule
 0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
@@
 \subsection{Definitions}
-
 We employ the following definitions:
 
-%\begin{tabularx}{\textwidth}{lrX}
-\begin{tabu}{X[,l]X[-r]X[-l]} 
+\begin{tabularx}{\textwidth}{lrX}
+%\begin{tabu}{X[,l]X[,r]X[,l]}
+
 \toprule
-Name & Value & Description \\
-\midrule 
-\hypertarget{Jwordbytes}{}
-$J_{\rm{wordbytes}}$ & 4  & This is the bytes in a word. \hypertarget{Jdatasetinit}{ \\
-$J_{\rm{{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
-$J_{\rm{{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
-$J_{\rm{{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis. }\hypertarget{Jcachegrowth}{\\
-$J_{\rm{{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch. }\hypertarget{Jepoch}{\\
-$J_{\rm{{epoch}}$ & 30000 & This is the blocks per epoch. }\hypertarget{Jmixbytes}{\\
-$J_{\rm{{mixbytes}}$ & 128 & This is the mix length in bytes. }\hypertarget{Jhashbytes}{\\
-$J_{\rm{{hashbytes}}$ & 64 & This is the hash length in bytes. }\hypertarget{Jparents}{\\
-$J_{\rm{\rm{{parents}}}$ & 256 & This is the number of parents of each dataset element. }\hypertarget{Jcacherounds}{\\
-$J_{\rm{{cacherounds}}$ & 3 & This is the number of rounds in cache production. }\hypertarget{Jaccesses}{\\
-$J_{\rm{{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop. }\\
+Name & Value & Description\hypertarget{Jwordbytes}{}\\
+\midrule 
+$J_{\rm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{\\
+$J_{\rm{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis.}\hypertarget{Jdatasetgrowth}{\\
+$J_{\rm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch.}\hypertarget{Jcacheinit}{\\
+$J_{\rm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
+$J_{\rm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
+$J_{\rm{epoch}}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
+$J_{\rm{mixbytes}}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
+$J_{\rm{hashbytes}}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
+$J_{\rm{parents}}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
+$J_{\rm{cacherounds}}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
+$J_{\rm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
 \bottomrule
-\end{tabu}
-%\end{tabularx}
+%\end{tabu}
+\end{tabularx}
 
@@
 \end{equation}
-The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\rm{{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_\rm{{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\rm{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\rm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_\rm{{hashbytes}}$} for the cache.\hypertarget{dsize}{
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\rm{hashbytes}}$} for the cache.\hypertarget{dsize}{
 Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
 \begin{equation}
- d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\rm{{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\rm{{datasetgrowth}} \cdot E_\rm{{epoch}} - \hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}, J_\rm{{mixbytes}})
+ d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\rm{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\rm{datasetgrowth}} \cdot E_{\rm{epoch}} - \hyperlink{Jmixbytes}{J_\{rm{mixbytes}}}, J_{\rm{mixbytes}})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_\rm{{cacheinit}}} + \hyperlink\rm{{cachegrowth}}{J_\rm{{cachegrowth}}} \cdot E_\rm{{epoch}} - \hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}, J_\rm{{hashbytes}})
+ c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\rm{cacheinit}}} + \hyperlink{\rm{cachegrowth}}{J_{\rm{cachegrowth}}} \cdot E_{\rm{epoch}} - \hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}, J_{\rm{hashbytes}})
 \end{equation}
@@
  C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_\rm{{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_\rm{{epoch}}})) & \text{otherwise}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\rm{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_\rm{{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\rm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing \hyperlink{Jcacherounds}{$J_\rm{{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_\rm{{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_\rm{{cacherounds}}})
-\end{equation}
-\begin{equation}
- E_\rm{{cacherounds}}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\rm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{\rm{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{\rm{cacherounds}}})
+\end{equation}
+\begin{equation}
+ E_{\rm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_\rm{{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{\rm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_\rm{{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_\rm{{hashbytes}}$} bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
+Essentially, we combine data from $J_{\rm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{\rm{hashbytes}}$} bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_\rm{{parents}}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_\rm{{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_\rm{{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_\rm{{parents}} -2 \\
+ E_{datasetitem}(\mathbf{c}, i) = E_{\rm{parents}}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{\rm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{\rm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\rm{parents}} -2 \\
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_\rm{{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\rm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\rm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\rm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-\hyperlink{Jmixbytes}{$J_\rm{{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+\hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_\rm{{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\rm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_\rm{{accesses}}$ function is used:
-\begin{equation}
- E_\rm{{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_\rm{{accesses}}} -2 \\
-E_\rm{{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+ n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\rm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{\rm{accesses}}$ function is used:
+\begin{equation}
+ E_{\rm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\rm{accesses}}} -2 \\
+E_{\rm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_\rm{{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\rm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_\rm{{hashbytes}}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\rm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\rm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}

### 2017-09-29 Paper.tex
Fixed cause of error: no closing brace for hyperlinks

--- 2017-09-29 Paper.tex
@@
-\documentclass[9pt,oneside]{amsart}
-\usepackage{cancel}
-%\usepackage{tweaklist}
-\usepackage{xspace}
-\usepackage{graphicx}
-\usepackage{multicol}
-\usepackage{subfig}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
-\usepackage{booktabs}
-\usepackage{array}
-\usepackage{verbatim}
-\usepackage{caption}
-\usepackage{natbib}
-\usepackage{float}
-\usepackage{pdflscape}
-\usepackage{mathtools}
-\usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{afterpage}
-\usepackage{tikz}
-\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage [autostyle]{csquotes}
-\MakeOuterQuote{"}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-%\usepackage{longtable}
-% https://stackoverflow.com/a/1918465/7438857
-%tabularx seems better
+\documentclass{amsart}
 
 \usepackage{tabularx}
-%https://stackoverflow.com/a/1919447/7438857
+\usepackage{booktabs}
+\usepackage{hyperref}
+\usepackage{amsmath}
 
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-%\PassOptionsToPackage{url}\g@addto@macro{\UrlBreaks}{\UrlOrds}
-% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
-Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
-If you force a URL to break over a page, the error still occurs.
-
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
-% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
-%\sloppy
-%\setlength{\emergencystretch}{2pt}
-% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
-% https://tex.stackexchange.com/a/59131/143781
-% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-%\usepackage{glossaries}
-
-\input{Version.tex}
-
-\newcommand{\hcancel}[1]{%
-    \tikz[baseline=(tocancel.base)]{
-        \node[inner sep=0pt,outer sep=0pt] (tocancel) {#1};
-        \draw[black] (tocancel.south west) -- (tocancel.north east);
-    }%
-}%
-
-\definecolor{lightyellow}{rgb}{1,0.98,0.9}
-\definecolor{lightpink}{rgb}{1,0.94,0.95}
-
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\newcommand*\eg{e.g.\@\xspace}
-\newcommand*\Eg{e.g.\@\xspace}
-\newcommand*\ie{i.e.\@\xspace}
-%\renewcommand{\itemhook}{\setlength{\topsep}{0pt}  \setlength{\itemsep}{0pt}\setlength{\leftmargin}{15pt}}
-
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
-\author{
-    Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
-}
 \begin{document}
-
-\pagecolor{lightyellow}
-%\pagecolor{lightpink}
-
-\begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not the least being Bitcoin.
-Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
-We can call this paradigm a transactional singleton machine with shared-state.
-
-Ethereum implements this paradigm in a generalised manner.
-Furthermore it provides a plurality of such resources, each with a distinct state and operating code but able to interact through a message-passing framework with others.
-We discuss its design, implementation issues, the opportunities it provides and the future hurdles we envisage.
-\end{abstract}
-
-\maketitle
-
-\setlength{\columnsep}{20pt}
-\begin{multicols}{2}
-
-\section{Introduction}\label{sec:introduction}
-
-With ubiquitous internet connections in most places of the world, global information transmission has become incredibly cheap.
-Technology-rooted movements like Bitcoin have demonstrated, through the power of the default, consensus mechanisms and voluntary respect of the social contract that it is possible to use the internet to make a decentralised value-transfer system, shared across the world and virtually free to use.
-This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
-Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications albeit rather simplistic ones.
-
-Ethereum is a project which attempts to build the generalised technology; technology on which all transaction-based state machine concepts may be built.
-Moreover it aims to provide to the end-developer a tightly integrated end-to-end system for building software on a hitherto unexplored compute paradigm in the mainstream: a trustful object messaging compute framework.
-
-\subsection{Driving Factors} \label{ch:driving}
-
-There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
-This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
-By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
-
-Dealings in this proposed system would have several attributes not often found in the real world.
-The incorruptibility of judgement, often difficult to find, comes naturally from a disinterested algorithmic interpreter.
-Transparency, or being able to see exactly how a state or judgement came about through the transaction log and rules or instructional codes, never happens perfectly in human-based systems since natural language is necessarily vague, information is often lacking, and plain old prejudices are difficult to shake.
-
-Overall, I wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
-
-\subsection{Previous Work} \label{ch:previous}
-
-\cite{buterin2013ethereum} first proposed the kernel of this work in late November, 2013.
-Though now evolved in many ways, the key functionality of a block-chain with a Turing-complete language and an effectively unlimited inter-transaction storage capability remains unchanged.
-
-\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet.
-The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}.
-\cite{back2002hashcash} later produced a system in a similar vein.
-
-The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, providing ``consumers'' with the ability to make micro-payments to ``suppliers'' for their services.
-The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
-Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope.
-The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.
-
-Other projects built on Bitcoin's success; the alt-coins introduced numerous other currencies through alteration to the protocol.
-Some of the best known are Litecoin and Primecoin, discussed by \cite{sprankel2013technical}.
-Other projects sought to take the core value content mechanism of the protocol and repurpose it; \cite{aron2012bitcoin} discusses, for example, the Namecoin project which aims to provide a decentralised name-resolution system.
-
-Other projects still aim to build upon the Bitcoin network itself, leveraging the large amount of value placed in the system and the vast amount of computation that goes into the consensus mechanism.
-The Mastercoin project, first proposed by \cite{mastercoin2013willett}, aims to build a richer protocol involving many additional high-level features on top of the Bitcoin protocol through utilisation of a number of auxiliary parts to the core protocol.
-The Coloured Coins project, proposed by \cite{colouredcoins2012rosenfeld}, takes a similar but more simplified strategy, embellishing the rules of a transaction in order to break the fungibility of Bitcoin's base currency and allow the creation and tracking of tokens through a special ``chroma-wallet''-protocol-aware piece of software.
-
-Additional work has been done in the area with discarding the decentralisation foundation; Ripple, discussed by \cite{boutellier2014pirates}, has sought to create a ``federated'' system for currency exchange, effectively creating a new financial clearing system.
-It has demonstrated that high efficiency gains can be made if the decentralisation premise is discarded.
-
-Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
-Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
-In this light, Ethereum may be seen as a general implementation of such a \textit{crypto-law} system.
-
-%E language/glossary?
-
-For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
-
-\section{The Blockchain Paradigm} \label{ch:overview}
-
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
-It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
-The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
-Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
-A valid state transition is one which comes about through a transaction.
-Formally:\hypertarget{Upsilon}{}
-\begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_\mathbf{t}, T)
-\end{equation}
-
-where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
-
-Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal or ledger, recording a series of transactions together with the previous block and an identifier for the final state (though blocks do not store the final state itself---that would be far too big).
-They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes place as a state-transition function, adding value to a nominated account.
-
-Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block.
-It is achieved thanks to a cryptographically secure proof.
-This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
-
-Formally, we expand to:
-\begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathbf{t}, B) \\
-B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
-
-This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
-
-\subsection{Value}
-
-In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
-The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
-One Ether is defined as being $10^{18}$ Wei.
-There exist other subdenominations of Ether:
-\par
-\begin{center}
-\begin{tabular}{rl}
-\toprule
-Multiplier & Name \\
-\midrule
-$10^0$ & Wei \\
-$10^{12}$ & Szabo \\
-$10^{15}$ & Finney \\
-$10^{18}$ & Ether \\
-\bottomrule
-\end{tabular}
-\end{center}
-\par
-
-Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.
-
-\subsection{Which History?}
-
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the \hyperlink{GenesisBlock}{genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
-If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
-
-This would mean that past a given point in time (block), multiple states of the system may coexist: some nodes believing one block to contain the canonical transactions, other nodes believing some other block to be canonical, potentially containing radically different or incompatible transactions.
-This is to be avoided at all costs as the uncertainty that would ensue would likely kill all confidence in the entire system.
-
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
-This process is described in detail in section \ref{ch:ghost}.
-
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
-In order to follow back the history of a path, one might need to reference multiple versions of this document.
-
-\section{Conventions}\label{ch:conventions}
-
-I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
-
-The two sets of highly structured, `top-level', state values, are denoted with bold lowercase Greek letters.
-They fall into those of world-state, which are denoted $\boldsymbol{\sigma}$ (or a variant thereupon) and those of machine-state, $\boldsymbol{\mu}$.
-
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg \hyperlink{Upsilon}{$\Upsilon$}, the Ethereum state transition function.
-
-For most functions, an uppercase letter is used, e.g.
-$C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
-Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
-
-Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_\mathbf{n}$, denotes the \hyperlink{transaction nonce}{nonce of said transaction}.
-The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
-
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
-Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
-For particularly important values, a bold uppercase letter may be used.
-
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
-The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \ref{ch:block}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
-
-In the case of the global state $\boldsymbol{\sigma}$, which is a sequence of accounts, themselves tuples, the square brackets are used to reference an individual account.
-
-When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
-
-I define a number of useful functions throughout. \hypertarget{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
-
-\begin{equation}
-\ell(\mathbf{x}) \equiv \mathbf{x}[\lVert \mathbf{x} \rVert - 1]
-\end{equation}
-
-\section{Blocks, State and Transactions} \label{ch:bst}
-
-Having introduced the basic concepts behind Ethereum, we will discuss the meaning of a transaction, a block and the state in more detail.
-
-\subsection{World State} \label{ch:state}
-
-The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
-Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
-This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
-Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly.
-Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.
-
-The account state comprises the following four fields:
-
-\begin{description}
-\hypertarget{account nonce}{}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_\mathbf{n}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_\mathbf{b}$.
-\item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_\mathbf{s}$.
-\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
-All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_\mathbf{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_\mathbf{c}$.
-\end{description}
-
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
-\begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_\mathbf{s}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
-\begin{equation}
-L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
-\end{equation}
-
-where:
-\begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
-\end{equation}
-
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
-
-If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_\mathbf{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
-\begin{equation}
-L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
-\end{equation}
-where
-\begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_\mathbf{n}, \boldsymbol{\sigma}[a]_\mathbf{b}, \boldsymbol{\sigma}[a]_\mathbf{s}, \boldsymbol{\sigma}[a]_\mathbf{c}) \big) \big)
-\end{equation}
-
-This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
-\begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
-\end{equation}
-\hypertarget{v x}{where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_\mathbf{n} \in \mathbb{P}_{256} \wedge x_\mathbf{b} \in \mathbb{P}_{256} \wedge x_\mathbf{s} \in \mathbb{B}_{32} \wedge x_\mathbf{c} \in \mathbb{B}_{32}
-}
-\end{equation}
-
-\subsection{The Transaction} \label{subsec:transaction}
-
-A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
-\eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
-There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
-
-\begin{description}
-\hypertarget{T n}{
-\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_\mathbf{n}$.}\hypertarget{T p}{
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_\mathbf{p}$.}\hypertarget{T g}{
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_\mathbf{g}$.}\hypertarget{T t}{
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathbf{t}$.}\hypertarget{T v}{
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathbf{v}$.}
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hyperlink{T w}{$T_\mathbf{w}$}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}.
-This is expanded in Appendix \ref{app:signing}.
-\end{description}
-
-Additionally, a contract creation transaction contains:
-
-\begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
-\end{description}
-
-\textbf{init} is an EVM-code fragment; it returns the \textbf{body}, a second fragment of code that executes each time the account receives a message call (either through a transaction or due to the internal execution of code).
-\textbf{init} is executed only once at account creation and gets discarded immediately thereafter.
-
-In contrast, a message call transaction contains:
-
-\begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
-\end{description}
-
-\begin{equation}
-L_\mathbf{T}(T) \equiv \begin{cases}
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}, \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) & \text{if} \; T_\mathbf{t} = \varnothing\\
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}, \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
-\begin{equation}
-\begin{array}[t]{lclclc}
-T_\mathbf{n} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_\mathbf{v}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{p} \in \mathbb{P}_{256} & \wedge \\
-T_\mathbf{g} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_\mathbf{w}} \in \mathbb{P}_5 & \wedge & T_\mathbf{r} \in \mathbb{P}_{256} & \wedge \\
-T_\mathbf{s} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
-\end{array}
-\end{equation}
-where
-\begin{equation}
-\mathbb{P}_\mathbf{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_\mathbf{t} \neq \varnothing \\
-\mathbb{B}_{0} & \text{otherwise}\end{cases}
-\end{equation}
-
-\subsection{The Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other \hypertarget{ommerheaders}{block headers $\mathbf{U}$} that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, as it is not a well-known word) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
-The block header contains several pieces of information:
-
-%\textit{TODO: Introduce logs}
-
-\begin{description}\hypertarget{H p}{
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathbf{p}$.}
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathbf{o}}$.\hypertarget{H c}{
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathbf{c}$.}
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarget{H d}{
-\item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
-\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}\hypertarget{H s}{
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathbf{s}$}.}\hypertarget{H x}{
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathbf{x}$.}\hypertarget{H m}{
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{m}$.}
-\hypertarget{block nonce}{\hypertarget{H n}{
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{n}$.}}
-\end{description}
-\hypertarget{B U}{
-The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathbf{U}$, and a series of the transactions, $B_\mathbf{T}$.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}}
-
-\hypertarget{Transaction Receipt}{\subsubsection{Transaction Receipt}}
-
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_\mathbf{e}$}.
-
-\hypertarget{R}{}
-The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_\mathbf{u}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathbf{u}}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.
-
-We assert $\hyperlink{R u}{R_\mathbf{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_\mathbf{b}$}, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-\hyperlink{R u}{R_\mathbf{u}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_\mathbf{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $\hyperlink{R l}{R_\mathbf{l}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_\mathbf{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_\mathbf{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_\mathbf{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
-\end{equation}
-
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
-\begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
-\end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence".} Formally:
-\begin{eqnarray}
-M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
-\mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
-m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
-\end{eqnarray}
-
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
-
-\subsubsection{Holistic Validity}
-
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathbf{r}$:
-\begin{equation}
-\begin{array}[t]{lclc}\hypertarget{H r}{
-H_\mathbf{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
-H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}}))) & \wedge }\hypertarget{H t}{\\
-H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
-H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
-H_\mathbf{b} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}}
-\end{array}
-\end{equation}
-
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
-\begin{equation}
-p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
-\end{equation}
-
-Furthermore:
-\begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_\mathbf{r}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the \hyperlink{transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{stateR-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
-
-\subsubsection{Serialisation}
-
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_\mathbf{p}}, \hyperlink{H o}{H_\mathbf{o}}, \hyperlink{H c}{H_\mathbf{c}}, \hyperlink{H r}{H_\mathbf{r}}, \hyperlink{H t}{H_\mathbf{t}}, \hyperlink{H e}{H_\mathbf{e}}, \hyperlink{H b}{H_\mathbf{b}}, \hyperlink{H d}{H_\mathbf{d}},\\ \hyperlink{H i}{H_\mathbf{i}}, \hyperlink{H l}{H_\mathbf{l}}, \hyperlink{H g}{H_\mathbf{g}}, \hyperlink{H s}{H_\mathbf{s}}, \hyperlink{H x}{H_\mathbf{x}}, \hyperlink{H m}{H_\mathbf{m}}, \hyperlink{H n}{H_\mathbf{n}} \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathbf{T}^*(B_\mathbf{T}), L_H^*(\hyperlink{B U}{B_\mathbf{U}}) \big)
-\end{eqnarray}
-
-With $L_\mathbf{T}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
-
-The component types are defined thus:
-\begin{equation}
-\begin{array}[t]{lclclcl}
-\hyperlink{H p}{H_\mathbf{p}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_\mathbf{o}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_\mathbf{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{H r}{H_\mathbf{r}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_\mathbf{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_\mathbf{e}} \in \mathbb{B}_{32} & \wedge \\
-H_\mathbf{b} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_\mathbf{d}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_\mathbf{i}} \in \mathbb{P} & \wedge \\
-\hyperlink{H l}{H_\mathbf{l}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_\mathbf{g}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_\mathbf{s}} \in \mathbb{P}_{256} & \wedge \\
-\hyperlink{H x}{H_\mathbf{x}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_\mathbf{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_\mathbf{n}} \in \mathbb{B}_{8}
-\end{array}
-\end{equation}
-
-where
-\begin{equation}
-\mathbb{B}_\mathbf{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
-\end{equation}
-
-We now have a rigorous specification for the construction of a formal block structure.
-The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
-
-\subsubsection{Block Header Validity}
-
-We define $P(B_H)$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_\mathbf{p}}
-\end{equation}
-
-The block number is the parent's block number incremented by one:
-\begin{equation}
-\hyperlink{H i}{H_\mathbf{i}} \equiv {{P(H)_H}_\mathbf{i}} + 1
-\end{equation}
-
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
-The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad \hyperlink{H i}{H_\mathbf{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_\mathbf{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
-where:
-\begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_\mathbf{d}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_\mathbf{s}} - {P(H)_H}_\mathbf{s}}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_\mathbf{i}} \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
-
-The canonical gas limit $\hyperlink{H l}{H_\mathbf{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000
-\end{eqnarray}
-
-\hyperlink{H s}{$H_\mathbf{s}$} is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-\hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s}
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
-
-The nonce of a block, $\hyperlink{H n}{H_\mathbf{n}}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathbf{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$.
-
-\hypertarget{h cancel n}{}
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_\mathbf{d}}$.
-
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the \hyperlink{block nonce}{block nonce} must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \wedge m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \\
-& & \hyperlink{H d}{H_\mathbf{d}} = D(H) \quad \wedge \\
-& & \hyperlink{H g}{H_\mathbf{g}} \le \hyperlink{H l}{H_\mathbf{l}}  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000  \quad \wedge \\
-& & \hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s} \quad \wedge \\
-& & \hyperlink{H i}{H_\mathbf{i}} = {P(H)_H}_\mathbf{i} +1 \quad \wedge \\
-& & \lVert \hyperlink{H x}{H_\mathbf{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$
-
-Noting additionally that \textbf{extraData} must be at most 32 bytes.
-
-\section{Gas and Payment} \label{ch:payment}
-
-In order to avoid issues of network abuse and to sidestep the inevitable questions stemming from Turing completeness, all programmable computation in Ethereum is subject to fees.
-The fee schedule is specified in units of \textit{gas} (see Appendix \ref{app:fees} for the fees associated with various computation).
-Thus any given fragment of programmable computation (this includes creating contracts, making message calls, utilising and accessing account storage and executing operations on the virtual machine) has a universally agreed cost in terms of gas.
-
-Every transaction has a specific amount of gas associated with it: \textbf{gasLimit}.
-This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according \textbf{gasPrice}, also specified in the transaction.
-The transaction is considered invalid if the account balance cannot support such a purchase.
-It is named \textbf{gasLimit} since any unused gas at the end of the transaction is refunded (at the same rate of purchase) to the sender's account.
-Gas does not exist outside of the execution of a transaction.
-Thus for accounts with trusted code associated, a relatively high gas limit may be set and left alone.
-
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
-Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
-Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
-Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.
-
-%\subsubsection{Determining Computation Costs}
-
-\section{Transaction Execution} \label{ch:transactions}
-
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon}{$\Upsilon$}.
-It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
-These include:
-
-\begin{enumerate}
-\item The transaction is well-formed RLP, with no additional trailing bytes;
-\item the transaction signature is valid;
-\item the \hyperlink{transaction nonce}{transaction nonce} is valid (equivalent to the  \hyperlink{account nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
-\end{enumerate}
-
-Formally, we consider the function \hyperlink{Upsilon}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
-\begin{equation}
-\boldsymbol{\sigma}' = \Upsilon(\boldsymbol{\sigma}, T)
-\end{equation}
-
-Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{Upsilong}{$\Upsilon^g$} to evaluate to the amount of gas used in the execution of a transaction and \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} to evaluate to the transaction's accrued log items, both to be formally defined later.
-
-\subsection{Substate}
-Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
-We call this \textit{transaction substate}, and represent it as $A$, which is a tuple:
-\begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{r})
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_\mathbf{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
-Though not immediately refunded, it is allowed to partially offset the total execution costs.
-
-For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), 0)
-\end{equation}
-
-\subsection{Execution}
-We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
-
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
-
-%todo Explain g_d reason?
-
-The up-front cost $v_0$ is calculated as:
-\begin{equation}
-v_0 \equiv T_\mathbf{g} T_\mathbf{p} + \hyperlink{T v}{T_\mathbf{v}}
-\end{equation}
-
-The validity is determined as:
-\begin{equation}
-\begin{array}[t]{rcl}
-S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_\mathbf{n} & = & \boldsymbol{\sigma}[S(T)]_\mathbf{n} \quad \wedge \\
-g_0 & \leqslant & T_\mathbf{g} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_\mathbf{b} \quad \wedge \\
-T_\mathbf{g} & \leqslant & {B_H}_\mathbf{l} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_\mathbf{R})_\mathbf{u}}
-\end{array}
-\end{equation}
-
-Note the final condition; the sum of the transaction's gas limit, $T_\mathbf{g}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathbf{R})_\mathbf{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_\mathbf{l}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_\mathbf{g}T_\mathbf{p}$.
-The gas available for the proceeding computation, $g$, is defined as $T_\mathbf{g} - g_0$.
-The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
-
-We define the checkpoint state $\boldsymbol{\sigma}_0$:
-\begin{eqnarray}
-\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\%
-\boldsymbol{\sigma}_0[S(T)]_\mathbf{n} & \equiv & \boldsymbol{\sigma}[S(T)]_\mathbf{n} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_\mathbf{o}, &\\ \quad\quad g, T_\mathbf{p}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}, 0) & \text{if} \quad T_\mathbf{t} = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_\mathbf{o}, &\\ \quad\quad T_\mathbf{t}, T_\mathbf{t}, g, T_\mathbf{p}, \hyperlink{T v}{T_\mathbf{v}}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}, 0) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
-\begin{equation}
-g \equiv T_\mathbf{g} - g_0
-\end{equation}
-and $T_\mathbf{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
-
-After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
-\begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_\mathbf{g} - g'}{2} \Big\rfloor, A_\mathbf{r} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_\mathbf{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_\mathbf{g} - g'$.
-
-The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_\mathbf{b} & \equiv & \boldsymbol{\sigma}_P[S(T)]_\mathbf{b} + g^* T_\mathbf{p} \\
-\boldsymbol{\sigma}^*[m]_\mathbf{b} & \equiv & \boldsymbol{\sigma}_P[m]_\mathbf{b} + (T_\mathbf{g} - g^*) T_\mathbf{p} \\
-m & \equiv & {B_H}_\mathbf{c}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-\hypertarget{Upsilong}{And finally, we specify $\Upsilon^g$, the total gas used in this transaction}\hypertarget{Upsilonl}{ and $\Upsilon^\mathbf{l}$, the logs created by this transaction:}
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_\mathbf{g} - \mathbf{g}' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction Receipt}{transaction receipt}.
-
-\section{Contract Creation} \label{ch:create}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_\mathbf{n} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the \hyperlink{account nonce}{sender's nonce value}; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
-
-The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
-Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
-\begin{equation}
-\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except:}
-\end{equation}
-\begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_\mathbf{b} &\equiv& \boldsymbol{\sigma}[s]_\mathbf{b} - v
-\end{eqnarray}
-
-where $v'$ is the account's pre-existing value, in the event it was previously in existence:
-\begin{equation}
-v' \equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_\mathbf{b} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-%It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
-
-Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
-Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
-\end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
-\begin{eqnarray}
-I_\mathbf{a} & \equiv & a \\
-I_\mathbf{o} & \equiv & o \\
-I_\mathbf{p} & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
-I_\mathbf{s} & \equiv & s \\
-\hyperlink{I v}{I_\mathbf{v}} & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_\mathbf{e} & \equiv & e
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
-
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
-g^{**} - \mathbf{c} & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_\mathbf{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  \mathbf{g}^{**} < \mathbf{c} \ \vee\  |\mathbf{o}| > 24576\big)
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-Note that the 24576 byte limit for $\mathbf{o}$ exists because a contract creation call can trigger O(n) cost in terms of reading the code from disk, preprocessing the code for VM execution, and also adding O(n) data to the Merkle proof for the block's proof-of-validity.
-With higher gas limits that can be caused by dynamic gas limit rules, this is a greater concern, and is especially inconvenient with light clients verifying proofs of validity or invalidity.
-
-Note that the intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-
-\subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code.
-Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
-For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
-
-\section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
-
-Aside from evaluating to a new state and transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
-This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
-\end{equation}
-Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
-
-We define $\boldsymbol{\sigma}_1$, the first transitional state as the original state but with the value transferred from sender to recipient:
-\begin{equation}
-\boldsymbol{\sigma}_1[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_\mathbf{b} \equiv \boldsymbol{\sigma}[s]_\mathbf{b} - v
-\end{equation}
-unless $s = r$.
-
-Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and \hyperlink{account nonce}{nonce}.
-Thus the previous equation should be taken to mean:
-\begin{equation}
-\boldsymbol{\sigma}_1 \equiv \boldsymbol{\sigma}_1' \quad \text{except:} \\
-\end{equation}
-\begin{equation}
-\boldsymbol{\sigma}_1[s]_\mathbf{b} \equiv \boldsymbol{\sigma}_1'[s]_\mathbf{b} - v
-\end{equation}
-\begin{equation}
-\text{and}\quad \boldsymbol{\sigma}_1' \equiv \boldsymbol{\sigma} \quad \text{except:} \\
-\end{equation}
-\begin{equation}
-\begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_\mathbf{c}$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
-due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
-$\boldsymbol{\sigma}$).
-
-\begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-\boldsymbol{\sigma}^{**} & \text{otherwise}
-\end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_\mathbf{a} & \equiv & r \\
-I_\mathbf{o} & \equiv & o \\
-I_\mathbf{p} & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_\mathbf{s} & \equiv & s \\
-\hyperlink{I v}{I_\mathbf{v}} & \equiv & \tilde{v} \\
-I_\mathbf{e} & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_\mathbf{c}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
-
-As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
-
-Their full formal definition is in Appendix \ref{app:precompiled}.
-
-\section{Execution Model} \label{ch:model}
-
-The execution model specifies how the system state is altered given a series of bytecode instructions and a small tuple of environmental data.
-This is specified through a formal model of a virtual state machine, known as the Ethereum Virtual Machine (EVM).
-It is a \textit{quasi-}Turing-complete machine; the \textit{quasi} qualification comes from the fact that the computation is intrinsically bounded through a parameter, \textit{gas}, which limits the total amount of computation done.
-
-\subsection{Basics}
-
-The EVM is a simple stack-based architecture.
-The word size of the machine (and thus size of stack item) is 256-bit.
-This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations.
-The memory model is a simple word-addressed byte array.
-The stack has a maximum size of $1024$.
-The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array.
-Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state.
-All locations in both storage and memory are well-defined initially as zero.
-
-The machine does not follow the standard von Neumann architecture.
-Rather than storing program code in generally-accessible memory or storage, it is stored separately in a virtual ROM interactable only through a specialised instruction.
-
-The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas exception, they do not leave state changes intact.
-Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
-
-\subsection{Fees Overview}
-
-Fees (denominated in gas) are charged under three distinct circumstances, all three as prerequisite to the execution of an operation.
-The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
-Finally, gas may be paid due to an increase in the usage of the memory.
-
-Over an account's execution, the total fee for memory-usage payable is proportional to the smallest multiple of 32 bytes that is required such that all memory indices (whether for read or write) are included in the range.
-This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
-Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds.
-That said, implementations must be able to manage this eventuality.
-
-Storage fees have a slightly nuanced behaviour---to incentivise minimisation of the use of storage (which corresponds directly to a larger state database on all nodes), the execution fee for an operation that clears an entry in the storage is not only waived, a qualified refund is given; in fact, this refund is effectively paid up-front since the initial usage of a storage location costs substantially more than normal usage.
-
-See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
-
-\subsection{Execution Environment}
-
-In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
-
-\begin{itemize}\hypertarget{I a}{
-\item $I_\mathbf{a}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
-\item $I_\mathbf{o}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
-\item $I_\mathbf{p}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
-\item $I_\mathbf{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
-\item $I_\mathbf{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.}\hypertarget{I e}{
-\item $I_\mathbf{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
-\end{itemize}
-
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will defined it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, r)
-\end{equation}
-
-\subsection{Execution Overview}
-
-We must now define the $\Xi$ function.
-In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$.
-Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-
-\hypertarget{emptySequence}{}
-The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_\mathbf{g}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_\mathbf{g} & \equiv & \mathbf{g} \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_\mathbf{i} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
-\end{eqnarray}
-\begin{equation}
-X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
-\end{cases}
-\end{equation}
-
-where
-\begin{eqnarray}
-\mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_\mathbf{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
-
-\subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
-
-For the ease of reading, the instruction mnemonics, written in small-caps (\eg \space {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.
-
-For the purposes of defining $Z$, $H$ and $O$, we define $w$ as the current operation to be executed:
-\begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
-\text{\small STOP} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
-\subsubsection{Exceptional Halting}
-
-The exceptional halting function $Z$ is defined as:
-\begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
-\begin{array}[t]{l}
-\boldsymbol{\mu}_\mathbf{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
-The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-
-\subsubsection{Jump Destination Validity}
-
-We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run.
-We define this as any position in the code occupied by a {\small JUMPDEST} instruction.
-
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined {\small STOP} operations that trail it).
-
-Formally:
-\begin{equation}
-D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
-\end{equation}
-
-where:
-\begin{equation}
-D_J(\mathbf{c}, i) \equiv \begin{cases}
-\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
-\begin{equation}
-N(i, w) \equiv \begin{cases}
-i + w - \text{\small PUSH1} + 2 & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
-i + 1 & \text{otherwise} \end{cases}
-\end{equation}
-
-\subsubsection{Normal Halting} \label{subsec:Normal Halting}
-
-The normal halting function $H$ is defined:
-\begin{equation} \label{eq:Normal Halting}
-H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
-\varnothing & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The data-returning halt operation, \hyperlink{RETURN}{\text{\small RETURN}}, has a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{emptySequence}{here}.
-
-\subsection{The Execution Cycle}
-
-Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:
-\begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
-\end{eqnarray}
-
-The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
-\begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)  \\
-\hypertarget{JUMP}{} \hypertarget{JUMPI}{} \label{eq:JUMP} \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-In general, we assume the memory, self-destruct set and system state don't change:
-\begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_\mathbf{i} & \equiv & \boldsymbol{\mu}_\mathbf{i} \\
-A' & \equiv & A \\
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
-\end{eqnarray}
-
-However, instructions do typically alter one or several components of these values.
-Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and a formal description of the gas requirements.
-
-\section{Blocktree to Blockchain} \label{ch:ghost}
-
-The canonical blockchain is a path from root to leaf through the entire block tree.
-In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{GenesisBlock}{genesis block}, in the path.
-The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
-This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
-
-Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.
-
-Thus we define the total difficulty of block $B$ recursively as:
-\begin{eqnarray}
-B_\mathbf{t} & \equiv & B'_\mathbf{t} + B_\mathbf{d} \\
-B' & \equiv & P(B_H)
-\end{eqnarray}
-
-As such given a block $B$, $B_\mathbf{t}$ is its total difficulty, $B'$ is its parent block and $B_\mathbf{d}$ is its difficulty.
-
-\section{Block Finalisation} \label{ch:finalisation}
-
-The process of finalising a block involves four stages:
-
-\begin{enumerate}
-\item Validate (or, if mining, determine) ommers;
-\item validate (or, if mining, determine) transactions;
-\item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{nonce}.
-\end{enumerate}
-
-\subsection{Ommer Validation}
-
-The validation of \hyperlink{ommerheaders}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
-The maximum of ommer headers is two.
-Formally:
-\begin{equation}
-\lVert \hyperlink{B U}{B_\mathbf{U}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_\mathbf{U}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
-\end{equation}
-
-where $k$ denotes the ``is-kin'' property:
-\begin{equation}
-k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
-s(U, H) &\\
-\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-and $s$ denotes the ``is-sibling'' property:
-\begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
-\end{equation}
-where $B(H)$ is the block of the corresponding header $H$.
-
-\subsection{Transaction Validation}
-
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_\mathbf{g}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H g}{{B_H}_\mathbf{g}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathbf{u}}
-\end{equation}
-
-\subsection{Reward Application}
-
-The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_\mathbf{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-\hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
-\begin{eqnarray}
-\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_\mathbf{c}}]_\mathbf{b} & = & \boldsymbol{\sigma}[{B_H}_\mathbf{c}]_\mathbf{b} + (1 + \frac{\lVert \hyperlink{B U}{B_\mathbf{U}}\rVert}{32})\hyperlink{block reward R b}{R_\mathbf{b}} \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[U_\mathbf{c}]_\mathbf{b} + (1 + \frac{1}{8} (U_\mathbf{i} - \hyperlink{H i}{{B_H}_\mathbf{i}})) \hyperlink{block reward R b}{R_\mathbf{b}}
-\end{eqnarray}
-
-If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.\hypertarget{block reward R b}{}
-
-We define the block reward as 5 Ether:
-\begin{equation}
-\text{Let} \quad R_\mathbf{b} = 5 \times 10^{18}
-\end{equation}
-
-\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
-
-\hypertarget{Gamma}{
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
-\begin{equation}
-\Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
-\end{cases}
-\end{equation}
-}
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-\hypertarget{Phi}{
-And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
-\begin{eqnarray}
-\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_\mathbf{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
-B'_\mathbf{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_\mathbf{r}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
-
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$ is the state-transition function}, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_\mathbf{u}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_\mathbf{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
-The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction Receipt}{${B_\mathbf{R}}[n]_\mathbf{u}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_\mathbf{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
-\begin{array}[b]{l}
-\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathbf{u}
-\end{array}
- & \text{otherwise} \end{cases}
-\end{equation}
-
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-\hypertarget{Pi}{
-Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
-\end{equation}}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
-
-\subsection{Mining Proof-of-Work} \label{ch:pow}
-
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
-However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
-
-For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of, or reward from, specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
-
-Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
-Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
-
-One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{ASICbitcoinwiki}).
-While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
-Because of this, a proof-of-work function that is ASIC-resistant (i.e.
-difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
-
-Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the \hyperlink{block nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum \textit{Frontier} and \textit{Homestead} we have chosen the first path.
-
-More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
-\begin{equation}
-m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathbf{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathbf{d}}$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being the \hyperlink{block nonce}{block nonce}, a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
-\subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum \textit{Frontier} and \textit{Homestead}.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{cacheinit}$} bytes in initial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{datasetinit}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every \hyperlink{Jepoch}{$J_{epoch}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
-
-\section{Implementing Contracts}
-
-There are several patterns of contracts engineering that allow particular useful behaviours; two of these that I will briefly discuss are data feeds and random numbers.
-
-\subsection{Data Feeds}
-A data feed contract is one which provides a single service: it gives access to information from the external world within Ethereum.
-The accuracy and timeliness of this information is not guaranteed and it is the task of a secondary contract author---the contract that utilises the data feed---to determine how much trust can be placed in any single data feed.
-
-The general pattern involves a single contract within Ethereum which, when given a message call, replies with some timely information concerning an external phenomenon.
-An example might be the local temperature of New York City.
-This would be implemented as a contract that returned that value of some known point in storage.
-Of course this point in storage must be maintained with the correct such temperature, and thus the second part of the pattern would be for an external server to run an Ethereum node, and immediately on discovery of a new block, creates a new valid transaction, sent to the contract, updating said value in storage.
-The contract's code would accept such updates only from the identity contained on said server.
-
-\subsection{Random Numbers}
-Providing random numbers within a deterministic system is, naturally, an impossible task.
-However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
-For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
-
-\section{Future Directions} \label{ch:future}
-
-The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
-
-Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{GenesisBlock}{genesis block}.
-This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
-
-Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
-
-\subsection{Scalability}
-
-Scalability remains an eternal concern.
-With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy.
-Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger.
-However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.
-
-Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidating each block and with duplicate or invalid transactions thrown out accordingly.
-
-Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
-
-\section{Conclusion} \label{ch:conclusion}
-
-I have introduced, discussed and formally defined the protocol of Ethereum.
-Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
-Contracts may be authored in order to algorithmically specify and autonomously enforce rules of interaction.
-
-\section{Acknowledgements}
-
-Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
-
-\bibliography{Biblio}
-\bibliographystyle{plainnat}
-
-\end{multicols}
-
-\appendix
-
-\section{Terminology} \label{ch:Terminology}
-
-\begin{description}
-\item[External Actor] A person or other entity able to interface to an Ethereum node, but external to the world of Ethereum.
-It can interact with Ethereum through depositing signed Transactions and inspecting the blockchain and associated state.
-Has one (or more) intrinsic Accounts.
-
-\item[Address] A 160-bit code used for identifying Accounts.
-
-\item[Account] Accounts have an intrinsic balance and transaction count maintained as part of the Ethereum state.
-They also have some (possibly empty) EVM Code and a (possibly empty) Storage State associated with them.
-Though homogenous, it makes sense to distinguish between two practical types of account: those with empty associated EVM Code (thus the account balance is controlled, if at all, by some external entity) and those with non-empty associated EVM Code (thus the account represents an Autonomous Object).
-Each Account has a single Address that identifies it.
-
-\item[Transaction] A piece of data, signed by an External Actor. It represents either a Message or a new Autonomous Object. Transactions are recorded into each block of the blockchain.
-
-\item[Autonomous Object] A notional object existent only within the hypothetical state of Ethereum.
-Has an intrinsic address and thus an associated account; the account will have non-empty associated EVM Code.
-Incorporated only as the Storage State of that account.
-
-\item[Storage State] The information particular to a given Account that is maintained between the times that the Account's associated EVM Code runs.
-
-\item[Message] Data (as a set of bytes) and Value (specified as Ether) that is passed between two Accounts, either through the deterministic operation of an Autonomous Object or the cryptographically secure signature of the Transaction.
-
-\item[Message Call] The act of passing a message from one Account to another.
-If the destination account is associated with non-empty EVM Code, then the VM will be started with the state of said Object and the Message acted upon.
-If the message sender is an Autonomous Object, then the Call passes any data returned from the VM operation.
-
-\item[Gas] The fundamental network cost unit.
-Paid for exclusively by Ether (as of PoC-4), which is converted freely to and from Gas as required.
-Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and miners are free to ignore Transactions whose Gas price is too low.
-
-\item[Contract] Informal term used to mean both a piece of EVM Code that may be associated with an Account or an Autonomous Object.
-
-\item[Object] Synonym for Autonomous Object.
-
-\item[App] An end-user-visible application hosted in the Ethereum Browser.
-
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
-
-\item[Ethereum Virtual Machine] (aka EVM) The virtual machine that forms the key part of the execution model for an Account's associated EVM Code.
-
-\item[Ethereum Runtime Environment] (aka ERE) The environment which is provided to an Autonomous Object executing in the EVM.
-Includes the EVM but also the structure of the world state on which the EVM relies for certain I/O instructions including CALL \& CREATE.
-
-\item[EVM Code] The bytecode that the EVM can natively execute. Used to formally specify the meaning and ramifications of a message to an Account.
-
-\item[EVM Assembly] The human-readable form of EVM-code.
-
-\item[LLL] The Lisp-like Low-level Language, a human-writable language used for authoring simple contracts and general low-level language toolkit for trans-compiling to.
-
-\end{description}
-
-\hypertarget{rlp}{}
-\section{Recursive Length Prefix}\label{app:rlp}
-This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
-
-We define the set of possible structures $\mathbb{T}$:
-\begin{eqnarray}
-\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
-Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
-
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_\mathbf{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_\mathbf{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
-
-\begin{itemize}
-\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
-\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
-\end{itemize}
-\hypertarget{R b}{
-Formally, we define $R_\mathbf{b}$:}
-\begin{eqnarray}
-R_\mathbf{b}(\mathbf{x}) & \equiv & \begin{cases}
-\mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
-(128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
-\end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
-\end{eqnarray}
-
-Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
-
-If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
-
-\begin{itemize}
-\item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
-\end{itemize}
-\hypertarget{R l}{
-Thus we finish by formally defining $R_\mathbf{l}$:
-\begin{eqnarray}
-R_\mathbf{l}(\mathbf{x}) & \equiv & \begin{cases}
-(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
-\end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
-\end{eqnarray}
-}
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
-Thus the RLP of some positive integer $i$ is defined as:
-\begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
-\end{equation}
-
-When interpreting RLP data, if an expected fragment is decoded as a scalar and leading zeroes are found in the byte sequence, clients are required to consider it non-canonical and treat it in the same manner as otherwise invalid RLP data, dismissing it completely.
-
-There is no specific canonical encoding format for signed or floating-point values.
-
-\section{Hex-Prefix Encoding}\label{app:hexprefix}
-Hex-prefix encoding is an efficient method of encoding an arbitrary number of nibbles as a byte array.
-It is able to store an additional flag which, when used in the context of the trie (the only context in which it is used), disambiguates between node types.
-
-It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
-(16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
-\text{if} \quad \lVert \mathbf{x} \rVert \; \text{is even} \\
-(16(f(t) + 1) + \mathbf{x}[0], 16\mathbf{x}[1] + \mathbf{x}[2], 16\mathbf{x}[3] + \mathbf{x}[4], ...) &
-\text{otherwise}
-\end{cases} \\
-f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \neq 0 \\ 0 & \text{otherwise} \end{cases}
-\end{eqnarray}
-
-Thus the high nibble of the first byte contains two flags; the lowest bit encoding the oddness of the length and the second-lowest encoding the flag $t$.
-The low nibble of the first byte is zero in the case of an even number of nibbles and the first nibble in the case of an odd number.
-All remaining nibbles (now an even number) fit properly into the remaining bytes.
-
-\hypertarget{trie}{}
-\section{Modified Merkle Patricia Tree}\label{app:trie}
-The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
-It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
-
-Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
-\begin{equation}
-\mathfrak{I} = \{ (k_0 \in \mathbb{B}, v_0 \in \mathbb{B}), k_1 \in \mathbb{B}, v_1 \in \mathbb{B}), ... \}
-\end{equation}
-
-When considering such a sequence, we use the common numeric subscript notation to refer to a tuple's key or value, thus:
-\begin{equation}
-\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
-\end{equation}
-
-Any series of bytes may also trivially be viewed as a series of nibbles, given an endian-specific notation; here we assume big-endian. Thus:
-\begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (k_0' \in \mathbb{Y}, v_0 \in \mathbb{B}), (k_1' \in \mathbb{Y}, v_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\ k_n\rVert} \quad k_n'[i] & \equiv &
-\begin{cases}
-\lfloor k_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-k_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
-\begin{equation}
-\texttt{\small TRIE}(\mathfrak{I}) \equiv \texttt{\small KEC}(c(\mathfrak{I}, 0))
-\end{equation}
-
-We also assume a function $n$, the trie's node cap function.
-When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
-Thus we define in terms of $c$, the node composition function:
-\begin{equation}
-n(\mathfrak{I}, i) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-In a manner similar to a radix tree, when the trie is traversed from root to leaf, one may build a single key-value pair.
-The key is accumulated through the traversal, acquiring a single nibble from each branch node (just as with a radix tree).
-Unlike a radix tree, in the case of multiple keys sharing the same prefix or in the case of a single key having a unique suffix, two optimising nodes are provided.
-Thus while traversing, one may potentially acquire multiple nibbles from each of the other two node types, extension and leaf.
-There are three kinds of nodes in the trie:
-\begin{description}
-\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
-\item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
-\item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
-The 17th item is used in the case of this being a terminator node and thus a key being ended at this point in its traversal.
-\end{description}
-
-A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
-\begin{equation}
-c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
-(x - 1)] = \mathbf{l} \\
-\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
-v & = & \begin{cases}
-I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
-() & \text{otherwise}
-\end{cases}
-\end{array}
-\end{cases}
-\end{equation}
-
-\subsection{Trie Database}
-Thus no explicit assumptions are made concerning what data is stored and what is not, since that is an implementation-specific consideration; we simply define the identity function mapping the key-value set $\mathfrak{I}$ to a 32-byte hash and assert that only a single such hash exists for any $\mathfrak{I}$, which though not strictly true is accurate within acceptable precision given the Keccak hash's collision resistance.
-In reality, a sensible implementation will not fully recompute the trie root hash for each set.
-
-A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
-This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
-Due to the dependency relationship, Merkle-proofs may be constructed with an $O(\log N)$ space requirement that can demonstrate a particular leaf must exist within a trie of a given root hash.
-
-\section{Precompiled Contracts}\label{app:precompiled}
-
-For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_\mathbf{r} \\
-(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_\mathbf{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_\mathbf{r}$, the gas requirements.
-
-For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
-\begin{eqnarray}
-\Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
-\text{if} \quad |\mathbf{o}| = 32: &&\\
-\mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
-\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
-h &=& \mathbf{d}[0..31]\\
-v &=& \mathbf{d}[32..63]\\
-r &=& \mathbf{d}[64..95]\\
-s &=& \mathbf{d}[96..127]
-\end{eqnarray}
-
-The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
-Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
-\begin{eqnarray}
-\Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
-\Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
-\end{eqnarray}
-
-For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
-\begin{eqnarray}
-\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
-\end{eqnarray}
-
-Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
-\begin{eqnarray}
-\Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
-\end{eqnarray}
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
-This method utilises the SECP-256k1 curve as described by \cite{secp256k1BitcoinWiki2016} and \cite{secp256k1StackExchange2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.
-9 of 15, para.
-3.
-
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom} and \cite{ECDSAWikipedia}.
-\footnote{For the former citation, refer to section 6.2 for ECDSAPUBKEY, and section 7 for ECDSASIGN and ECDSARECOVER.} %(assuming that the former is for signature generation and the latter is for signature verification)
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range)\hypertarget{v}{.
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28\}
-\end{align}
-where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
-
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_\mathbf{r}$, $T_\mathbf{s}$ and $\hyperlink{T w}{T_\mathbf{w}}$:
-\begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}) & \text{if} \; T_\mathbf{t} = 0\\
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}) & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
-\end{eqnarray}
-
-\hypertarget{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except:} \\
-(\hyperlink{T w}{T_\mathbf{w}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}}
-
-We may then define the sender function $S$ of the transaction as:
-\begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) \big) \big)
-\end{equation}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
-\begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
-\end{equation}
-
-\newpage
-\section{Fee Schedule}\label{app:fees}
-
-The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
 
@@
 \toprule
-Name & Value & Description* \\
+Foo \\
 \midrule 
-$G_{zero}$ & 0 & Nothing is paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & This is paid for an {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
-$R_{selfdestruct}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
-$G_{selfdestruct}$ & 5000 & This is the amount of gas to pay for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & This is paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
-$G_{call}$ & 700 & This is paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & This is a stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
-$G_{newaccount}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & This is paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & This is paid for every transaction. \\
-$G_{log}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & This is paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
-$G_{blockhash}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\
-
-%extern u256 const \mathbf{c}_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
-\bottomrule
-\end{tabularx}
-
-\section{Virtual Machine Specification}\label{app:vm}
-
-When interpreting 256-bit binary values as integers, the representation is big-endian.
-
-When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the left most 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
-
-\subsection{Gas Cost}
-
-The general gas cost function, $C$, is defined as:
-
-\begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathbf{i})-C_{mem}(\boldsymbol{\mu}_\mathbf{i}) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
-\text{\small STOP} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where:
-\begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with \hyperlink{tiny CALL}{$C_\text{\tiny CALL}$}, \hyperlink{C tiny SELFDESTRUCT}{$C_\text{\tiny SELFDESTRUCT}$} and \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$} as specified in the appropriate section below.
-We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{high}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write.
-Such accesses must be for non-zero number of bytes.
-
-Referencing a zero length range (e.g.
-by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_\mathbf{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
-
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
-
-While defining the instruction set, we defined the memory-expansion for range function, $M$, thus:
-
-\begin{equation}
-M(s, f, l) \equiv \begin{cases}
-s & \text{if} \quad l = 0 \\
-\max(s, \ceil{ (f + l) \div 32 }) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Another useful function is ``all but one 64th'' function~$L$ defined as:
-
-\begin{equation}
-L(n) \equiv n - \lfloor n / 64 \rfloor
-\end{equation}
-
-\subsection{Instruction Set}
-
-As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
-In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
-\begin{equation}
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \quad \text{with exceptions, as noted}
-\end{equation}
-
-Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from the stack, as defined in section \ref{ch:model}.
-
-%\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & This operation halts execution, outputting the empty sequence as per equation \ref{eq:Normal Halting}.
-\\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & This is the addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x02 & {\small MUL} & 2 & 1 & This is the multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x03 & {\small SUB} & 2 & 1 & This is the subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x04 & {\small DIV} & 2 & 1 & This is the integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-\midrule
-0x05 & {\small SDIV} & 2 & 1 & This is the signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-&&&& Note the overflow semantic when $-2^{255}$ is negated.\\
-\midrule
-0x06 & {\small MOD} & 2 & 1 & This is the modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
-\midrule
-0x07 & {\small SMOD} & 2 & 1 & This is the signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-\midrule
-0x08 & {\small ADDMOD} & 3 & 1 & This is the modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x09 & {\small MULMOD} & 3 & 1 & This is the modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x0a & {\small EXP} & 2 & 1 & This is the exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
-\midrule
-0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_\mathbf{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{10s: Comparison \& Bitwise Logic Operations}} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & This is the less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x11 & {\small GT} & 2 & 1 & This is the greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x12 & {\small SLT} & 2 & 1 & This is the signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-\midrule
-0x13 & {\small SGT} & 2 & 1 & This is the signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
-\midrule
-0x14 & {\small EQ} & 2 & 1 & This is the equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x15 & {\small ISZERO} & 1 & 1 & This is the logical negation operation, also called the logical complement or the NOT operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
-\midrule
-0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \vee \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
-\midrule
-0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
-\midrule
-0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute a Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{a}$ \\
-\midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_\mathbf{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
-\midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{o}$ \\
-&&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
-\midrule
-0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{s}$ \\
-&&&& This is the address of the account that is directly responsible for this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \hyperlink{I v}{I_\mathbf{v}}$ \\
-\midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get the input data of the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get the size of the input data in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy the input data in the current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get the size of code running in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
-\midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in the current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{p}$ \\
-&&&& This is the gas price specified by the originating transaction.\\
-\midrule
-0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{c} \rVert$ \\
-\midrule
-0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{c}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\hyperlink{H p}{H_\mathbf{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
-0 is left on the stack if the looked for block number is greater than the current block number or more than 256 blocks behind the current block.
-\\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_\mathbf{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_\mathbf{i}} \\ P(\hyperlink{H p}{H_\mathbf{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
-\midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{c}$ \\
-\midrule
-0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{s}$ \\
-\midrule
-0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{i}$ \\
-\midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{d}$ \\
-\midrule
-0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{l}$ \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove the top, first item from the stack. \\
-\midrule
-0x51 & {\small MLOAD} & 1 & 1 & Load the first word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x52 & {\small MSTORE} & 2 & 0 & Save the first word (which is the first item in the stack) to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x53 & {\small MSTORE8} & 2 & 0 & Save the first byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
-\midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
-&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
-\end{cases}$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\midrule
-0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
-\midrule
-0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
-\midrule
-0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
-\midrule
-0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
-\midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
-\midrule
-0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
-&&&& This operation has no effect on the machine state during execution. \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place a 1 byte item on the stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
-&&&& The bytes are read in line from the program code's bytes array. \\
-&&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
-&&&& The byte is right-aligned (\ie it takes the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place a 2-byte item on the stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
-&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
-&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
-&&&& where $\boldsymbol{c}$ is defined as above. \\
-&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{80s: Duplication Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x80 & {\small DUP1} & 1 & 2 & Duplicate the 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x81 & {\small DUP2} & 2 & 3 & Duplicate the 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x8f & {\small DUP16} & 16 & 17 & Duplicate the 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{90s: Exchange Operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x90 & {\small SWAP1} & 2 & 2 & Exchange the 1st and the 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x91 & {\small SWAP2} & 3 & 3 & Exchange the 1st and the 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0x9f & {\small SWAP16} & 17 & 17 & Exchange the 1st and the 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_\mathbf{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
-\multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
-\multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xa0 & {\small LOG0} & 2 & 0 & Append the log record with no topics. \\
-&&&& $\mathbf{t} \equiv ()$ \\
-\midrule
-0xa1 & {\small LOG1} & 3 & 0 & Append the log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
-\midrule
-\multicolumn{1}{c}{\vdots} & \multicolumn{1}{c}{\vdots} & \vdots & \vdots & \multicolumn{1}{c}{\vdots} \\
-\midrule
-0xa4 & {\small LOG4} & 6 & 0 & Append the log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
-\bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\toprule
-\multicolumn{5}{c}{\textbf{f0s: System operations}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_\mathbf{g}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_\mathbf{a}, I_\mathbf{o}, L(\boldsymbol{\mu}_\mathbf{g}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_\mathbf{e} + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_\mathbf{a}]_\mathbf{n} = \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{n} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_\mathbf{e} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_\mathbf{a}, \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{n})$, the address of the  \\
-&&&& newly created account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& Thus the operand order is: value, input offset, input size. \\
-\midrule
-0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_\mathbf{a}, I_\mathbf{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge \\ \quad\quad I_\mathbf{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{g} \equiv \boldsymbol{\mu}_\mathbf{g} + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ (not enough funds) or $I_\mathbf{e} = 1024$ (call depth limit reached); $x=1$ \\
-&&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_\mathbf{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_\mathbf{s}[0] \} & \text{if} \quad \boldsymbol{\mu}_\mathbf{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{a}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\\ \quad\quad{}I_\mathbf{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_\mathbf{a}$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
-\midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
-&&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See equation \ref{eq:Normal Halting}. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
-Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
-The omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$.
-Otherwise it is exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{s}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_\mathbf{p}, 0, \hyperlink{I v}{I_\mathbf{v}}, \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathbf{v}} \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
-This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{SELFDESTRUCT}{}\\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_\mathbf{a} \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} + \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ \\
-&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{b} \equiv 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
-0 & \text{otherwise} 
-\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
-&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{selfdestruct}} + \begin{cases}
-\hyperlink{Gnewaccount}{G_{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\bottomrule
-\end{tabularx}
-
-%\section{Low-level Lisp-like Language}\label{app:lll}
-%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-
-\hypertarget{GenesisBlock}{}
-\section{Genesis Block}\label{app:genesis}
-
-The genesis block is 15 items, and is specified thus:
-\begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the \hyperlink{H c}{beneficiary address}, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the \hyperlink{H b}{log bloom}, a 2048-bit sequence of all zeros; $2^{17}$ refers to the \hyperlink{H d}{difficulty}; the \hyperlink{H t}{transaction trie root}, \hyperlink{H e}{receipt trie root}, \hyperlink{H g}{gas used}, \hyperlink{H i}{block number} and \hyperlink{H x}{extradata} are all $0$ (listed in the order of decimal zeros in the above sequence), being equivalent to the empty byte array.
-The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the \hyperlink{block nonce}{nonce}.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in \hyperlink{rlp}{RLP}, both empty lists.
-
-The proof-of-concept series includes a development premine, making the state root hash some value $stateRoot$.
-Also \hyperlink{H s}{$time$} will be set to the initial timestamp of the genesis block.
-The latest documentation should be consulted for those values.
-
-\section{Ethash}\label{app:ethash}
-\subsection{Definitions}
-We employ the following definitions:
-
-\begin{tabularx}{\textwidth}{rlX}
-\toprule
-Name & Value & Description \\
-\midrule 
-\hypertarget{Jwordbytes}{
-$J_{wordbytes}$ & 4  & This is the bytes in a word. }\hypertarget{Jdatasetinit}{\\
-$J_{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
-$J_{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
-$J_{cacheinit}$ & $2^{24}$ & This is the bytes in the cache at genesis. }\hypertarget{Jcachegrowth}{\\
-$J_{cachegrowth}$ & $2^{17}$ & This is the cache growth per epoch. }\hypertarget{Jepoch}{\\
-$J_{epoch}$ & 30000 & This is the blocks per epoch. }\hypertarget{Jmixbytes}{\\
-$J_{mixbytes}$ & 128 & This is the mix length in bytes. }\hypertarget{Jhashbytes}{\\
-$J_{hashbytes}$ & 64 & This is the hash length in bytes. }\hypertarget{Jparents}{\\
-$J_{parents}$ & 256 & This is the number of parents of each dataset element. }\hypertarget{Jcacherounds}{\\
-$J_{cacherounds}$ & 3 & This is the number of rounds in cache production. }\hypertarget{Jaccesses}{\\
-$J_{accesses}$ & 64 & This is the number of accesses in a hashimoto loop. }\\
-\bottomrule
-\end{tabularx}
-
-\subsection{Size of dataset and cache}
-The size for Ethash's cache $\mathbf{c} \in \mathbb{B}$  and dataset $\mathbf{d} \in \mathbb{B}$ depend on the epoch, which in turn depends on the block number.
-\begin{equation}
- E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{epoch}}}\right\rfloor
-\end{equation}
-The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{cachegrowth}$} bytes, where the growth of each occurs every epoch.
-In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{mixbytes}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{hashbytes}$} for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{datasetgrowth}} \cdot E_{epoch} - \hyperlink{Jmixbytes}{J_{mixbytes}}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{cacheinit}} + \hyperlink{cachegrowth}{J_{cachegrowth}} \cdot E_{epoch} - \hyperlink{Jhashbytes}{J_{hashbytes}}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
-\end{cases}
-\end{equation}
-\subsection{Dataset generation}
-In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $s \in \mathbb{B}_{32}$.
-\subsubsection{Seed hash}
-The seed hash is different for every epoch.
-For the first epoch it is the Keccak-256 hash of a series of 32 bytes of zeros.
-For every other epoch it is always the Keccak-256 hash of the previous seed hash:
-\begin{equation}
- \mathbf{s} = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
-\end{equation}
-\begin{equation}
- C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{epoch}})) & \text{otherwise}
-\end{cases}
-\end{equation}
-With $\mathbf{0}_{32}$ being 32 bytes of zeros.
-
-\subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially fil up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{cacherounds}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
-The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
-
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
-\texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
-\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
-\end{cases}
-\end{equation}
-Therefore $ \mathbf{c'}$ can be defined as
-\begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{cacherounds}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{cacherounds}})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
-\mathbf{x} & \text{if} \quad y = 0 \quad  \\
-E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
-\end{cases}
-\end{equation}
-Where a single round modifies each subset of the cache as follows:
-\begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
-\end{equation}
-\begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
-\end{multline}
-
-\subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{hashbytes}$} bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
-\end{equation}
-In order to calculate the single item we use an algorithm inspired by the FNV hash (\cite{FowlerNollVo1991FNVHash}) in some cases as a non-associative substitute for XOR.
-\begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
-\end{equation}
-The single item of the dataset can now be calculated as:
-\begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{hashbytes}} / \hyperlink{Jwordbytes}{J_{wordbytes}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\subsection{Proof-of-work function}
-Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-\hyperlink{Jmixbytes}{$J_{mixbytes}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the \hyperlink{block nonce}{nonce} is valid.
-Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
-It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
-
-The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
-\begin{equation}
- \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
-\end{equation}
-
-The seed hash being:
-\begin{equation}
- \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
-\end{equation}
-We note that the ``$+$''-operator between two byte sequences results in the concatenation of both sequences.
-
-The dataset $\mathbf{d}$ is obtained as described in section \ref{dataset}.
-
-The number of replicated sequences in the mix is:
-\begin{equation}
- n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{mixbytes}}}{\hyperlink{Jhashbytes}{J_{hashbytes}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{accesses}} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{mixbytes}}}{\hyperlink{Jwordbytes}{J_{wordbytes}}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / \hyperlink{Jhashbytes}{J_{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
-\end{equation}
-The mix is compressed as follows:
-\begin{equation}
- E_{compress}(\mathbf{m}, i) = \begin{cases}
-\mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabularx}{\textwidth}{lcX}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\hypertarget{bigvee}{
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
-Thus it is the greatest element of such elements.
-More information is here: \cite{LatticeOrder}.}\\
+$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{\\
+Foo}\hyperlink{Jdatasetinit\\
+$ This doesn't have a closing bracket so will cause the 
 \bottomrule

### 2017-09-30 Paper.tex
This file actually builds and is the correct one;  the last one was an extract for testing purposes.  Changes made include changing \rm font to \mathrm,  fixing syntax errors from the previous failing build;  (e.g. uneven braces), adding cross-references,  extra citations (e.g. for Ethash and ASICs), a list of maths symbols which I've just added just one item so far—\bigvee.

--- 2017-09-29 Paper.tex
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{\rm{cacheinit}}$} bytes in initial size.
+From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{\mathrm{cacheinit}}$} bytes in initial size.
 Light clients store the cache.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every \hyperlink{Jepoch}{$J_{\rm{epoch}}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every \hyperlink{Jepoch}{$J_{\mathrm{epoch}}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 %\begin{tabu}{X[,l]X[,r]X[,l]} 
-
 \toprule
-Name & Value & Description\hypertarget{Jwordbytes}{}\\
-\midrule 
-$J_{\rm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{\\
-$J_{\rm{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis.}\hypertarget{Jdatasetgrowth}{\\
-$J_{\rm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch.}\hypertarget{Jcacheinit}{\\
-$J_{\rm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
-$J_{\rm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
-$J_{\rm{epoch}}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
-$J_{\rm{mixbytes}}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
-$J_{\rm{hashbytes}}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
-$J_{\rm{parents}}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
-$J_{\rm{cacherounds}}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
-$J_{\rm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
+Name & Value & Description \hypertarget{Jwordbytes}{} \\
+\midrule 
+$J_\mathrm{wordbytes}$ & 4  & This is the bytes in a word. \hypertarget{Jdatasetinit}{} \\
+$J_\mathrm{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{} \\
+$J_\mathrm{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch.  \hypertarget{Jcacheinit}{\\
+$J_\mathrm{cacheinit}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
+$J_\mathrm{cachegrowth}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
+$J_\mathrm{epoch}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
+$J_\mathrm{mixbytes}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
+$J_\mathrm{hashbytes}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
+$J_\mathrm{parents}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
+$J_\mathrm{cacherounds}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
+$J_\mathrm{accesses}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
 \bottomrule
@@
 \begin{equation}
-E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{\rm{epoch}}}}\right\rfloor
-\end{equation}
-The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\rm{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\rm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
+E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_\mathrm{epoch}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_\mathrm{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\rm{hashbytes}}$} for the cache.\hypertarget{dsize}{
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} for the cache.\hypertarget{dsize}{
 Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
 \begin{equation}
- d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\rm{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\rm{datasetgrowth}} \cdot E_{\rm{epoch}} - \hyperlink{Jmixbytes}{J_\{rm{mixbytes}}}, J_{\rm{mixbytes}})
+ d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_\{rm{mixbytes}}}, J_{\mathrm{mixbytes}})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\rm{cacheinit}}} + \hyperlink{\rm{cachegrowth}}{J_{\rm{cachegrowth}}} \cdot E_{\rm{epoch}} - \hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}, J_{\rm{hashbytes}})
+ c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{\mathrm{cachegrowth}}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
 \end{equation}
@@
  C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\rm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\rm{epoch}}})) & \text{otherwise}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\rm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\rm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{\rm{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{\rm{cacherounds}}})
-\end{equation}
-\begin{equation}
- E_{\rm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{\mathrm{cacherounds}}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{\rm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{\rm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{\rm{hashbytes}}$} bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
+Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{\rm{parents}}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{\rm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\rm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\rm{parents}} -2 \\
+ E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\rm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\rm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-\hyperlink{Jmixbytes}{$J_{\rm{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+\hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\rm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\rm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{\rm{accesses}}$ function is used:
-\begin{equation}
- E_{\rm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\rm{accesses}}} -2 \\
-E_{\rm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+ n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
+\begin{equation}
+ E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\mathrm{accesses}}} -2 \\
+E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\rm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\rm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\rm{hashbytes}}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}

### 2017-09-30 Paper.tex
The citations were rendering with a ?, so I fixed that by checking the output, finding that there were three errors with cross-references and fixed them. I also fixed some syntax for the Ethash functions, which while they haven't caused the build to fail, isn't good practice to have _\mathrm, so I changed this to _{\mathrm} and also fixed a couple of other errors.

@@
 
-The values stemming from the computation of transactions, specifically the \hyperlink{transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{stateR-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{stateR-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_\mathbf{n} & = & \boldsymbol{\sigma}[S(T)]_\mathbf{n} \quad \wedge \\
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction nonce}{\\
+T_\mathbf{n} & = & \boldsymbol{\sigma}[S(T)]_\mathbf{n} \quad \wedge \\}
 g_0 & \leqslant & T_\mathbf{g} \quad \wedge \\
@@
 Light clients store the cache.
-From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{\rm{datasetinit}}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{\mathrm{datasetinit}}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 \toprule
-Name & Value & Description \hypertarget{Jwordbytes}{} \\
-\midrule 
-$J_\mathrm{wordbytes}$ & 4  & This is the bytes in a word. \hypertarget{Jdatasetinit}{} \\
-$J_\mathrm{datasetinit}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{} \\
-$J_\mathrm{datasetgrowth}$ & $2^{23}$ & This is the dataset growth per epoch.  \hypertarget{Jcacheinit}{\\
-$J_\mathrm{cacheinit}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
-$J_\mathrm{cachegrowth}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
-$J_\mathrm{epoch}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
-$J_\mathrm{mixbytes}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
-$J_\mathrm{hashbytes}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
-$J_\mathrm{parents}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
-$J_\mathrm{cacherounds}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
-$J_\mathrm{accesses}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
+Name & Value & Description \hypertarget{Jwordbytes}{ \\
+\midrule 
+$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word. }\hypertarget{Jdatasetinit}{ \\
+$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{ \\
+$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
+$J_{\mathrm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
+$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
+$J_{\mathrm{epoch}}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
+$J_{\mathrm{mixbytes}}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
+$J_{\mathrm{hashbytes}}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
+$J_{\mathrm{parents}}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
+$J_{\mathrm{cacherounds}}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
+$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
 \bottomrule
@@
 \begin{equation}
-E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_\mathrm{epoch}}}\right\rfloor
-\end{equation}
-The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_\mathrm{datasetgrowth}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
+E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\mathrm{datasetgrowth}}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
@@
 \begin{equation}
- d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_\{rm{mixbytes}}}, J_{\mathrm{mixbytes}})
+ d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{\mathrm{cachegrowth}}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
+ c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
 \end{equation}
@@
 \begin{equation}
- \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, \hyperlink{Jcacherounds}{J_{\mathrm{cacherounds}}})
+ \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
 \end{equation}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-\hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch $J_{\mathrm{mixbytes}}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{\mathrm{mixbytes}}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 

### 2017-09-30 Paper.tex
Changed subscript to \mathrm e.g. for gas costs (using Word with Find and replace is much faster), added a citation for Mist.

@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{\mathrm{txdatazero}} & \text{if} \quad i = 0 \\ G_{\mathrm{txdatanonzero}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_{\mathrm{txcreate}} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{\mathrm{transaction}}
 \end{align}
@@
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 \begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
+c \equiv G_{\mathrm{codedeposit}} \times |\mathbf{o}|
 \end{equation}
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+If the initialisation execution ends with a \hyperlink{selfdestruct}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
+() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
 \varnothing & \text{otherwise}
@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, which is known as Mist since 8 July 2016 (\cite{Mist}).
 
@@
 \midrule
-$G_{zero}$ & 0 & Nothing is paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & This is paid for an {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
-$R_{selfdestruct}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
-$G_{selfdestruct}$ & 5000 & This is the amount of gas to pay for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & This is paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
-$G_{call}$ & 700 & This is paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & This is a stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
-$G_{newaccount}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & This is paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & This is paid for every transaction. \\
-$G_{log}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & This is paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
-$G_{blockhash}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\
+$G_{\mathrm{zero}}$ & 0 & Nothing is paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{base}}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
+$G_{\mathrm{low}}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{\mathrm{mid}}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
+$G_{\mathrm{\mathrm{high}}}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{\mathrm{high}}}$}. \\
+$G_{\mathrm{extcode}}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{\mathrm{balance}}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{\mathrm{sload}}$ & 200 & This is paid for an {\small SLOAD} operation. \\
+$G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{sset}}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
+$R_{\mathrm{selfdestruct}}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
+$G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \\
+$G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
+$G_{\mathrm{call}}$ & 700 & This is paid for a {\small CALL} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{\mathrm{callstipend}}$ & 2300 & This is a stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
+$G_{\mathrm{newaccount}}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \\
+$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
+$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
+$G_{\mathrm{transaction}}$ & 21000 & This is paid for every transaction. \\
+$G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
+$G_{\mathrm{sha3}}$ & 30 & This is paid for each {\small SHA3} operation. \\
+$G_{\mathrm{sha3word}}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{copy}}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\
 
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
+G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
+G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
+G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
+G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
+G_{\mathrm{\mathrm{high}}} & \text{if} \quad w \in W_{\mathrm{\mathrm{high}}}\\
+G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small BLOCKHASH}\\
 \end{cases}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{mem}(a) \equiv G_{\mathrm{memory}} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{high}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write.
+$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}\}
+
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{\mathrm{\mathrm{high}}}$ = \{{\small JUMPI}\}
+
+$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
+G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{\mathrm{sreset}} & \text{otherwise}
 \end{cases}$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{SELFDESTRUCT}{}\\
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{\mathrm{selfdestruct}}{}\\
 \midrule
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-\hyperlink{Rselfdestruct}{R_{selfdestruct}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
+\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
 0 & \text{otherwise} 
 \end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
-&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{selfdestruct}} + \begin{cases}
-\hyperlink{Gnewaccount}{G_{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{\mathrm{selfdestruct}}} + \begin{cases}
+\hyperlink{Gnewaccount}{G_{\mathrm{newaccount}}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 
+%Is in the terminology list.
 %\section{Low-level Lisp-like Language}\label{app:lll}
@@
 Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} for the cache.\hypertarget{dsize}{
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
-\begin{equation}
- d_{size} = E_{prime}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+Let $d_{\mathrm{\mathrm{size}}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
+\begin{equation}
+ d_{\mathrm{\mathrm{size}}} = E_{\mathrm{prime}}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
+\end{equation}
+The size of the cache, $c_{\mathrm{size}}$, is calculated using
+\begin{equation}
+ c_{\mathrm{size}} = E_{\mathrm{prime}}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
+E_{\mathrm{prime}}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $s \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $s \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- s = C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}})
-\end{equation}
-\begin{equation}
- C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
+ s = C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathbf{i}})
+\end{equation}
+\begin{equation}
+ C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
 \texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
+\texttt{KEC}(C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially fill up $c_{size}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fill up $c_{\mathrm{size}}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+ n = \left\lfloor\frac{c_{\mathrm{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+ \mathbf{d}[i] = E_{\mathrm{datasetitem}}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{\mathrm{size}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
+ E_{\mathrm{datasetitem}}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
 \end{equation}
@@
   E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{size}] \big) & \text{otherwise}
+ E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{\mathrm{size}}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{\mathrm{size}}] \big) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
+\end{equation}
+$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
+\begin{equation}
+ E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+ n_{\mathrm{mix}} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
 \end{equation}
@@
  E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\mathrm{accesses}}} -2 \\
-E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\mathrm{accesses}}} -2 \\
+E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{newdata}$ returns an array with $n_{\mathrm{mix}}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{\mathrm{size}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
 \end{equation}
@@
 \begin{equation}
- E_{compress}(\mathbf{m}, i) = \begin{cases}
+ E_{\mathrm{compress}}(\mathbf{m}, i) = \begin{cases}
 \mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+E_{\mathrm{compress}}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
 \end{cases}

### 2017-10-01 Paper.tex
Added citations for Keccak and cross-references.

@@
 
-%was using tabularx, but found tabu which seems more flexible, and generally better.
+% Tabu seems more flexible, and generally better.
 \usepackage{tabu}
@@
 \makeatother
-%\PassOptionsToPackage{url}\g@addto@macro{\UrlBreaks}{\UrlOrds}
+
 % info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
@@
 
+\usepackage[utf8]{inputenc}
+
 %\usepackage{glossaries}
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+These may be subscripted to denote specialised variants, \eg \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$}, the cost function for the \hyperlink{SSTORE}{\tiny SSTORE} operation.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases (\cite{Keccak}, \cite{KeccakWikipedia})) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_\mathbf{n}$, denotes the \hyperlink{transaction nonce}{nonce of said transaction}.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_\mathbf{n}$, denotes the nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
+When considering the use of existing functions, given a function $f$, the function \hyperlink{f*}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
 
@@
 
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the \hyperlink{f*}{element-wise transformation} of the base function $L_I$, given as:
 \begin{equation}
@@
 
+\hypertarget{f*}{
 With $L_\mathbf{T}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
@@
 f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
+\end{equation}}
 
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{\mathrm{txdatazero}} & \text{if} \quad i = 0 \\ G_{\mathrm{txdatanonzero}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_{\mathrm{txcreate}} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{\mathrm{transaction}}
-\end{align}
+\begin{align}\hypertarget{g0}{
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} \hyperlink{Gtxdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{Gtxdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+}\end{align}
 
@@
 The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
+Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
 
@@
 $G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \\
-$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
-$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
-$G_{\mathrm{transaction}}$ & 21000 & This is paid for every transaction. \\
+$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{txcreate}{\\
+$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.}\hypertarget{Gtxdatazero}{\\
+$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction.}\hypertarget{Gtxdatanonzero}{ \\
+$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction.}\hypertarget{Gtransaction}{ \\
+$G_{\mathrm{Gtransaction}}$ & 21000 & This is paid for every transaction.} \\
 $G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \hypertarget{SSTORE}{\\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
+&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
@@
 G_{\mathrm{sreset}} & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ }\\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
@@
 0 & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ }\\
 \midrule
@@
 \hypertarget{bigvee}{
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
-Thus it is the greatest element of such elements.
-More information is here: \cite{LatticeOrder}.}\\
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{LatticeOrder}).}\\
 \bottomrule

### 2017-10-01 Paper.tex
Fixed error with a space between \usepackage [autostyle]{csquotes}.
Fixed warnings for reference doesn't exist.
Trying to fix an error in build on l. 2111.
Trying to get build to pass.

@@
 %\usepackage{tweaklist}
+\usepackage[utf8]{inputenc}
+\DeclareUnicodeCharacter{C3}{\"{e}}
+\DeclareUnicodeCharacter{AD}{\Huge [ICI]}
 \usepackage{xspace}
@@
 \usepackage[english]{babel}
-\usepackage [autostyle]{csquotes}
+\usepackage[autostyle]{csquotes}
 \MakeOuterQuote{"}
@@
 % Tabu seems more flexible, and generally better.
-\usepackage{tabu}
+\usepackage{tabu} %requires array.
 
@@
 % <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-\usepackage[utf8]{inputenc}
 
@@
 $G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{txcreate}{\\
+$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{\\
 $G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.}\hypertarget{Gtxdatazero}{\\
@@
 \midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{\mathrm{selfdestruct}}{}\\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{\\
+\midrule
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. }\\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_\mathbf{a} \}$ \\
@@
 
-\begin{tabularx}{\textwidth}{lrX}
-%\begin{tabu}{X[,l]X[,r]X[,l]} 
+%\begin{tabularx}{\textwidth}{lrX}
+\begin{tabu}{X[,l]X[,r]X[,l]} 
 \toprule
-Name & Value & Description \hypertarget{Jwordbytes}{ \\
-\midrule 
-$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word. }\hypertarget{Jdatasetinit}{ \\
-$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{ \\
+Name & Value & Description \hypertarget{Jwordbytes}{\\
+\midrule 
+$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.}\hypertarget{Jdatasetinit}{\\
+$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
 $J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
@@
 \bottomrule
-%\end{tabu}
-\end{tabularx}
+\end{tabu}
+%\end{tabularx}
 

### 2017-10-02 Paper.tex
Removed e with Umlaut to see if it will still build successfully.
Minor edits after l. 2086 to 2111 for build to pass.
l. 2095 to 2111 retyping in case of copy paste causing the build error.
l. 2095 to 2111 retyping and deleting unnecesary white space in case copy and paste is causing a build error, plus changing hypertarget captions (shouldn't make a difference, but just in case).

@@
 \usepackage[utf8]{inputenc}
-\DeclareUnicodeCharacter{C3}{\"{e}}
-\DeclareUnicodeCharacter{AD}{\Huge [ICI]}
+%This character renders, so this shouldn't be needed: \DeclareUnicodeCharacter{C3}{\"{e}}
+%For testing: \DeclareUnicodeCharacter{AD}{\Huge [ICI]}
 \usepackage{xspace}
@@
 
-\section{Ethash}\label{app:ethash}
+\section{Ethash} \label{app:ethash}
 \subsection{Definitions}
@@
 \toprule
-Name & Value & Description \hypertarget{Jwordbytes}{\\
-\midrule 
-$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.}\hypertarget{Jdatasetinit}{\\
-$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. }\hypertarget{Jdatasetgrowth}{\\
-$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. }\hypertarget{Jcacheinit}{\\
-$J_{\mathrm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.}\hypertarget{Jcachegrowth}{\\
-$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.}\hypertarget{Jepoch}{\\
-$J_{\mathrm{epoch}}$ & 30000 & This is the blocks per epoch.}\hypertarget{Jmixbytes}{\\
-$J_{\mathrm{mixbytes}}$ & 128 & This is the mix length in bytes.}\hypertarget{Jhashbytes}{\\
-$J_{\mathrm{hashbytes}}$ & 64 & This is the hash length in bytes.}\hypertarget{Jparents}{\\
-$J_{\mathrm{parents}}$ & 256 & This is the number of parents of each dataset element.}\hypertarget{Jcacherounds}{\\
-$J_{\mathrm{cacherounds}}$ & 3 & This is the number of rounds in cache production.}\hypertarget{Jaccesses}{\\
-$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.}\\
+Name & Value & Description \hypertarget{Jwordbytes}{}\\
+\midrule
+$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.}\hypertarget{Jdatasetinit}{}\\
+$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{}\\
+$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. \hypertarget{Jcacheinit}{}\\
+$J_{\mathrm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.\hypertarget{Jcachegrowth}{}\\
+$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.\hypertarget{Jepoch}{}\\
+$J_{\mathrm{epoch}}$ & 30000 & This is the blocks per epoch.\hypertarget{Jmixbytes}{}\\
+$J_{\mathrm{mixbytes}}$ & 128 & This is the mix length in bytes.\hypertarget{Jhashbytes}{}\\
+$J_{\mathrm{hashbytes}}$ & 64 & This is the hash length in bytes.\hypertarget{Jparents}{}\\
+$J_{\mathrm{parents}}$ & 256 & This is the number of parents of each dataset element.\hypertarget{Jcacherounds}{}\\
+$J_{\mathrm{cacherounds}}$ & 3 & This is the number of rounds in cache production.\hypertarget{Jaccesses}{}\\
+$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.\\
 \bottomrule

### 2017-10-02 Paper.tex
Deleting extra }

@@
 \midrule
-$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.}\hypertarget{Jdatasetinit}{}\\
+$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{}\\
 $J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{}\\

### 2017-10-02 Paper.tex
Copied and pasted a formula with tiny in math mode to l. 2109 to see if it will pick up in the Travis CI build.

@@
 $J_{\mathrm{cacherounds}}$ & 3 & This is the number of rounds in cache production.\hypertarget{Jaccesses}{}\\
-$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.\\
+$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$\\
 \bottomrule

### 2017-10-02 Paper.tex
Copied and pasted a formula with tiny in math mode to l. 2088 to see if it will pick up in the Travis CI build.

--- 2017-10-02 Paper.tex
@@
 Also \hyperlink{H s}{$time$} will be set to the initial timestamp of the genesis block.
-The latest documentation should be consulted for those values.
-
-\section{Ethash} \label{app:ethash}
+The latest documentation should be consulted for those values.$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$
+
+\section{Ethash}\label{app:ethash}
 \subsection{Definitions}
@@
 \toprule
-Name & Value & Description \hypertarget{Jwordbytes}{}\\
+Name & Value & Description \hypertarget{Jwordbytes}{} \\
 \midrule
 $J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{}\\
-$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{}\\
+$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{} \\
 $J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. \hypertarget{Jcacheinit}{}\\

### 2017-10-02 Paper.tex
Copied and pasted a formula with tiny in math mode to l. 2097 to see if it will pick up in the Travis CI build.

@@
 Also \hyperlink{H s}{$time$} will be set to the initial timestamp of the genesis block.
-The latest documentation should be consulted for those values.$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$
+The latest documentation should be consulted for those values.
 
@@
 \subsection{Definitions}
-We employ the following definitions:
+We employ the following definitions:$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$
 

### 2017-10-05 Paper.tex
Copied and pasted a formula with tiny in math mode to l. 2097 to see if it will pick up in the Travis CI build.
Changing tabu to tabularx
Retyping lines 2095–2096, math with tiny at beginning of l. 2097.
Definition of e in signing transactions

@@
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\ e = \mathtt{\small HASH}(m) \\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range)\hypertarget{v}{.
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+$e$ is the hash of the message $m$, using some cryptographic hash function such as SHA-3.\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 
@@
 \subsection{Definitions}
-We employ the following definitions:$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$
-
-%\begin{tabularx}{\textwidth}{lrX}
-\begin{tabu}{X[,l]X[,r]X[,l]}
+We employ the following definitions:
+
+%\begin{tabu}{X[,l]X[,r]X[,l]}
+\begin{tabularx}{\textwidth}{lrX}
 \toprule
@@
 \bottomrule
-\end{tabu}
-%\end{tabularx}
+\end{tabularx}
+%\end{tabu}
 

### 2017-10-05 Paper.tex
Added a definition for e in signing transactions and moved the footnote to the reference note field.
Comment out \usepackage[utf8]{inputenc}
Typo no space

@@
 %\usepackage{tweaklist}
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
+%Commenting out as per this suggestion: https://tex.stackexchange.com/questions/4268/inputenc-error-unicode-char-u8-error-while-trying-to-write-a-degree-symbol#comment936894_4268.
+The Travis CI build shows that TeX live is loaded which should include XeTeX
 %This character renders, so this shouldn't be needed: \DeclareUnicodeCharacter{C3}{\"{e}}
@@
 These are formally defined in the literature, \eg by \cite{ECDSAcerticom} and \cite{ECDSAWikipedia}.
-\footnote{For the former citation, refer to section 6.2 for ECDSAPUBKEY, and section 7 for ECDSASIGN and ECDSARECOVER.} %(assuming that the former is for signature generation and the latter is for signature verification)
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\ e = \mathtt{\small HASH}(m) \\
+\ e & \equiv & \mathtt{\small HASH}(m) \\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-$e$ is the hash of the message $m$, using some cryptographic hash function such as SHA-3.\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
+$e$ is the hash of the message $m$, using some cryptographic hash function such as SHA-3.
+\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 

### 2017-10-05 Paper.tex
e was with a tilde not an umlaut. Updated hypertarget for T w.

--- 2017-10-05 Paper.tex
@@
 %\usepackage{tweaklist}
-\usepackage[utf8]{inputenc}
+%\usepackage[utf8]{inputenc}
 %This character renders, so this shouldn't be needed: \DeclareUnicodeCharacter{C3}{\"{e}}
@@
 \item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathbf{t}$.}\hypertarget{T v}{
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathbf{v}$.}
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hyperlink{T w}{$T_\mathbf{w}$}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathbf{v}$.}\hypertarget{T w}{
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathbf{w}$}}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 
-\hypertarget{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
+\hyperlink{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
 \begin{eqnarray}

### 2017-10-05 Paper.tex
Uncommenting \usepackage[utf8]{inputenc} so that ë shows in the PDF. Fixing definition of e in signing transactions to cross-reference to h(T).

@@
 %\usepackage{tweaklist}
-%\usepackage[utf8]{inputenc}
+\usepackage[utf8]{inputenc}
 %This character renders, so this shouldn't be needed: \DeclareUnicodeCharacter{C3}{\"{e}}
@@
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\ e & \equiv & \mathtt{\small HASH}(m) \\
+\ e & \equiv & \hyperlink{h T}{h(T)} \\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-$e$ is the hash of the message $m$, using some cryptographic hash function such as SHA-3.\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
+\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 
@@
 
+\hypertarget{h T}{
 The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_\mathbf{r}$, $T_\mathbf{s}$ and $\hyperlink{T w}{T_\mathbf{w}}$:
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )}
 \end{eqnarray}

### 2017-10-05 Paper.tex
Added commented lines which cause my local build to fail.

@@
 \usepackage[utf8]{inputenc}
-%This character renders, so this shouldn't be needed: \DeclareUnicodeCharacter{C3}{\"{e}}
+%\DeclareUnicodeCharacter{00A0}{ }
+%\DeclareUnicodeCharacter{C3 ab}{"e}
+%\DeclareUnicodeCharacter{C3 bc}{"u}
 %For testing: \DeclareUnicodeCharacter{AD}{\Huge [ICI]}
@@
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathbf{v}$.}\hypertarget{T w}{
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathbf{w}$}}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathbf{w}$}}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}, respectively.
 This is expanded in Appendix \ref{app:signing}.

### 2017-10-06 Paper.tex
Uncommenting \usepackage[utf8]{inputenc} again
It's perplexing that there is a difference between my local build and the repo Travis CI build, even though I have TeXLive installed.
grammar difficulty

@@
 
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
+This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation is required, lengthening the likely next period.
 Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.

### 2017-10-06 Paper.tex
Updating hypertarget for H_d to direct to "The canonical difficulty of a block of header..."

@@
 %\usepackage{tweaklist}
-\usepackage[utf8]{inputenc}
+
+%You don't use inputenc with XeTeX.
+%\usepackage[utf8]{inputenc}
 %\DeclareUnicodeCharacter{00A0}{ }
@@
 \item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarget{H d}{
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarlink{H d}{
 \item[difficulty] A scalar value corresponding to the difficulty level of this block.
@@
 
+\hypertarget{H d}{
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
@@
 \expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_\mathbf{i}} \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
+\end{equation}}
 

### 2017-10-06 Paper.tex
Biblio: moved dates to the note field, fixed author fields, added last  accessed (which also shows in the Wayback link), added filetea links for pages not able to be saved (or at least not properly) by the Wayback Machine. Removed some Wikipedia references, and I will remove more.
Paper.tex: added a note for the difficulty, with citations.

@@
 \item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hypertarlink{H d}{
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hyperlink{H d}{
 \item[difficulty] A scalar value corresponding to the difficulty level of this block.
 This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
-\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathbf{i}$.}\hypertarget{H l}{
+\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathbf{i}$.}\hyperlink{H l}{
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
@@
 
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{stateR-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 
+Note that \mindifficulty is the difficulty of the genesis block.
+The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+\expdiffsymb results in the difficulty increasing every 10000 blocks, which also stabilises the time difference between blocks, as computation power becomes more efficient, \eg with Moore's Law (\cite{Multichain}, \cite{Moore}).
+
+\hypertarget{H l}{
 The canonical gas limit $\hyperlink{H l}{H_\mathbf{l}}$ of a block of header $H$ must fulfil the relation:
@@
 & & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000
-\end{eqnarray}
+\end{eqnarray}}
 
@@
 These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{ASICbitcoinwiki}).
+In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
 While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, which is known as Mist since 8 July 2016 (\cite{Mist}).
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, which is known as Mist since 8 July 2016 (\citeauthor{Mist}).
 
@@
 $G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{\\
-$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.}\hypertarget{Gtxdatazero}{\\
+$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\textit{Homestead transition}}.}\hypertarget{Gtxdatazero}{\\
 $G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction.}\hypertarget{Gtxdatanonzero}{ \\

### 2017-10-06 Paper.tex
Added a ref. for Moore's Law: THOMPSON200620.

@@
 The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-\expdiffsymb results in the difficulty increasing every 10000 blocks, which also stabilises the time difference between blocks, as computation power becomes more efficient, \eg with Moore's Law (\cite{Multichain}, \cite{Moore}).
+\expdiffsymb results in the difficulty increasing every 10000 blocks, which also stabilises the time difference between blocks, as computation power becomes more efficient, \eg with Moore's Law (\cite{Multichain}, \cite{THOMPSON200620}).
 

### 2017-10-15 Paper.tex

--- 2017-10-06 Paper.tex
@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$}, the cost function for the \hyperlink{SSTORE}{\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases (\cite{Keccak}, \cite{KeccakWikipedia})) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases (\cite{Keccak}) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-\expdiffsymb results in the difficulty increasing every 10000 blocks, which also stabilises the time difference between blocks, as computation power becomes more efficient, \eg with Moore's Law (\cite{Multichain}, \cite{Moore}).
+\expdiffsymb results in the difficulty increasing every 10000 blocks.
+This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law (\cite{THOMPSON200620}).
 
@@
 The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
-This method utilises the SECP-256k1 curve as described by \cite{secp256k1BitcoinWiki2016} and \cite{secp256k1StackExchange2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.
+This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.
 9 of 15, para.
@@
 
-We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom} and \cite{ECDSAWikipedia}.
+We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
 \begin{eqnarray}
@@
 \hypertarget{bigvee}{
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{LatticeOrder}).}\\
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{zbMATH01748069}).}\\
 \bottomrule

### 2017-10-15 Paper.tex
\usepackage{hyperref}

@@
 \documentclass[9pt,oneside]{amsart}
+%\usepackage{tweaklist}
+\usepackage{hyperref}
+\usepackage{url}
 \usepackage{cancel}
-%\usepackage{tweaklist}
-
-%You don't use inputenc with XeTeX.
-%\usepackage[utf8]{inputenc}
-%\DeclareUnicodeCharacter{00A0}{ }
-%\DeclareUnicodeCharacter{C3 ab}{"e}
-%\DeclareUnicodeCharacter{C3 bc}{"u}
-%For testing: \DeclareUnicodeCharacter{AD}{\Huge [ICI]}
 \usepackage{xspace}
@@
 \usepackage{tikz}
-\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-%\usepackage{longtable}
-%https://stackoverflow.com/questions/1918366/automated-line-breaks-in-latex-tables/1918465#1918465
-%tabularx seems better, but we can include this if we need tables to break across pages automatically, as well as tabu and longtabu
-
-\usepackage{tabularx}
-%https://stackoverflow.com/a/1919447/7438857
-
-% Tabu seems more flexible, and generally better.
-\usepackage{tabu} %requires array.
-
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-
-% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
-Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
-If you force a URL to break over a page, the error still occurs.
-
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
-% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
-%\sloppy
-%\setlength{\emergencystretch}{2pt}
-% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
-% https://tex.stackexchange.com/a/59131/143781
-% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-%\usepackage{glossaries}
 
@@
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not the least being Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, providing ``consumers'' with the ability to make micro-payments to ``suppliers'' for their services.
+In this instance, the token was used to keep peer-to-peer file trading in check, ensuring ``consumers'' be able to make micro-payments to ``suppliers'' for their services.
 The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
@@
 
-%E language/glossary?
-
-For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
+%E language?
 
@@
 A valid state transition is one which comes about through a transaction.
-Formally:\hypertarget{Upsilon}{}
-\begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_\mathbf{t}, T)
+Formally:
+\begin{equation}
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal or ledger, recording a series of transactions together with the previous block and an identifier for the final state (though blocks do not store the final state itself---that would be far too big).
+Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
 They also punctuate the transaction series with incentives for nodes to \textit{mine}.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathbf{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (the \hyperlink{GenesisBlock}{genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 
-Functions operating on highly structured values are denoted with an upper-case greek letter, \eg \hyperlink{Upsilon}{$\Upsilon$}, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
 
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$}, the cost function for the \hyperlink{SSTORE}{\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases (\cite{Keccak}) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_\mathbf{n}$, denotes the nonce of said transaction}.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
 Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
-When considering the use of existing functions, given a function $f$, the function \hyperlink{f*}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-
-I define a number of useful functions throughout. \hypertarget{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
+It is formally defined in section \ref{ch:block}.
+
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
-\hypertarget{account nonce}{}
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_\mathbf{n}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_\mathbf{b}$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_\mathbf{s}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_\mathbf{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_\mathbf{c}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_\mathbf{s}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the \hyperlink{f*}{element-wise transformation} of the base function $L_I$, given as:
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
 \begin{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_\mathbf{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_\mathbf{n}, \boldsymbol{\sigma}[a]_\mathbf{b}, \boldsymbol{\sigma}[a]_\mathbf{s}, \boldsymbol{\sigma}[a]_\mathbf{c}) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
 \end{equation}
@@
 \end{equation}
-\hypertarget{v x}{where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_\mathbf{n} \in \mathbb{P}_{256} \wedge x_\mathbf{b} \in \mathbb{P}_{256} \wedge x_\mathbf{s} \in \mathbb{B}_{32} \wedge x_\mathbf{c} \in \mathbb{B}_{32}
-}
-\end{equation}
-
-\subsection{The Transaction} \label{subsec:transaction}
+where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+\subsection{The Transaction} \label{ch:transaction}
 
@@
 \begin{description}
-\hypertarget{T n}{
-\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_\mathbf{n}$.}\hypertarget{T p}{
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_\mathbf{p}$.}\hypertarget{T g}{
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_\mathbf{g}$.}\hypertarget{T t}{
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathbf{t}$.}\hypertarget{T v}{
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathbf{v}$.}\hypertarget{T w}{
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathbf{w}$}}, \hyperlink{T r}{$T_\mathbf{r}$} and \hyperlink{T s}{$T_\mathbf{s}$}, respectively.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
 This is expanded in Appendix \ref{app:signing}.
@@
 
-\begin{equation}
-L_\mathbf{T}(T) \equiv \begin{cases}
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}, \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) & \text{if} \; T_\mathbf{t} = \varnothing\\
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}, \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) & \text{otherwise}
+Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
+For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
+
+\begin{equation}
+L_T(T) \equiv \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_\mathbf{n} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_\mathbf{v}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{p} \in \mathbb{P}_{256} & \wedge \\
-T_\mathbf{g} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_\mathbf{w}} \in \mathbb{P}_5 & \wedge & T_\mathbf{r} \in \mathbb{P}_{256} & \wedge \\
-T_\mathbf{s} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_\mathbf{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_\mathbf{t} \neq \varnothing \\
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other \hypertarget{ommerheaders}{block headers $\mathbf{U}$} that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, as it is not a well-known word) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 
-\begin{description}\hypertarget{H p}{
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathbf{p}$.}
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathbf{o}}$.\hypertarget{H c}{
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathbf{c}$.}
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathbf{r}$}.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathbf{t}$}.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathbf{e}$}.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathbf{b}.$}\hyperlink{H d}{
-\item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathbf{d}$}.\hypertarget{H i}{
-\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathbf{i}$.}\hyperlink{H l}{
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathbf{l}$.}\hypertarget{H g}{
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathbf{g}$.}\hypertarget{H s}{
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathbf{s}$}.}\hypertarget{H x}{
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathbf{x}$.}\hypertarget{H m}{
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{m}$.}
-\hypertarget{block nonce}{\hypertarget{H n}{
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathbf{n}$.}}
+\begin{description}
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
-\hypertarget{B U}{
-The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathbf{U}$, and a series of the transactions, $B_\mathbf{T}$.
-Formally, we can refer to a block $B$:
+
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}}
-
-\hypertarget{Transaction Receipt}{\subsubsection{Transaction Receipt}}
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_\mathbf{e}$}.
-
-\hypertarget{R}{}
-The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathbf{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathbf{b}$}:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_\mathbf{u}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathbf{u}}, \hyperlink{R b}{R_\mathbf{b}}, \hyperlink{R l}{R_\mathbf{l}})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.
-
-We assert $\hyperlink{R u}{R_\mathbf{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_\mathbf{b}$}, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-\hyperlink{R u}{R_\mathbf{u}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_\mathbf{b}} \in \mathbb{B}_{256}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
 \end{equation}
@@
 
-The log entries, $\hyperlink{R l}{R_\mathbf{l}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_\mathbf{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_\mathbf{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_\mathbf{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathbf{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence".} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
+Formally:
 \begin{eqnarray}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathbf{r}$:
-\begin{equation}
-\begin{array}[t]{lclc}\hypertarget{H r}{
-H_\mathbf{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
-H_\mathbf{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathbf{U}}))) & \wedge }\hypertarget{H t}{\\
-H_\mathbf{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_\mathbf{T}(B_\mathbf{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
-H_\mathbf{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
-H_\mathbf{b} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_\mathbf{b} \big)}}
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\begin{equation}
+\begin{array}[t]{lclc}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_\mathbf{r}
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
 \end{equation}
@@
 
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathbf{R}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_\mathbf{p}}, \hyperlink{H o}{H_\mathbf{o}}, \hyperlink{H c}{H_\mathbf{c}}, \hyperlink{H r}{H_\mathbf{r}}, \hyperlink{H t}{H_\mathbf{t}}, \hyperlink{H e}{H_\mathbf{e}}, \hyperlink{H b}{H_\mathbf{b}}, \hyperlink{H d}{H_\mathbf{d}},\\ \hyperlink{H i}{H_\mathbf{i}}, \hyperlink{H l}{H_\mathbf{l}}, \hyperlink{H g}{H_\mathbf{g}}, \hyperlink{H s}{H_\mathbf{s}}, \hyperlink{H x}{H_\mathbf{x}}, \hyperlink{H m}{H_\mathbf{m}}, \hyperlink{H n}{H_\mathbf{n}} \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathbf{T}^*(B_\mathbf{T}), L_H^*(\hyperlink{B U}{B_\mathbf{U}}) \big)
-\end{eqnarray}
-
-\hypertarget{f*}{
-With $L_\mathbf{T}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
 f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}}
+\end{equation}
 
@@
 \begin{array}[t]{lclclcl}
-\hyperlink{H p}{H_\mathbf{p}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_\mathbf{o}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_\mathbf{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{H r}{H_\mathbf{r}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_\mathbf{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_\mathbf{e}} \in \mathbb{B}_{32} & \wedge \\
-H_\mathbf{b} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_\mathbf{d}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_\mathbf{i}} \in \mathbb{P} & \wedge \\
-\hyperlink{H l}{H_\mathbf{l}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_\mathbf{g}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_\mathbf{s}} \in \mathbb{P}_{256} & \wedge \\
-\hyperlink{H x}{H_\mathbf{x}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_\mathbf{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_\mathbf{n}} \in \mathbb{B}_{8}
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_\mathbf{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_\mathbf{p}}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
 \end{equation}
@@
 \begin{equation}
-\hyperlink{H i}{H_\mathbf{i}} \equiv {{P(H)_H}_\mathbf{i}} + 1
+H_i \equiv {{P(H)_H}_i} + 1
 \end{equation}
@@
 
-\hypertarget{H d}{
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad \hyperlink{H i}{H_\mathbf{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_\mathbf{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_\mathbf{d}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_\mathbf{s}} - {P(H)_H}_\mathbf{s}}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_\mathbf{i}} \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}}
-
-Note that \mindifficulty is the difficulty of the genesis block.
-The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-\expdiffsymb results in the difficulty increasing every 10000 blocks.
-This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law (\cite{THOMPSON200620}).
-
-\hypertarget{H l}{
-The canonical gas limit $\hyperlink{H l}{H_\mathbf{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000
-\end{eqnarray}}
-
-\hyperlink{H s}{$H_\mathbf{s}$} is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-\hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s}
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation is required, lengthening the likely next period.
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 125000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
+\end{equation}
+
+This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
 Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
 
-The nonce of a block, $\hyperlink{H n}{H_\mathbf{n}}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathbf{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$.
-
-\hypertarget{h cancel n}{}
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_\mathbf{d}}$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the \hyperlink{block nonce}{block nonce} must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \wedge m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \\
-& & \hyperlink{H d}{H_\mathbf{d}} = D(H) \quad \wedge \\
-& & \hyperlink{H g}{H_\mathbf{g}} \le \hyperlink{H l}{H_\mathbf{l}}  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} < {P(H)_H}_\mathbf{l} + \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} > {P(H)_H}_\mathbf{l} - \left\lfloor\frac{{P(H)_H}_\mathbf{l}}{1024}\right\rfloor  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathbf{l}} \geqslant 125000  \quad \wedge \\
-& & \hyperlink{H s}{H_\mathbf{s}} > {P(H)_H}_\mathbf{s} \quad \wedge \\
-& & \hyperlink{H i}{H_\mathbf{i}} = {P(H)_H}_\mathbf{i} +1 \quad \wedge \\
-& & \lVert \hyperlink{H x}{H_\mathbf{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})$
+Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+
+Thus we are able to define the block header validity function $V(H)$:
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
+& & H_d = D(H) \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l \geqslant 125000  \quad \wedge \\
+& & H_s > {P(H)_H}_s \quad \wedge \\
+& & H_i = {P(H)_H}_i +1 \quad \wedge \\
+& & \lVert H_x \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
 
@@
 
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon}{$\Upsilon$}.
+The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
 It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction nonce}{transaction nonce} is valid (equivalent to the  \hyperlink{account nonce}{sender account's current nonce});
+\item the transaction nonce is valid (equivalent to the sender account's current nonce);
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
@@
 
-Formally, we consider the function \hyperlink{Upsilon}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
+Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
 \begin{equation}
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{Upsilong}{$\Upsilon^g$} to evaluate to the amount of gas used in the execution of a transaction and \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} to evaluate to the transaction's accrued log items, both to be formally defined later.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{r})
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
 \end{equation}
@@
 $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_\mathbf{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}\hypertarget{g0}{
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} \hyperlink{Gtxdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{Gtxdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
-}\end{align}
+\begin{align}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
+\end{align}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 \begin{equation}
-v_0 \equiv T_\mathbf{g} T_\mathbf{p} + \hyperlink{T v}{T_\mathbf{v}}
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction nonce}{\\
-T_\mathbf{n} & = & \boldsymbol{\sigma}[S(T)]_\mathbf{n} \quad \wedge \\}
-g_0 & \leqslant & T_\mathbf{g} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_\mathbf{b} \quad \wedge \\
-T_\mathbf{g} & \leqslant & {B_H}_\mathbf{l} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_\mathbf{R})_\mathbf{u}}
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leqslant & T_g \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_\mathbf{g}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathbf{R})_\mathbf{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_\mathbf{l}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_\mathbf{g}T_\mathbf{p}$.
-The gas available for the proceeding computation, $g$, is defined as $T_\mathbf{g} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\%
-\boldsymbol{\sigma}_0[S(T)]_\mathbf{n} & \equiv & \boldsymbol{\sigma}[S(T)]_\mathbf{n} + 1
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_\mathbf{o}, &\\ \quad\quad g, T_\mathbf{p}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}, 0) & \text{if} \quad T_\mathbf{t} = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_\mathbf{o}, &\\ \quad\quad T_\mathbf{t}, T_\mathbf{t}, g, T_\mathbf{p}, \hyperlink{T v}{T_\mathbf{v}}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_\mathbf{g} - g_0
-\end{equation}
-and $T_\mathbf{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_g - g_0
+\end{equation}
+and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_\mathbf{g} - g'}{2} \Big\rfloor, A_\mathbf{r} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_\mathbf{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_\mathbf{g} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_\mathbf{b} & \equiv & \boldsymbol{\sigma}_P[S(T)]_\mathbf{b} + g^* T_\mathbf{p} \\
-\boldsymbol{\sigma}^*[m]_\mathbf{b} & \equiv & \boldsymbol{\sigma}_P[m]_\mathbf{b} + (T_\mathbf{g} - g^*) T_\mathbf{p} \\
-m & \equiv & {B_H}_\mathbf{c}
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
 \end{eqnarray}
@@
 
-\hypertarget{Upsilong}{And finally, we specify $\Upsilon^g$, the total gas used in this transaction}\hypertarget{Upsilonl}{ and $\Upsilon^\mathbf{l}$, the logs created by this transaction:}
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_\mathbf{g} - \mathbf{g}' \\
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
@@
 
-These are used to help define the \hyperlink{Transaction Receipt}{transaction receipt}.
+These are used to help define the transaction receipt, discussed later.
 
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_\mathbf{n} - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
-Note we use one fewer than the \hyperlink{account nonce}{sender's nonce value}; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_\mathbf{b} &\equiv& \boldsymbol{\sigma}[s]_\mathbf{b} - v
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_\mathbf{b} & \text{otherwise}
+\boldsymbol{\sigma}[a]_b & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_\mathbf{a} & \equiv & a \\
-I_\mathbf{o} & \equiv & o \\
-I_\mathbf{p} & \equiv & p \\
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_\mathbf{s} & \equiv & s \\
-\hyperlink{I v}{I_\mathbf{v}} & \equiv & v \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_\mathbf{e} & \equiv & e
+I_e & \equiv & e
 \end{eqnarray}
@@
 \begin{equation}
-c \equiv G_{\mathrm{codedeposit}} \times |\mathbf{o}|
+c \equiv G_{codedeposit} \times |\mathbf{o}|
 \end{equation}
@@
 0 & \text{if} \quad F \\
-g^{**} - \mathbf{c} & \text{otherwise} \\
+g^{**} - c & \text{otherwise} \\
 \end{cases} \\
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_\mathbf{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  \mathbf{g}^{**} < \mathbf{c} \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a \hyperlink{selfdestruct}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_\mathbf{b} \equiv \boldsymbol{\sigma}[s]_\mathbf{b} - v
+\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
 \end{equation}
@@
 
-Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and \hyperlink{account nonce}{nonce}.
+Throughout the present work, it is assumed that if $\boldsymbol{\sigma}_1[r]$ was originally undefined, it will be created as an account with no code or state and zero balance and nonce.
 Thus the previous equation should be taken to mean:
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_\mathbf{b} \equiv \boldsymbol{\sigma}_1'[s]_\mathbf{b} - v
+\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_\mathbf{b} \equiv \boldsymbol{\sigma}[r]_\mathbf{b} + v & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_\mathbf{c}$) is executed according to the execution model (see section \ref{ch:model}).
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_\mathbf{a} & \equiv & r \\
-I_\mathbf{o} & \equiv & o \\
-I_\mathbf{p} & \equiv & p \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_\mathbf{s} & \equiv & s \\
-\hyperlink{I v}{I_\mathbf{v}} & \equiv & \tilde{v} \\
-I_\mathbf{e} & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_\mathbf{c}
+I_s & \equiv & s \\
+I_v & \equiv & \tilde{v} \\
+I_e & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
 \end{eqnarray}
@@
 
-\begin{itemize}\hypertarget{I a}{
-\item $I_\mathbf{a}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
-\item $I_\mathbf{o}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
-\item $I_\mathbf{p}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
-\item $I_\mathbf{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
-\item $I_\mathbf{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
+\begin{itemize}
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.}\hypertarget{I e}{
-\item $I_\mathbf{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
@@
 
-\hypertarget{emptySequence}{}
-The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_\mathbf{g}, A, \mathbf{o}) \\
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_\mathbf{g} & \equiv & \mathbf{g} \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_\mathbf{i} & \equiv & 0 \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & ()
@@
 
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_\mathbf{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_\mathbf{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
@@
 
-\subsubsection{Normal Halting} \label{subsec:Normal Halting}
+\subsubsection{Normal Halting}
 
 The normal halting function $H$ is defined:
-\begin{equation} \label{eq:Normal Halting}
+\begin{equation}
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operation, \hyperlink{RETURN}{\text{\small RETURN}}, has a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{emptySequence}{here}.
+The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 \quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\hypertarget{JUMP}{} \hypertarget{JUMPI}{} \label{eq:JUMP} \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
 J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
@@
 \boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_\mathbf{i} & \equiv & \boldsymbol{\mu}_\mathbf{i} \\
+\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
 A' & \equiv & A \\
@@
 In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{GenesisBlock}{genesis block}, in the path.
+Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
 The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
@@
 \begin{eqnarray}
-B_\mathbf{t} & \equiv & B'_\mathbf{t} + B_\mathbf{d} \\
+B_t & \equiv & B'_t + B_d \\
 B' & \equiv & P(B_H)
@@
 
-As such given a block $B$, $B_\mathbf{t}$ is its total difficulty, $B'$ is its parent block and $B_\mathbf{d}$ is its difficulty.
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
 
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{nonce}.
+\item verify (or, if mining, compute a valid) state and nonce.
 \end{enumerate}
@@
 
-The validation of \hyperlink{ommerheaders}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
+The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
 The maximum of ommer headers is two.
@@
 \begin{equation}
-\lVert \hyperlink{B U}{B_\mathbf{U}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_\mathbf{U}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_\mathbf{g}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H g}{{B_H}_\mathbf{g}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathbf{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_\mathbf{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-\hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
 \begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_\mathbf{c}}]_\mathbf{b} & = & \boldsymbol{\sigma}[{B_H}_\mathbf{c}]_\mathbf{b} + (1 + \frac{\lVert \hyperlink{B U}{B_\mathbf{U}}\rVert}{32})\hyperlink{block reward R b}{R_\mathbf{b}} \\
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
 \forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_\mathbf{c}]_\mathbf{b} & = & \boldsymbol{\sigma}[U_\mathbf{c}]_\mathbf{b} + (1 + \frac{1}{8} (U_\mathbf{i} - \hyperlink{H i}{{B_H}_\mathbf{i}})) \hyperlink{block reward R b}{R_\mathbf{b}}
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.\hypertarget{block reward R b}{}
+two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
 
@@
 \begin{equation}
-\text{Let} \quad R_\mathbf{b} = 5 \times 10^{18}
+\text{Let} \quad R_b = 5 \times 10^{18}
 \end{equation}
@@
 
-\hypertarget{Gamma}{
 We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_\mathbf{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathbf{i})) = {P(B_H)_H}_\mathbf{r} & \text{otherwise}
-\end{cases}
-\end{equation}
-}
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathbf{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathbf{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-\hypertarget{Phi}{
-And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_\mathbf{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \\
-B'_\mathbf{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_\mathbf{r}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$ is the state-transition function}, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_\mathbf{u}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_\mathbf{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction Receipt}{${B_\mathbf{R}}[n]_\mathbf{u}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_\mathbf{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathbf{u}
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
 \begin{equation}
 \mathbf{R}[n]_\mathbf{l} =
-\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-\hypertarget{Pi}{
-Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
-\end{equation}}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
+\end{equation}
+
+Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
 
@@
 One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
+These are specialised pieces of compute hardware that exist only to do a single task.
+In Bitcoin's case the task is the SHA256 hash function.
 While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the \hyperlink{block nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum \textit{Frontier} and \textit{Homestead} we have chosen the first path.
+For Ethereum 1.0 we have chosen the first path.
 
@@
 \begin{equation}
-m = \hyperlink{H m}{H_\mathbf{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathbf{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathbf{n}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathbf{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathbf{d}}$ is the new block's difficulty value (i.e.
+m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being the \hyperlink{block nonce}{block nonce}, a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 The underlying algorithm is called Ethash and is described below.
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum \textit{Frontier} and \textit{Homestead}.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
+Ethash is the PoW algorithm for Ethereum 1.0.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
 The general route that the algorithm takes is as follows:
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{\mathrm{cacheinit}}$} bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{\mathrm{datasetinit}}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every \hyperlink{Jepoch}{$J_{\mathrm{epoch}}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 
-The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
-Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+The state database won't be forced to maintain all past state trie structures into the future.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
 
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{GenesisBlock}{genesis block}.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidating each block and with duplicate or invalid transactions thrown out accordingly.
+Parallelism could also come from a prioritised set of parallel blockchains, consolidated each block and with duplicate or invalid transactions thrown out accordingly.
 
@@
 
-Many thanks to Aeron Buchanan for authoring the \textit{Homestead} revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
+Many thanks to Aeron Buchanan for authoring the Homestead revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
 Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
@@
 
-\section{Terminology} \label{ch:Terminology}
+\section{Terminology}
 
@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, which is known as Mist since 8 July 2016 (\citeauthor{Mist}).
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
 
@@
 
-\hypertarget{rlp}{}
 \section{Recursive Length Prefix}\label{app:rlp}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_\mathbf{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_\mathbf{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-\hypertarget{R b}{
-Formally, we define $R_\mathbf{b}$:}
-\begin{eqnarray}
-R_\mathbf{b}(\mathbf{x}) & \equiv & \begin{cases}
+
+Formally, we define $R_b$:
+\begin{eqnarray}
+R_b(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{itemize}
-\hypertarget{R l}{
-Thus we finish by formally defining $R_\mathbf{l}$:
-\begin{eqnarray}
-R_\mathbf{l}(\mathbf{x}) & \equiv & \begin{cases}
+
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \end{eqnarray}
-}
+
 If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
@@
 
-\hypertarget{trie}{}
 \section{Modified Merkle Patricia Tree}\label{app:trie}
@@
 \begin{equation}
-\mathfrak{I} = \{ (k_0 \in \mathbb{B}, v_0 \in \mathbb{B}), k_1 \in \mathbb{B}, v_1 \in \mathbb{B}), ... \}
+\mathfrak{I} = \{ (\mathbf{k}_0 \in \mathbb{B}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1 \in \mathbb{B}, \mathbf{v}_1 \in \mathbb{B}), ... \}
 \end{equation}
@@
 \begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (k_0' \in \mathbb{Y}, v_0 \in \mathbb{B}), (k_1' \in \mathbb{Y}, v_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\ k_n\rVert} \quad k_n'[i] & \equiv &
+y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
+\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
 \begin{cases}
-\lfloor k_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-k_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_\mathbf{r} \\
-(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_\mathbf{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_\mathbf{r}$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 3000\\
+g_r &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathbf{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}
@@
 
-The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
-This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.
-9 of 15, para.
-3.
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
 
@@
 
-We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\ e & \equiv & \hyperlink{h T}{h(T)} \\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28\}
+v &\in \{27,28\}
 \end{align}
@@
 
-\hypertarget{h T}{
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_\mathbf{r}$, $T_\mathbf{s}$ and $\hyperlink{T w}{T_\mathbf{w}}$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{i}) & \text{if} \; T_\mathbf{t} = 0\\
-(T_\mathbf{n}, T_\mathbf{p}, T_\mathbf{g}, T_\mathbf{t}, \hyperlink{T v}{T_\mathbf{v}}, T_\mathbf{d}) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )}
-\end{eqnarray}
-
-\hyperlink{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
 \begin{eqnarray}
 G(T, p_r) \equiv T \quad \text{except:} \\
-(\hyperlink{T w}{T_\mathbf{w}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}}
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
 
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_\mathbf{w}}, T_\mathbf{r}, T_\mathbf{s}) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
 \end{equation}
@@
 
-\newpage
 \section{Fee Schedule}\label{app:fees}
@@
 
-\begin{tabularx}{\textwidth}{lrX}
+\begin{tabular*}{\columnwidth}[h]{lrl}
 \toprule
@@
 \midrule
-$G_{\mathrm{zero}}$ & 0 & Nothing is paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
-$G_{\mathrm{base}}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
-$G_{\mathrm{verylow}}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
-$G_{\mathrm{low}}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
-$G_{\mathrm{mid}}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
-$G_{\mathrm{\mathrm{high}}}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{\mathrm{high}}}$}. \\
-$G_{\mathrm{extcode}}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
-$G_{\mathrm{balance}}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{\mathrm{sload}}$ & 200 & This is paid for an {\small SLOAD} operation. \\
-$G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
-$G_{\mathrm{sset}}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
-$R_{\mathrm{selfdestruct}}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
-$G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
-$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \\
-$G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
-$G_{\mathrm{call}}$ & 700 & This is paid for a {\small CALL} operation. \\
-$G_{\mathrm{callvalue}}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{\mathrm{callstipend}}$ & 2300 & This is a stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
-$G_{\mathrm{newaccount}}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation which creates an account. \\
-$G_{\mathrm{exp}}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
-$G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{\\
-$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\textit{Homestead transition}}.}\hypertarget{Gtxdatazero}{\\
-$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction.}\hypertarget{Gtxdatanonzero}{ \\
-$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction.}\hypertarget{Gtransaction}{ \\
-$G_{\mathrm{Gtransaction}}$ & 21000 & This is paid for every transaction.} \\
-$G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
-$G_{\mathrm{logdata}}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
-$G_{\mathrm{logtopic}}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
-$G_{\mathrm{sha3}}$ & 30 & This is paid for each {\small SHA3} operation. \\
-$G_{\mathrm{sha3word}}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{\mathrm{copy}}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
-$G_{\mathrm{blockhash}}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\
-
-%extern u256 const \mathbf{c}_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
+$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
-\end{tabularx}
+\end{tabular*}
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathbf{i})-C_{mem}(\boldsymbol{\mu}_\mathbf{i}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
-G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
-G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
-G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
-G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
-G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
-G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
-G_{\mathrm{\mathrm{high}}} & \text{if} \quad w \in W_{\mathrm{\mathrm{high}}}\\
-G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
-G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+G_{create} & \text{if} \quad w = \text{\small CREATE}\\
+G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{zero} & \text{if} \quad w \in W_{zero}\\
+G_{base} & \text{if} \quad w \in W_{base}\\
+G_{verylow} & \text{if} \quad w \in W_{verylow}\\
+G_{low} & \text{if} \quad w \in W_{low}\\
+G_{mid} & \text{if} \quad w \in W_{mid}\\
+G_{high} & \text{if} \quad w \in W_{high}\\
+G_{extcode} & \text{if} \quad w \in W_{extcode}\\
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
 \end{cases}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{\mathrm{memory}} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with \hyperlink{tiny CALL}{$C_\text{\tiny CALL}$}, \hyperlink{C tiny SELFDESTRUCT}{$C_\text{\tiny SELFDESTRUCT}$} and \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$} as specified in the appropriate section below.
-We define the following subsets of instructions:
-
-$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{\mathrm{\mathrm{high}}}$ = \{{\small JUMPI}\}
-
-$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write.
+C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{high}$ = \{{\small JUMPI}\}
+
+$W_{extcode}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_\mathbf{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from the stack, as defined in section \ref{ch:model}.
-
-%\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\begin{tabularx}{\textwidth}{rlrrX}
+For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & This operation halts execution, outputting the empty sequence as per equation \ref{eq:Normal Halting}.
-\\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & This is the addition operation. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+\midrule
+0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
-0x02 & {\small MUL} & 2 & 1 & This is the multiplication operation. \\
+0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
-0x03 & {\small SUB} & 2 & 1 & This is the subtraction operation. \\
+0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
-0x04 & {\small DIV} & 2 & 1 & This is the integer division operation. \\
+0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
-0x05 & {\small SDIV} & 2 & 1 & This is the signed integer division operation (truncated). \\
+0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x06 & {\small MOD} & 2 & 1 & This is the modulo remainder operation. \\
+0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
-0x07 & {\small SMOD} & 2 & 1 & This is the signed modulo remainder operation. \\
+0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x08 & {\small ADDMOD} & 3 & 1 & This is the modulo addition operation. \\
+0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x09 & {\small MULMOD} & 3 & 1 & This is the modulo multiplication operation. \\
+0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
@@
 \midrule
-0x0a & {\small EXP} & 2 & 1 & This is the exponential operation. \\
+0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
 \midrule
-0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_\mathbf{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & This is the less-than comparison. \\
+0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x11 & {\small GT} & 2 & 1 & This is the greater-than comparison. \\
+0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x12 & {\small SLT} & 2 & 1 & This is the signed less-than comparison. \\
+0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x13 & {\small SGT} & 2 & 1 & This is the signed greater-than comparison. \\
+0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 \midrule
-0x14 & {\small EQ} & 2 & 1 & This is the equality comparison. \\
+0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x15 & {\small ISZERO} & 1 & 1 & This is the logical negation operation, also called the logical complement or the NOT operation. \\
+0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
-0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
-\midrule
-0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \vee \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
-\midrule
-0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_\mathbf{i}$ \\
-\midrule
-0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_\mathbf{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_\mathbf{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
+0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+\midrule
+0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+\midrule
+0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+\midrule
+0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute a Keccak-256 hash. \\
+0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{a}$ \\
-\midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_\mathbf{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
-\midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{o}$ \\
-&&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
-\midrule
-0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{s}$ \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+\midrule
+0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+\midrule
+0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& This is the sender of original transaction; it is never an account with non-empty \\
+&&&& associated code. \\
+\midrule
+0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
 \midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \hyperlink{I v}{I_\mathbf{v}}$ \\
-\midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get the input data of the current environment. \\
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+\midrule
+0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
@@
 \midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get the size of the input data in the current environment. \\
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
@@
 \midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy the input data in the current environment to memory. \\
+0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get the size of code running in the current environment. \\
+0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
 \midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in the current environment to memory. \\
+0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{p}$ \\
-&&&& This is the gas price specified by the originating transaction.\\
-\midrule
-0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{c} \rVert$ \\
+0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& This is gas price specified by the originating transaction.\\
+\midrule
+0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{c}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\hyperlink{H p}{H_\mathbf{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
-0 is left on the stack if the looked for block number is greater than the current block number or more than 256 blocks behind the current block.
-\\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_\mathbf{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_\mathbf{i}} \\ P(\hyperlink{H p}{H_\mathbf{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{c}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{s}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{d}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_\mathbf{l}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove the top, first item from the stack. \\
-\midrule
-0x51 & {\small MLOAD} & 1 & 1 & Load the first word from memory. \\
+0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
+\midrule
+0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x52 & {\small MSTORE} & 2 & 0 & Save the first word (which is the first item in the stack) to memory. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x53 & {\small MSTORE8} & 2 & 0 & Save the first byte to memory. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv \max(\boldsymbol{\mu}_\mathbf{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathbf{i}$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \hypertarget{SSTORE}{\\
-\midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{\\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+\midrule
+0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{\mathrm{sreset}} & \text{otherwise}
-\end{cases}$ }\\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sreset} & \text{otherwise}
+\end{cases}$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
-\end{cases}$ }\\
+\end{cases}$ \\
 \midrule
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
-\midrule
-0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment corresponding to this instruction. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+\midrule
+0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
+&&&&  corresponding to this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
@@
 \midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction for the cost of this instruction. \\
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
+&&&& for the cost of this instruction. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
@@
 0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
-&&&& This operation has no effect on the machine state during execution. \\
+&&&& This operation has no effect on machine state during execution. \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place a 1 byte item on the stack. \\
+0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
@@
 &&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
-&&&& The byte is right-aligned (\ie it takes the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place a 2-byte item on the stack. \\
+&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
+\midrule
+0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
-&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
+&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
 \midrule
@@
 &&&& where $\boldsymbol{c}$ is defined as above. \\
-&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
+&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x80 & {\small DUP1} & 1 & 2 & Duplicate the 1st stack item. \\
+0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
-0x81 & {\small DUP2} & 2 & 3 & Duplicate the 2nd stack item. \\
+0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
@@
 \midrule
-0x8f & {\small DUP16} & 16 & 17 & Duplicate the 16th stack item. \\
+0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x90 & {\small SWAP1} & 2 & 2 & Exchange the 1st and the 2nd stack items. \\
+0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
@@
 \midrule
-0x91 & {\small SWAP2} & 3 & 3 & Exchange the 1st and the 3rd stack items. \\
+0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
@@
 \midrule
-0x9f & {\small SWAP16} & 17 & 17 & Exchange the 1st and the 17th stack items. \\
+0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
@@
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_\mathbf{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xa0 & {\small LOG0} & 2 & 0 & Append the log record with no topics. \\
+0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
 &&&& $\mathbf{t} \equiv ()$ \\
 \midrule
-0xa1 & {\small LOG1} & 3 & 0 & Append the log record with one topic. \\
+0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
 &&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 \midrule
-0xa4 & {\small LOG4} & 6 & 0 & Append the log record with four topics. \\
+0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
 &&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
 \bottomrule
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_\mathbf{g}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_\mathbf{a}, I_\mathbf{o}, L(\boldsymbol{\mu}_\mathbf{g}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_\mathbf{e} + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_\mathbf{a}]_\mathbf{n} = \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{n} + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_\mathbf{e} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_\mathbf{a}, \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{n})$, the address of the  \\
-&&&& newly created account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_\mathbf{a}, I_\mathbf{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge \\ \quad\quad I_\mathbf{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{g} \equiv \boldsymbol{\mu}_\mathbf{g} + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ (not enough funds) or $I_\mathbf{e} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
 &&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_\mathbf{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_\mathbf{s}[0] \} & \text{if} \quad \boldsymbol{\mu}_\mathbf{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_\mathbf{s}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{a}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathbf{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\\ \quad\quad{}I_\mathbf{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_\mathbf{a}$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See equation \ref{eq:Normal Halting}. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{i} \equiv M(\boldsymbol{\mu}_\mathbf{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabularx}
-
-\begin{tabularx}{\textwidth}{rlrrX}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
-Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
-The omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$.
-Otherwise it is exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathbf{s}, I_\mathbf{o}, I_\mathbf{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_\mathbf{p}, 0, \hyperlink{I v}{I_\mathbf{v}}, \mathbf{i}, I_\mathbf{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathbf{v}} \leqslant \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b} \;\wedge\; I_\mathbf{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
-This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{\\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. }\\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_\mathbf{a} \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_\mathbf{b} + \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{b}$ \\
-&&&& $\boldsymbol{\sigma}'[I_\mathbf{a}]_\mathbf{b} \equiv 0$ \\
+&&&& See section \ref{ch:model}. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+&&&& the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& Otherwise exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
+&&&& and ninth parameters to the call $\Theta$.\\
+&&&& This means that the recipient is in fact the same account as at present, simply\\
+&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
+\midrule
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_\mathbf{a} \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
-&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{\mathrm{selfdestruct}}} + \begin{cases}
-\hyperlink{Gnewaccount}{G_{\mathrm{newaccount}}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+\end{cases}$ \\
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \bottomrule
-\end{tabularx}
-
-%Is in the terminology list.
+\end{tabular*}
+
 %\section{Low-level Lisp-like Language}\label{app:lll}
@@
 
-\hypertarget{GenesisBlock}{}
 \section{Genesis Block}\label{app:genesis}
@@
 
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the \hyperlink{H c}{beneficiary address}, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the \hyperlink{H b}{log bloom}, a 2048-bit sequence of all zeros; $2^{17}$ refers to the \hyperlink{H d}{difficulty}; the \hyperlink{H t}{transaction trie root}, \hyperlink{H e}{receipt trie root}, \hyperlink{H g}{gas used}, \hyperlink{H i}{block number} and \hyperlink{H x}{extradata} are all $0$ (listed in the order of decimal zeros in the above sequence), being equivalent to the empty byte array.
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the \hyperlink{block nonce}{nonce}.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in \hyperlink{rlp}{RLP}, both empty lists.
-
-The proof-of-concept series includes a development premine, making the state root hash some value $stateRoot$.
-Also \hyperlink{H s}{$time$} will be set to the initial timestamp of the genesis block.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+
+The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
+Also $time$ will be set to the initial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 
-%\begin{tabu}{X[,l]X[,r]X[,l]}
-\begin{tabularx}{\textwidth}{lrX}
+\begin{tabular*}{\columnwidth}[h]{lrl}
 \toprule
-Name & Value & Description \hypertarget{Jwordbytes}{} \\
-\midrule
-$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{}\\
-$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{} \\
-$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. \hypertarget{Jcacheinit}{}\\
-$J_{\mathrm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.\hypertarget{Jcachegrowth}{}\\
-$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.\hypertarget{Jepoch}{}\\
-$J_{\mathrm{epoch}}$ & 30000 & This is the blocks per epoch.\hypertarget{Jmixbytes}{}\\
-$J_{\mathrm{mixbytes}}$ & 128 & This is the mix length in bytes.\hypertarget{Jhashbytes}{}\\
-$J_{\mathrm{hashbytes}}$ & 64 & This is the hash length in bytes.\hypertarget{Jparents}{}\\
-$J_{\mathrm{parents}}$ & 256 & This is the number of parents of each dataset element.\hypertarget{Jcacherounds}{}\\
-$J_{\mathrm{cacherounds}}$ & 3 & This is the number of rounds in cache production.\hypertarget{Jaccesses}{}\\
-$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.\\
+Name & Value & Description \\
+\midrule
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
-\end{tabularx}
-%\end{tabu}
+\end{tabular*}
 
@@
 \begin{equation}
-E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathbf{i}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
-\end{equation}
-The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\mathrm{datasetgrowth}}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} for the cache.\hypertarget{dsize}{
-Let $d_{\mathrm{\mathrm{size}}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
-\begin{equation}
- d_{\mathrm{\mathrm{size}}} = E_{\mathrm{prime}}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
-\end{equation}
-The size of the cache, $c_{\mathrm{size}}$, is calculated using
-\begin{equation}
- c_{\mathrm{size}} = E_{\mathrm{prime}}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
-\end{equation}
-\begin{equation}
- E_{\mathrm{prime}}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{\mathrm{prime}}((x - 1) \cdot y, y) & \text{otherwise}
+E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $s \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- s = C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathbf{i}})
-\end{equation}
-\begin{equation}
- C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathbf{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathbf{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathbf{i}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially fill up $c_{\mathrm{size}}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{\mathrm{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
-\end{equation}
-\begin{equation}
- E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \end{equation}
-\hypertarget{Ecacherounds}{}
 \begin{multline}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{\mathrm{datasetitem}}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{\mathrm{size}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{\mathrm{datasetitem}}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
-E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{\mathrm{size}}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{\mathrm{size}}] \big) & \text{otherwise}
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch $J_{\mathrm{mixbytes}}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{\mathrm{mixbytes}}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the \hyperlink{block nonce}{nonce} is valid.
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the nonce is valid.
 Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
-\end{equation}
-$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
-\begin{equation}
- E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_{\mathrm{mix}} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
-\begin{equation}
- E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\mathrm{accesses}}} -2 \\
-E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{\mathrm{mix}}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{\mathrm{size}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{newdata}$ returns an array with $n_{mix}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}
@@
 \begin{equation}
- E_{\mathrm{compress}}(\mathbf{m}, i) = \begin{cases}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
 \mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{\mathrm{compress}}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabularx}{\textwidth}{lcX}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\hypertarget{bigvee}{
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{zbMATH01748069}).}\\
-\bottomrule
-\end{tabularx}
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+\end{cases}
+\end{equation}
 

### 2017-10-16 Paper.tex
Revert "Merge pull request #324 from lightning-li/modify-136"
This reverts commit 759dccde49c3a1a457196665d1db06baba220bf0, reversing
changes made to 030c1b56a6abc1f6bd3d4899b0f2a87edd6f424d.
Revert "Merge branch 'master' of https://github.com/jamesray1/yellowpaper"
This reverts commit d1f3adee4515e2d56962212548e9fcdb3776acfa, reversing
changes made to 999610d532ada9cbefd2a855543963625ceecf03.
hyperlinks and hypertargets for nonces and emptySequence
Originally from 9a4d05052e6c281525743e6c34f6c75f5b8004a5

--- 2017-10-15 Paper.tex
@@
 %\usepackage{tweaklist}
-\usepackage{hyperref}
 \usepackage{url}
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the \hyperlink{transaction nonce}{nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
 Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 \begin{description}
+\hypertarget{transaction nonce}{}
 \item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 
+\hypertarget{emptySequence}{}
 The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.

### 2017-10-19 Paper.tex
Note about the difficulty and a THOMPSON200620 ref

--- 2017-10-15 Paper.tex
@@
 \end{equation}
+
+Note that \mindifficulty is the difficulty of the genesis block.
+The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+\expdiffsymb results in the difficulty increasing every 10000 blocks.
+This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law, although the rate of increase in computational efficiency has slowed down and may probably continue to slow down (\cite{THOMPSON200620}.
 

### 2017-10-19 Paper.tex
Appendix list of symbols: only bigvee, plus zbMATH01748069 ref

@@
 \end{equation}
-
-Note that \mindifficulty is the difficulty of the genesis block.
-The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-\expdiffsymb results in the difficulty increasing every 10000 blocks.
-This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law, although the rate of increase in computational efficiency has slowed down and may probably continue to slow down (\cite{THOMPSON200620}.
 
@@
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabularx}{\textwidth}{lcX}
+\toprule
+Symbol & Latex Command & Description \\
+\midrule
+\hypertarget{bigvee}{
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{zbMATH01748069}).}\\
+\bottomrule
+\end{tabularx}
+
 \end{document}

### 2017-10-19 Paper.tex
$s around math in text

--- 2017-10-19 Paper.tex
@@
 
-Note that \mindifficulty is the difficulty of the genesis block.
-The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-\expdiffsymb results in the difficulty increasing every 10000 blocks.
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The effect of the difficulty parameter $\homesteadmod$ on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+$\expdiffsymb$ results in the difficulty increasing every 10000 blocks.
 This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law, although the rate of increase in computational efficiency has slowed down and may probably continue to slow down (\cite{THOMPSON200620}.

### 2017-10-19 Paper.tex
Rewording

@@
 Note that $\mindifficulty$ is the difficulty of the genesis block.
-The effect of the difficulty parameter $\homesteadmod$ on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+The difficulty parameter $\homesteadmod$ is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
 $\expdiffsymb$ results in the difficulty increasing every 10000 blocks.

### 2017-10-19 Paper.tex
\usepackage{tabularx}
Crossrefs for tiny SSTORE

--- 2017-10-19 Paper.tex
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
+These may be subscripted to denote specialised variants, \eg \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$}, the cost function for the \hyperlink{SSTORE}{\tiny SSTORE} operation.
 For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabularx}{\textwidth}{lcX}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\hypertarget{bigvee}{
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{zbMATH01748069}).}\\
-\bottomrule
-\end{tabularx}
-
 \end{document}

### 2017-10-19 Paper.tex
\textit{Homestead}

--- 2017-10-19 Paper.tex
@@
 
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The effect of the difficulty parameter $\homesteadmod$ on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-$\expdiffsymb$ results in the difficulty increasing every 10000 blocks.
-This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law, although the rate of increase in computational efficiency has slowed down and may probably continue to slow down (\cite{THOMPSON200620}.
-
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:

### 2017-10-19 Paper.tex

--- 2017-10-19 Paper.tex
@@
 \usepackage{tikz}
+\usepackage{tabularx}
+\usepackage{hyperref}
 
@@
 \end{equation}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The difficulty parameter $\homesteadmod$ is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-$\expdiffsymb$ results in the difficulty increasing every 10000 blocks.
-This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law, although the rate of increase in computational efficiency has slowed down and may probably continue to slow down (\cite{THOMPSON200620}.
 

### 2017-10-19 Paper.tex

--- 2017-10-19 Paper.tex
@@
 %\usepackage{tweaklist}
+\usepackage[utf8]{inputenc}
+%\DeclareUnicodeCharacter{00A0}{ }
+%\DeclareUnicodeCharacter{C3 ab}{"e}
+%\DeclareUnicodeCharacter{C3 bc}{"u}
+%For testing: \DeclareUnicodeCharacter{AD}{\Huge [ICI]}
 \usepackage{url}
@@
 \usepackage{tikz}
+\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+%,pagebackref=true
+
+%\usepackage{longtable}
+%https://stackoverflow.com/questions/1918366/automated-line-breaks-in-latex-tables/1918465#1918465
+%tabularx seems better, but we can include this if we need tables to break across pages automatically, as well as tabu and longtabu
+
+\usepackage{tabularx}
+%https://stackoverflow.com/a/1919447/7438857
+
+% Tabu seems more flexible, and generally better.
+\usepackage{tabu} %requires array.
+
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
+% "hyperref loads the url package internally.
+Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
+This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
+% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
+Try loading the url package earlier than the package that requires it.
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+%\usepackage{glossaries}
+\usepackage{tabu} %requires array.sty
+\usepackage{hyperref}
 
@@
 
-%E language?
+%E language/glossary?
+
+For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
 
@@
 A valid state transition is one which comes about through a transaction.
-Formally:
-\begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+Formally:\hypertarget{Upsilon}{}
+\begin{equation}
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_\mathrm{t}, T)
 \end{equation}
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\

### 2017-10-19 Paper.tex
Changed existing refs, added refs, grammar

--- 2017-10-19 Paper.tex
@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases [\cite{Keccak}]) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task.
-In Bitcoin's case the task is the SHA256 hash function.
+These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}).
+In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
 While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
+This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.
+9 of 15, para.
+3.
 
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
 \begin{eqnarray}

### 2017-10-19 Paper.tex
Hundreds of _\mathbf s replaced with _\mathrm

--- 2017-10-19 Paper.tex
@@
 \documentclass[9pt,oneside]{amsart}
+\usepackage{cancel}
 %\usepackage{tweaklist}
-\usepackage[utf8]{inputenc}
+
+%You don't use inputenc with XeTeX.
+%\usepackage[utf8]{inputenc}
 %\DeclareUnicodeCharacter{00A0}{ }
@@
 %For testing: \DeclareUnicodeCharacter{AD}{\Huge [ICI]}
-\usepackage{url}
-\usepackage{cancel}
 \usepackage{xspace}
@@
 %\usepackage{glossaries}
-\usepackage{tabu} %requires array.sty
-\usepackage{hyperref}
 
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathrm{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+These may be subscripted to denote specialised variants, \eg \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$}, the cost function for the \hyperlink{SSTORE}{\tiny SSTORE} operation.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases (\cite{Keccak}) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_\mathrm{n}$, denotes the nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
 Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathrm{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathrm{m}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
-
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+When considering the use of existing functions, given a function $f$, the function \hyperlink{f*}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
+
+I define a number of useful functions throughout. \hypertarget{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
+\hypertarget{account nonce}{}
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_\mathrm{n}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_\mathrm{b}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_\mathrm{s}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_\mathrm{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_\mathrm{c}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathrm{s})\big) \equiv \boldsymbol{\sigma}[a]_\mathrm{s}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the \hyperlink{f*}{element-wise transformation} of the base function $L_I$, given as:
 \begin{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathrm{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_\mathrm{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_\mathrm{n}, \boldsymbol{\sigma}[a]_\mathrm{b}, \boldsymbol{\sigma}[a]_\mathrm{s}, \boldsymbol{\sigma}[a]_\mathrm{c}) \big) \big)
 \end{equation}
@@
 \end{equation}
-where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{equation}
-
-\subsection{The Transaction} \label{ch:transaction}
+\hypertarget{v x}{where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_\mathrm{n} \in \mathbb{P}_{256} \wedge x_\mathrm{b} \in \mathbb{P}_{256} \wedge x_\mathrm{s} \in \mathbb{B}_{32} \wedge x_\mathrm{c} \in \mathbb{B}_{32}
+}
+\end{equation}
+
+\subsection{The Transaction} \label{subsec:transaction}
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\hypertarget{T n}{
+\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_\mathrm{n}$.}\hypertarget{T p}{
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_\mathrm{p}$.}\hypertarget{T g}{
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_\mathrm{g}$.}\hypertarget{T t}{
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathrm{t}$.}\hypertarget{T v}{
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathrm{v}$.}\hypertarget{T w}{
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathrm{w}$}}, \hyperlink{T r}{$T_\mathrm{r}$} and \hyperlink{T s}{$T_\mathrm{s}$}, respectively.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathrm{i}$.
 \end{description}
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathrm{d}$.
 \end{description}
 
-Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
-For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
-
-\begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+\begin{equation}
+L_\mathrm{T}(T) \equiv \begin{cases}
+(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}, \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) & \text{if} \; T_\mathrm{t} = \varnothing\\
+(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}, \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) & \text{otherwise}
 \end{cases}
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathrm{i}$ and $T_\mathrm{d}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_\mathrm{n} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_\mathrm{p} \in \mathbb{P}_{256} & \wedge \\
+T_\mathrm{g} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_\mathrm{r} \in \mathbb{P}_{256} & \wedge \\
+T_\mathrm{s} \in \mathbb{P}_{256} & \wedge & T_\mathrm{d} \in \mathbb{B} & \wedge & T_\mathrm{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+\mathbb{P}_\mathrm{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_\mathrm{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_\mathrm{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_\mathrm{t} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other \hypertarget{ommerheaders}{block headers $\mathbf{U}$} that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, as it is not a well-known word) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
 The block header contains several pieces of information:
@@
 
-\begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\begin{description}\hypertarget{H p}{
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathrm{p}$.}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathrm{o}}$.\hypertarget{H c}{
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathrm{c}$.}
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathrm{r}$}.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathrm{t}$}.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathrm{e}$}.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathrm{b}.$}\hyperlink{H d}{
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathrm{d}$}.\hypertarget{H i}{
+\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathrm{i}$.}\hyperlink{H l}{
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathrm{l}$.}\hypertarget{H g}{
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathrm{g}$.}\hypertarget{H s}{
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathrm{s}$}.}\hypertarget{H x}{
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathrm{x}$.}\hypertarget{H m}{
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathrm{m}$.}
+\hypertarget{block nonce}{\hypertarget{H n}{
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathrm{n}$.}}
 \end{description}
-
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}
+\hypertarget{B U}{
+The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathrm{U}$, and a series of the transactions, $B_\mathrm{T}$.
+Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathrm{T}, B_\mathrm{U})
+\end{equation}}
+
+\hypertarget{Transaction Receipt}{\subsubsection{Transaction Receipt}}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+Each receipt, denoted $B_\mathrm{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_\mathrm{e}$}.
+
+\hypertarget{R}{}
+The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathrm{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathrm{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathrm{b}$}:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_\mathrm{u}, \hyperlink{R b}{R_\mathrm{b}}, \hyperlink{R l}{R_\mathrm{l}})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathrm{u}}, \hyperlink{R b}{R_\mathrm{b}}, \hyperlink{R l}{R_\mathrm{l}})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.
+
+We assert $\hyperlink{R u}{R_\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_\mathrm{b}$}, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+\hyperlink{R u}{R_\mathrm{u}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_\mathrm{b}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathrm{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $\hyperlink{R l}{R_\mathrm{l}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_\mathrm{a}$, a series of 32-bytes log topics, $O_\mathrm{t}$ and some number of bytes of data, $O_\mathrm{d}$:
+\begin{equation}
+O \equiv (O_\mathrm{a}, ({O_\mathrm{t}}_0, {O_\mathrm{t}}_1, ...), O_\mathrm{d})
+\end{equation}
+\begin{equation}
+O_\mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathrm{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathrm{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathrm{a}\} \cup O_\mathrm{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
-Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence".} Formally:
 \begin{eqnarray}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
-\begin{equation}
-\begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathrm{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathrm{r}$:
+\begin{equation}
+\begin{array}[t]{lclc}\hypertarget{H r}{
+H_\mathrm{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
+H_\mathrm{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathrm{U}}))) & \wedge }\hypertarget{H t}{\\
+H_\mathrm{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathrm{T} \rVert, i \in \mathbb{P}: p(i, L_\mathrm{T}(B_\mathrm{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
+H_\mathrm{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_\mathrm{R}} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathrm{R}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
+H_\mathrm{b} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_\mathrm{R}} \big( \mathbf{r}_\mathrm{b} \big)}}
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_\mathrm{r}
 \end{equation}
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathrm{R}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_\mathrm{p}}, \hyperlink{H o}{H_\mathrm{o}}, \hyperlink{H c}{H_\mathrm{c}}, \hyperlink{H r}{H_\mathrm{r}}, \hyperlink{H t}{H_\mathrm{t}}, \hyperlink{H e}{H_\mathrm{e}}, \hyperlink{H b}{H_\mathrm{b}}, \hyperlink{H d}{H_\mathrm{d}},\\ \hyperlink{H i}{H_\mathrm{i}}, \hyperlink{H l}{H_\mathrm{l}}, \hyperlink{H g}{H_\mathrm{g}}, \hyperlink{H s}{H_\mathrm{s}}, \hyperlink{H x}{H_\mathrm{x}}, \hyperlink{H m}{H_\mathrm{m}}, \hyperlink{H n}{H_\mathrm{n}} \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathrm{T}^*(B_\mathrm{T}), L_H^*(\hyperlink{B U}{B_\mathrm{U}}) \big)
+\end{eqnarray}
+
+\hypertarget{f*}{
+With $L_\mathrm{T}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
 f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
+\end{equation}}
 
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+\hyperlink{H p}{H_\mathrm{p}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_\mathrm{o}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{H r}{H_\mathrm{r}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
+H_\mathrm{b} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_\mathrm{i}} \in \mathbb{P} & \wedge \\
+\hyperlink{H l}{H_\mathrm{l}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{H x}{H_\mathrm{x}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_\mathrm{n}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_\mathrm{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_\mathrm{p}}
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+\hyperlink{H i}{H_\mathrm{i}} \equiv {{P(H)_H}_\mathrm{i}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad \hyperlink{H i}{H_\mathrm{i}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_\mathrm{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
-
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 125000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_\mathrm{d}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_\mathrm{s}} - {P(H)_H}_\mathrm{s}}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}}
+
+Note that \mindifficulty is the difficulty of the genesis block.
+The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+\expdiffsymb results in the difficulty increasing every 10000 blocks.
+This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law (\cite{THOMPSON200620}).
+
+\hypertarget{H l}{
+The canonical gas limit $\hyperlink{H l}{H_\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & \hyperlink{H l}{H_\mathrm{l}} < {P(H)_H}_\mathrm{l} + \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_\mathrm{l}} > {P(H)_H}_\mathrm{l} - \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_\mathrm{l}} \geqslant 125000
+\end{eqnarray}}
+
+\hyperlink{H s}{$H_\mathrm{s}$} is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+\hyperlink{H s}{H_\mathrm{s}} > {P(H)_H}_\mathrm{s}
+\end{equation}
+
+This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation is required, lengthening the likely next period.
 Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+The nonce of a block, $\hyperlink{H n}{H_\mathrm{n}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})$.
+
+\hypertarget{h cancel n}{}
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_\mathrm{d}}$.
 
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 125000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+Because the \hyperlink{block nonce}{block nonce} must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+
+Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \wedge m = \hyperlink{H m}{H_\mathrm{m}} \quad \wedge \\
+& & \hyperlink{H d}{H_\mathrm{d}} = D(H) \quad \wedge \\
+& & \hyperlink{H g}{H_\mathrm{g}} \le \hyperlink{H l}{H_\mathrm{l}}  \quad \wedge \\
+& & \hyperlink{H l}{H_\mathrm{l}} < {P(H)_H}_\mathrm{l} + \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_\mathrm{l}} > {P(H)_H}_\mathrm{l} - \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_\mathrm{l}} \geqslant 125000  \quad \wedge \\
+& & \hyperlink{H s}{H_\mathrm{s}} > {P(H)_H}_\mathrm{s} \quad \wedge \\
+& & \hyperlink{H i}{H_\mathrm{i}} = {P(H)_H}_\mathrm{i} +1 \quad \wedge \\
+& & \lVert \hyperlink{H x}{H_\mathrm{x}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathrm{s}, A_\mathrm{l}, A_\mathrm{r})
+\end{equation}
+
+The tuple contents include $A_\mathrm{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathrm{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_\mathrm{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
-
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+\begin{align}\hypertarget{g0}{
+g_0 \equiv {} & \sum_{i \in T_\mathrm{i}, T_\mathrm{d}} \begin{cases} \hyperlink{Gtxdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{Gtxdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_\mathrm{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+}\end{align}
+
+where $T_\mathrm{i},T_\mathrm{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv T_\mathrm{g} T_\mathrm{p} + \hyperlink{T v}{T_\mathrm{v}}
 \end{equation}
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction nonce}{\\
+T_\mathrm{n} & = & \boldsymbol{\sigma}[S(T)]_\mathrm{n} \quad \wedge \\}
+g_0 & \leqslant & T_\mathrm{g} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_\mathrm{b} \quad \wedge \\
+T_\mathrm{g} & \leqslant & {B_H}_\mathrm{l} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_\mathrm{R})_\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_\mathrm{g}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathrm{R})_\mathrm{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_\mathrm{l}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_\mathrm{g}T_\mathrm{p}$.
+The gas available for the proceeding computation, $g$, is defined as $T_\mathrm{g} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\%
+\boldsymbol{\sigma}_0[S(T)]_\mathrm{n} & \equiv & \boldsymbol{\sigma}[S(T)]_\mathrm{n} + 1
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_\mathrm{o}, &\\ \quad\quad g, T_\mathrm{p}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}, 0) & \text{if} \quad T_\mathrm{t} = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_\mathrm{o}, &\\ \quad\quad T_\mathrm{t}, T_\mathrm{t}, g, T_\mathrm{p}, \hyperlink{T v}{T_\mathrm{v}}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_\mathrm{g} - g_0
+\end{equation}
+and $T_\mathrm{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_\mathrm{g} - g'}{2} \Big\rfloor, A_\mathrm{r} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_\mathrm{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_\mathrm{g} - g'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+\boldsymbol{\sigma}^*[S(T)]_\mathrm{b} & \equiv & \boldsymbol{\sigma}_P[S(T)]_\mathrm{b} + g^* T_\mathrm{p} \\
+\boldsymbol{\sigma}^*[m]_\mathrm{b} & \equiv & \boldsymbol{\sigma}_P[m]_\mathrm{b} + (T_\mathrm{g} - g^*) T_\mathrm{p} \\
+m & \equiv & {B_H}_\mathrm{c}
 \end{eqnarray}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
-\end{eqnarray}
-
-These are used to help define the transaction receipt, discussed later.
+\forall i \in A_\mathrm{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+\hypertarget{Upsilong}{And finally, we specify $\Upsilon^g$, the total gas used in this transaction}\hypertarget{Upsilonl}{ and $\Upsilon^\mathbf{l}$, the logs created by this transaction:}
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_\mathrm{g} - \mathbf{g}' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathrm{l}
+\end{eqnarray}
+
+These are used to help define the \hyperlink{Transaction Receipt}{transaction receipt}.
 
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_\mathrm{n} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_\mathrm{b} &\equiv& \boldsymbol{\sigma}[s]_\mathrm{b} - v
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_b & \text{otherwise}
+\boldsymbol{\sigma}[a]_\mathrm{b} & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+I_\mathrm{a} & \equiv & a \\
+I_\mathrm{o} & \equiv & o \\
+I_\mathrm{p} & \equiv & p \\
+I_\mathrm{d} & \equiv & () \\
+I_\mathrm{s} & \equiv & s \\
+\hyperlink{I v}{I_\mathrm{v}} & \equiv & v \\
+I_\mathrm{b} & \equiv & \mathbf{i} \\
+I_\mathrm{e} & \equiv & e
+\end{eqnarray}
+
+$I_\mathrm{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_\mathrm{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  \mathbf{g}^{**} < \mathbf{c} \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+If the initialisation execution ends with a \hyperlink{selfdestruct}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
 For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}_1[r]_\mathrm{b} \equiv \boldsymbol{\sigma}[r]_\mathrm{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_\mathrm{b} \equiv \boldsymbol{\sigma}[s]_\mathrm{b} - v
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1[s]_\mathrm{b} \equiv \boldsymbol{\sigma}_1'[s]_\mathrm{b} - v
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_\mathrm{b} \equiv \boldsymbol{\sigma}[r]_\mathrm{b} + v & \text{otherwise}
 \end{cases}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_\mathrm{c}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+I_\mathrm{a} & \equiv & r \\
+I_\mathrm{o} & \equiv & o \\
+I_\mathrm{p} & \equiv & p \\
+I_\mathrm{d} & \equiv & \mathbf{d} \\
+I_\mathrm{s} & \equiv & s \\
+\hyperlink{I v}{I_\mathrm{v}} & \equiv & \tilde{v} \\
+I_\mathrm{e} & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathrm{b}) & = & \boldsymbol{\sigma}[c]_\mathrm{c}
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathrm{b}), I_\mathrm{b})$ at some point prior in order to make the determination of $I_\mathrm{b}$ feasible.
 
@@
 
-\begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\begin{itemize}\hypertarget{I a}{
+\item $I_\mathrm{a}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
+\item $I_\mathrm{o}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
+\item $I_\mathrm{p}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
+\item $I_\mathrm{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
+\item $I_\mathrm{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
+\item $I_\mathrm{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
+\item $I_\mathrm{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.}\hypertarget{I e}{
+\item $I_\mathrm{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
 \end{itemize}
@@
 
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\hypertarget{emptySequence}{}
+The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_\mathrm{g}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
+\boldsymbol{\mu}_\mathrm{g} & \equiv & \mathbf{g} \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\boldsymbol{\mu}_\mathrm{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_\mathrm{i} & \equiv & 0 \\
+\boldsymbol{\mu}_\mathrm{s} & \equiv & ()
 \end{eqnarray}
@@
 
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_\mathrm{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
+The memory contents $\boldsymbol{\mu}_\mathrm{m}$ are a series of zeroes of size $2^{256}$.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
+w \equiv \begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathrm{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_\mathrm{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+\lVert\boldsymbol{\mu}_\mathrm{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathrm{s}[0] \notin D(I_\mathrm{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathrm{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
@@
 
-\subsubsection{Normal Halting}
+\subsubsection{Normal Halting} \label{subsec:Normal Halting}
 
 The normal halting function $H$ is defined:
-\begin{equation}
+\begin{equation} \label{eq:Normal Halting}
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operation, \hyperlink{RETURN}{\text{\small RETURN}}, has a special function $H_{\text{\tiny RETURN}}$.
+Note also the difference between the empty sequence and the empty set as discussed \hyperlink{emptySequence}{here}.
 
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\lVert\boldsymbol{\mu}'_\mathrm{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathrm{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathrm{s}\rVert): \boldsymbol{\mu}'_\mathrm{s}[x] & \equiv & \boldsymbol{\mu}_\mathrm{s}[x-\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\mu}'_\mathrm{m} & \equiv & \boldsymbol{\mu}_\mathrm{m} \\
+\boldsymbol{\mu}'_\mathrm{i} & \equiv & \boldsymbol{\mu}_\mathrm{i} \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_t & \equiv & B'_t + B_d \\
+B_\mathrm{t} & \equiv & B'_\mathrm{t} + B_\mathrm{d} \\
 B' & \equiv & P(B_H)
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
+As such given a block $B$, $B_\mathrm{t}$ is its total difficulty, $B'$ is its parent block and $B_\mathrm{d}$ is its difficulty.
 
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert \hyperlink{B U}{B_\mathrm{U}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_\mathrm{U}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathrm{U})
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathrm{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_\mathrm{g}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H g}{{B_H}_\mathrm{g}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathrm{u}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_\mathrm{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+\hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
 \begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_\mathrm{c}}]_\mathrm{b} & = & \boldsymbol{\sigma}[{B_H}_\mathrm{c}]_\mathrm{b} + (1 + \frac{\lVert \hyperlink{B U}{B_\mathrm{U}}\rVert}{32})\hyperlink{block reward R b}{R_\mathrm{b}} \\
+\forall_{U \in B_\mathrm{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_\mathrm{c}]_\mathrm{b} & = & \boldsymbol{\sigma}[U_\mathrm{c}]_\mathrm{b} + (1 + \frac{1}{8} (U_\mathrm{i} - \hyperlink{H i}{{B_H}_\mathrm{i}})) \hyperlink{block reward R b}{R_\mathrm{b}}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_b = 5 \times 10^{18}
+\text{Let} \quad R_\mathrm{b} = 5 \times 10^{18}
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\boldsymbol{\sigma}_\mathrm{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathrm{i})) = {P(B_H)_H}_\mathrm{r} & \text{otherwise}
 \end{cases}
 \end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+}
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathrm{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathrm{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\hypertarget{Phi}{
+And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+B'_\mathrm{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \\
+B'_\mathrm{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_\mathrm{r}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$ is the state-transition function}, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathrm{l}$ and $\mathbf{R}[n]_\mathrm{u}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_\mathrm{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathrm{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of \hyperlink{Transaction Receipt}{${B_\mathrm{R}}[n]_\mathrm{u}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_\mathrm{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathrm{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathrm{u}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
-\end{equation}
-
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
+For $\mathbf{R}[n]_\mathrm{l}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathrm{l} =
+\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathrm{T}[n])
+\end{equation}
+
+\hypertarget{Pi}{
+Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathrm{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
+\end{equation}}
+
+Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
 
@@
 One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task.
-In Bitcoin's case the task is the SHA256 hash function.
+These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}).
+In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
 While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the \hyperlink{block nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
+For Ethereum \textit{Frontier} and \textit{Homestead} we have chosen the first path.
 
@@
 \begin{equation}
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+m = \hyperlink{H m}{H_\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathrm{n}, \mathbf{d})
+\end{equation}
+
+Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathrm{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathrm{d}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being the \hyperlink{block nonce}{block nonce}, a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 The underlying algorithm is called Ethash and is described below.
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+Ethash is the PoW algorithm for Ethereum \textit{Frontier} and \textit{Homestead}.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
 The general route that the algorithm takes is as follows:
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, \hyperlink{Jcacheinit}{$J_{\mathrm{cacheinit}}$} bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, \hyperlink{Jdatasetinit}{$J_{\mathrm{datasetinit}}$} bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every \hyperlink{Jepoch}{$J_{\mathrm{epoch}}$} blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 
-The state database won't be forced to maintain all past state trie structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
+Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
 
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{GenesisBlock}{genesis block}.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, which is known as Mist since 8 July 2016 (\citeauthor{Mist}).
 
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R b}{
+Formally, we define $R_\mathrm{b}$:}
+\begin{eqnarray}
+R_\mathrm{b}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{itemize}
-
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R l}{
+Thus we finish by formally defining $R_\mathrm{l}$:
+\begin{eqnarray}
+R_\mathrm{l}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_\mathrm{r} \\
+(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_\mathrm{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_\mathrm{r}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+\mathbf{g}_\mathrm{r} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
-\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|I_\mathrm{d}|-1)] &=& I_\mathrm{d}\\
+\mathbf{d}[|I_\mathrm{d}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+\mathbf{g}_\mathrm{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathrm{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{g}_\mathrm{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathrm{d})\\
 \end{eqnarray}
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
+\mathbf{g}_\mathrm{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathrm{d}
 \end{eqnarray}
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\ e & \equiv & \hyperlink{h T}{h(T)} \\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\hypertarget{h T}{
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_\mathrm{r}$, $T_\mathrm{s}$ and $\hyperlink{T w}{T_\mathrm{w}}$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}) & \text{if} \; T_\mathrm{t} = 0\\
+(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}) & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )}
+\end{eqnarray}
+
+\hyperlink{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
 \begin{eqnarray}
 G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}
+(\hyperlink{T w}{T_\mathrm{w}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}}
 
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) \big) \big)
 \end{equation}
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{\mathrm{zero}}$ & 0 & Nothing is paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{base}}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
+$G_{\mathrm{low}}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{\mathrm{mid}}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
+$G_{\mathrm{\mathrm{high}}}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{\mathrm{high}}}$}. \\
+$G_{\mathrm{extcode}}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{\mathrm{balance}}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{\mathrm{sload}}$ & 200 & This is paid for an {\small SLOAD} operation. \\
+$G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{sset}}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
+$R_{\mathrm{selfdestruct}}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
+$G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \\
+$G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
+$G_{\mathrm{call}}$ & 700 & This is paid for a {\small CALL} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{\mathrm{callstipend}}$ & 2300 & This is a stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
+$G_{\mathrm{newaccount}}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{\\
+$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\textit{Homestead transition}}.}\hypertarget{Gtxdatazero}{\\
+$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction.}\hypertarget{Gtxdatanonzero}{ \\
+$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction.}\hypertarget{Gtransaction}{ \\
+$G_{\mathrm{Gtransaction}}$ & 21000 & This is paid for every transaction.} \\
+$G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
+$G_{\mathrm{sha3}}$ & 30 & This is paid for each {\small SHA3} operation. \\
+$G_{\mathrm{sha3word}}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{copy}}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\
+
+%extern u256 const \mathbf{c}_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
-\end{tabular*}
+\end{tabularx}
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathrm{i})-C_{mem}(\boldsymbol{\mu}_\mathrm{i}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathrm{s}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathrm{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathrm{s}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathrm{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathrm{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
+G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
+G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
+G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
+G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
+G_{\mathrm{\mathrm{high}}} & \text{if} \quad w \in W_{\mathrm{\mathrm{high}}}\\
+G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small BLOCKHASH}\\
 \end{cases}
@@
 \begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
+w \equiv \begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathrm{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{high}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+C_{mem}(a) \equiv G_{\mathrm{memory}} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with \hyperlink{tiny CALL}{$C_\text{\tiny CALL}$}, \hyperlink{C tiny SELFDESTRUCT}{$C_\text{\tiny SELFDESTRUCT}$} and \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$} as specified in the appropriate section below.
+We define the following subsets of instructions:
+
+$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}\}
+
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{\mathrm{\mathrm{high}}}$ = \{{\small JUMPI}\}
+
+$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathrm{i}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_\mathrm{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 Here given are the various exceptions to the state transition rules given in section \ref{ch:model} specified for each instruction, together with the additional instruction-specific definitions of $J$ and $C$.
-For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from stack, as defined in section \ref{ch:model}.
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+For each instruction, also specified is $\alpha$, the additional items placed on the stack and $\delta$, the items removed from the stack, as defined in section \ref{ch:model}.
+
+%\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-\midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+0x00 & {\small STOP} & 0 & 0 & This operation halts execution, outputting the empty sequence as per equation \ref{eq:Normal Halting}.
+\\
+\midrule
+0x01 & {\small ADD} & 2 & 1 & This is the addition operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+\midrule
+0x02 & {\small MUL} & 2 & 1 & This is the multiplication operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] \times \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+\midrule
+0x03 & {\small SUB} & 2 & 1 & This is the subtraction operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] - \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+\midrule
+0x04 & {\small DIV} & 2 & 1 & This is the integer division operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+\midrule
+0x05 & {\small SDIV} & 2 & 1 & This is the signed integer division operation (truncated). \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathrm{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]) \lfloor |\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 \midrule
-0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
-\midrule
-0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
+0x06 & {\small MOD} & 2 & 1 & This is the modulo remainder operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \boldsymbol{\mu}_\mathrm{s}[0] \bmod \boldsymbol{\mu}_\mathrm{s}[1] & \text{otherwise}\end{cases}$  \\
+\midrule
+0x07 & {\small SMOD} & 2 & 1 & This is the signed modulo remainder operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathrm{s}[0]) (|\boldsymbol{\mu}_\mathrm{s}[0]| \bmod |\boldsymbol{\mu}_\mathrm{s}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+0x08 & {\small ADDMOD} & 3 & 1 & This is the modulo addition operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] = 0\\ (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1]) \mod \boldsymbol{\mu}_\mathrm{s}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
-0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+0x09 & {\small MULMOD} & 3 & 1 & This is the modulo multiplication operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] = 0\\ (\boldsymbol{\mu}_\mathrm{s}[0] \times \boldsymbol{\mu}_\mathrm{s}[1]) \mod \boldsymbol{\mu}_\mathrm{s}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
-0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
-\midrule
-0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+0x0a & {\small EXP} & 2 & 1 & This is the exponential operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] ^ {\boldsymbol{\mu}_\mathrm{s}[1] }$ \\
+\midrule
+0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathrm{s}[0] + 1) \\ \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathrm{s}[x]_\mathrm{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathrm{s}[x]$} \vspace{5pt} \\
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x10 & {\small LT} & 2 & 1 & This is the less-than comparison. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] < \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x11 & {\small GT} & 2 & 1 & This is the greater-than comparison. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x12 & {\small SLT} & 2 & 1 & This is the signed less-than comparison. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] < \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x13 & {\small SGT} & 2 & 1 & This is the signed greater-than comparison. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+0x14 & {\small EQ} & 2 & 1 & This is the equality comparison. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x15 & {\small ISZERO} & 1 & 1 & This is the logical negation operation, also called the logical complement or the NOT operation. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} \wedge \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i}$ \\
+\midrule
+0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} \vee \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i}$ \\
+\midrule
+0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} \oplus \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i}$ \\
+\midrule
+0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[1]_{(i + 8\boldsymbol{\mu}_\mathrm{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathrm{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+0x20 & {\small SHA3} & 2 & 1 & Compute a Keccak-256 hash. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
-\midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
-\midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty \\
-&&&& associated code. \\
-\midrule
-0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{a}$ \\
+\midrule
+0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0]]_\mathrm{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+\midrule
+0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{o}$ \\
+&&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
+\midrule
+0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{s}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
 \midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
-\midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
+0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \hyperlink{I v}{I_\mathrm{v}}$ \\
+\midrule
+0x35 & {\small CALLDATALOAD} & 1 & 1 & Get the input data of the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{d}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ] \quad \text{with} \quad I_\mathrm{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathrm{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get the size of the input data in the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert I_\mathrm{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy the input data in the current environment to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
+\begin{cases} I_\mathrm{d}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert I_\mathrm{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
-\midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
-&&&& This is gas price specified by the originating transaction.\\
-\midrule
-0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+0x38 & {\small CODESIZE} & 0 & 1 & Get the size of code running in the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert I_\mathrm{b} \rVert$ \\
+\midrule
+0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in the current environment to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
+\begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert I_\mathrm{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+\midrule
+0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{p}$ \\
+&&&& This is the gas price specified by the originating transaction.\\
+\midrule
+0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{c} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[3] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathrm{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{c}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[1], \boldsymbol{\mu}_\mathrm{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv P(I_{\hyperlink{H p}{H_\mathrm{p}}}, \boldsymbol{\mu}_\mathrm{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
+0 is left on the stack if the looked for block number is greater than the current block number or more than 256 blocks behind the current block.
+\\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_\mathrm{i}} \\ P(\hyperlink{H p}{H_\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{c}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{s}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{i}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{d}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{l}$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
-\midrule
-0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
-\midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+0x50 & {\small POP} & 1 & 0 & Remove the top, first item from the stack. \\
+\midrule
+0x51 & {\small MLOAD} & 1 & 1 & Load the first word from memory. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv \max(\boldsymbol{\mu}_\mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x52 & {\small MSTORE} & 2 & 0 & Save the first word (which is the first item in the stack) to memory. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv \max(\boldsymbol{\mu}_\mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x53 & {\small MSTORE8} & 2 & 0 & Save the first byte to memory. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] ] \equiv (\boldsymbol{\mu}_\mathrm{s}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv \max(\boldsymbol{\mu}_\mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]]$ \hypertarget{SSTORE}{\\
+\midrule
+0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_\mathrm{a}]_\mathrm{s}[ \boldsymbol{\mu}_\mathrm{s}[0] ] \equiv \boldsymbol{\mu}_\mathrm{s}[1] $ \hypertarget{C tiny SSTORE}{\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
-\end{cases}$ \\
+G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]] = 0 \\
+G_{\mathrm{sreset}} & \text{otherwise}
+\end{cases}$ }\\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]] \neq 0 \\
 0 & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ }\\
 \midrule
 0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathrm{s}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
 \midrule
 0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
-&&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
+\midrule
+0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment corresponding to this instruction. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
-\midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
-&&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+\midrule
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction for the cost of this instruction. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \midrule
 0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
-&&&& This operation has no effect on machine state during execution. \\
+&&&& This operation has no effect on the machine state during execution. \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+0x60 & {\small PUSH1} & 0 & 1 & Place a 1 byte item on the stack. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathrm{b}[x] & \text{if} \quad x < \lVert I_\mathrm{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
 &&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
-&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& The byte is right-aligned (\ie it takes the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
+\midrule
+0x61 & {\small PUSH2} & 0 & 1 & Place a 2-byte item on the stack. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
 \midrule
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+0x80 & {\small DUP1} & 1 & 2 & Duplicate the 1st stack item. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
+\midrule
+0x81 & {\small DUP2} & 2 & 3 & Duplicate the 2nd stack item. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
 \midrule
@@
 \midrule
-0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
+0x8f & {\small DUP16} & 16 & 17 & Duplicate the 16th stack item. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[15]$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+0x90 & {\small SWAP1} & 2 & 2 & Exchange the 1st and the 2nd stack items. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[1] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
+\midrule
+0x91 & {\small SWAP2} & 3 & 3 & Exchange the 1st and the 3rd stack items. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[2]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[2] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
 \midrule
@@
 \midrule
-0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+0x9f & {\small SWAP16} & 17 & 17 & Exchange the 1st and the 17th stack items. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[16]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[16] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathrm{l} \equiv A_\mathrm{l} \cdot (I_\mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
+0xa0 & {\small LOG0} & 2 & 0 & Append the log record with no topics. \\
 &&&& $\mathbf{t} \equiv ()$ \\
 \midrule
-0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
+0xa1 & {\small LOG1} & 3 & 0 & Append the log record with one topic. \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathrm{s}[2])$ \\
 \midrule
@@
 \midrule
-0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
+0xa4 & {\small LOG4} & 6 & 0 & Append the log record with four topics. \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[3], \boldsymbol{\mu}_\mathrm{s}[4], \boldsymbol{\mu}_\mathrm{s}[5])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[1] \dots (\boldsymbol{\mu}_\mathrm{s}[1] + \boldsymbol{\mu}_\mathrm{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_\mathrm{g}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_\mathrm{a}, I_\mathrm{o}, L(\boldsymbol{\mu}_\mathrm{g}), I_\mathrm{p}, \boldsymbol{\mu}_\mathrm{s}[0], \mathbf{i}, I_\mathrm{e} + 1) & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge\; I_\mathrm{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathrm{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_\mathrm{a}]_\mathrm{n} = \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{n} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathrm{s} \equiv A_\mathrm{s} \cup A^+_\mathrm{s} \quad \wedge \quad A'_\mathrm{l} \equiv A_\mathrm{l} \cdot A^+_\mathrm{l} \quad \wedge \quad A'_\mathrm{r} \equiv A_\mathrm{r} + A^+_\mathrm{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_\mathrm{e} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_\mathrm{a}, \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{n})$, the address of the  \\
+&&&& newly created account. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[1], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[3] \dots (\boldsymbol{\mu}_\mathrm{s}[3] + \boldsymbol{\mu}_\mathrm{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_\mathrm{a}, I_\mathrm{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathrm{p}, \boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[2], \mathbf{i}, I_\mathrm{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge \\ \quad\quad I_\mathrm{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathrm{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[5] \dots (\boldsymbol{\mu}_\mathrm{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{g} \equiv \boldsymbol{\mu}_\mathrm{g} + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathrm{s}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
 &&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathrm{s}[2] > \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b}$ (not enough funds) or $I_\mathrm{e} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[3], \boldsymbol{\mu}_\mathrm{s}[4]), \boldsymbol{\mu}_\mathrm{s}[5], \boldsymbol{\mu}_\mathrm{s}[6])$ \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
 &&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_\mathrm{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_\mathrm{s}[0] \} & \text{if} \quad \boldsymbol{\mu}_\mathrm{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_\mathrm{s}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathrm{a}, I_\mathrm{o}, I_\mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathrm{p}, \boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[2], \mathbf{i}, I_\mathrm{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge\\ \quad\quad{}I_\mathrm{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathrm{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_\mathrm{a}$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
 \midrule
 0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots ( \boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& See equation \ref{eq:Normal Halting}. \\
+&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
+Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
+The omitted argument is $\boldsymbol{\mu}_\mathrm{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$, $\boldsymbol{\mu}_\mathrm{s}[5]$ and $\boldsymbol{\mu}_\mathrm{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathrm{s}[2]$, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$ and $\boldsymbol{\mu}_\mathrm{s}[5]$.
+Otherwise it is exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathrm{s}, I_\mathrm{o}, I_\mathrm{a}, t,\\\quad \boldsymbol{\mu}_\mathrm{s}[0], I_\mathrm{p}, 0, \hyperlink{I v}{I_\mathrm{v}}, \mathbf{i}, I_\mathrm{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge\; I_\mathrm{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
+This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{\\
+\midrule
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. }\\
+&&&& $A'_\mathrm{s} \equiv A_\mathrm{s} \cup \{ I_\mathrm{a} \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{b} + \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b}$ \\
+&&&& $\boldsymbol{\sigma}'[I_\mathrm{a}]_\mathrm{b} \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_\mathrm{a} \notin A_\mathrm{s} \\
 0 & \text{otherwise} 
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
+&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{\mathrm{selfdestruct}}} + \begin{cases}
+\hyperlink{Gnewaccount}{G_{\mathrm{newaccount}}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \bottomrule
-\end{tabular*}
-
+\end{tabularx}
+
+%Is in the terminology list.
 %\section{Low-level Lisp-like Language}\label{app:lll}
@@
 
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the \hyperlink{H c}{beneficiary address}, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the \hyperlink{H b}{log bloom}, a 2048-bit sequence of all zeros; $2^{17}$ refers to the \hyperlink{H d}{difficulty}; the \hyperlink{H t}{transaction trie root}, \hyperlink{H e}{receipt trie root}, \hyperlink{H g}{gas used}, \hyperlink{H i}{block number} and \hyperlink{H x}{extradata} are all $0$ (listed in the order of decimal zeros in the above sequence), being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
-
-The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the initial timestamp of the genesis block.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the \hyperlink{block nonce}{nonce}.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in \hyperlink{rlp}{RLP}, both empty lists.
+
+The proof-of-concept series includes a development premine, making the state root hash some value $stateRoot$.
+Also \hyperlink{H s}{$time$} will be set to the initial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 
-\begin{tabular*}{\columnwidth}[h]{lrl}
+%\begin{tabu}{X[,l]X[,r]X[,l]}
+\begin{tabularx}{\textwidth}{lrX}
 \toprule
-Name & Value & Description \\
-\midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+Name & Value & Description \hypertarget{Jwordbytes}{} \\
+\midrule
+$J_{\mathrm{wordbytes}}$ & 4  & This is the bytes in a word.\hypertarget{Jdatasetinit}{}\\
+$J_{\mathrm{datasetinit}}$ & $2^{30}$ & This is the bytes in the dataset at genesis. \hypertarget{Jdatasetgrowth}{} \\
+$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & This is the dataset growth per epoch. \hypertarget{Jcacheinit}{}\\
+$J_{\mathrm{cacheinit}}$ & $2^{24}$ & This is the bytes in the cache at genesis.\hypertarget{Jcachegrowth}{}\\
+$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & This is the cache growth per epoch.\hypertarget{Jepoch}{}\\
+$J_{\mathrm{epoch}}$ & 30000 & This is the blocks per epoch.\hypertarget{Jmixbytes}{}\\
+$J_{\mathrm{mixbytes}}$ & 128 & This is the mix length in bytes.\hypertarget{Jhashbytes}{}\\
+$J_{\mathrm{hashbytes}}$ & 64 & This is the hash length in bytes.\hypertarget{Jparents}{}\\
+$J_{\mathrm{parents}}$ & 256 & This is the number of parents of each dataset element.\hypertarget{Jcacherounds}{}\\
+$J_{\mathrm{cacherounds}}$ & 3 & This is the number of rounds in cache production.\hypertarget{Jaccesses}{}\\
+$J_{\mathrm{accesses}}$ & 64 & This is the number of accesses in a hashimoto loop.\\
 \bottomrule
-\end{tabular*}
+\end{tabularx}
+%\end{tabu}
 
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathrm{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathrm{i}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\mathrm{datasetgrowth}}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} for the cache.\hypertarget{dsize}{
+Let $d_{\mathrm{\mathrm{size}}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
+\begin{equation}
+ d_{\mathrm{\mathrm{size}}} = E_{\mathrm{prime}}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
+\end{equation}
+The size of the cache, $c_{\mathrm{size}}$, is calculated using
+\begin{equation}
+ c_{\mathrm{size}} = E_{\mathrm{prime}}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
+E_{\mathrm{prime}}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $s \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ s = C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathrm{i}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathrm{i}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fill up $c_{\mathrm{size}}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{\mathrm{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{\mathrm{datasetitem}}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{\hyperlink{dsize}{d_{\mathrm{size}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+ E_{\mathrm{datasetitem}}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+ E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod \mathbf{c}_{\mathrm{size}}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}} / \hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}} \rfloor]) \mod \mathbf{c}_{\mathrm{size}}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the nonce is valid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch $J_{\mathrm{mixbytes}}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{\mathrm{mixbytes}}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the \hyperlink{block nonce}{nonce} is valid.
 Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathrm{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathrm{n}}) + \mathbf{m}_\mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathrm{c}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_\mathrm{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_\mathrm{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathrm{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_\mathrm{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
+\end{equation}
+$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
+\begin{equation}
+ E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+ n_{\mathrm{mix}} =  \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
+\begin{equation}
+ E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = \hyperlink{Jaccesses}{J_{\mathrm{accesses}}} -2 \\
+E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{newdata}$ returns an array with $n_{\mathrm{mix}}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{\hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}}{\hyperlink{Jwordbytes}{J_{\mathrm{wordbytes}}}}\right\rfloor]) \mod \left\lfloor\frac{\hyperlink{dsize}{d_{\mathrm{size}}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
 \end{equation}

### 2017-10-19 Paper.tex
fix ommer reference

--- 2017-10-19 Paper.tex
@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases [\cite{Keccak}]) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
+These are specialised pieces of compute hardware that exist only to do a single task.
+In Bitcoin's case the task is the SHA256 hash function.
 While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}

### 2017-10-19 Paper.tex

--- 2017-10-19 Paper.tex
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_\mathrm{t}, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_\mathrm{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{\mathrm{t}}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_\mathrm{n}$, denotes the nonce of said transaction}.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_{\mathrm{n}}$, denotes the nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathrm{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathrm{m}[0..31]$ denotes the first 32 items of the machine's memory.
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathrm{s}}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathrm{m}}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_\mathrm{n}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_\mathrm{b}$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_\mathrm{s}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_\mathrm{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_\mathrm{c}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathrm{s})\big) \equiv \boldsymbol{\sigma}[a]_\mathrm{s}
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_{\mathrm{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
 \end{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathrm{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_{\mathrm{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_\mathrm{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_\mathrm{n}, \boldsymbol{\sigma}[a]_\mathrm{b}, \boldsymbol{\sigma}[a]_\mathrm{s}, \boldsymbol{\sigma}[a]_\mathrm{c}) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
 \end{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_\mathrm{n} \in \mathbb{P}_{256} \wedge x_\mathrm{b} \in \mathbb{P}_{256} \wedge x_\mathrm{s} \in \mathbb{B}_{32} \wedge x_\mathrm{c} \in \mathbb{B}_{32}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
 }
@@
 \hypertarget{T n}{
-\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_\mathrm{n}$.}\hypertarget{T p}{
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_\mathrm{p}$.}\hypertarget{T g}{
+\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.}\hypertarget{T p}{
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.}\hypertarget{T g}{
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_\mathrm{g}$.}\hypertarget{T t}{
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_\mathrm{t}$.}\hypertarget{T v}{
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_\mathrm{v}$.}\hypertarget{T w}{
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_\mathrm{w}$}}, \hyperlink{T r}{$T_\mathrm{r}$} and \hyperlink{T s}{$T_\mathrm{s}$}, respectively.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.}\hypertarget{T t}{
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.}\hypertarget{T v}{
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.}\hypertarget{T w}{
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{w}}$}}, \hyperlink{T r}{$T_{\mathrm{r}}$} and \hyperlink{T s}{$T_{\mathrm{s}}$}, respectively.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathrm{i}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathrm{i}}$.
 \end{description}
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathrm{d}$.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathrm{d}}$.
 \end{description}
@@
 \begin{equation}
-L_\mathrm{T}(T) \equiv \begin{cases}
-(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}, \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) & \text{if} \; T_\mathrm{t} = \varnothing\\
-(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}, \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) & \text{otherwise}
+L_{\mathrm{t}}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}, \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}, \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
 \end{cases}
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathrm{i}$ and $T_\mathrm{d}$.
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathrm{i}}$ and $T_{\mathrm{d}}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_\mathrm{n} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_\mathrm{p} \in \mathbb{P}_{256} & \wedge \\
-T_\mathrm{g} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_\mathrm{r} \in \mathbb{P}_{256} & \wedge \\
-T_\mathrm{s} \in \mathbb{P}_{256} & \wedge & T_\mathrm{d} \in \mathbb{B} & \wedge & T_\mathrm{i} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_{\mathrm{v}}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_{\mathrm{w}}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{d}} \in \mathbb{B} & \wedge & T_{\mathrm{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_\mathrm{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_\mathrm{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathrm{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_\mathrm{t} \neq \varnothing \\
+\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_{\mathrm{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_{\mathrm{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{description}\hypertarget{H p}{
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_\mathrm{p}$.}
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_\mathrm{o}}$.\hypertarget{H c}{
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_\mathrm{c}$.}
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_\mathrm{r}$}.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_\mathrm{t}$}.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_\mathrm{e}$}.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_\mathrm{b}.$}\hyperlink{H d}{
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_{\mathrm{o}}}$.\hypertarget{H c}{
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_{\mathrm{c}}$.}
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_{\mathrm{r}}$}.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_{\mathrm{t}}$}.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_{\mathrm{e}}$}.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_{\mathrm{b}}.$}\hyperlink{H d}{
 \item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_\mathrm{d}$}.\hypertarget{H i}{
-\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_\mathrm{i}$.}\hyperlink{H l}{
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_\mathrm{l}$.}\hypertarget{H g}{
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_\mathrm{g}$.}\hypertarget{H s}{
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_\mathrm{s}$}.}\hypertarget{H x}{
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_\mathrm{x}$.}\hypertarget{H m}{
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_\mathrm{m}$.}
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$}.\hypertarget{H i}{
+\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_{\mathrm{i}}$.}\hyperlink{H l}{
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.}\hypertarget{H g}{
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.}\hypertarget{H s}{
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_{\mathrm{s}}$}.}\hypertarget{H x}{
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.}\hypertarget{H m}{
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.}
 \hypertarget{block nonce}{\hypertarget{H n}{
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_\mathrm{n}$.}}
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.}}
 \end{description}
 \hypertarget{B U}{
-The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_\mathrm{U}$, and a series of the transactions, $B_\mathrm{T}$.
+The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathrm{u}}$, and a series of the transactions, $B_{\mathrm{t}}$.
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (B_H, B_\mathrm{T}, B_\mathrm{U})
+B \equiv (B_H, B_{\mathrm{t}}, B_{\mathrm{u}})
 \end{equation}}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathrm{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_\mathrm{e}$}.
+Each receipt, denoted $B_{\mathrm{r}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_{\mathrm{e}}$}.
 
 \hypertarget{R}{}
-The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_\mathrm{u}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_\mathrm{l}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_\mathrm{b}$}:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_\mathrm{u}, \hyperlink{R b}{R_\mathrm{b}}, \hyperlink{R l}{R_\mathrm{l}})
+The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_{\mathrm{l}}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_{\mathrm{b}}$}:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_{\mathrm{u}}, \hyperlink{R b}{R_{\mathrm{b}}}, \hyperlink{R l}{R_{\mathrm{l}}})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_\mathrm{u}}, \hyperlink{R b}{R_\mathrm{b}}, \hyperlink{R l}{R_\mathrm{l}})
+L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_{\mathrm{u}}}, \hyperlink{R b}{R_{\mathrm{b}}}, \hyperlink{R l}{R_{\mathrm{l}}})
 \end{equation}
@@
 
-We assert $\hyperlink{R u}{R_\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_\mathrm{b}$}, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-\hyperlink{R u}{R_\mathrm{u}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathrm{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $\hyperlink{R l}{R_\mathrm{l}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_\mathrm{a}$, a series of 32-bytes log topics, $O_\mathrm{t}$ and some number of bytes of data, $O_\mathrm{d}$:
-\begin{equation}
-O \equiv (O_\mathrm{a}, ({O_\mathrm{t}}_0, {O_\mathrm{t}}_1, ...), O_\mathrm{d})
-\end{equation}
-\begin{equation}
-O_\mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathrm{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathrm{d} \in \mathbb{B}
+We assert $\hyperlink{R u}{R_{\mathrm{u}}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_{\mathrm{b}}$}, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+\hyperlink{R u}{R_{\mathrm{u}}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_{\mathrm{b}}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_{\mathrm{t}}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $\hyperlink{R l}{R_{\mathrm{l}}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_{\mathrm{t}}$ and some number of bytes of data, $O_{\mathrm{d}}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_{\mathrm{t}}}_0, {O_{\mathrm{t}}}_1, ...), O_{\mathrm{d}})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathrm{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathrm{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_\mathrm{a}\} \cup O_\mathrm{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathrm{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathrm{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_\mathrm{r}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathrm{t}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}\hypertarget{H r}{
-H_\mathrm{r} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
-H_\mathrm{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_\mathrm{U}}))) & \wedge }\hypertarget{H t}{\\
-H_\mathrm{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathrm{T} \rVert, i \in \mathbb{P}: p(i, L_\mathrm{T}(B_\mathrm{T}[i]))\}) }& \wedge \hypertarget{H e}{\\
-H_\mathrm{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_\mathrm{R}} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathrm{R}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
-H_\mathrm{b} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_\mathrm{R}} \big( \mathbf{r}_\mathrm{b} \big)}}
+H_{\mathrm{r}} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
+H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_{\mathrm{u}}}))) & \wedge }\hypertarget{H t}{\\
+H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathrm{t}} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{t}}(B_{\mathrm{t}}[i]))\}) }& \wedge \hypertarget{H e}{\\
+H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_{\mathrm{r}}} \rVert, i \in \mathbb{P}: p(i, L_R(B_{\mathrm{r}}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
+H_{\mathrm{b}} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_{\mathrm{r}}} \big( \mathbf{r}_{\mathrm{b}} \big)}}
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_\mathrm{r}
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
 \end{equation}
@@
 
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_\mathrm{R}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_{\mathrm{r}}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_\mathrm{p}}, \hyperlink{H o}{H_\mathrm{o}}, \hyperlink{H c}{H_\mathrm{c}}, \hyperlink{H r}{H_\mathrm{r}}, \hyperlink{H t}{H_\mathrm{t}}, \hyperlink{H e}{H_\mathrm{e}}, \hyperlink{H b}{H_\mathrm{b}}, \hyperlink{H d}{H_\mathrm{d}},\\ \hyperlink{H i}{H_\mathrm{i}}, \hyperlink{H l}{H_\mathrm{l}}, \hyperlink{H g}{H_\mathrm{g}}, \hyperlink{H s}{H_\mathrm{s}}, \hyperlink{H x}{H_\mathrm{x}}, \hyperlink{H m}{H_\mathrm{m}}, \hyperlink{H n}{H_\mathrm{n}} \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_\mathrm{T}^*(B_\mathrm{T}), L_H^*(\hyperlink{B U}{B_\mathrm{U}}) \big)
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_{\mathrm{p}}}, \hyperlink{H o}{H_{\mathrm{o}}}, \hyperlink{H c}{H_{\mathrm{c}}}, \hyperlink{H r}{H_{\mathrm{r}}}, \hyperlink{H t}{H_{\mathrm{t}}}, \hyperlink{H e}{H_{\mathrm{e}}}, \hyperlink{H b}{H_{\mathrm{b}}}, \hyperlink{H d}{H_{\mathrm{d}}},\\ \hyperlink{H i}{H_{\mathrm{i}}}, \hyperlink{H l}{H_{\mathrm{l}}}, \hyperlink{H g}{H_{\mathrm{g}}}, \hyperlink{H s}{H_{\mathrm{s}}}, \hyperlink{H x}{H_{\mathrm{x}}}, \hyperlink{H m}{H_{\mathrm{m}}}, \hyperlink{H n}{H_{\mathrm{n}}} \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_{\mathrm{t}}^*(B_{\mathrm{t}}), L_H^*(\hyperlink{B U}{B_{\mathrm{u}}}) \big)
 \end{eqnarray}
@@
 \hypertarget{f*}{
-With $L_\mathrm{T}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+With $L_{\mathrm{t}}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-\hyperlink{H p}{H_\mathrm{p}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_\mathrm{o}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{H r}{H_\mathrm{r}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
-H_\mathrm{b} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_\mathrm{i}} \in \mathbb{P} & \wedge \\
-\hyperlink{H l}{H_\mathrm{l}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
-\hyperlink{H x}{H_\mathrm{x}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_\mathrm{n}} \in \mathbb{B}_{8}
+\hyperlink{H p}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_{\mathrm{o}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_{\mathrm{c}}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{H r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_{\mathrm{t}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
+H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_{\mathrm{d}}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
+\hyperlink{H l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_{\mathrm{g}}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{H x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_{\mathrm{m}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_\mathrm{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_\mathrm{p}}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_{\mathrm{p}}}
 \end{equation}
@@
 \begin{equation}
-\hyperlink{H i}{H_\mathrm{i}} \equiv {{P(H)_H}_\mathrm{i}} + 1
+\hyperlink{H i}{H_{\mathrm{i}}} \equiv {{P(H)_H}_{\mathrm{i}}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad \hyperlink{H i}{H_\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_\mathrm{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad \hyperlink{H i}{H_{\mathrm{i}}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_\mathrm{d}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_\mathrm{s}} - {P(H)_H}_\mathrm{s}}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_{\mathrm{s}}} - {P(H)_H}_{\mathrm{s}}}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_{\mathrm{i}}} \div 100000 \right\rfloor - 2 } \right\rfloor
 \end{equation}}
@@
 \hypertarget{H l}{
-The canonical gas limit $\hyperlink{H l}{H_\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & \hyperlink{H l}{H_\mathrm{l}} < {P(H)_H}_\mathrm{l} + \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor \quad \wedge \\
-& & \hyperlink{H l}{H_\mathrm{l}} > {P(H)_H}_\mathrm{l} - \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor \quad \wedge \\
-& & \hyperlink{H l}{H_\mathrm{l}} \geqslant 125000
+The canonical gas limit $\hyperlink{H l}{H_{\mathrm{l}}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & \hyperlink{H l}{H_{\mathrm{l}}} < {P(H)_H}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} > {P(H)_H}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} \geqslant 125000
 \end{eqnarray}}
 
-\hyperlink{H s}{$H_\mathrm{s}$} is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-\hyperlink{H s}{H_\mathrm{s}} > {P(H)_H}_\mathrm{s}
+\hyperlink{H s}{$H_{\mathrm{s}}$} is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+\hyperlink{H s}{H_{\mathrm{s}}} > {P(H)_H}_{\mathrm{s}}
 \end{equation}
@@
 
-The nonce of a block, $\hyperlink{H n}{H_\mathrm{n}}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \quad \wedge \quad m = \hyperlink{H m}{H_\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})$.
+The nonce of a block, $\hyperlink{H n}{H_{\mathrm{n}}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \quad \wedge \quad m = \hyperlink{H m}{H_{\mathrm{m}}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})$.
 
@@
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_\mathrm{d}}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_{\mathrm{d}}}$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \wedge m = \hyperlink{H m}{H_\mathrm{m}} \quad \wedge \\
-& & \hyperlink{H d}{H_\mathrm{d}} = D(H) \quad \wedge \\
-& & \hyperlink{H g}{H_\mathrm{g}} \le \hyperlink{H l}{H_\mathrm{l}}  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathrm{l}} < {P(H)_H}_\mathrm{l} + \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathrm{l}} > {P(H)_H}_\mathrm{l} - \left\lfloor\frac{{P(H)_H}_\mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
-& & \hyperlink{H l}{H_\mathrm{l}} \geqslant 125000  \quad \wedge \\
-& & \hyperlink{H s}{H_\mathrm{s}} > {P(H)_H}_\mathrm{s} \quad \wedge \\
-& & \hyperlink{H i}{H_\mathrm{i}} = {P(H)_H}_\mathrm{i} +1 \quad \wedge \\
-& & \lVert \hyperlink{H x}{H_\mathrm{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})$
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \wedge m = \hyperlink{H m}{H_{\mathrm{m}}} \quad \wedge \\
+& & \hyperlink{H d}{H_{\mathrm{d}}} = D(H) \quad \wedge \\
+& & \hyperlink{H g}{H_{\mathrm{g}}} \le \hyperlink{H l}{H_{\mathrm{l}}}  \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} < {P(H)_H}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} > {P(H)_H}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} \geqslant 125000  \quad \wedge \\
+& & \hyperlink{H s}{H_{\mathrm{s}}} > {P(H)_H}_{\mathrm{s}} \quad \wedge \\
+& & \hyperlink{H i}{H_{\mathrm{i}}} = {P(H)_H}_{\mathrm{i}} +1 \quad \wedge \\
+& & \lVert \hyperlink{H x}{H_{\mathrm{x}}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathrm{s}, A_\mathrm{l}, A_\mathrm{r})
-\end{equation}
-
-The tuple contents include $A_\mathrm{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathrm{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_\mathrm{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathrm{s}}, A_{\mathrm{l}}, A_{\mathrm{r}})
+\end{equation}
+
+The tuple contents include $A_{\mathrm{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_{\mathrm{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{align}\hypertarget{g0}{
-g_0 \equiv {} & \sum_{i \in T_\mathrm{i}, T_\mathrm{d}} \begin{cases} \hyperlink{Gtxdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{Gtxdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_\mathrm{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_{i \in T_{\mathrm{i}}, T_{\mathrm{d}}} \begin{cases} \hyperlink{Gtxdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{Gtxdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
@@
 
-where $T_\mathrm{i},T_\mathrm{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+where $T_{\mathrm{i}},T_{\mathrm{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
 $G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
@@
 \begin{equation}
-v_0 \equiv T_\mathrm{g} T_\mathrm{p} + \hyperlink{T v}{T_\mathrm{v}}
+v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + \hyperlink{T v}{T_{\mathrm{v}}}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction nonce}{\\
-T_\mathrm{n} & = & \boldsymbol{\sigma}[S(T)]_\mathrm{n} \quad \wedge \\}
-g_0 & \leqslant & T_\mathrm{g} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_\mathrm{b} \quad \wedge \\
-T_\mathrm{g} & \leqslant & {B_H}_\mathrm{l} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_\mathrm{R})_\mathrm{u}}
+T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\}
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_H}_{\mathrm{l}} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_{\mathrm{r}})_{\mathrm{u}}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_\mathrm{g}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_\mathrm{R})_\mathrm{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_\mathrm{l}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_\mathrm{g}T_\mathrm{p}$.
-The gas available for the proceeding computation, $g$, is defined as $T_\mathrm{g} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_{\mathrm{r}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\%
-\boldsymbol{\sigma}_0[S(T)]_\mathrm{n} & \equiv & \boldsymbol{\sigma}[S(T)]_\mathrm{n} + 1
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_{\mathrm{g}} T_{\mathrm{p}} \\%
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_\mathrm{o}, &\\ \quad\quad g, T_\mathrm{p}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}, 0) & \text{if} \quad T_\mathrm{t} = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_\mathrm{o}, &\\ \quad\quad T_\mathrm{t}, T_\mathrm{t}, g, T_\mathrm{p}, \hyperlink{T v}{T_\mathrm{v}}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}, 0) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}, 0) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, \hyperlink{T v}{T_{\mathrm{v}}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_\mathrm{g} - g_0
-\end{equation}
-and $T_\mathrm{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_{\mathrm{g}} - g_0
+\end{equation}
+and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_\mathrm{g} - g'}{2} \Big\rfloor, A_\mathrm{r} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_\mathrm{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_\mathrm{g} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_\mathrm{b} & \equiv & \boldsymbol{\sigma}_P[S(T)]_\mathrm{b} + g^* T_\mathrm{p} \\
-\boldsymbol{\sigma}^*[m]_\mathrm{b} & \equiv & \boldsymbol{\sigma}_P[m]_\mathrm{b} + (T_\mathrm{g} - g^*) T_\mathrm{p} \\
-m & \equiv & {B_H}_\mathrm{c}
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_P[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_P[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_H}_{\mathrm{c}}
 \end{eqnarray}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathrm{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\forall i \in A_{\mathrm{s}}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_\mathrm{g} - \mathbf{g}' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathrm{l}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - \mathbf{g}' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathrm{l}}
 \end{eqnarray}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_\mathrm{n} - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_\mathrm{b} &\equiv& \boldsymbol{\sigma}[s]_\mathrm{b} - v
+\boldsymbol{\sigma}^*[s]_{\mathrm{b}} &\equiv& \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_\mathrm{b} & \text{otherwise}
+\boldsymbol{\sigma}[a]_{\mathrm{b}} & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_\mathrm{a} & \equiv & a \\
-I_\mathrm{o} & \equiv & o \\
-I_\mathrm{p} & \equiv & p \\
-I_\mathrm{d} & \equiv & () \\
-I_\mathrm{s} & \equiv & s \\
-\hyperlink{I v}{I_\mathrm{v}} & \equiv & v \\
-I_\mathrm{b} & \equiv & \mathbf{i} \\
-I_\mathrm{e} & \equiv & e
-\end{eqnarray}
-
-$I_\mathrm{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+I_{\mathrm{a}} & \equiv & a \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathrm{d}} & \equiv & () \\
+I_{\mathrm{s}} & \equiv & s \\
+\hyperlink{I v}{I_{\mathrm{v}}} & \equiv & v \\
+I_{\mathrm{b}} & \equiv & \mathbf{i} \\
+I_{\mathrm{e}} & \equiv & e
+\end{eqnarray}
+
+$I_{\mathrm{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_\mathrm{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_\mathrm{b} \equiv \boldsymbol{\sigma}[r]_\mathrm{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_\mathrm{b} \equiv \boldsymbol{\sigma}[s]_\mathrm{b} - v
+\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_\mathrm{b} \equiv \boldsymbol{\sigma}_1'[s]_\mathrm{b} - v
+\boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v
 \end{equation}
@@
 \boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_\mathrm{b} \equiv \boldsymbol{\sigma}[r]_\mathrm{b} + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v & \text{otherwise}
 \end{cases}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_\mathrm{c}$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_\mathrm{a} & \equiv & r \\
-I_\mathrm{o} & \equiv & o \\
-I_\mathrm{p} & \equiv & p \\
-I_\mathrm{d} & \equiv & \mathbf{d} \\
-I_\mathrm{s} & \equiv & s \\
-\hyperlink{I v}{I_\mathrm{v}} & \equiv & \tilde{v} \\
-I_\mathrm{e} & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathrm{b}) & = & \boldsymbol{\sigma}[c]_\mathrm{c}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathrm{b}), I_\mathrm{b})$ at some point prior in order to make the determination of $I_\mathrm{b}$ feasible.
+I_{\mathrm{a}} & \equiv & r \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathrm{d}} & \equiv & \mathbf{d} \\
+I_{\mathrm{s}} & \equiv & s \\
+\hyperlink{I v}{I_{\mathrm{v}}} & \equiv & \tilde{v} \\
+I_{\mathrm{e}} & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathrm{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathrm{b}}), I_{\mathrm{b}})$ at some point prior in order to make the determination of $I_{\mathrm{b}}$ feasible.
 
@@
 \begin{itemize}\hypertarget{I a}{
-\item $I_\mathrm{a}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
-\item $I_\mathrm{o}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
-\item $I_\mathrm{p}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
-\item $I_\mathrm{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
-\item $I_\mathrm{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
-\item $I_\mathrm{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
-\item $I_\mathrm{b}$, the byte array that is the machine code to be executed.
+\item $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
+\item $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
+\item $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
+\item $I_{\mathrm{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
+\item $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
+\item $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
+\item $I_{\mathrm{b}}$, the byte array that is the machine code to be executed.
 \item $I_H$, the block header of the present block.}\hypertarget{I e}{
-\item $I_\mathrm{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
+\item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_\mathrm{g}, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_\mathrm{g} & \equiv & \mathbf{g} \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathrm{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_\mathrm{i} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathrm{s} & \equiv & ()
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & \mathbf{g} \\
+\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{m}} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{s}} & \equiv & ()
 \end{eqnarray}
@@
 
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_\mathrm{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathrm{m}$ are a series of zeroes of size $2^{256}$.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(\mathrm{g}, \mathrm{pc}, \mathbf{m}, \mathrm{i}, \mathbf{s})$ which are the gas available, the program counter $\mathrm{pc} \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The memory contents $\boldsymbol{\mu}_{\mathrm{m}}$ are a series of zeroes of size $2^{256}$.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathrm{b} \rVert \\
+w \equiv \begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathrm{b}} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_\mathrm{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathrm{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathrm{s}[0] \notin D(I_\mathrm{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathrm{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+\lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathrm{s}}[0] \notin D(I_{\mathrm{b}}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathrm{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathrm{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathrm{s}\rVert): \boldsymbol{\mu}'_\mathrm{s}[x] & \equiv & \boldsymbol{\mu}_\mathrm{s}[x-\Delta]
+\lVert\boldsymbol{\mu}'_{\mathrm{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_{\mathrm{s}}\rVert): \boldsymbol{\mu}'_{\mathrm{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathrm{s}}[x-\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)  \\
-\hypertarget{JUMP}{} \hypertarget{JUMPI}{} \label{eq:JUMP} \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)  \\
+\hypertarget{JUMP}{} \hypertarget{JUMPI}{} \label{eq:JUMP} \quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
 J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_\mathrm{m} & \equiv & \boldsymbol{\mu}_\mathrm{m} \\
-\boldsymbol{\mu}'_\mathrm{i} & \equiv & \boldsymbol{\mu}_\mathrm{i} \\
+\boldsymbol{\mu}'_{\mathrm{m}} & \equiv & \boldsymbol{\mu}_{\mathrm{m}} \\
+\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_\mathrm{t} & \equiv & B'_\mathrm{t} + B_\mathrm{d} \\
+B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
 B' & \equiv & P(B_H)
@@
 
-As such given a block $B$, $B_\mathrm{t}$ is its total difficulty, $B'$ is its parent block and $B_\mathrm{d}$ is its difficulty.
+As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
 
@@
 \begin{equation}
-\lVert \hyperlink{B U}{B_\mathrm{U}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_\mathrm{U}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert \hyperlink{B U}{B_{\mathrm{u}}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_{\mathrm{u}}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathrm{U})
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathrm{u}})
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathrm{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_\mathrm{g}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H g}{{B_H}_\mathrm{g}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_\mathrm{u}}
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathrm{t}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H g}{{B_H}_{\mathrm{g}}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_\mathrm{b}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_{\mathrm{b}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 \hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_\mathrm{c}}]_\mathrm{b} & = & \boldsymbol{\sigma}[{B_H}_\mathrm{c}]_\mathrm{b} + (1 + \frac{\lVert \hyperlink{B U}{B_\mathrm{U}}\rVert}{32})\hyperlink{block reward R b}{R_\mathrm{b}} \\
-\forall_{U \in B_\mathrm{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_\mathrm{c}]_\mathrm{b} & = & \boldsymbol{\sigma}[U_\mathrm{c}]_\mathrm{b} + (1 + \frac{1}{8} (U_\mathrm{i} - \hyperlink{H i}{{B_H}_\mathrm{i}})) \hyperlink{block reward R b}{R_\mathrm{b}}
+\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_{\mathrm{c}}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_H}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \hyperlink{B U}{B_{\mathrm{u}}}\rVert}{32})\hyperlink{block reward R b}{R_{\mathrm{b}}} \\
+\forall_{U \in B_{\mathrm{u}}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{1}{8} (U_{\mathrm{i}} - \hyperlink{H i}{{B_H}_{\mathrm{i}}})) \hyperlink{block reward R b}{R_{\mathrm{b}}}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_\mathrm{b} = 5 \times 10^{18}
+\text{Let} \quad R_{\mathrm{b}} = 5 \times 10^{18}
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_\mathrm{i}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_\mathrm{i})) = {P(B_H)_H}_\mathrm{r} & \text{otherwise}
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_H)_H}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
@@
 }
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_\mathrm{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_\mathrm{i}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 \hypertarget{Phi}{
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_\mathrm{n} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \\
-B'_\mathrm{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_\mathrm{r}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 
-\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathrm{l}$ and $\mathbf{R}[n]_\mathrm{u}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_\mathrm{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
+\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_{\mathrm{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathrm{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction Receipt}{${B_\mathrm{R}}[n]_\mathrm{u}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_\mathrm{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_{\mathrm{t}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of \hyperlink{Transaction Receipt}{${B_{\mathrm{r}}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathrm{T}[n])\\ \quad + \mathbf{R}[n-1]_\mathrm{u}
+\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_{\mathrm{t}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathrm{l}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathrm{l} =
-\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathrm{T}[n])
+For $\mathbf{R}[n]_{\mathrm{l}}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{l}} =
+\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_{\mathrm{t}}[n])
 \end{equation}
@@
 \hypertarget{Pi}{
-Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_\mathrm{R})_{\boldsymbol{\sigma}}$:
+Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_{\mathrm{r}})_{\boldsymbol{\sigma}}$:
 \begin{equation}
@@
 \begin{equation}
-m = \hyperlink{H m}{H_\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_\mathrm{n}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_\mathrm{n}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_\mathrm{d}}$ is the new block's difficulty value (i.e.
+m = \hyperlink{H m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
+\end{equation}
+
+Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_{\mathrm{n}}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_{\mathrm{d}}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_{\mathrm{b}}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_{\mathrm{l}}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \hypertarget{R b}{
-Formally, we define $R_\mathrm{b}$:}
-\begin{eqnarray}
-R_\mathrm{b}(\mathbf{x}) & \equiv & \begin{cases}
+Formally, we define $R_{\mathrm{b}}$:}
+\begin{eqnarray}
+R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \hypertarget{R l}{
-Thus we finish by formally defining $R_\mathrm{l}$:
-\begin{eqnarray}
-R_\mathrm{l}(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_{\mathrm{l}}$:
+\begin{eqnarray}
+R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_\mathrm{r} \\
-(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_\mathrm{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_\mathrm{r}$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_{\mathrm{r}} \\
+(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_{\mathrm{r}}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathrm{r} &=& 3000\\
+\mathbf{g}_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_\mathrm{d}|-1)] &=& I_\mathrm{d}\\
-\mathbf{d}[|I_\mathrm{d}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|I_{\mathrm{d}}|-1)] &=& I_{\mathrm{d}}\\
+\mathbf{d}[|I_{\mathrm{d}}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathrm{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathrm{d})\\
+\mathbf{g}_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathrm{d}}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathrm{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathrm{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
+\mathbf{g}_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathrm{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathrm{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathrm{d}})\\
 \end{eqnarray}
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-\mathbf{g}_\mathrm{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathrm{d}
+\mathbf{g}_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathrm{d}}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_{\mathrm{d}}
 \end{eqnarray}
@@
 \hypertarget{h T}{
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_\mathrm{r}$, $T_\mathrm{s}$ and $\hyperlink{T w}{T_\mathrm{w}}$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $\hyperlink{T w}{T_{\mathrm{w}}}$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{i}) & \text{if} \; T_\mathrm{t} = 0\\
-(T_\mathrm{n}, T_\mathrm{p}, T_\mathrm{g}, T_\mathrm{t}, \hyperlink{T v}{T_\mathrm{v}}, T_\mathrm{d}) & \text{otherwise}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}) & \text{if} \; T_{\mathrm{t}} = 0\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}) & \text{otherwise}
 \end{cases} \\
@@
 G(T, p_r) \equiv T \quad \text{except:} \\
-(\hyperlink{T w}{T_\mathrm{w}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+(\hyperlink{T w}{T_{\mathrm{w}}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}}
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_\mathrm{w}}, T_\mathrm{r}, T_\mathrm{s}) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big)
 \end{equation}
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathrm{i})-C_{mem}(\boldsymbol{\mu}_\mathrm{i}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathrm{s}[1] = 0 \\
-G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathrm{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathrm{s}[1] > 0 \\
-G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathrm{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathrm{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathrm{s}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathrm{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathrm{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathrm{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 \begin{equation}
-w \equiv \begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathrm{b} \rVert\\
+w \equiv \begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathrm{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 
-Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathrm{i}$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_\mathrm{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 0x01 & {\small ADD} & 2 & 1 & This is the addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & This is the multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] \times \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] \times \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & This is the subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] - \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] - \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & This is the integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathrm{s}}[0] \div \boldsymbol{\mu}_{\mathrm{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & This is the signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathrm{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]) \lfloor |\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathrm{s}}[0] \div \boldsymbol{\mu}_{\mathrm{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathrm{s}}[0] \div \boldsymbol{\mu}_{\mathrm{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x06 & {\small MOD} & 2 & 1 & This is the modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \boldsymbol{\mu}_\mathrm{s}[0] \bmod \boldsymbol{\mu}_\mathrm{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathrm{s}}[0] \bmod \boldsymbol{\mu}_{\mathrm{s}}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & This is the signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathrm{s}[0]) (|\boldsymbol{\mu}_\mathrm{s}[0]| \bmod |\boldsymbol{\mu}_\mathrm{s}[1]|) & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathrm{s}}[0]) (|\boldsymbol{\mu}_{\mathrm{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathrm{s}}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & This is the modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] = 0\\ (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1]) \mod \boldsymbol{\mu}_\mathrm{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1]) \mod \boldsymbol{\mu}_{\mathrm{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x09 & {\small MULMOD} & 3 & 1 & This is the modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] = 0\\ (\boldsymbol{\mu}_\mathrm{s}[0] \times \boldsymbol{\mu}_\mathrm{s}[1]) \mod \boldsymbol{\mu}_\mathrm{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathrm{s}}[0] \times \boldsymbol{\mu}_{\mathrm{s}}[1]) \mod \boldsymbol{\mu}_{\mathrm{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x0a & {\small EXP} & 2 & 1 & This is the exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] ^ {\boldsymbol{\mu}_\mathrm{s}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] ^ {\boldsymbol{\mu}_{\mathrm{s}}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathrm{s}[0] + 1) \\ \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathrm{s}[x]_\mathrm{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathrm{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathrm{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathrm{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathrm{s}}[x]$} \vspace{5pt} \\
 \end{tabularx}
@@
 0x10 & {\small LT} & 2 & 1 & This is the less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] < \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] < \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x11 & {\small GT} & 2 & 1 & This is the greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x12 & {\small SLT} & 2 & 1 & This is the signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] < \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] < \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x13 & {\small SGT} & 2 & 1 & This is the signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x14 & {\small EQ} & 2 & 1 & This is the equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] = \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & This is the logical negation operation, also called the logical complement or the NOT operation. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} \wedge \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} \vee \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} \oplus \boldsymbol{\mu}_\mathrm{s}[1]_\mathrm{i}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0]_\mathrm{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_\mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[1]_{(i + 8\boldsymbol{\mu}_\mathrm{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathrm{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathrm{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathrm{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathrm{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute a Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{a}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0]]_\mathrm{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{o}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{o}}$ \\
 &&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{s}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \hyperlink{I v}{I_\mathrm{v}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \hyperlink{I v}{I_{\mathrm{v}}}$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get the input data of the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{d}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ] \quad \text{with} \quad I_\mathrm{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathrm{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{d}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathrm{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathrm{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get the size of the input data in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert I_\mathrm{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \lVert I_{\mathrm{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy the input data in the current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
-\begin{cases} I_\mathrm{d}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert I_\mathrm{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathrm{d}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] + i < \lVert I_{\mathrm{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathrm{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get the size of code running in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert I_\mathrm{b} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \lVert I_{\mathrm{b}} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in the current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
-\begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert I_\mathrm{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] + i < \lVert I_{\mathrm{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathrm{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{p}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is the gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{c} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[3] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathrm{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{c}$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[1], \boldsymbol{\mu}_\mathrm{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathrm{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[1], \boldsymbol{\mu}_{\mathrm{s}}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathrm{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv P(I_{\hyperlink{H p}{H_\mathrm{p}}}, \boldsymbol{\mu}_\mathrm{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv P(I_{\hyperlink{H p}{H_{\mathrm{p}}}}, \boldsymbol{\mu}_{\mathrm{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
@@
 \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_\mathrm{i}} \\ P(\hyperlink{H p}{H_\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_{\mathrm{i}}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_{\mathrm{i}}} \\ P(\hyperlink{H p}{H_{\mathrm{p}}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{c}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{s}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{d}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_\mathrm{l}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{l}}$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load the first word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv \max(\boldsymbol{\mu}_\mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathrm{s}}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x52 & {\small MSTORE} & 2 & 0 & Save the first word (which is the first item in the stack) to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv \max(\boldsymbol{\mu}_\mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathrm{s}}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x53 & {\small MSTORE8} & 2 & 0 & Save the first byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] ] \equiv (\boldsymbol{\mu}_\mathrm{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv \max(\boldsymbol{\mu}_\mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_\mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathrm{s}}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathrm{s}}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]]$ \hypertarget{SSTORE}{\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]]$ \hypertarget{SSTORE}{\\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_\mathrm{a}]_\mathrm{s}[ \boldsymbol{\mu}_\mathrm{s}[0] ] \equiv \boldsymbol{\mu}_\mathrm{s}[1] $ \hypertarget{C tiny SSTORE}{\\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{s}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1] $ \hypertarget{C tiny SSTORE}{\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]] = 0 \\
+G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]] = 0 \\
 G_{\mathrm{sreset}} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]] \neq 0 \\
+R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathrm{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
 \midrule
 0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathrm{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
 \midrule
 0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv 32\boldsymbol{\mu}_{\mathrm{i}}$ \\
 \midrule
 0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{g}}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place a 1 byte item on the stack. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathrm{b}[x] & \text{if} \quad x < \lVert I_\mathrm{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathrm{b}}[x] & \text{if} \quad x < \lVert I_{\mathrm{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place a 2-byte item on the stack. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 0x80 & {\small DUP1} & 1 & 2 & Duplicate the 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
 \midrule
 0x81 & {\small DUP2} & 2 & 3 & Duplicate the 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
 \midrule
@@
 0x8f & {\small DUP16} & 16 & 17 & Duplicate the 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[15]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[15]$ \\
 \bottomrule
@@
 0x90 & {\small SWAP1} & 2 & 2 & Exchange the 1st and the 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[1] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[1] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
 \midrule
 0x91 & {\small SWAP2} & 3 & 3 & Exchange the 1st and the 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[2] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[2]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[2] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
 \midrule
@@
 0x9f & {\small SWAP16} & 17 & 17 & Exchange the 1st and the 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[16] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[16]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[16] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
 \bottomrule
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathrm{l} \equiv A_\mathrm{l} \cdot (I_\mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_{\mathrm{l}} \equiv A_{\mathrm{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xa1 & {\small LOG1} & 3 & 0 & Append the log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathrm{s}[2])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 \midrule
@@
 0xa4 & {\small LOG4} & 6 & 0 & Append the log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[3], \boldsymbol{\mu}_\mathrm{s}[4], \boldsymbol{\mu}_\mathrm{s}[5])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathrm{s}}[2], \boldsymbol{\mu}_{\mathrm{s}}[3], \boldsymbol{\mu}_{\mathrm{s}}[4], \boldsymbol{\mu}_{\mathrm{s}}[5])$ \\
 \bottomrule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[1] \dots (\boldsymbol{\mu}_\mathrm{s}[1] + \boldsymbol{\mu}_\mathrm{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_\mathrm{g}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_\mathrm{a}, I_\mathrm{o}, L(\boldsymbol{\mu}_\mathrm{g}), I_\mathrm{p}, \boldsymbol{\mu}_\mathrm{s}[0], \mathbf{i}, I_\mathrm{e} + 1) & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge\; I_\mathrm{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathrm{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_\mathrm{a}]_\mathrm{n} = \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{n} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathrm{s} \equiv A_\mathrm{s} \cup A^+_\mathrm{s} \quad \wedge \quad A'_\mathrm{l} \equiv A_\mathrm{l} \cdot A^+_\mathrm{l} \quad \wedge \quad A'_\mathrm{r} \equiv A_\mathrm{r} + A^+_\mathrm{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[1] \dots (\boldsymbol{\mu}_{\mathrm{s}}[1] + \boldsymbol{\mu}_{\mathrm{s}}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1) & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_{\mathrm{s}} \equiv A_{\mathrm{s}} \cup A^+_{\mathrm{s}} \quad \wedge \quad A'_{\mathrm{l}} \equiv A_{\mathrm{l}} \cdot A^+_{\mathrm{l}} \quad \wedge \quad A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + A^+_{\mathrm{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_\mathrm{e} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_\mathrm{a}, \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{n})$, the address of the  \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the  \\
 &&&& newly created account. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[1], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[1], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[3] \dots (\boldsymbol{\mu}_\mathrm{s}[3] + \boldsymbol{\mu}_\mathrm{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_\mathrm{a}, I_\mathrm{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathrm{p}, \boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[2], \mathbf{i}, I_\mathrm{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge \\ \quad\quad I_\mathrm{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathrm{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[5] \dots (\boldsymbol{\mu}_\mathrm{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{g} \equiv \boldsymbol{\mu}_\mathrm{g} + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[3] \dots (\boldsymbol{\mu}_{\mathrm{s}}[3] + \boldsymbol{\mu}_{\mathrm{s}}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[2], \boldsymbol{\mu}_{\mathrm{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathrm{s}}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[5] \dots (\boldsymbol{\mu}_{\mathrm{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathrm{s}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathrm{s}}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
 &&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathrm{s}[2] > \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b}$ (not enough funds) or $I_\mathrm{e} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_{\mathrm{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[3], \boldsymbol{\mu}_\mathrm{s}[4]), \boldsymbol{\mu}_\mathrm{s}[5], \boldsymbol{\mu}_\mathrm{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[3], \boldsymbol{\mu}_{\mathrm{s}}[4]), \boldsymbol{\mu}_{\mathrm{s}}[5], \boldsymbol{\mu}_{\mathrm{s}}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_\mathrm{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_\mathrm{s}[0] \} & \text{if} \quad \boldsymbol{\mu}_\mathrm{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_\mathrm{s}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathrm{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_{\mathrm{s}}[0] & \text{otherwise}
 \end{cases}$\\
@@
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
+G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[1] \mod 2^{160}] = \varnothing \\
+G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathrm{a}, I_\mathrm{o}, I_\mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_\mathrm{p}, \boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[2], \mathbf{i}, I_\mathrm{e} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge\\ \quad\quad{}I_\mathrm{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathrm{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_\mathrm{a}$. This means that the recipient is in fact the\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[2], \boldsymbol{\mu}_{\mathrm{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathrm{s}}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
@@
 0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots ( \boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots ( \boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See equation \ref{eq:Normal Halting}. \\
-&&&& $\boldsymbol{\mu}'_\mathrm{i} \equiv M(\boldsymbol{\mu}_\mathrm{i}, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$ \\
 \end{tabularx}
@@
 Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
-The omitted argument is $\boldsymbol{\mu}_\mathrm{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$, $\boldsymbol{\mu}_\mathrm{s}[5]$ and $\boldsymbol{\mu}_\mathrm{s}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathrm{s}[2]$, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$ and $\boldsymbol{\mu}_\mathrm{s}[5]$.
+The omitted argument is $\boldsymbol{\mu}_{\mathrm{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathrm{s}}[3]$, $\boldsymbol{\mu}_{\mathrm{s}}[4]$, $\boldsymbol{\mu}_{\mathrm{s}}[5]$ and $\boldsymbol{\mu}_{\mathrm{s}}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathrm{s}}[2]$, $\boldsymbol{\mu}_{\mathrm{s}}[3]$, $\boldsymbol{\mu}_{\mathrm{s}}[4]$ and $\boldsymbol{\mu}_{\mathrm{s}}[5]$.
 Otherwise it is exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_\mathrm{s}, I_\mathrm{o}, I_\mathrm{a}, t,\\\quad \boldsymbol{\mu}_\mathrm{s}[0], I_\mathrm{p}, 0, \hyperlink{I v}{I_\mathrm{v}}, \mathbf{i}, I_\mathrm{e} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b} \;\wedge\; I_\mathrm{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathrm{s}}[0], I_{\mathrm{p}}, 0, \hyperlink{I v}{I_{\mathrm{v}}}, \mathbf{i}, I_{\mathrm{e}} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_{\mathrm{v}}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. }\\
-&&&& $A'_\mathrm{s} \equiv A_\mathrm{s} \cup \{ I_\mathrm{a} \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{b} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_\mathrm{b} + \boldsymbol{\sigma}[I_\mathrm{a}]_\mathrm{b}$ \\
-&&&& $\boldsymbol{\sigma}'[I_\mathrm{a}]_\mathrm{b} \equiv 0$ \\
+&&&& $A'_{\mathrm{s}} \equiv A_{\mathrm{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_\mathrm{a} \notin A_\mathrm{s} \\
+\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathrm{s}} \\
 0 & \text{otherwise} 
@@
 &&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{\mathrm{selfdestruct}}} + \begin{cases}
-\hyperlink{Gnewaccount}{G_{\mathrm{newaccount}}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}] = \varnothing \\
+\hyperlink{Gnewaccount}{G_{\mathrm{newaccount}}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \begin{equation}
-E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathrm{i}}) = \left\lfloor\frac{\hyperlink{H i}{H_\mathrm{i}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
+E_{\mathrm{epoch}}(\hyperlink{H i}{H_{\mathrm{i}}}) = \left\lfloor\frac{\hyperlink{H i}{H_{\mathrm{i}}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- s = C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathrm{i}})
-\end{equation}
-\begin{equation}
- C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathrm{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_\mathrm{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_\mathrm{i}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
+ s = C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_{\mathrm{i}}})
+\end{equation}
+\begin{equation}
+ C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_{\mathrm{i}}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_{\mathrm{i}}}) = 0 \quad  \\
+\texttt{KEC}(C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_{\mathrm{i}}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
+We define the array $\mathbf{c}_{\mathrm{i}}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
+\begin{equation}
+ \mathbf{c}_{\mathrm{i}} = \begin{cases}
 \texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+ \mathbf{c'}[i] = \mathbf{c}_{\mathrm{i}} \quad \forall \quad i < n
 \end{equation}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathrm{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathrm{n}}) + \mathbf{m}_\mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathrm{n}}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathrm{c}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_\mathrm{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_\mathrm{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathrm{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_{\mathrm{n}}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_\mathrm{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
+ \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
 \end{equation}

### 2017-10-20 Paper.tex

@@
+%Note that the Travis CI build is using Trusty, thus the latest version of LaTeX can't be used, which is not desirable.
 \documentclass[9pt,oneside]{amsart}
@@
 \usepackage{tikz}
-\usepackage{underscore}
+%\usepackage{underscore}
 %\usepackage{lipsum} % to generate filler text for testing of document rendering
@@
 In order to follow back the history of a path, one might need to reference multiple versions of this document.
+The Ethereum protocol underwent a Byzantium hard fork, the first phase of Metropolis, at block 4370000 on Oct 16 2017 (\cite{EthblogByzantiumHFAnnouncement}, \cite{Codetract}, \cite{EtherscanBlock4370000}).
+Thus, this document version is currently not the current implementation of the Ethereum protocol, and it will be updated soon.
+For further information about previous protocol release revisions, see \cite{EthdocsPRroadmap} and \cite{EthblogDevRoadmap}.
 
@@
 \item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.}\hypertarget{T v}{
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.}\hypertarget{T w}{
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{w}}$}}, \hyperlink{T r}{$T_{\mathrm{r}}$} and \hyperlink{T s}{$T_{\mathrm{s}}$}, respectively.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.}
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hyperlink{T w}{$T_{\mathrm{w}}$}, \hyperlink{T r}{$T_{\mathrm{r}}$} and \hyperlink{T s}{$T_{\mathrm{s}}$}, respectively.
 This is expanded in Appendix \ref{app:signing}.
@@
 \item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_{\mathrm{e}}$}.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_{\mathrm{b}}.$}\hyperlink{H d}{
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_{\mathrm{b}}.$}
 \item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$}.\hypertarget{H i}{
-\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_{\mathrm{i}}$.}\hyperlink{H l}{
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.}\hypertarget{H g}{
+This can be calculated from the previous block's difficulty level and the timestamp; formally \hyperlink{H d}{$H_{\mathrm{d}}$}.\hypertarget{H i}{
+\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_{\mathrm{i}}$.}
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally \hyperlink{H l}{$H_{\mathrm{l}}$}.\hypertarget{H g}{
 \item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.}\hypertarget{H s}{
@@
 
-Note that \mindifficulty is the difficulty of the genesis block.
-The effect of the difficulty parameter \homesteadmod on a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-\expdiffsymb results in the difficulty increasing every 10000 blocks.
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The difficulty parameter $\homesteadmod$ is designed to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+$\expdiffsymb$ results in the difficulty increasing every 10000 blocks.
 This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law (\cite{THOMPSON200620}).
@@
 
-The tuple contents include $A_{\mathrm{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_{\mathrm{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+The tuple contents include \hyperlink{A s}{$A_{\mathrm{s}}$}, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$\hyperlink{A l}{A_{\mathrm{l}}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $\hyperlink{A r}{A_{\mathrm{r}}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{A r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $\hyperlink{A r}{A_{\mathrm{r}}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \hypertarget{A s}{}\\
 \forall i \in A_{\mathrm{s}}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
@@
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - \mathbf{g}' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathrm{l}}
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{A l}{A_{\mathrm{l}}}
 \end{eqnarray}
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+\hypertarget{endow}{}There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
 Thus we define the resultant address for the new account $a$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, \hyperlink{endow}{v} + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s]_{\mathrm{b}} &\equiv& \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
@@
 
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
+Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a \hyperlink{hhalt}{natural halting} state.
 In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
@@
 \begin{equation}
-c \equiv G_{\mathrm{codedeposit}} \times |\mathbf{o}|
+c \equiv \hyperlink{codedeposit}{G_{\mathrm{codedeposit}}} \times |\mathbf{o}|
 \end{equation}
@@
 If the initialisation execution ends with a \hyperlink{selfdestruct}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
-For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
+For a normal \hyperlink{stop}{\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
 
@@
 The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
+Just as with contract creation, if the execution \hyperlink{zhalt}{halts in an exceptional fashion} (i.e.
 due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
@@
 Like the out-of-gas exception, they do not leave state changes intact.
-Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
+Rather, the machine \\hyperlink{zhalt}{halts} immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
 
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
 
 \hypertarget{emptySequence}{}
-The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of \hyperlink{hhalt}{$H$}, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
 \begin{eqnarray}
@@
 
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
@@
 w \equiv \begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathrm{b}} \rVert \\
-\text{\small STOP} & \text{otherwise}
+\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
 \end{cases}
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
+\hypertarget{zhalt}{}
 \subsubsection{Exceptional Halting}
@@
 
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined {\small STOP} operations that trail it).
+All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{\small STOP} operations that trail it).
 
@@
 \end{equation}
-
+\hypertarget{hhalt}{}
 \subsubsection{Normal Halting} \label{subsec:Normal Halting}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \hyperlink{RETURN}{\text{\small RETURN}} \\
+() & \text{if} \quad w \in \text{\hyperlink{stop}{\small STOP}}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
 \varnothing & \text{otherwise}
@@
 One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}, \cite{ASICWiki}).
+These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}).
 In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
@@
 
-\hypertarget{rlp}{}
+\hypertarget{rlp}{}W
 \section{Recursive Length Prefix}\label{app:rlp}
@@
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\ e & \equiv & \hyperlink{h T}{h(T)} \\
+\ e & \equiv & \hyperlink{h T}{h(T)} \hypertarget{ECDSASIGN}{}\\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-\hypertarget{v}{It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 
@@
 \hypertarget{h T}{
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $\hyperlink{T w}{T_{\mathrm{w}}}$:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$:
 \begin{eqnarray}
@@
 
-\hyperlink{T w}{}\hypertarget{T r}{}\hypertarget{T s}{The signed transaction $G(T, p_r)$ is defined as:
+The signed transaction $G(T, p_r)$ is defined as:
 \begin{eqnarray}
 G(T, p_r) \equiv T \quad \text{except:} \\
-(\hyperlink{T w}{T_{\mathrm{w}}}, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}}
-
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
+
+Where the output of $\mathtt{\small ECDSASIGN}$ is also \hyperlink{ECDSASIGN}{defined previously}. Thus:
+\begin{eqnarray}\hypertarget{T w}{}\hypertarget{T r}{}\hypertarget{T s}{}
+T_{\mathrm{w}} = \hyperlink{v}{v}\\
+T_{\mathrm{r}} = \hyperlink{r}{r}\\
+T_{\mathrm{s}} = \hyperlink{s}{s}
+\end{eqnarray}
 We may then define the sender function $S$ of the transaction as:
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big)
 \end{equation}
@@
 $G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
-$G_{\mathrm{sset}}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$G_{\mathrm{sset}}$ & 20000 & This is paid for an \hyperlink{SSTORE}{{\small SSTORE}} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an \hyperlink{SSTORE}{{\small SSTORE}} operation when the storage value's zeroness remains unchanged or is set to zero. \\
 $R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
@@
 $G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
-$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \\
+$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \hypertarget{codedeposit}{}\\
 $G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
@@
 C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
+C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{SSTORE}{{\small SSTORE}}} \\
 G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1] = 0 \\
@@
 w \equiv \begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathrm{b}} \rVert\\
-\text{\small STOP} & \text{otherwise}
+\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
 \end{cases}
@@
 
-$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}\}
+$W_{\mathrm{zero}}$ = \{{\hyperlink{stop}{\small STOP}, \hyperlink{RETURN}{\text{\small RETURN}}\}
 
@@
 \multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & This operation halts execution, outputting the empty sequence as per equation \ref{eq:Normal Halting}.
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{stop}{}\\
+0x00 & {\small STOP} & 0 & 0 & This operation \hyperlink{hhalt}{halts execution}.
 \\
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] + i < \lVert I_{\mathrm{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] + i < \lVert I_{\mathrm{b}} \rVert \\ \text{\hyperlink{stop}{\small STOP}} & \text{otherwise} \end{cases}$\\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
@@
 &&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathrm{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathrm{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\hyperlink{stop}{\small STOP}} & \text{otherwise} \end{cases}$\\
 &&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
@@
 G_{\mathrm{sreset}} & \text{otherwise}
-\end{cases}$ }\\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+\end{cases}$ }\hypertarget{A r}{}\\
+&&&& $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
 R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]] \neq 0 \\
@@
 \multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
+\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\hypertarget{A l}{}\\
 \multicolumn{5}{l}{$A'_{\mathrm{l}} \equiv A_{\mathrm{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1) ])$}\\
@@
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_{\mathrm{s}} \equiv A_{\mathrm{s}} \cup A^+_{\mathrm{s}} \quad \wedge \quad A'_{\mathrm{l}} \equiv A_{\mathrm{l}} \cdot A^+_{\mathrm{l}} \quad \wedge \quad A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + A^+_{\mathrm{r}}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_{\mathrm{s}} \equiv \hyperlink{A s}{A_{\mathrm{s}}} \cup A^+_{\mathrm{s}} \quad \wedge \quad A'_{\mathrm{l}} \equiv \hyperlink{A l}{A_{\mathrm{l}}} \cdot A^+_{\mathrm{l}} \quad \wedge \quad A'_{\mathrm{r}} \equiv \hyperlink{A r}{A_{\mathrm{r}}} + A^+_{\mathrm{r}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting: \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hypertarget{zhalt}{exceptional halting}: \\
 &&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_{\mathrm{e}} = 1024$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_{\mathrm{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hypertarget{zhalt}{exceptional halting} \\
 &&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
@@
 \midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+0xf3 & {\small RETURN} & 2 & 0 & \hyperlink{hhalt}{Halt execution} returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots ( \boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See equation \ref{eq:Normal Halting}. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$ \\
@@
 \midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. }\\
-&&&& $A'_{\mathrm{s}} \equiv A_{\mathrm{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+0xff & {\small SELFDESTRUCT} & 1 & 0 & \hyperlink{hhalt}{Halt execution} and register account for later deletion. }\\
+&&&& $A'_{\mathrm{s}} \equiv \hyperlink{A s}{A_{\mathrm{s}}} \cup \{ I_{\mathrm{a}} \}$ \\
 &&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathrm{s}} \\
+\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_{\mathrm{a}} \notin \hyperlink{A s}{A_{\mathrm{s}}} \\
 0 & \text{otherwise} 
@@
 
-%Is in the terminology list.
+%This is in the terminology list.
 %\section{Low-level Lisp-like Language}\label{app:lll}

### 2018-01-02 Paper.tex

--- 2017-11-18 Paper.tex
@@
+%Note that the Travis CI build is using Trusty, thus the latest version of LaTeX can't be used, which is not desirable.
 \documentclass[9pt,oneside]{amsart}
+\usepackage{cancel}
 %\usepackage{tweaklist}
-\usepackage{url}
-\usepackage{cancel}
+
+%You don't use inputenc with XeTeX.
+%\usepackage[utf8]{inputenc}
+%\DeclareUnicodeCharacter{00A0}{ }
+%\DeclareUnicodeCharacter{C3 ab}{"e}
+%\DeclareUnicodeCharacter{C3 bc}{"u}
+%For testing: \DeclareUnicodeCharacter{AD}{\Huge [ICI]}
 \usepackage{xspace}
@@
 \usepackage{tikz}
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+%,pagebackref=true
+
+%\usepackage{longtable}
+%https://stackoverflow.com/questions/1918366/automated-line-breaks-in-latex-tables/1918465#1918465
+%tabularx seems better, but we can include this if we need tables to break across pages automatically, as well as tabu and longtabu
+
+\usepackage{tabularx}
+%https://stackoverflow.com/a/1919447/7438857
+
+% Tabu seems more flexible, and generally better.
+\usepackage{tabu} %requires array.
+
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
+% "hyperref loads the url package internally.
+Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
+This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
+% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
+Try loading the url package earlier than the package that requires it.
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+%\usepackage{glossaries}
 
@@
 A valid state transition is one which comes about through a transaction.
-Formally:
-\begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+Formally:\hypertarget{Upsilon}{}
+\begin{equation}
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{\mathrm{t}}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_{\mathrm{n}}$, denotes the nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
 Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathrm{s}}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathrm{m}}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 \begin{description}
+\hypertarget{account nonce}{}
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_{\mathrm{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the \hyperlink{f*}{element-wise transformation} of the base function $L_I$, given as:
 \begin{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_{\mathrm{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
 \end{equation}
@@
 \end{equation}
-where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{equation}
-
-\subsection{The Transaction} \label{ch:transaction}
+\hypertarget{v x}{where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+}
+\end{equation}
+
+\subsection{The Transaction} \label{subsec:transaction}
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\hypertarget{T n}{
+\item[nonce] \label{itm:nonce} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.}\hypertarget{T p}{
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.}\hypertarget{T g}{
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.}\hypertarget{T t}{
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.}\hypertarget{T v}{
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.}
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hyperlink{T w}{$T_{\mathrm{w}}$}, \hyperlink{T r}{$T_{\mathrm{r}}$} and \hyperlink{T s}{$T_{\mathrm{s}}$}, respectively.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathrm{i}}$.
 \end{description}
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathrm{d}}$.
 \end{description}
 
-Appendix \ref{app:signing} specifies the function, $S$, which maps transactions to the sender, and happens through the ECDSA of the SECP-256k1 curve, using the hash of the transaction (excepting the latter three signature fields) as the datum to sign.
-For the present we simply assert that the sender of a given transaction $T$ can be represented with $S(T)$.
-
-\begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+\begin{equation}
+L_{\mathrm{t}}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}, \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}, \hyperlink{T w}{T_{\mathrm{w}}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
 \end{cases}
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathrm{i}}$ and $T_{\mathrm{d}}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & \hyperlink{T v}{T_{\mathrm{v}}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & \hyperlink{T w}{T_{\mathrm{w}}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{d}} \in \mathbb{B} & \wedge & T_{\mathrm{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_{\mathrm{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_{\mathrm{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\begin{description}\hypertarget{H p}{
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $\hyperlink{H o}{H_{\mathrm{o}}}$.\hypertarget{H c}{
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block shall be transferred; formally $H_{\mathrm{c}}$.}
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state \hyperlink{trie}{trie}, after all transactions are executed and finalisations applied; formally \hyperlink{H r}{$H_{\mathrm{r}}$}.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with each transaction in the transactions list portion of the block; formally \hyperlink{H t}{$H_{\mathrm{t}}$}.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the \hyperlink{trie}{trie} structure populated with the receipts of each transaction in the transactions list portion of the block; formally \hyperlink{H e}{$H_{\mathrm{e}}$}.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally \hyperlink{H b}{$H_{\mathrm{b}}.$}
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally \hyperlink{H d}{$H_{\mathrm{d}}$}.\hypertarget{H i}{
+\item[number] A scalar value equal to the number of ancestor blocks. The \hyperlink{GenesisBlock}{genesis block} has a number of zero; formally $H_{\mathrm{i}}$.}
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally \hyperlink{H l}{$H_{\mathrm{l}}$}.\hypertarget{H g}{
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.}\hypertarget{H s}{
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally {$H_{\mathrm{s}}$}.}\hypertarget{H x}{
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.}\hypertarget{H m}{
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.}
+\hypertarget{block nonce}{\hypertarget{H n}{
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.}}
 \end{description}
-
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}
+\hypertarget{B U}{
+The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathrm{u}}$, and a series of the transactions, $B_{\mathrm{t}}$.
+Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_{\mathrm{t}}, B_{\mathrm{u}})
+\end{equation}}
+
+\hypertarget{Transaction Receipt}{\subsubsection{Transaction Receipt}}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
-\begin{equation}
-R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+Each receipt, denoted $B_{\mathrm{r}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{H e}{$H_{\mathrm{e}}$}.
+
+\hypertarget{R}{}
+The transaction receipt, $R$, is a tuple of four items \hypertarget{R sigma}{comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$}, \hypertarget{R u}{the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$}, the set of logs created through execution of the transaction, \hyperlink{R l}{$R_{\mathrm{l}}$} and the Bloom filter composed from information in those logs, \hyperlink{R b}{$R_{\mathrm{b}}$}:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_{\mathrm{u}}, \hyperlink{R b}{R_{\mathrm{b}}}, \hyperlink{R l}{R_{\mathrm{l}}})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
-\end{equation}
-thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+L_R(R) \equiv (\mathtt{\small \hyperlink{trie}{TRIE}}(L_S(R_{\boldsymbol{\sigma}})), \hyperlink{R u}{R_{\mathrm{u}}}, \hyperlink{R b}{R_{\mathrm{b}}}, \hyperlink{R l}{R_{\mathrm{l}}})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a \hyperlink{trie}{trie} structure, the root of which forms the first item.
+
+We assert $\hyperlink{R u}{R_{\mathrm{u}}}$, the cumulative gas used is a positive integer and that the logs Bloom, \hyperlink{R b}{$R_{\mathrm{b}}$}, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+\hyperlink{R u}{R_{\mathrm{u}}} \in \mathbb{P} \quad \wedge \quad \hyperlink{R b}{R_{\mathrm{b}}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_{\mathrm{t}}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $\hyperlink{R l}{R_{\mathrm{l}}}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_{\mathrm{t}}$ and some number of bytes of data, $O_{\mathrm{d}}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_{\mathrm{t}}}_0, {O_{\mathrm{t}}}_1, ...), O_{\mathrm{d}})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathrm{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathrm{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathrm{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
-\begin{equation}
-\begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathrm{t}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+\begin{equation}
+\begin{array}[t]{lclc}\hypertarget{H r}{
+H_{\mathrm{r}} &\equiv& \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}, B))) & \wedge }\hypertarget{H o}{\\
+H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(\hyperlink{B U}{B_{\mathrm{u}}}))) & \wedge }\hypertarget{H t}{\\
+H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathrm{t}} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{t}}(B_{\mathrm{t}}[i]))\}) }& \wedge \hypertarget{H e}{\\
+H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert \hyperlink{Transaction Receipt}{B_{\mathrm{r}}} \rVert, i \in \mathbb{P}: p(i, L_R(B_{\mathrm{r}}[i]))\}) & \wedge }\hypertarget{H b}{\hypertarget{r}{\\
+H_{\mathrm{b}} &\equiv& \hyperlink{bigvee}{\bigvee}_{\mathbf{r} \in B_{\mathrm{r}}} \big( \mathbf{r}_{\mathrm{b}} \big)}}
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
 \end{equation}
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction Receipt}{transaction receipts, $B_{\mathrm{r}}$}, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}\hyperlink{H p}{H_{\mathrm{p}}}, \hyperlink{H o}{H_{\mathrm{o}}}, \hyperlink{H c}{H_{\mathrm{c}}}, \hyperlink{H r}{H_{\mathrm{r}}}, \hyperlink{H t}{H_{\mathrm{t}}}, \hyperlink{H e}{H_{\mathrm{e}}}, \hyperlink{H b}{H_{\mathrm{b}}}, \hyperlink{H d}{H_{\mathrm{d}}},\\ \hyperlink{H i}{H_{\mathrm{i}}}, \hyperlink{H l}{H_{\mathrm{l}}}, \hyperlink{H g}{H_{\mathrm{g}}}, \hyperlink{H s}{H_{\mathrm{s}}}, \hyperlink{H x}{H_{\mathrm{x}}}, \hyperlink{H m}{H_{\mathrm{m}}}, \hyperlink{H n}{H_{\mathrm{n}}} \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_{\mathrm{t}}^*(B_{\mathrm{t}}), L_H^*(\hyperlink{B U}{B_{\mathrm{u}}}) \big)
+\end{eqnarray}
+
+\hypertarget{f*}{
+With $L_{\mathrm{t}}^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
 f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
-\end{equation}
+\end{equation}}
 
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+\hyperlink{H p}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H o}{H_{\mathrm{o}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H c}{H_{\mathrm{c}}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{H r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H t}{H_{\mathrm{t}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
+H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & \hyperlink{H d}{H_{\mathrm{d}}} \in \mathbb{P} & \wedge & \hyperlink{H i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
+\hyperlink{H l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & \hyperlink{H g}{H_{\mathrm{g}}} \in \mathbb{P} & \wedge & \hyperlink{H s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{H x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & \hyperlink{H m}{H_{\mathrm{m}}} \in \mathbb{B}_{32} & \wedge & \hyperlink{H n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = \hyperlink{H p}{H_{\mathrm{p}}}
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+\hyperlink{H i}{H_{\mathrm{i}}} \equiv {{P(H)_H}_{\mathrm{i}}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad \hyperlink{H i}{H_{\mathrm{i}}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
-
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{\hyperlink{H s}{H_{\mathrm{s}}} - {P(H)_H}_{\mathrm{s}}}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor \hyperlink{H i}{H_{\mathrm{i}}} \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}}
+
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The difficulty parameter $\homesteadmod$ is designed to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+$\expdiffsymb$ results in the difficulty increasing every 10000 blocks.
+This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law (\cite{THOMPSON200620}).
+
+\hypertarget{H l}{
+The canonical gas limit $\hyperlink{H l}{H_{\mathrm{l}}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & \hyperlink{H l}{H_{\mathrm{l}}} < {P(H)_H}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} > {P(H)_H}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} \geqslant 125000
+\end{eqnarray}}
+
+\hyperlink{H s}{$H_{\mathrm{s}}$} is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+\hyperlink{H s}{H_{\mathrm{s}}} > {P(H)_H}_{\mathrm{s}}
+\end{equation}
+
+This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation is required, lengthening the likely next period.
 Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+The nonce of a block, $\hyperlink{H n}{H_{\mathrm{n}}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \quad \wedge \quad m = \hyperlink{H m}{H_{\mathrm{m}}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})$.
+
+\hypertarget{h cancel n}{}
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the \hyperlink{block nonce}{nonce} and mix-hash block components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $\hyperlink{H d}{H_{\mathrm{d}}}$.
 
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
-
-Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 125000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+Because the \hyperlink{block nonce}{block nonce} must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+
+Thus we are able to define the \hypertarget{BlockHeaderValidityFunction}{block header validity function $V(H)$}:\label{eq:BlockHeaderValidityFunction}
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \wedge m = \hyperlink{H m}{H_{\mathrm{m}}} \quad \wedge \\
+& & \hyperlink{H d}{H_{\mathrm{d}}} = D(H) \quad \wedge \\
+& & \hyperlink{H g}{H_{\mathrm{g}}} \le \hyperlink{H l}{H_{\mathrm{l}}}  \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} < {P(H)_H}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} > {P(H)_H}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_H}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & \hyperlink{H l}{H_{\mathrm{l}}} \geqslant 125000  \quad \wedge \\
+& & \hyperlink{H s}{H_{\mathrm{s}}} > {P(H)_H}_{\mathrm{s}} \quad \wedge \\
+& & \hyperlink{H i}{H_{\mathrm{i}}} = {P(H)_H}_{\mathrm{i}} +1 \quad \wedge \\
+& & \lVert \hyperlink{H x}{H_{\mathrm{x}}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathrm{s}}, A_{\mathrm{l}}, A_{\mathrm{r}})
+\end{equation}
+
+The tuple contents include \hyperlink{A s}{$A_{\mathrm{s}}$}, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$\hyperlink{A l}{A_{\mathrm{l}}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $\hyperlink{A r}{A_{\mathrm{r}}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
-
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+\begin{align}\hypertarget{g0}{
+g_0 \equiv {} & \sum_{i \in T_{\mathrm{i}}, T_{\mathrm{d}}} \begin{cases} \hyperlink{Gtxdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{Gtxdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+}\end{align}
+
+where $T_{\mathrm{i}},T_{\mathrm{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + \hyperlink{T v}{T_{\mathrm{v}}}
 \end{equation}
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction nonce}{\\
+T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\}
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_H}_{\mathrm{l}} - \hyperlink{ell}{\ell}\hyperlink{R u}{(B_{\mathrm{r}})_{\mathrm{u}}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $ \hyperlink{ell}{\ell}(B_{\mathrm{r}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_{\mathrm{g}} T_{\mathrm{p}} \\%
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
 \end{eqnarray}
@@
 (\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}, 0) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, \hyperlink{T v}{T_{\mathrm{v}}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_{\mathrm{g}} - g_0
+\end{equation}
+and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{A r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $\hyperlink{A r}{A_{\mathrm{r}}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_P[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_P[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_H}_{\mathrm{c}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
-\end{eqnarray}
-
-These are used to help define the transaction receipt, discussed later.
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \hypertarget{A s}{}\\
+\forall i \in A_{\mathrm{s}}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+\hypertarget{Upsilong}{And finally, we specify $\Upsilon^g$, the total gas used in this transaction}\hypertarget{Upsilonl}{ and $\Upsilon^\mathbf{l}$, the logs created by this transaction:}
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - \mathbf{g}' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{A l}{A_{\mathrm{l}}}
+\end{eqnarray}
+
+These are used to help define the \hyperlink{Transaction Receipt}{transaction receipt}.
 
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{nonce}.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, \hyperlink{endow}{v} + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_{\mathrm{b}} &\equiv& \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
 \end{eqnarray}
@@
 \begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
+I_{\mathrm{a}} & \equiv & a \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathrm{d}} & \equiv & () \\
+I_{\mathrm{s}} & \equiv & s \\
+\hyperlink{I v}{I_{\mathrm{v}}} & \equiv & v \\
+I_{\mathrm{b}} & \equiv & \mathbf{i} \\
+I_{\mathrm{e}} & \equiv & e
+\end{eqnarray}
+
+$I_{\mathrm{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+
+Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a \hyperlink{hhalt}{natural halting} state.
 In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
@@
 \begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
+c \equiv \hyperlink{codedeposit}{G_{\mathrm{codedeposit}}} \times |\mathbf{o}|
 \end{equation}
@@
 \quad g' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} - c & \text{otherwise} \\
+0 & \text{if} \quad F \\
+g^{**} - \mathbf{c} & \text{otherwise} \\
 \end{cases} \\
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  \mathbf{g}^{**} < \mathbf{c} \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 Thus any message call received by it during this time causes no code to be executed.
-If the initialisation execution ends with a {\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
-For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
+If the initialisation execution ends with a \hyperlink{selfdestruct}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed.
+For a normal \hyperlink{stop}{\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
+\boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v & \text{otherwise}
 \end{cases}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
-Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
+Just as with contract creation, if the execution \hyperlink{zhalt}{halts in an exceptional fashion} (i.e.
 due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+I_{\mathrm{a}} & \equiv & r \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathrm{d}} & \equiv & \mathbf{d} \\
+I_{\mathrm{s}} & \equiv & s \\
+\hyperlink{I v}{I_{\mathrm{v}}} & \equiv & \tilde{v} \\
+I_{\mathrm{e}} & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathrm{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathrm{b}}), I_{\mathrm{b}})$ at some point prior in order to make the determination of $I_{\mathrm{b}}$ feasible.
 
@@
 
-\begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\begin{itemize}\hypertarget{I a}{
+\item $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.}\hypertarget{I o}{
+\item $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.}\hypertarget{I p}{
+\item $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.}\hypertarget{I d}{
+\item $I_{\mathrm{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.}\hypertarget{I s}{
+\item $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.}\hypertarget{I v}{
+\item $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.}\hypertarget{I b}{
+\item $I_{\mathrm{b}}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.}\hypertarget{I e}{
+\item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).}
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
+
+\hypertarget{emptySequence}{}
+The empty sequence, denoted {{$()$}}, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of \hyperlink{hhalt}{$H$}, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & \mathbf{g} \\
+\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{m}} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{s}} & \equiv & ()
 \end{eqnarray}
@@
 
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(\mathrm{g}, \mathrm{pc}, \mathbf{m}, \mathrm{i}, \mathbf{s})$ which are the gas available, the program counter $\mathrm{pc} \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The memory contents $\boldsymbol{\mu}_{\mathrm{m}}$ are a series of zeroes of size $2^{256}$.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
-\text{\small STOP} & \text{otherwise}
+w \equiv \begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathrm{b}} \rVert \\
+\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+\lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathrm{s}}[0] \notin D(I_{\mathrm{b}}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
@@
 \end{equation}
-
-\subsubsection{Normal Halting}
+\hypertarget{hhalt}{}
+\subsubsection{Normal Halting} \label{subsec:Normal Halting}
 
 The normal halting function $H$ is defined:
-\begin{equation}
+\begin{equation} \label{eq:Normal Halting}
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \hyperlink{RETURN}{\text{\small RETURN}} \\
+() & \text{if} \quad w \in \text{\hyperlink{stop}{\small STOP}}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
 \varnothing & \text{otherwise}
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
+\lVert\boldsymbol{\mu}'_{\mathrm{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_{\mathrm{s}}\rVert): \boldsymbol{\mu}'_{\mathrm{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathrm{s}}[x-\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)  \\
+\hypertarget{JUMP}{} \hypertarget{JUMPI}{} \label{eq:JUMP} \quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
 J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\mu}'_{\mathrm{m}} & \equiv & \boldsymbol{\mu}_{\mathrm{m}} \\
+\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_t & \equiv & B'_t + B_d \\
+B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
 B' & \equiv & P(B_H)
@@
 
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
+As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
 
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert \hyperlink{B U}{B_{\mathrm{u}}} \rVert \leqslant 2 \bigwedge_{U \in \hyperlink{B U}{B_{\mathrm{u}}}} \hyperlink{BlockHeaderValidityFunction}{V(U)} \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathrm{u}})
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathrm{t}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H g}{${B_H}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H g}{{B_H}_{\mathrm{g}}} = \hyperlink{ell}{\ell}\hyperlink{R u}{(\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+We raise the block's beneficiary account by \hyperlink{block reward R b}{$R_{\mathrm{b}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+\hypertarget{Omega}{Formally we define the function $\Omega$, the block-finalisation state transition function (a function that rewards a nominated party):}
 \begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\forall_{U \in B_\mathbf{U}}: \\ \nonumber
- \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\boldsymbol{\sigma}'[\hyperlink{H c}{{B_H}_{\mathrm{c}}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_H}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \hyperlink{B U}{B_{\mathrm{u}}}\rVert}{32})\hyperlink{block reward R b}{R_{\mathrm{b}}} \\
+\forall_{U \in B_{\mathrm{u}}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{1}{8} (U_{\mathrm{i}} - \hyperlink{H i}{{B_H}_{\mathrm{i}}})) \hyperlink{block reward R b}{R_{\mathrm{b}}}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_b = 5 \times 10^{18}
+\text{Let} \quad R_{\mathrm{b}} = 5 \times 10^{18}
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_S(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_H)_H}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
 \end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+}
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\hypertarget{Phi}{
+And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:}
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{h cancel n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$ is the state-transition function}, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+\hyperlink{Transaction Receipt}{As previously detailed}, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_{\mathrm{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_{\mathrm{t}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of \hyperlink{Transaction Receipt}{${B_{\mathrm{r}}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\hyperlink{Upsilong}{\Upsilon^g}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_{\mathrm{t}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
-\end{equation}
-
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
+For $\mathbf{R}[n]_{\mathrm{l}}$, we utilise the \hyperlink{Upsilonl}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{l}} =
+\hyperlink{Upsilonl}{\Upsilon^\mathbf{l}}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_{\mathrm{t}}[n])
+\end{equation}
+
+\hypertarget{Pi}{
+Finally, we define \hyperlink{Pi}{$\Pi$} as the new state given the block reward function \hyperlink{Omega}{$\Omega$} applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(B_{\mathrm{r}})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\hyperlink{R sigma}{\mathbf{R})_{\boldsymbol{\sigma}}})
+\end{equation}}
+
+Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
 
@@
 \begin{equation}
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+m = \hyperlink{H m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
+\end{equation}
+
+Where $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $\hyperlink{H n}{H_{\mathrm{n}}}$ is the nonce of the block header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $\hyperlink{H d}{H_{\mathrm{d}}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being the \hyperlink{block nonce}{block nonce}, a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 The underlying algorithm is called Ethash and is described below.
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
+Ethash is the PoW algorithm for Ethereum \textit{Frontier} and \textit{Homestead}.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
 The general route that the algorithm takes is as follows:
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miner to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
+\hypertarget{rlp}{}W
 \section{Recursive Length Prefix}\label{app:rlp}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} \hyperlink{R b}{R_{\mathrm{b}}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ \hyperlink{R l}{R_{\mathrm{l}}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R b}{
+Formally, we define $R_{\mathrm{b}}$:}
+\begin{eqnarray}
+R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{itemize}
-
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R l}{
+Thus we finish by formally defining $R_{\mathrm{l}}$:
+\begin{eqnarray}
+R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad \mathbf{g} < \mathbf{g}_{\mathrm{r}} \\
+(\boldsymbol\sigma, \mathbf{g} - \mathbf{g}_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathbf{g}_{\mathrm{r}}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+\mathbf{g}_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
-\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|I_{\mathrm{d}}|-1)] &=& I_{\mathrm{d}}\\
+\mathbf{d}[|I_{\mathrm{d}}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+\mathbf{g}_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathrm{d}}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathrm{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{g}_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathrm{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathrm{d}})\\
 \end{eqnarray}
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
+\mathbf{g}_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathrm{d}}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_{\mathrm{d}}
 \end{eqnarray}
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\ e & \equiv & \hyperlink{h T}{h(T)} \hypertarget{ECDSASIGN}{}\\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.}
 
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\hypertarget{h T}{
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{i}}) & \text{if} \; T_{\mathrm{t}} = 0\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, \hyperlink{T v}{T_{\mathrm{v}}}, T_{\mathrm{d}}) & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )}
 \end{eqnarray}
@@
 G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}
-
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
+
+Where the output of $\mathtt{\small ECDSASIGN}$ is also \hyperlink{ECDSASIGN}{defined previously}. Thus:
+\begin{eqnarray}\hypertarget{T w}{}\hypertarget{T r}{}\hypertarget{T s}{}
+T_{\mathrm{w}} = \hyperlink{v}{v}\\
+T_{\mathrm{r}} = \hyperlink{r}{r}\\
+T_{\mathrm{s}} = \hyperlink{s}{s}
+\end{eqnarray}
 We may then define the sender function $S$ of the transaction as:
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big)
 \end{equation}
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{\mathrm{zero}}$ & 0 & Nothing is paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{base}}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
+$G_{\mathrm{low}}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{\mathrm{mid}}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
+$G_{\mathrm{\mathrm{high}}}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{\mathrm{high}}}$}. \\
+$G_{\mathrm{extcode}}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{\mathrm{balance}}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{\mathrm{sload}}$ & 200 & This is paid for an {\small SLOAD} operation. \\
+$G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{sset}}$ & 20000 & This is paid for an \hyperlink{SSTORE}{{\small SSTORE}} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an \hyperlink{SSTORE}{{\small SSTORE}} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
+$R_{\mathrm{selfdestruct}}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
+$G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \hypertarget{codedeposit}{}\\
+$G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
+$G_{\mathrm{call}}$ & 700 & This is paid for a {\small CALL} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{\mathrm{callstipend}}$ & 2300 & This is a stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
+$G_{\mathrm{newaccount}}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{\\
+$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\textit{Homestead transition}}.}\hypertarget{Gtxdatazero}{\\
+$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction.}\hypertarget{Gtxdatanonzero}{ \\
+$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction.}\hypertarget{Gtransaction}{ \\
+$G_{\mathrm{Gtransaction}}$ & 21000 & This is paid for every transaction.} \\
+$G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
+$G_{\mathrm{sha3}}$ & 30 & This is paid for each {\small SHA3} operation. \\
+$G_{\mathrm{sha3word}}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{copy}}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & This is a payment for a {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
+
+%extern u256 const \mathbf{c}_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
-\end{tabular*}
+\end{tabularx}
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{SSTORE}{{\small SSTORE}}} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathrm{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathrm{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathrm{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathrm{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
+G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
+G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
+G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
+G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
+G_{\mathrm{\mathrm{high}}} & \text{if} \quad w \in W_{\mathrm{\mathrm{high}}}\\
+G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
 \end{cases}
@@
 \begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
-\text{\small STOP} & \text{otherwise}
+w \equiv \begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathrm{b}} \rVert\\
+\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{high}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+C_{mem}(a) \equiv G_{\mathrm{memory}} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with \hyperlink{tiny CALL}{$C_\text{\tiny CALL}$}, \hyperlink{C tiny SELFDESTRUCT}{$C_\text{\tiny SELFDESTRUCT}$} and \hyperlink{C tiny SSTORE}{$C_\text{\tiny SSTORE}$} as specified in the appropriate section below.
+We define the following subsets of instructions:
+
+$W_{\mathrm{zero}}$ = \{{\hyperlink{stop}{\small STOP}, \hyperlink{RETURN}{\text{\small RETURN}}\}
+
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{\mathrm{\mathrm{high}}}$ = \{{\small JUMPI}\}
+
+$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 \multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
-\midrule
-0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-\midrule
-0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
-\midrule
-0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{stop}{}\\
+0x00 & {\small STOP} & 0 & 0 & This operation \hyperlink{hhalt}{halts execution}.
+\\
+\midrule
+0x01 & {\small ADD} & 2 & 1 & This is the addition operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+\midrule
+0x02 & {\small MUL} & 2 & 1 & This is the multiplication operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] \times \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+\midrule
+0x03 & {\small SUB} & 2 & 1 & This is the subtraction operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] - \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+\midrule
+0x04 & {\small DIV} & 2 & 1 & This is the integer division operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathrm{s}}[0] \div \boldsymbol{\mu}_{\mathrm{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+\midrule
+0x05 & {\small SDIV} & 2 & 1 & This is the signed integer division operation (truncated). \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathrm{s}}[0] \div \boldsymbol{\mu}_{\mathrm{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathrm{s}}[0] \div \boldsymbol{\mu}_{\mathrm{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 \midrule
-0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
-\midrule
-0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
+0x06 & {\small MOD} & 2 & 1 & This is the modulo remainder operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathrm{s}}[0] \bmod \boldsymbol{\mu}_{\mathrm{s}}[1] & \text{otherwise}\end{cases}$  \\
+\midrule
+0x07 & {\small SMOD} & 2 & 1 & This is the signed modulo remainder operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathrm{s}}[0]) (|\boldsymbol{\mu}_{\mathrm{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathrm{s}}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+0x08 & {\small ADDMOD} & 3 & 1 & This is the modulo addition operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1]) \mod \boldsymbol{\mu}_{\mathrm{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
-0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+0x09 & {\small MULMOD} & 3 & 1 & This is the modulo multiplication operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathrm{s}}[0] \times \boldsymbol{\mu}_{\mathrm{s}}[1]) \mod \boldsymbol{\mu}_{\mathrm{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
-0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
-\midrule
-0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+0x0a & {\small EXP} & 2 & 1 & This is the exponential operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] ^ {\boldsymbol{\mu}_{\mathrm{s}}[1] }$ \\
+\midrule
+0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend the length of a two's complement signed integer. \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathrm{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathrm{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathrm{s}}[x]$} \vspace{5pt} \\
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x10 & {\small LT} & 2 & 1 & This is the less-than comparison. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] < \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x11 & {\small GT} & 2 & 1 & This is the greater-than comparison. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x12 & {\small SLT} & 2 & 1 & This is the signed less-than comparison. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] < \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x13 & {\small SGT} & 2 & 1 & This is the signed greater-than comparison. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
-\midrule
-0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+0x14 & {\small EQ} & 2 & 1 & This is the equality comparison. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] = \boldsymbol{\mu}_{\mathrm{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x15 & {\small ISZERO} & 1 & 1 & This is the logical negation operation, also called the logical complement or the NOT operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x16 & {\small AND} & 2 & 1 & This is the bitwise AND operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}}$ \\
+\midrule
+0x17 & {\small OR} & 2 & 1 & This is the bitwise OR operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}}$ \\
+\midrule
+0x18 & {\small XOR} & 2 & 1 & This is the bitwise XOR operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathrm{s}}[1]_{\mathrm{i}}$ \\
+\midrule
+0x19 & {\small NOT} & 1 & 1 & This is the bitwise NOT operation. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x1a & {\small BYTE} & 2 & 1 & Retrieve a single byte from a word. \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathrm{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathrm{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathrm{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathrm{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& For the Nth byte, we count from the left (i.e. $N=0$ would be the most significant in big endian). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+0x20 & {\small SHA3} & 2 & 1 & Compute a Keccak-256 hash. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
-\midrule
-0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
-\midrule
-0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty \\
-&&&& associated code. \\
-\midrule
-0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get the address of the currently executing account. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{a}}$ \\
+\midrule
+0x31 & {\small BALANCE} & 1 & 1 & Get the balance of the given account. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+\midrule
+0x32 & {\small ORIGIN} & 0 & 1 & Get the address that the execution originated from. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{o}}$ \\
+&&&& This is the sender of the original transaction; it is never an account with non-empty associated code. \\
+\midrule
+0x33 & {\small CALLER} & 0 & 1 & Get the caller address. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
 \midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
-\midrule
-0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
+0x34 & {\small CALLVALUE} & 0 & 1 & Get the deposited value by the instruction/transaction responsible for this execution. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \hyperlink{I v}{I_{\mathrm{v}}}$ \\
+\midrule
+0x35 & {\small CALLDATALOAD} & 1 & 1 & Get the input data of the current environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{d}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathrm{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathrm{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get the size of the input data in the current environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \lVert I_{\mathrm{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy the input data in the current environment to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathrm{d}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] + i < \lVert I_{\mathrm{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathrm{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
-0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
-\midrule
-0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
-&&&& This is gas price specified by the originating transaction.\\
-\midrule
-0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+0x38 & {\small CODESIZE} & 0 & 1 & Get the size of code running in the current environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \lVert I_{\mathrm{b}} \rVert$ \\
+\midrule
+0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in the current environment to memory. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathrm{b}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] + i < \lVert I_{\mathrm{b}} \rVert \\ \text{\hyperlink{stop}{\small STOP}} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathrm{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+\midrule
+0x3a & {\small GASPRICE} & 0 & 1 & Get the price of gas in the current environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& This is the gas price specified by the originating transaction.\\
+\midrule
+0x3b & {\small EXTCODESIZE} & 1 & 1 & Get the size of an account's code. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathrm{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathrm{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\hyperlink{stop}{\small STOP}} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[1], \boldsymbol{\mu}_{\mathrm{s}}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathrm{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
 \multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{blockhash}{BLOCKHASH}\\
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv P(I_{\hyperlink{H p}{H_{\mathrm{p}}}}, \boldsymbol{\mu}_{\mathrm{s}}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.
+0 is left on the stack if the looked for block number is greater than the current block number or more than 256 blocks behind the current block.
+\\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > \hyperlink{H i}{H_{\mathrm{i}}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = \hyperlink{H i}{H_{\mathrm{i}}} \\ P(\hyperlink{H p}{H_{\mathrm{p}}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert that the reason why the header $H$ can be determined is because its hash is the parent hash in the block following it. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv {I_H}_{\mathrm{l}}$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
-\midrule
-0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
-\midrule
-0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
-\midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+0x50 & {\small POP} & 1 & 0 & Remove the top, first item from the stack. \\
+\midrule
+0x51 & {\small MLOAD} & 1 & 1 & Load the first word from memory. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{m}}[\boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathrm{s}}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x52 & {\small MSTORE} & 2 & 0 & Save the first word (which is the first item in the stack) to memory. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathrm{s}}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x53 & {\small MSTORE8} & 2 & 0 & Save the first byte to memory. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathrm{s}}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathrm{s}}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+\midrule
+0x54 & {\small SLOAD} & 1 & 1 & Load the first word from storage. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]]$ \hypertarget{SSTORE}{\\
+\midrule
+0x55 & {\small SSTORE} & 2 & 0 & Save the first word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{s}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1] $ \hypertarget{C tiny SSTORE}{\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
-\end{cases}$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]] = 0 \\
+G_{\mathrm{sreset}} & \text{otherwise}
+\end{cases}$ }\hypertarget{A r}{}\\
+&&&& $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
+R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{s}}[\boldsymbol{\mu}_{\mathrm{s}}[0]] \neq 0 \\
 0 & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ }\\
 \midrule
 0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathrm{s}}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See equation \hyperlink{JUMP}{\ref{eq:JUMP}}. \\
 \midrule
 0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment \\
-&&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathrm{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \hyperlink{JUMPI}{\ref{eq:JUMP}}. \\
+\midrule
+0x58 & {\small PC} & 0 & 1 & Get the value of the program counter \textit{prior} to the increment corresponding to this instruction. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
-\midrule
-0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction \\
-&&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv 32\boldsymbol{\mu}_{\mathrm{i}}$ \\
+\midrule
+0x5a & {\small GAS} & 0 & 1 & Get the amount of available gas, including the corresponding reduction for the cost of this instruction. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{g}}$ \\
 \midrule
 0x5b & {\small JUMPDEST} & 0 & 0 & Mark a valid destination for jumps. \\
-&&&& This operation has no effect on machine state during execution. \\
+&&&& This operation has no effect on the machine state during execution. \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+0x60 & {\small PUSH1} & 0 & 1 & Place a 1 byte item on the stack. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathrm{b}}[x] & \text{if} \quad x < \lVert I_{\mathrm{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
 &&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
-&&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
-\midrule
-0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& The byte is right-aligned (\ie it takes the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
+\midrule
+0x61 & {\small PUSH2} & 0 & 1 & Place a 2-byte item on the stack. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
 \midrule
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
-&&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
+&&&& Similarly, the bytes are right-aligned (\ie they take the lowest significant place and the last, highest address, which is the big-endian interpretation). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+0x80 & {\small DUP1} & 1 & 2 & Duplicate the 1st stack item. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
+\midrule
+0x81 & {\small DUP2} & 2 & 3 & Duplicate the 2nd stack item. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
 \midrule
@@
 \midrule
-0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
+0x8f & {\small DUP16} & 16 & 17 & Duplicate the 16th stack item. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[15]$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
-\midrule
-0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+0x90 & {\small SWAP1} & 2 & 2 & Exchange the 1st and the 2nd stack items. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[1] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
+\midrule
+0x91 & {\small SWAP2} & 3 & 3 & Exchange the 1st and the 3rd stack items. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[2]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[2] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
 \midrule
@@
 \midrule
-0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+0x9f & {\small SWAP16} & 17 & 17 & Exchange the 1st and the 17th stack items. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{s}}[16]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[16] \equiv \boldsymbol{\mu}_{\mathrm{s}}[0]$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
 \multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\hypertarget{A l}{}\\
+\multicolumn{5}{l}{$A'_{\mathrm{l}} \equiv A_{\mathrm{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots (\boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0xa0 & {\small LOG0} & 2 & 0 & Append log record with no topics. \\
+0xa0 & {\small LOG0} & 2 & 0 & Append the log record with no topics. \\
 &&&& $\mathbf{t} \equiv ()$ \\
 \midrule
-0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
+0xa1 & {\small LOG1} & 3 & 0 & Append the log record with one topic. \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 \midrule
@@
 \midrule
-0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
+0xa4 & {\small LOG4} & 6 & 0 & Append the log record with four topics. \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathrm{s}}[2], \boldsymbol{\mu}_{\mathrm{s}}[3], \boldsymbol{\mu}_{\mathrm{s}}[4], \boldsymbol{\mu}_{\mathrm{s}}[5])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
 \toprule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[1] \dots (\boldsymbol{\mu}_{\mathrm{s}}[1] + \boldsymbol{\mu}_{\mathrm{s}}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1) & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_{\mathrm{s}} \equiv \hyperlink{A s}{A_{\mathrm{s}}} \cup A^+_{\mathrm{s}} \quad \wedge \quad A'_{\mathrm{l}} \equiv \hyperlink{A l}{A_{\mathrm{l}}} \cdot A^+_{\mathrm{l}} \quad \wedge \quad A'_{\mathrm{r}} \equiv \hyperlink{A r}{A_{\mathrm{r}}} + A^+_{\mathrm{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hypertarget{zhalt}{exceptional halting}: \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the  \\
+&&&& newly created account. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[1], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[3] \dots (\boldsymbol{\mu}_{\mathrm{s}}[3] + \boldsymbol{\mu}_{\mathrm{s}}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[2], \boldsymbol{\mu}_{\mathrm{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathrm{s}}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[5] \dots (\boldsymbol{\mu}_{\mathrm{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathrm{s}}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hypertarget{zhalt}{exceptional halting} \\
 &&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_{\mathrm{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[3], \boldsymbol{\mu}_{\mathrm{s}}[4]), \boldsymbol{\mu}_{\mathrm{s}}[5], \boldsymbol{\mu}_{\mathrm{s}}[6])$ \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
 &&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathrm{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_{\mathrm{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
-\midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[2], \boldsymbol{\mu}_{\mathrm{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathrm{s}}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
+\midrule
+0xf3 & {\small RETURN} & 2 & 0 & \hyperlink{hhalt}{Halt execution} returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[0] \dots ( \boldsymbol{\mu}_{\mathrm{s}}[0] + \boldsymbol{\mu}_{\mathrm{s}}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
-&&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
-&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \boldsymbol{\mu}_{\mathrm{s}}[1])$ \\
+\end{tabularx}
+
+\begin{tabularx}{\textwidth}{rlrrX}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting the current values for {\it sender} and {\it value}.
+Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.
+The omitted argument is $\boldsymbol{\mu}_{\mathrm{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathrm{s}}[3]$, $\boldsymbol{\mu}_{\mathrm{s}}[4]$, $\boldsymbol{\mu}_{\mathrm{s}}[5]$ and $\boldsymbol{\mu}_{\mathrm{s}}[6]$ in the definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathrm{s}}[2]$, $\boldsymbol{\mu}_{\mathrm{s}}[3]$, $\boldsymbol{\mu}_{\mathrm{s}}[4]$ and $\boldsymbol{\mu}_{\mathrm{s}}[5]$.
+Otherwise it is exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathrm{s}}[0], I_{\mathrm{p}}, 0, \hyperlink{I v}{I_{\mathrm{v}}}, \mathbf{i}, I_{\mathrm{e}} + 1)\end{array} & \text{if} \quad \hyperlink{I v}{I_{\mathrm{v}}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second and ninth parameters to the call $\Theta$.
+This means that the recipient is in fact the same account as at present, simply that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{\\
+\midrule
+0xff & {\small SELFDESTRUCT} & 1 & 0 & \hyperlink{hhalt}{Halt execution} and register account for later deletion. }\\
+&&&& $A'_{\mathrm{s}} \equiv \hyperlink{A s}{A_{\mathrm{s}}} \cup \{ I_{\mathrm{a}} \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+\hyperlink{Rselfdestruct}{R_{\mathrm{selfdestruct}}} & \text{if} \quad I_{\mathrm{a}} \notin \hyperlink{A s}{A_{\mathrm{s}}} \\
 0 & \text{otherwise} 
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
+&&&& $ C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \hyperlink{Gselfdestruct}{G_{\mathrm{selfdestruct}}} + \begin{cases}
+\hyperlink{Gnewaccount}{G_{\mathrm{newaccount}}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathrm{s}}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 \bottomrule
-\end{tabular*}
-
+\end{tabularx}
+
+%This is in the terminology list.
 %\section{Low-level Lisp-like Language}\label{app:lll}
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+E_{\mathrm{epoch}}(\hyperlink{H i}{H_{\mathrm{i}}}) = \left\lfloor\frac{\hyperlink{H i}{H_{\mathrm{i}}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\mathrm{datasetgrowth}}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} for the cache.\hypertarget{dsize}{
+Let $d_{\mathrm{\mathrm{size}}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
+\begin{equation}
+ d_{\mathrm{\mathrm{size}}} = E_{\mathrm{prime}}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
+\end{equation}
+The size of the cache, $c_{\mathrm{size}}$, is calculated using
+\begin{equation}
+ c_{\mathrm{size}} = E_{\mathrm{prime}}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
+E_{\mathrm{prime}}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ s = C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_{\mathrm{i}}})
+\end{equation}
+\begin{equation}
+ C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_{\mathrm{i}}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{\mathrm{epoch}}(\hyperlink{H i}{H_{\mathrm{i}}}) = 0 \quad  \\
+\texttt{KEC}(C_{\mathrm{\mathrm{\mathrm{seedhash}}}}(\hyperlink{H i}{H_{\mathrm{i}}} - \hyperlink{Jepoch}{J_{\mathrm{epoch}}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially fill up $c_{\mathrm{size}}$ bytes of memory, then performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
 
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
+We define the array $\mathbf{c}_{\mathrm{i}}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
+\begin{equation}
+ \mathbf{c}_{\mathrm{i}} = \begin{cases}
 \texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ \mathbf{c'}[i] = \mathbf{c}_{\mathrm{i}} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{\mathrm{size}}}{\hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing \hyperlink{Jcacherounds}{$J_{\mathrm{cacherounds}}$} rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_{\mathrm{n}}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_{\mathrm{n}}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
+\end{equation}
+$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
+\begin{equation}
+ E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}

### 2018-01-08 Paper.tex

--- 2018-01-07 Paper.tex
@@
 
-Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case greek letter, \eg \hyperlink{Upsilon}{$\Upsilon$}, the Ethereum state transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg  \hyperlink{transaction nonce}{$T_{\mathrm{n}}$, denotes the nonce of said transaction}.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction nonce}{transaction nonce}.
+Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
 
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other \hypertarget{ommerheaders}{block headers $\mathbf{U}$} that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, as it is not a well-known word) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language\#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  \mathbf{g}^{**} < \mathbf{c} \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
-\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v & \text{otherwise}
 \end{cases}
@@
 \Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\lVert\boldsymbol{\mu}'_{\mathrm{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathrm{s}}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_{\mathrm{s}}\rVert): \boldsymbol{\mu}'_{\mathrm{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathrm{s}}[x-\Delta]
 \end{eqnarray}
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSASIGN}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\ e & \equiv & \hyperlink{h T}{h(T)} \hypertarget{ECDSASIGN}{}\\
 \mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{\mathrm{zero}}$ & 0 & Nothing is paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{base}}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
+$G_{\mathrm{low}}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{\mathrm{mid}}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
+$G_{\mathrm{\mathrm{high}}}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{\mathrm{high}}}$}. \\
+$G_{\mathrm{extcode}}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{\mathrm{balance}}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{\mathrm{sload}}$ & 200 & This is paid for an {\small SLOAD} operation. \\
+$G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{sset}}$ & 20000 & This is paid for an \hyperlink{SSTORE}{{\small SSTORE}} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an \hyperlink{SSTORE}{{\small SSTORE}} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
+$R_{\mathrm{selfdestruct}}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
+$G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \hypertarget{codedeposit}{}\\
+$G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
+$G_{\mathrm{call}}$ & 700 & This is paid for a {\small CALL} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{\mathrm{callstipend}}$ & 2300 & This is a stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
+$G_{\mathrm{newaccount}}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{\\
+$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\textit{Homestead transition}}.}\hypertarget{Gtxdatazero}{\\
+$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction.}\hypertarget{Gtxdatanonzero}{ \\
+$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction.}\hypertarget{Gtransaction}{ \\
+$G_{\mathrm{Gtransaction}}$ & 21000 & This is paid for every transaction.} \\
+$G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
+$G_{\mathrm{sha3}}$ & 30 & This is paid for each {\small SHA3} operation. \\
+$G_{\mathrm{sha3word}}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{copy}}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & This is a payment for a {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
+
+%extern u256 const \mathbf{c}_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
-\end{tabular*}
+\end{tabularx}
 
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathrm{m}}[ \boldsymbol{\mu}_{\mathrm{s}}[1] \dots (\boldsymbol{\mu}_{\mathrm{s}}[1] + \boldsymbol{\mu}_{\mathrm{s}}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathrm{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1) & \text{if} \quad \boldsymbol{\mu}_{\mathrm{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_{\mathrm{s}} \equiv \hyperlink{A s}{A_{\mathrm{s}}} \cup A^+_{\mathrm{s}} \quad \wedge \quad A'_{\mathrm{l}} \equiv \hyperlink{A l}{A_{\mathrm{l}}} \cdot A^+_{\mathrm{l}} \quad \wedge \quad A'_{\mathrm{r}} \equiv \hyperlink{A r}{A_{\mathrm{r}}} + A^+_{\mathrm{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{s}}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hypertarget{zhalt}{exceptional halting}: \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathrm{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (the balance of the caller is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the  \\
+&&&& newly created account. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathrm{s}}[1], \boldsymbol{\mu}_{\mathrm{s}}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+E_{\mathrm{epoch}}(\hyperlink{H i}{H_{\mathrm{i}}}) = \left\lfloor\frac{\hyperlink{H i}{H_{\mathrm{i}}}}{\hyperlink{Jepoch}{J_{\mathrm{epoch}}}}\right\rfloor
+\end{equation}
+The size of the dataset growth is \hyperlink{Jdatasetgrowth}{$J_{\mathrm{datasetgrowth}}$} bytes, and the size of the cache growth is \hyperlink{Jcachegrowth}{$J_{\mathrm{cachegrowth}}$} bytes, where the growth of each occurs every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$}, for the dataset, and \hyperlink{Jhashbytes}{$J_{\mathrm{hashbytes}}$} for the cache.\hypertarget{dsize}{
+Let $d_{\mathrm{\mathrm{size}}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using:}
+\begin{equation}
+ d_{\mathrm{\mathrm{size}}} = E_{\mathrm{prime}}(\hyperlink{Jdatasetinit}{J_{\mathrm{datasetinit}}} + \hyperlink{Jdatasetgrowth}{J_{\mathrm{datasetgrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jmixbytes}{J_{\mathrm{mixbytes}}}, J_{\mathrm{mixbytes}})
+\end{equation}
+The size of the cache, $c_{\mathrm{size}}$, is calculated using
+\begin{equation}
+ c_{\mathrm{size}} = E_{\mathrm{prime}}(\hyperlink{Jcacheinit}{J_{\mathrm{cacheinit}}} + \hyperlink{Jcachegrowth}{J_{\mathrm{cachegrowth}}} \cdot E_{\mathrm{epoch}} - \hyperlink{Jhashbytes}{J_{\mathrm{hashbytes}}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
+
+E_{\mathrm{prime}}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $c$, which is an array of bytes. It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $s \in \mathbb{B}_{32}$.
+
 \subsubsection{Seed hash}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
-
-If the output of this algorithm is below the desired target, then the nonce is valid.
+Essentially, we maintain a "mix" \hyperlink{Jmixbytes}{$J_{\mathrm{mixbytes}}$} bytes wide, and repeatedly sequentially fetch $J_{\mathrm{mixbytes}}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{\mathrm{mixbytes}}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+
+If the output of this algorithm is below the desired target, then the \hyperlink{block nonce}{nonce} is valid.
 Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.

### 2018-01-13 Paper.tex

--- 2018-01-07 Paper.tex
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point for which $r$ is the x value.
+This value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 

### 2018-01-16 Paper.tex
Update Paper.tex

@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point for which $r$ is the x value.
-This value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
 

### 2018-01-17 Paper.tex

--- 2018-01-13 Paper.tex
@@
 \Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point for which $r$ is the x value.
-This value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 

### 2018-01-18 Paper.tex
\section{List of mathematical symbols} \bigvee
\cite{Davey2002_zbMATH01748069}
\textit{Homestead}

@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the \textit{Homestead} transition.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\

### 2018-01-19 Paper.tex
hyperref; backslash before hash

@@
 \usepackage{tikz}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:

### 2018-01-19 Paper.tex
Options for hyperref, and \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

--- 2018-01-18 Paper.tex
@@
 \usepackage{tikz}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+%,pagebackref=true
+
+% I will insert other table packages in here after this is merged.
+
+%This should be the last package before \input{Version.tex}
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
+% "hyperref loads the url package internally.
+Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
+This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
+% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
+Try loading the url package earlier than the package that requires it.
+If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
 

### 2018-01-19 Paper.tex
Formatting packages

--- 2018-01-19 Paper.tex
@@
 %\usepackage{tweaklist}
+\usepackage{url}
 \usepackage{cancel}
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-% I will insert other table packages in here after this is merged.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+% Put before hyperref
 

### 2018-01-19 Paper.tex
Formatting fixes for URLs breaking over lines
Edit details for Gavin Wood author header
Footnote for Gavin's email
May as well comment out \lightpink if it's unused for performance reasons
not the least being Bitcoin

@@
 
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-% Put before hyperref
-
 \input{Version.tex}

### 2018-01-19 Paper.tex
The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead} difficulty parameter

--- 2018-01-19 Paper.tex
@@
 Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
-Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
+Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
 
@@
 
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+$\expdiffsymb$ results in the difficulty increasing every 10000 blocks.
+This is necessary to stabilise the time difference between blocks as  computation power becomes more efficient, with Moore's Law, although the rate of increase in computational efficiency has slowed down and will probably continue to slow down (\cite{THOMPSON200620}.
+
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
@@
 \Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account. \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\

### 2018-01-19 Paper.tex

--- 2018-01-19 Paper.tex
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is the `recovery id', a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The recovery identifier is a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The chain identifier $\mathtt{\tiny chain\_id}$ is a constant natural number.
 

### 2018-01-19 Paper.tex

--- 2018-01-19 Paper.tex
@@
 I_w & \equiv & w \\
+\mathbf{t} & \equiv & \{s, r\} \\
+\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c

### 2018-01-19 Paper.tex

@@
 \end{cases}
+\end{eqnarray}
+
+\subsection{zkSNARK Related Precompiled Contracts}
+
+We choose two numbers, both of which are prime.
+\begin{eqnarray}
+p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
+q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
+\end{eqnarray}
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+
+We define a set~$C_1$ with
+\begin{equation}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-addition}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ X_1\neq X_2\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\begin{equation}\label{eq:ec-scalar-multiplication}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+\end{equation}
+for a natural number $n$ and a point $P$ in $C_1$.
+
+We define $P_1$ to be a point $(1,2)$ on $C_1$.
+Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
+$G_1$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+$\log_{P_1}(P)$ is at most $q-1$.
+
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+$(C_2,+)$ is also known to be a group.
+We define $P_2$ in $C_2$ with
+\begin{eqnarray}
+P_2&\equiv&
+(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
+\end{eqnarray}
+We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
+$G_2$ is known to be a cyclic group of order $q$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
+\varnothing&\text{otherwise}
+\end{cases}
+\end{equation}
+
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
+g_1&\text{if}\ g_1\in G_1\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_1&\equiv&\begin{cases}
+(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
+g_2&\text{if}\ g_2\in G_2\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+g_2&\equiv&\begin{cases}
+((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
+\varnothing&\text{otherwise}
+\end{cases}\\
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
+\end{eqnarray}
+
+We define a precompiled contract for zkSNARK verification.
+\begin{eqnarray}
+\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+g_r&=& 60000k + 40000 \\
+\mathbf o[0..31]&\equiv&\begin{cases}
+0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
+0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
+\end{cases}\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+\vdots\\
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+\end{eqnarray}
+
+We define a precompiled contract for addition on $G_1$.
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_r &=& 500\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+0&\text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_r &=& 40000\\
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}

### 2018-01-19 Paper.tex

@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
-\Lambda_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\quad s' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
+\end{cases}
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\ \nonumber
+s' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+1 & \text{otherwise}
+\end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}

### 2018-01-19 Paper.tex

@@
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Founder, Ethereum \& Parity (formerly Ethcore)\\
+    i@gavwood.com\footnote{Please talk to maintainers on Github such as @pirapira and @nicksavers, only contact Gavin with their permission.}
 }
@@
 \end{equation*}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
+\begin{align}
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_i &\equiv \max(H_i - 3000000, 0)
+\end{align}
 

### 2018-01-19 Paper.tex
Byzantium revision

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity (formerly Ethcore)\\
-    i@gavwood.com\footnote{Please talk to maintainers on Github such as @pirapira and @nicksavers, only contact Gavin with their permission.}
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }

### 2018-01-20 Paper.tex
Remove lightpink lines

--- 2018-01-19 Paper.tex
@@
 \pagecolor{lightyellow}
-%\pagecolor{lightpink}
 
@@
 \quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
-\end{equation}
-
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
 \end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
 \end{equation}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\quad s' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases}
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
 \end{equation}
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-s' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
 \end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger than 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_r&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
-\end{eqnarray}
-
 \section{Signing Transactions}\label{app:signing}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\

### 2018-01-20 Paper.tex
Update Paper.tex

@@
 \pagecolor{lightyellow}
+%\pagecolor{lightpink}
 
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.

### 2018-01-20 Paper.tex
Update \expdiffsymb explanation

--- 2018-01-20 Paper.tex
@@
 
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+$\expdiffsymb$ results in the difficulty increasing every 10000 blocks, which puts upward pressure on the time between blocks, causing it to slowly accelerate, and putting time pressure on transitioning to proof-of-stake.
+
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:

### 2018-01-20 Paper.tex
\expdiffsymb, cite EIP-649 and EIP-2
Add a missing closing brace

@@
 Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead}} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
 $\expdiffsymb$ results in the difficulty increasing every 10000 blocks, which puts upward pressure on the time between blocks, causing it to slowly accelerate, and putting time pressure on transitioning to proof-of-stake.
+This effect was explained in EIP-649\cite{EIP-649} and implemented earlier in EIP-2\cite{EIP-2}.
 

### 2018-01-20 Paper.tex

--- 2018-01-20 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 \quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation*}
+x \equiv \begin{cases}
+1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+2 & \text{otherwise}
+\end{cases}
+\end{equation*}
+\begin{align}
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_i &\equiv \max(H_i - 3000000, 0)
+\end{align}
+
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead}} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
+$\expdiffsymb$ results in the difficulty increasing every 10000 blocks, which puts upward pressure on the time between blocks, causing it to slowly accelerate, and putting time pressure on transitioning to proof-of-stake.
+This effect was explained in EIP-649\cite{EIP-649} and implemented earlier in EIP-2\cite{EIP-2}.
 
@@
 g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+s' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
+1 & \text{otherwise}
 \end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
 \end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger than 1024.
+where
+\begin{equation}
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+\end{array}
+\end{equation}
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.

### 2018-01-20 Paper.tex
Update explanations of \homesteadmod and \expdiffsymb with citations for EIPs

@@
 Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead}} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-$\expdiffsymb$ results in the difficulty increasing every 10000 blocks, which puts upward pressure on the time between blocks, causing it to slowly accelerate, and putting time pressure on transitioning to proof-of-stake.
-This effect was explained in EIP-649\cite{EIP-649} and implemented earlier in EIP-2\cite{EIP-2}.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, $\expdiffsymb$ results in the difficulty increasing every 10000 blocks, which puts upward pressure on the time between blocks, causing it to slowly accelerate, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
 

### 2018-01-22 Paper.tex

--- 2018-01-20 Paper.tex
@@
 \end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{\href{https://github.com/ethereum/EIPs/blob/master/EIPS/eip-606.md}{Homestead}} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below.
-$\expdiffsymb$ results in the difficulty increasing every 10000 blocks, which puts upward pressure on the time between blocks, causing it to slowly accelerate, and putting time pressure on transitioning to proof-of-stake.
-This effect was explained in EIP-649\cite{EIP-649} and implemented earlier in EIP-2\cite{EIP-2}.
 

### 2018-01-22 Paper.tex

@@
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.

### 2018-01-22 Paper.tex
Clean up formatting of REVERT opcode description.

@@
 \end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \quad s' &\equiv \begin{cases}
@@
 1 & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}

### 2018-01-22 Paper.tex
Fix formatting of s' definition.
Add a line break for maths
"where" following an equation does not introduce a new paragprah

--- 2018-01-22 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
+Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 %,pagebackref=true
 
-% I will insert other table packages in here after this is merged.
+\usepackage{tabu} %requires array.
 
@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
+
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\usepackage{seqsplit}
+
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+
+\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
 
 \input{Version.tex}
+% Default rendering options
+\definecolor{pagecolor}{rgb}{1,0.98,0.9}
+\def\YellowPaperVersionNumber{unknown revision}
+\IfFileExists{Options.tex}{\input{Options.tex}}
 
@@
 
-\definecolor{lightyellow}{rgb}{1,0.98,0.9}
-
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Founder, Ethereum \& Parity\\
+    gavin@parity.io
 }
@@
 
-\pagecolor{lightyellow}
+\pagecolor{pagecolor}
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
-\end{equation}
-
+\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
+\end{equation}
 where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
+In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful than any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
 
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 This document describes one version of the protocol.
-In order to follow back the history of a path, one might need to reference multiple versions of this document.
+In order to follow back the history of a path, one must reference multiple versions of this document.
 
@@
 
-Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg $\Upsilon$, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the Ethereum state transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
 Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \ref{ch:block}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
+This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
+
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-The account state comprises the following four fields:
+The account state, $\boldsymbol{\sigma}[a]$, comprises the following four fields:
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
 
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
-\begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
-\begin{equation}
-L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
-\end{equation}
-
+Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
+\begin{equation}
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
+\begin{equation}
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\end{equation}
 where:
 \begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
-\end{equation}
-
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
+\end{equation}
+
+It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
-\begin{equation}
-L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_{S}$:
+\begin{equation}
+L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \end{equation}
-where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
 \end{equation}
@@
 \begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+L_{T}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\subsection{The Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
+\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
+\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
+\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
+The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
+\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
 \end{description}
-
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}
-
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
+\end{equation}
+
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
+Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
+\end{equation}
+
+\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
 \end{equation}
@@
 
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
-\end{equation}
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\begin{equation}
+R_{\mathrm{z}} \in \mathbb{N}
+\end{equation}
+
+\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+
+The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
-\end{equation}
-
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+\end{equation}
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
 \end{equation}
-
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
+\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
 \end{array}
 \end{equation}
-
 where
 \begin{equation}
-\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 
-The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
 \end{equation}
-where:
+where
 \begin{equation}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
 \begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+y \equiv \begin{cases}
+1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
 2 & \text{otherwise}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
 \end{align}
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} \geqslant 5000
+\end{eqnarray}
+
+\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
+\begin{equation}
+H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
 \end{equation}
@@
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
-
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Above, $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
 
@@
 
-Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 5000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
+\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
+\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
+\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0)
+A^0 \equiv (\varnothing,(), \varnothing, 0)
 \end{equation}
@@
 \subsection{Execution}
-We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
+g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
 \end{align}
-
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\end{cases}
+\end{equation}
 where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
 \begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_{\mathrm{g}} - g_0
+\end{equation}
+and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 
-After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
-\begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
+\begin{equation}
+	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
+\end{equation}
+
+Then the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
+\begin{equation}
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_{H}}_{\mathrm{c}}
 \end{eqnarray}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
-\end{eqnarray}
-
-These are used to help define the transaction receipt, discussed later.
-
-\section{Contract Creation} \label{ch:create}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
+
+\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
+
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
+\end{equation}
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s] &=& \begin{cases}
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
-\end{eqnarray}
-
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\end{eqnarray}
 where $v'$ is the account's pre-existing value, in the event it was previously in existence:
@@
 \end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
-\begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
+\begin{eqnarray}
+I_{\mathrm{a}} & \equiv & a \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathbf{d}} & \equiv & () \\
+I_{\mathrm{s}} & \equiv & s \\
+\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
+I_{\mathbf{b}} & \equiv & \mathbf{i} \\
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w
+\end{eqnarray}
+
+$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, s')$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
-\quad s' &\equiv \begin{cases}
+\quad z &\equiv \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
@@
 \subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code.
+Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address should return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
 Thus any message call received by it during this time causes no code to be executed.
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\mathbf{a}_1 \equiv \left(\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}}\right)
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
+&\quad \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
 \end{cases} \\ \nonumber
-s' & \equiv & \begin{cases}
+z & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
@@
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
+I_{\mathrm{a}} & \equiv & r \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
+I_{\mathbf{d}} & \equiv & \mathbf{d} \\
+I_{\mathrm{s}} & \equiv & s \\
+I_{\mathrm{v}} & \equiv & \tilde{v} \\
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w \\
+\mathbf{t} & \equiv & \{s, r\} \\
+\end{eqnarray}
+\nopagebreak[1]where
+\begin{equation}
+\Xi \equiv \begin{cases}
 \Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
@@
 \Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-I_w & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
+\end{equation}
+
+\begin{equation}
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
 
@@
 The EVM is a simple stack-based architecture.
-The word size of the machine (and thus size of stack item) is 256-bit.
+The word size of the machine (and thus size of stack items) is 256-bit.
 This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations.
@@
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
-
-The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
+
+\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
+\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
 \end{eqnarray}
@@
 \end{equation}
-
 where
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
 
@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
+\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
+ (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
 \end{array}
@@
 \begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+W(w, \boldsymbol{\mu}) \equiv \\ 
+\begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
-\end{equation}
-
-where:
-\begin{equation}
-D_J(\mathbf{c}, i) \equiv \begin{cases}
+D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
+\end{equation}
+where
+\begin{equation}
+D_{J}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
+\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
+\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\end{cases}
+\end{equation}
 where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
@@
 N(i, w) \equiv \begin{cases}
-i + w - \text{\small PUSH1} + 2 & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
+i + w - \text{\small PUSH1} + 2 & \\
+\quad\quad\text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
 i + 1 & \text{otherwise} \end{cases}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
-\varnothing & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
+() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
+\varnothing \quad\quad\ \text{otherwise}&
+\end{cases}
+\end{equation}
+
+The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
+Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
 
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
+\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
 \end{eqnarray}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
-\end{equation}
-
+\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
+\end{equation}
 where $k$ denotes the ``is-kin'' property:
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
+\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
+\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
 \end{eqnarray}
@@
 
-We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_b = 3 \times 10^{18}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
 \end{equation}
@@
 
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+
+\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
-\end{eqnarray}
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+\end{eqnarray}
+
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
-\end{equation}
-
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_{\mathbf{l}} =
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\end{equation}
+
+We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{z}} =
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\end{equation}
+
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
+\end{equation}
+
+Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
 
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the \hyperlink{block_nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
@@
 \begin{equation}
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
+\end{equation}
+where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
+where $\mathbb{O}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
 \begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\mathtt{\tiny RLP}(i : i \in \mathbb{N}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
 \end{equation}
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
-
-For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
+
+We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
+
 \begin{eqnarray}
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+g_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_{\mathbf{d}}
 \end{eqnarray}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
 f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+x^2 & \text{if}\ x \le 64 \\[0.5em]
+\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
+\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
 \end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
+\ell'_{E} &=& \begin{cases}
+0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
+8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
+8(\ell_{E} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{P}_{8\ell_{M}} \\
+\ell_{B} &\equiv& i[0..31] \\
+\ell_{E} &\equiv& i[32..63] \\
+\ell_{M} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{B})] \\
+E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
+M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
 i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+I_{\mathbf{d}}[x] &\text{if}\ x < |I_{\mathbf{d}}| \\
 0 &\text{otherwise}
@@
 \end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
 
@@
 \begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
+C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ on $C_1$ for distinct elements $(X_1, Y_1), (X_2, Y_2)$ with
 \begin{eqnarray}\label{eq:ec-addition}
@@
 \end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+\nonumber \lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
+\nonumber X&\equiv&\lambda^2-X_1-X_2\\
+\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
+\end{eqnarray}
+
+In the case where $(X_1, Y_1) = (X_2, Y_2)$, we define $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-doubling}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ Y_1\neq 0\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+\nonumber \lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
+\nonumber X&\equiv&\lambda^2-2X_1\\
+\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define scalar multiplication $\cdot$ with
 \begin{equation}\label{eq:ec-scalar-multiplication}
@@
 
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+Let $F_{p^2}$ be a field $F_{p}[i]/(i^2+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3(i+9)^{-1}\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.
@@
 We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
+$G_2$ is known to be the only cyclic group of order $q$ on $C_2$.
 For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
+Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$.
+It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
+This bilinear map is a type three pairing.
+There are several such bilinear maps, it does not matter which is chosen to be $e$.
+Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
+\begin{eqnarray} \label{eq:pairing-check}
+\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
+\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
+\end{eqnarray}
+Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).
+
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
+\begin{equation}
+\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
+\mathbf{x}&\text{if}\ \mathbf{x}<p\\
 \varnothing&\text{otherwise}
@@
 
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
+A 64 byte data $\mathbf{x}\in\mathbf{B}_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf{x})&\equiv&\begin{cases}
 g_1&\text{if}\ g_1\in G_1\\
@@
 \end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
+x&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
+y&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf{x}\in\mathbf{B}_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf{x})&\equiv&\begin{cases}
 g_2&\text{if}\ g_2\in G_2\\
@@
 \end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
+\end{eqnarray}
+
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.
+
 \begin{eqnarray}
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_r&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
+F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
+k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
+g_{\mathrm{r}}&=& 60000k + 40000 \\
+\mathbf{o}[0..31]&\equiv&\begin{cases}
 0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
@@
 v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf{d}}[64..191])\\\nonumber
 \vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_k&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
 \end{eqnarray}
@@
 We define a precompiled contract for addition on $G_1$.
+
 \begin{eqnarray}
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_{\mathrm{r}} &=& 500\\
+\mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
+y&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[64..127]\right)\\
+\label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
+I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
 0&\text{otherwise}
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf{d}}$ is defined in (\ref{eq:complemented_input}).
 \begin{eqnarray}
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
+g_{\mathrm{r}} &=& 40000\\
+\mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf{d}}[0..31]\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[32..95]\right)
 \end{eqnarray}
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
-
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
+This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
+
+It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+
+We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
+These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
+It is assumed that \hypertarget{v}{}$v$ is the `recovery identifier'.
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
@@
 
-We declare that a signature is invalid unless all the following conditions are true:
+\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true\footnote{A signature of a transaction can be valid not only with a recovery identifier but with some other numbers.
+See \hyperlink{T__w}{how the component $T_w$ of a transaction is interpreted.}}:
 \begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
+0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+\hyperlink{v}{v} &\in \{27,28\}
 \end{align}
@@
 
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+\end{equation}
+
+\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+L_{S}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
-\end{eqnarray}
+h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
+\begin{eqnarray}
+G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+\end{eqnarray}
+
+\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
+\begin{eqnarray}
+\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
+\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
+\end{eqnarray}
+$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.
+In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
 v_0 &\equiv& \begin{cases}
-T_w &\text{if}\ T_w\in\{27, 28\} \\
-28 - (T_w \bmod 2) &\text{otherwise}
+T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
+28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
 \end{cases}
@@
 
-\begin{tabular*}{\columnwidth}[h]{lrl}
+\begin{tabu}{l r l}
 \toprule
@@
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+&&is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
+&&non-zero.\\
+\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
 $G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
+\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
-
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
+&&contract. \\
+
+%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
-\end{tabular*}
+\end{tabu}
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
+&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
+&\quad\text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 \end{equation}
-
 where:
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
 \end{equation}
@@
 \subsection{Instruction Set}
+\label{subsec:instruction-set}
 
@@
 
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\begin{tabu}{r l r r l} \savetabu{opcodes}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
+&&&& modulo. \\
 \midrule
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
+&&&& modulo. \\
 \midrule
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
-\midrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
+\midrule
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
+&&&& in big endian). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty \\
-&&&& associated code. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
+&&&& This is the sender of original transaction; it is never an account with\\
+&&&& non-empty associated code. \\
 \midrule
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
-&&&& This is the address of the account that is directly responsible for this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
+&&&& This is the address of the account that is directly responsible for\\
+&&&& this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
+&&&& this execution. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
+&&&& This pertains to the input data passed with the message call\\
+&&&& instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
+&&&& environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
+&&&& This pertains to the input data passed with the message call\\
+&&&& instruction or transaction. \\
 \midrule
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& This pertains to the input data passed with the message call instruction\\
+&&&& or transaction. \\
 \midrule
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+\midrule
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
+&&&& environment. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
 \midrule
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than\\
+&&&& the current block number or more than 256 blocks behind the current block.\\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
+&&&& $h$ is zero.\\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[15]$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[16]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[16] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4], \boldsymbol{\mu}_{\mathbf{s}}[5])$ \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
+&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\ 
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
+&&&& address of the newly created account. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
-\midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
+&&&& fact the same account as at present, simply that the code is overwritten.\\
+\midrule
+\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
+&&&& persisting the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
+&&&& omitted argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the\\
+&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and\\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced \\
+&&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \midrule
@@
 \midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+\end{cases}$\\ \\
+&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad n \\
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
-\end{tabular*}
-
-%\section{Low-level Lisp-like Language}\label{app:lll}
-%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-
-\section{Genesis Block}\label{app:genesis}
+\end{tabu}
+
+\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
 
@@
 \end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
 
@@
 
-\begin{tabular*}{\columnwidth}[h]{lrl}
+\begin{tabu*}{lcl}
 \toprule
@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
+\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
-\end{tabular*}
+\end{tabu*}
 
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
+ c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
+x & \text{if} \quad x / y \in \mathbb{N} \\
+E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
+\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \end{equation}
-Where a single round modifies each subset of the cache as follows:
-\begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
+where a single round modifies each subset of the cache as follows:
+\begin{equation}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
@@
 
-\subsection{Deletion of an Account Dispite Out-of-gas}
-
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\subsection{Deletion of an Account Despite Out-of-gas}
+
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{c c X[j]} \savetabu{col3}
+\toprule
+Symbol & Latex Command & Description \\
+\midrule
+\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
+Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+\bottomrule
+\end{tabu*}
 \end{document}

### 2018-01-22 Paper.tex
"where" following an equation does not introduce a new paragprah

--- 2018-01-22 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-% I will insert other table packages in here after this is merged.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-
-\input{Version.tex}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 \pagecolor{lightyellow}
+%\pagecolor{lightpink}
 
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 \end{equation}
-
 where $\Upsilon$ is the Ethereum state transition function.
@@
 \end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 \end{equation}
-
 where:
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
@@
 
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
-where:
-\begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
+\end{equation}
+where
+\begin{equation}
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
 
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Above, $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
-
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
+$G$ is defined in Appendix \ref{app:fees}.
+
+%todo Explain g_d reason?
 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\end{cases}
+\end{equation}
 where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
 \end{eqnarray}
@@
 
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
 \section{Contract Creation} \label{ch:create}
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
@@
 \end{equation}
-
 where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
-\end{eqnarray}
-
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\end{eqnarray}
 where $v'$ is the account's pre-existing value, in the event it was previously in existence:
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**} < c \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
 \begin{eqnarray}
@@
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+I_e & \equiv & e
+\end{eqnarray}
+where $c$ is the code-deposit cost:
 \begin{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, s')$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad s' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
-\end{equation}
-Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \mathbf{d}, e)
+\end{equation}
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\boldsymbol{\sigma}_1'[s]_b \equiv \boldsymbol{\sigma}_1[s]_b - v
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-s' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
+I_v & \equiv & v \\
 I_e & \equiv & e \\
-I_w & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
 \text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
@@
 
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will define it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
+\end{cases}
+\end{equation}
 where
@@
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 \end{equation}
-
-where:
+where
 \begin{equation}
@@
 \end{equation}
-
 where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
 \varnothing & \text{otherwise}
@@
 \end{equation}
-
 where $k$ denotes the ``is-kin'' property:
@@
 \begin{eqnarray}
-\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 
-We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_b = 3 \times 10^{18}
+We define the block reward as 5 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 5 \times 10^{18}
 \end{equation}
@@
 \mathbf{R}[n]_\mathbf{l} = 
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}
@@
 \end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 
-Many thanks to Aeron Buchanan for authoring the \textit{Homestead} revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
 
@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+where $\mathbb{Y}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
@@
 
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
-\end{align}
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
 where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+\begin{eqnarray}
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
-The other operates on nine elements:
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
 \begin{eqnarray}
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
 \end{cases} \\
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_w &\text{if}\ T_w\in\{27, 28\} \\
-28 - (T_w \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
 \begin{equation}
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
 $G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 \end{equation}
-
 where:
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 \bottomrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account. \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& same account as at present, simply that the code is overridden altered.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
-\bottomrule
+\end{cases}$ \\\bottomrule
 \end{tabular*}
@@
 \end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order the generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$. 
 \subsubsection{Seed hash}
@@
 \end{equation}
-Where a single round modifies each subset of the cache as follows:
+where a single round modifies each subset of the cache as follows:
 \begin{equation}
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Dispite Out-of-gas}
-
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2018-01-22 Paper.tex
Rephrase a sentence on the $\expdiffsymb$ and add an explanation of the difficulty bomb delay

--- 2018-01-22 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-% I will insert other table packages in here after this is merged.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
 
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \end{align}
+
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up due to network congestion.
 

### 2018-01-22 Paper.tex

@@
 
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up due to network congestion.
-
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:

### 2018-01-22 Paper.tex
Change Ethcore to Parity.

@@
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Founder, Ethereum \& Parity\\
+    gavin@parity.io
 }
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
+&&&& For the gas calculation, we need to define the memory usage: \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\

### 2018-01-22 Paper.tex

@@
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 
-\subsection{Deletion of an Account Dispite Out-of-gas}
+\subsection{Deletion of an Account Despite Out-of-gas}
 
@@
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabularx}{\textwidth}{lcX}
+\toprule
+Symbol & Latex Command & Description \\
+\midrule
+\hypertarget{bigvee}{
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).}\\
+\bottomrule
+\end{tabularx}
+
 \end{document}

### 2018-01-23 Paper.tex
... refer to Appendix \ref{ch:Terminology}.
Unix's [sic] time

--- 2018-01-22 Paper.tex
@@
 \item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix time() at this block's inception; formally $H_s$.
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.

### 2018-01-23 Paper.tex
Remove seemingly out-of-date to-do.
*that* are touched and empty

@@
 \item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix time() at this block's inception; formally $H_s$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or that are touched and empty:
 \begin{eqnarray}

### 2018-01-23 Paper.tex
hypertargets lines 100–199

--- 2018-01-23 Paper.tex
@@
 A valid state transition is one which comes about through a transaction.
-Formally:
+Formally:\hypertarget{Upsilon_state_transition}{Upsilon_state_transition}
 \begin{equation}
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \hypertarget{ell}{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-\begin{description}
+\begin{description}\hypertarget{account_nonce}{account_nonce}
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.

### 2018-01-23 Paper.tex
hypertargets lines 200–299

--- 2018-01-23 Paper.tex
@@
 \end{equation}
-where $v$ is the account validity function:
+where $v$ is the account validity function:\hypertarget{v x}{}
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
-
-\begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+Both types specify a number of common fields:\hypertarget{T n}{}
+
+\begin{description}\hypertarget{T n}{}
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{T p}{}
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{T g}{}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{T t}{}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{T v}{}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-23 Paper.tex
Update labels and targets

@@
 \end{equation}
-where $v$ is the account validity function:\hypertarget{v x}{}
+where $v$ is the account validity function:\hypertarget{account_validity_function_v__x}{account_validity_function_v__x}
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:\hypertarget{T n}{}
-
-\begin{description}\hypertarget{T n}{}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{T p}{}
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{T g}{}
+Both types specify a number of common fields:
+
+\begin{description}\hypertarget{tx_nonce}{tx_nonce_T__n}
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx_gas_price_T__p}
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{tx_gas_limit_T__g}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{T t}{}
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{T v}{}
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx_to_address_T__t}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx_value_T__v}{tx_value_T__v}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-23 Paper.tex
hypertargets lines 300–349

--- 2018-01-23 Paper.tex
@@
 A valid state transition is one which comes about through a transaction.
-Formally:\hypertarget{Upsilon_state_transition}{Upsilon_state_transition}
+Formally:
 \begin{equation}
@@
 
-I define a number of useful functions throughout. \hypertarget{ell}{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-\begin{description}\hypertarget{account_nonce}{account_nonce}
+\begin{description}
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{ommerheaders} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-\begin{description}
+\begin{description}\hypertarget{H p}{parentHash}
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{H c}{beneficiary}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts-Root-def}{receipts-Root-def}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs-Bloom-def}{logs-Bloom-def}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{H_i}{blocknumber}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{H_g}{gasUsed}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{H_s}{timestamp}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{H_x}{extraData}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{H_m}{mixHash}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block nonce}{block_nonce}\hypertarget{H_n}{block_nonce}
 \item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
-\end{description}
+\end{description}\hypertarget{B U}{}
 
@@
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
+\end{equation}\hypertarget{Transaction Receipt}{Transaction_Receipt}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{R}{transaction_receipt}\hypertarget{R u}{tx_receipt_gas_used}
 

### 2018-01-23 Paper.tex
Update labels=targets

@@
 
-\begin{description}\hypertarget{H p}{parentHash}
+\begin{description}\hypertarget{parent_Hash_H__p_def_words}{parent_Hash_H__p_def_words}
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{H c}{beneficiary}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{beneficiary_H__c}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts-Root-def}{receipts-Root-def}
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs-Bloom-def}{logs-Bloom-def}
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{receipts_Root_def_words}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{logs_Bloom_def_words}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{H_i}{blocknumber}
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{block_number_H_i}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{H_g}{gasUsed}
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{H_s}{timestamp}
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{H_x}{extraData}
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{H_m}{mixHash}
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block nonce}{block_nonce}\hypertarget{H_n}{block_nonce}
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{block_gas_used_H__g}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{block_timestamp_H__s}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{block_extraData_H__x}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{mixHash_H__m}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{block_nonce_H__n}
 \item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
-\end{description}\hypertarget{B U}{}
+\end{description}\hypertarget{ommer_block_headers_B__U}{ommer_block_headers_B__U}\hypertarget{block_B}{block_B}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{R}{transaction_receipt}\hypertarget{R u}{tx_receipt_gas_used}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{transaction_receipt_R}\hypertarget{tx_receipt_gas_used_R__u}{tx_receipt_gas_used_R__u}
 

### 2018-01-23 Paper.tex
hypertargets lines 350–399

--- 2018-01-23 Paper.tex
@@
 \begin{equation}
-\begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+\begin{array}[t]{lclc}\hypertarget{H_r}{block_validity}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \hypertarget{H_o}{Ommer_block_hash}\\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \hypertarget{H_t}{tx_block_hash}\\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \hypertarget{H_e}{Receipts-Root-H_e}\\
 H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
@@
 
+\hypertarget{f*}{general-element-wise-sequence-transformation}
 With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
@@
 
+\hypertarget{H d}{block-difficulty}
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
@@
 
+\hypertarget{H l}{block-gas-limit}
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
@@
 
+\hypertarget{h cancel n}{}
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
@@
 
+\hypertarget{block-header-validity-function}{block-header-validity-function}
 Thus we are able to define the block header validity function $V(H)$:
@@
 \end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+\hypertarget{self_destruct_set_wordy_defn}{self_destruct_set_wordy_defn}
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the\hypertarget{tx_log_series_wordy_defn}{tx_log_series_wordy_defn} transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as\hypertarget{tx_touched_accounts_wordy_defn} decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end\hypertarget{refund_balance_defn_words}{refund_balance_defn_words} of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
@@
 
-\subsection{Execution}
+\subsection{Execution}\hypertarget{intrinsic_gas_g_0}{intrinsic_gas_g_0}
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction_nonce}{transaction_nonce}\\
 T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \hypertarget{self_destruct_list_A__s}{self_destruct_list_A__s}\\
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \hypertarget{touched_A__t}{touched_A__t}\\
 \forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
@@
 
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{tx_total_gas_used_Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{tx_logs_Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_s}
 And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
@@
 \section{Contract Creation} \label{ch:create}
+\hypertarget{endow}{endow}
 

### 2018-01-23 Paper.tex
hypertargets lines 350–399

--- 2018-01-23 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{ommerheaders} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-\begin{description}\hypertarget{H p}{parentHash}
+\begin{description}
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{H c}{beneficiary}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts-Root-def}{receipts-Root-def}
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs-Bloom-def}{logs-Bloom-def}
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{H_i}{blocknumber}
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{H_g}{gasUsed}
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{H_s}{timestamp}
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{H_x}{extraData}
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{H_m}{mixHash}
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block nonce}{block_nonce}\hypertarget{H_n}{block_nonce}
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
 \item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
-\end{description}\hypertarget{B U}{}
+\end{description}
 
@@
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}\hypertarget{Transaction Receipt}{Transaction_Receipt}
+\end{equation}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{R}{transaction_receipt}\hypertarget{R u}{tx_receipt_gas_used}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
@@
 \begin{equation}
-\begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+\begin{array}[t]{lclc}\hypertarget{block_validity_H__r}{block_validity_H__r}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \hypertarget{Ommer_block_hash_H__o}{Ommer_block_hash_H__o}\\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \hypertarget{tx_block_hash_H__t}{tx_block_hash_H__t}\\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \hypertarget{Receipts_Root_H__e}{Receipts_Root_H__e}\\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \hypertarget{logs_Bloom_filter_H__b}{logs_Bloom_filter_H__b}\\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)

### 2018-01-23 Paper.tex
Hypertargets lines 400–499

@@
 \begin{equation}
-\begin{array}[t]{lclc}\hypertarget{block_validity_H__r}{block_validity_H__r}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \hypertarget{Ommer_block_hash_H__o}{Ommer_block_hash_H__o}\\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \hypertarget{tx_block_hash_H__t}{tx_block_hash_H__t}\\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \hypertarget{Receipts_Root_H__e}{Receipts_Root_H__e}\\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \hypertarget{logs_Bloom_filter_H__b}{logs_Bloom_filter_H__b}\\
+\begin{array}[t]{lclc}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
@@
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
+\end{eqnarray}\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{general_element_wise_sequence_transformation_f_pow_asterisk}
 
@@
 \newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
+\newcommand{\diffadjustment}{x}\hypertarget{block_difficulty_H__d}{block_difficulty_H__d}
 
@@
 H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}
+\end{align}\hypertarget{block_gas_limit_H__l}{block_gas_limit_H__l}
 

### 2018-01-23 Paper.tex
hypertargets lines 500–599

@@
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{general_element_wise_sequence_transformation_f_pow_asterisk}
+\end{eqnarray}
 
@@
 \newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}\hypertarget{block_difficulty_H__d}{block_difficulty_H__d}
+\newcommand{\diffadjustment}{x}
 
@@
 H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}\hypertarget{block_gas_limit_H__l}{block_gas_limit_H__l}
+\end{align}
 
@@
 
-$H_s$ is the timestamp (in Unix time) of block $H$ at the block's inception, and must fulfil the relation:
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
 \begin{equation}
@@
 \end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{block_header_without_nonce_and_mixmash_h__cancel_n}
 
@@
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.\hypertarget{block_header_validity_function}{block_header_validity_function}
 
@@
 A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+\end{equation}\hypertarget{self_destruct_set_wordy_defn}{self_destruct_set_wordy_defn}
+
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the\hypertarget{tx_log_series_wordy_defn}{tx_log_series_wordy_defn} transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as\hypertarget{tx_touched_accounts_wordy_defn} decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end\hypertarget{refund_balance_defn_words}{refund_balance_defn_words} of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.

### 2018-01-23 Paper.tex
hypertargets lines 871–882 exec params

--- 2018-01-23 Paper.tex
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \hypertarget{self_destruct_list_A__s}{self_destruct_list_A__s}\\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \hypertarget{touched_A__t}{touched_A__t}\\
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
 \forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{tx_total_gas_used_Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{tx_logs_Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_s}
 And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
@@
 \section{Contract Creation} \label{ch:create}
-\hypertarget{endow}{endow}
 
@@
 
-\begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\begin{itemize}\hypertarget{address_of_account_which_owns_code_I__a_def_words}{address_of_account_which_owns_code_I__a_def_words}
+\item $I_a$, the address of the account which owns the code that is executing.\hypertarget{address_of_sender_of_tx_that_originated_execution_I__o}{address_of_sender_of_tx_that_originated_execution_I__o}
+\item $I_o$, the sender address of the transaction that originated this execution.\hypertarget{gas_price_in_tx_that_originated_execution_I__p}{gas_price_in_tx_that_originated_execution_I__p}
+\item $I_p$, the price of gas in the transaction that originated this execution.\hypertarget{execution_input_data_I__d}{execution_input_data_I__d}
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.\hypertarget{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.\hypertarget{Wei_value_exec}{Wei_value_exec_I__v}
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.\hypertarget{exec_machine_code_I__b}{exec_machine_code_I__b}
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.\hypertarget{exec_block_header_I__H}{exec_block_header_I__H}
+\item $I_H$, the block header of the present block.\hypertarget{exec_call_or_create_depth_I__e}{exec_call_or_create_depth_I__e}
+\item $I_e$, the depth of the present message-call or contract-creation (i.e.
+the number of {\small CALL}s or {\small CREATE}s being executed at present).\hypertarget{exec_permission_to_modify_state_I__w}{exec_permission_to_modify_state_I__w}
 \item $I_w$, the permission to make modifications to the state.

### 2018-01-23 Paper.tex
\hypertarget{empty_sequence_vs_empty_set}{empty_sequence_vs_empty_set}

--- 2018-01-23 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-\begin{description}
+\begin{description}\hypertarget{H p}{}
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{H c}{}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
@@
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{H i}{}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{H g}{}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{H s}{}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{H x}{
 \item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.\hypertarget{empty_sequence_vs_empty_set}{empty_sequence_vs_empty_set}
 

### 2018-01-23 Paper.tex
\hypertarget{Exceptional_Halting_function_Z}{Exceptional_Halting_function_Z}

--- 2018-01-23 Paper.tex
@@
 
-\begin{itemize}\hypertarget{address_of_account_which_owns_code_I__a_def_words}{address_of_account_which_owns_code_I__a_def_words}
-\item $I_a$, the address of the account which owns the code that is executing.\hypertarget{address_of_sender_of_tx_that_originated_execution_I__o}{address_of_sender_of_tx_that_originated_execution_I__o}
-\item $I_o$, the sender address of the transaction that originated this execution.\hypertarget{gas_price_in_tx_that_originated_execution_I__p}{gas_price_in_tx_that_originated_execution_I__p}
-\item $I_p$, the price of gas in the transaction that originated this execution.\hypertarget{execution_input_data_I__d}{execution_input_data_I__d}
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.\hypertarget{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.\hypertarget{Wei_value_exec}{Wei_value_exec_I__v}
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.\hypertarget{exec_machine_code_I__b}{exec_machine_code_I__b}
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.\hypertarget{exec_block_header_I__H}{exec_block_header_I__H}
-\item $I_H$, the block header of the present block.\hypertarget{exec_call_or_create_depth_I__e}{exec_call_or_create_depth_I__e}
-\item $I_e$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).\hypertarget{exec_permission_to_modify_state_I__w}{exec_permission_to_modify_state_I__w}
+\begin{itemize}
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \item $I_w$, the permission to make modifications to the state.
@@
 
-\subsubsection{Exceptional Halting}
+\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{Exceptional_Halting_function_Z}
 

### 2018-01-23 Paper.tex
\hypertarget{normal_halting_function_H}{normal_halting_function_H}
\hypertarget{JUMP}{JUMP} \hypertarget{JUMPI}{JUMPI} \label{eq:mu_pc}
Comment out tabularx
Remove duplicate REVERT opcode

@@
 
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{Exceptional_Halting_function_Z}
+\subsubsection{Exceptional Halting}
 
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
-                 but returning data and gas. \\
+\midrule
 0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\

### 2018-01-23 Paper.tex
changed ordering or receipt items to correct client implementation

@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the status code of the transaction, $R_{s'}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$, and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{s'}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 $G$ is fully defined in Appendix \ref{app:fees}.
-
-%todo Explain g_d reason?
 

### 2018-01-24 Paper.tex

--- 2018-01-24 Paper.tex
@@
 \midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we need to define the memory usage: \\
+&&&& For the gas calculation, we use the memory expansion function, \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\

### 2018-01-24 Paper.tex
Update Paper.tex

--- 2018-01-24 Paper.tex
@@
 $G$ is fully defined in Appendix \ref{app:fees}.
+
+%todo Explain g_d reason?
 
@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small 
+} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
@@
 0 & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ \hypertarget{JUMP}{JUMP} \\
 \midrule
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \hypertarget{JUMPI}{JUMPI}\\
 \midrule
@@
 &&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes
+                 but returning data and gas. \\
 0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and gas. \\

### 2018-01-24 Paper.tex
Omit comments on longtable and tabularx

--- 2018-01-24 Paper.tex
@@
 
-% I will insert other table packages in here after this is merged.
+% Tabu seems more flexible, and generally better.
+\usepackage{tabu} %requires array.
 
@@
 
+%todo Explain g_d reason?
+
 The up-front cost $v_0$ is calculated as:
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or that are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}

### 2018-01-25 Paper.tex
Omit comment on tabu
Add a space before \hypertarget in text mode to see if that will cause the build to pass.
omit network congestion

@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-% Tabu seems more flexible, and generally better.
-\usepackage{tabu} %requires array.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
 
@@
 
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
@@
 \end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \quad s' &\equiv \begin{cases}
@@
 1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\end{cases}
 \end{align}

### 2018-01-25 Paper.tex
Testing without underscores in the captions
Move \hypertarget{account_nonce}{account nonce}

--- 2018-01-24 Paper.tex
@@
 
-% Tabu seems more flexible, and generally better.
-\usepackage{tabu} %requires array.
+% I will insert other table packages in here after this is merged.
 
@@
 A valid state transition is one which comes about through a transaction.
-Formally:
+Formally: \hypertarget{Upsilon_state_transition}{Upsilon state transition}
 \begin{equation}
@@
 
-The account state comprises the following four fields:
+The account state comprises the following four fields:\hypertarget{account_nonce}{account nonce}
 

### 2018-01-25 Paper.tex
Move \hypertarget{tx_nonce}{tx_nonce_T__n}

--- 2018-01-24 Paper.tex
@@
 \end{equation}
-where $v$ is the account validity function:
+where $v$ is the account validity function:\hypertarget{account_validity_function_v__x}{account_validity_function_v__x}
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+Both types specify a number of common fields:\hypertarget{tx_nonce}{tx_nonce_T__n}
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx_gas_price_T__p}
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{tx_gas_limit_T__g}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx_to_address_T__t}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx_value_T__v}{tx_value_T__v}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-25 Paper.tex
Move \hypertarget{parent_Hash_H__p_def_words}{parent_Hash_H__p_def_words}

--- 2018-01-25 Paper.tex
@@
 A valid state transition is one which comes about through a transaction.
-Formally: \hypertarget{Upsilon_state_transition}{Upsilon state transition}
+Formally:
 \begin{equation}
@@
 
-The account state comprises the following four fields:\hypertarget{account_nonce}{account nonce}
+The account state comprises the following four fields:
 
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
-The block header contains several pieces of information:
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{ommerheaders} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block header contains several pieces of information:\hypertarget{parent_Hash_H__p_def_words}{parent_Hash_H__p_def_words}
 
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{beneficiary_H__c}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{receipts_Root_def_words}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{logs_Bloom_def_words}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{block_number_H_i}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{block_gas_used_H__g}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{block_timestamp_H__s}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{block_extraData_H__x}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{mixHash_H__m}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{block_nonce_H__n}
 \item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
-\end{description}
+\end{description}\hypertarget{ommer_block_headers_B__U}{ommer_block_headers_B__U}\hypertarget{block_B}{block_B}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{transaction_receipt_R}\hypertarget{tx_receipt_gas_used_R__u}{tx_receipt_gas_used_R__u}
 

### 2018-01-25 Paper.tex
Move \hypertarget{ommer_block_headers_B__U}{ommer_block_headers_B__U}\hypertarget{block_B}{block_B}

@@
 \item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{block_nonce_H__n}
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
-\end{description}\hypertarget{ommer_block_headers_B__U}{ommer_block_headers_B__U}\hypertarget{block_B}{block_B}
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.\hypertarget{ommer_block_headers_B__U}{ommer_block_headers_B__U}\hypertarget{block_B}{block_B}
+\end{description}
 

### 2018-01-25 Paper.tex
Omit underscores in captions

--- 2018-01-25 Paper.tex
@@
 \end{equation}
-where $v$ is the account validity function:\hypertarget{account_validity_function_v__x}{account_validity_function_v__x}
+where $v$ is the account validity function:\hypertarget{account_validity_function_v__x}{account validity function v x}
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:\hypertarget{tx_nonce}{tx_nonce_T__n}
+Both types specify a number of common fields:\hypertarget{tx_nonce}{tx nonce T n}
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx_gas_price_T__p}
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx gas price T p}
 \item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{tx_gas_limit_T__g}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx_to_address_T__t}
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx_value_T__v}{tx_value_T__v}
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx to address T t}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx value T v}{tx value T v}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-25 Paper.tex
Test a hypen in a caption

@@
 A valid state transition is one which comes about through a transaction.
-Formally:
+Formally: \hypertarget{Upsilon_state_transition}{Upsilon-state transition}
 \begin{equation}
@@
 \end{equation}
-where $v$ is the account validity function:\hypertarget{account_validity_function_v__x}{account validity function v x}
+where $v$ is the account validity function:
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:\hypertarget{tx_nonce}{tx nonce T n}
+Both types specify a number of common fields:
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx gas price T p}
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{tx_gas_limit_T__g}
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx to address T t}
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx value T v}{tx value T v}
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-25 Paper.tex
Hypertargets for Omega, block-reward-R-b, Gamma, Phi and Pi

@@
 A valid state transition is one which comes about through a transaction.
-Formally: \hypertarget{Upsilon_state_transition}{Upsilon-state transition}
+Formally:
 \begin{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{Omega} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
+two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.\hypertarget{block-reward-R-b}{block-reward-R-b}
 
@@
 
-\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
+\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}\hypertarget{Gamma}{Gamma}
 
@@
 
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.\hypertarget{Phi}{Phi}
 
@@
 \mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\hypertarget{Pi}{Pi}
 \end{equation}

### 2018-01-25 Paper.tex
Change underscores to spaces in hypertarget captions

--- 2018-01-25 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
-The block header contains several pieces of information:
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{ommerheaders} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block header contains several pieces of information:\hypertarget{parent_Hash_H__p_def_words}{parent Hash H p def words}
 
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{beneficiary H c}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{receipts_Root_def_words}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{logs Bloom def words}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{block number H i}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{block gas used H g}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{block timestamp H s}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{block extraData H x}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{mixHash_H__m}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{block nonce H n}
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.\hypertarget{ommer block headers B U}{ommer_block_headers_B__U}\hypertarget{block_B}{block B}
 \end{description}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{transaction_receipt_R}\hypertarget{tx_receipt_gas_used_R__u}{tx receipt gas used R u}
 

### 2018-01-25 Paper.tex
Change underscores to spaces in hypertarget captions

--- 2018-01-25 Paper.tex
@@
 \end{equation}
-where $v$ is the account validity function:
+where $v$ is the account validity function:\hypertarget{account validity function v x}{account validity function v x}
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+Both types specify a number of common fields:\hypertarget{tx_nonce}{tx nonce T n}
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx gas price T p}
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{tx gas limit T g}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx to address T t}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx value T v}{tx value T v}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-25 Paper.tex
Change underscores to spaces in hypertarget captions

--- 2018-01-25 Paper.tex
@@
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{receipts_Root_def_words}
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{receipts Root def words}
 \item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{logs Bloom def words}
@@
 \item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{block extraData H x}
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{mixHash_H__m}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{mixHash H m}
 \item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{block nonce H n}
@@
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}\hypertarget{Transaction Receipt}{Transaction_Receipt}
+\end{equation}\hypertarget{Transaction Receipt}{Transaction Receipt}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{transaction_receipt_R}\hypertarget{tx_receipt_gas_used_R__u}{tx receipt gas used R u}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{transaction receipt R}\hypertarget{tx_receipt_gas_used_R__u}{tx receipt gas used R u}
 

### 2018-01-25 Paper.tex
Change captions to empty.

--- 2018-01-25 Paper.tex
@@
 \end{equation}
-where $v$ is the account validity function:\hypertarget{account validity function v x}{account validity function v x}
+where $v$ is the account validity function:
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:\hypertarget{tx_nonce}{tx nonce T n}
+Both types specify a number of common fields:
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{tx gas price T p}
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{tx gas limit T g}
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{tx to address T t}
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx value T v}{tx value T v}
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
@@
 \begin{equation}
-\begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+\begin{array}[t]{lclc}\hypertarget{block_validity_H__r}{}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \hypertarget{Ommer_block_hash_H__o}{}\\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \hypertarget{tx_block_hash_H__t}{}\\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \hypertarget{Receipts_Root_H__e}{}\\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \hypertarget{logs_Bloom_filter_H__b}{}\\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)

### 2018-01-25 Paper.tex
Empty caption

@@
 A valid state transition is one which comes about through a transaction.
-Formally:
+Formally: \hypertarget{Upsilon_state_transition}{}
 \begin{equation}
@@
 \begin{equation}
-\begin{array}[t]{lclc}\hypertarget{block_validity_H__r}{}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \hypertarget{Ommer_block_hash_H__o}{}\\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \hypertarget{tx_block_hash_H__t}{}\\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \hypertarget{Receipts_Root_H__e}{}\\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \hypertarget{logs_Bloom_filter_H__b}{}\\
+\begin{array}[t]{lclc}
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

@@
 
-I define a number of useful functions throughout. \hypertarget{ell}{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \hypertarget{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-The account state comprises the following four fields:\hypertarget{account_nonce}{account nonce}
+The account state comprises the following four fields:\hypertarget{account_nonce}{}
 

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

@@
 A valid state transition is one which comes about through a transaction.
-Formally: \hypertarget{Upsilon_state_transition}{}
+Formally:
 \begin{equation}
@@
 
-I define a number of useful functions throughout. \hypertarget{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-The account state comprises the following four fields:\hypertarget{account_nonce}{}
+The account state comprises the following four fields:
 
@@
 \begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
+\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))\hypertarget{account validity function v x}{}
 \end{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+Both types specify a number of common fields:\hypertarget{tx_nonce}{}
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\hypertarget{tx_gas_price_T__p}{}
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\hypertarget{tx_gas_limit_T__g}{}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\hypertarget{tx_to_address_T__t}{}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\hypertarget{tx value T v}{}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

--- 2018-01-25 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
-The block header contains several pieces of information:
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block header contains several pieces of information:\hypertarget{parent_Hash_H__p_def_words}{}
 
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{}
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.\hypertarget{}{ommer_block_headers_B__U}\hypertarget{block_B}{}
 \end{description}
@@
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
+\end{equation}\hypertarget{Transaction Receipt}{}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{}\hypertarget{tx_receipt_gas_used_R__u}{}
 

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

--- 2018-01-25 Paper.tex
@@
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
+\end{eqnarray}\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}
 
@@
 \newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
+\newcommand{\diffadjustment}{x}\hypertarget{block_difficulty_H__d}{}
 
@@
 H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}
+\end{align}\hypertarget{block_gas_limit_H__l}{}
 

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

@@
 \quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}
+\end{eqnarray}
 
@@
 \newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}\hypertarget{block_difficulty_H__d}{}
+\newcommand{\diffadjustment}{x}
 
@@
 H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}\hypertarget{block_gas_limit_H__l}{}
+\end{align}
 
@@
 \end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}
 
@@
 This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.\hypertarget{block_header_validity_function}{}
 
@@
 A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+\end{equation}\hypertarget{self_destruct_set_wordy_defn}{}
+
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the\hypertarget{tx_log_series_wordy_defn}{} transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as\hypertarget{tx_touched_accounts_wordy_defn}{} decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end\hypertarget{refund_balance_defn_words}{} of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
@@
 
-\subsection{Execution}
+\subsection{Execution}\hypertarget{intrinsic_gas_g_0}{}
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
+\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \hypertarget{transaction_nonce}{}\\
 T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

--- 2018-01-25 Paper.tex
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \hypertarget{self_destruct_list_A__s}{}\\
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \hypertarget{touched_A__t}{}\\
 \forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
@@
 
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}
 And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
@@
 
-These are used to help define the transaction receipt, discussed later.
+These are used to help define the transaction receipt, discussed later.\hypertarget{endow}{}
 

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings

--- 2018-01-25 Paper.tex
@@
 
-\begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\begin{itemize}\hypertarget{address_of_account_which_owns_code_I__a_def_words}{}
+\item $I_a$, the address of the account which owns the code that is executing.\hypertarget{address_of_sender_of_tx_that_originated_execution_I__o}{}
+\item $I_o$, the sender address of the transaction that originated this execution.\hypertarget{gas_price_in_tx_that_originated_execution_I__p}{}
+\item $I_p$, the price of gas in the transaction that originated this execution.\hypertarget{execution_input_data_I__d}{}
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.\hypertarget{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{}
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.\hypertarget{Wei_value_exec}{}
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.\hypertarget{exec_machine_code_I__b}{}
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.\hypertarget{exec_block_header_I__H}{}
+\item $I_H$, the block header of the present block.\hypertarget{exec_call_or_create_depth_I__e}{exec_call_or_create_depth_I__e}
+\item $I_e$, the depth of the present message-call or contract-creation (i.e.
+the number of {\small CALL}s or {\small CREATE}s being executed at present).\hypertarget{exec_permission_to_modify_state_I__w}{}
 \item $I_w$, the permission to make modifications to the state.

### 2018-01-25 Paper.tex
Add a missing {}

--- 2018-01-25 Paper.tex
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{}
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.\hypertarget{}{ommer_block_headers_B__U}{}\hypertarget{block_B}{}
 \end{description}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \hypertarget{self_destruct_list_A__s}{}\\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \hypertarget{touched_A__t}{}\\
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
 \forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}
 And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
@@
 
-These are used to help define the transaction receipt, discussed later.\hypertarget{endow}{}
+These are used to help define the transaction receipt, discussed later.
 

### 2018-01-25 Paper.tex
Change hypertarget captions to empty strings
move \hypertarget{account nonce}{}

--- 2018-01-25 Paper.tex
@@
 
-The account state comprises the following four fields:
+The account state comprises the following four fields:\hypertarget{account nonce}{}
 

### 2018-01-25 Paper.tex
Change hypertarget caption to empty string

--- 2018-01-25 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding\hypertarget{ommerheaders}{} to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
-The block header contains several pieces of information:\hypertarget{parent_Hash_H__p_def_words}{}
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block header contains several pieces of information:
 
@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{}
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{}
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{}
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block.
-This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{}
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{}
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{}
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{}
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{}
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{}
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.\hypertarget{}{ommer_block_headers_B__U}{}\hypertarget{block_B}{}
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}\hypertarget{Transaction Receipt}{}
+\end{equation}
 
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.\hypertarget{transaction_receipt_R}{}\hypertarget{tx_receipt_gas_used_R__u}{}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
@@
 
-\subsubsection{Exceptional Halting}
+\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}
 

### 2018-01-25 Paper.tex
Change hypertarget caption to empty string normal_halting_function_H
Fix ommer_block_headers_B__U hypertarget

@@
 \item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.\hypertarget{beneficiary_H__c}{}
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.\hypertarget{receipts_Root_def_words}{}
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.\hypertarget{logs_Bloom_def_words}{}
 \item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block.
+This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.\hypertarget{block_number_H_i}{}
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.\hypertarget{block_gas_used_H__g}{}
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.\hypertarget{block_timestamp_H__s}{}
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.\hypertarget{block_extraData_H__x}{}
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.\hypertarget{mixHash_H__m}{}
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.\hypertarget{block_nonce_H__n}{}
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.\hypertarget{ommer_block_headers_B__U}{}\hypertarget{block_B}{}
 \end{description}
@@
 
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}
+\subsubsection{Exceptional Halting}
 

### 2018-01-25 Paper.tex
\hypertarget{empty_sequence_vs_empty_set}{}

--- 2018-01-25 Paper.tex
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.\hypertarget{empty_sequence_vs_empty_set}{}
 

### 2018-01-25 Paper.tex
Add a comment on using the hyperref package

@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
 
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.\hypertarget{empty_sequence_vs_empty_set}{}
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
 

### 2018-01-25 Paper.tex
\hypertarget{rlp}{}
\hypertarget{R b}{
\hypertarget{R l}{}
\hypertarget{trie}{}
\hypertarget{ECDSASIGN}{}
\hypertarget{v}{}

@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
 

### 2018-01-25 Paper.tex
Add definitions for T_w, T_r and T_s (commented out, to be uncommented later)
\hypertarget{h T}{}
hypertargets for gas costs

--- 2018-01-25 Paper.tex
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
 $G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \hypertarget{codedeposit}{}\\
 $G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
 $G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
@@
 $G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \hypertarget{Gtxcreate}{}\\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\hypertarget{Gtxdatazero}{}\\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \hypertarget{Gtxdatanonzero}{}\\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \hypertarget{Gtransaction}{}\\
 $G_{transaction}$ & 21000 & Paid for every transaction. \\

### 2018-01-25 Paper.tex
hypertargets for operations

--- 2018-01-25 Paper.tex
@@
 
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 \multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{stop}{}\\
 0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
@@
 \multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{blockhash}{BLOCKHASH}\\
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \hypertarget{SSTORE}{}\\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
@@
 G_{sreset} & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ \hypertarget{A r}{}\\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
@@
 \multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
+\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\hypertarget{A l}{}\\
 \multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
@@
 \midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{}\\
 \midrule
@@
 0 & \text{otherwise}
-\end{cases}$ \\
+\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}

### 2018-01-25 Paper.tex
Delete comments on LLL
This is already in the terminology list.

--- 2018-01-25 Paper.tex
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-\hypertarget{Exceptional_Halting_function_Z}{}
+
 \subsubsection{Exceptional Halting}
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\

### 2018-01-25 Paper.tex
\hypertarget{Jwordbytes}{}
hypertargets for Ethash parameters

@@
 
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
 \hypertarget{GenesisBlock}{}
@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{wordbytes}$ & 4  & Bytes in word. \hypertarget{Jdatasetinit}{}\\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \hypertarget{Jdatasetgrowth}{}\\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \hypertarget{Jcacheinit}{}\\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \hypertarget{Jcachegrowth}{}\\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \hypertarget{Jepoch}{}\\
+$J_{epoch}$ & 30000 & Blocks per epoch. \hypertarget{Jmixbytes}{}\\
+$J_{mixbytes}$ & 128 & mix length in bytes. \hypertarget{Jhashbytes}{}\\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \hypertarget{Jparents}{}\\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\hypertarget{Jcacherounds}{}\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \hypertarget{Jaccesses}{}\\
 $J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
@@
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.\hypertarget{dsize}{}
 Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\hypertarget{Ecacherounds}{}
 \end{equation}

### 2018-01-25 Paper.tex
\hypertarget{bigvee}{} move closing brace

--- 2018-01-25 Paper.tex
@@
 
-\hypertarget{GenesisBlock}{}
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
 \section{Genesis Block}\label{app:genesis}
@@
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabularx}{\textwidth}{lcX}
+\toprule
+Symbol & Latex Command & Description \\
+\midrule
+\hypertarget{bigvee}{}
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+\bottomrule
+\end{tabularx}
+
 \end{document}

### 2018-01-25 Paper.tex
Delete redundant empty line
Add commands to manage URL breaks + glossaries package
Replace all _\mathbf with _\mathrm

--- 2018-01-25 Paper.tex
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathrm{s})\big) \equiv \boldsymbol{\sigma}[a]_s
 \end{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathrm{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathrm{d}$.
 \end{description}
@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+(T_n, T_p, T_g, T_t, T_v, T_\mathrm{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathrm{d}, T_w, T_r, T_s) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathrm{i}$ and $T_\mathrm{d}$.
 \begin{equation}
@@
 T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathrm{d} \in \mathbb{B} & \wedge & T_\mathrm{i} \in \mathbb{B}
 \end{array}
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+The address hash $T_\mathrm{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_\mathrm{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_H, B_\mathrm{T}, B_\mathrm{U})
 \end{equation}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+Each receipt, denoted $B_\mathrm{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathrm{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathrm{l}, R_{s'})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathrm{l})
 \end{equation}
@@
 
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
-
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+%Notably $B_\mathrm{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathrm{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathrm{t}$ and some number of bytes of data, $O_\mathrm{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathrm{t}}_0, {O_\mathrm{t}}_1, ...), O_\mathrm{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathrm{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathrm{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathrm{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathrm{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
@@
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathrm{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathrm{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathrm{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathrm{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathrm{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathrm{R}} \big( \mathbf{r}_b \big)
 \end{array}
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathrm{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathrm{T}), L_H^*(B_\mathrm{U}) \big)
 \end{eqnarray}
@@
 x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+1 & \text{if} \, \lVert P(H)_\mathrm{U}\rVert = 0 \\
 2 & \text{otherwise}
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+A \equiv (A_\mathrm{s}, A_\mathrm{l}, A_\mathrm{t}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathrm{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathrm{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+$A_\mathrm{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
 Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_{i \in T_\mathrm{i}, T_\mathrm{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
@@
 
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+where $T_\mathrm{i},T_\mathrm{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
 $G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+T_g & \leqslant & {B_H}_l - \ell(B_\mathrm{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathrm{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
 
@@
 (\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathrm{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathrm{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\forall i \in A_\mathrm{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathrm{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathrm{l} \\
 \Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathrm{s}, \boldsymbol{\sigma}[s]_c)
 \end{eqnarray}
@@
 I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
+I_\mathrm{d} & \equiv & () \\
 I_s & \equiv & s \\
 I_v & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_\mathrm{b} & \equiv & \mathbf{i} \\
 I_e & \equiv & e \\
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+$I_\mathrm{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathrm{s}, \boldsymbol{\sigma}_1'[s]_c)
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathrm{s}, \boldsymbol{\sigma}[r]_c)
 \end{equation}
@@
 I_p & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_\mathrm{d} & \equiv & \mathbf{d} \\
 I_s & \equiv & s \\
@@
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathrm{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathrm{b}), I_\mathrm{b})$ at some point prior in order to make the determination of $I_\mathrm{b}$ feasible.
 
@@
 \item $I_p$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_\mathrm{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
 \item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_\mathrm{b}$, the byte array that is the machine code to be executed.
 \item $I_H$, the block header of the present block.
@@
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_\mathrm{m} & \equiv & (0, 0, ...) \\
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
+\boldsymbol{\mu}_\mathrm{s} & \equiv & () \\
+\boldsymbol{\mu}_\mathrm{o} & \equiv & ()
 \end{eqnarray}
@@
 \mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathrm{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathrm{s}[0] \notin D(I_\mathrm{b}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathrm{s}[1] + \boldsymbol{\mu}_\mathrm{s}[2] > \lVert\boldsymbol{\mu}_\mathrm{o}\rVert) \quad \vee \\
+ \lVert\boldsymbol{\mu}_\mathrm{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
  \neg I_w \wedge W(w, \boldsymbol{\mu})
@@
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathrm{s}[2] \neq 0
 \end{array}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
+\boldsymbol{\mu}'_\mathrm{m} & \equiv & \boldsymbol{\mu}_\mathrm{m} \\
 \boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathrm{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathrm{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathrm{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
 
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathrm{U}\rVert}{32})R_b \\
+\qquad\forall_{U \in B_\mathrm{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathrm{s}, \boldsymbol{\sigma}[U_c]_c) \\
 R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathrm{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathrm{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathrm{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathrm{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+For $\mathbf{R}[n]_\mathrm{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathrm{l} =
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathrm{T}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathrm{T}[n])
 \end{equation}
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathrm{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathrm{d})\\
 \end{eqnarray}
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathrm{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathrm{d}
 \end{eqnarray}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+F&\equiv&(|I_\mathrm{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+k &=& \dfrac{|I_\mathrm{d}|}{192} \\
 g_r&=& 60000k + 40000 \\
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathrm{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathrm{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathrm{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathrm{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathrm{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathrm{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathrm{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathrm{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathrm{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathrm{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] \times \boldsymbol{\mu}_\mathrm{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] - \boldsymbol{\mu}_\mathrm{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathrm{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]) \lfloor |\boldsymbol{\mu}_\mathrm{s}[0] \div \boldsymbol{\mu}_\mathrm{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \boldsymbol{\mu}_\mathrm{s}[0] \bmod \boldsymbol{\mu}_\mathrm{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathrm{s}[0]) (|\boldsymbol{\mu}_\mathrm{s}[0]| \bmod |\boldsymbol{\mu}_\mathrm{s}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] = 0\\ (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1]) \mod \boldsymbol{\mu}_\mathrm{s}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] = 0\\ (\boldsymbol{\mu}_\mathrm{s}[0] \times \boldsymbol{\mu}_\mathrm{s}[1]) \mod \boldsymbol{\mu}_\mathrm{s}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0] ^ {\boldsymbol{\mu}_\mathrm{s}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathrm{s}[0] + 1) \\ \boldsymbol{\mu}_\mathrm{s}[1]_i &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathrm{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathrm{s}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] < \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] < \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = \boldsymbol{\mu}_\mathrm{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_i \equiv \boldsymbol{\mu}_\mathrm{s}[0]_i \wedge \boldsymbol{\mu}_\mathrm{s}[1]_i$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_i \equiv \boldsymbol{\mu}_\mathrm{s}[0]_i \vee \boldsymbol{\mu}_\mathrm{s}[1]_i$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_i \equiv \boldsymbol{\mu}_\mathrm{s}[0]_i \oplus \boldsymbol{\mu}_\mathrm{s}[1]_i$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathrm{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[1]_{(i + 8\boldsymbol{\mu}_\mathrm{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathrm{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_o$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_s$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_v$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_\mathrm{d}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ] \quad \text{with} \quad I_\mathrm{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathrm{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert I_\mathrm{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
+\begin{cases} I_\mathrm{d}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert I_\mathrm{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert I_\mathrm{b} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
+\begin{cases} I_\mathrm{b}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert I_\mathrm{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[3] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathrm{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[1], \boldsymbol{\mu}_\mathrm{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathrm{o} \rVert$ \\
 \midrule
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathrm{s}[2] - 1\} } \boldsymbol{\mu}'_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_\mathrm{o}[\boldsymbol{\mu}_\mathrm{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] + i < \lVert \boldsymbol{\mu}_\mathrm{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathrm{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathrm{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{m}[\boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 32) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
@@
 0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 32) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
@@
 0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] ] \equiv (\boldsymbol{\mu}_\mathrm{s}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathrm{s}[0] + 1) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathrm{s}[ \boldsymbol{\mu}_\mathrm{s}[0] ] \equiv \boldsymbol{\mu}_\mathrm{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathrm{s}[\boldsymbol{\mu}_\mathrm{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathrm{s}[0] $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
@@
 0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathrm{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathrm{b}[x] & \text{if} \quad x < \lVert I_\mathrm{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
 \midrule
 0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
 \midrule
@@
 0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[15]$ \\
 \bottomrule
@@
 0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[1] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
 \midrule
 0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[2]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[2] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
 \midrule
@@
 0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv \boldsymbol{\mu}_\mathrm{s}[16]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[16] \equiv \boldsymbol{\mu}_\mathrm{s}[0]$ \\
 \bottomrule
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathrm{l} \equiv A_\mathrm{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots (\boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathrm{s}[2])$ \\
 \midrule
@@
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[3], \boldsymbol{\mu}_\mathrm{s}[4], \boldsymbol{\mu}_\mathrm{s}[5])$ \\
 \bottomrule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[1] \dots (\boldsymbol{\mu}_\mathrm{s}[1] + \boldsymbol{\mu}_\mathrm{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathrm{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathrm{s} \equiv A_\mathrm{s} \cup A^+_\mathrm{s} \quad \wedge \quad A'_\mathrm{l} \equiv A_\mathrm{l} \cdot A^+_\mathrm{l} \quad \wedge \quad A'_\mathrm{t} \equiv A_\mathrm{t} \cup A^+_\mathrm{t}$ \\
+&&&& $ \wedge \quad A'_\mathrm{r} \equiv A_\mathrm{r} + A^+_\mathrm{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathrm{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[1], \boldsymbol{\mu}_\mathrm{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[3] \dots (\boldsymbol{\mu}_\mathrm{s}[3] + \boldsymbol{\mu}_\mathrm{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathrm{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[5] \dots (\boldsymbol{\mu}_\mathrm{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\boldsymbol{\mu}'_\mathrm{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathrm{s}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathrm{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[3], \boldsymbol{\mu}_\mathrm{s}[4]), \boldsymbol{\mu}_\mathrm{s}[5], \boldsymbol{\mu}_\mathrm{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathrm{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathrm{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathrm{s}[2], \boldsymbol{\mu}_\mathrm{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathrm{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathrm{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
@@
 0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathrm{m}[ \boldsymbol{\mu}_\mathrm{s}[0] \dots ( \boldsymbol{\mu}_\mathrm{s}[0] + \boldsymbol{\mu}_\mathrm{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& argument is $\boldsymbol{\mu}_\mathrm{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$, $\boldsymbol{\mu}_\mathrm{s}[5]$ and $\boldsymbol{\mu}_\mathrm{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathrm{s}[2]$, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$ and $\boldsymbol{\mu}_\mathrm{s}[5]$. \\
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathrm{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The argument $\boldsymbol{\mu}_\mathrm{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$, $\boldsymbol{\mu}_\mathrm{s}[5]$ and $\boldsymbol{\mu}_\mathrm{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathrm{s}[2]$, $\boldsymbol{\mu}_\mathrm{s}[3]$, $\boldsymbol{\mu}_\mathrm{s}[4]$ and $\boldsymbol{\mu}_\mathrm{s}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathrm{s}[0], \boldsymbol{\mu}_\mathrm{s}[1])$ \\
 \midrule
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_\mathrm{s} \equiv A_\mathrm{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
 \varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathrm{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
+(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathrm{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
 \end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
+&&&& where $r = \boldsymbol{\mu}_\mathrm{s}[0] \bmod 2^{160}$\\
 &&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathrm{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathrm{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule

### 2018-01-25 Paper.tex
Find and replace all _[:alpha][:alnum]* with _mathrm{$1} ...
... using Libre Open Office with regular expressions enabled as an option

--- 2018-01-25 Paper.tex
@@
 %E language?
-For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
 
@@
 \item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_mathrm{p}$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+(T_n, T_mathrm{p}, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_mathrm{p}, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_mathrm{p} \in \mathbb{P}_{256} & \wedge \\
 T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_mathrm{p}$.
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
@@
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_mathrm{l}$.
 \item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
@@
 \begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_mathrm{H}, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+A log entry, $O$, is a tuple of a logger's address, $O_mathrm{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_mathrm{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_mathrm{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}^*(B_\mathbf{U}))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_mathrm{H})_mathrm{H}}_r
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_mathrm{H})$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+The function $L_B$ and $L_mathrm{H}$ are the preparation functions for a block and block header respectively.
 Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\quad L_mathrm{H}(H) & \equiv & (\begin{array}[t]{l}H_mathrm{p}, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_mathrm{l}, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_mathrm{H}(B_mathrm{H}), L_T^*(B_\mathbf{T}), L_mathrm{H}^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_mathrm{H}^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_mathrm{p} \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
 H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_mathrm{l} \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
 H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
@@
 
-We define $P(B_H)$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+We define $P(B_mathrm{H})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_mathrm{H})) = H_mathrm{p}
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+H_i \equiv {{P(H)_mathrm{H}}_i} + 1
 \end{equation}
@@
 \mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\text{max}\!\left(\mindifficulty, {P(H)_mathrm{H}}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_mathrm{H}}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_mathrm{H}}_s}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
+The canonical gas limit $H_mathrm{l}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & H_mathrm{l} \geqslant 5000
 \end{eqnarray}
@@
 \begin{equation}
-H_s > {P(H)_H}_s
+H_s > {P(H)_mathrm{H}}_s
 \end{equation}
@@
 & & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 5000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
+& & H_g \le H_mathrm{l}  \quad \wedge \\
+& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
+& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
+& & H_mathrm{l} \geqslant 5000  \quad \wedge \\
+& & H_s > {P(H)_mathrm{H}}_s \quad \wedge \\
+& & H_i = {P(H)_mathrm{H}}_i +1 \quad \wedge \\
 & & \lVert H_x \rVert \le 32
@@
 
-%todo Explain g_d reason?
-
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv T_g T_mathrm{p} + T_v
 \end{equation}
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+T_g & \leqslant & {B_mathrm{H}}_mathrm{l} - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_mathrm{H}}_mathrm{l}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_mathrm{p}$.
 The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_mathrm{p} \\
 \boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_mathrm{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_mathrm{P}$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_mathrm{P}, g', A, s') \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_mathrm{p}, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_mathrm{p}, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_mathrm{P}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_mathrm{P} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_mathrm{P}[S(T)]_b + g^* T_mathrm{p} \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_mathrm{P}[m]_b + (T_g - g^*) T_mathrm{p} \\
+m & \equiv & {B_mathrm{H}}_c
 \end{eqnarray}
@@
 \begin{eqnarray}
-I_a & \equiv & a \\
+I_mathrm{a} & \equiv & a \\
 I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_mathrm{p} & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_mathrm{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
+I_mathrm{a} & \equiv & r \\
 I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_mathrm{p} & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
@@
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_mathrm{a}$, the address of the account which owns the code that is executing.
 \item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_mathrm{p}$, the price of gas in the transaction that originated this execution.
 \item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
@@
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
+\item $I_mathrm{H}$, the block header of the present block.
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
@@
 B_t & \equiv & B'_t + B_d \\
-B' & \equiv & P(B_H)
+B' & \equiv & P(B_mathrm{H})
 \end{eqnarray}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_mathrm{H})_mathrm{H}, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_mathrm{H}, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_mathrm{H}}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_mathrm{H}}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\boldsymbol{\sigma}'[{B_mathrm{H}}_c]_b & = & \boldsymbol{\sigma}[{B_mathrm{H}}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
 \qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_mathrm{H}}_i)) R_b
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_mathrm{H}) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_mathrm{H})_mathrm{H}}_r & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_mathrm{l}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_mathrm{l}$:
+\begin{eqnarray}
+R_mathrm{l}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_mathrm{p}$.
 
@@
 \begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+C_1\equiv\{(X,Y)\in F_mathrm{p}\times F_mathrm{p}\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_mathrm{p}[i]/(i+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_mathrm{p}$.
+\begin{equation}
+\delta_mathrm{p}(\mathbf x)\equiv\begin{cases}
 \mathbf x&\text{if}\ \mathbf x<p\\
@@
 \end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
+x&\equiv&\delta_mathrm{p}(\mathbf x[0..31])\\
+y&\equiv&\delta_mathrm{p}(\mathbf x[32..63])
 \end{eqnarray}
@@
 \end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
+x_0&\equiv&\delta_mathrm{p}(\mathbf x[0..31])\\
+y_0&\equiv&\delta_mathrm{p}(\mathbf x[32..63])\\
+x_1&\equiv&\delta_mathrm{p}(\mathbf x[64..95])\\
+y_1&\equiv&\delta_mathrm{p}(\mathbf x[96..127])
 \end{eqnarray}
@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+(T_n, T_mathrm{p}, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_mathrm{p}, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{a}$ \\
 \midrule
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{p}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_mathrm{p}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{l}$ \\
 \bottomrule
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_mathrm{a}, I_o, L(\boldsymbol{\mu}_g), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_mathrm{a}]_n = \boldsymbol{\sigma}[I_mathrm{a}]_n + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_mathrm{a}]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_mathrm{a}, \boldsymbol{\sigma}[I_mathrm{a}]_n)$, the address of the newly \\
 &&&& created account. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_mathrm{a}, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_mathrm{a}]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{a}, I_o, I_mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_mathrm{a}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_mathrm{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_mathrm{p}, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_mathrm{a} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_mathrm{a}]_b = 0\\
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_mathrm{a}]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_mathrm{a} \\
 (\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
@@
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_mathrm{a} \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_mathrm{a}]_b \neq 0$ \\
 \bottomrule
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_mathrm{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+ \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
 \end{equation}
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabularx}{\textwidth}{lcX}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\hypertarget{bigvee}{}
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabularx}
-
 \end{document}

### 2018-01-25 Paper.tex
Replace a double quote with two single quotes

--- 2018-01-25 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
-\quad s' &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabularx}{\textwidth}{lcX}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\hypertarget{bigvee}{}
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabularx}
-
 \end{document}

### 2018-01-25 Paper.tex

--- 2018-01-25 Paper.tex
@@
 
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+\usepackage{glossaries}
+
 \input{Version.tex}
@@
 \item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_mathrm{p}$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
@@
 L_T(T) \equiv \begin{cases}
-(T_n, T_mathrm{p}, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_mathrm{p}, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_mathrm{p} \in \mathbb{P}_{256} & \wedge \\
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
 T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_mathrm{p}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
@@
 \item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_mathrm{l}$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
 \item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
@@
 \begin{equation}
-B \equiv (B_mathrm{H}, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_mathrm{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_mathrm{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_mathrm{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}^*(B_\mathbf{U}))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
 H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_mathrm{H})_mathrm{H}}_r
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_mathrm{H})$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_B$ and $L_mathrm{H}$ are the preparation functions for a block and block header respectively.
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
 Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
-\quad L_mathrm{H}(H) & \equiv & (\begin{array}[t]{l}H_mathrm{p}, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_mathrm{l}, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_mathrm{H}(B_mathrm{H}), L_T^*(B_\mathbf{T}), L_mathrm{H}^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_mathrm{H}^*$ being element-wise sequence transformations, thus:
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_mathrm{p} \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
 H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
 H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_mathrm{l} \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
 H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
@@
 
-We define $P(B_mathrm{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_mathrm{H})) = H_mathrm{p}
+We define $P(B_H)$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_mathrm{H}}_i} + 1
+H_i \equiv {{P(H)_H}_i} + 1
 \end{equation}
@@
 \mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_mathrm{H}}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_mathrm{H}}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_mathrm{H}}_s}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 
-The canonical gas limit $H_mathrm{l}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
-& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
-& & H_mathrm{l} \geqslant 5000
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 5000
 \end{eqnarray}
@@
 \begin{equation}
-H_s > {P(H)_mathrm{H}}_s
+H_s > {P(H)_H}_s
 \end{equation}
@@
 & & H_d = D(H) \quad \wedge \\
-& & H_g \le H_mathrm{l}  \quad \wedge \\
-& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
-& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
-& & H_mathrm{l} \geqslant 5000  \quad \wedge \\
-& & H_s > {P(H)_mathrm{H}}_s \quad \wedge \\
-& & H_i = {P(H)_mathrm{H}}_i +1 \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l \geqslant 5000  \quad \wedge \\
+& & H_s > {P(H)_H}_s \quad \wedge \\
+& & H_i = {P(H)_H}_i +1 \quad \wedge \\
 & & \lVert H_x \rVert \le 32
@@
 \begin{equation}
-v_0 \equiv T_g T_mathrm{p} + T_v
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_mathrm{H}}_mathrm{l} - \ell(B_\mathbf{R})_u
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_mathrm{H}}_mathrm{l}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_mathrm{p}$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
 The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_mathrm{p} \\
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
 \boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
@@
 
-Evaluating $\boldsymbol{\sigma}_mathrm{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_mathrm{P}$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_mathrm{P}, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_mathrm{p}, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_mathrm{p}, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_mathrm{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_mathrm{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_mathrm{P}[S(T)]_b + g^* T_mathrm{p} \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_mathrm{P}[m]_b + (T_g - g^*) T_mathrm{p} \\
-m & \equiv & {B_mathrm{H}}_c
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
 \end{eqnarray}
@@
 \begin{eqnarray}
-I_mathrm{a} & \equiv & a \\
+I_a & \equiv & a \\
 I_o & \equiv & o \\
-I_mathrm{p} & \equiv & p \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_mathrm{H}$ has no special treatment and is determined from the blockchain.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_mathrm{a} & \equiv & r \\
+I_a & \equiv & r \\
 I_o & \equiv & o \\
-I_mathrm{p} & \equiv & p \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
@@
 \begin{itemize}
-\item $I_mathrm{a}$, the address of the account which owns the code that is executing.
+\item $I_a$, the address of the account which owns the code that is executing.
 \item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_mathrm{p}$, the price of gas in the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
 \item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
@@
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_mathrm{H}$, the block header of the present block.
+\item $I_H$, the block header of the present block.
 \item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
@@
 B_t & \equiv & B'_t + B_d \\
-B' & \equiv & P(B_mathrm{H})
+B' & \equiv & P(B_H)
 \end{eqnarray}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_mathrm{H})_mathrm{H}, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_mathrm{H}, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_mathrm{H}}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_mathrm{H}}_g = \ell(\mathbf{R})_u
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_mathrm{H}}_c]_b & = & \boldsymbol{\sigma}[{B_mathrm{H}}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
 \qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_mathrm{H}}_i)) R_b
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_mathrm{H}) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_mathrm{H})_mathrm{H}}_r & \text{otherwise}
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_mathrm{l}(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 
-Thus we finish by formally defining $R_mathrm{l}$:
-\begin{eqnarray}
-R_mathrm{l}(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_mathrm{p}$.
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
 
@@
 \begin{equation}
-C_1\equiv\{(X,Y)\in F_mathrm{p}\times F_mathrm{p}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 
-Let $F_{p^2}$ be a field $F_mathrm{p}[i]/(i+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_mathrm{p}$.
-\begin{equation}
-\delta_mathrm{p}(\mathbf x)\equiv\begin{cases}
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
 \mathbf x&\text{if}\ \mathbf x<p\\
@@
 \end{cases}\\
-x&\equiv&\delta_mathrm{p}(\mathbf x[0..31])\\
-y&\equiv&\delta_mathrm{p}(\mathbf x[32..63])
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
 \end{eqnarray}
@@
 \end{cases}\\
-x_0&\equiv&\delta_mathrm{p}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_mathrm{p}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_mathrm{p}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_mathrm{p}(\mathbf x[96..127])
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
 \end{eqnarray}
@@
 L_S(T) & \equiv & \begin{cases}
-(T_n, T_mathrm{p}, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_mathrm{p}, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{a}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{p}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_mathrm{p}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{l}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_mathrm{a}, I_o, L(\boldsymbol{\mu}_g), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_mathrm{a}]_n = \boldsymbol{\sigma}[I_mathrm{a}]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_mathrm{a}]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_mathrm{a}, \boldsymbol{\sigma}[I_mathrm{a}]_n)$, the address of the newly \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
 &&&& created account. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_mathrm{a}, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_mathrm{a}]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{a}, I_o, I_mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_mathrm{a}$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_mathrm{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_mathrm{p}, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_mathrm{a} \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_mathrm{a}]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_mathrm{a}]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_mathrm{a} \\
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
 (\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
@@
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_mathrm{a} \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_mathrm{a}]_b \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_mathrm{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
 \end{equation}

### 2018-01-25 Paper.tex
Add a line break before a tx address so that it doesn't run over the margin.

@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-
-% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
-Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
-If you force a URL to break over a page, the error still occurs.
-
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
-% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
-%\sloppy
-%\setlength{\emergencystretch}{2pt}
-% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
-% https://tex.stackexchange.com/a/59131/143781
-% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-\usepackage{glossaries}
 
@@
 
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction 
+\texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-25 Paper.tex
Remove parentheses surrounding the revision number in the title so that the text doesn't clash with the page number

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
 \author{
@@
 
-At block 2675119, in the transaction 
-\texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-25 Paper.tex

@@
 
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+% Put before hyperref
+
 \input{Version.tex}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
 \author{

### 2018-01-25 Paper.tex

@@
 \end{align}
+
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
 

### 2018-01-25 Paper.tex

@@
 
-% I will insert other table packages in here after this is merged.
+\usepackage{tabu} %requires array.
 

### 2018-01-26 Paper.tex

--- 2018-01-25 Paper.tex
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_mathrm{t}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_mathrm{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_mathrm{n}$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 \item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_mathrm{n}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_mathrm{b}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_mathrm{s}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_mathrm{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_mathrm{c}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
-\begin{equation}
-L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_mathrm{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_mathrm{s}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_mathrm{I}^*$, is defined as the element-wise transformation of the base function $L_mathrm{I}$, given as:
+\begin{equation}
+L_mathrm{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
-\begin{equation}
-L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_mathrm{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_mathrm{S}$:
+\begin{equation}
+L_mathrm{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
-\end{equation}
-
-This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_mathrm{n}, \boldsymbol{\sigma}[a]_mathrm{b}, \boldsymbol{\sigma}[a]_mathrm{s}, \boldsymbol{\sigma}[a]_mathrm{c}) \big) \big)
+\end{equation}
+
+This function, $L_mathrm{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
 \begin{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\quad v(x) \equiv x_mathrm{n} \in \mathbb{P}_{256} \wedge x_mathrm{b} \in \mathbb{P}_{256} \wedge x_mathrm{s} \in \mathbb{B}_{32} \wedge x_mathrm{c} \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_mathrm{c} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_mathrm{n} = 0 \wedge \boldsymbol{\sigma}[a]_mathrm{b} = 0
 \end{equation}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_mathrm{n}$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_mathrm{p}$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_mathrm{g}$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_mathrm{t}$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_mathrm{v}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_mathrm{w}$, $T_mathrm{r}$ and $T_mathrm{s}$.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+L_mathrm{T}(T) \equiv \begin{cases}
+(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, T_\mathbf{i}, T_mathrm{w}, T_mathrm{r}, T_mathrm{s}) & \text{if} \; T_mathrm{t} = \varnothing\\
+(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, T_\mathbf{d}, T_mathrm{w}, T_mathrm{r}, T_mathrm{s}) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_mathrm{n} \in \mathbb{P}_{256} & \wedge & T_mathrm{v} \in \mathbb{P}_{256} & \wedge & T_mathrm{p} \in \mathbb{P}_{256} & \wedge \\
+T_mathrm{g} \in \mathbb{P}_{256} & \wedge & T_mathrm{w} \in \mathbb{P}_5 & \wedge & T_mathrm{r} \in \mathbb{P}_{256} & \wedge \\
+T_mathrm{s} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{P}_mathrm{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_mathrm{t} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_mathrm{neutral}_mathrm{language}\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_mathrm{p}$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_mathrm{o}$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_mathrm{c}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_mathrm{r}$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_mathrm{t}$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_mathrm{e}$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_mathrm{b}$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_mathrm{d}$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_mathrm{i}$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_mathrm{l}$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_mathrm{g}$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_mathrm{s}$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_mathrm{x}$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_mathrm{m}$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_mathrm{n}$.
 \end{description}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_mathrm{e}$.
+
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_mathrm{u}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_mathrm{b}$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_mathrm{u}, R_mathrm{b}, R_\mathbf{l}, R_{s'})
+\end{equation}
+
+The function $L_mathrm{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_mathrm{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_mathrm{u}, R_mathrm{b}, R_\mathbf{l})
 \end{equation}
@@
 
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+We assert $R_mathrm{u}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_mathrm{b}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_mathrm{u} \in \mathbb{P} \quad \wedge \quad R_mathrm{b} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_mathrm{B}$; it is merely a convenience equivalence.
 
 The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+A log entry, $O$, is a tuple of a logger's address, $O_mathrm{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_mathrm{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_mathrm{j}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_mathrm{r}$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+H_mathrm{r} &\equiv& \mathtt{\small TRIE}(L_mathrm{S}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_mathrm{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}^*(B_\mathbf{U}))) & \wedge \\
+H_mathrm{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_mathrm{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+H_mathrm{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_mathrm{R}(B_\mathbf{R}[i]))\}) & \wedge \\
+H_mathrm{b} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_mathrm{b} \big)
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma})) = {P(B_mathrm{H})_mathrm{H}}_mathrm{r}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_mathrm{H})$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+The function $L_mathrm{B}$ and $L_mathrm{H}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_mathrm{R}$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_mathrm{H}(H) & \equiv & (\begin{array}[t]{l}H_mathrm{p}, H_mathrm{o}, H_mathrm{c}, H_mathrm{r}, H_mathrm{t}, H_mathrm{e}, H_mathrm{b}, H_mathrm{d},\\ H_mathrm{i}, H_mathrm{l}, H_mathrm{g}, H_mathrm{s}, H_mathrm{x}, H_mathrm{m}, H_mathrm{n} \; )\end{array} \\
+\quad L_mathrm{B}(B) & \equiv & \big( L_mathrm{H}(B_mathrm{H}), L_mathrm{T}^*(B_\mathbf{T}), L_mathrm{H}^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_mathrm{T}^*$ and $L_mathrm{H}^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+H_mathrm{p} \in \mathbb{B}_{32} & \wedge & H_mathrm{o} \in \mathbb{B}_{32} & \wedge & H_mathrm{c} \in \mathbb{B}_{20} & \wedge \\
+H_mathrm{r} \in \mathbb{B}_{32} & \wedge & H_mathrm{t} \in \mathbb{B}_{32} & \wedge & H_mathrm{e} \in \mathbb{B}_{32} & \wedge \\
+H_mathrm{b} \in \mathbb{B}_{256} & \wedge & H_mathrm{d} \in \mathbb{P} & \wedge & H_mathrm{i} \in \mathbb{P} & \wedge \\
+H_mathrm{l} \in \mathbb{P} & \wedge & H_mathrm{g} \in \mathbb{P} & \wedge & H_mathrm{s} \in \mathbb{P}_{256} & \wedge \\
+H_mathrm{x} \in \mathbb{B} & \wedge & H_mathrm{m} \in \mathbb{B}_{32} & \wedge & H_mathrm{n} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_mathrm{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+H_mathrm{i} \equiv {{P(H)_mathrm{H}}_mathrm{i}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_mathrm{i} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_mathrm{H}}_mathrm{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{d}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_mathrm{s} - {P(H)_mathrm{H}}_mathrm{s}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_mathrm{i} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_mathrm{i} &\equiv \max(H_mathrm{i} - 3000000, 0)
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_mathrm{i}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+The canonical gas limit $H_mathrm{l}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & H_mathrm{l} \geqslant 5000
+\end{eqnarray}
+
+$H_mathrm{s}$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_mathrm{s} > {P(H)_mathrm{H}}_mathrm{s}
 \end{equation}
@@
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+The nonce, $H_mathrm{n}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_mathrm{d}} \quad \wedge \quad m = H_mathrm{m}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d})$.
 
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_mathrm{d}$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 5000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_mathrm{d}} \wedge m = H_mathrm{m} \quad \wedge \\
+& & H_mathrm{d} = D(H) \quad \wedge \\
+& & H_mathrm{g} \le H_mathrm{l}  \quad \wedge \\
+& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
+& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
+& & H_mathrm{l} \geqslant 5000  \quad \wedge \\
+& & H_mathrm{s} > {P(H)_mathrm{H}}_mathrm{s} \quad \wedge \\
+& & H_mathrm{i} = {P(H)_mathrm{H}}_mathrm{i} +1 \quad \wedge \\
+& & \lVert H_mathrm{x} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_mathrm{r})
 \end{equation}
@@
 $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Finally there is $A_mathrm{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_mathrm{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv T_mathrm{g} T_mathrm{p} + T_mathrm{v}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+T_mathrm{n} & = & \boldsymbol{\sigma}[S(T)]_mathrm{n} \quad \wedge \\
+g_0 & \leqslant & T_mathrm{g} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_mathrm{b} \quad \wedge \\
+T_mathrm{g} & \leqslant & {B_mathrm{H}}_mathrm{l} - \ell(B_\mathbf{R})_mathrm{u}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_mathrm{g}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_mathrm{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_mathrm{H}}_mathrm{l}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_mathrm{gT}_mathrm{p}$.
+The gas available for the proceeding computation, $g$, is defined as $T_mathrm{g} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_mathrm{b} & \equiv & \boldsymbol{\sigma}[S(T)]_mathrm{b} - T_mathrm{g} T_mathrm{p} \\
+\boldsymbol{\sigma}_0[S(T)]_mathrm{n} & \equiv & \boldsymbol{\sigma}[S(T)]_mathrm{n} + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_mathrm{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_mathrm{P}$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_mathrm{P}, g', A, s') \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_mathrm{o}, &\\ \quad\quad g, T_mathrm{p}, T_mathrm{v}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_mathrm{t} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_mathrm{o}, &\\ \quad\quad T_mathrm{t}, T_mathrm{t}, g, T_mathrm{p}, T_mathrm{v}, T_mathrm{v}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_g - g_0
-\end{equation}
-and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_mathrm{g} - g_0
+\end{equation}
+and $T_mathrm{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_mathrm{g} - g'}{2} \Big\rfloor, A_mathrm{r} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_mathrm{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_mathrm{g} - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_mathrm{P}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_mathrm{P} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_mathrm{b} & \equiv & \boldsymbol{\sigma}_mathrm{P}[S(T)]_mathrm{b} + g^* T_mathrm{p} \\
+\boldsymbol{\sigma}^*[m]_mathrm{b} & \equiv & \boldsymbol{\sigma}_mathrm{P}[m]_mathrm{b} + (T_mathrm{g} - g^*) T_mathrm{p} \\
+m & \equiv & {B_mathrm{H}}_mathrm{c}
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_mathrm{g} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_mathrm{n} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_mathrm{n}, \boldsymbol{\sigma}[s]_mathrm{b} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_mathrm{c})
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_b & \text{otherwise}
+\boldsymbol{\sigma}[a]_mathrm{b} & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_mathrm{a} & \equiv & a \\
+I_mathrm{o} & \equiv & o \\
+I_mathrm{p} & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
+I_mathrm{s} & \equiv & s \\
+I_mathrm{v} & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+I_mathrm{e} & \equiv & e \\
+I_mathrm{w} & \equiv & w
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_mathrm{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_mathrm{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
+\boldsymbol{\sigma}_1[r]_mathrm{b} \equiv \boldsymbol{\sigma}[r]_mathrm{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_mathrm{b} \equiv \boldsymbol{\sigma}[s]_mathrm{b} - v
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_mathrm{n}, \boldsymbol{\sigma}_1'[s]_mathrm{b} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_mathrm{c})
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_mathrm{n}, \boldsymbol{\sigma}[r]_mathrm{b} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c})
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_mathrm{c}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{BN\_mathrm{ADD}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_mathrm{MUL}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
 \Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_mathrm{a} & \equiv & r \\
+I_mathrm{o} & \equiv & o \\
+I_mathrm{p} & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-I_w & \equiv & w \\
+I_mathrm{s} & \equiv & s \\
+I_mathrm{v} & \equiv & \tilde{v} \\
+I_mathrm{e} & \equiv & e \\
+I_mathrm{w} & \equiv & w \\
 \mathbf{t} & \equiv & \{s, r\} \\
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_mathrm{c}
 \end{eqnarray}
@@
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_mathrm{a}$, the address of the account which owns the code that is executing.
+\item $I_mathrm{o}$, the sender address of the transaction that originated this execution.
+\item $I_mathrm{p}$, the price of gas in the transaction that originated this execution.
 \item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_mathrm{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_mathrm{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
+\item $I_mathrm{H}$, the block header of the present block.
+\item $I_mathrm{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_mathrm{w}$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_mathrm{g}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
+\boldsymbol{\mu}_mathrm{g} & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_mathrm{i} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+\boldsymbol{\mu}'_mathrm{g} & \equiv & \boldsymbol{\mu}_mathrm{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_mathrm{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+\boldsymbol{\mu}_mathrm{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_mathrm{w} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_mathrm{w} \quad \vee \\
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_mathrm{w} + \mathbf{\alpha}_mathrm{w} > 1024 \quad \vee \\
+ \neg I_mathrm{w} \wedge W(w, \boldsymbol{\mu})
 \end{array}
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_mathrm{J}(\mathbf{c}, 0)
 \end{equation}
@@
 \begin{equation}
-D_J(\mathbf{c}, i) \equiv \begin{cases}
+D_mathrm{J}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_mathrm{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_mathrm{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
+\Delta & \equiv & \mathbf{\alpha}_mathrm{w} - \mathbf{\delta}_mathrm{w} \\
 \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_mathrm{w}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 \boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\mu}'_mathrm{i} & \equiv & \boldsymbol{\mu}_mathrm{i} \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_t & \equiv & B'_t + B_d \\
-B' & \equiv & P(B_H)
-\end{eqnarray}
-
-As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
+B_mathrm{t} & \equiv & B'_mathrm{t} + B_mathrm{d} \\
+B' & \equiv & P(B_mathrm{H})
+\end{eqnarray}
+
+As such given a block $B$, $B_mathrm{t}$ is its total difficulty, $B'$ is its parent block and $B_mathrm{d}$ is its difficulty.
 
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_mathrm{H}}_mathrm{g}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_mathrm{H}}_mathrm{g} = \ell(\mathbf{R})_mathrm{u}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by $R_mathrm{b}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\boldsymbol{\sigma}'[{B_mathrm{H}}_mathrm{c}]_mathrm{b} & = & \boldsymbol{\sigma}[{B_mathrm{H}}_mathrm{c}]_mathrm{b} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_mathrm{b} \\
 \qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
+\boldsymbol{\sigma}'[U_mathrm{c}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_mathrm{c}] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_mathrm{c}]_mathrm{n}, \boldsymbol{\sigma}[U_mathrm{c}]_mathrm{b} + R, \boldsymbol{\sigma}[U_mathrm{c}]_\mathbf{s}, \boldsymbol{\sigma}[U_mathrm{c}]_mathrm{c}) \\
+R & \equiv & (1 + \frac{1}{8} (U_mathrm{i} - {B_mathrm{H}}_mathrm{i})) R_mathrm{b}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_b = 3 \times 10^{18}
+\text{Let} \quad R_mathrm{b} = 3 \times 10^{18}
 \end{equation}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_mathrm{H}) = \varnothing \\
+\boldsymbol{\sigma}_mathrm{i}: \mathtt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma}_mathrm{i})) = {P(B_mathrm{H})_mathrm{H}}_mathrm{r} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma}_mathrm{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_mathrm{i}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+B'_mathrm{n} & = & n: \quad x \leqslant \frac{2^{256}}{H_mathrm{d}} \\
+B'_mathrm{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_mathrm{r} = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_mathrm{u}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_mathrm{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+In the case of $B_\mathbf{R}[n]_mathrm{u}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_mathrm{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_mathrm{u}
 \end{array}
@@
 \begin{equation}
-m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+m = H_mathrm{m} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_mathrm{d}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_mathrm{n}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_mathrm{d}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_mathrm{b}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_mathrm{l}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+Formally, we define $R_mathrm{b}$:
+\begin{eqnarray}
+R_mathrm{b}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_mathrm{n} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_mathrm{x}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+\forall_mathrm{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_mathrm{n}\rVert} \quad \mathbf{k}_mathrm{n}'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_mathrm{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_mathrm{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_mathrm{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_mathrm{r} \\
+(\boldsymbol\sigma, g - g_mathrm{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_mathrm{r}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+g_mathrm{r} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_mathrm{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_mathrm{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_mathrm{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+g_mathrm{r} &=& \Big\lfloor\frac{f\big(\max(\ell_mathrm{M},\ell_mathrm{B})\big)\max(\ell'_mathrm{E},1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
+\ell'_mathrm{E} &=& \begin{cases}
+0 & \text{if}\ \ell_mathrm{E}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_mathrm{E}\le 32 \wedge E \neq 0 \\
+8(\ell_mathrm{E} - 32) + \lfloor \log_2(i[(96+\ell_mathrm{B})..(127+\ell_mathrm{B})]) \rfloor & \text{if}\ 32 < \ell_mathrm{E} \wedge i[(96 + \ell_mathrm{B})..(127 + \ell_mathrm{B})]\neq 0 \\
+8(\ell_mathrm{E} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_mathrm{M}} \\
+\ell_mathrm{B} &\equiv& i[0..31] \\
+\ell_mathrm{E} &\equiv& i[32..63] \\
+\ell_mathrm{M} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_mathrm{B})] \\
+E &\equiv& i[(96+\ell_mathrm{B})..(95+\ell_mathrm{B}+\ell_mathrm{E})] \\
+M &\equiv& i[(96+\ell_mathrm{B}+\ell_mathrm{E})..(95+\ell_mathrm{B}+\ell_mathrm{E}+\ell_mathrm{M})] \\
 i[x] &\equiv& \begin{cases}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_mathrm{j}=\varnothing\vee b_mathrm{j}=\varnothing))\\
 k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_r&=& 60000k + 40000 \\
+g_mathrm{r}&=& 60000k + 40000 \\
 \mathbf o[0..31]&\equiv&\begin{cases}
@@
 \end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_mathrm{k})\times\log_{P_2}(b_mathrm{k})\equiv 1\mod q)\\
 a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
@@
 \vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_mathrm{k}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_mathrm{k}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r &=& 500\\
+\Xi_{\mathtt{BN\_mathrm{ADD}}}&\equiv&\Xi_{\mathtt{BN\_mathrm{PRE}}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_mathrm{ADD}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_mathrm{r} &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+\label{eq:complemented_mathrm{input}}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
 I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r &=& 40000\\
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_mathrm{input}}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_mathrm{MUL}}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_mathrm{MUL}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_mathrm{r} &=& 40000\\
 \mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+It is assumed that the sender has a valid private key $p_mathrm{r}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
 
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+\mathtt{\small ECDSAPUBKEY}(p_mathrm{r} \in \mathbb{B}_{32}) & \equiv & p_mathrm{u} \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_mathrm{r} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_mathrm{u} \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_mathrm{u}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_mathrm{r}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+v &\in \{27,28,\mathtt{\tiny chain\_mathrm{id}} \times 2 + 35, \mathtt{\tiny chain\_mathrm{id}} \times 2 + 36\}
 \end{align}
@@
 
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+For a given private key, $p_mathrm{r}$, the Ethereum address $A(p_mathrm{r})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_mathrm{r}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_mathrm{r}) \big) \big)
 \end{equation}
@@
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+One operates without the latter three signature components, formally described as $T_mathrm{r}$, $T_mathrm{s}$ and $T_mathrm{w}$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+L_mathrm{S}(T) & \equiv & \begin{cases}
+(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, \mathbf{p}, \mathtt{\tiny chain\_mathrm{id}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_mathrm{t} = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+h(T) & \equiv & \mathtt{\small KEC}( L_mathrm{S}(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_mathrm{r})$ is defined as:
+\begin{eqnarray}
+G(T, p_mathrm{r}) \equiv T \quad \text{except:} \\
+(T_mathrm{w}, T_mathrm{r}, T_mathrm{s}) = \mathtt{\small ECDSASIGN}(h(T), p_mathrm{r})
 \end{eqnarray}
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_mathrm{r}, T_mathrm{s}) \big) \big) \\
 v_0 &\equiv& \begin{cases}
-T_w &\text{if}\ T_w\in\{27, 28\} \\
-28 - (T_w \bmod 2) &\text{otherwise}
+T_mathrm{w} &\text{if}\ T_mathrm{w}\in\{27, 28\} \\
+28 - (T_mathrm{w} \bmod 2) &\text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
+\forall T: \forall p_mathrm{r}: S(G(T, p_mathrm{r})) \equiv A(p_mathrm{r})
 \end{equation}
@@
 
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+%extern u256 const c_mathrm{copyGas};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_mathrm{i})-C_{mem}(\boldsymbol{\mu}_mathrm{i}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_mathrm{i}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_mathrm{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_mathrm{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} \vee \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{a}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_mathrm{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{o}$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{s}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{v}$ \\
 \midrule
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_mathrm{c} \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_mathrm{c}$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_mathrm{i} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_mathrm{i} \\ P(H_mathrm{p}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{c}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{s}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{i}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{d}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{l}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_mathrm{g}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_mathrm{a}, I_mathrm{o}, L(\boldsymbol{\mu}_mathrm{g}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\; I_mathrm{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_mathrm{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_mathrm{a}]_mathrm{n} = \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{n} + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_mathrm{e} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_mathrm{a}, \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{n})$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_mathrm{a}, I_mathrm{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge \\ \quad\quad I_mathrm{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{g} \equiv \boldsymbol{\mu}_mathrm{g} + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}$ (not enough funds) or $I_mathrm{e} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(\boldsymbol{\mu}_mathrm{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_mathrm{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{a}, I_mathrm{o}, I_mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\\ \quad\quad{}I_mathrm{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_mathrm{a}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
@@
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{s}, I_mathrm{o}, I_mathrm{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_mathrm{p}, 0, I_mathrm{v}, \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \text{if} \quad I_mathrm{v} \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\; I_mathrm{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_mathrm{a} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} = 0\\
+(\boldsymbol{\sigma}[r]_mathrm{n}, \boldsymbol{\sigma}[r]_mathrm{b} + \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c}) & \text{if}\ r \neq I_mathrm{a} \\
+(\boldsymbol{\sigma}[r]_mathrm{n}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c}) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_mathrm{b} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_mathrm{a} \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \neq 0$ \\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+ E_{epoch}(H_mathrm{i}) = \left\lfloor\frac{H_mathrm{i}}{J_{epoch}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_mathrm{i})
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_mathrm{i}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_mathrm{i}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_mathrm{i} - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d}) = \lbrace \mathbf{m}_mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_mathrm{n}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_mathrm{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_mathrm{n}) + \mathbf{m}_mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_mathrm{n}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_mathrm{c}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_mathrm{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}

### 2018-01-26 Paper.tex
Replace all _(\{[:alpha:]+\}) with _mathrm{$1}
using Libre Open Office with regular expressions ticked as an option.

--- 2018-01-25 Paper.tex
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_mathrm{txdatazero} & \text{if} \quad i = 0 \\ G_mathrm{txdatanonzero} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
+{} & + G_mathrm{transaction}
 \end{align}
@@
 \begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
+c \equiv G_mathrm{codedeposit} \times |\mathbf{o}|
 \end{equation}
@@
 \boldsymbol{\mu}_g & \equiv & g \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
+\boldsymbol{\mu}_mathrm{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_mathrm{pc}] & \text{if} \quad \boldsymbol{\mu}_mathrm{pc} < \lVert I_\mathbf{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_mathrm{g} & \equiv & \boldsymbol{\mu}_mathrm{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\quad \boldsymbol{\mu}'_mathrm{pc} & \equiv & \begin{cases}
 J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_mathrm{pc}, w) & \text{otherwise}
 \end{cases}
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_mathrm{cacheinit}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $J_mathrm{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $J_mathrm{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
+
 v & = & \begin{cases}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_mathrm{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_mathrm{n}
 \end{equation}
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_mathrm{zero}$ & 0 & Nothing paid for operations of the set {\small $W_mathrm{zero}$}. \\
+$G_mathrm{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_mathrm{base}$}. \\
+$G_mathrm{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_mathrm{verylow}$}. \\
+$G_mathrm{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_mathrm{low}$}. \\
+$G_mathrm{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_mathrm{mid}$}. \\
+$G_mathrm{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_mathrm{high}$}. \\
+$G_mathrm{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_mathrm{extcode}$}. \\
+$G_mathrm{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_mathrm{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_mathrm{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_mathrm{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_mathrm{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_mathrm{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_mathrm{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_mathrm{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_mathrm{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+$G_mathrm{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_mathrm{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_mathrm{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_mathrm{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_mathrm{callvalue}$ for a non-zero value transfer. \\
+$G_mathrm{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_mathrm{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_mathrm{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_mathrm{memory}$ & 3 & Paid for every additional word when expanding memory. \\
 $G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_mathrm{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_mathrm{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_mathrm{transaction}$ & 21000 & Paid for every transaction. \\
+$G_mathrm{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_mathrm{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_mathrm{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
 $G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+$G_mathrm{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_mathrm{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_mathrm{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_mathrm{mem}(\boldsymbol{\mu}'_i)-C_mathrm{mem}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_mathrm{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_mathrm{exp} + G_mathrm{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_mathrm{verylow} + G_mathrm{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_mathrm{extcode} + G_mathrm{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
+G_mathrm{create} & \text{if} \quad w = \text{\small CREATE}\\
 G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
+G_mathrm{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_mathrm{sload} & \text{if} \quad w = \text{\small SLOAD}\\
+G_mathrm{zero} & \text{if} \quad w \in W_mathrm{zero}\\
+G_mathrm{base} & \text{if} \quad w \in W_mathrm{base}\\
+G_mathrm{verylow} & \text{if} \quad w \in W_mathrm{verylow}\\
+G_mathrm{low} & \text{if} \quad w \in W_mathrm{low}\\
+G_mathrm{mid} & \text{if} \quad w \in W_mathrm{mid}\\
+G_mathrm{high} & \text{if} \quad w \in W_mathrm{high}\\
+G_mathrm{extcode} & \text{if} \quad w \in W_mathrm{extcode}\\
+G_mathrm{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_mathrm{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_mathrm{pc}] & \text{if} \quad \boldsymbol{\mu}_mathrm{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_mathrm{mem}(a) \equiv G_mathrm{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{high}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+$W_mathrm{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_mathrm{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_mathrm{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_mathrm{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_mathrm{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_mathrm{high}$ = \{{\small JUMPI}\}
+
+$W_mathrm{extcode}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_mathrm{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_mathrm{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
+G_mathrm{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_mathrm{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+&&&& $A'_mathrm{r} \equiv A_mathrm{r} + \begin{cases}
+R_mathrm{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_mathrm{pc}$. See section \ref{ch:model}. \\
 \midrule
 0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_mathrm{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_mathrm{pc}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_mathrm{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_mathrm{i}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_mathrm{g}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_mathrm{pc} + 1)$ \\
 &&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_mathrm{pc} + 1) \dots (\boldsymbol{\mu}_mathrm{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_mathrm{pc} + 1) \dots (\boldsymbol{\mu}_mathrm{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_mathrm{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_mathrm{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_mathrm{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_mathrm{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+&&&& $A'_mathrm{r} \equiv A_mathrm{r} + \begin{cases}
+R_mathrm{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_mathrm{selfdestruct} + \begin{cases}
+G_mathrm{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+$J_mathrm{wordbytes}$ & 4  & Bytes in word. \\
+$J_mathrm{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_mathrm{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_mathrm{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_mathrm{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_mathrm{epoch}$ & 30000 & Blocks per epoch. \\
+$J_mathrm{mixbytes}$ & 128 & mix length in bytes. \\
+$J_mathrm{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_mathrm{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_mathrm{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_mathrm{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+ E_mathrm{epoch}(H_i) = \left\lfloor\frac{H_i}{J_mathrm{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_mathrm{datasetgrowth}$ bytes, and the size of the cache by $J_mathrm{cachegrowth}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of $J_mathrm{mixbytes}$, for the dataset, and $J_mathrm{hashbytes}$ for the cache.
+Let $d_mathrm{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_mathrm{size} = E_mathrm{prime}(J_mathrm{datasetinit} + J_mathrm{datasetgrowth} \cdot E_mathrm{epoch} - J_mathrm{mixbytes}, J_mathrm{mixbytes})
+\end{equation}
+The size of the cache, $c_mathrm{size}$, is calculated using
+\begin{equation}
+ c_mathrm{size} = E_mathrm{prime}(J_mathrm{cacheinit} + J_mathrm{cachegrowth} \cdot E_mathrm{epoch} - J_mathrm{hashbytes}, J_mathrm{hashbytes})
+\end{equation}
+\begin{equation}
+ E_mathrm{prime}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
+E_mathrm{prime}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_mathrm{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_mathrm{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_mathrm{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_mathrm{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_mathrm{seedhash}(H_i - J_mathrm{epoch})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially filling up $c_mathrm{size}$ bytes of memory, then performing $J_mathrm{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
 
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
+We define the array $\mathbf{c}_mathrm{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
+\begin{equation}
+ \mathbf{c}_mathrm{i} = \begin{cases}
 \texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ \mathbf{c'}[i] = \mathbf{c}_mathrm{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_mathrm{size}}{J_mathrm{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_mathrm{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_mathrm{cacherounds}(\mathbf{c'}, J_mathrm{cacherounds})
+\end{equation}
+\begin{equation}
+ E_mathrm{cacherounds}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_mathrm{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_mathrm{rmh}(\mathbf{x}, 0), E_mathrm{rmh}(\mathbf{x}, 1), ... , E_mathrm{rmh}(\mathbf{x}, n - 1) \big)
 \end{equation}
 \begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+  E_mathrm{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_mathrm{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
 \end{multline}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+Essentially, we combine data from $J_mathrm{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_mathrm{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_mathrm{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_mathrm{size}}{J_mathrm{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+ E_mathrm{datasetitem}(\mathbf{c}, i) = E_mathrm{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_mathrm{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_mathrm{parents}(\mathbf{c}, i, p +1, E_mathrm{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_mathrm{parents} -2 \\
+E_mathrm{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_mathrm{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_mathrm{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_mathrm{hashbytes} / J_mathrm{wordbytes} \rfloor]) \mod c_mathrm{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a ``mix'' $J_mathrm{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_mathrm{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_mathrm{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_mathrm{compress}(E_mathrm{accesses}(\mathbf{d}, \sum_{i = 0}^{n_mathrm{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_mathrm{revert}(\mathbf{n}))
+\end{equation}
+$E_mathrm{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_mathrm{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ n_mathrm{mix} =  \left\lfloor\frac{J_mathrm{mixbytes}}{J_mathrm{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_mathrm{accesses}$ function is used:
+\begin{equation}
+ E_mathrm{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_mathrm{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_mathrm{accesses} -2 \\
+E_mathrm{accesses}(E_mathrm{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_mathrm{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_mathrm{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_mathrm{newdata}$ returns an array with $n_mathrm{mix}$ elements:
+\begin{equation}
+ E_mathrm{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_mathrm{mixbytes}}{J_mathrm{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_mathrm{size} / J_mathrm{hashbytes}}{n_mathrm{mix}}\right\rfloor \cdot n_mathrm{mix} + j] \quad \forall \quad j < n_mathrm{mix}
 \end{equation}
@@
 \begin{equation}
- E_{compress}(\mathbf{m}, i) = \begin{cases}
+ E_mathrm{compress}(\mathbf{m}, i) = \begin{cases}
 \mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+E_mathrm{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
 \end{cases}

### 2018-01-26 Paper.tex

@@
 u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
-
 v & = & \begin{cases}

### 2018-01-26 Paper.tex

@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\boldsymbol{\sigma}_mathrm{t}+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_mathrm{t}+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_mathrm{s}'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_mathrm{s}'})
 \end{equation}
@@
 
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
+We assert that the status code $R_mathrm{s}'}$ is an integer.
+\begin{equation}
+R_mathrm{s}'} \in \mathbb{P}
 \end{equation}
@@
 \begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_mathrm{t} \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_mathrm{t} \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
+\forall_mathrm{i} \in \{0, 2, 4\}}&:& \mathcal{B}_mathrm{m}(\mathbf{x}, i)}(\mathbf{y}) = 1\\
 m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_mathrm{txdatazero} & \text{if} \quad i = 0 \\ G_mathrm{txdatanonzero} & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_mathrm{i} \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_mathrm{transaction}
+{} & + G_{transaction}
 \end{align}
@@
 
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_mathrm{a}..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{equation}
-c \equiv G_mathrm{codedeposit} \times |\mathbf{o}|
+c \equiv G_{codedeposit} \times |\mathbf{o}|
 \end{equation}
@@
 \boldsymbol{\mu}_g & \equiv & g \\
-\boldsymbol{\mu}_mathrm{pc} & \equiv & 0 \\
+\boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_mathrm{pc}] & \text{if} \quad \boldsymbol{\mu}_mathrm{pc} < \lVert I_\mathbf{b} \rVert \\
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_mathrm{g} & \equiv & \boldsymbol{\mu}_mathrm{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\quad \boldsymbol{\mu}'_mathrm{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
 J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_mathrm{pc}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
 \end{cases}
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_mathrm{s}'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
+We define $\mathbf{R}[n]_mathrm{s}'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_mathrm{s}'} =
 \Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_mathrm{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $J_mathrm{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_mathrm{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 \mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_mathrm{n} < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_mathrm{n} < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
 \end{eqnarray}
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_mathrm{n} = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+\forall_n \quad \forall_mathrm{i}: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
 \begin{cases}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_mathrm{quaddivisor}}\Big\rfloor \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_mathrm{n}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
 \end{equation}
@@
 
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+Let $F_mathrm{p}^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_mathrm{p}^2}\times F_mathrm{p}^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 \midrule
-$G_mathrm{zero}$ & 0 & Nothing paid for operations of the set {\small $W_mathrm{zero}$}. \\
-$G_mathrm{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_mathrm{base}$}. \\
-$G_mathrm{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_mathrm{verylow}$}. \\
-$G_mathrm{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_mathrm{low}$}. \\
-$G_mathrm{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_mathrm{mid}$}. \\
-$G_mathrm{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_mathrm{high}$}. \\
-$G_mathrm{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_mathrm{extcode}$}. \\
-$G_mathrm{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_mathrm{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_mathrm{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_mathrm{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_mathrm{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_mathrm{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_mathrm{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-$G_mathrm{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_mathrm{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-$G_mathrm{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_mathrm{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_mathrm{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_mathrm{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_mathrm{callvalue}$ for a non-zero value transfer. \\
-$G_mathrm{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_mathrm{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_mathrm{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_mathrm{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
+$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
+$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
 $G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-$G_mathrm{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_mathrm{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_mathrm{transaction}$ & 21000 & Paid for every transaction. \\
-$G_mathrm{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_mathrm{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_mathrm{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
 $G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_mathrm{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_mathrm{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_mathrm{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_mathrm{mem}(\boldsymbol{\mu}'_i)-C_mathrm{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_mathrm{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_mathrm{exp} + G_mathrm{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_mathrm{verylow} + G_mathrm{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_mathrm{extcode} + G_mathrm{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_mathrm{log}+G_mathrm{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_mathrm{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_mathrm{create} & \text{if} \quad w = \text{\small CREATE}\\
+G_{create} & \text{if} \quad w = \text{\small CREATE}\\
 G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_mathrm{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_mathrm{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_mathrm{zero} & \text{if} \quad w \in W_mathrm{zero}\\
-G_mathrm{base} & \text{if} \quad w \in W_mathrm{base}\\
-G_mathrm{verylow} & \text{if} \quad w \in W_mathrm{verylow}\\
-G_mathrm{low} & \text{if} \quad w \in W_mathrm{low}\\
-G_mathrm{mid} & \text{if} \quad w \in W_mathrm{mid}\\
-G_mathrm{high} & \text{if} \quad w \in W_mathrm{high}\\
-G_mathrm{extcode} & \text{if} \quad w \in W_mathrm{extcode}\\
-G_mathrm{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_mathrm{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_mathrm{pc}] & \text{if} \quad \boldsymbol{\mu}_mathrm{pc} < \lVert I_\mathbf{b} \rVert\\
+G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{zero} & \text{if} \quad w \in W_{zero}\\
+G_{base} & \text{if} \quad w \in W_{base}\\
+G_{verylow} & \text{if} \quad w \in W_{verylow}\\
+G_{low} & \text{if} \quad w \in W_{low}\\
+G_{mid} & \text{if} \quad w \in W_{mid}\\
+G_{high} & \text{if} \quad w \in W_{high}\\
+G_{extcode} & \text{if} \quad w \in W_{extcode}\\
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{equation}
-C_mathrm{mem}(a) \equiv G_mathrm{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}
@@
 
-$W_mathrm{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_mathrm{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_mathrm{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_mathrm{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_mathrm{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_mathrm{high}$ = \{{\small JUMPI}\}
-
-$W_mathrm{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_mathrm{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{high}$ = \{{\small JUMPI}\}
+
+$W_{extcode}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 
-Note also that $C_mathrm{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_mathrm{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
-G_mathrm{sreset} & \text{otherwise}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&& $A'_mathrm{r} \equiv A_mathrm{r} + \begin{cases}
-R_mathrm{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_mathrm{pc}$. See section \ref{ch:model}. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
 0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_mathrm{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_mathrm{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_mathrm{pc}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_mathrm{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_mathrm{g}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_mathrm{pc} + 1)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
 &&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_mathrm{pc} + 1) \dots (\boldsymbol{\mu}_mathrm{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_mathrm{pc} + 1) \dots (\boldsymbol{\mu}_mathrm{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_mathrm{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_mathrm{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_mathrm{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_mathrm{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
-&&&& $A'_mathrm{r} \equiv A_mathrm{r} + \begin{cases}
-R_mathrm{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_mathrm{selfdestruct} + \begin{cases}
-G_mathrm{newaccount} & \text{if} \quad n \\
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
@@
 \midrule
-$J_mathrm{wordbytes}$ & 4  & Bytes in word. \\
-$J_mathrm{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_mathrm{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_mathrm{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_mathrm{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_mathrm{epoch}$ & 30000 & Blocks per epoch. \\
-$J_mathrm{mixbytes}$ & 128 & mix length in bytes. \\
-$J_mathrm{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_mathrm{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_mathrm{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_mathrm{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 \begin{equation}
- E_mathrm{epoch}(H_i) = \left\lfloor\frac{H_i}{J_mathrm{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_mathrm{datasetgrowth}$ bytes, and the size of the cache by $J_mathrm{cachegrowth}$ bytes, every epoch.
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_mathrm{mixbytes}$, for the dataset, and $J_mathrm{hashbytes}$ for the cache.
-Let $d_mathrm{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_mathrm{size} = E_mathrm{prime}(J_mathrm{datasetinit} + J_mathrm{datasetgrowth} \cdot E_mathrm{epoch} - J_mathrm{mixbytes}, J_mathrm{mixbytes})
-\end{equation}
-The size of the cache, $c_mathrm{size}$, is calculated using
-\begin{equation}
- c_mathrm{size} = E_mathrm{prime}(J_mathrm{cacheinit} + J_mathrm{cachegrowth} \cdot E_mathrm{epoch} - J_mathrm{hashbytes}, J_mathrm{hashbytes})
-\end{equation}
-\begin{equation}
- E_mathrm{prime}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_mathrm{prime}((x - 1) \cdot y, y) & \text{otherwise}
+E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_mathrm{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- \mathbf{s} = C_mathrm{seedhash}(H_i)
-\end{equation}
-\begin{equation}
- C_mathrm{seedhash}(H_i) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_mathrm{epoch}(H_i) = 0 \quad  \\
-\texttt{KEC}(C_mathrm{seedhash}(H_i - J_mathrm{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_mathrm{size}$ bytes of memory, then performing $J_mathrm{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
 
-We define the array $\mathbf{c}_mathrm{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_mathrm{i} = \begin{cases}
+We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
+\begin{equation}
+ \mathbf{c}_{i} = \begin{cases}
 \texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
-\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
+\texttt{KEC512}(\mathbf{c}_mathrm{i}-1}) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_mathrm{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_mathrm{size}}{J_mathrm{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_mathrm{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_mathrm{cacherounds}(\mathbf{c'}, J_mathrm{cacherounds})
-\end{equation}
-\begin{equation}
- E_mathrm{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_mathrm{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_mathrm{rmh}(\mathbf{x}, 0), E_mathrm{rmh}(\mathbf{x}, 1), ... , E_mathrm{rmh}(\mathbf{x}, n - 1) \big)
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
 \end{equation}
 \begin{multline}
-  E_mathrm{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_mathrm{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
 \end{multline}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_mathrm{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_mathrm{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_mathrm{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_mathrm{size}}{J_mathrm{hashbytes}}\right\rfloor
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_mathrm{datasetitem}(\mathbf{c}, i) = E_mathrm{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_mathrm{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_mathrm{parents}(\mathbf{c}, i, p +1, E_mathrm{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_mathrm{parents} -2 \\
-E_mathrm{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_mathrm{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_mathrm{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_mathrm{hashbytes} / J_mathrm{wordbytes} \rfloor]) \mod c_mathrm{size}] \big) & \text{otherwise}
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_mathrm{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_mathrm{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_mathrm{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_mathrm{compress}(E_mathrm{accesses}(\mathbf{d}, \sum_{i = 0}^{n_mathrm{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_mathrm{i} = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_mathrm{revert}(\mathbf{n}))
-\end{equation}
-$E_mathrm{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_mathrm{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_mathrm{mix} =  \left\lfloor\frac{J_mathrm{mixbytes}}{J_mathrm{hashbytes}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_mathrm{accesses}$ function is used:
-\begin{equation}
- E_mathrm{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_mathrm{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_mathrm{accesses} -2 \\
-E_mathrm{accesses}(E_mathrm{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_mathrm{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_mathrm{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_mathrm{newdata}$ returns an array with $n_mathrm{mix}$ elements:
-\begin{equation}
- E_mathrm{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_mathrm{mixbytes}}{J_mathrm{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_mathrm{size} / J_mathrm{hashbytes}}{n_mathrm{mix}}\right\rfloor \cdot n_mathrm{mix} + j] \quad \forall \quad j < n_mathrm{mix}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{newdata}$ returns an array with $n_{mix}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}
@@
 \begin{equation}
- E_mathrm{compress}(\mathbf{m}, i) = \begin{cases}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
 \mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_mathrm{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
 \end{cases}

### 2018-01-26 Paper.tex
Replace all _(\{[A-Z]+\b)([^\}]) with _mathbf$1}$2...
...using Libre Open Office with match case and regular expressions ticked as options.

@@
 
-\usepackage{tabu} %requires array.
+% I will insert other table packages in here after this is merged.
 
@@
 
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-% Put before hyperref
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+\usepackage{glossaries}
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_mathrm{t}+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_mathrm{t}+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_mathrm{s}'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_mathrm{s}'})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
 \end{equation}
@@
 
-We assert that the status code $R_mathrm{s}'}$ is an integer.
-\begin{equation}
-R_mathrm{s}'} \in \mathbb{P}
+We assert that the status code $R_{s'}$ is an integer.
+\begin{equation}
+R_{s'} \in \mathbb{P}
 \end{equation}
@@
 \begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_mathrm{t} \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_mathrm{t} \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall_mathrm{i} \in \{0, 2, 4\}}&:& \mathcal{B}_mathrm{m}(\mathbf{x}, i)}(\mathbf{y}) = 1\\
+\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
 m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
@@
 \end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
 
@@
 \begin{align}
-g_0 \equiv {} & \sum_mathrm{i} \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
@@
 
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_mathrm{a}..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_mathbf{U} \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\qquad\forall_mathbf{U} \in B_\mathbf{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 
-As previously detailed, $\mathbf{R}[n]_mathrm{s}'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_mathrm{s}'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_mathrm{s}'} =
+We define $\mathbf{R}[n]_{s'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{s'} =
 \Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
@@
 \mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_mathrm{n} < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_mathrm{n} < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
 \end{eqnarray}
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_mathrm{n} = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 \begin{equation}
-\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_n \quad \forall_mathrm{i}: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
+\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
 \begin{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_mathbf{I} \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 
-Let $F_mathrm{p}^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_mathrm{p}^2}\times F_mathrm{p}^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_mathrm{i} \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
 \begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 \texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
-\texttt{KEC512}(\mathbf{c}_mathrm{i}-1}) & \text{otherwise}
+\texttt{KEC512}(\mathbf{c}_{i-1}) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+Essentially, we maintain a ``mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
 $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
@@
 \begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_mathrm{i} = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}

### 2018-01-26 Paper.tex
Replace all _(\{[A-Z]+\b)([^\}]) with _mathbf$1}$2...
...using Libre Open Office with match case and regular expressions ticked as options.

@@
 
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-
-% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
-Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
-If you force a URL to break over a page, the error still occurs.
-
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
-% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
-%\sloppy
-%\setlength{\emergencystretch}{2pt}
-% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
-% https://tex.stackexchange.com/a/59131/143781
-% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-\usepackage{glossaries}
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
+% Put before hyperref
 

### 2018-01-26 Paper.tex
Add commands to manage URL breaks + glossaries package
Blocks function as a ledger

@@
 Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
+Blocks function as a ledger, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
 They also punctuate the transaction series with incentives for nodes to \textit{mine}.
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_mathbf{U} \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_mathbf{U} \in B_\mathbf{U}}: \\ \nonumber
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 \begin{equation}
-\forall_mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_mathbf{I} \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\

### 2018-01-26 Paper.tex
Replace all _(\{[A-Z]+\b)([^\}]) with _{mathbf$1}$2...
...using Libre Open Office with match case and regular expressions ticked as options.

@@
 Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a ledger, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
+Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
 They also punctuate the transaction series with incentives for nodes to \textit{mine}.

### 2018-01-26 Paper.tex
\PassOptionsToPackage{tabularx}\usepackage{tabu}
Comment out \usepackage{glossaries}

--- 2018-01-26 Paper.tex
@@
 
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+%\usepackage{glossaries}
+
 \input{Version.tex}

### 2018-01-26 Paper.tex
\begin{tabu}{X[l]X[c]X} \savetabu{col3}
Move \hypertarget{bigvee}{} to preceding line
Remove parentheses

--- 2018-01-26 Paper.tex
@@
 \section{List of mathematical symbols}\label{app:symbols}
-\begin{tabularx}{\textwidth}{lcX}
+\begin{tabu}{X[l]X[c]X} \savetabu{col3}
 \toprule
-Symbol & Latex Command & Description \\
-\midrule
-\hypertarget{bigvee}{}
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
 \bottomrule
-\end{tabularx}
+\end{tabu}
 

### 2018-01-26 Paper.tex
tabu*
\hyperlink{GenesisBlock}{the genesis block}

--- 2018-01-26 Paper.tex
@@
 \toprule
-Name & Value & Description \\
+Name & Value & Description \hypertarget{Jwordbytes}{}\\
 \midrule

### 2018-01-26 Paper.tex
\end{tabu*}
\usepackage[ocgcolorlinks]{ocgx2}

--- 2018-01-26 Paper.tex
@@
 %,pagebackref=true
-
+\usepackage[ocgcolorlinks]{ocgx2} %For getting links to line and page break.
+https://tex.stackexchange.com/a/409479/143781https://tex.stackexchange.com/a/409479/143781https://tex.stackexchange.com/a/409479/143781
 \usepackage{tabu} %requires array.
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
-\bottomrule
-\end{tabu}
-
 \end{document}

### 2018-01-26 Paper.tex
Add \nop command to modify hypertargets, and replace all \\hypertarget(\{.*\})\{\} with \nop$1

--- 2018-01-26 Paper.tex
@@
 
+\makeatletter
+ \newcommand{\nop}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+
 \input{Version.tex}
@@
 A valid state transition is one which comes about through a transaction.
-Formally:
+Formally: \nop{Upsilon_state_transition}
 \begin{equation}
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (\hyperlink{GenesisBlock}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. 
+op{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 The account state comprises the following four fields:
+op{account_nonce}
 
@@
 %The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-\hypertarget{GenesisBlock}{}
+
 \section{Genesis Block}\label{app:genesis}
@@
 \toprule
-Name & Value & Description \hypertarget{Jwordbytes}{}\\
+Name & Value & Description \\
 \midrule

### 2018-01-26 Paper.tex
Change linebreak op{ to \nop, and move \nop

@@
 A valid state transition is one which comes about through a transaction.
-Formally: \nop{Upsilon_state_transition}
-\begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+Formally:
+\begin{equation}
+\nop{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 
-I define a number of useful functions throughout. 
-op{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \nop{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 The account state comprises the following four fields:
-op{account_nonce}
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+\item[nonce] \nop{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.

### 2018-01-26 Paper.tex
Add line breaks which removes many empty lines

--- 2018-01-26 Paper.tex
@@
 %,pagebackref=true
-\usepackage[ocgcolorlinks]{ocgx2} %For getting links to line and page break.
-https://tex.stackexchange.com/a/409479/143781https://tex.stackexchange.com/a/409479/143781https://tex.stackexchange.com/a/409479/143781
+
 \usepackage{tabu} %requires array.
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which g last item in the given sequence:
 
@@
 \end{equation}
+
 where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+
 \begin{eqnarray}

### 2018-01-26 Paper.tex
Fix accident
Remove unnecessary comment
Given that these commands are after the hyperref package and haven't caused an issue, clearly this comment (to put them before) can be removed.

@@
 \MakeOuterQuote{"}
-% Put before hyperref
 
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which g last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \end{equation}
-
 where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
-
 \begin{eqnarray}

### 2018-01-26 Paper.tex

@@
 
+\makeatletter
+ \newcommand{\nop}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
 % For formatting
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\nop{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 
-I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \nop{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+\item[nonce] \nop{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.

### 2018-01-26 Paper.tex
Add a \\ as a manual break

--- 2018-01-26 Paper.tex
@@
 
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-
 \input{Version.tex}
@@
 
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction 
+\texttt{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-26 Paper.tex
Replace \nop with \linkdest

--- 2018-01-26 Paper.tex
@@
 
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
 % For formatting

### 2018-01-26 Paper.tex
\usepackage{seqsplit} & \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}

--- 2018-01-26 Paper.tex
@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+
+\usepackage{seqsplit}
 
@@
 At block 2675119, in the transaction 
-\texttt{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-26 Paper.tex
Remove a line break

@@
 
-At block 2675119, in the transaction 
-\seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-26 Paper.tex

--- 2018-01-26 Paper.tex
@@
 
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
 % For formatting
@@
 \MakeOuterQuote{"}
+
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+%\usepackage{glossaries}
 
@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 
-I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.

### 2018-01-26 Paper.tex
Change s' to z for the status code.

@@
 
+\usepackage{seqsplit}
+
 % For formatting
@@
 \MakeOuterQuote{"}
-
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-
-% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
-Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
-If you force a URL to break over a page, the error still occurs.
-
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
-% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
-%\sloppy
-%\setlength{\emergencystretch}{2pt}
-% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter.
-% https://tex.stackexchange.com/a/59131/143781
-% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-%\usepackage{glossaries}
 
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{z}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{z})
 \end{equation}
@@
 
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
+We assert that the status code $R_{z}$ is an integer.
+\begin{equation}
+R_{z} \in \mathbb{P}
 \end{equation}
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, z) \equiv \begin{cases}
 \Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{z}$, the status code of this transaction:
 \begin{eqnarray}
@@
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \end{cases} \\
-\quad s' &\equiv \begin{cases}
+\quad z &\equiv \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
@@
 \end{cases} \\ \nonumber
-s' & \equiv & \begin{cases}
+z & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{z}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_{z}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{z} =
+\Upsilon^{z}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 
-At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-26 Paper.tex
Add missing introductory text for $\Upsilon^s$.

--- 2018-01-26 Paper.tex
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items and $\Upsilon^s$ to evaluate to the status code resulting from the transaction.
+These will be formally defined later.
 

### 2018-01-29 Paper.tex

--- 2018-01-26 Paper.tex
@@
 
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
 \usepackage{seqsplit}

### 2018-01-29 Paper.tex

--- 2018-01-29 Paper.tex
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{z}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{z})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_z)
 \end{equation}
@@
 
-We assert that the status code $R_{z}$ is an integer.
-\begin{equation}
-R_{z} \in \mathbb{P}
+We assert that the status code $R_z$ is an integer.
+\begin{equation}
+R_z \in \mathbb{P}
 \end{equation}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{z}$, the status code of this transaction:
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
 \begin{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{z}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_z$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_{z}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{z} =
-\Upsilon^{z}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_z$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_z =
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}

### 2018-01-29 Paper.tex
\label{ch:Terminology}

--- 2018-01-29 Paper.tex
@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
 
@@
 \MakeOuterQuote{"}
+% Put before hyperref
 
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{z}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{z})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
 \end{equation}
@@
 
-We assert that the status code $R_{z}$ is an integer.
-\begin{equation}
-R_{z} \in \mathbb{P}
+We assert that the status code $R_{s'}$ is an integer.
+\begin{equation}
+R_{s'} \in \mathbb{P}
 \end{equation}
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, z) \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
 \Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{z}$, the status code of this transaction:
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}
@@
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, s')$ where:
 
@@
 \end{cases} \\
-\quad z &\equiv \begin{cases}
+\quad s' &\equiv \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, s', \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
+s' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
@@
 
-As previously detailed, $\mathbf{R}[n]_{z}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_{z}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{z} =
-\Upsilon^{z}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_{s'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{s'} =
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \end{description}
-\hypertarget{rlp}{}
+
 \section{Recursive Length Prefix}\label{app:rlp}
@@
 \end{itemize}
-\hypertarget{R b}{}
+
 Formally, we define $R_b$:
@@
 All remaining nibbles (now an even number) fit properly into the remaining bytes.
-\hypertarget{trie}{}
+
 \section{Modified Merkle Patricia Tree}\label{app:trie}
@@
 
-\subsection{Deletion of an Account Dispite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\subsection{Deletion of an Account Despite Out-of-gas}
+
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\toprule
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
+\bottomrule
+\end{tabu*}
+
 \end{document}

### 2018-01-29 Paper.tex
\linkdest, underscores

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 \begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
+\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))\linkdest{account_validity_function_v__x}{}
 \end{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+Both types specify a number of common fields:\linkdest{tx_nonce}{}
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\linkdest{tx_gas_price_T__p}{}
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\linkdest{tx_gas_limit_T__g}{}
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\linkdest{tx_to_address_T__t}{}
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\linkdest{tx_value_T__v}{}
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
@@
 
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
 The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
-\bottomrule
-\end{tabu*}
-
 \end{document}

### 2018-01-29 Paper.tex
move linkdest to same line

@@
 \begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))\linkdest{account_validity_function_v__x}{}
-\end{equation}
-where $v$ is the account validity function:
+\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
+\end{equation}
+\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
 \begin{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:\linkdest{tx_nonce}{}
+Both types specify a number of common fields:
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\linkdest{tx_gas_price_T__p}{}
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\linkdest{tx_gas_limit_T__g}{}
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\linkdest{tx_to_address_T__t}{}
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\linkdest{tx_value_T__v}{}
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.{}
+\item[value]\linkdest{tx_value_T__v} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.

### 2018-01-29 Paper.tex
hypertarget => linkdest, move to same line

--- 2018-01-29 Paper.tex
@@
 \begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))\linkdest{account_validity_function_v__x}{}
+\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
 \end{equation}
@@
 There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:\linkdest{tx_nonce}{}
+Both types specify a number of common fields:
 
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.\linkdest{tx_gas_price_T__p}{}
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.\linkdest{tx_gas_limit_T__g}{}
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.\linkdest{tx_to_address_T__t}{}
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.\linkdest{tx_value_T__v}{}
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
+Formally, we can refer to a block $B$:
 \begin{equation}
 B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
+\end{equation}\linkdest{Transaction Receipt}{}
 
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
 \begin{equation}

### 2018-01-29 Paper.tex

@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
+
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\usepackage{seqsplit}
+
+% For formatting
+%\usepackage{underscore}
+%\usepackage{lipsum} % to generate filler text for testing of document rendering
+\usepackage[english]{babel}
+\usepackage[autostyle]{csquotes}
+\MakeOuterQuote{"}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Ethcore\\
-    gavin@ethcore.io
+    Founder, Ethereum \& Parity\\
+    gavin@parity.io
 }
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_z)
 \end{equation}
@@
 
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
+We assert that the status code $R_z$ is an integer.
+\begin{equation}
+R_z \in \mathbb{P}
 \end{equation}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
 \begin{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_z$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_z$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_z =
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}

### 2018-01-29 Paper.tex
hypertarget => linkdest, move to same line

--- 2018-01-29 Paper.tex
@@
 \end{equation}
-\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
+where $v$ is the account validity function:
 \begin{equation}
@@
 \begin{description}
-\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
-\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.{}
-\item[value]\linkdest{tx_value_T__v} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
@@
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+\linkdest{block_validity_H__r}{}H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}

### 2018-01-29 Paper.tex
hypertarget: move to same line

@@
 \begin{array}[t]{lclc}
-\linkdest{block_validity_H__r}{}H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
-
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}

### 2018-01-29 Paper.tex
hypertarget: move to the same line as targeted text, change to linkdest where not in text mode

@@
 
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 
-\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
@@
 
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}

### 2018-01-29 Paper.tex
hypertarget: move to the same line as targeted text, change to linkdest where not in text mode

@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}
@@
 
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
+\section{Contract Creation} \label{ch:create}
 

### 2018-01-30 Paper.tex
hypertarget: move to the same line as the targeted text, change to linkdest where not in text mode

@@
 
-\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}
+\subsubsection{Normal Halting}
 
@@
 
-We define the block reward as 3 Ether:
+\hypertarget{block-reward-R-b}{block-reward-R-b}We define the block reward as 3 Ether:
 \begin{equation}
@@
 
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+\hypertarget{Gamma}{Gamma}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
@@
 
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\hypertarget{Phi}{Phi}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
@@
 
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+\hypertarget{Pi}{Pi}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}

### 2018-01-30 Paper.tex
hypertarget: move to the same line as the targeted text, change to linkdest where not in text mode

@@
 
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
 \begin{eqnarray}

### 2018-01-30 Paper.tex
hypertarget: move to the same line as the targeted text, change to linkdest where not in text mode

@@
 \mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
 \mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-$G_{transaction}$ & 21000 & Paid for every transaction. \\
+\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
+\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\

### 2018-01-30 Paper.tex
hypertarget: move to the same line as the targeted text, change to linkdest where not in text mode

@@
 \multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{blockhash}{BLOCKHASH}\\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+\linkdest{blockhash}{BLOCKHASH}0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \hypertarget{SSTORE}{}\\
-\midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
-&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+\midrule
+\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
-\end{cases}$ \hypertarget{A r}{}\\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+\end{cases}$ \\
+&&&& \linkdest{A r}{}$A'_{r} \equiv A_{r} + \begin{cases}
 R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
@@
 \multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\hypertarget{A l}{}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
-&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
+&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
@@
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
-\midrule
-0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& same account as at present, simply that the code is overwritten.\\
+\midrule
+\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
@@
 \midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{}\\
-\midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
+\midrule
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 0 & \text{otherwise}
-\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
-&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+\end{cases}$ \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad n \\

### 2018-01-30 Paper.tex
hypertarget: move to the same line as the targeted text, change to linkdest where not in text mode Jwordbytes

--- 2018-01-30 Paper.tex
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. 
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
 $G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 %The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
-
+\hypertarget{GenesisBlock}{}
 \section{Genesis Block}\label{app:genesis}
@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
 $J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\

### 2018-01-30 Paper.tex
Remove `mathrm`s in a URL

--- 2018-01-30 Paper.tex
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_mathrm{i} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_mathrm{i} &\equiv \max(H_mathrm{i} - 3000000, 0)
 \end{align}
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_mathrm{i}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+The canonical gas limit $H_mathrm{l}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
+& & H_mathrm{l} \geqslant 5000
+\end{eqnarray}
+
+$H_mathrm{s}$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_mathrm{s} > {P(H)_mathrm{H}}_mathrm{s}
 \end{equation}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
@@
 \multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
-\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \hypertarget{blockhash}{BLOCKHASH}\\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \hypertarget{SSTORE}{}\\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \hypertarget{C tiny SSTORE}{}\\
+&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
-\end{cases}$ \hypertarget{A r}{}\\
+\end{cases}$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{c}{\textbf{a0s: Logging Operations}} \vspace{5pt} \\
-\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\hypertarget{A l}{}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}$ (not enough funds) or $I_mathrm{e} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
-&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \hypertarget{tiny CALL}{}\\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
 &&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{a}, I_mathrm{o}, I_mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\\ \quad\quad{}I_mathrm{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\hypertarget{RETURN}{}\\
+&&&& (as in {\small CALL}) to the present address $I_mathrm{a}$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \hypertarget{selfdestruct}{}\\
+&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
 \midrule
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_mathrm{a} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} = 0\\
+(\boldsymbol{\sigma}[r]_mathrm{n}, \boldsymbol{\sigma}[r]_mathrm{b} + \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c}) & \text{if}\ r \neq I_mathrm{a} \\
+(\boldsymbol{\sigma}[r]_mathrm{n}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c}) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_mathrm{b} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_mathrm{a} \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \hypertarget{C tiny SELFDESTRUCT}{}\\
+\end{cases}$ \\
 &&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}

### 2018-01-30 Paper.tex
Find and replace all _([:alpha:][:alnum:]*) with _{\mathrm{$1}}

@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_{\mathrm{and}}_{\mathrm{Hypertarget}}.
+
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\usepackage{seqsplit}
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_mathrm{t}, T)
+\linkdest{Upsilon_{\mathrm{state}}_{\mathrm{transition}}}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_mathrm{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{\mathrm{t}}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_mathrm{n}$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_mathrm{n}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_mathrm{b}$.
+\item[nonce] \linkdest{account_{\mathrm{nonce}}}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_mathrm{s}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_mathrm{c}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_mathrm{c}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_mathrm{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_mathrm{s}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_mathrm{I}^*$, is defined as the element-wise transformation of the base function $L_mathrm{I}$, given as:
-\begin{equation}
-L_mathrm{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
+\begin{equation}
+L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_mathrm{c} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_mathrm{S}$:
-\begin{equation}
-L_mathrm{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_{\mathrm{S}}$:
+\begin{equation}
+L_{\mathrm{S}}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_mathrm{n}, \boldsymbol{\sigma}[a]_mathrm{b}, \boldsymbol{\sigma}[a]_mathrm{s}, \boldsymbol{\sigma}[a]_mathrm{c}) \big) \big)
-\end{equation}
-
-This function, $L_mathrm{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
+\end{equation}
+
+This function, $L_{\mathrm{S}}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
 \begin{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_mathrm{n} \in \mathbb{P}_{256} \wedge x_mathrm{b} \in \mathbb{P}_{256} \wedge x_mathrm{s} \in \mathbb{B}_{32} \wedge x_mathrm{c} \in \mathbb{B}_{32}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_mathrm{c} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_mathrm{n} = 0 \wedge \boldsymbol{\sigma}[a]_mathrm{b} = 0
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
 \end{equation}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_mathrm{n}$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_mathrm{p}$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_mathrm{g}$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_mathrm{t}$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_mathrm{v}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_mathrm{w}$, $T_mathrm{r}$ and $T_mathrm{s}$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-L_mathrm{T}(T) \equiv \begin{cases}
-(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, T_\mathbf{i}, T_mathrm{w}, T_mathrm{r}, T_mathrm{s}) & \text{if} \; T_mathrm{t} = \varnothing\\
-(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, T_\mathbf{d}, T_mathrm{w}, T_mathrm{r}, T_mathrm{s}) & \text{otherwise}
+L_{\mathrm{T}}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{i}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{d}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_mathrm{n} \in \mathbb{P}_{256} & \wedge & T_mathrm{v} \in \mathbb{P}_{256} & \wedge & T_mathrm{p} \in \mathbb{P}_{256} & \wedge \\
-T_mathrm{g} \in \mathbb{P}_{256} & \wedge & T_mathrm{w} \in \mathbb{P}_5 & \wedge & T_mathrm{r} \in \mathbb{P}_{256} & \wedge \\
-T_mathrm{s} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_mathrm{n} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_mathrm{t} \neq \varnothing \\
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_mathrm{p}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_mathrm{o}$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_mathrm{c}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_mathrm{r}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_mathrm{t}$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_mathrm{e}$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_mathrm{b}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_mathrm{d}$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_mathrm{i}$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_mathrm{l}$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_mathrm{g}$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_mathrm{s}$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_mathrm{x}$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_mathrm{m}$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_mathrm{n}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 \begin{equation}
-B \equiv (B_mathrm{H}, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_{\mathrm{H}}, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_mathrm{e}$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_mathrm{u}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_mathrm{b}$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_mathrm{u}, R_mathrm{b}, R_\mathbf{l}, R_{s'})
-\end{equation}
-
-The function $L_mathrm{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_mathrm{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_mathrm{u}, R_mathrm{b}, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
+\end{equation}
+
+The function $L_{\mathrm{R}}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{\mathrm{R}}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
 \end{equation}
@@
 
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
-\end{equation}
-
-We assert $R_mathrm{u}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_mathrm{b}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_mathrm{u} \in \mathbb{P} \quad \wedge \quad R_mathrm{b} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_mathrm{B}$; it is merely a convenience equivalence.
+We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\begin{equation}
+R_{\mathrm{z}} \in \mathbb{P}
+\end{equation}
+
+We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{\mathrm{B}}$; it is merely a convenience equivalence.
 
 The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_mathrm{a}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_mathrm{a}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_mathrm{a} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_mathrm{a}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_mathrm{j}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_mathrm{r}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_mathrm{r} &\equiv& \mathtt{\small TRIE}(L_mathrm{S}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_mathrm{o} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}^*(B_\mathbf{U}))) & \wedge \\
-H_mathrm{t} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_mathrm{T}(B_\mathbf{T}[i]))\}) & \wedge \\
-H_mathrm{e} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_mathrm{R}(B_\mathbf{R}[i]))\}) & \wedge \\
-H_mathrm{b} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_mathrm{b} \big)
+H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_{\mathrm{S}}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}^*(B_\mathbf{U}))) & \wedge \\
+H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{T}}(B_\mathbf{T}[i]))\}) & \wedge \\
+H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{R}}(B_\mathbf{R}[i]))\}) & \wedge \\
+H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma})) = {P(B_mathrm{H})_mathrm{H}}_mathrm{r}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_mathrm{H})$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{\mathrm{H}})$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_mathrm{B}$ and $L_mathrm{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_mathrm{R}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_mathrm{H}(H) & \equiv & (\begin{array}[t]{l}H_mathrm{p}, H_mathrm{o}, H_mathrm{c}, H_mathrm{r}, H_mathrm{t}, H_mathrm{e}, H_mathrm{b}, H_mathrm{d},\\ H_mathrm{i}, H_mathrm{l}, H_mathrm{g}, H_mathrm{s}, H_mathrm{x}, H_mathrm{m}, H_mathrm{n} \; )\end{array} \\
-\quad L_mathrm{B}(B) & \equiv & \big( L_mathrm{H}(B_mathrm{H}), L_mathrm{T}^*(B_\mathbf{T}), L_mathrm{H}^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_mathrm{T}^*$ and $L_mathrm{H}^*$ being element-wise sequence transformations, thus:
+The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_{\mathrm{R}}$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_\mathbf{T}), L_{\mathrm{H}}^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_mathrm{p} \in \mathbb{B}_{32} & \wedge & H_mathrm{o} \in \mathbb{B}_{32} & \wedge & H_mathrm{c} \in \mathbb{B}_{20} & \wedge \\
-H_mathrm{r} \in \mathbb{B}_{32} & \wedge & H_mathrm{t} \in \mathbb{B}_{32} & \wedge & H_mathrm{e} \in \mathbb{B}_{32} & \wedge \\
-H_mathrm{b} \in \mathbb{B}_{256} & \wedge & H_mathrm{d} \in \mathbb{P} & \wedge & H_mathrm{i} \in \mathbb{P} & \wedge \\
-H_mathrm{l} \in \mathbb{P} & \wedge & H_mathrm{g} \in \mathbb{P} & \wedge & H_mathrm{s} \in \mathbb{P}_{256} & \wedge \\
-H_mathrm{x} \in \mathbb{B} & \wedge & H_mathrm{m} \in \mathbb{B}_{32} & \wedge & H_mathrm{n} \in \mathbb{B}_{8}
+H_{\mathrm{p}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+H_{\mathrm{r}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
+H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & H_{\mathrm{i}} \in \mathbb{P} & \wedge \\
+H_{\mathrm{l}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & H_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
+H_{\mathrm{x}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{n}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_mathrm{n} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 
-We define $P(B_mathrm{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_mathrm{H})) = H_mathrm{p}
+We define $P(B_{\mathrm{H}})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{\mathrm{H}})) = H_{\mathrm{p}}
 \end{equation}
@@
 \begin{equation}
-H_mathrm{i} \equiv {{P(H)_mathrm{H}}_mathrm{i}} + 1
+H_{\mathrm{i}} \equiv {{P(H)_{\mathrm{H}}}_{\mathrm{i}}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_mathrm{i} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_mathrm{H}}_mathrm{d} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{d}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_mathrm{s} - {P(H)_mathrm{H}}_mathrm{s}}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_mathrm{i} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_mathrm{i} &\equiv \max(H_mathrm{i} - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_mathrm{i}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-The canonical gas limit $H_mathrm{l}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
-& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor \quad \wedge \\
-& & H_mathrm{l} \geqslant 5000
-\end{eqnarray}
-
-$H_mathrm{s}$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_mathrm{s} > {P(H)_mathrm{H}}_mathrm{s}
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000
+\end{eqnarray}
+
+$H_{\mathrm{s}}$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}}
 \end{equation}
@@
 
-The nonce, $H_mathrm{n}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_mathrm{d}} \quad \wedge \quad m = H_mathrm{m}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d})$.
+The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
 
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_mathrm{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_mathrm{d}} \wedge m = H_mathrm{m} \quad \wedge \\
-& & H_mathrm{d} = D(H) \quad \wedge \\
-& & H_mathrm{g} \le H_mathrm{l}  \quad \wedge \\
-& & H_mathrm{l} < {P(H)_mathrm{H}}_mathrm{l} + \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
-& & H_mathrm{l} > {P(H)_mathrm{H}}_mathrm{l} - \left\lfloor\frac{{P(H)_mathrm{H}}_mathrm{l}}{1024}\right\rfloor  \quad \wedge \\
-& & H_mathrm{l} \geqslant 5000  \quad \wedge \\
-& & H_mathrm{s} > {P(H)_mathrm{H}}_mathrm{s} \quad \wedge \\
-& & H_mathrm{i} = {P(H)_mathrm{H}}_mathrm{i} +1 \quad \wedge \\
-& & \lVert H_mathrm{x} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d})$
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
+& & H_{\mathrm{d}} = D(H) \quad \wedge \\
+& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
+& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
+& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
+& & \lVert H_{\mathrm{x}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_mathrm{r})
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
 \end{equation}
@@
 $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_mathrm{r}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_mathrm{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 \begin{equation}
-v_0 \equiv T_mathrm{g} T_mathrm{p} + T_mathrm{v}
+v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_mathrm{n} & = & \boldsymbol{\sigma}[S(T)]_mathrm{n} \quad \wedge \\
-g_0 & \leqslant & T_mathrm{g} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_mathrm{b} \quad \wedge \\
-T_mathrm{g} & \leqslant & {B_mathrm{H}}_mathrm{l} - \ell(B_\mathbf{R})_mathrm{u}
+T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_mathrm{g}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_mathrm{u}$, must be no greater than the block's \textbf{gasLimit}, ${B_mathrm{H}}_mathrm{l}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_mathrm{gT}_mathrm{p}$.
-The gas available for the proceeding computation, $g$, is defined as $T_mathrm{g} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{gT}}_{\mathrm{p}}$.
+The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_mathrm{b} & \equiv & \boldsymbol{\sigma}[S(T)]_mathrm{b} - T_mathrm{g} T_mathrm{p} \\
-\boldsymbol{\sigma}_0[S(T)]_mathrm{n} & \equiv & \boldsymbol{\sigma}[S(T)]_mathrm{n} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_mathrm{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_mathrm{P}$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_mathrm{P}, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_mathrm{o}, &\\ \quad\quad g, T_mathrm{p}, T_mathrm{v}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_mathrm{t} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_mathrm{o}, &\\ \quad\quad T_mathrm{t}, T_mathrm{t}, g, T_mathrm{p}, T_mathrm{v}, T_mathrm{v}, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_mathrm{g} - g_0
-\end{equation}
-and $T_mathrm{o}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_{\mathrm{g}} - g_0
+\end{equation}
+and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_mathrm{g} - g'}{2} \Big\rfloor, A_mathrm{r} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_mathrm{r}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_mathrm{g} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_mathrm{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_mathrm{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_mathrm{b} & \equiv & \boldsymbol{\sigma}_mathrm{P}[S(T)]_mathrm{b} + g^* T_mathrm{p} \\
-\boldsymbol{\sigma}^*[m]_mathrm{b} & \equiv & \boldsymbol{\sigma}_mathrm{P}[m]_mathrm{b} + (T_mathrm{g} - g^*) T_mathrm{p} \\
-m & \equiv & {B_mathrm{H}}_mathrm{c}
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{\mathrm{P}} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}}
 \end{eqnarray}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_mathrm{g} - g' \\
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
 \end{eqnarray}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_mathrm{n} - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_mathrm{n}, \boldsymbol{\sigma}[s]_mathrm{b} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_mathrm{c})
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_mathrm{b} & \text{otherwise}
+\boldsymbol{\sigma}[a]_{\mathrm{b}} & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_mathrm{a} & \equiv & a \\
-I_mathrm{o} & \equiv & o \\
-I_mathrm{p} & \equiv & p \\
+I_{\mathrm{a}} & \equiv & a \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_mathrm{s} & \equiv & s \\
-I_mathrm{v} & \equiv & v \\
+I_{\mathrm{s}} & \equiv & s \\
+I_{\mathrm{v}} & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_mathrm{e} & \equiv & e \\
-I_mathrm{w} & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_mathrm{H}$ has no special treatment and is determined from the blockchain.
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{\mathrm{H}}$ has no special treatment and is determined from the blockchain.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_mathrm{c} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
-\quad s' &\equiv \begin{cases}
+\quad z &\equiv \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_mathrm{b} \equiv \boldsymbol{\sigma}[r]_mathrm{b} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_mathrm{b} \equiv \boldsymbol{\sigma}[s]_mathrm{b} - v
+\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_mathrm{n}, \boldsymbol{\sigma}_1'[s]_mathrm{b} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_mathrm{c})
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_mathrm{n}, \boldsymbol{\sigma}[r]_mathrm{b} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c})
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_mathrm{c}$) is executed according to the execution model (see section \ref{ch:model}).
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_mathrm{ADD}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_mathrm{MUL}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{BN\_{\mathrm{ADD}}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_{\mathrm{MUL}}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
 \Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_mathrm{a} & \equiv & r \\
-I_mathrm{o} & \equiv & o \\
-I_mathrm{p} & \equiv & p \\
+I_{\mathrm{a}} & \equiv & r \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_mathrm{s} & \equiv & s \\
-I_mathrm{v} & \equiv & \tilde{v} \\
-I_mathrm{e} & \equiv & e \\
-I_mathrm{w} & \equiv & w \\
+I_{\mathrm{s}} & \equiv & s \\
+I_{\mathrm{v}} & \equiv & \tilde{v} \\
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w \\
 \mathbf{t} & \equiv & \{s, r\} \\
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_mathrm{c}
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
 \end{eqnarray}
@@
 \begin{itemize}
-\item $I_mathrm{a}$, the address of the account which owns the code that is executing.
-\item $I_mathrm{o}$, the sender address of the transaction that originated this execution.
-\item $I_mathrm{p}$, the price of gas in the transaction that originated this execution.
+\item $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
 \item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_mathrm{s}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_mathrm{v}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_mathrm{H}$, the block header of the present block.
-\item $I_mathrm{e}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_mathrm{w}$, the permission to make modifications to the state.
+\item $I_{\mathrm{H}}$, the block header of the present block.
+\item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_mathrm{g}, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_mathrm{g} & \equiv & g \\
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_mathrm{i} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_mathrm{g} & \equiv & \boldsymbol{\mu}_mathrm{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_mathrm{g}$ from the resultant machine state $\boldsymbol{\mu}'$.
+\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
+\hypertarget{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{\mathrm{Z}}}{}
 \subsubsection{Exceptional Halting}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_mathrm{g} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_mathrm{w} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_mathrm{w} \quad \vee \\
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_mathrm{w} + \mathbf{\alpha}_mathrm{w} > 1024 \quad \vee \\
- \neg I_mathrm{w} \wedge W(w, \boldsymbol{\mu})
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+ \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
 \end{array}
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_mathrm{J}(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_{\mathrm{J}}(\mathbf{c}, 0)
 \end{equation}
@@
 \begin{equation}
-D_mathrm{J}(\mathbf{c}, i) \equiv \begin{cases}
+D_{\mathrm{J}}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_mathrm{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_mathrm{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_mathrm{w} - \mathbf{\delta}_mathrm{w} \\
+\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
 \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_mathrm{w}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 \boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_mathrm{i} & \equiv & \boldsymbol{\mu}_mathrm{i} \\
+\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_mathrm{t} & \equiv & B'_mathrm{t} + B_mathrm{d} \\
-B' & \equiv & P(B_mathrm{H})
-\end{eqnarray}
-
-As such given a block $B$, $B_mathrm{t}$ is its total difficulty, $B'$ is its parent block and $B_mathrm{d}$ is its difficulty.
+B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
+B' & \equiv & P(B_{\mathrm{H}})
+\end{eqnarray}
+
+As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
 
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_mathrm{H})_mathrm{H}, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_{\mathrm{H}})_{\mathrm{H}}, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_mathrm{H}, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_{\mathrm{H}}, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_mathrm{H}}_mathrm{g}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_mathrm{H}}_mathrm{g} = \ell(\mathbf{R})_mathrm{u}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{\mathrm{H}}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{\mathrm{H}}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_mathrm{b}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_mathrm{H}}_mathrm{c}]_mathrm{b} & = & \boldsymbol{\sigma}[{B_mathrm{H}}_mathrm{c}]_mathrm{b} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_mathrm{b} \\
+\qquad\boldsymbol{\sigma}'[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_mathrm{c}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_mathrm{c}] = \varnothing\ \wedge\ R = 0 \\
+\boldsymbol{\sigma}'[U_{\mathrm{c}}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_mathrm{c}]_mathrm{n}, \boldsymbol{\sigma}[U_mathrm{c}]_mathrm{b} + R, \boldsymbol{\sigma}[U_mathrm{c}]_\mathbf{s}, \boldsymbol{\sigma}[U_mathrm{c}]_mathrm{c}) \\
-R & \equiv & (1 + \frac{1}{8} (U_mathrm{i} - {B_mathrm{H}}_mathrm{i})) R_mathrm{b}
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{\mathrm{H}}}_{\mathrm{i}})) R_{\mathrm{b}}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_mathrm{b} = 3 \times 10^{18}
+\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
 \end{equation}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_mathrm{H}) = \varnothing \\
-\boldsymbol{\sigma}_mathrm{i}: \mathtt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma}_mathrm{i})) = {P(B_mathrm{H})_mathrm{H}}_mathrm{r} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_mathrm{S}(\boldsymbol{\sigma}_mathrm{i}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_mathrm{i}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{\mathrm{H}}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_mathrm{n} & = & n: \quad x \leqslant \frac{2^{256}}{H_mathrm{d}} \\
-B'_mathrm{m} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_mathrm{r} = r(\Pi(\Gamma(B), B))
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_mathrm{u}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_mathrm{b}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $B_\mathbf{R}[n]_mathrm{u}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_mathrm{u} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+In the case of $B_\mathbf{R}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_mathrm{u}
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-We define $\mathbf{R}[n]_{s'}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{s'} =
-\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{z}} =
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \begin{equation}
-m = H_mathrm{m} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_mathrm{d}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_mathrm{n}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_mathrm{d}$ is the new block's difficulty value (i.e.
+m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_mathrm{b}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_mathrm{l}(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-
-Formally, we define $R_mathrm{b}$:
-\begin{eqnarray}
-R_mathrm{b}(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R b}{}
+Formally, we define $R_{\mathrm{b}}$:
+\begin{eqnarray}
+R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_mathrm{n} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus we finish by formally defining $R_mathrm{l}$:
-\begin{eqnarray}
-R_mathrm{l}(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_{\mathrm{l}}$:
+\begin{eqnarray}
+R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_mathrm{x}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_mathrm{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_mathrm{n}\rVert} \quad \mathbf{k}_mathrm{n}'[i] & \equiv &
+\forall_{\mathrm{n}} \quad \forall_{i: i < 2\lVert\mathbf{k}_{\mathrm{n}}\rVert} \quad \mathbf{k}_{\mathrm{n}}'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_mathrm{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_mathrm{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_{\mathrm{n}}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_{\mathrm{n}}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_mathrm{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{\mathrm{x}} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_mathrm{r} \\
-(\boldsymbol\sigma, g - g_mathrm{r}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_mathrm{r}$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_mathrm{r} &=& 3000\\
+g_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_mathrm{r} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_mathrm{r} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_mathrm{r} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_mathrm{r} &=& \Big\lfloor\frac{f\big(\max(\ell_mathrm{M},\ell_mathrm{B})\big)\max(\ell'_mathrm{E},1)}{G_{quaddivisor}}\Big\rfloor \\
+g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_mathrm{E} &=& \begin{cases}
-0 & \text{if}\ \ell_mathrm{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_mathrm{E}\le 32 \wedge E \neq 0 \\
-8(\ell_mathrm{E} - 32) + \lfloor \log_2(i[(96+\ell_mathrm{B})..(127+\ell_mathrm{B})]) \rfloor & \text{if}\ 32 < \ell_mathrm{E} \wedge i[(96 + \ell_mathrm{B})..(127 + \ell_mathrm{B})]\neq 0 \\
-8(\ell_mathrm{E} - 32) & \text{otherwise} \\
+\ell'_{\mathrm{E}} &=& \begin{cases}
+0 & \text{if}\ \ell_{\mathrm{E}}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{\mathrm{E}}\le 32 \wedge E \neq 0 \\
+8(\ell_{\mathrm{E}} - 32) + \lfloor \log_2(i[(96+\ell_{\mathrm{B}})..(127+\ell_{\mathrm{B}})]) \rfloor & \text{if}\ 32 < \ell_{\mathrm{E}} \wedge i[(96 + \ell_{\mathrm{B}})..(127 + \ell_{\mathrm{B}})]\neq 0 \\
+8(\ell_{\mathrm{E}} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_mathrm{M}} \\
-\ell_mathrm{B} &\equiv& i[0..31] \\
-\ell_mathrm{E} &\equiv& i[32..63] \\
-\ell_mathrm{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_mathrm{B})] \\
-E &\equiv& i[(96+\ell_mathrm{B})..(95+\ell_mathrm{B}+\ell_mathrm{E})] \\
-M &\equiv& i[(96+\ell_mathrm{B}+\ell_mathrm{E})..(95+\ell_mathrm{B}+\ell_mathrm{E}+\ell_mathrm{M})] \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{\mathrm{M}}} \\
+\ell_{\mathrm{B}} &\equiv& i[0..31] \\
+\ell_{\mathrm{E}} &\equiv& i[32..63] \\
+\ell_{\mathrm{M}} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{\mathrm{B}})] \\
+E &\equiv& i[(96+\ell_{\mathrm{B}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})] \\
+M &\equiv& i[(96+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}}+\ell_{\mathrm{M}})] \\
 i[x] &\equiv& \begin{cases}
@@
 \end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_mathrm{p}$.
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
 
@@
 \begin{equation}
-C_1\equiv\{(X,Y)\in F_mathrm{p}\times F_mathrm{p}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 
-Let $F_{p^2}$ be a field $F_mathrm{p}[i]/(i+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_mathrm{p}$.
-\begin{equation}
-\delta_mathrm{p}(\mathbf x)\equiv\begin{cases}
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
+\begin{equation}
+\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
 \mathbf x&\text{if}\ \mathbf x<p\\
@@
 \end{cases}\\
-x&\equiv&\delta_mathrm{p}(\mathbf x[0..31])\\
-y&\equiv&\delta_mathrm{p}(\mathbf x[32..63])
+x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
 \end{eqnarray}
@@
 \end{cases}\\
-x_0&\equiv&\delta_mathrm{p}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_mathrm{p}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_mathrm{p}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_mathrm{p}(\mathbf x[96..127])
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
 \end{eqnarray}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_mathrm{j}=\varnothing\vee b_mathrm{j}=\varnothing))\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
 k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_mathrm{r}&=& 60000k + 40000 \\
+g_{\mathrm{r}}&=& 60000k + 40000 \\
 \mathbf o[0..31]&\equiv&\begin{cases}
@@
 \end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_mathrm{k})\times\log_{P_2}(b_mathrm{k})\equiv 1\mod q)\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
 a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
@@
 \vdots\\
-a_mathrm{k}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_mathrm{k}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
 \end{eqnarray}
@@
 \begin{eqnarray}
-\Xi_{\mathtt{BN\_mathrm{ADD}}}&\equiv&\Xi_{\mathtt{BN\_mathrm{PRE}}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_mathrm{ADD}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_mathrm{r} &=& 500\\
+\Xi_{\mathtt{BN\_{\mathrm{ADD}}}}&\equiv&\Xi_{\mathtt{BN\_{\mathrm{PRE}}}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_{\mathrm{ADD}}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_{\mathrm{r}} &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_mathrm{input}}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+\label{eq:complemented_{\mathrm{input}}}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
 I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_mathrm{input}}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_mathrm{MUL}}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_mathrm{MUL}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_mathrm{r} &=& 40000\\
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_{\mathrm{input}}}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_{\mathrm{MUL}}}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_{\mathrm{MUL}}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+g_{\mathrm{r}} &=& 40000\\
 \mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 
-It is assumed that the sender has a valid private key $p_mathrm{r}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
 
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_mathrm{r} \in \mathbb{B}_{32}) & \equiv & p_mathrm{u} \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_mathrm{r} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_mathrm{u} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_mathrm{u}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_mathrm{r}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_mathrm{id}} \times 2 + 35, \mathtt{\tiny chain\_mathrm{id}} \times 2 + 36\}
+v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
 \end{align}
@@
 
-For a given private key, $p_mathrm{r}$, the Ethereum address $A(p_mathrm{r})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_mathrm{r}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_mathrm{r}) \big) \big)
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
 \end{equation}
@@
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_mathrm{r}$, $T_mathrm{s}$ and $T_mathrm{w}$.
+One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_mathrm{S}(T) & \equiv & \begin{cases}
-(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_mathrm{n}, T_mathrm{p}, T_mathrm{g}, T_mathrm{t}, T_mathrm{v}, \mathbf{p}, \mathtt{\tiny chain\_mathrm{id}}, (), ()) & \text{otherwise} \\
+L_{\mathrm{S}}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_mathrm{t} = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_mathrm{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_mathrm{r})$ is defined as:
-\begin{eqnarray}
-G(T, p_mathrm{r}) \equiv T \quad \text{except:} \\
-(T_mathrm{w}, T_mathrm{r}, T_mathrm{s}) = \mathtt{\small ECDSASIGN}(h(T), p_mathrm{r})
+h(T) & \equiv & \mathtt{\small KEC}( L_{\mathrm{S}}(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
+\begin{eqnarray}
+G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
 \end{eqnarray}
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_mathrm{r}, T_mathrm{s}) \big) \big) \\
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
 v_0 &\equiv& \begin{cases}
-T_mathrm{w} &\text{if}\ T_mathrm{w}\in\{27, 28\} \\
-28 - (T_mathrm{w} \bmod 2) &\text{otherwise}
+T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
+28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\forall T: \forall p_mathrm{r}: S(G(T, p_mathrm{r})) \equiv A(p_mathrm{r})
+\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
 \end{equation}
@@
 
-%extern u256 const c_mathrm{copyGas};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_mathrm{i})-C_{mem}(\boldsymbol{\mu}_mathrm{i}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_mathrm{i}$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_mathrm{i}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{t} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_mathrm{i}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} \vee \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_mathrm{i}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_mathrm{i} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_mathrm{i} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{a}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_mathrm{b}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{o}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{s}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{v}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
 \midrule
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_mathrm{p}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_mathrm{c} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_mathrm{c}$ \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_mathrm{i} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_mathrm{i} \\ P(H_mathrm{p}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{c}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{s}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{d}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_mathrm{H}}_mathrm{l}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv \max(\boldsymbol{\mu}_mathrm{i}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_mathrm{i}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_mathrm{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_mathrm{a}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_mathrm{g}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_mathrm{a}, I_mathrm{o}, L(\boldsymbol{\mu}_mathrm{g}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\; I_mathrm{e} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_mathrm{g}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_mathrm{a}]_mathrm{n} = \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{n} + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_mathrm{e} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_mathrm{a}, \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{n})$, the address of the newly \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{\mathrm{Z}}}{exceptional halting}\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_mathrm{a}, I_mathrm{o}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge \\ \quad\quad I_mathrm{e} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_mathrm{g} \equiv \boldsymbol{\mu}_mathrm{g} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{\mathrm{Z}}}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}$ (not enough funds) or $I_mathrm{e} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_mathrm{g} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_mathrm{g} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{a}, I_mathrm{o}, I_mathrm{a}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_mathrm{p}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\\ \quad\quad{}I_mathrm{e} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_mathrm{a}$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
@@
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_mathrm{s}, I_mathrm{o}, I_mathrm{a}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_mathrm{p}, 0, I_mathrm{v}, \mathbf{i}, I_mathrm{e} + 1, I_mathrm{w})\end{array} & \text{if} \quad I_mathrm{v} \leqslant \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \;\wedge\; I_mathrm{e} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_mathrm{i} \equiv M(\boldsymbol{\mu}_mathrm{i}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_mathrm{a} \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_{\mathrm{a}} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} = 0\\
-(\boldsymbol{\sigma}[r]_mathrm{n}, \boldsymbol{\sigma}[r]_mathrm{b} + \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c}) & \text{if}\ r \neq I_mathrm{a} \\
-(\boldsymbol{\sigma}[r]_mathrm{n}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_mathrm{c}) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_mathrm{a}]_mathrm{b} = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_mathrm{a} \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_mathrm{a}]_mathrm{b} \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(H_mathrm{i}) = \left\lfloor\frac{H_mathrm{i}}{J_{epoch}}\right\rfloor
+ E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_mathrm{i})
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_mathrm{i}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_mathrm{i}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_mathrm{i} - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_mathrm{n}, \mathbf{d}) = \lbrace \mathbf{m}_mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_mathrm{n}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_mathrm{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_mathrm{n}) + \mathbf{m}_mathrm{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_mathrm{H}(H_{\hcancel{n}}))), H_mathrm{n}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_mathrm{c}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_mathrm{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_mathrm{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+ \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
 \end{equation}

### 2018-01-30 Paper.tex
Remove underscores in a URL
Remove mathrms that have been added in the wrong place.
Remove mathrms that have been added in the wrong place
Replace all mathrm\\\{([A-Z]\} with $1
Using Libre Open Office with match case and regex.

@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_{\mathrm{and}}_{\mathrm{Hypertarget}}.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
 
@@
 \begin{equation}
-\linkdest{Upsilon_{\mathrm{state}}_{\mathrm{transition}}}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
+\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
 \end{equation}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
-\begin{equation}
-L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
+\begin{equation}
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 
-Thus we may define a world-state collapse function $L_{\mathrm{S}}$:
-\begin{equation}
-L_{\mathrm{S}}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+Thus we may define a world-state collapse function $L_{S}$:
+\begin{equation}
+L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 
-This function, $L_{\mathrm{S}}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
 \begin{equation}
@@
 \begin{equation}
-L_{\mathrm{T}}(T) \equiv \begin{cases}
+L_{T}(T) \equiv \begin{cases}
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{i}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
@@
 \begin{equation}
-B \equiv (B_{\mathrm{H}}, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 
-The function $L_{\mathrm{R}}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{\mathrm{R}}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
+The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
 \end{equation}
@@
 
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{\mathrm{B}}$; it is merely a convenience equivalence.
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
 
@@
 \begin{array}[t]{lclc}
-H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_{\mathrm{S}}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}^*(B_\mathbf{U}))) & \wedge \\
-H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{T}}(B_\mathbf{T}[i]))\}) & \wedge \\
-H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{R}}(B_\mathbf{R}[i]))\}) & \wedge \\
+H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_{S}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}^*(B_\mathbf{U}))) & \wedge \\
+H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
 H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{\mathrm{H}})$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{\mathrm{R}}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_\mathbf{T}), L_{\mathrm{H}}^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
+The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_{T}^*$ and $L_{H}^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 
-We define $P(B_{\mathrm{H}})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{\mathrm{H}})) = H_{\mathrm{p}}
+We define $P(B_{H})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = H_{\mathrm{p}}
 \end{equation}
@@
 \begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{\mathrm{H}}}_{\mathrm{i}}} + 1
+H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
 \end{equation}
@@
 \mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
 & & H_{\mathrm{l}} \geqslant 5000
@@
 \begin{equation}
-H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}}
+H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
 \end{equation}
@@
 & & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
 & & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
-& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
+& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
 & & \lVert H_{\mathrm{x}} \rVert \le 32
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
 
@@
 
-Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
 \Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
@@
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{\mathrm{P}} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}}
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_{H}}_{\mathrm{c}}
 \end{eqnarray}
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{\mathrm{H}}$ has no special treatment and is determined from the blockchain.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_{\mathrm{H}}$, the block header of the present block.
+\item $I_{H}$, the block header of the present block.
 \item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-\hypertarget{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{\mathrm{Z}}}{}
+\hypertarget{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{Z}}{}
 \subsubsection{Exceptional Halting}
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_{\mathrm{J}}(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
 \end{equation}
@@
 \begin{equation}
-D_{\mathrm{J}}(\mathbf{c}, i) \equiv \begin{cases}
+D_{J}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
-B' & \equiv & P(B_{\mathrm{H}})
+B' & \equiv & P(B_{H})
 \end{eqnarray}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_{\mathrm{H}})_{\mathrm{H}}, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_{H})_{H}, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_{\mathrm{H}}, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{\mathrm{H}}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{\mathrm{H}}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
+\qquad\boldsymbol{\sigma}'[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{\mathrm{H}}}_{\mathrm{i}})) R_{\mathrm{b}}
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{\mathrm{H}}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{quaddivisor}}\Big\rfloor \\
+g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_{\mathrm{E}} &=& \begin{cases}
-0 & \text{if}\ \ell_{\mathrm{E}}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{\mathrm{E}}\le 32 \wedge E \neq 0 \\
-8(\ell_{\mathrm{E}} - 32) + \lfloor \log_2(i[(96+\ell_{\mathrm{B}})..(127+\ell_{\mathrm{B}})]) \rfloor & \text{if}\ 32 < \ell_{\mathrm{E}} \wedge i[(96 + \ell_{\mathrm{B}})..(127 + \ell_{\mathrm{B}})]\neq 0 \\
-8(\ell_{\mathrm{E}} - 32) & \text{otherwise} \\
+\ell'_{E} &=& \begin{cases}
+0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
+8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
+8(\ell_{E} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{\mathrm{M}}} \\
-\ell_{\mathrm{B}} &\equiv& i[0..31] \\
-\ell_{\mathrm{E}} &\equiv& i[32..63] \\
-\ell_{\mathrm{M}} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{\mathrm{B}})] \\
-E &\equiv& i[(96+\ell_{\mathrm{B}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})] \\
-M &\equiv& i[(96+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}}+\ell_{\mathrm{M}})] \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
+\ell_{B} &\equiv& i[0..31] \\
+\ell_{E} &\equiv& i[32..63] \\
+\ell_{M} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{B})] \\
+E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
+M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
 i[x] &\equiv& \begin{cases}
@@
 \begin{eqnarray}
-L_{\mathrm{S}}(T) & \equiv & \begin{cases}
+L_{S}(T) & \equiv & \begin{cases}
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{\mathrm{S}}(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
 \end{eqnarray}
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{\mathrm{Z}}}{exceptional halting}\\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{Z}}{exceptional halting}\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{\mathrm{Z}}}{exceptional halting} \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{Z}}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
 \end{equation}

### 2018-01-30 Paper.tex
Remove mathrms where wrongly added
Remove wrong mathrms

@@
 \Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_{\mathrm{ADD}}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_{\mathrm{MUL}}}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
 \Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
@@
 \begin{eqnarray}
-\Xi_{\mathtt{BN\_{\mathrm{ADD}}}}&\equiv&\Xi_{\mathtt{BN\_{\mathrm{PRE}}}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_{\mathrm{ADD}}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_{\mathrm{PRE}}}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 500\\
@@
 y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_{\mathrm{input}}}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
 I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_{\mathrm{input}}}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_{\mathrm{MUL}}}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_{\mathrm{MUL}}}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+\begin{eqnarray}
+\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 40000\\

### 2018-01-30 Paper.tex
Remove wrong mathrms

@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-\hypertarget{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{Z}}{}
+\hypertarget{Exceptional_Halting_function_Z}{}
 \subsubsection{Exceptional Halting}
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_{\mathrm{n}} \quad \forall_{i: i < 2\lVert\mathbf{k}_{\mathrm{n}}\rVert} \quad \mathbf{k}_{\mathrm{n}}'[i] & \equiv &
+\forall_{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_{n}\rVert} \quad \mathbf{k}_{n}'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_{\mathrm{n}}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_{\mathrm{n}}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{\mathrm{x}} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_{\mathrm{PRE}}}}\quad\text{except:}\\
+\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
 \Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{Z}}{exceptional halting}\\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_{\mathrm{Halting}}_{\mathrm{function}}_{Z}}{exceptional halting} \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}_{Z}}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\

### 2018-01-30 Paper.tex
Add a missing forward slash

@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{\mathrm{t}}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 \begin{description}
-\item[nonce] \linkdest{account_{\mathrm{nonce}}}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
+\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\begin{equation}
+L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_{S}$:
-\begin{equation}
-L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_S$:
+\begin{equation}
+L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
-\end{equation}
-
-This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_n, \boldsymbol{\sigma}[a]_b, \boldsymbol{\sigma}[a]_s, \boldsymbol{\sigma}[a]_c) \big) \big)
+\end{equation}
+
+This function, $L_S$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
 \begin{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
 \end{equation}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{i}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{d}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
+L_T(T) \equiv \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
@@
 \begin{equation}
-B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
-\end{equation}
-
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 
-We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+We assert that the status code $R_{s'}$ is an integer.
+\begin{equation}
+R_{s'} \in \mathbb{P}
+\end{equation}
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
 
 The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_{S}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}^*(B_\mathbf{U}))) & \wedge \\
-H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
-H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
-H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_{T}^*$ and $L_{H}^*$ being element-wise sequence transformations, thus:
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_{\mathrm{p}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-H_{\mathrm{r}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
-H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & H_{\mathrm{i}} \in \mathbb{P} & \wedge \\
-H_{\mathrm{l}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & H_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
-H_{\mathrm{x}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{n}} \in \mathbb{B}_{8}
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 
-We define $P(B_{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = H_{\mathrm{p}}
+We define $P(B_H)$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
 \end{equation}
@@
 \begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
+H_i \equiv {{P(H)_H}_i} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_i &\equiv \max(H_i - 3000000, 0)
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-$H_{\mathrm{s}}$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 5000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 
-The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
 
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-& & H_{\mathrm{d}} = D(H) \quad \wedge \\
-& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-& & \lVert H_{\mathrm{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
+& & H_d = D(H) \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l \geqslant 5000  \quad \wedge \\
+& & H_s > {P(H)_H}_s \quad \wedge \\
+& & H_i = {P(H)_H}_i +1 \quad \wedge \\
+& & \lVert H_x \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
 
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items and $\Upsilon^z$ to evaluate to the status code resulting from the transaction.
-These will be formally defined later.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
 \end{equation}
@@
 $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 \begin{equation}
-v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leqslant & T_g \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{gT}}_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_{\mathrm{g}} - g_0
-\end{equation}
-and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+g \equiv T_g - g_0
+\end{equation}
+and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
 \end{eqnarray}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
 \end{eqnarray}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
 \end{eqnarray}
@@
 0 & \text{if} \quad \boldsymbol{\sigma}[a] = \varnothing\\
-\boldsymbol{\sigma}[a]_{\mathrm{b}} & \text{otherwise}
+\boldsymbol{\sigma}[a]_b & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & v \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+I_e & \equiv & e \\
+I_w & \equiv & w
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
-\quad z &\equiv \begin{cases}
+\quad s' &\equiv \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
+\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
+I_s & \equiv & s \\
+I_v & \equiv & \tilde{v} \\
+I_e & \equiv & e \\
+I_w & \equiv & w \\
 \mathbf{t} & \equiv & \{s, r\} \\
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
 \end{eqnarray}
@@
 \begin{itemize}
-\item $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
 \item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_{H}$, the block header of the present block.
-\item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+ \neg I_w \wedge W(w, \boldsymbol{\mu})
 \end{array}
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
 \end{equation}
@@
 \begin{equation}
-D_{J}(\mathbf{c}, i) \equiv \begin{cases}
+D_J(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
+\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
 \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 \boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
+\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
 A' & \equiv & A \\
@@
 \begin{eqnarray}
-B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
-B' & \equiv & P(B_{H})
-\end{eqnarray}
-
-As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
+B_t & \equiv & B'_t + B_d \\
+B' & \equiv & P(B_H)
+\end{eqnarray}
+
+As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block and $B_d$ is its difficulty.
 
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_{H})_{H}, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_H, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\forall_{mathbf{U} \in B_\mathbf{U}}: \\ \nonumber
+\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
+\text{Let} \quad R_b = 3 \times 10^{18}
 \end{equation}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $B_\mathbf{R}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_{s'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{s'} =
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
+m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-\hypertarget{R b}{}
-Formally, we define $R_{\mathrm{b}}$:
-\begin{eqnarray}
-R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
+
+Formally, we define $R_b$:
+\begin{eqnarray}
+R_b(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus we finish by formally defining $R_{\mathrm{l}}$:
-\begin{eqnarray}
-R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
@@
 \begin{equation}
-\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_{mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_{n}\rVert} \quad \mathbf{k}_{n}'[i] & \equiv &
+\forall_n \quad \forall_{i: i < 2\lVert\mathbf{k}_n\rVert} \quad \mathbf{k}_n'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_n[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_n[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{mathbf{I} \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
+g_r &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
+g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+\ell'_E &=& \begin{cases}
+0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
+8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
+8(\ell_E - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
+\ell_B &\equiv& i[0..31] \\
+\ell_E &\equiv& i[32..63] \\
+\ell_M &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_B)] \\
+E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
+M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
 i[x] &\equiv& \begin{cases}
@@
 \end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
+Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
 
@@
 \begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
+\begin{equation}
+\delta_p(\mathbf x)\equiv\begin{cases}
 \mathbf x&\text{if}\ \mathbf x<p\\
@@
 \end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
+x&\equiv&\delta_p(\mathbf x[0..31])\\
+y&\equiv&\delta_p(\mathbf x[32..63])
 \end{eqnarray}
@@
 \end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
+x_0&\equiv&\delta_p(\mathbf x[0..31])\\
+y_0&\equiv&\delta_p(\mathbf x[32..63])\\
+x_1&\equiv&\delta_p(\mathbf x[64..95])\\
+y_1&\equiv&\delta_p(\mathbf x[96..127])
 \end{eqnarray}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
 k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
+g_r&=& 60000k + 40000 \\
 \mathbf o[0..31]&\equiv&\begin{cases}
@@
 \end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
 a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
@@
 \vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
 \end{eqnarray}
@@
 \Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
+g_r &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 \Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
+g_r &=& 40000\\
 \mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
 
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
 \end{equation}
@@
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
+One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
 v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
+T_w &\text{if}\ T_w\in\{27, 28\} \\
+28 - (T_w \bmod 2) &\text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
 \end{equation}
@@
 
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
 
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{a}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
 \midrule
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}_{Z}}{exceptional halting} \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
@@
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_{\mathrm{a}} \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
+(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
+ E_{epoch}(H_i) = \left\lfloor\frac{H_i}{J_{epoch}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_i)
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_i) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_i) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_i - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+ \mathbf{s}_h(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
 \end{equation}
@@
 
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-01-30 Paper.tex
Add missing forward slashes before mathbfs
Fix error from auto find and replace T_gT_p => T_{\mathrm{gT}}_{\mathrm{p}} => T_{\mathrm{g}}T_{\mathrm{p}}

--- 2018-01-30 Paper.tex
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{gT}}_{\mathrm{p}}$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.

### 2018-01-30 Paper.tex
Remove repetitions of a URL.
Add \pagebreak[1]s
Fix an error with \hyperlink{Exceptional_Halting_function_Z}
hypertarget: move to the same line as the targeted text, change to linkdest where not in text mode

--- 2018-01-30 Paper.tex
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 \qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{mathbf{U} \in B_\mathbf{U}}: \\ \nonumber
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 \begin{equation}
-\forall_{mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{mathbf{I} \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 $J_{wordbytes}$ & 4  & Bytes in word. \\
-$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-$J_{epoch}$ & 30000 & Blocks per epoch. \\
-$J_{mixbytes}$ & 128 & mix length in bytes. \\
-$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
+\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule

### 2018-01-30 Paper.tex

--- 2018-01-30 Paper.tex
@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
+% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
+
+\makeatletter
+ \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
+\makeatother
+\usepackage{seqsplit}
 
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_z)
 \end{equation}

### 2018-01-30 Paper.tex

@@
 
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\toprule
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
+\bottomrule
+\end{tabu*}
+
 \end{document}

### 2018-01-30 Paper.tex

@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
 \begin{eqnarray}

### 2018-01-30 Paper.tex

@@
 \begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_z)
+R \equiv (R_z, R_u, R_b, R_\mathbf{l})
 \end{equation}

### 2018-01-30 Paper.tex

@@
 \midrule
-$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
 \linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\

### 2018-01-30 Paper.tex
\hypertarget{R l}{}Thus

--- 2018-01-30 Paper.tex
@@
 If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
-
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
 
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
-\begin{equation}
-R \equiv (R_z, R_u, R_b, R_\mathbf{l})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
 \end{equation}
@@
 
-We assert that the status code $R_z$ is an integer.
-\begin{equation}
-R_z \in \mathbb{P}
+We assert that the status code $R_{s'}$ is an integer.
+\begin{equation}
+R_{s'} \in \mathbb{P}
 \end{equation}
@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 
-\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
+Thus we are able to define the block header validity function $V(H)$:
 \begin{eqnarray}
@@
 
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
 g_0 & \leqslant & T_g \quad \wedge \\
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
 \begin{eqnarray}
@@
 \end{equation}
-\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
 \begin{eqnarray}
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_a$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_o$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_p$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_H$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_w$, the permission to make modifications to the state.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
@@
 
-As previously detailed, $\mathbf{R}[n]_z$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_z$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_z =
-\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+We define $\mathbf{R}[n]_{s'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{s'} =
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \midrule
-\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{wordbytes}$ & 4  & Bytes in word. \\
 $J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
@@
 
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\subsection{Deletion of an Account Dispite Out-of-gas}
+
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
-\bottomrule
-\end{tabu*}
-
 \end{document}

### 2018-01-30 Paper.tex
Move \linkdest{Transaction Receipt}{}

@@
 
-\subsubsection{Transaction Receipt}
+\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
 
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_z)
 \end{equation}
@@
 
-\hypertarget{R l}{}Thus we finish by formally defining $R_l$:
+Thus we finish by formally defining $R_l$:
 \begin{eqnarray}

### 2018-01-30 Paper.tex
Readd JUMPI

--- 2018-01-30 Paper.tex
@@
 \midrule
-0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+\linkdest{JUMP}{JUMP}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
-\midrule
-0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
+\midrule
+\linkdest{JUMPI}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\

### 2018-01-30 Paper.tex
Append {} after \linkdest{JUMP}, remove JUMP in second braces

--- 2018-01-30 Paper.tex
@@
 \midrule
-\linkdest{JUMP}{JUMP}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
@@
 \midrule
-\linkdest{JUMPI}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\

### 2018-01-30 Paper.tex

@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+\begin{equation}
+R \equiv (R_z, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
 \begin{eqnarray}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
 \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
@@
 \midrule
-\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
-\midrule
-\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+\midrule
+0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
 &&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\

### 2018-01-30 Paper.tex

@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.

### 2018-01-30 Paper.tex

@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+
 \begin{equation}

### 2018-01-30 Paper.tex

@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.

### 2018-01-31 Paper.tex

@@
 
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
 Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$, the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
 

### 2018-01-31 Paper.tex
Fix the calculation of cache size in ethash
See https://github.com/ethereum/yellowpaper/issues/517

This has to be ported to all protocol versions.
Remove \small and \tiny in math mode because they have no effects
and caused error messages like
```
LaTeX Font Warning: Command \tiny invalid in math mode on input line 623.
```

--- 2018-01-30 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-\usepackage{tabu} %requires array.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
-
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-
-\input{Version.tex}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
-Formally, we can refer to a block $B$:
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
-
-\begin{equation}
-R \equiv (R_z, R_u, R_b, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_z$ is an integer.
-\begin{equation}
-R_z \in \mathbb{P}
-\end{equation}
+L_R(R) \equiv (\mathtt{TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
 
@@
 
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
 \begin{equation}
-p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
+p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
 \end{equation}
@@
 
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
-\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\begin{equation}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
 \end{equation}
@@
 \begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}
@@
 & & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp (in Unix time) of block $H$ at the block's inception, and must fulfil the relation:
+& & H_l \geqslant 125000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
 \begin{equation}
@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
+$G$ is defined in Appendix \ref{app:fees}.
+
+%todo Explain g_d reason?
 
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
 g_0 & \leqslant & T_g \quad \wedge \\ 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
 \begin{eqnarray}
 \Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
 \end{eqnarray}
@@
 
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
+\section{Contract Creation} \label{ch:create}
+
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{TRIE}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**} < c \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+Where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
 \begin{eqnarray}
@@
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+I_e & \equiv & e
+\end{eqnarray}
+
+where $c$ is the code-deposit cost:
 \begin{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
-g^{**} - c & \text{otherwise} \\
-\end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
-\end{equation}
-Note that we need to differentiate between the value that is to be transferred, $v$, from the value apparent in the execution context, $\tilde{v}$, for the {\small DELEGATECALL} instruction.
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \mathbf{d}, e)
+\end{equation}
 
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{KEC}(()), \mathtt{TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
+I_v & \equiv & v \\
 I_e & \equiv & e \\
-I_w & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+\text{Let} \; \mathtt{KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 
+%More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
+%
+%\begin{itemize}
+%\item {\small SHA3} costs $G_{sha3}$ gas
+%\item {\small SLOAD} costs $G_{sload}$ gas
+%\item {\small BALANCE} costs $G_{balance}$ gas
+%\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
+%\begin{itemize}
+%\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
+%\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
+%\item $d = 1$ otherwise.
+%\end{itemize}
+%\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
+%\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
+%\item {\small STOP} costs $G_{stop}$ gas
+%\item {\small SUICIDE} costs $G_{suicide}$ gas
+%\item All other operations cost $G_{step}$ gas.
+%\end{itemize}
+%
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+
 See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
 
+%Whenever a higher memory index is referenced, the fee difference to take it to the higher usage from the original (lower) usage is charged.
+Notably, because {\small MSTORE} and {\small MLOAD} operate on word lengths, they implicitly increase the highest-accessed index to 31 greater than their target index.
+
 \subsection{Execution Environment}
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_a$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_o$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_p$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_H$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_w$, the permission to make modifications to the state.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will define it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
 \boldsymbol{\mu}_g & \equiv & g \\
@@
 \begin{eqnarray}
-\\ \nonumber
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{\mathbf{U} \in B_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_z$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \mathbf{R}[n]_\mathbf{l} = 
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-We define $\mathbf{R}[n]_z$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_z =
-\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}
@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+
+Where $\mathbb{Y}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
 
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{equation}
-\forall_{\mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall_{I \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{\mathbf{I} \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
+\section{Signing Transactions}\label{app:signing}
+
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{ECDSASIGN}$, $\mathtt{ECDSARESTORE}$ and $\mathtt{ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
+where:
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
+
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_p$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
-k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_r&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
-\end{eqnarray}
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
-
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
-\end{align}
-where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
-
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+h(T) & \equiv & \mathtt{KEC}( L_S(T) )
 \end{eqnarray}
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_w &\text{if}\ T_w\in\{27, 28\} \\
-28 - (T_w \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
 \begin{equation}
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. 
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{blockhash}{}0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
@@
 \midrule
-\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
 &&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
@@
 \end{cases}$ \\
-&&&& \linkdest{A r}{}$A'_{r} \equiv A_{r} + \begin{cases}
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
 R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
-\multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account. \\
+&&&& created account, otherwise. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_g - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_g \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
 G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
-\midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& same account as at present, simply that the code is overridden altered.\\
+\midrule
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+0 & \text{otherwise}
+\end{cases}$ \\\bottomrule
 \end{tabular*}
 
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
-\bottomrule
-\end{tabular*}
-
-\hypertarget{GenesisBlock}{}
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
 \section{Genesis Block}\label{app:genesis}
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
 The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the initial timestamp of the genesis block.
+Also $time$ will be set to the intial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 \midrule
-\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
 \begin{equation}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
 \end{equation}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+Essentially, we maintain a "mix" $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
 $J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
-\bottomrule
-\end{tabu*}
-
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2018-01-31 Paper.tex
Fix the cache size computation in ethash
See https://github.com/ethereum/yellowpaper/issues/517

@@
 \begin{equation}
-L_R(R) \equiv (\mathtt{TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 \forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
-m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
+m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
 \end{eqnarray}
@@
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
 H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
@@
 \begin{equation}
-p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
+p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
 \end{equation}
@@
 \begin{equation}
-\mathtt{TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
 \end{equation}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_H)) = H_p
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
 \end{equation}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{TRIE}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{KEC}(()), \mathtt{TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
 \boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
@@
 I_e & \equiv & e \\
-\text{Let} \; \mathtt{KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 
-We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus $\mathtt{BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
 
@@
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
 \end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) ...
+s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
 \end{eqnarray}
@@
 \begin{equation}
-\mathtt{RLP}(i : i \in \mathbb{P}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
+\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
 \end{equation}
@@
 
-It is defined as the function $\mathtt{HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
-
-\begin{eqnarray}
-\mathtt{HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
+It is defined as the function $\mathtt{\tiny HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
+
+\begin{eqnarray}
+\mathtt{\tiny HP}(\mathbf{x}, t): \mathbf{x} \in \mathbb{Y} & \equiv & \begin{cases}
 (16f(t), 16\mathbf{x}[0] + \mathbf{x}[1], 16\mathbf{x}[2] + \mathbf{x}[3], ...) &
@@
 For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-Importantly in the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
+Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
 \begin{eqnarray}
@@
 g_r &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
 \text{if} \quad |\mathbf{o}| = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
 \mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
@@
 g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_\mathbf{d})\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
@@
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
 \end{eqnarray}
@@
 \begin{eqnarray}
-\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
+\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
 \end{eqnarray}
@@
 
-in the range $(1, \mathtt{secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
-
-We assert the functions $\mathtt{ECDSASIGN}$, $\mathtt{ECDSARESTORE}$ and $\mathtt{ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 \begin{eqnarray}
-0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
-0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
 v \in \{27,28\}
@@
 \begin{eqnarray}
-\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
 \end{eqnarray}
@@
 \begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_r) \big) \big)
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
 \end{equation}
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_S(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
 \end{eqnarray}
@@
 G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{ECDSASIGN}(h(T), p_r)
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 \begin{equation}
-S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
 \end{equation}
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
+\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
 \end{equation}
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
 \end{equation}

### 2018-01-31 Paper.tex

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
     Dr. Gavin Wood\\
-    Co-Founder \& Lead, Ethereum Project\\
-    gavin@ethereum.org
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 
+\newcommand{\mindifficulty}{D_0}
+\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
+\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
+\newcommand{\diffadjustment}{x}
+
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{cases}
-131072 & \text{if} \quad H_i = 0\\
-{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
-\max \{ 131072, x \} & \text{otherwise}\\
-\end{cases}
+D(H) \equiv \begin{dcases}
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\end{dcases}
 \end{equation}
@@
 \begin{equation}
-x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\mindifficulty \equiv 131072
+\end{equation}
+\begin{equation}
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
 \end{equation}
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{equation}
-g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
-\end{equation}
+\begin{align}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
+\end{align}
 
 where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G$ is defined in Appendix \ref{app:fees}.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G$ is fully defined in Appendix \ref{app:fees}.
 
@@
 
-%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
-%\begin{eqnarray}
-%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
-%\end{eqnarray}
-
 \section{Contract Creation} \label{ch:create}
 
-There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 
-where $c$ is the code-deposit cost:
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+
+Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+
+If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
 \begin{equation}
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
-$I_H$ has no special treatment and is determined from the blockchain.
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
-
-No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
+
+The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
+However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{align}
+\quad g' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{align}
+
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
+
+Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
 
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
-\begin{equation}
-(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+For the present context, we will defined it as:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+\end{equation}
+
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, r)
 \end{equation}
@@
 
-We declare that a signature is invalid unless the following is true:
-\begin{eqnarray}
-0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
-v \in \{27,28\}
-\end{eqnarray}
-
+\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
+
+We declare that a signature is invalid unless all the following conditions are true:
+\begin{align}
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28\}
+\end{align}
 where:
-\begin{eqnarray}
-\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
-\end{eqnarray}
+\begin{align}
+\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
+\end{align}
 
@@
 $R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
-$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
 $G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
+$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 
+Another useful function is ``all but one 64th'' function~$L$ defined as:
+
+\begin{equation}
+L(n) \equiv n - \lfloor n / 64 \rfloor
+\end{equation}
+
 \subsection{Instruction Set}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}_g' \equiv g' + \boldsymbol{\mu}_g - L(\boldsymbol{\mu}_g)$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 \end{cases}$ \\
-&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
+&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
-\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
-0 & \text{otherwise}
-\end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overridden altered.\\
+&&&& same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\midrule
-0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+&&&& the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& Otherwise exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
+&&&& and ninth parameters to the call $\Theta$.\\
+&&&& This means that the recipient is in fact the same account as at present, simply\\
+&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
+\midrule
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \\\bottomrule
+\end{cases}$ \\
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\bottomrule
 \end{tabular*}
@@
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+E_{prime}(x - 1 \cdot y, y) & \text{otherwise}
 \end{cases}

### 2018-02-01 Paper.tex
Remove {}

@@
 \end{equation}
-where $v$ is the account validity function:
+\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
 \begin{equation}
 \quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+An account is \textit{empty} when it has no code, zero nonce and zero balance:
+\begin{equation}
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_n = 0 \wedge \boldsymbol{\sigma}[a]_b = 0
+\end{equation}
+Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
+
+An account is \textit{dead} when its account state is non-existent or empty:
+\begin{equation}
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+\item[value]\linkdest{tx_value_T__v} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation}
-\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
-\end{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation*}
+x \equiv \begin{cases}
+1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+2 & \text{otherwise}
+\end{cases}
+\end{equation*}
+\begin{align}
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_i &\equiv \max(H_i - 3000000, 0)
+\end{align}
 
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
 \quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases}
+\end{cases} \\
+\quad s' &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+1 & \text{otherwise}
+\end{cases} \\
+\nonumber \text{where} \\
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+g^{**} & \text{otherwise}
+\end{cases} \\ \nonumber
+s' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-g^{**} & \text{otherwise}
+1 & \text{otherwise}
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
-\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_a & \equiv & r \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\mu}_g' \equiv g' + \boldsymbol{\mu}_g - L(\boldsymbol{\mu}_g)$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
-&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
-&&&& created account, otherwise. \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\

### 2018-02-01 Paper.tex
\linkdest{block_validity}{}, \linkdest{new_state_H__r}{}

@@
 \begin{description}
-\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
-\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value]\linkdest{tx_value_T__v} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+\linkdest{new_state_H__r}{}H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}

### 2018-02-01 Paper.tex
\hypertarget{Omega}{}

--- 2018-02-01 Paper.tex
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
-Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}

### 2018-02-01 Paper.tex
\hypertarget{block_reward_R__b}{}
\hypertarget{Gamma}{}

@@
 
-\hypertarget{block-reward-R-b}{block-reward-R-b}We define the block reward as 3 Ether:
+\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
 \begin{equation}
@@
 
-\hypertarget{Gamma}{Gamma}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}

### 2018-02-01 Paper.tex
\hypertarget{Phi}{}
\hypertarget{Pi}{}

@@
 
-\hypertarget{Phi}{Phi}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
@@
 
-\hypertarget{Pi}{Pi}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}

### 2018-02-01 Paper.tex
Move \hypertarget{h T}{}

@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 
-\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
+We define the block reward as 3 Ether:
 \begin{equation}
@@
 
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
@@
 
-\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}
@@
 
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.

### 2018-02-01 Paper.tex
_{\mathrm{t}} => _{t}

@@
 \begin{equation}
-\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+\item[nonce] \linkdest{account_{\mathrm{nonce}}}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
-\begin{equation}
-L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
+\begin{equation}
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_S$:
-\begin{equation}
-L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_{S}$:
+\begin{equation}
+L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
 \item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-L_T(T) \equiv \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
-(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+L_{T}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{i}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{d}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
 \end{cases}
@@
 \begin{array}[t]{lclclc}
-T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
-T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
-T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 \begin{equation}
-B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
 \end{equation}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
+\end{equation}
+
+The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
 \end{equation}
@@
 
-We assert that the status code $R_{s'}$ is an integer.
-\begin{equation}
-R_{s'} \in \mathbb{P}
-\end{equation}
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\begin{equation}
+R_{\mathrm{z}} \in \mathbb{P}
+\end{equation}
+
+We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
 
 The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}
-H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_{S}(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}^*(B_\mathbf{U}))) & \wedge \\
+H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
+H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
 
@@
 
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_{T}^*$ and $L_{H}^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \begin{array}[t]{lclclcl}
-H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
-H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
-H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
-H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
-H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
+H_{\mathrm{p}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+H_{\mathrm{r}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
+H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & H_{\mathrm{i}} \in \mathbb{P} & \wedge \\
+H_{\mathrm{l}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & H_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
+H_{\mathrm{x}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{n}} \in \mathbb{B}_{8}
 \end{array}
@@
 
-We define $P(B_H)$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+We define $P(B_{H})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = H_{\mathrm{p}}
 \end{equation}
@@
 \begin{equation}
-H_i \equiv {{P(H)_H}_i} + 1
+H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
 \end{equation}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
 \end{align}
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp of block $H$ and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000
+\end{eqnarray}
+
+$H_{\mathrm{s}}$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
 \end{equation}
@@
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
 
 Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
-& & H_d = D(H) \quad \wedge \\
-& & H_g \le H_l  \quad \wedge \\
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
-& & H_l \geqslant 5000  \quad \wedge \\
-& & H_s > {P(H)_H}_s \quad \wedge \\
-& & H_i = {P(H)_H}_i +1 \quad \wedge \\
-& & \lVert H_x \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
+& & H_{\mathrm{d}} = D(H) \quad \wedge \\
+& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
+& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+& & \lVert H_{\mathrm{x}} \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
 \end{equation}
@@
 $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 \begin{equation}
-v_0 \equiv T_g T_p + T_v
+v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
-The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
-\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
-m & \equiv & {B_H}_c
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_{H}}_{\mathrm{c}}
 \end{eqnarray}
@@
 
-And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
 \end{eqnarray}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
 \end{equation}
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
 \end{eqnarray}
@@
 \begin{eqnarray}
-I_a & \equiv & a \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_{\mathrm{a}} & \equiv & a \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_s & \equiv & s \\
-I_v & \equiv & v \\
+I_{\mathrm{s}} & \equiv & s \\
+I_{\mathrm{v}} & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_a & \equiv & r \\
-I_o & \equiv & o \\
-I_p & \equiv & p \\
+I_{\mathrm{a}} & \equiv & r \\
+I_{\mathrm{o}} & \equiv & o \\
+I_{\mathrm{p}} & \equiv & p \\
 I_\mathbf{d} & \equiv & \mathbf{d} \\
-I_s & \equiv & s \\
-I_v & \equiv & \tilde{v} \\
-I_e & \equiv & e \\
-I_w & \equiv & w \\
+I_{\mathrm{s}} & \equiv & s \\
+I_{\mathrm{v}} & \equiv & \tilde{v} \\
+I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{w}} & \equiv & w \\
 \mathbf{t} & \equiv & \{s, r\} \\
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
 \end{eqnarray}
@@
 \begin{itemize}
-\item $I_a$, the address of the account which owns the code that is executing.
-\item $I_o$, the sender address of the transaction that originated this execution.
-\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
 \item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
 \item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_H$, the block header of the present block.
-\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_w$, the permission to make modifications to the state.
+\item $I_{H}$, the block header of the present block.
+\item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_g & \equiv & g \\
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
- \neg I_w \wedge W(w, \boldsymbol{\mu})
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+ \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
 \end{array}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
+\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
 \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 \boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
-\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
+\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
 A' & \equiv & A \\
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_{H})_{H}, 6)
 \end{equation}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_H}_g = \ell(\mathbf{R})_u
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
 Formally we define the function $\Omega$:
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\boldsymbol{\sigma}'[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
+\boldsymbol{\sigma}'[U_{\mathrm{c}}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
 \end{eqnarray}
@@
 \begin{equation}
-\text{Let} \quad R_b = 3 \times 10^{18}
+\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
 \end{equation}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
-\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
-B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_{s'}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
+In the case of $B_\mathbf{R}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \end{itemize}
-
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R b}{}
+Formally, we define $R_{\mathrm{b}}$:
+\begin{eqnarray}
+R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 
-Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_{\mathrm{l}}$:
+\begin{eqnarray}
+R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
-(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
 
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 3000\\
+g_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_\mathbf{d}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_r &=& \Big\lfloor\frac{f\big(\max(\ell_M,\ell_B)\big)\max(\ell'_E,1)}{G_{quaddivisor}}\Big\rfloor \\
+g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_E &=& \begin{cases}
-0 & \text{if}\ \ell_E\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_E\le 32 \wedge E \neq 0 \\
-8(\ell_E - 32) + \lfloor \log_2(i[(96+\ell_B)..(127+\ell_B)]) \rfloor & \text{if}\ 32 < \ell_E \wedge i[(96 + \ell_B)..(127 + \ell_B)]\neq 0 \\
-8(\ell_E - 32) & \text{otherwise} \\
+\ell'_{E} &=& \begin{cases}
+0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
+8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
+8(\ell_{E} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_M} \\
-\ell_B &\equiv& i[0..31] \\
-\ell_E &\equiv& i[32..63] \\
-\ell_M &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_B)] \\
-E &\equiv& i[(96+\ell_B)..(95+\ell_B+\ell_E)] \\
-M &\equiv& i[(96+\ell_B+\ell_E)..(95+\ell_B+\ell_E+\ell_M)] \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
+\ell_{B} &\equiv& i[0..31] \\
+\ell_{E} &\equiv& i[32..63] \\
+\ell_{M} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{B})] \\
+E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
+M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
 i[x] &\equiv& \begin{cases}
@@
 \begin{equation}
-C_1\equiv\{(X,Y)\in F_p\times F_p\mid Y^2=X^3+3\}\cup\{(0,0)\}
+C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
 \end{equation}
@@
 
-Let $F_{p^2}$ be a field $F_p[i]/(i+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_p$.
-\begin{equation}
-\delta_p(\mathbf x)\equiv\begin{cases}
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
+\begin{equation}
+\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
 \mathbf x&\text{if}\ \mathbf x<p\\
@@
 \end{cases}\\
-x&\equiv&\delta_p(\mathbf x[0..31])\\
-y&\equiv&\delta_p(\mathbf x[32..63])
+x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
 \end{eqnarray}
@@
 \end{cases}\\
-x_0&\equiv&\delta_p(\mathbf x[0..31])\\
-y_0&\equiv&\delta_p(\mathbf x[32..63])\\
-x_1&\equiv&\delta_p(\mathbf x[64..95])\\
-y_1&\equiv&\delta_p(\mathbf x[96..127])
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
 \end{eqnarray}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_j=\varnothing\vee b_j=\varnothing))\\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
 k &=& \dfrac{|I_\mathbf{d}|}{192} \\
-g_r&=& 60000k + 40000 \\
+g_{\mathrm{r}}&=& 60000k + 40000 \\
 \mathbf o[0..31]&\equiv&\begin{cases}
@@
 \end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
 a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
@@
 \vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
 \end{eqnarray}
@@
 \Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_r &=& 500\\
+g_{\mathrm{r}} &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 \Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_r &=& 40000\\
+g_{\mathrm{r}} &=& 40000\\
 \mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 
-It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
 
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
 \end{align}
@@
 
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+L_{S}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_r)$ is defined as:
-\begin{eqnarray}
-G(T, p_r) \equiv T \quad \text{except:} \\
-(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
+\begin{eqnarray}
+G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
 \end{eqnarray}
@@
 
-%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
@@
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
 \midrule
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
@@
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 &&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1, I_w) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
 &&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
@@
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1, I_w)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_{\mathrm{a}} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}

### 2018-02-02 Paper.tex

@@
 \begin{description}
-\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
+\item[ommersHash]\linkdest{beneficiary_H__c}{} The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
 \item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
@@
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
+\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
+\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
+\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
-\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
+\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
 
-The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
+Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 
-The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of a logger's address, $O_{\mathrm{a}}$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
 \begin{equation}
@@
 
-With $L_{T}^*$ and $L_{H}^*$ being element-wise sequence transformations, thus:
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 
-The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}
@@
 
-$H_{\mathrm{s}}$ is the timestamp of block $H$ and must fulfil the relation:
+$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
 \begin{equation}
@@
 
-Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
@@
 
-The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
-$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
-$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
-Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
+\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
 g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
@@
 \begin{itemize}
-\item $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item $I_{H}$, the block header of the present block.
-\item $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_{H})_{H}, 6)
+\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} V({\mathbf{U}}) \; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{B_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
-\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
+\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
+\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
@@
 
-Thus we finish by formally defining $R_{\mathrm{l}}$:
+\hypertarget{R l}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
 \begin{eqnarray}
@@
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
 \mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 \midrule
-0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
@@
 \end{cases}$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
 R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 \midrule
-0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_{\mathrm{a}} \}$ \\
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}

### 2018-02-02 Paper.tex
Fix in-text citations: *by* \cite{EIP-#}

--- 2018-02-01 Paper.tex
@@
 
-\subsubsection{Transaction Receipt}
-
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
+\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
 Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
-The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
-\begin{equation}
-R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$, the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
+
+\begin{equation}
+R \equiv (R_z, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 
-The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+Note that $\mindifficulty$ is the difficulty of the genesis block.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
+In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}
@@
 
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.

### 2018-02-02 Paper.tex
Add parentheses around \cite{Davey2002_zbMATH01748069}
Remove \\ in 0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba

@@
 Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
 In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
 Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
@@
 
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-02-02 Paper.tex
Move \hypertarget{Exceptional_Halting_function_Z}{}
The column actually broke on the preceding line, causing hyperlinks to direct to the bottom of the column.

@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-\hypertarget{Exceptional_Halting_function_Z}{}
-\subsubsection{Exceptional Halting}
+
+\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}
 
@@
 
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e\\4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.

### 2018-02-02 Paper.tex
Move hypertargets to the same line as section headings, and rename some targets

@@
 \end{description}
-\hypertarget{rlp}{}
-\section{Recursive Length Prefix}\label{app:rlp}
+
+\section{Recursive Length Prefix}\label{app:rlp}\hypertarget{rlp}{}
 This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
@@
 \end{itemize}
-\hypertarget{R b}{}
-Formally, we define $R_b$:
+
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_b$:
 \begin{eqnarray}
@@
 
-\hypertarget{R l}{}Thus we finish by formally defining $R_l$:
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_l$:
 \begin{eqnarray}
@@
 All remaining nibbles (now an even number) fit properly into the remaining bytes.
-\hypertarget{trie}{}
-\section{Modified Merkle Patricia Tree}\label{app:trie}
+
+\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
@@
 
-\hypertarget{GenesisBlock}{}
-\section{Genesis Block}\label{app:genesis}
+\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
 

### 2018-02-02 Paper.tex
Add RLP R_b and R_l word defs.

@@
 
-If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 
-If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 

### 2018-02-02 Paper.tex
Remove duplicate hypertargets

--- 2018-02-02 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
-Formally, we can refer to a block $B$:
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
 \begin{equation}
@@
 
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
-
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
+\subsubsection{Transaction Receipt}
+
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
 Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$, the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
-
-\begin{equation}
-R \equiv (R_z, R_u, R_b, R_\mathbf{l})
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
 \end{equation}
@@
 
-The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
 \begin{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
 \begin{eqnarray}
@@
 \qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{\mathbf{U} \in B_{\mathbf{U}}}: \\ \nonumber
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[U_c] & = & \begin{cases}
@@
 \end{eqnarray}
+
+%Uncomment out after merging \hypertargets for v, r, and s.
+%Thus: 
+%\begin{eqnarray}\hypertarget{T__w}{}\hypertarget{T__r}{}\hypertarget{T__s}{}
+%T_{\mathrm{w}} = \hyperlink{v}{v}\\
+%T_{\mathrm{r}} = \hyperlink{r}{r}\\
+%T_{\mathrm{s}} = \hyperlink{s}{s}
+%\end{eqnarray}
 
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
@@
 \midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
@@
 \midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
 &&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
@@
 \end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad n \\

### 2018-02-02 Paper.tex

--- 2018-02-02 Paper.tex
@@
 We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction.
-
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}
+\hypertarget{Exceptional_Halting_function_Z}{}
+\subsubsection{Exceptional Halting}
 
@@
 \end{description}
-
-\section{Recursive Length Prefix}\label{app:rlp}\hypertarget{rlp}{}
+\hypertarget{rlp}{}
+\section{Recursive Length Prefix}\label{app:rlp}
 This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
@@
 \end{itemize}
-
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_b$:
+\hypertarget{R b}{}
+Formally, we define $R_b$:
 \begin{eqnarray}
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_l$:
+\hypertarget{R l}{}Thus we finish by formally defining $R_l$:
 \begin{eqnarray}
@@
 All remaining nibbles (now an even number) fit properly into the remaining bytes.
-
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
+\hypertarget{trie}{}
+\section{Modified Merkle Patricia Tree}\label{app:trie}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
@@
 
-\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
+\hypertarget{GenesisBlock}{}
+\section{Genesis Block}\label{app:genesis}
 

### 2018-02-02 Paper.tex

@@
 
-We define the block reward as 3 Ether:
+\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
 \begin{equation}
@@
 
-We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
@@
 
-And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
@@
 
-Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}

### 2018-02-02 Paper.tex

@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
 \end{equation}
@@
 \begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
+\item[nonce] \linkdest{account_{\mathrm{nonce}}}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
-\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
-\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
+\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
+\item[ommersHash]\linkdest{beneficiary_H__c}{} The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
+\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
+\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
+\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
+\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$, the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_z$:
-
-\begin{equation}
-R \equiv (R_z, R_u, R_b, R_\mathbf{l})
-\end{equation}
-
-The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
+\end{equation}
+
+The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
 \end{equation}
@@
 
-We assert that the status code $R_z$ is an integer.
-\begin{equation}
-R_z \in \mathbb{P}
-\end{equation}
-
-We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\begin{equation}
+R_{\mathrm{z}} \in \mathbb{P}
+\end{equation}
+
+We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
 
@@
 \begin{equation}
-O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
+\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
@@
 
-The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
-\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
 \end{eqnarray}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_i$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
-& & H_l \geqslant 5000
-\end{eqnarray}
-
-$H_s$ is the timestamp (in Unix time) of block $H$ at the block's inception, and must fulfil the relation:
-\begin{equation}
-H_s > {P(H)_H}_s
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000
+\end{eqnarray}
+
+$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
+\begin{equation}
+H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
 \end{equation}
@@
 
-The nonce, $H_n$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
+The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
-g_0 & \leqslant & T_g \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
-T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
+\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
+g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
 \end{array}
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
-\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_a$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_o$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_p$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_\mathbf{b}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_H$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_w$, the permission to make modifications to the state.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 \begin{equation}
-\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} V({\mathbf{U}}) \; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
-\qquad\forall_{\mathbf{U} \in B_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
+\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
+\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
-R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
 \end{eqnarray}
@@
 
-As previously detailed, $\mathbf{R}[n]_z$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-We define $\mathbf{R}[n]_z$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_z =
+We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{\mathrm{z}} =
 \Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
@@
 \hypertarget{R b}{}
-Formally, we define $R_b$:
-\begin{eqnarray}
-R_b(\mathbf{x}) & \equiv & \begin{cases}
+Formally, we define $R_{\mathrm{b}}$:
+\begin{eqnarray}
+R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 
-\hypertarget{R l}{}Thus we finish by formally defining $R_l$:
-\begin{eqnarray}
-R_l(\mathbf{x}) & \equiv & \begin{cases}
+\hypertarget{R l}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
+\begin{eqnarray}
+R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{\mathbf{I} \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 
-For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
-\end{equation}
-
-The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+\end{equation}
+
+\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
+One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_S(T) & \equiv & \begin{cases}
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+L_{S}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{blockhash}{}0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 \linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&& \linkdest{A r}{}$A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_e = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
@@
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 \linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
-(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
-(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
 0 & \text{otherwise}

### 2018-02-02 Paper.tex
Final touches
See PR https://github.com/ethereum/yellowpaper/pull/515

@@
 \begin{description}
-\item[nonce] \linkdest{account_{\mathrm{nonce}}}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
@@
 \item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash]\linkdest{beneficiary_H__c}{} The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.

### 2018-02-02 Paper.tex

@@
 Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 \cite{EIP-2}.
+The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
 In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 \cite{EIP-649} and delayed and implemented earlier in EIP-2 \cite{EIP-2}.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks \cite{EIP-100}.
+This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
 Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
@@
 \midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements \cite{Davey2002_zbMATH01748069}.\\
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
 \bottomrule

### 2018-02-02 Paper.tex

@@
 \end{itemize}
-\hypertarget{R b}{}
-Formally, we define $R_{\mathrm{b}}$:
+
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
 \begin{eqnarray}
@@
 
-\hypertarget{R l}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
 \begin{eqnarray}

### 2018-02-02 Paper.tex

@@
 \end{equation}
-where $v$ is the account validity function:
+\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
 \begin{equation}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value]\linkdest{tx_value_T__v} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.

### 2018-02-04 Paper.tex
2 \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}

--- 2018-02-02 Paper.tex
@@
 
-When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
+When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
 It is formally defined in section \ref{ch:block}.
@@
 \begin{equation}
-f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}

### 2018-02-04 Paper.tex
\linkdest{r}{r} \linkdest{s}{s}

--- 2018-02-02 Paper.tex
@@
 \begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
+0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
 v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}

### 2018-02-04 Paper.tex
\hypertarget{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}

@@
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hypertarget{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
 v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}

### 2018-02-04 Paper.tex
hypertarget => linkdest
I tested the former and it direct to the next line. Thus linkdest must be used whenever it encapsulates math mode, or any part, or is included in it.

@@
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \hypertarget{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.

### 2018-02-04 Paper.tex
Thus => Reiterating from \hyperlink{T__w_T__r_T__s}{previously}

--- 2018-02-04 Paper.tex
@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 \begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
+\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
-\end{equation}
-
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$ and the status code of the transaction, $R_{s'}$:
+\begin{equation}
+R \equiv (R_u, R_b, R_\mathbf{l}, R_{s'})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (0 \in \mathbb{B}_{256}, R_u, R_b, R_\mathbf{l})
 \end{equation}
@@
 
-We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+We assert that the status code $R_{s'}$ is an integer.
+\begin{equation}
+R_{s'} \in \mathbb{P}
+\end{equation}
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 
-The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 
-\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
+\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
+H'_i &\equiv \max(H_i - 3000000, 0)
 \end{align}
 
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 5000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 
-The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 
-\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-& & H_{\mathrm{d}} = D(H) \quad \wedge \\
-& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-& & \lVert H_{\mathrm{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+Thus we are able to define the block header validity function $V(H)$:
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
+& & H_d = D(H) \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l \geqslant 5000  \quad \wedge \\
+& & H_s > {P(H)_H}_s \quad \wedge \\
+& & H_i = {P(H)_H}_i +1 \quad \wedge \\
+& & \lVert H_x \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
 
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
 g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 {} & + G_{transaction}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leqslant & T_g \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{s}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\Upsilon^s(\boldsymbol{\sigma}, T) & \equiv & s'
 \end{eqnarray}
@@
 \end{equation}
-\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+\begin{eqnarray}
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
 I_\mathbf{d} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & v \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+I_e & \equiv & e \\
+I_w & \equiv & w
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
+\item $I_w$, the permission to make modifications to the state.
 \end{itemize}
@@
 
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
 \boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
@@
 \begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} V({\mathbf{U}}) \; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 
-\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
+We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 3 \times 10^{18}
 \end{equation}
@@
 
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+We define $\mathbf{R}[n]_{s'}$ in a similar manner.
+\begin{equation}
+\mathbf{R}[n]_{s'} =
+\Upsilon^{s}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}
@@
 
-\section{Recursive Length Prefix}\label{app:rlp}\hypertarget{rlp}{}
+\section{Recursive Length Prefix}\label{app:rlp}
 This is a serialisation method for encoding arbitrarily structured binary data (byte arrays).
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\end{equation}
+
+If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
-\begin{eqnarray}
-R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
+Formally, we define $R_b$:
+\begin{eqnarray}
+R_b(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
-\begin{eqnarray}
-R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
@@
 
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
+\section{Modified Merkle Patricia Tree}\label{app:trie}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
 It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 \begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
+One operates without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$.
 The other operates on nine elements:
 \begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_n, T_p, T_g, T_t, T_v, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
+
+Reiterating from \hyperlink{T__w_T__r_T__s}{previously}: 
+\begin{eqnarray}
+\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
+\linkdest{T__r}{T_{\mathrm{r}} = \hyperlink{r}{r}\\
+\linkdest{T__s}{T_{\mathrm{s}} = \hyperlink{s}{s}
 \end{eqnarray}
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
-\midrule
-\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+\midrule
+0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1, I_w)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
 \midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
 &&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
@@
 \midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
+(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
 \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad n \\
@@
 
-\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
+\section{Genesis Block}\label{app:genesis}
 
@@
 
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\subsection{Deletion of an Account Dispite Out-of-gas}
+
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
-
 \end{document}

### 2018-02-04 Paper.tex
\hyperlink{trie}{trie} & \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}

--- 2018-02-04 Paper.tex
@@
 
-\subsection{The Transaction} \label{subsec:transaction}
+\subsection{The Transaction} \label{ch:transaction}
 
@@
 \item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
 This is expanded in Appendix \ref{app:signing}.
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
 

### 2018-02-04 Paper.tex
Ommer footnote, link words to maths.

@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, in that it isn't commonly used) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}}})
+\end{equation}
+
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
 

### 2018-02-05 Paper.tex

--- 2018-02-04 Paper.tex
@@
 
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
 \end{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
@@
 L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{i}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{d}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
 \begin{equation}
@@
 T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 
-The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{equation}
-B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
+\end{equation}
+
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
+Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_{\mathbf{l}}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
 \end{equation}
@@
 \begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
 \end{equation}
@@
 
-%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+
+The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
@@
 \linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_{\mathbf{R}}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(B_{\mathbf{U}}) \big)
 \end{eqnarray}
@@
 x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
+1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
 2 & \text{otherwise}
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items and $\Upsilon^z$ to evaluate to the status code resulting from the transaction.
+We also define $\Upsilon^{\mathrm{g}}$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^{\mathbf{l}}$ to evaluate to the transaction's accrued log items and $\Upsilon^{\mathrm{z}}$ to evaluate to the status code resulting from the transaction.
 These will be formally defined later.
@@
 \begin{equation}
-A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
@@
 
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
 $G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
 
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the transaction receipt, discussed later.
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
 
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
 \end{eqnarray}
@@
 I_{\mathrm{p}} & \equiv & p \\
-I_\mathbf{d} & \equiv & () \\
+I_{\mathbf{d}} & \equiv & () \\
 I_{\mathrm{s}} & \equiv & s \\
 I_{\mathrm{v}} & \equiv & v \\
-I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_{\mathbf{b}} & \equiv & \mathbf{i} \\
 I_{\mathrm{e}} & \equiv & e \\
@@
 
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
 \end{equation}
@@
 I_{\mathrm{p}} & \equiv & p \\
-I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_{\mathbf{d}} & \equiv & \mathbf{d} \\
 I_{\mathrm{s}} & \equiv & s \\
@@
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
 
@@
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
 \boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
+\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
+\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
 \end{eqnarray}
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 \mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
+ \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
@@
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 \Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
-\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
+\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
+\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
 \boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
 
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
 R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
@@
 
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_{\mathbf{l}} =
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
 \end{equation}
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
-\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|I_{\mathbf{d}}|-1)] &=& I_{\mathbf{d}}\\
+\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
+g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})\\
 \end{eqnarray}
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_\mathbf{d}
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_{\mathbf{d}}
 \end{eqnarray}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_\mathbf{d}|}{192} \\
+F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
+k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
 g_{\mathrm{r}}&=& 60000k + 40000 \\
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 \begin{equation}
-w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{a}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
 \midrule
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
 \midrule
 \linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
 &&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
@@
 \linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
+&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
@@
 \linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \end{tabular*}
@@
 &&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
 \midrule
@@
 \varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
 \end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
+&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule

### 2018-02-05 Paper.tex
Transaction receipt changes
\Upsilon^z => \Upsilon{\mathrm{z}}, \Upsilon^g => \Upsilon{\mathrm{g}} 
\linkdest{Transaction_Receipt}{}
\hyperlink{Transaction_Receipt}{transaction receipt}
Other targets for parameters

--- 2018-02-04 Paper.tex
@@
 
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 
-The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_\mathbf{d}$:
 \begin{equation}
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the transaction receipt, discussed later.
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
 
@@
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 \mathbf{R}[n]_\mathbf{l} =
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}

### 2018-02-05 Paper.tex
Remove braces around \mathbf

@@
 
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_\mathbf{d}$:
+The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
 \begin{equation}

### 2018-02-05 Paper.tex
Replace _(\\mathbf\{[:alnum:]+\}) with _{$1}
...using Libre Open Office with regular expressions ticked as an option.

--- 2018-02-05 Paper.tex
@@
 
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
 
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_{\mathbf{l}}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_{\mathbf{l}}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^{\mathrm{g}}$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^{\mathbf{l}}$ to evaluate to the transaction's accrued log items and $\Upsilon^{\mathrm{z}}$ to evaluate to the status code resulting from the transaction.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items and $\Upsilon^z$ to evaluate to the status code resulting from the transaction.
 These will be formally defined later.
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
 
@@
 \begin{array}[b]{l}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 \mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
 \end{equation}
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\

### 2018-02-05 Paper.tex
\hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}

--- 2018-02-05 Paper.tex
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
 
@@
 \begin{array}[b]{l}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 \mathbf{R}[n]_\mathbf{l} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}

### 2018-02-05 Paper.tex
2 more \hyperlink{transaction_nonce}{transaction nonce}

@@
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
 Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 \item the transaction signature is valid;
-\item the transaction nonce is valid (equivalent to the sender account's current nonce);
+\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the sender account's current nonce);
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;

### 2018-02-05 Paper.tex
hyperlinks for state transition functions

--- 2018-02-05 Paper.tex
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \linkdest{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
@@
 
-\subsection{The Block} \label{ch:block}
+\subsection{The Block} \linkdest{block}{}
 

### 2018-02-05 Paper.tex
\hyperref{https://medium.com/@VitalikButerin/the-meaning-of-decentralization-a0c92b76a274}{decentralised}
This probably doesn't need to be a citation/reference.

--- 2018-02-05 Paper.tex
@@
 
-Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
+Since the system is \hyperref{https://medium.com/@VitalikButerin/the-meaning-of-decentralization-a0c92b76a274}{decentralised} and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
 In order to form a consensus as to which path, from root (\hyperlink{GenesisBlock}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.

### 2018-02-05 Paper.tex
\hyperlink{Upsilon_state_transition}{$\Upsilon$}

@@
 
-Since the system is \hyperref{https://medium.com/@VitalikButerin/the-meaning-of-decentralization-a0c92b76a274}{decentralised} and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
+Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
 In order to form a consensus as to which path, from root (\hyperlink{GenesisBlock}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
 

### 2018-02-05 Paper.tex
\hyperlink{C__SSTORE} \hyperlink{SSTORE}

--- 2018-02-05 Paper.tex
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
+These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
 For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\

### 2018-02-05 Paper.tex
\cite{Keccak}

--- 2018-02-05 Paper.tex
@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases, by \cite{Keccak}) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.

### 2018-02-05 Paper.tex
\linkdest{ell}{} (add {})

@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest, rather than later releases, by \cite{Keccak}) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 
-I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \linkdest{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 

### 2018-02-05 Paper.tex
Add missing closing braces

@@
 
-I define a number of useful functions throughout. \linkdest{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_t = 0 \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
+\end{eqnarray}
+
+\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
+\begin{eqnarray}
+\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
+\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
+\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
 \end{eqnarray}

### 2018-02-05 Paper.tex
Test: Swap bibliographystyle and bibliography
Add \pagebreak[1] before where... in case a column or page needs to break
or to prevent extra white lines from appearing, as happened with #509.

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium revision (\YellowPaperVersionNumber{})}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 \end{equation}
-
-where $\Upsilon$ is the Ethereum state transition function.
+\pagebreak[1]where $\Upsilon$ is the Ethereum state transition function.
 In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
@@
 \end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\pagebreak[1]where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 \end{equation}
-
-where:
+\pagebreak[1]where:
 \begin{equation}
@@
 \end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence./footnote{11 bits $= 2^2048$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+\pagebreak[1]where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+\pagebreak[1]where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 \end{equation}
-
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
+\pagebreak[1]where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
 \begin{equation}
@@
 \end{equation}
-
-where
+\pagebreak[1]where
 \begin{equation}
@@
 
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
 The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_i = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
-where:
-\begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_s - {P(H)_H}_s}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_i \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_i &\equiv \max(H_i - 3000000, 0)
-\end{align}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
+\end{equation}
+\pagebreak[1]where
+\begin{equation}
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
 
@@
 We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
-
-where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
+\pagebreak[1]where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G$ is defined in Appendix \ref{app:fees}.
+
+%todo Explain g_d reason?
 
@@
 
-Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$, substate $A$ and status code $s'$:
-\begin{equation}
-(\boldsymbol{\sigma}_P, g', A, s') \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0, \top) & \text{if} \quad T_t = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\end{cases}
+\end{equation}
+\pagebreak[1]where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
 \begin{equation}
@@
 \end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+\pagebreak[1]where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
-\end{eqnarray}
-
-where $v'$ is the account's pre-existing value, in the event it was previously in existence:
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\end{eqnarray}
+\pagebreak[1]where $v'$ is the account's pre-existing value, in the event it was previously in existence:
 \begin{equation}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**} < c \\
+g^{**} - c & \text{otherwise} \\
+\end{cases} \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
+\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
 \begin{eqnarray}
@@
 I_\mathbf{b} & \equiv & \mathbf{i} \\
-I_e & \equiv & e \\
-I_w & \equiv & w
-\end{eqnarray}
-
-$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+I_e & \equiv & e
+\end{eqnarray}
+\pagebreak[1]where $c$ is the code-deposit cost:
 \begin{equation}
@@
 \boldsymbol{\mu}_i & \equiv & 0 \\
-\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
-\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
+\end{cases}
+\pagebreak[1]\end{equation}
+\pagebreak[1]where
 \begin{eqnarray}
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_g & \equiv & \boldsymbol{\mu}_g - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
+\end{eqnarray}
+
+Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
 
@@
 \end{equation}
-
-where:
+\pagebreak[1]where
 \begin{equation}
@@
 \end{equation}
-
-where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
+\pagebreak[1]where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
 \begin{equation}
@@
 \end{equation}
-
-where $k$ denotes the ``is-kin'' property:
+\pagebreak[1]where $k$ denotes the ``is-kin'' property:
 \begin{equation}
@@
 \end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
+\pagebreak[1]where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+\pagebreak[1]where $\mathbb{Y}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
@@
 Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < r &< \mathtt{\tiny secp256k1n} \\
-0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
-\end{align}
-where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+\pagebreak[1]where:
+\begin{eqnarray}
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
 
@@
 
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
-\begin{eqnarray}
-\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
-\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
-\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
-\end{eqnarray}
-
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_r, T_s) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_w &\text{if}\ T_w\in\{27, 28\} \\
-28 - (T_w \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
 \begin{equation}
@@
 \end{equation}
-
-where:
-\begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+\pagebreak[1]where:
+\begin{equation}
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+\pagebreak[1]where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
@@
 \end{equation}
-Where a single round modifies each subset of the cache as follows:
+\pagebreak[1]where a single round modifies each subset of the cache as follows:
 \begin{equation}

### 2018-02-05 Paper.tex
Remove accidentally added \pagebreak[1]
Adjust line breaks to fix overrunning the margin / overfull hboxes
Fixing many, but not all, overfull hbox warnings
Since there are many changes, I may fix more overfull hbox warnings in a separate PR (although it is easier to do it in one PR, or wait until this is merged before making more related fixes).

--- 2018-02-05 Paper.tex
@@
 
+\usepackage{microtype}
+
 \input{Version.tex}
@@
 
-I define a number of useful functions throughout. \linkdest{ell}{}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
+&\quad \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
@@
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
 I_{\mathrm{a}} & \equiv & r \\
@@
 It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+where
+\begin{equation}
+\Xi \equiv \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
+\end{equation}
 
@@
 \begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+W(w, \boldsymbol{\mu}) \equiv \\ 
+\begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
@@
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
+\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
 D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
-\varnothing & \text{otherwise}
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \ w \in \{\text{\small RETURN}, \text{\small REVERT}\} &\\
+() \quad\quad\ \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} &\\
+\varnothing \quad\quad\ \text{otherwise}&
 \end{cases}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
 \end{cases}
@@
 \midrule
+\everyrow{\tabucline[on 2pt]-}
 $G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
@@
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. 
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+&&is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
+&&non-zero.\\
+\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
 \linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled\\
+&&contract. \\
 
@@
 \bottomrule
-\end{tabular*}
+\end{tabu}
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
+&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
+&\quad\text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
+&&&& modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
+&&&& modulo. \\
 \midrule
@@
 \midrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
+&&&& in big endian). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty \\
-&&&& associated code. \\
+&&&& This is the sender of original transaction; it is never an account with\\
+&&&& non-empty associated code. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
-&&&& This is the address of the account that is directly responsible for this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
+&&&& This is the address of the account that is directly responsible for\\
+&&&& this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
+&&&& this execution. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
+&&&& This pertains to the input data passed with the message call\\
+&&&& instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
+&&&& environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+&&&& This pertains to the input data passed with the message call\\
+&&&& instruction or transaction. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+&&&& This pertains to the input data passed with the message call instruction\\
+&&&& or transaction. \\
 \midrule
@@
 \midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
+&&&& environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
@@
 \linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than\\
+&&&& the current block number or more than 256 blocks behind the current block.\\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent\\
+&&&& hash in the block following it. \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
-&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge$ \\
+&&&& $A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t} \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
-&&&& created account. \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\ 
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
+&&&& address of the newly created account. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
+&&&& fact the same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
+&&&& persisting the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
+&&&& omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the\\
+&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and\\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced \\
+&&&& with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 \varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\ (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise} \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
@@
 \bottomrule
-\end{tabular*}
+\end{tabu*}
 
@@
 \section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\begin{tabu*}{c c X[j]} \savetabu{col3}
 \toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+Symbol & Latex Command & Description \\
+\midrule
+\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
+Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
 \bottomrule
 \end{tabu*}
-
 \end{document}

### 2018-02-07 Paper.tex
J__wordbytes
\usepackage[final]{microtype} (add a final option)
https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142

@@
 
-\usepackage{microtype}
-
 \input{Version.tex}
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
 \item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
-&\quad \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
 g^{**} & \text{otherwise}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_{\mathrm{a}} & \equiv & r \\
@@
 It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
-
-where
-\begin{equation}
-\Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
 
@@
 \begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \\ 
-\begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
@@
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
-\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
 D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
@@
 N(i, w) \equiv \begin{cases}
-i + w - \text{\small PUSH1} + 2 & \\
-\quad\quad\text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
+i + w - \text{\small PUSH1} + 2 & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
 i + 1 & \text{otherwise} \end{cases}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \ w \in \{\text{\small RETURN}, \text{\small REVERT}\} &\\
-() \quad\quad\ \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} &\\
-\varnothing \quad\quad\ \text{otherwise}&
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+\varnothing & \text{otherwise}
 \end{cases}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
@@
 
-\begin{tabu}{l r l}
+\begin{tabular*}{\columnwidth}[h]{lrl}
 \toprule
@@
 \midrule
-\everyrow{\tabucline[on 2pt]-}
 $G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
@@
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
-&&is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
-&&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. 
+\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
 \linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled\\
-&&contract. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 \bottomrule
-\end{tabu}
+\end{tabular*}
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
-&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 
-\begin{tabu}{r l r r l} \savetabu{opcodes}
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 \midrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
-&&&& in big endian). \\
+&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
-&&&& This is the sender of original transaction; it is never an account with\\
-&&&& non-empty associated code. \\
+&&&& This is the sender of original transaction; it is never an account with non-empty \\
+&&&& associated code. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
-&&&& This is the address of the account that is directly responsible for\\
-&&&& this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
-&&&& this execution. \\
+&&&& This is the address of the account that is directly responsible for this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
@@
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
-&&&& environment. \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction\\
-&&&& or transaction. \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
@@
 \midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
-&&&& environment. \\
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
@@
 \linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent\\
-&&&& hash in the block following it. \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge$ \\
-&&&& $A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t} \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
 &&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\ 
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
-&&&& address of the newly created account. \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
 &&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
-&&&& fact the same account as at present, simply that the code is overwritten.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
-&&&& persisting the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
-&&&& omitted argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
-As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the\\
-&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and\\
-&&&& $\boldsymbol{\mu}_\mathbf{s}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+&&&& the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& Otherwise exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced \\
-&&&& with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 \varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\ (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise} \end{cases}$\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+\end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
@@
 
-\begin{tabu*}{lcl}
+\begin{tabular*}{\columnwidth}[h]{lrl}
 \toprule
@@
 \bottomrule
-\end{tabu*}
+\end{tabular*}
 
@@
 \section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{c c X[j]} \savetabu{col3}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
 \toprule
-Symbol & Latex Command & Description \\
-\midrule
-\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
-Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
 \bottomrule
 \end{tabu*}
+
 \end{document}

### 2018-02-07 Paper.tex

@@
 \item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.

### 2018-02-07 Paper.tex

@@
 
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
 
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}

### 2018-02-07 Paper.tex

@@
 \begin{equation}
-B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
 \end{equation}
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
+Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
 
@@
 \begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
 \end{equation}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 \begin{array}[b]{l}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_\mathbf{l} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_{\mathbf{l}} =
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
 \end{equation}

### 2018-02-07 Paper.tex

@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
 Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.

### 2018-02-07 Paper.tex

@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\

### 2018-02-07 Paper.tex
Cleanup ECREC definition and a few other things

@@
 \end{eqnarray}
+
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
@@
 
-For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
+We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
+See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
+
 \begin{eqnarray}
@@
 
-The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
+We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively.
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
+
 \begin{eqnarray}
@@
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
 \end{eqnarray}
@@
 
-We define a precompiled contract for zkSNARK verification.
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
 \begin{eqnarray}
@@
 We define a precompiled contract for addition on $G_1$.
+
 \begin{eqnarray}
@@
 
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assume the existence of functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}

### 2018-02-08 Paper.tex
Remove an extra }

--- 2018-02-07 Paper.tex
@@
 
-Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg $\Upsilon$, the Ethereum state transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_{\mathrm{n}}$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
 Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
+When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
 It is formally defined in section \ref{ch:block}.
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
 \end{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
@@
 \begin{description}
-\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
-\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{description}
-\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_{\mathbf{i}}$.
+\item[init] An unlimited size byte array specifying the EVM-code for the account initialisation procedure, formally $T_\mathbf{i}$.
 \end{description}
@@
 \begin{description}
-\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_{\mathbf{d}}$.
+\item[data] An unlimited size byte array specifying the input data of the message call, formally $T_\mathbf{d}$.
 \end{description}
@@
 L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{i}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_\mathbf{d}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
 \begin{equation}
@@
 T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 
-The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
 Formally, we can refer to a block $B$:
@@
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
 \end{equation}
@@
 \begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l})
 \end{equation}
@@
 
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+
+The sequence $R_\mathbf{l}$ is a series of log entries, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_{\mathrm{a}}, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
@@
 \linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_\mathbf{T}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_\mathbf{R}[i]))\}) & \wedge \\
+\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_{\mathrm{b}} \big)
 \end{array}
@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_{\mathbf{R}}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(B_{\mathbf{U}}) \big)
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
 \end{eqnarray}
@@
 \begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
+1 & \text{if} \, \lVert P(H)_\mathbf{U}\rVert = 0 \\
 2 & \text{otherwise}
@@
 
-The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function \hyperlink{Upsilon_state_transition}{$\Upsilon$}.
+The execution of a transaction is the most complex part of the Ethereum protocol: it defines the state transition function $\Upsilon$.
 It is assumed that any transactions executed first pass the initial tests of intrinsic validity.
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the sender account's current nonce);
+\item the transaction nonce is valid (equivalent to the sender account's current nonce);
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
@@
 
-Formally, we consider the function \hyperlink{Upsilon_state_transition}{$\Upsilon$}, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
+Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\boldsymbol{\sigma}$ the state:
 \begin{equation}
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
@@
 
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
 $G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
 
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_\mathbf{i}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_\mathbf{d}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l} \\
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
 
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
 \end{eqnarray}
@@
 I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
+I_\mathbf{d} & \equiv & () \\
 I_{\mathrm{s}} & \equiv & s \\
 I_{\mathrm{v}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
+I_\mathbf{b} & \equiv & \mathbf{i} \\
 I_{\mathrm{e}} & \equiv & e \\
@@
 
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
 \end{equation}
@@
 I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
 I_{\mathrm{s}} & \equiv & s \\
@@
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
 
@@
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
 \boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
+\boldsymbol{\mu}_\mathbf{s} & \equiv & () \\
+\boldsymbol{\mu}_\mathbf{o} & \equiv & ()
 \end{eqnarray}
@@
 The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
+The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
 \text{\small STOP} & \text{otherwise}
@@
 \mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] > \lVert\boldsymbol{\mu}_\mathbf{o}\rVert) \quad \vee \\
+ \lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
@@
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0
 \end{array}
@@
 \Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
-\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
+\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
+\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
 \boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
@@
 \begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathbf{U}})
+s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_\mathbf{U})
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
 
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{b}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
 R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathbf{R}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathbf{l} =
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n])
 \end{equation}
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_{\mathbf{d}}|-1)] &=& I_{\mathbf{d}}\\
-\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
+\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
+g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
 \end{eqnarray}
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
 \end{eqnarray}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
+F&\equiv&(|I_\mathbf{d}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
+k &=& \dfrac{|I_\mathbf{d}|}{192} \\
 g_{\mathrm{r}}&=& 60000k + 40000 \\
@@
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 \begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod \boldsymbol{\mu}_{\mathbf{s}}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0]) (|\boldsymbol{\mu}_{\mathbf{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathbf{s}}[1]|) & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) (|\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]|) & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
@@
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \midrule
@@
 0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_\mathbf{s}[1]_{\mathrm{i}}$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{a}}$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{o}}$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{s}}$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{v}}$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_{\mathrm{p}}$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\mu}_\mathbf{o} \rVert$ \\
 \midrule
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} \boldsymbol{\mu}_\mathbf{o}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert \boldsymbol{\mu}_\mathbf{o} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 \linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
 &&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 &&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
@@
 \linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 0x80 & {\small DUP1} & 1 & 2 & Duplicate 1st stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
 0x81 & {\small DUP2} & 2 & 3 & Duplicate 2nd stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
@@
 0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[15]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
 \bottomrule
@@
 0x90 & {\small SWAP1} & 2 & 2 & Exchange 1st and 2nd stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[1] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[1] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
 0x91 & {\small SWAP2} & 3 & 3 & Exchange 1st and 3rd stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[2]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[2] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[2]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[2] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \midrule
@@
 0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[16]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[16] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \bottomrule
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
+\multicolumn{5}{l}\linkdest{A l}{}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xa1 & {\small LOG1} & 3 & 0 & Append log record with one topic. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2])$ \\
 \midrule
@@
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4], \boldsymbol{\mu}_{\mathbf{s}}[5])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
 \bottomrule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
-&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
 &&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
 &&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} \equiv ()$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{o} = \mathbf{o}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
 &&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
 &&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
@@
 \linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& The argument $\boldsymbol{\mu}_\mathbf{s}[2]$ is replaced with $0$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 &&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \midrule
@@
 \varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
 \end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
+&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_\mathbf{s} \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule

### 2018-02-08 Paper.tex
ref{ch:block} => \hyperlink{block}

@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \ref{ch:block}.
+This is formally defined in section \hyperlink{block}.
 
@@
 When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
+It is formally defined in section \hyperlink{block}.
 
@@
 
-\subsection{The Block} \label{ch:block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (which is not saying much, in that it isn't commonly used) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+\subsection{The Block} \linkdest{block}{}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
 

### 2018-02-08 Paper.tex
move by \cite{Keccak}

@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \linkdest{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases),  is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \hyperlink{block}.
+This is formally defined in section \ref{ch:block}.
 
@@
 When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}.
+It is formally defined in section \ref{ch:block}.
 
@@
 
-\subsection{The Block} \linkdest{block}{}
+\subsection{The Block} \label{ch:block}
 

### 2018-02-08 Paper.tex
Move Xi precompiles to line. 845

@@
 
+\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
+
 \input{Version.tex}
@@
 These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases),  is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 \end{eqnarray}
+where
+\begin{equation}
+\Xi \equiv \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
+\end{equation}
 
@@
 \midrule
-\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
 \linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\

### 2018-02-08 Paper.tex
Move equations around
Fix spacing of table of maths symbols

--- 2018-02-08 Paper.tex
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases),  is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 \begin{description}
-\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
-\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
 This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$.
+\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
+\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-B \equiv (B_{H}, B_\mathbf{T}, B_\mathbf{U})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
+\end{equation}
+
+\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_\mathbf{l}, R_{\mathrm{z}})
+Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\begin{equation}
+R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
 \end{equation}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_\mathbf{R}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
 \midrule
 \linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
-&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
+&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\begin{tabu*} spread 0pt{X[l] X[c] X} \savetabu{col3}
 \toprule
@@
 \end{tabu*}
-
 \end{document}

### 2018-02-08 Paper.tex
\hyperlink{Transaction Receipt} \hyperlink{Pi}{

@@
 
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_{\mathbf{R}}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 
-\begin{tabu*} spread 0pt{X[l] X[c] X} \savetabu{col3}
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
 \toprule
@@
 \end{tabu*}
+
 \end{document}

### 2018-02-08 Paper.tex
\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}
L_{R} hypertargets and links

--- 2018-02-08 Paper.tex
@@
 
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
 \begin{equation}
@@
 \linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
+Much like the \hyperlink{transacation_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}

### 2018-02-08 Paper.tex
\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}
These aren't actually used elsewhere.

--- 2018-02-08 Paper.tex
@@
 
-The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
 Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:

### 2018-02-08 Paper.tex
Links for component types (plus update elsewhere) in block and block header preparation functions.

@@
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
+\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
+The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
 \item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
+\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
 Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
@@
 \begin{array}[t]{lclclcl}
-H_{\mathrm{p}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-H_{\mathrm{r}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
-H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & H_{\mathrm{i}} \in \mathbb{P} & \wedge \\
-H_{\mathrm{l}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & H_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
-H_{\mathrm{x}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{n}} \in \mathbb{B}_{8}
+\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
+\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
 \end{array}
@@
 
-$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
+\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
 \begin{equation}

### 2018-02-08 Paper.tex
\hypertarget{block_number_H__i}{}
\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
More hyperlinks: upsilon_pow_g,l,z, account_nonce, block_nonce_H__n

@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = H_{\mathrm{p}}
-\end{equation}
-
-The block number is the parent's block number incremented by one:
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+\end{equation}
+
+\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
 \begin{equation}
@@
 
-The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
+The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
 \begin{equation}
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the sender account's current nonce);
+\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^{\mathrm{g}}$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^{\mathbf{l}}$ to evaluate to the transaction's accrued log items and $\Upsilon^{\mathrm{z}}$ to evaluate to the status code resulting from the transaction.
+We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
 These will be formally defined later.

### 2018-02-08 Paper.tex
\hyperlink{SSTORE}{{\small SSTORE}}

--- 2018-02-08 Paper.tex
@@
 
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.

### 2018-02-08 Paper.tex
Intrinsic gas hyperlinks and mathrm: \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}}, etc.

@@
 
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
+g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
 \end{align}
@@
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.

### 2018-02-08 Paper.tex
The up-front cost hyperlinks
It's pretty simple anyway but nevertheless...

@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
 \end{align}
@@
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
 
-The up-front cost $v_0$ is calculated as:
-\begin{equation}
-v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
+The up-front cost $v_0$ is calculated as the transaction gas limit multiplied by the gas price, plus the transaction value:
+\begin{equation}
+v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
 \end{equation}

### 2018-02-08 Paper.tex
\hyperlink{ell}{\ell}

@@
 
-The up-front cost $v_0$ is calculated as the transaction gas limit multiplied by the gas price, plus the transaction value:
-\begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+The up-front cost $v_0$ is calculated as:
+\begin{equation}
+v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
 \end{equation}
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
 

### 2018-02-08 Paper.tex
\hyperlink{account_nonce}{nonce of the account of the sender}

--- 2018-02-08 Paper.tex
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.

### 2018-02-08 Paper.tex
\hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}}

--- 2018-02-08 Paper.tex
@@
 
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
 \end{equation}

### 2018-02-08 Paper.tex
\hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} etc.

@@
 \usepackage{amsmath}
-\usepackage{amssymb}
+\usepackage{amssymb}Th
 \usepackage[a4paper,width=170mm,top=18mm,bottom=22mm,includeheadfoot]{geometry}
@@
 
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z

### 2018-02-08 Paper.tex
Remove a mistake "Th"
\hyperlink{rlp}{RLP} etc. and grammar

--- 2018-02-08 Paper.tex
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{account nonce}.
 Thus we define the resultant address for the new account $a$:
@@
 
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.

### 2018-02-08 Paper.tex
\hyperlink{trie}{TRIE}

@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{account nonce}.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
@@
 
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.

### 2018-02-08 Paper.tex
\hyperlink{execenv}{execution environment} and linkdest

@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s] &=& \begin{cases}
@@
 \end{equation}
-\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
+\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
 \begin{eqnarray}
@@
 I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
+\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & e \\
 I_{\mathrm{w}} & \equiv & w
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
 the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__e} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}

### 2018-02-08 Paper.tex
\hyperlink{hhalt}{normal halting} \hyperlink{zhalt}{exceptional halting} + linkdests

--- 2018-02-08 Paper.tex
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
 
@@
 
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}
+\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}
 
@@
 
-\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}
+\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}\linkdest{hhalt}
 

### 2018-02-08 Paper.tex
\hyperlink{RETURN} selfdestruct

@@
 
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).
+All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined {\small STOP} operations that trail it).
 
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
 

### 2018-02-08 Paper.tex
discussed \hyperlink{empty_sequence_vs_empty_set}{here}
\hyperlink{JUMP}{J_{\text{JUMP}}} \hyperlink{JUMPI}{J_{\text{JUMPI}}}

--- 2018-02-08 Paper.tex
@@
 \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
+\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
+\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
 N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}

### 2018-02-08 Paper.tex
\hyperlink{Genesis_Block}{genesis block}

@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (\hyperlink{GenesisBlock}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (\hyperlink{Genesis_Block}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
+J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
+J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
 N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
@@
 
-\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
+\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
 

### 2018-02-08 Paper.tex
hyperlinks block_header_validity_function ommer_block_header

--- 2018-02-08 Paper.tex
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and nonce.
+\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{block nonce}.
 \end{enumerate}
@@
 \begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} V({\mathbf{U}}) \; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
+\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
 \end{equation}

### 2018-02-08 Paper.tex
\hyperlink{H__g}

--- 2018-02-08 Paper.tex
@@
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
@@
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
 \item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\mathbf{R})_{\mathrm{u}}
 \end{equation}

### 2018-02-08 Paper.tex
\hyperlink{R__u}

@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 \begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\mathbf{R})_{\mathrm{u}}
+{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}

### 2018-02-08 Paper.tex
R_{\mathrm{b}} as the block reward => R_{\mathrm{block}} + links

--- 2018-02-08 Paper.tex
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{block nonce}.
+\item verify (or, if mining, compute a valid) state and nonce.
 \end{enumerate}
@@
 
-The validation of \hyperlink{ommer_block_headers_B__U}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
+The validation of ommer headers means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
 The maximum of ommer headers is two.
@@
 \begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
+\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} V({\mathbf{U}}) \; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{b}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
 \end{eqnarray}
@@
 
-\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
+\hypertarget{block_reward_R__block}{}We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
 \end{equation}
@@
 
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 \hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
-\begin{eqnarray}
+\begin{eqnarray}R_\{\\mathrm\{b\}\}
 R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 

### 2018-02-08 Paper.tex
Remove accidentally added R_\{\\mathrm\{b\}\}
\hyperlink{H_cancel_n}{\hcancel{n}}, etc. + grammar

--- 2018-02-08 Paper.tex
@@
 \item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
@@
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
 \item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
@@
 
-\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\mathbf{R})_{\mathrm{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
@@
 
-\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))

### 2018-02-08 Paper.tex
which is, more links in State and Nonce Evaluation

@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
+\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
 \end{equation}
@@
 
-\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
 \begin{eqnarray}
@@
 
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
 
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
 \begin{equation}
@@
 
-We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
+We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
 \begin{equation}
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
 \begin{equation}
@@
 
-Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
+Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
 

### 2018-02-08 Paper.tex
Remove an extra }

--- 2018-02-08 Paper.tex
@@
 \begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
@@
 
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))

### 2018-02-08 Paper.tex

@@
 \quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(B_{\mathbf{U}}) \big)
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
 \end{eqnarray}

### 2018-02-08 Paper.tex

@@
 w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
-\text{\small STOP} & \text{otherwise}
+\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
 \end{cases}

### 2018-02-08 Paper.tex

@@
 
-The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
+The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
+Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
 

### 2018-02-08 Paper.tex

@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}

### 2018-02-09 Paper.tex
but isn't a common word

--- 2018-02-08 Paper.tex
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (\hyperlink{Genesis_Block}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (\hyperlink{GenesisBlock}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
+These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
 For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (but isn't a common word) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
+\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
 \item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
+\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_\mathbf{T}), L_{H}^*(B_\mathbf{U}) \big)
 \end{eqnarray}
@@
 \begin{array}[t]{lclclcl}
-\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
+H_{\mathrm{p}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
+H_{\mathrm{r}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
+H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & H_{\mathrm{i}} \in \mathbb{P} & \wedge \\
+H_{\mathrm{l}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & H_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
+H_{\mathrm{x}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{n}} \in \mathbb{B}_{8}
 \end{array}
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
-\end{equation}
-
-\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = H_{\mathrm{p}}
+\end{equation}
+
+The block number is the parent's block number incremented by one:
 \begin{equation}
@@
 
-\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
+$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
 \begin{equation}
@@
 
-The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
+The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
 \begin{equation}
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
+\item the transaction nonce is valid (equivalent to the sender account's current nonce);
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items and $\Upsilon^z$ to evaluate to the status code resulting from the transaction.
 These will be formally defined later.
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_{\mathrm{r}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
 \end{align}
 
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_\mathbf{R})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
-\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
+\text{\small STOP} & \text{otherwise}
 \end{cases}
@@
 
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}
+\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}
 
@@
 
-\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}\linkdest{hhalt}
+\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}
 
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small RETURN}, \text{\small REVERT}\} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
+The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{Genesis_Block}{genesis block}, in the path.
+Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined genesis block, in the path.
 The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
@@
 
+\bibliography{Biblio}
 \bibliographystyle{plainnat}
-\bibliography{Biblio}
 
@@
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
 \midrule
 \linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
+&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
@@
 
-\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
+\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
 

### 2018-02-09 Paper.tex

--- 2018-02-08 Paper.tex
@@
 When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
+It is formally defined in section \hyperlink{block}.
 

### 2018-02-09 Paper.tex

--- 2018-02-08 Paper.tex
@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.

### 2018-02-09 Paper.tex
Add parity bit explanation for EIP 155 signatures
T_w describe the parity of the y value of the curve point of the signature. This is true for both the original (pre-EIP 155) signature scheme and for the EIP 155 signature scheme.

@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+The values 27 and 35 represent an even $y$ value and 28 and 36 represent an odd $y$ value.
 

### 2018-02-10 Paper.tex

@@
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The values 27 and 35 represent an even $y$ value and 28 and 36 represent an odd $y$ value.
+The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
 
@@
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
+E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}

### 2018-02-10 Paper.tex

@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-\usepackage{tabu} %requires array.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
-
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks.
-In order to form a consensus as to which path, from root (\hyperlink{Genesis_Block}{the genesis block}) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
+In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme.
 If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.
@@
 
-Functions operating on highly structured values are denoted with an upper-case Greek letter, \eg \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the Ethereum state transition function.
+Functions operating on highly structured values are denoted with an upper-case greek letter, \eg $\Upsilon$, the Ethereum state transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
-Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
 For particularly important values, a bold uppercase letter may be used.
@@
 
-When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
+When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
 It is formally defined in section \ref{ch:block}.
 
-I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\begin{equation}
+L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 
-It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_{S}$:
-\begin{equation}
-L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_S$:
+\begin{equation}
+L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \end{equation}
-\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
-\end{equation}
-
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-\subsection{The Transaction} \label{subsec:transaction}
+where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+\subsection{The Transaction} \label{ch:transaction}
 
@@
 \begin{description}
-\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
-\item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
-\item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
-\item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[nonce] A scalar value equal to the number of transactions sent by the sender; formally $T_n$.
+\item[gasPrice] A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_p$.
+\item[gasLimit] A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_g$.
+\item[to] The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_t$.
+\item[value] A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_v$.
+\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally $T_w$, $T_r$ and $T_s$.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{equation}
-L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+L_T(T) \equiv \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_t \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
-\end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+\end{equation}
+
+where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 \begin{array}[t]{lclclcl}
-\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
@@
 
-We define $P(B_{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
-\end{equation}
-
-\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
-\begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
+We define $P(B_H)$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_H)) = H_p
+\end{equation}
+
+The block number is the parent's block number incremented by one:
+\begin{equation}
+H_i \equiv {{P(H)_H}_i} + 1
 \end{equation}
@@
 
-\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
 \begin{equation}
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\mindifficulty & \text{if} \quad H_i = 0\\
+\text{max}\!\left(\mindifficulty, {P(H)_H}_d + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 125000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 
-The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 
-\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
-\begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-& & H_{\mathrm{d}} = D(H) \quad \wedge \\
-& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-& & \lVert H_{\mathrm{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+Thus we are able to define the block header validity function $V(H)$:
+\begin{eqnarray}
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_d} \wedge m = H_m \quad \wedge \\
+& & H_d = D(H) \quad \wedge \\
+& & H_g \le H_l  \quad \wedge \\
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor  \quad \wedge \\
+& & H_l \geqslant 125000  \quad \wedge \\
+& & H_s > {P(H)_H}_s \quad \wedge \\
+& & H_i = {P(H)_H}_i +1 \quad \wedge \\
+& & \lVert H_x \rVert \le 32
+\end{eqnarray}
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$
 
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
-These will be formally defined later.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
 
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
 \end{align}
 
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
 
+%todo Explain g_d reason?
+
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leqslant & T_g \quad \wedge \\
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g \equiv T_{\mathrm{g}} - g_0
-\end{equation}
-and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+g \equiv T_g - g_0
+\end{equation}
+and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g^* \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
+
+\section{Contract Creation} \label{ch:create}
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{account nonce}.
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+\begin{eqnarray}
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & () \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_e & \equiv & e
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
 
@@
 \quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e)
 \end{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}_1'[s]_b - v
 \end{equation}
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{\tiny KEC}(()), \mathtt{\tiny TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+(\boldsymbol{\sigma}^{**}, g^{**}, \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_s & \equiv & s \\
+I_v & \equiv & \tilde{v} \\
+I_e & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
-
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}
+\subsubsection{Exceptional Halting}
 
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 
-\subsubsection{Normal Halting}\hypertarget{normal_halting_function_H}{}\linkdest{hhalt}
+\subsubsection{Normal Halting}
 
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} \\
 \varnothing & \text{otherwise}
@@
 
-The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
+The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
-\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
+\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
+\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
 \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
-\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
+J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
+J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
 N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
@@
 \begin{eqnarray}
-\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
-\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
+\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
+\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
 A' & \equiv & A \\
@@
 \begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
 \end{equation}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
+\begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 
-\hypertarget{block_reward_R__block}{}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
+We define the block reward as 5 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 5 \times 10^{18}
 \end{equation}
@@
 
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
@@
 
-As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathbf{l} =
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\end{equation}
+
+If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
+\begin{equation}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
 
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
 \end{eqnarray}
@@
 
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
 \end{eqnarray}
@@
 
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 \begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28\}
 \end{align}
@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
 
@@
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
-
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
-\midrule
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
 \end{tabular*}
@@
 0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
 &&&& This is the sender of original transaction; it is never an account with non-empty \\
@@
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
 &&&& This is the address of the account that is directly responsible for this execution. \\
@@
 0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[1] + i$ are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_c$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[3])$ \\
+&&&& The additions in $\boldsymbol{\mu}_\mathbf{s}[2] + i$ are not subject to the $2^{256}$ modulo. \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
+&&&& The addition in the calculation of $\boldsymbol{\mu}'_i$ is not subject to the $2^{256}$ modulo. \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
-\midrule
-\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+\midrule
+0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \midrule
-\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
-\midrule
-\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+\midrule
+0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
 &&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
+\multicolumn{5}{l}{$\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
-&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
-&&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge \\ \quad\quad I_e < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\\ \quad\quad{}I_e < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
 &&&& same account as at present, simply that the code is overwritten.\\
 \midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \end{tabular*}
@@
 &&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& argument is $\boldsymbol{\mu}_\mathbf{s}[2]$.
+As a result, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$, $\boldsymbol{\mu}_\mathbf{s}[5]$ and $\boldsymbol{\mu}_\mathbf{s}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_\mathbf{s}[2]$, $\boldsymbol{\mu}_\mathbf{s}[3]$, $\boldsymbol{\mu}_\mathbf{s}[4]$ and $\boldsymbol{\mu}_\mathbf{s}[5]$. \\
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_s, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, 0, I_v, \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad I_v \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 \midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
 \midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 
-\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
+\section{Genesis Block}\label{app:genesis}
 
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
-
 \end{document}

### 2018-02-12 build.sh
Add option to build white paper

@@ -1,18 +1,31 @@
 #!/usr/bin/env bash
 
+##
+## Usage: build.sh [white]
+##
+
+
 set -e
+
+rm -rf Options.tex
 
 if [ -d ".git" ]; then
 
 SHA=`git rev-parse --short --verify HEAD`
 DATE=`git show -s --format="%cd" --date=short HEAD`
 REV="$SHA - $DATE"
-
-else
-
-REV="unknown revision"
+echo "\def\YellowPaperVersionNumber{$REV}" >> Options.tex
 
 fi
+
+
+if [ "$1" == "white" ]; then
+
+echo "\definecolor{pagecolor}{rgb}{1,1,1}" >> Options.tex
+
+fi
+
+
 
 echo "\newcommand{\YellowPaperVersionNumber}{$REV}" > Version.tex
 
@@ -21,3 +34,4 @@ bibtex Paper && \
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex && \
 pdflatex -interaction=errorstopmode -halt-on-error Paper.tex
+rm -rf Options.tex

### 2018-02-13 Paper.tex
Add option to build white paper
Update ommer footnote

--- 2018-02-10 Paper.tex
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_\mathbf{U}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_\mathbf{s}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
 \end{eqnarray}
@@
 
-\hypertarget{block_reward_R__block}{}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
+\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
@@
 
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 
@@
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+E_{prime}((x - 1) \cdot y, y) & \text{otherwise}
 \end{cases}

### 2018-02-13 Paper.tex
Add a space

--- 2018-02-12 Paper.tex
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_{\mathbf{R}}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(B_{\mathbf{U}}) \big)
 \end{eqnarray}
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
+\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the sender account's current nonce);
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
+We also define $\Upsilon^{\mathrm{g}}$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^{\mathbf{l}}$ to evaluate to the transaction's accrued log items and $\Upsilon^{\mathrm{z}}$ to evaluate to the status code resulting from the transaction.
 These will be formally defined later.
@@
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
@@
 w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
-\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
+\text{\small STOP} & \text{otherwise}
 \end{cases}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{b}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
 \end{eqnarray}
@@
 
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
 \end{eqnarray}
-
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
@@
 
-We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
-We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
-
+For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
 \begin{eqnarray}
@@
 
-We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively.
+The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
-
 \begin{eqnarray}
@@
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})\\
 \end{eqnarray}
@@
 
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
+We define a precompiled contract for zkSNARK verification.
 \begin{eqnarray}
@@
 We define a precompiled contract for addition on $G_1$.
-
 \begin{eqnarray}
@@
 
-We assume the existence of functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
 \toprule

### 2018-02-13 Paper.tex
Re-add removed line
Remove maths in English

--- 2018-02-12 Paper.tex
@@
 
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 

### 2018-02-13 Paper.tex
Fix an error

--- 2018-02-13 Paper.tex
@@
 I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & v \\
+\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
 I_{\mathbf{b}} & \equiv & \mathbf{i} \\
@@
 
-\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
 \toprule

### 2018-02-13 Paper.tex

--- 2018-02-13 Paper.tex
@@
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
 \item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
@@
 \end{description}
-
 \linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
@@
 
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{equation}
-A^0 \equiv (\varnothing,(), \varnothing, 0)
+A^0 \equiv (\varnothing, (), \varnothing, 0)
 \end{equation}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}

### 2018-02-13 Paper.tex

@@
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
 \item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
@@
 \end{description}
+
 \linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
 \end{equation}
@@
 
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
+
 B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
 B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}

### 2018-02-13 Paper.tex
Delete an empty line

@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-
 B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\

### 2018-02-13 Paper.tex

--- 2018-02-13 Paper.tex
@@
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
+
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
 Formally, we can refer to a block $B$:

### 2018-02-13 Paper.tex

@@
 \begin{equation}
-v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
+v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
 \end{equation}
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
+\section{List of mathematical symbols} \label{app:symbols}
+\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
 \toprule

### 2018-02-16 Paper.tex
Use serial comma

@@
 
-\input{Version.tex}
+% Default rendering options
+\definecolor{pagecolor}{rgb}{1,0.98,0.9}
+\def\YellowPaperVersionNumber{unknown revision}
+\IfFileExists{Options.tex}{\input{Options.tex}}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
 \author{

### 2018-02-16 Paper.tex
Typos and grammar fixes

@@
 This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
 
@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
@@
 \item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
 \item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miners to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
@@
 \end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
 \begin{equation}

### 2018-02-16 build.sh
Fix the build, #583

@@ -29,7 +29,7 @@ fi
 
 echo "\newcommand{\YellowPaperVersionNumber}{$REV}" > Version.tex
 
-mkdir build
+mkdir -p build
 pdflatex -output-directory=build -interaction=errorstopmode -halt-on-error Paper.tex && \
 bibtex build/Paper && \
 pdflatex -output-directory=build -interaction=errorstopmode -halt-on-error Paper.tex && \

### 2018-02-16 Paper.tex
Clearer OOG exception wording, fixes #84
From @SergioDemianLerner

--- 2018-02-16 Paper.tex
@@
 There are many goals of this project; one key goal is to facilitate transactions between consenting individuals who would otherwise have no means to trust one another.
-This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience, or corruption of existing legal systems.
+This may be due to geographical separation, interfacing difficulty, or perhaps the incompatibility, incompetence, unwillingness, expense, uncertainty, inconvenience or corruption of existing legal systems.
 By specifying a state-change system through a rich and unambiguous language, and furthermore architecting a system such that we can reasonably expect that an agreement will be thus enforced autonomously, we can provide a means to this end.
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas exception, they do not leave state changes intact.
+Like the out-of-gas exception, they leave the state intact.
 Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.

### 2018-02-16 Paper.tex
Replace _(\\mathbf\{[:alnum:]+\}) with _{$1}
...using Libre Open Office with match case and regular expressions ticked as options.

@@
 
-% Default rendering options
-\definecolor{pagecolor}{rgb}{1,0.98,0.9}
-\def\YellowPaperVersionNumber{unknown revision}
-\IfFileExists{Options.tex}{\input{Options.tex}}
+\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
+
+\input{Version.tex}
 
@@
 
+\definecolor{lightyellow}{rgb}{1,0.98,0.9}
+
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber{}}}}
 \author{
@@
 
-\pagecolor{pagecolor}
+\pagecolor{lightyellow}
 
@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
+\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
 \end{equation}
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
 Formally, we can refer to a block $B$:
@@
 
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+\subsubsection{Transaction Receipt}\linkdest{Transaction Receipt}{}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_{\mathrm{e}}$.
+
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, $R_{\mathbf{l}}$ and the Bloom filter composed from information in those logs, $R_{\mathrm{b}}$ and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+v_0 \equiv T_{\mathrm{g}} T_{\mathrm{p}} + T_{\mathrm{v}}
 \end{equation}
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_s}{}And finally, we specify $\Upsilon^g$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^z$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
+\Upsilon^z(\boldsymbol{\sigma}, T) & \equiv & z
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
 
@@
 \mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
+\end{eqnarray}
+\nopagebreak[1]where
+\begin{equation}
+\Xi \equiv \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
+\end{equation}
+
+\begin{equation}
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
 
@@
 The machine can have exceptional execution for several reasons, including stack underflows and invalid instructions.
-Like the out-of-gas exception, they leave the state intact.
+Like the out-of-gas exception, they do not leave state changes intact.
 Rather, the machine halts immediately and reports the issue to the execution agent (either the transaction processor or, recursively, the spawning execution environment) which will deal with it separately.
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
-\varnothing & \text{otherwise}
-\end{cases}
-\end{equation}
-
-The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \ w \in \{\text{\small RETURN}, \text{\small REVERT}\} &\\
+() \quad\quad\ \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SELFDESTRUCT} \} &\\
+\varnothing \quad\quad\ \text{otherwise}&
+\end{cases}
+\end{equation}
+
+The data-returning halt operations, \text{\small RETURN} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+\end{eqnarray}
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
 \begin{equation}
 \mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\end{equation}
+
+We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
 \begin{equation}
 \mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\Upsilon^z(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
+\end{equation}
+
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}
@@
 
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
+Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
 
@@
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+&&is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
+&&non-zero.\\
+\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
 \linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
+&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
+&\quad\text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
+&&&& modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
+&&&& modulo. \\
 \midrule
@@
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
+&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
+&&&& in big endian). \\
 \bottomrule
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \toprule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
-&&&& This is the sender of original transaction; it is never an account with non-empty \\
-&&&& associated code. \\
+&&&& This is the sender of original transaction; it is never an account with\\
+&&&& non-empty associated code. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
-&&&& This is the address of the account that is directly responsible for this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
+&&&& This is the address of the account that is directly responsible for\\
+&&&& this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
+&&&& this execution. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
+&&&& This pertains to the input data passed with the message call\\
+&&&& instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
+&&&& environment. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+&&&& This pertains to the input data passed with the message call\\
+&&&& instruction or transaction. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+&&&& This pertains to the input data passed with the message call instruction\\
+&&&& or transaction. \\
 \midrule
@@
 \midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
+&&&& environment. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
@@
 \linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than\\
+&&&& the current block number or more than 256 blocks behind the current block.\\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent\\
+&&&& hash in the block following it. \\
 \midrule
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+&&&&\linkdest{C tiny SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
 G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
-&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
+&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
-&&&& created account. \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\ 
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
+&&&& address of the newly created account. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
+&&&& fact the same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\end{tabular*}
-
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
+&&&& persisting the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
+&&&& omitted argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the\\
+&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and\\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced \\
+&&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 \varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\ (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise} \end{cases}$\\
 &&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
@@
 \bottomrule
-\end{tabular*}
-
-\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
+\end{tabu}
+
+\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
 
@@
 
-\section{List of mathematical symbols} \label{app:symbols}
-\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{c c X[j]} \savetabu{col3}
 \toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+Symbol & Latex Command & Description \\
+\midrule
+\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
+Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
 \bottomrule

### 2018-02-16 Paper.tex
Replaces #548

--- 2018-02-16 Paper.tex
@@
 
+\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
+
+\input{Version.tex}
 % Default rendering options
@@
 M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+>>>>>>> master
 \end{equation}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
-\varnothing & \text{otherwise}
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
+() \quad\quad\ \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
+\varnothing \quad\quad\ \text{otherwise}&
 \end{cases}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
 \end{cases}
@@
 
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}

### 2018-02-16 Paper.tex

--- 2018-02-16 Paper.tex
@@
 \end{description}
-
 \linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.

### 2018-02-16 Paper.tex

@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 
-When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
+When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
 
@@
 
-I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
+Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
 \begin{equation}
@@
 \end{description}
+
 \linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.

### 2018-02-16 Paper.tex
\hyperlink{hhalt} zhalt

@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
 

### 2018-02-16 Paper.tex
\hyperlink{stop}{STOP}

--- 2018-02-16 Paper.tex
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small text{\hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}

### 2018-02-16 Paper.tex
link{H__g}

@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small text{\hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_H}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
 \begin{equation}

### 2018-02-16 Paper.tex
\hyperlink{H g}{${B_H}_{\mathrm{g}}$} link{R__u}

@@
 \begin{equation}
-{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+\hyperlink{H g}{${B_H}_{\mathrm{g}}$} = \hyperlink{ell}{\ell}\hyperlink{R__u}{(\mathbf{R})_{\mathrm{u}}}
 \end{equation}

### 2018-02-16 Paper.tex
link{R__block}

--- 2018-02-16 Paper.tex
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by \hyperlink{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 
-\hypertarget{block_reward_R__block}{}We define the block reward as 3 Ether:
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
 \begin{equation}

### 2018-02-16 Paper.tex

--- 2018-02-16 Paper.tex
@@
 
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
 \begin{equation}
@@
 
-We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
 \begin{equation}
@@
 
-We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
 \begin{equation}
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_H}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
 \begin{equation}

### 2018-02-16 Paper.tex
\hyperlink{Gamma}{\Gamma}

--- 2018-02-16 Paper.tex
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
+\hypertarget{block_reward_R__block}{}We define the block reward as 3 Ether:
 \begin{equation}
@@
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
 \end{equation}

### 2018-02-16 Paper.tex
\hyperlink{Omega}{\Omega}

@@
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
 \end{equation}

### 2018-02-16 Paper.tex
Footnote clarifying Ethereum 1.0 and link to wiki FAQs on PoS & sharding

--- 2018-02-16 Paper.tex
@@
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
+For Ethereum 1.0 we have chosen the first path.\footnote{Ethereum 1.0 uses proof-of-work, while it is planned that Ethereum 2.0 will include \href{https://github.com/ethereum/wiki/wiki/Proof-of-Stake-FAQ}{proof-of-stake} and \href{https://github.com/ethereum/wiki/wiki/Sharding-FAQ}{sharding}.}
 

### 2018-02-16 Paper.tex
\hyperlink{block nonce}{nonce}

--- 2018-02-16 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
+Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 %,pagebackref=true
@@
 where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
+In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful than any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
 
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \ref{ch:block}.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
+This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
 
@@
 When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \ref{ch:block}.
+It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
 
@@
 
-The account state comprises the following four fields:
+The account state, $\boldsymbol{\sigma}[a]$, comprises the following four fields:
 
@@
 \begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
+k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
 \end{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
 \end{equation}
@@
 
-\subsection{The Block} \label{ch:block}
+\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
 
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 
-We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\begin{equation}
+R_{\mathrm{z}} \in \mathbb{N}
+\end{equation}
+
+\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
 \end{equation}
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
 \hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
 \begin{equation*}
-x \equiv \begin{cases}
+y \equiv \begin{cases}
 1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
@@
 & & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
+\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} \geqslant 5000
 \end{eqnarray}
@@
 V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-& & H_{\mathrm{d}} = D(H) \quad \wedge \\
-& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-& & \lVert H_{\mathrm{x}} \rVert \le 32
+\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
+\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
+\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
 \end{eqnarray}
@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
 \end{align}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \right\}
 \end{equation}
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{account nonce}.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
 Thus we define the resultant address for the new account $a$:
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\mathbf{a}_1 \equiv \left(\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}}\right)
 \end{equation}
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
 the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
-\boldsymbol{\mu}_{pc} & \equiv & 0 \\
+\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
 \boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
@@
 
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert \\
 \text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
+\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
+ (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
 \end{array}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
-\varnothing & \text{otherwise}
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
+() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
+\varnothing \quad\quad\ \text{otherwise}&
 \end{cases}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
+\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
 \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
 \end{eqnarray}
@@
 \quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
 \hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 \hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
 \end{cases}
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{block nonce}.
+\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
 \end{enumerate}
@@
 \begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
 \end{eqnarray}
@@
 
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
 \begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\mathtt{\tiny RLP}(i : i \in \mathbb{N}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
 \end{equation}
@@
 
-We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|I_{\mathbf{d}}|-1)] &=& I_{\mathbf{d}}\\
+\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
 \mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
+g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
 f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
+x^2 & \text{if}\ x \le 64 \\[0.5em]
+\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
+\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
 \end{cases}\\
@@
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{P}_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\
@@
 i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
+I_{\mathbf{d}}[x] &\text{if}\ x < |I_{\mathbf{d}}| \\
 0 &\text{otherwise}
@@
 \end{equation}
-We define a binary operation $+$ on $C_1$ with
+We define a binary operation $+$ on $C_1$ for distinct elements $(X_1, Y_1), (X_2, Y_2)$ with
 \begin{eqnarray}\label{eq:ec-addition}
@@
 
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+Let $F_{p^2}$ be a field $F_{p}[i]/(i^2+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3(i+9)^{-1}\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.
@@
 
+Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$.
+It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
+This bilinear map is a type three pairing.
+There are several such bilinear maps, it does not matter which is chosen to be $e$.
+Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
+\begin{eqnarray} \label{eq:pairing-check}
+\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
+\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
+\end{eqnarray}
+Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).
+
 A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
 \begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
+\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
+\mathbf{x}&\text{if}\ \mathbf{x}<p\\
 \varnothing&\text{otherwise}
@@
 \end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
+x&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
+y&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf{x}\in\mathbf{B}_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf{x})&\equiv&\begin{cases}
 g_2&\text{if}\ g_2\in G_2\\
@@
 \end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
+\end{eqnarray}
+
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.
+
 \begin{eqnarray}
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
 F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
@@
 g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
+\mathbf{o}[0..31]&\equiv&\begin{cases}
 0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
@@
 \end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
+a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf{d}}[64..191])\\\nonumber
 \vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_k&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
 \end{eqnarray}
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+\mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
+y&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[64..127]\right)\\
+\label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
+I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
 0&\text{otherwise}
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf{d}}$ is defined in (\ref{eq:complemented_input}).
 \begin{eqnarray}
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+\mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf{d}}[0..31]\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[32..95]\right)
 \end{eqnarray}
@@
 
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
+This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
 
@@
 
-We assume the existence of functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
+These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
 \mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
+It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 
@@
 
-We declare that a signature is invalid unless all the following conditions are true:
+\linkdest{invalidsig}We declare that a signature is invalid unless all the following conditions are true:
 \begin{align}
@@
 0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 
+\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
+\begin{eqnarray}
+\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
+\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
+\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
+\end{eqnarray}
+
 We may then define the sender function $S$ of the transaction as:
@@
 \linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
+&&contract. \\
 
@@
 G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
 G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
+G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
 G_{extcode} & \text{if} \quad w \in W_{extcode}\\
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
+G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
 \end{equation}
@@
 \linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than\\
+&&&& the current block number or more than 256 blocks behind the current block.\\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
+&&&& $h$ is zero.\\
 \midrule
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}.\\
 \midrule
 \linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
 &&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 \linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
 \linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
-\end{tabular*}
+\end{tabu*}
 
@@
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
 \linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
 \begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+ d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+ c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
+x & \text{if} \quad x / y \in \mathbb{N} \\
+E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+ n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}

### 2018-02-16 Paper.tex
link{H__m}

--- 2018-02-16 Paper.tex
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
 \item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}

### 2018-02-16 Paper.tex
\hyperlink{H__d}{H_{\mathrm{d} \hyperlink{H_cancel_n}{\hcancel{n}
were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash})

@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
 \item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
@@
 \begin{equation}
-m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}

### 2018-02-16 Paper.tex

@@
 Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
 The general route that the algorithm takes is as follows:

### 2018-02-16 Paper.tex
links: J__datasetinit, J__datasetinit, J__epoch

--- 2018-02-16 Paper.tex
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.

### 2018-02-16 Paper.tex
\hyperlink{blockhash}{BLOCKHASH}

--- 2018-02-16 Paper.tex
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
 $G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
@@
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
 \end{cases}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+\linkdest{blockhash}{}0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\

### 2018-02-16 Paper.tex
consolidating

@@
 
-The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
-Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+The state database won't be forced to maintain all past state trie structures into the future.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
 
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{Genesis_Block}{genesis block}.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the genesis block.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.

### 2018-02-16 Paper.tex
Mist

@@
 
-\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
+\item[Mist] A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, and is developed by \cite{Mist}.
 

### 2018-02-16 Paper.tex
the keys of nibbles and the keys of branches as

--- 2018-02-16 Paper.tex
@@
 \begin{description}
-\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
+\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
 The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.

### 2018-02-16 Paper.tex
\mathrm{g}

--- 2018-02-16 Paper.tex
@@
 \begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \mathrm{g}, I, T) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad \mathrm{g} < \mathbf{g}_{\mathrm{r}} \\
+(\boldsymbol\sigma, \mathrm{g} - \mathrm{g}_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $\mathrm{g}_{\mathrm{r}}$, the gas requirements.
 

### 2018-02-16 Paper.tex
\mathrm{g}
\hyperlink{I__d}{I_{\mathbf{d}

--- 2018-02-16 Paper.tex
@@
 \item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
@@
 \begin{description}
-\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
+\item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
 The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
@@
 (T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+\end{eqnarray}
+
+\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
+\begin{eqnarray}
+\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
+\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
+\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
 \end{eqnarray}

### 2018-02-16 Paper.tex
More \mathrm{g}s
Citations for signing transactions

@@
 \item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
+\mathbf{d}[0..(|I_{\mathbf{d}}|-1)] &=& I_{\mathbf{d}}\\
 \mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\

### 2018-02-16 Paper.tex
\ e & \equiv & \hyperlink{h T}{h(T)}\\

--- 2018-02-16 Paper.tex
@@
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\ e & \equiv & \hyperlink{h T}{h(T)}\\
 \linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\

### 2018-02-16 Paper.tex
Move \hypertarget{v}{}

@@
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\ e & \equiv & \hyperlink{h T}{h(T)}\\
 \linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.

### 2018-02-16 Paper.tex
Replace all _(\{[:alnum:]+\}) with _{\mathrm$1}...
...using Libre Open Office with regular expressions ticked as an option.

@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
+\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{\mathrm{t}}, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{\mathrm{t}}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
@@
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{\mathrm{32}}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{\mathrm{256}}$.
 This is formally defined in section \hyperlink{block}.
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\texttt{\small TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
+\begin{equation}
+L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 \begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
+k \in \mathbb{B}_{\mathrm{32}} \quad \wedge \quad v \in \mathbb{P}
 \end{equation}
@@
 
-Thus we may define a world-state collapse function $L_{S}$:
-\begin{equation}
-L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+Thus we may define a world-state collapse function $L_{\mathrm{S}}$:
+\begin{equation}
+L_{\mathrm{S}}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 
-This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
-\begin{equation}
-\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
+This function, $L_{\mathrm{S}}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+\begin{equation}
+\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{\mathrm{20}} \; \wedge \; v(\boldsymbol{\sigma}[a]))
 \end{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{\mathrm{256}} \wedge x_{\mathrm{b}} \in \mathbb{P}_{\mathrm{256}} \wedge x_{\mathrm{s}} \in \mathbb{B}_{\mathrm{32}} \wedge x_{\mathrm{c}} \in \mathbb{B}_{\mathrm{32}}
 \end{equation}
@@
 \begin{equation}
-L_{T}(T) \equiv \begin{cases}
+L_{\mathrm{T}}(T) \equiv \begin{cases}
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
@@
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{\mathrm{256}} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{\mathrm{256}} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{\mathrm{256}} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{\mathrm{256}} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{\mathrm{256}} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{\mathrm{256}} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
-\mathbb{B}_{0} & \text{otherwise}\end{cases}
+T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{\mathrm{20}} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+\mathbb{B}_{\mathrm{0}} & \text{otherwise}\end{cases}
 \end{equation}
@@
 \begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
+B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
 \end{equation}
@@
 
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
+The function $L_{\mathrm{R}}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{\mathrm{R}}(R) \equiv (0 \in \mathbb{B}_{\mathrm{256}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+\end{equation}
+where $0 \in \mathbb{B}_{\mathrm{256}}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
 
@@
 \begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{\mathrm{256}}
+\end{equation}
+
+%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{\mathrm{B}}$; it is merely a convenience equivalence.
 
@@
 \begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+O_{\mathrm{a}} \in \mathbb{B}_{\mathrm{20}} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{\mathrm{32}} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{eqnarray}
-M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{256} \quad \text{where:}\\
+M_{3:2048}(\mathbf{x}: \mathbf{x} \in \mathbb{B}) & \equiv & \mathbf{y}: \mathbf{y} \in \mathbb{B}_{\mathrm{256}} \quad \text{where:}\\
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{T}}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{\mathrm{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{\mathrm{H}})$ is the parent block of $B$, defined directly.
 
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_{\mathrm{R}}$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
 \end{eqnarray}
@@
 \begin{array}[t]{lclclcl}
-\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
+\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{\mathrm{32}} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{\mathrm{32}} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{\mathrm{20}} & \wedge \\
+\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{\mathrm{32}} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{\mathrm{32}} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{\mathrm{32}} & \wedge \\
+\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{\mathrm{256}} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{\mathrm{256}} & \wedge \\
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{\mathrm{32}} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{\mathrm{8}}
 \end{array}
@@
 
-We define $P(B_{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+We define $P(B_{\mathrm{H}})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{\mathrm{H}})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
 \end{equation}
@@
 \begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
+H_{\mathrm{i}} \equiv {{P(H)_{\mathrm{H}}}_{\mathrm{i}}} + 1
 \end{equation}
@@
 \mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
+\text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
 \end{dcases}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
 & & H_{\mathrm{l}} \geqslant 5000
@@
 \begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}}
 \end{equation}
@@
 & & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
 & & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
+& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
 & & \lVert H_{\mathrm{x}} \rVert \le 32
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
 
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
+\Lambda_{\mathrm{4}}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{\mathrm{4}}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\Theta_{\mathrm{4}}$ and $\Lambda_{\mathrm{4}}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{\mathrm{P}} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}}
 \end{eqnarray}
@@
 
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{\mathrm{H}}$ has no special treatment and is determined from the blockchain.
 
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{\mathrm{256}}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
+D(\mathbf{c}) \equiv D_{\mathrm{J}}(\mathbf{c}, 0)
 \end{equation}
@@
 \begin{equation}
-D_{J}(\mathbf{c}, i) \equiv \begin{cases}
+D_{\mathrm{J}}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
-\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
 \hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 \hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
 \end{cases}
@@
 B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
-B' & \equiv & P(B_{H})
+B' & \equiv & P(B_{\mathrm{H}})
 \end{eqnarray}
@@
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_{\mathrm{H}}, n - 1) & \text{otherwise}
 \end{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{\mathrm{H}}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{\mathrm{H}}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{\mathrm{H}}}_{\mathrm{i}})) R_{\mathrm{block}}
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{\mathrm{H}}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_{\mathrm{cacheinit}}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $J_{\mathrm{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $J_{\mathrm{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_{n}\rVert} \quad \mathbf{k}_{n}'[i] & \equiv &
+\forall_{\mathrm{n}} \quad \forall_{i: i < 2\lVert\mathbf{k}_{\mathrm{n}}\rVert} \quad \mathbf{k}_{\mathrm{n}}'[i] & \equiv &
 \begin{cases}
-\lfloor \mathbf{k}_{n}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
-\mathbf{k}_{n}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
+\lfloor \mathbf{k}_{\mathrm{n}}[i \div 2] \div 16 \rfloor & \text{if} \; i \; \text{is even} \\
+\mathbf{k}_{\mathrm{n}}[\lfloor i \div 2 \rfloor] \bmod 16 & \text{otherwise}
 \end{cases}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{\mathrm{x}} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
@@
 \begin{eqnarray}
-\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{\mathrm{32}} \\
+\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{\mathrm{20}}
 \end{eqnarray}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
+g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{\mathrm{quaddivisor}}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+\ell'_{\mathrm{E}} &=& \begin{cases}
+0 & \text{if}\ \ell_{\mathrm{E}}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{\mathrm{E}}\le 32 \wedge E \neq 0 \\
+8(\ell_{\mathrm{E}} - 32) + \lfloor \log_2(i[(96+\ell_{\mathrm{B}})..(127+\ell_{\mathrm{B}})]) \rfloor & \text{if}\ 32 < \ell_{\mathrm{E}} \wedge i[(96 + \ell_{\mathrm{B}})..(127 + \ell_{\mathrm{B}})]\neq 0 \\
+8(\ell_{\mathrm{E}} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{\mathrm{M}}} \\
+\ell_{\mathrm{B}} &\equiv& i[0..31] \\
+\ell_{\mathrm{E}} &\equiv& i[32..63] \\
+\ell_{\mathrm{M}} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{\mathrm{B}})] \\
+E &\equiv& i[(96+\ell_{\mathrm{B}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})] \\
+M &\equiv& i[(96+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}}+\ell_{\mathrm{M}})] \\
 i[x] &\equiv& \begin{cases}
@@
 \begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
+n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{\mathrm{n}}
 \end{equation}
@@
 
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
+A 32 byte number $\mathbf{x}\in\mathbf{P}_{\mathrm{256}}$ might and might not represent an element of $F_{\mathrm{p}}$.
 \begin{equation}
@@
 
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+A 64 byte data $\mathbf x\in\mathbf B_{\mathrm{512}}$ might and might not represent an element of $G_1$.
 \begin{eqnarray}
@@
 
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+A 128 byte data $\mathbf x\in\mathbf B_{\mathrm{1024}}$ might and might not represent an element of $G_2$.
 \begin{eqnarray}
@@
 \begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{\mathrm{32}}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{\mathrm{64}} \\
+\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{\mathrm{32}}, p_{\mathrm{r}} \in \mathbb{B}_{\mathrm{32}}) & \equiv & (v \in \mathbb{B}_{\mathrm{1}}, r \in \mathbb{B}_{\mathrm{32}}, s \in \mathbb{B}_{\mathrm{32}}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{\mathrm{32}}, v \in \mathbb{B}_{\mathrm{1}}, r \in \mathbb{B}_{\mathrm{32}}, s \in \mathbb{B}_{\mathrm{32}}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{\mathrm{64}}
 \end{eqnarray}
@@
 \begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
+L_{\mathrm{S}}(T) & \equiv & \begin{cases}
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_{\mathrm{S}}(T) )
 \end{eqnarray}
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
+$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
+$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
+$G_{\mathrm{extcode}}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{\mathrm{balance}}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{\mathrm{sload}}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{\mathrm{jumpdest}}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
+\linkdest{R__selfdestruct}{}\\$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
+\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{\mathrm{call}}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
+\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by $\lceil\log_{\mathrm{256}}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
 \linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
+$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{\mathrm{sha3}}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{\mathrm{sha3word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{\mathrm{quaddivisor}}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{\mathrm{mem}}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{\mathrm{mem}}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{\mathrm{256}}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
+G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
+G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
+G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
+G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
+G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
+G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+C_{\mathrm{mem}}(a) \equiv G_{\mathrm{memory}} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
-
-$W_{high}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+
+$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+
+$W_{\mathrm{high}}$ = \{{\small JUMPI}\}
+
+$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}\}
+
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_{\mathrm{mem}}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{l}}$ \\
 \bottomrule
@@
 &&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
+G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
+G_{\mathrm{sreset}} & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
+&&&&\linkdest{A r}{} $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
+R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{\mathrm{i}}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{g}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{g}}$ \\
 \midrule
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
+&&&& $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
+R_{\mathrm{selfdestruct}} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{selfdestruct}} + \begin{cases}
+G_{\mathrm{newaccount}} & \text{if} \quad n \\
 0 & \text{otherwise}
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{\mathrm{256}}, \mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big), 0_{\mathrm{160}}, stateRoot, 0, 0, 0_{\mathrm{2048}}, 2^{17}, 0, 0, 3141592, time, 0, 0_{\mathrm{256}},  \mathtt{\tiny KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{\mathrm{256}}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{\mathrm{160}}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{\mathrm{2048}}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
@@
 \midrule
-\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{Jwordbytes}{}$J_{\mathrm{wordbytes}}$ & 4  & Bytes in word. \\
+\linkdest{J__datasetinit}{}$J_{\mathrm{datasetinit}}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{\mathrm{cacheinit}}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{\mathrm{epoch}}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{\mathrm{mixbytes}}$ & 128 & mix length in bytes. \\
+\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__cacherounds}{}$J_{\mathrm{cacherounds}}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 \begin{equation}
- E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+ E_{\mathrm{epoch}}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{\mathrm{epoch}}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{\mathrm{datasetgrowth}}$ bytes, and the size of the cache by $J_{\mathrm{cachegrowth}}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
-\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+Therefore the size is reduced by a multiple of $J_{\mathrm{mixbytes}}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
+\linkdest{d__size}{}Let $d_{\mathrm{size}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{\mathrm{size}} = E_{\mathrm{prime}}(J_{\mathrm{datasetinit}} + J_{\mathrm{datasetgrowth}} \cdot E_{\mathrm{epoch}} - J_{\mathrm{mixbytes}}, J_{\mathrm{mixbytes}})
+\end{equation}
+The size of the cache, $c_{\mathrm{size}}$, is calculated using
+\begin{equation}
+ c_{\mathrm{size}} = E_{\mathrm{prime}}(J_{\mathrm{cacheinit}} + J_{\mathrm{cachegrowth}} \cdot E_{\mathrm{epoch}} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes.
+It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{\mathrm{32}}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
-\end{cases}
-\end{equation}
-With $\mathbf{0}_{32}$ being 32 bytes of zeros.
+ \mathbf{s} = C_{\mathrm{seedhash}}(H_{\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{\mathrm{seedhash}}(H_{\mathrm{i}}) = \begin{cases}
+\texttt{KEC}(\mathbf{0}_{\mathrm{32}}) & \text{if} \quad E_{\mathrm{epoch}}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{\mathrm{seedhash}}(H_{\mathrm{i}} - J_{\mathrm{epoch}})) & \text{otherwise}
+\end{cases}
+\end{equation}
+With $\mathbf{0}_{\mathrm{32}}$ being 32 bytes of zeros.
 
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially filling up $c_{\mathrm{size}}$ bytes of memory, then performing $J_{\mathrm{cacherounds}}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
 
-We define the array $\mathbf{c}_{i}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
-\begin{equation}
- \mathbf{c}_{i} = \begin{cases}
+We define the array $\mathbf{c}_{\mathrm{i}}$, consisting of 64 single bytes,  as the $i$th element of the initial cache:
+\begin{equation}
+ \mathbf{c}_{\mathrm{i}} = \begin{cases}
 \texttt{KEC512}(\mathbf{s}) & \text{if} \quad i = 0 \quad  \\
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ \mathbf{c'}[i] = \mathbf{c}_{\mathrm{i}} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{\mathrm{size}}}{J_{\mathrm{hashbytes}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{\mathrm{cacherounds}}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{\mathrm{rmh}}(\mathbf{x}, 0), E_{\mathrm{rmh}}(\mathbf{x}, 1), ... , E_{\mathrm{rmh}}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
 \begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+  E_{\mathrm{rmh}}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{\mathrm{rmh}}(\mathbf{x}, j) \quad \forall \quad j < i
 \end{multline}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
+Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_{\mathrm{hashbytes}}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{\mathrm{datasetitem}}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{\mathrm{size}}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+ E_{\mathrm{datasetitem}}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{\mathrm{size}}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{\mathrm{wordbytes}} \rfloor]) \mod c_{\mathrm{size}}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a ``mix'' $J_{\mathrm{mixbytes}}$ bytes wide, and repeatedly sequentially fetch $J_{\mathrm{mixbytes}}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{\mathrm{mixbytes}}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
-\end{equation}
-$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
+\end{equation}
+$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
-\begin{equation}
- E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
-E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ n_{\mathrm{mix}} =  \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{hashbytes}}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
+\begin{equation}
+ E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
+E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+\end{equation}
+$E_{\mathrm{newdata}}$ returns an array with $n_{\mathrm{mix}}$ elements:
+\begin{equation}
+ E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{wordbytes}}}\right\rfloor]) \mod \left\lfloor\frac{d_{\mathrm{size}} / J_{\mathrm{hashbytes}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
 \end{equation}

### 2018-02-16 Paper.tex

--- 2018-02-16 Paper.tex
@@
 When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}.
+It is formally defined in section \ref{ch:block}.
 
@@
 (T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
-\end{eqnarray}
-
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
-\begin{eqnarray}
-\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
-\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
-\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
 \end{eqnarray}

### 2018-02-16 Paper.tex
I__e => I__w

--- 2018-02-16 Paper.tex
@@
 
-% Default rendering options
-\definecolor{pagecolor}{rgb}{1,0.98,0.9}
-\def\YellowPaperVersionNumber{unknown revision}
-\IfFileExists{Options.tex}{\input{Options.tex}}
+\input{Version.tex}
 
@@
 
-Overall, we wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
+Overall, I wish to provide a system such that users can be guaranteed that no matter with which other individuals, systems or organisations they interact, they can do so with absolute confidence in the possible outcomes and how those outcomes might come about.
 
@@
 
-We use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
+I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
 
@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 
-When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
+When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
 
@@
 
-We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+I define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 
-Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
 \begin{equation}
@@
 \item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
+\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
@@
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
 Formally, we can refer to a block $B$:
@@
 This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
 
@@
 
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
 Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
+\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the sender account's current nonce);
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
 \item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
 \begin{eqnarray}
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
 
@@
 \boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+\hypertarget{Phi}{}And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_{\mathrm{r}} = r(\Pi(\Gamma(B), B))
+\end{eqnarray}
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\boldsymbol{\sigma}[n] = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
+For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
 \begin{equation}
@@
 
-We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
+We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
 \begin{equation}
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
 \begin{equation}
@@
 
-There are several patterns of contracts engineering that allow particular useful behaviours; two of these that we will briefly discuss are data feeds and random numbers.
+There are several patterns of contracts engineering that allow particular useful behaviours; two of these that I will briefly discuss are data feeds and random numbers.
 
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
-We have introduced, discussed and formally defined the protocol of Ethereum.
+I have introduced, discussed and formally defined the protocol of Ethereum.
 Through this protocol the reader may implement a node on the Ethereum network and join others in a decentralised secure social operating system.
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
@@
 $\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
@@
 \end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
 \begin{equation}
@@
 
-\section{List of mathematical symbols} \label{app:symbols}
-\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
+\section{List of mathematical symbols}\label{app:symbols}
+\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
 \toprule

### 2018-02-16 Paper.tex

@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
-\end{equation}
-
-where $\Upsilon$ is the Ethereum state transition function.
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
+\end{equation}
+\pagebreak[1]where $\Upsilon$ is the Ethereum state transition function.
 In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
-\end{equation}
-
-where:
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\begin{equation}
+L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+\end{equation}
+\pagebreak[1]where:
 \begin{equation}
@@
 \end{equation}
-\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
-\end{equation}
-
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-\subsection{The Transaction} \label{subsec:transaction}
+where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+\subsection{The Transaction} \label{ch:transaction}
 
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+While is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
 \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
@@
 \begin{equation}
-L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+L_T(T) \equiv \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_{\mathrm{i}}$.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_{\mathrm{s}}$.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
 
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
-\end{equation}
-
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
-\begin{equation}
-M(O) \equiv \bigvee_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
-\end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+\end{equation}
+
+We define the Bloom filter function, $M$, to reduce a log entry include a single 256-byte hash:
+\begin{equation}
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+\end{equation}
+\pagebreak[1]where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+\pagebreak[1]where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
+\pagebreak[1]where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
 \begin{equation}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the transaction receipts, $B_{\mathbf{R}}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
+\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 
-The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(B_{\mathbf{U}}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 \begin{array}[t]{lclclcl}
-H_{\mathrm{p}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-H_{\mathrm{r}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{e}} \in \mathbb{B}_{32} & \wedge \\
-H_{\mathrm{b}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & H_{\mathrm{i}} \in \mathbb{P} & \wedge \\
-H_{\mathrm{l}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & H_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge \\
-H_{\mathrm{x}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{n}} \in \mathbb{B}_{8}
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
 \end{equation}
-
-where
-\begin{equation}
-\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\pagebreak[1]where
+\begin{equation}
+\mathbb{B}_n = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
-\end{equation}
-
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
-\hypertarget{block_difficulty_H__d}{}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
-where:
-\begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the demoninator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by substracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+H_i \equiv {{P(H)_H}_i} + 1
+\end{equation}
+
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\begin{equation}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
+\end{equation}
+\pagebreak[1]where
+\begin{equation}
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 125000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 
-The nonce, $H_{\mathrm{n}}$, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+
+Above, $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define $\Upsilon^{\mathrm{g}}$ to evaluate to the amount of gas used in the execution of a transaction, $\Upsilon^{\mathbf{l}}$ to evaluate to the transaction's accrued log items and $\Upsilon^{\mathrm{z}}$ to evaluate to the status code resulting from the transaction.
-These will be formally defined later.
+We also define $\Upsilon^g$ to evaluate to the amount of gas used in the execution of a transaction and $\Upsilon^\mathbf{l}$ to evaluate to the transaction's accrued log items, both to be formally defined later.
 
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + G_{transaction}
-\end{align}
-
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
+\pagebreak[1]where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G$ is defined in Appendix \ref{app:fees}.
+
+%todo Explain g_d reason?
 
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \ell(B_{\mathbf{R}})_{\mathrm{u}}
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leqslant & T_g \quad \wedge \\ 
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\ell(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_g$, and the gas utilised in this block prior, given by $\ell(B_\mathbf{R})_u$, must be no greater than the block's \textbf{gasLimit}, ${B_H}_l$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the nonce of the account of the sender, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_gT_p$.
+The gas available for the proceeding computation, $g$, is defined as $T_g - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
-\begin{equation}
-g \equiv T_{\mathrm{g}} - g_0
-\end{equation}
-and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
+\end{cases}
+\end{equation}
+\pagebreak[1]where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+\begin{equation}
+g \equiv T_g - g_0
+\end{equation}
+and $T_o$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
+
+Note we use $\Theta_{3}$ to denote the fact that only the first three components of the function's value are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A_{\mathrm{r}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to $A_{\mathrm{r}}$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\hypertarget{tx_logs_Upsilon_pow_l}{}\hypertarget{tx_status_Upsilon_pow_z}{}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_{\mathbf{l}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt}.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
+
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
+\section{Contract Creation} \label{ch:create}
+
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+\pagebreak[1]where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
-\end{eqnarray}
-
-where $v'$ is the account's pre-existing value, in the event it was previously in existence:
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
+\end{eqnarray}
+\pagebreak[1]where $v'$ is the account's pre-existing value, in the event it was previously in existence:
 \begin{equation}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
-
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**} < c \\
 g^{**} - c & \text{otherwise} \\
 \end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
+\begin{eqnarray}
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & () \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_e & \equiv & e
+\end{eqnarray}
+\pagebreak[1]where $c$ is the code-deposit cost:
+\begin{equation}
+c \equiv G_{codedeposit} \times |\mathbf{o}|
+\end{equation}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[r]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[r]_{\mathrm{b}} + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_{\mathrm{b}} \equiv \boldsymbol{\sigma}[s]_{\mathrm{b}} - v
-\end{equation}
-unless $s = r$.
+\boldsymbol{\sigma}_1[r]_b \equiv \boldsymbol{\sigma}[r]_b + v \quad\wedge\quad \boldsymbol{\sigma}_1[s]_b \equiv \boldsymbol{\sigma}[s]_b - v
+\end{equation}
 
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\boldsymbol{\sigma}_1'[s]_b \equiv \boldsymbol{\sigma}_1[s]_b - v
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_e & \equiv & e \\
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 
+%More formally, given an instruction, it is possible to calculate the gas cost of executing it as follows:
+%
+%\begin{itemize}
+%\item {\small SHA3} costs $G_{sha3}$ gas
+%\item {\small SLOAD} costs $G_{sload}$ gas
+%\item {\small BALANCE} costs $G_{balance}$ gas
+%\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
+%\begin{itemize}
+%\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
+%\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
+%\item $d = 1$ otherwise.
+%\end{itemize}
+%\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of the account's associated code, if non-empty.
+%\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of the account initialisation code.
+%\item {\small STOP} costs $G_{stop}$ gas
+%\item {\small SUICIDE} costs $G_{suicide}$ gas
+%\item All other operations cost $G_{step}$ gas.
+%\end{itemize}
+%
+%Additionally, when memory is accessed with {\small MSTORE}, {\small MSTORE8}, {\small MLOAD}, {\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURN}, {\small SHA3}, {\small CREATE} or {\small CALL}, the memory should be enlarged to the smallest multiple of words such that all addressed bytes now fit in it.
+
 See Appendix \ref{app:vm} for a rigorous definition of the EVM gas cost.
 
-\subsection{Execution Environment}\linkdest{exec_env}
+%Whenever a higher memory index is referenced, the fee difference to take it to the higher usage from the original (lower) usage is charged.
+Notably, because {\small MSTORE} and {\small MLOAD} operate on word lengths, they implicitly increase the highest-accessed index to 31 greater than their target index.
+
+\subsection{Execution Environment}
 
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
-For the present context, we will define it as:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the suicide list $\mathbf{s}$, the log series $\mathbf{l}$, the refunds $r$ and the resultant output, $\mathbf{o}$, given these definitions:
+\begin{equation}
+(\boldsymbol{\sigma}', g', \mathbf{s}, \mathbf{l}, r, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
+\end{cases}
+\end{equation}
+\pagebreak[1]where
 \begin{eqnarray}
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
+\end{eqnarray}
+
+Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
 
@@
 \begin{equation}
-D(\mathbf{c}) \equiv D_{J}(\mathbf{c}, 0)
-\end{equation}
-
-where:
-\begin{equation}
-D_{J}(\mathbf{c}, i) \equiv \begin{cases}
+D(\mathbf{c}) \equiv D_J(\mathbf{c}, 0)
+\end{equation}
+\pagebreak[1]where
+\begin{equation}
+D_J(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
+\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\end{cases}
+\end{equation}
+\pagebreak[1]where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
 \begin{equation}
@@
 
-In general, we assume the memory, self-destruct set and system state don't change:
-\begin{eqnarray}
-\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
-\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
+In general, we assume the memory, suicide list and system state don't change:
+\begin{eqnarray}
+\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
+\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
 A' & \equiv & A \\
@@
 \begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} V({\mathbf{U}}) \; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
-\end{equation}
-
-where $k$ denotes the ``is-kin'' property:
+\lVert B_\mathbf{U} \rVert \leqslant 2 \bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; k(U, P(B_H)_H, 6)
+\end{equation}
+\pagebreak[1]where $k$ denotes the ``is-kin'' property:
 \begin{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by $R_{\mathrm{b}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
+\begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{b}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{b}}
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 
-\hypertarget{block_reward_R__b}{}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{b}} = 3 \times 10^{18}
+We define the block reward as 5 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 5 \times 10^{18}
 \end{equation}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-We define $\mathbf{R}[n]_{\mathrm{z}}$ in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathbf{l} = 
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
 \end{equation}
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
+m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
+\end{equation}
+\pagebreak[1]where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 
-Many thanks to Aeron Buchanan for authoring the \textit{Homestead} revisions, Christoph Jentzsch for authoring the Ethash algorithm and Yoichi Hirai for doing most of the EIP-150 changes.
-Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Gustav Simonsson, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic, Giacomo Tazzari and, of course, Vitalik Buterin.
-
-\section{Availability}
-
-The source of this paper is maintained at \url{https://github.com/ethereum/yellowpaper/}. An auto-generated PDF is located at \url{https://ethereum.github.io/yellowpaper/paper.pdf}.
-
+Important maintenance, useful corrections and suggestions were provided by a number of others from the Ethereum DEV organisation and Ethereum community at large including Christoph Jentzsch, Gustav Simonsson, Aeron Buchanan, Pawe\l{} Bylica, Jutta Steiner, Nick Savers, Viktor Tr\'{o}n, Marko Simovic and, of course, Vitalik Buterin.
+
+\bibliography{Biblio}
 \bibliographystyle{plainnat}
-\bibliography{Biblio}
 
@@
 
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
+\section{Modified Merkle Patricia Tree}\label{app:trie}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
 
@@
 
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+\section{Signing Transactions}\label{app:signing}
+
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{\tiny secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{\tiny secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+\pagebreak[1]where:
+\begin{eqnarray}
+\mathtt{\tiny secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
+
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
-\end{eqnarray}
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
-
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
-\end{align}
-where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
-
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
-\begin{equation}
-\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
 \end{equation}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
-
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{memory}(\boldsymbol{\mu}'_i)-C_{memory}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 G_{high} & \text{if} \quad w \in W_{high}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
+G_{ext} & \text{if} \quad w \in W_{ext}
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+$W_{ext}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small BLOCKHASH}\}
+
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
-
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+$\boldsymbol{\mu}'_i$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+
+Note also that $C_{memory}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod \boldsymbol{\mu}_{\mathbf{s}}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0]) (|\boldsymbol{\mu}_{\mathbf{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathbf{s}}[1]|) & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x10 & {\small LT} & 2 & 1 & Less-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x11 & {\small GT} & 2 & 1 & Greater-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x12 & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x14 & {\small EQ} & 2 & 1 & Equality comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
 &&&& This pertains to the input data passed with the message call instruction or transaction. \\
@@
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 \bottomrule
@@
 0x51 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
 \midrule
 0x52 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \equiv \boldsymbol{\mu}_\mathbf{s}[1]$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 32) \div 32 })$ \\
 \midrule
 0x53 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 256) $ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \div 32 })$ \\
-&&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv (\boldsymbol{\mu}_\mathbf{s}[1] \bmod 256) $ \\
+&&&& $\boldsymbol{\mu}'_i \equiv \max(\boldsymbol{\mu}_i, \ceil{ (\boldsymbol{\mu}_\mathbf{s}[0] + 1) \div 32 })$ \\
 \midrule
 0x54 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
-\midrule
-\linkdest{SSTORE}{}0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]]$ \\
+\midrule
+0x55 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
+&&&& $\boldsymbol{\sigma}'[I_a]_\mathbf{s}[ \boldsymbol{\mu}_\mathbf{s}[0] ] \equiv \boldsymbol{\mu}_\mathbf{s}[1] $ \\
+&&&& $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
 G_{sreset} & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_a]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\
 0 & \text{otherwise}
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
-\multicolumn{5}{l}{and to update the memory consumption counter:}\\
-\multicolumn{5}{l}{$\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$}\\
+\multicolumn{5}{l}{$A'_\mathbf{l} \equiv A_\mathbf{l} \cdot (I_a, \mathbf{t}, \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$}\\
 \multicolumn{5}{l}{The entry's topic series, $\mathbf{t}$, differs accordingly:}\vspace{5pt} \\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
-&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
-&&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
-\midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overridden altered.\\
+\midrule
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+0 & \text{otherwise}
+\end{cases}$ \\\bottomrule
 \end{tabular*}
 
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
-\bottomrule
-\end{tabular*}
-
-\section{Genesis Block}\label{app:genesis}\hypertarget{GenesisBlock}{}
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
+\section{Genesis Block}\label{app:genesis}
 
@@
 \end{equation}
-Where a single round modifies each subset of the cache as follows:
-\begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
+\pagebreak[1]where a single round modifies each subset of the cache as follows:
+\begin{equation}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)
 \end{equation}
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{X[l]X[c]X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
-
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2018-02-17 Paper.tex

@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quadp (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}

### 2018-02-17 Paper.tex
Add a missing \

--- 2018-02-16 Paper.tex
@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
 \begin{equation}
@@
 
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
+We assert that the status code $R_{\mathrm{z}}$ is an integer.
 \begin{equation}
@@
 
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
 \begin{equation}
@@
 \linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, L_{R}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_{R}$, we assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
@@
 \end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
+\pagebreak[1]where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:\pagebreak[1]
 \begin{eqnarray}
@@
 I_{\mathrm{s}} & \equiv & s \\
-\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
+I_{\mathrm{v}} & \equiv & v \\
 I_{\mathbf{b}} & \equiv & \mathbf{i} \\
@@
 
-\subsection{Execution Environment}\linkdest{exec_env}
+\subsection{Execution Environment}
 
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
+\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{execution_input_data_I__d}{} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
 \item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
+\item\linkdest{Wei_value_exec}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item\linkdest{exec_machine_code_I__b}{} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
+\item\linkdest{exec_block_header_I__H}{} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_call_or_create_depth_I__e}{} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
 the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item\linkdest{exec_permission_to_modify_state_I__w}{} $I_{\mathrm{w}}$, the permission to make modifications to the state.
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
 
@@
 
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small \text{\hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}

### 2018-02-17 Paper.tex
Remove $s

@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small \text{\hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_{H}}_{\mathrm{g}}$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \ell(\mathbf{R})_{\mathrm{u}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_H}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H g}{{B_H}_{\mathrm{g}}} = \hyperlink{ell}{\ell}\hyperlink{R__u}{(\mathbf{R})_{\mathrm{u}}}
 \end{equation}

### 2018-02-17 Paper.tex

--- 2018-02-17 Paper.tex
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_H}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H g}{{B_H}_{\mathrm{g}}} = \hyperlink{ell}{\ell}\hyperlink{R__u}{(\mathbf{R})_{\mathrm{u}}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}

### 2018-02-17 Paper.tex
Fix a link, add links H__n

@@
 
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \linkdest{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \hyperlink{block}.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
+This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
 
@@
 When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}.
+It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
 
@@
 
-\subsection{The Block} \linkdest{block}{}
+\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
 
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
 \end{description}
@@
 \begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
+R_{\mathrm{z}} \in \mathbb{N}
 \end{equation}
@@
 \begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
 \end{equation}
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 
-The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
+The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
 \begin{equation}
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.

### 2018-02-17 Paper.tex

--- 2018-02-17 Paper.tex
@@
 % For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
+
+%Put after \PassOptionsToPackage PR
+% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
+Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
+\makeatletter
+\g@addto@macro{\UrlBreaks}{\UrlOrds}
+\makeatother
+
+%The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
+If you force a URL to break over a page, the error still occurs.
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
+% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
+% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter:
+%\sloppy
+%\setlength{\emergencystretch}{2pt}
+% https://tex.stackexchange.com/a/59131/143781
+% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
+
+% You can use tables instead of this package, but it might be worthwhile considering.
+%\usepackage{glossaries}
 

### 2018-02-17 Paper.tex
Pull changes from https://github.com/ethereum/yellowpaper/pull/151/files#diff-9f702e1491c55da9d76a68d651278764R919

--- 2018-02-17 Paper.tex
@@
 The exceptional halting function $Z$ is defined as:
-\begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
-\begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+\begin{gather} 
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\ 
+\mathbf{\delta}_w = \varnothing \quad \vee   \nonumber \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \nonumber \\
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \nonumber
+\\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
-\end{array}
-\end{equation}
+\end{gather}
 where

### 2018-02-18 Paper.tex

--- 2018-02-17 Paper.tex
@@
 % For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-%Put after \PassOptionsToPackage PR
-% info on the above command: <https://tex.stackexchange.com/posts/10419/revisions>.
-Note that the three preceding lines, as well as \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} did not fix the error where if a URL breaks over a page, it shows a snippet in the header as shown here: <https://i.stack.imgur.com/tYUe2.png>.
-\makeatletter
-\g@addto@macro{\UrlBreaks}{\UrlOrds}
-\makeatother
-
-%The following command did fix this error, however only because it happened to compact the text, preventing a URL from breaking over a page.
-If you force a URL to break over a page, the error still occurs.
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
-% More info: <https://tex.stackexchange.com/questions/392832/how-to-ensure-that-a-url-break-over-a-page-doesnt-add-a-tiny-url-box-in-the-hea>.
-% These don't appear to make a difference with the preceding use of the \hypens package and \expandafter:
-%\sloppy
-%\setlength{\emergencystretch}{2pt}
-% https://tex.stackexchange.com/a/59131/143781
-% <https://tex.stackexchange.com/questions/241343/what-is-the-meaning-of-fussy-sloppy-emergencystretch-tolerance-hbadness>
-
-% You can use tables instead of this package, but it might be worthwhile considering.
-%\usepackage{glossaries}
 

### 2018-02-18 Paper.tex

--- 2018-02-17 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-\usepackage{tabu} %requires array.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
-
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-
-% Default rendering options
-\definecolor{pagecolor}{rgb}{1,0.98,0.9}
-\def\YellowPaperVersionNumber{unknown revision}
-\IfFileExists{Options.tex}{\input{Options.tex}}
+
+\input{Version.tex}
 
@@
 
+\definecolor{lightyellow}{rgb}{1,0.98,0.9}
+\definecolor{lightpink}{rgb}{1,0.94,0.95}
+
 \DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Founder, Ethereum \& Ethcore\\
+    gavin@ethcore.io
 }
@@
 
-\pagecolor{pagecolor}
+\pagecolor{lightyellow}
+%\pagecolor{lightpink}
 
 \begin{abstract}
-The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, with Bitcoin being one of the most notable ones.
+The blockchain paradigm when coupled with cryptographically-secured transactions has demonstrated its utility through a number of projects, not least Bitcoin.
 Each such project can be seen as a simple application on a decentralised, but singleton, compute resource.
@@
 \begin{equation}
-\linkdest{Upsilon_state_transition}\linkdest{Upsilon}\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_{t}, T)
+\boldsymbol{\sigma}_{t+1} \equiv \Upsilon(\boldsymbol{\sigma}_t, T)
 \end{equation}
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \linkdest{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 $C$, the general cost function.
-These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+These may be subscripted to denote specialised variants, \eg $C_\text{\tiny SSTORE}$, the cost function for the {\tiny SSTORE} operation.
+For specialised and possibly externally defined functions, I may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest) is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
 Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \hyperlink{block}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
+This is formally defined in section \ref{ch:block}.
+
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}.
-
-We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
+
+When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
+It is formally defined in section \ref{ch:block}.
+
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
+\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
 
-Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
-\begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
+\begin{equation}
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\begin{equation}
+L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 
-\subsection{The Block} \linkdest{block}{}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+\subsection{The Block} \label{ch:block}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+The function $L_B$ and $L_H$ are the preparation functions for a block and block header respectively.
+Much like the transaction receipt preparation function $L_R$, we assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_H(H) & \equiv & (\begin{array}[t]{l}H_p, H_o, H_c, H_r, H_t, H_e, H_b, H_d,\\ H_i, H_l, H_g, H_s, H_x, H_m, H_n \; )\end{array} \\
+\quad L_B(B) & \equiv & \big( L_H(B_H), L_T^*(B_\mathbf{T}), L_H^*(B_\mathbf{U}) \big)
+\end{eqnarray}
+
+With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 125000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 
-The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
+The nonce, $H_n$, must satisfy the relations:
+\begin{equation}
+n \leqslant \frac{2^{256}}{H_d} \quad \wedge \quad m = H_m
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})$.
+
+Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to proof that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
+Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_d$.
 
@@
 \item the transaction signature is valid;
-\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
+\item the transaction nonce is valid (equivalent to the sender account's current nonce);
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
 \item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing,(), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no self-destructs, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
 \end{align}
 
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
 
+%todo Explain g_d reason?
+
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the self-destruct set:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
+
+\section{Contract Creation} \label{ch:create}
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+\begin{eqnarray}
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & () \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_e & \equiv & e
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{itemize}
-\item\linkdest{address_of_account_which_owns_code_I__a_def_words}{}\linkdest{I__a} $I_{\mathrm{a}}$, the address of the account which owns the code that is executing.
-\item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
-\item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
-\item\linkdest{the_address_of_the_account_which_caused_the_code_to_be_executing_I__s}{} $I_{\mathrm{s}}$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
-\item\linkdest{Wei_value_exec_I__v}{}\linkdest{I__v}{} $I_{\mathrm{v}}$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
-\item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
-\item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
-\item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
+\item $I_a$, the address of the account which owns the code that is executing.
+\item $I_o$, the sender address of the transaction that originated this execution.
+\item $I_p$, the price of gas in the transaction that originated this execution.
+\item $I_\mathbf{d}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
+\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
+\item $I_v$, the value, in Wei, passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
+\item $I_\mathbf{b}$, the byte array that is the machine code to be executed.
+\item $I_H$, the block header of the present block.
+\item $I_e$, the depth of the present message-call or contract-creation (i.e. the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \end{itemize}
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
-
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_g$ from the resultant machine state $\boldsymbol{\mu}'$.
+
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
@@
 
-\subsubsection{Exceptional Halting}\hypertarget{Exceptional_Halting_function_Z}{}\linkdest{zhalt}
+\subsubsection{Exceptional Halting}
 
 The exceptional halting function $Z$ is defined as:
-\begin{gather} 
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\ 
-\mathbf{\delta}_w = \varnothing \quad \vee   \nonumber \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \nonumber \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \nonumber
-\\
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
-\end{gather}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+\begin{equation}
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+\begin{array}[t]{l}
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
@@
 
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+\end{eqnarray}
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
-\end{equation}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathbf{l} =
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
+\end{equation}
+
+Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
 
@@
 
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
+\section{Modified Merkle Patricia Tree}\label{app:trie}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
 
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
-
-We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
-We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
-
+\begin{equation}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+
+For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
 \begin{eqnarray}
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
+g_r &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 
-We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively.
+The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
-
 \begin{eqnarray}
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_\mathbf{d})\\
 \end{eqnarray}
@@
 
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
-\end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
 \end{eqnarray}
@@
 
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
-It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARESTORE}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise}
 \end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
-\end{eqnarray}
-
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
-\begin{eqnarray}
-\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
-\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
-\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
+h(T) & \equiv & \mathtt{\small KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{\small ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
 $G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
-&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
-&&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}_g' \equiv g' + \boldsymbol{\mu}_g - L(\boldsymbol{\mu}_g)$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(g^* - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad g^* \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
+&&&& $g^* \equiv \boldsymbol{\mu}_{g} - (C_{mem}(\boldsymbol{\mu}'_i)-C_{mem}(\boldsymbol{\mu}_i))$ \\
 &&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols} \label{app:symbols}
-\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
 \end{document}

### 2018-02-18 Paper.tex
Create a branch to upload a file
fix-links-block-nonce-account-nonce-Gtxcreate-etc

--- 2018-02-18 Paper.tex
@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \hyperlink{block}.
+This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
 
@@
 \item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}

### 2018-02-18 Paper.tex
Fix more links: \hyperlink{block}{\ref{subsec:The_Block}}
\linkdset{block_nonce}

@@
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
 This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdset{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
 \end{description}

### 2018-02-19 Paper.tex

@@
 
-\item[Mist] A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, and is developed by \cite{Mist}.
+\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.
 

### 2018-02-19 Paper.tex
quad p

--- 2018-02-19 Paper.tex
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_{\mathrm{block}}$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
 \begin{eqnarray}

### 2018-02-19 Paper.tex
Remove \text within \text

--- 2018-02-19 Paper.tex
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small {\hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\boldsymbol{\sigma}))
 \end{equation}
@@
 
-We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+We assume the existence of functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
 \begin{eqnarray}
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\linkdest{ECDSASIGN}{}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
 \mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}

### 2018-02-19 Paper.tex

--- 2018-02-19 Paper.tex
@@
 0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
 \end{align}

### 2018-02-19 Paper.tex

@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miners to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
 $G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
@@
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
 \end{cases}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{blockhash}{}0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\

### 2018-02-19 Paper.tex
Replace all \bg\b with mathrm{g}...
...using Libre Open Office with match case and regular expressions ticked as options.

@@
 \item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
-This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{g}}$.
+This is paid up-front, before any computation is done and may not be increased later; formally $T_{\mathrm{\mathrm{g}}}$.
 \item[to]\linkdest{tx_to_address_T__t}{} The 160-bit address of the message call's recipient or, for a contract creation transaction, $\varnothing$, used here to denote the only member of $\mathbb{B}_0$ ; formally $T_{\mathrm{t}}$.
@@
 L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{\mathrm{g}}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{\mathrm{g}}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
 \end{cases}
@@
 T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{\mathrm{g}}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
 T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
@@
 \item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
+\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{\mathrm{g}}}$.
 \item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
@@
 \begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{\mathrm{g}}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
 \quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
@@
 \hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{\mathrm{g}}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
 \hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
@@
 & & H_{\mathrm{d}} = D(H) \quad \wedge \\
-& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+& & H_{\mathrm{\mathrm{g}}} \le H_{\mathrm{l}}  \quad \wedge \\
 & & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
@@
 \item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
+\item the gas limit is no smaller than the intrinsic gas, $\mathrm{\mathrm{g}}_0$, used by the transaction; and
 \item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
@@
 Thus $\boldsymbol{\sigma}'$ is the post-transactional state.
-We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{g}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
+We also define \hyperlink{tx_total_gas_used_Upsilon_pow_g}{$\Upsilon^{\mathrm{\mathrm{g}}}$} to evaluate to the amount of gas used in the execution of a transaction, \hyperlink{tx_logs_Upsilon_pow_l}{$\Upsilon^{\mathbf{l}}$} to evaluate to the transaction's accrued log items and \hyperlink{tx_status_Upsilon_pow_z}{$\Upsilon^{\mathrm{z}}$} to evaluate to the status code resulting from the transaction.
 These will be formally defined later.
@@
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
+\mathrm{\mathrm{g}}_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
 {} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
@@
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{\mathrm{g}}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
 \end{equation}
@@
 \linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
+\mathrm{\mathrm{g}}_0 & \leqslant & T_{\mathrm{\mathrm{g}}} \quad \wedge \\
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{\mathrm{g}}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{\mathrm{g}}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{\mathrm{g}}}T_{\mathrm{p}}$.
+The gas available for the proceeding computation, $\mathrm{g}$, is defined as $T_{\mathrm{\mathrm{g}}} - \mathrm{\mathrm{g}}_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{\mathrm{g}}} T_{\mathrm{p}} \\
 \boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
-\begin{equation}
-g \equiv T_{\mathrm{g}} - g_0
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $\mathrm{g}'$, substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, \mathrm{g}', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\end{cases}
+\end{equation}
+
+where $\mathrm{g}$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+\begin{equation}
+\mathrm{g} \equiv T_{\mathrm{\mathrm{g}}} - g_0
 \end{equation}
@@
 
-After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
-\begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $\mathrm{g}^*$ from the remaining gas, $\mathrm{g}'$, plus some allowance from the refund counter, to the sender at the original rate.
+\begin{equation}
+\mathrm{g}^* \equiv \mathrm{g}' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{\mathrm{g}}} - \mathrm{g}'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $\mathrm{g}'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{\mathrm{g}}} - \mathrm{g}'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + \mathrm{g}^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{\mathrm{g}}} - \mathrm{g}^*) T_{\mathrm{p}} \\
 m & \equiv & {B_{H}}_{\mathrm{c}}
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{\mathrm{g}}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{\mathrm{g}}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{\mathrm{g}}} - \mathrm{g}' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($\mathrm{g}$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', \mathrm{g}', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', \mathrm{g}', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, \mathrm{g}, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $\mathrm{g}^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, \mathrm{g}^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, \mathrm{g}, I, \{s, a\}) \\
 \end{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
+If there is not enough gas remaining to pay this, \ie $\mathrm{g}^{**} < c$, then we also declare an out-of-gas exception.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', \mathrm{g}', A, z)$ where:
 
 \begin{align}
-\quad g' &\equiv \begin{cases}
+\quad \mathrm{g}' &\equiv \begin{cases}
 0 & \text{if} \quad F \\
-g^{**} - c & \text{otherwise} \\
+\mathrm{g}^{**} - c & \text{otherwise} \\
 \end{cases} \\
@@
 \quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor \mathrm{g}^{**} < c \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  \mathrm{g}^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($\mathrm{g}$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', \mathrm{g}', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, \mathrm{g}, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
+\mathrm{g}' & \equiv & \begin{cases}
 0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
+\mathrm{g}^{**} & \text{otherwise}
 \end{cases} \\ \nonumber
@@
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+\qquad (\boldsymbol{\sigma}^{**}, \mathrm{g}^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_{\mathrm{a}} & \equiv & r \\
@@
 
-In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
+In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $\mathrm{g}$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
 
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $\mathrm{g}'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
 For the present context, we will define it as:
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+(\boldsymbol{\sigma}', \mathrm{g}', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, \mathrm{g}, I)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, \mathrm{g}, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{\mathrm{g}}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+\boldsymbol{\mu}_{\mathrm{\mathrm{g}}} & \equiv & \mathrm{g} \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
+\boldsymbol{\mu}'_{\mathrm{\mathrm{g}}} & \equiv & \boldsymbol{\mu}_{\mathrm{\mathrm{g}}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\end{eqnarray}
+
+Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{\mathrm{g}}}$ from the resultant machine state $\boldsymbol{\mu}'$.
 
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(\mathrm{g}, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_{\mathrm{\mathrm{g}}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
 \quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $\mathrm{g}[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{\mathrm{g}}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_{H}}_{\mathrm{\mathrm{g}}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 \begin{array}[b]{l}
-\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{\mathrm{g}}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 \begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \mathrm{g}, I, T) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad \mathrm{g} < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, \mathrm{g} - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
+\mathrm{\mathrm{g}}_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathrm{\mathrm{g}}_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
@@
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,\mathrm{g},I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
 F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,\mathrm{g},I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 500\\
@@
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,\mathrm{g},I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 40000\\
@@
 L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{\mathrm{g}}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{\mathrm{g}}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{\mathrm{g}}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{\mathrm{g}}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{\mathrm{g}}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{\mathrm{g}}} \equiv \boldsymbol{\mu}_{\mathrm{\mathrm{g}}} + \mathrm{g}'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+\min\{ L(\boldsymbol{\mu}_{\mathrm{\mathrm{g}}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{\mathrm{g}}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2018-02-20 Paper.tex

@@
 \begin{equation}
-\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
+\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = \mathrm{n} \}
 \end{equation}
@@
 \begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad \mathrm{m} = H_{\mathrm{m}}
+\end{equation}
+with $(n, \mathrm{m}) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
+V(H) & \equiv &  \mathrm{n} \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge \mathrm{m} = H_{\mathrm{m}} \quad \wedge \\
 & & H_{\mathrm{d}} = D(H) \quad \wedge \\
@@
 \end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+where $(n, \mathrm{m}) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}_{P}, \mathrm{g}', A, z) \equiv \begin{cases}
+(\boldsymbol{\sigma}_{P}, \mathrm{g}', A, \mathrm{z}) \equiv \begin{cases}
 \Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
@@
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & \mathrm{z}
 \end{eqnarray}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', \mathrm{g}', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, \mathrm{g}, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z}, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, \mathrm{g}, p, v, \mathbf{i}, \mathrm{e}, w)
 \end{equation}
@@
 I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{e}} & \equiv & \mathrm{e} \\
 I_{\mathrm{w}} & \equiv & w
@@
 
-If there is not enough gas remaining to pay this, \ie $\mathrm{g}^{**} < c$, then we also declare an out-of-gas exception.
+If there is not enough gas remaining to pay this, \ie $\mathrm{g}^{**} < \mathrm{c}$, then we also declare an out-of-gas exception.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', \mathrm{g}', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z})$ where:
 
@@
 0 & \text{if} \quad F \\
-\mathrm{g}^{**} - c & \text{otherwise} \\
+\mathrm{g}^{**} - \mathrm{c} & \text{otherwise} \\
 \end{cases} \\
@@
 \end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor \mathrm{g}^{**} < c \\
+\quad \mathrm{z} &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor \mathrm{g}^{**} < \mathrm{c} \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  \mathrm{g}^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  \mathrm{g}^{**} < \mathrm{c} \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', \mathrm{g}', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, \mathrm{g}, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z}, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, \mathrm{c}, \mathrm{g}, p, v, \tilde{v}, \mathbf{d}, \mathrm{e}, w)
 \end{equation}
@@
 I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
+I_{\mathrm{e}} & \equiv & \mathrm{e} \\
 I_{\mathrm{w}} & \equiv & w \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
+(a, b, \mathrm{c}, d) \cdot \mathrm{e} & \equiv & (a, b, \mathrm{c}, d, \mathrm{e}) \\
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
@@
 \begin{equation}
-k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
+k(U, H, \mathrm{n}) \equiv \begin{cases} false & \text{if} \quad \mathrm{n} = 0 \\
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_{H}, \mathrm{n} - 1) & \text{otherwise}
 \end{cases}
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B'_{\mathrm{n}} & = & \mathrm{n}: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & \mathrm{m} \quad \text{with } (x, \mathrm{m}) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, \mathrm{n}, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
@@
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad \mathrm{n} < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad \mathrm{n} < 0 \\
 \begin{array}[b]{l}
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+m = H_{\mathrm{m}} \quad \wedge \quad \mathrm{n} \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, \mathrm{n}) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - \mathrm{n}} \\
+(a) \cdot (b, \mathrm{c}) \cdot (d, \mathrm{e}) & = & (a, b, \mathrm{c}, d, \mathrm{e})
 \end{eqnarray}
@@
 () & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
+c(\mathfrak{I}, i) & \text{if} \quad \lVert \mathrm{c}(\mathfrak{I}, i)\rVert < 32 \\
 \texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), \mathrm{n}(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
 \texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
+u(j) & \equiv & \mathrm{n}(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
 v & = & \begin{cases}
@@
 \begin{equation}
-L(n) \equiv n - \lfloor n / 64 \rfloor
+L(n) \equiv \mathrm{n} - \lfloor \mathrm{n} / 64 \rfloor
 \end{equation}
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, \mathrm{n}, a) \equiv \begin{cases} 0 & \text{if} \quad \mathrm{n} > H_{\mathrm{i}} \vee a = 256 \vee \mathrm{h} = 0 \\ \mathrm{h} & \text{if} \quad \mathrm{n} = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, \mathrm{n}, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathrm{c}(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad \mathrm{c}(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
-&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
+&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., \mathrm{c}(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + \mathrm{n} - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
@@
 &&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+G_{newaccount} & \text{if} \quad \mathrm{n} \\
 0 & \text{otherwise}
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
-\end{equation}
-\begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < \mathrm{n}
+\end{equation}
+\begin{equation}
+ \mathrm{n} = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, \mathrm{n} - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
 \begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + \mathrm{n}) \mod \mathrm{n}] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod \mathrm{n}]) \\
   \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i

### 2018-02-20 Paper.tex

--- 2018-02-19 Paper.tex
@@
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $\mathrm{n}$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
 Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $\mathrm{g}'$, substate $A$ and status code $z$:
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $\mathrm{g}'$, substate $A$ and status code $\mathrm{z}$:
 \begin{equation}
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($\mathrm{g}$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($\mathrm{g}$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($\mathrm{e}$).
 
@@
 
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
+If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $\mathrm{c}$, proportional to the size of the created contract's code:
 \begin{equation}
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($\mathrm{g}$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($\mathrm{c}$, usually the same as recipient), available gas ($\mathrm{g}$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($\mathrm{e}$).
 
@@
 
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $\mathrm{n}$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+We also define the $\mathrm{n}$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
 \begin{equation}
@@
 
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
+The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $\mathrm{n}$.
 It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
@@
 
-We also assume a function $n$, the trie's node cap function.
+We also assume a function $\mathrm{n}$, the trie's node cap function.
 When composing a node, we use RLP to encode the structure.
 As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
-Thus we define in terms of $c$, the node composition function:
+Thus we define in terms of $\mathrm{c}$, the node composition function:
 \begin{equation}
@@
 
-A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry. We may formally define this structure with the structural composition function $c$:
+A branch is then only used when necessary; no branch nodes may exist that contain only a single non-zero entry.
+We may formally define this structure with the structural composition function $\mathrm{c}$:
 \begin{equation}
@@
 
-A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
+A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $\mathrm{c}$.
 This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.
@@
 \end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
+for a natural number $\mathrm{n}$ and a point $P$ in $C_1$.
 
@@
 $G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
+For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $\mathrm{n}$ satisfying $n\cdot P_1=P$.
 $\log_{P_1}(P)$ is at most $q-1$.
@@
 $G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
+For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $\mathrm{n}$ satisfying $n\cdot P_2=P$.
 With this definition, $\log_{P_2}(P)$ is at most $q-1$.
@@
 &&&& The bytes are read in line from the program code's bytes array. \\
-&&&& The function $c$ ensures the bytes default to zero if they extend past the limits.\\
+&&&& The function $\mathrm{c}$ ensures the bytes default to zero if they extend past the limits.\\
 &&&& The byte is right-aligned (takes the lowest significant place in big endian). \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
-&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
+&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $\mathrm{c}$ as defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\

### 2018-02-20 Paper.tex

--- 2018-02-20 Paper.tex
@@
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
+\mathrm{g}_{\mathrm{r}} &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathrm{g}_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathrm{g}_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+\mathrm{g}_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
 \mathbf{o} &=& I_{\mathbf{d}}

### 2018-02-20 Paper.tex
dummy

--- 2018-02-20 Paper.tex
@@
-\documentclass[9pt,oneside]{amsart}
+%\documentclass[9pt,oneside]{amsart}
 %\usepackage{tweaklist}

### 2018-02-20 Paper.tex
Replace all \$([ceg-hmnz])\$ with \$\mathrm{$1}\$ using Libre Open Office with regular expressions ticked as an option.

@@
-%\documentclass[9pt,oneside]{amsart}
+\documentclass[9pt,oneside]{amsart}
 %\usepackage{tweaklist}
@@
 The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
-The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
+The gas available for the proceeding computation, $\mathrm{g}$, is defined as $T_{\mathrm{g}} - g_0$.
 The computation, whether contract creation or a message call, results in an eventual state (which may legally be equivalent to the current state), the change to which is deterministic and never invalid: there can be no invalid transactions from this point.
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $\mathrm{z}$:
 \begin{equation}
@@
 
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+where $\mathrm{g}$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
 \begin{equation}
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($\mathrm{g}$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($\mathrm{e}$).
 
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($\mathrm{c}$, usually the same as recipient), available gas ($\mathrm{g}$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($\mathrm{e}$).
 

### 2018-02-20 Paper.tex
dummy
Replace all \s\b([ceg-hmnz])\b with " \mathrm{$1}" using Libre Open Office with match case and regular expressions ticked as options.

--- 2018-02-20 Paper.tex
@@
 \begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+(\boldsymbol{\sigma}_{P}, \mathrm{g}', A, \mathrm{z}) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv \mathrm{g}' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - \mathrm{g}'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - \mathrm{g}'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + \mathrm{g}^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - \mathrm{g}^*) T_{\mathrm{p}} \\
 m & \equiv & {B_{H}}_{\mathrm{c}}
@@
 \begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - \mathrm{g}' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & \mathrm{z}
 \end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', \mathrm{g}', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z}, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, \mathrm{g}, p, v, \mathbf{i}, \mathrm{e}, w)
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
+(\boldsymbol{\sigma}^{**}, \mathrm{g}^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, \mathrm{g}, I, \{s, a\}) \\
 \end{equation}
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
+If there is not enough gas remaining to pay this, \ie $g^{**} < \mathrm{c}$, then we also declare an out-of-gas exception.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z})$ where:
 
 \begin{align}
-\quad g' &\equiv \begin{cases}
+\quad \mathrm{g}' &\equiv \begin{cases}
 0 & \text{if} \quad F \\
-g^{**} - c & \text{otherwise} \\
+g^{**} - \mathrm{c} & \text{otherwise} \\
 \end{cases} \\
@@
 \end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+\quad \mathrm{z} &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor \mathrm{g}^{**} < \mathrm{c} \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  \mathrm{g}^{**} < \mathrm{c} \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z}, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, \mathrm{c}, \mathrm{g}, p, v, \tilde{v}, \mathbf{d}, \mathrm{e}, w)
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
+(\boldsymbol{\sigma}', \mathrm{g}', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, \mathrm{g}, I)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, \mathrm{g}, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & \mathrm{g} \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + \mathrm{n} - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + \mathrm{g}'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\

### 2018-02-20 Paper.tex
Add a missing \

--- 2018-02-20 Paper.tex
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the \hypertarget{v}{}aforementioned range).
+It is assumed that $v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
@@
 0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+v &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
 \end{align}

### 2018-02-20 Paper.tex
Update equation presentation

@@
 \begin{equation} 
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee   \nonumber \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \nonumber \\
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \nonumber
+\\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
 \end{array}
-\end{equation}
+\end{gather}
 where
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $J_{epoch}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.

### 2018-02-20 Paper.tex

--- 2018-02-20 Paper.tex
@@
 \begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv \boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
 \begin{array}[t]{l}
-\mathbf{\delta}_w = \varnothing \quad \vee   \nonumber \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \nonumber \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \nonumber \\
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \nonumber
-\\
+ \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
 \end{array}
-\end{gather}
+\end{equation}
 where

### 2018-02-20 Paper.tex

@@
 
-The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
-This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
 

### 2018-02-20 Paper.tex

@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\

### 2018-02-20 Paper.tex
Revert "Spurious Dragon: EIP155 replay protection"

--- 2018-02-20 Paper.tex
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_t, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
 Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
+This is formally defined in section \ref{ch:block}.
+
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-When considering variants of existing values, we follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
-On very particular occasions, in order to maximise readability and only if unambiguous in meaning, we may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
-
-When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
-
-We define a number of useful functions throughout. \linkdest{ell}One of the more common is $\ell$, which evaluates to the last item in the given sequence:
+When considering variants of existing values, I follow the rule that within a given scope for definition, if we assume that the unmodified `input' value be denoted by the placeholder $\Box$ then the modified and utilisable value is denoted as $\Box'$, and intermediate values would be $\Box^*$,  $\Box^{**}$ \&c.
+On very particular occasions, in order to maximise readability and only if unambiguous in meaning, I may use alpha-numeric subscripts to denote intermediate values, especially those of particular note.
+
+When considering the use of existing functions, given a function $f$, the function $f^*$ denotes a similar, element-wise version of the function mapping instead between sequences.
+It is formally defined in section \ref{ch:block}.
+
+I define a number of useful functions throughout. One of the more common is $\ell$, which evaluates to the last item in the given sequence:
 
@@
 \end{equation}
-\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{equation}
-\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_\mathbf{t} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+\subsection{The Block} \label{ch:block}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language#Aunt.2FUncle}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{P}
-\end{equation}
-
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction, is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (\mathtt{\small TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.
+Formally:
 \begin{eqnarray}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-x \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( 1 - \left\lfloor\frac{H_s - {P(H)_H}_s}{10}\right\rfloor, -99 \right)
+\end{equation}
+\begin{equation}
+\expdiffsymb \equiv \left\lfloor 2^{ \left\lfloor H_i \div 100000 \right\rfloor - 2 } \right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 5000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_\mathbf{t}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+$A_\mathbf{t}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
 \begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
+g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
+{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_t = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + G_{transaction}
 \end{align}
 
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
 $G$ is fully defined in Appendix \ref{app:fees}.
 
+%todo Explain g_d reason?
+
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
 \end{eqnarray}
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\forall i \in A_\mathbf{t}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
+
+\section{Contract Creation} \label{ch:create}
 
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_n, \boldsymbol{\sigma}[s]_b - v, \boldsymbol{\sigma}[s]_\mathbf{s}, \boldsymbol{\sigma}[s]_c)
 \end{eqnarray}
@@
 \end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
+where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}, that is:
+\begin{eqnarray}
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & () \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_e & \equiv & e
+\end{eqnarray}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call. $I_H$ has no special treatment and is determined from the blockchain.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big(\boldsymbol{\sigma}^{**} = \varnothing \ \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_n, \boldsymbol{\sigma}_1'[s]_b - v, \boldsymbol{\sigma}_1'[s]_\mathbf{s}, \boldsymbol{\sigma}_1'[s]_c)
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
+\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + v, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c)
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
 g^{**} & \text{otherwise}
 \end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
 \qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
@@
 \Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_s & \equiv & s \\
+I_v & \equiv & \tilde{v} \\
+I_e & \equiv & e \\
 \mathbf{t} & \equiv & \{s, r\} \\
 \\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+\text{Let} \; \mathtt{\tiny KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
-
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_g, A, \mathbf{o}) \\
+(\boldsymbol{\sigma}, \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\boldsymbol{\mu}_g & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger than 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
-\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
+\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
+\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x-\Delta]
 \end{eqnarray}
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
+\item verify (or, if mining, compute a valid) state and nonce.
 \end{enumerate}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
 \begin{eqnarray}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
+\qquad\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\qquad\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+\boldsymbol{\sigma}'[U_c] & = & \begin{cases}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[U_c] = \varnothing\ \wedge\ R = 0 \\
 \mathbf{a}' &\text{otherwise}
 \end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_c]_n, \boldsymbol{\sigma}[U_c]_b + R, \boldsymbol{\sigma}[U_c]_\mathbf{s}, \boldsymbol{\sigma}[U_c]_c) \\
+R & \equiv & (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b
 \end{eqnarray}
@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
+We define the block reward as 5 Ether:
+\begin{equation}
+\text{Let} \quad R_b = 5 \times 10^{18}
 \end{equation}
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious miner to control those values, one should use the {\small BLOCKHASH} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
+\section{Modified Merkle Patricia Tree}\label{app:trie}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32 byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
-
-We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
-We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
-
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+
+For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+Importantly in the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
 \begin{eqnarray}
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
+g_r &=& 3000\\
 |\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
@@
 \mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
-\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
+\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
-It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+It is assumed that the sender has a valid private key $p_r$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+
+We assert the functions $\mathtt{\small ECDSASIGN}$, $\mathtt{\small ECDSARECOVER}$ and $\mathtt{\small ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 
@@
 \begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+0 < r &< \mathtt{\tiny secp256k1n} \\
+0 < s &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+v &\in \{27,28\}
 \end{align}
@@
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\it Homestead transition}.\\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
-
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
+G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
-&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
-&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
-&&&& created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, L(\boldsymbol{\mu}_g), I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{t} \equiv A_\mathbf{t} \cup A^+_\mathbf{t}$ \\
+&&&& $ \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 \midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\midrule
 0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
 \midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
+\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_a]_b = 0\\
+(\boldsymbol{\sigma}[r]_n, \boldsymbol{\sigma}[r]_b + \boldsymbol{\sigma}[I_a]_b, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{if}\ r \neq I_a \\
+(\boldsymbol{\sigma}[r]_n, 0, \boldsymbol{\sigma}[r]_\mathbf{s}, \boldsymbol{\sigma}[r]_c) & \text{otherwise}
 \end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+&&&& where $r = \boldsymbol{\mu}_\mathbf{s}[0] \bmod 2^{160}$\\
+&&&& $\boldsymbol{\sigma}'[I_a]_b = 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
+R_{selfdestruct} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+&&&& $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
 G_{newaccount} & \text{if} \quad n \\
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_a]_b \neq 0$ \\
 \bottomrule
@@
 
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
+\subsection{Deletion of an Account Dispite Out-of-gas}
+
+At block 2675119, in the transaction \texttt{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
 Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
 
-\section{List of mathematical symbols} \label{app:symbols}
-\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
 \end{document}

### 2018-02-20 Paper.tex
Add braces

--- 2018-02-20 Paper.tex
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
+\boldsymbol{\sigma}_{t+1} & \equiv & \Pi(\boldsymbol{\sigma}_{t}, B) \\
 B & \equiv & (..., (T_0, T_1, ...) ) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
+\end{eqnarray}
+
+Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.
 
@@
 If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
-This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
+This is formally defined in section \ref{ch:block}.
 
@@
 When considering the use of existing functions, given a function $f$, the function \hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{$f^*$} denotes a similar, element-wise version of the function mapping instead between sequences.
-It is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
+It is formally defined in section \ref{ch:block}.
 
@@
 
-\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
+\subsection{The Block} \label{ch:block}
 
@@
 \item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
+
 \linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
+{} & + \begin{cases} \hyperlink{Gtxcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+{} & + \hyperlink{Gtransaction}{G_{\mathrm{transaction}}}
 \end{align}
@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account nonce}{account nonce}.
 Thus we define the resultant address for the new account $a$:
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small {\hyperlink{stop}{STOP}}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}} \} \\
 \varnothing & \text{otherwise}
@@
 \item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
+\item verify (or, if mining, compute a valid) state and \hyperlink{block nonce}{block nonce}.
 \end{enumerate}
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\

### 2018-02-20 Paper.tex
(\cite{ASICSmithMJS}) (\cite{Courtois2014})

--- 2018-02-20 Paper.tex
@@
 \begin{equation}
-\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = \mathrm{n} \}
+\mathbb{B}_{\mathrm{n}} = \{ B: B \in \mathbb{B} \wedge \lVert B \rVert = n \}
 \end{equation}
@@
 \begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad \mathrm{m} = H_{\mathrm{m}}
-\end{equation}
-with $(n, \mathrm{m}) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
+\end{equation}
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  \mathrm{n} \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge \mathrm{m} = H_{\mathrm{m}} \quad \wedge \\
+V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
 & & H_{\mathrm{d}} = D(H) \quad \wedge \\
@@
 \end{eqnarray}
-where $(n, \mathrm{m}) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}_{P}, \mathrm{g}', A, \mathrm{z}) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, \mathrm{g}, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g^* \equiv \mathrm{g}' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - \mathrm{g}'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - \mathrm{g}'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 
@@
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + \mathrm{g}^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - \mathrm{g}^*) T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
 m & \equiv & {B_{H}}_{\mathrm{c}}
@@
 \begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - \mathrm{g}' \\
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & \mathrm{z}
+\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
 \end{eqnarray}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', \mathrm{g}', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z}, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, \mathrm{g}, p, v, \mathbf{i}, \mathrm{e}, w)
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, \mathrm{g}^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, \mathrm{g}, I, \{s, a\}) \\
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
 \end{equation}
@@
 I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & \mathrm{e} \\
+I_{\mathrm{e}} & \equiv & e \\
 I_{\mathrm{w}} & \equiv & w
@@
 
-If there is not enough gas remaining to pay this, \ie $g^{**} < \mathrm{c}$, then we also declare an out-of-gas exception.
+If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z})$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
 
 \begin{align}
-\quad \mathrm{g}' &\equiv \begin{cases}
+\quad g' &\equiv \begin{cases}
 0 & \text{if} \quad F \\
-g^{**} - \mathrm{c} & \text{otherwise} \\
+g^{**} - c & \text{otherwise} \\
 \end{cases} \\
@@
 \end{cases} \\
-\quad \mathrm{z} &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor \mathrm{g}^{**} < \mathrm{c} \\
+\quad z &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  \mathrm{g}^{**} < \mathrm{c} \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
 \end{align}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', \mathrm{g}', A, \mathrm{z}, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, \mathrm{c}, \mathrm{g}, p, v, \tilde{v}, \mathbf{d}, \mathrm{e}, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, \mathrm{g}^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, \mathrm{g}, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
+\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
 I_{\mathrm{a}} & \equiv & r \\
@@
 I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & \mathrm{e} \\
+I_{\mathrm{e}} & \equiv & e \\
 I_{\mathrm{w}} & \equiv & w \\
@@
 \begin{equation}
-(\boldsymbol{\sigma}', \mathrm{g}', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, \mathrm{g}, I)
+(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, \mathrm{g}, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
+\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
 (\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & \mathrm{g} \\
+\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
 \boldsymbol{\mu}_{pc} & \equiv & 0 \\
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, \mathrm{c}, d) \cdot \mathrm{e} & \equiv & (a, b, \mathrm{c}, d, \mathrm{e}) \\
+(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
@@
 \begin{equation}
-k(U, H, \mathrm{n}) \equiv \begin{cases} false & \text{if} \quad \mathrm{n} = 0 \\
+k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
 s(U, H) &\\
-\quad \vee \; k(U, P(H)_{H}, \mathrm{n} - 1) & \text{otherwise}
+\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
 \end{cases}
@@
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & \mathrm{n}: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & \mathrm{m} \quad \text{with } (x, \mathrm{m}) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, \mathrm{n}, \mathbf{d}) \\
+B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
@@
 \begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad \mathrm{n} < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
+\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad \mathrm{n} < 0 \\
+\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad \mathrm{n} \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, \mathrm{n}) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}
@@
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - \mathrm{n}} \\
-(a) \cdot (b, \mathrm{c}) \cdot (d, \mathrm{e}) & = & (a, b, \mathrm{c}, d, \mathrm{e})
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
 \end{eqnarray}
@@
 () & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert \mathrm{c}(\mathfrak{I}, i)\rVert < 32 \\
+c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
 \texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
@@
 \texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(j - 1)], false), \mathrm{n}(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
 \texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
-u(j) & \equiv & \mathrm{n}(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
+u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
 v & = & \begin{cases}
@@
 \begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \mathrm{g}, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad \mathrm{g} < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, \mathrm{g} - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 \begin{equation}
-L(n) \equiv \mathrm{n} - \lfloor \mathrm{n} / 64 \rfloor
+L(n) \equiv n - \lfloor n / 64 \rfloor
 \end{equation}
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, \mathrm{n}, a) \equiv \begin{cases} 0 & \text{if} \quad \mathrm{n} > H_{\mathrm{i}} \vee a = 256 \vee \mathrm{h} = 0 \\ \mathrm{h} & \text{if} \quad \mathrm{n} = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, \mathrm{n}, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
 &&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathrm{c}(\boldsymbol{\mu}_{pc} + 1)$ \\
-&&&& $\text{where} \quad \mathrm{c}(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
-&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., \mathrm{c}(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
+&&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
 &&&& The bytes are right-aligned (takes the lowest significant place in big endian). \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + \mathrm{n} - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + \mathrm{g}'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
@@
 &&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', \mathrm{g}', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, \mathrm{g}, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad \mathrm{n} \\
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
@@
 \begin{equation}
- \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < \mathrm{n}
-\end{equation}
-\begin{equation}
- \mathrm{n} = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+ \mathbf{c'}[i] = \mathbf{c}_{i} \quad \forall \quad i < n
+\end{equation}
+\begin{equation}
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, \mathrm{n} - 1) \big)\linkdest{E__cacherounds}{}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
 \begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + \mathrm{n}) \mod \mathrm{n}] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod \mathrm{n}]) \\
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
   \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i

### 2018-02-20 Paper.tex
Fixing bracing

--- 2018-02-20 Paper.tex
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small {\hyperlink{stop}{STOP}}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}} \} \\
+() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
 \varnothing & \text{otherwise}

### 2018-02-21 Paper.tex
Remove an extra }

--- 2018-02-20 Paper.tex
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
+() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
 \varnothing & \text{otherwise}

### 2018-02-21 Paper.tex
Add Appendix with See *Appendix \ref{app:signing}
Other instances of \ref{app:foo} have Appendix prepended.

--- 2018-02-20 Paper.tex
@@
 We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.

### 2018-02-21 Paper.tex

@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
+() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
 \varnothing & \text{otherwise}
@@
 We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+See \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.

### 2018-02-22 Paper.tex
\linkdest{invalidsig}
Update link label h T to h_of_T

--- 2018-02-21 Paper.tex
@@
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\ e & \equiv & \hyperlink{h_of_T}{h(T)}\\
 \linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\

### 2018-02-22 Paper.tex
Update Paper.tex
Replace all _(\{[alpha][:alnum:]+\}) with " \mathrm{$1}" using Libre Open Office with regular expressions ticked as an option.

--- 2018-02-21 Paper.tex
@@
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{low} & \text{if} \quad w \in W_{low}\\
+G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
 G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{high} & \text{if} \quad w \in W_{high}\\
+G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
 G_{extcode} & \text{if} \quad w \in W_{extcode}\\
@@
 \begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 \linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
 \linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 \begin{equation}
- c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
-\end{equation}
-\begin{equation}
- E_{prime}(x, y) = \begin{cases}
+ c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{prime}}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
+E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}

### 2018-02-22 Paper.tex
Pull in the changes from #616 that are from #151

--- 2018-02-22 Paper.tex
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
+\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
@@
 W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+w \in \{\text{\small CREATE}, \, \text{\small SSTORE}, \\
+\text{\small SELFDESTRUCT}\} \quad \vee \\
+(\text{\small LOG0} \le w \wedge w \le \text{\small LOG4}) \quad \vee \\
+(w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0)
 \end{array}
@@
 \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\ e & \equiv & \hyperlink{h_of_T}{h(T)}\\
 \linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
@@
 
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+\hypertarget{h T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
 Two different flavors of signing schemes are available.

### 2018-02-22 Paper.tex
Fix the font of \alpha_w and \delta_w
Since w is a variable for an opcode, it should be typeset in italic.

@@
 \boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+\mathbf{\delta}_{w} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{w} \quad \vee \\
 ( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
+ \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{w} + \mathbf{\alpha}_{w} > 1024 \quad \vee \\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})

### 2018-02-23 Paper.tex
Update author in header to match pdftitle wording
Move e definition

--- 2018-02-22 Paper.tex
@@
 H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} \\
 \varnothing & \text{otherwise}
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e \equiv \hyperlink{h_of_T}{h(T)}$, i.e.
+$e$ is the hash of the transaction.
 It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.

### 2018-02-23 Paper.tex
Re add Gavin contact info
Fix new line

--- 2018-02-22 Paper.tex
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_{w} = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{w} \quad \vee \\
-( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_w = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
+( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
+\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
- \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{w} + \mathbf{\alpha}_{w} > 1024 \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})

### 2018-02-23 Paper.tex
One must reference multiple versions of this document
Use well-known \mathbb{N for natural numbers

@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
  E_{\mathrm{prime}}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{P} \\
+x & \text{if} \quad x / y \in \mathbb{N} \\
 E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}

### 2018-02-24 Paper.tex
Introduce account state variable when referenced
Remove verifiable computation
Reword definition of e.

--- 2018-02-23 Paper.tex
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e \equiv \hyperlink{h_of_T}{h(T)}$, i.e.
-$e$ is the hash of the transaction.
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
 It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.

### 2018-02-25 Paper.tex

@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
-&&&& in the block following it. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H(h)_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H(h)_{\mathrm{i}} \\ P(H(h)_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H(h)$ can be determined from its hash~$h$ unless $h$ is zero. \\
 \midrule

### 2018-02-27 Paper.tex
Add math $

--- 2018-02-25 Paper.tex
@@
 \item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 The EVM is a simple stack-based architecture.
-The word size of the machine (and thus size of stack items) is 256-bit.
+The word size of the machine (and thus size of stack item) is 256-bit.
 This was chosen to facilitate the Keccak-256 hash scheme and elliptic-curve computations.

### 2018-02-27 Paper.tex
Use a different variable in eq44 to avoid ambiguity
Use a space character that LaTeX can recognize

@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
-Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Gavin Wood and others as per https://github.com/ethereum/yellowpaper/commits/master},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 %,pagebackref=true
@@
 
-The account state, $\boldsymbol{\sigma}[a]$, comprises the following four fields:
+The account state comprises the following four fields:
 
@@
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
-\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
+\boldsymbol{\mu}_{pc} & \equiv & 0 \\
 \boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert \\
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert \\
 \text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
@@
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
-\mathbf{\delta}_w = \varnothing \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
-\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\boldsymbol{\mu}_{\mathrm{g}} < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\mathbf{\delta}_{\mathrm{w}} = \varnothing \quad \vee \\
+\lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert < \mathbf{\delta}_{\mathrm{w}} \quad \vee \\
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \notin D(I_{\mathbf{b}}) ) \quad \vee \\
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
+ \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert - \mathbf{\delta}_{\mathrm{w}} + \mathbf{\alpha}_{\mathrm{w}} > 1024 \quad \vee \\
  \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
@@
 W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \, \text{\small SSTORE}, \\
-\text{\small SELFDESTRUCT}\} \quad \vee \\
-(\text{\small LOG0} \le w \wedge w \le \text{\small LOG4}) \quad \vee \\
-(w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0)
+w \in \{\text{\small CREATE}, \text{\small SSTORE}, \text{\small SELFDESTRUCT}\} \quad \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
+\Delta & \equiv & \mathbf{\alpha}_{\mathrm{w}} - \mathbf{\delta}_{\mathrm{w}} \\
 \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
+\quad \forall x \in [\mathbf{\alpha}_{\mathrm{w}}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
 \end{eqnarray}
@@
 \quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
-\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
+\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
 \hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
 \hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
+N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
 \end{cases}
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
 It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 
-\linkdest{invalidsig}We declare that a signature is invalid unless all the following conditions are true:
+We declare that a signature is invalid unless all the following conditions are true:
 \begin{align}
@@
 $G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
 $G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
 $G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
@@
 \linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
@@
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
@@
 G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
+G_{low} & \text{if} \quad w \in W_{low}\\
 G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
+G_{high} & \text{if} \quad w \in W_{high}\\
 G_{extcode} & \text{if} \quad w \in W_{extcode}\\
@@
 \begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
+w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_{\mathbf{b}} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 
-$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
 
@@
 
-$W_{\mathrm{high}}$ = \{{\small JUMPI}\}
+$W_{high}$ = \{{\small JUMPI}\}
 
@@
 &&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}.\\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}.\\
 \midrule
 \linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
 &&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 \linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
+\linkdest{J__hashbytes}{}$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+\linkdest{J__parents}{}$J_{parents}$ & 256 & Number of parents of each dataset element. \\
 \linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
+\linkdest{J__accesses}{}$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
 \linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
 \begin{equation}
- d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
-\end{equation}
-\begin{equation}
- E_{\mathrm{prime}}(x, y) = \begin{cases}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
 x & \text{if} \quad x / y \in \mathbb{P} \\
-E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
+E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \subsubsection{Full dataset calculation} \label{dataset}
-Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{\mathrm{hashbytes}}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
-\end{equation}
-\begin{equation}
-  E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+ E_{datasetitem}(\mathbf{c}, i) = E_{parents}(\mathbf{c}, i, -1, \varnothing)
+\end{equation}
+\begin{equation}
+  E_{parents}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
+E_{parents}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{parents} -2 \\
 E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
@@
 \texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{hashbytes} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-In order to add random dataset nodes to the mix, the $E_{\mathrm{accesses}}$ function is used:
-\begin{equation}
- E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+In order to add random dataset nodes to the mix, the $E_{accesses}$ function is used:
+\begin{equation}
+ E_{accesses}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{accesses} -2 \\
+E_{accesses}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{hashbytes}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}

### 2018-03-01 Paper.tex

@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \, \text{\small SSTORE}, \\
-\text{\small SELFDESTRUCT}\} \quad \vee \\
-(\text{\small LOG0} \le w \wedge w \le \text{\small LOG4}) \quad \vee \\
-(w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0)
+W(w, \boldsymbol{\mu}) \equiv \\ 
+\begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
+\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
+w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
-\varnothing & \text{otherwise}
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
+() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
+\varnothing \quad\quad\ \text{otherwise}&
 \end{cases}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
+&&contract. \\
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
+&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
+&\quad\text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
-&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
-&&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H(h)_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H(h)_{\mathrm{i}} \\ P(H(h)_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H(h)$ can be determined from its hash~$h$ unless $h$ is zero. \\
+0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than\\
+&&&& the current block number or more than 256 blocks behind the current block.\\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
+&&&& $h$ is zero.\\
 \midrule

### 2018-03-01 Paper.tex

--- 2018-02-27 Paper.tex
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 \begin{equation}
-m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
+m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{h_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}

### 2018-03-02 Paper.tex
hash -> value

--- 2018-03-01 Paper.tex
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}

### 2018-03-02 Paper.tex

@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
-Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-\usepackage{tabu} %requires array.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
-
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-
-% Default rendering options
-\definecolor{pagecolor}{rgb}{1,0.98,0.9}
-\def\YellowPaperVersionNumber{unknown revision}
-\IfFileExists{Options.tex}{\input{Options.tex}}
 
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Final Draft - UNDER REVIEW}}}
 \author{
     Dr. Gavin Wood\\
-    Founder, Ethereum \& Parity\\
-    gavin@parity.io
+    Co-Founder \& Lead, Ethereum Project\\
+    gavin@ethereum.org
 }
@@
 
-\cite{dwork92pricingvia} provided the first work into the usage of a cryptographic proof of computational expenditure (``proof-of-work'') as a means of transmitting a value signal over the Internet.
-The value-signal was utilised here as a spam deterrence mechanism rather than any kind of currency, but critically demonstrated the potential for a basic data channel to carry a \textit{strong economic signal}, allowing a receiver to make a physical assertion without having to rely upon \textit{trust}.
-\cite{back2002hashcash} later produced a system in a similar vein.
-
-The first example of utilising the proof-of-work as a strong economic signal to secure a currency was by \cite{vishnumurthy03karma:a}.
-In this instance, the token was used to keep peer-to-peer file trading in check, providing ``consumers'' with the ability to make micro-payments to ``suppliers'' for their services.
-The security model afforded by the proof-of-work was augmented with digital signatures and a ledger in order to ensure that the historical record couldn't be corrupted and that malicious actors could not spoof payment or unjustly complain about service delivery.
-Five years later, \cite{nakamoto2008bitcoin} introduced another such proof-of-work-secured value token, somewhat wider in scope.
-The fruits of this project, Bitcoin, became the first widely adopted global decentralised transaction ledger.
+Hashcash, introduced by \cite{back2002hashcash} (in a five-year retrospective), provided the first work into the usage of a cryptographic proof of computational expenditure as a means of transmitting a value signal over the Internet.
+Though not widely adopted, the work was later utilised and expanded upon by \cite{nakamoto2008bitcoin} in order to devise a cryptographically secure mechanism for coming to a decentralised social consensus over the order and contents of a series of cryptographically signed financial transactions.
+The fruits of this project, Bitcoin, provided a first glimpse into a decentralised transaction ledger.
 
@@
 
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
-In order to follow back the history of a path, one must reference multiple versions of this document.
-
 \section{Conventions}\label{ch:conventions}
 
-We use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
+I use a number of typographical conventions for the formal notation, some of which are quite particular to the present work:
 
@@
 Tuples are typically denoted with an upper-case letter, \eg $T$, is used to denote an Ethereum transaction.
-This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg \hyperlink{transaction_nonce}{$T_{\mathrm{n}}$}, denotes the nonce of said transaction.
+This symbol may, if accordingly defined, be subscripted to refer to an individual component, \eg $T_n$, denotes the nonce of said transaction.
 The form of the subscript is used to denote its type; \eg uppercase subscripts refer to tuples with subscriptable components.
 
-Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a \hyperlink{transaction_nonce}{transaction nonce}.
-Those with a particularly special meaning may be Greek, \eg $\delta$, the number of items required on the stack for a given operation.
-
-Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte sequence given as the output data of a message call.
+Scalars and fixed-size byte sequences (or, synonymously, arrays) are denoted with a normal lower-case letter, \eg $n$ is used in the document to denote a transaction nonce.
+Those with a particularly special meaning may be greek, \eg $\delta$, the number of items required on the stack for a given operation.
+
+Arbitrary-length sequences are typically denoted as a bold lower-case letter, \eg $\mathbf{o}$ is used to denote the byte-sequence given as the output data of a message call.
 For particularly important values, a bold uppercase letter may be used.
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
-This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
-
-Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_{\mathbf{s}}[0]$ denotes the first item on the machine's stack.
-For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_{\mathbf{m}}[0..31]$ denotes the first 32 items of the machine's memory.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
+This is formally defined in section \ref{ch:block}.
+
+Square brackets are used to index into and reference individual components or subsequences of sequences, \eg $\boldsymbol{\mu}_\mathbf{s}[0]$ denotes the first item on the machine's stack.
+For subsequences, ellipses are used to specify the intended range, to include elements at both limits, \eg $\boldsymbol{\mu}_\mathbf{m}[0..31]$ denotes the first 32 items of the machine's memory.
 
@@
 
-The account state, $\boldsymbol{\sigma}[a]$, comprises the following four fields:
+The account state comprises the following four fields:
 
 \begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
-For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
-\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{b}}$.
+\item[nonce] A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_n$.
+\item[balance] A scalar value equal to the number of Wei owned by this address. Formally denoted $\boldsymbol{\sigma}[a]_b$.
 \item[storageRoot] A 256-bit hash of the root node of a Merkle Patricia tree that encodes the storage contents of the account (a mapping between 256-bit integer values), encoded into the trie as a mapping from the Keccak 256-bit hash of the  256-bit integer keys to the RLP-encoded 256-bit integer values.
-The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
+The hash is formally denoted $\boldsymbol{\sigma}[a]_s$.
 \item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_c$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_c$.
 \end{description}
 
-Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
-\begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+Since I typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, I define a convenient equivalence:
+\begin{equation}
+\texttt{\small TRIE}\big(L_I^*(\boldsymbol{\sigma}[a]_\mathbf{s})\big) \equiv \boldsymbol{\sigma}[a]_s
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_I^*$, is defined as the element-wise transformation of the base function $L_I$, given as:
+\begin{equation}
+L_I\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 \begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
-\end{equation}
-
-It shall be understood that $\boldsymbol{\sigma}[a]_{\mathbf{s}}$ is not a `physical' member of the account and does not contribute to its later serialisation.
+k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
+\end{equation}
+
+It shall be understood that $\boldsymbol{\sigma}[a]_\mathbf{s}$ is not a `physical' member of the account and does not contribute to its later serialisation.
 
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
-
-Thus we may define a world-state collapse function $L_{S}$:
-\begin{equation}
-L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
+$\boldsymbol{\sigma}[a]_c = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+
+Thus we may define a world-state collapse function $L_S$:
+\begin{equation}
+L_S(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
 \end{equation}
@@
 \end{equation}
-\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
-\begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
-\end{equation}
-
-An account is \textit{empty} when it has no code, zero nonce and zero balance:
-\begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
-\end{equation}
-Even callable precompiled contracts can have an empty account state. This is because their account states do not usually contain the code describing its behavior.
-
-An account is \textit{dead} when its account state is non-existent or empty:
-\begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
-\end{equation}
-
-\subsection{The Transaction} \label{subsec:transaction}
+where $v$ is the account validity function:
+\begin{equation}
+\quad v(x) \equiv x_n \in \mathbb{P}_{256} \wedge x_b \in \mathbb{P}_{256} \wedge x_s \in \mathbb{B}_{32} \wedge x_c \in \mathbb{B}_{32}
+\end{equation}
+
+\subsection{The Transaction} \label{ch:transaction}
 
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+While is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
 \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
@@
 \begin{equation}
-L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+L_T(T) \equiv \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}, T_w, T_r, T_s) & \text{if} \; T_t = \varnothing\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}, T_w, T_r, T_s) & \text{otherwise} 
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_\mathbf{i}$ and $T_\mathbf{d}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_n \in \mathbb{P}_{256} & \wedge & T_v \in \mathbb{P}_{256} & \wedge & T_p \in \mathbb{P}_{256} & \wedge \\
+T_g \in \mathbb{P}_{256} & \wedge & T_w \in \mathbb{P}_5 & \wedge & T_r \in \mathbb{P}_{256} & \wedge \\
+T_s \in \mathbb{P}_{256} & \wedge & T_\mathbf{d} \in \mathbb{B} & \wedge & T_\mathbf{i} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
-\end{equation}
-
-The address hash $T_{\mathbf{t}}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte sequence and thus the member of $\mathbb{B}_0$:
-\begin{equation}
-T_{\mathbf{t}} \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+\mathbb{P}_n = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
+\end{equation}
+
+The address hash $T_\mathbf{t}$ is slightly different: it is either a 20-byte address hash or, in the case of being a contract-creation transaction (and thus formally equal to $\varnothing$), it is the RLP empty byte-series and thus the member of $\mathbb{B}_0$:
+\begin{equation}
+T_t \in \begin{cases} \mathbb{B}_{20} & \text{if} \quad T_t \neq \varnothing \\
 \mathbb{B}_{0} & \text{otherwise}\end{cases}
@@
 
-\subsection{The Block}\linkdest{block}\label{subsec:The_Block}
-
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+\subsection{The Block} \label{ch:block}
+
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is the most prevalent (not saying much) gender-neutral term to mean ``sibling of parent''; see \url{http://nonbinary.org/wiki/Gender_neutral_language#Family_Terms}}).
 The block header contains several pieces of information:
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{N}
-\end{equation}
-
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
-\end{equation}
-
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
-\begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
-\end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (\mathtt{TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+\end{equation}
+
+We define the Bloom filter function, $M$, to reduce a log entry include a single 256-byte hash:
+\begin{equation}
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+\end{equation}
+
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 \forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
-m(\mathbf{x}, i) &\equiv& \mathtt{\tiny KEC}(\mathbf{x})[i, i + 1] \bmod 2048
-\end{eqnarray}
-
-where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
+\end{eqnarray}
+
+where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_j(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
 
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
 \begin{equation}
-p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
+p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
 \end{equation}
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
-
-The values stemming from the computation of transactions, specifically the \hyperlink{Transaction_Receipt}{transaction receipts}, $B_{\mathbf{R}}$, and that defined through the transaction's \hyperlink{Pi}{state-accumulation function, $\Pi$}, are formalised later in section \ref{sec:statenoncevalidation}.
+\mathtt{TRIE}(L_S(\boldsymbol{\sigma})) = {P(B_H)_H}_r
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_S(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_H)$ is the parent block of $B$, defined directly.
+
+The values stemming from the computation of transactions, specifically the transaction receipts, $B_\mathbf{R}$, and that defined through the transactions state-accumulation function, $\Pi$, are formalised later in section \ref{sec:statenoncevalidation}.
 
@@
 \begin{array}[t]{lclclcl}
-\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
-\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
+H_p \in \mathbb{B}_{32} & \wedge & H_o \in \mathbb{B}_{32} & \wedge & H_c \in \mathbb{B}_{20} & \wedge \\
+H_r \in \mathbb{B}_{32} & \wedge & H_t \in \mathbb{B}_{32} & \wedge & H_e \in \mathbb{B}_{32} & \wedge \\
+H_b \in \mathbb{B}_{256} & \wedge & H_d \in \mathbb{P} & \wedge & H_i \in \mathbb{P} & \wedge \\
+H_l \in \mathbb{P} & \wedge & H_g \in \mathbb{P} & \wedge & H_s \in \mathbb{P}_{256} & \wedge \\
+H_x \in \mathbb{B} & \wedge & H_m \in \mathbb{B}_{32} & \wedge & H_n \in \mathbb{B}_{8}
 \end{array}
@@
 
-We define $P(B_{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
-\end{equation}
-
-\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
-\begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
-\end{equation}
-
-\newcommand{\mindifficulty}{D_0}
-\newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
-\newcommand{\expdiffsymb}{\ensuremath{\epsilon}}
-\newcommand{\diffadjustment}{x}
-
-\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
+We define $P(B_H)$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_H)) = H_p
+\end{equation}
+
+The block number is the parent's block number incremented by one:
+\begin{equation}
+H_i \equiv {{P(H)_H}_i} + 1
+\end{equation}
+
+The canonical difficulty of a block of header $H$ is defined as $D(H)$:
+\begin{equation}
+D(H) \equiv \begin{cases}
+131072 & \text{if} \quad H_i = 0\\
+{P(H)_H}_d + \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor & \text{if} \quad H_s < {P(H)_H}_s + 13\\
+\max \{ 131072, x \} & \text{otherwise}\\
+\end{cases}
 \end{equation}
@@
 \begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-y \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
-\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
-\begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
+x = {P(H)_H}_d - \left\lfloor\frac{{P(H)_H}_d}{2048}\right\rfloor
+\end{equation}
+
+The canonical gas limit $H_l$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_l < {P(H)_H}_l + \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l > {P(H)_H}_l - \left\lfloor\frac{{P(H)_H}_l}{1024}\right\rfloor \quad \wedge \\
+& & H_l \geqslant 125000
+\end{eqnarray}
+
+$H_s$ is the timestamp of block $H$ and must fulfil the relation:
+\begin{equation}
+H_s > {P(H)_H}_s
 \end{equation}
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_\mathbf{s}, A_\mathbf{l}, A_r)
+\end{equation}
+
+The tuple contents include $A_\mathbf{s}$, the suicide set: a set of accounts that will be discarded following the transaction's completion.
+$A_\mathbf{l}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+Finally there is $A_r$, the refund balance, increased through using the {\small SSTORE} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing,(), \varnothing, 0)
+For brevity, we define the empty substate $A^0$ to have no suicides, no logs and a zero refund balance:
+\begin{equation}
+A^0 \equiv (\varnothing, (), 0)
 \end{equation}
@@
 \subsection{Execution}
-\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
-\begin{align}
-g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
-\end{align}
-
-where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
-$G$ is fully defined in Appendix \ref{app:fees}.
+We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
+\begin{equation}
+g_0 \equiv \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} + G_{transaction}\\
+\end{equation}
+
+where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
+$G$ is defined in Appendix \ref{app:fees}.
+
+%todo Explain g_d reason?
 
@@
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+v_0 \equiv T_g T_p + T_v
 \end{equation}
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
-\linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
-g_0 & \leqslant & T_{\mathrm{g}} \quad \wedge \\
-v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_n & = & \boldsymbol{\sigma}[S(T)]_n \quad \wedge \\
+g_0 & \leqslant & T_g \quad \wedge \\ 
+v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_b \quad \wedge \\
+T_g & \leqslant & {B_H}_l - \ell(B_\mathbf{R})_u
 \end{array}
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
+
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
+\section{Contract Creation} \label{ch:create}
+
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
-\boldsymbol{\sigma}^*[s] &=& \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}^* & \text{otherwise}
-\end{cases} \\
-\mathbf{a}^* &\equiv& (\boldsymbol{\sigma}[s]_{\mathrm{n}}, \boldsymbol{\sigma}[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}[s]_{\mathbf{s}}, \boldsymbol{\sigma}[s]_{\mathrm{c}})
+\boldsymbol{\sigma}^*[a] &\equiv& \big( 0, v + v', \mathtt{TRIE}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[s]_b &\equiv& \boldsymbol{\sigma}[s]_b - v
 \end{eqnarray}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
-\begin{eqnarray}
-I_{\mathrm{a}} & \equiv & a \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & () \\
-I_{\mathrm{s}} & \equiv & s \\
-\hyperlink{I__v}{I_{\mathrm{v}}} & \equiv & v \\
-I_{\mathbf{b}} & \equiv & \mathbf{i} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w
-\end{eqnarray}
-
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
-
-Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
-
-If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
-\begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
-\end{equation}
-
-If there is not enough gas remaining to pay this, \ie $g^{**} < c$, then we also declare an out-of-gas exception.
-
-The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
-
-If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
-
-\begin{align}
-\quad g' &\equiv \begin{cases}
-0 & \text{if} \quad F \\
+As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{b}$.
+
+Code execution depletes gas; thus it may exit before the code has come to a natural halting state.
+In this (and several other) exceptional cases we say an Out-of-Gas exception has occurred: The evaluated state is defined as being the empty set $\varnothing$ and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+The gas remaining should be zero in any such exceptional condition.
+If the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost: it is paid regardless.
+
+If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persevere.
+Thus formally, we may specify the resultant state, gas and substate as $(\boldsymbol{\sigma}', g', A)$ where:
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I) \\
+\end{equation}
+\begin{eqnarray}
+\quad g' & \equiv & \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+g^{**} & \text{if} \quad g^{**} < c \\
 g^{**} - c & \text{otherwise} \\
 \end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\quad \boldsymbol{\sigma}' & \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\sigma}^{**} & \text{if} \quad g^{**} < c \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
-\boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
-\end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
-1 & \text{otherwise}
-\end{cases} \\
-\nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
-\end{align}
-
-The exception in the determination of $\boldsymbol{\sigma}'$ dictates that $\mathbf{o}$, the resultant byte sequence from the execution of the initialisation code, specifies the final body code for the newly-created account.
-
-Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+\quad\boldsymbol{\sigma}'[a]_c = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+Where $I$ contains the parameters of the execution environment as defined in section \ref{ch:model}.
+\begin{eqnarray}
+I_a & \equiv & a \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & () \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_\mathbf{b} & \equiv & \mathbf{i} \\
+I_e & \equiv & e
+\end{eqnarray}
+
+where $c$ is the code-deposit cost:
+\begin{equation}
+c \equiv G_{codedeposit} \times |\mathbf{o}|
+\end{equation}
+
+$I_\mathbf{d}$ evaluates to the empty tuple as there is no input data to this call.
+$I_H$ has no special treatment and is determined from the blockchain.
+The exception in the determination of $\boldsymbol{\sigma}'$ dictates that the resultant byte sequence from the execution of the initialisation code specifies the final body code for the newly-created account, with $\boldsymbol{\sigma}'[a]_c$ being the Keccak 256-bit hash of the newly created account's body code and $\mathbf{o}$ the output byte sequence of the code execution.
+
+No code is deposited in the state if the gas does not cover the additional per-byte contract deposit fee, however, the value is still transferred and the execution side-effects take place.
 
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
-\boldsymbol{\sigma}_1'[r] \equiv \mathbf{a}_1' & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1' \equiv (\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + v, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}})
-\end{equation}
-
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
+\boldsymbol{\sigma}_1'[r] \equiv (v, 0, \mathtt{KEC}(()), \mathtt{TRIE}(\varnothing)) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \\
+\boldsymbol{\sigma}_1'[r]_b \equiv \boldsymbol{\sigma}[r]_b + v & \text{otherwise}
+\end{cases}
+\end{equation}
+
+The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_c$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \ \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
-\qquad (\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) & \equiv & \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases} \\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\\ \nonumber
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{eqnarray}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_e & \equiv & e \\
+\text{Let} \; \mathtt{KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 Formally, we define it recursively with a function $X$.
-This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions \hyperlink{zhalt}{$Z$} which determines if the present state is an \hyperlink{zhalt}{exceptional halting} state of the machine and \hyperlink{hhalt}{$H$}, specifying the output data of the instruction if and only if the present state is a \hyperlink{hhalt}{normal halting} state of the machine.
-
-\hypertarget{empty_sequence_vs_empty_set}{}The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
-\begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
-\boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
-\boldsymbol{\mu}_{\mathrm{pc}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{m}} & \equiv & (0, 0, ...) \\
-\boldsymbol{\mu}_{\mathrm{i}} & \equiv & 0 \\
-\boldsymbol{\mu}_{\mathbf{s}} & \equiv & () \\
-\boldsymbol{\mu}_{\mathbf{o}} & \equiv & ()
-\end{eqnarray}
-\begin{equation} \label{eq:X-def}
+This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying the output data of the instruction if and only if the present state is a normal halting state of the machine.
+
+The empty sequence, denoted $()$, is not equal to the empty set, denoted $\varnothing$; this is important when interpreting the output of $H$, which evaluates to $\varnothing$ when execution is to continue but a series (potentially empty) when execution should halt.
+\begin{eqnarray}
+\Xi(\boldsymbol{\sigma}, g, I) & \equiv & X_{0,1,2,4}\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0, I)\big) \\
+\boldsymbol{\mu}_g & \equiv & g \\
+\boldsymbol{\mu}_{pc} & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{m} & \equiv & (0, 0, ...) \\
+\boldsymbol{\mu}_i & \equiv & 0 \\
+\boldsymbol{\mu}_\mathbf{s} & \equiv & ()
+\end{eqnarray}
+\begin{equation}
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+\big(\varnothing, \boldsymbol{\mu}, A^0, I, ()\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)\\
+O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing\\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise}\\
 \end{cases}
@@
 \mathbf{o} & \equiv & H(\boldsymbol{\mu}, I) \\
-(a, b, c, d) \cdot e & \equiv & (a, b, c, d, e) \\
-\boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
-\end{eqnarray}
-
-Note that, when we evaluate $\Xi$, we drop the fourth element $I'$ and extract the remaining gas $\boldsymbol{\mu}'_{\mathrm{g}}$ from the resultant machine state $\boldsymbol{\mu}'$.
-
-$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either \hyperlink{zhalt}{$Z$} becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until \hyperlink{hhalt}{$H$} becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
+(a, b, c) \cdot d & \equiv & (a, b, c, d)
+\end{eqnarray}
+
+Note that we must drop the fourth value in the tuple returned by $X$ to correctly evaluate $\Xi$, hence the subscript $X_{0,1,2,4}$.
+
+$X$ is thus cycled (recursively here, but implementations are generally expected to use a simple iterative loop) until either $Z$ becomes true indicating that the present state is exceptional and that the machine must be halted and any changes discarded or until $H$ becomes a series (rather than the empty set) indicating that the machine has reached a controlled halt.
 
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
-The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The memory contents $\boldsymbol{\mu}_\mathbf{m}$ are a series of zeroes of size $2^{256}$.
 
@@
 \begin{equation}\label{eq:currentoperation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert \\
-\text{\hyperlink{stop}{\small STOP}} & \text{otherwise}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert \\
+\text{\small STOP} & \text{otherwise}
 \end{cases}
@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
-\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \, \text{\small SSTORE}, \\
-\text{\small SELFDESTRUCT}\} \quad \vee \\
-(\text{\small LOG0} \le w \wedge w \le \text{\small LOG4}) \quad \vee \\
-(w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0)
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
-() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
 \varnothing & \text{otherwise}
@@
 O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
-\Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
-\lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert & \equiv & \lVert\boldsymbol{\mu}_{\mathbf{s}}\rVert + \Delta \\
-\quad \forall x \in [\mathbf{\alpha}_{w}, \lVert\boldsymbol{\mu}'_{\mathbf{s}}\rVert): \boldsymbol{\mu}'_{\mathbf{s}}[x] & \equiv & \boldsymbol{\mu}_{\mathbf{s}}[x-\Delta]
+\Delta & \equiv & \mathbf{\alpha}_w - \mathbf{\delta}_w \\
+\lVert\boldsymbol{\mu}'_\mathbf{s}\rVert & \equiv & \lVert\boldsymbol{\mu}_\mathbf{s}\rVert + \Delta \\
+\quad \forall x \in [\mathbf{\alpha}_w, \lVert\boldsymbol{\mu}'_\mathbf{s}\rVert): \boldsymbol{\mu}'_\mathbf{s}[x] & \equiv & \boldsymbol{\mu}_\mathbf{s}[x+\Delta]
 \end{eqnarray}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
-\quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
-\hyperlink{JUMP}{J_{\text{JUMP}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
-\hyperlink{JUMPI}{J_{\text{JUMPI}}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
-N(\boldsymbol{\mu}_{\mathrm{pc}}, w) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-In general, we assume the memory, self-destruct set and system state don't change:
-\begin{eqnarray}
-\boldsymbol{\mu}'_{\mathbf{m}} & \equiv & \boldsymbol{\mu}_{\mathbf{m}} \\
-\boldsymbol{\mu}'_{\mathrm{i}} & \equiv & \boldsymbol{\mu}_{\mathrm{i}} \\
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
+\quad \boldsymbol{\mu}'_{pc} & \equiv & \begin{cases}
+J_{\text{JUMP}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMP} \\
+J_{\text{JUMPI}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small JUMPI} \\
+N(\boldsymbol{\mu}_{pc}, w) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+In general, we assume the memory, suicide list and system state don't change:
+\begin{eqnarray}
+\boldsymbol{\mu}'_\mathbf{m} & \equiv & \boldsymbol{\mu}_\mathbf{m} \\
+\boldsymbol{\mu}'_i & \equiv & \boldsymbol{\mu}_i \\
 A' & \equiv & A \\
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
+\begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+\end{eqnarray}
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 
-For $\mathbf{R}[n]_{\mathbf{l}}$, we utilise the \hyperlink{Upsilon_pow_l}{$\Upsilon^\mathbf{l}$} function that we conveniently defined in the transaction execution function.
-\begin{equation}
-\mathbf{R}[n]_{\mathbf{l}} =
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-We define \hyperlink{Upsilon_pow_z}{$\mathbf{R}[n]_{\mathrm{z}}$} in a similar manner.
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{z}} =
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])
-\end{equation}
-
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
-\end{equation}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
+For $\mathbf{R}[n]_\mathbf{l}$, we utilise the $\Upsilon^\mathbf{l}$ function that we conveniently defined in the transaction execution function.
+\begin{equation}
+\mathbf{R}[n]_\mathbf{l} = 
+\Upsilon^\mathbf{l}(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])
+\end{equation}
+
+Finally, we define $\Pi$ as the new state given the block reward function $\Omega$ applied to the final transaction's resultant state, $\ell(B_\mathbf{R})_{\boldsymbol{\sigma}}$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \Omega(B, \ell(\mathbf{R})_{\boldsymbol{\sigma}})
+\end{equation}
+
+Thus the complete block-transition mechanism, less $\mathtt{PoW}$, the proof-of-work function is defined.
 
@@
 \begin{equation}
-m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{h_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
+m = H_m \quad \wedge \quad n \leqslant \frac{2^{256}}{H_d} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d})
+\end{equation}
+
+Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_n$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_d$ is the new block's difficulty value (i.e.
 the block difficulty from section \ref{ch:ghost}).
@@
 \subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
+Ethash is the planned PoW algorithm for Ethereum 1.0.
+It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms have been drastically changed in the last month of research and development.
 The general route that the algorithm takes is as follows:
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $J_{cacheinit}$ bytes in intial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $J_{datasetinit}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 
-The state database won't be forced to maintain all past state \hyperlink{trie}{trie} structures into the future.
-It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints.
-Checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
+The state database won't be forced to maintain all past state trie structures into the future.
+It should maintain an age for each node and eventually discard nodes that are neither recent enough nor checkpoints; checkpoints, or a set of nodes in the database that allow a particular block's state trie to be traversed, could be used to place a maximum limit on the amount of computation needed in order to retrieve any state throughout the blockchain.
 
 Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
-A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{Genesis_Block}{genesis block}.
+A compressed archive of the trie structure at given points in time (perhaps one in every 10000th block) could be maintained by the peer network, effectively recasting the genesis block.
 This would reduce the amount to be downloaded to a single archive plus a hard maximum limit  of blocks.
@@
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
-\end{eqnarray}
-
-Where $\mathbb{O}$ is the set of bytes.
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{Y} \}
+\end{eqnarray}
+
+Where $\mathbb{Y}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
 
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_b(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_l(\mathbf{x}) & \text{otherwise} \end{cases}
+\end{equation}
+
+If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
@@
 
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_math_def}{}Formally, we define $R_{\mathrm{b}}$:
-\begin{eqnarray}
-R_{\mathrm{b}}(\mathbf{x}) & \equiv & \begin{cases}
+Formally, we define $R_b$:
+\begin{eqnarray}
+R_b(\mathbf{x}) & \equiv & \begin{cases}
 \mathbf{x} & \text{if} \quad \lVert \mathbf{x} \rVert = 1 \wedge \mathbf{x}[0] < 128 \\
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
-
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_word_def}{}\linkdest{R__l}If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
+Thus $\mathtt{BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+
+If instead, the value to be serialised is a sequence of other items then the RLP serialisation takes one of two forms:
 
@@
 
-\hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
-\begin{eqnarray}
-R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
+Thus we finish by formally defining $R_l$:
+\begin{eqnarray}
+R_l(\mathbf{x}) & \equiv & \begin{cases}
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
 \end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
-\end{eqnarray}
-
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+s(\mathbf{x}) & \equiv & \mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) ...
+\end{eqnarray}
+
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
 \begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{N}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\mathtt{RLP}(i : i \in \mathbb{P}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
 \end{equation}
@@
 
-\section{Modified Merkle Patricia Tree}\label{app:trie}\hypertarget{trie}{}
+\section{Modified Merkle Patricia Tree}\label{app:trie}
 The modified Merkle Patricia tree (trie) provides a persistent data structure to map between arbitrary-length binary data (byte arrays).
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
-It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may either a 32 byte sequence or the empty byte sequence.
+It is left as an implementation consideration to store and maintain the structure of the trie in a manner the allows effective and efficient realisation of the protocol.
 
@@
 For each precompiled contract, we make use of a template function, $\Xi_{\mathtt{PRE}}$, which implements the out-of-gas checking.
-\begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
-\end{equation}
-
-The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_{\mathrm{r}}$, the gas requirements.
-
-We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
-We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
-
+\begin{equation}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I) \equiv \begin{cases}
+(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_r \\
+(\boldsymbol\sigma, g - g_r, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\end{equation}
+
+The precompiled contracts each use these definitions and provide specifications for the $\mathbf{o}$ (the output data) and $g_r$, the gas requirements.
+
+For the elliptic curve DSA recover VM execution function, we also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
+Importantly in the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), then we have no output.
 \begin{eqnarray}
 \Xi_{\mathtt{ECREC}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+g_r &=& 3000\\
+|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
 \text{if} \quad |\mathbf{o}| = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
-\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
+\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{d}[0..(|I_\mathbf{d}|-1)] &=& I_\mathbf{d}\\
+\mathbf{d}[|I_\mathbf{d}|..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 
-We define $\Xi_{\mathtt{SHA256}}$ and $\Xi_{\mathtt{RIP160}}$ as precompiled contracts implementing the SHA2-256 and RIPEMD-160 hash functions respectively.
+The two hash functions, RIPEMD-160 and SHA2-256 are more trivially defined as an almost pass-through operation.
 Their gas usage is dependent on the input data size, a factor rounded up to the nearest number of words.
-
 \begin{eqnarray}
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
+g_r &=& 60 + 12\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_\mathbf{d})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_r &=& 600 + 120\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
+\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_\mathbf{d})\\
 \end{eqnarray}
@@
 For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
-
-\begin{eqnarray}
-\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
-\end{eqnarray}
-
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+\begin{eqnarray}
+\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
+\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\end{eqnarray}
+
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\
-\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
-\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+\section{Signing Transactions}\label{app:signing}
+
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{ECDSASIGN}$, $\mathtt{ECDSARESTORE}$ and $\mathtt{ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
+where:
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
+
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
-\end{eqnarray}
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
-This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
-
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
-It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-\linkdest{invalidsig}We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
-\end{align}
-where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
-
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
-\end{eqnarray}
-
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
-\begin{eqnarray}
-\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
-\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
-\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
+h(T) & \equiv & \mathtt{KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 We may then define the sender function $S$ of the transaction as:
-\begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-The assertion that the sender of a signed transaction equals the address of the signer should be self-evident:
-\begin{equation}
-\forall T: \forall p_{\mathrm{r}}: S(G(T, p_{\mathrm{r}})) \equiv A(p_{\mathrm{r}})
+\begin{equation}
+S(T) \equiv \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), T_w, T_r, T_s) \big) \big)
+\end{equation}
+
+The assertion that the sender of the a signed transaction equals the address of the signer should be self-evident:
+\begin{equation}
+\forall T: \forall p_r: S(G(T, p_r)) \equiv A(p_r)
 \end{equation}
@@
 $G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
 $G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
 $G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
@@
 $G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
-\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
-
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{memory}(\boldsymbol{\mu}'_i)-C_{memory}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 G_{verylow} & \text{if} \quad w \in W_{verylow}\\
-G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
+G_{low} & \text{if} \quad w \in W_{low}\\
 G_{mid} & \text{if} \quad w \in W_{mid}\\
-G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
-\end{cases}
-\end{equation}
-\begin{equation}
-w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
+G_{high} & \text{if} \quad w \in W_{high}\\
+G_{ext} & \text{if} \quad w \in W_{ext}
+\end{cases}
+\end{equation}
+\begin{equation}
+w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
 \text{\small STOP} & \text{otherwise}
@@
 
-$W_{\mathrm{high}}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+$W_{high}$ = \{{\small JUMPI}\}
+
+$W_{ext}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small BLOCKHASH}\}
+
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_i$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
@@
 &&&& or more than 256 blocks behind the current block. \\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H(h)_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H(h)_{\mathrm{i}} \\ P(H(h)_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H(h)$ can be determined from its hash~$h$ unless $h$ is zero. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
@@
 \midrule
-\linkdest{JUMP}{}0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
-&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}.\\
-\midrule
-\linkdest{JUMPI}{}0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
-&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \\ \boldsymbol{\mu}_{\mathrm{pc}} + 1 & \text{otherwise} \end{cases} $ \\
-&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{\mathrm{pc}}$. See section \ref{ch:model}. \\
+0x56 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
+&&&& $J_{\text{\tiny JUMP}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{s}[0] $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
+\midrule
+0x57 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
+&&&& $J_{\text{\tiny JUMPI}}(\boldsymbol{\mu}) \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[0] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \\ \boldsymbol{\mu}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
+&&&& This has the effect of writing said value to $\boldsymbol{\mu}_{pc}$. See section \ref{ch:model}. \\
 \midrule
@@
 &&&&  corresponding to this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_{pc}$ \\
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv c(\boldsymbol{\mu}_{\mathrm{pc}} + 1)$ \\
-&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_{\mathbf{b}}[x] & \text{if} \quad x < \lVert I_{\mathbf{b}} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv c(\boldsymbol{\mu}_{pc} + 1)$ \\
+&&&& $\text{where} \quad c(x) \equiv \begin{cases} I_\mathbf{b}[x] & \text{if} \quad x < \lVert I_\mathbf{b} \rVert \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& The bytes are read in line from the program code's bytes array. \\
@@
 0x61 & {\small PUSH2} & 0 & 1 & Place 2-byte item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 2) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big( (\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 2) \big)$ \\
 &&&& with $\boldsymbol{c}(\boldsymbol{x}) \equiv (c(\boldsymbol{x}_0), ..., c(\boldsymbol{x}_{\lVert x \rVert -1})) $ with $c$ as defined as above. \\
@@
 0x7f & {\small PUSH32} & 0 & 1 & Place 32-byte (full word) item on stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{\mathrm{pc}} + 1) \dots (\boldsymbol{\mu}_{\mathrm{pc}} + 32) \big)$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{c}\big((\boldsymbol{\mu}_{pc} + 1) \dots (\boldsymbol{\mu}_{pc} + 32) \big)$ \\
 &&&& where $\boldsymbol{c}$ is defined as above. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
-&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
-&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
-&&&& same account as at present, simply that the code is overwritten.\\
-\midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overridden altered.\\
+\midrule
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
+0 & \text{otherwise}
+\end{cases}$ \\\bottomrule
 \end{tabular*}
 
-\begin{tabular*}{\columnwidth}[h]{rlrrl}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
-&&&& the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
-&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
-&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& Otherwise exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
-\bottomrule
-\end{tabular*}
-
-\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
+\section{Genesis Block}\label{app:genesis}
 
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer lists in RLP, both empty lists.
 
 The proof-of-concept series include a development premine, making the state root hash some value $stateRoot$.
-Also $time$ will be set to the initial timestamp of the genesis block.
+Also $time$ will be set to the intial timestamp of the genesis block.
 The latest documentation should be consulted for those values.
@@
 \midrule
-\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
@@
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
-\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{hashbytes}$ for the cache.
+Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{prime}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
 \end{equation}
@@
 \begin{equation}
- c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
-\end{equation}
-\begin{equation}
- E_{\mathrm{prime}}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{N} \\
-E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}
+ c_{size} = E_{prime}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{hashbytes}, J_{hashbytes})
+\end{equation}
+\begin{equation}
+ E_{prime}(x, y) = \begin{cases}
+x & \text{if} \quad x / y \in \mathbb{P} \\
+E_{prime}(x - 2 \cdot y, y) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
+ n = \left\lfloor\frac{c_{size}}{J_{hashbytes}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the inital cache $\mathbf{c'}$:
 \begin{equation}
@@
 \subsubsection{Full dataset calculation} \label{dataset} 
-Essentially, we combine data from $J_{\mathrm{parents}}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
-The entire dataset is then generated by a number of items, each $J_{\mathrm{hashbytes}}$ bytes in size:
-\begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
+Essentially, we combine data from $J_{parents}$ pseudorandomly selected cache nodes, and hash that to compute the dataset.
+The entire dataset is then generated by a number of items, each $J_{hashbytes}$ bytes in size:
+\begin{equation}
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{hashbytes}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
-\end{equation}
-With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_{\mathrm{c}}$ is obtained as follows:
-\begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathtt{PoW}(H_{\hcancel{n}}, H_n, \mathbf{d}) = \lbrace \mathbf{m}_c(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d}), \texttt{KEC}(\mathbf{s}_h(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n) + \mathbf{m}_c(\mathtt{KEC}(\mathtt{RLP}(L_H(H_{\hcancel{n}}))), H_n, \mathbf{d})) \rbrace
+\end{equation}
+With $H_{\hcancel{n}}$ being the hash of the header without the nonce. The compressed mix $\mathbf{m}_c$ is obtained as follows:
+\begin{equation}
+ \mathbf{m}_c(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{accesses}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_h(\mathbf{h}, \mathbf{n}), \mathbf{s}_h(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols} \label{app:symbols}
-\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
-\midrule
-$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2018-03-03 Paper.tex
Add \left \right to various EQs
Add \left \right to EQs, leave \big on ones that use it for bigger brackets when \left\right do not apply.

Edit spacing on (231) to reflect.

--- 2018-03-01 Paper.tex
@@
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.\footnote{Ethereum 1.0 uses proof-of-work, while it is planned that Ethereum 2.0 will include \href{https://github.com/ethereum/wiki/wiki/Proof-of-Stake-FAQ}{proof-of-stake} and \href{https://github.com/ethereum/wiki/wiki/Sharding-FAQ}{sharding}.}
+For Ethereum 1.0 we have chosen the first path.
 
@@
 \end{cases} \\
-\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
+\mathbf o &=& \left(B^E\bmod M\right)\in\mathbb P_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\
@@
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
 F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
-y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf d}[0..63]\right)\\
+y&\equiv&\delta_1\left(\bar I_{\mathbf d}[64..127]\right)\\
 \label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
@@
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 40000\\
@@
 n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
+x&\equiv&\delta_1\left(\bar I_{\mathbf d}[32..95]\right)
 \end{eqnarray}

### 2018-03-03 Paper.tex
Add \nonumber to lines in multi-line equations in equation arrays

@@
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \right\}
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \}
 \end{equation}
@@
 \begin{equation}
-\mathbf{a}_1 \equiv \left(\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}}\right)
+\mathbf{a}_1 \equiv (\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}})
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + (1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32})R_{\mathrm{block}} \\
 \qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
+R & \equiv & (1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})) R_{\mathrm{block}}
 \end{eqnarray}
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
+g_{\mathrm{r}} &=& \Big\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\Big\rfloor \\
 f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\[0.5em]
-\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
-\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
+x^2 & \text{if}\ x \le 64 \\
+\Big\lfloor\dfrac{x^2}{4}\Big\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\
+\Big\lfloor\dfrac{x^2}{16}\Big\rfloor + 480x - 199680 & \text{otherwise}
 \end{cases}\\
@@
 \end{cases} \\
-\mathbf o &=& \left(B^E\bmod M\right)\in\mathbb P_{8\ell_{M}} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\
@@
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
 F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 500\\
 \mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf d}[0..63]\right)\\
-y&\equiv&\delta_1\left(\bar I_{\mathbf d}[64..127]\right)\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
 \label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
@@
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 40000\\
@@
 n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf d}[32..95]\right)
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
+C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
 \end{equation}

### 2018-03-04 Paper.tex
Corrected constant coefficient of G2
Fix Elliptic Curve group addition equations

@@
 & & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} \geqslant 5000
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000
 \end{eqnarray}
@@
 V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
-\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
+& & H_{\mathrm{d}} = D(H) \quad \wedge \\
+& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
+& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+& & \lVert H_{\mathrm{x}} \rVert \le 32
 \end{eqnarray}
@@
 \end{cases}\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
 X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+Y&\equiv&\lambda(X_1-X)-Y_1
+\end{eqnarray}
+
+In the case where $(X_1, Y_1) = (X_2, Y_2)$, we define $+$ on $C_1$ with
+\begin{eqnarray}\label{eq:ec-doubling}
+(X_1, Y_1) + (X_1, Y_1)&\equiv&\begin{cases}
+(X,Y)&\text{if}\ Y_1\neq 0\\
+(0,0)&\text{otherwise}
+\end{cases}\\
+\lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
+X&\equiv&\lambda^2-2X_1\\
+Y&\equiv&\lambda(X_1-X)-Y_1
 \end{eqnarray}
@@
 \end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.

### 2018-03-05 Paper.tex

@@
 \end{equation}
-We define a binary operation $+$ on $C_1$ for distinct elements $(X_1, Y_1), (X_2, Y_2)$ with
+We define a binary operation $+$ on $C_1$ with
 \begin{eqnarray}\label{eq:ec-addition}
@@
 \end{cases}\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
 X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1
-\end{eqnarray}
-
-In the case where $(X_1, Y_1) = (X_2, Y_2)$, we define $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-doubling}
-(X_1, Y_1) + (X_1, Y_1)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ Y_1\neq 0\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-\lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
-X&\equiv&\lambda^2-2X_1\\
-Y&\equiv&\lambda(X_1-X)-Y_1
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
 \end{eqnarray}
@@
 \end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.

### 2018-03-05 Paper.tex
For more mathematical symbols, see \href{https://en.wikipedia.org/wiki/List_of_mathematical_symbols}{here}.

@@
 \begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3(i+9)^{-1}\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.

### 2018-03-06 Paper.tex
Add \nonumber to lines in elliptic curve group law

@@
 
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
+$(C_1,+)$ is known to form a group. We define scalar multiplication $\cdot$ with
 \begin{equation}\label{eq:ec-scalar-multiplication}
@@
 
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i^2+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 \end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
+We define a binary operation $+$ and scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.

### 2018-03-06 Paper.tex

--- 2018-03-05 Paper.tex
@@
 \usepackage{tikz}
-\usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
-Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
-%,pagebackref=true
-
-\usepackage{tabu} %requires array.
-
-%This should be the last package before \input{Version.tex}
-\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
-% "hyperref loads the url package internally.
-Use \PassOptionsToPackage{hyphens}{url}\usepackage{hyperref} to pass the option to the url package when it is loaded by hyperref.
-This avoids any package option clashes." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment44478_3034>.
-% Note also this: "If the \PassOptionsToPackage{hyphens}{url} approach does not work, maybe it's "because you're trying to load the url package with a specific option, but it's being loaded by one of your packages before that with a different set of options.
-Try loading the url package earlier than the package that requires it.
-If it's loaded by the document class, try using \RequirePackage[hyphens]{url} before the document class." Source: <https://tex.stackexchange.com/questions/3033/forcing-linebreaks-in-url/3034#comment555944_3034>.
-% For more information on using the hyperref package, refer to e.g. https://en.wikibooks.org/w/index.php?title=LaTeX/Hyperlinks&stable=0#Hyperlink_and_Hypertarget.
-
-\makeatletter
- \newcommand{\linkdest}[1]{\Hy@raisedlink{\hypertarget{#1}{}}}
-\makeatother
-\usepackage{seqsplit}
-
-% For formatting
-%\usepackage{underscore}
-%\usepackage{lipsum} % to generate filler text for testing of document rendering
-\usepackage[english]{babel}
-\usepackage[autostyle]{csquotes}
-\MakeOuterQuote{"}
-
-\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
-
-\input{Version.tex}
-% Default rendering options
-\definecolor{pagecolor}{rgb}{1,0.98,0.9}
-\def\YellowPaperVersionNumber{unknown revision}
-\IfFileExists{Options.tex}{\input{Options.tex}}
 
@@
 \begin{description}
-\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
-\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
-\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
-\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
-\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
-The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
-\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
-\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
-\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
-\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[parentHash] The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_p$.
+\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_o$.
+\item[beneficiary] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_c$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
+\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_t$.
+\item[receiptsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_e$.
+\item[logsBloom] The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_b$.
+\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
+\item[number] A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally $H_i$.
+\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_l$.
+\item[gasUsed] A scalar value equal to the total gas used in transactions in this block; formally $H_g$.
+\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
+\item[extraData] An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_x$.
+\item[mixHash] A 256-bit hash which proves combined with the nonce that a sufficient amount of computation has been carried out on this block; formally $H_m$.
+\item[nonce] A 64-bit hash which proves combined with the mix-hash that a sufficient amount of computation has been carried out on this block; formally $H_n$.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
-Formally, we can refer to a block $B$:
-\begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
-\end{equation}
-
-\subsubsection{Transaction Receipt}\linkdest{Transaction_Receipt}{}
+
+The other two components in the block are simply a list of ommer block headers (of the same format as above) and a series of the transactions. Formally, we can refer to a block $B$:
+\begin{equation}
+B \equiv (B_H, B_\mathbf{T}, B_\mathbf{U})
+\end{equation}
+
+\subsubsection{Transaction Receipt}
 
 In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
-Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
-
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
-\begin{equation}
-R_{\mathrm{z}} \in \mathbb{N}
-\end{equation}
-
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
-\begin{equation}
-R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
-\end{equation}
-
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
-
-The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
-\begin{equation}
-O \equiv (O_{\mathrm{a}}, ({O_{\mathbf{t}}}_0, {O_{\mathbf{t}}}_1, ...), O_{\mathbf{d}})
-\end{equation}
-\begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
-\end{equation}
-
-We define the Bloom filter function, $M$, to reduce a log entry into a single 256-byte hash:
-\begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
-\end{equation}
-
-where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+Each receipt, denoted $B_\mathbf{R}[i]$ for the $i$th transaction) is placed in an index-keyed trie and the root recorded in the header as $H_e$.
+
+The transaction receipt is a tuple of four items comprising the post-transaction state, $R_{\boldsymbol{\sigma}}$, the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_u$, the set of logs created through execution of the transaction, $R_\mathbf{l}$ and the Bloom filter composed from information in those logs, $R_b$:
+\begin{equation}
+R \equiv (R_{\boldsymbol{\sigma}}, R_u, R_b, R_\mathbf{l})
+\end{equation}
+
+The function $L_R$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_R(R) \equiv (\mathtt{TRIE}(L_S(R_{\boldsymbol{\sigma}})), R_u, R_b, R_\mathbf{l})
+\end{equation}
+thus the post-transaction state, $R_{\boldsymbol{\sigma}}$ is encoded into a trie structure, the root of which forms the first item.
+
+We assert $R_u$, the cumulative gas used is a positive integer and that the logs Bloom, $R_b$, is a hash of size 2048 bits (256 bytes):
+\begin{equation}
+R_u \in \mathbb{P} \quad \wedge \quad R_b \in \mathbb{B}_{256}
+\end{equation}
+
+%Notably $B_\mathbf{T}$ does not get serialised into the block by the block preparation function $L_B$; it is merely a convenience equivalence.
+
+The log entries, $R_\mathbf{l}$, is a series of log entries, termed, for example, $(O_0, O_1, ...)$.
+A log entry, $O$, is a tuple of a logger's address, $O_a$, a series of 32-bytes log topics, $O_\mathbf{t}$ and some number of bytes of data, $O_\mathbf{d}$:
+\begin{equation}
+O \equiv (O_a, ({O_\mathbf{t}}_0, {O_\mathbf{t}}_1, ...), O_\mathbf{d})
+\end{equation}
+\begin{equation}
+O_a \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_\mathbf{t}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_\mathbf{d} \in \mathbb{B}
+\end{equation}
+
+We define the Bloom filter function, $M$, to reduce a log entry include a single 256-byte hash:
+\begin{equation}
+M(O) \equiv \bigvee_{t \in \{O_a\} \cup O_\mathbf{t}} \big( M_{3:2048}(t) \big)
+\end{equation}
+
+where $M_{3:512}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte series.
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte series.
+Formally:
 \begin{eqnarray}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_\mathbf{T}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_r$:
 \begin{equation}
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
-\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
+H_r &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+H_o &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_\mathbf{U}))) & \wedge \\
+H_t &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{T} \rVert, i \in \mathbb{P}: p(i, L_T(B_\mathbf{T}[i]))\}) & \wedge \\
+H_e &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_\mathbf{R} \rVert, i \in \mathbb{P}: p(i, L_R(B_\mathbf{R}[i]))\}) & \wedge \\
+H_b &\equiv& \bigvee_{\mathbf{r} \in B_\mathbf{R}} \big( \mathbf{r}_b \big)
 \end{array}
 \end{equation}
+
 where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
 \begin{equation}
-p(k, v) \equiv \big( \mathtt{\small RLP}(k), \mathtt{\small RLP}(v) \big)
+p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
 \end{equation}
@@
 \boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
-\end{eqnarray}
-
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\boldsymbol{\sigma}_0[S(T)]_b & \equiv & \boldsymbol{\sigma}[S(T)]_b - T_g T_p \\
+\boldsymbol{\sigma}_0[S(T)]_n & \equiv & \boldsymbol{\sigma}[S(T)]_n + 1
+\end{eqnarray}
+
+Evaluating $\boldsymbol{\sigma}_P$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_P$, remaining gas $g'$ and substate $A$:
+\begin{equation}
+(\boldsymbol{\sigma}_P, g', A) \equiv \begin{cases}
+\Lambda(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad g, T_p, T_v, T_\mathbf{i}, 0) & \text{if} \quad T_t = \varnothing \\
+\Theta_{3}(\boldsymbol{\sigma}_0, S(T), T_o, &\\ \quad\quad T_t, T_t, g, T_p, T_v, T_\mathbf{d}, 0) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \right\}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \{ \Big\lfloor \dfrac{T_g - g'}{2} \Big\rfloor, A_r \}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to $A_r$, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_g - g'$.
 
 The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
-\end{eqnarray}
-
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_P$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_P \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_b & \equiv & \boldsymbol{\sigma}_P[S(T)]_b + g^* T_p \\
+\boldsymbol{\sigma}^*[m]_b & \equiv & \boldsymbol{\sigma}_P[m]_b + (T_g - g^*) T_p \\
+m & \equiv & {B_H}_c
+\end{eqnarray}
+
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that appear in the suicide list:
 \begin{eqnarray}
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
-\Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
-\end{eqnarray}
-
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
-
-\section{Contract Creation}\label{ch:create}\hypertarget{endow}{}
-
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
+\forall i \in A_\mathbf{s}: \boldsymbol{\sigma}'[i] & \equiv & \varnothing
+\end{eqnarray}
+
+And finally, we specify $\Upsilon^g$, the total gas used in this transaction and $\Upsilon^\mathbf{l}$, the logs created by this transaction:
+\begin{eqnarray}
+\Upsilon^g(\boldsymbol{\sigma}, T) & \equiv & T_g - g' \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & A_\mathbf{l}
+\end{eqnarray}
+
+These are used to help define the transaction receipt, discussed later.
+
+%In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
+%\begin{eqnarray}
+%\boldsymbol{\sigma}'[s]_b & \equiv & \boldsymbol{\sigma}_P[s]_b + g
+%\end{eqnarray}
+
+\section{Contract Creation} \label{ch:create}
+
+There are number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+
+We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas and accrued transaction substate $(\boldsymbol{\sigma}', g', A)$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of RLP encoding of the structure containing only the sender and the nonce.
 Thus we define the resultant address for the new account $a$:
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_n - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$ and $\boldsymbol{\sigma}[x]$ is the address state of $x$ or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
 
-The account's nonce is initially defined as one, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
+The account's nonce is initially defined as zero, the balance as the value passed, the storage as empty and the code hash as the Keccak 256-bit hash of the empty string; the sender's balance is also reduced by the value passed.
 Thus the mutated state becomes $\boldsymbol{\sigma}^*$:
@@
 \begin{equation}
-\boldsymbol{\sigma}_1[s] \equiv \begin{cases}
-\varnothing & \text{if}\ \boldsymbol{\sigma}_1'[s] = \varnothing \ \wedge\ v = 0 \\
-\mathbf{a}_1 &\text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
-\mathbf{a}_1 \equiv \left(\boldsymbol{\sigma}_1'[s]_{\mathrm{n}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{b}} - v, \boldsymbol{\sigma}_1'[s]_{\mathbf{s}}, \boldsymbol{\sigma}_1'[s]_{\mathrm{c}}\right)
+\boldsymbol{\sigma}_1'[s]_b \equiv \boldsymbol{\sigma}_1[s]_b - v
 \end{equation}
@@
 \end{cases} \\
-g' & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \ \wedge \\
-&\quad \mathbf{o} = \varnothing \\
-g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
-z & \equiv & \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
-1 & \text{otherwise}
-\end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
-I_{\mathrm{a}} & \equiv & r \\
-I_{\mathrm{o}} & \equiv & o \\
-I_{\mathrm{p}} & \equiv & p \\
-I_{\mathbf{d}} & \equiv & \mathbf{d} \\
-I_{\mathrm{s}} & \equiv & s \\
-I_{\mathrm{v}} & \equiv & \tilde{v} \\
-I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
-\end{eqnarray}
-\nopagebreak[1]where
-\begin{equation}
-\Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\begin{equation}
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
-
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+(\boldsymbol{\sigma}^{**}, g', \mathbf{s}, \mathbf{o}) & \equiv & \begin{cases}
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I) & \text{if} \quad r = 4 \\
+\Xi(\boldsymbol{\sigma}_1, g, I) & \text{otherwise} \end{cases} \\
+I_a & \equiv & r \\
+I_o & \equiv & o \\
+I_p & \equiv & p \\
+I_\mathbf{d} & \equiv & \mathbf{d} \\
+I_s & \equiv & s \\
+I_v & \equiv & v \\
+I_e & \equiv & e \\
+\text{Let} \; \mathtt{KEC}(I_\mathbf{b}) & = & \boldsymbol{\sigma}[c]_c
+\end{eqnarray}
+
+It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_\mathbf{b}), I_\mathbf{b})$ at some point prior in order to make the determination of $I_\mathbf{b}$ feasible.
+
+As can be seen, there are four exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are four so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The four contracts in addresses 1, 2, 3 and 4 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme and the identity function respectively.
 
@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
-\quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+( w \in \{ \text{\small JUMP}, \text{\small JUMPI} \} \quad \wedge \\ \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad
 \end{array}
 \end{equation}
-where
-\begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \\ 
-\begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
-\end{array}
-\end{equation}
-
-This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid, the new stack size would be larger than 1024 or state modification is attempted during a static call.
+
+This states that the execution is in an exceptional halting state if there is insufficient gas, if the instruction is invalid (and therefore its $\delta$ subscript is undefined), if there are insufficient stack items, if a {\small JUMP}/{\small JUMPI} destination is invalid or the new stack size would be larger then 1024.
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
@@
 \begin{equation}
-D_{J}(\mathbf{c}, i) \equiv \begin{cases}
+D_J(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
-\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+\{ i \} \cup D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
+D_J(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
-() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
-\varnothing \quad\quad\ \text{otherwise}&
-\end{cases}
-\end{equation}
-
-The data-returning halt operations, \hyperlink{RETURN}{\text{\small RETURN}} and \text{\small REVERT}, have a special function $H_{\text{\tiny RETURN}}$.
-Note also the difference between the empty sequence and the empty set as discussed \hyperlink{empty_sequence_vs_empty_set}{here}.
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
+() & \text{if} \quad w \in \{ \text{\small STOP}, \text{\small SUICIDE} \} \\
+\varnothing & \text{otherwise}
+\end{cases}
+\end{equation}
+
+The data-returning halt operation, \text{\small RETURN}, has a special function $H_{\text{\tiny RETURN}}$, defined in Appendix \ref{app:vm}.
 
@@
 
-%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_{\mathbf{T}}[i]$, and $g[i]$ the total gas used immediately after said transaction.
-
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+%where $s[i]$ equals the root of the state trie immediately after the execution of the transaction $B_\mathbf{T}[i]$, and $g[i]$ the total gas used immediately after said transaction.
+
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: ${B_H}_g$, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+{B_H}_g = \ell(\mathbf{R})_u
 \end{equation}
@@
 The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
+We raise the block's beneficiary account by $R_b$; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.
+Formally we define the function $\Omega$:
+\begin{eqnarray}
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
+\boldsymbol{\sigma}'[{B_H}_c]_b & = & \boldsymbol{\sigma}[{B_H}_c]_b + (1 + \frac{\lVert B_\mathbf{U}\rVert}{32})R_b \\
+\forall_{U \in B_\mathbf{U}}: \\ \nonumber
+ \boldsymbol{\sigma}'[U_c]_b & = & \boldsymbol{\sigma}[U_c]_b + (1 + \frac{1}{8} (U_i - {B_H}_i)) R_b 
 \end{eqnarray}
@@
 
-\hypertarget{Gamma}{}We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
+We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
 \begin{equation}
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
-
-\hypertarget{Phi}{}And finally we define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_H) = \varnothing \\
+\boldsymbol{\sigma}_i: \mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i)) = {P(B_H)_H}_r & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_S(\boldsymbol{\sigma}_i))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_i$; it is assumed that implementations will store this in the state database, trivial and efficient since the trie is by nature an immutable data structure.
+
+And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
 \begin{eqnarray}
 \Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
+B'_n & = & n: \quad x \leqslant \frac{2^{256}}{H_d} \\
+B'_m & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hcancel{n}}, n, \mathbf{d}) \\
+B^* & \equiv & B \quad \text{except:} \quad B^*_r = r(\Pi(\Gamma(B), B))
+\end{eqnarray}
 With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
 
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
-
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
-We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
-\boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
-\end{equation}
-
-In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
-\begin{equation}
-\mathbf{R}[n]_{\mathrm{u}} = \begin{cases} 0 & \text{if} \quad n < 0 \\
+As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined.
+
+As previously detailed, $\mathbf{R}[n]_{\boldsymbol{\sigma}}$, $\mathbf{R}[n]_\mathbf{l}$ and $\mathbf{R}[n]_u$ are the $n$th corresponding states, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_b$, the fourth component in the tuple, has already been defined in terms of the logs).
+The former is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
+\begin{equation}
+\mathbf{R}[n]_{\boldsymbol{\sigma}} = \begin{cases} \Gamma(B) & \text{if} \quad n < 0 \\ \Upsilon(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n]) & \text{otherwise} \end{cases}
+\end{equation}
+
+In the case of $B_\mathbf{R}[n]_u$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+\begin{equation}
+\mathbf{R}[n]_u = \begin{cases} 0 & \text{if} \quad n < 0 \\
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\Upsilon^g(\mathbf{R}[n - 1]_{\boldsymbol{\sigma}}, B_\mathbf{T}[n])\\ \quad + \mathbf{R}[n-1]_u
 \end{array}
@@
 For the purposes here, we assume we have well-defined standard cryptographic functions for RIPEMD-160 and SHA2-256 of the form:
-
-\begin{eqnarray}
-\mathtt{\small SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
-\mathtt{\small RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
-\end{eqnarray}
-
-The fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
+\begin{eqnarray}
+\mathtt{SHA256}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{32} \\
+\mathtt{RIPEMD160}(\mathbf{i} \in \mathbb{B}) & \equiv & o \in \mathbb{B}_{20}
+\end{eqnarray}
+
+Finally, the fourth contract, the identity function $\Xi_{\mathtt{ID}}$ simply defines the output as the input:
 \begin{eqnarray}
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o} &=& I_{\mathbf{d}}
-\end{eqnarray}
-
-The fifth contract performs arbitrary-precision exponentiation under modulo.
-Here, $0 ^ 0$ is taken to be one, and $x \bmod 0$ is zero for all $x$.
-The first word in the input specifies the number of bytes that the first non-negative integer $B$ occupies.
-The second word in the input specifies the number of bytes that the second non-negative integer $E$ occupies.
-The third word in the input specifies the number of bytes that the third non-negative integer $M$ occupies.
-These three words are followed by $B$, $E$ and $M$.
-The rest of the input is discarded.
-Whenever the input is too short, the missing bytes are considered to be zero.
-The output is encoded big-endian into the same format as $M$'s.
-
-\begin{eqnarray}
-\Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\[0.5em]
-\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
-\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+g_r &=& 15 + 3\Big\lceil \dfrac{|I_\mathbf{d}|}{32} \Big\rceil\\
+\mathbf{o} &=& I_\mathbf{d}
+\end{eqnarray}
+
+\section{Signing Transactions}\label{app:signing}
+
+The method of signing transactions is similar to the `Electrum style signatures'; it utilises the SECP-256k1 curve as described by \cite{gura2004comparing}.
+
+It is assumed that the sender has a valid private key $p_r$, a randomly selected positive integer
+% to avoid line break in range definition
+
+in the range $(1, \mathtt{secp256k1n} - 1)$ represented as a byte array of length 32 in big-endian form.
+
+We assert the functions $\mathtt{ECDSASIGN}$, $\mathtt{ECDSARESTORE}$ and $\mathtt{ECDSAPUBKEY}$. These are formally defined in the literature.
+\begin{eqnarray}
+\mathtt{ECDSAPUBKEY}(p_r \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64} \\
+\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_r \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_u \in \mathbb{B}_{64}
+\end{eqnarray}
+
+Where $p_u$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_r$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
+It is assumed that $v$ is the `recovery id', a 1 byte value specifying the sign and finiteness of the curve point; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
+
+We declare that a signature is invalid unless the following is true:
+\begin{eqnarray}
+0 < r < \mathtt{secp256k1n} \quad\wedge\quad \\
+0 < s < \mathtt{secp256k1n} \quad\wedge\quad \\
+v \in \{27,28\}
+\end{eqnarray}
+
+where:
+\begin{eqnarray}
+\mathtt{secp256k1n} &=& 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &=& 2^{256} - 2^{32} - 977\\
+\end{eqnarray}
+
+For a given private key, $p_r$, the Ethereum address $A(p_r)$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+\begin{equation}
+A(p_r) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_r) \big) \big)
+\end{equation}
+
+The message hash, $h(T)$, to be signed is the Keccak hash of the transaction without the latter three signature components, formally described as $T_r$, $T_s$ and $T_w$:
+\begin{eqnarray}
+L_S(T) & \equiv & \begin{cases}
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{i}) & \text{if} \; T_t = 0\\
+(T_n, T_p, T_g, T_t, T_v, T_\mathbf{d}) & \text{otherwise} 
 \end{cases} \\
-\mathbf o &=& \left(B^E\bmod M\right)\in\mathbb P_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
-i[x] &\equiv& \begin{cases}
-I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
-0 &\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-\subsection{zkSNARK Related Precompiled Contracts}
-
-We choose two numbers, both of which are prime.
-\begin{eqnarray}
-p &\equiv& 21888242871839275222246405745257275088696311157297823662689037894645226208583 \\
-q &\equiv& 21888242871839275222246405745257275088548364400416034343698204186575808495617
-\end{eqnarray}
-Since $p$ is a prime number, $\{0, 1, \ldots, p - 1\}$ forms a field with addition and multiplication modulo $p$. We call this field $F_{\mathrm{p}}$.
-
-We define a set~$C_1$ with
-\begin{equation}
-C_1\equiv\{(X,Y)\in F_{\mathrm{p}}\times F_{\mathrm{p}}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-addition}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ X_1\neq X_2\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
-\begin{equation}\label{eq:ec-scalar-multiplication}
-n\cdot P\equiv(0,0)+\underbrace{P+\cdots+P}_{n}
-\end{equation}
-for a natural number $n$ and a point $P$ in $C_1$.
-
-We define $P_1$ to be a point $(1,2)$ on $C_1$.
-Let $G_1$ be the subgroup of $(C_1,+)$ generated by $P_1$.
-$G_1$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_1$, we define $\log_{P_1}(P)$ to be the smallest natural number $n$ satisfying $n\cdot P_1=P$.
-$\log_{P_1}(P)$ is at most $q-1$.
-
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
-$(C_2,+)$ is also known to be a group.
-We define $P_2$ in $C_2$ with
-\begin{eqnarray}
-P_2&\equiv&
-(11559732032986387107991004021392285783925812861821192530917403151452391805634 \times i\\\nonumber &&+ 10857046999023057135944570762232829481370756359578518086990519993285655852781,\\\nonumber && 4082367875863433681332203403145435568316851327593401208105741076214120093531 \times i\\\nonumber &&+ 8495653923123431417604973247489272438418190587263600148770280649306958101930)
-\end{eqnarray}
-We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be a cyclic group of order $q$.
-For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
-With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
-\begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
-\varnothing&\text{otherwise}
-\end{cases}
-\end{equation}
-
-A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf x)&\equiv&\begin{cases}
-g_1&\text{if}\ g_1\in G_1\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_1&\equiv&\begin{cases}
-(x,y)&\text{if}\ x\neq\varnothing\wedge y\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf x)&\equiv&\begin{cases}
-g_2&\text{if}\ g_2\in G_2\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-g_2&\equiv&\begin{cases}
-((x_0i+y_0),(x_1i+y_1))&\text{if}\ x_0\neq\varnothing\wedge y_0\neq\varnothing\wedge x_1\neq\varnothing\wedge y_1\neq\varnothing\\
-\varnothing&\text{otherwise}
-\end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
-\begin{eqnarray}
-\Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf o[0..31]&\equiv&\begin{cases}
-0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
-0x0000000000000000000000000000000000000000000000000000000000000000&\text{if}\ \neg v\wedge\neg F
-\end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
-\vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
-\end{eqnarray}
-
-We define a precompiled contract for addition on $G_1$.
-
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf d}[0..63]\right)\\
-y&\equiv&\delta_1\left(\bar I_{\mathbf d}[64..127]\right)\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
-0&\text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
-\begin{eqnarray}
-\Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf d}[32..95]\right)
-\end{eqnarray}
-
-\section{Signing Transactions}\label{app:signing}
-
-The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
-This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
-
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
-\end{eqnarray}
-
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
-It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-
-\newcommand{\slimit}{\ensuremath{\text{s-limit}}}
-
-\linkdest{invalidsig}We declare that a signature is invalid unless all the following conditions are true:
-\begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
-\end{align}
-where:
-\begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
-\end{align}
-
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
-\begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
-\end{equation}
-
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
-\end{cases} \\
-\nonumber \text{where} \\
-\nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
-T_{\mathbf{d}} & \text{otherwise}
-\end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
-\end{eqnarray}
-
-The signed transaction $G(T, p_{\mathrm{r}})$ is defined as:
-\begin{eqnarray}
-G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
-\end{eqnarray}
-
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}: 
-\begin{eqnarray}
-\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
-\linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
-\linkdest{T__s}{T_{\mathrm{s}}} = \hyperlink{s}{s}
+h(T) & \equiv & \mathtt{KEC}( L_S(T) )
+\end{eqnarray}
+
+The signed transaction $G(T, p_r)$ is defined as:
+\begin{eqnarray}
+G(T, p_r) \equiv T \quad \text{except:} \\
+(T_w, T_r, T_s) = \mathtt{ECDSASIGN}(h(T), p_r)
 \end{eqnarray}
@@
 $G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
-$G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
+$G_{low}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{low}$}. \\
 $G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
-$G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{high}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{high}$}. \\
+$G_{ext}$ & 20 & Amount of gas to pay for operations of the set {\small $W_{ext}$}. \\
+$G_{sload}$ & 50 & Paid for a {\small SLOAD} operation. \\
 $G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
-&&is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
-&&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero. \\
+$R_{suicide}$ & 24000 & Refund given (added into refund counter) for suiciding an account. \\
 $G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 40 & Paid for a {\small CALL} operation. \\
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
 $G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{callnewaccount}$ & 25000 & Paid for a {\small CALL} operation to a not previously excisting account. \\
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 10 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
-\linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
-$G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
-$G_{\mathrm{logdata}}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
-$G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
+$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+$G_{transaction}$ & 21000 & Paid for every transaction. \\
+$G_{log}$ & 375 & Partial payment for a {\small LOG} operation. \\
+$G_{logdata}$ & 8 & Paid for each byte in a {\small LOG} operation's data. \\
+$G_{logtopic}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
 $G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
-&&contract. \\
-
-%extern u256 const c_{\mathrm{copyGas}};			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
+
+%extern u256 const c_copyGas;			///< Multiplied by the number of 32-byte words that are copied (round up) for any *COPY operation and added.
 \bottomrule
-\end{tabu}
+\end{tabular*}
 
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{memory}(\boldsymbol{\mu}'_i)-C_{memory}(\boldsymbol{\mu}_i) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
-&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
-C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
+G_{ext} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \\
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+C_{memory}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
+\end{equation}
+
+with $C_\text{\tiny CALL}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+
+$W_{zero}$ = \{{\small STOP}, {\small SUICIDE}, {\small RETURN}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 
-\begin{tabu}{r l r r l} \savetabu{opcodes}
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
 \multicolumn{5}{c}{\textbf{0s: Stop and Arithmetic Operations}} \\
-\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted. The zero-th power of zero $0^0$ is defined to be one.} \vspace{5pt} \\
+\multicolumn{5}{l}{All arithmetic is modulo $2^{256}$ unless otherwise noted.} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-\linkdest{stop}{}0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
+0x00 & {\small STOP} & 0 & 0 & Halts execution. \\
 \midrule
 0x01 & {\small ADD} & 2 & 1 & Addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x02 & {\small MUL} & 2 & 1 & Multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x03 & {\small SUB} & 2 & 1 & Subtraction operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] - \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] - \boldsymbol{\mu}_\mathbf{s}[1]$ \\
 \midrule
 0x04 & {\small DIV} & 2 & 1 & Integer division operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \lfloor\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \lfloor\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]\rfloor & \text{otherwise}\end{cases}$  \\
+&&&& where division is not subject to modulo $2^{256}$. \\
 \midrule
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]) \lfloor |\boldsymbol{\mu}_\mathbf{s}[0] \div \boldsymbol{\mu}_\mathbf{s}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& Where all values are treated as two's complement signed 256-bit integers and division is not subject to modulo $2^{256}.$ \\
 &&&& Note the overflow semantic when $-2^{255}$ is negated.\\
@@
 0x06 & {\small MOD} & 2 & 1 & Modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod \boldsymbol{\mu}_{\mathbf{s}}[1] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \boldsymbol{\mu}_\mathbf{s}[0] \bmod \boldsymbol{\mu}_\mathbf{s}[1] & \text{otherwise}\end{cases}$  \\
+&&&& where $\mod$ is not subject to modulo $2^{256}$. \\
 \midrule
 0x07 & {\small SMOD} & 2 & 1 & Signed modulo remainder operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0]) (|\boldsymbol{\mu}_{\mathbf{s}}[0]| \bmod |\boldsymbol{\mu}_{\mathbf{s}}[1]|) & \text{otherwise}\end{cases}$  \\
-&&&& Where all values are treated as two's complement signed 256-bit integers. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0\\ \mathbf{sgn} (\boldsymbol{\mu}_\mathbf{s}[0]) |\boldsymbol{\mu}_\mathbf{s}[0]| \bmod |\boldsymbol{\mu}_\mathbf{s}[1]| & \text{otherwise}\end{cases}$  \\
+&&&& Where all values are treated as two's complement signed 256-bit integers and $\mod$ is not subject to modulo $2^{256}$. \\
 \midrule
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] = 0\\ (\boldsymbol{\mu}_\mathbf{s}[0] \times \boldsymbol{\mu}_\mathbf{s}[1]) \mod \boldsymbol{\mu}_\mathbf{s}[2] & \text{otherwise}\end{cases}$  \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
 0x0a & {\small EXP} & 2 & 1 & Exponential operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0] ^ {\boldsymbol{\mu}_{\mathbf{s}}[1] }$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[0] ^ {\boldsymbol{\mu}_\mathbf{s}[1] }$ \\
 \midrule
 0x0b & {\small SIGNEXTEND} & 2 & 1 & Extend length of two's complement signed integer. \\
-&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
-\midrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+&&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_t &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_\mathbf{s}[0] + 1) \\ \boldsymbol{\mu}_\mathbf{s}[1]_i &\text{otherwise} \end{cases}$ \\
+\multicolumn{5}{l}{$\boldsymbol{\mu}_\mathbf{s}[x]_i$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_\mathbf{s}[x]$} \vspace{5pt} \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x10 & {\small LT} & 2 & 1 & Less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x11 & {\small GT} & 2 & 1 & Greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
-\midrule
-0x12 & {\small SLT} & 2 & 1 & Signed less-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] < \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x10 & {\small LT} & 2 & 1 & Less-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x11 & {\small GT} & 2 & 1 & Greater-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+\midrule
+0x12 & {\small SLT} & 2 & 1 & Signed less-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] < \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x13 & {\small SGT} & 2 & 1 & Signed greater-than comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
 \midrule
-0x14 & {\small EQ} & 2 & 1 & Equality comparison. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = \boldsymbol{\mu}_{\mathbf{s}}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
+0x14 & {\small EQ} & 2 & 1 & Equality comparision. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = \boldsymbol{\mu}_\mathbf{s}[1] \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x15 & {\small ISZERO} & 1 & 1 & Simple not operator. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x16 & {\small AND} & 2 & 1 & Bitwise AND operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \wedge \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x17 & {\small OR} & 2 & 1 & Bitwise OR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \vee \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \vee \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x18 & {\small XOR} & 2 & 1 & Bitwise XOR operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} \oplus \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \boldsymbol{\mu}_\mathbf{s}[0]_i \oplus \boldsymbol{\mu}_\mathbf{s}[1]_i$ \\
 \midrule
 0x19 & {\small NOT} & 1 & 1 & Bitwise NOT operation. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0]_{\mathrm{i}} = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} 1 & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0]_i = 0 \\ 0 & \text{otherwise} \end{cases}$ \\
 \midrule
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
-&&&& in big endian). \\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_\mathbf{s}[0]_i \equiv \begin{cases} \boldsymbol{\mu}_\mathbf{s}[1]_{(i + 8\boldsymbol{\mu}_\mathbf{s}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_\mathbf{s}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_a$ \\
 \midrule
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0]]_b& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
 0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
-&&&& This is the sender of original transaction; it is never an account with\\
-&&&& non-empty associated code. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_o$ \\
+&&&& This is the sender of original transaction; it is never an account with non-empty \\
+&&&& associated code. \\
 \midrule
 0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
-&&&& This is the address of the account that is directly responsible for\\
-&&&& this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
-&&&& this execution. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_s$ \\
+&&&& This is the address of the account that is directly responsible for this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_v$ \\
 \midrule
 0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
-&&&& environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_\mathbf{d}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots (\boldsymbol{\mu}_\mathbf{s}[0] + 31) ] \quad \text{with} \quad I_\mathbf{d}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_\mathbf{d} \rVert$ \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction\\
-&&&& or transaction. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{d}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{d} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
 0x38 & {\small CODESIZE} & 0 & 1 & Get size of code running in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{b}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert I_\mathbf{b} \rVert$ \\
 \midrule
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[2] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[0] + i ] \equiv
+\begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_\mathbf{s}[1] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] + i < \lVert I_\mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
 \midrule
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv I_p$ \\
 &&&& This is gas price specified by the originating transaction.\\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c \rVert$ \\
 \midrule
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
-&&&& environment. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
-\midrule
-0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
-&&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_\mathbf{s}[3] - 1\} } \boldsymbol{\mu}'_\mathbf{m}[\boldsymbol{\mu}_\mathbf{s}[1] + i ] \equiv
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_\mathbf{s}[2] + i] & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_s[0] \mod 2^{160}]_c$ \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv P(I_{H_p}, \boldsymbol{\mu}_\mathbf{s}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_i \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_i \\ P(H_p, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule
 0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_c$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_s$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_i$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_d$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv {I_H}_l$ \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x8f & {\small DUP16} & 16 & 17 & Duplicate 16th stack item. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[15]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[15]$ \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0x9f & {\small SWAP16} & 17 & 17 & Exchange 1st and 17th stack items. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{s}}[16]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[16] \equiv \boldsymbol{\mu}_{\mathbf{s}}[0]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv \boldsymbol{\mu}_\mathbf{s}[16]$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[16] \equiv \boldsymbol{\mu}_\mathbf{s}[0]$ \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0xa4 & {\small LOG4} & 6 & 0 & Append log record with four topics. \\
-&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4], \boldsymbol{\mu}_{\mathbf{s}}[5])$ \\
+&&&& $\mathbf{t} \equiv (\boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4], \boldsymbol{\mu}_\mathbf{s}[5])$ \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\ 
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
-&&&& address of the newly created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[1] \dots (\boldsymbol{\mu}_\mathbf{s}[1] + \boldsymbol{\mu}_\mathbf{s}[2] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_g, A^+) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_a, I_o, \boldsymbol{\mu}_g, I_p, \boldsymbol{\mu}_\mathbf{s}[0], \mathbf{i}, I_e + 1) & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[0] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_g, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_a]_n = \boldsymbol{\sigma}[I_a]_n + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which implies: $A'_\mathbf{s} \equiv A_\mathbf{s} \cup A^+_\mathbf{s} \quad \wedge \quad A'_\mathbf{l} \equiv A_\mathbf{l} \cdot A^+_\mathbf{l} \quad \wedge \quad A'_\mathbf{r} \equiv A_\mathbf{r} + A^+_\mathbf{r}$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}^*, \boldsymbol{\mu}, I) = \top$ or $I_e = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_\mathbf{s}[0] > \boldsymbol{\sigma}[I_a]_b$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_a, \boldsymbol{\sigma}[I_a]_n)$, the address of the newly \\
+&&&& created account, otherwise. \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[1], \boldsymbol{\mu}_\mathbf{s}[2])$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 0xf1 & {\small CALL} & 7 & 1 & Message-call into an account. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[3] \dots (\boldsymbol{\mu}_\mathbf{s}[3] + \boldsymbol{\mu}_\mathbf{s}[4] - 1) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_a, I_o, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_p, \boldsymbol{\mu}_\mathbf{s}[2], \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024\\ (\boldsymbol{\sigma}, 0, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_\mathbf{s}[6], |\mathbf{o}|\})$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[5] \dots (\boldsymbol{\mu}_\mathbf{s}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
+&&&& $\boldsymbol{\mu}'_g \equiv \boldsymbol{\mu}_g - C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) + g'$ \\
+&&&& $\boldsymbol{\mu}'_\mathbf{s}[0] \equiv x$ \\
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
+&&&& $t \equiv \boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an exceptional halting \\
+&&&& $Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) = \top$ or if  \\
+&&&& $\boldsymbol{\mu}_\mathbf{s}[2] > \boldsymbol{\sigma}[I_a]_b$ (not enough funds) or $I_e = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[3], \boldsymbol{\mu}_\mathbf{s}[4]), \boldsymbol{\mu}_\mathbf{s}[5], \boldsymbol{\mu}_\mathbf{s}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
-\end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
-\boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
-\end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + \boldsymbol{\mu}_\mathbf{s}[0] + C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) + C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
+&&&& $C_{\text{\tiny CALLXFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+&&&& $C_{\text{\tiny CALLNEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{callnewaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\
 0 & \text{otherwise}
 \end{cases}$ \\
-\midrule
-0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with an alternative account's code. \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}) \equiv  \begin{cases}
+\boldsymbol{\mu}_\mathbf{s}[0] + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\
+0 & \text{otherwise}
+\end{cases}$ \\
+\midrule
+0xf2 & {\small CALLCODE} & 7 & 1 & Message-call into this account with alternative account's code. \\
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
-&&&& fact the same account as at present, simply that the code is overwritten.\\
-\midrule
-\linkdest{RETURN}{}0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
-&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_a, I_o, I_a, t,\\\quad \boldsymbol{\mu}_\mathbf{s}[0], I_p, \boldsymbol{\mu}_\mathbf{s}[2], \mathbf{i}, I_e + 1)\end{array} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \leqslant \boldsymbol{\sigma}[I_a]_b \;\wedge\; I_e < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, \mathbf{o}) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_\mathbf{s}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_a$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overridden altered.\\
+\midrule
+0xf3 & {\small RETURN} & 2 & 0 & Halt execution returning output data. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_\mathbf{m}[ \boldsymbol{\mu}_\mathbf{s}[0] \dots ( \boldsymbol{\mu}_\mathbf{s}[0] + \boldsymbol{\mu}_\mathbf{s}[1] - 1 ) ]$ \\
 &&&& This has the effect of halting the execution at this point with output defined.\\
 &&&& See section \ref{ch:model}. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
-&&&& persisting the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
-&&&& omitted argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the\\
-&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and\\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
-&&&& This means that the recipient is in fact the same account as at present, simply\\
-&&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\midrule
-0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\
-&&&& Exactly equivalent to {\small CALL} except: \\
-&&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced \\
-&&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
-&&&& The effect of this operation is described in (\ref{eq:X-def}). \\
-&&&& For the gas calculation, we use the memory expansion function, \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\midrule
-0xfe & {\small INVALID} & $\varnothing$ & $\varnothing$ & Designated invalid instruction. \\
-\midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
-&&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
-&&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
-&&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+&&&& $\boldsymbol{\mu}'_i \equiv M(\boldsymbol{\mu}_i, \boldsymbol{\mu}_\mathbf{s}[0], \boldsymbol{\mu}_\mathbf{s}[1])$ \\
+\midrule
+0xff & {\small SUICIDE} & 1 & 0 & Halt execution and register account for later deletion. \\
+&&&& $A'_\mathbf{s} \equiv A_\mathbf{s} \cup \{ I_a \}$ \\
+&&&& $\boldsymbol{\sigma}'[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[0] \mod 2^{160}]_b + \boldsymbol{\sigma}[I_a]_b$ \\
+&&&& $\boldsymbol{\sigma}'[I_a]_b \equiv 0$ \\
 &&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
+R_{suicide} & \text{if} \quad I_a \notin A_\mathbf{s} \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
-0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
-\bottomrule
-\end{tabu}
-
-\section{Genesis Block}\label{app:genesis}\hypertarget{Genesis_Block}{}
+\end{cases}$ \\\bottomrule
+\end{tabular*}
+
+%\section{Low-level Lisp-like Language}\label{app:lll}
+%The Low-level Lisp-like Language is a language created in order to efficiently author low-level programs (contracts) without having to resort to EVM-Assembly.
+
+\section{Genesis Block}\label{app:genesis}
 
@@
 \midrule
-\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
-\linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
-\linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
-\linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
+$J_{wordbytes}$ & 4  & Bytes in word. \\
+$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+$J_{epoch}$ & 30000 & Blocks per epoch. \\
+$J_{mixbytes}$ & 128 & mix length in bytes. \\
+$J_{hashbytes}$ & 64 & Hash length in bytes. \\
+$J_{parents}$ & 256 & Number of parents of each dataset element. \\
+$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+$J_{accesses}$ & 64 & Number of accesses in hashimoto loop. \\
 \bottomrule
-\end{tabu*}
+\end{tabular*}
 
@@
 
-\section{Anomalies on the Main Network}
-
-\subsection{Deletion of an Account Despite Out-of-gas}
-
-At block 2675119, in the transaction \seqsplit{0xcf416c536ec1a19ed1fb89e4ec7ffb3cf73aa413b3aa9b77d60e4fd81a4296ba}, an account at address 0x03 was called and an out-of-gas occurred during the call.
-Against the equation (\ref{eq:pre}), this added 0x03 in the set of touched addresses, and this transaction turned $\boldsymbol{\sigma}[0x03]$ into $\varnothing$.
-
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{c c X[j]} \savetabu{col3}
-\toprule
-Symbol & Latex Command & Description \\
-\midrule
-\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
-Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
-\bottomrule
-\end{tabu*}
+\subsection{Mining}
+In order to mine a valid block, one repeatedly chooses a random nonce $\mathbf{n}_{rand} \in \mathbb{B}_{8}$ and calculates the $\mathtt{PoW}$ function until
+\begin{equation}
+ \mathtt{PoW}(H_{\hcancel{n}}, \mathbf{n}_{rand}, \mathbf{d})[1] \leqslant \frac{2^{256}}{H_d}\quad.
+\end{equation}
+$H_d$ being the current difficulty of the block.
 \end{document}

### 2018-03-09 Paper.tex
Document what the SNARKV precompiled contract is doing

--- 2018-03-06 Paper.tex
@@
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
- \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
+ (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
 \end{array}
@@
 \end{cases}\\
-\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
-X&\equiv&\lambda^2-X_1-X_2\\
-Y&\equiv&\lambda(X_1-X)-Y_1
+\nonumber \lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
+\nonumber X&\equiv&\lambda^2-X_1-X_2\\
+\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
 \end{eqnarray}
@@
 \end{cases}\\
-\lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
-X&\equiv&\lambda^2-2X_1\\
-Y&\equiv&\lambda(X_1-X)-Y_1
+\nonumber \lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
+\nonumber X&\equiv&\lambda^2-2X_1\\
+\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
 \end{eqnarray}
@@
 With this definition, $\log_{P_2}(P)$ is at most $q-1$.
+
+Let $G_T$ be the multiplicative abelian group underlying $F_{p^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
+This bilinear map is a type three pairing. The ZK-Snark verification contract takes a set of points $a_i$ that are in $G_1$, and a set of points $b_i$ that are in $G_2$. It checks that
+\begin{equation}
+\prod_{i=0}^{n}e\left(a_i, b_i\right) = 1
+\end{equation}
+If this is true, the contract returns $1$, otherwise it returns $0$, unless any of the points $a_i$ or $b_i$ are not on their respective curves, in which case the transaction is reverted.
 

### 2018-03-12 Paper.tex
merge definitions of the SNARKV contract

@@
 \end{cases} \\
-\mathbf o &=& \left(B^E\bmod M\right)\in\mathbb P_{8\ell_{M}} \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{P}_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\
@@
 
-Let $G_T$ be the multiplicative abelian group underlying $F_{p^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
-This bilinear map is a type three pairing. The ZK-Snark verification contract takes a set of points $a_i$ that are in $G_1$, and a set of points $b_i$ that are in $G_2$. It checks that
-\begin{equation}
-\prod_{i=0}^{n}e\left(a_i, b_i\right) = 1
-\end{equation}
-If this is true, the contract returns $1$, otherwise it returns $0$, unless any of the points $a_i$ or $b_i$ are not on their respective curves, in which case the transaction is reverted.
-
 A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
 \begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
+\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
+\mathbf{x}&\text{if}\ \mathbf{x}<p\\
 \varnothing&\text{otherwise}
@@
 \end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
 \end{eqnarray}
@@
 v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
+a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf{d}}[64..191])\\\nonumber
 \vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
+a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
+b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
 \end{eqnarray}
@@
 g_{\mathrm{r}} &=& 500\\
-\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf d}[0..63]\right)\\
-y&\equiv&\delta_1\left(\bar I_{\mathbf d}[64..127]\right)\\
-\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
-I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
+\mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
+y&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[64..127]\right)\\
+\label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
+I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
 0&\text{otherwise}
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf{d}}$ is defined in (\ref{eq:complemented_input}).
 \begin{eqnarray}
@@
 g_{\mathrm{r}} &=& 40000\\
-\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf d}[0..31]\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf d}[32..95]\right)
+\mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf{d}}[0..31]\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[32..95]\right)
 \end{eqnarray}

### 2018-03-13 Paper.tex
\linkdest{block_nonce}

--- 2018-03-09 Paper.tex
@@
 
-\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
-
-\input{Version.tex}
 % Default rendering options
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
 \item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, &\\ \quad\quad T_{\mathrm{t}}, T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \mathbf{t} & \equiv & \{s, r\} \\
-\end{eqnarray}
-\nopagebreak[1]where
-\begin{equation}
-\Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\begin{equation}
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
+\\ \nonumber
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) & = & \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{eqnarray}
 
@@
 ( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
- (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024
+ \neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu})
 \end{array}
@@
 \begin{equation}
-W(w, \boldsymbol{\mu}) \equiv \\ 
-\begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+W(w, \boldsymbol{\mu}) \equiv \begin{array}[t]{l}
+w \in \{\text{\small CREATE}, \, \text{\small SSTORE}, \\
+\text{\small SELFDESTRUCT}\} \quad \vee \\
+(\text{\small LOG0} \le w \wedge w \le \text{\small LOG4}) \quad \vee \\
+(w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0)
 \end{array}
@@
 H(\boldsymbol{\mu}, I) \equiv \begin{cases}
-H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \ \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} &\\
-() \quad\quad\ \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}} \} &\\
-\varnothing \quad\quad\ \text{otherwise}&
+H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w \in \{\text{\small \hyperlink{RETURN}{RETURN}}, \text{\small REVERT}\} \\
+() & \text{if} \quad w \in \{ \text{\small \hyperlink{stop}{STOP}}, \text{\small \hyperlink{selfdestruct}{SELFDESTRUCT}}\} \\
+\varnothing & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}
@@
 
-As previously detailed, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
+\hyperlink{Transaction_Receipt}{As previously detailed}, $\mathbf{R}[n]_{\mathrm{z}}$, $\mathbf{R}[n]_{\mathbf{l}}$ and $\mathbf{R}[n]_{\mathrm{u}}$ are the $n$th corresponding status code, logs and cumulative gas used after each transaction ($\mathbf{R}[n]_{\mathrm{b}}$, the fourth component in the tuple, has already been defined in terms of the logs).
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
@@
 
-In the case of $B_{\mathbf{R}}[n]_{\mathrm{u}}$, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
+In the case of \hyperlink{Transaction_Receipt}{$B_{\mathbf{R}}[n]_{\mathrm{u}}$}, we take a similar approach defining each item as the gas used in evaluating the corresponding transaction summed with the previous item (or zero, if it is the first), giving us a running total:
 \begin{equation}
@@
 \begin{array}[b]{l}
-\Upsilon^g(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
+\linkdest{Upsilon_pow_g2}\hyperlink{Upsilon_pow_g}{\Upsilon^{\mathrm{g}}}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n])\\ \quad + \mathbf{R}[n-1]_{\mathrm{u}}
 \end{array}
@@
 \end{cases}\\
-\nonumber \lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}\\
-\nonumber X&\equiv&\lambda^2-X_1-X_2\\
-\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
-\end{eqnarray}
-
-In the case where $(X_1, Y_1) = (X_2, Y_2)$, we define $+$ on $C_1$ with
-\begin{eqnarray}\label{eq:ec-doubling}
-(X_1, Y_1) + (X_1, Y_1)&\equiv&\begin{cases}
-(X,Y)&\text{if}\ Y_1\neq 0\\
-(0,0)&\text{otherwise}
-\end{cases}\\
-\nonumber \lambda&\equiv&\frac{3X_1^2}{2Y_1}\\
-\nonumber X&\equiv&\lambda^2-2X_1\\
-\nonumber Y&\equiv&\lambda(X_1-X)-Y_1
-\end{eqnarray}
-
-$(C_1,+)$ is known to form a group. We define scalar multiplication $\cdot$ with
+X&\equiv&\lambda^2-X_1-X_2\\
+Y&\equiv&\lambda(X_1-X)-Y_1\\
+\lambda&\equiv&\frac{Y_2-Y_1}{X_2-X_1}
+\end{eqnarray}
+
+$(C_1,+)$ is known to form a group. We define the scalar multiplication $\cdot$ with
 \begin{equation}\label{eq:ec-scalar-multiplication}
@@
 
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i^2+1)$. We define a set $C_2$ with
-\begin{equation}
-C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3(i+9)^{-1}\}\cup\{(0,0)\}
-\end{equation}
-We define a binary operation $+$ and scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}), (\ref{eq:ec-doubling}) and (\ref{eq:ec-scalar-multiplication}).
+Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i+1)$. We define a set $C_2$ with
+\begin{equation}
+C_2\equiv\{(X,Y)\in F_{p^2}\times F_{p^2}\mid Y^2=X^3+3\}\cup\{(0,0)\}
+\end{equation}
+We define a binary operation $+$ and a scalar multiplication $\cdot$ with the same equations (\ref{eq:ec-addition}) and (\ref{eq:ec-scalar-multiplication}).
 $(C_2,+)$ is also known to be a group.
@@
 We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be the only cyclic group of order $q$ on $C_2$.
+$G_2$ is known to be a cyclic group of order $q$.
 For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
 With this definition, $\log_{P_2}(P)$ is at most $q-1$.
-
-Let $G_T$ be the multiplicative abelian group underlying $F_{p^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
-This bilinear map is a type three pairing. The ZK-Snark verification contract takes a set of points $a_i$ that are in $G_1$, and a set of points $b_i$ that are in $G_2$. It checks that
-\begin{equation}
-\prod_{i=0}^{n}e\left(a_i, b_i\right) = 1
-\end{equation}
-If this is true, the contract returns $1$, otherwise it returns $0$, unless any of the points $a_i$ or $b_i$ are not on their respective curves, in which case the transaction is reverted.
 
@@
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
-&&is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
-&&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from non-zero.
+\linkdest{R__selfdestruct}{}\\$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
 \linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
-&&contract. \\
+$G_{blockhash}$ & 20 & Payment for {\small \hyperlink{blockhash}{BLOCKHASH}} operation. \\
+$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponation-over-modulo precompiled contract. \\
 
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
-&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
-&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
-&&&& modulo. \\
+&&&& All intermediate calculations of this operation are not subject to the $2^{256}$ modulo. \\
 \midrule
@@
 &&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
-&&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
-&&&& in big endian). \\
+&&&& For Nth byte, we count from the left (i.e. N=0 would be the most significant in big endian). \\
 \bottomrule
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
 \toprule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{o}}$ \\
-&&&& This is the sender of original transaction; it is never an account with\\
-&&&& non-empty associated code. \\
+&&&& This is the sender of original transaction; it is never an account with non-empty \\
+&&&& associated code. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{s}}$ \\
-&&&& This is the address of the account that is directly responsible for\\
-&&&& this execution. \\
-\midrule
-0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for\\
-&&&& this execution. \\
+&&&& This is the address of the account that is directly responsible for this execution. \\
+\midrule
+0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{v}}$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathbf{d}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \quad \text{with} \quad I_{\mathbf{d}}[x] = 0 \quad \text{if} \quad x \geqslant \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
-\midrule
-0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current\\
-&&&& environment. \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
+\midrule
+0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert I_{\mathbf{d}} \rVert$ \\
-&&&& This pertains to the input data passed with the message call\\
-&&&& instruction or transaction. \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& This pertains to the input data passed with the message call instruction\\
-&&&& or transaction. \\
+&&&& This pertains to the input data passed with the message call instruction or transaction. \\
 \midrule
@@
 \midrule
-0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current\\
-&&&& environment. \\
+0x3d & {\small RETURNDATASIZE} & 0 & 1 & Get size of output data from the previous call from the current environment. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
-&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
+0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
+&&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H(h)_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H(h)_{\mathrm{i}} \\ P(H(h)_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
+&&&& and we assert the header $H(h)$ can be determined from its hash~$h$ unless $h$ is zero. \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge \quad A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}}$ \\
+&&&& $ \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\ 
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
-&&&& address of the newly created account. \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting}\\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too \\
+&&&& low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the address of the newly \\
+&&&& created account. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t,\\ \quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& where $x=0$ if the code execution for this operation failed due to an \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} \\
+&&&& (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
-&&&& fact the same account as at present, simply that the code is overwritten.\\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value $\boldsymbol{\mu}_{\mathbf{s}}[1]$\\
+&&&& (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in fact the\\
+&&&& same account as at present, simply that the code is overwritten.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
-\midrule
-0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but\\
-&&&& persisting the current values for {\it sender} and {\it value}. \\
-&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments. The\\
-&&&& omitted argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
-As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the\\
-&&&& definition of {\small CALL} should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and\\
-&&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+\end{tabular*}
+
+\begin{tabular*}{\columnwidth}[h]{rlrrl}
+\midrule
+0xf4 & {\small DELEGATECALL} & 6 & 1 & Message-call into this account with an alternative account's code, but persisting\\
+&&&& the current values for {\it sender} and {\it value}. \\
+&&&& Compared with {\small CALL}, {\small DELEGATECALL} takes one fewer arguments.  The omitted\\
+&&&& argument is $\boldsymbol{\mu}_{\mathbf{s}}[2]$.
+As a result, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ in the definition of {\small CALL} \\
+&&&& should respectively be replaced with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& Otherwise exactly equivalent to {\small CALL} except: \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\; I_{\mathrm{e}} < 1024 \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& The argument $\boldsymbol{\mu}_{\mathbf{s}}[2]$ is replaced with $0$. \\
-&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced \\
-&&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
+&&&& The deeper argument $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$, $\boldsymbol{\mu}_{\mathbf{s}}[5]$ and $\boldsymbol{\mu}_{\mathbf{s}}[6]$ are respectively replaced with \\
+&&&& $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
 &&&& The last argument of $\Theta$ is $\bot$. \\
@@
 \midrule
-\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{Jwordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
 \linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
@@
 
-\section{List of mathematical symbols}\label{app:symbols}
-\begin{tabu*}{c c X[j]} \savetabu{col3}
+\section{List of mathematical symbols} \label{app:symbols}
+\begin{tabu*} spread 0pt {X[l] X[c] X} \savetabu{col3}
 \toprule
-Symbol & Latex Command & Description \\
-\midrule
-\linkdest{bigvee}$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on.
-Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
+Symbol & Latex Command & Description \hypertarget{bigvee}{}\\
+\midrule
+$\bigvee$ & \verb|\bigvee| & This is the least upper bound, supremum, or join of all elements operated on. Thus it is the greatest element of such elements (\cite{Davey2002_zbMATH01748069}).\\
 \bottomrule

### 2018-03-15 Paper.tex
a_k and b_k should be all italic when k is an index
Make ZkSnark definitions match

--- 2018-03-09 Paper.tex
@@
 
-Let $G_T$ be the multiplicative abelian group underlying $F_{p^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
-This bilinear map is a type three pairing. The ZK-Snark verification contract takes a set of points $a_i$ that are in $G_1$, and a set of points $b_i$ that are in $G_2$. It checks that
-\begin{equation}
-\prod_{i=0}^{n}e\left(a_i, b_i\right) = 1
-\end{equation}
-If this is true, the contract returns $1$, otherwise it returns $0$, unless any of the points $a_i$ or $b_i$ are not on their respective curves, in which case the transaction is reverted.
+Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
+This bilinear map is a type three pairing.
+Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points that are in $G_2$.
+It follows from the definition of a pairing that the following are equivalent
+\begin{eqnarray} \label{eq:pairing-check}
+\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
+\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
+\end{eqnarray}
+Thus the pairing check provides a method to verify \ref{eq:pairing-check}.
 
@@
 
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if \ref{eq:pairing-check} holds, for intended use in zkSNARK verification.
+
 \begin{eqnarray}

### 2018-03-15 Paper.tex
Add parenthesis to ref's

@@
 
-Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
+Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$.
+It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
 This bilinear map is a type three pairing.
-Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points that are in $G_2$.
-It follows from the definition of a pairing that the following are equivalent
+There are several such bilinear maps, it does not matter which is chosen to be $e$.
+Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
 \begin{eqnarray} \label{eq:pairing-check}
@@
 \end{eqnarray}
-Thus the pairing check provides a method to verify \ref{eq:pairing-check}.
+Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).
 
@@
 
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if \ref{eq:pairing-check} holds, for intended use in zkSNARK verification.
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.
 

### 2018-03-16 Paper.tex

--- 2018-03-15 Paper.tex
@@
 
-Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$. It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
-This bilinear map is a type three pairing.
-Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points that are in $G_2$.
-It follows from the definition of a pairing that the following are equivalent
-\begin{eqnarray} \label{eq:pairing-check}
-\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
-\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
-\end{eqnarray}
-Thus the pairing check provides a method to verify \ref{eq:pairing-check}.
-
 A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
 \begin{equation}
-\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
-\mathbf x&\text{if}\ \mathbf x<p\\
+\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
+\mathbf{x}&\text{if}\ \mathbf{x}<p\\
 \varnothing&\text{otherwise}
@@
 \end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if \ref{eq:pairing-check} holds, for intended use in zkSNARK verification.
-
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
+\end{eqnarray}
+
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
 \begin{eqnarray}
@@
 0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
+\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}

### 2018-03-16 Paper.tex

@@
 
+Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$.
+It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
+This bilinear map is a type three pairing.
+There are several such bilinear maps, it does not matter which is chosen to be $e$.
+Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
+\begin{eqnarray} \label{eq:pairing-check}
+\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
+\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
+\end{eqnarray}
+Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).
+
 A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
@@
 
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.
+
 \begin{eqnarray}

### 2018-03-20 Paper.tex
Fix typo ERCEC to ECREC

@@
 
-We define $\Xi_{\mathtt{ERCEC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
+We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
 See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
@@
 \end{cases}\\
-v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
+v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_k)\times\log_{P_2}(b_k)\equiv 1\mod q)\\
 a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
@@
 \vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
+a_k&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
 \end{eqnarray}

### 2018-03-24 Paper.tex
Correct to 2^64

--- 2018-03-16 Paper.tex
@@
 This document describes one version of the protocol.
-In order to follow back the history of a path, one must reference multiple versions of this document.
+In order to follow back the history of a path, one might need to reference multiple versions of this document.
 
@@
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{P}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{P}_{256}$.
 This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
@@
 \begin{equation}
-k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
+k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{P}
 \end{equation}
@@
 \begin{equation}
-\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
+\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{P}_{256} \wedge x_{\mathrm{b}} \in \mathbb{P}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
 \end{equation}
@@
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{n}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{P}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{P}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{P}_{256} & \wedge \\
+T_{\mathrm{s}} \in \mathbb{P}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 \begin{equation}
-\mathbb{N}_{\mathrm{n}} = \{ P: P \in \mathbb{N} \wedge P < 2^n \}
+\mathbb{P}_{\mathrm{n}} = \{ P: P \in \mathbb{P} \wedge P < 2^n \}
 \end{equation}
@@
 \begin{equation}
-R_{\mathrm{z}} \in \mathbb{N}
+R_{\mathrm{z}} \in \mathbb{P}
 \end{equation}
@@
 \begin{equation}
-R_{\mathrm{u}} \in \mathbb{N} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
+R_{\mathrm{u}} \in \mathbb{P} \quad \wedge \quad R_{\mathrm{b}} \in \mathbb{B}_{256}
 \end{equation}
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: p(i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
-\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
-\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
+\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{P} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{P} & \wedge \\
+\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{P} & \wedge & H_{\mathrm{g}} \in \mathbb{P} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{P}_{256} & \wedge \\
 \hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
@@
 \begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\homesteadmod \equiv \text{max}\left( x - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
 \begin{equation*}
-y \equiv \begin{cases}
+x \equiv \begin{cases}
 1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^\mathbf{l}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{P}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 \mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ...
+) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \; \wedge \; \lVert \mathbf{t} \rVert < 2^{256} \} \\
 \mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
@@
 
-Where $\mathbb{O}$ is the set of bytes.
+where $\mathbb{O}$ is the set of bytes, and an additional restriction is placed on $\mathbb{L}$ as discussed below.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
@@
 \item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\item If the concatenated serialisations of each contained item is less than $2^{64}$ bytes in length, then the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\item Otherwise, this sequence cannot be encoded.
+
 \end{itemize}
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{P}$ or any $x$ for $\mathbb{P}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some positive integer $i$ is defined as:
 \begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{N}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\mathtt{\tiny RLP}(i : i \in \mathbb{P}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
 \end{equation}
@@
 \end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{P}_{8\ell_{M}} \\
+\mathbf o &=& (B^E\bmod M)\in\mathbb P_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\
@@
 i[x] &\equiv& \begin{cases}
-I_{\mathbf{d}}[x] &\text{if}\ x < |I_{\mathbf{d}}| \\
+I_{\mathbf d}[x] &\text{if}\ x < |I_{\mathbf d}| \\
 0 &\text{otherwise}
@@
 We define $G_2$ to be the subgroup of $(C_2,+)$ generated by $P_2$.
-$G_2$ is known to be the only cyclic group of order $q$ on $C_2$.
+$G_2$ is known to be a cyclic group of order $q$.
 For a point $P$ in $G_2$, we define $\log_{P_2}(P)$ be the smallest natural number $n$ satisfying $n\cdot P_2=P$.
@@
 
-Let $G_T$ be the multiplicative abelian group underlying $F_{q^{12}}$.
-It is known that a non-degenerate bilinear map $e : G_1\times G_2 \to G_T$ exists.
-This bilinear map is a type three pairing.
-There are several such bilinear maps, it does not matter which is chosen to be $e$.
-Let $P_T = e(P_1, P_2)$, $a$ be a set of $k$ points in $G_1$, and $b$ be a set of $k$ points in $G_2$. It follows from the definition of a pairing that the following are equivalent
-\begin{eqnarray} \label{eq:pairing-check}
-\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{k})\times\log_{P_2}(b_{k})&\equiv& 1\mod q\\
-\prod_{i=0}^{k}e\left(a_i, b_i\right) &=& P_T
-\end{eqnarray}
-Thus the pairing operation provides a method to verify (\ref{eq:pairing-check}).
-
 A 32 byte number $\mathbf{x}\in\mathbf{P}_{256}$ might and might not represent an element of $F_{\mathrm{p}}$.
 \begin{equation}
-\delta_{\mathrm{p}}(\mathbf{x})\equiv\begin{cases}
-\mathbf{x}&\text{if}\ \mathbf{x}<p\\
+\delta_{\mathrm{p}}(\mathbf x)\equiv\begin{cases}
+\mathbf x&\text{if}\ \mathbf x<p\\
 \varnothing&\text{otherwise}
@@
 
-A 64 byte data $\mathbf{x}\in\mathbf{B}_{512}$ might and might not represent an element of $G_1$.
-\begin{eqnarray}
-\delta_1(\mathbf{x})&\equiv&\begin{cases}
+A 64 byte data $\mathbf x\in\mathbf B_{512}$ might and might not represent an element of $G_1$.
+\begin{eqnarray}
+\delta_1(\mathbf x)&\equiv&\begin{cases}
 g_1&\text{if}\ g_1\in G_1\\
@@
 \end{cases}\\
-x&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
-y&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])
-\end{eqnarray}
-
-A 128 byte data $\mathbf{x}\in\mathbf{B}_{1024}$ might and might not represent an element of $G_2$.
-\begin{eqnarray}
-\delta_2(\mathbf{x})&\equiv&\begin{cases}
+x&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])
+\end{eqnarray}
+
+A 128 byte data $\mathbf x\in\mathbf B_{1024}$ might and might not represent an element of $G_2$.
+\begin{eqnarray}
+\delta_2(\mathbf x)&\equiv&\begin{cases}
 g_2&\text{if}\ g_2\in G_2\\
@@
 \end{cases}\\
-x_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[0..31])\\
-y_0&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[32..63])\\
-x_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[64..95])\\
-y_1&\equiv&\delta_{\mathrm{p}}(\mathbf{x}[96..127])
-\end{eqnarray}
-
-We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract which checks if (\ref{eq:pairing-check}) holds, for intended use in zkSNARK verification.
-
+x_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[0..31])\\
+y_0&\equiv&\delta_{\mathrm{p}}(\mathbf x[32..63])\\
+x_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[64..95])\\
+y_1&\equiv&\delta_{\mathrm{p}}(\mathbf x[96..127])
+\end{eqnarray}
+
+We define $\Xi_{\mathtt{SNARKV}}$ as a precompiled contract for zkSNARK verification.
 \begin{eqnarray}
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ F\\
 F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
@@
 g_{\mathrm{r}}&=& 60000k + 40000 \\
-\mathbf{o}[0..31]&\equiv&\begin{cases}
+\mathbf o[0..31]&\equiv&\begin{cases}
 0x0000000000000000000000000000000000000000000000000000000000000001&\text{if}\ v\wedge\neg F\\
@@
 v&\equiv&(\log_{P_1}(a_1)\times\log_{P_2}(b_1)+\cdots+\log_{P_1}(a_{\mathrm{k}})\times\log_{P_2}(b_{\mathrm{k}})\equiv 1\mod q)\\
-a_1&\equiv&\delta_1(I_{\mathbf{d}}[0..63])\\
-b_1&\equiv&\delta_2(I_{\mathbf{d}}[64..191])\\\nonumber
+a_1&\equiv&\delta_1(I_{\mathbf d}[0..63])\\
+b_1&\equiv&\delta_2(I_{\mathbf d}[64..191])\\\nonumber
 \vdots\\
-a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
-b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
+a_{\mathrm{k}}&\equiv&\delta_1(I_{\mathbf d}[(|I_{\mathbf d}|-192)..(|I_{\mathbf d}|-129)])\\
+b_{\mathrm{k}}&\equiv&\delta_2(I_{\mathbf d}[(|I_{\mathbf d}|-128)..(|I_{\mathbf d}|-1)])
 \end{eqnarray}
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 500\\
-\mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
-y&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[64..127]\right)\\
-\label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
-I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
+\mathbf o&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[0..63])\\
+y&\equiv&\delta_1(\bar I_{\mathbf d}[64..127])\\
+\label{eq:complemented_input}\bar I_{\mathbf d}[x]&\equiv&\begin{cases}
+I_{\mathbf d}[x]&\text{if}\ x < |I_{\mathbf d}|\\
 0&\text{otherwise}
@@
 
-We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf{d}}$ is defined in (\ref{eq:complemented_input}).
+We define a precompiled contract for scalar multiplication on $G_1$, where $\bar I_{\mathbf d}$ is defined in (\ref{eq:complemented_input}).
 \begin{eqnarray}
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&(\varnothing,0,A^0,())\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 40000\\
-\mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf{d}}[0..31]\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[32..95]\right)
+\mathbf o&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
+n&\equiv&\bar I_{\mathbf d}[0..31]\\
+x&\equiv&\delta_1(\bar I_{\mathbf d}[32..95])
 \end{eqnarray}
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$) and $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range).
 It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
@@
 0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 35, \mathtt{\tiny chain\_{\mathrm{id}}} \times 2 + 36\}
 \end{align}
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_{\mathrm{id}}}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
-&&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
+0x40 & \linkdest{blockhash}{}{\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+&&&& where $P$ is the hash of a block of a particular number, up to a maximum age.\\
+&&&& 0 is left on the stack if the looked for block number is greater than the current block number \\
+&&&& or more than 256 blocks behind the current block. \\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
+&&&& and we assert the header $H$ can be determined as its hash is the parent hash \\
+&&&& in the block following it. \\
 \midrule
@@
  E_{\mathrm{prime}}(x, y) = \begin{cases}
-x & \text{if} \quad x / y \in \mathbb{N} \\
+x & \text{if} \quad x / y \in \mathbb{P} \\
 E_{\mathrm{prime}}(x - 2 \cdot y, y) & \text{otherwise}

### 2018-03-26 Paper.tex
Add bitwise shifting instructions EIP-145
https://github.com/ethereum/EIPs/blob/f6a0ec6c37b878b8f063b54a44c60013007f5b99/EIPS/eip-145.md

--- 2018-03-20 Paper.tex
@@
 &&&& $ \forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{t}} &\text{if} \quad i \leqslant t \quad \text{where} \; t = 256 - 8(\boldsymbol{\mu}_{\mathbf{s}}[0] + 1) \\ \boldsymbol{\mu}_{\mathbf{s}}[1]_{\mathrm{i}} &\text{otherwise} \end{cases}$ \\
-\multicolumn{5}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt} \\
+& \multicolumn{4}{l}{$\boldsymbol{\mu}_{\mathbf{s}}[x]_{\mathrm{i}}$ gives the $i$th bit (counting from zero) of $\boldsymbol{\mu}_{\mathbf{s}}[x]$} \vspace{5pt}  \\
+\midrule
+0x1b & {\small SHL} & 2 & 1 & Bitwise shifting to the left. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 2 ^ {\boldsymbol{\mu}_{\mathbf{s}}[0]} \times \boldsymbol{\mu}_{\mathbf{s}}[1].$ \\
+\midrule
+0x1c & {\small SHR} & 2 & 1 & Bitwise logical shifting to the right. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \left\lfloor \dfrac{ \boldsymbol{\mu}_{\mathbf{s}}[1]}{ 2 ^ {\boldsymbol{\mu}_{\mathbf{s}}[0]} } \right\rfloor$
+\\
+\midrule
+0x1d & {\small SAR} & 2 & 1 & Bitwise arithmetic shifting to the right. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \left\lfloor \dfrac{ \boldsymbol{\mu}_{\mathbf{s}}[1]}{ 2 ^ {\boldsymbol{\mu}_{\mathbf{s}}[0]} } \right\rfloor$ \\
+& \multicolumn{4}{l}{where $\boldsymbol{\mu}_{\mathbf{s}}[1]$ is treated as two's complement signed 256-bit integer.}  \\
 \midrule

### 2018-04-03 Paper.tex
Fix typo formula no.235
Add explanation of recovery identifier for EIP155 signatures
This commit adds explanation of how to recover the public key
from the signature in the case of signed EIP155 transaction. Previously,
only the pre-EIP155 signatures were explained.
Change "2nd" to second in EIP155 sign. explanation
Requested change since this is more consistent with the rest of the yellow paper.

--- 2018-03-20 Paper.tex
@@
 \begin{eqnarray}\label{eq:ec-doubling}
-(X_1, Y_1) + (X_1, Y_1)&\equiv&\begin{cases}
+(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
 (X,Y)&\text{if}\ Y_1\neq 0\\
@@
 The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 

### 2018-04-04 Paper.tex
Remove a mathrm on a variable
as pointed out by @DanielRX
capitalize h H_cancel_n

--- 2018-03-24 Paper.tex
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit hash which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 \mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ...
-) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \; \wedge \; \lVert \mathbf{t} \rVert < 2^{256} \} \\
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
 \mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
@@
 
-where $\mathbb{O}$ is the set of bytes, and an additional restriction is placed on $\mathbb{L}$ as discussed below.
+Where $\mathbb{O}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
@@
 \item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item If the concatenated serialisations of each contained item is less than $2^{64}$ bytes in length, then the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
-\item Otherwise, this sequence cannot be encoded.
-
+\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
 \end{itemize}

### 2018-04-04 Paper.tex

--- 2018-04-03 Paper.tex
@@
 
-Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i^2+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_{p}[i]/(i^2+1)$. We define a set $C_2$ with
 \begin{equation}

### 2018-04-05 Paper.tex
Remove dotted lines from the gas value table

@@
 \item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
 \end{description}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
 $G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\

### 2018-04-13 Paper.tex
Correct transaction execution gas used equation
This commit seeks to address a disconnect between the Yellow Paper
specification and what is implemented in the Ethereum clients in
practice for calculating the correct transaction receipt field.

Currently the Yellow Paper specifies that the total gas by a
transaction is the difference between the transaction gas limit
and the gas remaining from the root-level message call or contract
creation within the transaction.

This neglects the fact that the transaction sender can be refunded
additional gas after the root-level message call or contract c

--- 2018-04-05 Paper.tex
@@
 
-After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
+After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate\footnote{Note that the total refundable gas equation prevents the gas refund from exceeding the transaction gas limit $T_{\mathrm{g}}$.}.
 \begin{equation}
@@
 The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
 
@@
 
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
 \Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\

### 2018-04-13 Paper.tex
Correct self-destruct refund semantics
The current Yellow Paper specification for the self-destruct refunds can
lead to double refunds and is not consistent with the logic clients use
to fulfill refunds in practice.

The crux of the issue is that the current Yellow Paper performs the
self-destruct refund within the SELFDESTRUCT operation of a specific EVM
invocation and each EVM invocation creates its own transaction substate
(rather than having it passed in from its parent). Therefore, although
the transaction substate is checked to see if the account has already
been included in its self-des

@@
 
-After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate\footnote{Note that the total refundable gas equation prevents the gas refund from exceeding the transaction gas limit $T_{\mathrm{g}}$.}.
-\begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \right\}
+After the message call or contract creation is processed, the refund counter has to be incremented for the account that were self-destructed throughout its invocation.
+\begin{equation}
+	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
+\end{equation}
+
+Then the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
+\begin{equation}
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
 \end{equation}
@@
 (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
+\end{cases}$\\ \\
+&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\ \\
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&& $A'_{r} \equiv A_{r} + \begin{cases}
-R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
-0 & \text{otherwise}
-\end{cases}$ \\
 &&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}

### 2018-04-18 Paper.tex

@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
 

### 2018-04-23 Paper.tex
Seed hash is just zero for epoch 0 (#682)
Fixes https://github.com/ethereum/yellowpaper/issues/312

@@
 Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
-It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
+It is assumed that \hypertarget{v}{}$v$ is the `recovery identifier'.
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
 The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
-In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 
@@
 
-\linkdest{invalidsig}We declare that a signature is invalid unless all the following conditions are true:
+\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true\footnote{A signature of a transaction can be valid not only with a recovery identifier but with some other numbers.
+See \hyperlink{T__w}{how the component $T_w$ of a transaction is interpreted.}}:
 \begin{align}
@@
 0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
+\hyperlink{v}{v} &\in \{27,28\}
 \end{align}
@@
 \begin{eqnarray}
-\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
 \linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
@@
 \end{eqnarray}
+$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.
+In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 
@@
  C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
-\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
+\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
 \texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}

### 2018-04-26 Paper.tex
Correct wording in a couple of places

@@
 
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
 Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
@@
 
-After the message call or contract creation is processed, the refund counter has to be incremented for the account that were self-destructed throughout its invocation.
+After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
 \begin{equation}

### 2018-05-22 Paper.tex
Update Paper.tex
- fix `B` subscript
- add `P` function description
Remove "kern 10pc" visible in the PDF
Replace remaining \mathbb{P} with \mathbb{N}.
It looks like, when \mathbb{P} was replaced with \mathbb{N} in most of the
paper, three occurrences were not replaced. This commit replaces them.

Some trailing whitespace was also automatically removed.

@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \end{equation}
-where $B(H)$ is the block of the corresponding header $H$.
+where $B(H)$ and $P(H)$ are the block and the parent block of the corresponding header $H$ respectively.
 
@@
 \end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{P}_{8\ell_{M}} \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\

### 2018-05-24 Paper.tex
Change 'positive' to 'non-negative' in some places.
Since 'positive', referred to numbers, typically means greater than 0, it may be
clearer to say 'non-negative'. This commit performs this replacement where
values in \mathbb{N} are mentioned.

There are still three remaining occurrences of 'positive', all in Appendix F,
which refer to EC private keys and coordinates of EC public keys.

@@
 
-Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
+Throughout, we assume scalars are non-negative integers and thus belong to the set $\mathbb{N}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all non-negative integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
 This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
@@
 
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a non-negative integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
 \begin{equation}
@@
 
-Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+Thus $\mathtt{\tiny BE}$ is the function that expands a non-negative integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
 
@@
 
-If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
-Thus the RLP of some positive integer $i$ is defined as:
+If RLP is used to encode a scalar, defined only as a non-negative integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+Thus the RLP of some non-negative integer $i$ is defined as:
 \begin{equation}

### 2018-05-25 Paper.tex
status code is non-negative
Add colon ":" to help parse description of the transaction receipt R.  Add "non-negative" to "integer" to make it clear the status code cannot be negative.   Add "that" and comma to assertion about R_u.

@@
 \boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
-B & \equiv & (..., (T_0, T_1, ...) ) \\
+B & \equiv & (..., (T_0, T_1, ...), ...) \\
 \Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer.
 \begin{equation}
@@
 
-\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a non-negative integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\linkdest{R__u_assert}We assert that $R_{\mathrm{u}}$, the cumulative gas used, is a non-negative integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
 \begin{equation}

### 2018-05-31 Paper.tex
Clarify that a byte is 8 bits.
This is quite often the case, but not always, e.g. see ISO/IEC 9899:2011
standard for C. This added parenthetical remark avoids any possible confusion.
Allow series of log topics to be "possibly empty", and change some instances of "t" to "x" to avoid confusing with bold "t".

@@
 The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a possibly empty series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
 \begin{equation}
@@
 \begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{x \in O_{\mathbf{t}}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{x \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(x) \big)
 \end{equation}
@@
 
-Where $\mathbb{O}$ is the set of bytes.
+Where $\mathbb{O}$ is the set of (8-bit) bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.

### 2018-06-01 Paper.tex
Remove \tiny and \small from math text.
These cause LaTeX warnings about being invalid in math mode, and do not seem
needed, e.g. just \mathtt{KEC} seems fine. There were occurrences of the same
function name with \tiny, with \small, and with nothing.

--- 2018-05-25 Paper.tex
@@
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{equation}
-\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
 \end{equation}
@@
 \begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
+p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
+\mathtt{EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
 \end{equation}
@@
 \begin{equation}
-\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
+\mathtt{DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}
@@
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \begin{equation}
-\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
+\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
 
@@
 We now have a rigorous specification for the construction of a formal block structure.
-The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
+The RLP function $\texttt{RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
 
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
 \end{equation}
@@
 \linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\hyperlink{trie}{TRIE}}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s] &=& \begin{cases}
@@
 
-%It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
+%It is asserted that the state database will also change such that it defines the pair $(\mathtt{KEC}(\mathbf{b}), \mathbf{b})$.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{TRIE}(\varnothing), \mathtt{KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
 \boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
@@
 \begin{equation}
-\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
-It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
+\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
+
+It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
 
@@
 \boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 
-Where $\mathbb{O}$ is the set of bytes.
+Where $\mathbb{O}$ is the set of (8-bit) bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
 
-We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
+We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus $\mathtt{\tiny BE}$ is the function that expands a non-negative integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+Thus $\mathtt{BE}$ is the function that expands a non-negative integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
 
@@
 \begin{equation}
-\mathtt{\tiny RLP}(i : i \in \mathbb{N}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
+\mathtt{RLP}(i : i \in \mathbb{N}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
 \end{equation}
@@
 
-We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
-\begin{equation}
-\texttt{\small TRIE}(\mathfrak{I}) \equiv \texttt{\small KEC}(c(\mathfrak{I}, 0))
+We define the function $\texttt{TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
+\begin{equation}
+\texttt{TRIE}(\mathfrak{I}) \equiv \texttt{KEC}(c(\mathfrak{I}, 0))
 \end{equation}
@@
 c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
+\texttt{KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
 \end{cases}
@@
 c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
-(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
+\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
 (j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
-\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
+\texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
 u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
@@
 We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
+In the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
 
@@
 g_{\mathrm{r}} &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
 \text{if} \quad |\mathbf{o}| = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
 \mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
@@
 g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
+\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_{\mathbf{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
@@
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
+\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_{\mathbf{d}})
 \end{eqnarray}
@@
 
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
-These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
-\begin{eqnarray}
-\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{secp256k1n} - 1]$}.
+
+We assume the existence of functions $\mathtt{ECDSAPUBKEY}$, $\mathtt{ECDSASIGN}$ and $\mathtt{ECDSARECOVER}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+\begin{eqnarray}
+\mathtt{ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\linkdest{ECDSASIGN}\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 \begin{align}
-0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+0 < \linkdest{r}{r} &< \mathtt{secp256k1n} \\
+0 < \linkdest{s}{s} &< \mathtt{secp256k1n} \div 2 + 1 \\
 \hyperlink{v}{v} &\in \{27,28\}
@@
 \begin{align}
-\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
+\mathtt{secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{secp256k1p} &= 2^{256} - 2^{32} - 977\\
 \end{align}
@@
 \begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
 \end{equation}
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
+h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
 \end{eqnarray}
@@
 G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
 \end{eqnarray}
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
 v_0 &\equiv& \begin{cases}
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
 
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
 \end{equation}

### 2018-06-11 Paper.tex
Clarify that the \mathfrak{I} set of pairs is a map.
add staticcall function cost

@@
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
+&\quad\text{\small DELEGATECALL} \lor \text{\small STATICCALL} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\

### 2018-07-06 Paper.tex
Some fixes to section 2

--- 2018-06-05 Paper.tex
@@
 
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
+Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to transform it into some final state.
 It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
 The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
-Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
-Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
+Transactions thus represent a valid transition between two states.
+The `valid' part is important---there exist far more invalid state transitions than valid state transitions.
+An invalid state transitions might, for \eg, reduce an account balance without increasing another by the same amount.
 A valid state transition is one which comes about through a transaction.
-Formally:
+
+Formally, we define a well-known genesis state $\boldsymbol{\sigma}_0$, a subsequent series of states $\boldsymbol{\sigma}_t$, and a state transition function $\Upsilon$:
 \begin{equation}
@@
 
-where $\Upsilon$ is the Ethereum state transition function.
-In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful than any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
-
-Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
-Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
-They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes place as a state-transition function, adding value to a nominated account.
+The state and state transition function of Ethereum are considerably more powerful than those of existing comparable system; $\Upsilon$ is allowed to carry out arbitrary computation, and $\boldsymbol{\sigma}$ is allowed to store arbitrary state between transactions.
+
+Transactions are collated into blocks; a block consists of a list of transactions together with a cryptographic hash of the previous block that it ``builds upon'', as well as a cryptographic commitment to the final state (blocks do not store the final state itself---that would be far too big).
+Blocks also incentivize nodes to \textit{mine} by allowing the state transition function to add value to a nominated account.
 
@@
 
-Formally, we expand to:
+Formally, we define a block \hyperlink{block}{$B$} as a list of transactions as well as some other components:
+\begin{eqnarray}
+B & \equiv & (..., (T_0, T_1, ...), ...)
+\end{eqnarray}
+
+We define the block-level state transition function $\hyperlink{Pi}{\Pi}$, the block-level analogue to the transaction-level state transition function $\hyperlink{Upsilon}{\Upsilon}$.
+$\Pi$ applies $\Upsilon$ to the transactions of a block in order and then applies an additional block-finalisation state transition $\Omega$.
 \begin{eqnarray}
 \boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
-B & \equiv & (..., (T_0, T_1, ...), ...) \\
 \Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
@@
 
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
-
 This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
@@
 
-In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
-To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
-The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
+In order to incentivise computation within the network, there needs to be a method to transmit value.
+For this purpose, Ethereum provides an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
+The smallest denomination of Ether is the Wei; all other denominations must be an integer multiple of a Wei.
 One Ether is defined as being $10^{18}$ Wei.

### 2018-10-11 Paper.tex
Correct minor inconsistency in output of PoW wrt Sec 11.5 and Appendix J.4

@@
 \end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+with $(m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
 
@@
 \end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+where $(m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (m, x) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))

### 2018-11-08 Paper.tex

--- 2018-08-04 Paper.tex
@@
 
-where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
+where we will remember that $A$, the accrued substate is defined as the tuple of the selfdestructs set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
 

### 2018-11-12 Paper.tex
Fix EXTCODESIZE description on account code
`\mu(a)_c` is defined as account code hash above. `b`, where
`KEC(b) == \mu(a)_c`, is the raw code bytes.

--- 2018-08-04 Paper.tex
@@
 \boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
-B & \equiv & (..., (T_0, T_1, ...), ...) \\
+B & \equiv & (..., (T_0, T_1, ...) ) \\
 \Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
@@
 
-Throughout, we assume scalars are non-negative integers and thus belong to the set $\mathbb{N}$.
+Throughout, we assume scalars are positive integers and thus belong to the set $\mathbb{N}$.
 The set of all byte sequences is $\mathbb{B}$, formally defined in Appendix \ref{app:rlp}.
-If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all non-negative integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
+If such a set of sequences is restricted to those of a particular length, it is denoted with a subscript, thus the set of all byte sequences of length $32$ is named $\mathbb{B}_{32}$ and the set of all positive integers smaller than $2^{256}$ is named $\mathbb{N}_{256}$.
 This is formally defined in section \hyperlink{block}{\ref{subsec:The_Block}}.
@@
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
-This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
+This hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{c}}$, and thus the code may be denoted as $\mathbf{b}$, given that $\texttt{\small KEC}(\mathbf{b}) = \boldsymbol{\sigma}[a]_{\mathrm{c}}$.
 \end{description}
@@
 \begin{equation}
-\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\texttt{\small TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
 \end{equation}
@@
 \begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{\small KEC}(k), \texttt{\small RLP}(v)\big)
 \end{equation}
@@
 If the \textbf{codeHash} field is the Keccak-256 hash of the empty string, i.e.
-$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
+$\boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big)$, then the node represents a simple account, sometimes referred to as a ``non-contract'' account.
 
@@
 \begin{equation}
-p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
+p(a) \equiv  \big(\texttt{\small KEC}(a), \texttt{\small RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
 \end{equation}
@@
 \begin{equation}
-\mathtt{EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
+\mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a]_{\mathrm{c}} = \texttt{\small KEC}\big(()\big) \wedge \boldsymbol{\sigma}[a]_{\mathrm{n}} = 0 \wedge \boldsymbol{\sigma}[a]_{\mathrm{b}} = 0
 \end{equation}
@@
 \begin{equation}
-\mathtt{DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{EMPTY}(\boldsymbol{\sigma}, a)
+\mathtt{\tiny DEAD}(\boldsymbol{\sigma}, a) \quad\equiv\quad \boldsymbol{\sigma}[a] = \varnothing \vee \mathtt{\tiny EMPTY}(\boldsymbol{\sigma}, a)
 \end{equation}
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
+\item[mixHash]\linkdest{mixHash_H__m}{} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{n}}$.
 \end{description}
@@
 
-In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from its execution.
+In order to encode information about a transaction concerning which it may be useful to form a zero-knowledge proof, or index and search, we encode a receipt of each transaction containing certain information from concerning its execution.
 Each receipt, denoted $B_{\mathbf{R}}[i]$ for the $i$th transaction, is placed in an index-keyed \hyperlink{trie}{trie} and the root recorded in the header as \hyperlink{Receipts_Root_H__e}{$H_{\mathrm{e}}$}.
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
 \begin{equation}
@@
 
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer.
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is an integer.
 \begin{equation}
@@
 
-\linkdest{R__u_assert}We assert that $R_{\mathrm{u}}$, the cumulative gas used, is a non-negative integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
+\linkdest{R__u_assert}We assert $R_{\mathrm{u}}$, the cumulative gas used is a positive integer and that the logs Bloom, $R_{\mathrm{b}}$, is a hash of size 2048 bits (256 bytes):
 \begin{equation}
@@
 The sequence $R_{\mathbf{l}}$ is a series of log entries, $(O_0, O_1, ...)$.
-A log entry, $O$, is a tuple of the logger's address, $O_a$, a possibly empty series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
+A log entry, $O$, is a tuple of the logger's address, $O_a$, a series of 32-byte log topics, $O_{\mathbf{t}}$ and some number of bytes of data, $O_{\mathbf{d}}$:
 \begin{equation}
@@
 \begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{x \in O_{\mathbf{t}}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{t \in O_{\mathbf{t}}}: t \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \begin{equation}
-M(O) \equiv \hyperlink{bigvee}{\bigvee}_{x \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(x) \big)
+M(O) \equiv \hyperlink{bigvee}{\bigvee}_{t \in \{O_{\mathrm{a}}\} \cup O_{\mathbf{t}}} \big( M_{3:2048}(t) \big)
 \end{equation}
@@
 \begin{array}[t]{lclc}
-\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{\small TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
+\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{\small KEC}(\mathtt{\small RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{\small TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{P}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \begin{equation}
-\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
+\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{\small TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
 
@@
 We now have a rigorous specification for the construction of a formal block structure.
-The RLP function $\texttt{RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
+The RLP function $\texttt{\small RLP}$ (see Appendix \ref{app:rlp}) provides the canonical method for transforming this structure into a sequence of bytes ready for transmission over the wire or storage locally.
 
@@
 \begin{equation}
-P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+P(H) \equiv B': \mathtt{\tiny KEC}(\mathtt{\tiny RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
 \end{equation}
@@
 
-The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
+The nonce, \hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
 \begin{equation}
@@
 
-After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
-\begin{equation}
-	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
-\end{equation}
-
-Then the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
-\begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
+After the message call or contract creation is processed, the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
+\begin{equation}
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A_{\mathrm{r}}} \right\}
 \end{equation}
@@
 The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
-Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
 
@@
 \linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
-\end{eqnarray}
-
-\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
-\begin{eqnarray}
-\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\end{eqnarray}
+
+\hypertarget{tx_total_gas_used_Upsilon_pow_g}{}\linkdest{Upsilon_pow_g}\hypertarget{tx_logs_Upsilon_pow_l}{}\linkdest{Upsilon_pow_l}\hypertarget{tx_status_Upsilon_pow_z}{}\linkdest{Upsilon_pow_z}And finally, we specify $\Upsilon^{\mathrm{g}}$, the total gas used in this transaction, $\Upsilon^\mathbf{l}$, the logs created by this transaction and $\Upsilon^{\mathrm{z}}$, the status code of this transaction:
+\begin{eqnarray}
+\Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g' \\
 \Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 \begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{\tiny KEC}\Big(\mathtt{\tiny RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{\tiny KEC}$ is the Keccak 256-bit hash function, $\mathtt{\tiny RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\hyperlink{trie}{TRIE}}(\varnothing), \mathtt{KEC}\big(()\big) \big) \\
+\boldsymbol{\sigma}^*[a] &=& \big( 1, v + v', \mathtt{\tiny \hyperlink{trie}{TRIE}}(\varnothing), \mathtt{\tiny KEC}\big(()\big) \big) \\
 \boldsymbol{\sigma}^*[s] &=& \begin{cases}
@@
 
-%It is asserted that the state database will also change such that it defines the pair $(\mathtt{KEC}(\mathbf{b}), \mathbf{b})$.
+%It is asserted that the state database will also change such that it defines the pair $(\mathtt{\tiny KEC}(\mathbf{b}), \mathbf{b})$.
 
@@
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}^{**}, a) \\
+\quad\boldsymbol{\sigma}'[a] = \varnothing & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}^{**}, a) \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
-\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
+\quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{\small KEC}(\mathbf{o}) & \text{otherwise}
 \end{cases} \\
@@
 \subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address should return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
+Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code.
 Thus any message call received by it during this time causes no code to be executed.
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call and finally the present depth of the message-call/contract-creation stack ($e$).
 
@@
 \begin{cases}
-\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{TRIE}(\varnothing), \mathtt{KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
+\boldsymbol{\sigma}_1'[r] \equiv (0, v, \mathtt{\tiny TRIE}(\varnothing), \mathtt{\tiny KEC}(())) & \text{if} \quad \boldsymbol{\sigma}[r] = \varnothing \wedge v \neq 0 \\
 \boldsymbol{\sigma}_1'[r] \equiv \varnothing & \text{if}\quad \boldsymbol{\sigma}[r] = \varnothing \wedge v = 0 \\
@@
 \begin{equation}
-\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
-It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
+\text{Let} \; \mathtt{\tiny KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
+
+It is assumed that the client will have stored the pair $(\mathtt{\tiny KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
 
@@
 \end{equation}
-where $B(H)$ and $P(H)$ are the block and the parent block of the corresponding header $H$ respectively.
+where $B(H)$ is the block of the corresponding header $H$.
 
@@
 \boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\small \hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{\small TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the \hyperlink{block_nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
+engineer the function such that the determination of the nonce requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
 The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
@@
 \begin{equation}
-m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
+m = H_{\mathrm{m}} \quad \wedge \quad n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})
 \end{equation}
@@
 
-Where $\mathbb{O}$ is the set of (8-bit) bytes.
+Where $\mathbb{O}$ is the set of bytes.
 Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
 
-We define the RLP function as $\mathtt{RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
-\begin{equation}
-\mathtt{RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
+We define the RLP function as $\mathtt{\tiny RLP}$ through two sub-functions, the first handling the instance when the value is a byte array, the second when it is a sequence of further values:
+\begin{equation}
+\mathtt{\tiny RLP}(\mathbf{x}) \equiv \begin{cases} R_{\mathrm{b}}(\mathbf{x}) & \text{if} \quad \mathbf{x} \in \mathbb{B} \\ R_{\mathrm{l}}(\mathbf{x}) & \text{otherwise} \end{cases}
 \end{equation}
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{\tiny BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 
-Thus $\mathtt{BE}$ is the function that expands a non-negative integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
+Thus $\mathtt{\tiny BE}$ is the function that expands a positive integer value to a big-endian byte array of minimal length and the dot operator performs sequence concatenation.
 
@@
 (192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+\big(247 + \big\lVert \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{\tiny BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
 \end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) ...
-\end{eqnarray}
-
-If RLP is used to encode a scalar, defined only as a non-negative integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
-Thus the RLP of some non-negative integer $i$ is defined as:
-\begin{equation}
-\mathtt{RLP}(i : i \in \mathbb{N}) \equiv \mathtt{RLP}(\mathtt{BE}(i))
+s(\mathbf{x}) & \equiv & \mathtt{\tiny RLP}(\mathbf{x}_0) \cdot \mathtt{\tiny RLP}(\mathbf{x}_1) ...
+\end{eqnarray}
+
+If RLP is used to encode a scalar, defined only as a positive integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+Thus the RLP of some positive integer $i$ is defined as:
+\begin{equation}
+\mathtt{\tiny RLP}(i : i \in \mathbb{N}) \equiv \mathtt{\tiny RLP}(\mathtt{\tiny BE}(i))
 \end{equation}
@@
 
-We define the function $\texttt{TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
-\begin{equation}
-\texttt{TRIE}(\mathfrak{I}) \equiv \texttt{KEC}(c(\mathfrak{I}, 0))
+We define the function $\texttt{\small TRIE}$, which evaluates to the root of the trie that represents this set when encoded in this structure:
+\begin{equation}
+\texttt{\small TRIE}(\mathfrak{I}) \equiv \texttt{\small KEC}(c(\mathfrak{I}, 0))
 \end{equation}
@@
 c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
+\texttt{\small KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
 \end{cases}
@@
 c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
+\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
+(\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{\small RLP}\Big( \big(\texttt{\small HP}(I_0[i ..
 (j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
 (x - 1)] = \mathbf{l} \\
-\texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
+\texttt{\small RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
 u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
@@
 We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$.
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{\tiny ECDSARECOVER}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
+In the case of an invalid signature ($\mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
 
@@
 g_{\mathrm{r}} &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{\tiny ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
 \text{if} \quad |\mathbf{o}| = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{o}[12..31] &=& \mathtt{\tiny KEC}\big(\mathtt{\tiny ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
 \mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
@@
 g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
-\mathbf{o}[0..31] &=& \mathtt{SHA256}(I_{\mathbf{d}})\\
+\mathbf{o}[0..31] &=& \mathtt{\tiny SHA256}(I_{\mathbf{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
@@
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{RIPEMD160}(I_{\mathbf{d}})
+\mathbf{o}[12..31] &=& \mathtt{\tiny RIPEMD160}(I_{\mathbf{d}})
 \end{eqnarray}
@@
 \end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{P}_{8\ell_{M}} \\
 \ell_{B} &\equiv& i[0..31] \\
@@
 \begin{eqnarray}\label{eq:ec-doubling}
-(X_1, Y_1) + (X_2, Y_2)&\equiv&\begin{cases}
+(X_1, Y_1) + (X_1, Y_1)&\equiv&\begin{cases}
 (X,Y)&\text{if}\ Y_1\neq 0\\
@@
 
-Let $F_{p^2}$ be a field $F_{p}[i]/(i^2+1)$. We define a set $C_2$ with
+Let $F_{p^2}$ be a field $F_{\mathrm{p}}[i]/(i^2+1)$. We define a set $C_2$ with
 \begin{equation}
@@
 
-It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{secp256k1n} - 1]$}.
-
-We assume the existence of functions $\mathtt{ECDSAPUBKEY}$, $\mathtt{ECDSASIGN}$ and $\mathtt{ECDSARECOVER}$. These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
-\begin{eqnarray}
-\mathtt{ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
-\linkdest{ECDSASIGN}\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+It is assumed that the sender has a valid private key $p_{\mathrm{r}}$, which is a randomly selected positive integer (represented as a byte array of length 32 in big-endian form) in the range \hbox{$[1, \mathtt{\tiny secp256k1n} - 1]$}.
+
+We assume the existence of functions $\mathtt{\small ECDSAPUBKEY}$, $\mathtt{\small ECDSASIGN}$ and $\mathtt{\small ECDSARECOVER}$.
+These are formally defined in the literature, \eg by \cite{ECDSAcerticom}.
+\begin{eqnarray}
+\mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64} \\
+\linkdest{ECDSASIGN}\mathtt{\small ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
+\mathtt{\small ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
-It is assumed that \hypertarget{v}{}$v$ is the `recovery identifier'.
+It is assumed that \hypertarget{v}{}$v$ is either the `recovery identifier' or `chain identifier doubled plus 35 or 36'.
 The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
@@
 
-\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true\footnote{A signature of a transaction can be valid not only with a recovery identifier but with some other numbers.
-See \hyperlink{T__w}{how the component $T_w$ of a transaction is interpreted.}}:
+\linkdest{invalidsig}We declare that a signature is invalid unless all the following conditions are true:
 \begin{align}
-0 < \linkdest{r}{r} &< \mathtt{secp256k1n} \\
-0 < \linkdest{s}{s} &< \mathtt{secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28\}
+0 < \linkdest{r}{r} &< \mathtt{\tiny secp256k1n} \\
+0 < \linkdest{s}{s} &< \mathtt{\tiny secp256k1n} \div 2 + 1 \\
+\hyperlink{v}{v} &\in \{27,28,\mathtt{\tiny chain\_id} \times 2 + 35, \mathtt{\tiny chain\_id} \times 2 + 36\}
 \end{align}
@@
 \begin{align}
-\mathtt{secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
-%\mathtt{secp256k1p} &= 2^{256} - 2^{32} - 977\\
+\mathtt{\tiny secp256k1n} &= 115792089237316195423570985008687907852837564279074904382605163141518161494337
+%\mathtt{\tiny secp256k1p} &= 2^{256} - 2^{32} - 977\\
 \end{align}
@@
 \begin{equation}
-A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
+A(p_{\mathrm{r}}) = \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSAPUBKEY}(p_{\mathrm{r}}) \big) \big)
 \end{equation}
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{chain\_id}, (), ()) & \text{otherwise} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{\tiny chain\_id}, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
+h(T) & \equiv & \mathtt{\small KEC}( L_{S}(T) )
 \end{eqnarray}
@@
 G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
+(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{\small ECDSASIGN}(h(T), p_{\mathrm{r}})
 \end{eqnarray}
@@
 \begin{eqnarray}
+\linkdest{T__w}{T_{\mathrm{w}}} = \hyperlink{v}{v}\\
 \linkdest{T__r}{T_{\mathrm{r}}} = \hyperlink{r}{r}\\
@@
 \end{eqnarray}
-$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.
-In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{\tiny KEC}\big( \mathtt{\small ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
 v_0 &\equiv& \begin{cases}
@@
 $G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\ \everyrow{}
 $G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
@@
 \subsection{Instruction Set}
-\label{subsec:instruction-set}
 
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{\tiny Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\
+&&&& where $\texttt{\small KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 \midrule
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
-\end{cases}$\\ \\
-&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\ \\
+\end{cases}$\\
+&&&& where $r = \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}$\\
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
+&&&& $A'_{r} \equiv A_{r} + \begin{cases}
+R_{selfdestruct} & \text{if} \quad I_{\mathrm{a}} \notin A_{\mathbf{s}} \\
+0 & \text{otherwise}
+\end{cases}$ \\
 &&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{\tiny DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
-$\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
+$\mathtt{\tiny KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{\tiny KEC}\big(\mathtt{\tiny RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
 
@@
  C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
-\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(\mathbf{0}_{32}) & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
 \texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
 \end{equation}

### 2018-11-13 Paper.tex

@@
 \toprule
-\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{30s: Environmental Information (0x30 - 0x39)}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
-\midrule
+\bottomrule
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
+\toprule
+\multicolumn{5}{c}{\textbf{30s: Environmental Information (0x3a - 0x3f)}} \vspace{5pt} \\
+\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+\midrule
+0x3f & {\small EXTCODEHASH} & 1 & 1 & Get hash of an account's code. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 \bottomrule
@@
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
+&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its\\
+&&&& hash is the parent $h$ is zero.\\
 \midrule

### 2018-12-10 Paper.tex
Fix confusing and irrelevant ECDSA citation (#712)
Fixes part of #704 

Appendix F contains a reference to "Electrum style signatures", but the citation points to a JavaScript frontend library with no mention of ECDSA.
Change version in title to Constantinople
Clarify that the \mathfrak{I} set of pairs is a map. (#705)

--- 2018-11-08 Paper.tex
@@
 
-Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
+Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences with unique keys:
 \begin{equation}
@@
 
-The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
+Transactions are signed using recoverable ECDSA signatures.
 This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.

### 2019-01-27 Paper.tex
Fix small notation error

@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{sha3}+G_{sha3word}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
 G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\

### 2019-03-29 Paper.tex
Paper.tex: fix gas forwarded in DELEGATECALL definition
Change arg max to max in Appendix D.
Here j is the maximum x such that there exists an \mathbf{l} such that etc. etc.
That is, j is the length of the longest common prefix of all the keys (nibble
arrays) of the \mathfrak{I} map.

Note that j is not the argument x that maximizes the value of a function over x.
The expression \exists \mathbf{l} ... following \arg \max is boolean-valued.

Also change a colon into a \wedge, which conjoins the two constraints over x and
\mathbf{l} in the set comprehension.
Explicate non-RLP-encodable structure

--- 2018-12-10 Paper.tex
@@
 \item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
+\item Otherwise, the output is equal to the input, provided that it contains fewer than $2^{64}$ bytes, prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
 \end{itemize}
+
+Byte arrays containing $2^{64}$ or more bytes cannot be encoded.
+This restriction ensures that the first byte of the encoding of a byte array is always below 192, and thus it can be readily distinguished from the encodings of sequences in $\mathbb{L}$.
 
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
+\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 2^{64} \\
+\varnothing & \text{otherwise}
 \end{cases} \\
@@
 \item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\item Otherwise, the output is equal to the concatenated serialisations, provided that they contain fewer than $2^{64}$ bytes, prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
 \end{itemize}
 
+Sequences whose concatenated serialized items contain $2^{64}$ or more bytes cannot be encoded.
+This restriction ensures that the first byte of the encoding does not exceed 255 (otherwise it would not be a byte).
+
 \hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
@@
 R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
-(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
+(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 56 \\
+\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{else if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 2^{64} \\
+\varnothing & \text{otherwise}
 \end{cases} \\
-s(\mathbf{x}) & \equiv & \mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) ...
+s(\mathbf{x}) & \equiv & \begin{cases}
+\mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) \cdot ... & \text{if} \quad \forall i: \mathtt{RLP}(\mathbf{x}_i) \neq \varnothing \\
+\varnothing & \text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 \texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
-(x - 1)] = \mathbf{l} \\
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(x - 1)] = \mathbf{l} \} \\
 \texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}

### 2019-04-13 Paper.tex
Paper.tex: fix gas forwarded in DELEGATECALL definition (#733)
Fix typo.
The \mathrm{n} should be just n here. It is the index that the sum ranges over.
Add missing comma.

@@
 \end{cases} \\
-\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification, is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2019-04-17 Paper.tex
Clarify the definition of RLP trees. (#739)
Consistently use 'byte array'.
A few occurrences of 'bytearray' and 'byte-array' in certain parts of the paper
have been changed to 'byte array'.

@@
 Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
+The trie requires a simple database backend that maintains a mapping of byte arrays to byte arrays; we name this underlying database the state database.
 This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
@@
 \begin{eqnarray}
-\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
+\mathbb{T} & \equiv & \mathbb{L} \uplus \mathbb{B} \\
 \mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
@@
 Where $\mathbb{O}$ is the set of (8-bit) bytes.
-Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
+Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte arrays and such structural sequences.
+The disjoint union $\uplus$ is needed only to distinguish the empty byte array $()\in\mathbb{B}$ from the empty list $()\in\mathbb{L}$, which are encoded differently as defined below; as common, we will abuse notation and leave the disjoint union indices implicit, inferable from context.
 
@@
 
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte array, the RLP serialisation takes one of three forms:
 
 \begin{itemize}
-\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
-\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input, provided that it contains fewer than $2^{64}$ bytes, prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
+\item If the byte array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
+\item If the byte array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
+\item Otherwise, the output is equal to the input, provided that it contains fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
 \end{itemize}
@@
 \item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations, provided that they contain fewer than $2^{64}$ bytes, prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\item Otherwise, the output is equal to the concatenated serialisations, provided that they contain fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
 \end{itemize}

### 2019-04-18 Paper.tex
Fix font and improve wording.
The \mathtm{x} used as subscript to \mathbb{N} was not matching the plain
(\mathit) x in the phrase.

The wording was also not very clear. The new wording should be better.

@@
 
-If RLP is used to encode a scalar, defined only as a non-negative integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some non-negative integer $i$ is defined as:

### 2019-05-04 Paper.tex
Fix font and improve wording.
The \mathtm{x} used as subscript to \mathbb{N} was not matching the plain
(\mathit) x in the phrase.

The wording was also not very clear. The new wording should be better.
Fix notation for sequence elements.
According to eqs. (177) and (178), the elements of the sequence are selected
with the notation ...[...]. While eq. (180) uses that notation for x[0],
eq. (184) was using a subscript notation x_0, x_1, and x_i. This commit turns
these into x[0], x[1], and x[i].
Improve summation notation in BE definition.
Summation normally has one of the following forms:

\su

@@
 \end{cases} \\
-\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
 (a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
@@
 s(\mathbf{x}) & \equiv & \begin{cases}
-\mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) \cdot ... & \text{if} \quad \forall i: \mathtt{RLP}(\mathbf{x}_i) \neq \varnothing \\
+\mathtt{RLP}(\mathbf{x}[0]) \cdot \mathtt{RLP}(\mathbf{x}[1]) \cdot ... & \text{if} \quad \forall i: \mathtt{RLP}(\mathbf{x}[i]) \neq \varnothing \\
 \varnothing & \text{otherwise}
@@
 
-If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
+If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be encoded as the shortest byte array whose big-endian interpretation is the scalar.
 Thus the RLP of some non-negative integer $i$ is defined as:

### 2019-05-05 Paper.tex
Improve sentence clarity and explicitness.
Treat t argument of HP (and f) as boolean.
The paragraph just before eq. (186) says that the second argument, t, of the
function HP, is a boolean value. The function HP is passed the boolean values
true and false as t in eq. (194). However, eq. (187) was treating t as a number,
in the condition 't \neq 0'. Even though booleans are isomorphic to {0, 1}, it's
clearer to change that condition to 't = true'.

@@
 
-If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be encoded as the shortest byte array whose big-endian interpretation is the scalar.
+If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some non-negative integer $i$ is defined as:
@@
 \end{cases} \\
-f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t \neq 0 \\ 0 & \text{otherwise} \end{cases}
+f(t) & \equiv & \begin{cases} 2 & \text{if} \quad t = \mathit{true} \\ 0 & \text{otherwise} \end{cases}
 \end{eqnarray}

### 2019-05-06 Paper.tex
Improve appearance of true and false.
In LaTeX's math mode, multi-letter items like true and false should be enclosed
in \mathit{...}, otherwise, each letter is considered a separate item and the
multi-letter item is rendered as an arithmetic product of its letters, with a
bit of additional spacing between the letters.

This commit adds \mathit{...} around all the math-mode occurrences of true and
false.

--- 2019-05-04 Paper.tex
@@
 \begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{x \in O_{\mathbf{t}}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall x \in O_{\mathbf{t}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
+\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
 m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
@@
 \begin{equation}
-k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
+k(U, H, n) \equiv \begin{cases} \mathit{false} & \text{if} \quad n = 0 \\
 s(U, H) &\\
@@
 \qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
-\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
+\\ \nonumber
+\forall \mathbf{U} \in \mathbf{B}_{\mathbf{U}}: \\ \nonumber
 \boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
@@
 \mathbb{T} & \equiv & \mathbb{L} \uplus \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{t} \rVert : \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{b} \rVert : \mathbf{b}[n] \in \mathbb{O} \}
 \end{eqnarray}
@@
 \mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
+(x_1, ..., x_n) \cdot (y_1, ..., y_m) & = & (x_1, ..., x_n, y_1, ..., y_m)
 \end{eqnarray}
@@
 \begin{equation}
-\forall_{\mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
+\forall I \in \mathfrak{I}: I \equiv (I_0, I_1)
 \end{equation}
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall_{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_{n}\rVert} \quad \mathbf{k}_{n}'[i] & \equiv &
+\forall n: \quad \forall i < 2\lVert\mathbf{k}_{n}\rVert: \quad \mathbf{k}_{n}'[i] & \equiv &
 \begin{cases}
@@
 \item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $\mathit{true}$.
 \item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $\mathit{false}$.
 \item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
@@
 c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
 \texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(\lVert I_0\rVert - 1)], \mathit{true}), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
+(j - 1)], \mathit{false}), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
 (x - 1)] = \mathbf{l} \} \\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
 \begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
 \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
+&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
 \begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\

### 2019-05-08 Paper.tex
Pass a bit as the second argument of HP.
Appendix D was passing true and false as the second argument t of HP. However,
HP (and f) in Appendix C treat t as a bit, not as a boolean. Even though bits
and booleans are isomoprhic, it's clearer to change Appendix D to pass bits. The
mention of true and false in the text that precedes eq. (194) has also been
changes to 1 and 0.

@@
 \item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $\mathit{true}$.
+The hex-prefix encoding method is used and the second parameter to the function is required to be 1.
 \item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be $\mathit{false}$.
+The hex-prefix encoding method is used and the second parameter to the function is required to be 0.
 \item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
@@
 c(\mathfrak{I}, i) \equiv \begin{cases}
+\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], 1), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
 \texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(\lVert I_0\rVert - 1)], \mathit{true}), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(j - 1)], \mathit{false}), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
+(j - 1)], 0), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
 (x - 1)] = \mathbf{l} \} \\

### 2019-05-13 Paper.tex
Fix two obvious typos.

@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak hash of the corresponding ECDSA public key:
 \begin{equation}
@@
 
-When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the left most 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
+When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the leftmost 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
 

### 2019-05-13 Paper.tex
Paper.tex: correct layout of input data to BN_MUL fixes #751

@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
-Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
+For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per version 3 of the winning entry to the SHA-3 contest by \cite{Keccak}, rather than the final SHA-3 specification), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+Also, $\texttt{KEC512}$ refers to the Keccak-512 hash function.
 
@@
 Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
-We call this \textit{transaction substate}, and represent it as $A$, which is a tuple:
+We call this the \textit{accrued transaction substate}, or \textit{accrued substate} for short, and represent it as $A$, which is a tuple:
 \begin{equation}
@@
 
-We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
+We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
 \begin{equation}
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
 \begin{equation}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct set or are touched and empty:
 \begin{eqnarray}
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
 
@@
 
-Aside from evaluating to a new state and transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
@@
 
-where we will remember that $A$, the accrued substate is defined as the tuple of the selfdestructs set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
-
-\begin{equation}
-A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
+where we will remember that $A$, the accrued substate, is defined as the tuple of the self-destructs set $A_{\mathbf{s}}$, the log series $A_{\mathbf{l}}$, the touched accounts $A_{\mathbf{t}}$ and the refund balance $A_{\mathrm{r}}$:
+\begin{equation}
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
 \end{equation}
@@
 \mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-n&\equiv&\bar I_{\mathbf{d}}[0..31]\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[32..95]\right)
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
+n&\equiv&\bar I_{\mathbf{d}}[64..95]
 \end{eqnarray}

### 2019-05-31 Paper.tex
Update Paper.tex

@@
 These may be subscripted to denote specialised variants, \eg \hyperlink{C__SSTORE}{$C_\text{SSTORE}$}, the cost function for the \hyperlink{SSTORE}{{\tiny SSTORE}} operation.
-For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per version 3 of the winning entry to the SHA-3 contest by \cite{Keccak}, rather than the final SHA-3 specification), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
-Also, $\texttt{KEC512}$ refers to the Keccak-512 hash function.
+For specialised and possibly externally defined functions, we may format as typewriter text, \eg the Keccak-256 hash function (as per the winning entry to the SHA-3 contest by \cite{Keccak}, rather than later releases), is denoted $\texttt{KEC}$ (and generally referred to as plain Keccak).
+Also $\texttt{KEC512}$ is referring to the Keccak 512 hash function.
 
@@
 \mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
-x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[0..63]\right)\\
-n&\equiv&\bar I_{\mathbf{d}}[64..95]
+n&\equiv&\bar I_{\mathbf{d}}[0..31]\\
+x&\equiv&\delta_1\left(\bar I_{\mathbf{d}}[32..95]\right)
 \end{eqnarray}

### 2019-07-28 Paper.tex
Clarify what `ADDRESS` does
`executing` can refer to the code being executed.
This is misleading as this is not the case with `DELEGATECALL` and `CALLCODE`.
Paper.tex: fix gas cost of ECPAIRING
Add missing param

@@
 E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+E_{\mathrm{accesses}}(\mathbf{d}, E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}

### 2019-07-30 Paper.tex
Paper.tex: fix G_quaddivisor (should be 20)
Paper.tex: fix gas cost of ECPAIRING (#760)
Paper.tex: correct layout of input data to BN_MUL (#752)
fixes #751

--- 2019-07-02 Paper.tex
@@
 k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
-g_{\mathrm{r}}&=& 60000k + 40000 \\
+g_{\mathrm{r}}&=& 80000k + 100000 \\
 \mathbf{o}[0..31]&\equiv&\begin{cases}
@@
 $G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 100 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
+$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
 &&contract. \\

### 2019-08-16 Paper.tex
add missing right parenthesis （添加缺失的右括号）
Fix mathematical symbols.
This should be the norm of o (|| o ||), not the absolute value (| o |).
Fix mathematical symbols (17 modifications).
These variables (c, o, Id, d) are all arrays. The intention here is to get the length of the array, so apply the norm symbol ( || ) instead of the absolute value symbol ( | ).
The modified content includes the following formulas: 140, 198, 199, 202, 203, 209, 212, 218, 231, 255, 256, 262, 263, 270. And a partial description of the CALL instruction in the appendix "Virtual Machine Specification".

@@
 \begin{equation}
-c \equiv G_{codedeposit} \times |\mathbf{o}|
+c \equiv G_{codedeposit} \times \lVert \mathbf{o} \rVert
 \end{equation}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  |\mathbf{o}| > 24576\big)
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  \lVert \mathbf{o} \rVert > 24576\big)
 \end{align}
@@
 D_{J}(\mathbf{c}, i) \equiv \begin{cases}
-\{\} & \text{if} \quad i \geqslant |\mathbf{c}|  \\
+\{\} & \text{if} \quad i \geqslant \lVert \mathbf{c} \rVert  \\
 \{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
@@
 g_{\mathrm{r}} &=& 3000\\
-|\mathbf{o}| &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
-\text{if} \quad |\mathbf{o}| = 32: &&\\
+\lVert \mathbf{o} \rVert &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+\text{if} \quad \lVert \mathbf{o} \rVert = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
 \mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
-\mathbf{d}[0..(|\hyperlink{I__d}{I_{\mathbf{d}}}|-1)] &=& I_{\mathbf{d}}\\
-\mathbf{d}[|I_{\mathbf{d}}|..] &=& (0, 0, ...) \\
+\mathbf{d}[0..(\lVert \hyperlink{I__d}{I_{\mathbf{d}}} \rVert-1)] &=& I_{\mathbf{d}}\\
+\mathbf{d}[\lVert I_{\mathbf{d}} \rVert..] &=& (0, 0, ...) \\
 h &=& \mathbf{d}[0..31]\\
@@
 \Xi_{\mathtt{SHA256}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 60 + 12\Big\lceil \dfrac{\lVert I_{\mathbf{d}} \rVert}{32} \Big\rceil\\
 \mathbf{o}[0..31] &=& \mathtt{SHA256}(I_{\mathbf{d}})\\
 \Xi_{\mathtt{RIP160}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 600 + 120\Big\lceil \dfrac{\lVert I_{\mathbf{d}} \rVert}{32} \Big\rceil\\
 \mathbf{o}[0..11] &=& 0 \\
@@
 \Xi_{\mathtt{ID}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{where:} \\
-g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{|I_{\mathbf{d}}|}{32} \Big\rceil\\
+g_{\mathrm{r}} &=& 15 + 3\Big\lceil \dfrac{\lVert I_{\mathbf{d}} \rVert}{32} \Big\rceil\\
 \mathbf{o} &=& I_{\mathbf{d}}
@@
 i[x] &\equiv& \begin{cases}
-I_{\mathbf{d}}[x] &\text{if}\ x < |I_{\mathbf{d}}| \\
+I_{\mathbf{d}}[x] &\text{if}\ x < \lVert I_{\mathbf{d}} \rVert \\
 0 &\text{otherwise}
@@
 \qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
-F&\equiv&(|I_{\mathbf{d}}|\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
-k &=& \dfrac{|I_{\mathbf{d}}|}{192} \\
+F&\equiv&(\lVert I_{\mathbf{d}} \rVert\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
+k &=& \dfrac{\lVert I_{\mathbf{d}} \rVert}{192} \\
 g_{\mathrm{r}}&=& 80000k + 100000 \\
@@
 \vdots\\
-a_k&\equiv&\delta_1(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-192)..(|I_{\mathbf{d}}|-129)])\\
-b_k&\equiv&\delta_2(I_{\mathbf{d}}[(|I_{\mathbf{d}}|-128)..(|I_{\mathbf{d}}|-1)])
+a_k&\equiv&\delta_1(I_{\mathbf{d}}[(\lVert I_{\mathbf{d}} \rVert-192)..(\lVert I_{\mathbf{d}} \rVert-129)])\\
+b_k&\equiv&\delta_2(I_{\mathbf{d}}[(\lVert I_{\mathbf{d}} \rVert-128)..(\lVert I_{\mathbf{d}} \rVert-1)])
 \end{eqnarray}
@@
 \label{eq:complemented_input}\bar I_{\mathbf{d}}[x]&\equiv&\begin{cases}
-I_{\mathbf{d}}[x]&\text{if}\ x < |I_{\mathbf{d}}|\\
+I_{\mathbf{d}}[x]&\text{if}\ x < \lVert I_{\mathbf{d}} \rVert\\
 0&\text{otherwise}
@@
 &&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], |\mathbf{o}|\})$ \\
+&&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[5] \dots (\boldsymbol{\mu}_{\mathbf{s}}[5] + n - 1) ] = \mathbf{o}[0 \dots (n - 1)]$ \\
@@
 \begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
 \end{equation}

### 2019-10-10 Paper.tex

--- 2019-07-28 Paper.tex
@@
 Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
-The trie requires a simple database backend that maintains a mapping of byte arrays to byte arrays; we name this underlying database the state database.
+The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database.
 This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state.
@@
 \begin{equation}
-O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall x \in O_{\mathbf{t}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
+O_{\mathrm{a}} \in \mathbb{B}_{20} \quad \wedge \quad \forall_{x \in O_{\mathbf{t}}}: x \in \mathbb{B}_{32} \quad \wedge \quad O_{\mathbf{d}} \in \mathbb{B}
 \end{equation}
@@
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
+\forall_{i \in \{0, 2, 4\}}&:& \mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
 m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
@@
 Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
-We call this the \textit{accrued transaction substate}, or \textit{accrued substate} for short, and represent it as $A$, which is a tuple:
+We call this \textit{transaction substate}, and represent it as $A$, which is a tuple:
 \begin{equation}
@@
 
-We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
+We define the empty substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
 \begin{equation}
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, substate $A$ and status code $z$:
 \begin{equation}
@@
 
-The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct set or are touched and empty:
+The final state, $\boldsymbol{\sigma}'$, is reached after deleting all accounts that either appear in the self-destruct list or are touched and empty:
 \begin{eqnarray}
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
 
@@
 
-Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+Aside from evaluating to a new state and transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
@@
 
-where we will remember that $A$, the accrued substate, is defined as the tuple of the self-destructs set $A_{\mathbf{s}}$, the log series $A_{\mathbf{l}}$, the touched accounts $A_{\mathbf{t}}$ and the refund balance $A_{\mathrm{r}}$:
-\begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
+where we will remember that $A$, the accrued substate is defined as the tuple of the suicides set $\mathbf{s}$, the log series $\mathbf{l}$, the touched accounts $\mathbf{t}$ and the refunds $r$:
+
+\begin{equation}
+A \equiv (\mathbf{s}, \mathbf{l}, \mathbf{t}, r)
 \end{equation}
@@
 \begin{equation}
-k(U, H, n) \equiv \begin{cases} \mathit{false} & \text{if} \quad n = 0 \\
+k(U, H, n) \equiv \begin{cases} false & \text{if} \quad n = 0 \\
 s(U, H) &\\
@@
 \qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
-\\ \nonumber
-\forall \mathbf{U} \in \mathbf{B}_{\mathbf{U}}: \\ \nonumber
+\qquad\forall_{\mathbf{U} \in \mathbf{B}_{\mathbf{U}}}: \\ \nonumber
 \boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
@@
 \begin{eqnarray}
-\mathbb{T} & \equiv & \mathbb{L} \uplus \mathbb{B} \\
-\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{t} \rVert : \mathbf{t}[n] \in \mathbb{T} \} \\
-\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall n < \lVert \mathbf{b} \rVert : \mathbf{b}[n] \in \mathbb{O} \}
+\mathbb{T} & \equiv & \mathbb{L} \cup \mathbb{B} \\
+\mathbb{L} & \equiv & \{ \mathbf{t}: \mathbf{t} = ( \mathbf{t}[0], \mathbf{t}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{t} \rVert} \; \mathbf{t}[n] \in \mathbb{T} \} \\
+\mathbb{B} & \equiv & \{ \mathbf{b}: \mathbf{b} = ( \mathbf{b}[0], \mathbf{b}[1], ... ) \; \wedge \; \forall_{n < \lVert \mathbf{b} \rVert} \; \mathbf{b}[n] \in \mathbb{O} \}
 \end{eqnarray}
@@
 Where $\mathbb{O}$ is the set of (8-bit) bytes.
-Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte arrays and such structural sequences.
-The disjoint union $\uplus$ is needed only to distinguish the empty byte array $()\in\mathbb{B}$ from the empty list $()\in\mathbb{L}$, which are encoded differently as defined below; as common, we will abuse notation and leave the disjoint union indices implicit, inferable from context.
+Thus $\mathbb{B}$ is the set of all sequences of bytes (otherwise known as byte-arrays, and a leaf if imagined as a tree), $\mathbb{L}$ is the set of all tree-like (sub-)structures that are not a single leaf (a branch node if imagined as a tree) and $\mathbb{T}$ is the set of all byte-arrays and such structural sequences.
 
@@
 
-\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte array, the RLP serialisation takes one of three forms:
+\hypertarget{RLP_serialisation_of_a_byte_array_R__b_word_def}{}\linkdest{R__b}If the value to be serialised is a byte-array, the RLP serialisation takes one of three forms:
 
 \begin{itemize}
-\item If the byte array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
-\item If the byte array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
-\item Otherwise, the output is equal to the input, provided that it contains fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
+\item If the byte-array contains solely a single byte and that single byte is less than 128, then the input is exactly equal to the output.
+\item If the byte-array contains fewer than 56 bytes, then the output is equal to the input prefixed by the byte equal to the length of the byte array plus 128.
+\item Otherwise, the output is equal to the input prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the input byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 183.
 \end{itemize}
-
-Byte arrays containing $2^{64}$ or more bytes cannot be encoded.
-This restriction ensures that the first byte of the encoding of a byte array is always below 192, and thus it can be readily distinguished from the encodings of sequences in $\mathbb{L}$.
 
@@
 (128 + \lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 56 \\
-\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{else if} \quad \lVert \mathbf{x} \rVert < 2^{64} \\
-\varnothing & \text{otherwise}
+\big(183 + \big\lVert \mathtt{BE}(\lVert \mathbf{x} \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert \mathbf{x} \rVert) \cdot \mathbf{x} & \text{otherwise}
 \end{cases} \\
-\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
-(x_1, ..., x_n) \cdot (y_1, ..., y_m) & = & (x_1, ..., x_n, y_1, ..., y_m)
+\mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{n < \lVert \mathbf{b} \rVert} b_{\mathrm{n}} \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
+(a) \cdot (b, c) \cdot (d, e) & = & (a, b, c, d, e)
 \end{eqnarray}
@@
 \item If the concatenated serialisations of each contained item is less than 56 bytes in length, then the output is equal to that concatenation prefixed by the byte equal to the length of this byte array plus 192.
-\item Otherwise, the output is equal to the concatenated serialisations, provided that they contain fewer than $2^{64}$ bytes, prefixed by the minimal-length byte array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
+\item Otherwise, the output is equal to the concatenated serialisations prefixed by the minimal-length byte-array which when interpreted as a big-endian integer is equal to the length of the concatenated serialisations byte array, which is itself prefixed by the number of bytes required to faithfully encode this length value plus 247.
 \end{itemize}
 
-Sequences whose concatenated serialized items contain $2^{64}$ or more bytes cannot be encoded.
-This restriction ensures that the first byte of the encoding does not exceed 255 (otherwise it would not be a byte).
-
 \hypertarget{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{}Thus we finish by formally defining $R_{\mathrm{l}}$:
@@
 R_{\mathrm{l}}(\mathbf{x}) & \equiv & \begin{cases}
-(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 56 \\
-\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{else if} \quad s(\mathbf{x}) \neq \varnothing \wedge \lVert s(\mathbf{x}) \rVert < 2^{64} \\
-\varnothing & \text{otherwise}
+(192 + \lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{if} \quad \lVert s(\mathbf{x}) \rVert < 56 \\
+\big(247 + \big\lVert \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \big\rVert \big) \cdot \mathtt{BE}(\lVert s(\mathbf{x}) \rVert) \cdot s(\mathbf{x}) & \text{otherwise}
 \end{cases} \\
-s(\mathbf{x}) & \equiv & \begin{cases}
-\mathtt{RLP}(\mathbf{x}[0]) \cdot \mathtt{RLP}(\mathbf{x}[1]) \cdot ... & \text{if} \quad \forall i: \mathtt{RLP}(\mathbf{x}[i]) \neq \varnothing \\
-\varnothing & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-If RLP is used to encode a scalar, defined only as a non-negative integer (in $\mathbb{N}$, or in $\mathbb{N}_x$ for any $x$), it must be encoded as the shortest byte array whose big-endian interpretation is the scalar.
+s(\mathbf{x}) & \equiv & \mathtt{RLP}(\mathbf{x}_0) \cdot \mathtt{RLP}(\mathbf{x}_1) ...
+\end{eqnarray}
+
+If RLP is used to encode a scalar, defined only as a non-negative integer ($\mathbb{N}$ or any $x$ for $\mathbb{N}_{\mathrm{x}}$), it must be specified as the shortest byte array such that the big-endian interpretation of it is equal.
 Thus the RLP of some non-negative integer $i$ is defined as:
@@
 It is defined in terms of a mutable data structure to map between 256-bit binary fragments and arbitrary-length binary data, typically implemented as a database.
-The core of the trie, and its sole requirement in terms of the protocol specification, is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
+The core of the trie, and its sole requirement in terms of the protocol specification is to provide a single value that identifies a given set of key-value pairs, which may be either a 32-byte sequence or the empty byte sequence.
 It is left as an implementation consideration to store and maintain the structure of the trie in a manner that allows effective and efficient realisation of the protocol.
 
-Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences with unique keys:
+Formally, we assume the input value $\mathfrak{I}$, a set containing pairs of byte sequences:
 \begin{equation}
@@
 \begin{equation}
-\forall I \in \mathfrak{I}: I \equiv (I_0, I_1)
+\forall_{\mathbf{I} \in \mathfrak{I}} I \equiv (I_0, I_1)
 \end{equation}
@@
 y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
-\forall n: \quad \forall i < 2\lVert\mathbf{k}_{n}\rVert: \quad \mathbf{k}_{n}'[i] & \equiv &
+\forall_{n} \quad \forall_{i: i < 2\lVert\mathbf{k}_{n}\rVert} \quad \mathbf{k}_{n}'[i] & \equiv &
 \begin{cases}
@@
 \item[Leaf] A two-item structure whose first item corresponds to the nibbles in the key not already accounted for by the accumulation of keys and branches traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be 1.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $true$.
 \item[Extension] A two-item structure whose first item corresponds to a series of nibbles of size greater than one that are shared by at least two distinct keys past the accumulation of the keys of nibbles and the keys of branches as traversed from the root.
-The hex-prefix encoding method is used and the second parameter to the function is required to be 0.
+The hex-prefix encoding method is used and the second parameter to the function is required to be $false$.
 \item[Branch] A 17-item structure whose first sixteen items correspond to each of the sixteen possible nibble values for the keys at this point in their traversal.
@@
 c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], 1), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], true), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
 \texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(j - 1)], 0), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
-(x - 1)] = \mathbf{l} \} \\
+(j - 1)], false), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \arg \max_{x} : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x : \forall_{I \in \mathfrak{I}}: I_0[0 ..
+(x - 1)] = \mathbf{l} \\
 \texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
@@
 
-Transactions are signed using recoverable ECDSA signatures.
+The method of signing transactions is similar to the `Electrum style signatures' as defined by \cite{npmElectrum2017}, heading "Managing styles with Radium" in the bullet point list.
 This method utilises the SECP-256k1 curve as described by \cite{Courtois2014}, and is implemented similarly to as described by \cite{gura2004comparing} on p.~9 of 15, para.~3.
@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak hash of the corresponding ECDSA public key:
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the right most 160-bits of the Keccak hash of the corresponding ECDSA public key:
 \begin{equation}
@@
 
-When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the leftmost 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
+When a 256-bit machine datum is converted to and from a 160-bit address or hash, the rightwards (low-order for BE) 20 bytes are used and the left most 12 are discarded or filled with zeroes, thus the integer values (when the bytes are interpreted as big-endian) are equivalent.
 
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{sha3}+G_{sha3word}\times\lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
 G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
@@
 0x37 & {\small CALLDATACOPY} & 3 & 0 & Copy input data in current environment to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
 \begin{cases} I_{\mathbf{d}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{d}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 0x39 & {\small CODECOPY} & 3 & 0 & Copy code running in current environment to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
 \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert I_{\mathbf{b}} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3c & {\small EXTCODECOPY} & 4 & 0 & Copy an account's code to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
 \begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
@@
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
+&&&& $\forall_{i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\} } \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
 \begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t,\\\quad \boldsymbol{\mu}_{\mathbf{s}}[0], I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(\mathbf{d}, E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
 \end{cases}

### 2019-11-16 Paper.tex
Fix syntax error.
Prefer MUST to SHOULD for requirements

--- 2019-08-16 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Shanghai version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Shanghai version \YellowPaperVersionNumber}}}
 \author{
@@
 This system can be said to be a very specialised version of a cryptographically secure, transaction-based state machine.
-Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications albeit rather simplistic ones.
+Follow-up systems such as Namecoin adapted this original ``currency application'' of the technology into other applications, albeit rather simplistic ones.
 
@@
 %E language?
-For a list of terms used in this paper, refer to Appendix \ref{ch:Terminology}.
+For a list of terms used in this paper, refer to Appendix~\ref{ch:Terminology}.
 
@@
 
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
-It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
+Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some current state.
+It is this current state which we accept as the canonical ``version'' of the world of Ethereum.
 The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
@@
 Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
-They also punctuate the transaction series with incentives for nodes to \textit{mine}.
-This incentivisation takes place as a state-transition function, adding value to a nominated account.
-
-Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block.
-It is achieved thanks to a cryptographically secure proof.
-This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.
 
@@
 B & \equiv & (..., (T_0, T_1, ...), ...) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...
+\end{eqnarray}
+
+Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function for transactions\footnote{Note that since the Shanghai fork, blocks also needs to process \textit{withdrawal operations} in order to reach their final state.
+Withdrawal operations are defined in sub-section \ref{subsec:The_Withdrawal}, and block final state is discussed in greater detail in section \ref{ch:finalisation}.}.
 
@@
 $10^0$ & Wei \\
+$10^9$ & Gwei \\
 $10^{12}$ & Szabo \\
@@
 
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
-This process is described in detail in section \ref{ch:ghost}.
-
-Sometimes, a path follows a new protocol from a particular height.
-This document describes one version of the protocol.
+Since the \textit{Paris} hard fork, reaching consensus on new blocks is managed by a protocol called the \textit{Beacon Chain}.
+It is known as the \textit{consensus layer} of Ethereum, and it defines the rules for determining the canonical history of Ethereum blocks.
+This document describes the \textit{execution layer} of Ethereum.
+The execution layer defines the rules for interacting with and updating the state of the Ethereum virtual machine.
+The consensus layer is described in greater detail in the \href{https://github.com/ethereum/consensus-specs}{consensus specifications}.
+How the consensus layer is used to determine the canonical state of Ethereum is discussed in section \ref{ch:blocktree_to_blockchain}.
+
+There are many versions of Ethereum, as the protocol has undergone a number of updates.
+These updates can be specified to occur:
+\begin{itemize}
+  \item at a particular block number in the case of pre-\textit{Paris} updates,
+  \item after reaching a \textit{terminal total difficulty} in the case of the \textit{Paris} update, or
+  \item at a particular block timestamp in the case of post-\textit{Paris} updates.
+\end{itemize}
+This document describes the \textit{Shanghai} version.
+
 In order to follow back the history of a path, one must reference multiple versions of this document.
+Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris, Shanghai, and every upcoming forks activated at a given block number (e.g.
+15,537,394 for Paris), the trigger was not the block number, but rather reaching a specified \textit{timestamp} (or \textit{total difficulty} for Paris).
+The trigger for the \textit{Paris} and subsequent hard fork are discussed in greater detail in section \ref{ch:pos_transition}.}
+\par
+\begin{center}
+\begin{tabular}{lr}
+\toprule
+Name & First Block Number \\
+\midrule
+$F_{\mathrm{Homestead}}$         &  1150000 \\
+$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
+$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
+$F_{\mathrm{Byzantium}}$         &  4370000 \\
+$F_{\mathrm{Constantinople}}$    &  7280000 \\
+$F_{\mathrm{Petersburg}}$        &  7280000 \\
+$F_{\mathrm{Istanbul}}$          &  9069000 \\
+$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
+$F_{\mathrm{Berlin}}$            & 12244000 \\
+$F_{\mathrm{London}}$            & 12965000 \\
+$F_{\mathrm{Arrow Glacier}}$     & 13773000 \\
+$F_{\mathrm{Gray Glacier}}$      & 15050000 \\
+$F_{\mathrm{Paris}}$             & 15537394 \\
+$F_{\mathrm{Shanghai}}$          & 17034870 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\par
+
+Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
+In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
+For the Ethereum main network
+\begin{equation}
+  \linkdest{chain_id}
+  \beta = 1
+\end{equation}
 
@@
 
-The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}).
+The world state (\textit{state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix~\ref{app:rlp}).
 Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}).
@@
 The hash is formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{s}}$.
-\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call; it is immutable and thus, unlike all other fields, cannot be changed after construction.
+\item[codeHash] The hash of the EVM code of this account---this is the code that gets executed should this address receive a message call.
 All such code fragments are contained in the state database under their corresponding hashes for later retrieval.
@@
 \begin{equation}
-\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
-\begin{equation}
-L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
+\texttt{TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
+\begin{equation}
+L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
 \end{equation}
@@
 
-This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
+This function, $L_{\mathrm{S}}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
 \begin{equation}
@@
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+The sender of a transaction cannot be a contract.
+While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably,
+such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
 \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
-There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+EIP-2718 by \cite{EIP-2718} introduced the notion of different transaction types.
+As of the \textit{London} version of the protocol, there are three transaction types: 0 (legacy), 1 (EIP-2930 by \cite{EIP-2930}), and 2 (EIP-1559 by \cite{EIP-1559}).
+Further, there are two subtypes of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
+All transaction types specify a number of common fields:
 
 \begin{description}
+\item[type]\linkdest{tx_type}{} EIP-2718 transaction type; formally $T_{\mathrm{x}}$.
 \item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
 \item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
@@
 \item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[r, s]\linkdest{T__r_T__s}{} Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 
-Additionally, a contract creation transaction contains:
+EIP-2930 (type 1) and EIP-1559 (type 2) transactions also have:
+\begin{description}
+\item[accessList]\linkdest{tx_access_list}{} List of access entries to warm up; formally $T_{\mathbf{A}}$.
+\linkdest{access_list_entry}{}Each access list entry $E$ is a tuple of an account address and a list of storage keys: $E \equiv (E_{\mathrm{a}}, E_{\mathbf{s}})$. 
+\item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
+\item[yParity]\linkdest{tx_y_parity}{} Signature Y parity; formally $T_{\mathrm{y}}$.
+\end{description}
+
+Legacy transactions do not have an \textbf{accessList} ($T_{\mathbf{A}}=()$), while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
+\begin{description}
+\item[w]\linkdest{T__w}{} A scalar value encoding Y parity and possibly chain ID; formally $T_{\mathrm{w}}$.
+$T_{\mathrm{w}} = 27 + T_{\mathrm{y}}$ or $T_{\mathrm{w}} = 2\beta + 35 + T_{\mathrm{y}}$ (see EIP-155 by \cite{EIP-155}).
+\end{description}
+
+There are differences in how one's acceptable gas price is specified in type 2 transactions versus type 0 and type 1 transactions.
+Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee}\footnote{The \textit{priority fee} is discussed in greater detail in sections \ref{ch:payment} and \ref{ch:transactions}.} that is paid.
+Type 2 transactions have the following two fields related to gas:
+
+\begin{description}
+  \item[maxFeePerGas]\linkdest{max_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{m}}$.
+  \item[maxPriorityFeePerGas]\linkdest{max_priority_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid to the block's fee recipient as an incentive to include the transaction; formally $T_{\mathrm{f}}$.
+\end{description}
+
+In contrast, type 0 and type 1 transactions specify gas price as a single value:
+
+\begin{description}
+  \item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.\footnote{Type 0 and type 1 transactions will get the same gas price behavior as a type 2 transaction with $T_{\mathrm{m}}$ and $T_{\mathrm{f}}$ set to the value of $T_{\mathrm{p}}$.}
+\end{description}
+
+Additionally, a contract creation transaction (regardless of transaction type) contains:
 
@@
 \begin{equation}
-L_{T}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+\linkdest{L_transaction} L_{\mathrm{T}}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 0 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{f}}, T_{\mathrm{m}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
+\end{cases}
+\end{equation}
+where
+\begin{equation}
+\mathbf{p} \equiv \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
+T_{\mathbf{d}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{A}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{x}} \in \{0, 1, 2\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B} & \wedge \\
+T_{\mathrm{m}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{f}} \in \mathbb{N}_{256}
 \end{array}
@@
 
+\subsection{The Withdrawal}\linkdest{withdrawal}\label{subsec:The_Withdrawal}
+A withdrawal (formally, \textit{W}) is a tuple of data describing a consensus layer validator's withdrawal of some amount of its staked Ether.
+A withdrawal is created and validated in the consensus layer of the blockchain and then pushed to the execution layer.
+A withdrawal is composed of the following fields:
+
+\begin{description}
+\item[globalIndex] zero based incrementing withdrawal index that acts as a unique identifier for this withdrawal; formally $W_{\mathrm{g}}$.
+\item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
+\item[recipient] the 20-byte address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
+\item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
+\end{description}
+
+Withdrawal serialisation is defined as:
+\begin{equation}
+\linkdest{L_Withdrawal} L_{\mathrm{W}}(W) \equiv 
+(W_{\mathrm{g}}, W_{\mathrm{v}}, W_{\mathrm{r}}, T_{\mathrm{a}})
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values except for $W_\mathbf{r}$ which is a 20-byte address:
+\begin{equation}
+\begin{array}[t]{lclc}
+W_{\mathrm{g}} \in \mathbb{N}_{64} & \wedge & W_{\mathrm{v}} \in \mathbb{N}_{64} & \wedge \\
+W_{\mathrm{r}} \in \mathbb{B}_{20} & \wedge & W_{\mathrm{a}} \in \mathbb{N}_{64}
+\end{array}
+\end{equation}
+
 \subsection{The Block}\linkdest{block}\label{subsec:The_Block}
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
-The block header contains several pieces of information:
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions,
+$\mathbf{T}$,\hypertarget{ommerheaders}{} a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}),
+and, since the \textit{Shanghai} hard fork, $\mathbf{W}$, a collection of validator's withdrawal pushed by the consensus layer. The block header contains several pieces of information:
 
@@
 \item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[ommersHash] A 256-bit hash field that is now deprecated due to the replacement of proof of work consensus. It is now to a constant, $\texttt{KEC}(\texttt{RLP}(()))$; formally $H_{\mathrm{o}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions and withdrawals are executed and finalisations applied; formally $H_{\mathrm{r}}$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
@@
 \item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
-\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
+\item[difficulty] A scalar field that is now deprecated due to the replacement of proof of work consensus. It is set to 0; formally $H_{\mathrm{d}}$.
 \item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks.
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
+\item[prevRandao]\linkdest{prevRandao_H__a}{}\linkdest{H__a} the latest RANDAO mix\footnote{RANDAO is a pseudorandom value generated by validators on the Ethereum consensus layer.
+Refer to the consensus layer specs (\url{https://github.com/ethereum/consensus-specs}) for more detail on RANDAO.} of the post beacon state of the previous block; formally $H_{\mathrm{a}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacement of proof of work consensus.
+It is set to \texttt{0x0000000000000000}; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
+\item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
+\item[withdrawalsRoot]\linkdest{withdrawals_root_H__w}{} The Keccak 256-bit hash of the root node of the trie structure populated with each withdrawal operations pushed by the consensus layer for this block; formally \hyperlink{H__w}{$H_{\mathrm{w}}$}.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other three components in the block are a series of transactions, $B_{\mathbf{T}}$, an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$, and a series of withdrawals, $B_{\mathbf{W}}$.
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
+B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}}, B_{\mathbf{W}})
 \end{equation}
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer.
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of five items comprising:
+the type of the transaction, $R_{\mathrm{x}}$,
+the status code of the transaction, $R_{\mathrm{z}}$,
+the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$,
+the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$}:
+\begin{equation}
+R \equiv (R_{\mathrm{x}}, R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+\end{equation}
+
+$R_{\mathrm{x}}$ is equal to the \hyperlink{tx_type}{type} of the corresponding transaction.
+
+\linkdest{L__R}The function $L_{\mathrm{R}}$ prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{\mathrm{R}}(R) \equiv (R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+\end{equation}
+
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer:
 \begin{equation}
@@
 
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
+%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{\mathrm{B}}$; it is merely a convenience equivalence.
 
@@
 \end{equation}
-
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{2048 $= 2^{11}$(11 bits), and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 \mathbf{y} & = & (0, 0, ..., 0) \quad \text{except:}\\
-\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
+\forall i \in \{0, 2, 4\}&:&\mathcal{B}_{2047 - m(\mathbf{x}, i)}(\mathbf{y}) = 1\\
 m(\mathbf{x}, i) &\equiv& \mathtt{KEC}(\mathbf{x})[i, i + 1] \bmod 2048
 \end{eqnarray}
-
 where $\mathcal{B}$ is the bit reference function such that $\mathcal{B}_{\mathrm{j}}(\mathbf{x})$ equals the bit of index $j$ (indexed from 0) in the byte array $\mathbf{x}$.
+Notably, it treats $\mathbf{x}$ as big-endian (more significant bits will have smaller indices).
 
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions:
+the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$ and withdrawals $B_{\mathbf{W}}$.
+For the header to be consistent with the transactions $B_{\mathbf{T}}$ and withdrawals $B_{\mathbf{W}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions, then all withdrawals, in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}),
+and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), \textbf{logsBloom} ($H_{\mathrm{b}}$) and \textbf{withdrawalsRoot} ($H_{\mathrm{w}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, the resulting logs, and the withdrawals, respectively.
 \begin{equation}
 \begin{array}[t]{lclc}
+B_{\mathbf{U}} &\equiv& () & \wedge \\
 \linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
+\linkdest{Withdrawals_Root_H__w}{}H_{\mathrm{w}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{W}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{W}}(i, B_{\mathbf{W}}[i])\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \end{equation}
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
-\begin{equation}
-p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
-\end{equation}
+
+where $p_{\mathrm{W}}(k, v)$ is a pairwise RLP transformation for withdrawals:
+\begin{equation}
+p_{\mathrm{W}}(k, W) \equiv \left( \mathtt{RLP}(k), \mathtt{RLP}(\hyperlink{L_Withdrawal}{L_{\mathrm{W}}}(W)) \right)
+\end{equation}
+
+similarly, $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
+\begin{equation}
+p_{\mathrm{T}}(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
+\mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
+(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+\end{cases}
+\right)
+\end{equation}
+and
+\begin{equation}
+p_{\mathrm{R}}(k, R) \equiv \left( \mathtt{RLP}(k), \begin{cases}
+\mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
+(R_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
+\end{cases}
+\right)
+\end{equation}
+($\cdot$ is the concatenation of byte arrays).
 
@@
 \begin{equation}
-\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
+\mathtt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{\mathrm{H}})$ is the parent block of $B$, defined directly.
 
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
-\begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
+We assert the types and order of the structure for when the RLP transformation is required:
+\begin{eqnarray}
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}}, H_{\mathrm{w}} \; )\end{array} \\
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}), L_{\mathrm{W}}^*(B_{\mathbf{W}}) \big)
+\end{eqnarray}
+where $\widetilde{L}_{\mathrm{T}}$ takes a special care of EIP-2718 transactions:
+\begin{equation}
+\widetilde{L}_{\mathrm{T}}(T) = \begin{cases}
+\hyperlink{L_transaction}{L_{\mathrm{T}}}(T) & \text{if} \quad T_{\mathrm{x}} = 0 \\
+(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+\end{cases}
+\end{equation}
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$, $L_{\mathrm{H}}^*$, and $L_{\mathrm{W}}^*$, being element-wise sequence transformations, thus:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 \hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{a}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
+\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N} & \wedge & \hyperlink{withdrawals_root_H__w}{H_{\mathrm{w}}} \in \mathbb{B}_{32}
+
 \end{array}
@@
 
-We define $P(B_{H})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+We define $P(B_{\mathrm{H}})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{\mathrm{H}})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
 \end{equation}
@@
 \begin{equation}
-H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
-\end{equation}
-
-\newcommand{\mindifficulty}{D_0}
+H_{\mathrm{i}} \equiv {{P(H)_{\mathrm{H}}}_{\mathrm{i}}} + 1
+\end{equation}
+
+\newcommand{\mindifficulty}{D_\mathrm{min}}
 \newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
@@
 
-\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
+\linkdest{H__f}The \textit{London} release introduced the block attribute \textit{base fee per gas} \hyperlink{block_baseFeePerGas_H__f}{$H_{\mathrm{f}}$} (see EIP-1559 by \cite{EIP-1559}).
+The base fee is the amount of wei burned per unit of gas consumed while executing transactions within the block.
+The value of the base fee is a function of the difference between the gas used by the parent block and the parent block's \textit{gas target}.
+
+The expected base fee per gas is defined as $F(H)$:
+
+\begin{equation}
+  F(H) \equiv \begin{cases} 
+    1000000000  & \text{if} \quad H_{\mathrm{i}} = F_{\mathrm{London}} \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}}  & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} = \tau \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}} - \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}} + \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau
+  \end{cases}
+  \end{equation}
+
 where:
-\begin{equation}
-\mindifficulty \equiv 131072
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-y \equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases}
-\end{equation*}
   \begin{align}
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
+    \tau &\equiv \lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{\rho} \rfloor \\
+    \rho &\equiv 2 \\
+    \nu^* &\equiv \begin{cases}
+      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({\tau - P(H)_{\mathrm{H}}}_{\mathrm{g}})}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({P(H)_{\mathrm{H}}}_{\mathrm{g}} - \tau)}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+  \end{cases} \\
+  \nu &\equiv \begin{cases}
+    \lfloor \frac{\nu^*}{\xi} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+    max(\lfloor \frac{\nu^*}{\xi} \rfloor, 1) &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+  \end{cases} \\
+  \xi &\equiv 8
 \end{align}
 
-Note that $\mindifficulty$ is the difficulty of the genesis block.
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+The \textit{gas target}, $\tau$, is defined as the gas limit $H_{\mathrm{l}}$ divided by the \textit{elasticity multiplier}, $\rho$, a global constant set to 2.
+So while blocks can consume as much gas as the gas limit, the base fee is adjusted so that on average, blocks consume as much gas as the gas target.
+The base fee is increased in the current block when the parent block's gas usage exceeds the gas target, and, conversely, the base fee is decreased in the current block when the parent block's gas usage is less than the gas target.
+
+The magnitude of the increase or decrease in the base fee, defined as $\nu$, is proportional to the difference between the amount of gas the parent block consumed and the parent block's gas target.
+The effect on the base fee is dampened by a global constant called the \textit{base fee max change denominator}, formally $\xi$, set to 8.
+A value of 8 entails that the base fee can increase or decrease by at most 12.5\% from one block to the next.
 
@@
 \begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000
@@
 
+where:
+  \begin{align}
+    {P(H)_{\mathrm{H}}}_{\mathrm{l}'} &\equiv \begin{cases}
+      {P(H)_{\mathrm{H}}}_{\mathrm{l}} \times \rho &  \text{if} \quad H_{\mathrm{i}} = F_{\mathrm{London}} \\
+      {P(H)_{\mathrm{H}}}_{\mathrm{l}} &  \text{if} \quad H_{\mathrm{i}} > F_{\mathrm{London}} \\
+  \end{cases}
+\end{align}
+
+To avoid a discontinuity in gas usage, the value of the parent block gas limit for the purpose of validating the current block's gas limit is modified at the \textit{London} fork block by multiplying it by the \textit{elasticity multiplier}, $\rho$.
+We call this modified value ${P(H)_{\mathrm{H}}}_{\mathrm{l}'}$.
+This ensures that the gas target for post-\textit{London} blocks can be set roughly in line with the gas limit of pre-\textit{London} blocks.
+
 \hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
 \begin{equation}
-H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
-\end{equation}
-
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
-
-The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
-
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}}
+\end{equation}
+
+The \textit{Paris} hard fork changed Ethereum's consensus from proof of work to proof of stake, and thus deprecated many block header properties related to proof of work.
+These deprecated properties include \textbf{nonce} ($H_{\mathbf{n}}$), \textbf{ommersHash} ($H_{\mathbf{o}}$), \textbf{difficulty} ($H_{\mathbf{d}}$), and \textbf{mixHash} ($H_{\mathbf{m}}$).
+
+\textbf{mixHash} has been replaced with a new field \textbf{prevRandao} ($H_{\mathbf{a}}$). The other header fields related to proof of work have been replaced with constants:
+
+\begin{eqnarray}
+  H_{\mathrm{o}} & \equiv & \texttt{KEC}(\texttt{RLP}(())) \\
+  H_{\mathrm{d}} & \equiv  & 0 \\
+  H_{\mathrm{n}}  & \equiv & \texttt{0x0000000000000000}
+\end{eqnarray}
+
+The value of \textbf{prevRandao} must be determined using information from the Beacon Chain.
+While the details of generating the \textit{RANDAO} value on the Beacon Chain is beyond the scope of this paper, we refer to the expected \textit{RANDAO} value for the previous block as $\mathtt{PREVRANDAO}()$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
-\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+V(H) & \equiv & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
-\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
-
-Noting additionally that \textbf{extraData} must be at most 32 bytes.
+\nonumber& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
+\nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
+\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32 \quad \wedge \\
+\nonumber& & H_{\mathrm{f}} = F(H) \wedge \\
+\nonumber& & H_{\mathrm{o}} = \texttt{KEC}(\texttt{RLP}(())) \quad \wedge \\
+\nonumber& & H_{\mathrm{d}} = 0 \quad \wedge \\
+\nonumber& & H_{\mathrm{n}} = \texttt{0x0000000000000000} \quad \wedge \\
+\nonumber& & H_{\mathrm{a}} = \mathtt{PREVRANDAO}()
+\end{eqnarray}
+
+Note additionally that \textbf{extraData} must be at most 32 bytes.
 
@@
 This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according \textbf{gasPrice}, also specified in the transaction.
+The purchase happens at the \textit{effective gas price} defined in section \ref{ch:transactions}.
 The transaction is considered invalid if the account balance cannot support such a purchase.
@@
 
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
-Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
-Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
-Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.
+Since the introduction of EIP-1559 by \cite{EIP-1559} in the \textit{London} hard fork, every transaction included in a block must pay a \textit{base fee}, which is specified as wei per unit of gas consumed and is constant for each transaction within a block.
+The ether that is paid to meet the base fee is burned (taken out of circulation).
+The base fee adjusts dynamically as a function of the previous block's gas consumption relative to its \textit{gas target} (a value that is currently half the block's gas limit, which can be adjusted by validators).
+If the previous block's total gas consumption exceeds the gas target, this indicates excess demand for block space at the current base fee, and the base fee is increased.
+Conversely, if the gas consumed in the previous block is lower than the gas target, demand for block space is lower than the gas target at the current base fee, and thus the base fee is decreased.
+This process of adjusting the base fee should bring the average block's gas consumption in line with the gas target.
+Refer to section \ref{subsec:The_Block} for greater detail on how the base fee is set.
+
+To incentivize validators to include transactions, there is an additional fee known as a \textit{priority fee}, which is also specified as wei per unit of gas consumed.
+The total fee paid by the transactor therefore is the sum of the base fee per gas and the priority fee per gas multiplied by the total gas consumed.
+Ether used to satisfy the priority fee is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the validator.
+
+Transactors using type 2 transactions can specify the maximum priority fee they are willing to pay (\textbf{maxPriorityFeePerGas}) as well as the max total fee they are willing to pay (\textbf{maxFeePerGas}), inclusive of both the priority fee and the base fee.
+\textbf{maxFeePerGas} must be at least as high as the base fee for the transaction to be included in a block.
+Type 0 and type 1 transactions have only one field for specifiying a gas price--\textbf{gasPrice}--which also must be at least as high as the base fee for inclusion in a block.
+The amount by which \textbf{gasPrice} is higher than the base fee constitutes the priority fee in the case of a type 0 or type 1 transaction.
+
+Transactors are free to select any priority fee that they wish, however validators are free to ignore transactions as they choose.
+A higher priority fee on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the validator and thus will more likely be selected for inclusion.
+Since there will be a (weighted) distribution of minimum acceptable priority fees, transactors will necessarily have a trade-off to make between lowering the priority fee and maximising the chance that their transaction will be included in a block in a timely manner.
 
@@
 \item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
+\item the sender account has no contract code deployed (see EIP-3607 by \cite{EIP-3607});
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
+\item the sender account balance contains at least the cost, $v_0$, required in up-front payment;
+\item the \textbf{maxFeePerGas}, $T_m$, in the case of type 2 transactions, or \textbf{gasPrice}, $T_p$, in the case of type 0 and type 1 transactions, is greater than or equal to the block's base fee, $H_{\mathrm{f}}$; and
+\item for type 2 transactions, \textbf{maxPriorityFeePerGas}, $T_f$, must be no larger than \textbf{maxFeePerGas}, $T_m$.
 \end{enumerate}
@@
 
-\subsection{Substate}
+\subsection{Substate} \label{ch:substate}
 Throughout transaction execution, we accrue certain information that is acted upon immediately following the transaction.
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}}, A_{\mathbf{a}}, A_{\mathbf{K}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+\hypertarget{refund_balance_defn_words_A__r}{}$A_{\mathrm{r}}$ is the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
-
-We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing,(), \varnothing, 0)
-\end{equation}
+Finally, EIP-2929 by \cite{EIP-2929} introduced \hypertarget{accessed_addresses_defn_words_A__a}{}$A_{\mathbf{a}}$, the set of accessed account addresses, and \hypertarget{accessed_storage_keys_defn_words_A__k}{}$A_{\mathbf{K}}$, the set of accessed storage keys
+(more accurately, each element of $A_{\mathbf{K}}$ is a tuple of a 20-byte account address and a 32-byte storage slot).
+
+We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts, zero refund balance, all precompiled contracts in the accessed addresses, and no accessed storage:
+\begin{equation}
+A^0 \equiv (\varnothing, (), \varnothing, 0, \pi, \varnothing)
+\end{equation}
+where $\hyperlink{precompiled_set}{\pi}$ is the set of all precompiled addresses.
 
@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
+\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{R({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
+\nonumber {} & + \sum_{j = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \big( G_{\mathrm{accesslistaddress}} + \lVert T_{\mathbf{A}}[j]_{\mathbf{s}} \rVert G_{\mathrm{accessliststorage}} \big)
 \end{align}
-
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if it is a message call.
+$\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
 $G$ is fully defined in Appendix \ref{app:fees}.
 
+\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $R$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:
+
+\begin{equation}
+  R(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
+\end{equation}
+
+\hypertarget{effective_gas_price_p}{}We define the \textit{effective gas price}, formally $p$, as the amount of wei the transaction signer will pay per unit of gas consumed during the transaction's execution.
+It is calculated as follows:
+
+\begin{equation}
+  p \equiv \begin{cases}
+  T_{\mathrm{p}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  f + H_{\mathrm{f}} & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
+\end{equation}
+
+\hypertarget{priority_fee_f}{}where $f$ is the \textit{priority fee}---the amount of wei the block's beneficiary address will receive per unit of gas consumed during the transaction's execution.
+It is calculated as:
+
+\begin{equation}
+  f \equiv \begin{cases}
+  T_{\mathrm{p}} - H_{\mathrm{f}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  min(T_{\mathrm{f}}, T_{\mathrm{m}} - H_{\mathrm{f}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
+\end{equation}
+
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+  v_0 \equiv \begin{cases}
+  \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}}T_{\mathrm{p}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}}T_{\mathrm{m}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}} & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
 \end{equation}
@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
+\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} & = & \texttt{KEC}\big( () \big) \quad \wedge \\
 \linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+m & \geqslant & H_{\mathrm{f}} \quad \wedge \\
+n & \leqslant & 49152 \quad \wedge \\
+T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
 \end{equation}
-
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
-
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+where
+\begin{equation}
+  m \equiv \begin{cases}
+  T_{\mathrm{p}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  T_{\mathrm{m}} & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
+\end{equation}
+and
+\begin{equation}
+  n \equiv \begin{cases}
+    \lVert T_{\mathrm{i}} \rVert & \text{if} \; T_{\mathrm{t}} \neq \varnothing \\
+    0 & \text{otherwise}
+  \end{cases}
+\end{equation}
+
+The penultimate condition ensures that, for create transactions, the length of the initcode is no greater than 49152 bytes.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
+Also, with a slight abuse of notation, we assume that $\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} = \texttt{KEC}\big( () \big)$, $\boldsymbol{\sigma}[S(T)]_{\mathrm{n}} = 0$, and $\boldsymbol{\sigma}[S(T)]_{\mathrm{b}} = 0$ if $\boldsymbol{\sigma}[S(T)] = \varnothing$.
+
+For type 2 transactions, we add an additional check that \textbf{maxPriorityFeePerGas} is no larger than \textbf{maxFeePerGas}:
+\begin{equation}
+  T_{\mathrm{m}} \geqslant T_{\mathrm{f}}
+\end{equation}
+
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}} p$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\linkdest{sigma_0}{}\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} p \\
 \boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
+\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; p, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, p, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\end{cases}
+\end{equation}
+where
+\begin{align}
+A^* & \equiv A^0 \quad \text{except} \\
+A^*_{\mathbf{K}} & \equiv \bigcup_{E \in T_{\mathbf{A}}} \big\{ \forall i < \lVert E_{\mathbf{s}} \rVert, i \in \mathbb{N}: \; (E_{\mathrm{a}}, E_{\mathbf{s}}[i]) \big\} \\
+A^*_{\mathbf{a}} & \equiv \begin{cases}
+a \cup T_{\mathrm{t}} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+a & \text{otherwise}
+\end{cases} \\
+a & \equiv A^0_{\mathbf{a}} \cup S(T) \cup H_c \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
+\end{align}
+and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
 \begin{equation}
@@
 \end{equation}
-and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
-
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
-
-After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
-\begin{equation}
-	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
-\end{equation}
+
+Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{5} \Big\rfloor, A_{\mathrm{r}} \right\}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of one fifth\footnote{The max refundable proportion of gas was reduced from one half to one fifth by EIP-3529 by \cite{EIP-3529} in the \textit{London} release} (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
 
-The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{P}$:
-\begin{eqnarray}
-\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{P} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{P}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{H}}_{\mathrm{c}}
+The validator, whose address is specified as the beneficiary of the present block $B$, receives the gas consumed multiplied by the transaction's \textit{priority fee per gas}, defined as \hyperlink{priority_fee_f}{$f$} in this section.
+The ether that is paid by the transactor that goes toward the base fee is debited from the transactor's account but credited to no other account, so it is burned.
+
+We define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$:
+\begin{eqnarray}
+\boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{\mathrm{P}} \quad \text{except} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* p \\
+\boldsymbol{\sigma}^*[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) f
 \end{eqnarray}
@@
 
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{sigma_n}{later} for state validation.
 
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
-
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
-\end{equation}
-
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor\footnote{which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code} ($o$),
+available gas ($g$), effective gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
+The salt \linkdest{salt}{$\zeta$} might be missing ($\zeta = \varnothing$); formally,
+\begin{equation}
+\zeta \in \mathbb{B}_{32} \cup \mathbb{B}_{0}
+\end{equation}
+If the creation was caused by {\small \hyperlink{create2}{CREATE2}}, then $\zeta \neq \varnothing$.
+
+We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued substate $A$, to the tuple containing the new state, remaining gas, new accrued substate, status code and output $(\boldsymbol{\sigma}', g', A', z, \mathbf{o})$:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, A, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
+\end{equation}
+
+The address of the new account is defined as being the rightmost 160 bits of the Keccak-256 hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
+For {\small \hyperlink{create2}{CREATE2}} the rule is different and is described in EIP-1014 by \cite{EIP-1014}.
+Combining the two cases, we define the resultant address for the new account $a$:
+\begin{align}
+a & \equiv \mathtt{ADDR}(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
+\label{eq:new-address} \mathtt{ADDR}(s, n, \zeta, \mathbf{i}) & \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big( L_{\mathrm{A}}(s, n, \zeta, \mathbf{i})\big) \Big) \\
+L_{\mathrm{A}}(s, n, \zeta, \mathbf{i}) & \equiv \begin{cases}
+\mathtt{RLP}\big(\;(s, n)\;\big) & \text{if}\ \zeta = \varnothing \\
+(255) \cdot s \cdot \zeta \cdot \mathtt{KEC}(\mathbf{i}) & \text{otherwise}
+\end{cases}
+\end{align}
+where $\cdot$ is the concatenation of byte arrays, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+
+The address of the new account is added to the set of accessed accounts:
+\begin{equation}
+A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{a\}
+\end{equation}
 
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
+As such, the code execution function $\hyperlink{xi_def}{\Xi}$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the resultant accrued substate $A^{**}$ and the body code of the account $\mathbf{o}$.
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, A^*, I) \\
 \end{equation}
@@
 \begin{equation}
-c \equiv G_{codedeposit} \times \lVert \mathbf{o} \rVert
+c \equiv G_{\mathrm{codedeposit}} \times \lVert \mathbf{o} \rVert
 \end{equation}
@@
 The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
+However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas, thus the contract's code is not stored.
 
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A', z)$ where:
 
 \begin{align}
-\quad g' &\equiv \begin{cases}
+\quad g' \equiv & \begin{cases}
 0 & \text{if} \quad F \\
@@
 \end{cases} \\
-\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \\
+\quad \boldsymbol{\sigma}' \equiv & \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+\quad A' \equiv & \begin{cases}
+A^* & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
+A^{**} & \text{otherwise} \\
+\end{cases} \\
+\quad z \equiv & \begin{cases}
+0 & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  \lVert \mathbf{o} \rVert > 24576\big)
+F \equiv  & \big( \boldsymbol{\sigma}[a] \neq \varnothing \ \wedge\ \big(\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}\big(()\big) \vee \boldsymbol{\sigma}[a]_n \neq 0 \big) \big) \quad \vee \\
+\nonumber &(\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \quad \vee \\
+\nonumber &g^{**} < c \quad \vee \\
+\nonumber &\lVert \mathbf{o} \rVert > 24576 \quad \vee \\
+\nonumber &\mathbf{o}[0] = \texttt{0xef}
 \end{align}
+\hypertarget{contract_creation_result}{}
+
+Note the last condition of $F$ indicates that contract code cannot begin with the byte \texttt{0xef} (refer to EIP-3541 by \cite{EIP-3541}),
 
@@
 Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+In addition, observe that if the execution of the initialising code \hyperlink{REVERT}{reverts} ($\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} \neq \varnothing$), the resultant gas $g'$ is not depleted (provided there was no other exception), but no new account is created.
 
 \subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address should return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
+Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address must return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
 Thus any message call received by it during this time causes no code to be executed.
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
-
-Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and effective gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
+
+Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array~$\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, A, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 
-The account's associated code (identified as the fragment whose Keccak hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
+The account's associated code (identified as the fragment whose Keccak-256 hash is $\boldsymbol{\sigma}[c]_{\mathrm{c}}$) is executed according to the execution model (see section \ref{ch:model}).
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
@@
 g^{**} & \text{otherwise}
-\end{cases} \\ \nonumber
+\end{cases} \\
+A' & \equiv & \begin{cases}
+A & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+A^{**} & \text{otherwise}
+\end{cases} \\
 z & \equiv & \begin{cases}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
+\hypertarget{code_execution_result}{}
+(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) & \equiv & \Xi\\
 I_{\mathrm{a}} & \equiv & r \\
@@
 I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
+I_{\mathrm{w}} & \equiv & w
 \end{eqnarray}
@@
 \Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\begin{equation}
-\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
+\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 1 \\
+\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 2 \\
+\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 3 \\
+\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 4 \\
+\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 5 \\
+\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 6 \\
+\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 7 \\
+\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 9 \\
+\Xi                     (\boldsymbol{\sigma}_1, g, A, I) & \text{otherwise} \end{cases}
+\end{equation}
+and
+\begin{equation}
+\mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
 It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
 
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
-
+As can be seen, there are nine exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The contracts in addresses 1 to 9 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication, an elliptic curve pairing check, and the BLAKE2 compression function $\mathtt{F}$ respectively.
 Their full formal definition is in Appendix \ref{app:precompiled}.
+\hypertarget{precompiled_set}{}We denote the set of the addresses of the precompiled contracts by $\pi$:
+\begin{equation}
+\pi \equiv \{1, 2, 3, 4, 5, 6, 7, 8, 9 \}
+\end{equation}
 
@@
 The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CREATE2}, {\small CALL} and {\small CALLCODE}.
 Finally, gas may be paid due to an increase in the usage of the memory.
@@
 
-In addition to the system state $\boldsymbol{\sigma}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
+In addition to the system state $\boldsymbol{\sigma}$, the remaining gas for computation $g$, and the accrued substate $A$,
+there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
 
@@
 \item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas paid by the signer of the transaction that originated this execution.
+This is defined as the effective gas price \hyperlink{effective_gas_price_p}{$p$} in section \ref{ch:transactions}.
 \item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
@@
 \item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathrm{H}}$, the block header of the present block.
 \item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
+the number of {\small CALL}s or {\small CREATE(2)}s being executed at present).
 \item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the accrued substate $A$ and the resultant output, $\mathbf{o}$, given these definitions.
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the resultant accrued substate $A'$ and the resultant output, $\mathbf{o}$, given these definitions.
 For the present context, we will define it as:
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, I)
-\end{equation}
-
-where we will remember that $A$, the accrued substate, is defined as the tuple of the self-destructs set $A_{\mathbf{s}}$, the log series $A_{\mathbf{l}}$, the touched accounts $A_{\mathbf{t}}$ and the refund balance $A_{\mathrm{r}}$:
-\begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
+(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, A, I)
+\end{equation}
+where we will remember that $A$, the accrued substate, is defined in section \ref{ch:substate}.
 
@@
 
+\linkdest{xi_def}
 We must now define the $\Xi$ function.
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) \\
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A^0, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A^0, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+\big(\varnothing, \boldsymbol{\mu}, A, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
+\big(\varnothing, \boldsymbol{\mu}', A, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
 O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
@@
 \boldsymbol{\mu}' & \equiv & \boldsymbol{\mu}\ \text{except:} \\
-\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I)
+\boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)
 \end{eqnarray}
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s}, \mathbf{o})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the stack contents, and the returndata buffer.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 \begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
+( w = \text{\small JUMP} \; \wedge \; \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small JUMPI} \; \wedge \; \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \\
 \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
- (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
+( w = \text{\small RETURNDATACOPY} \; \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+( \neg I_{\mathrm{w}} \; \wedge \; W(w, \boldsymbol{\mu}) ) \quad \vee \\
+( w = \text{\small SSTORE} \; \wedge \; \boldsymbol{\mu}_g \leqslant G_{\mathrm{callstipend}} )
 \end{array}
@@
 \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+w \in \{\text{\small CREATE}, \text{\small CREATE2}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
+\text{\small LOG0} \le w \; \wedge \; w \le \text{\small LOG4} \quad \vee \\
+w = \text{\small CALL} \; \wedge \; \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{SSTORE}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$
+-- see EIP-2200 by \cite{EIP-2200} for more information.
 
@@
 \begin{equation}
-D_{J}(\mathbf{c}, i) \equiv \begin{cases}
+D_{\mathrm{J}}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant \lVert \mathbf{c} \rVert  \\
-\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
+\{ i \} \cup D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
 \quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \label{eq:mu_pc}\\
 \quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
@@
 
-In general, we assume the memory, self-destruct set and system state don't change:
+In general, we assume the memory, accrued substate and system state do not change:
 \begin{eqnarray}
@@
 
-\section{Blocktree to Blockchain} \label{ch:ghost}
-
-The canonical blockchain is a path from root to leaf through the entire block tree.
-In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{Genesis_Block}{genesis block}, in the path.
-The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
-This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
-
-Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.
-
-Thus we define the total difficulty of block $B$ recursively as:
-\begin{eqnarray}
-B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
-B' & \equiv & P(B_{H})
-\end{eqnarray}
-
-As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
+\section{Transition to Proof of Stake} \label{ch:pos_transition}
+The \textit{Paris} hard fork changed the underlying consensus mechanism of Ethereum from proof of work to proof of stake.
+
+Unlike all previous hard forks of Ethereum, \textit{Paris} was not defined to occur at a particular block height, but rather after a specified \textit{terminal total difficulty} was reached.
+Total difficulty was used instead of block height to avoid a scenario in which a minority of hash power could create a malicious fork that could race to satisfy the block height requirement and claim the first proof of stake block.
+
+Thus the \textit{terminal block}, the last proof of work block before the \textit{Paris} fork takes effect, is defined as having:
+\begin{eqnarray}
+  B_{\mathrm{t}} & \geqslant & 58750000000000000000000 \\
+  P(B_{\mathrm{H}})_{\mathrm{t}} & < & 58750000000000000000000
+\end{eqnarray}
+
+where $B_{\mathrm{t}}$ is the total difficulty of block $B$ and $P(B_{\mathrm{H}})_{\mathrm{t}}$ is the total difficulty of its parent.
+
+Total difficulty for a proof of work (pre-\textit{Paris}) block $B$ is defined recursively as:
+\begin{eqnarray}
+B_{\mathrm{t}} & \equiv & P(B_{\mathrm{H}})_{\mathrm{t}} + H_{\mathrm{d}}
+\end{eqnarray}
+
+where $H_{\mathrm{d}}$ is the difficulty of the current block $B$.
+
+Upon reaching the terminal block, new blocks are processed by the Beacon Chain.
+
+\subsection{Post-Paris Updates}
+Because the Beacon Chain generate a new slot every 12 seconds, post-\textit{Paris} updates can be scheduled to occur at a specific timestamp.
+At the execution layer, the update will then happen in the first produced block after the scheduled timestamp.
+For example the \textit{Shanghai} hard fork was scheduled to occur at 2023-04-12 10:27:35 UTC, on Epoch 194,048.
+Validators failed to propose a block during the two first slots of this Epoch, but at slot 6,209,538 a validator finally proposed the block 17,034,870, marking the transition to \textit{Shanghai} on the execution layer.
+
+\section{Blocktree to Blockchain} \label{ch:blocktree_to_blockchain}
+
+Prior to the transition to proof of stake at the \textit{Paris} hard fork, the canonical blockchain was defined as the block path with the greatest \textit{total difficulty}, defined in section \ref{ch:pos_transition} as $B_{\mathrm{t}}$.
+
+After reaching the \textit{terminal block} described in section \ref{ch:pos_transition}, the greatest \textit{total difficulty} rule must be removed in favor of a rule known as \textit{LMD Ghost}.\footnote{\textit{LMD GHOST} comprises two acronyms, "Latest Message Driven", and "Greedy Heaviest-Observed Sub-Tree".}
+
+Note that in order to determine what blocks comprise the canonical Ethereum blockchain after \textit{Paris}, one must have additional information from the Beacon Chain, which is not described herein.
+We denote events emitted by the Beacon Chain with the prefix $\mathtt{POS\_}$.
+
+On each occurrence of a $\mathtt{POS\_FORKCHOICE\_UPDATED}$ event, starting with the first at the \textit{transition block} described in section \ref{ch:pos_transition}, the canonical chain is defined as the chain beginning with the genesis block and ending at the block nominated by the event as the head of the chain.
+
+The head of the chain should be updated if and only if a $\mathtt{POS\_FORKCHOICE\_UPDATED}$ is emitted, in which case the head should set to the block specified by the event.
+No optimistic updates to the head of the chain should be made.
+
+The $\mathtt{POS\_FORKCHOICE\_UPDATED}$ event additionally references a finalized block.
+The most recent finalized block should be set to this block.
+
+The canonical blockchain must also contain a block with the hash and number of the \textit{terminal block} defined in section \ref{ch:pos_transition}.
 
@@
 
-The process of finalising a block involves four stages:
+The process of finalising a block involves three stages:
 
 \begin{enumerate}
-\item Validate (or, if mining, determine) ommers;
-\item validate (or, if mining, determine) transactions;
-\item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
+\item executing withdrawals;
+\item validate transactions;
+\item verify state.
 \end{enumerate}
 
-\subsection{Ommer Validation}
-
-The validation of \hyperlink{ommer_block_headers_B__U}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
-The maximum of ommer headers is two.
-Formally:
-\begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
-\end{equation}
-
-where $k$ denotes the ``is-kin'' property:
-\begin{equation}
-k(U, H, n) \equiv \begin{cases} \mathit{false} & \text{if} \quad n = 0 \\
-s(U, H) &\\
-\quad \vee \; k(U, P(H)_{H}, n - 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-and $s$ denotes the ``is-sibling'' property:
-\begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathbf{U}})
-\end{equation}
-where $B(H)$ and $P(H)$ are the block and the parent block of the corresponding header $H$ respectively.
+\subsection{Executing Withdrawals}
+After processing the block's transactions, the withdrawals are executed.
+A withdrawal is simply an increase of the recipient account's balance of the specified Gwei amount.
+No other balances are decreased, a withdrawal is not a transfer but a creation of funds.
+A withdrawal operation cannot fail and has no gas cost. We define the function $E$ as the withdrawal state transition function:
+
+\begin{eqnarray}
+E(\boldsymbol{\sigma}_w, W) & \equiv & \boldsymbol{\sigma}_{w+1} \\
+\boldsymbol{\sigma}_{w+1} & \equiv & \boldsymbol{\sigma}_w \quad \text{except:} \\
+\boldsymbol{\sigma}_{w+1}[W_r]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{w}[W_r]_{\mathrm{b}} + (W_a \times 10^9)
+\end{eqnarray}
+
+\hypertarget{Kappa}{}Finally, we define $K$, the block-level state transition function for withdrawals:
+
+\begin{equation}
+K(\boldsymbol{\sigma}, B) \equiv E(E(\boldsymbol{\sigma}, W_0), W_1) ...
+\end{equation}
 
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
-\end{equation}
-
-\subsection{Reward Application}
-
-The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
-\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
-\\ \nonumber
-\forall \mathbf{U} \in \mathbf{B}_{\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
-\end{eqnarray}
-
-If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
-
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
-\end{equation}
-
-\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{\mathrm{H}}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H__g}{B_{\mathrm{H}}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+\end{equation}
+
+\subsection{State Validation}\label{sec:statenoncevalidation}
 
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{\mathrm{H}}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 \begin{eqnarray}
-\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
-
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+\Phi(B) & \equiv & B': \quad B' = B \quad \text{except:} \\
+{B'_{\mathrm{H}}}_{\mathrm{r}} & = & \mathtt{TRIE}(L_{\mathrm{S}}(\hyperlink{Kappa}{K}(\hyperlink{Pi}{\Pi}(\Gamma(B), B), B)))
+\end{eqnarray}
+
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function.
 
@@
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}
+\begin{equation}\linkdest{sigma_n}{}
 \boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
-\end{equation}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
-
-\subsection{Mining Proof-of-Work} \label{ch:pow}
-
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
-However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
-
-For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of, or reward from, specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
-
-Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
-Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
-
-One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
-While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
-Because of this, a proof-of-work function that is ASIC-resistant (i.e.
-difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
-
-Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the \hyperlink{block_nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
-
-More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
-\begin{equation}
-m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
-\subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \ell(\boldsymbol{\sigma})
+\end{equation}
+
+Thus the complete block-transition mechanism (before consensus) is defined.
 
@@
 However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+Such data might include the block's hash and the block's beneficiary address.
+In order to make it hard for malicious validators to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
-Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full, mining, node.
+Blockchain consolidation could be used in order to reduce the amount of blocks a client would need to download to act as a full node.
 A compressed archive of the trie structure at given points in time (perhaps one in every 10,000th block) could be maintained by the peer network, effectively recasting the \hyperlink{Genesis_Block}{genesis block}.
@@
 Finally, blockchain compression could perhaps be conducted: nodes in state trie that haven't sent/received a transaction in some constant amount of blocks could be thrown out, reducing both Ether-leakage and the growth of the state database.
-
-\subsection{Scalability}
-
-Scalability remains an eternal concern.
-With a generalised state transition function, it becomes difficult to partition and parallelise transactions to apply the divide-and-conquer strategy.
-Unaddressed, the dynamic value-range of the system remains essentially fixed and as the average transaction value increases, the less valuable of them become ignored, being economically pointless to include in the main ledger.
-However, several strategies exist that may potentially be exploited to provide a considerably more scalable protocol.
-
-Some form of hierarchical structure, achieved by either consolidating smaller lighter-weight chains into the main block or building the main block through the incremental combination and adhesion (through proof-of-work) of smaller transaction sets may allow parallelisation of transaction combination and block-building.
-Parallelism could also come from a prioritised set of parallel blockchains, consolidating each block and with duplicate or invalid transactions thrown out accordingly.
-
-Finally, verifiable computation, if made generally available and efficient enough, may provide a route to allow the proof-of-work to be the verification of final state.
 
@@
 Paid for exclusively by Ether (as of PoC-4), which is converted freely to and from Gas as required.
-Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and miners are free to ignore Transactions whose Gas price is too low.
+Gas does not exist outside of the internal Ethereum computation engine; its price is set by the Transaction and validators are free to ignore Transactions whose Gas price is too low.
 
@@
 \end{cases} \\
+\label{eq:BE}
 \mathtt{BE}(x) & \equiv & (b_0, b_1, ...): b_0 \neq 0 \wedge x = \sum_{n = 0}^{\lVert \mathbf{b} \rVert - 1} b_n \cdot 256^{\lVert \mathbf{b} \rVert - 1 - n} \\
@@
 
-It is defined as the function $\mathtt{HP}$ which maps from a sequence of nibbles (represented by the set $\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
+It is defined as the function $\mathtt{HP}$ which maps from a sequence of nibbles (represented by the set \linkdest{set_of_sequence_of_nibbles}$\mathbb{Y}$) together with a boolean value to a sequence of bytes (represented by the set $\mathbb{B}$):
 
@@
 \begin{eqnarray}
-y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \mathbb{Y}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
+y(\mathfrak{I}) & = & \{ (\mathbf{k}_0' \in \hyperlink{set_of_sequence_of_nibbles}{\mathbb{Y}}, \mathbf{v}_0 \in \mathbb{B}), (\mathbf{k}_1' \in \mathbb{Y}, \mathbf{v}_1 \in \mathbb{B}), ... \} \\
 \forall n: \quad \forall i < 2\lVert\mathbf{k}_{n}\rVert: \quad \mathbf{k}_{n}'[i] & \equiv &
@@
 \begin{equation}
-\texttt{TRIE}(\mathfrak{I}) \equiv \texttt{KEC}(c(\mathfrak{I}, 0))
+\texttt{TRIE}(\mathfrak{I}) \equiv \texttt{KEC}\big(\texttt{RLP} (c(\mathfrak{I}, 0))\big)
 \end{equation}
@@
 When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
+As a means of reducing storage complexity, we store nodes whose composed RLP is fewer than 32 bytes directly; for those larger we assert prescience of the byte array whose Keccak-256 hash evaluates to our reference.
 Thus we define in terms of $c$, the node composition function:
@@
 n(\mathfrak{I}, i) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
+() \in \mathbb{B} & \text{if} \quad \mathfrak{I} = \varnothing \\
+c(\mathfrak{I}, i) & \text{if} \quad \lVert \, \texttt{RLP} \big( c(\mathfrak{I}, i) \big) \rVert < 32 \\
+\texttt{KEC}\big(\texttt{RLP}( c(\mathfrak{I}, i)) \big) & \text{otherwise}
 \end{cases}
@@
 c(\mathfrak{I}, i) \equiv \begin{cases}
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], 1), I_1 \big) \Big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
-\texttt{RLP}\Big( \big(\texttt{HP}(I_0[i ..
-(j - 1)], 0), n(\mathfrak{I}, j) \big) \Big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
+ \big(\texttt{HP}(I_0[i .. (\lVert I_0\rVert - 1)], 1), I_1 \big) & \text{if} \quad \lVert \mathfrak{I} \rVert = 1 \quad \text{where} \; \exists I: I \in \mathfrak{I} \\
+\big(\texttt{HP}(I_0[i ..
+(j - 1)], 0), n(\mathfrak{I}, j) \big) & \text{if} \quad i \ne j \quad \text{where} \; j = \max \{ x : \exists \mathbf{l}: \lVert \mathbf{l} \rVert = x \wedge \forall I \in \mathfrak{I}: I_0[0 ..
 (x - 1)] = \mathbf{l} \} \\
-\texttt{RLP}\Big( (u(0), u(1), ..., u(15), v) \Big) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
+(u(0), u(1), ..., u(15), v) & \text{otherwise} \quad \text{where} \begin{array}[t]{rcl}
 u(j) & \equiv & n(\{ I : I \in \mathfrak{I} \wedge I_0[i] = j \}, i + 1) \\
@@
 I_1 & \text{if} \quad \exists I: I \in \mathfrak{I} \wedge \lVert I_0 \rVert = i \\
-() & \text{otherwise}
+() \in \mathbb{B} & \text{otherwise}
 \end{cases}
@@
 \begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, I, T) \equiv \begin{cases}
-(\varnothing, 0, A^0, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A^0, \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, A, I) \equiv \begin{cases}
+(\varnothing, 0, A, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A, \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$.
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$ and the constant $\mathtt{secp256k1n}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
+In the case of an invalid signature, we return no output.
 
@@
 g_{\mathrm{r}} &=& 3000\\
-\lVert \mathbf{o} \rVert &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+\lVert \mathbf{o} \rVert &=& \begin{cases} \label{eq:ecrec_precompile_validity}
+  0 & \text{if} \quad v \notin \{27, 28\} \,\vee\, r = 0 \,\vee\, r \ge \mathtt{secp256k1n} \,\vee\, s = 0 \,\vee\, s \ge \mathtt{secp256k1n} \\
+  0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v - 27, r, s) = \varnothing\\
+  32 & \text{otherwise}
+\end{cases}\\
 \text{if} \quad \lVert \mathbf{o} \rVert = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v - 27, r, s)\big)[12..31] \quad \text{where:}\\
 \mathbf{d}[0..(\lVert \hyperlink{I__d}{I_{\mathbf{d}}} \rVert-1)] &=& I_{\mathbf{d}}\\
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\[0.5em]
-\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
-\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+g_{\mathrm{r}} &=& \max \left(200, \left\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{\mathrm{quaddivisor}}}\right\rfloor \right) \\
+G_{\mathrm{quaddivisor}} &\equiv& 3 \\
+f(x) &\equiv& \left\lceil \frac{x}{8} \right\rceil ^2 \\
+\ell'_{\mathrm{E}} &=& \begin{cases}
+0 & \text{if}\ \ell_{\mathrm{E}}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{\mathrm{E}}\le 32 \wedge E \neq 0 \\
+8(\ell_{\mathrm{E}} - 32) + \lfloor \log_2(i[(96+\ell_{\mathrm{B}})..(127+\ell_{\mathrm{B}})]) \rfloor & \text{if}\ 32 < \ell_{\mathrm{E}} \wedge i[(96 + \ell_{\mathrm{B}})..(127 + \ell_{\mathrm{B}})]\neq 0 \\
+8(\ell_{\mathrm{E}} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{\mathrm{M}}} \\
+\ell_{\mathrm{B}} &\equiv& i[0..31] \\
+\ell_{\mathrm{E}} &\equiv& i[32..63] \\
+\ell_{\mathrm{M}} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{\mathrm{B}})] \\
+E &\equiv& i[(96+\ell_{\mathrm{B}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})] \\
+M &\equiv& i[(96+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}}+\ell_{\mathrm{M}})] \\
 i[x] &\equiv& \begin{cases}
@@
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ F\\
 F&\equiv&(\lVert I_{\mathbf{d}} \rVert\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
 k &=& \dfrac{\lVert I_{\mathbf{d}} \rVert}{192} \\
-g_{\mathrm{r}}&=& 80000k + 100000 \\
+g_{\mathrm{r}}&=& 34000k + 45000 \\
 \mathbf{o}[0..31]&\equiv&\begin{cases}
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+g_{\mathrm{r}} &=& 150\\
 \mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ x=\varnothing\\
+g_{\mathrm{r}} &=& 6000\\
 \mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 
+\subsection{BLAKE2 Precompiled Contract}
+
+EIP-152 by \cite{EIP-152} defines $\Xi_{\mathtt{BLAKE2\_F}}$ as a precompiled contract implementing the compression function $\mathtt{F}$ used in the BLAKE2 cryptographic hashing algorithm.
+The $\mathtt{F}$ compression function is specified in RFC 7693 by \cite{RFC-7693}.
+\begin{eqnarray}
+  \Xi_{\mathtt{BLAKE2\_F}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+  \Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ \lVert I_{\mathbf{d}}\rVert \neq 213 \vee f \notin \{0, 1\} \\
+  g_{\mathrm{r}} &=& r\\
+  \mathbf{o} &\equiv& \mathtt{LE}_8(h'_0)\cdot ... \cdot \mathtt{LE}_8(h'_7) \\
+  (h'_0,\dots,h'_7) &\equiv& \mathtt{F}(h, m, t_\mathrm{low}, t_\mathrm{high}, f) \quad\text{with } r \text{ rounds and } w = 64 \\
+  \mathtt{BE}_4(r) &\equiv& I_{\mathbf{d}}[0..4] \\
+  \mathtt{LE}_8(h_0) &\equiv& I_{\mathbf{d}}[4..12] \\
+  &\dots& \\
+  \mathtt{LE}_8(h_7) &\equiv& I_{\mathbf{d}}[60..68] \\
+  \mathtt{LE}_8(m_0) &\equiv& I_{\mathbf{d}}[68..76] \\
+  &\dots& \\
+  \mathtt{LE}_8(m_{15}) &\equiv& I_{\mathbf{d}}[188..196] \\
+  \mathtt{LE}_8(t_\mathrm{low}) &\equiv& I_{\mathbf{d}}[196..204] \\
+  \mathtt{LE}_8(t_\mathrm{high}) &\equiv& I_{\mathbf{d}}[204..212] \\
+  f &\equiv& I_{\mathbf{d}}[212]
+\end{eqnarray}
+where $r \in \mathbb{B}_{32}$, $\forall i \in 0..7: h_i \in \mathbb{B}_{64}$, $\forall i \in 0..15: m_i \in \mathbb{B}_{64}$,
+$t_\mathrm{low} \in \mathbb{B}_{64}$, $t_\mathrm{high} \in \mathbb{B}_{64}$, $f \in \mathbb{B}_8$,
+$\mathtt{BE}_k$ is the $k$-byte big-endian representation---compare with(\ref{eq:BE}):
+\begin{equation}
+  \mathtt{BE}_k(x) \equiv (b_0, b_1, ..., b_{k-1}): x = \sum_{n = 0}^{k-1} b_n \cdot 256^{k-1-n}
+\end{equation}
+and $\mathtt{LE}_k$ is the $k$-byte little-endian representation:
+\begin{equation}
+  \mathtt{LE}_k(x) \equiv (b_0, b_1, ..., b_{k-1}): x = \sum_{n = 0}^{k-1} b_n \cdot 256^n
+\end{equation}
+
 \section{Signing Transactions}\label{app:signing}
@@
 
-Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$), $p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
+Where $p_{\mathrm{u}}$ is the public key, assumed to be a byte array of size 64 (formed from the concatenation of two positive integers each $< 2^{256}$),
+$p_{\mathrm{r}}$ is the private key, a byte array of size 32 (or a single positive integer in the aforementioned range) and $e$ is the hash of the transaction, \hyperlink{h_of_T}{$h(T)$}.
 It is assumed that \hypertarget{v}{}$v$ is the `recovery identifier'.
-The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[27, 30]$, however we declare the upper two possibilities, representing infinite values, invalid.
-The value 27 represents an even $y$ value and 28 represents an odd $y$ value.
+The recovery identifier is a 1 byte value specifying the parity and finiteness of the coordinates of the curve point for which $r$ is the x-value; this value is in the range of $[0, 3]$, however we declare the upper two possibilities, representing infinite values, invalid.
+The value 0 represents an even $y$ value and 1 represents an odd $y$ value.
 
@@
 
-\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true\footnote{A signature of a transaction can be valid not only with a recovery identifier but with some other numbers.
-See \hyperlink{T__w}{how the component $T_w$ of a transaction is interpreted.}}:
+\linkdest{invalidsig}We declare that an ECDSA signature is invalid unless all the following conditions are true:
 \begin{align}
@@
 0 < \linkdest{s}{s} &< \mathtt{secp256k1n} \div 2 + 1 \\
-\hyperlink{v}{v} &\in \{27,28\}
+\hyperlink{v}{v} &\in \{0,1\}
 \end{align}
@@
 \end{align}
-
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak hash of the corresponding ECDSA public key:
+Note that this restriction on $s$ is more stringent than restriction \ref{eq:ecrec_precompile_validity} in the $\Xi_{\mathtt{ECREC}}$ precompile; see EIP-2 by \cite{EIP-2} for more detail.
+
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak-256 hash of the corresponding ECDSA public key:
 \begin{equation}
@@
 
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{S}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{chain\_id}, (), ()) & \text{otherwise} \\
+\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction. Four different flavours of signing schemes are available:
+\begin{eqnarray}
+L_{\mathrm{X}}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{f}}, T_{\mathrm{m}},  T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 2
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
+h(T) & \equiv & \begin{cases}
+\mathtt{KEC}( \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{if} \; T_{\mathrm{x}} = 0 \\
+\mathtt{KEC}( T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
-\end{eqnarray}
-
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}:
+(T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
+\end{eqnarray}
+
+\hyperlink{T__r_T__s}{Reiterating from previously}:
 \begin{eqnarray}
@@
 \end{eqnarray}
-$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.
-In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
+and $\hyperlink{T__w}{T_{\mathrm{w}}}$ of legacy transactions is either $27 + T_{\mathrm{y}}$ or $2\hyperlink{chain_id}{\beta} + 35 + T_{\mathrm{y}}$.
 
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
+v &\equiv& \begin{cases}
+T_{\mathrm{w}} - 27 & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{27, 28\} \\
+(T_{\mathrm{w}}-35) \bmod 2 & \text{if} \ T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
+T_{\mathrm{y}} & \text{if} \ T_{\mathrm{x}} = 1 \lor T_{\mathrm{x}} = 2
 \end{cases}
@@
 
-The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
-
+\nopagebreak
+The fee schedule $G$ is a tuple of scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
+
+\nopagebreak
 \begin{tabu}{l r l}
 \toprule
-Name & Value & Description* \\
-\midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+Name & Value & Description \\
+\midrule
+$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
 $G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+$G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
+\linkdest{G__accesslistaddress}{}$G_{\mathrm{accesslistaddress}}$ & 2400 & Cost of warming up an account with the access list. \\
+\linkdest{G__accessliststorage}{}$G_{\mathrm{accessliststorage}}$ & 1900 & Cost of warming up a storage with the access list. \\
+$G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
+$G_{\mathrm{coldsload}}$ & 2100 & Cost of a cold storage access. \\
+$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 2900 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
 &&is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
-&&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+$R_{\mathrm{sclear}}$ & 4800 & Refund given (added into refund counter) when the storage value is set to zero from\\
+&&non-zero. The refund amount is defined as $G_{\mathrm{sreset}} + G_{\mathrm{accessliststorage}}$. \\
+\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
+\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+\linkdest{G__initcodeword}{}$G_{\mathrm{initcodeword}}$ & 2 & Paid per word of the initcode at the beginning of a deploy transaction, a {\small CREATE},\\
+&&or a {\small CREATE2} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
+\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
 \linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
 $G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 $G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
-&&contract. \\
+$G_{\mathrm{keccak256}}$ & 30 & Paid for each {\small KECCAK256} operation. \\
+$G_{\mathrm{keccak256word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small KECCAK256} operation. \\
+$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & Payment for each {\small BLOCKHASH} operation. \\
 
@@
 
-\begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
-&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
+\nopagebreak
+\begin{equation}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv C_{\mathrm{mem}}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{\mathrm{mem}}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SSTORE} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w \in W_{\mathrm{copy}} \\
+
+C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) & \text{if} \quad w \in W_{\mathrm{extaccount}}\\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{if} \quad w \in W_{\mathrm{call}} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
+G_{\mathrm{create}} + \hyperlink{initcode_cost}{R(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{R(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
+G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
+C_\text{\tiny SLOAD}(\boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SLOAD}\\
+G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
+G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
+G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
 G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
+G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
 \end{cases}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
-\end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
-
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
-
-$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
-
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+C_{\mathrm{mem}}(a) \equiv G_{\mathrm{memory}} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
+\end{equation}
+\begin{equation}
+\linkdest{C_aaccess}{}C_{\mathrm{aaccess}}(x, A) \equiv
+\begin{cases}
+G_{\mathrm{warmaccess}}        & \text{if} \quad x \in \hyperlink{accessed_addresses_defn_words_A__a}{A_{\mathbf{a}}}\\
+G_{\mathrm{coldaccountaccess}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$, $C_\text{\tiny SLOAD}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below.
+We define the following subsets of instructions:
+
+$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}, {\small PUSH0}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH1}, ..., {\small PUSH32}, {\small DUP*}, {\small SWAP*}\}
+
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}, {\small SELFBALANCE}\}
+
+$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
 
@@
 
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
-
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+$W_{\mathrm{copy}}$ = \{{\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURNDATACOPY}\}
+
+$W_{\mathrm{call}}$ = \{{\small CALL}, {\small CALLCODE}, {\small DELEGATECALL}, {\small STATICCALL}\}
+
+$W_{\mathrm{extaccount}}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small EXTCODEHASH}\}
+
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
-
-Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
+$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special cases are given where these two are not equal.
+
+Note also that $C_{\mathrm{mem}}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 704B of memory used, after which it costs substantially more.
 
@@
 
+\nopagebreak
 \begin{equation}
@@
 \begin{equation}
+\linkdest{L_but_64}{}
 L(n) \equiv n - \lfloor n / 64 \rfloor
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \bmod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
@@
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \bmod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
@@
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i - 248 + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i \geq 248 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\
 &&&& in big endian). \\
+\midrule
+0x1b & {\small SHL} & 2 & 1 & Left shift operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \bmod 2^{256}$ \\
+\midrule
+0x1c & {\small SHR} & 2 & 1 & Logical right shift operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
+\midrule
+0x1d & {\small SAR} & 2 & 1 & Arithmetic (signed) right shift operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
+&&&& Where $\boldsymbol{\mu}'_{\mathbf{s}}[0]$ and $\boldsymbol{\mu}_{\mathbf{s}}[1]$ are treated as two's complement signed 256-bit integers, \\
+&&&& while $\boldsymbol{\mu}_{\mathbf{s}}[0]$ is treated as unsigned. \\
 \bottomrule
@@
 \toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{20s: KECCAK256}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+0x20 & {\small KECCAK256} & 2 & 1 & Compute Keccak-256 hash. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{KEC}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
@@
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \midrule
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
+&&&& This is the \textit{effective gas price} defined in section \ref{ch:transactions}. \\
+&&&& Note that as of the \textit{London} hard fork, this value no longer \\ 
+&&&& represents what is received by the validator, \\
+&&&& but rather just what is paid by the sender. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
-&&&& This is gas price specified by the originating transaction.\\
 \midrule
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
+\begin{cases}
+\lVert \mathbf{b} \rVert & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing \\
+0                        & \text{otherwise}
+\end{cases}$ \\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \midrule
@@
 &&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+\begin{cases} \mathbf{b}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
+&&&& We assume $\mathbf{b} \equiv ()$ if $\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] = \varnothing$. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+\midrule
+\linkdest{extcodehash}{}0x3f & {\small EXTCODEHASH} & 1 & 1 & Get hash of an account's code. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
+\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathrm{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than or\\
+&&&& equal to the current block number or more than 256 blocks behind the current block.\\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
-\midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
-\midrule
-0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
-\midrule
-0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
-\midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
-\midrule
-0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& and we assert the header $H$ can be determined from its hash~$h$ unless $h$ is zero\\
+&&&& (as is the case for the parent hash of the genesis block).\\
+\midrule
+0x41 & {\small COINBASE} & 0 & 1 & Get the current block's beneficiary address. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{c}}$ \\
+\midrule
+0x42 & {\small TIMESTAMP} & 0 & 1 & Get the current block's timestamp. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{s}}$ \\
+\midrule
+0x43 & {\small NUMBER} & 0 & 1 & Get the current block's number. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{i}}$ \\
+\midrule
+0x44 & {\small PREVRANDAO} & 0 & 1 & Get the latest RANDAO mix of the post beacon state of the previous block. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{a}}$ \\
+\midrule
+0x45 & {\small GASLIMIT} & 0 & 1 & Get the current block's gas limit. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{l}}$ \\
+\midrule
+0x46 & {\small CHAINID} & 0 & 1 & Get the \hyperlink{chain_id}{chain ID}. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \beta$ \\
+\midrule
+0x47 & {\small SELFBALANCE} & 0 & 1 & Get balance of currently executing account. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} $ \\
+\midrule
+0x48 & {\small BASEFEE} & 0 & 1 & Get the current block's base fee. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{f}}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
+&&&& $A'_{\mathbf{K}} \equiv A_{\mathbf{K}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
+&&&& $C_{\text{\tiny SLOAD}}(\boldsymbol{\mu}, A, I) \equiv
+\begin{cases}
+G_{\mathrm{warmaccess}} & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{K}} \\
+G_{\mathrm{coldsload}}  & \text{otherwise}
+\end{cases}$ \\
 \midrule
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
+&&&& $A'_{\mathbf{K}} \equiv A_{\mathbf{K}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
+&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{\mathrm{r}}$ are specified by EIP-2200 as follows. \\
+&&&& We remind the reader that the checkpoint (``original'') state $\hyperlink{sigma_0}{\boldsymbol{\sigma}_0}$ is the state \\
+&&&& if the current transaction were to revert. \\
+&&&& Let $v_0 = \boldsymbol{\sigma}_0[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the original value of the storage slot. \\
+&&&& Let $v = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the current value. \\
+&&&& Let $v' = \boldsymbol{\mu}_{\mathbf{s}}[1]$ be the new value. \\
+&&&& Then: \\
+&&&& $\!\begin{aligned}
+C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) &\equiv
+\begin{cases}
+0                      & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{K}} \\
+G_{\mathrm{coldsload}} & \text{otherwise} \\
+\end{cases} \\
+&+
+\begin{cases}
+G_{\mathrm{warmaccess}}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
+G_{\mathrm{sset}}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
+G_{\mathrm{sreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
+\end{cases}
+\end{aligned}$ \\
+&&&& \linkdest{A r}{}$A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
+R_{\mathrm{sclear}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
+r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 \neq v \\
+0 & \text{otherwise} \\
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
-0 & \text{otherwise}
+&&&& where \\
+&&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
+-R_{\mathrm{sclear}} & \text{if} \quad v_0 \neq 0 \; \wedge \; v = 0 \\
+R_{\mathrm{sclear}} & \text{if} \quad v_0 \neq 0 \; \wedge \; v' = 0 \\
+0 & \text{otherwise} \\
+\end{cases}$ \\
+&&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
+G_{\mathrm{sset}} - G_{\mathrm{warmaccess}}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
+G_{\mathrm{sreset}} - G_{\mathrm{warmaccess}} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
+0 & \text{otherwise} \\
 \end{cases}$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{\mathrm{i}}$ \\
 \midrule
@@
 \toprule
-\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{5f, 60s \& 70s: Push Operations}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
+0x5f & {\small PUSH0} & 0 & 1 & Place 0 on the stack. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 0$ \\
+\midrule
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
@@
 \multicolumn{5}{l}{For all logging operations, the state change is to append an additional log entry on to the substate's log series:}\\
-\multicolumn{5}{l}\linkdest{A l}{}{$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
+\multicolumn{5}{l}{\linkdest{A l}{}$A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot (I_{\mathrm{a}}, \mathbf{t}, \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$}\\
 \multicolumn{5}{l}{and to update the memory consumption counter:}\\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\hyperlink{salt}{\zeta} \equiv \varnothing$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \begin{cases}
+\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \wedge \\ \quad &\; I_{\mathrm{e}} < 1024 \wedge \lVert \mathbf{i} \rVert \leqslant 49152\\
+\big(\boldsymbol{\sigma}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} - \hyperlink{L_but_64}{L}(\boldsymbol{\mu}_{\mathrm{g}}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
-&&&& address of the newly created account. \\
+&&&& where $x=0$ if $z = 0$, i.e., the \hyperlink{contract_creation_result}{contract creation process failed},\\
+&&&& or $\lVert \mathbf{i} \rVert > 49152$ (initcode is longer than 49152 bytes),\\
+&&&& or $I_{\mathrm{e}} = 1024$ (the maximum call depth limit is reached),\\
+&&&& or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too low to fulfil the value transfer);\\
+&&&& and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the address of the newly created account (\ref{eq:new-address}). \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv \begin{cases}
+() & \text{if} \quad z = 1 \\
+\mathbf{o} & \text{otherwise}
+\end{cases}$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}
+\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+  & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A) + g'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& $A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{t\}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& where $x=0$ if the \hyperlink{code_execution_result}{code execution for this operation failed}, or if \\
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
+&&&& \linkdest{tiny CALL}{}$C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv  \begin{cases}
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv \begin{cases}
+\min\{ \hyperlink{L_but_64}{L}(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv \hyperlink{C_aaccess}{C_{\mathrm{aaccess}}}(t, A) + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, t) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+  & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
+&&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+  & \text{if} \quad I_{\mathrm{e}} < 1024 \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
+&&&& and ninth parameters to the call $\hyperlink{theta}{\Theta}$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\midrule
+\linkdest{create2}{} 0xf5 & {\small CREATE2} & 4 & 1 & Create a new account with associated code. \\
+&&&& Exactly equivalent to {\small CREATE} except: \\
+&&&& The salt $\zeta \equiv \boldsymbol{\mu}_{\mathbf{s}}[3]$.\\
 \midrule
@@
 &&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
-\midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+&&&& The last argument of $\hyperlink{theta}{\Theta}$ is $\bot$. \\
+\midrule
+\linkdest{REVERT}{}0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\
@@
 \midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. Readers should be warned, \\
+&&&& since the \textit{Shanghai} update, this opcode has been marked as deprecated. \\
+&&&& Its behavior might be subject to breaking change in an upcoming update. \\
 &&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ r \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+\varnothing &\text{if}\quad \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\quad r \neq I_{\mathrm{a}} \\
 (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $\begin{aligned}
+C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) &\equiv G_{\mathrm{selfdestruct}} 
++ \begin{cases}
+  0 & \text{if}\quad r \in \hyperlink{accessed_addresses_defn_words_A__a}{A_{\mathbf{a}}}\\
+  G_{\mathrm{coldaccountaccess}} & \text{otherwise}
+\end{cases}\\
+&+ \begin{cases}
+G_{\mathrm{newaccount}} & \text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}, r) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0 \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+\end{cases}
+\end{aligned}$ \\
 \bottomrule
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{34}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{34}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak-256 hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
 $\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.
@@
 \section{Ethash}\label{app:ethash}
+\subsection{Deprecation}
+This section is kept for historical purposes, but the Ethash algorithm is no longer used for consensus as of the \textit{Paris} hard fork.
+
 \subsection{Definitions}
@@
 
+\nopagebreak
 \begin{tabu*}{lcl}
@@
 \midrule
-\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
+\linkdest{J__wordbytes}{}$J_{\mathrm{wordbytes}}$ & 4  & Bytes in word. \\
+\linkdest{J__datasetinit}{}$J_{\mathrm{datasetinit}}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{\mathrm{cacheinit}}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{\mathrm{epoch}}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{\mathrm{mixbytes}}$ & 128 & mix length in bytes. \\
 \linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
 \linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__cacherounds}{}$J_{\mathrm{cacherounds}}$ & 3 & Number of rounds in cache production. \\
 \linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
@@
 \begin{equation}
- E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
+ E_{\mathrm{epoch}}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{\mathrm{epoch}}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{\mathrm{datasetgrowth}}$ bytes, and the size of the cache by $J_{\mathrm{cachegrowth}}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
-\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
-\end{equation}
-The size of the cache, $c_{size}$, is calculated using
-\begin{equation}
- c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+Therefore the size is reduced by a multiple of $J_{\mathrm{mixbytes}}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
+\linkdest{d__size}{}Let $d_{\mathrm{size}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{\mathrm{size}} = E_{\mathrm{prime}}(J_{\mathrm{datasetinit}} + J_{\mathrm{datasetgrowth}} \cdot E_{\mathrm{epoch}} - J_{\mathrm{mixbytes}}, J_{\mathrm{mixbytes}})
+\end{equation}
+The size of the cache, $c_{\mathrm{size}}$, is calculated using
+\begin{equation}
+ c_{\mathrm{size}} = E_{\mathrm{prime}}(J_{\mathrm{cacheinit}} + J_{\mathrm{cachegrowth}} \cdot E_{\mathrm{epoch}} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
 \end{equation}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
-\end{equation}
-\begin{equation}
- C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
-\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
+ \mathbf{s} = C_{\mathrm{seedhash}}(H_{\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{\mathrm{seedhash}}(H_{\mathrm{i}}) = \begin{cases}
+\mathbf{0}_{32} & \text{if} \quad E_{\mathrm{epoch}}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{\mathrm{seedhash}}(H_{\mathrm{i}} - J_{\mathrm{epoch}})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially filling up $c_{\mathrm{size}}$ bytes of memory, then performing $J_{\mathrm{cacherounds}}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
-\end{equation}
-\begin{equation}
- E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{\mathrm{size}}}{J_{\mathrm{hashbytes}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{\mathrm{cacherounds}}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
+\end{equation}
+\begin{equation}
+ E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{\mathrm{rmh}}(\mathbf{x}, 0), E_{\mathrm{rmh}}(\mathbf{x}, 1), ... , E_{\mathrm{rmh}}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
 \begin{multline}
-  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
+  E_{\mathrm{rmh}}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \bmod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \bmod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{\mathrm{rmh}}(\mathbf{x}, j) \quad \forall \quad j < i
 \end{multline}
@@
 \begin{equation}
- \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
+ \mathbf{d}[i] = E_{\mathrm{datasetitem}}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{\mathrm{size}}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \bmod 2^{32}
 \end{equation}
@@
   E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \bmod c_\mathrm{{size}}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \bmod \lfloor J_{\mathrm{hashbytes}} / J_{\mathrm{wordbytes}} \rfloor]) \bmod c_{\mathrm{size}}] \big) & \text{otherwise}
 \end{cases}
@@
 Note that the extra application of \texttt{KEC} at the end ensures that there exists an intermediate nonce which can be provided to prove that at least a small amount of work was done; this quick outer PoW verification can be used for anti-DDoS purposes.
-It also serves to provide statistical assurance that the result is an unbiased, 256 bit number.
+It also serves to provide statistical assurance that the result is an unbiased, 256-bit number.
 
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{\mathrm{hashbytes}}}\right\rfloor
+ n_{\mathrm{mix}} =  \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
  E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
+E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
+\end{equation}
+$E_{\mathrm{newdata}}$ returns an array with $n_{\mathrm{mix}}$ elements:
+\begin{equation}
+ E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \bmod \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{wordbytes}}}\right\rfloor]) \bmod \left\lfloor\frac{d_{\mathrm{size}} / J_{\mathrm{hashbytes}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
 \end{equation}

### 2019-12-11 Paper.tex

@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{sha3}+G_{sha3word}\times\lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
 G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\

### 2020-05-08 Paper.tex
Small corrections

@@
 
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to transform it into some final state.
-It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
+Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some current state.
+It is this current state which we accept as the canonical ``version'' of the world of Ethereum.
 The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
-Transactions thus represent a valid transition between two states.
-The `valid' part is important---there exist far more invalid state transitions than valid state transitions.
-An invalid state transitions might, for \eg, reduce an account balance without increasing another by the same amount.
+Transactions thus represent a valid arc between two states; the `valid' part is important---there exist far more invalid state changes than valid state changes.
+Invalid state changes might, \eg, be things such as reducing an account balance without an equal and opposite increase elsewhere.
 A valid state transition is one which comes about through a transaction.
-
-Formally, we define a well-known genesis state $\boldsymbol{\sigma}_0$, a subsequent series of states $\boldsymbol{\sigma}_t$, and a state transition function $\Upsilon$:
+Formally:
 \begin{equation}
@@
 
-The state and state transition function of Ethereum are considerably more powerful than those of existing comparable system; $\Upsilon$ is allowed to carry out arbitrary computation, and $\boldsymbol{\sigma}$ is allowed to store arbitrary state between transactions.
-
-Transactions are collated into blocks; a block consists of a list of transactions together with a cryptographic hash of the previous block that it ``builds upon'', as well as a cryptographic commitment to the final state (blocks do not store the final state itself---that would be far too big).
-Blocks also incentivize nodes to \textit{mine} by allowing the state transition function to add value to a nominated account.
+where $\Upsilon$ is the Ethereum state transition function.
+In Ethereum, $\Upsilon$, together with $\boldsymbol{\sigma}$ are considerably more powerful than any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\boldsymbol{\sigma}$ allows components to store arbitrary state between transactions.
+
+Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference.
+Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).
+They also punctuate the transaction series with incentives for nodes to \textit{mine}.
+This incentivisation takes place as a state-transition function, adding value to a nominated account.
 
@@
 
-Formally, we define a block \hyperlink{block}{$B$} as a list of transactions as well as some other components:
-\begin{eqnarray}
-B & \equiv & (..., (T_0, T_1, ...), ...)
-\end{eqnarray}
-
-We define the block-level state transition function $\hyperlink{Pi}{\Pi}$, the block-level analogue to the transaction-level state transition function $\hyperlink{Upsilon}{\Upsilon}$.
-$\Pi$ applies $\Upsilon$ to the transactions of a block in order and then applies an additional block-finalisation state transition $\Omega$.
+Formally, we expand to:
 \begin{eqnarray}
 \boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
+B & \equiv & (..., (T_0, T_1, ...), ...) \\
 \Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
@@
 
+Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+
 This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.
@@
 
-In order to incentivise computation within the network, there needs to be a method to transmit value.
-For this purpose, Ethereum provides an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
-The smallest denomination of Ether is the Wei; all other denominations must be an integer multiple of a Wei.
+In order to incentivise computation within the network, there needs to be an agreed method for transmitting value.
+To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH} and sometimes referred to by the Old English \DH{}.
+The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei.
 One Ether is defined as being $10^{18}$ Wei.
@@
 
-Sometimes, a path follows a new protocol from a particular height.
+Sometimes, a path follows a new protocol from a particular height (block number).
 This document describes one version of the protocol.
@@
 I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\} \\
+\mathbf{t} & \equiv & \{s, r\}
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
 \end{cases}

### 2020-05-09 Paper.tex
EIP 1052: EXTCODEHASH opcode

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Constantinople version \YellowPaperVersionNumber}}}
 \author{
@@
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
+$G_{extcodehash}$ & 400 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
 $G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
@@
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{extcode} & \text{if} \quad w \in W_{extcode}\\
+G_{extcode} & \text{if} \quad w = \text{\small EXTCODESIZE}\\
+G_{extcodehash} & \text{if} \quad w = \text{\small \hyperlink{extcodehash}{EXTCODEHASH}}\\
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
@@
 $W_{\mathrm{high}}$ = \{{\small JUMPI}\}
-
-$W_{extcode}$ = \{{\small EXTCODESIZE}\}
 
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{KEC}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
+\midrule
+\linkdest{extcodehash}{}0x3f & {\small EXTCODEHASH} & 1 & 1 & Get hash of an account's code. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
+\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \\ \mathtt{KEC}(\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}) & \text{otherwise} \end{cases}$ \\
 \bottomrule

### 2020-05-10 Paper.tex
EIP 1234: Constantinople Difficulty Bomb Delay and Block Reward Adjustment

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Constantinople version \YellowPaperVersionNumber}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Petersburg version \YellowPaperVersionNumber}}}
 \author{
@@
 \expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 5000000, 0)
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+In the \textit{Byzantium} release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million\footnote{five million since the \textit{Constantinople} release as per EIP-1234} from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
 
@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 2 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 2 \times 10^{18}
 \end{equation}

### 2020-05-10 Paper.tex
EIP 145: Bitwise shifting instructions in EVM

@@
 
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
 
@@
 &&&& in big endian). \\
+\midrule
+0x1b & {\small SHL} & 2 & 1 & Left shift operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \mod 2^{256}$ \\
+\midrule
+0x1c & {\small SHR} & 2 & 1 & Logical right shift operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
+\midrule
+0x1d & {\small SAR} & 2 & 1 & Arithmetic (signed) right shift operation. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \geqslant 0 \\ -\lceil |\boldsymbol{\mu}_{\mathbf{s}}[1]| \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rceil & \text{otherwise} \end{cases} $ \\
+&&&& Where $\boldsymbol{\mu}'_{\mathbf{s}}[0]$ and $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (but not $\boldsymbol{\mu}_{\mathbf{s}}[0]$) are treated as two's complement signed 256-bit integers. \\
 \bottomrule

### 2020-05-10 Paper.tex
Simplify the defenition of SAR

@@
 
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
 
@@
 0x1d & {\small SAR} & 2 & 1 & Arithmetic (signed) right shift operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \geqslant 0 \\ -\lceil |\boldsymbol{\mu}_{\mathbf{s}}[1]| \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rceil & \text{otherwise} \end{cases} $ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
 &&&& Where $\boldsymbol{\mu}'_{\mathbf{s}}[0]$ and $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (but not $\boldsymbol{\mu}_{\mathbf{s}}[0]$) are treated as two's complement signed 256-bit integers. \\

### 2020-05-10 Paper.tex
EIP 1014: Skinny CREATE2

@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
 \Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
+The salt $\zeta$ might be missing ($\zeta = \varnothing$); formally,
+\begin{equation}
+\zeta \in \mathbb{B}_{32} \cup \mathbb{B}_{0}
+\end{equation}
+If the creation was caused by {\small \hyperlink{create2}{CREATE2}}, then $\zeta \neq \varnothing$.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
 \end{equation}
@@
 The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
-Thus we define the resultant address for the new account $a$:
-\begin{equation}
-a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
-\end{equation}
-
-where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+For {\small \hyperlink{create2}{CREATE2}} the rule is different and is described in EIP-1014.
+Combining the two cases, we define the resultant address for the new account $a$:
+\begin{eqnarray}
+a & \equiv & A(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
+\label{eq:new-address} A(s, n, \zeta, \mathbf{i}) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big(B(s, n, \zeta, \mathbf{i})\big)\Big) \\
+B(s, n, \zeta, \mathbf{i}) & \equiv & \begin{cases}
+\mathtt{RLP}\big(\;(s, n)\;\big) & \text{if}\ \zeta = \varnothing \\
+(255) \cdot s \cdot \zeta \cdot \mathtt{KEC}(\mathbf{i}) & \text{otherwise}
+\end{cases}
+\end{eqnarray}
+
+where $\cdot$ is the concatenation of byte arrays, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 \item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE}s being executed at present).
+the number of {\small CALL}s or {\small CREATE(2)}s being executed at present).
 \item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
@@
 \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
+w \in \{\text{\small CREATE}, \text{\small CREATE2}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
+G_{create}+G_{sha3word} \lceil \mathbf{s}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
 G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
-&&&& address of the newly created account. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
-&&&& Thus the operand order is: value, input offset, input size. \\
+&&&& Exactly equivalent to {\small \hyperlink{create2}{CREATE2}} except: \\
+&&&& The salt $\zeta = \varnothing$ instead of $\boldsymbol{\mu}_{\mathbf{s}}[3]$. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
+\midrule
+\linkdest{create2}{} 0xf5 & {\small CREATE2} & 4 & 1 & Create a new account with associated code. \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
+&&&& $\zeta \equiv \boldsymbol{\mu}_{\mathbf{s}}[3]$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
+&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
+&&&& address of the newly created account (\ref{eq:new-address}). \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& Thus the operand order is: value, input offset, input size, salt. \\
 \midrule

### 2020-05-11 Paper.tex
Define CREATE2 in terms of CREATE rather than the other way round
Correct notation for the cost of SHA3 and CREATE2

@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{create}+G_{sha3word} \lceil \mathbf{s}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
-G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{create}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{sha3}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
 G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
@@
 0xf0 & {\small CREATE} & 3 & 1 & Create a new account with associated code. \\
-&&&& Exactly equivalent to {\small \hyperlink{create2}{CREATE2}} except: \\
-&&&& The salt $\zeta = \varnothing$ instead of $\boldsymbol{\mu}_{\mathbf{s}}[3]$. \\
+&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
+&&&& $\zeta \equiv \varnothing$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
+&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& where $x=0$ if the code execution for this operation failed due to an\\
+&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
+&&&& address of the newly created account (\ref{eq:new-address}). \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& Thus the operand order is: value, input offset, input size. \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
 \midrule
 \linkdest{create2}{} 0xf5 & {\small CREATE2} & 4 & 1 & Create a new account with associated code. \\
-&&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $\zeta \equiv \boldsymbol{\mu}_{\mathbf{s}}[3]$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
-&&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
-&&&& address of the newly created account (\ref{eq:new-address}). \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
-&&&& Thus the operand order is: value, input offset, input size, salt. \\
+&&&& Exactly equivalent to {\small CREATE} except: \\
+&&&& The salt $\zeta \equiv \boldsymbol{\mu}_{\mathbf{s}}[3]$.\\
 \midrule

### 2020-05-11 Paper.tex
Fix code vs its hash for EXTCODESIZE, EXTCODECOPY, EXTCODEHASH

@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 \midrule
@@
 &&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+\begin{cases} \mathbf{b}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
-\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \\ \mathtt{KEC}(\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}) & \text{otherwise} \end{cases}$ \\
+\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
 \bottomrule

### 2020-05-11 Paper.tex
Small comment improvement for SAR

@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
-&&&& Where $\boldsymbol{\mu}'_{\mathbf{s}}[0]$ and $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (but not $\boldsymbol{\mu}_{\mathbf{s}}[0]$) are treated as two's complement signed 256-bit integers. \\
+&&&& Where $\boldsymbol{\mu}'_{\mathbf{s}}[0]$ and $\boldsymbol{\mu}_{\mathbf{s}}[1]$ are treated as two's complement signed 256-bit integers, \\
+&&&& while $\boldsymbol{\mu}_{\mathbf{s}}[0]$ is treated as unsigned. \\
 \bottomrule

### 2020-05-11 Paper.tex
Don't show version on every page header as it obscures page numbers

@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Petersburg version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller{Petersburg version \YellowPaperVersionNumber}}
 \author{

### 2020-05-11 build.sh

@@ -13,7 +13,7 @@ if [ -d ".git" ]; then
 
 SHA=`git rev-parse --short --verify HEAD`
 DATE=`git show -s --format="%cd" --date=short HEAD`
-REV="$SHA - $DATE"
+REV="$SHA -- $DATE"
 echo "\def\YellowPaperVersionNumber{$REV}" >> Options.tex
 
 fi

### 2020-05-27 Paper.tex
Update the version to Istanbul + minor corrections

--- 2020-05-11 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Petersburg version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller{Istanbul version \YellowPaperVersionNumber}}
 \author{
@@
 
-The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
+The fee schedule $G$ is a tuple of scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
 
@@
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \> \lor \\
-&\quad\text{\small DELEGATECALL} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w \in W_{\mathrm{call}} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 
+$W_{\mathrm{call}}$ = \{{\small CALL}, {\small CALLCODE}, {\small DELEGATECALL}, {\small STATICCALL}\}
+
 Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
@@
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\
-&&&& and we assert the header $H$ can be determined from its hash~$h$ unless as its hash is the parent\\
-&&&& $h$ is zero.\\
+&&&& and we assert the header $H$ can be determined from its hash~$h$ unless $h$ is zero\\
+&&&& (as is the case for the parent hash of the genesis block).\\
 \midrule

### 2020-06-03 Paper.tex
Improve spec of MLOAD, MSTORE, and MSIZE.
This is in reference to Issue #767, which has all the relevant discussion.

https://github.com/ethereum/yellowpaper/issues/767

--- 2020-05-11 Paper.tex
@@
 
-\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Petersburg version \YellowPaperVersionNumber}}}
+\title{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{Byzantium version \YellowPaperVersionNumber}}}
 \author{
@@
 
-Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some current state.
-It is this current state which we accept as the canonical ``version'' of the world of Ethereum.
+Ethereum, taken as a whole, can be viewed as a transaction-based state machine: we begin with a genesis state and incrementally execute transactions to morph it into some final state.
+It is this final state which we accept as the canonical ``version'' of the world of Ethereum.
 The state can include such information as account balances, reputations, trust arrangements, data pertaining to information of the physical world; in short, anything that can currently be represented by a computer is admissible.
@@
 
-Sometimes, a path follows a new protocol from a particular height (block number).
+Sometimes, a path follows a new protocol from a particular height.
 This document describes one version of the protocol.
@@
 \expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 5000000, 0)
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 3000000, 0)
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-In the \textit{Byzantium} release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million\footnote{five million since the \textit{Constantinople} release as per EIP-1234} from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+Finally, in the Byzantium release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
 
@@
 (\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
 \Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
-The salt $\zeta$ might be missing ($\zeta = \varnothing$); formally,
-\begin{equation}
-\zeta \in \mathbb{B}_{32} \cup \mathbb{B}_{0}
-\end{equation}
-If the creation was caused by {\small \hyperlink{create2}{CREATE2}}, then $\zeta \neq \varnothing$.
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, w)
 \end{equation}
@@
 The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
-For {\small \hyperlink{create2}{CREATE2}} the rule is different and is described in EIP-1014.
-Combining the two cases, we define the resultant address for the new account $a$:
-\begin{eqnarray}
-a & \equiv & A(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
-\label{eq:new-address} A(s, n, \zeta, \mathbf{i}) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big(B(s, n, \zeta, \mathbf{i})\big)\Big) \\
-B(s, n, \zeta, \mathbf{i}) & \equiv & \begin{cases}
-\mathtt{RLP}\big(\;(s, n)\;\big) & \text{if}\ \zeta = \varnothing \\
-(255) \cdot s \cdot \zeta \cdot \mathtt{KEC}(\mathbf{i}) & \text{otherwise}
-\end{cases}
-\end{eqnarray}
-
-where $\cdot$ is the concatenation of byte arrays, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
+Thus we define the resultant address for the new account $a$:
+\begin{equation}
+a \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\Big(\mathtt{RLP}\big(\;(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1)\;\big)\Big)\Big)
+\end{equation}
+
+where $\mathtt{KEC}$ is the Keccak 256-bit hash function, $\mathtt{RLP}$ is the RLP encoding function, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
@@
 I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\}
+\mathbf{t} & \equiv & \{s, r\} \\
 \end{eqnarray}
@@
 The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CREATE2}, {\small CALL} and {\small CALLCODE}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CALL} and {\small CALLCODE}.
 Finally, gas may be paid due to an increase in the usage of the memory.
@@
 \item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
-the number of {\small CALL}s or {\small CREATE(2)}s being executed at present).
+the number of {\small CALL}s or {\small CREATE}s being executed at present).
 \item\linkdest{exec_permission_to_modify_state_I__w}{}\linkdest{I__w} $I_{\mathrm{w}}$, the permission to make modifications to the state.
@@
 \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small CREATE2}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
+w \in \{\text{\small CREATE}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 2 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 2 \times 10^{18}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 3 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
 \end{equation}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\boldsymbol{\sigma}_0 kern 10pc \ \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} \quad\text{otherwise}
 \end{cases}
@@
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
-$G_{extcodehash}$ & 400 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
 $G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
@@
 G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{create}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
-G_{sha3}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
 G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
@@
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{extcode} & \text{if} \quad w = \text{\small EXTCODESIZE}\\
-G_{extcodehash} & \text{if} \quad w = \text{\small \hyperlink{extcodehash}{EXTCODEHASH}}\\
+G_{extcode} & \text{if} \quad w \in W_{extcode}\\
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
@@
 
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
 
@@
 $W_{\mathrm{high}}$ = \{{\small JUMPI}\}
+
+$W_{extcode}$ = \{{\small EXTCODESIZE}\}
 
@@
 &&&& in big endian). \\
-\midrule
-0x1b & {\small SHL} & 2 & 1 & Left shift operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \mod 2^{256}$ \\
-\midrule
-0x1c & {\small SHR} & 2 & 1 & Logical right shift operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
-\midrule
-0x1d & {\small SAR} & 2 & 1 & Arithmetic (signed) right shift operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lfloor \boldsymbol{\mu}_{\mathbf{s}}[1] \div 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]} \rfloor$ \\
-&&&& Where $\boldsymbol{\mu}'_{\mathbf{s}}[0]$ and $\boldsymbol{\mu}_{\mathbf{s}}[1]$ are treated as two's complement signed 256-bit integers, \\
-&&&& while $\boldsymbol{\mu}_{\mathbf{s}}[0]$ is treated as unsigned. \\
 \bottomrule
@@
 0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{KEC}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{Keccak}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[1])$ \\
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\
-&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} \rVert$ \\
 \midrule
@@
 &&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[3] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i ] \equiv
-\begin{cases} \mathbf{b}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+\begin{cases} \mathbf{c}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{c} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
+&&&& where $\mathbf{c} \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert$ \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-\midrule
-\linkdest{extcodehash}{}0x3f & {\small EXTCODEHASH} & 1 & 1 & Get hash of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
-\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\min(\boldsymbol{\mu}_{\mathbf{s}}[0] + 32, 2^{256}) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + 31) ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1]$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\boldsymbol{\mu}_{\mathbf{s}}[0] + 32) \div 32 })$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv \max(\boldsymbol{\mu}_{\mathrm{i}}, \ceil{ (\min(\boldsymbol{\mu}_{\mathbf{s}}[0] + 32, 2^{256}) \div 32 })$ \\
 &&&& The addition in the calculation of $\boldsymbol{\mu}'_{\mathrm{i}}$ is not subject to the $2^{256}$ modulo. \\
@@
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases} 32\boldsymbol{\mu}_{i} & \text{if} \quad \boldsymbol{\mu}_{i} < 2^{251} \\ 2^{251}-1 & \text{otherwise} \end{cases}$ \\
 \midrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $\zeta \equiv \varnothing$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
@@
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
-&&&& address of the newly created account (\ref{eq:new-address}). \\
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}})$, the\\
+&&&& address of the newly created account. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
@@
 \midrule
-\linkdest{create2}{} 0xf5 & {\small CREATE2} & 4 & 1 & Create a new account with associated code. \\
-&&&& Exactly equivalent to {\small CREATE} except: \\
-&&&& The salt $\zeta \equiv \boldsymbol{\mu}_{\mathbf{s}}[3]$.\\
-\midrule
 0xfa & {\small STATICCALL} & 6 & 1 & Static message-call into an account. \\

### 2020-06-09 Paper.tex
Update authors.
I think it makes sense to amend authors with new names at this point. See https://twitter.com/paulmillr/status/1270270652738605056

@@
 \usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
-Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+Gavin Wood, James Ray, Yoichi Hirai, Christoph Jentzsch & Ethereum community},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 %,pagebackref=true

### 2020-06-18 Paper.tex
Status in receipt RLP may be != 0
EIP1679 - EIP2200 + EIP2387

--- 2020-06-08 Paper.tex
@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller{Petersburg version \YellowPaperVersionNumber}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Istanbul version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Istanbul version \YellowPaperVersionNumber}}}
 \author{
@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol.
+This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
 In order to follow back the history of a path, one must reference multiple versions of this document.
+
+Occasionally actors cannot agree on a protocol change.
+For instance, in 2016 a fraction of the Ethereum community disagreed with the DAO bailout, which led to an alternative blockchain dubbed "Ethereum Classic".
+In order to distinguish between different blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
+For the Ethereum main network
+\begin{equation}
+  \linkdest{chain_id}
+  \beta = 1
+\end{equation}
 
@@
 T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_5 & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
 T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
@@
 \expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 5000000, 0)
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0) \\
+\kappa &\equiv \begin{cases} 
+  0  & \text{if} \quad H_{\mathrm{i}} < 4370000 \\
+  3000000  & \text{if} \quad 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
+  5000000  & \text{if} \quad 7280000 \leqslant H_{\mathrm{i}} < 9200000 \\
+  9000000  & \text{otherwise} \\
+\end{cases}
 \end{align}
@@
 $\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-In the \textit{Byzantium} release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million\footnote{five million since the \textit{Constantinople} release as per EIP-1234} from the actual block number, which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+In the \textit{Byzantium} release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number,
+which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
+Subsequently, with EIP-1234 by \cite{EIP-1234} and EIP-2384 by \cite{EIP-2384} the subtrahend $\kappa$ was increased to five and nine million respectively.
 
@@
 
-Note we use $\Theta_{4}$ and $\Lambda_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
 
@@
 There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
-The salt $\zeta$ might be missing ($\zeta = \varnothing$); formally,
+The salt \linkdest{salt}{$\zeta$} might be missing ($\zeta = \varnothing$); formally,
 \begin{equation}
@@
 
-We define the creation function formally as the function $\Lambda$, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
 \begin{equation}
@@
 The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
-For {\small \hyperlink{create2}{CREATE2}} the rule is different and is described in EIP-1014.
+For {\small \hyperlink{create2}{CREATE2}} the rule is different and is described in EIP-1014 by \cite{EIP-1014}.
 Combining the two cases, we define the resultant address for the new account $a$:
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Theta(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 9 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
@@
 
-As can be seen, there are eight exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are eight so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
-The eight contracts in addresses 1 to 8 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication and an elliptic curve pairing check respectively.
+As can be seen, there are nine exceptions to the usage of the general execution framework $\Xi$ for evaluation of the message call: these are so-called `precompiled' contracts, meant as a preliminary piece of architecture that may later become \textit{native extensions}.
+The contracts in addresses 1 to 9 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication, an elliptic curve pairing check, and the BLAKE2 compression function $\mathtt{F}$ respectively.
 
@@
 
-With $\mathbf{d}$ being a dataset as specified in appendix \ref{app:ethash}.
+With $\mathbf{d}$ being a dataset as specified in Appendix \ref{app:ethash}.
 
@@
 The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in appendix \ref{app:ethash}.
+The mentioned parameters as well as the algorithm is explained in detail in Appendix \ref{app:ethash}.
 
@@
 k &=& \dfrac{\lVert I_{\mathbf{d}} \rVert}{192} \\
-g_{\mathrm{r}}&=& 80000k + 100000 \\
+g_{\mathrm{r}}&=& 34000k + 45000 \\
 \mathbf{o}[0..31]&\equiv&\begin{cases}
@@
 \Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
-g_{\mathrm{r}} &=& 500\\
+g_{\mathrm{r}} &=& 150\\
 \mathbf{o}&\equiv&\delta_1^{-1}(x+y)\quad\text{where $+$ is the group operation in $G_1$}\\
@@
 \Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
-g_{\mathrm{r}} &=& 40000\\
+g_{\mathrm{r}} &=& 6000\\
 \mathbf{o}&\equiv&\delta_1^{-1}(n\cdot x)\quad\text{where $\cdot$ is the scalar multiplication in $G_1$}\\
@@
 \end{eqnarray}
+
+\subsection{BLAKE2 Precompiled Contract}
+
+EIP-152 by \cite{EIP-152} defines $\Xi_{\mathtt{BLAKE2\_F}}$ as a precompiled contract implementing the compression function $\mathtt{F}$ used in the BLAKE2 cryptographic hashing algorithm.
+The $\mathtt{F}$ compression function is specified in RFC 7693 by \cite{RFC-7693}.
+\begin{eqnarray}
+  \Xi_{\mathtt{BLAKE2\_F}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
+  \Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ \lVert I_{\mathbf{d}}\rVert \neq 213 \vee f \notin \{0, 1\} \\
+  g_{\mathrm{r}} &=& r\\
+  \mathbf{o} &\equiv& \mathtt{LE}_8(h'_0)\cdot ... \cdot \mathtt{LE}_8(h'_7) \\
+  (h'_0,\dots,h'_7) &\equiv& \mathtt{F}(h, m, t_\mathrm{low}, t_\mathrm{high}, f) \quad\text{with } r \text{ rounds and } w = 64 \\
+  \mathtt{BE}_4(r) &\equiv& I_{\mathbf{d}}[0..4] \\
+  \mathtt{LE}_8(h_0) &\equiv& I_{\mathbf{d}}[4..12] \\
+  &\dots& \\
+  \mathtt{LE}_8(h_7) &\equiv& I_{\mathbf{d}}[60..68] \\
+  \mathtt{LE}_8(m_0) &\equiv& I_{\mathbf{d}}[68..76] \\
+  &\dots& \\
+  \mathtt{LE}_8(m_{15}) &\equiv& I_{\mathbf{d}}[188..196] \\
+  \mathtt{LE}_8(t_\mathrm{low}) &\equiv& I_{\mathbf{d}}[196..204] \\
+  \mathtt{LE}_8(t_\mathrm{high}) &\equiv& I_{\mathbf{d}}[204..212] \\
+  f &\equiv& I_{\mathbf{d}}[212]
+\end{eqnarray}
+where $r \in \mathbb{B}_{32}$, $\forall i \in 0..7: h_i \in \mathbb{B}_{64}$, $\forall i \in 0..15: m_i \in \mathbb{B}_{64}$,
+$t_\mathrm{low} \in \mathbb{B}_{64}$, $t_\mathrm{high} \in \mathbb{B}_{64}$, $f \in \mathbb{B}_8$,
+$\mathtt{BE}_k$ is the $k$-byte big-endian representation---compare with(\ref{eq:BE}):
+\begin{equation}
+  \mathtt{BE}_k(x) \equiv (b_0, b_1, ..., b_{k-1}): x = \sum_{n = 0}^{k-1} b_n \cdot 256^{k-1-n}
+\end{equation}
+and $\mathtt{LE}_k$ is the $k$-byte little-endian representation:
+\begin{equation}
+  \mathtt{LE}_k(x) \equiv (b_0, b_1, ..., b_{k-1}): x = \sum_{n = 0}^{k-1} b_n \cdot 256^n
+\end{equation}
 
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \mathtt{chain\_id}, (), ()) & \text{otherwise} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{otherwise} \\
 \end{cases} \\
@@
 \end{eqnarray}
-$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier doubled plus 35 or 36'.
-In the second case, where \hypertarget{v}{}$v$ is the chain identifier doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
+$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier \hyperlink{chain_id}{$\beta$} doubled plus 35 or 36'.
+In the second case, where \hypertarget{v}{}$v$ is the chain identifier $\beta$ doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
 
@@
 
-The fee schedule $G$ is a tuple of 31 scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
-
+\nopagebreak
+The fee schedule $G$ is a tuple of scalar values corresponding to the relative costs, in gas, of a number of abstract operations that a transaction may effect.
+
+\nopagebreak
 \begin{tabu}{l r l}
 \toprule
-Name & Value & Description* \\
+Name & Value & Description \\
 \midrule
 $G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{jumpdest}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
 $G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
@@
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
-$G_{extcodehash}$ & 400 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{balance}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 800 & Paid for an {\small SLOAD} operation. \\
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
@@
 $G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+\linkdest{G__callstipend}{}$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
 \linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
@@
 \linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 68 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
 \linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
@@
 G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \> \lor \\
-&\quad\text{\small CODECOPY} \lor \text{\small RETURNDATACOPY} \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w \in W_{\mathrm{copy}} \\
+
 G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{extcode} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
 G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{extcode} & \text{if} \quad w = \text{\small EXTCODESIZE}\\
-G_{extcodehash} & \text{if} \quad w = \text{\small \hyperlink{extcodehash}{EXTCODEHASH}}\\
 G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
@@
 
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 
-$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}
+$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}, {\small SELFBALANCE}\}
 
@@
 $W_{\mathrm{high}}$ = \{{\small JUMPI}\}
+
+$W_{\mathrm{copy}}$ = \{{\small CALLDATACOPY}, {\small CODECOPY}, {\small RETURNDATACOPY}\}
+
+$W_{\mathrm{call}}$ = \{{\small CALL}, {\small CALLCODE}, {\small DELEGATECALL}, {\small STATICCALL}\}
+
+$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}, {\small EXTCODEHASH}\}
 
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+\midrule
+0x46 & {\small CHAINID} & 0 & 1 & Get the \hyperlink{chain_id}{chain ID}. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \beta$ \\
+\midrule
+0x47 & {\small SELFBALANCE} & 0 & 1 & Get balance of currently executing account. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} $ \\
 \bottomrule
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[1] \dots (\boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] - 1) ]$ \\
-&&&& $\zeta \equiv \varnothing$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\Lambda(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $\hyperlink{salt}{\zeta} \equiv \varnothing$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
-&&&& Note the change in the fourth parameter to the call $\Theta$ from the 2nd stack value \\
+&&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[1]$ (as in {\small CALL}) to the present address $I_{\mathrm{a}}$. This means that the recipient is in\\
@@
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
-&&&& and ninth parameters to the call $\Theta$.\\
+&&&& and ninth parameters to the call $\hyperlink{theta}{\Theta}$.\\
 &&&& This means that the recipient is in fact the same account as at present, simply\\
@@
 &&&& with $\boldsymbol{\mu}_{\mathbf{s}}[2]$, $\boldsymbol{\mu}_{\mathbf{s}}[3]$, $\boldsymbol{\mu}_{\mathbf{s}}[4]$ and $\boldsymbol{\mu}_{\mathbf{s}}[5]$. \\
-&&&& The last argument of $\Theta$ is $\bot$. \\
+&&&& The last argument of $\hyperlink{theta}{\Theta}$ is $\bot$. \\
 \midrule

### 2020-06-30 Paper.tex
CALLCODE may be readonly executed when value != 0
There's no need to check balance in DELEGATECALL as there's no value transfer
CALLCODE & DELEGATECALL don't bump nonce unlike CREATE, so they should use sigma rather than sigma* in line with CALL
nopagebreak in Appendix J. Ethash
transaction_receipt_preparation_function_for_RLP_serialisation is unnecesary

--- 2020-06-08 Paper.tex
@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising: the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$, the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$} and the status code of the transaction, $R_{\mathrm{z}}$:
-\begin{equation}
-R \equiv (R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}}, R_{\mathrm{z}})
-\end{equation}
-
-\hypertarget{transaction_receipt_preparation_function_for_RLP_serialisation}{}\linkdest{L__R}The function $L_{R}$ trivially prepares a transaction receipt for being transformed into an RLP-serialised byte array:
-\begin{equation}
-L_{R}(R) \equiv (0 \in \mathbb{B}_{256}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
-\end{equation}
-where $0 \in \mathbb{B}_{256}$ replaces the pre-transaction state root that existed in previous versions of the protocol.
-
-\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer.
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising:
+the status code of the transaction, $R_{\mathrm{z}}$,
+the cumulative gas used in the block containing the transaction receipt as of immediately after the transaction has happened, $R_{\mathrm{u}}$,
+the set of logs created through execution of the transaction, \hyperlink{RLP_serialisation_of_a_sequence_of_other_items_R__l_math_def}{$R_\mathbf{l}$} and the Bloom filter composed from information in those logs, \hyperlink{RLP_serialisation_of_a_byte_array_R__b_math_def}{$R_{\mathrm{b}}$}:
+\begin{equation}
+R \equiv (R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+\end{equation}
+
+\linkdest{R__z_assert}We assert that the status code $R_{\mathrm{z}}$ is a non-negative integer:
 \begin{equation}
@@
 \linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{L_{R}}(B_{\mathbf{R}}[i]))\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
-Much like the \hyperlink{transaction_receipt_preparation_function_for_RLP_serialisation}{transaction receipt preparation function $L_{R}$}, we assert the types and order of the structure for when the RLP transformation is required:
+We assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}

### 2020-07-03 Paper.tex
Clarify reward rounding

--- 2020-06-09 Paper.tex
@@
 \usepackage[bookmarks=true, unicode=true, pdftitle={Ethereum Yellow Paper: a formal specification of Ethereum, a programmable blockchain}, pdfauthor={Dr.
-Gavin Wood, James Ray, Yoichi Hirai, Christoph Jentzsch & Ethereum community},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
+Gavin Wood},pdfkeywords={Ethereum, Yellow Paper, blockchain, virtual machine, cryptography, decentralised, singleton, transaction, generalised},pdfborder={0 0 0.5 [1 3]}]{hyperref}
 %,pagebackref=true
@@
 \begin{eqnarray}
-\\ \nonumber
-\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
+\Omega(B, \boldsymbol{\sigma}) \!\!& \equiv &\!\! \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} \!\!& = &\!\!
+\boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + R_{\mathrm{block}} + \lVert\mathbf{B}_{\mathbf{U}}\rVert \left\lfloor\frac{R_{\mathrm{block}}}{32}\right\rfloor \\
 \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
+R & \equiv & \left\lfloor\frac{(8 + U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}}) R_{\mathrm{block}}}{8}\right\rfloor
 \end{eqnarray}

### 2020-07-05 Paper.tex
Don't mention Ethereum Classic

--- 2020-06-18 Paper.tex
@@
 
-Occasionally actors cannot agree on a protocol change.
-For instance, in 2016 a fraction of the Ethereum community disagreed with the DAO bailout, which led to an alternative blockchain dubbed "Ethereum Classic".
-In order to distinguish between different blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
+Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
+In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
 For the Ethereum main network

### 2020-07-25 Paper.tex
Fix hyperlink

@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
 The block header contains several pieces of information:
@@
 \kappa &\equiv \begin{cases} 
-  0  & \text{if} \quad H_{\mathrm{i}} < 4370000 \\
   3000000  & \text{if} \quad 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
   5000000  & \text{if} \quad 7280000 \leqslant H_{\mathrm{i}} < 9200000 \\
-  9000000  & \text{otherwise} \\
+  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant 9200000 \\
 \end{cases}
@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 2 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 2 \times 10^{18}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward in Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 10^{18} \times \begin{cases}
+5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
+3 &\text{if}\ H_{\mathrm{i}} < 7280000 \\
+2 &\text{otherwise}
+\end{cases} \\
 \end{equation}
@@
 B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction_Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}
@@
 \midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
+0x41 & {\small COINBASE} & 0 & 1 & Get the current block's beneficiary address. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
-0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
+0x42 & {\small TIMESTAMP} & 0 & 1 & Get the current block's timestamp. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
-0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
+0x43 & {\small NUMBER} & 0 & 1 & Get the current block's number. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
+0x44 & {\small DIFFICULTY} & 0 & 1 & Get the current block's difficulty. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
-0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
+0x45 & {\small GASLIMIT} & 0 & 1 & Get the current block's gas limit. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\

### 2020-07-25 Paper.tex
Fix hyperlink

--- 2020-07-03 Paper.tex
@@
 \begin{eqnarray}
-\Omega(B, \boldsymbol{\sigma}) \!\!& \equiv &\!\! \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} \!\!& = &\!\!
-\boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + R_{\mathrm{block}} + \lVert\mathbf{B}_{\mathbf{U}}\rVert \left\lfloor\frac{R_{\mathrm{block}}}{32}\right\rfloor \\
+\\ \nonumber
+\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
 \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left\lfloor\frac{(8 + U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}}) R_{\mathrm{block}}}{8}\right\rfloor
+R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
 \end{eqnarray}

### 2020-08-27 Paper.tex
fix: Add Block Reward application for pre and post Byzantium cases

@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 2 Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 2 \times 10^{18}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward in Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
+\text{Let} \quad R_{\mathrm{block}} = 10^{18} \times \begin{cases}
+5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
+3 &\text{otherwise}
+\end{cases} \\
 \end{equation}
@@
 B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction_Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
 \end{eqnarray}

### 2020-08-31 Paper.tex
fix: Add Block Reward application for pre and post Constantinople cases
EIP 2028: Transaction data gas cost reduction
This change was released in Istanbul.

@@
 \begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
 \text{Let} \quad R_{\mathrm{block}} = 10^{18} \times \begin{cases}
 5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
-3 &\text{otherwise}
+3 &\text{if}\ H_{\mathrm{i}} < 7280000 \\
+2 &\text{otherwise}
 \end{cases} \\
@@
 $G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
-$G_{extcodehash}$ & 400 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
-$G_{balance}$ & 400 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 200 & Paid for a {\small SLOAD} operation. \\
+$G_{extcodehash}$ & 700 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
+$G_{balance}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 800 & Paid for a {\small SLOAD} operation. \\
 $G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\

### 2020-08-31 Paper.tex
EIP 2028: Transaction data gas cost reduction
This change was released in Istanbul.

--- 2020-08-27 Paper.tex
@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward in Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 3 \times 10^{18}
-\text{Let} \quad R_{\mathrm{block}} = 10^{18} \times \begin{cases}
-5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
-3 &\text{otherwise}
-\end{cases} \\
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward as 2 Ether:
+\begin{equation}
+\text{Let} \quad R_{\mathrm{block}} = 2 \times 10^{18}
 \end{equation}

### 2020-09-17 Paper.tex
Difficulty formulae in Homestead and Frontier were different from Byzantium

@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol.
+This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
 In order to follow back the history of a path, one must reference multiple versions of this document.
+
+Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
+In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
+For the Ethereum main network
+\begin{equation}
+  \linkdest{chain_id}
+  \beta = 1
+\end{equation}
 
@@
 \expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - 5000000, 0)
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0) \\
+\kappa &\equiv \begin{cases} 
+  3000000  & \text{if} \quad 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
+  5000000  & \text{if} \quad 7280000 \leqslant H_{\mathrm{i}} < 9200000 \\
+  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant 9200000 \\
+\end{cases}
 \end{align}

### 2021-04-11 Paper.tex
Update `SSTORE` cost and refund for EIP-2200
This patch ports the decision tree from the specification of [EIP-2200]
into the instruction set reference. Fixes #788.

[EIP-2200]: https://eips.ethereum.org/EIPS/eip-2200

wchargin-branch: eip-2200
wchargin-source: 92c2af76d68bd0da255770fa4578091342835ef0

--- 2020-08-31 Paper.tex
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{} $C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sset} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] = 0 \\
-G_{sreset} & \text{otherwise}
+&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{r}$ are specified by EIP-2200 as follows. \\
+&&&& Let $\boldsymbol{\sigma}^0$ be the state if the current transaction were to revert (the ``original state''). \\
+&&&& Let $v^0 = \boldsymbol{\sigma}^0[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the original value of the storage slot. \\
+&&&& Let $v = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the current value. \\
+&&&& Let $v' = \boldsymbol{\mu}_{\mathbf{s}}[1]$ be the new value. \\
+&&&& Then: \\
+&&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
+G_{sload} & \text{if} \quad v = v' \\
+G_{sset}  & \text{if} \quad v^0 = v \; \wedge \; v^0 = 0 \\
+G_{sreset}  & \text{if} \quad v^0 = v \; \wedge \; v^0 \neq 0 \\
+G_{sload} & \text{otherwise}
 \end{cases}$ \\
 &&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]] \neq 0 \\
-0 & \text{otherwise}
+0 & \text{if} \quad v = v' \\
+0 & \text{if} \quad v^0 = v \; \wedge \; v^0 = 0 \\
+R_{sclear} & \text{if} \quad v^0 = v \; \wedge \; v' = 0 \\
+r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{otherwise}
+\end{cases}$ \\
+&&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
+-R_{sclear} & \text{if} \quad v^0 \neq 0 \; \wedge \; v = 0 \\
+R_{sclear} & \text{if} \quad v^0 \neq 0 \; \wedge \; v' = 0 \\
+0 & \text{otherwise} \\
+\end{cases}$ \\
+&&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
+G_{sset} - G_{sload}   & \text{if} \quad v^0 = v' \; \wedge \; v^0 = 0 \\
+G_{sreset} - G_{sload} & \text{if} \quad v^0 = v' \; \wedge \; v^0 \neq 0 \\
+0 & \text{otherwise} \\
 \end{cases}$ \\

### 2021-05-04 Paper.tex
update nonbinary.miraheze.org url

--- 2021-03-07 Paper.tex
@@
 \end{equation}
-with $(m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
+with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
 
@@
 \end{eqnarray}
-where $(m, n) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
 
@@
 B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (m, x) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
+B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
 B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))

### 2021-05-11 Paper.tex
Specify which block is affected

@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language\#Aunt.2FUncle}}).
 The block header contains several pieces of information:

### 2021-05-11 Paper.tex
Fix off-by-one error in BLOCKHASH domain specification

@@
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
-&&&& age. 0 is left on the stack if the looked for block number is greater than\\
-&&&& the current block number or more than 256 blocks behind the current block.\\
+&&&& age. 0 is left on the stack if the looked for block number is greater than or\\
+&&&& equal to the current block number or more than 256 blocks behind the current block.\\
 &&&& $P(h, n, a) \equiv \begin{cases} 0 & \text{if} \quad n > H_{\mathrm{i}} \vee a = 256 \vee h = 0 \\ h & \text{if} \quad n = H_{\mathrm{i}} \\ P(H_{\mathrm{p}}, n, a + 1) & \text{otherwise} \end{cases}$ \\

### 2021-05-29 Paper.tex
Add EIP 1344's ChainID opcode

--- 2021-05-04 Paper.tex
@@
 
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, {\small CHAINID} \}
 
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+\midrule
+0x46 & {\small CHAINID} & 0 & 1 & Get the chain's id. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{chain\_id}$ \\
 \bottomrule

### 2021-06-07 Paper.tex

@@
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for an {\small EXTCODESIZE} operation. \\
-$G_{extcodehash}$ & 700 & Amount of gas to pay for an {\small EXTCODEHASH} operation. \\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
 $G_{balance}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 800 & Paid for a {\small SLOAD} operation. \\
-$G_{jumpdest}$ & 1 & Paid for a {\small JUMPDEST} operation. \\
+$G_{sload}$ & 800 & Paid for an {\small SLOAD} operation. \\
 $G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
@@
 
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, {\small CHAINID} \}
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \midrule
-0x46 & {\small CHAINID} & 0 & 1 & Get the chain's id. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{chain\_id}$ \\
+0x46 & {\small CHAINID} & 0 & 1 & Get the \hyperlink{chain_id}{chain ID}. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \beta$ \\
+\midrule
+0x47 & {\small SELFBALANCE} & 0 & 1 & Get balance of currently executing account. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} $ \\
 \bottomrule

### 2021-06-10 Paper.tex

--- 2021-05-29 Paper.tex
@@
 
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, {\small CHAINID} \}
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \midrule
-0x41 & {\small COINBASE} & 0 & 1 & Get the block's beneficiary address. \\
+0x41 & {\small COINBASE} & 0 & 1 & Get the current block's beneficiary address. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
-0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
+0x42 & {\small TIMESTAMP} & 0 & 1 & Get the current block's timestamp. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
-0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
+0x43 & {\small NUMBER} & 0 & 1 & Get the current block's number. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
+0x44 & {\small DIFFICULTY} & 0 & 1 & Get the current block's difficulty. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
-0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
+0x45 & {\small GASLIMIT} & 0 & 1 & Get the current block's gas limit. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
-\midrule
-0x46 & {\small CHAINID} & 0 & 1 & Get the chain's id. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{chain\_id}$ \\
 \bottomrule

### 2021-06-11 Paper.tex

@@
 \text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w \in \{\text{\small CALL}, \text{\small CALLCODE}\} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+w = \text{\small CALL} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{v}} \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \; I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2021-06-11 Paper.tex
Change description of G_expbyte (#787)

--- 2021-06-11 Paper.tex
@@
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by the number of bytes in the $exponent$ for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\

### 2021-06-11 Paper.tex
Block reward corrections

@@
 
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward in Ether:
-\begin{equation}
-\text{Let} \quad R_{\mathrm{block}} = 10^{18} \times \begin{cases}
+\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward in Wei:
+\begin{equation}
+R_{\mathrm{block}} = 10^{18} \times \begin{cases}
 5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
-3 &\text{if}\ H_{\mathrm{i}} < 7280000 \\
-2 &\text{otherwise}
+3 &\text{if}\ 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
+2 &\text{if}\ H_{\mathrm{i}} \geqslant 7280000 \\
 \end{cases} \\

### 2021-06-13 Paper.tex
Clarify EIP-2200 cases to be mutually exclusive (#798)
Corrections for CALLCODE and DELEGATECALL (#778)
* CALLCODE may be readonly executed when value != 0

* There's no need to check balance in DELEGATECALL as there's no value transfer

* CALLCODE & DELEGATECALL don't bump nonce unlike CREATE, so they should use sigma rather than sigma* in line with CALL
Block reward corrections (#797)

@@
 &&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sload} & \text{if} \quad v = v' \\
-G_{sset}  & \text{if} \quad v^0 = v \; \wedge \; v^0 = 0 \\
-G_{sreset}  & \text{if} \quad v^0 = v \; \wedge \; v^0 \neq 0 \\
-G_{sload} & \text{otherwise}
+G_{sload}  & \text{if} \quad v = v' \; \vee \; v^0 \neq v \\
+G_{sset}   & \text{if} \quad v \neq v' \; \wedge \; v^0 = v \; \wedge \; v^0 = 0 \\
+G_{sreset} & \text{if} \quad v \neq v' \; \wedge \; v^0 = v \; \wedge \; v^0 \neq 0 \\
 \end{cases}$ \\
 &&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-0 & \text{if} \quad v = v' \\
-0 & \text{if} \quad v^0 = v \; \wedge \; v^0 = 0 \\
-R_{sclear} & \text{if} \quad v^0 = v \; \wedge \; v' = 0 \\
-r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{otherwise}
+R_{sclear} & \text{if} \quad v \neq v' \; \wedge \; v^0 = v \; \wedge \; v' = 0 \\
+r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{if} \quad v \neq v' \; \wedge \; v^0 \neq v \\
+0 & \text{otherwise} \\
 \end{cases}$ \\

### 2021-06-13 Paper.tex
Cosmetic change
EIP-1706: Disable SSTORE with gasleft lower than call stipend (part of EIP-2200)

@@
 \lVert\boldsymbol{\mu}_\mathbf{s}\rVert < \mathbf{\delta}_w \quad \vee \\
-( w = \text{\small JUMP} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small JUMPI} \quad \wedge \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \quad \wedge \\
+( w = \text{\small JUMP} \; \wedge \; \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
+( w = \text{\small JUMPI} \; \wedge \; \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \\
 \quad \boldsymbol{\mu}_\mathbf{s}[0] \notin D(I_\mathbf{b}) ) \quad \vee \\
-( w = \text{\small RETURNDATACOPY} \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
-\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \vee
- (\neg I_{\mathrm{w}} \wedge W(w, \boldsymbol{\mu}))
+( w = \text{\small RETURNDATACOPY} \; \wedge \\ \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert) \quad \vee \\
+\lVert\boldsymbol{\mu}_\mathbf{s}\rVert - \mathbf{\delta}_w + \mathbf{\alpha}_w > 1024 \quad \vee \\
+( \neg I_{\mathrm{w}} \; \wedge \; W(w, \boldsymbol{\mu}) ) \quad \vee \\
+( w = \text{\small SSTORE} \; \wedge \; \boldsymbol{\mu}_g \leqslant G_{callstipend} )
 \end{array}
@@
 w \in \{\text{\small CREATE}, \text{\small CREATE2}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
-\text{\small LOG0} \le w \wedge w \le \text{\small LOG4} \quad \vee \\
-w = \text{\small CALL} \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
+\text{\small LOG0} \le w \; \wedge \; w \le \text{\small LOG4} \quad \vee \\
+w = \text{\small CALL} \; \wedge \; \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0
 \end{array}
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
+Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{callstipend}}$.
+The last condition was introduced in EIP-2200 by \cite{EIP-2200}.
 
@@
 $G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by the number of bytes in the $exponent$ for the {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
 $G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\

### 2021-06-13 Paper.tex
Cite EIP-1706
Fix a typo
Fix off-by-one error in BLOCKHASH domain specification (#794)

@@
 Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{callstipend}}$.
-The last condition was introduced in EIP-2200 by \cite{EIP-2200}.
+The last condition was introduced in EIP-1706 by \cite{EIP-1706} (part of EIP-2200 by \cite{EIP-2200}).
 

### 2021-06-13 Paper.tex
EIP-2200 original state is the same as the checkpoint state

@@
 \begin{eqnarray}
-\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
+\linkdest{sigma_0}{}\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
 \boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
@@
 &&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{r}$ are specified by EIP-2200 as follows. \\
-&&&& Let $\boldsymbol{\sigma}^0$ be the state if the current transaction were to revert (the ``original state''). \\
-&&&& Let $v^0 = \boldsymbol{\sigma}^0[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the original value of the storage slot. \\
+&&&& The checkpoint (``original'') state $\hyperlink{sigma_0}{\boldsymbol{\sigma}_0}$ is the state if the current transaction were to revert. \\
+&&&& Let $v_0 = \boldsymbol{\sigma}_0[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the original value of the storage slot. \\
 &&&& Let $v = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the current value. \\
@@
 &&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sload}  & \text{if} \quad v = v' \; \vee \; v^0 \neq v \\
-G_{sset}   & \text{if} \quad v \neq v' \; \wedge \; v^0 = v \; \wedge \; v^0 = 0 \\
-G_{sreset} & \text{if} \quad v \neq v' \; \wedge \; v^0 = v \; \wedge \; v^0 \neq 0 \\
+G_{sload}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
+G_{sset}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
+G_{sreset} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
 \end{cases}$ \\
 &&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad v \neq v' \; \wedge \; v^0 = v \; \wedge \; v' = 0 \\
-r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{if} \quad v \neq v' \; \wedge \; v^0 \neq v \\
+R_{sclear} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
+r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 \neq v \\
 0 & \text{otherwise} \\
@@
 &&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
--R_{sclear} & \text{if} \quad v^0 \neq 0 \; \wedge \; v = 0 \\
-R_{sclear} & \text{if} \quad v^0 \neq 0 \; \wedge \; v' = 0 \\
+-R_{sclear} & \text{if} \quad v_0 \neq 0 \; \wedge \; v = 0 \\
+R_{sclear} & \text{if} \quad v_0 \neq 0 \; \wedge \; v' = 0 \\
 0 & \text{otherwise} \\
@@
 &&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
-G_{sset} - G_{sload}   & \text{if} \quad v^0 = v' \; \wedge \; v^0 = 0 \\
-G_{sreset} - G_{sload} & \text{if} \quad v^0 = v' \; \wedge \; v^0 \neq 0 \\
+G_{sset} - G_{sload}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
+G_{sreset} - G_{sload} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
 0 & \text{otherwise} \\

### 2021-06-14 Paper.tex
EIP-3607: transactions from senders with deployed contract are forbidden
Creation on non-empty code or non-zero nonce throws an exception (#323)
Co-authored-by: Andrew Ashikhmin <34320705+yperbasis@users.noreply.github.com>
Amend contract creation when initialising code reverts

@@
 \begin{align}
-\quad g' &\equiv \begin{cases}
+\quad g' \equiv & \begin{cases}
 0 & \text{if} \quad F \\
@@
 \end{cases} \\
-\quad z &\equiv \begin{cases}
-0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
+\quad z \equiv & \begin{cases}
+0 & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  \lVert \mathbf{o} \rVert > 24576\big)
+F \equiv  & \big( \boldsymbol{\sigma}[a] \neq \varnothing \ \wedge\ \big(\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}\big(()\big) \vee \boldsymbol{\sigma}[a]_n \neq 0 \big) \big) \quad \vee \\
+\nonumber &(\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \quad \vee \\
+\nonumber &g^{**} < c \quad \vee \\
+\nonumber &\lVert \mathbf{o} \rVert > 24576
 \end{align}
@@
 Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
+In addition, observe that if the execution of the initialising code \hyperlink{REVERT}{reverts} ($\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} \neq \varnothing$), the resultant gas $g'$ is not depleted (provided there was no other exception), but no new account is created.
 
@@
 
-Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
+Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array~$\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
@@
 \midrule
-0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+\linkdest{REVERT}{}0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\

### 2021-06-15 Paper.tex
\mathrm for non-variable subscripts (#803)

@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{\mathrm{T}}(B_{\mathbf{T}}[i]))\}) & \wedge \\
 \linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
 We assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
-\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
 \end{eqnarray}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000
@@
 \nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
-\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
+\nonumber& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
+\nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
 \nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
@@
 
-Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
 \Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
@@
 \begin{equation}
-	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
+	A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \sum_{i \in A_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
 \end{equation}
@@
 ( \neg I_{\mathrm{w}} \; \wedge \; W(w, \boldsymbol{\mu}) ) \quad \vee \\
-( w = \text{\small SSTORE} \; \wedge \; \boldsymbol{\mu}_g \leqslant G_{callstipend} )
+( w = \text{\small SSTORE} \; \wedge \; \boldsymbol{\mu}_g \leqslant G_{\mathrm{callstipend}} )
 \end{array}
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{callstipend}}$.
+Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$.
 The last condition was introduced in EIP-1706 by \cite{EIP-1706} (part of EIP-2200 by \cite{EIP-2200}).
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
 \quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{\mathrm{H}}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H__g}{B_{\mathrm{H}}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
 \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
+R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{\mathrm{H}}}_{\mathrm{i}})\right) R_{\mathrm{block}}
 \end{eqnarray}
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{\mathrm{H}}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{\mathrm{cacheinit}}}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{\mathrm{datasetinit}}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $\hyperlink{J__epoch}{J_{\mathrm{epoch}}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in Appendix \ref{app:ethash}.
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
+g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{\mathrm{quaddivisor}}}\right\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_{E} &=& \begin{cases}
-0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
-8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
-8(\ell_{E} - 32) & \text{otherwise} \\
+\ell'_{\mathrm{E}} &=& \begin{cases}
+0 & \text{if}\ \ell_{\mathrm{E}}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{\mathrm{E}}\le 32 \wedge E \neq 0 \\
+8(\ell_{\mathrm{E}} - 32) + \lfloor \log_2(i[(96+\ell_{\mathrm{B}})..(127+\ell_{\mathrm{B}})]) \rfloor & \text{if}\ 32 < \ell_{\mathrm{E}} \wedge i[(96 + \ell_{\mathrm{B}})..(127 + \ell_{\mathrm{B}})]\neq 0 \\
+8(\ell_{\mathrm{E}} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
-\ell_{B} &\equiv& i[0..31] \\
-\ell_{E} &\equiv& i[32..63] \\
-\ell_{M} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{B})] \\
-E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
-M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{\mathrm{M}}} \\
+\ell_{\mathrm{B}} &\equiv& i[0..31] \\
+\ell_{\mathrm{E}} &\equiv& i[32..63] \\
+\ell_{\mathrm{M}} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{\mathrm{B}})] \\
+E &\equiv& i[(96+\ell_{\mathrm{B}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})] \\
+M &\equiv& i[(96+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}}+\ell_{\mathrm{M}})] \\
 i[x] &\equiv& \begin{cases}
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
+h(T) & \equiv & \mathtt{KEC}( L_{\mathrm{S}}(T) )
 \end{eqnarray}
@@
 \midrule
-$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
-$G_{jumpdest}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
-$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
-$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
+$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
+$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
+$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
+$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
 $G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
-$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
+$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
-$G_{balance}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{sload}$ & 800 & Paid for an {\small SLOAD} operation. \\
-$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+$G_{\mathrm{extcode}}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{\mathrm{balance}}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{\mathrm{sload}}$ & 800 & Paid for an {\small SLOAD} operation. \\
+$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{\mathrm{sreset}}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
 &&is set to zero. \\
-$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
+$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
 &&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-\linkdest{G__callstipend}{}$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{expbyte}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
-$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
+\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
+\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{\mathrm{call}}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
+\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
+$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
 \linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
+\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
 $G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 $G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
+$G_{\mathrm{sha3}}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{\mathrm{sha3word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{\mathrm{quaddivisor}}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
 &&contract. \\
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{\mathrm{mem}}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{\mathrm{mem}}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w \in W_{\mathrm{copy}} \\
-
-G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{extcode} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
-G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
+G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w \in W_{\mathrm{copy}} \\
+
+G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
+G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w \in W_{\mathrm{call}} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{create} & \text{if} \quad w = \text{\small CREATE}\\
-G_{create}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
-G_{sha3}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{zero} & \text{if} \quad w \in W_{zero}\\
-G_{base} & \text{if} \quad w \in W_{base}\\
-G_{verylow} & \text{if} \quad w \in W_{verylow}\\
+G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{create}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
+G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
+G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
 G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
-G_{mid} & \text{if} \quad w \in W_{mid}\\
+G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
 \end{cases}
@@
 \begin{equation}
-C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
+C_{\mathrm{mem}}(a) \equiv G_{\mathrm{memory}} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
 \end{equation}
@@
 
-$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
 
@@
 
-$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
 
@@
 
-Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the current block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the current block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the current block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the current block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the current block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{l}}$ \\
 \midrule
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{r}$ are specified by EIP-2200 as follows. \\
+&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{\mathrm{r}}$ are specified by EIP-2200 as follows. \\
 &&&& The checkpoint (``original'') state $\hyperlink{sigma_0}{\boldsymbol{\sigma}_0}$ is the state if the current transaction were to revert. \\
@@
 &&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{sload}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
-G_{sset}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
-G_{sreset} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
+G_{\mathrm{sload}}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
+G_{\mathrm{sset}}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
+G_{\mathrm{sreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
-R_{sclear} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
+&&&&\linkdest{A r}{} $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
+R_{\mathrm{sclear}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
 r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 \neq v \\
@@
 &&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
--R_{sclear} & \text{if} \quad v_0 \neq 0 \; \wedge \; v = 0 \\
-R_{sclear} & \text{if} \quad v_0 \neq 0 \; \wedge \; v' = 0 \\
+-R_{\mathrm{sclear}} & \text{if} \quad v_0 \neq 0 \; \wedge \; v = 0 \\
+R_{\mathrm{sclear}} & \text{if} \quad v_0 \neq 0 \; \wedge \; v' = 0 \\
 0 & \text{otherwise} \\
@@
 &&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
-G_{sset} - G_{sload}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
-G_{sreset} - G_{sload} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
+G_{\mathrm{sset}} - G_{\mathrm{sload}}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
+G_{\mathrm{sreset}} - G_{\mathrm{sload}} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
 0 & \text{otherwise} \\
@@
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{\mathrm{i}}$ \\
 \midrule
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
-G_{newaccount} & \text{if} \quad n \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{selfdestruct}} + \begin{cases}
+G_{\mathrm{newaccount}} & \text{if} \quad n \\
 0 & \text{otherwise}
@@
   E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
-E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_\mathrm{{size}}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{\mathrm{wordbytes}} \rfloor]) \mod c_{\mathrm{size}}] \big) & \text{otherwise}
 \end{cases}
@@
  E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
-\end{equation}
-$E_{newdata}$ returns an array with $n_{mix}$ elements:
-\begin{equation}
- E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
+E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
+E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
+\end{equation}
+$E_{\mathrm{newdata}}$ returns an array with $n_{\mathrm{mix}}$ elements:
+\begin{equation}
+ E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{wordbytes}}}\right\rfloor]) \mod \left\lfloor\frac{d_{\mathrm{size}} / J_{\mathrm{hashbytes}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
 \end{equation}

### 2021-06-16 Paper.tex
Fix broken hyperlink

@@
 
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{sigma_n}{later} for state and nonce validation.
 

### 2021-06-16 Paper.tex
Introduce fork block numbers

@@
 In order to follow back the history of a path, one must reference multiple versions of this document.
-
-Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
+Here are the block numbers of protocol updates (forks) on the Ethereum main network:
+\par
+\begin{center}
+\begin{tabular}{lr}
+\toprule
+Fork Name & First Block Number \\
+\midrule
+$F_{\mathrm{Homestead}}$         &  1150000 \\
+$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
+$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
+$F_{\mathrm{Byzantium}}$         &  4370000 \\
+$F_{\mathrm{Constantinople}}$    &  7280000 \\
+$F_{\mathrm{Petersburg}}$        &  7280000 \\
+$F_{\mathrm{Istanbul}}$          &  9069000 \\
+$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
+$F_{\mathrm{Berlin}}$            & 12244000 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\par
+
+Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
 In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
@@
 \kappa &\equiv \begin{cases} 
-  3000000  & \text{if} \quad 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
-  5000000  & \text{if} \quad 7280000 \leqslant H_{\mathrm{i}} < 9200000 \\
-  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant 9200000 \\
+  3000000  & \text{if} \quad F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
+  5000000  & \text{if} \quad F_{\mathrm{Constantinople}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Muir Glacier}} \\
+  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Muir Glacier}} \\
 \end{cases}
@@
 R_{\mathrm{block}} = 10^{18} \times \begin{cases}
-5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
-3 &\text{if}\ 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
-2 &\text{if}\ H_{\mathrm{i}} \geqslant 7280000 \\
+5 &\text{if}\ H_{\mathrm{i}} < F_{\mathrm{Byzantium}} \\
+3 &\text{if}\ F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
+2 &\text{if}\ H_{\mathrm{i}} \geqslant F_{\mathrm{Constantinople}} \\
 \end{cases} \\

### 2021-06-17 Paper.tex
Don't use fork for protocol updates
Update the version to Berlin

--- 2021-06-16 Paper.tex
@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Istanbul version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Istanbul version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Berlin version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
 \author{
@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
+This document describes one version of the protocol, namely the \textit{Berlin} version defined by \cite{BeikoLondon}.
 In order to follow back the history of a path, one must reference multiple versions of this document.

### 2021-06-17 Paper.tex
EIP-2565: ModExp Gas Cost

@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{\mathrm{quaddivisor}}}\right\rfloor \\
-f(x) &\equiv& \begin{cases}
-x^2 & \text{if}\ x \le 64 \\[0.5em]
-\left\lfloor\dfrac{x^2}{4}\right\rfloor + 96 x - 3072 & \text{if}\ 64 < x \le 1024 \\[1em]
-\left\lfloor\dfrac{x^2}{16}\right\rfloor + 480x - 199680 & \text{otherwise}
-\end{cases}\\
+g_{\mathrm{r}} &=& \max \left(200, \left\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{\mathrm{quaddivisor}}}\right\rfloor \right) \\
+G_{\mathrm{quaddivisor}} &\equiv& 3 \\
+f(x) &\equiv& \left\lceil \frac{x}{8} \right\rceil ^2 \\
 \ell'_{\mathrm{E}} &=& \begin{cases}
@@
 $G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{\mathrm{quaddivisor}}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
-&&contract. \\
 

### 2021-06-18 Paper.tex
ADDR instead of A for CREATE(2) new address function to avoid confusion with accrued substate

--- 2021-06-17 Paper.tex
@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Berlin version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Istanbul version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Istanbul version \YellowPaperVersionNumber}}}
 \author{
@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol, namely the \textit{Berlin} version defined by \cite{BeikoLondon}.
+This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
 In order to follow back the history of a path, one must reference multiple versions of this document.
@@
 Combining the two cases, we define the resultant address for the new account $a$:
-\begin{eqnarray}
-a & \equiv & A(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
-\label{eq:new-address} A(s, n, \zeta, \mathbf{i}) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big(B(s, n, \zeta, \mathbf{i})\big)\Big) \\
-B(s, n, \zeta, \mathbf{i}) & \equiv & \begin{cases}
+\begin{align}
+a & \equiv \mathtt{ADDR}(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
+\label{eq:new-address} \mathtt{ADDR}(s, n, \zeta, \mathbf{i}) & \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big( L_{\mathrm{A}}(s, n, \zeta, \mathbf{i})\big) \Big) \\
+L_{\mathrm{A}}(s, n, \zeta, \mathbf{i}) & \equiv \begin{cases}
 \mathtt{RLP}\big(\;(s, n)\;\big) & \text{if}\ \zeta = \varnothing \\
@@
 \end{cases}
-\end{eqnarray}
+\end{align}
 
@@
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
+&&&& is too low to fulfil the value transfer); and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
 &&&& address of the newly created account (\ref{eq:new-address}). \\

### 2021-06-18 Paper.tex
Cosmetic issue with I_H

@@
 \item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathrm{H}}$, the block header of the present block.
 \item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathrm{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum\\

### 2021-06-18 Paper.tex
BALANCE clarification

@@
 0x1b & {\small SHL} & 2 & 1 & Left shift operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \mod 2^{256}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \bmod 2^{256}$ \\
 \midrule
@@
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\
-&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
 \midrule
@@
 \begin{cases} \mathbf{b}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
-\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
+\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
 \bottomrule
@@
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an\\
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, t) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 \begin{multline}
-  E_{\mathrm{rmh}}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  E_{\mathrm{rmh}}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \bmod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \bmod n]) \\
   \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{\mathrm{rmh}}(\mathbf{x}, j) \quad \forall \quad j < i
@@
  E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \mod c_\mathrm{{size}}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{\mathrm{wordbytes}} \rfloor]) \mod c_{\mathrm{size}}] \big) & \text{otherwise}
+\texttt{KEC512}(\mathbf{c}[i \bmod c_\mathrm{{size}}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \bmod \lfloor J_{\mathrm{hashbytes}} / J_{\mathrm{wordbytes}} \rfloor]) \bmod c_{\mathrm{size}}] \big) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{wordbytes}}}\right\rfloor]) \mod \left\lfloor\frac{d_{\mathrm{size}} / J_{\mathrm{hashbytes}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
+ E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \bmod \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{wordbytes}}}\right\rfloor]) \bmod \left\lfloor\frac{d_{\mathrm{size}} / J_{\mathrm{hashbytes}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
 \end{equation}

### 2021-06-21 Paper.tex
Specify the behaviors of EXTCODESIZE and EXTCODECOPY when the target account does not exist

@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
+\begin{cases}
+\lVert \mathbf{b} \rVert & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing \\
+0                        & \text{otherwise}
+\end{cases}$ \\
 &&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
@@
 &&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
+&&&& We assume $\mathbf{b} \equiv ()$ if $\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] = \varnothing$. \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule

### 2021-06-22 Paper.tex
EIP-2929 part 1

@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}})
-\end{equation}
-
-\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.\hypertarget{refund_balance_defn_words_A__r}{} Finally there is $A_{\mathrm{r}}$, the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}}, A_{\mathbf{a}}, A_{\mathbf{k}})
+\end{equation}
+
+\hypertarget{self_destruct_set_wordy_defn_A__s}{}The tuple contents include $A_{\mathbf{s}}$, the self-destruct set: a set of accounts that will be discarded following the transaction's completion.
+\hypertarget{tx_log_series_wordy_defn_A__l}{} $A_{\mathbf{l}}$ is the log series: this is a series of archived and indexable `checkpoints' in VM code execution that allow for contract-calls to be easily tracked by onlookers external to the Ethereum world (such as decentralised application front-ends).
+\hypertarget{tx_touched_accounts_wordy_defn_A__t}{} $A_{\mathbf{t}}$ is the set of touched accounts, of which the empty ones are deleted at the end of a transaction.
+\hypertarget{refund_balance_defn_words_A__r}{}$A_{\mathrm{r}}$ is the refund balance, increased through using the \hyperlink{SSTORE}{{\small SSTORE}} instruction in order to reset contract storage to zero from some non-zero value.
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
-
-We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts and a zero refund balance:
-\begin{equation}
-A^0 \equiv (\varnothing,(), \varnothing, 0)
+Finally, EIP-2929 by \cite{EIP-2929} introduced \hypertarget{accessed_addresses_defn_words_A__a}{}$A_{\mathbf{a}}$, the set of accessed account addresses, and \hypertarget{accessed_storage_keys_defn_words_A__k}{}$A_{\mathbf{k}}$, the set of accessed storage keys.
+
+We define the empty accrued substate $A^0$ to have no self-destructs, no logs, no touched accounts, zero refund balance, all precompiled contracts in the accessed addresses, and no accessed storage:
+\begin{equation}
+A^0 \equiv (\varnothing, (), \varnothing, 0, \hyperlink{precompiled_set}{\pi}, \varnothing)
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
-\Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\Theta_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, T_{\mathrm{t}}, T_{\mathrm{t}}, &\\ \quad\quad g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-where $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
+(\boldsymbol{\sigma}_{\mathrm{P}}, g', A', z) \equiv \begin{cases}
+\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^+, S(T), S(T), g, &\\ \quad\; T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^+, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\end{cases}
+\end{equation}
+where
+\begin{equation}
+A^+ \equiv A^0 \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A^0_{\mathbf{a}} \cup \{S(T)\}
+\end{equation}
+and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
 \begin{equation}
@@
 \end{equation}
-and $T_{\mathrm{o}}$ is the original transactor, which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code.
 
@@
 \begin{equation}
-	A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \sum_{i \in A_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
+A^*_{\mathrm{r}} \equiv A'_{\mathrm{r}} + \sum_{i \in A'_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
 \end{equation}
@@
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A^*_{\mathrm{r}}} \right\}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A'_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\linkdest{self_destruct_list_A__s}{}\forall i \in A'_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A'_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A'_{\mathbf{l}}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 
-There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor ($o$), available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
+There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor\footnote{which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code} ($o$),
+available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
 The salt \linkdest{salt}{$\zeta$} might be missing ($\zeta = \varnothing$); formally,
@@
 
-We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ to the tuple containing the new state, remaining gas, accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
+We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued transaction substate $A$ to the tuple containing the new state, remaining gas, new accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, A, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
 \end{equation}
@@
 \end{align}
-
 where $\cdot$ is the concatenation of byte arrays, $\mathcal{B}_{a..b}(X)$ evaluates to a binary value containing the bits of indices in the range $[a, b]$ of the binary data $X$, and $\boldsymbol{\sigma}[x]$ is the address state of $x$, or $\varnothing$ if none exists.
 Note we use one fewer than the sender's nonce value; we assert that we have incremented the sender account's nonce prior to this call, and so the value used is the sender's nonce at the beginning of the responsible transaction or VM operation.
+
+The address of the new account is added to the set of accessed accounts:
+\begin{equation}
+A^+ \equiv A \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{a\}
+\end{equation}
 
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\Xi$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the accrued substate $A$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, I, \{s, a\}) \\
+As such, the code execution function $\hyperlink{xi_def}{\Xi}$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the new accrued substate $A'$ and the body code of the account $\mathbf{o}$.
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, A^+, I) \\
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', g', A, z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, A, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**},A, \mathbf{o}) & \equiv & \Xi\\
+(\boldsymbol{\sigma}^{**}, g^{**}, A', \mathbf{o}) & \equiv & \Xi\\
 I_{\mathrm{a}} & \equiv & r \\
@@
 I_{\mathrm{e}} & \equiv & e \\
-I_{\mathrm{w}} & \equiv & w \\
-\mathbf{t} & \equiv & \{s, r\}
+I_{\mathrm{w}} & \equiv & w
 \end{eqnarray}
@@
 \Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 9 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\begin{equation}
-\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
+\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 1 \\
+\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 2 \\
+\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 3 \\
+\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 4 \\
+\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 5 \\
+\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 6 \\
+\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 7 \\
+\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 9 \\
+\Xi                     (\boldsymbol{\sigma}_1, g, A^+, I) & \text{otherwise} \end{cases}
+\end{equation}
+\begin{equation}
+A^+ \equiv A \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{c\}
+\end{equation}
+and
+\begin{equation}
+\mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
 It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
@@
 The contracts in addresses 1 to 9 execute the elliptic curve public key recovery function, the SHA2 256-bit hash scheme, the RIPEMD 160-bit hash scheme, the identity function, arbitrary precision modular exponentiation, elliptic curve addition, elliptic curve scalar multiplication, an elliptic curve pairing check, and the BLAKE2 compression function $\mathtt{F}$ respectively.
-
 Their full formal definition is in Appendix \ref{app:precompiled}.
+
+We denote the set of the addresses of the precompiled contracts by $\pi$:
+\begin{equation}\hypertarget{precompiled_set}{}
+\pi \equiv \{1, 2, 3, 4, 5, 6, 7, 8, 9 \}
+\end{equation}
 
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, I, T) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A, \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A, ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+\Xi(\boldsymbol{\sigma}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
@@
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{\mathrm{extcode}}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
-$G_{\mathrm{balance}}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{\mathrm{sload}}$ & 800 & Paid for an {\small SLOAD} operation. \\
+$G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
+$G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
+$G_{\mathrm{coldsload}}$ & 2100 & Cost of a cold storage access. \\
 $G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{\mathrm{sreset}}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+$G_{\mathrm{sreset}}$ & 2900 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
 &&is set to zero. \\
@@
 \linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{\mathrm{call}}$ & 700 & Paid for a {\small CALL} operation. \\
 $G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{\mathrm{mem}}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{\mathrm{mem}}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
-C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv C_{\mathrm{mem}}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{\mathrm{mem}}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SSTORE} \\
 G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
@@
 
-G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) & \text{if} \quad w \in W_{\mathrm{extaccount}}\\
 G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
@@
 G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
-C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w \in W_{\mathrm{call}} \\
+C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{if} \quad w \in W_{\mathrm{call}} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
@@
 G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
+C_\text{\tiny SLOAD}(\boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SLOAD}\\
 G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
@@
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
 G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
@@
 \end{equation}
-
-with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below. We define the following subsets of instructions:
+\begin{equation}
+C_{\mathrm{aaccess}}(x, A) \equiv
+\begin{cases}
+G_{\mathrm{warmaccess}}        & \text{if} \quad x \in \hyperlink{accessed_addresses_defn_words_A__a}{A_{\mathbf{a}}}\\
+G_{\mathrm{coldaccountaccess}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+with $C_\text{\tiny CALL}$, $C_\text{\tiny SELFDESTRUCT}$, $C_\text{\tiny SLOAD}$ and $C_\text{\tiny SSTORE}$ as specified in the appropriate section below.
+We define the following subsets of instructions:
 
@@
 
-$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}, {\small EXTCODEHASH}\}
+$W_{\mathrm{extaccount}}$ = \{{\small BALANCE}, {\small EXTCODESIZE}, {\small EXTCODEHASH}\}
 
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \midrule
@@
 &&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \midrule
@@
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \end{tabu}
@@
 \begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160} \}$ \\
 \bottomrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
+&&&& $A'_{\mathbf{k}} \equiv A_{\mathbf{k}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
+&&&& $C_{\text{\tiny SLOAD}}(\boldsymbol{\mu}, A, I) \equiv
+\begin{cases}
+G_{\mathrm{warmaccess}} & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{k}} \\
+G_{\mathrm{coldsload}}  & \text{otherwise}
+\end{cases}$ \\
 \midrule
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
+&&&& $A'_{\mathbf{k}} \equiv A_{\mathbf{k}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
 &&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{\mathrm{r}}$ are specified by EIP-2200 as follows. \\
-&&&& The checkpoint (``original'') state $\hyperlink{sigma_0}{\boldsymbol{\sigma}_0}$ is the state if the current transaction were to revert. \\
+&&&& We remind the reader that the checkpoint (``original'') state $\hyperlink{sigma_0}{\boldsymbol{\sigma}_0}$ is the state \\
+&&&& if the current transaction were to revert. \\
 &&&& Let $v_0 = \boldsymbol{\sigma}_0[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ be the original value of the storage slot. \\
@@
 &&&& Then: \\
-&&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{sload}}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
+&&&& $\!\begin{aligned}
+C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) &\equiv
+\begin{cases}
+0                      & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{k}} \\
+G_{\mathrm{coldsload}} & \text{otherwise} \\
+\end{cases} \\
+&+
+\begin{cases}
+G_{\mathrm{warmaccess}}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
 G_{\mathrm{sset}}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
 G_{\mathrm{sreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
-\end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
+\end{cases}
+\end{aligned}$ \\
+&&&& \linkdest{A r}{}$A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
 R_{\mathrm{sclear}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
@@
 \end{cases}$ \\
+&&&& where \\
 &&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
@@
 &&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
-G_{\mathrm{sset}} - G_{\mathrm{sload}}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
-G_{\mathrm{sreset}} - G_{\mathrm{sload}} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
+G_{\mathrm{sset}} - G_{\mathrm{warmaccess}}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
+G_{\mathrm{sreset}} - G_{\mathrm{warmaccess}} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
 0 & \text{otherwise} \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{pc}}$ \\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& $\hyperlink{salt}{\zeta} \equiv \varnothing$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A^+, \mathbf{o}) \equiv \begin{cases}\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\ \big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, \varnothing\big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \equiv \begin{cases}
+\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\
+\big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, A, () \big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
-&&&& $A' \equiv A \Cup A^+$ which abbreviates: $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup A^+_{\mathbf{s}} \quad \wedge \quad A'_{\mathbf{l}} \equiv A_{\mathbf{l}} \cdot A^+_{\mathbf{l}} \quad \wedge$ \\
-&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup A^+_{\mathbf{t}} \wedge \quad A'_{\mathbf{r}} \equiv A_{\mathbf{r}} + A^+_{\mathbf{r}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}
+\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+& \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
+(\boldsymbol{\sigma}, g, A, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& $A' \equiv A \Cup A^+$ \\
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
@@
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
-&&&&\linkdest{tiny CALL}{} $C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ \\
-&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
+&&&& \linkdest{tiny CALL}{}$C_{\text{\tiny CALL}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) + C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)$ \\
+&&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv  \begin{cases}
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{otherwise}
 \end{cases}$ \\
-&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu})\\
+&&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv \begin{cases}
+\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv C_{\text{\tiny ACCESS}}(\boldsymbol{\mu}, A) + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny ACCESS}}(\boldsymbol{\mu}, A) \equiv \begin{cases}
+G_{\mathrm{warmaccess}}        & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160} \in A_{\mathbf{a}} \\
+G_{\mathrm{coldaccountaccess}} & \text{otherwise}
+\end{cases}$ \\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}

### 2021-06-22 Paper.tex
Revert an unintended change from PR #805
Introduce protocol update block numbers (#804)
* Introduce fork block numbers

* Don't use fork for protocol updates

--- 2021-06-18 Paper.tex
@@
 In order to follow back the history of a path, one must reference multiple versions of this document.
-
-Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
+Here are the block numbers of protocol updates on the Ethereum main network:
+\par
+\begin{center}
+\begin{tabular}{lr}
+\toprule
+Name & First Block Number \\
+\midrule
+$F_{\mathrm{Homestead}}$         &  1150000 \\
+$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
+$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
+$F_{\mathrm{Byzantium}}$         &  4370000 \\
+$F_{\mathrm{Constantinople}}$    &  7280000 \\
+$F_{\mathrm{Petersburg}}$        &  7280000 \\
+$F_{\mathrm{Istanbul}}$          &  9069000 \\
+$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
+$F_{\mathrm{Berlin}}$            & 12244000 \\
+\bottomrule
+\end{tabular}
+\end{center}
+\par
+
+Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
 In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, t) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}

### 2021-06-23 Paper.tex
Fix a hyperlink

--- 2021-06-22 Paper.tex
@@
 \Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 8 \\
-\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad r = 9 \\
+\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 1 \\
+\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 2 \\
+\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 3 \\
+\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 4 \\
+\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 5 \\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 6 \\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 7 \\
+\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 9 \\
 \Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$.
+Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{SSTORE}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$.
 The last condition was introduced in EIP-1706 by \cite{EIP-1706} (part of EIP-2200 by \cite{EIP-2200}).

### 2021-06-28 Paper.tex
3607: update to use empty code hash instead of codesize

--- 2021-06-22 Paper.tex
@@
 In order to follow back the history of a path, one must reference multiple versions of this document.
-Here are the block numbers of protocol updates on the Ethereum main network:
-\par
-\begin{center}
-\begin{tabular}{lr}
-\toprule
-Name & First Block Number \\
-\midrule
-$F_{\mathrm{Homestead}}$         &  1150000 \\
-$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
-$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
-$F_{\mathrm{Byzantium}}$         &  4370000 \\
-$F_{\mathrm{Constantinople}}$    &  7280000 \\
-$F_{\mathrm{Petersburg}}$        &  7280000 \\
-$F_{\mathrm{Istanbul}}$          &  9069000 \\
-$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
-$F_{\mathrm{Berlin}}$            & 12244000 \\
-\bottomrule
-\end{tabular}
-\end{center}
-\par
-
-Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
+
+Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
 In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
@@
 \begin{equation}
-\texttt{TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
-\end{equation}
-
-The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
-\begin{equation}
-L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
+\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
+\end{equation}
+
+The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
+\begin{equation}
+L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
 \end{equation}
@@
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
+The sender of a transaction can not be a contract.
 While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
@@
 \begin{equation}
-L_{\mathrm{T}}(T) \equiv \begin{cases}
+L_{T}(T) \equiv \begin{cases}
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
@@
 \begin{equation}
-B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
+B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
 \end{equation}
@@
 
-%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{\mathrm{B}}$; it is merely a convenience equivalence.
+%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.
 
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{\mathrm{T}}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
 \linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
@@
 \begin{equation}
-\mathtt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
-\end{equation}
-
-Thus $\texttt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{\mathrm{H}})$ is the parent block of $B$, defined directly.
+\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
+\end{equation}
+
+Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.
 
@@
 
-\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
+\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
 We assert the types and order of the structure for when the RLP transformation is required:
 \begin{eqnarray}
-\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
 \end{eqnarray}
@@
 
-We define $P(B_{\mathrm{H}})$ to be the parent block of $B$, formally:
-\begin{equation}
-P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{\mathrm{H}})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
+We define $P(B_{H})$ to be the parent block of $B$, formally:
+\begin{equation}
+P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
 \end{equation}
@@
 \begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
+\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
+\end{equation}
+\begin{equation}
+\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
 \end{equation}
@@
 \kappa &\equiv \begin{cases} 
-  3000000  & \text{if} \quad F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
-  5000000  & \text{if} \quad F_{\mathrm{Constantinople}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Muir Glacier}} \\
-  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Muir Glacier}} \\
+  3000000  & \text{if} \quad 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
+  5000000  & \text{if} \quad 7280000 \leqslant H_{\mathrm{i}} < 9200000 \\
+  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant 9200000 \\
 \end{cases}
@@
 \begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000
@@
 \nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
-\nonumber& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
-\nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
+\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
+\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
 \nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
@@
 \item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
+\item the sender account has no contract code deployed ($\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} = \texttt{KEC}\big( () \big)$).
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
@@
 \boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
+\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} & = & \texttt{KEC}\big( () \big) \quad \wedge \\
 \linkdest{transaction_nonce}{}T_{\mathrm{n}} & = & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} \quad \wedge \\
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
+T_{\mathrm{g}} & \leqslant & {B_{H}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
 \end{array}
@@
 
-Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
+Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{H}}_{\mathrm{l}}$.
 
@@
 
-Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
-\begin{equation}
-(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
+Evaluating $\boldsymbol{\sigma}_{P}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{P}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
+\begin{equation}
+(\boldsymbol{\sigma}_{P}, g', A, z) \equiv \begin{cases}
 \Lambda_{4}(\boldsymbol{\sigma}_0, S(T), T_{\mathrm{o}}, g, &\\ \quad\quad T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
@@
 \begin{equation}
-	A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \sum_{i \in A_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
+	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
 \end{equation}
@@
 
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{sigma_n}{later} for state and nonce validation.
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.
 
@@
 Combining the two cases, we define the resultant address for the new account $a$:
-\begin{align}
-a & \equiv \mathtt{ADDR}(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
-\label{eq:new-address} \mathtt{ADDR}(s, n, \zeta, \mathbf{i}) & \equiv \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big( L_{\mathrm{A}}(s, n, \zeta, \mathbf{i})\big) \Big) \\
-L_{\mathrm{A}}(s, n, \zeta, \mathbf{i}) & \equiv \begin{cases}
+\begin{eqnarray}
+a & \equiv & A(s, \boldsymbol{\sigma}[s]_{\mathrm{n}} - 1, \zeta, \mathbf{i}) \\
+\label{eq:new-address} A(s, n, \zeta, \mathbf{i}) & \equiv & \mathcal{B}_{96..255}\Big(\mathtt{KEC}\big(B(s, n, \zeta, \mathbf{i})\big)\Big) \\
+B(s, n, \zeta, \mathbf{i}) & \equiv & \begin{cases}
 \mathtt{RLP}\big(\;(s, n)\;\big) & \text{if}\ \zeta = \varnothing \\
@@
 \end{cases}
-\end{align}
+\end{eqnarray}
 
@@
 
-$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{\mathrm{H}}$ has no special treatment and is determined from the blockchain.
+$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
 
@@
 \begin{equation}
-c \equiv G_{\mathrm{codedeposit}} \times \lVert \mathbf{o} \rVert
+c \equiv G_{codedeposit} \times \lVert \mathbf{o} \rVert
 \end{equation}
@@
 \begin{align}
-\quad g' \equiv & \begin{cases}
+\quad g' &\equiv \begin{cases}
 0 & \text{if} \quad F \\
@@
 \end{cases} \\
-\quad \boldsymbol{\sigma}' \equiv & \begin{cases}
-\boldsymbol{\sigma} & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
+\quad \boldsymbol{\sigma}' &\equiv  \begin{cases}
+\boldsymbol{\sigma} & \text{if} \quad F \\
 \boldsymbol{\sigma}^{**} \quad \text{except:} & \\
@@
 \end{cases} \\
-\quad z \equiv & \begin{cases}
-0 & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
+\quad z &\equiv \begin{cases}
+0 & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \lor g^{**} < c \\
 1 & \text{otherwise}
@@
 \nonumber \text{where} \\
-F \equiv  & \big( \boldsymbol{\sigma}[a] \neq \varnothing \ \wedge\ \big(\boldsymbol{\sigma}[a]_c \neq \texttt{\small KEC}\big(()\big) \vee \boldsymbol{\sigma}[a]_n \neq 0 \big) \big) \quad \vee \\
-\nonumber &(\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \quad \vee \\
-\nonumber &g^{**} < c \quad \vee \\
-\nonumber &\lVert \mathbf{o} \rVert > 24576
+F &\equiv \big((\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} = \varnothing) \vee\  g^{**} < c \ \vee\  \lVert \mathbf{o} \rVert > 24576\big)
 \end{align}
@@
 Note that intention is that the result is either a successfully created new contract with its endowment, or no new contract with no transfer of value.
-In addition, observe that if the execution of the initialising code \hyperlink{REVERT}{reverts} ($\boldsymbol{\sigma}^{**} = \varnothing \ \wedge\ \mathbf{o} \neq \varnothing$), the resultant gas $g'$ is not depleted (provided there was no other exception), but no new account is created.
 
@@
 
-Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array~$\mathbf{o}$.
+Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array $\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
@@
 \item\linkdest{exec_machine_code_I__b}{}\linkdest{I__b} $I_{\mathbf{b}}$, the byte array that is the machine code to be executed.
-\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathrm{H}}$, the block header of the present block.
+\item\linkdest{exec_block_header_I__H}{}\linkdest{I__H} $I_{\mathbf{H}}$, the block header of the present block.
 \item\linkdest{exec_call_or_create_depth_I__e}{}\linkdest{I__e} $I_{\mathrm{e}}$, the depth of the present message-call or contract-creation (i.e.
@@
 ( \neg I_{\mathrm{w}} \; \wedge \; W(w, \boldsymbol{\mu}) ) \quad \vee \\
-( w = \text{\small SSTORE} \; \wedge \; \boldsymbol{\mu}_g \leqslant G_{\mathrm{callstipend}} )
+( w = \text{\small SSTORE} \; \wedge \; \boldsymbol{\mu}_g \leqslant G_{callstipend} )
 \end{array}
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$.
+Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{sstore}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{callstipend}}$.
 The last condition was introduced in EIP-1706 by \cite{EIP-1706} (part of EIP-2200 by \cite{EIP-2200}).
@@
 \begin{equation}
-D_{\mathrm{J}}(\mathbf{c}, i) \equiv \begin{cases}
+D_{J}(\mathbf{c}, i) \equiv \begin{cases}
 \{\} & \text{if} \quad i \geqslant \lVert \mathbf{c} \rVert  \\
-\{ i \} \cup D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
+\{ i \} \cup D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
 \quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
+D_{J}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
 \end{cases}
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{g} & \equiv & \boldsymbol{\mu}_{g} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
 \quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
@@
 
-The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{\mathrm{H}}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
-\begin{equation}
-\hyperlink{H__g}{B_{\mathrm{H}}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
+The given \textbf{gasUsed} must correspond faithfully to the transactions listed: \hyperlink{H__g}{${B_{H}}_{\mathrm{g}}$}, the total gas used in the block, must be equal to the accumulated gas used according to the final transaction:
+\begin{equation}
+\hyperlink{H__g}{B_{H}}_{\mathrm{g}} = \hyperlink{ell}{\ell}(\hyperlink{R__u}{\mathbf{R})_{\mathrm{u}}}
 \end{equation}
@@
 \Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
+\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{H}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
 \\ \nonumber
@@
 \mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{\mathrm{H}}}_{\mathrm{i}})\right) R_{\mathrm{block}}
+R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{H}}_{\mathrm{i}})\right) R_{\mathrm{block}}
 \end{eqnarray}
@@
 R_{\mathrm{block}} = 10^{18} \times \begin{cases}
-5 &\text{if}\ H_{\mathrm{i}} < F_{\mathrm{Byzantium}} \\
-3 &\text{if}\ F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
-2 &\text{if}\ H_{\mathrm{i}} \geqslant F_{\mathrm{Constantinople}} \\
+5 &\text{if}\ H_{\mathrm{i}} < 4370000 \\
+3 &\text{if}\ 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
+2 &\text{if}\ H_{\mathrm{i}} \geqslant 7280000 \\
 \end{cases} \\
@@
 \Gamma(B) \equiv \begin{cases}
-\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{\mathrm{H}}) = \varnothing \\
-\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{\mathrm{H}})_{\mathrm{H}}}_{\mathrm{r}} & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, $\mathtt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
+\boldsymbol{\sigma}_0 & \text{if} \quad P(B_{H}) = \varnothing \\
+\boldsymbol{\sigma}_{\mathrm{i}}: \mathtt{\hyperlink{trie}{TRIE}}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}})) = {P(B_{H})_{H}}_{\mathrm{r}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, $\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma}_{\mathrm{i}}))$ means the hash of the root node of a trie of state $\boldsymbol{\sigma}_{\mathrm{i}}$; it is assumed that implementations will store this in the state database, which is trivial and efficient since the trie is by nature an immutable data structure.
 
@@
 We also define the $n$th state $\boldsymbol{\sigma}[n]$, which is defined simply as the state resulting from applying the corresponding transaction to the state resulting from the previous transaction (or the block's initial state in the case of the first such transaction):
-\begin{equation}\linkdest{sigma_n}{}
+\begin{equation}
 \boldsymbol{\sigma}[n] = \begin{cases} \hyperlink{Gamma}{\Gamma}(B) & \text{if} \quad n < 0 \\ \hyperlink{Upsilon}{\Upsilon}(\boldsymbol{\sigma}[n - 1], B_{\mathbf{T}}[n]) & \text{otherwise} \end{cases}
@@
 There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{\mathrm{cacheinit}}}$ bytes in initial size.
+From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{cacheinit}}$ bytes in initial size.
 Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{\mathrm{datasetinit}}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
+From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{datasetinit}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
 Full clients and miners store the dataset.
@@
 Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $\hyperlink{J__epoch}{J_{\mathrm{epoch}}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
+The large dataset is updated once every $\hyperlink{J__epoch}{J_{epoch}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
 The mentioned parameters as well as the algorithm is explained in detail in Appendix \ref{app:ethash}.
@@
 \Xi_{\mathtt{EXPMOD}} &\equiv& \Xi_{\mathtt{PRE}} \quad \text{except:} \\
-g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{\mathrm{M}},\ell_{\mathrm{B}})\big)\max(\ell'_{\mathrm{E}},1)}{G_{\mathrm{quaddivisor}}}\right\rfloor \\
+g_{\mathrm{r}} &=& \left\lfloor\frac{f\big(\max(\ell_{M},\ell_{B})\big)\max(\ell'_{E},1)}{G_{quaddivisor}}\right\rfloor \\
 f(x) &\equiv& \begin{cases}
@@
 \end{cases}\\
-\ell'_{\mathrm{E}} &=& \begin{cases}
-0 & \text{if}\ \ell_{\mathrm{E}}\le 32\wedge E=0 \\
-\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{\mathrm{E}}\le 32 \wedge E \neq 0 \\
-8(\ell_{\mathrm{E}} - 32) + \lfloor \log_2(i[(96+\ell_{\mathrm{B}})..(127+\ell_{\mathrm{B}})]) \rfloor & \text{if}\ 32 < \ell_{\mathrm{E}} \wedge i[(96 + \ell_{\mathrm{B}})..(127 + \ell_{\mathrm{B}})]\neq 0 \\
-8(\ell_{\mathrm{E}} - 32) & \text{otherwise} \\
+\ell'_{E} &=& \begin{cases}
+0 & \text{if}\ \ell_{E}\le 32\wedge E=0 \\
+\lfloor \log_2(E)\rfloor &\text{if}\ \ell_{E}\le 32 \wedge E \neq 0 \\
+8(\ell_{E} - 32) + \lfloor \log_2(i[(96+\ell_{B})..(127+\ell_{B})]) \rfloor & \text{if}\ 32 < \ell_{E} \wedge i[(96 + \ell_{B})..(127 + \ell_{B})]\neq 0 \\
+8(\ell_{E} - 32) & \text{otherwise} \\
 \end{cases} \\
-\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{\mathrm{M}}} \\
-\ell_{\mathrm{B}} &\equiv& i[0..31] \\
-\ell_{\mathrm{E}} &\equiv& i[32..63] \\
-\ell_{\mathrm{M}} &\equiv& i[64..95] \\
-B &\equiv& i[96..(95+\ell_{\mathrm{B}})] \\
-E &\equiv& i[(96+\ell_{\mathrm{B}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})] \\
-M &\equiv& i[(96+\ell_{\mathrm{B}}+\ell_{\mathrm{E}})..(95+\ell_{\mathrm{B}}+\ell_{\mathrm{E}}+\ell_{\mathrm{M}})] \\
+\mathbf{o} &=& \left(B^E\bmod M\right)\in\mathbb{N}_{8\ell_{M}} \\
+\ell_{B} &\equiv& i[0..31] \\
+\ell_{E} &\equiv& i[32..63] \\
+\ell_{M} &\equiv& i[64..95] \\
+B &\equiv& i[96..(95+\ell_{B})] \\
+E &\equiv& i[(96+\ell_{B})..(95+\ell_{B}+\ell_{E})] \\
+M &\equiv& i[(96+\ell_{B}+\ell_{E})..(95+\ell_{B}+\ell_{E}+\ell_{M})] \\
 i[x] &\equiv& \begin{cases}
@@
 \end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_{\mathrm{S}}(T) )
+h(T) & \equiv & \mathtt{KEC}( L_{S}(T) )
 \end{eqnarray}
@@
 \midrule
-$G_{\mathrm{zero}}$ & 0 & Nothing paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
-$G_{\mathrm{jumpdest}}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
-$G_{\mathrm{base}}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
-$G_{\mathrm{verylow}}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
+$G_{zero}$ & 0 & Nothing paid for operations of the set {\small $W_{zero}$}. \\
+$G_{jumpdest}$ & 1 & Amount of gas to pay for a {\small JUMPDEST} operation. \\
+$G_{base}$ & 2 & Amount of gas to pay for operations of the set {\small $W_{base}$}. \\
+$G_{verylow}$ & 3 & Amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
 $G_{\mathrm{low}}$ & 5 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
-$G_{\mathrm{mid}}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
+$G_{mid}$ & 8 & Amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
-$G_{\mathrm{extcode}}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
-$G_{\mathrm{balance}}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
-$G_{\mathrm{sload}}$ & 800 & Paid for an {\small SLOAD} operation. \\
-$G_{\mathrm{sset}}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
-$G_{\mathrm{sreset}}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
+$G_{extcode}$ & 700 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
+$G_{balance}$ & 700 & Amount of gas to pay for a {\small BALANCE} operation. \\
+$G_{sload}$ & 800 & Paid for an {\small SLOAD} operation. \\
+$G_{sset}$ & 20000 & Paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
+$G_{sreset}$ & 5000 & Paid for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or\\
 &&is set to zero. \\
-$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
+$R_{sclear}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
 &&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
-\linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
-$G_{\mathrm{create}}$ & 32000 & Paid for a {\small CREATE} operation. \\
-\linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-$G_{\mathrm{call}}$ & 700 & Paid for a {\small CALL} operation. \\
-$G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
-\linkdest{G__callstipend}{}$G_{\mathrm{callstipend}}$ & 2300 & A stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \\
-\linkdest{G__newaccount}{}$G_{\mathrm{newaccount}}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
-$G_{\mathrm{exp}}$ & 10 & Partial payment for an {\small EXP} operation. \\
-$G_{\mathrm{expbyte}}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
-$G_{\mathrm{memory}}$ & 3 & Paid for every additional word when expanding memory. \\
+\linkdest{R__selfdestruct}{}$R_{selfdestruct}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+\linkdest{G__selfdestruct}{}$G_{selfdestruct}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
+$G_{create}$ & 32000 & Paid for a {\small CREATE} operation. \\
+\linkdest{G__codedeposit}{}$G_{codedeposit}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+$G_{call}$ & 700 & Paid for a {\small CALL} operation. \\
+$G_{callvalue}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
+\linkdest{G__callstipend}{}$G_{callstipend}$ & 2300 & A stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \\
+\linkdest{G__newaccount}{}$G_{newaccount}$ & 25000 & Paid for a {\small CALL} or {\small SELFDESTRUCT} operation which creates an account. \\
+$G_{exp}$ & 10 & Partial payment for an {\small EXP} operation. \\
+$G_{expbyte}$ & 50 & Partial payment when multiplied by the number of bytes in the exponent for the {\small EXP} operation. \\
+$G_{memory}$ & 3 & Paid for every additional word when expanding memory. \\
 \linkdest{G__txcreate}{}$G_\text{txcreate}$ & 32000 & Paid by all contract-creating transactions after the {\textit{Homestead} transition}.\\
-\linkdest{G__txdatazero}{}$G_{\mathrm{txdatazero}}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
-\linkdest{G__txdatanonzero}{}$G_{\mathrm{txdatanonzero}}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
-\linkdest{G__transaction}{}$G_{\mathrm{transaction}}$ & 21000 & Paid for every transaction. \\
+\linkdest{G__txdatazero}{}$G_{txdatazero}$ & 4 & Paid for every zero byte of data or code for a transaction. \\
+\linkdest{G__txdatanonzero}{}$G_{txdatanonzero}$ & 16 & Paid for every non-zero byte of data or code for a transaction. \\
+\linkdest{G__transaction}{}$G_{transaction}$ & 21000 & Paid for every transaction. \\
 $G_{\mathrm{log}}$ & 375 & Partial payment for a {\small LOG} operation. \\
@@
 $G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{\mathrm{sha3}}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{\mathrm{sha3word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
-$G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
-$G_{\mathrm{quaddivisor}}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
+$G_{sha3}$ & 30 & Paid for each {\small SHA3} operation. \\
+$G_{sha3word}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{copy}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
+$G_{blockhash}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{quaddivisor}$ & 20 & The quadratic coefficient of the input sizes of the exponentiation-over-modulo precompiled\\
 &&contract. \\
@@
 \begin{equation}
-C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{\mathrm{mem}}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{\mathrm{mem}}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
+C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_{\mathrm{i}})-C_{mem}(\boldsymbol{\mu}_{\mathrm{i}}) + \begin{cases}
 C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
-G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
-G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
-G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w \in W_{\mathrm{copy}} \\
-
-G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
-G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
-G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
+G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] = 0 \\
+G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_{\mathbf{s}}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_{\mathbf{s}}[1] > 0 \\
+G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[2] \div 32\rceil & \text{if} \quad w \in W_{\mathrm{copy}} \\
+
+G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
+G_{extcode} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
+G_{log}+G_{logdata}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
 C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w \in W_{\mathrm{call}} \\
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
-G_{\mathrm{create}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
-G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
-G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
-G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
-G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
-G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
-G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
+G_{create} & \text{if} \quad w = \text{\small CREATE}\\
+G_{create}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{sha3}+G_{sha3word} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
+G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
+G_{zero} & \text{if} \quad w \in W_{zero}\\
+G_{base} & \text{if} \quad w \in W_{base}\\
+G_{verylow} & \text{if} \quad w \in W_{verylow}\\
 G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
-G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
+G_{mid} & \text{if} \quad w \in W_{mid}\\
 G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
-G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
-G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
+G_{blockhash} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
 \end{cases}
@@
 \begin{equation}
-C_{\mathrm{mem}}(a) \equiv G_{\mathrm{memory}} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
+C_{mem}(a) \equiv G_{memory} \cdot a + \left\lfloor \dfrac{a^2}{512} \right\rfloor
 \end{equation}
@@
 
-$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
-
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
-
-$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+$W_{zero}$ = \{{\small STOP}, {\small RETURN}, {\small REVERT}\}
+
+$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+
+$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
 
@@
 
-$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
+$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}
 
@@
 
-Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
+Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
 Such accesses must be for non-zero number of bytes.
@@
 
-Note also that $C_{\mathrm{mem}}$ is the memory cost function (the expansion function being the difference between the cost before and after).
+Note also that $C_{mem}$ is the memory cost function (the expansion function being the difference between the cost before and after).
 It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
@@
 0x05 & {\small SDIV} & 2 & 1 & Signed integer division operation (truncated). \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[1] = 0\\ -2^{255} & \text{if} \> \boldsymbol{\mu}_{\mathbf{s}}[0] = -2^{255} \wedge \, \boldsymbol{\mu}_{\mathbf{s}}[1] = -1\\ \mathbf{sgn} (\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]) \lfloor |\boldsymbol{\mu}_{\mathbf{s}}[0] \div \boldsymbol{\mu}_{\mathbf{s}}[1]| \rfloor & \text{otherwise}\end{cases}$  \\
 &&&& Where all values are treated as two's complement signed 256-bit integers. \\
@@
 0x08 & {\small ADDMOD} & 3 & 1 & Modulo addition operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \bmod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
@@
 0x09 & {\small MULMOD} & 3 & 1 & Modulo multiplication operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \bmod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}0 & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] = 0\\ (\boldsymbol{\mu}_{\mathbf{s}}[0] \times \boldsymbol{\mu}_{\mathbf{s}}[1]) \mod \boldsymbol{\mu}_{\mathbf{s}}[2] & \text{otherwise}\end{cases}$  \\
 &&&& All intermediate calculations of this operation are not subject to the $2^{256}$ \\
@@
 0x1b & {\small SHL} & 2 & 1 & Left shift operation. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \bmod 2^{256}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv (\boldsymbol{\mu}_{\mathbf{s}}[1] \times 2^{\boldsymbol{\mu}_{\mathbf{s}}[0]}) \mod 2^{256}$ \\
 \midrule
@@
 0x31 & {\small BALANCE} & 1 & 1 & Get balance of the given account. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \begin{cases}\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0]]_{\mathrm{b}}& \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}] \neq \varnothing\\0&\text{otherwise}\end{cases}$ \\
 \midrule
@@
 0x3b & {\small EXTCODESIZE} & 1 & 1 & Get size of an account's code. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
-\begin{cases}
-\lVert \mathbf{b} \rVert & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] \neq \varnothing \\
-0                        & \text{otherwise}
-\end{cases}$ \\
-&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \lVert \mathbf{b} \rVert$ \\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 \midrule
@@
 \begin{cases} \mathbf{b}[\boldsymbol{\mu}_{\mathbf{s}}[2] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] + i < \lVert \mathbf{b} \rVert \\ \text{\small STOP} & \text{otherwise} \end{cases}$\\
-&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}}$ \\
-&&&& We assume $\mathbf{b} \equiv ()$ if $\boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}] = \varnothing$. \\
+&&&& where $\mathtt{KEC}(\mathbf{b}) \equiv \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}}$ \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[3])$ \\
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[2] + i$ are not subject to the $2^{256}$ modulo. \\
-\end{tabu}
-
-\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv
-\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
+\begin{cases} 0 & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \\ \boldsymbol{\sigma}[\boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}]_{\mathrm{c}} & \text{otherwise} \end{cases}$ \\
 \bottomrule
@@
 0x40 & {\small BLOCKHASH} & 1 & 1 & Get the hash of one of the 256 most recent complete blocks. \\
-\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathrm{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
+\linkdest{blockhash}{}&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv P(I_{\mathbf{H}_{\mathrm{p}}}, \boldsymbol{\mu}_{\mathbf{s}}[0], 0)$ \\
 &&&& where $P$ is the hash of a block of a particular number, up to a maximum\\
@@
 0x41 & {\small COINBASE} & 0 & 1 & Get the current block's beneficiary address. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{c}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{c}}$ \\
 \midrule
 0x42 & {\small TIMESTAMP} & 0 & 1 & Get the current block's timestamp. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{s}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{s}}$ \\
 \midrule
 0x43 & {\small NUMBER} & 0 & 1 & Get the current block's number. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{i}}$ \\
 \midrule
 0x44 & {\small DIFFICULTY} & 0 & 1 & Get the current block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{d}}$ \\
 \midrule
 0x45 & {\small GASLIMIT} & 0 & 1 & Get the current block's gas limit. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{l}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{H}}_{\mathrm{l}}$ \\
 \midrule
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{\mathrm{r}}$ are specified by EIP-2200 as follows. \\
+&&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{r}$ are specified by EIP-2200 as follows. \\
 &&&& The checkpoint (``original'') state $\hyperlink{sigma_0}{\boldsymbol{\sigma}_0}$ is the state if the current transaction were to revert. \\
@@
 &&&&$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{sload}}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
-G_{\mathrm{sset}}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
-G_{\mathrm{sreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
+G_{sload}  & \text{if} \quad v = v' \; \vee \; v_0 \neq v \\
+G_{sset}   & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 = 0 \\
+G_{sreset} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v_0 \neq 0 \\
 \end{cases}$ \\
-&&&&\linkdest{A r}{} $A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \begin{cases}
-R_{\mathrm{sclear}} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
+&&&&\linkdest{A r}{} $A'_{r} \equiv A_{r} + \begin{cases}
+R_{sclear} & \text{if} \quad v \neq v' \; \wedge \; v_0 = v \; \wedge \; v' = 0 \\
 r_{\text{dirtyclear}} + r_{\text{dirtyreset}} & \text{if} \quad v \neq v' \; \wedge \; v_0 \neq v \\
@@
 &&&&$r_{\text{dirtyclear}} \equiv \begin{cases}
--R_{\mathrm{sclear}} & \text{if} \quad v_0 \neq 0 \; \wedge \; v = 0 \\
-R_{\mathrm{sclear}} & \text{if} \quad v_0 \neq 0 \; \wedge \; v' = 0 \\
+-R_{sclear} & \text{if} \quad v_0 \neq 0 \; \wedge \; v = 0 \\
+R_{sclear} & \text{if} \quad v_0 \neq 0 \; \wedge \; v' = 0 \\
 0 & \text{otherwise} \\
@@
 &&&&$r_{\text{dirtyreset}} \equiv \begin{cases}
-G_{\mathrm{sset}} - G_{\mathrm{sload}}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
-G_{\mathrm{sreset}} - G_{\mathrm{sload}} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
+G_{sset} - G_{sload}   & \text{if} \quad v_0 = v' \; \wedge \; v_0 = 0 \\
+G_{sreset} - G_{sload} & \text{if} \quad v_0 = v' \; \wedge \; v_0 \neq 0 \\
 0 & \text{otherwise} \\
@@
 0x59 & {\small MSIZE} & 0 & 1 & Get the size of active memory in bytes. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{\mathrm{i}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 32\boldsymbol{\mu}_{i}$ \\
 \midrule
@@
 &&&& for the cost of this instruction. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{\mathrm{g}}$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\mu}_{g}$ \\
 \midrule
@@
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
+&&&& is too low to fulfil the value transfer); and otherwise $x=A(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
 &&&& address of the newly created account (\ref{eq:new-address}). \\
@@
 &&&& $A' \equiv A \Cup A^+$ \\
-&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
+&&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}$ \\
 &&&& where $x=0$ if the code execution for this operation failed due to an\\
@@
 &&&& $C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv  \begin{cases}
-C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{otherwise}
@@
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{newaccount} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \mod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 \midrule
-\linkdest{REVERT}{}0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{selfdestruct}} + \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad n \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{selfdestruct} + \begin{cases}
+G_{newaccount} & \text{if} \quad n \\
 0 & \text{otherwise}
 \end{cases}$ \\
-&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \mod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
 \bottomrule
@@
 \midrule
-\linkdest{J__wordbytes}{}$J_{\mathrm{wordbytes}}$ & 4  & Bytes in word. \\
-\linkdest{J__datasetinit}{}$J_{\mathrm{datasetinit}}$ & $2^{30}$ & Bytes in dataset at genesis. \\
-\linkdest{J__datasetgrowth}{}$J_{\mathrm{datasetgrowth}}$ & $2^{23}$ & Dataset growth per epoch. \\
-\linkdest{J__cacheinit}{}$J_{\mathrm{cacheinit}}$ & $2^{24}$ & Bytes in cache at genesis. \\
-\linkdest{J__cachegrowth}{}$J_{\mathrm{cachegrowth}}$ & $2^{17}$ & Cache growth per epoch. \\
-\linkdest{J__epoch}{}$J_{\mathrm{epoch}}$ & 30000 & Blocks per epoch. \\
-\linkdest{J__mixbytes}{}$J_{\mathrm{mixbytes}}$ & 128 & mix length in bytes. \\
+\linkdest{J__wordbytes}{}$J_{wordbytes}$ & 4  & Bytes in word. \\
+\linkdest{J__datasetinit}{}$J_{datasetinit}$ & $2^{30}$ & Bytes in dataset at genesis. \\
+\linkdest{J__datasetgrowth}{}$J_{datasetgrowth}$ & $2^{23}$ & Dataset growth per epoch. \\
+\linkdest{J__cacheinit}{}$J_{cacheinit}$ & $2^{24}$ & Bytes in cache at genesis. \\
+\linkdest{J__cachegrowth}{}$J_{cachegrowth}$ & $2^{17}$ & Cache growth per epoch. \\
+\linkdest{J__epoch}{}$J_{epoch}$ & 30000 & Blocks per epoch. \\
+\linkdest{J__mixbytes}{}$J_{mixbytes}$ & 128 & mix length in bytes. \\
 \linkdest{J__hashbytes}{}$J_{\mathrm{hashbytes}}$ & 64 & Hash length in bytes. \\
 \linkdest{J__parents}{}$J_{\mathrm{parents}}$ & 256 & Number of parents of each dataset element. \\
-\linkdest{J__cacherounds}{}$J_{\mathrm{cacherounds}}$ & 3 & Number of rounds in cache production. \\
+\linkdest{J__cacherounds}{}$J_{cacherounds}$ & 3 & Number of rounds in cache production. \\
 \linkdest{J__accesses}{}$J_{\mathrm{accesses}}$ & 64 & Number of accesses in hashimoto loop. \\
@@
 \begin{equation}
- E_{\mathrm{epoch}}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{\mathrm{epoch}}}\right\rfloor
-\end{equation}
-The size of the dataset growth by $J_{\mathrm{datasetgrowth}}$ bytes, and the size of the cache by $J_{\mathrm{cachegrowth}}$ bytes, every epoch.
+ E_{epoch}(H_{\mathrm{i}}) = \left\lfloor\frac{H_{\mathrm{i}}}{J_{epoch}}\right\rfloor
+\end{equation}
+The size of the dataset growth by $J_{datasetgrowth}$ bytes, and the size of the cache by $J_{cachegrowth}$ bytes, every epoch.
 In order to avoid regularity leading to cyclic behavior, the size must be a prime number.
-Therefore the size is reduced by a multiple of $J_{\mathrm{mixbytes}}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
-\linkdest{d__size}{}Let $d_{\mathrm{size}} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
-\begin{equation}
- d_{\mathrm{size}} = E_{\mathrm{prime}}(J_{\mathrm{datasetinit}} + J_{\mathrm{datasetgrowth}} \cdot E_{\mathrm{epoch}} - J_{\mathrm{mixbytes}}, J_{\mathrm{mixbytes}})
-\end{equation}
-The size of the cache, $c_{\mathrm{size}}$, is calculated using
-\begin{equation}
- c_{\mathrm{size}} = E_{\mathrm{prime}}(J_{\mathrm{cacheinit}} + J_{\mathrm{cachegrowth}} \cdot E_{\mathrm{epoch}} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
+Therefore the size is reduced by a multiple of $J_{mixbytes}$, for the dataset, and $J_{\mathrm{hashbytes}}$ for the cache.
+\linkdest{d__size}{}Let $d_{size} = \lVert \mathbf{d} \rVert$ be the size of the dataset. Which is calculated using
+\begin{equation}
+ d_{size} = E_{\mathrm{prime}}(J_{datasetinit} + J_{datasetgrowth} \cdot E_{epoch} - J_{mixbytes}, J_{mixbytes})
+\end{equation}
+The size of the cache, $c_{size}$, is calculated using
+\begin{equation}
+ c_{size} = E_{\mathrm{prime}}(J_{cacheinit} + J_{cachegrowth} \cdot E_{epoch} - J_{\mathrm{hashbytes}}, J_{\mathrm{hashbytes}})
 \end{equation}
@@
 \subsection{Dataset generation}
-In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{\mathrm{size}}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
+In order to generate the dataset we need the cache $\mathbf{c}$, which is an array of bytes. It depends on the cache size  $c_{size}$ and the seed hash $\mathbf{s} \in \mathbb{B}_{32}$.
 \subsubsection{Seed hash}
@@
 \begin{equation}
- \mathbf{s} = C_{\mathrm{seedhash}}(H_{\mathrm{i}})
-\end{equation}
-\begin{equation}
- C_{\mathrm{seedhash}}(H_{\mathrm{i}}) = \begin{cases}
-\mathbf{0}_{32} & \text{if} \quad E_{\mathrm{epoch}}(H_{\mathrm{i}}) = 0 \quad  \\
-\texttt{KEC}(C_{\mathrm{seedhash}}(H_{\mathrm{i}} - J_{\mathrm{epoch}})) & \text{otherwise}
+ \mathbf{s} = C_{seedhash}(H_{\mathrm{i}})
+\end{equation}
+\begin{equation}
+ C_{seedhash}(H_{\mathrm{i}}) = \begin{cases}
+\mathbf{0}_{32} & \text{if} \quad E_{epoch}(H_{\mathrm{i}}) = 0 \quad  \\
+\texttt{KEC}(C_{seedhash}(H_{\mathrm{i}} - J_{epoch})) & \text{otherwise}
 \end{cases}
@@
 \subsubsection{Cache}
-The cache production process involves using the seed hash to first sequentially filling up $c_{\mathrm{size}}$ bytes of memory, then performing $J_{\mathrm{cacherounds}}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
+The cache production process involves using the seed hash to first sequentially filling up $c_{size}$ bytes of memory, then performing $J_{cacherounds}$ passes of the RandMemoHash algorithm created by \cite{lerner2014randmemohash}.
 The initial cache $\mathbf{c'}$, being an array of arrays of single bytes, will be constructed as follows.
@@
 \begin{equation}
- n = \left\lfloor\frac{c_{\mathrm{size}}}{J_{\mathrm{hashbytes}}}\right\rfloor
-\end{equation}
-The cache is calculated by performing $J_{\mathrm{cacherounds}}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
-\begin{equation}
- \mathbf{c} = E_{\mathrm{cacherounds}}(\mathbf{c'}, J_{\mathrm{cacherounds}})
-\end{equation}
-\begin{equation}
- E_{\mathrm{cacherounds}}(\mathbf{x}, y) = \begin{cases}
+ n = \left\lfloor\frac{c_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
+\end{equation}
+The cache is calculated by performing $J_{cacherounds}$ rounds of the RandMemoHash algorithm to the initial cache $\mathbf{c'}$:
+\begin{equation}
+ \mathbf{c} = E_{cacherounds}(\mathbf{c'}, J_{cacherounds})
+\end{equation}
+\begin{equation}
+ E_{cacherounds}(\mathbf{x}, y) = \begin{cases}
 \mathbf{x} & \text{if} \quad y = 0 \quad  \\
 E_\text{\tiny RMH}(\mathbf{x}) & \text{if} \quad y = 1 \quad  \\
-E_{\mathrm{cacherounds}}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
+E_{cacherounds}(E_\text{\tiny RMH}(\mathbf{x}), y -1 ) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
- E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{\mathrm{rmh}}(\mathbf{x}, 0), E_{\mathrm{rmh}}(\mathbf{x}, 1), ... , E_{\mathrm{rmh}}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
+ E_\text{\tiny RMH}(\mathbf{x}) = \big( E_{rmh}(\mathbf{x}, 0), E_{rmh}(\mathbf{x}, 1), ... , E_{rmh}(\mathbf{x}, n - 1) \big)\linkdest{E__cacherounds}{}
 \end{equation}
 \begin{multline}
-  E_{\mathrm{rmh}}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \bmod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \bmod n]) \\
-  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{\mathrm{rmh}}(\mathbf{x}, j) \quad \forall \quad j < i
+  E_{rmh}(\mathbf{x}, i) = \texttt{KEC512}(\mathbf{x'}[(i - 1 + n) \mod n] \oplus \mathbf{x'}[\mathbf{x'}[i][0] \mod n]) \\
+  \text{with} \quad \mathbf{x'} = \mathbf{x} \quad \text{except} \quad \mathbf{x'}[j] = E_{rmh}(\mathbf{x}, j) \quad \forall \quad j < i
 \end{multline}
@@
 \begin{equation}
- \mathbf{d}[i] = E_{\mathrm{datasetitem}}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{\mathrm{size}}}{J_{\mathrm{hashbytes}}}\right\rfloor
+ \mathbf{d}[i] = E_{datasetitem}(\mathbf{c}, i) \quad \forall \quad i < \left\lfloor\frac{d_{size}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
 \begin{equation}
- E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \bmod 2^{32}
+ E_\text{\tiny FNV}(\mathbf{x}, \mathbf{y}) = (\mathbf{x} \cdot (\mathrm{0x01000193} \oplus \mathbf{y})) \mod 2^{32}
 \end{equation}
@@
 \begin{equation}
- E_{\mathrm{datasetitem}}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
+ E_{datasetitem}(\mathbf{c}, i) = E_{\mathrm{parents}}(\mathbf{c}, i, -1, \varnothing)
 \end{equation}
@@
   E_{\mathrm{parents}}(\mathbf{c}, i, p, \mathbf{m}) = \begin{cases}
-E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
-E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{\mathrm{mix}}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
-\texttt{KEC512}(\mathbf{c}[i \bmod c_\mathrm{{size}}] \oplus i) & \text{if} \quad p = 0 \\
-E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \bmod \lfloor J_{\mathrm{hashbytes}} / J_{\mathrm{wordbytes}} \rfloor]) \bmod c_{\mathrm{size}}] \big) & \text{otherwise}
+E_{\mathrm{parents}}(\mathbf{c}, i, p +1, E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1)) & \text{if} \quad p < J_{\mathrm{parents}} -2 \\
+E_{mix}(\mathbf{m}, \mathbf{c}, i, p + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mix}(\mathbf{m}, \mathbf{c}, i, p) = \begin{cases}
+\texttt{KEC512}(\mathbf{c}[i \mod c_{size}] \oplus i) & \text{if} \quad p = 0 \\
+E_\text{\tiny FNV}\big(\mathbf{m}, \mathbf{c}[E_\text{\tiny FNV}(i \oplus p, \mathbf{m}[p \mod \lfloor J_{\mathrm{hashbytes}} / J_{wordbytes} \rfloor]) \mod c_{size}] \big) & \text{otherwise}
 \end{cases}
@@
 \subsection{Proof-of-work function}
-Essentially, we maintain a ``mix'' $J_{\mathrm{mixbytes}}$ bytes wide, and repeatedly sequentially fetch $J_{\mathrm{mixbytes}}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
-$J_{\mathrm{mixbytes}}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
+Essentially, we maintain a ``mix'' $J_{mixbytes}$ bytes wide, and repeatedly sequentially fetch $J_{mixbytes}$ bytes from the full dataset and use the $E_\text{\tiny FNV}$ function to combine it with the mix.
+$J_{mixbytes}$ bytes of sequential access are used so that each round of the algorithm always fetches a full page from RAM, minimizing translation lookaside buffer misses which ASICs would theoretically be able to avoid.
 
@@
 \begin{equation}
- \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{\mathrm{H}}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
+ \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d}) = \lbrace \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_{\mathrm{h}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}) + \mathbf{m}_{\mathrm{c}}(\mathtt{KEC}(\mathtt{RLP}(L_{H}(H_{\hcancel{n}}))), H_{\mathrm{n}}, \mathbf{d})) \rbrace
 \end{equation}
@@
 \begin{equation}
- \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
+ \mathbf{m}_{\mathrm{c}}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}), -1), -4)
 \end{equation}
@@
 \begin{equation}
- \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
-\end{equation}
-$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
-\begin{equation}
- E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
+ \mathbf{s}_{\mathrm{h}}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
+\end{equation}
+$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the nonce $\mathbf{n}$:
+\begin{equation}
+ E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
 \end{equation}
@@
 \begin{equation}
- n_{\mathrm{mix}} =  \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{hashbytes}}}\right\rfloor
+ n_{mix} =  \left\lfloor\frac{J_{mixbytes}}{J_{\mathrm{hashbytes}}}\right\rfloor
 \end{equation}
@@
  E_{\mathrm{accesses}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = \begin{cases}
-E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
-E_{\mathrm{accesses}}(E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-\begin{equation}
- E_{\mathrm{mixdataset}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
-\end{equation}
-$E_{\mathrm{newdata}}$ returns an array with $n_{\mathrm{mix}}$ elements:
-\begin{equation}
- E_{\mathrm{newdata}}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \bmod \left\lfloor\frac{J_{\mathrm{mixbytes}}}{J_{\mathrm{wordbytes}}}\right\rfloor]) \bmod \left\lfloor\frac{d_{\mathrm{size}} / J_{\mathrm{hashbytes}}}{n_{\mathrm{mix}}}\right\rfloor \cdot n_{\mathrm{mix}} + j] \quad \forall \quad j < n_{\mathrm{mix}}
+E_{mixdataset}(\mathbf{d}, \mathbf{m},  \mathbf{s}, i) & \text{if} \quad i = J_{\mathrm{accesses}} -2 \\
+E_{\mathrm{accesses}}(E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i), \mathbf{s}, i + 1) & \text{otherwise}
+\end{cases}
+\end{equation}
+\begin{equation}
+ E_{mixdataset}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i) = E_\text{\tiny FNV}(\mathbf{m}, E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i))
+\end{equation}
+$E_{newdata}$ returns an array with $n_{mix}$ elements:
+\begin{equation}
+ E_{newdata}(\mathbf{d}, \mathbf{m}, \mathbf{s}, i)[j] = \mathbf{d}[E_\text{\tiny FNV}(i \oplus \mathbf{s}[0], \mathbf{m}[i \mod \left\lfloor\frac{J_{mixbytes}}{J_{wordbytes}}\right\rfloor]) \mod \left\lfloor\frac{d_{size} / J_{\mathrm{hashbytes}}}{n_{mix}}\right\rfloor \cdot n_{mix} + j] \quad \forall \quad j < n_{mix}
 \end{equation}
@@
 \begin{equation}
- E_{\mathrm{compress}}(\mathbf{m}, i) = \begin{cases}
+ E_{compress}(\mathbf{m}, i) = \begin{cases}
 \mathbf{m} & \text{if} \quad i \geqslant \lVert \mathbf{m} \rVert - 8 \\
-E_{\mathrm{compress}}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
+E_{compress}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(E_\text{\tiny FNV}(\mathbf{m}[i + 4], \mathbf{m}[i + 5]), \mathbf{m}[i + 6]), \mathbf{m}[i + 7]), i + 8) & \text{otherwise}
 \end{cases}

### 2021-07-07 Paper.tex
Remove duplicated T_t
Correct merkle trie RLP details

--- 2021-06-23 Paper.tex
@@
 When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, for nodes whose composed RLP is fewer than 32 bytes, we store the RLP directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
+As a means of reducing storage complexity, we store nodes whose composed RLP is fewer than 32 bytes directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
 Thus we define in terms of $c$, the node composition function:
@@
 () & \text{if} \quad \mathfrak{I} = \varnothing \\
-c(\mathfrak{I}, i) & \text{if} \quad \lVert c(\mathfrak{I}, i)\rVert < 32 \\
-\texttt{KEC}(c(\mathfrak{I}, i)) & \text{otherwise}
+c(\mathfrak{I}, i) & \text{if} \quad \lVert \, \texttt{RLP} \big( c(\mathfrak{I}, i) \big) \rVert < 32 \\
+\texttt{KEC}\big(\texttt{RLP}( c(\mathfrak{I}, i)) \big) & \text{otherwise}
 \end{cases}

### 2021-10-10 Paper.tex
change (i, i < 8) to (i - 248, i >= 248)
EIP-3607: transactions from senders with deployed contract are forbidden (#801)
* EIP-3607: transactions from senders with deployed contract are forbidden

* 3607: update to use empty code hash instead of codesize

@@
 \Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 1 \\
-\Xi_{\mathtt{SHA256}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 2 \\
-\Xi_{\mathtt{RIP160}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 3 \\
-\Xi_{\mathtt{ID}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 4 \\
-\Xi_{\mathtt{EXPMOD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 5 \\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 6 \\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 7 \\
-\Xi_{\mathtt{SNARKV}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 8 \\
-\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{if} \quad c = 9 \\
-\Xi(\boldsymbol{\sigma}_1, g, I, \mathbf{t}) & \text{otherwise} \end{cases}
-\end{equation}
-
-\begin{equation}
-\text{Let} \; \mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
-\end{equation}
-
+\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 1 \\
+\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 2 \\
+\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 3 \\
+\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 4 \\
+\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 5 \\
+\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 6 \\
+\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 7 \\
+\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 9 \\
+\Xi                     (\boldsymbol{\sigma}_1, g, A^+, I) & \text{otherwise} \end{cases}
+\end{equation}
+\begin{equation}
+A^+ \equiv A \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{c\}
+\end{equation}
+and
+\begin{equation}
+\mathtt{KEC}(I_{\mathbf{b}}) = \boldsymbol{\sigma}[c]_{\mathrm{c}}
+\end{equation}
 It is assumed that the client will have stored the pair $(\mathtt{KEC}(I_{\mathbf{b}}), I_{\mathbf{b}})$ at some point prior in order to make the determination of $I_{\mathbf{b}}$ feasible.
@@
 0x1a & {\small BYTE} & 2 & 1 & Retrieve single byte from word. \\
-&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i < 8 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
+&&&& $\forall i \in [0..255]: \boldsymbol{\mu}'_{\mathbf{s}}[0]_{\mathrm{i}} \equiv \begin{cases} \boldsymbol{\mu}_{\mathbf{s}}[1]_{(i - 248 + 8\boldsymbol{\mu}_{\mathbf{s}}[0])} & \text{if} \quad i \geq 248 \wedge \boldsymbol{\mu}_{\mathbf{s}}[0] < 32 \\ 0 & \text{otherwise} \end{cases} $\\
 &&&& For the Nth byte, we count from the left (i.e. N=0 would be the most significant\\

### 2021-10-10 Paper.tex
EIP-3554: Difficulty Bomb Delay to December 2021

@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol, namely the \textit{Berlin} version defined by \cite{BeikoLondon}.
+This document describes one version of the protocol, namely the \textit{Berlin} version defined by \cite{BeikoBerlin}.
 In order to follow back the history of a path, one must reference multiple versions of this document.
@@
 $F_{\mathrm{Berlin}}$            & 12244000 \\
+$F_{\mathrm{London}}$            & 12965000 \\
 \bottomrule
@@
   5000000  & \text{if} \quad F_{\mathrm{Constantinople}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Muir Glacier}} \\
-  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Muir Glacier}} \\
+  9000000  & \text{if} \quad F_{\mathrm{Muir Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{London}} \\
+  9700000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{London}} \\
 \end{cases}

### 2021-10-14 Paper.tex
EIP-2929: CALL & SELFDESTRUCT
More work on EIP-2929

@@
 \begin{equation}
-A^0 \equiv (\varnothing, (), \varnothing, 0, \hyperlink{precompiled_set}{\pi}, \varnothing)
-\end{equation}
+A^0 \equiv (\varnothing, (), \varnothing, 0, \pi, \varnothing)
+\end{equation}
+where $\hyperlink{precompiled_set}{\pi}$ is the set of all precompiled addresses.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}_{\mathrm{P}}, g', A', z) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^+, S(T), S(T), g, &\\ \quad\; T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^+, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
+\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 \begin{equation}
-A^+ \equiv A^0 \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A^0_{\mathbf{a}} \cup \{S(T)\}
+A^* \equiv A^0 \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A^0_{\mathbf{a}} \cup \{S(T)\}
 \end{equation}
@@
 \begin{equation}
-A^*_{\mathrm{r}} \equiv A'_{\mathrm{r}} + \sum_{i \in A'_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
+	A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \sum_{i \in A_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
 \end{equation}
@@
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A^*_{\mathrm{r}}} \right\}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A'_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
@@
 \boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}^* \quad \text{except} \\
-\linkdest{self_destruct_list_A__s}{}\forall i \in A'_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A'_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
+\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 \Upsilon^{\mathrm{g}}(\boldsymbol{\sigma}, T) & \equiv & T_{\mathrm{g}} - g^* \\
-\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A'_{\mathbf{l}}} \\
+\Upsilon^{\mathbf{l}}(\boldsymbol{\sigma}, T) & \equiv & \hyperlink{tx_log_series_wordy_defn_A__l}{A_{\mathbf{l}}} \\
 \Upsilon^{\mathrm{z}}(\boldsymbol{\sigma}, T) & \equiv & z
@@
 
-We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued transaction substate $A$ to the tuple containing the new state, remaining gas, new accrued transaction substate and an error message $(\boldsymbol{\sigma}', g', A, \mathbf{o})$, as in section \ref{ch:transactions}:
+We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued transaction substate $A$ to the tuple containing the new state, remaining gas, new accrued transaction substate, status code and output $(\boldsymbol{\sigma}', g', A', z, \mathbf{o})$:
 \begin{equation}
@@
 \begin{equation}
-A^+ \equiv A \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{a\}
+A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{a\}
 \end{equation}
@@
 Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\hyperlink{xi_def}{\Xi}$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the new accrued substate $A'$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, A^+, I) \\
+As such, the code execution function $\hyperlink{xi_def}{\Xi}$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the resultant accrued substate $A^{**}$ and the body code of the account $\mathbf{o}$.
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, A^*, I) \\
 \end{equation}
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A, z)$ where:
+Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A', z)$ where:
 
@@
 \end{cases} \\
+\quad A' \equiv & \begin{cases}
+A^* & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
+A^{**} & \text{otherwise} \\
+\end{cases} \\
 \quad z \equiv & \begin{cases}
@@
 \end{cases} \\
-(\boldsymbol{\sigma}^{**}, g^{**}, A', \mathbf{o}) & \equiv & \Xi\\
+(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) & \equiv & \Xi\\
 I_{\mathrm{a}} & \equiv & r \\
@@
 \Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 1 \\
-\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 2 \\
-\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 3 \\
-\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 4 \\
-\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 5 \\
-\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 6 \\
-\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 7 \\
-\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 8 \\
-\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, A^+, I) & \text{if} \quad c = 9 \\
-\Xi                     (\boldsymbol{\sigma}_1, g, A^+, I) & \text{otherwise} \end{cases}
-\end{equation}
-\begin{equation}
-A^+ \equiv A \quad \text{except} \quad A^+_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{c\}
+\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 1 \\
+\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 2 \\
+\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 3 \\
+\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 4 \\
+\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 5 \\
+\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 6 \\
+\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 7 \\
+\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 9 \\
+\Xi                     (\boldsymbol{\sigma}_1, g, A, I) & \text{otherwise} \end{cases}
 \end{equation}
@@
 \begin{equation}
-C_{\mathrm{aaccess}}(x, A) \equiv
+\linkdest{C_aaccess}{}C_{\mathrm{aaccess}}(x, A) \equiv
 \begin{cases}
@@
 &&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}
-\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
 & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
+&&&& $A^* \equiv A$ except $A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{t\}$ \\
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
@@
 &&&& $C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv \begin{cases}
-\min\{ L(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)\\
+\min\{ \hyperlink{L_but_64}{L}(\boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)), \boldsymbol{\mu}_{\mathbf{s}}[0] \} & \text{if} \quad \boldsymbol{\mu}_{\mathrm{g}} \ge C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A)\\
 \boldsymbol{\mu}_{\mathbf{s}}[0] & \text{otherwise}
 \end{cases}$\\
-&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv C_{\text{\tiny ACCESS}}(\boldsymbol{\mu}, A) + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
-&&&& $C_{\text{\tiny ACCESS}}(\boldsymbol{\mu}, A) \equiv \begin{cases}
-G_{\mathrm{warmaccess}}        & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160} \in A_{\mathbf{a}} \\
-G_{\mathrm{coldaccountaccess}} & \text{otherwise}
-\end{cases}$ \\
+&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) \equiv \hyperlink{C_aaccess}{C_{\mathrm{aaccess}}}(t, A) + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\
 &&&& $C_{\text{\tiny XFER}}(\boldsymbol{\mu}) \equiv \begin{cases}
@@
 &&&& $C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if} \quad \mathtt{DEAD}(\boldsymbol{\sigma}, t) \wedge \boldsymbol{\mu}_{\mathbf{s}}[2] \neq 0 \\
 0 & \text{otherwise}
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A^+, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\
+&&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ r \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathrm{b}} = 0$ \\
-&&&&\linkdest{C tiny SELFDESTRUCT}{} $C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{selfdestruct}} + \begin{cases}
-G_{\mathrm{newaccount}} & \text{if} \quad n \\
+&&&&\linkdest{C tiny SELFDESTRUCT}{} $\begin{aligned}
+C_{\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) &\equiv G_{\mathrm{selfdestruct}} 
++ \begin{cases}
+  0 & \text{if}\ r \in \hyperlink{accessed_addresses_defn_words_A__a}{A_{\mathbf{a}}}\\
+  G_{\mathrm{coldaccountaccess}} & \text{otherwise}
+\end{cases}\\
+&+ \begin{cases}
+G_{\mathrm{newaccount}} & \text{if}\ \mathtt{DEAD}(\boldsymbol{\sigma}, r) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0 \\
 0 & \text{otherwise}
-\end{cases}$ \\
-&&&& $n \equiv \mathtt{DEAD}(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0$ \\
+\end{cases}
+\end{aligned}$ \\
 \bottomrule

### 2021-10-14 Paper.tex
Wrapping up EIP-2929

@@
 Their full formal definition is in Appendix \ref{app:precompiled}.
-
 We denote the set of the addresses of the precompiled contracts by $\pi$:
@@
 \Xi(\boldsymbol{\sigma}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A^0\!, I)\big) \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) \\
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
@@
 \begin{equation}
-Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv
+Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv
 \begin{array}[t]{l}
-\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \quad \vee \\
+\boldsymbol{\mu}_g < C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \quad \vee \\
 \mathbf{\delta}_w = \varnothing \quad \vee \\
@@
 \begin{eqnarray}
-\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \label{eq:mu_pc}\\
+\quad \boldsymbol{\mu}'_{\mathrm{g}} & \equiv & \boldsymbol{\mu}_{\mathrm{g}} - C(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \label{eq:mu_pc}\\
 \quad \boldsymbol{\mu}'_{\mathrm{pc}} & \equiv & \begin{cases}
@@
 \Xi_{\mathtt{SNARKV}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ F\\
+\qquad\Xi_{\mathtt{SNARKV}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ F\\
 F&\equiv&(\lVert I_{\mathbf{d}} \rVert\bmod 192\neq 0\vee(\exists j.\ a_{\mathrm{j}}=\varnothing\vee b_{\mathrm{j}}=\varnothing))\\
@@
 \Xi_{\mathtt{BN\_ADD}}&\equiv&\Xi_{\mathtt{BN\_PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
+\Xi_{\mathtt{BN\_ADD}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ x=\varnothing\vee y=\varnothing\\
 g_{\mathrm{r}} &=& 150\\
@@
 \Xi_{\mathtt{BN\_MUL}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ x=\varnothing\\
+\Xi_{\mathtt{BN\_MUL}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ x=\varnothing\\
 g_{\mathrm{r}} &=& 6000\\
@@
   \Xi_{\mathtt{BLAKE2\_F}}&\equiv&\Xi_{\mathtt{PRE}}\quad\text{except:}\\
-  \Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol\sigma,g,I)&=&\left(\varnothing,0,A^0,()\right)\quad\text{if}\ \lVert I_{\mathbf{d}}\rVert \neq 213 \vee f \notin \{0, 1\} \\
+  \Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol\sigma,g,A,I)&=&\left(\varnothing,0,A,()\right)\quad\text{if}\ \lVert I_{\mathbf{d}}\rVert \neq 213 \vee f \notin \{0, 1\} \\
   g_{\mathrm{r}} &=& r\\

### 2021-10-14 Paper.tex
Cite EIP-3607
EIP-2929 finishing touches

@@
 \item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
-\item the sender account has no contract code deployed ($\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} = \texttt{KEC}\big( () \big)$).
+\item the sender account has no contract code deployed (see EIP-3607 by \cite{EIP-3607});
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
@@
 
-We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued transaction substate $A$ to the tuple containing the new state, remaining gas, new accrued transaction substate, status code and output $(\boldsymbol{\sigma}', g', A', z, \mathbf{o})$:
+We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued substate $A$, to the tuple containing the new state, remaining gas, new accrued substate, status code and output $(\boldsymbol{\sigma}', g', A', z, \mathbf{o})$:
 \begin{equation}
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& $A^* \equiv A$ except $A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{t\}$ \\
+&&&& $A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{t\}$ \\
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, A, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, \varnothing, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, A, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
-(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\ r \neq I_{\mathrm{a}} \\
+\varnothing &\text{if}\quad \boldsymbol{\sigma}[r] = \varnothing\ \wedge\ \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} = 0\\
+(\boldsymbol{\sigma}[r]_{\mathrm{n}}, \boldsymbol{\sigma}[r]_{\mathrm{b}} + \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{if}\quad r \neq I_{\mathrm{a}} \\
 (\boldsymbol{\sigma}[r]_{\mathrm{n}}, 0, \boldsymbol{\sigma}[r]_{\mathbf{s}}, \boldsymbol{\sigma}[r]_{\mathrm{c}}) & \text{otherwise}
@@
 + \begin{cases}
-  0 & \text{if}\ r \in \hyperlink{accessed_addresses_defn_words_A__a}{A_{\mathbf{a}}}\\
+  0 & \text{if}\quad r \in \hyperlink{accessed_addresses_defn_words_A__a}{A_{\mathbf{a}}}\\
   G_{\mathrm{coldaccountaccess}} & \text{otherwise}
@@
 &+ \begin{cases}
-G_{\mathrm{newaccount}} & \text{if}\ \mathtt{DEAD}(\boldsymbol{\sigma}, r) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0 \\
+G_{\mathrm{newaccount}} & \text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}, r) \wedge \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \neq 0 \\
 0 & \text{otherwise}

### 2021-11-01 Paper.tex
change m(x,i) to be the complement of 2047
Amendment: transaction sender might not exists

@@
 S(T) & \neq & \varnothing \quad \wedge \\
-\boldsymbol{\sigma}[S(T)] & \neq & \varnothing \quad \wedge \\
 \boldsymbol{\sigma}[S(T)]_{\mathrm{c}} & = & \texttt{KEC}\big( () \big) \quad \wedge \\
@@
 \end{equation}
-
 Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
+Also, with a slight abuse of notation, we assume that $\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} = \texttt{KEC}\big( () \big)$, $\boldsymbol{\sigma}[S(T)]_{\mathrm{n}} = 0$, and $\boldsymbol{\sigma}[S(T)]_{\mathrm{b}} = 0$ if $\boldsymbol{\sigma}[S(T)] = \varnothing$.
 

### 2021-11-02 Paper.tex
add specification about $\mathcal{B}$
Fix bit-selection in bloom filter function (#820)
* change m(x,i) to be the complement of 2047

* add specification about $\mathcal{B}$
Fix rendering of Bloom filter formula (#823)

@@
 \end{equation}
-
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.

### 2021-11-02 Paper.tex
Arrow Glacier network upgrade

@@
 $F_{\mathrm{London}}$            & 12965000 \\
+$F_{\mathrm{Arrow Glacier}}$     & 13773000 \\
 \bottomrule
@@
   9000000  & \text{if} \quad F_{\mathrm{Muir Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{London}} \\
-  9700000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{London}} \\
+  9700000  & \text{if} \quad F_{\mathrm{London}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Arrow Glacier}} \\
+ 10700000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Arrow Glacier}} \\
 \end{cases}
@@
 which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-Subsequently, with EIP-1234 by \cite{EIP-1234} and EIP-2384 by \cite{EIP-2384} the subtrahend $\kappa$ was increased to five and nine million respectively.
+Subsequently, EIP-1234 by \cite{EIP-1234}, EIP-2384 by \cite{EIP-2384}, EIP-3554 by \cite{EIP-3554}, and EIP-4345 by \cite{EIP-4345} increased the subtrahend $\kappa$.
 

### 2021-11-26 Paper.tex
Replace SHA3 opcode with KECCAK256 (#825)

@@
 $G_{\mathrm{logtopic}}$ & 375 & Paid for each topic of a {\small LOG} operation. \\
-$G_{\mathrm{sha3}}$ & 30 & Paid for each {\small SHA3} operation. \\
-$G_{\mathrm{sha3word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
+$G_{\mathrm{keccak256}}$ & 30 & Paid for each {\small KECCAK256} operation. \\
+$G_{\mathrm{keccak256word}}$ & 6 & Paid for each word (rounded up) for input data to a {\small KECCAK256} operation. \\
 $G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
@@
 G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
-G_{\mathrm{create}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
-G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
+G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
 G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
@@
 \toprule
-\multicolumn{5}{c}{\textbf{20s: SHA3}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{20s: KECCAK256}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x20 & {\small SHA3} & 2 & 1 & Compute Keccak-256 hash. \\
+0x20 & {\small KECCAK256} & 2 & 1 & Compute Keccak-256 hash. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \mathtt{KEC}(\boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots (\boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1) ])$ \\

### 2021-11-26 Paper.tex
Use keccak-256 consistently

@@
 
-The address of the new account is defined as being the rightmost 160 bits of the Keccak hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
+The address of the new account is defined as being the rightmost 160 bits of the Keccak-256 hash of the \hyperlink{rlp}{RLP} encoding of the structure containing only the sender and the \hyperlink{account_nonce}{account nonce}.
 For {\small \hyperlink{create2}{CREATE2}} the rule is different and is described in EIP-1014 by \cite{EIP-1014}.
@@
 When composing a node, we use RLP to encode the structure.
-As a means of reducing storage complexity, we store nodes whose composed RLP is fewer than 32 bytes directly; for those larger we assert prescience of the byte array whose Keccak hash evaluates to our reference.
+As a means of reducing storage complexity, we store nodes whose composed RLP is fewer than 32 bytes directly; for those larger we assert prescience of the byte array whose Keccak-256 hash evaluates to our reference.
 Thus we define in terms of $c$, the node composition function:
@@
 
-For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak hash of the corresponding ECDSA public key:
+For a given private key, $p_{\mathrm{r}}$, the Ethereum address $A(p_{\mathrm{r}})$ (a 160-bit value) to which it corresponds is defined as the rightmost 160-bits of the Keccak-256 hash of the corresponding ECDSA public key:
 \begin{equation}
@@
 
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak hash of the transaction.
+\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction.
 Two different flavors of signing schemes are available.
@@
 The sequences of both ommers and transactions are empty and represented by $()$.
-$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
+$\mathtt{KEC}\big( (42) \big)$ refers to the Keccak-256 hash of a byte array of length one whose first and only byte is of value 42, used for the nonce.
 $\mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big)$ value refers to the hash of the ommer list in RLP, both empty lists.

### 2021-11-26 Paper.tex
Implement EIP-2681: Limit account nonce to 2^64-1

--- 2021-11-26 Paper.tex
@@
 \begin{description}
-\item[nonce] \linkdest{account_nonce}A scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
+\item[nonce] \linkdest{account_nonce}A 64-bit scalar value equal to the number of transactions sent from this address or, in the case of accounts with associated code, the number of contract-creations made by this account.
 For account of address $a$ in state $\boldsymbol{\sigma}$, this would be formally denoted $\boldsymbol{\sigma}[a]_{\mathrm{n}}$.
@@
 \begin{description}
-\item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
+\item[nonce]\linkdest{tx_nonce}{} A 64-bit scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
 \item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
@@
 \item the transaction signature is valid;
+\item the \hyperlink{transaction_nonce}{transaction nonce} does not equal or exceed $2^64-1$ (see EIP-2681 by \cite{EIP-2681});
 \item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});

### 2021-12-01 Paper.tex
corrected the definition of N_n so that it correctly reflects byte representation
Genesis block difficulty is 2^34, not 2^17

--- 2021-11-26 Paper.tex
@@
 
-\newcommand{\mindifficulty}{D_0}
+\newcommand{\mindifficulty}{D_\mathrm{min}}
 \newcommand{\homesteadmod}{\ensuremath{\varsigma_2}}
@@
 D(H) \equiv \begin{dcases}
-\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
+2^{34} & \text{if} \quad H_{\mathrm{i}} = 0\\
 \text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
@@
 \begin{equation}
-\mindifficulty \equiv 131072
+\mindifficulty \equiv 2^{17}
 \end{equation}
@@
 \begin{equation}
-\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{17}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
-\end{equation}
-
-Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{17}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
+\big( \big( 0_{256}, \mathtt{KEC}\big(\mathtt{RLP}\big( () \big)\big), 0_{160}, stateRoot, 0, 0, 0_{2048}, 2^{34}, 0, 0, 3141592, time, 0, 0_{256},  \mathtt{KEC}\big( (42) \big) \big), (), () \big)
+\end{equation}
+
+Where $0_{256}$ refers to the parent hash, a 256-bit hash which is all zeroes; $0_{160}$ refers to the beneficiary address, a 160-bit hash which is all zeroes; $0_{2048}$ refers to the log bloom, 2048-bit of all zeros; $2^{34}$ refers to the difficulty; the transaction trie root, receipt trie root, gas used, block number and extradata are both $0$, being equivalent to the empty byte array.
 The sequences of both ommers and transactions are empty and represented by $()$.

### 2021-12-01 Paper.tex
mindifficulty != the difficulty of the genesis block
Improve formula alignment

@@
 where:
-\begin{equation}
-\mindifficulty \equiv 2^{17}
-\end{equation}
-\begin{equation}
-\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
-\end{equation}
-\begin{equation}
-\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
-\end{equation}
-\begin{equation*}
-y \equiv \begin{cases}
+\begin{align}
+\mindifficulty &\equiv 2^{17} \\
+\diffadjustment &\equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor \\
+\homesteadmod &\equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right) \\
+y &\equiv \begin{cases}
 1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
 2 & \text{otherwise}
-\end{cases}
-\end{equation*}
-\begin{align}
+\end{cases} \\
 \expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0) \\
-\kappa &\equiv \begin{cases} 
+H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0)
+\end{align}
+\begin{equation}
+\kappa \equiv \begin{cases} 
   3000000  & \text{if} \quad F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
@@
 \end{cases}
-\end{align}
-
-Note that $\mindifficulty$ is the difficulty of the genesis block.
+\end{equation}
+
 The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.

### 2021-12-06 Paper.tex
Genesis block difficulty is 2^34, not 2^17 (#830)
* Genesis block difficulty is 2^34, not 2^17

* mindifficulty != the difficulty of the genesis block

* Improve formula alignment

* tiny formatting improvement

@@
 y &\equiv \begin{cases}
-1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
+1 & \text{if} \quad \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
 2 & \text{otherwise}

### 2022-01-09 Paper.tex
Update Rsclear from 15000 to 4800
EIP-2718: Typed Transaction Envelope

--- 2021-12-01 Paper.tex
@@
 A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum.
-The sender of a transaction can not be a contract.
-While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
+The sender of a transaction cannot be a contract.
+While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably,
+such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents.
 \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}.
-There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
-Both types specify a number of common fields:
+EIP-2718 by \cite{EIP-2718} introduced the notion of different transaction types.
+As of the Berlin version of the protocol, there are two transaction types: 0 (legacy) and 1 (EIP-2930 by \cite{EIP-2930}).
+Further, there are two subtypes of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
+All transaction types specify a number of common fields:
 
 \begin{description}
+\item[type]\linkdest{tx_type}{} EIP-2718 transaction type; formally $T_{\mathrm{x}}$.
 \item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
@@
 \item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[v, r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally \linkdest{T__w_T__r_T__s}{$T_{\mathrm{w}}$, $T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.
+\end{description}
+
+EIP-2930 (type 1) transactions also have:
+\begin{description}
+\item[accessList]\linkdest{tx_access_list}{} TODO; formally $T_{\mathbf{a}}$.
+\item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
+\item[yParity]\linkdest{tx_y_parity}{} Signature Y parity; formally $T_{\mathrm{y}}$.
+\end{description}
+
+Legacy transactions do not have an \textbf{accessList}, while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
+\begin{description}
+\item[w] A scalar value encoding Y parity and possibly chain ID; \linkdest{T__w_T__r_T__s}formally $T_{\mathrm{w}}$.
+$T_{\mathrm{w}} = 27 + T_{\mathrm{y}}$ or $T_{\mathrm{w}} = 2\beta + 35 + T_{\mathrm{y}}$ (see EIP-155 by \cite{EIP-155}).
 \end{description}
@@
 \begin{equation}
-L_{\mathrm{T}}(T) \equiv \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+\linkdest{L_transaction} L_{\mathrm{T}}(T) \equiv \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 0 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{a}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
+\end{cases}
+\end{equation}
+where
+\begin{equation}
+\mathbf{p} \equiv \begin{cases}
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
+T_{\mathbf{d}} & \text{otherwise}
+\end{cases}
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{a}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
 \begin{equation}
 \begin{array}[t]{lclclc}
-T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{x}} \in \{0, 1\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
 \end{array}
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{\mathrm{T}}(B_{\mathbf{T}}[i]))\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p'(i, B_{\mathbf{T}}[i])\}) & \wedge \\
 \linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
@@
 \end{equation}
+and $p'(k, T)$ is similar, but has a special treatment for EIP-2718 transactions:
+\begin{equation}
+p'(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
+\mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
+T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+\end{cases}
+\right)
+\end{equation}
+($\cdot$ is the concatenation of byte arrays).
 
@@
 \quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
-\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
-\end{eqnarray}
-
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}With $L_T^*$ and $L_H^*$ being element-wise sequence transformations, thus:
-\begin{equation}
-\hyperlink{general_element_wise_sequence_transformation_f_pow_asterisk}{f^*}\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\end{eqnarray}
+where $\widetilde{L}_{\mathrm{T}}$ takes a special care of EIP-2718 transactions:
+\begin{equation}
+\widetilde{L}_{\mathrm{T}}(T) = \begin{cases}
+\hyperlink{L_transaction}{L_{\mathrm{T}}}(T) & \text{if} \quad T_{\mathrm{x}} = 0 \\
+T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+\end{cases}
+\end{equation}
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
+\begin{equation}
+f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
 \end{equation}
@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-{} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-{} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
+\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
 \end{align}
-
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
@@
 
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction.
-Two different flavors of signing schemes are available.
-One operates without the latter three signature components, formally described as $T_{\mathrm{r}}$, $T_{\mathrm{s}}$ and $T_{\mathrm{w}}$.
-The other operates on nine elements:
-\begin{eqnarray}
-L_{\mathrm{S}}(T) & \equiv & \begin{cases}
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; v \in \{27, 28\} \\
-(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{otherwise} \\
+\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction. Three different flavours of signing schemes are available:
+\begin{eqnarray}
+L_{\mathrm{X}}(T) & \equiv & \begin{cases}
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{27, 28\} \\
+(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{a}}) & \text{if} \; T_{\mathrm{x}} = 1
 \end{cases} \\
@@
 \nonumber \mathbf{p} & \equiv & \begin{cases}
-T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = 0 \\
+T_{\mathbf{i}} & \text{if}\ T_{\mathrm{t}} = \varnothing \\
 T_{\mathbf{d}} & \text{otherwise}
 \end{cases} \\
-h(T) & \equiv & \mathtt{KEC}( L_{\mathrm{S}}(T) )
+h(T) & \equiv & \begin{cases}
+\mathtt{KEC}( \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{if} \; T_{\mathrm{x}} = 0 \\
+\mathtt{KEC}( T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{X}}(T)) ) & \text{otherwise}
+\end{cases}
 \end{eqnarray}
@@
 G(T, p_{\mathrm{r}}) \equiv T \quad \text{except:} \\
-(T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
+(T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) = \mathtt{ECDSASIGN}(h(T), p_{\mathrm{r}})
 \end{eqnarray}
@@
 \end{eqnarray}
-$\linkdest{T__w}{T_{\mathrm{w}}}$ is either the recovery identifier or `chain identifier \hyperlink{chain_id}{$\beta$} doubled plus 35 or 36'.
-In the second case, where \hypertarget{v}{}$v$ is the chain identifier $\beta$ doubled plus 35 or 36, the values 35 and 36 assume the role of the `recovery identifier' by specifying the parity of $y$, with the value 35 representing an even value and 36 representing an odd value.
+and $\linkdest{T__w}{T_{\mathrm{w}}}$ of legacy transcations is either $27 + T_{\mathrm{y}}$ or $2\hyperlink{chain_id}{\beta} + 35 + T_{\mathrm{y}}$.
 
@@
 \begin{eqnarray}
-S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v_0, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
-v_0 &\equiv& \begin{cases}
-T_{\mathrm{w}} &\text{if}\ T_{\mathrm{w}}\in\{27, 28\} \\
-28 - (T_{\mathrm{w}} \bmod 2) &\text{otherwise}
+S(T) &\equiv& \mathcal{B}_{96..255}\big(\mathtt{KEC}\big( \mathtt{ECDSARECOVER}(h(T), v, T_{\mathrm{r}}, T_{\mathrm{s}}) \big) \big) \\
+v &\equiv& \begin{cases}
+T_{\mathrm{w}} - 27 & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{27, 28\} \\
+(T_{\mathrm{w}}-35) \bmod 2 & \text{if} \ T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
+T_{\mathrm{y}} & \text{if} \ T_{\mathrm{x}} = 1
 \end{cases}

### 2022-01-16 Paper.tex
Add access lists costs to the fee schedule
EIP-2718 receipts

@@
 
-\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of four items comprising:
+\linkdest{transaction_receipt_R}{}\linkdest{tx_receipt_gas_used_R__u}{}\linkdest{R__u}The transaction receipt, $R$, is a tuple of five items comprising:
+the type of the transaction, $R_{\mathrm{x}}$,
 the status code of the transaction, $R_{\mathrm{z}}$,
@@
 \begin{equation}
-R \equiv (R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+R \equiv (R_{\mathrm{x}}, R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
+\end{equation}
+
+$R_{\mathrm{x}}$ is equal to the \hyperlink{tx_type}{type} of the corresponding transaction.
+
+\linkdest{L__R}The function $L_{\mathrm{R}}$ prepares a transaction receipt for being transformed into an RLP-serialised byte array:
+\begin{equation}
+L_{\mathrm{R}}(R) \equiv (R_{\mathrm{z}}, R_{\mathrm{u}}, R_{\mathrm{b}}, R_{\mathbf{l}})
 \end{equation}
@@
 \linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
-\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p'(i, B_{\mathbf{T}}[i])\}) & \wedge \\
-\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
+\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
+\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \end{equation}
-where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
-\begin{equation}
-p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
-\end{equation}
-and $p'(k, T)$ is similar, but has a special treatment for EIP-2718 transactions:
-\begin{equation}
-p'(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
-\mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
+where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
+\begin{equation}
+p_{\mathrm{T}}(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
+\mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
 T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+\end{cases}
+\right)
+\end{equation}
+and
+\begin{equation}
+p_{\mathrm{R}}(k, R) \equiv \left( \mathtt{RLP}(k), \begin{cases}
+\mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
+R_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
 \end{cases}
@@
 $G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
+$G_{\mathrm{accesslistaddress}}$ & 2400 & Cost of warming up an account with the access list. \\
+$G_{\mathrm{accessliststorage}}$ & 1900 & Cost of warming up a storage with the access list. \\
 $G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
@@
 $G_{\mathrm{copy}}$ & 3 & Partial payment for {\small *COPY} operations, multiplied by words copied, rounded up. \\
-$G_{\mathrm{blockhash}}$ & 20 & Payment for {\small BLOCKHASH} operation. \\
+$G_{\mathrm{blockhash}}$ & 20 & Payment for each {\small BLOCKHASH} operation. \\
 

### 2022-01-16 Paper.tex
Improve hyperlinks

@@
 \item[value]\linkdest{tx_value_T__v}{} A scalar value equal to the number of Wei to be transferred to the message call's recipient or, in the case of contract creation, as an endowment to the newly created account; formally $T_{\mathrm{v}}$.
-\item[r, s] Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
+\item[r, s]\linkdest{T__r_T__s}{} Values corresponding to the signature of the transaction and used to determine the sender of the transaction; formally {$T_{\mathrm{r}}$ and $T_{\mathrm{s}}$}.
 This is expanded in Appendix \ref{app:signing}.
@@
 \begin{description}
-\item[w] A scalar value encoding Y parity and possibly chain ID; \linkdest{T__w_T__r_T__s}formally $T_{\mathrm{w}}$.
+\item[w]\linkdest{T__w}{} A scalar value encoding Y parity and possibly chain ID; formally $T_{\mathrm{w}}$.
 $T_{\mathrm{w}} = 27 + T_{\mathrm{y}}$ or $T_{\mathrm{w}} = 2\beta + 35 + T_{\mathrm{y}}$ (see EIP-155 by \cite{EIP-155}).
@@
 
-\hyperlink{T__w_T__r_T__s}{Reiterating from previously}:
+\hyperlink{T__r_T__s}{Reiterating from previously}:
 \begin{eqnarray}
@@
 \end{eqnarray}
-and $\linkdest{T__w}{T_{\mathrm{w}}}$ of legacy transcations is either $27 + T_{\mathrm{y}}$ or $2\hyperlink{chain_id}{\beta} + 35 + T_{\mathrm{y}}$.
+and $\hyperlink{T__w}{T_{\mathrm{w}}}$ of legacy transcations is either $27 + T_{\mathrm{y}}$ or $2\hyperlink{chain_id}{\beta} + 35 + T_{\mathrm{y}}$.
 

### 2022-01-16 Paper.tex
EIP-2930: Optional access lists

@@
 \begin{description}
-\item[accessList]\linkdest{tx_access_list}{} TODO; formally $T_{\mathbf{a}}$.
+\item[accessList]\linkdest{tx_access_list}{} List of access entries to warm up; formally $T_{\mathbf{A}}$.
+Each access list entry $A$ is a tuple of an account address and a list of storage keys: $A \equiv (A_{\mathrm{a}}, A_{\mathbf{s}})$. 
 \item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
@@
 
-Legacy transactions do not have an \textbf{accessList}, while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
+Legacy transactions do not have an \textbf{accessList} ($T_{\mathbf{A}}=()$), while \textbf{chainId} and \textbf{yParity} for legacy transactions are combined into a single value:
 \begin{description}
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 0 \\
-(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{a}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
 \end{cases}
@@
 
-Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{a}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
+Here, we assume all components are interpreted by the RLP as integer values, with the exception of the access list $T_{\mathbf{A}}$ and the arbitrary length byte arrays $T_{\mathbf{i}}$ and $T_{\mathbf{d}}$.
 \begin{equation}
@@
 \begin{equation}
-A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}}, A_{\mathbf{a}}, A_{\mathbf{k}})
+A \equiv (A_{\mathbf{s}}, A_{\mathbf{l}}, A_{\mathbf{t}}, A_{\mathrm{r}}, A_{\mathbf{a}}, A_{\mathbf{K}})
 \end{equation}
@@
 Though not immediately refunded, it is allowed to partially offset the total execution costs.
-Finally, EIP-2929 by \cite{EIP-2929} introduced \hypertarget{accessed_addresses_defn_words_A__a}{}$A_{\mathbf{a}}$, the set of accessed account addresses, and \hypertarget{accessed_storage_keys_defn_words_A__k}{}$A_{\mathbf{k}}$, the set of accessed storage keys.
+Finally, EIP-2929 by \cite{EIP-2929} introduced \hypertarget{accessed_addresses_defn_words_A__a}{}$A_{\mathbf{a}}$, the set of accessed account addresses, and \hypertarget{accessed_storage_keys_defn_words_A__k}{}$A_{\mathbf{K}}$, the set of accessed storage keys
+(more accurately, each element of $A_{\mathbf{K}}$ is a tuple of a 20-byte account address and a 32-byte storage slot).
 
@@
 \nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
-\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}}
+\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
+\nonumber {} & + \sum_{j = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \big( G_{\mathrm{accesslistaddress}} + \lVert T_{\mathbf{A}}[j]_{\mathbf{s}} \rVert G_{\mathrm{accessliststorage}} \big)
 \end{align}
@@
 $G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
 $G$ is fully defined in Appendix \ref{app:fees}.
@@
 where
-\begin{equation}
-A^* \equiv A^0 \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A^0_{\mathbf{a}} \cup \{S(T)\}
-\end{equation}
+\begin{align}
+A^* & \equiv A^0 \quad \text{except} \\
+A^*_{\mathbf{a}} & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{ T_{\mathbf{A}}[i]_{\mathrm{a}}, \; i < \lVert T_{\mathbf{A}} \rVert \} \\
+A^*_{\mathbf{K}} & \equiv A^0_{\mathbf{K}} \bigcup_{i = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \{ (T_{\mathbf{A}}[i]_{\mathrm{a}}, T_{\mathbf{A}}[i]_{\mathbf{s}}[j]), \; j < \lVert T_{\mathbf{A}}[i]_{\mathbf{s}} \rVert \}
+\end{align}
 and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
-(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{a}}) & \text{if} \; T_{\mathrm{x}} = 1
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 1
 \end{cases} \\
@@
 $G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
-$G_{\mathrm{accesslistaddress}}$ & 2400 & Cost of warming up an account with the access list. \\
-$G_{\mathrm{accessliststorage}}$ & 1900 & Cost of warming up a storage with the access list. \\
+\linkdest{G__accesslistaddress}{}$G_{\mathrm{accesslistaddress}}$ & 2400 & Cost of warming up an account with the access list. \\
+\linkdest{G__accessliststorage}{}$G_{\mathrm{accessliststorage}}$ & 1900 & Cost of warming up a storage with the access list. \\
 $G_{\mathrm{coldaccountaccess}}$ & 2600 & Cost of a cold account access. \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
-&&&& $A'_{\mathbf{k}} \equiv A_{\mathbf{k}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
+&&&& $A'_{\mathbf{K}} \equiv A_{\mathbf{K}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
 &&&& $C_{\text{\tiny SLOAD}}(\boldsymbol{\mu}, A, I) \equiv
 \begin{cases}
-G_{\mathrm{warmaccess}} & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{k}} \\
+G_{\mathrm{warmaccess}} & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{K}} \\
 G_{\mathrm{coldsload}}  & \text{otherwise}
@@
 &&&& $\boldsymbol{\sigma}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
-&&&& $A'_{\mathbf{k}} \equiv A_{\mathbf{k}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
+&&&& $A'_{\mathbf{K}} \equiv A_{\mathbf{K}} \cup \{(I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0])\}$ \\
 &&&&\linkdest{C__SSTORE}{}$C_{\text{\tiny SSTORE}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$ and \linkdest{A r}{}$A'_{\mathrm{r}}$ are specified by EIP-2200 as follows. \\
@@
 \begin{cases}
-0                      & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{k}} \\
+0                      & \text{if} \quad (I_{\mathrm{a}}, \boldsymbol{\mu}_{\mathbf{s}}[0]) \in A_{\mathbf{K}} \\
 G_{\mathrm{coldsload}} & \text{otherwise} \\

### 2022-01-17 Paper.tex
Small improvements
Small correction
Small clarification
More consistent set notation

@@
 \item[accessList]\linkdest{tx_access_list}{} List of access entries to warm up; formally $T_{\mathbf{A}}$.
-Each access list entry $A$ is a tuple of an account address and a list of storage keys: $A \equiv (A_{\mathrm{a}}, A_{\mathbf{s}})$. 
+\linkdest{access_list_entry}{}Each access list entry $E$ is a tuple of an account address and a list of storage keys: $E \equiv (E_{\mathrm{a}}, E_{\mathbf{s}})$. 
 \item[chainId]\linkdest{tx_chain_id}{} Chain ID; formally $T_{\mathrm{c}}$. Must be equal to the network chain ID \hyperlink{chain_id}{$\beta$}.
@@
 
-Additionally, a contract creation transaction contains:
+Additionally, a contract creation transaction (regardless whether legacy or EIP-2930) contains:
 
@@
 \mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
-T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
 \end{cases}
@@
 \mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
-R_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
+(R_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
 \end{cases}
@@
 \hyperlink{L_transaction}{L_{\mathrm{T}}}(T) & \text{if} \quad T_{\mathrm{x}} = 0 \\
-T_{\mathrm{x}} \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
+(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
 \end{cases}
@@
 A^* & \equiv A^0 \quad \text{except} \\
-A^*_{\mathbf{a}} & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{ T_{\mathbf{A}}[i]_{\mathrm{a}}, \; i < \lVert T_{\mathbf{A}} \rVert \} \\
-A^*_{\mathbf{K}} & \equiv A^0_{\mathbf{K}} \bigcup_{i = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \{ (T_{\mathbf{A}}[i]_{\mathrm{a}}, T_{\mathbf{A}}[i]_{\mathbf{s}}[j]), \; j < \lVert T_{\mathbf{A}}[i]_{\mathbf{s}} \rVert \}
+A^*_{\mathbf{a}} & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \} \\
+A^*_{\mathbf{K}} & \equiv \bigcup_{E \in T_{\mathbf{A}}} \big\{ \forall i < \lVert E_{\mathbf{s}} \rVert, i \in \mathbb{N}: \; (E_{\mathrm{a}}, E_{\mathbf{s}}[i]) \big\}
 \end{align}
@@
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, \hyperlink{refund_balance_defn_words_A__r}{A'_{\mathrm{r}}} \right\}
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A'_{\mathrm{r}} \right\}
 \end{equation}
@@
 Their full formal definition is in Appendix \ref{app:precompiled}.
-We denote the set of the addresses of the precompiled contracts by $\pi$:
-\begin{equation}\hypertarget{precompiled_set}{}
+\hypertarget{precompiled_set}{}We denote the set of the addresses of the precompiled contracts by $\pi$:
+\begin{equation}
 \pi \equiv \{1, 2, 3, 4, 5, 6, 7, 8, 9 \}

### 2022-02-12 Paper.tex
9.4.3. Jump Destination Validity - treat code as reachable by PC

@@
 
-We previously used $D$ as the function to determine the set of valid jump destinations given the code that is being run.
 We define this as any position in the code occupied by a {\small JUMPDEST} instruction.
 
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).
-
-Formally:
-\begin{equation}
-D(\mathbf{c}) \equiv D_{\mathrm{J}}(\mathbf{c}, 0)
-\end{equation}
-
-where:
-\begin{equation}
-D_{\mathrm{J}}(\mathbf{c}, i) \equiv \begin{cases}
-\{\} & \text{if} \quad i \geqslant \lVert \mathbf{c} \rVert  \\
-\{ i \} \cup D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \\
-\quad\quad\text{if} \quad \mathbf{c}[i] = \text{\small JUMPDEST} \\
-D_{\mathrm{J}}(\mathbf{c}, N(i, \mathbf{c}[i])) & \text{otherwise} \\
-\end{cases}
-\end{equation}
-
-where $N$ is the next valid instruction position in the code, skipping the data of a {\small PUSH} instruction, if any:
-\begin{equation}
-N(i, w) \equiv \begin{cases}
-i + w - \text{\small PUSH1} + 2 & \\
-\quad\quad\text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
-i + 1 & \text{otherwise} \end{cases}
-\end{equation}
+All such positions must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).
 

### 2022-02-12 Paper.tex
9.4.3. Jump Destination Validity - be explicit about validation
The sentence "must appear within the explicitly defined portion of the code"
leads the reader to think that only executable code is checked for JUMPDEST validity.
In practice, all clients validate jump destinations prior to execution.

--- 2022-01-17 Paper.tex
@@
 
-All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it).
+All such positions must be on valid instruction boundaries, rather than sitting in the data portion of {\small PUSH} operations and must appear within the explicitly defined portion of the code (rather than in the implicitly defined \hyperlink{stop}{{\small STOP}} operations that trail it), where code denotes any executable code or data (non-reachable by Program Counter) between {\small PUSH} operations.
+Jump destinations must be validated prior to execution.
 

### 2022-02-13 Paper.tex
improved consistency of the definitions of `CALL` and `CREATE` instructions (#841)

--- 2022-01-17 Paper.tex
@@
 \end{align}
+\hypertarget{contract_creation_result}{}
 
@@
 \end{cases} \\
+\hypertarget{code_execution_result}{}
 (\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) & \equiv & \Xi\\
@@
 &&&& $\hyperlink{salt}{\zeta} \equiv \varnothing$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \equiv \begin{cases}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A', z, \mathbf{o}) \equiv \begin{cases}
 \hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\
-\big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, A, () \big) & \text{otherwise} \end{cases}$ \\
+\big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, A, 0, () \big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$, or $I_{\mathrm{e}} = 1024$ \\
+&&&& where $x=0$ if $z = 0$, i.e., the \hyperlink{contract_creation_result}{contract creation process failed}, or $I_{\mathrm{e}} = 1024$ \\
 &&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}
 \begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
 & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
-(\boldsymbol{\sigma}, g, A, ()) & \text{otherwise} \end{cases}$ \\
+(\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
-&&&& where $x=0$ if the code execution for this operation failed due to an\\
-&&&& \hyperlink{Exceptional_Halting_function_Z}{exceptional halting} (or for a \text{\small REVERT}) $\boldsymbol{\sigma}' = \varnothing$ or if \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
+&&&& where $x=0$ if the \hyperlink{code_execution_result}{code execution for this operation failed}, or if \\
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[2] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (not enough funds) or $I_{\mathrm{e}} = 1024$ (call depth limit reached); $x=1$ \\
 &&&& otherwise. \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[3], \boldsymbol{\mu}_{\mathbf{s}}[4]), \boldsymbol{\mu}_{\mathbf{s}}[5], \boldsymbol{\mu}_{\mathbf{s}}[6])$ \\
 &&&& Thus the operand order is: gas, to, value, in offset, in size, out offset, out size. \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, A, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, A, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2022-02-21 Paper.tex
Fix gas specification in CALL* operations
Fix the upsidedown multiplier in the bloom filter footnote

@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{2048 $= 2^{11}$(11 bits), and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}
@@
 &&&& $\hyperlink{salt}{\zeta} \equiv \varnothing$ \\
-&&&& $(\boldsymbol{\sigma}', \boldsymbol{\mu}'_{\mathrm{g}}, A', z, \mathbf{o}) \equiv \begin{cases}
+&&&& $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \begin{cases}
 \hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\
-\big(\boldsymbol{\sigma}, \boldsymbol{\mu}_{\mathrm{g}}, A, 0, () \big) & \text{otherwise} \end{cases}$ \\
+\big(\boldsymbol{\sigma}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} - \hyperlink{L_but_64}{L}(\boldsymbol{\mu}_{\mathrm{g}}) + g'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
@@
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
-(\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{o}} = \mathbf{o}$ \\
-&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} + g'$ \\
+&&&& $\boldsymbol{\mu}'_{\mathrm{g}} \equiv \boldsymbol{\mu}_{\mathrm{g}} - C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A) + g'$ \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\ (\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+  & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\mu}), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array} & \text{if} \quad I_{\mathrm{e}} < 1024 \\(\boldsymbol{\sigma}, g, A, 0, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+  & \text{if} \quad I_{\mathrm{e}} < 1024 \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2022-03-06 Paper.tex
Revert "Fix the upsidedown multiplier in the bloom filter footnote (#847)"
This reverts commit b8ffc51b68d53916b688028f0793588a3fdf563f.

@@
 where $M_{3:2048}$ is a specialised Bloom filter that sets three bits out of 2048, given an arbitrary byte sequence.
-It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{2048 $= 2^{11}$(11 bits), and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
+It does this through taking the low-order 11 bits of each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence.\footnote{11 bits $= 2^{2048}$, and the low-order 11 bits is the modulo 2048 of the operand, which is in this case is "each of the first three pairs of bytes in a Keccak-256 hash of the byte sequence."} Formally:
 \begin{eqnarray}

### 2022-03-09 Paper.tex
Don't cite withdrawn EIP-1706

--- 2022-02-21 Paper.tex
@@
 The astute reader will realise that this implies that no instruction can, through its execution, cause an exceptional halt.
-Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{SSTORE}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$.
-The last condition was introduced in EIP-1706 by \cite{EIP-1706} (part of EIP-2200 by \cite{EIP-2200}).
+Also, the execution is in an exceptional halting state if the gas left prior to executing an \hyperlink{SSTORE}{{\small SSTORE}} instruction is less than or equal to the call stipend $\hyperlink{G__callstipend}{G_{\mathrm{callstipend}}}$
+-- see EIP-2200 by \cite{EIP-2200} for more information.
 

### 2022-04-05 Paper.tex
Specify empty round brackets in Appendix D

@@
 n(\mathfrak{I}, i) \equiv \begin{cases}
-() & \text{if} \quad \mathfrak{I} = \varnothing \\
+() \in \mathbb{B} & \text{if} \quad \mathfrak{I} = \varnothing \\
 c(\mathfrak{I}, i) & \text{if} \quad \lVert \, \texttt{RLP} \big( c(\mathfrak{I}, i) \big) \rVert < 32 \\

### 2022-04-15 Paper.tex
Specify when touched accounts are updated

@@
 \boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
-m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}}
+m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}} \\
+A'_{\mathbf{t}} & \equiv & A_{\mathbf{t}} \cup \{ m, S(T) \}
 \end{eqnarray}
@@
 \linkdest{self_destruct_list_A__s}{}\forall i \in A_{\mathbf{s}}: \boldsymbol{\sigma}'[i] & = & \varnothing \\
-\linkdest{touched_A__t}{}\forall i \in A_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
+\linkdest{touched_A__t}{}\forall i \in A'_{\mathbf{t}}: \boldsymbol{\sigma}'[i] & = & \varnothing \quad\text{if}\quad \mathtt{DEAD}(\boldsymbol{\sigma}^*\kern -2pt, i)
 \end{eqnarray}
@@
 The address of the new account is added to the set of accessed accounts:
-\begin{equation}
-A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{a\}
-\end{equation}
+\begin{eqnarray}
+A^* & \equiv & A \quad \text{except:} \\
+A^*_{\mathbf{a}} & \equiv & A_{\mathbf{a}} \cup \{ a \} \\
+A^*_{\mathbf{t}} & \equiv & A_{\mathbf{t}} \cup \{ a, s \} \\
+\end{eqnarray}
 
@@
 \end{eqnarray}
-
 where $v'$ is the account's pre-existing value, in the event it was previously in existence:
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& $A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{t\}$ \\
+&&&& $A^* \equiv A \quad \text{except} \quad A^*_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{t\} \quad \text{and} \quad A^*_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup \{  I_{\mathrm{a}}, t \}$ \\
 &&&& $t \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] \bmod 2^{160}$ \\
@@
   (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $A^*_{\mathbf{t}} \equiv A_{\mathbf{t}}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& fact the same account as at present, simply that the code is overwritten.\\
+&&&& Also note that, on itself, {\small CALLCODE} does not update the set of touched accounts $\hyperlink{tx_touched_accounts_wordy_defn_A__t}{A_{\mathbf{t}}}$\\
+&&&& (unlike {\small CALL}).\\
+\end{tabu}
+
+\begin{tabu}{\usetabu{opcodes}}
 \midrule
@@
   (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+&&&& $A^*_{\mathbf{t}} \equiv A_{\mathbf{t}}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\
@@
 &&&& that the code is overwritten {\it and} the context is almost entirely identical.\\
+&&&& Also note that, on itself, {\small DELEGATECALL} does not update the set of touched\\
+&&&& accounts $\hyperlink{tx_touched_accounts_wordy_defn_A__t}{A_{\mathbf{t}}}$ (unlike {\small CALL}).\\
 \midrule
@@
 &&&& $A'_{\mathbf{a}} \equiv A_{\mathbf{a}} \cup \{ r \}$ \\
+&&&& $A'_{\mathbf{t}} \equiv A_{\mathbf{t}} \cup \{ r \}$ \\
 &&&& $\boldsymbol{\sigma}'[r] \equiv \begin{cases}

### 2022-04-24 Paper.tex
Add more clear explanation for out-of-gas exception
7. Contract Creation
https://github.com/ethereum/yellowpaper/blob/934279c87ab7e58c4d06f4042642afcfdab1f412/Paper.tex#L877
Add more clear explanation of exception
fix(sec. 9.2): Added two of the CALL opcodes
All CALLs caused a subordinate message call.

--- 2022-04-05 Paper.tex
@@
 The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CREATE2}, {\small CALL} and {\small CALLCODE}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CREATE2}, {\small CALL}, {\small CALLCODE}, {\small DELEGATECALL}, and {\small STATICCALL}.
 Finally, gas may be paid due to an increase in the usage of the memory.

### 2022-04-24 Paper.tex
fix(sec 9.2): Updated the memory expansion cost
The cost of memory expansion is not always proportional to the memory used. For memories above 704 bytes, it approximates the square of the memory used (see eq. 326).

@@
 
-Over an account's execution, the total fee for memory-usage payable is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
-This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will certainly result in an additional memory usage fee.
+Over an account's execution, the total fee for memory-usage payable is specified in \ref{app:vm}.
+If you use a small amount of memory, this fee is proportional to smallest multiple of 32 bytes that are required such that all memory indices (whether for read or write) are included in the range.
+For larger amounts of memory the fee is approximately the square of that number.
+See Appendix \label{app:vm} for the exact expression.
+This is paid for on a just-in-time basis; as such, referencing an area of memory at least 32 bytes greater than any previously indexed memory will result in an additional memory usage fee.
 Due to this fee it is highly unlikely addresses will ever go above 32-bit bounds.

### 2022-05-06 Paper.tex
fix(eq. 142): The actual initial value is ∅
See 2nd paragraph in section 9.4.
add REVERT data return semantics (#858)

--- 2022-04-24 Paper.tex
@@
 The gas remaining will be zero in any such exceptional condition, \ie if the creation was conducted as the reception of a transaction, then this doesn't affect payment of the intrinsic cost of contract creation; it is paid regardless.
-However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas.
+However, the value of the transaction is not transferred to the aborted contract's address when we are out-of-gas, thus the contract's code is not stored.
 
@@
 The first and most common is the fee intrinsic to the computation of the operation (see Appendix \ref{app:fees}).
-Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CREATE2}, {\small CALL}, {\small CALLCODE}, {\small DELEGATECALL}, and {\small STATICCALL}.
+Secondly, gas may be deducted in order to form the payment for a subordinate message call or contract creation; this forms part of the payment for {\small CREATE}, {\small CREATE2}, {\small CALL} and {\small CALLCODE}.
 Finally, gas may be paid due to an increase in the usage of the memory.
@@
 \linkdest{REVERT}{}0xfd & {\small REVERT} & 2 & 0 & Halt execution reverting state changes but returning data and remaining gas. \\
+&&&& $H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] \dots ( \boldsymbol{\mu}_{\mathbf{s}}[0] + \boldsymbol{\mu}_{\mathbf{s}}[1] - 1 ) ]$ \\
 &&&& The effect of this operation is described in (\ref{eq:X-def}). \\

### 2022-06-08 Paper.tex
Explain that ecrecover only accepts v={27,28}
Update Paper.tex
Co-authored-by: Alex Beregszaszi <alex@rtfs.hu>
Explain that ecrecover only accepts v={27,28} (#860)
* Explain that ecrecover only accepts v={27,28}

* Update Paper.tex

Co-authored-by: Alex Beregszaszi <alex@rtfs.hu>

* Update Paper.tex

Co-authored-by: Andrew Ashikhmin <34320705+yperbasis@users.noreply.github.com>

Co-authored-by: Andrew Ashikhmin <34320705+yperbasis@users.noreply.github.com>
Restrictions on r & s in ECDSARECOVER precompile

@@
 g_{\mathrm{r}} &=& 3000\\
-\lVert \mathbf{o} \rVert &=& \begin{cases} 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
+\lVert \mathbf{o} \rVert &=& \begin{cases}
+  0 & \text{if} \quad v \notin \{27, 28\} \,\vee\, r = 0 \,\vee\, r \ge \mathtt{secp256k1n} \,\vee\, s = 0 \,\vee\, s \ge \mathtt{secp256k1n} \\
+  0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v - 27, r, s) = \varnothing\\
+  32 & \text{otherwise}
+\end{cases}\\
 \text{if} \quad \lVert \mathbf{o} \rVert = 32: &&\\
 \mathbf{o}[0..11] &=& 0 \\
-\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v, r, s)\big)[12..31] \quad \text{where:}\\
+\mathbf{o}[12..31] &=& \mathtt{KEC}\big(\mathtt{ECDSARECOVER}(h, v - 27, r, s)\big)[12..31] \quad \text{where:}\\
 \mathbf{d}[0..(\lVert \hyperlink{I__d}{I_{\mathbf{d}}} \rVert-1)] &=& I_{\mathbf{d}}\\

### 2022-06-08 Paper.tex
Contract s restriction between precompile and signature recovery

@@
 g_{\mathrm{r}} &=& 3000\\
-\lVert \mathbf{o} \rVert &=& \begin{cases}
+\lVert \mathbf{o} \rVert &=& \begin{cases} \label{eq:ECDSARECOVER_precompile_validity}
   0 & \text{if} \quad v \notin \{27, 28\} \,\vee\, r = 0 \,\vee\, r \ge \mathtt{secp256k1n} \,\vee\, s = 0 \,\vee\, s \ge \mathtt{secp256k1n} \\
@@
 \linkdest{ECDSASIGN}\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 \end{align}
+Note that this restriction on $s$ is more stringent than restriction \ref{eq:ECDSARECOVER_precompile_validity} in the ECREC precompile; see EIP-2 by \cite{EIP-2} for more detail.
 

### 2022-06-08 Paper.tex
Cosmetic change

@@
 g_{\mathrm{r}} &=& 3000\\
-\lVert \mathbf{o} \rVert &=& \begin{cases} \label{eq:ECDSARECOVER_precompile_validity}
+\lVert \mathbf{o} \rVert &=& \begin{cases} \label{eq:ecrec_precompile_validity}
   0 & \text{if} \quad v \notin \{27, 28\} \,\vee\, r = 0 \,\vee\, r \ge \mathtt{secp256k1n} \,\vee\, s = 0 \,\vee\, s \ge \mathtt{secp256k1n} \\
@@
 \end{align}
-Note that this restriction on $s$ is more stringent than restriction \ref{eq:ECDSARECOVER_precompile_validity} in the ECREC precompile; see EIP-2 by \cite{EIP-2} for more detail.
+Note that this restriction on $s$ is more stringent than restriction \ref{eq:ecrec_precompile_validity} in the ECREC precompile; see EIP-2 by \cite{EIP-2} for more detail.
 

### 2022-06-08 Paper.tex
Update Paper.tex
Co-authored-by: Alex Beregszaszi <alex@rtfs.hu>

@@
 \end{align}
-Note that this restriction on $s$ is more stringent than restriction \ref{eq:ecrec_precompile_validity} in the ECREC precompile; see EIP-2 by \cite{EIP-2} for more detail.
+Note that this restriction on $s$ is more stringent than restriction \ref{eq:ecrec_precompile_validity} in the $\Xi_{\mathtt{ECREC}}$ precompile; see EIP-2 by \cite{EIP-2} for more detail.
 

### 2022-06-09 Paper.tex
Fix: revert message should be available in the returndata buffer in case a CREATE reverts

--- 2022-06-08 Paper.tex
@@
 We define $\Xi_{\mathtt{ECREC}}$ as a precompiled contract for the elliptic curve digital signature algorithm (ECDSA) public key recovery function (ecrecover).
-See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$ and the constant $\mathtt{secp256k1n}$.
+See Appendix \ref{app:signing} for the definition of the function $\mathtt{ECDSARECOVER}$.
 We also define $\mathbf{d}$ to be the input data, well-defined for an infinite length by appending zeroes as required.
-In the case of an invalid signature, we return no output.
+In the case of an invalid signature ($\mathtt{ECDSARECOVER}(h, v, r, s) = \varnothing$), we return no output.
 
@@
 g_{\mathrm{r}} &=& 3000\\
-\lVert \mathbf{o} \rVert &=& \begin{cases} \label{eq:ecrec_precompile_validity}
-  0 & \text{if} \quad v \notin \{27, 28\} \,\vee\, r = 0 \,\vee\, r \ge \mathtt{secp256k1n} \,\vee\, s = 0 \,\vee\, s \ge \mathtt{secp256k1n} \\
-  0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v - 27, r, s) = \varnothing\\
-  32 & \text{otherwise}
-\end{cases}\\
+\lVert \mathbf{o} \rVert &=& \begin{cases} 0 & \text{if} \quad \mathtt{v} \neq 27 \wedge \mathtt{v} \neq 28 \\ 0 & \text{if} \quad \mathtt{ECDSARECOVER}(h, v - 27, r, s) = \varnothing\\ 32 & \text{otherwise} \end{cases}\\
 \text{if} \quad \lVert \mathbf{o} \rVert = 32: &&\\
@@
 \linkdest{ECDSASIGN}\mathtt{ECDSASIGN}(e \in \mathbb{B}_{32}, p_{\mathrm{r}} \in \mathbb{B}_{32}) & \equiv & (v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) \\
-\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{2}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
+\mathtt{ECDSARECOVER}(e \in \mathbb{B}_{32}, v \in \mathbb{B}_{1}, r \in \mathbb{B}_{32}, s \in \mathbb{B}_{32}) & \equiv & p_{\mathrm{u}} \in \mathbb{B}_{64}
 \end{eqnarray}
@@
 \end{align}
-Note that this restriction on $s$ is more stringent than restriction \ref{eq:ecrec_precompile_validity} in the ECREC precompile; see EIP-2 by \cite{EIP-2} for more detail.
 
@@
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv ()$ \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{o}} \equiv \begin{cases}
+() & \text{if} \quad z = 1 \\
+\mathbf{o} & \text{otherwise}
+\end{cases}$ \\
 &&&& Thus the operand order is: value, input offset, input size. \\

### 2022-07-01 Paper.tex
Gray Glacier bomb delay (#866)

@@
 $F_{\mathrm{Arrow Glacier}}$     & 13773000 \\
+$F_{\mathrm{Gray Glacier}}$      & 15050000 \\
 \bottomrule
@@
   9700000  & \text{if} \quad F_{\mathrm{London}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Arrow Glacier}} \\
- 10700000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Arrow Glacier}} \\
+ 10700000  & \text{if} \quad F_{\mathrm{Arrow Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Gray Glacier}} \\
+ 11400000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Gray Glacier}} \\
 \end{cases}
@@
 which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-Subsequently, EIP-1234 by \cite{EIP-1234}, EIP-2384 by \cite{EIP-2384}, EIP-3554 by \cite{EIP-3554}, and EIP-4345 by \cite{EIP-4345} increased the subtrahend $\kappa$.
+Subsequently, EIP-1234 by \cite{EIP-1234}, EIP-2384 by \cite{EIP-2384}, EIP-3554 by \cite{EIP-3554}, EIP-4345 by \cite{EIP-4345}, and EIP-5133 by \cite{EIP-5133} increased the subtrahend $\kappa$.
 

### 2022-07-12 Paper.tex
test

@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Berlin version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
+\title[イーサリアム: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Berlin version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
 \author{

### 2022-07-13 Paper.tex
Fix memory linear cost limit: 724B -> 704B
The "a" in the C_mem function is number of *words*.

If a * a >= 512, the cost will become non-linear.

1 word = 32 bytes

>>> bytes_used = 704
>>> a = int(math.ceil(float(bytes_used) / 32))
>>> a
22
>>> a * a >= 512
False

>>> bytes_used = 705
>>> a = int(math.ceil(float(bytes_used) / 32))
>>> a
23
>>> a * a >= 512
True

The cost is linear up to 704B.

--- 2022-07-01 Paper.tex
@@
 Note also that $C_{\mathrm{mem}}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
+It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 704B of memory used, after which it costs substantially more.
 

### 2022-07-13 Paper.tex
add: first commit

--- 2022-07-12 Paper.tex
@@
 
-\title[イーサリアム: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Berlin version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
+\title[イーサリアム: 安全な分散型汎用トランザクション台帳\\ \smaller \textbf{{Berlin version}}]{イーサリアム: 安全な分散型汎用トランザクション台帳\\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
 \author{

### 2022-07-13 Paper.tex
Contract code is effectively muttable with CREATE2 (#869)

--- 2022-07-13 Paper.tex
@@
 Note also that $C_{\mathrm{mem}}$ is the memory cost function (the expansion function being the difference between the cost before and after).
-It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 704B of memory used, after which it costs substantially more.
+It is a polynomial, with the higher-order coefficient divided and floored, and thus linear up to 724B of memory used, after which it costs substantially more.
 

### 2022-09-28 Paper.tex
add: title
Fix DIFFICULTY opcode with PREVRANDAO
Reference: https://eips.ethereum.org/EIPS/eip-4399

Since the Merge, opcode 0x44 returns the output of the randomness beacon provided by the beacon chain

@@
 
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
 
@@
 \midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the current block's difficulty. \\
+0x44 & {\small PREVRANDAO} & 0 & 1 & Get the randomness beacon. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\

### 2023-11-08 Paper.tex
incorporate london EIPs

--- 2022-10-23 Paper.tex
@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Berlin version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Berlin version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{London version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{London version \YellowPaperVersionNumber}}}
 \author{
@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol, namely the \textit{Berlin} version defined by \cite{BeikoBerlin}.
+This document describes one version of the protocol, namely the \textit{London} version defined by \cite{BeikoLondon}.
 In order to follow back the history of a path, one must reference multiple versions of this document.
@@
 EIP-2718 by \cite{EIP-2718} introduced the notion of different transaction types.
-As of the Berlin version of the protocol, there are two transaction types: 0 (legacy) and 1 (EIP-2930 by \cite{EIP-2930}).
+As of the \textit{London} version of the protocol, there are three transaction types: 0 (legacy), 1 (EIP-2930 by \cite{EIP-2930}), and 2 (EIP-1559 by \cite{EIP-1559}).
 Further, there are two subtypes of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation').
@@
 \item[nonce]\linkdest{tx_nonce}{} A scalar value equal to the number of transactions sent by the sender; formally $T_{\mathrm{n}}$.
-\item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
 \item[gasLimit]\linkdest{tx_gas_limit_T__g}{} A scalar value equal to the maximum amount of gas that should be used in executing this transaction.
@@
 
-EIP-2930 (type 1) transactions also have:
+EIP-2930 (type 1) and EIP-1559 (type 2) transactions also have:
 \begin{description}
@@
 
-Additionally, a contract creation transaction (regardless whether legacy or EIP-2930) contains:
+There are differences in how one's acceptable gas price is specified in type 2 transactions versus type 0 and type 1 transactions.
+Type 2 transactions take advantage of the gas market improvements introduced in EIP-1559 and thus have the following fields related to gas price:
+
+\begin{description}
+  \item[maxFeePerGas]\linkdest{max_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{m}}$.
+  \item[maxPriorityFeePerGas]\linkdest{max_priority_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid to the block's fee recicipient as an incentive to include the transaction; formally $T_{\mathrm{f}}$.
+\end{description}
+
+In contrast, type 0 and type 1 transactions specify gas price as a single value:
+
+\begin{description}
+  \item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+\end{description}
+
+Additionally, a contract creation transaction (regardless of transaction type) contains:
 
@@
 (T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{m}}, T_{\mathrm{f}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
 \end{cases}
@@
 T_{\mathrm{w}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{r}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{s}} \in \mathbb{N}_{256} & \wedge \\
-T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B}
+T_{\mathrm{y}} \in \mathbb{N}_{1} & \wedge & T_{\mathbf{d}} \in \mathbb{B} & \wedge & T_{\mathbf{i}} \in \mathbb{B} & \wedge \\
+T_{\mathrm{m}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{f}} \in \mathbb{N}_{256}
 \end{array}
@@
 \item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
+\item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
 \end{description}
@@
 \begin{eqnarray}
-\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
 \quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
@@
 \hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8}
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
+\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N}
+
 \end{array}
@@
 
+\linkdest{H__f}The \textit{London} release introduced the block attribute \textit{base fee per gas} \hyperlink{block_baseFeePerGas_H__f}{$H_{\mathrm{f}}$} (see EIP-1559 by \cite{EIP-1559}).
+The base fee is the amount of wei burned per unit of gas consumed while executing transactions within the block.
+The value of the base fee is a function of the difference between the gas used by the parent block and the parent block's gas target.
+
+The expected base fee per gas is defined as $F(H)$:
+
+\begin{equation}
+  F(H) \equiv \begin{cases} 
+    1000000000  & \text{if} \quad H_{\mathrm{i}} = F_{\mathrm{London}} \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}}  & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} = \tau \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}} - \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}} + \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+  \end{cases}
+  \end{equation}
+
+where:
+  \begin{align}
+    \tau &\equiv \frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{\rho} \\
+    \rho &\equiv 2 \\
+    \nu &\equiv \begin{cases}
+      \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({\tau - P(H)_{\mathrm{H}}}_{\mathrm{g}})}{\tau \times \xi} &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+      \max(\frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({P(H)_{\mathrm{H}}}_{\mathrm{g}} - \tau)}{\tau \times \xi},1) &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+  \end{cases} \\
+    \xi &\equiv 8
+\end{align}
+
+The \textit{gas target}, $\tau$, is defined as the gas limit $H_{\mathrm{l}}$ divided by the \textit{elasticity multiplier}, $\rho$, a global constant set to 2.
+So while blocks can consume as much gas as the gas limit, the base fee is adjusted so that on average, blocks consume as much gas as the gas target.
+The base fee is increased in the current block when the parent block's gas usage exceeds the gas target, and, conversely, the base fee is decreased in the current block when the parent block's gas usage is less than the gas target.
+
+The magnitude of the increase or decrease in the base fee, defined as $\nu$, is proportional to the difference between the amount of gas the parent block consumed and the parent block's gas target.
+The effect on the base fee is dampened by a global constant called the \textit{base fee max change denominator}, formally $\xi$, set to 8.
+A value of 8 entails that the base fee can increase or decrease by at most 12.5\% from one block to the next.
+
 \hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
@@
 \nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
-\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
+\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32 \quad \wedge \\
+\nonumber& & H_{\mathrm{f}} = F(H)
 \end{eqnarray}
@@
 
-Noting additionally that \textbf{extraData} must be at most 32 bytes.
+Note additionally that \textbf{extraData} must be at most 32 bytes.
 
@@
 This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according \textbf{gasPrice}, also specified in the transaction.
+The purchase happens at the according to the \textit{effective gas price} defined in section \ref{ch:transactions}.
 The transaction is considered invalid if the account balance cannot support such a purchase.
@@
 
-In general, Ether used to purchase gas that is not refunded is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
-Transactors are free to specify any \textbf{gasPrice} that they wish, however miners are free to ignore transactions as they choose.
-A higher gas price on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion by more miners.
-Miners, in general, will choose to advertise the minimum gas price for which they will execute transactions and transactors will be free to canvas these prices in determining what gas price to offer.
-Since there will be a (weighted) distribution of minimum acceptable gas prices, transactors will necessarily have a trade-off to make between lowering the gas price and maximising the chance that their transaction will be mined in a timely manner.
+Since the introduction of EIP-1559 by \cite{EIP-1559} in the \textit{London} hard fork, every transaction included in a block must pay a \textit{base fee}, which is specified as wei per unit of gas consumed and is constant for each transaction within a block.
+The ether that is paid to meet the base fee is burned (taken out of circulation).
+The base fee adjusts dynamically as a function of the previous block's gas consumption relative to its \textit{gas target} (a value that is currently half the block's gas limit, which can be adjusted by miners).
+If the previous block's total gas consumption exceeds the gas target, this indicates excess demand for block space at the current base fee, and the base fee is increased.
+Conversely, if the gas consumed in the previous block is lower than the gas target, demand for block space is lower than the gas target at the current base fee, and thus the base fee is decreased.
+This process of adjusting the base fee should bring the average block's gas consumption in line with the gas target.
+Refer to section \ref{subsec:The_Block} for greater detail on how the base fee is set.
+
+To incentivize miners to include transactions, there is an additional fee known as a \textit{priority fee}, which is also specified as wei per unit of gas consumed.
+The total fee fee paid by the transactor therefore is the sum of the base fee per gas and the priority fee per gas multiplied by the total gas consumed.
+Ether used to satisfy the priority fee is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
+
+Transactors using type 2 transactions can specify the maximum priority fee they are willing to pay (\textbf{maxPriorityFeePerGas}) as well as the max total fee they are willing to pay (\textbf{maxFeePerGas}), inclusive of both the priority fee and the base fee.
+\textbf{maxFeePerGas} must at least as high as the base fee for the transaction to be included in a block.
+Type 0 and type 1 transactions have only one field for specifiying a gas price--\textbf{gasPrice}--which also must be at least as high as the base fee for inclusion in a block.
+The amount by which \textbf{gasPrice} is higher than the base fee constitutes the priority fee in the case of a type 0 or type 1 transaction.
+
+Transactors are free to select any priority fee that they wish, however miners are free to ignore transactions as they choose.
+A higher priority fee on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion.
+Since there will be a (weighted) distribution of minimum acceptable priority fees, transactors will necessarily have a trade-off to make between lowering the priority fee and maximising the chance that their transaction will be mined in a timely manner.
 
@@
 \item the sender account has no contract code deployed (see EIP-3607 by \cite{EIP-3607});
-\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
+\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
+\item the sender account balance contains at least the cost, $v_0$, required in up-front payment; and
+\item the effective gas price of the transaction is greater than or equal to the block's base fee $H_{\mathrm{f}}$.
 \end{enumerate}
@@
 
+We define the \textit{effective gas price}, formally $p$, as the amount of wei the transaction signer will pay per unit of gas consumed during the transaction's execution. It is calculated as follows:
+
+\begin{equation}
+  p \equiv \begin{cases}
+  T_{\mathrm{p}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  f + H_{\mathrm{f}} & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
+\end{equation}
+
+where $f$ is the \textit{priority fee}---the amount of wei the block's beneficiary address will receive per unit of gas consumed during the transaction's execution. It is calculated as:
+
+\begin{equation}
+  f \equiv \begin{cases}
+  T_{\mathrm{p}} - H_{\mathrm{f}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  min(T_{\mathrm{f}}, T_{\mathrm{m}} - H_{\mathrm{f}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
+\end{equation}
+
 The up-front cost $v_0$ is calculated as:
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} \hyperlink{tx_gas_price_T__p}{T_{\mathrm{p}}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} p + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
 \end{equation}
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
+p & \geqslant & H_{\mathrm{f}} \quad \wedge \\
 T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
@@
 
-The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}}T_{\mathrm{p}}$.
+The execution of a valid transaction begins with an irrevocable change made to the state: the \hyperlink{account_nonce}{nonce of the account of the sender}, $S(T)$, is incremented by one and the balance is reduced by part of the up-front cost, $T_{\mathrm{g}} p$.
 The gas available for the proceeding computation, $g$, is defined as $T_{\mathrm{g}} - g_0$.
@@
 \linkdest{sigma_0}{}\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{p}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{e}} \\
 \boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
@@
 (\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{p}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; T_{\mathrm{e}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{e}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
-\begin{equation}
-	A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \sum_{i \in A_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
-\end{equation}
-
 Then the state is finalised by determining the amount to be refunded, $g^*$ from the remaining gas, $g'$, plus some allowance from the refund counter, to the sender at the original rate.
 \begin{equation}
-g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{2} \Big\rfloor, A'_{\mathrm{r}} \right\}
-\end{equation}
-
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of half (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+g^* \equiv g' + \min \left\{ \Big\lfloor \dfrac{T_{\mathrm{g}} - g'}{5} \Big\rfloor, A_{\mathrm{r}} \right\}
+\end{equation}
+
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of one fifth\footnote{The max refundable proportion of gas was changed by EIP-3529 by \cite{EIP-3529}} (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
 
-The Ether for the gas is given to the miner, whose address is specified as the beneficiary of the present block $B$.
-So we define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$:
+The miner, whose address is specified as the beneficiary of the present block $B$, receives the gas consumed multiplied by the transaction's \textit{priority fee per gas}, defined as $f$ defined in this section.
+The ether that is paid by the transactor that goes toward the base fee is debited from the transactor's account but credited to no other account, so it is burned.
+
+We define the pre-final state $\boldsymbol{\sigma}^*$ in terms of the provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$:
 \begin{eqnarray}
 \boldsymbol{\sigma}^* & \equiv & \boldsymbol{\sigma}_{\mathrm{P}} \quad \text{except} \\
-\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* T_{\mathrm{p}} \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) T_{\mathrm{p}} \\
+\boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* p \\
+\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) f \\
 m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}}
@@
 There are a number of intrinsic parameters used when creating an account: sender ($s$), original transactor\footnote{which can differ from the sender in the case of a message call or contract creation not directly triggered by a transaction but coming from the execution of EVM-code} ($o$),
-available gas ($g$), gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
+available gas ($g$), effective gas price ($p$), endowment ($v$) together with an arbitrary length byte array, $\mathbf{i}$, the initialisation EVM code, the present depth of the message-call/contract-creation stack ($e$), the salt for new account's address ($\zeta$) and finally the permission to make modifications to the state ($w$).
 The salt \linkdest{salt}{$\zeta$} might be missing ($\zeta = \varnothing$); formally,
@@
 \nonumber &g^{**} < c \quad \vee \\
-\nonumber &\lVert \mathbf{o} \rVert > 24576
+\nonumber &\lVert \mathbf{o} \rVert > 24576 \quad \vee \\
+\nonumber &\mathbf{o}[0] = \texttt{0xef}
 \end{align}
@@
 \section{Message Call} \label{ch:call}
-In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
+In the case of executing a message call, several parameters are required: sender ($s$), transaction originator ($o$), recipient ($r$), the account whose code is to be executed ($c$, usually the same as recipient), available gas ($g$), value ($v$) and effective gas price ($p$) together with an arbitrary length byte array, $\mathbf{d}$, the input data of the call, the present depth of the message-call/contract-creation stack ($e$) and finally the permission to make modifications to the state ($w$).
 
@@
 \item\linkdest{address_of_sender_of_tx_that_originated_execution_I__o}{}\linkdest{I__o} $I_{\mathrm{o}}$, the sender address of the transaction that originated this execution.
-\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas in the transaction that originated this execution.
+\item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas paid by the signer of the transaction that originated this execution.
+This is defined as the effective gas price $p$ in section \ref{ch:transactions}.
 \item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
@@
 
-\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction. Three different flavours of signing schemes are available:
+\hypertarget{h_of_T}{}The message hash, $h(T)$, to be signed is the Keccak-256 hash of the transaction. Four different flavours of signing schemes are available:
 \begin{eqnarray}
@@
 (T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, \beta, (), ()) & \text{if} \; T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
-(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 1
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{m}}, T_{\mathrm{f}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 2
 \end{cases} \\
@@
 (T_{\mathrm{w}}-35) \bmod 2 & \text{if} \ T_{\mathrm{x}} = 0 \land T_{\mathrm{w}} \in \{2\beta + 35, 2\beta + 36\} \\
-T_{\mathrm{y}} & \text{if} \ T_{\mathrm{x}} = 1
+T_{\mathrm{y}} & \text{if} \ T_{\mathrm{x}} = 1 \lor T_{\mathrm{x}} = 2
 \end{cases}
@@
 &&is set to zero. \\
-$R_{\mathrm{sclear}}$ & 15000 & Refund given (added into refund counter) when the storage value is set to zero from\\
-&&non-zero.\\
-\linkdest{R__selfdestruct}{}$R_{\mathrm{selfdestruct}}$ & 24000 & Refund given (added into refund counter) for self-destructing an account. \\
+$R_{\mathrm{sclear}}$ & 4800 & Refund given (added into refund counter) when the storage value is set to zero from\\
+&&non-zero. The refund amount is defined as $G_{\mathrm{sreset}} + G_{\mathrm{accessliststorage}}$. \\
 \linkdest{G__selfdestruct}{}$G_{\mathrm{selfdestruct}}$ & 5000 & Amount of gas to pay for a {\small SELFDESTRUCT} operation. \\
@@
 
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}\}
 
@@
 0x3a & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
+&&&& This is the \textit{effective gas price} defined in section \ref{ch:transactions}. \\
+&&&& Note that as of the \textit{London} hard fork, this value no longer \\ 
+&&&& represents what is received by the miner, \\
+&&&& but rather just what is paid by the sender. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{p}}$ \\
-&&&& This is gas price specified by the originating transaction.\\
 \midrule
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} $ \\
+\midrule
+0x48 & {\small BASEFEE} & 0 & 1 & Get the current block's base fee. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{f}}$ \\
 \bottomrule

### 2023-11-08 Paper.tex
various grammar and reference edits

@@
 There are differences in how one's acceptable gas price is specified in type 2 transactions versus type 0 and type 1 transactions.
-Type 2 transactions take advantage of the gas market improvements introduced in EIP-1559 and thus have the following fields related to gas price:
+Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee} that is paid.
+Type two transactions have the following two fields related to gas:
 
@@
 The base fee is the amount of wei burned per unit of gas consumed while executing transactions within the block.
-The value of the base fee is a function of the difference between the gas used by the parent block and the parent block's gas target.
+The value of the base fee is a function of the difference between the gas used by the parent block and the parent block's \textit{gas target}.
 
@@
 This is the amount of gas which is implicitly purchased from the sender's account balance.
-The purchase happens at the according to the \textit{effective gas price} defined in section \ref{ch:transactions}.
+The purchase happens at the \textit{effective gas price} defined in section \ref{ch:transactions}.
 The transaction is considered invalid if the account balance cannot support such a purchase.
@@
 
-We define the \textit{effective gas price}, formally $p$, as the amount of wei the transaction signer will pay per unit of gas consumed during the transaction's execution. It is calculated as follows:
+\hypertarget{effective_gas_price_p}{}We define the \textit{effective gas price}, formally $p$, as the amount of wei the transaction signer will pay per unit of gas consumed during the transaction's execution.
+It is calculated as follows:
 
@@
 
-where $f$ is the \textit{priority fee}---the amount of wei the block's beneficiary address will receive per unit of gas consumed during the transaction's execution. It is calculated as:
+\hypertarget{priority_fee_f}{}where $f$ is the \textit{priority fee}---the amount of wei the block's beneficiary address will receive per unit of gas consumed during the transaction's execution.
+It is calculated as:
 
@@
 \linkdest{sigma_0}{}\boldsymbol{\sigma}_0 & \equiv & \boldsymbol{\sigma} \quad \text{except:} \\
-\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} T_{\mathrm{e}} \\
+\boldsymbol{\sigma}_0[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} - T_{\mathrm{g}} p \\
 \boldsymbol{\sigma}_0[S(T)]_{\mathrm{n}} & \equiv & \boldsymbol{\sigma}[S(T)]_{\mathrm{n}} + 1
@@
 (\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; T_{\mathrm{e}}, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, T_{\mathrm{e}}, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; p, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, p, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-The miner, whose address is specified as the beneficiary of the present block $B$, receives the gas consumed multiplied by the transaction's \textit{priority fee per gas}, defined as $f$ defined in this section.
+The miner, whose address is specified as the beneficiary of the present block $B$, receives the gas consumed multiplied by the transaction's \textit{priority fee per gas}, defined as \hyperlink{priority_fee_f}{$f$} in this section.
 The ether that is paid by the transactor that goes toward the base fee is debited from the transactor's account but credited to no other account, so it is burned.
@@
 \item\linkdest{gas_price_in_tx_that_originated_execution_I__p}{}\linkdest{I__p} $I_{\mathrm{p}}$, the price of gas paid by the signer of the transaction that originated this execution.
-This is defined as the effective gas price $p$ in section \ref{ch:transactions}.
+This is defined as the effective gas price \hyperlink{effective_gas_price_p}{$p$} in section \ref{ch:transactions}.
 \item\linkdest{execution_input_data_I__d}{}\linkdest{I__d} $I_{\mathbf{d}}$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.

### 2023-11-08 Paper.tex
grammar edits and clarifications

@@
   \item[maxFeePerGas]\linkdest{max_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{m}}$.
-  \item[maxPriorityFeePerGas]\linkdest{max_priority_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid to the block's fee recicipient as an incentive to include the transaction; formally $T_{\mathrm{f}}$.
+  \item[maxPriorityFeePerGas]\linkdest{max_priority_fee_per_gas}{} A scalar value equal to the maximum number of Wei to be paid to the block's fee recipient as an incentive to include the transaction; formally $T_{\mathrm{f}}$.
 \end{description}
@@
 
-The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of one fifth\footnote{The max refundable proportion of gas was changed by EIP-3529 by \cite{EIP-3529}} (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
+The total refundable amount is the legitimately remaining gas $g'$, added to \hyperlink{refund_balance_defn_words_A__r}{$A_{\mathrm{r}}$}, with the latter component being capped up to a maximum of one fifth\footnote{The max refundable proportion of gas was reduced from one half to one fifth by EIP-3529 by \cite{EIP-3529} in the \textit{London} release} (rounded down) of the total amount used $T_{\mathrm{g}} - g'$.
 Therefore, $g^*$ is the total gas that remains after the transaction has been executed.
@@
 \hypertarget{contract_creation_result}{}
+
+Note the last condition of $F$ indicates that contract code cannot begin with the byte \texttt{0xef} (refer to EIP-3541 by \cite{EIP-3541}),
 

### 2023-11-08 Paper.tex
fix inconsistency with transaction type
add footnote for priority fee

@@
 There are differences in how one's acceptable gas price is specified in type 2 transactions versus type 0 and type 1 transactions.
-Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee} that is paid.
-Type two transactions have the following two fields related to gas:
+Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee}\footnote{The \textit{priority fee} is discussed in greater detail in sections \ref{ch:payment} and \ref{ch:transactions}} that is paid.
+Type 2 transactions have the following two fields related to gas:
 

### 2023-11-08 Paper.tex
missing period

@@
 There are differences in how one's acceptable gas price is specified in type 2 transactions versus type 0 and type 1 transactions.
-Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee}\footnote{The \textit{priority fee} is discussed in greater detail in sections \ref{ch:payment} and \ref{ch:transactions}} that is paid.
+Type 2 transactions take better advantage of the gas market improvements introduced in EIP-1559 by explicitly limiting the \textit{priority fee}\footnote{The \textit{priority fee} is discussed in greater detail in sections \ref{ch:payment} and \ref{ch:transactions}.} that is paid.
 Type 2 transactions have the following two fields related to gas:

### 2023-11-09 Paper.tex
add arrow glacier and gray glacier forks

@@
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes one version of the protocol, namely the \textit{London} version defined by \cite{BeikoLondon}.
+This document describes the \textit{London}, \textit{Arrow Glacier}, and \textit{Gray Glacier} versions of the protocol.\footnote{Note that the \textit{Arrow Glacier} and \textit{Gray Glacier} forks only affected the \hyperlink{difficulty_bomb}{"difficulty bomb"} described in section \ref{subsec:The_Block}.}
+Note that this document also covers the  and  versions of Ethereum, .
 In order to follow back the history of a path, one must reference multiple versions of this document.
@@
 
-The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
+\hypertarget{difficulty_bomb}{}The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
 In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.

### 2023-11-09 Paper.tex
get rid of extraneous text

@@
 This document describes the \textit{London}, \textit{Arrow Glacier}, and \textit{Gray Glacier} versions of the protocol.\footnote{Note that the \textit{Arrow Glacier} and \textit{Gray Glacier} forks only affected the \hyperlink{difficulty_bomb}{"difficulty bomb"} described in section \ref{subsec:The_Block}.}
-Note that this document also covers the  and  versions of Ethereum, .
 In order to follow back the history of a path, one must reference multiple versions of this document.

### 2023-11-09 Paper.tex
fix order of items in the transaction tuple

@@
 (T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
-(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{m}}, T_{\mathrm{f}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{f}}, T_{\mathrm{m}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}, T_{\mathrm{y}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{x}} = 2 \\
 \end{cases}
@@
 (T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 1 \\
-(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{m}}, T_{\mathrm{f}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 2
+(T_{\mathrm{c}}, T_{\mathrm{n}}, T_{\mathrm{f}}, T_{\mathrm{m}},  T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, \mathbf{p}, T_{\mathbf{A}}) & \text{if} \; T_{\mathrm{x}} = 2
 \end{cases} \\

### 2023-11-27 Paper.tex
clarify the division behavior for the change in base fee

@@
     {P(H)_{\mathrm{H}}}_{\mathrm{f}} - \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
-    {P(H)_{\mathrm{H}}}_{\mathrm{f}} + \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+    {P(H)_{\mathrm{H}}}_{\mathrm{f}} + \nu & \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau
   \end{cases}
@@
     \rho &\equiv 2 \\
+    \nu^* &\equiv \begin{cases}
+      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({\tau - P(H)_{\mathrm{H}}}_{\mathrm{g}})}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+      \lfloor \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({P(H)_{\mathrm{H}}}_{\mathrm{g}} - \tau)}{\tau} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+  \end{cases} \\
   \nu &\equiv \begin{cases}
-      \frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({\tau - P(H)_{\mathrm{H}}}_{\mathrm{g}})}{\tau \times \xi} &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
-      \max(\frac{{P(H)_{\mathrm{H}}}_{\mathrm{f}} \times ({P(H)_{\mathrm{H}}}_{\mathrm{g}} - \tau)}{\tau \times \xi},1) &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
+    \lfloor \frac{\nu^*}{\xi} \rfloor &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} < \tau \\
+    max(\lfloor \frac{\nu^*}{\xi} \rfloor, 1) &  \text{if} \quad {P(H)_{\mathrm{H}}}_{\mathrm{g}} > \tau \\
   \end{cases} \\

### 2023-11-27 Paper.tex
add floor to gas target calculation
adjust gas limit verification at the London fork block

@@
 
-\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
-\begin{eqnarray}
-& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} \geqslant 5000
-\end{eqnarray}
-
 \linkdest{H__f}The \textit{London} release introduced the block attribute \textit{base fee per gas} \hyperlink{block_baseFeePerGas_H__f}{$H_{\mathrm{f}}$} (see EIP-1559 by \cite{EIP-1559}).
@@
   \begin{align}
-    \tau &\equiv \frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{\rho} \\
+    \tau &\equiv \lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{\rho} \rfloor \\
     \rho &\equiv 2 \\
@@
 
+\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
+\begin{eqnarray}
+& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} \geqslant 5000
+\end{eqnarray}
+
+where:
+  \begin{align}
+    {P(H)_{\mathrm{H}}}_{\mathrm{l}'} &\equiv \begin{cases}
+      {P(H)_{\mathrm{H}}}_{\mathrm{l}} \times \rho &  \text{if} \quad H_{\mathrm{i}} = F_{\mathrm{London}} \\
+      {P(H)_{\mathrm{H}}}_{\mathrm{l}} &  \text{if} \quad H_{\mathrm{i}} > F_{\mathrm{London}} \\
+  \end{cases}
+\end{align}
+
+To avoid a discontinuity in gas usage, the value of the parent block gas limit for the purpose of validating the current block's gas limit is modified at the \textit{London} fork block by multiplying it by the \textit{elasticity multiplier}, $\rho$.
+We call this modified value ${P(H)_{\mathrm{H}}}_{\mathrm{l}'}$.
+This ensures that the gas target for post-\textit{London} blocks can be set roughly in line with the gas limit of pre-\textit{London} blocks.
+
 \hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
@@
 \nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l'}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\

### 2023-11-28 Paper.tex
fix typo in block header validity function
add a footnote regarding the behavior of Tp as it relates to type 2 txs

@@
 \begin{description}
-  \item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.
+  \item[gasPrice]\linkdest{tx_gas_price_T__p}{} A scalar value equal to the number of Wei to be paid per unit of \textit{gas} for all computation costs incurred as a result of the execution of this transaction; formally $T_{\mathrm{p}}$.\footnote{Type 0 and type 1 transactions will get the same gas price behavior as a type 2 transaction with $T_{\mathrm{m}}$ and $T_{\mathrm{f}}$ set to the value of $T_{\mathrm{p}}$.}
 \end{description}
@@
 \nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
-\nonumber& & H_{\mathrm{l'}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
+\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}'} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\

### 2023-11-29 Paper.tex
fix transaction verification for gas values

@@
 \begin{equation}
-v_0 \equiv \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}} p + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}}
+  v_0 \equiv \begin{cases}
+  \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}}T_{\mathrm{p}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  \hyperlink{tx_gas_limit_T__g}{T_{\mathrm{g}}}T_{\mathrm{m}} + \hyperlink{tx_value_T__v}{T_{\mathrm{v}}} & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
 \end{equation}
@@
 v_0 & \leqslant & \boldsymbol{\sigma}[S(T)]_{\mathrm{b}} \quad \wedge \\
-p & \geqslant & H_{\mathrm{f}} \quad \wedge \\
+m & \geqslant & H_{\mathrm{f}} \quad \wedge \\
 T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
@@
 \end{equation}
+where
+\begin{equation}
+  m \equiv \begin{cases}
+  T_{\mathrm{p}} & \text{if} \; T_{\mathrm{x}} = 0 \lor T_{\mathrm{x}} = 1 \\
+  T_{\mathrm{m}} & \text{if} \; T_{\mathrm{x}} = 2 \\
+  \end{cases}
+\end{equation}
+
 Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
 Also, with a slight abuse of notation, we assume that $\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} = \texttt{KEC}\big( () \big)$, $\boldsymbol{\sigma}[S(T)]_{\mathrm{n}} = 0$, and $\boldsymbol{\sigma}[S(T)]_{\mathrm{b}} = 0$ if $\boldsymbol{\sigma}[S(T)] = \varnothing$.
+
+For type 2 transactions, we add an additional check that \textbf{maxPriorityFeePerGas} is no larger than \textbf{maxFeePerGas}:
+\begin{equation}
+  T_{\mathrm{m}} \geqslant T_{\mathrm{f}}
+\end{equation}
 

### 2023-12-07 Paper.tex
correction to transaction validity, and disambiguate the variable m in transaction execution

@@
 \item the sender account balance contains at least the cost, $v_0$, required in up-front payment; and
-\item the effective gas price of the transaction is greater than or equal to the block's base fee $H_{\mathrm{f}}$.
+\item the \textbf{maxFeePerGas}, $T_m$, in the case of type 2 transactions, or \textbf{gasPrice}, $T_p$, in the case of type 0 and type 1 transactions, is greater than or equal to the block's base fee, $H_{\mathrm{f}}$.
 \end{enumerate}
@@
 \boldsymbol{\sigma}^*[S(T)]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[S(T)]_{\mathrm{b}} + g^* p \\
-\boldsymbol{\sigma}^*[m]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[m]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) f \\
-m & \equiv & {B_{\mathrm{H}}}_{\mathrm{c}}
+\boldsymbol{\sigma}^*[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{\mathrm{P}}[{B_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + (T_{\mathrm{g}} - g^*) f
 \end{eqnarray}

### 2023-12-07 Paper.tex
add the condition that maxPriorityFee must be less than or equal to maxFeePerGas

@@
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment; and
-\item the \textbf{maxFeePerGas}, $T_m$, in the case of type 2 transactions, or \textbf{gasPrice}, $T_p$, in the case of type 0 and type 1 transactions, is greater than or equal to the block's base fee, $H_{\mathrm{f}}$.
+\item the sender account balance contains at least the cost, $v_0$, required in up-front payment;
+\item the \textbf{maxFeePerGas}, $T_m$, in the case of type 2 transactions, or \textbf{gasPrice}, $T_p$, in the case of type 0 and type 1 transactions, is greater than or equal to the block's base fee, $H_{\mathrm{f}}$; and
+\item for type 2 transactions, \textbf{maxPriorityFeePerGas}, $T_f$, must be no larger than \textbf{maxFeePerGas}, $T_m$.
 \end{enumerate}

### 2023-12-07 Paper.tex
incorporate Paris EIPs

--- 2023-11-29 Paper.tex
@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{London version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{London version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Paris version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Paris version \YellowPaperVersionNumber}}}
 \author{
@@
 B & \equiv & (..., (T_0, T_1, ...), ...) \\
-\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
-\end{eqnarray}
-
-Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...
+\end{eqnarray}
+
+Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 
-The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}.
-This process is described in detail in section \ref{ch:ghost}.
+Since the \textit{Paris} hard fork, reaching consensus on new blocks is managed by a protocol called the \textit{Beacon Chain}.
+It is known as the \textit{consensus layer} of Ethereum, and it defines the rules for determining the canonical history of Ethereum blocks.
+This document describes the \textit{execution layer} of Ethereum.
+The execution layer defines the rules for interacting with and updating the state of the Ethereum virtual machine.
+The consensus layer is described in greater detail in the \href{https://github.com/ethereum/consensus-specs}{consensus specifications}.
+How the consensus layer is used to determine the canonical state of Ethereum is discussed in section \ref{ch:blocktree_to_blockchain}.
 
 Sometimes, a path follows a new protocol from a particular height (block number).
-This document describes the \textit{London}, \textit{Arrow Glacier}, and \textit{Gray Glacier} versions of the protocol.\footnote{Note that the \textit{Arrow Glacier} and \textit{Gray Glacier} forks only affected the \hyperlink{difficulty_bomb}{"difficulty bomb"} described in section \ref{subsec:The_Block}.}
+This document describes the \textit{Paris} version.
+
 In order to follow back the history of a path, one must reference multiple versions of this document.
-Here are the block numbers of protocol updates on the Ethereum main network:
+Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris fork activated at block 15,537,393, the trigger was not the block number, but rather reaching a specified \textit{total difficulty}.
+The trigger for the \textit{Paris} hard fork is discussed in greater detail in section \ref{ch:pos_transition}.}
 \par
@@
 $F_{\mathrm{Gray Glacier}}$      & 15050000 \\
+$F_{\mathrm{Paris}}$      & 15537393 \\
 \bottomrule
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
 The block header contains several pieces of information:
@@
 \item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
-\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
+\item[ommersHash] A 256-bit hash field that is now deprecated due the replacement of proof of work consensus. It is now to a constant, \texttt{Keccak256(RLP([]))}; formally $H_{\mathrm{o}}$.
+\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
 \item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
@@
 \item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
-\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
+\item[prevRandao]\linkdest{prevRandao_H__a}{}\linkdest{H__a} the latest RANDAO mix\footnote{RANDAO is a pseudorandom value generated by validators on the Ethereum consensus layer.
+Refer to the consensus layer specs (\url{https://github.com/ethereum/consensus-specs}) for more detail on RANDAO.} of the post beacon state of the previous block; formally $H_{\mathrm{a}}$.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacment of proof of work consensus.
+It is set to \texttt{0x0000000000000000}; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
 \item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are a series of transactions, $B_{\mathbf{T}}$, and an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$.
 Formally, we can refer to a block $B$:
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: its ommers field $B_{\mathbf{U}}$ must be an empty array and it must be internally consistent with the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
 \begin{equation}
 \begin{array}[t]{lclc}
+B_{\mathbf{U}} &\equiv& [] & \wedge \\
 \linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
-\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
 \linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
@@
 \begin{eqnarray}
-\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
 \quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
@@
 \hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
-\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
+\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{a}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
 \hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N}
@@
 \newcommand{\diffadjustment}{x}
-
-\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
-\begin{equation}
-D(H) \equiv \begin{dcases}
-2^{34} & \text{if} \quad H_{\mathrm{i}} = 0\\
-\text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
-\end{dcases}
-\end{equation}
-where:
-\begin{align}
-\mindifficulty &\equiv 2^{17} \\
-\diffadjustment &\equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor \\
-\homesteadmod &\equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right) \\
-y &\equiv \begin{cases}
-1 & \text{if} \quad \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
-2 & \text{otherwise}
-\end{cases} \\
-\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
-H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0)
-\end{align}
-\begin{equation}
-\kappa \equiv \begin{cases} 
-  3000000  & \text{if} \quad F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
-  5000000  & \text{if} \quad F_{\mathrm{Constantinople}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Muir Glacier}} \\
-  9000000  & \text{if} \quad F_{\mathrm{Muir Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{London}} \\
-  9700000  & \text{if} \quad F_{\mathrm{London}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Arrow Glacier}} \\
- 10700000  & \text{if} \quad F_{\mathrm{Arrow Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Gray Glacier}} \\
- 11400000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Gray Glacier}} \\
-\end{cases}
-\end{equation}
-
-\hypertarget{difficulty_bomb}{}The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
-In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
-This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
-$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
-In the \textit{Byzantium} release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number,
-which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
-Subsequently, EIP-1234 by \cite{EIP-1234}, EIP-2384 by \cite{EIP-2384}, EIP-3554 by \cite{EIP-3554}, EIP-4345 by \cite{EIP-4345}, and EIP-5133 by \cite{EIP-5133} increased the subtrahend $\kappa$.
 
@@
 
-This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period.
-Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.
-
-The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
-\begin{equation}
-n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
-\end{equation}
-with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.
-
-\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.
-
-This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propagate newly created blocks that would otherwise overwrite (``rewrite'') history.
-Because the nonce must satisfy this requirement, and because its satisfaction depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.
+The \textit{Paris} hard fork changed Ethereum's consensus from proof of work to proof of stake, and thus deprecated many block header properties related to proof of work.
+These deprecated properties include \textbf{nonce} ($H_{\mathbf{n}}$), \textbf{ommersHash} ($H_{\mathbf{o}}$), \textbf{difficulty} ($H_{\mathbf{d}}$), and \textbf{mixHash} ($H_{\mathbf{m}}$).
+
+\textbf{mixHash} has been replaced with a new field \textbf{prevRandao} ($H_{\mathbf{a}}$). The other header fields related to proof of work have been replaced with constants:
+
+\begin{eqnarray}
+  H_{\mathrm{o}} & \equiv & \mathtt{KEC}( \mathtt{RLP}([])) \\
+  H_{\mathrm{d}} & \equiv  & 0 \\
+  H_{\mathrm{n}}  & \equiv & \texttt{0x0000000000000000}
+\end{eqnarray}
+
+The value of \textbf{prevRandao} must be determined using information from the Beacon Chain.
+While the details of generating the \textit{RANDAO} value on the Beacon Chain is beyond the scope of this paper, we refer to the expected \textit{RANDAO} value for the previous block as $\mathtt{PREVRANDAO}()$.
 
@@
 \begin{eqnarray}
-V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
-\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
-\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
+V(H) & \equiv & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
 \nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}'} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}'}}{1024}\right\rfloor  \quad \wedge \\
@@
 \nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32 \quad \wedge \\
-\nonumber& & H_{\mathrm{f}} = F(H)
-\end{eqnarray}
-where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$
+\nonumber& & H_{\mathrm{f}} = F(H) \wedge \\
+\nonumber& & H_{\mathrm{o}} = \mathtt{KEC}( \mathtt{RLP}([]) ) \quad \wedge \\
+\nonumber& & H_{\mathrm{d}} = 0 \quad \wedge \\
+\nonumber& & H_{\mathrm{n}} = \texttt{0x0000000000000000} \quad \wedge \\
+\nonumber& & H_{\mathrm{a}} = \mathtt{PREVRANDAO}() \\
+\end{eqnarray}
 
@@
 The ether that is paid to meet the base fee is burned (taken out of circulation).
-The base fee adjusts dynamically as a function of the previous block's gas consumption relative to its \textit{gas target} (a value that is currently half the block's gas limit, which can be adjusted by miners).
+The base fee adjusts dynamically as a function of the previous block's gas consumption relative to its \textit{gas target} (a value that is currently half the block's gas limit, which can be adjusted by validators).
 If the previous block's total gas consumption exceeds the gas target, this indicates excess demand for block space at the current base fee, and the base fee is increased.
@@
 
-To incentivize miners to include transactions, there is an additional fee known as a \textit{priority fee}, which is also specified as wei per unit of gas consumed.
+To incentivize validators to include transactions, there is an additional fee known as a \textit{priority fee}, which is also specified as wei per unit of gas consumed.
 The total fee fee paid by the transactor therefore is the sum of the base fee per gas and the priority fee per gas multiplied by the total gas consumed.
-Ether used to satisfy the priority fee is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the miner.
+Ether used to satisfy the priority fee is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the validator.
 
@@
 
-Transactors are free to select any priority fee that they wish, however miners are free to ignore transactions as they choose.
-A higher priority fee on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the miner and thus will more likely be selected for inclusion.
-Since there will be a (weighted) distribution of minimum acceptable priority fees, transactors will necessarily have a trade-off to make between lowering the priority fee and maximising the chance that their transaction will be mined in a timely manner.
+Transactors are free to select any priority fee that they wish, however validators are free to ignore transactions as they choose.
+A higher priority fee on a transaction will therefore cost the sender more in terms of Ether and deliver a greater value to the validator and thus will more likely be selected for inclusion.
+Since there will be a (weighted) distribution of minimum acceptable priority fees, transactors will necessarily have a trade-off to make between lowering the priority fee and maximising the chance that their transaction will be included in a block in a timely manner.
 
@@
 
-The miner, whose address is specified as the beneficiary of the present block $B$, receives the gas consumed multiplied by the transaction's \textit{priority fee per gas}, defined as \hyperlink{priority_fee_f}{$f$} in this section.
+The validator, whose address is specified as the beneficiary of the present block $B$, receives the gas consumed multiplied by the transaction's \textit{priority fee per gas}, defined as \hyperlink{priority_fee_f}{$f$} in this section.
 The ether that is paid by the transactor that goes toward the base fee is debited from the transactor's account but credited to no other account, so it is burned.
@@
 
-These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{sigma_n}{later} for state and nonce validation.
+These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{sigma_n}{later} for state validation.
 
@@
 
-\section{Blocktree to Blockchain} \label{ch:ghost}
-
-The canonical blockchain is a path from root to leaf through the entire block tree.
-In order to have consensus over which path it is, conceptually we identify the path that has had the most computation done upon it, or, the \textit{heaviest} path.
-Clearly one factor that helps determine the heaviest path is the block number of the leaf, equivalent to the number of blocks, not counting the unmined \hyperlink{Genesis_Block}{genesis block}, in the path.
-The longer the path, the greater the total mining effort that must have been done in order to arrive at the leaf.
-This is akin to existing schemes, such as that employed in Bitcoin-derived protocols.
-
-Since a block header includes the difficulty, the header alone is enough to validate the computation done. Any block contributes toward the total computation or \textit{total difficulty} of a chain.
-
-Thus we define the total difficulty of block $B$ recursively as:
-\begin{eqnarray}
-B_{\mathrm{t}} & \equiv & B'_{\mathrm{t}} + B_{\mathrm{d}} \\
-B' & \equiv & P(B_{\mathrm{H}})
-\end{eqnarray}
-
-As such given a block $B$, $B_{\mathrm{t}}$ is its total difficulty, $B'$ is its parent block and $B_{\mathrm{d}}$ is its difficulty.
+\section{Transition to Proof of Stake} \label{ch:pos_transition}
+The \textit{Paris} hard fork changed the underlying consensus mechanism of Ethereum from proof of work to proof of stake.
+
+Unlike all previous hard forks of Ethereum, \textit{Paris} was not defined to occur at a particular block height, but rather at a \textit{terminal total difficulty}.
+Total difficulty was used to avoid a scenario in which a minority of hash power could create a malicious fork that could race to satisfy the block height requirement and claim the first proof of stake block.
+
+Thus the \textit{terminal block}, the last proof of work block before the \textit{Paris} fork takes effect, is defined as having:
+\begin{eqnarray}
+  B_{\mathrm{t}} & \geqslant & 58750000000000000000000 \\
+  P(B_{\mathrm{H}})_{\mathrm{t}} & < & 58750000000000000000000
+\end{eqnarray}
+
+where $B_{\mathrm{t}}$ is the total difficulty of block $B$ and $P(B_{\mathrm{H}})_{\mathrm{t}}$ is the total difficulty of its parent.
+
+Total difficulty for a proof of work (pre-\textit{Paris}) block $B$ is defined recursively as:
+\begin{eqnarray}
+B_{\mathrm{t}} & \equiv & P(B_{\mathrm{H}})_{\mathrm{t}} + H_{\mathrm{d}}
+\end{eqnarray}
+
+where $H_{\mathrm{d}}$ is the difficulty of the current block $B$.
+
+Upon reaching the terminal block, new blocks are processed by the Beacon Chain.
+
+\section{Blocktree to Blockchain} \label{ch:blocktree_to_blockchain}
+
+Prior to the transition to proof of stake at the \textit{Paris} hard fork, the canonical blockchain was defined as the block path with the greatest \textit{total difficulty}, defined in section \ref{ch:pos_transition} as $B_{\mathrm{t}}$.
+
+After reaching the \textit{terminal block} described in section \ref{ch:pos_transition}, the greatest \textit{total difficulty} rule must be removed in favor of a rule known as \textit{LMD Ghost}\footnote{\textit{LMD GHOST} comprises two acronyms, "Latest Message Driven", and "Greedy Heaviest-Observed Sub-Tree".}.
+
+Note that in order to determine what blocks comprise the canonical Ethereum blockchain after \textit{Paris}, one must have additional information from the beacon chain, which is not described herein.
+We denote events emitted by the beacon chain with the prefix $\mathtt{POS\_}$.
+
+On each occurence of a $\mathtt{POS\_FORKCHOICE\_UPDATED}$ event, starting with the first at the \textit{transition block} described in section \ref{ch:pos_transition}, the canonical chain is defined as the chain beginning with the genesis block and ending at the block nominated by the event as the head of the chain.
+
+The head of the chain should be updated if and only if a $\mathtt{POS\_FORKCHOICE\_UPDATED}$ is emitted, in which case the head should set to the block specified by the event.
+No optimistic updates to the head of the chain should be made.
+
+The $\mathtt{POS\_FORKCHOICE\_UPDATED}$ event additionally references a finalized block.
+The most recent finalized block should be set to this block.
+
+The canonical blockchain must also contain a block with the hash and number of the \textit{terminal block} defined in section \ref{ch:pos_transition}.
 
@@
 
-The process of finalising a block involves four stages:
+The process of finalising a block involves two stages:
 
 \begin{enumerate}
-\item Validate (or, if mining, determine) ommers;
-\item validate (or, if mining, determine) transactions;
-\item apply rewards;
-\item verify (or, if mining, compute a valid) state and \hyperlink{block_nonce}{block nonce}.
+\item validate transactions;
+\item verify state.
 \end{enumerate}
 
-\subsection{Ommer Validation}
-
-The validation of \hyperlink{ommer_block_headers_B__U}{ommer headers} means nothing more than verifying that each ommer header is both a valid header and satisfies the relation of $N$th-generation ommer to the present block where $N \leq 6$.
-The maximum of ommer headers is two.
-Formally:
-\begin{equation}
-\lVert B_{\mathbf{U}} \rVert \leqslant 2 \bigwedge_{\mathbf{U} \in B_{\mathbf{U}}} \hyperlink{block_header_validity_function}{V({\mathbf{U}}})\; \wedge \; k({\mathbf{U}}, P(\mathbf{B}_{\mathbf{H}})_{\mathbf{H}}, 6)
-\end{equation}
-
-where $k$ denotes the ``is-kin'' property:
-\begin{equation}
-k(U, H, n) \equiv \begin{cases} \mathit{false} & \text{if} \quad n = 0 \\
-s(U, H) &\\
-\quad \vee \; k(U, P(H)_{\mathrm{H}}, n - 1) & \text{otherwise}
-\end{cases}
-\end{equation}
-
-and $s$ denotes the ``is-sibling'' property:
-\begin{equation}
-s(U, H) \equiv (P(H) = P(U)\; \wedge \; H \neq U \; \wedge \; U \notin B(H)_{\mathbf{U}})
-\end{equation}
-where $B(H)$ and $P(H)$ are the block and the parent block of the corresponding header $H$ respectively.
-
 \subsection{Transaction Validation}
@@
 
-\subsection{Reward Application}
-
-The application of rewards to a block involves raising the balance of the accounts of the beneficiary address of the block and each ommer by a certain amount.
-We raise the block's beneficiary account by \hyperlink{R__block}{$R_{\mathrm{block}}$}; for each ommer, we raise the block's beneficiary by an additional $\frac{1}{32}$ of the block reward and the beneficiary of the ommer gets rewarded depending on the block number.\hypertarget{Omega}{} Formally we define the function $\Omega$:
-\begin{eqnarray}
-\\ \nonumber
-\Omega(B, \boldsymbol{\sigma}) & \equiv & \boldsymbol{\sigma}': \boldsymbol{\sigma}' = \boldsymbol{\sigma} \quad \text{except:} \\
-\qquad\boldsymbol{\sigma}'[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} & = & \boldsymbol{\sigma}[{\mathbf{B}_{\mathrm{H}}}_{\mathrm{c}}]_{\mathrm{b}} + \left(1 + \frac{\lVert \mathbf{B}_{\mathbf{U}}\rVert}{32}\right)R_{\mathrm{block}} \\
-\\ \nonumber
-\forall \mathbf{U} \in \mathbf{B}_{\mathbf{U}}: \\ \nonumber
-\boldsymbol{\sigma}'[\mathbf{U}_{\mathrm{c}}] & = & \begin{cases}
-\varnothing &\text{if}\ \boldsymbol{\sigma}[\mathbf{U}_{\mathrm{c}}] = \varnothing\ \wedge\ R = 0 \\
-\mathbf{a}' &\text{otherwise}
-\end{cases} \\
-\mathbf{a}' &\equiv& (\boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{n}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{b}} + R, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathbf{s}}, \boldsymbol{\sigma}[U_{\mathrm{c}}]_{\mathrm{c}}) \\
-R & \equiv & \left(1 + \frac{1}{8} (U_{\mathrm{i}} - {B_{\mathrm{H}}}_{\mathrm{i}})\right) R_{\mathrm{block}}
-\end{eqnarray}
-
-If there are collisions of the beneficiary addresses between ommers and the block (i.e.
-two ommers with the same beneficiary address or an ommer with the same beneficiary address as the present block), additions are applied cumulatively.
-
-\hypertarget{block_reward_R__block}{}\linkdest{R__block}We define the block reward in Wei:
-\begin{equation}
-R_{\mathrm{block}} = 10^{18} \times \begin{cases}
-5 &\text{if}\ H_{\mathrm{i}} < F_{\mathrm{Byzantium}} \\
-3 &\text{if}\ F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
-2 &\text{if}\ H_{\mathrm{i}} \geqslant F_{\mathrm{Constantinople}} \\
-\end{cases} \\
-\end{equation}
-
-\subsection{State \& Nonce Validation}\label{sec:statenoncevalidation}
+\subsection{State Validation}\label{sec:statenoncevalidation}
 
@@
 \begin{eqnarray}
-\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
-B'_{\mathrm{n}} & = & n: \quad x \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \\
-B'_{\mathrm{m}} & = & m \quad \text{with } (x, m) = \mathtt{PoW}(B^*_{\hyperlink{H_cancel_n}{\hcancel{n}}}, n, \mathbf{d}) \\
-B^* & \equiv & B \quad \text{except:} \quad {\hyperlink{Transaction_Receipt}{B^*_{\mathrm{r}}}} = \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
-\end{eqnarray}
-
-With $\mathbf{d}$ being a dataset as specified in Appendix \ref{app:ethash}.
-
-As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Omega}{$\Omega$}, the block finalisation function and \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function, both now well-defined.
+\Phi(B) & \equiv & B': \quad B' = B \quad \text{except:} \\
+{\hyperlink{Transaction_Receipt}{{B'_{\mathrm{H}}}_{\mathrm{r}}}} & = & \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+\end{eqnarray}
+
+As specified at the beginning of the present work, \hyperlink{Pi}{$\Pi$} is the state-transition function, which is defined in terms of \hyperlink{Upsilon_state_transition}{$\Upsilon$}, the transaction-evaluation function.
 
@@
 
-\hypertarget{Pi}{}Finally, we define $\Pi$ as the new state given the block reward function $\hyperlink{Omega}{\Omega}$ applied to the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
-\begin{equation}
-\Pi(\boldsymbol{\sigma}, B) \equiv \hyperlink{Omega}{\Omega}(B, \ell(\boldsymbol{\sigma}))
-\end{equation}
-
-Thus the complete block-transition mechanism is defined, except for $\mathtt{PoW}$, the proof-of-work function.
-
-\subsection{Mining Proof-of-Work} \label{ch:pow}
-
-The mining proof-of-work (PoW) exists as a cryptographically secure nonce that proves beyond reasonable doubt that a particular amount of computation has been expended in the determination of some token value $n$.
-It is utilised to enforce the blockchain security by giving meaning and credence to the notion of difficulty (and, by extension, total difficulty).
-However, since mining new blocks comes with an attached reward, the proof-of-work not only functions as a method of securing confidence that the blockchain will remain canonical into the future, but also as a wealth distribution mechanism.
-
-For both reasons, there are two important goals of the proof-of-work function; firstly, it should be as accessible as possible to as many people as possible.
-The requirement of, or reward from, specialised and uncommon hardware should be minimised.
-This makes the distribution model as open as possible, and, ideally, makes the act of mining a simple swap from electricity to Ether at roughly the same rate for anyone around the world.
-
-Secondly, it should not be possible to make super-linear profits, and especially not so with a high initial barrier.
-Such a mechanism allows a well-funded adversary to gain a troublesome amount of the network's total mining power and as such gives them a super-linear reward (thus skewing distribution in their favour) as well as reducing the network security.
-
-One plague of the Bitcoin world is ASICs.
-These are specialised pieces of compute hardware that exist only to do a single task (\cite{ASICSmithMJS}).
-In Bitcoin's case the task is the SHA256 hash function (\cite{Courtois2014}).
-While ASICs exist for a proof-of-work function, both goals are placed in jeopardy.
-Because of this, a proof-of-work function that is ASIC-resistant (i.e.
-difficult or economically inefficient to implement in specialised compute hardware) has been identified as the proverbial silver bullet.
-
-Two directions exist for ASIC resistance; firstly make it sequential memory-hard, i.e.
-engineer the function such that the determination of the \hyperlink{block_nonce}{nonce} requires a lot of memory and bandwidth such that the memory cannot be used in parallel to discover multiple nonces simultaneously.
-The second is to make the type of computation it would need to do general-purpose; the meaning of ``specialised hardware''  for a general-purpose task set is, naturally, general purpose hardware and as such commodity desktop computers are likely to be pretty close to ``specialised hardware'' for the task.
-For Ethereum 1.0 we have chosen the first path.
-
-More formally, the proof-of-work function takes the form of $\mathtt{PoW}$:
-\begin{equation}
-m = \hyperlink{H__m}{H_{\mathrm{m}}} \quad \wedge \quad n \leqslant \frac{2^{256}}{\hyperlink{H__d}{H_{\mathrm{d}}}} \quad \text{with} \quad (m, n) = \mathtt{PoW}(\hyperlink{H_cancel_n}{H_{\hcancel{n}}}, H_{\mathrm{n}}, \mathbf{d})
-\end{equation}
-
-Where $H_{\hcancel{n}}$ is the new block's header but \textit{without} the nonce and mix-hash components; $H_{\mathrm{n}}$ is the nonce of the header; $\mathbf{d}$ is a large data set needed to compute the mixHash and $H_{\mathrm{d}}$ is the new block's difficulty value (i.e.
-the block difficulty from section \ref{ch:ghost}).
-$\mathtt{PoW}$ is the proof-of-work function which evaluates to an array with the first item being the mixHash and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$.
-The underlying algorithm is called Ethash and is described below.
-\subsubsection{Ethash}
-Ethash is the PoW algorithm for Ethereum 1.0.
-It is the latest version of Dagger-Hashimoto, introduced by \cite{dagger} and \cite{hashimoto}, although it can no longer appropriately be called that since many of the original features of both algorithms were drastically changed with R\&D from February 2015 until May 4 2015 (\cite{commitdateforEthash}).
-The general route that the algorithm takes is as follows:
-
-There exists a seed which can be computed for each block by scanning through the block headers up until that point.
-From the seed, one can compute a pseudorandom cache, $\hyperlink{J__cacheinit}{J_{\mathrm{cacheinit}}}$ bytes in initial size.
-Light clients store the cache.
-From the cache, we can generate a dataset, $\hyperlink{J__datasetinit}{J_{\mathrm{datasetinit}}}$ bytes in initial size, with the property that each item in the dataset depends on only a small number of items from the cache.
-Full clients and miners store the dataset.
-The dataset grows linearly with time.
-
-Mining involves grabbing random slices of the dataset and hashing them together.
-Verification can be done with low memory by using the cache to regenerate the specific pieces of the dataset that you need, so you only need to store the cache.
-The large dataset is updated once every $\hyperlink{J__epoch}{J_{\mathrm{epoch}}}$ blocks, so the vast majority of a miner's effort will be reading the dataset, not making changes to it.
-The mentioned parameters as well as the algorithm is explained in detail in Appendix \ref{app:ethash}.
+\hypertarget{Pi}{}Finally, we define $\Pi$ as the final transaction's resultant state, $\hyperlink{ell}{\ell}(\boldsymbol{\sigma})$:
+\begin{equation}
+\Pi(\boldsymbol{\sigma}, B) \equiv \ell(\boldsymbol{\sigma})
+\end{equation}
+
+Thus the complete block-transition mechanism (before consensus) is defined.
 
@@
 Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
-In order to make it hard for malicious miners to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
+In order to make it hard for malicious validators to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.
 For a series of such numbers, a trivial solution would be to add some constant amount and hashing the result.
@@
 
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}\}
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}\}
 
@@
 &&&& Note that as of the \textit{London} hard fork, this value no longer \\ 
-&&&& represents what is received by the miner, \\
+&&&& represents what is received by the validator, \\
 &&&& but rather just what is paid by the sender. \\
@@
 \midrule
-0x44 & {\small DIFFICULTY} & 0 & 1 & Get the current block's difficulty. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{d}}$ \\
+0x44 & {\small PREVRANDAO} & 0 & 1 & Get the latest RANDAO mix of the post beacon state of the previous block. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv {I_{\mathrm{H}}}_{\mathrm{a}}$ \\
 \midrule

### 2023-12-15 Paper.tex
wording, and remove extra line in equation

@@
 \nonumber& & H_{\mathrm{n}} = \texttt{0x0000000000000000} \quad \wedge \\
-\nonumber& & H_{\mathrm{a}} = \mathtt{PREVRANDAO}() \\
+\nonumber& & H_{\mathrm{a}} = \mathtt{PREVRANDAO}()
 \end{eqnarray}
@@
 
-Unlike all previous hard forks of Ethereum, \textit{Paris} was not defined to occur at a particular block height, but rather at a \textit{terminal total difficulty}.
-Total difficulty was used to avoid a scenario in which a minority of hash power could create a malicious fork that could race to satisfy the block height requirement and claim the first proof of stake block.
+Unlike all previous hard forks of Ethereum, \textit{Paris} was not defined to occur at a particular block height, but rather after a specified \textit{terminal total difficulty} was reached.
+Total difficulty was used instead of block height to avoid a scenario in which a minority of hash power could create a malicious fork that could race to satisfy the block height requirement and claim the first proof of stake block.
 

### 2023-12-18 Paper.tex
address comments round 1

@@
 
-Sometimes, a path follows a new protocol from a particular height (block number).
+There are many versions of Ethereum, as the protocol has undergone a number of updates.
+These updates can be specified to occur:
+\begin{itemize}
+  \item at a particular block number in the case of pre-\textit{Paris} updates,
+  \item after reaching a \textit{terminal total difficulty} in the case of the \textit{Paris} update, or
+  \item at a particular block timestamp in the case of post-\textit{Paris} updates.
+\end{itemize}
 This document describes the \textit{Paris} version.
@@
 $F_{\mathrm{Gray Glacier}}$      & 15050000 \\
-$F_{\mathrm{Paris}}$      & 15537393 \\
+$F_{\mathrm{Paris}}$             & 15537394 \\
 \bottomrule
@@
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
 The block header contains several pieces of information:
@@
 \item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] A 256-bit hash field that is now deprecated due the replacement of proof of work consensus. It is now to a constant, \texttt{Keccak256(RLP([]))}; formally $H_{\mathrm{o}}$.
+\item[ommersHash] A 256-bit hash field that is now deprecated due the replacement of proof of work consensus. It is now to a constant, $\texttt{KEC}(\texttt{RLP}(()))$; formally $H_{\mathrm{o}}$.
 \item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
@@
 \begin{eqnarray}
-  H_{\mathrm{o}} & \equiv & \mathtt{KEC}( \mathtt{RLP}([])) \\
+  H_{\mathrm{o}} & \equiv & \texttt{KEC}(\texttt{RLP}(())) \\
   H_{\mathrm{d}} & \equiv  & 0 \\
@@
 \nonumber& & H_{\mathrm{f}} = F(H) \wedge \\
-\nonumber& & H_{\mathrm{o}} = \mathtt{KEC}( \mathtt{RLP}([]) ) \quad \wedge \\
+\nonumber& & H_{\mathrm{o}} = \texttt{KEC}(\texttt{RLP}(())) \quad \wedge \\
 \nonumber& & H_{\mathrm{d}} = 0 \quad \wedge \\
@@
 However, we can approximate with pseudo-random numbers by utilising data which is generally unknowable at the time of transacting.
-Such data might include the block's hash, the block's timestamp and the block's beneficiary address.
+Such data might include the block's hash and the block's beneficiary address.
 In order to make it hard for malicious validators to control those values, one should use the {\small \hyperlink{blockhash}{BLOCKHASH}} operation in order to use hashes of the previous 256 blocks as pseudo-random numbers.

### 2023-12-21 Paper.tex
grammar edit
address comments round 2

@@
 \item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
-\item[ommersHash] A 256-bit hash field that is now deprecated due the replacement of proof of work consensus. It is now to a constant, $\texttt{KEC}(\texttt{RLP}(()))$; formally $H_{\mathrm{o}}$.
+\item[ommersHash] A 256-bit hash field that is now deprecated due to the replacement of proof of work consensus. It is now to a constant, $\texttt{KEC}(\texttt{RLP}(()))$; formally $H_{\mathrm{o}}$.
 \item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: its ommers field $B_{\mathbf{U}}$ must be an empty array and it must be internally consistent with the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$.
+For the header to be consistent with the transactions $B_{\mathbf{T}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}), and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), and \textbf{logsBloom} ($H_{\mathrm{b}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, and the resulting logs, respectively.
 \begin{equation}
 \begin{array}[t]{lclc}
-B_{\mathbf{U}} &\equiv& [] & \wedge \\
+B_{\mathbf{U}} &\equiv& () & \wedge \\
 \linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\

### 2024-01-08 Paper.tex
fix typo

@@
 Refer to the consensus layer specs (\url{https://github.com/ethereum/consensus-specs}) for more detail on RANDAO.} of the post beacon state of the previous block; formally $H_{\mathrm{a}}$.
-\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacment of proof of work consensus.
+\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacement of proof of work consensus.
 It is set to \texttt{0x0000000000000000}; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.

### 2024-01-09 Paper.tex
fix state root calculation in the block finalization section
fix Paris fork block number in footnote

@@
 In order to follow back the history of a path, one must reference multiple versions of this document.
-Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris fork activated at block 15,537,393, the trigger was not the block number, but rather reaching a specified \textit{total difficulty}.
+Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris fork activated at block 15,537,394, the trigger was not the block number, but rather reaching a specified \textit{total difficulty}.
 The trigger for the \textit{Paris} hard fork is discussed in greater detail in section \ref{ch:pos_transition}.}
@@
 \Phi(B) & \equiv & B': \quad B' = B \quad \text{except:} \\
-{\hyperlink{Transaction_Receipt}{{B'_{\mathrm{H}}}_{\mathrm{r}}}} & = & \hyperlink{r}{r}(\hyperlink{Pi}{\Pi}(\Gamma(B), B))
+{B'_{\mathrm{H}}}_{\mathrm{r}} & = & \mathtt{TRIE}(L_{\mathrm{S}}(\hyperlink{Pi}{\Pi}(\Gamma(B), B)))
 \end{eqnarray}

### 2024-01-09 Paper.tex
Incorporate Paris EIPs (#894)
* incorporate Paris EIPs

* wording, and remove extra line in equation

* address comments round 1

* grammar edit

* address comments round 2

* fix typo

* fix state root calculation in the block finalization section

* fix Paris fork block number in footnote

@@
 \item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
-\item the sender account balance contains at least the cost, $v_0$, required in up-front payment; and
-\item the effective gas price of the transaction is greater than or equal to the block's base fee $H_{\mathrm{f}}$.
+\item the sender account balance contains at least the cost, $v_0$, required in up-front payment;
+\item the \textbf{maxFeePerGas}, $T_m$, in the case of type 2 transactions, or \textbf{gasPrice}, $T_p$, in the case of type 0 and type 1 transactions, is greater than or equal to the block's base fee, $H_{\mathrm{f}}$; and
+\item for type 2 transactions, \textbf{maxPriorityFeePerGas}, $T_f$, must be no larger than \textbf{maxFeePerGas}, $T_m$.
 \end{enumerate}

### 2024-01-10 Paper.tex
grammar edits

@@
 
-After reaching the \textit{terminal block} described in section \ref{ch:pos_transition}, the greatest \textit{total difficulty} rule must be removed in favor of a rule known as \textit{LMD Ghost}\footnote{\textit{LMD GHOST} comprises two acronyms, "Latest Message Driven", and "Greedy Heaviest-Observed Sub-Tree".}.
-
-Note that in order to determine what blocks comprise the canonical Ethereum blockchain after \textit{Paris}, one must have additional information from the beacon chain, which is not described herein.
-We denote events emitted by the beacon chain with the prefix $\mathtt{POS\_}$.
+After reaching the \textit{terminal block} described in section \ref{ch:pos_transition}, the greatest \textit{total difficulty} rule must be removed in favor of a rule known as \textit{LMD Ghost}.\footnote{\textit{LMD GHOST} comprises two acronyms, "Latest Message Driven", and "Greedy Heaviest-Observed Sub-Tree".}
+
+Note that in order to determine what blocks comprise the canonical Ethereum blockchain after \textit{Paris}, one must have additional information from the Beacon Chain, which is not described herein.
+We denote events emitted by the Beacon Chain with the prefix $\mathtt{POS\_}$.
 

### 2024-02-20 Paper.tex
add tx type 2 (#898)
fix: typo

@@
 \begin{array}[t]{lclclc}
-T_{\mathrm{x}} \in \{0, 1\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
+T_{\mathrm{x}} \in \{0, 1, 2\} & \wedge & T_{\mathrm{c}} = \beta & \wedge & T_{\mathrm{n}} \in \mathbb{N}_{256} & \wedge \\
 T_{\mathrm{p}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{g}} \in \mathbb{N}_{256} & \wedge & T_{\mathrm{v}} \in \mathbb{N}_{256} & \wedge \\
@@
 To incentivize validators to include transactions, there is an additional fee known as a \textit{priority fee}, which is also specified as wei per unit of gas consumed.
-The total fee fee paid by the transactor therefore is the sum of the base fee per gas and the priority fee per gas multiplied by the total gas consumed.
+The total fee paid by the transactor therefore is the sum of the base fee per gas and the priority fee per gas multiplied by the total gas consumed.
 Ether used to satisfy the priority fee is delivered to the \textit{beneficiary} address, the address of an account typically under the control of the validator.

### 2024-02-26 Paper.tex
Sentence made clearer and simpler
fix: word typo

@@
 
-A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memoise the function $c$.
+A reasonable implementation will maintain a database of nodes determined from the computation of various tries or, more formally, it will memorise the function $c$.
 This strategy uses the nature of the trie to both easily recall the contents of any previous key-value set and to store multiple such sets in a very efficient manner.

### 2024-03-03 Paper.tex
fix: warm tx recipient (#903)
Update Paper.tex

--- 2024-02-20 Paper.tex
@@
 A^* & \equiv A^0 \quad \text{except} \\
-A^*_{\mathbf{a}} & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \} \\
-A^*_{\mathbf{K}} & \equiv \bigcup_{E \in T_{\mathbf{A}}} \big\{ \forall i < \lVert E_{\mathbf{s}} \rVert, i \in \mathbb{N}: \; (E_{\mathrm{a}}, E_{\mathbf{s}}[i]) \big\}
+A^*_{\mathbf{K}} & \equiv \bigcup_{E \in T_{\mathbf{A}}} \big\{ \forall i < \lVert E_{\mathbf{s}} \rVert, i \in \mathbb{N}: \; (E_{\mathrm{a}}, E_{\mathbf{s}}[i]) \big\} \\
+A^*_{\mathbf{a}} & \equiv \begin{cases}
+a \cup T_{\mathrm{t}} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
+a & \text{otherwise}
+\end{cases} \\
+a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
 \end{align}
@@
 
-On each occurence of a $\mathtt{POS\_FORKCHOICE\_UPDATED}$ event, starting with the first at the \textit{transition block} described in section \ref{ch:pos_transition}, the canonical chain is defined as the chain beginning with the genesis block and ending at the block nominated by the event as the head of the chain.
+On each occurrence of a $\mathtt{POS\_FORKCHOICE\_UPDATED}$ event, starting with the first at the \textit{transition block} described in section \ref{ch:pos_transition}, the canonical chain is defined as the chain beginning with the genesis block and ending at the block nominated by the event as the head of the chain.
 

### 2024-04-13 Paper.tex
eip-3651

@@
 \end{cases} \\
-a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
+a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{H_c\}\cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
 \end{align}

### 2024-04-13 Paper.tex
eip-3855

@@
 
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}\}
-
-$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}, {\small PUSH0}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH1}, ..., {\small PUSH32}, {\small DUP*}, {\small SWAP*}\}
 

### 2024-04-14 Paper.tex
eip-3860

@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{N({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 \nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
@@
 
+\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $N$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:
+
+\begin{equation}
+  N(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
+\end{equation}
+
 \hypertarget{effective_gas_price_p}{}We define the \textit{effective gas price}, formally $p$, as the amount of wei the transaction signer will pay per unit of gas consumed during the transaction's execution.
@@
 m & \geqslant & H_{\mathrm{f}} \quad \wedge \\
+n & \leqslant & 49152 \quad \wedge \\
 T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
@@
 \end{equation}
-
+and
+\begin{equation}
+  n \equiv \begin{cases}
+    \lVert T_{\mathrm{i}} \rVert & \text{if} \; T_{\mathrm{t}} \neq \varnothing \\
+    0 & \text{otherwise}
+  \end{cases}
+\end{equation}
+
+The penultimate condition ensures that, for create transactions, the length of the initcode is no greater than 49152 bytes.
 Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
@@
 \linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
+\linkdest{G__initcodeword}{}$G_{\mathrm{initcodeword}}$ & 2 & Paid per word of the initcode at the beginning of a deploy transaction, a {\small CREATE},\\
+&&or a {\small CREATE2} operation. \\
 $G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
@@
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
-G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{\mathrm{create}} + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
 G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
@@
 &&&& $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\
+\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \wedge \\ \quad &\; I_{\mathrm{e}} < 1024 \wedge \lVert \mathbf{i} \rVert \leqslant 49152\\
 \big(\boldsymbol{\sigma}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if $z = 0$, i.e., the \hyperlink{contract_creation_result}{contract creation process failed}, or $I_{\mathrm{e}} = 1024$ \\
-&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
-&&&& is too low to fulfil the value transfer); and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
-&&&& address of the newly created account (\ref{eq:new-address}). \\
+&&&& where $x=0$ if $z = 0$, i.e., the \hyperlink{contract_creation_result}{contract creation process failed},\\
+&&&& or $\lVert \mathbf{i} \rVert > 49152$ (initcode is longer than 49152 bytes),\\
+&&&& or $I_{\mathrm{e}} = 1024$ (the maximum call depth limit is reached),\\
+&&&& or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too low to fulfil the value transfer);\\
+&&&& and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the address of the newly created account (\ref{eq:new-address}). \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\

### 2024-04-21 Paper.tex
eip-4895

@@
 
-Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
+Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function for transactions\footnote{Note that since the Shanghai fork, blocks also needs to process \textit{withdrawal operations} in order to reach their final state.
+Withdrawal operations are defined in sub-section \ref{subsec:The_Withdrawal}, and block final state is discussed in greater detail in section \ref{ch:finalisation}.}.
 
@@
 
+\subsection{The Withdrawal}\linkdest{withdrawal}\label{subsec:The_Withdrawal}
+A withdrawal (formally, \textit{W}) is a tuple of data describing a consensus layer validator's withdrawal of some amount of its staked Ether.
+A withdrawal is created and validated in the consensus layer of the blockchain and then pushed to the execution layer.
+A withdrawal is composed of the following fields:
+
+\begin{description}
+\item[globalIndex] zero based incrementing withdrawal index that acts as a unique identifier for this withdrawal; formally $W_{\mathrm{g}}$.
+\item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
+\item[recipient] the 20-bytes address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
+\item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
+\end{description}
+
+Withdrawal serialisation is defined as:
+\begin{equation}
+\linkdest{L_Withdrawal} L_{\mathrm{W}}(W) \equiv 
+(W_{\mathrm{g}}, W_{\mathrm{v}}, W_{\mathrm{r}}, T_{\mathrm{a}})
+\end{equation}
+
+Here, we assume all components are interpreted by the RLP as integer values except for $W_\mathbf{r}$ which is a 20-bytes address:
+\begin{equation}
+\begin{array}[t]{lclc}
+W_{\mathrm{g}} \in \mathbb{N}_{64} & \wedge & W_{\mathrm{v}} \in \mathbb{N}_{64} & \wedge \\
+W_{\mathrm{r}} \in \mathbb{B}_{20} & \wedge & W_{\mathrm{a}} \in \mathbb{N}_{64}
+\end{array}
+\end{equation}
+
 \subsection{The Block}\linkdest{block}\label{subsec:The_Block}
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
-The block header contains several pieces of information:
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions,
+$\mathbf{T}$,\hypertarget{ommerheaders}{} a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}),
+and, since the \textit{Shanghai} hard fork, $\mathbf{W}$, a collection of validator's withdrawal pushed by the consensus layer. The block header contains several pieces of information:
 
@@
 \item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions and withdrawals are executed and finalisations applied; formally $H_{\mathrm{r}}$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
@@
 \item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
+\item[withdrawalsRoot]\linkdest{withdrawals_root_H__w}{} The Keccak 256-bit hash of the root node of the trie structure populated with each withdrawal operations pushed by the consensus layer for this block; formally \hyperlink{H__w}{$H_{\mathrm{w}}$}.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are a series of transactions, $B_{\mathbf{T}}$, and an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other three components in the block are a series of transactions, $B_{\mathbf{T}}$, an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$, and a series of withdrawals, $B_{\mathbf{W}}$.
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
+B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}}, B_{\mathbf{W}})
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$.
-For the header to be consistent with the transactions $B_{\mathbf{T}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}), and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), and \textbf{logsBloom} ($H_{\mathrm{b}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, and the resulting logs, respectively.
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions:
+the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$ and withdrawals $B_{\mathbf{W}}$.
+For the header to be consistent with the transactions $B_{\mathbf{T}}$ and withdrawals $B_{\mathbf{W}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions, then all withdrawals, in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}),
+and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), \textbf{logsBloom} ($H_{\mathrm{b}}$) and \textbf{withdrawalsRoot} ($H_{\mathrm{w}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, the resulting logs, and the withdrawals, respectively.
 \begin{equation}
@@
 \linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
+\linkdest{Withdrawals_Root_H__w}{}H_{\mathrm{w}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{W}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{W}}(i, B_{\mathbf{W}}[i])\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \end{equation}
-where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
+
+where $p_{\mathrm{W}}(k, v)$ is a pairwise RLP transformation for withdrawals:
+\begin{equation}
+p_{\mathrm{W}}(k, W) \equiv \left( \mathtt{RLP}(k), \mathtt{RLP}(\hyperlink{L_Withdrawal}{L_{\mathrm{W}}}(W)) \right)
+\end{equation}
+
+similarly, $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
 \begin{equation}
@@
 \begin{eqnarray}
-\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
-\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}}, H_{\mathrm{w}} \; )\end{array} \\
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}), L_{\mathrm{W}}^*(B_{\mathbf{W}}) \big)
 \end{eqnarray}
@@
 \end{equation}
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$, $L_{\mathrm{H}}^*$, and $L_{\mathrm{W}}^*$, being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{a}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
-\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N}
+\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N} & \wedge & \hyperlink{withdrawals_root_H__w}{H_{\mathrm{w}}} \in \mathbb{B}_{32}
 
@@
 \begin{enumerate}
+\item executing withdrawals;
 \item validate transactions;
@@
 
+\subsection{Executing Withdrawals}
+After processing the block's transactions, the withdrawals are executed.
+A withdrawal is simply an increase of the recipient account's balance of the specified amount.
+No other balances are decreased, a withdrawal is not a transfer but a creation of funds.
+A withdrawal operation cannot fail and has no gas cost. We define the function $P$ as the withdrawal state transition function:
+
+\begin{eqnarray}
+P(\boldsymbol{\sigma}_w, W) & \equiv & \boldsymbol{\sigma}_{w+1} \\
+\boldsymbol{\sigma}_{w+1} & \equiv & \boldsymbol{\sigma}_w \quad \text{except:} \\
+\boldsymbol{\sigma}_{w+1}[W_r]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{w}[W_r]_{\mathrm{b}} + (W_a \times 10^9)
+\end{eqnarray}
+
+\hypertarget{Kappa}{}Finally, we define $K$, the block-level state transition function for withdrawals:
+
+\begin{equation}
+K(\boldsymbol{\sigma}, B) \equiv P(P(\boldsymbol{\sigma}, W_0), W_1) ...
+\end{equation}
+
 \subsection{Transaction Validation}
@@
 \Phi(B) & \equiv & B': \quad B' = B \quad \text{except:} \\
-{B'_{\mathrm{H}}}_{\mathrm{r}} & = & \mathtt{TRIE}(L_{\mathrm{S}}(\hyperlink{Pi}{\Pi}(\Gamma(B), B)))
+{B'_{\mathrm{H}}}_{\mathrm{r}} & = & \mathtt{TRIE}(L_{\mathrm{S}}(\hyperlink{Kappa}{K}(\hyperlink{Pi}{\Pi}(\Gamma(B), B), B)))
 \end{eqnarray}

### 2024-04-23 Paper.tex
eip-6049
shanghai

@@
 
-\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Paris version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Paris version \YellowPaperVersionNumber}}}
+\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Shanghai version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Shanghai version \YellowPaperVersionNumber}}}
 \author{
@@
 \end{itemize}
-This document describes the \textit{Paris} version.
+This document describes the \textit{Shanghai} version.
 
 In order to follow back the history of a path, one must reference multiple versions of this document.
-Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris fork activated at block 15,537,394, the trigger was not the block number, but rather reaching a specified \textit{total difficulty}.
-The trigger for the \textit{Paris} hard fork is discussed in greater detail in section \ref{ch:pos_transition}.}
+Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris, Shanghai, and every upcoming forks activated at a given block number (e.g.
+15,537,394 for Paris), the trigger was not the block number, but rather reaching a specified \textit{timestamp} (or \textit{total difficulty} for Paris).
+The trigger for the \textit{Paris} and subsequent hard fork are discussed in greater detail in section \ref{ch:pos_transition}.}
 \par
@@
 $F_{\mathrm{Paris}}$             & 15537394 \\
+$F_{\mathrm{Shanghai}}$          & 17034870 \\
 \bottomrule
@@
 Upon reaching the terminal block, new blocks are processed by the Beacon Chain.
+
+\subsection{Post-Paris Updates}
+Because the Beacon Chain generate a new slot every 12 seconds, post-\textit{Paris} updates can be scheduled to occur at a specific timestamp.
+At the execution layer, the update will then happen in the first produced block after the scheduled timestamp.
+For example the \textit{Shanghai} hard fork was scheduled to occur at 2023-04-12 10:27:35 UTC, on Epoch 194,048.
+Validators failed to propose a block during the two first slots of this Epoch, but at slot 6,209,538 a validator finally proposed a the block 17,034,870, marking the transition to \textit{Shanghai} on the execution layer.
 
@@
 \midrule
-\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. \\
+\linkdest{selfdestruct}{}0xff & {\small SELFDESTRUCT} & 1 & 0 & Halt execution and register account for later deletion. Readers should be warned, \\
+&&&& since the \textit{Shanghai} update, this opcode has been marked as deprecated. \\
+&&&& Its behavior might be subject to change in an upcoming update. \\
 &&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\

### 2024-05-25 Paper.tex
eip-3651-alexprut-review
eip-3860-alexprut-review

@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{N({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{R({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 \nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
@@
 
-\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $N$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:
-
-\begin{equation}
-  N(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
+\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $R$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:
+
+\begin{equation}
+  R(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
 \end{equation}
@@
 \end{cases} \\
-a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{H_c\}\cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
+a & \equiv A^0_{\mathbf{a}} \cup S(T) \cup H_c \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
 \end{align}
@@
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{\mathrm{create}} + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
-G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{\mathrm{create}} + \hyperlink{initcode_cost}{R(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{R(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
 G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\

### 2024-06-10 Paper.tex
Fix returndata details

--- 2024-04-21 Paper.tex
@@
 
-Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function for transactions\footnote{Note that since the Shanghai fork, blocks also needs to process \textit{withdrawal operations} in order to reach their final state.
-Withdrawal operations are defined in sub-section \ref{subsec:The_Withdrawal}, and block final state is discussed in greater detail in section \ref{ch:finalisation}.}.
+Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.
 
@@
 
-\subsection{The Withdrawal}\linkdest{withdrawal}\label{subsec:The_Withdrawal}
-A withdrawal (formally, \textit{W}) is a tuple of data describing a consensus layer validator's withdrawal of some amount of its staked Ether.
-A withdrawal is created and validated in the consensus layer of the blockchain and then pushed to the execution layer.
-A withdrawal is composed of the following fields:
-
-\begin{description}
-\item[globalIndex] zero based incrementing withdrawal index that acts as a unique identifier for this withdrawal; formally $W_{\mathrm{g}}$.
-\item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
-\item[recipient] the 20-bytes address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
-\item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
-\end{description}
-
-Withdrawal serialisation is defined as:
-\begin{equation}
-\linkdest{L_Withdrawal} L_{\mathrm{W}}(W) \equiv 
-(W_{\mathrm{g}}, W_{\mathrm{v}}, W_{\mathrm{r}}, T_{\mathrm{a}})
-\end{equation}
-
-Here, we assume all components are interpreted by the RLP as integer values except for $W_\mathbf{r}$ which is a 20-bytes address:
-\begin{equation}
-\begin{array}[t]{lclc}
-W_{\mathrm{g}} \in \mathbb{N}_{64} & \wedge & W_{\mathrm{v}} \in \mathbb{N}_{64} & \wedge \\
-W_{\mathrm{r}} \in \mathbb{B}_{20} & \wedge & W_{\mathrm{a}} \in \mathbb{N}_{64}
-\end{array}
-\end{equation}
-
 \subsection{The Block}\linkdest{block}\label{subsec:The_Block}
 
-The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions,
-$\mathbf{T}$,\hypertarget{ommerheaders}{} a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}),
-and, since the \textit{Shanghai} hard fork, $\mathbf{W}$, a collection of validator's withdrawal pushed by the consensus layer. The block header contains several pieces of information:
+The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers of blocks whose parents were equal to the present block's parent's parent (such blocks were known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}).
+The block header contains several pieces of information:
 
@@
 \item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
-\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions and withdrawals are executed and finalisations applied; formally $H_{\mathrm{r}}$.
+\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
 \item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
@@
 \item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
-\item[withdrawalsRoot]\linkdest{withdrawals_root_H__w}{} The Keccak 256-bit hash of the root node of the trie structure populated with each withdrawal operations pushed by the consensus layer for this block; formally \hyperlink{H__w}{$H_{\mathrm{w}}$}.
 \end{description}
-\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other three components in the block are a series of transactions, $B_{\mathbf{T}}$, an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$, and a series of withdrawals, $B_{\mathbf{W}}$.
+\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are a series of transactions, $B_{\mathbf{T}}$, and an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$.
 Formally, we can refer to a block $B$:
 \begin{equation}
-B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}}, B_{\mathbf{W}})
+B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
 \end{equation}
@@
 
-\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions:
-the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$ and withdrawals $B_{\mathbf{W}}$.
-For the header to be consistent with the transactions $B_{\mathbf{T}}$ and withdrawals $B_{\mathbf{W}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions, then all withdrawals, in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}),
-and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), \textbf{logsBloom} ($H_{\mathrm{b}}$) and \textbf{withdrawalsRoot} ($H_{\mathrm{w}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, the resulting logs, and the withdrawals, respectively.
+\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: the block's ommers field $B_{\mathbf{U}}$ must be an empty array and the block's header must be consistent with the given transactions $B_{\mathbf{T}}$.
+For the header to be consistent with the transactions $B_{\mathbf{T}}$, \textbf{stateRoot} ($H_{\mathrm{r}}$) must match the resultant state after executing all transactions in order on the base state $\boldsymbol{\sigma}$ (as specified in section \ref{ch:finalisation}), and \textbf{transactionsRoot} ($H_{\mathrm{t}}$), \textbf{receiptsRoot} ($H_{\mathrm{e}}$), and \textbf{logsBloom} ($H_{\mathrm{b}}$) must be correctly derived from the transactions themselves, the transaction receipts resulting from execution, and the resulting logs, respectively.
 \begin{equation}
@@
 \linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
-\linkdest{Withdrawals_Root_H__w}{}H_{\mathrm{w}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{W}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{W}}(i, B_{\mathbf{W}}[i])\}) & \wedge \\
 \linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
@@
 \end{equation}
-
-where $p_{\mathrm{W}}(k, v)$ is a pairwise RLP transformation for withdrawals:
-\begin{equation}
-p_{\mathrm{W}}(k, W) \equiv \left( \mathtt{RLP}(k), \mathtt{RLP}(\hyperlink{L_Withdrawal}{L_{\mathrm{W}}}(W)) \right)
-\end{equation}
-
-similarly, $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
+where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
 \begin{equation}
@@
 \begin{eqnarray}
-\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}}, H_{\mathrm{w}} \; )\end{array} \\
-\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}), L_{\mathrm{W}}^*(B_{\mathbf{W}}) \big)
+\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
+\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
 \end{eqnarray}
@@
 \end{equation}
-\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$, $L_{\mathrm{H}}^*$, and $L_{\mathrm{W}}^*$, being element-wise sequence transformations, thus:
+\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
 \begin{equation}
@@
 \hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{a}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
-\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N} & \wedge & \hyperlink{withdrawals_root_H__w}{H_{\mathrm{w}}} \in \mathbb{B}_{32}
+\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N}
 
@@
 g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
-\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{N({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
+\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
 \nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
@@
 
-\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $N$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:
-
-\begin{equation}
-  N(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
-\end{equation}
-
 \hypertarget{effective_gas_price_p}{}We define the \textit{effective gas price}, formally $p$, as the amount of wei the transaction signer will pay per unit of gas consumed during the transaction's execution.
@@
 m & \geqslant & H_{\mathrm{f}} \quad \wedge \\
-n & \leqslant & 49152 \quad \wedge \\
 T_{\mathrm{g}} & \leqslant & {B_{\mathrm{H}}}_{\mathrm{l}} - \hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}
@@
 \end{equation}
-and
-\begin{equation}
-  n \equiv \begin{cases}
-    \lVert T_{\mathrm{i}} \rVert & \text{if} \; T_{\mathrm{t}} \neq \varnothing \\
-    0 & \text{otherwise}
-  \end{cases}
-\end{equation}
-
-The penultimate condition ensures that, for create transactions, the length of the initcode is no greater than 49152 bytes.
+
 Note the final condition; the sum of the transaction's gas limit, $T_{\mathrm{g}}$, and the gas utilised in this block prior, given by $\hyperlink{ell}{\ell}(B_{\mathbf{R}})_{\mathrm{u}}$, must be no greater than the block's \textbf{gasLimit}, ${B_{\mathrm{H}}}_{\mathrm{l}}$.
@@
 \end{cases} \\
-a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{H_c\}\cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
+a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
 \end{align}
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s}, \mathbf{o})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the stack contents, and the returndata buffer.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{o}, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the returndata cache,and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 \begin{enumerate}
-\item executing withdrawals;
 \item validate transactions;
@@
 
-\subsection{Executing Withdrawals}
-After processing the block's transactions, the withdrawals are executed.
-A withdrawal is simply an increase of the recipient account's balance of the specified amount.
-No other balances are decreased, a withdrawal is not a transfer but a creation of funds.
-A withdrawal operation cannot fail and has no gas cost. We define the function $P$ as the withdrawal state transition function:
-
-\begin{eqnarray}
-P(\boldsymbol{\sigma}_w, W) & \equiv & \boldsymbol{\sigma}_{w+1} \\
-\boldsymbol{\sigma}_{w+1} & \equiv & \boldsymbol{\sigma}_w \quad \text{except:} \\
-\boldsymbol{\sigma}_{w+1}[W_r]_{\mathrm{b}} & \equiv & \boldsymbol{\sigma}_{w}[W_r]_{\mathrm{b}} + (W_a \times 10^9)
-\end{eqnarray}
-
-\hypertarget{Kappa}{}Finally, we define $K$, the block-level state transition function for withdrawals:
-
-\begin{equation}
-K(\boldsymbol{\sigma}, B) \equiv P(P(\boldsymbol{\sigma}, W_0), W_1) ...
-\end{equation}
-
 \subsection{Transaction Validation}
@@
 \Phi(B) & \equiv & B': \quad B' = B \quad \text{except:} \\
-{B'_{\mathrm{H}}}_{\mathrm{r}} & = & \mathtt{TRIE}(L_{\mathrm{S}}(\hyperlink{Kappa}{K}(\hyperlink{Pi}{\Pi}(\Gamma(B), B), B)))
+{B'_{\mathrm{H}}}_{\mathrm{r}} & = & \mathtt{TRIE}(L_{\mathrm{S}}(\hyperlink{Pi}{\Pi}(\Gamma(B), B)))
 \end{eqnarray}
@@
 \linkdest{G__codedeposit}{}$G_{\mathrm{codedeposit}}$ & 200 & Paid per byte for a {\small CREATE} operation to succeed in placing code into state. \\
-\linkdest{G__initcodeword}{}$G_{\mathrm{initcodeword}}$ & 2 & Paid per word of the initcode at the beginning of a deploy transaction, a {\small CREATE},\\
-&&or a {\small CREATE2} operation. \\
 $G_{\mathrm{callvalue}}$ & 9000 & Paid for a non-zero value transfer as part of the {\small CALL} operation. \\
@@
 C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
-G_{\mathrm{create}} + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
-G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
+G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
+G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
 G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
@@
 
-$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}, {\small PUSH0}\}
-
-$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH1}, ..., {\small PUSH32}, {\small DUP*}, {\small SWAP*}\}
+$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small PREVRANDAO}, {\small GASLIMIT}, {\small CHAINID}, {\small RETURNDATASIZE}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}, \newline \noindent\hspace*{1cm} {\small BASEFEE}\}
+
+$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small SHL}, {\small SHR}, {\small SAR}, \newline \noindent\hspace*{1cm} {\small CALLDATALOAD}, {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}
 
@@
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
+&&&& reverts if $\quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert$\\
 &&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
+\boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i]$\\
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\
@@
 \toprule
-\multicolumn{5}{c}{\textbf{5f, 60s \& 70s: Push Operations}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{60s \& 70s: Push Operations}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x5f & {\small PUSH0} & 0 & 1 & Place 0 on the stack. \\
-&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv 0$ \\
-\midrule
 0x60 & {\small PUSH1} & 0 & 1 & Place 1 byte item on stack. \\
@@
 &&&& $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \wedge \\ \quad &\; I_{\mathrm{e}} < 1024 \wedge \lVert \mathbf{i} \rVert \leqslant 49152\\
+\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \\ \quad &\wedge\; I_{\mathrm{e}} < 1024\\
 \big(\boldsymbol{\sigma}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\
@@
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv x$ \\
-&&&& where $x=0$ if $z = 0$, i.e., the \hyperlink{contract_creation_result}{contract creation process failed},\\
-&&&& or $\lVert \mathbf{i} \rVert > 49152$ (initcode is longer than 49152 bytes),\\
-&&&& or $I_{\mathrm{e}} = 1024$ (the maximum call depth limit is reached),\\
-&&&& or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller is too low to fulfil the value transfer);\\
-&&&& and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the address of the newly created account (\ref{eq:new-address}). \\
+&&&& where $x=0$ if $z = 0$, i.e., the \hyperlink{contract_creation_result}{contract creation process failed}, or $I_{\mathrm{e}} = 1024$ \\
+&&&& (the maximum call depth limit is reached) or $\boldsymbol{\mu}_{\mathbf{s}}[0] > \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}}$ (balance of the caller\\
+&&&& is too low to fulfil the value transfer); and otherwise $x=\mathtt{ADDR}(I_{\mathrm{a}}, \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}}, \zeta, \mathbf{i} )$, the\\
+&&&& address of the newly created account (\ref{eq:new-address}). \\
 &&&& $\boldsymbol{\mu}'_{\mathrm{i}} \equiv M(\boldsymbol{\mu}_{\mathrm{i}}, \boldsymbol{\mu}_{\mathbf{s}}[1], \boldsymbol{\mu}_{\mathbf{s}}[2])$ \\

### 2024-06-10 Paper.tex
Replace 'cache' with 'buffer'

@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{o}, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the returndata cache,and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{o}, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the returndata cache, and the stack contents.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.

### 2024-06-14 Paper.tex
eip-4895-alexprut-review

--- 2024-05-25 Paper.tex
@@
 
-The process of finalising a block involves two stages:
+The process of finalising a block involves three stages:
 
@@
 After processing the block's transactions, the withdrawals are executed.
-A withdrawal is simply an increase of the recipient account's balance of the specified amount.
+A withdrawal is simply an increase of the recipient account's balance of the specified Gwei amount.
 No other balances are decreased, a withdrawal is not a transfer but a creation of funds.
-A withdrawal operation cannot fail and has no gas cost. We define the function $P$ as the withdrawal state transition function:
-
-\begin{eqnarray}
-P(\boldsymbol{\sigma}_w, W) & \equiv & \boldsymbol{\sigma}_{w+1} \\
+A withdrawal operation cannot fail and has no gas cost. We define the function $E$ as the withdrawal state transition function:
+
+\begin{eqnarray}
+E(\boldsymbol{\sigma}_w, W) & \equiv & \boldsymbol{\sigma}_{w+1} \\
 \boldsymbol{\sigma}_{w+1} & \equiv & \boldsymbol{\sigma}_w \quad \text{except:} \\
@@
 \begin{equation}
-K(\boldsymbol{\sigma}, B) \equiv P(P(\boldsymbol{\sigma}, W_0), W_1) ...
+K(\boldsymbol{\sigma}, B) \equiv E(E(\boldsymbol{\sigma}, W_0), W_1) ...
 \end{equation}

### 2024-06-17 Paper.tex
eip-6049-alexprut-review
Reorder tuple elements

--- 2024-06-10 Paper.tex
@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{o}, \mathbf{s})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the returndata cache,and the stack contents.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s}, \mathbf{o})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the stack contents, and the returndata cache.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.

### 2024-06-26 Paper.tex
Fix returndata details (#909)
* Fix returndata details

* Replace 'cache' with 'buffer'

* Reorder tuple elements

* Rename 'cache' to 'buffer', revert some changes

@@
 \subsubsection{Machine State}
-The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s}, \mathbf{o})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the stack contents, and the returndata cache.
+The machine state $\boldsymbol{\mu}$ is defined as the tuple $(g, pc, \mathbf{m}, i, \mathbf{s}, \mathbf{o})$ which are the gas available, the program counter $pc \in \mathbb{N}_{256}$ , the memory contents, the active number of words in memory (counting continuously from position 0), the stack contents, and the returndata buffer.
 The memory contents $\boldsymbol{\mu}_{\mathbf{m}}$ are a series of zeroes of size $2^{256}$.
@@
 0x3e & {\small RETURNDATACOPY} & 3 & 0 & Copy output data from the previous call to memory. \\
-&&&& reverts if $\quad \boldsymbol{\mu}_{\mathbf{s}}[1] + \boldsymbol{\mu}_{\mathbf{s}}[2] > \lVert\boldsymbol{\mu}_{\mathbf{o}}\rVert$\\
 &&&& $\forall i \in \{ 0 \dots \boldsymbol{\mu}_{\mathbf{s}}[2] - 1\}: \boldsymbol{\mu}'_{\mathbf{m}}[\boldsymbol{\mu}_{\mathbf{s}}[0] + i ] \equiv
-\boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i]$\\
+\begin{cases} \boldsymbol{\mu}_{\mathbf{o}}[\boldsymbol{\mu}_{\mathbf{s}}[1] + i] & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[1] + i < \lVert \boldsymbol{\mu}_{\mathbf{o}} \rVert \\ 0 & \text{otherwise} \end{cases}$\\
 &&&& The additions in $\boldsymbol{\mu}_{\mathbf{s}}[1] + i$ are not subject to the $2^{256}$ modulo. \\

### 2024-07-01 Paper.tex

--- 2024-06-14 Paper.tex
@@
 For example the \textit{Shanghai} hard fork was scheduled to occur at 2023-04-12 10:27:35 UTC, on Epoch 194,048.
-Validators failed to propose a block during the two first slots of this Epoch, but at slot 6,209,538 a validator finally proposed a the block 17,034,870, marking the transition to \textit{Shanghai} on the execution layer.
+Validators failed to propose a block during the two first slots of this Epoch, but at slot 6,209,538 a validator finally proposed the block 17,034,870, marking the transition to \textit{Shanghai} on the execution layer.
 
@@
 &&&& since the \textit{Shanghai} update, this opcode has been marked as deprecated. \\
-&&&& Its behavior might be subject to change in an upcoming update. \\
+&&&& Its behavior might be subject to breaking change in an upcoming update. \\
 &&&& $A'_{\mathbf{s}} \equiv A_{\mathbf{s}} \cup \{ I_{\mathrm{a}} \}$ \\

### 2024-07-22 Paper.tex
typo

@@
 \item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
-\item[recipient] the 20-bytes address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
+\item[recipient] the 20-byte address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
 \item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
@@
 
-Here, we assume all components are interpreted by the RLP as integer values except for $W_\mathbf{r}$ which is a 20-bytes address:
+Here, we assume all components are interpreted by the RLP as integer values except for $W_\mathbf{r}$ which is a 20-byte address:
 \begin{equation}

### 2024-07-22 Paper.tex
Minor Typos

@@
 by attempting to pass it as the input range to a CALL) does not require memory to be extended to the beginning of the range.
-$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special-cases are given where these two are not equal.
+$\boldsymbol{\mu}'_{\mathrm{i}}$ is defined as this new maximum number of words of active memory; special cases are given where these two are not equal.
 

### 2024-08-25 Paper.tex
Target address should be kept warm even if CALL is aborted
transient storage

@@
 Transactors using type 2 transactions can specify the maximum priority fee they are willing to pay (\textbf{maxPriorityFeePerGas}) as well as the max total fee they are willing to pay (\textbf{maxFeePerGas}), inclusive of both the priority fee and the base fee.
-\textbf{maxFeePerGas} must at least as high as the base fee for the transaction to be included in a block.
+\textbf{maxFeePerGas} must be at least as high as the base fee for the transaction to be included in a block.
 Type 0 and type 1 transactions have only one field for specifiying a gas price--\textbf{gasPrice}--which also must be at least as high as the base fee for inclusion in a block.
@@
 
+\subsection{Transient Storage}
+Introduced by EIP-1153 (\cite{EIP-1153}) during the Cancun update, the transient storage, formally $\boldsymbol{\theta}$, is a volatile version of the world state account storage $\boldsymbol{\sigma}[a]_s$.
+It's a mapping between addresses (160-bit identifiers) and storage (another mapping between 256-bit integer values).
+
+\begin{equation}
+\forall a, \forall k: \boldsymbol{\theta}[a][k] = v
+\end{equation}
+where:
+\begin{equation}
+a \in \mathbb{B}_{160} \quad \wedge \quad k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
+\end{equation}
+
+At the beginning of each transaction, a new empty transient storage, $\boldsymbol{\theta}^0$, is created.
+
+\begin{equation}
+\boldsymbol{\theta}^0 \equiv \varnothing
+\end{equation}
+
+The transient storage is then destroyed at the end of the transaction, making it a volatile storage alternative for data that doesn't need to be persisted after the transaction ends.
+Since the transient storage is volatile, it doesn't need to be serialized into a Merkle Patricia tree, nor to be saved to disk, making it cheaper to use than the regular storage $\boldsymbol{\sigma}[a]_s$.
+
 \subsection{Execution}
@@
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if it is a message call.
 $\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
@@
   n \equiv \begin{cases}
-    \lVert T_{\mathrm{i}} \rVert & \text{if} \; T_{\mathrm{t}} \neq \varnothing \\
+    \lVert T_{\mathrm{i}} \rVert & \text{if} \; T_{\mathrm{t}} = \varnothing \\
     0 & \text{otherwise}
@@
 (\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; p, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
-\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, p, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
+\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, \boldsymbol{\theta}_0, A^*, S(T), S(T), g, &\\ \quad\; p, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
+\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, \boldsymbol{\theta}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, p, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
 \end{cases}
@@
 
-Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.
+Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the finals represents the message-call's output value (a byte array), and the new transient storage and are unused in the context of transaction evaluation.
 
@@
 
-We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$ and the accrued substate $A$, to the tuple containing the new state, remaining gas, new accrued substate, status code and output $(\boldsymbol{\sigma}', g', A', z, \mathbf{o})$:
-\begin{equation}
-(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \Lambda(\boldsymbol{\sigma}, A, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
+We define the creation function formally as the function \linkdest{lambda}{$\Lambda$}, which evaluates from these values, together with the state $\boldsymbol{\sigma}$, the transient storage $\boldsymbol{\theta}$ and the accrued substate $A$, to the tuple containing the new state, remaining gas, new accrued substate, status code, output and new transient storage $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}, \boldsymbol{\theta}')$:
+\begin{equation}
+(\boldsymbol{\sigma}', g', A', z, \mathbf{o}, \boldsymbol{\theta}') \equiv \Lambda(\boldsymbol{\sigma}, \boldsymbol{\theta}, A, s, o, g, p, v, \mathbf{i}, e, \zeta, w)
 \end{equation}
@@
 Finally, the account is initialised through the execution of the initialising EVM code $\mathbf{i}$ according to the execution model (see section \ref{ch:model}).
-Code execution can effect several events that are not internal to the execution state: the account's storage can be altered, further accounts can be created and further message calls can be made.
-As such, the code execution function $\hyperlink{xi_def}{\Xi}$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, available gas remaining $g^{**}$, the resultant accrued substate $A^{**}$ and the body code of the account $\mathbf{o}$.
-
-\begin{equation}
-(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, g, A^*, I) \\
+Code execution can effect several events that are not internal to the execution state: the account's storage and transient storage can be altered, further accounts can be created and further message calls can be made.
+As such, the code execution function $\hyperlink{xi_def}{\Xi}$ evaluates to a tuple of the resultant state $\boldsymbol{\sigma}^{**}$, resultant transient storage $\boldsymbol{\theta}^{**}$, available gas remaining $g^{**}$, the resultant accrued substate $A^{**}$ and the body code of the account $\mathbf{o}$.
+
+\begin{equation}
+(\boldsymbol{\sigma}^{**}, \boldsymbol{\theta}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, \boldsymbol{\theta}^*, g, A^*, I) \\
 \end{equation}
@@
 Code execution depletes gas, and gas may not go below zero, thus execution may exit before the code has come to a natural halting state.
-In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state, effectively leaving it as it was immediately prior to attempting the creation.
+In this (and several other) exceptional cases we say an out-of-gas (OOG) exception has occurred: The evaluated state is defined as being the empty set, $\varnothing$, and the entire create operation should have no effect on the state and transient storage, effectively leaving both of them as they were immediately prior to attempting the creation.
 
@@
 If such an exception does not occur, then the remaining gas is refunded to the originator and the now-altered state is allowed to persist.
-Thus formally, we may specify the resultant state, gas, accrued substate and status code as $(\boldsymbol{\sigma}', g', A', z)$ where:
+Thus formally, we may specify the resultant state, resultant transient storage, gas, accrued substate and status code as $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', z)$ where:
 
@@
 \quad\boldsymbol{\sigma}'[a]_{\mathrm{c}} = \texttt{KEC}(\mathbf{o}) & \text{otherwise}
+\end{cases} \\
+\quad \boldsymbol{\theta}' \equiv & \begin{cases}
+\boldsymbol{\theta} & \text{if} \quad F \ \lor\ \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\theta}^{**} & \text{otherwise} \\
 \end{cases} \\
@@
 
-Aside from evaluating to a new state and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array~$\mathbf{o}$.
+Aside from evaluating to a new state, new transient storage and accrued transaction substate, message calls also have an extra component---the output data denoted by the byte array~$\mathbf{o}$.
 This is ignored when executing transactions, however message calls can be initiated due to VM-code execution and in this case this information is used.
 \begin{equation}
-(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, A, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
+(\boldsymbol{\sigma}', g', A', z, \mathbf{o}, \boldsymbol{\theta}') \equiv \linkdest{theta}{\Theta}(\boldsymbol{\sigma}, \boldsymbol{\theta}, A, s, o, r, c, g, p, v, \tilde{v}, \mathbf{d}, e, w)
 \end{equation}
@@
 Just as with contract creation, if the execution halts in an exceptional fashion (i.e.
-due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state is reverted to the point immediately prior to balance transfer (i.e.
-$\boldsymbol{\sigma}$).
+due to an exhausted gas supply, stack underflow, invalid jump destination or invalid instruction), then no gas is refunded to the caller and the state and transient storage are reverted to the point immediately prior to balance transfer (i.e.
+$\boldsymbol{\sigma}$, $\boldsymbol{\theta}$).
 
@@
 \boldsymbol{\sigma}^{**} & \text{otherwise}
+\end{cases} \\
+\boldsymbol{\theta}' & \equiv & \begin{cases}
+\boldsymbol{\theta} & \text{if} \quad \boldsymbol{\sigma}^{**} = \varnothing \\
+\boldsymbol{\theta}^{**} & \text{otherwise}
 \end{cases} \\
@@
 1 & \text{otherwise}
-\end{cases} \\
+\end{cases}
+\end{eqnarray}
+\begin{equation}
 \hypertarget{code_execution_result}{}
-(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}) & \equiv & \Xi\\
+(\boldsymbol{\sigma}^{**}, \boldsymbol{\theta}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi
+\end{equation}
+\begin{eqnarray}
 I_{\mathrm{a}} & \equiv & r \\
@@
 \Xi \equiv \begin{cases}
-\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 1 \\
-\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 2 \\
-\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 3 \\
-\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 4 \\
-\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 5 \\
-\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 6 \\
-\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 7 \\
-\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 8 \\
-\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, g, A, I) & \text{if} \quad c = 9 \\
-\Xi                     (\boldsymbol{\sigma}_1, g, A, I) & \text{otherwise} \end{cases}
+\Xi_{\mathtt{ECREC}}    (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 1 \\
+\Xi_{\mathtt{SHA256}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 2 \\
+\Xi_{\mathtt{RIP160}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 3 \\
+\Xi_{\mathtt{ID}}       (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 4 \\
+\Xi_{\mathtt{EXPMOD}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 5 \\
+\Xi_{\mathtt{BN\_ADD}}  (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 6 \\
+\Xi_{\mathtt{BN\_MUL}}  (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 7 \\
+\Xi_{\mathtt{SNARKV}}   (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 8 \\
+\Xi_{\mathtt{BLAKE2\_F}}(\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{if} \quad c = 9 \\
+\Xi                     (\boldsymbol{\sigma}_1, \boldsymbol{\theta}, g, A, I) & \text{otherwise} \end{cases}
 \end{equation}
@@
 
-In addition to the system state $\boldsymbol{\sigma}$, the remaining gas for computation $g$, and the accrued substate $A$,
+In addition to the system state $\boldsymbol{\sigma}$, the transient storage $\boldsymbol{\theta}$, the remaining gas for computation $g$, and the accrued substate $A$,
 there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:
@@
 
-The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the remaining gas $g'$, the resultant accrued substate $A'$ and the resultant output, $\mathbf{o}$, given these definitions.
+The execution model defines the function $\Xi$, which can compute the resultant state $\boldsymbol{\sigma}'$, the resultant transient storage $\boldsymbol{\theta}'$, the remaining gas $g'$, the resultant accrued substate $A'$ and the resultant output, $\mathbf{o}$, given these definitions.
 For the present context, we will define it as:
 \begin{equation}
-(\boldsymbol{\sigma}', g', A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, g, A, I)
+(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I)
 \end{equation}
@@
 We must now define the $\Xi$ function.
-In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$ and the machine state, $\boldsymbol{\mu}$.
+In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\boldsymbol{\sigma}$, transient storage, $\boldsymbol{\theta}$, and the machine state, $\boldsymbol{\mu}$.
 Formally, we define it recursively with a function $X$.
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) \\
+\Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\theta}', \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
+(\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) \\
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
@@
 \begin{equation} \label{eq:X-def}
-X\big( (\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\mu}, A, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
-\big(\varnothing, \boldsymbol{\mu}', A, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
-X\big(O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
+X\big( (\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
+\big(\varnothing, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
+\big(\varnothing, \boldsymbol{\theta}, \boldsymbol{\mu}', A, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
+X\big(O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \text{otherwise} \\
 \end{cases}
@@
 \begin{array}[t]{l}
-w \in \{\text{\small CREATE}, \text{\small CREATE2}, \text{\small SSTORE},\\ \text{\small SELFDESTRUCT}\} \ \vee \\
+w \in \{\text{\small CREATE}, \text{\small CREATE2}, \text{\small SSTORE},\\ \text{\small TSTORE}, \text{\small SELFDESTRUCT}\} \ \vee \\
 \text{\small LOG0} \le w \; \wedge \; w \le \text{\small LOG4} \quad \vee \\
@@
 \begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \\
+O\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}', A', I) \\
 \Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
@@
 
-In general, we assume the memory, accrued substate and system state do not change:
+In general, we assume the memory, accrued substate, system state and transient strorage do not change:
 \begin{eqnarray}
@@
 A' & \equiv & A \\
-\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma}
+\boldsymbol{\sigma}' & \equiv & \boldsymbol{\sigma} \\
+\boldsymbol{\theta}' & \equiv & \boldsymbol{\theta}
 \end{eqnarray}
@@
 \begin{equation} \label{eq:pre}
-\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, g, A, I) \equiv \begin{cases}
-(\varnothing, 0, A, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, g - g_{\mathrm{r}}, A, \mathbf{o}) & \text{otherwise}\end{cases}
+\Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) \equiv \begin{cases}
+(\varnothing, \boldsymbol{\theta}, 0, A, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, \boldsymbol{\theta}, g - g_{\mathrm{r}}, A, \mathbf{o}) & \text{otherwise}\end{cases}
 \end{equation}
@@
 \end{eqnarray}
-and $\hyperlink{T__w}{T_{\mathrm{w}}}$ of legacy transcations is either $27 + T_{\mathrm{y}}$ or $2\hyperlink{chain_id}{\beta} + 35 + T_{\mathrm{y}}$.
+and $\hyperlink{T__w}{T_{\mathrm{w}}}$ of legacy transactions is either $27 + T_{\mathrm{y}}$ or $2\hyperlink{chain_id}{\beta} + 35 + T_{\mathrm{y}}$.
 
@@
 $G_{\mathrm{high}}$ & 10 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{high}}$}. \\
+$G_{\mathrm{transientaccess}}$ & 100 & Amount of gas to pay for operations of the set {\small $W_{\mathrm{transient}}$}. \\
 $G_{\mathrm{warmaccess}}$ & 100 & Cost of a warm account or storage access. \\
@@
 G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+G_{\mathrm{transientaccess}} & \text{if} \quad w \in W_{\mathrm{transient}}\\
 \end{cases}
@@
 
+$W_{\mathrm{transient}}$ = \{{\small TLOAD}, {\small TSTORE}\}
+
 Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_{\mathrm{i}}$ in order that all accesses reference valid memory whether for read or write.
@@
 As previously specified in section \ref{ch:model}, these definitions take place in the final context there.
-In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\mu}')$ such that:
-\begin{equation}
-O(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I) \quad \text{with exceptions, as noted}
+In particular we assume $O$ is the EVM state-progression function and define the terms pertaining to the next cycle's state $(\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}')$ such that:
+\begin{equation}
+O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \equiv (\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}', A', I) \quad \text{with exceptions, as noted}
 \end{equation}
@@
 \toprule
-\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
+\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage, Transient Storage and Flow Operations}} \vspace{5pt} \\
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
@@
 &&&& This operation has no effect on machine state during execution. \\
+\midrule
+0x5c & {\small TLOAD} & 1 & 1 & Load word from transient storage. \\
+&&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv \boldsymbol{\theta}[I_{\mathrm{a}}]_{\mathbf{s}}[\boldsymbol{\mu}_{\mathbf{s}}[0]]$ \\
+\midrule
+0x5d & {\small TSTORE} & 2 & 0 & Save word to transient storage. \\
+&&&& $\boldsymbol{\theta}'[I_{\mathrm{a}}]_{\mathbf{s}}[ \boldsymbol{\mu}_{\mathbf{s}}[0] ] \equiv \boldsymbol{\mu}_{\mathbf{s}}[1] $ \\
 \bottomrule
@@
 &&&& $\hyperlink{salt}{\zeta} \equiv \varnothing$ \\
-&&&& $(\boldsymbol{\sigma}', g', A', z, \mathbf{o}) \equiv \begin{cases}
-\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \wedge \\ \quad &\; I_{\mathrm{e}} < 1024 \wedge \lVert \mathbf{i} \rVert \leqslant 49152\\
-\big(\boldsymbol{\sigma}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', z, \mathbf{o}) \equiv \begin{cases}
+\hyperlink{lambda}{\Lambda}(\boldsymbol{\sigma}^*, \boldsymbol{\theta}, A, I_{\mathrm{a}}, I_{\mathrm{o}}, L(\boldsymbol{\mu}_{\mathrm{g}}), I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[0], \mathbf{i}, I_{\mathrm{e}} + 1, \zeta, I_{\mathrm{w}}) & \text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[0] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \; \wedge \\ \quad &\; I_{\mathrm{e}} < 1024 \wedge \lVert \mathbf{i} \rVert \leqslant 49152\\
+\big(\boldsymbol{\sigma}, \boldsymbol{\theta}, L(\boldsymbol{\mu}_{\mathrm{g}}), A, 0, () \big) & \text{otherwise} \end{cases}$ \\
 &&&& $\boldsymbol{\sigma}^* \equiv \boldsymbol{\sigma} \quad \text{except} \quad \boldsymbol{\sigma}^*[I_{\mathrm{a}}]_{\mathrm{n}} = \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{n}} + 1$ \\
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}
-\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', x, \mathbf{o}) \equiv \begin{cases}
+\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, \boldsymbol{\theta}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, \boldsymbol{\theta}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, \boldsymbol{\theta}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, \boldsymbol{\theta}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, \boldsymbol{\theta}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
   & \text{if} \quad I_{\mathrm{e}} < 1024 \\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, \boldsymbol{\theta}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2024-08-29 Paper.tex
add missing A*
style: add a missing "be".
style: add a missing "it is".

--- 2024-07-22 Paper.tex
@@
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if it is a result of EVM-code.
 $\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
@@
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
   & \text{if} \quad I_{\mathrm{e}} < 1024 \\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2024-11-04 Paper.tex
refactor: clearer sentence
Fix transactions typo (#919)

@@
 where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
-$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if it is a result of EVM-code.
+$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if it is a message call.
 $\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.

### 2025-01-16 Paper.tex
🐞 fix: Formula for initcode limit (#920)
Co-authored-by: rahul <raxhvl@users.noreply.github.com>
fix: add missing punctuation
simpler definition

--- 2024-08-25 Paper.tex
@@
 \subsection{Transient Storage}
-Introduced by EIP-1153 (\cite{EIP-1153}) during the Cancun update, the transient storage, formally $\boldsymbol{\theta}$, is a volatile version of the world state account storage $\boldsymbol{\sigma}[a]_s$.
-It's a mapping between addresses (160-bit identifiers) and storage (another mapping between 256-bit integer values).
+Formally defined as $\boldsymbol{\theta}$, the transient storage (EIP-1153: \cite{EIP-1153}), is a volatile version of the world state account storage $\boldsymbol{\sigma}[a]_s$.
+It's a bi-dimensional mapping between addresses and storage (i.e. a mapping of mappings between addresses, storage keys, and storage values).
 

### 2025-02-05 Paper.tex
🐞 fix: Correct 'form' to 'from' in withdrawal description (#923)

--- 2024-11-04 Paper.tex
@@
 \item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
-\item[recipient] the 20-byte address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
+\item[recipient] the 20-byte address that will receives Ether from this withdrawal; formally $W_{\mathrm{r}}$.
 \item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get address of the current account. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\

### 2025-02-07 Paper.tex

@@
 \textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
-0x30 & {\small ADDRESS} & 0 & 1 & Get address of the current account. \\
+0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
 &&&& $\boldsymbol{\mu}'_{\mathbf{s}}[0] \equiv I_{\mathrm{a}}$ \\

### 2025-02-10 Paper.tex

--- 2025-02-05 Paper.tex
@@
 Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
+Around the 1990s, it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
 Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.

### 2025-02-21 Paper.tex

@@
 Early work on smart contracts has been done by \cite{szabo1997formalizing} and \cite{miller1997future}.
-Around the 1990s, it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
+Around the 1990s it became clear that algorithmic enforcement of agreements could become a significant force in human cooperation.
 Though no specific system was proposed to implement such a system, it was proposed that the future of law would be heavily affected by such systems.
@@
 \item[validatorIndex] index of consensus layer's validator this withdrawal corresponds to; formally $W_{\mathrm{v}}$.
-\item[recipient] the 20-byte address that will receives Ether from this withdrawal; formally $W_{\mathrm{r}}$.
+\item[recipient] the 20-byte address that will receives Ether form this withdrawal; formally $W_{\mathrm{r}}$.
 \item[amount] a nonzero amount of Ether denominated in Gwei ($10^9$ Wei); formally $W_{\mathrm{a}}$.
@@
 
+\subsection{Transient Storage}
+Formally defined as $\boldsymbol{\theta}$, the transient storage (EIP-1153: \cite{EIP-1153}), is a volatile version of the world state account storage $\boldsymbol{\sigma}[a]_s$.
+It's a bi-dimensional mapping between addresses and storage (i.e. a mapping of mappings between addresses, storage keys, and storage values).
+
+\begin{equation}
+\forall a, \forall k: \boldsymbol{\theta}[a][k] = v
+\end{equation}
+where:
+\begin{equation}
+a \in \mathbb{B}_{160} \quad \wedge \quad k \in \mathbb{B}_{32} \quad \wedge \quad v \in \mathbb{N}
+\end{equation}
+
+At the beginning of each transaction, a new empty transient storage, $\boldsymbol{\theta}^0$, is created.
+
+\begin{equation}
+\boldsymbol{\theta}^0 \equiv \varnothing
+\end{equation}
+
+The transient storage is then destroyed at the end of the transaction, making it a volatile storage alternative for data that doesn't need to be persisted after the transaction ends.
+Since the transient storage is volatile, it doesn't need to be serialized into a Merkle Patricia tree, nor to be saved to disk, making it cheaper to use than the regular storage $\boldsymbol{\sigma}[a]_s$.
+
 \subsection{Execution}
@@
 G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
+G_{\mathrm{warmaccess}} & \text{if} \quad w \in W_{\mathrm{transient}}\\
 \end{cases}
@@
 &&&& $\mathbf{i} \equiv \boldsymbol{\mu}_{\mathbf{m}}[ \boldsymbol{\mu}_{\mathbf{s}}[3] \dots (\boldsymbol{\mu}_{\mathbf{s}}[3] + \boldsymbol{\mu}_{\mathbf{s}}[4] - 1) ]$ \\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}
-\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', x, \mathbf{o}) \equiv \begin{cases}
+\begin{array}{l}\hyperlink{theta}{\Theta}(\boldsymbol{\sigma}, \boldsymbol{\theta}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, t, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A),\\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge \\ \quad\quad I_{\mathrm{e}} < 1024\end{array}\\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, \boldsymbol{\theta}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& $n \equiv \min(\{ \boldsymbol{\mu}_{\mathbf{s}}[6], \lVert \mathbf{o} \rVert\})$ \\
@@
 &&&& Exactly equivalent to {\small CALL} except: \\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, \boldsymbol{\theta}, A^*, I_{\mathrm{a}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\ \quad I_{\mathrm{p}}, \boldsymbol{\mu}_{\mathbf{s}}[2], \boldsymbol{\mu}_{\mathbf{s}}[2], \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
   & \begin{array}{l}\text{if} \quad \boldsymbol{\mu}_{\mathbf{s}}[2] \leqslant \boldsymbol{\sigma}[I_{\mathrm{a}}]_{\mathrm{b}} \;\wedge\\ \quad\quad{}I_{\mathrm{e}} < 1024\end{array} \\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, \boldsymbol{\theta}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the change in the fourth parameter to the call $\hyperlink{theta}{\Theta}$ from the 2nd stack value \\
@@
 &&&& $\boldsymbol{\mu}_{\mathbf{s}}[5]$. Otherwise it is equivalent to {\small CALL} except:\\
-&&&& $(\boldsymbol{\sigma}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
+&&&& $(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', x, \mathbf{o}) \equiv \begin{cases}\begin{array}{l}\Theta(\boldsymbol{\sigma}, \boldsymbol{\theta}, A^*, I_{\mathrm{s}}, I_{\mathrm{o}}, I_{\mathrm{a}}, t, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), \\\quad I_{\mathrm{p}}, 0, I_{\mathrm{v}}, \mathbf{i}, I_{\mathrm{e}} + 1, I_{\mathrm{w}})\end{array}
   & \text{if} \quad I_{\mathrm{e}} < 1024 \\
-  (\boldsymbol{\sigma}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
+  (\boldsymbol{\sigma}, \boldsymbol{\theta}, C_{\text{\tiny CALLGAS}}(\boldsymbol{\sigma},\boldsymbol{\mu},A), A^*, 0, ()) & \text{otherwise} \end{cases}$ \\
 &&&& Note the changes (in addition to that of the fourth parameter) to the second \\

### 2025-02-25 Paper.tex
Xi theta return order

--- 2025-02-10 Paper.tex
@@
 \subsection{Subtleties}
-Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address should return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
+Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code\footnote{During initialization code execution, \texttt{EXTCODESIZE} on the address must return zero, which is the length of the code of the account while \texttt{CODESIZE} should return the length of the initialization code (as defined in \ref{subsec:instruction-set}).}.
 Thus any message call received by it during this time causes no code to be executed.

### 2025-02-26 Paper.tex
typo

--- 2025-02-21 Paper.tex
@@
 
-Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the finals represents the message-call's output value (a byte array), and the new transient storage and are unused in the context of transaction evaluation.
+Note we use $\hyperlink{theta}{\Theta}_{4}$ and $\hyperlink{lambda}{\Lambda}_{4}$ to denote the fact that only the first four components of the functions' values are taken; the final represents the message-call's output value (a byte array), and the new transient storage and are unused in the context of transaction evaluation.
 
@@
 \begin{equation}
-(\boldsymbol{\sigma}^{**}, \boldsymbol{\theta}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}^*, \boldsymbol{\theta}^*, g, A^*, I) \\
-\end{equation}
-\pagebreak[1]where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:\pagebreak[1]
+(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}, \boldsymbol{\theta}^{**}) \equiv \Xi(\boldsymbol{\sigma}^*, \boldsymbol{\theta}^*, g, A^*, I) \\
+\end{equation}
+
+where $I$ contains the parameters of the \hyperlink{exec_env}{execution environment}, that is:
 \begin{eqnarray}
@@
 \hypertarget{code_execution_result}{}
-(\boldsymbol{\sigma}^{**}, \boldsymbol{\theta}^{**}, g^{**}, A^{**}, \mathbf{o}) \equiv \Xi
+(\boldsymbol{\sigma}^{**}, g^{**}, A^{**}, \mathbf{o}, \boldsymbol{\theta}^{**}) \equiv \Xi
 \end{equation}
@@
 \begin{equation}
-(\boldsymbol{\sigma}', \boldsymbol{\theta}', g', A', \mathbf{o}) \equiv \Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I)
+(\boldsymbol{\sigma}', g', A', \mathbf{o}, \boldsymbol{\theta}') \equiv \Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I)
 \end{equation}
@@
 \begin{eqnarray}
-\Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\theta}', \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}) \\
-(\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}) & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) \\
+\Xi(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) & \equiv & (\boldsymbol{\sigma}'\!, \boldsymbol{\mu}'_{\mathrm{g}}, A', \mathbf{o}, \boldsymbol{\theta}') \\
+(\boldsymbol{\sigma}', \boldsymbol{\mu}'\!, A', ..., \mathbf{o}, \boldsymbol{\theta}') & \equiv & X\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) \\
 \boldsymbol{\mu}_{\mathrm{g}} & \equiv & g \\
@@
 X\big( (\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \big) \equiv \begin{cases}
-\big(\varnothing, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I, \varnothing\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
-\big(\varnothing, \boldsymbol{\theta}, \boldsymbol{\mu}', A, I, \mathbf{o}\big) & \text{if} \quad w = \text{\small REVERT} \\
+\big(\varnothing, \boldsymbol{\mu}, A, I, \varnothing, \boldsymbol{\theta}\big) & \text{if} \quad Z(\boldsymbol{\sigma}, \boldsymbol{\mu}, A, I) \\
+\big(\varnothing, \boldsymbol{\mu}', A, I, \mathbf{o}, \boldsymbol{\theta}\big) & \text{if} \quad w = \text{\small REVERT} \\
 O(\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I) \cdot \mathbf{o} & \text{if} \quad \mathbf{o} \neq \varnothing \\
@@
 \begin{eqnarray}
-O\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\theta}', \boldsymbol{\mu}', A', I) \\
+O\big((\boldsymbol{\sigma}, \boldsymbol{\theta}, \boldsymbol{\mu}, A, I)\big) & \equiv & (\boldsymbol{\sigma}', \boldsymbol{\mu}', A', I, \boldsymbol{\theta}') \\
 \Delta & \equiv & \mathbf{\alpha}_{w} - \mathbf{\delta}_{w} \\
@@
 \Xi_{\mathtt{PRE}}(\boldsymbol{\sigma}, \boldsymbol{\theta}, g, A, I) \equiv \begin{cases}
-(\varnothing, \boldsymbol{\theta}, 0, A, ()) & \text{if} \quad g < g_{\mathrm{r}} \\
-(\boldsymbol\sigma, \boldsymbol{\theta}, g - g_{\mathrm{r}}, A, \mathbf{o}) & \text{otherwise}\end{cases}
+(\varnothing, 0, A, (), \boldsymbol{\theta}) & \text{if} \quad g < g_{\mathrm{r}} \\
+(\boldsymbol\sigma, g - g_{\mathrm{r}}, A, \mathbf{o}, \boldsymbol{\theta}) & \text{otherwise}\end{cases}
 \end{equation}
